### 2020-12-19 paper/abstract.tex
Add initial version of the paper.

\begin{abstract}

Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases with minimal social and economic costs. Epidemiological models have a hard time predicting the effects of such fine grained policies. We propose a novel simulation-based model to address this shortcoming. We build on state-of-the-art agent-based simulation models but replace the way contacts between susceptible and infected people take place. Firstly, we allow for heterogeneity in the types of contacts (e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we strictly separate the number of contacts from the probabilities that a contact leads to an infection. The number of contacts changes with social distancing policies, the infection probabilities remain invariant. This allows us to model many types of fine grained policies that cannot easily be incorporated into other models. To validate our model, we show that it can accurately predict the effect of the German November lockdown even if no similar policy has been observed in the time series that were used to estimate the model parameters.

\vspace{1cm}
JEL Classification: C63, I18

Keywords: Covid-19, agent based simulation model, public health measures

\end{abstract}

### 2020-12-19 paper/calibration_and_estimation.tex
Add initial version of the paper.

\section{Calibration and Estimation}
\label{sec:calibration}

The model is described by a large number of parameters that govern the number of contacts a person has, the likelihood of becoming infected on each contact, the likelihood of developing light or strong symptoms or even dying from the disease as well as the duration each stage of the disease takes.

Most of these parameters can be calibrated from existing datasets or the medical literature. Only the infection probabilities have to be estimated inside the model by fitting it to time series data of case numbers and fatality rates.

\subsection{Medical Parameters}

\subsubsection{Length of Presymptomatic Stage / Incubation Period}

Estimates of the incubation period usually give a range from 2 to 12 days. A meta analysis by \cite{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modelled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95 \% CIs) of 1.63 (95 \% CI 1.51 to 1.75) and 0.50 (95 \% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.

\subsubsection{Begin of Infectiousness}

The period between infection and onset of infectiousness is called latent or latency period. However, the latency period is rarely given in epidemiological reports on Covid-19. Instead, scientists and agencies usually report the incubation period, the period from infection to the onset of symptoms. A few studies used measurements of virus shedding to estimate infectiousness during the course of the disease. When measurements started before the onset of symptoms the development of the viral load before symptoms gives us an indication of number of days between the onset of infectiousness and symptoms.

The European Centre for Disease Prevention and Control estimates that people become infectious between one and two days before the symptoms set in. This is similar to \cite{He2020} who estimate this to take 2.3 days and is in line with \cite{Peak2020}.

Given these numbers and the length of the incubation period we can calculate the latency period for symptomatic people. To our knowledge no estimates for the latency period of asymptomatic cases of COVID-19 exist. We assume it to be the same for symptomatic and asymptomatic cases.

Thus, we arrive at the following distribution for latency periods: 40\% have one day. 35\% have two days. 20\% have three days and 5\% have 5 days.

\subsection{Duration of Infectiousness}

We assume that the duration of infectiousness is the same for both symptomatic and asymptomatic individuals as evidence suggests little differences in the transmission rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients (\cite{Yin2020}) and that the viral load between symptomatic and asymptomatic individuals are similar (\cite{Zou2020}, \cite{Byrne2020}, \cite{Singanayagam2020}).

Our distribution of the duration of infectiousness is based on \cite{Byrne2020}.

For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8 days of infectiousness afterwards.\footnote{Viral loads may be detected much later but 8 days seems to be the time after which most people are culture negative, as also reported by \cite{Singanayagam2020}} Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic (see also figure 5). This duration range is very much in line with the meta-analysis’ reported evidence for asymptomatic individuals (see their figure 1). Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic. This duration range is very much in line with the meta-analysis' reported evidence for asymptomatic individuals.

Following this evidence we assume the following discretized distribution of the infectiousness period: 10\% of individuals are infectious for three days, 25\% for five days, another 25\% for seven days, 20\% for nine days and 20\% for eleven days.

\subsection{Duration of Symptoms}

We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3, Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring symptomatic cases.

We collapse the data to the following distribution: 10\% recover after 15 days and 30\% require 18, 22 or 27 days respectively.

% These long symptom durations align with reports by \cite{Tenforde2020}.

These numbers are only used for mild cases. We do not disaggregate by age. Note that the length of symptoms is not very important in our model given that individuals stop being infectious before their symptoms cease.

\subsubsection{Time from Symptom Onset to Admission to ICU}

The data on how many percent of symptomatic patients will require ICU is pretty thin. We rely on data by the US CDC (\cite{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/master/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project (2020-09-14)}.

From these sources we arrive at the following probabilities of requiring intensive care:

\begin{table}[!ht]
    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
    \label{tab:symptomatic-to-ICU}
    \centering

    \begin{tabular}{ll}
        \toprule
        Age Group & Share \\
        \midrule
        0-9 & 0.00005 \\
        10-19 & 0.00030 \\
        20-29 & 0.00075 \\
        30-39 & 0.00345 \\
        40-49 & 0.01380 \\
        50-59 & 0.03404 \\
        60-69 & 0.10138 \\
        70-79 & 0.16891 \\
        80-100 & 0.26871 \\
        \bottomrule
    \end{tabular}

\end{table}

For those who will require intensive care we follow \cite{Chen2020} who estimate the time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.

This aligns well with numbers reported for the time from first symptoms to hospitalization: \cite{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also in line with the durations collected by \href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the Robert Koch Institut}.

We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with equal probabilities. These times mostly matter for the ICU capacities.

\subsubsection{Death and Recovery from ICU}

We take the survival probabilities and time to death and time until recovery from intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.

They report time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. Approximating this with the normal distribution, we have nearly 10\% probability mass below 0. We use it nevertheless as several other distributions (such as chi squared and uniform) were unable to match the variance.
Discretizing this leads to 41\% of individuals who die from Covid-19 to die after one day in intensive care. 22\% day after 12 days, 29\% after 20 days and 7\% after 32 days. Again, we rescale this for every age group among those that will not survive.

They report time until recovery to have a mean of 18.8 days and a standard deviation of 12.21 days. Approximating this with the normal distribution, we have over 5\% probability mass below 0. Discretizing this of those who recover in intensive care 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

\subsection{Number of Contacts}

% Mossong data
We calibrate the parameters for the predicted numbers of contacts from contact diaries of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg \citep{Mossong2008}. Each contact diary contains all contacts an individual had throughout one day, including information on the other person (such as age and gender) and information on the contact. Importantly, for each contact individuals entered of which type the contact (school, leisure, work etc.) was and how frequent the contact with the other person is.

Thus, we can use the empirical distributions from this data as pre-pandemic number of contacts.

\subsection{Assortative Matching}

As mentioned in section \ref{sec:matching}, the probability that two individuals are matched can depend on background characteristics. In particular, we allow this probability to depend on age and county of residence. While we do not have good data on geographical assortativeness and just roughly calibrate it such that 60 \% of contacts are within the same county,\footnote{We are working on improving this estimation with mobility data.} we can calibrate the age assortativeness from the same data we use to calibrate the number of contacts.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/assortative_matching_probability_example.png}
    \caption{Distribution of random non-work contacts by age of participants}
    \label{fig:assortativeness}
\end{figure}

Figure~\ref{fig:assortativeness} shows that overall contacts within age groups or age groups that are closer to each other are more frequent.

\subsection{Infection Probabilities}
\label{sec:estimation}

To calibrate infection probabilities outside of the model, it would be important to know the exact duration and distance of each contact type as well as virus loads. Since this is not available in any dataset, we instead estimate those parameters inside the model by minimizing the distance between simulated and observed infection rates. Since our model includes a lot of randomness, we average simulated infection rates over several model runs.

Currently, we use data for Germany from August to October (both inclusive). We do not use earlier periods to save computational time. Moreover, we would be worried that the infection probabilities have small seasonal variation that we currently cannot model. However, we plan to expand the estimation period soon.

To avoid overfitting and simplify the numerical optimization problem, we only allow for four different probabilities: 1) for contacts in schools, preschools and nurseries. 2) for work contacts. 3) for households. 4) for leisure activities.

\section{Model Validation}

We validate our model in two ways: 1) We look at the in-sample fit over the estimation period. 2) We look at the out-of-sample fit for November. The last one is a challenging test for our model because there was a strong policy change between the estimation period and November. The model convincingly passes both tests.

\subsection{In-Sample Fit}

Despite fitting only four free parameters, the in-sample fit is very good. The best fit is achieved in the largest age groups. This is so mechanically, because we weight the deviations between simulated and observed infection rates by group sizes. The worst fit is achieved for the 80 to 100 years old. The model predicts too few infections for these groups because they have very few contacts in all contact types we have included so far. We plan to address this issue soon by adding another contact type that captures all contacts in the data by \cite{Mossong2008} that we have not included so far. Moreover, we expect an improved model fit when we allow for more free parameters.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/goodness_of_fit_by_age_group}
    \caption{Actual vs. simulated infection and fatality rates}
    \label{fig:goodness_of_fit}
\end{figure}

\subsection{Out-of-Sample Fit}

We can assess the out-of-sample fit by projecting the effect of the lockdown light and comparing it to case numbers until now. It is important to note that this is not just a simple extrapolation of a time trend because the lockdown light only started after the estimation period. The out-of-sample fit can be assessed in Figure~\ref{fig:out-of-sample-fit}.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/out_of_sample_validation}
    \caption{Projected effect of the lockdown-light}
    \label{fig:out-of-sample-fit}
\end{figure}

The model correctly predicts the effect of the lockdown light with reasonable accuracy. In particular, the actual case numbers are between our neutral and pessimistic projection. The plot also shows that ending the lockdown light as planned on November 30 would lead to an explosive growth in case numbers in all scenarios.

### 2020-12-19 paper/conclusion.tex
Add initial version of the paper.

\section{Conclusion}
\label{sec:conclusion}

We propose a simulation based model of infectious disease transmission that is designed to predict the effects of fine grained social distancing policies. In particular, the model can be used to model policies such as several ways of splitting school classes or work reduction policies that affect essential and non-essential workers differently. Both policies would be hard to implement in standard SEIR or agent based simulation models.

To predict the effects of such policies, it is not only important to have a way of expressing such flexible policies in terms of model quantities, but also to incorporate heterogeneity in disease progression as well as meeting patterns. We calibrate age dependent disease progression parameters from the medical literature and age dependent contact frequencies from contact diaries. Moreover, we distinguish eight types of contacts out of which some are only relevant for certain age groups.

The model has a good fit on past German case numbers and passes an out of sample validation despite a drastic change in the policy environment between the estimation period and the validation period.

Despite these encouraging results we still see the model as work in progress and plan to implement more features such as a detailed model of testing and contact tracing. Moreover, the estimation of the infection probabilities and the model fit will improve as more data becomes available.

We invite researchers from any discipline, but particularly epidemiologists to provide feedback on the model and welcome collaborations.

### 2020-12-19 paper/introduction.tex
Add initial version of the paper.

\section{Introduction}
\label{sec:introduction}

The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions across the world. As a result, many countries were able to reduce the number of positive cases and could relax restrictions imposed on their societies. At the same time, the social and economic costs were enormous. In the second wave, countries are trying to implement more nuanced policies. For example, Germany imposed a ``lockdown-light'' in November. Some businesses are closed and certain leisure activities are prohibited but schools and nurseries stay open.\blfootnote{Gabler and Röhrl are grateful for financial support by the German Research Foundation (DFG) through CRC-TR 224 (Projects C01 and A02, respectively). Gabler is grateful for funding by IZA Institute of Labor Economics.}

Current epidemiological models have not been designed to predict the effect of such fine-grained policies. These models need to be extended for each new policy proposal and cannot be easily adapted to fast changing environments. This report describes a model that has been designed from the ground up to predict the effects of contact reducing policies in real time. At the time of this writing, it has the following features:

\setstretch{1.0}
\begin{enumerate}
    \item At the core of the model, people meet people based on a matching algorithm. We distinguish various types of contacts. Currently, these are households, leisure activities, schools, nurseries and several types of contacts at the workplace. Contact types can be random or recurrent and vary in frequency.
    \item Policies can be implemented as shutting down contact types entirely or partially. The reduction of contacts can be random or systematic, e.g., to allow for essential workers.
    \item Infection probabilities vary across contact types, but are invariant to policies which reduce contacts.
    \item The model achieves a good fit on German infection and fatality rate data even if only the infection probabilities are fit to the data and the remaining parameters are calibrated from the medical literature and datasets on contact frequencies.
    \item High quality Python code for the model is freely available on \href{https://github.com/covid-19-impact-lab/sid}{https://github.com/covid-19-impact-lab/sid}, well documented and very flexible. We are actively looking for researchers who want to use our model for their projects.
\end{enumerate}
\setstretch{1.25}

After a brief literature review we describe our model in detail and validate its in-sample and out-of-sample fit on German data.

This report describes the model in an abstract way, but uses many realistic examples from a version that is specialized to Germany. It is important to note that this specialization is not baked into the model or the Python code. It is easy to adjust the model to other countries if data on the number of contacts and a dataset with background characteristics are available.

More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\cite{Dorn2020a}).

### 2020-12-19 paper/literature_review.tex
Add initial version of the paper.

\section{Literature Review}
\label{sec:literature_review}

We build on two strands of literature: Recent extensions of the epidemiological SEIR model and agent-based simulation models.

The traditional SEIR model is not fine-grained enough to model nuanced policies. This has motivated a large number of researchers to extend the standard model to allow for more heterogeneity and flexibility. Examples are \cite{Grimm2020}, \cite{Donsimoni2020} and \cite{Acemoglu2020} who develop multi group SEIR models to analyze the effects of targeted lockdowns and \cite{Berger2020} who extend the SEIR model to analyze testing and conditional quarantines. For a more comprehensive review see \cite{Avery2020}. Others have used the results of a standard SEIR model as input for economic models that estimate the cost of policies (e.g. \cite{Dorn2020}).

While the popularity of the SEIR model is mainly due to its simplicity, the extensions are quite complex. It is unlikely that there will be a SEIR model that combines all proposed extensions. Moreover, the extensions do not address other key issues: The main parameter of the SEIR model, the basic reproduction number ($R_0$), is not policy-invariant. It is a composite of the number of contacts each person has and the infection probability of the contacts. In fact, policy simulations are done by setting $R_0$ to a different value but it is hard to translate a real policy into the value of $R_0$ it will induce.

Another commonly used model class in epidemiology are agent-based simulation models. In these models individuals are simulated as moving particles. Infections take place when two particles come closer than a certain contact radius (e.g. \cite{Silva2020} and \cite{Cuevas2020}). While the simulation approach makes it easy to incorporate heterogeneity in disease progression, it is hard to incorporate heterogeneity in meeting patterns. Moreover, policies are modeled as changes in the contact radius or momentum equation of the particles. The translation from real policies to corresponding model parameters is a hard task.

\cite{Hinch2020} is a recent extension of the prototypical agent-based simulation model that replaces moving particles by contact networks for households, work and random contacts. This model is similar in spirit to ours but focuses on contact tracing rather than social distancing policies.

The above assessment of epidemiological models is not meant as a critique. We are aware that those models were not designed to predict the effect of fine grained social distancing policies in real time and are very well suited to their purpose. We invite epidemiologists to provide feedback and collaborate to improve our model.

### 2020-12-19 paper/model.tex
Add initial version of the paper.

\section{Model}
\label{sec:model}

\subsection{Summary}

To alleviate these problems we propose a different model structure. Our model inherits many features from agent-based simulation models but replaces the contacts between moving particles by contacts between individuals who work, go to school, live in a household and enjoy leisure activities. The structure of the model is depicted in Figure~\ref{fig:model_graph}.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/model_detailed.png}
    \caption{Simplified graph of the model}
    \label{fig:model_graph}
\end{figure}

The background characteristics include age, county and occupation of each simulated individual. Contact models are functions that map individual characteristics into a predicted number of contacts. Currently we distinguish between eight types of contact models which are all listed in Figure \ref{fig:model_graph}: households, recurrent and random work contacts, recurrent and random leisure contacts, and nursery, preschool, and school contacts.

The predicted number of contacts is translated into infections by a matching algorithm. There are different matching algorithms for recurrent contacts (e.g. classmates, family members) and non-recurrent contacts (e.g. clients, contacts in supermarkets). The infection probability can differ for each contact type. All types of contacts can be assortative with respect to geographic and demographic characteristics.

Once a person is infected, the disease progresses in a fairly standard way which is also depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild symptoms are infectious for some time and recover eventually. Cases with severe symptoms additionally require hospitalization and lead to either recovery or death.

People who experience symptoms as well as people who had a risk contact can demand a test. Test allocation models that take into account the testing capacity determine which individuals actually receive a test. Test processing models determine in which order the tests are processed and when infected individuals are notified.

In addition, people who have symptoms, received a positive test, or had a risk contact can reduce their number of contacts across all contact types endogenously.

The model makes it very simple to translate policies into model quantities. For example, school closures imply the complete suspension of school contacts. A strict lockdown implies shutting down work contacts of all people who are not employed in a systemically relevant sector. It is also possible to have more sophisticated policies that condition the number of contacts on observable characteristics, risk contacts or health states.

Another key advantage of the model is that the number of contacts an individual has of each contact type can be calibrated from publicly available data \citep{Mossong2008}. This in turn allows us to estimate policy-invariant infection probabilities from time series of infection and death rates using the method of simulated moments \citep{McFadden1989}. Since the infection probabilities are time-invariant, data collected since the beginning of the pandemic can be used for estimation. Moreover, since we can model the testing strategies that were in place at each point in time, we can correct the estimates for the fact that not all infections are observed.

In the following sections we describe each of the model components in more detail.

\subsection{Modeling Numbers of Contacts}
\label{sec:number_of_contacts}

Consider a hypothetical population of 1,000 individuals in which 50 were infected with a novel infectious disease. From this data alone, it is impossible to say whether only those 50 people had contact with an infectious person and the disease has an infection probability of 1 in each contact or whether everyone met an infectious person but the disease has an infection probability of only 5 percent per contact. SEIR models do not even try to distinguish contact frequency from the infectiousness of each contact and combine the two in one parameter that is not invariant to social distancing policies.

To model social distancing policies, we need to disentangle the effects of the number of contacts of each individual and the effect of policy-invariant infection probabilities specific to each contact type. Since not all contacts are equally infectious, we distinguish different contact types.

The number and type of contacts in our model can be easily extended. Each type of contacts is described by a function that maps individual characteristics, health states and the date into a number of planned contacts for each individual. This allows to model a wide range of contact types.

Currently, there are the following contact types:

\begin{itemize}
    \item Households: Each household member meets all other household members every day. The household sizes and structures are calibrated to be representative for Germany.
    \item Random non-work contacts: Each person has contacts with randomly drawn other people. This contact type reflects contacts during pure leisure activities as well as non leisure activities such as grocery shopping or medical appointments.
    \item Random work contacts: Each working adult has contact to randomly drawn other people.
    \item Recurrent daily work contacts: Each working adult meets other workers every day. This is meant to capture work colleagues.
    \item Recurrent weekly work contacts: Each working adult meets other workers once per week. We randomize over the days on which the meetings take place. This is meant to capture meetings with clients, superiors or other colleagues which happen infrequently.
    \item Schools: Each student meets all of his classmates every day. Class sizes are calibrated to be representative for Germany. Schools are closed on weekends and during vacations, which vary by states.
    \item Preschools: Similar to schools but for younger children.
    \item Nurseries: Similar to schools but for very young children.
\end{itemize}

The number of random and recurrent contacts at the workplace and at home is calibrated with data provided by \cite{Mossong2008}. For details see Section~\ref{sec:calibration}. In particular, we sample the number of contacts or group sizes from empirical distributions that sometimes depend on age. Instead it would also be possible to use economic or other behavioral models to predict the number of contacts.

Theoretically, each contact type can have its own infection probability. However, to reduce the number of free parameters and thus avoid a potential over-fitting we impose some constraints. For now, infection probabilities in schools, preschools and nurseries are equal. Moreover, we restrict all work contacts to have the same infection probability.

\subsection{Reducing Numbers of Contacts Through Policies}
\label{sec:policies}

The main motivation of our model is to predict the effect of policies that affect the number of contacts people have. Examples range from school closures and lockdowns to more nuanced policies such as a mandatory quarantines for symptomatic individuals or a class splitting policy where only half of the students come to school in person and the other half joins digitally with weekly rotation.

Instead of thinking of policies as completely replacing how many contacts people have, it is often more helpful to think of them as adjusting the pre-pandemic number of contacts.

Therefore, we implement policies as a step that happens after the number of contacts is calculated but before individuals are matched.

On an abstract level, a policy is a functions that modifies the number of contacts of one contact type. For example, school closures simply set all school contacts to zero. A lockdown where only essential workers are allowed to work means that approximately two thirds of the working population have zero work contacts and the rest has the same number of contacts as before.

% importance of fine-grained contact types and having the entire population
This, in conjunction with our fine grained contact types, allows us to easily implement a wide variety of policies. Allowing policies to depend on the health states of the entire population means that adaptive lockdowns where, for example, schools close when a certain threshold of infections is surpassed at the county level would be as simple as determining which counties are above the threshold and then setting all school contacts in these counties to zero.

The dependency of policies on health states also makes it possible to model contact tracing. For example, a policy could check whether each child has a classmate who's received a positive test result and then bar all children of that class from attending school.

% contact tracing and how including the states means the sky is the limit
Some policies can be easily implemented if the background characteristics are suitably extended. For example, a schooling policy with split classes, where each half attends school every other week can be implemented by storing whether the child would attend in even or odd weeks in the background characteristics and then using that information in the policy function.

For some policies the exact effect on each contact type is not easy to determine. If this refers to a policy during the estimation period, it is possible to estimate such parameters by fitting the model to time series data of infection rates. This is only possible if the policy was not active during the whole estimation period and thus the infection probabilities can be identified separately. If instead it refers to a policy that we want to simulate, we make a scenario analysis in which the model is simulated with several assumptions about how the policy affects the number of contacts.

\subsection{Endogenous Contact Reductions}
\label{sec:endogenous_contact_reductions}

Policies are not the only way in which the number of contacts are reduced compared to the pre-pandemic level. It is important to model those other channels. Otherwise, the effect of policies would be overestimated and policy recommendations based on the model would be biased.

Examples of endogenous contact reductions are manifold: symptomatic people stay at home; Members of risk groups try to reduce their number of contacts more strongly than others; People self-isolate if they know they had a risk contact.

Since we model the number of contacts as arbitrary functions of background characteristics and health states, it is easy to implement such considerations.

In our current empirical application we only model that symptomatic people reduce their number of contacts across all contact types (except for households) by 70 \%. Within households they reduce contacts by 50\%. We are working on extending this to allow for formal and informal contact tracing as well as quarantines after positive test results.

\subsection{Matching Individuals}
\label{sec:matching}

The empirical data described above only allows to estimate the number of contacts each person has. In order to simulate transmissions of Covid-19, the numbers of contacts has to be translated into actual meetings between people. This is achieved by matching algorithms:

As described in section \ref{sec:number_of_contacts}, some contact types are recurrent (i.e. the same people meet regularly), others are non-recurrent (i.e. it would only be by accident that two people meet twice). The matching process is different for recurrent and non recurrent contact models.

Recurrent contacts are described by two components: 1) A variable in the background characteristics. An example would be a school class identifier which could come from actual data or be drawn randomly to achieve representative class sizes. 2) A deterministic or random function that takes the value 0 (non-participating) and 1 (participating) and can depend on the weekday, date and health state. This can be used to model vacations, weekends or symptomatic people who stay home (see section \ref{sec:endogenous_contact_reductions} for details).

The matching process for recurrent contacts is then extremely simple: On each simulated day, every person who does not stay home meets all other group members who do not stay home. The assumption that all group members have contacts with all other group members is not fully realistic, but seems like a good approximation to reality, especially in light of the suspected role of aerosol transmission for Covid-19 (\cite{Morawska2020, Anderson2020}).

The matching in non-recurrent contact models is more difficult and implemented in a two stage sampling procedure to allow for assortative matching. Currently most contact models are assortative with respect to age (it is more likely to meet people from the same age group) and county (it is more likely to meet people from the same county) but in principle any set of discrete variables can be used. This set of variables that influence matching probabilities introduce a discrete partition of the population into groups. The first stage of the two stage sampling process samples on the group level. The second stage on the individual level.

Below, we first show pseudo code for the non-recurrent matching algorithm and then describe how the algorithm works in words.

%%% ToDo: Get syntax highlighting for the pseudo code. Try out different languages.

\begin{lstlisting}
while unmatched contacts remain:
    model, id = draw contact model and individual
    for contact in remaining_contacts[id, model]:
        draw group of other person
        draw other person from that group
        determine if id infects other or vice versa
        remaining_contacts[id, model] -= 1
        remaining_contacts[other, model] -= 1
\end{lstlisting}

We first randomly draw a contact type and individual. For each contact oth the drawn contact type that person has, we first draw the group of the other person (first stage). Next, we calculate the probability to be drawn for each member of the group, based on the number of remaining contacts. I.e. people who have more remaining contacts are drawn with a higher probability. This has to be re-calculated each time because with each matched contact, the number of remaining contacts changes. We then draw the other individual, determine whether an infection takes place and if so update the health states of the newly infected person. Finally, we reduce the number of remaining contacts of the two matched individuals by one.

The recalculation of matching probabilities in the second stage is computationally intensive because it requires summing up all remaining contacts in that group. Using a two stage sampling process where the first stage probabilities remain constant over time makes the matching computationally much more tractable because the number of computations increases quadratically in the second stage group size.

\FloatBarrier

\subsection{Course of the Disease}

The following medical parameters describing the progression of the disease are taken from systematic reviews (e.g. \cite{He2020}). After an infection occurs, the disease progresses in the way depicted in Figure~\ref{fig:course_of_disease}.

\begin{figure}[!ht]
    \includegraphics[width=0.7\textwidth]{../figures/disease_progression.png}
    \caption{Course of Disease in the model}
    \label{fig:course_of_disease}
\end{figure}

First, infected individuals will become infectious after one to five days. About one third of people stays asymptomatic. The rest develops symptoms about one to two days after they become infectious. Modeling asymptomatic and pre-symptomatic cases is important because those people do not reduce their contacts or demand a test and can potentially infect many other people \citep{Donsimoni2020}.

A small share of symptomatic people will develop strong symptoms that require intensive care. The exact share and time span is age-dependent. An age-dependent share of intensive care unit (ICU) patients will die after spending up to 32 days in intensive care. Moreover, if the ICU capacity was reached, all patients who require intensive care but do not receive it die.

It would be easy to make the course of disease even more fine grained. For example, we could model people who require hospitalization but not intensive care. So far we opted against that because only the ICU capacity might become a bottleneck in Germany.

We allow the progression of the disease to be stochastic in two ways: Firstly, state changes only occur with a certain probability (e.g. only a fraction of infected individuals develops symptoms). Secondly, the number of periods for which an individual remains in a state is drawn randomly. The parameters that govern these processes are taken from the literature. They can vary with the age of an individual.

Detailed information on the calibration of the disease parameters is available as part of our \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.

### 2020-12-19 paper/model.tex
Split up the section on the model in subfiles, added a file on testing.

@@
 
-\subsection{Summary}
+\input{model/summary}
 
-To alleviate these problems we propose a different model structure.
-Our model inherits many features from agent-based simulation models but replaces the contacts between moving particles by contacts between individuals who work, go to school, live in a household and enjoy leisure activities.
-The structure of the model is depicted in Figure~\ref{fig:model_graph}.
+\input{model/number_of_contacts}
 
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[width=0.7\textwidth]{../figures/model_detailed.png}
-    \caption{Simplified graph of the model}
-    \label{fig:model_graph}
-\end{figure}
+\input{model/policies}
 
-The background characteristics include age, county and occupation of each simulated individual.
-Contact models are functions that map individual characteristics into a predicted number of contacts.
-Currently we distinguish between eight types of contact models which are all listed in Figure \ref{fig:model_graph}: households, recurrent and random work contacts, recurrent and random leisure contacts, and nursery, preschool, and school contacts.
+\input{model/endogenous_contact_reductions}
 
-The predicted number of contacts is translated into infections by a matching algorithm.
-There are different matching algorithms for recurrent contacts (e.g.
-classmates, family members) and non-recurrent contacts (e.g.
-clients, contacts in supermarkets).
-The infection probability can differ for each contact type.
-All types of contacts can be assortative with respect to geographic and demographic characteristics.
-
-Once a person is infected, the disease progresses in a fairly standard way which is also depicted in Figure~\ref{fig:course_of_disease}.
-Asymptomatic cases and cases with mild symptoms are infectious for some time and recover eventually.
-Cases with severe symptoms additionally require hospitalization and lead to either recovery or death.
-
-People who experience symptoms as well as people who had a risk contact can demand a test.
-Test allocation models that take into account the testing capacity determine which individuals actually receive a test.
-Test processing models determine in which order the tests are processed and when infected individuals are notified.
-
-In addition, people who have symptoms, received a positive test, or had a risk contact can reduce their number of contacts across all contact types endogenously.
-
-The model makes it very simple to translate policies into model quantities.
-For example, school closures imply the complete suspension of school contacts.
-A strict lockdown implies shutting down work contacts of all people who are not employed in a systemically relevant sector.
-It is also possible to have more sophisticated policies that condition the number of contacts on observable characteristics, risk contacts or health states.
-
-Another key advantage of the model is that the number of contacts an individual has of each contact type can be calibrated from publicly available data \citep{Mossong2008}.
-This in turn allows us to estimate policy-invariant infection probabilities from time series of infection and death rates using the method of simulated moments \citep{McFadden1989}.
-Since the infection probabilities are time-invariant, data collected since the beginning of the pandemic can be used for estimation.
-Moreover, since we can model the testing strategies that were in place at each point in time, we can correct the estimates for the fact that not all infections are observed.
-
-In the following sections we describe each of the model components in more detail.
-
-\subsection{Modeling Numbers of Contacts}
-\label{sec:number_of_contacts}
-
-Consider a hypothetical population of 1,000 individuals in which 50 were infected with a novel infectious disease.
-From this data alone, it is impossible to say whether only those 50 people had contact with an infectious person and the disease has an infection probability of 1 in each contact or whether everyone met an infectious person but the disease has an infection probability of only 5 percent per contact.
-SEIR models do not even try to distinguish contact frequency from the infectiousness of each contact and combine the two in one parameter that is not invariant to social distancing policies.
-
-To model social distancing policies, we need to disentangle the effects of the number of contacts of each individual and the effect of policy-invariant infection probabilities specific to each contact type.
-Since not all contacts are equally infectious, we distinguish different contact types.
-
-The number and type of contacts in our model can be easily extended.
-Each type of contacts is described by a function that maps individual characteristics, health states and the date into a number of planned contacts for each individual.
-This allows to model a wide range of contact types.
-
-Currently, there are the following contact types:
-
-\begin{itemize}
-    \item Households: Each household member meets all other household members every day. The household sizes and structures are calibrated to be representative for Germany.
-    \item Random non-work contacts: Each person has contacts with randomly drawn other people.
-This contact type reflects contacts during pure leisure activities as well as non leisure activities such as grocery shopping or medical appointments.
-    \item Random work contacts: Each working adult has contact to randomly drawn other people.
-    \item Recurrent daily work contacts: Each working adult meets other workers every day. This is meant to capture work colleagues.
-    \item Recurrent weekly work contacts: Each working adult meets other workers once per week.
-We randomize over the days on which the meetings take place.
-This is meant to capture meetings with clients, superiors or other colleagues which happen infrequently.
-    \item Schools: Each student meets all of his classmates every day.
-Class sizes are calibrated to be representative for Germany.
-Schools are closed on weekends and during vacations, which vary by states.
-    \item Preschools: Similar to schools but for younger children.
-    \item Nurseries: Similar to schools but for very young children.
-\end{itemize}
-
-The number of random and recurrent contacts at the workplace and at home is calibrated with data provided by \cite{Mossong2008}.
-For details see Section~\ref{sec:calibration}.
-In particular, we sample the number of contacts or group sizes from empirical distributions that sometimes depend on age.
-Instead it would also be possible to use economic or other behavioral models to predict the number of contacts.
-
-Theoretically, each contact type can have its own infection probability.
-However, to reduce the number of free parameters and thus avoid a potential over-fitting we impose some constraints.
-For now, infection probabilities in schools, preschools and nurseries are equal.
-Moreover, we restrict all work contacts to have the same infection probability.
-
-\subsection{Reducing Numbers of Contacts Through Policies}
-\label{sec:policies}
-
-The main motivation of our model is to predict the effect of policies that affect the number of contacts people have.
-Examples range from school closures and lockdowns to more nuanced policies such as a mandatory quarantines for symptomatic individuals or a class splitting policy where only half of the students come to school in person and the other half joins digitally with weekly rotation.
-
-Instead of thinking of policies as completely replacing how many contacts people have, it is often more helpful to think of them as adjusting the pre-pandemic number of contacts.
-
-Therefore, we implement policies as a step that happens after the number of contacts is calculated but before individuals are matched.
-
-On an abstract level, a policy is a functions that modifies the number of contacts of one contact type.
-For example, school closures simply set all school contacts to zero.
-A lockdown where only essential workers are allowed to work means that approximately two thirds of the working population have zero work contacts and the rest has the same number of contacts as before.
-
-% importance of fine-grained contact types and having the entire population
-This, in conjunction with our fine grained contact types, allows us to easily implement a wide variety of policies.
-Allowing policies to depend on the health states of the entire population means that adaptive lockdowns where, for example, schools close when a certain threshold of infections is surpassed at the county level would be as simple as determining which counties are above the threshold and then setting all school contacts in these counties to zero.
-
-The dependency of policies on health states also makes it possible to model contact tracing.
-For example, a policy could check whether each child has a classmate who's received a positive test result and then bar all children of that class from attending school.
-
-% contact tracing and how including the states means the sky is the limit
-Some policies can be easily implemented if the background characteristics are suitably extended.
-For example, a schooling policy with split classes, where each half attends school every other week can be implemented by storing whether the child would attend in even or odd weeks in the background characteristics and then using that information in the policy function.
-
-For some policies the exact effect on each contact type is not easy to determine.
-If this refers to a policy during the estimation period, it is possible to estimate such parameters by fitting the model to time series data of infection rates.
-This is only possible if the policy was not active during the whole estimation period and thus the infection probabilities can be identified separately.
-If instead it refers to a policy that we want to simulate, we make a scenario analysis in which the model is simulated with several assumptions about how the policy affects the number of contacts.
-
-\subsection{Endogenous Contact Reductions}
-\label{sec:endogenous_contact_reductions}
-
-Policies are not the only way in which the number of contacts are reduced compared to the pre-pandemic level.
-It is important to model those other channels.
-Otherwise, the effect of policies would be overestimated and policy recommendations based on the model would be biased.
-
-Examples of endogenous contact reductions are manifold: symptomatic people stay at home; Members of risk groups try to reduce their number of contacts more strongly than others; People self-isolate if they know they had a risk contact.
-
-Since we model the number of contacts as arbitrary functions of background characteristics and health states, it is easy to implement such considerations.
-
-In our current empirical application we only model that symptomatic people reduce their number of contacts across all contact types (except for households) by 70 \%.
-Within households they reduce contacts by 50\%.
-We are working on extending this to allow for formal and informal contact tracing as well as quarantines after positive test results.
-
-\subsection{Matching Individuals}
-\label{sec:matching}
-
-The empirical data described above only allows to estimate the number of contacts each person has.
-In order to simulate transmissions of Covid-19, the numbers of contacts has to be translated into actual meetings between people.
-This is achieved by matching algorithms:
-
-As described in section \ref{sec:number_of_contacts}, some contact types are recurrent (i.e.
-the same people meet regularly), others are non-recurrent (i.e.
-it would only be by accident that two people meet twice).
-The matching process is different for recurrent and non recurrent contact models.
-
-Recurrent contacts are described by two components: 1) A variable in the background characteristics.
-An example would be a school class identifier which could come from actual data or be drawn randomly to achieve representative class sizes.
-2) A deterministic or random function that takes the value 0 (non-participating) and 1 (participating) and can depend on the weekday, date and health state.
-This can be used to model vacations, weekends or symptomatic people who stay home (see section \ref{sec:endogenous_contact_reductions} for details).
-
-The matching process for recurrent contacts is then extremely simple: On each simulated day, every person who does not stay home meets all other group members who do not stay home.
-The assumption that all group members have contacts with all other group members is not fully realistic, but seems like a good approximation to reality, especially in light of the suspected role of aerosol transmission for Covid-19 (\cite{Morawska2020, Anderson2020}).
-
-The matching in non-recurrent contact models is more difficult and implemented in a two stage sampling procedure to allow for assortative matching.
-Currently most contact models are assortative with respect to age (it is more likely to meet people from the same age group) and county (it is more likely to meet people from the same county) but in principle any set of discrete variables can be used.
-This set of variables that influence matching probabilities introduce a discrete partition of the population into groups.
-The first stage of the two stage sampling process samples on the group level.
-The second stage on the individual level.
-
-Below, we first show pseudo code for the non-recurrent matching algorithm and then describe how the algorithm works in words.
-
-%%% ToDo: Get syntax highlighting for the pseudo code. Try out different languages.
-
-\begin{lstlisting}
-while unmatched contacts remain:
-    model, id = draw contact model and individual
-    for contact in remaining_contacts[id, model]:
-        draw group of other person
-        draw other person from that group
-        determine if id infects other or vice versa
-        remaining_contacts[id, model] -= 1
-        remaining_contacts[other, model] -= 1
-\end{lstlisting}
-
-We first randomly draw a contact type and individual.
-For each contact oth the drawn contact type that person has, we first draw the group of the other person (first stage).
-Next, we calculate the probability to be drawn for each member of the group, based on the number of remaining contacts.
-I.e.
-people who have more remaining contacts are drawn with a higher probability.
-This has to be re-calculated each time because with each matched contact, the number of remaining contacts changes.
-We then draw the other individual, determine whether an infection takes place and if so update the health states of the newly infected person.
-Finally, we reduce the number of remaining contacts of the two matched individuals by one.
-
-The recalculation of matching probabilities in the second stage is computationally intensive because it requires summing up all remaining contacts in that group.
-Using a two stage sampling process where the first stage probabilities remain constant over time makes the matching computationally much more tractable because the number of computations increases quadratically in the second stage group size.
+\input{model/matching}
 
@@
 
-\subsection{Course of the Disease}
+\input{model/course_of_disease}
 
-The following medical parameters describing the progression of the disease are taken from systematic reviews (e.g.
-\cite{He2020}).
-After an infection occurs, the disease progresses in the way depicted in Figure~\ref{fig:course_of_disease}.
-
-\begin{figure}[!ht]
-    \includegraphics[width=0.7\textwidth]{../figures/disease_progression.png}
-    \caption{Course of Disease in the model}
-    \label{fig:course_of_disease}
-\end{figure}
-
-First, infected individuals will become infectious after one to five days.
-About one third of people stays asymptomatic.
-The rest develops symptoms about one to two days after they become infectious.
-Modeling asymptomatic and pre-symptomatic cases is important because those people do not reduce their contacts or demand a test and can potentially infect many other people \citep{Donsimoni2020}.
-
-A small share of symptomatic people will develop strong symptoms that require intensive care.
-The exact share and time span is age-dependent.
-An age-dependent share of intensive care unit (ICU) patients will die after spending up to 32 days in intensive care.
-Moreover, if the ICU capacity was reached, all patients who require intensive care but do not receive it die.
-
-It would be easy to make the course of disease even more fine grained.
-For example, we could model people who require hospitalization but not intensive care.
-So far we opted against that because only the ICU capacity might become a bottleneck in Germany.
-
-We allow the progression of the disease to be stochastic in two ways: Firstly, state changes only occur with a certain probability (e.g.
-only a fraction of infected individuals develops symptoms).
-Secondly, the number of periods for which an individual remains in a state is drawn randomly.
-The parameters that govern these processes are taken from the literature.
-They can vary with the age of an individual.
-
-Detailed information on the calibration of the disease parameters is available as part of our \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.
+\input{model/testing}

### 2020-12-19 paper/model/course_of_disease.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Course of the Disease}
\label{sub:course_of_disease}

The following medical parameters describing the progression of the disease are taken from systematic reviews (e.g. \cite{He2020}). After an infection occurs, the disease progresses in the way depicted in Figure~\ref{fig:course_of_disease}.

\begin{figure}[!ht]
    \includegraphics[width=0.7\textwidth]{../figures/disease_progression.png}
    \caption{Course of Disease in the model}
    \label{fig:course_of_disease}
\end{figure}

First, infected individuals will become infectious after one to five days. About one third of people stays asymptomatic. The rest develops symptoms about one to two days after they become infectious. Modeling asymptomatic and pre-symptomatic cases is important because those people do not reduce their contacts or demand a test and can potentially infect many other people \citep{Donsimoni2020}.

A small share of symptomatic people will develop strong symptoms that require intensive care. The exact share and time span is age-dependent. An age-dependent share of intensive care unit (ICU) patients will die after spending up to 32 days in intensive care. Moreover, if the ICU capacity was reached, all patients who require intensive care but do not receive it die.

It would be easy to make the course of disease even more fine grained. For example, we could model people who require hospitalization but not intensive care. So far we opted against that because only the ICU capacity might become a bottleneck in Germany.

We allow the progression of the disease to be stochastic in two ways: Firstly, state changes only occur with a certain probability (e.g. only a fraction of infected individuals develops symptoms). Secondly, the number of periods for which an individual remains in a state is drawn randomly. The parameters that govern these processes are taken from the literature. They can vary with the age of an individual.

Detailed information on the calibration of the disease parameters is available as part of our \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.

### 2020-12-19 paper/model/endogenous_contact_reductions.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Endogenous Contact Reductions}
\label{sec:endogenous_contact_reductions}

Policies are not the only way in which the number of contacts are reduced compared to the pre-pandemic level. It is important to model those other channels. Otherwise, the effect of policies would be overestimated and policy recommendations based on the model would be biased.

Examples of endogenous contact reductions are manifold: symptomatic people stay at home; Members of risk groups try to reduce their number of contacts more strongly than others; People self-isolate if they know they had a risk contact.

Since we model the number of contacts as arbitrary functions of background characteristics and health states, it is easy to implement such considerations.

In our current empirical application we only model that symptomatic people reduce their number of contacts across all contact types (except for households) by 70 \%. Within households they reduce contacts by 50\%. We are working on extending this to allow for formal and informal contact tracing as well as quarantines after positive test results.

### 2020-12-19 paper/model/matching.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Matching Individuals}
\label{sec:matching}

The empirical data described above only allows to estimate the number of contacts each person has. In order to simulate transmissions of Covid-19, the numbers of contacts has to be translated into actual meetings between people. This is achieved by matching algorithms:

As described in section \ref{sec:number_of_contacts}, some contact types are recurrent (i.e. the same people meet regularly), others are non-recurrent (i.e. it would only be by accident that two people meet twice). The matching process is different for recurrent and non recurrent contact models.

Recurrent contacts are described by two components: 1) A variable in the background characteristics. An example would be a school class identifier which could come from actual data or be drawn randomly to achieve representative class sizes. 2) A deterministic or random function that takes the value 0 (non-participating) and 1 (participating) and can depend on the weekday, date and health state. This can be used to model vacations, weekends or symptomatic people who stay home (see section \ref{sec:endogenous_contact_reductions} for details).

The matching process for recurrent contacts is then extremely simple: On each simulated day, every person who does not stay home meets all other group members who do not stay home. The assumption that all group members have contacts with all other group members is not fully realistic, but seems like a good approximation to reality, especially in light of the suspected role of aerosol transmission for Covid-19 (\cite{Morawska2020, Anderson2020}).

The matching in non-recurrent contact models is more difficult and implemented in a two stage sampling procedure to allow for assortative matching. Currently most contact models are assortative with respect to age (it is more likely to meet people from the same age group) and county (it is more likely to meet people from the same county) but in principle any set of discrete variables can be used. This set of variables that influence matching probabilities introduce a discrete partition of the population into groups. The first stage of the two stage sampling process samples on the group level. The second stage on the individual level.

Below, we first show pseudo code for the non-recurrent matching algorithm and then describe how the algorithm works in words.

%%% ToDo: Get syntax highlighting for the pseudo code. Try out different languages.

\begin{lstlisting}
while unmatched contacts remain:
    model, id = draw contact model and individual
    for contact in remaining_contacts[id, model]:
        draw group of other person
        draw other person from that group
        determine if id infects other or vice versa
        remaining_contacts[id, model] -= 1
        remaining_contacts[other, model] -= 1
\end{lstlisting}

We first randomly draw a contact type and individual. For each contact oth the drawn contact type that person has, we first draw the group of the other person (first stage). Next, we calculate the probability to be drawn for each member of the group, based on the number of remaining contacts. I.e. people who have more remaining contacts are drawn with a higher probability. This has to be re-calculated each time because with each matched contact, the number of remaining contacts changes. We then draw the other individual, determine whether an infection takes place and if so update the health states of the newly infected person. Finally, we reduce the number of remaining contacts of the two matched individuals by one.

The recalculation of matching probabilities in the second stage is computationally intensive because it requires summing up all remaining contacts in that group. Using a two stage sampling process where the first stage probabilities remain constant over time makes the matching computationally much more tractable because the number of computations increases quadratically in the second stage group size.

### 2020-12-19 paper/model/number_of_contacts.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Modeling Numbers of Contacts}
\label{sec:number_of_contacts}

Consider a hypothetical population of 1,000 individuals in which 50 were infected with a novel infectious disease. From this data alone, it is impossible to say whether only those 50 people had contact with an infectious person and the disease has an infection probability of 1 in each contact or whether everyone met an infectious person but the disease has an infection probability of only 5 percent per contact. SEIR models do not even try to distinguish contact frequency from the infectiousness of each contact and combine the two in one parameter that is not invariant to social distancing policies.

To model social distancing policies, we need to disentangle the effects of the number of contacts of each individual and the effect of policy-invariant infection probabilities specific to each contact type. Since not all contacts are equally infectious, we distinguish different contact types.

The number and type of contacts in our model can be easily extended. Each type of contacts is described by a function that maps individual characteristics, health states and the date into a number of planned contacts for each individual. This allows to model a wide range of contact types.

Currently, there are the following contact types:

\begin{itemize}
    \item Households: Each household member meets all other household members every day. The household sizes and structures are calibrated to be representative for Germany.
    \item Random non-work contacts: Each person has contacts with randomly drawn other people. This contact type reflects contacts during pure leisure activities as well as non leisure activities such as grocery shopping or medical appointments.
    \item Random work contacts: Each working adult has contact to randomly drawn other people.
    \item Recurrent daily work contacts: Each working adult meets other workers every day. This is meant to capture work colleagues.
    \item Recurrent weekly work contacts: Each working adult meets other workers once per week. We randomize over the days on which the meetings take place. This is meant to capture meetings with clients, superiors or other colleagues which happen infrequently.
    \item Schools: Each student meets all of his classmates every day. Class sizes are calibrated to be representative for Germany. Schools are closed on weekends and during vacations, which vary by states.
    \item Preschools: Similar to schools but for younger children.
    \item Nurseries: Similar to schools but for very young children.
\end{itemize}

The number of random and recurrent contacts at the workplace and at home is calibrated with data provided by \cite{Mossong2008}. For details see Section~\ref{sec:calibration}. In particular, we sample the number of contacts or group sizes from empirical distributions that sometimes depend on age. Instead it would also be possible to use economic or other behavioral models to predict the number of contacts.

Theoretically, each contact type can have its own infection probability. However, to reduce the number of free parameters and thus avoid a potential over-fitting we impose some constraints. For now, infection probabilities in schools, preschools and nurseries are equal. Moreover, we restrict all work contacts to have the same infection probability.

### 2020-12-19 paper/model/policies.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Reducing Numbers of Contacts Through Policies}
\label{sec:policies}

The main motivation of our model is to predict the effect of policies that affect the number of contacts people have. Examples range from school closures and lockdowns to more nuanced policies such as a mandatory quarantines for symptomatic individuals or a class splitting policy where only half of the students come to school in person and the other half joins digitally with weekly rotation.

Instead of thinking of policies as completely replacing how many contacts people have, it is often more helpful to think of them as adjusting the pre-pandemic number of contacts.

Therefore, we implement policies as a step that happens after the number of contacts is calculated but before individuals are matched.

On an abstract level, a policy is a functions that modifies the number of contacts of one contact type. For example, school closures simply set all school contacts to zero. A lockdown where only essential workers are allowed to work means that approximately two thirds of the working population have zero work contacts and the rest has the same number of contacts as before.

% importance of fine-grained contact types and having the entire population
This, in conjunction with our fine grained contact types, allows us to easily implement a wide variety of policies. Allowing policies to depend on the health states of the entire population means that adaptive lockdowns where, for example, schools close when a certain threshold of infections is surpassed at the county level would be as simple as determining which counties are above the threshold and then setting all school contacts in these counties to zero.

The dependency of policies on health states also makes it possible to model contact tracing. For example, a policy could check whether each child has a classmate who's received a positive test result and then bar all children of that class from attending school.

% contact tracing and how including the states means the sky is the limit
Some policies can be easily implemented if the background characteristics are suitably extended. For example, a schooling policy with split classes, where each half attends school every other week can be implemented by storing whether the child would attend in even or odd weeks in the background characteristics and then using that information in the policy function.

For some policies the exact effect on each contact type is not easy to determine. If this refers to a policy during the estimation period, it is possible to estimate such parameters by fitting the model to time series data of infection rates. This is only possible if the policy was not active during the whole estimation period and thus the infection probabilities can be identified separately. If instead it refers to a policy that we want to simulate, we make a scenario analysis in which the model is simulated with several assumptions about how the policy affects the number of contacts.

### 2020-12-19 paper/model/summary.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Summary}
\label{sub:model_summary}

To alleviate these problems we propose a different model structure. Our model inherits many features from agent-based simulation models but replaces the contacts between moving particles by contacts between individuals who work, go to school, live in a household and enjoy leisure activities. The structure of the model is depicted in Figure~\ref{fig:model_graph}.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/model_detailed.png}
    \caption{Simplified graph of the model}
    \label{fig:model_graph}
\end{figure}

\pdfcommentmarkup[markup=Highlight]{

The background characteristics include age, county and occupation of each simulated individual. Contact models are functions that map individual characteristics into a predicted number of contacts. Currently we distinguish between eight types of contact models which are all listed in Figure \ref{fig:model_graph}: households, recurrent and random work contacts, recurrent and random leisure contacts, and nursery, preschool, and school contacts.

}{Are these contact models still up-to-date.}

The predicted number of contacts is translated into infections by a matching algorithm. There are different matching algorithms for recurrent contacts (e.g. classmates, family members) and non-recurrent contacts (e.g. clients, contacts in supermarkets). The infection probability can differ for each contact type. All types of contacts can be assortative with respect to geographic and demographic characteristics.

Once a person is infected, the disease progresses in a fairly standard way which is also depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild symptoms are infectious for some time and recover eventually. Cases with severe symptoms additionally require hospitalization and lead to either recovery or death.

There are two possible ways to enable testing for Covid-19 in the model. The first way allows to specify the ratio of known to unknown infections for each day. A random sample of all newly infected individuals will then receive a test result in the following days. The second approach allows people who experience symptoms as well as people who had a risk contact can demand a test. Test allocation models that take into account the testing capacity determine which individuals actually receive a test. Test processing models determine in which order the tests are processed and when infected individuals are notified.

In addition, people who have symptoms, received a positive test, or had a risk contact can reduce their number of contacts across all contact types endogenously.

The model makes it very simple to translate policies into model quantities. For example, school closures imply the complete suspension of school contacts. A strict lockdown implies shutting down work contacts of all people who are not employed in a systemically relevant sector. It is also possible to have more sophisticated policies that condition the number of contacts on observable characteristics, risk contacts or health states.

Another key advantage of the model is that the number of contacts an individual has of each contact type can be calibrated from publicly available data \citep{Mossong2008}. This in turn allows us to estimate policy-invariant infection probabilities from time series of infection and death rates using the method of simulated moments \citep{McFadden1989}. Since the infection probabilities are time-invariant, data collected since the beginning of the pandemic can be used for estimation. Moreover, since we can model the testing strategies that were in place at each point in time, we can correct the estimates for the fact that not all infections are observed.

In the following sections we describe each of the model components in more detail.

### 2020-12-19 paper/model/testing.tex
Split up the section on the model in subfiles, added a file on testing.

\subsection{Testing} % (fold)
\label{sub:testing}

The model offers two approaches to implement testing for Covid-19.

The first way can be described as an top-down approach. Using data on estimates for the ratio of known to all infections like from the DunkelzifferRadar project\footnotemark, a random sample of individuals who is newly infected is sampled and assigned a test. The result of the test will be revealed after some duration which must take into account the time it takes for a person to develop symptoms, the availability of tests and time to process the tests.

\footnotetext[dunkelzifferradar]{The Dunkelzifferradar can be found here: \url{https://covid19.dunkelzifferradar.de/}}

The second way is an bottom-up approach which consists of three phases. In the first phase, the demand for tests is modeled. Demand functions map from individual characteristics to a probability which is the probability for this individual to demand a test. There can be multiple demand functions where each function may describe a different channel. For example, individuals who experience symptoms or have a risk contact may ask for a test. Or, the ministry of education requires a negative test result from every teacher every second week. After the probabilities for each individual and every demand model are computed, individuals who demand a test as well as the channel is sampled.

The second stage is the allocation phase in which demand and supply for tests are matched. The number of available tests can be inferred from official data and used to model shortages in supply. When demand exceeds supply, some individuals might be given preferred access to tests because of their own vulnerability or their potential to become a super-spreader.

In the last and third phase, administered tests are determined to be processed. This step can become a bottleneck in the testing process if there are not enough laboratories or necessary resources available to evaluate all tests.

The strength of the first approach is its simplicity, but it relies on estimates for the ratio of known to unknown cases for which not much data exist in the absence of systematic and randomized testing. The second approach is much more detailed and in principle able to reflect individual responses and governmental restrictions on the access to tests. At the same time, individual behavior and restrictions are complex and the determinants of the demand, allocation and processing functions require a lot of calibration.

% subsection testing (end)

### 2020-12-19 paper/model/number_of_contacts.tex
Some extensions to the description of educational facilities.

@@
     \item Households: Each household member meets all other household members every day. The household sizes and structures are calibrated to be representative for Germany.
+
     \item Random non-work contacts: Each person has contacts with randomly drawn other people.
 This contact type reflects contacts during pure leisure activities as well as non leisure activities such as grocery shopping or medical appointments.
+
     \item Random work contacts: Each working adult has contact to randomly drawn other people.
+
     \item Recurrent daily work contacts: Each working adult meets other workers every day. This is meant to capture work colleagues.
+
     \item Recurrent weekly work contacts: Each working adult meets other workers once per week.
@@
 This is meant to capture meetings with clients, superiors or other colleagues which happen infrequently.
+
     \item Schools: Each student meets all of his classmates every day.
-Class sizes are calibrated to be representative for Germany.
+Class sizes are calibrated to be representative for Germany and students have the same age.
 Schools are closed on weekends and during vacations, which vary by states.
-    \item Preschools: Similar to schools but for younger children.
-    \item Nurseries: Similar to schools but for very young children.
+School classes also meet three pairs of teachers every school day.
+The pairs are meant to represent interactions between teachers.
+
+    \item Preschools: Children who are at least three years old and younger than six may attend preschool.
+Each group of nine children interacts with the same two adults every day.
+The children in each group are of the same age.
+The remaining mechanics are similar to schools.
+
+    \item Nurseries: Children younger than three years may attend a nursery and interact with one adult. The age of the children varies within groups. The remaining mechanics are similar to schools.
 \end{itemize}

### 2020-12-19 paper/model/testing.tex
Some extensions to the description of educational facilities.

@@
 The first way can be described as an top-down approach.
-Using data on estimates for the ratio of known to all infections like from the DunkelzifferRadar project\footnotemark, a random sample of individuals who is newly infected is sampled and assigned a test.
+Using data on estimates for the ratio of known to all infections\footnotemark a random sample of individuals who is newly infected is sampled and assigned a test.
 The result of the test will be revealed after some duration which must take into account the time it takes for a person to develop symptoms, the availability of tests and time to process the tests.
 
-\footnotetext[dunkelzifferradar]{The Dunkelzifferradar can be found here: \url{https://covid19.dunkelzifferradar.de/}}
+\footnotetext{The Dunkelzifferradar project publishes daily estimates of the dark figure of infections under \url{https://covid19.dunkelzifferradar.de/}}
 

### 2020-12-19 paper/abstract.tex
Clean up preamble, fix typos and add section on initial conditions.

@@
 
-Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases with minimal social and economic costs.
-Epidemiological models have a hard time predicting the effects of such fine grained policies.
+Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs.
+Epidemiological models have a hard time predicting the effects of such fine-grained policies.
 We propose a novel simulation-based model to address this shortcoming.
@@
 The number of contacts changes with social distancing policies, the infection probabilities remain invariant.
-This allows us to model many types of fine grained policies that cannot easily be incorporated into other models.
+This allows us to model many types of fine-grained policies that cannot easily be incorporated into other models.
 To validate our model, we show that it can accurately predict the effect of the German November lockdown even if no similar policy has been observed in the time series that were used to estimate the model parameters.

### 2020-12-19 paper/conclusion.tex
Clean up preamble, fix typos and add section on initial conditions.

@@
 
-We propose a simulation based model of infectious disease transmission that is designed to predict the effects of fine grained social distancing policies.
+We propose a simulation based model of infectious disease transmission that is designed to predict the effects of fine-grained social distancing policies.
 In particular, the model can be used to model policies such as several ways of splitting school classes or work reduction policies that affect essential and non-essential workers differently.
@@
 We calibrate age dependent disease progression parameters from the medical literature and age dependent contact frequencies from contact diaries.
-Moreover, we distinguish eight types of contacts out of which some are only relevant for certain age groups.
+Moreover, we distinguish ten types of contacts out of which some are only relevant for certain age groups.
 

### 2020-12-19 paper/literature_review.tex
Split up the section on the model in subfiles, added a file on testing.
Clean up preamble, fix typos and add section on initial conditions.

@@
 In fact, policy simulations are done by setting $R_0$ to a different value but it is hard to translate a real policy into the value of $R_0$ it will induce.
+In other words, SEIR models are not suited for evaluating the effect of policies which have never been experienced before.
 
@@
 The above assessment of epidemiological models is not meant as a critique.
-We are aware that those models were not designed to predict the effect of fine grained social distancing policies in real time and are very well suited to their purpose.
+We are aware that those models were not designed to predict the effect of fine-grained social distancing policies in real time and are very well suited to their purpose.
 We invite epidemiologists to provide feedback and collaborate to improve our model.

### 2020-12-19 paper/model/initial_conditions.tex
Clean up preamble, fix typos and add section on initial conditions.

\subsection{Initial Conditions} % (fold)
\label{sub:initial_conditions}

Consider a situation where you want to start a simulation with the beginning set amidst the pandemic. It means that several thousands of individuals have already recovered from the disease, are infectious, symptomatic or in ICUs. Additionally, the sample of infectious people who will determine the course of the pandemic in the following periods is likely not representative of the whole population because of differences in behavior (number of contacts, assortativeness), past policies (school closures), etc.. The distribution of courses of diseases in the population at the begin of the simulation are called initial conditions.

To come up with realistic initial conditions, we match reported infections from official data to simulated individuals by available characteristics like age and geographic information. The matching must be done for each day of a longer time frame like a month to have individuals in all possible health states. Then, health statuses evolve until the begin of the simulation period without simulating infections by contacts. It is also possible to correct reported infections for a reporting lag and scale them up to arrive at the true number of infections.

% subsection initial_conditions (end)

### 2020-12-19 paper/model/number_of_contacts.tex
Clean up preamble, fix typos and add section on initial conditions.

@@
 
-    \item Random non-work contacts: Each person has contacts with randomly drawn other people.
+    \item Random non-work contacts: Each person has contacts with randomly drawn other people which are assortative with respect to region and age group.
 This contact type reflects contacts during pure leisure activities as well as non leisure activities such as grocery shopping or medical appointments.
 
-    \item Random work contacts: Each working adult has contact to randomly drawn other people.
+    \item Recurrent daily non-work contacts: Each person has daily recurring contacts which allows to model close relationships other than families between individuals.
+
+    \item Recurrent weekly non-work contacts: Each person has weekly recurring contacts like sports groups or other weekly activities.
+
+    \item Random work contacts: Each working adult has contact to randomly drawn other people at work.
 
@@
 Moreover, we restrict all work contacts to have the same infection probability.
+???STILL TRUE???

### 2020-12-19 paper/model/policies.tex
Clean up preamble, fix typos and add section on initial conditions.

@@
 The main motivation of our model is to predict the effect of policies that affect the number of contacts people have.
-Examples range from school closures and lockdowns to more nuanced policies such as a mandatory quarantines for symptomatic individuals or a class splitting policy where only half of the students come to school in person and the other half joins digitally with weekly rotation.
+Examples range from school closures and lockdowns to more nuanced policies such as mandatory quarantines for symptomatic individuals or a class splitting policy where only half of the students come to school in person and the other half joins digitally with weekly rotation.
 
@@
 % importance of fine-grained contact types and having the entire population
-This, in conjunction with our fine grained contact types, allows us to easily implement a wide variety of policies.
+This, in conjunction with our fine-grained contact types, allows us to easily implement a wide variety of policies.
 Allowing policies to depend on the health states of the entire population means that adaptive lockdowns where, for example, schools close when a certain threshold of infections is surpassed at the county level would be as simple as determining which counties are above the threshold and then setting all school contacts in these counties to zero.

### 2020-12-19 paper/model/summary.tex
Clean up preamble, fix typos and add section on initial conditions.

@@
 
-\pdfcommentmarkup[markup=Highlight]{
-
 The background characteristics include age, county and occupation of each simulated individual.
 Contact models are functions that map individual characteristics into a predicted number of contacts.
-Currently we distinguish between eight types of contact models which are all listed in Figure \ref{fig:model_graph}: households, recurrent and random work contacts, recurrent and random leisure contacts, and nursery, preschool, and school contacts.
-
-}{Are these contact models still up-to-date.}
+Currently we distinguish between eight types of contact models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and random work contacts, recurrent and random leisure contacts, and nursery, preschool, and school contacts.
 
@@
 A random sample of all newly infected individuals will then receive a test result in the following days.
-The second approach allows people who experience symptoms as well as people who had a risk contact can demand a test.
-Test allocation models that take into account the testing capacity determine which individuals actually receive a test.
+The second approach allows people who experience symptoms or people who had a risk contact can demand a test.
+Test allocation models assign tests to individuals while also respecting the capacities.
 Test processing models determine in which order the tests are processed and when infected individuals are notified.

### 2020-12-20 paper/model/initial_conditions.tex
minor ameliorations.

@@
 Additionally, the sample of infectious people who will determine the course of the pandemic in the following periods is likely not representative of the whole population because of differences in behavior (number of contacts, assortativeness), past policies (school closures), etc..
-The distribution of courses of diseases in the population at the begin of the simulation are called initial conditions.
+The distribution of courses of diseases in the population at the begin of the simulation is called initial conditions.
 
 To come up with realistic initial conditions, we match reported infections from official data to simulated individuals by available characteristics like age and geographic information.
-The matching must be done for each day of a longer time frame like a month to have individuals in all possible health states.
+The matching must be done for each day of a longer time frame like a month to have individuals with possible health states.
 Then, health statuses evolve until the begin of the simulation period without simulating infections by contacts.

### 2020-12-20 paper/model/summary.tex
minor ameliorations.

@@
 A random sample of all newly infected individuals will then receive a test result in the following days.
-The second approach allows people who experience symptoms or people who had a risk contact can demand a test.
-Test allocation models assign tests to individuals while also respecting the capacities.
-Test processing models determine in which order the tests are processed and when infected individuals are notified.
+The second approach consists of three steps.
+First, individuals demand a test because they, e.g., experience symptoms.
+Secondly, tests are allocated to individuals while respecting governmental access restrictions to test.
+Thirdly, depending of the capacities of laboratories, tests are processed for some time until the individual receives her test result.
 
@@
 
+Last but not least, performing simulations whose starting point is set amidst the pandemic requires special adjustments to arrive at a realistic distribution of courses of diseases.
+We solve the initial conditions problem by matching reported infections to individuals in our data while also correcting for reporting lag and undetected cases.
+
 In the following sections we describe each of the model components in more detail.

### 2020-12-20 paper/abstract.tex
Move to Holger's templates and some other - cannot remember - fixes.

@@
-\begin{abstract}
-
-Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs.
+\noindent Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs.
 Epidemiological models have a hard time predicting the effects of such fine-grained policies.
@@
 \vspace{1cm}
-JEL Classification: C63, I18
+\noindent JEL Classification: C63, I18
 
-Keywords: Covid-19, agent based simulation model, public health measures
-
-\end{abstract}
+\noindent Keywords: Covid-19, agent based simulation model, public health measures

### 2020-12-20 paper/calibration_and_estimation.tex
Clean up preamble, fix typos and add section on initial conditions.
Move to Holger's templates and some other - cannot remember - fixes.

@@
 Estimates of the incubation period usually give a range from 2 to 12 days.
-A meta analysis by \cite{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modelled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95 \% CIs) of 1.63 (95 \% CI 1.51 to 1.75) and 0.50 (95 \% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.
+A meta analysis by \cite{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modelled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.
 
@@
 
-\subsection{Duration of Infectiousness}
+\subsubsection{Duration of Infectiousness}
 
@@
 
-\subsection{Duration of Symptoms}
+\subsubsection{Duration of Symptoms}
 
@@
 The data on how many percent of symptomatic patients will require ICU is pretty thin.
-We rely on data by the US CDC (\cite{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/master/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project (2020-09-14)}.
+We rely on data by the US CDC (\cite{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.
+Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of requiring intensive care per age group.
 
-From these sources we arrive at the following probabilities of requiring intensive care:
-
-\begin{table}[!ht]
+\begin{table}[tb]
     \caption{Shares of symptomatic patients who will require ICU care by age groups.}
@@
     \end{tabular}
+
+    \tablenotes{The data is taken from \cite{Stokes2020} and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.}
 
@@
 
-% Mossong data
 We calibrate the parameters for the predicted numbers of contacts from contact diaries of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg \citep{Mossong2008}.
@@
 In particular, we allow this probability to depend on age and county of residence.
-While we do not have good data on geographical assortativeness and just roughly calibrate it such that 60 \% of contacts are within the same county,\footnote{We are working on improving this estimation with mobility data.} we can calibrate the age assortativeness from the same data we use to calibrate the number of contacts.
+While we do not have good data on geographical assortativeness and just roughly calibrate it such that 60 \% of contacts are within the same county\footnotemark, we can calibrate the age assortativeness from the same data we use to calibrate the number of contacts.
 
-\begin{figure}[!ht]
+\footnotetext{We are working on improving this estimation with mobility data.}
+
+\begin{figure}[!tp]
     \centering
-    \includegraphics[width=0.7\textwidth]{../figures/assortative_matching_probability_example.png}
+    \includegraphics[width=\textwidth]{../figures/assortative_matching_probability_example.png}
     \caption{Distribution of random non-work contacts by age of participants}
@@
 4) for leisure activities.
-
-\section{Model Validation}
-
-We validate our model in two ways: 1) We look at the in-sample fit over the estimation period.
-2) We look at the out-of-sample fit for November.
-The last one is a challenging test for our model because there was a strong policy change between the estimation period and November.
-The model convincingly passes both tests.
-
-\subsection{In-Sample Fit}
-
-Despite fitting only four free parameters, the in-sample fit is very good.
-The best fit is achieved in the largest age groups.
-This is so mechanically, because we weight the deviations between simulated and observed infection rates by group sizes.
-The worst fit is achieved for the 80 to 100 years old.
-The model predicts too few infections for these groups because they have very few contacts in all contact types we have included so far.
-We plan to address this issue soon by adding another contact type that captures all contacts in the data by \cite{Mossong2008} that we have not included so far.
-Moreover, we expect an improved model fit when we allow for more free parameters.
-
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[width=0.7\textwidth]{../figures/goodness_of_fit_by_age_group}
-    \caption{Actual vs. simulated infection and fatality rates}
-    \label{fig:goodness_of_fit}
-\end{figure}
-
-\subsection{Out-of-Sample Fit}
-
-We can assess the out-of-sample fit by projecting the effect of the lockdown light and comparing it to case numbers until now.
-It is important to note that this is not just a simple extrapolation of a time trend because the lockdown light only started after the estimation period.
-The out-of-sample fit can be assessed in Figure~\ref{fig:out-of-sample-fit}.
-
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[width=0.7\textwidth]{../figures/out_of_sample_validation}
-    \caption{Projected effect of the lockdown-light}
-    \label{fig:out-of-sample-fit}
-\end{figure}
-
-The model correctly predicts the effect of the lockdown light with reasonable accuracy.
-In particular, the actual case numbers are between our neutral and pessimistic projection.
-The plot also shows that ending the lockdown light as planned on November 30 would lead to an explosive growth in case numbers in all scenarios.

### 2020-12-20 paper/introduction.tex
Move to Holger's templates and some other - cannot remember - fixes.

@@
 For example, Germany imposed a ``lockdown-light'' in November.
-Some businesses are closed and certain leisure activities are prohibited but schools and nurseries stay open.\blfootnote{Gabler and Röhrl are grateful for financial support by the German Research Foundation (DFG) through CRC-TR 224 (Projects C01 and A02, respectively).
-Gabler is grateful for funding by IZA Institute of Labor Economics.}
+Some businesses are closed and certain leisure activities are prohibited but schools and nurseries stay open.
 
@@
 
-\setstretch{1.0}
 \begin{enumerate}
@@
     \item The model achieves a good fit on German infection and fatality rate data even if only the infection probabilities are fit to the data and the remaining parameters are calibrated from the medical literature and datasets on contact frequencies.
-    \item High quality Python code for the model is freely available on \href{https://github.com/covid-19-impact-lab/sid}{https://github.com/covid-19-impact-lab/sid}, well documented and very flexible.
+    \item High quality Python code for the model is freely available on Github, well documented and very flexible\footnotemark.
 We are actively looking for researchers who want to use our model for their projects.
 \end{enumerate}
-\setstretch{1.25}
 
-After a brief literature review we describe our model in detail and validate its in-sample and out-of-sample fit on German data.
+\footnotetext{The code can be found under \url{https://github.com/covid-19-impact-lab/sid} and the documentation with tutorials and background information under \url{https://sid-dev.readthedocs.io/}.}
+
+In Section~\ref{sec:literature_review}, we give a short overview of epidemiological models and how they relate to our model.
+We continue with a detailed model description in Section~\ref{sec:model} and proceed with a description of model parameters and the estimation in Section~\ref{sec:calibration_and_estimation}.
+The model is validated in Section~\ref{sec:model_validation} by assessing the in-sample fit for reported infections from August to October and by comparing the out-of-sample fit for the period from November until the beginning of Christmas for different lockdown scenarios.
+We conclude in Section~\ref{sec:conclusion}.
 
@@
 
-More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\cite{Dorn2020a}).
+More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\cite{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new report.}

### 2020-12-20 paper/model_validation.tex
Move to Holger's templates and some other - cannot remember - fixes.

\section{Model Validation}
\label{sec:model_validation}

We validate our model in two ways: 1) We look at the in-sample fit over the estimation period. 2) We look at the out-of-sample fit for November. The last one is a challenging test for our model because there was a strong policy change between the estimation period and November. The model convincingly passes both tests.

\subsection{In-Sample Fit}

Despite fitting only four free parameters, the in-sample fit is very good. The best fit is achieved in the largest age groups. This is so mechanically, because we weight the deviations between simulated and observed infection rates by group sizes. The worst fit is achieved for the 80 to 100 years old. The model predicts too few infections for these groups because they have very few contacts in all contact types we have included so far. We plan to address this issue soon by adding another contact type that captures all contacts in the data by \cite{Mossong2008} that we have not included so far. Moreover, we expect an improved model fit when we allow for more free parameters.

\begin{figure}[!tp]
    \centering
    \includegraphics[width=\textwidth]{../figures/goodness_of_fit_by_age_group}
    \caption{Actual vs. simulated infection and fatality rates}
    \label{fig:goodness_of_fit}
\end{figure}

\subsection{Out-of-Sample Fit}

We can assess the out-of-sample fit by projecting the effect of the lockdown light and comparing it to case numbers until now. It is important to note that this is not just a simple extrapolation of a time trend because the lockdown light only started after the estimation period. The out-of-sample fit can be assessed in Figure~\ref{fig:out-of-sample-fit}.

\begin{figure}[!tp]
    \centering
    \includegraphics[width=\textwidth]{../figures/out_of_sample_validation}
    \caption{Projected effect of the lockdown-light}
    \label{fig:out-of-sample-fit}
\end{figure}

The model correctly predicts the effect of the lockdown light with reasonable accuracy. In particular, the actual case numbers are between our neutral and pessimistic projection. The plot also shows that ending the lockdown light as planned on November 30 would lead to an explosive growth in case numbers in all scenarios.

### 2020-12-20 task_compile_documents.py
Add pytask.

@@ -1,0 +1,13 @@
+import pytask
+from pathlib import Path
+
+
+ROOT = Path(__file__).parent
+BLD = ROOT / "bld"
+
+
+@pytask.mark.latex
+@pytask.mark.depends_on(ROOT / "paper" / "paper.tex")
+@pytask.mark.produces(BLD / "paper.pdf")
+def task_compile_documents():
+    pass

### 2020-12-20 paper/calibration_and_estimation.tex
Add pytask.
more fixes.

@@
 \section{Calibration and Estimation}
-\label{sec:calibration}
+\label{sec:calibration_and_estimation}
 
@@
 
-\subsection{Medical Parameters}
-
-\subsubsection{Length of Presymptomatic Stage / Incubation Period}
-
-Estimates of the incubation period usually give a range from 2 to 12 days.
-A meta analysis by \cite{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modelled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.
-
-\subsubsection{Begin of Infectiousness}
-
-The period between infection and onset of infectiousness is called latent or latency period.
-However, the latency period is rarely given in epidemiological reports on Covid-19.
-Instead, scientists and agencies usually report the incubation period, the period from infection to the onset of symptoms.
-A few studies used measurements of virus shedding to estimate infectiousness during the course of the disease.
-When measurements started before the onset of symptoms the development of the viral load before symptoms gives us an indication of number of days between the onset of infectiousness and symptoms.
-
-The European Centre for Disease Prevention and Control estimates that people become infectious between one and two days before the symptoms set in.
-This is similar to \cite{He2020} who estimate this to take 2.3 days and is in line with \cite{Peak2020}.
-
-Given these numbers and the length of the incubation period we can calculate the latency period for symptomatic people.
-To our knowledge no estimates for the latency period of asymptomatic cases of COVID-19 exist.
-We assume it to be the same for symptomatic and asymptomatic cases.
-
-Thus, we arrive at the following distribution for latency periods: 40\% have one day. 35\% have two days. 20\% have three days and 5\% have 5 days.
-
-\subsubsection{Duration of Infectiousness}
-
-We assume that the duration of infectiousness is the same for both symptomatic and asymptomatic individuals as evidence suggests little differences in the transmission rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients (\cite{Yin2020}) and that the viral load between symptomatic and asymptomatic individuals are similar (\cite{Zou2020}, \cite{Byrne2020}, \cite{Singanayagam2020}).
-
-Our distribution of the duration of infectiousness is based on \cite{Byrne2020}.
-
-For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8 days of infectiousness afterwards.\footnote{Viral loads may be detected much later but 8 days seems to be the time after which most people are culture negative, as also reported by \cite{Singanayagam2020}} Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic (see also figure 5).
-This duration range is very much in line with the meta-analysis’ reported evidence for asymptomatic individuals (see their figure 1).
-Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic.
-This duration range is very much in line with the meta-analysis' reported evidence for asymptomatic individuals.
-
-Following this evidence we assume the following discretized distribution of the infectiousness period: 10\% of individuals are infectious for three days, 25\% for five days, another 25\% for seven days, 20\% for nine days and 20\% for eleven days.
-
-\subsubsection{Duration of Symptoms}
-
-We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3, Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring symptomatic cases.
-
-We collapse the data to the following distribution: 10\% recover after 15 days and 30\% require 18, 22 or 27 days respectively.
-
-% These long symptom durations align with reports by \cite{Tenforde2020}.
-
-These numbers are only used for mild cases.
-We do not disaggregate by age.
-Note that the length of symptoms is not very important in our model given that individuals stop being infectious before their symptoms cease.
-
-\subsubsection{Time from Symptom Onset to Admission to ICU}
-
-The data on how many percent of symptomatic patients will require ICU is pretty thin.
-We rely on data by the US CDC (\cite{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.
-Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of requiring intensive care per age group.
-
-\begin{table}[tb]
-    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
-    \label{tab:symptomatic-to-ICU}
-    \centering
-
-    \begin{tabular}{ll}
-        \toprule
-        Age Group & Share \\
-        \midrule
-        0-9 & 0.00005 \\
-        10-19 & 0.00030 \\
-        20-29 & 0.00075 \\
-        30-39 & 0.00345 \\
-        40-49 & 0.01380 \\
-        50-59 & 0.03404 \\
-        60-69 & 0.10138 \\
-        70-79 & 0.16891 \\
-        80-100 & 0.26871 \\
-        \bottomrule
-    \end{tabular}
-
-    \tablenotes{The data is taken from \cite{Stokes2020} and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.}
-
-\end{table}
-
-For those who will require intensive care we follow \cite{Chen2020} who estimate the time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.
-
-This aligns well with numbers reported for the time from first symptoms to hospitalization: \cite{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4.
-This is also in line with the durations collected by \href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the Robert Koch Institut}.
-
-We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with equal probabilities. These times mostly matter for the ICU capacities.
-
-\subsubsection{Death and Recovery from ICU}
-
-We take the survival probabilities and time to death and time until recovery from intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.
-
-They report time until death to have a mean of 11.74 days and a standard deviation of 8.79 days.
-Approximating this with the normal distribution, we have nearly 10\% probability mass below 0.
-We use it nevertheless as several other distributions (such as chi squared and uniform) were unable to match the variance.
-Discretizing this leads to 41\% of individuals who die from Covid-19 to die after one day in intensive care.
-22\% day after 12 days, 29\% after 20 days and 7\% after 32 days.
-Again, we rescale this for every age group among those that will not survive.
-
-They report time until recovery to have a mean of 18.8 days and a standard deviation of 12.21 days.
-Approximating this with the normal distribution, we have over 5\% probability mass below 0.
-Discretizing this of those who recover in intensive care 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.
+\input{medical_parameters}
 
@@
 To calibrate infection probabilities outside of the model, it would be important to know the exact duration and distance of each contact type as well as virus loads.
-Since this is not available in any dataset, we instead estimate those parameters inside the model by minimizing the distance between simulated and observed infection rates.
+Since this is not available in any dataset, we estimate those parameters inside the model with the method of simulated moments \citep{McFadden1989} by minimizing the distance between simulated and observed infection rates.
 Since our model includes a lot of randomness, we average simulated infection rates over several model runs.

### 2020-12-20 paper/introduction.tex
more fixes.

@@
 
-More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\cite{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new report.}
+More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\citet{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new report.}

### 2020-12-20 paper/literature_review.tex
more fixes.

@@
 This has motivated a large number of researchers to extend the standard model to allow for more heterogeneity and flexibility.
-Examples are \cite{Grimm2020}, \cite{Donsimoni2020} and \cite{Acemoglu2020} who develop multi group SEIR models to analyze the effects of targeted lockdowns and \cite{Berger2020} who extend the SEIR model to analyze testing and conditional quarantines.
-For a more comprehensive review see \cite{Avery2020}.
+Examples are \citet{Grimm2020}, \citet{Donsimoni2020} and \citet{Acemoglu2020} who develop multi group SEIR models to analyze the effects of targeted lockdowns and \citet{Berger2020} who extend the SEIR model to analyze testing and conditional quarantines.
+For a more comprehensive review see \citet{Avery2020}.
 Others have used the results of a standard SEIR model as input for economic models that estimate the cost of policies (e.g.
-\cite{Dorn2020}).
+\citet{Dorn2020}).
 
@@
 Infections take place when two particles come closer than a certain contact radius (e.g.
-\cite{Silva2020} and \cite{Cuevas2020}).
+\citet{Silva2020} and \citet{Cuevas2020}).
 While the simulation approach makes it easy to incorporate heterogeneity in disease progression, it is hard to incorporate heterogeneity in meeting patterns.
@@
 
-\cite{Hinch2020} is a recent extension of the prototypical agent-based simulation model that replaces moving particles by contact networks for households, work and random contacts.
+\citet{Hinch2020} is a recent extension of the prototypical agent-based simulation model that replaces moving particles by contact networks for households, work and random contacts.
 This model is similar in spirit to ours but focuses on contact tracing rather than social distancing policies.

### 2020-12-20 paper/medical_parameters.tex
more fixes.

\subsection{Medical Parameters}

This section discusses the medical parameters used in the model, their sources and how we arrived at the distributions used in the model.\footnotemark

\footnotetext{Additional information can be found in the \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.}

\subsubsection{Length of Presymptomatic Stage / Incubation Period}

Estimates of the incubation period usually give a range from 2 to 12 days. A meta analysis by \citet{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.

\subsubsection{Begin of Infectiousness}

The period between infection and onset of infectiousness is called latent or latency period. However, the latency period is rarely given in epidemiological reports on Covid-19. Instead, scientists and agencies usually report the incubation period, the period from infection to the onset of symptoms. A few studies used measurements of virus shedding to estimate infectiousness during the course of the disease. When measurements started before the onset of symptoms the development of the viral load before symptoms gives us an indication of number of days between the onset of infectiousness and symptoms.

The European Centre for Disease Prevention and Control estimates that people become infectious between one and two days before the symptoms set in. This is similar to \citet{He2020} who estimate this to take 2.3 days and is in line with \citet{Peak2020}.

Given these numbers and the length of the incubation period we can calculate the latency period for symptomatic people. To our knowledge no estimates for the latency period of asymptomatic cases of COVID-19 exist. We assume it to be the same for symptomatic and asymptomatic cases.

Thus, we arrive at the following distribution for latency periods: 40\% have one day. 35\% have two days. 20\% have three days and 5\% have 5 days.

\subsubsection{Duration of Infectiousness}

We assume that the duration of infectiousness is the same for both symptomatic and asymptomatic individuals as evidence suggests little differences in the transmission rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).

Our distribution of the duration of infectiousness is based on \citet{Byrne2020}.

For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8 days of infectiousness afterwards.\footnote{Viral loads may be detected much later but 8 days seems to be the time after which most people are culture negative, as also reported by \citet{Singanayagam2020}} Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic (see also figure 5). This duration range is very much in line with the meta-analysis’ reported evidence for asymptomatic individuals (see their figure 1). Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic. This duration range is very much in line with the meta-analysis' reported evidence for asymptomatic individuals.

Following this evidence we assume the following discretized distribution of the infectiousness period: 10\% of individuals are infectious for three days, 25\% for five days, another 25\% for seven days, 20\% for nine days and 20\% for eleven days.

\subsubsection{Duration of Symptoms}

We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3, Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring symptomatic cases.

We collapse the data to the following distribution: 10\% recover after 15 days and 30\% require 18, 22 or 27 days respectively.

These numbers are only used for mild cases. We do not disaggregate by age. Note that the length of symptoms is not very important in our model given that individuals stop being infectious before their symptoms cease.

\subsubsection{Time from Symptom Onset to Admission to ICU}

The data on how many percent of symptomatic patients will require ICU is pretty thin. We rely on data by the US CDC (\citet{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of requiring intensive care per age group.

\begin{table}[tb]
    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
    \label{tab:symptomatic-to-ICU}
    \centering

    \begin{tabular}{ll}
        \toprule
        Age Group & Share \\
        \midrule
        0-9 & 0.00005 \\
        10-19 & 0.00030 \\
        20-29 & 0.00075 \\
        30-39 & 0.00345 \\
        40-49 & 0.01380 \\
        50-59 & 0.03404 \\
        60-69 & 0.10138 \\
        70-79 & 0.16891 \\
        80-100 & 0.26871 \\
        \bottomrule
    \end{tabular}

    \tablenotes{The data is taken from \citet{Stokes2020} and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.}

\end{table}

For those who will require intensive care we follow \citet{Chen2020} who estimate the time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.

This aligns well with numbers reported for the time from first symptoms to hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also in line with the durations collected by \href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the Robert Koch Institut}.

We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with equal probabilities. These times mostly matter for the ICU capacities.

\subsubsection{Death and Recovery from ICU}

We take the survival probabilities and time to death and time until recovery from intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.

They report time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. Approximating this with the normal distribution, we have nearly 10\% probability mass below 0. We use it nevertheless as several other distributions (such as chi squared and uniform) were unable to match the variance.
Discretizing this leads to 41\% of individuals who die from Covid-19 to die after one day in intensive care. 22\% day after 12 days, 29\% after 20 days and 7\% after 32 days. Again, we rescale this for every age group among those that will not survive.

They report time until recovery to have a mean of 18.8 days and a standard deviation of 12.21 days. Approximating this with the normal distribution, we have over 5\% probability mass below 0. Discretizing this of those who recover in intensive care 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

### 2020-12-20 paper/model/matching.tex
more fixes.

@@
 The matching process for recurrent contacts is then extremely simple: On each simulated day, every person who does not stay home meets all other group members who do not stay home.
-The assumption that all group members have contacts with all other group members is not fully realistic, but seems like a good approximation to reality, especially in light of the suspected role of aerosol transmission for Covid-19 (\cite{Morawska2020, Anderson2020}).
+The assumption that all group members have contacts with all other group members is not fully realistic, but seems like a good approximation to reality, especially in light of the suspected role of aerosol transmission for Covid-19 (\citet{Morawska2020, Anderson2020}).
 

### 2020-12-20 paper/model/number_of_contacts.tex
Move to Holger's templates and some other - cannot remember - fixes.
more fixes.

@@
 
-The number of random and recurrent contacts at the workplace and at home is calibrated with data provided by \cite{Mossong2008}.
+The number of random and recurrent contacts at the workplace and at home is calibrated with data provided by \citet{Mossong2008}.
 For details see Section~\ref{sec:calibration}.
@@
 For now, infection probabilities in schools, preschools and nurseries are equal.
-Moreover, we restrict all work contacts to have the same infection probability.
-???STILL TRUE???
+Moreover, we restrict all work contacts to have the same infection probability.\comment[id=T]{I have no idea about the current infection probabilities.
+Does it need an update?}

### 2020-12-20 paper/model_validation.tex
more fixes.

@@
 The model predicts too few infections for these groups because they have very few contacts in all contact types we have included so far.
-We plan to address this issue soon by adding another contact type that captures all contacts in the data by \cite{Mossong2008} that we have not included so far.
+We plan to address this issue soon by adding another contact type that captures all contacts in the data by \citet{Mossong2008} that we have not included so far.
 Moreover, we expect an improved model fit when we allow for more free parameters.

### 2020-12-20 task_compile_documents.py

@@ -7,7 +7,9 @@ BLD = ROOT / "bld"
 
 
 @pytask.mark.latex
-@pytask.mark.depends_on(ROOT / "paper" / "paper.tex")
+@pytask.mark.depends_on(
+    [ROOT / "paper" / "paper.tex"] + list(ROOT.joinpath("paper").rglob("*.tex"))
+)
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass

### 2020-12-20 paper/calibration_and_estimation.tex
Enforced 88 character limit.

@@
 
-The model is described by a large number of parameters that govern the number of contacts a person has, the likelihood of becoming infected on each contact, the likelihood of developing light or strong symptoms or even dying from the disease as well as the duration each stage of the disease takes.
+The model is described by a large number of parameters that govern the number of
+contacts a person has, the likelihood of becoming infected on each contact, the
+likelihood of developing light or strong symptoms or even dying from the disease as well
+as the duration each stage of the disease takes.
 
-Most of these parameters can be calibrated from existing datasets or the medical literature.
-Only the infection probabilities have to be estimated inside the model by fitting it to time series data of case numbers and fatality rates.
+Most of these parameters can be calibrated from existing datasets or the medical
+literature. Only the infection probabilities have to be estimated inside the model by
+fitting it to time series data of case numbers and fatality rates.
 
@@
 
-We calibrate the parameters for the predicted numbers of contacts from contact diaries of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg \citep{Mossong2008}.
-Each contact diary contains all contacts an individual had throughout one day, including information on the other person (such as age and gender) and information on the contact.
-Importantly, for each contact individuals entered of which type the contact (school, leisure, work etc.) was and how frequent the contact with the other person is.
+We calibrate the parameters for the predicted numbers of contacts from contact diaries
+of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
+\citep{Mossong2008}. Each contact diary contains all contacts an individual had
+throughout one day, including information on the other person (such as age and gender)
+and information on the contact. Importantly, for each contact individuals entered of
+which type the contact (school, leisure, work etc.) was and how frequent the contact
+with the other person is.
 
-Thus, we can use the empirical distributions from this data as pre-pandemic number of contacts.
+Thus, we can use the empirical distributions from this data as pre-pandemic number of
+contacts.
 
@@
 
-As mentioned in section \ref{sec:matching}, the probability that two individuals are matched can depend on background characteristics.
-In particular, we allow this probability to depend on age and county of residence.
-While we do not have good data on geographical assortativeness and just roughly calibrate it such that 60 \% of contacts are within the same county\footnotemark, we can calibrate the age assortativeness from the same data we use to calibrate the number of contacts.
+As mentioned in section \ref{sec:matching}, the probability that two individuals are
+matched can depend on background characteristics. In particular, we allow this
+probability to depend on age and county of residence. While we do not have good data on
+geographical assortativeness and just roughly calibrate it such that 60 \% of contacts
+are within the same county\footnotemark, we can calibrate the age assortativeness from
+the same data we use to calibrate the number of contacts.
 
@@
 
-Figure~\ref{fig:assortativeness} shows that overall contacts within age groups or age groups that are closer to each other are more frequent.
+Figure~\ref{fig:assortativeness} shows that overall contacts within age groups or age
+groups that are closer to each other are more frequent.
 
@@
 
-To calibrate infection probabilities outside of the model, it would be important to know the exact duration and distance of each contact type as well as virus loads.
-Since this is not available in any dataset, we estimate those parameters inside the model with the method of simulated moments \citep{McFadden1989} by minimizing the distance between simulated and observed infection rates.
-Since our model includes a lot of randomness, we average simulated infection rates over several model runs.
+To calibrate infection probabilities outside of the model, it would be important to know
+the exact duration and distance of each contact type as well as virus loads. Since this
+is not available in any dataset, we estimate those parameters inside the model with the
+method of simulated moments \citep{McFadden1989} by minimizing the distance between
+simulated and observed infection rates. Since our model includes a lot of randomness, we
+average simulated infection rates over several model runs.
 
-Currently, we use data for Germany from August to October (both inclusive).
-We do not use earlier periods to save computational time.
-Moreover, we would be worried that the infection probabilities have small seasonal variation that we currently cannot model.
+Currently, we use data for Germany from August to October (both inclusive). We do not
+use earlier periods to save computational time. Moreover, we would be worried that the
+infection probabilities have small seasonal variation that we currently cannot model.
 However, we plan to expand the estimation period soon.
 
-To avoid overfitting and simplify the numerical optimization problem, we only allow for four different probabilities: 1) for contacts in schools, preschools and nurseries.
-2) for work contacts.
-3) for households.
-4) for leisure activities.
+To avoid overfitting and simplify the numerical optimization problem, we only allow for
+four different probabilities: 1) for contacts in schools, preschools and nurseries. 2)
+for work contacts. 3) for households. 4) for leisure activities.

### 2020-12-20 paper/introduction.tex
Enforced 88 character limit.

@@
 
-The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions across the world.
-As a result, many countries were able to reduce the number of positive cases and could relax restrictions imposed on their societies.
-At the same time, the social and economic costs were enormous.
-In the second wave, countries are trying to implement more nuanced policies.
-For example, Germany imposed a ``lockdown-light'' in November.
-Some businesses are closed and certain leisure activities are prohibited but schools and nurseries stay open.
+The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions
+across the world. As a result, many countries were able to reduce the number of positive
+cases and could relax restrictions imposed on their societies. At the same time, the
+social and economic costs were enormous. In the second wave, countries are trying to
+implement more nuanced policies. For example, Germany imposed a ``lockdown-light'' in
+November. Some businesses are closed and certain leisure activities are prohibited but
+schools and nurseries stay open.
 
-Current epidemiological models have not been designed to predict the effect of such fine-grained policies.
-These models need to be extended for each new policy proposal and cannot be easily adapted to fast changing environments.
-This report describes a model that has been designed from the ground up to predict the effects of contact reducing policies in real time.
-At the time of this writing, it has the following features:
+Current epidemiological models have not been designed to predict the effect of such
+fine-grained policies. These models need to be extended for each new policy proposal and
+cannot be easily adapted to fast changing environments. This report describes a model
+that has been designed from the ground up to predict the effects of contact reducing
+policies in real time. At the time of this writing, it has the following features:
 
 \begin{enumerate}
-    \item At the core of the model, people meet people based on a matching algorithm.
-We distinguish various types of contacts.
-Currently, these are households, leisure activities, schools, nurseries and several types of contacts at the workplace.
-Contact types can be random or recurrent and vary in frequency.
-    \item Policies can be implemented as shutting down contact types entirely or partially. The reduction of contacts can be random or systematic, e.g., to allow for essential workers.
-    \item Infection probabilities vary across contact types, but are invariant to policies which reduce contacts.
-    \item The model achieves a good fit on German infection and fatality rate data even if only the infection probabilities are fit to the data and the remaining parameters are calibrated from the medical literature and datasets on contact frequencies.
-    \item High quality Python code for the model is freely available on Github, well documented and very flexible\footnotemark.
-We are actively looking for researchers who want to use our model for their projects.
+    \item At the core of the model, people meet people based on a matching algorithm. We
+          distinguish various types of contacts. Currently, these are households,
+          leisure activities, schools, nurseries and several types of contacts at the
+          workplace. Contact types can be random or recurrent and vary in frequency.
+
+    \item Policies can be implemented as shutting down contact types entirely or
+          partially. The reduction of contacts can be random or systematic, e.g., to
+          allow for essential workers.
+
+    \item Infection probabilities vary across contact types, but are invariant to
+          policies which reduce contacts.
+
+    \item The model achieves a good fit on German infection and fatality rate data even
+          if only the infection probabilities are fit to the data and the remaining
+          parameters are calibrated from the medical literature and datasets on contact
+          frequencies.
+
+    \item High quality Python code for the model is freely available on Github, well
+          documented and very flexible\footnotemark. We are actively looking for
+          researchers who want to use our model for their projects.
 \end{enumerate}
 
-\footnotetext{The code can be found under \url{https://github.com/covid-19-impact-lab/sid} and the documentation with tutorials and background information under \url{https://sid-dev.readthedocs.io/}.}
+\footnotetext{
+    The code can be found under \url{https://github.com/covid-19-impact-lab/sid} and the
+    documentation with tutorials and background information under
+    \url{https://sid-dev.readthedocs.io/}.
+}
 
-In Section~\ref{sec:literature_review}, we give a short overview of epidemiological models and how they relate to our model.
-We continue with a detailed model description in Section~\ref{sec:model} and proceed with a description of model parameters and the estimation in Section~\ref{sec:calibration_and_estimation}.
-The model is validated in Section~\ref{sec:model_validation} by assessing the in-sample fit for reported infections from August to October and by comparing the out-of-sample fit for the period from November until the beginning of Christmas for different lockdown scenarios.
-We conclude in Section~\ref{sec:conclusion}.
+In Section~\ref{sec:literature_review}, we give a short overview of epidemiological
+models and how they relate to our model. We continue with a detailed model description
+in Section~\ref{sec:model} and proceed with a description of model parameters and the
+estimation in Section~\ref{sec:calibration_and_estimation}. The model is validated in
+Section~\ref{sec:model_validation} by assessing the in-sample fit for reported
+infections from August to October and by comparing the out-of-sample fit for the period
+from November until the beginning of Christmas for different lockdown scenarios. We
+conclude in Section~\ref{sec:conclusion}.
 
-This report describes the model in an abstract way, but uses many realistic examples from a version that is specialized to Germany.
-It is important to note that this specialization is not baked into the model or the Python code.
-It is easy to adjust the model to other countries if data on the number of contacts and a dataset with background characteristics are available.
+This report describes the model in an abstract way, but uses many realistic examples
+from a version that is specialized to Germany. It is important to note that this
+specialization is not baked into the model or the Python code. It is easy to adjust the
+model to other countries if data on the number of contacts and a dataset with background
+characteristics are available.
 
-More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\citet{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new report.}
+More details about the German model as well as applications to currently discussed
+German policies can be found in a companion paper
+(\citet{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new
+report.}

### 2020-12-20 paper/literature_review.tex
Enforced 88 character limit.

@@
 
-We build on two strands of literature: Recent extensions of the epidemiological SEIR model and agent-based simulation models.
+We build on two strands of literature: Recent extensions of the epidemiological SEIR
+model and agent-based simulation models.
 
-The traditional SEIR model is not fine-grained enough to model nuanced policies.
-This has motivated a large number of researchers to extend the standard model to allow for more heterogeneity and flexibility.
-Examples are \citet{Grimm2020}, \citet{Donsimoni2020} and \citet{Acemoglu2020} who develop multi group SEIR models to analyze the effects of targeted lockdowns and \citet{Berger2020} who extend the SEIR model to analyze testing and conditional quarantines.
-For a more comprehensive review see \citet{Avery2020}.
-Others have used the results of a standard SEIR model as input for economic models that estimate the cost of policies (e.g.
-\citet{Dorn2020}).
+The traditional SEIR model is not fine-grained enough to model nuanced policies. This
+has motivated a large number of researchers to extend the standard model to allow for
+more heterogeneity and flexibility. Examples are \citet{Grimm2020},
+\citet{Donsimoni2020} and \citet{Acemoglu2020} who develop multi group SEIR models to
+analyze the effects of targeted lockdowns and \citet{Berger2020} who extend the SEIR
+model to analyze testing and conditional quarantines. For a more comprehensive review
+see \citet{Avery2020}. Others have used the results of a standard SEIR model as input
+for economic models that estimate the cost of policies (e.g. \citet{Dorn2020}).
 
-While the popularity of the SEIR model is mainly due to its simplicity, the extensions are quite complex.
-It is unlikely that there will be a SEIR model that combines all proposed extensions.
-Moreover, the extensions do not address other key issues: The main parameter of the SEIR model, the basic reproduction number ($R_0$), is not policy-invariant.
-It is a composite of the number of contacts each person has and the infection probability of the contacts.
-In fact, policy simulations are done by setting $R_0$ to a different value but it is hard to translate a real policy into the value of $R_0$ it will induce.
-In other words, SEIR models are not suited for evaluating the effect of policies which have never been experienced before.
+While the popularity of the SEIR model is mainly due to its simplicity, the extensions
+are quite complex. It is unlikely that there will be a SEIR model that combines all
+proposed extensions. Moreover, the extensions do not address other key issues: The main
+parameter of the SEIR model, the basic reproduction number ($R_0$), is not
+policy-invariant. It is a composite of the number of contacts each person has and the
+infection probability of the contacts. In fact, policy simulations are done by setting
+$R_0$ to a different value but it is hard to translate a real policy into the value of
+$R_0$ it will induce. In other words, SEIR models are not suited for evaluating the
+effect of policies which have never been experienced before.
 
-Another commonly used model class in epidemiology are agent-based simulation models.
-In these models individuals are simulated as moving particles.
-Infections take place when two particles come closer than a certain contact radius (e.g.
-\citet{Silva2020} and \citet{Cuevas2020}).
-While the simulation approach makes it easy to incorporate heterogeneity in disease progression, it is hard to incorporate heterogeneity in meeting patterns.
-Moreover, policies are modeled as changes in the contact radius or momentum equation of the particles.
-The translation from real policies to corresponding model parameters is a hard task.
+Another commonly used model class in epidemiology are agent-based simulation models. In
+these models individuals are simulated as moving particles. Infections take place when
+two particles come closer than a certain contact radius (e.g. \citet{Silva2020} and
+\citet{Cuevas2020}). While the simulation approach makes it easy to incorporate
+heterogeneity in disease progression, it is hard to incorporate heterogeneity in meeting
+patterns. Moreover, policies are modeled as changes in the contact radius or momentum
+equation of the particles. The translation from real policies to corresponding model
+parameters is a hard task.
 
-\citet{Hinch2020} is a recent extension of the prototypical agent-based simulation model that replaces moving particles by contact networks for households, work and random contacts.
-This model is similar in spirit to ours but focuses on contact tracing rather than social distancing policies.
+\citet{Hinch2020} is a recent extension of the prototypical agent-based simulation model
+that replaces moving particles by contact networks for households, work and random
+contacts. This model is similar in spirit to ours but focuses on contact tracing rather
+than social distancing policies.
 
-The above assessment of epidemiological models is not meant as a critique.
-We are aware that those models were not designed to predict the effect of fine-grained social distancing policies in real time and are very well suited to their purpose.
-We invite epidemiologists to provide feedback and collaborate to improve our model.
+The above assessment of epidemiological models is not meant as a critique. We are aware
+that those models were not designed to predict the effect of fine-grained social
+distancing policies in real time and are very well suited to their purpose. We invite
+epidemiologists to provide feedback and collaborate to improve our model.

### 2020-12-20 paper/medical_parameters.tex
Enforced 88 character limit.

@@
 
-This section discusses the medical parameters used in the model, their sources and how we arrived at the distributions used in the model.\footnotemark
+This section discusses the medical parameters used in the model, their sources and how
+we arrived at the distributions used in the model.\footnotemark
 
-\footnotetext{Additional information can be found in the \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.}
+\footnotetext{Additional information can be found in the
+\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
+documentation}.}
 
@@
 
-Estimates of the incubation period usually give a range from 2 to 12 days.
-A meta analysis by \citet{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.
+Estimates of the incubation period usually give a range from 2 to 12 days. A meta
+analysis by \citet{McAloon2020} comes to the conclusion that ``The incubation period
+distribution may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$
+parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55),
+respectively.'' For simplicity we discretize this distribution into four bins.
 
@@
 
-The period between infection and onset of infectiousness is called latent or latency period.
-However, the latency period is rarely given in epidemiological reports on Covid-19.
-Instead, scientists and agencies usually report the incubation period, the period from infection to the onset of symptoms.
-A few studies used measurements of virus shedding to estimate infectiousness during the course of the disease.
-When measurements started before the onset of symptoms the development of the viral load before symptoms gives us an indication of number of days between the onset of infectiousness and symptoms.
+The period between infection and onset of infectiousness is called latent or latency
+period. However, the latency period is rarely given in epidemiological reports on
+Covid-19. Instead, scientists and agencies usually report the incubation period, the
+period from infection to the onset of symptoms. A few studies used measurements of virus
+shedding to estimate infectiousness during the course of the disease. When measurements
+started before the onset of symptoms the development of the viral load before symptoms
+gives us an indication of number of days between the onset of infectiousness and
+symptoms.
 
-The European Centre for Disease Prevention and Control estimates that people become infectious between one and two days before the symptoms set in.
-This is similar to \citet{He2020} who estimate this to take 2.3 days and is in line with \citet{Peak2020}.
+The European Centre for Disease Prevention and Control estimates that people become
+infectious between one and two days before the symptoms set in. This is similar to
+\citet{He2020} who estimate this to take 2.3 days and is in line with \citet{Peak2020}.
 
-Given these numbers and the length of the incubation period we can calculate the latency period for symptomatic people.
-To our knowledge no estimates for the latency period of asymptomatic cases of COVID-19 exist.
-We assume it to be the same for symptomatic and asymptomatic cases.
+Given these numbers and the length of the incubation period we can calculate the latency
+period for symptomatic people. To our knowledge no estimates for the latency period of
+asymptomatic cases of COVID-19 exist. We assume it to be the same for symptomatic and
+asymptomatic cases.
 
-Thus, we arrive at the following distribution for latency periods: 40\% have one day. 35\% have two days. 20\% have three days and 5\% have 5 days.
+Thus, we arrive at the following distribution for latency periods: 40\% have one day.
+35\% have two days. 20\% have three days and 5\% have 5 days.
 
@@
 
-We assume that the duration of infectiousness is the same for both symptomatic and asymptomatic individuals as evidence suggests little differences in the transmission rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).
+We assume that the duration of infectiousness is the same for both symptomatic and
+asymptomatic individuals as evidence suggests little differences in the transmission
+rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients
+(\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic
+individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).
 
@@
 
-For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8 days of infectiousness afterwards.\footnote{Viral loads may be detected much later but 8 days seems to be the time after which most people are culture negative, as also reported by \citet{Singanayagam2020}} Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic (see also figure 5).
-This duration range is very much in line with the meta-analysis’ reported evidence for asymptomatic individuals (see their figure 1).
-Thus, we arrive at 0 to 13 days as the range for infectiousness among individuals who become symptomatic.
-This duration range is very much in line with the meta-analysis' reported evidence for asymptomatic individuals.
+For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8
+days of infectiousness afterwards.\footnote{Viral loads may be detected much later but 8
+days seems to be the time after which most people are culture negative, as also reported
+by \citet{Singanayagam2020}} Thus, we arrive at 0 to 13 days as the range for
+infectiousness among individuals who become symptomatic (see also figure 5). This
+duration range is very much in line with the meta-analysis’ reported evidence for
+asymptomatic individuals (see their figure 1). Thus, we arrive at 0 to 13 days as the
+range for infectiousness among individuals who become symptomatic. This duration range
+is very much in line with the meta-analysis' reported evidence for asymptomatic
+individuals.
 
-Following this evidence we assume the following discretized distribution of the infectiousness period: 10\% of individuals are infectious for three days, 25\% for five days, another 25\% for seven days, 20\% for nine days and 20\% for eleven days.
+Following this evidence we assume the following discretized distribution of the
+infectiousness period: 10\% of individuals are infectious for three days, 25\% for five
+days, another 25\% for seven days, 20\% for nine days and 20\% for eleven days.
 
@@
 
-We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3, Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring symptomatic cases.
+We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
+Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring
+symptomatic cases.
 
-We collapse the data to the following distribution: 10\% recover after 15 days and 30\% require 18, 22 or 27 days respectively.
+We collapse the data to the following distribution: 10\% recover after 15 days and 30\%
+require 18, 22 or 27 days respectively.
 
-These numbers are only used for mild cases.
-We do not disaggregate by age.
-Note that the length of symptoms is not very important in our model given that individuals stop being infectious before their symptoms cease.
+These numbers are only used for mild cases. We do not disaggregate by age. Note that the
+length of symptoms is not very important in our model given that individuals stop being
+infectious before their symptoms cease.
 
@@
 
-The data on how many percent of symptomatic patients will require ICU is pretty thin.
-We rely on data by the US CDC (\citet{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.
-Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of requiring intensive care per age group.
+The data on how many percent of symptomatic patients will require ICU is pretty thin. We
+rely on data by the US CDC (\citet{Stokes2020}) and
+\href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the
+OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the
+probabilities of requiring intensive care per age group.
 
@@
 
-    \tablenotes{The data is taken from \citet{Stokes2020} and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.}
+    \tablenotes{
+        The data is taken from \citet{Stokes2020} and
+        \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the
+        OpenABM-Project}. }
 
@@
 
-For those who will require intensive care we follow \citet{Chen2020} who estimate the time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.
+For those who will require intensive care we follow \citet{Chen2020} who estimate the
+time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.
 
-This aligns well with numbers reported for the time from first symptoms to hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4.
-This is also in line with the durations collected by \href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the Robert Koch Institut}.
+This aligns well with numbers reported for the time from first symptoms to
+hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation
+of 4. This is also in line with the durations collected by
+\href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the
+Robert Koch Institut}.
 
-We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with equal probabilities. These times mostly matter for the ICU capacities.
+We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with
+equal probabilities. These times mostly matter for the ICU capacities.
 
@@
 
-We take the survival probabilities and time to death and time until recovery from intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.
+We take the survival probabilities and time to death and time until recovery from
+intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.
 
-They report time until death to have a mean of 11.74 days and a standard deviation of 8.79 days.
-Approximating this with the normal distribution, we have nearly 10\% probability mass below 0.
-We use it nevertheless as several other distributions (such as chi squared and uniform) were unable to match the variance.
-Discretizing this leads to 41\% of individuals who die from Covid-19 to die after one day in intensive care.
-22\% day after 12 days, 29\% after 20 days and 7\% after 32 days.
-Again, we rescale this for every age group among those that will not survive.
+They report time until death to have a mean of 11.74 days and a standard deviation of
+8.79 days. Approximating this with the normal distribution, we have nearly 10\%
+probability mass below 0. We use it nevertheless as several other distributions (such as
+chi squared and uniform) were unable to match the variance. Discretizing this leads to
+41\% of individuals who die from Covid-19 to die after one day in intensive care. 22\%
+day after 12 days, 29\% after 20 days and 7\% after 32 days. Again, we rescale this for
+every age group among those that will not survive.
 
-They report time until recovery to have a mean of 18.8 days and a standard deviation of 12.21 days.
-Approximating this with the normal distribution, we have over 5\% probability mass below 0.
-Discretizing this of those who recover in intensive care 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.
+They report time until recovery to have a mean of 18.8 days and a standard deviation of
+12.21 days. Approximating this with the normal distribution, we have over 5\%
+probability mass below 0. Discretizing this of those who recover in intensive care 22\%
+do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

### 2020-12-20 paper/model/course_of_disease.tex
Clean up preamble, fix typos and add section on initial conditions.
Move to Holger's templates and some other - cannot remember - fixes.
more fixes.
Enforced 88 character limit.

@@
 
-The following medical parameters describing the progression of the disease are taken from systematic reviews (e.g.
-\cite{He2020}).
-After an infection occurs, the disease progresses in the way depicted in Figure~\ref{fig:course_of_disease}.
+The following medical parameters describing the progression of the disease are taken
+from systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
+progresses in the way depicted in Figure~\ref{fig:course_of_disease}.
 
-\begin{figure}[!ht]
-    \includegraphics[width=0.7\textwidth]{../figures/disease_progression.png}
+\begin{figure}[!tp]
+    \includegraphics[width=\textwidth]{../figures/disease_progression.png}
     \caption{Course of Disease in the model}
@@
 
-First, infected individuals will become infectious after one to five days.
-About one third of people stays asymptomatic.
-The rest develops symptoms about one to two days after they become infectious.
-Modeling asymptomatic and pre-symptomatic cases is important because those people do not reduce their contacts or demand a test and can potentially infect many other people \citep{Donsimoni2020}.
+First, infected individuals will become infectious after one to five days. About one
+third of people stays asymptomatic. The rest develops symptoms about one to two days
+after they become infectious. Modeling asymptomatic and pre-symptomatic cases is
+important because those people do not reduce their contacts or demand a test and can
+potentially infect many other people \citep{Donsimoni2020}.
 
-A small share of symptomatic people will develop strong symptoms that require intensive care.
-The exact share and time span is age-dependent.
-An age-dependent share of intensive care unit (ICU) patients will die after spending up to 32 days in intensive care.
-Moreover, if the ICU capacity was reached, all patients who require intensive care but do not receive it die.
+A small share of symptomatic people will develop strong symptoms that require intensive
+care. The exact share and time span is age-dependent. An age-dependent share of
+intensive care unit (ICU) patients will die after spending up to 32 days in intensive
+care. Moreover, if the ICU capacity was reached, all patients who require intensive care
+but do not receive it die.
 
-It would be easy to make the course of disease even more fine grained.
-For example, we could model people who require hospitalization but not intensive care.
-So far we opted against that because only the ICU capacity might become a bottleneck in Germany.
+It would be easy to make the course of disease even more fine-grained. For example, we
+could model people who require hospitalization but not intensive care. So far we opted
+against that because only the ICU capacity might become a bottleneck in Germany.
 
-We allow the progression of the disease to be stochastic in two ways: Firstly, state changes only occur with a certain probability (e.g.
-only a fraction of infected individuals develops symptoms).
-Secondly, the number of periods for which an individual remains in a state is drawn randomly.
-The parameters that govern these processes are taken from the literature.
-They can vary with the age of an individual.
+We allow the progression of the disease to be stochastic in two ways: Firstly, state
+changes only occur with a certain probability (e.g. only a fraction of infected
+individuals develops symptoms). Secondly, the number of periods for which an individual
+remains in a state is drawn randomly. The parameters that govern these processes are
+taken from the literature. They can vary with the age of an individual.
 
-Detailed information on the calibration of the disease parameters is available as part of our \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.
+Detailed information on the calibration of the disease parameters is available as part
+of our
+\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
+documentation}.

### 2020-12-20 paper/model/summary.tex
Move to Holger's templates and some other - cannot remember - fixes.
Enforced 88 character limit.

@@
 
-To alleviate these problems we propose a different model structure.
-Our model inherits many features from agent-based simulation models but replaces the contacts between moving particles by contacts between individuals who work, go to school, live in a household and enjoy leisure activities.
-The structure of the model is depicted in Figure~\ref{fig:model_graph}.
+To alleviate these problems we propose a different model structure. Our model inherits
+many features from agent-based simulation models but replaces the contacts between
+moving particles by contacts between individuals who work, go to school, live in a
+household and enjoy leisure activities. The structure of the model is depicted in
+Figure~\ref{fig:model_graph}.
 
-\begin{figure}[!ht]
+\begin{figure}[!tp]
     \centering
-    \includegraphics[width=0.7\textwidth]{../figures/model_detailed.png}
+    \includegraphics[width=\textwidth]{../figures/model_detailed.png}
     \caption{Simplified graph of the model}
@@
 
-The background characteristics include age, county and occupation of each simulated individual.
-Contact models are functions that map individual characteristics into a predicted number of contacts.
-Currently we distinguish between eight types of contact models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and random work contacts, recurrent and random leisure contacts, and nursery, preschool, and school contacts.
+The background characteristics include age, county and occupation of each simulated
+individual. Contact models are functions that map individual characteristics into a
+predicted number of contacts. Currently we distinguish between eight types of contact
+models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and
+random work contacts, recurrent and random leisure contacts, and nursery, preschool, and
+school contacts.
 
 The predicted number of contacts is translated into infections by a matching algorithm.
-There are different matching algorithms for recurrent contacts (e.g.
-classmates, family members) and non-recurrent contacts (e.g.
-clients, contacts in supermarkets).
-The infection probability can differ for each contact type.
-All types of contacts can be assortative with respect to geographic and demographic characteristics.
+There are different matching algorithms for recurrent contacts (e.g. classmates, family
+members) and non-recurrent contacts (e.g. clients, contacts in supermarkets). The
+infection probability can differ for each contact type. All types of contacts can be
+assortative with respect to geographic and demographic characteristics.
 
-Once a person is infected, the disease progresses in a fairly standard way which is also depicted in Figure~\ref{fig:course_of_disease}.
-Asymptomatic cases and cases with mild symptoms are infectious for some time and recover eventually.
-Cases with severe symptoms additionally require hospitalization and lead to either recovery or death.
+Once a person is infected, the disease progresses in a fairly standard way which is also
+depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild
+symptoms are infectious for some time and recover eventually. Cases with severe symptoms
+additionally require hospitalization and lead to either recovery or death.
 
-There are two possible ways to enable testing for Covid-19 in the model.
-The first way allows to specify the ratio of known to unknown infections for each day.
-A random sample of all newly infected individuals will then receive a test result in the following days.
-The second approach consists of three steps.
-First, individuals demand a test because they, e.g., experience symptoms.
-Secondly, tests are allocated to individuals while respecting governmental access restrictions to test.
-Thirdly, depending of the capacities of laboratories, tests are processed for some time until the individual receives her test result.
+There are two possible ways to enable testing for Covid-19 in the model. The first way
+allows to specify the ratio of known to unknown infections for each day. A random sample
+of all newly infected individuals will then receive a test result in the following days.
+The second approach consists of three steps. First, individuals demand a test because
+they, e.g., experience symptoms. Secondly, tests are allocated to individuals while
+respecting governmental access restrictions to test. Thirdly, depending of the
+capacities of laboratories, tests are processed for some time until the individual
+receives her test result.
 
-In addition, people who have symptoms, received a positive test, or had a risk contact can reduce their number of contacts across all contact types endogenously.
+In addition, people who have symptoms, received a positive test, or had a risk contact
+can reduce their number of contacts across all contact types endogenously.
 
-The model makes it very simple to translate policies into model quantities.
-For example, school closures imply the complete suspension of school contacts.
-A strict lockdown implies shutting down work contacts of all people who are not employed in a systemically relevant sector.
-It is also possible to have more sophisticated policies that condition the number of contacts on observable characteristics, risk contacts or health states.
+The model makes it very simple to translate policies into model quantities. For example,
+school closures imply the complete suspension of school contacts. A strict lockdown
+implies shutting down work contacts of all people who are not employed in a systemically
+relevant sector. It is also possible to have more sophisticated policies that condition
+the number of contacts on observable characteristics, risk contacts or health states.
 
-Another key advantage of the model is that the number of contacts an individual has of each contact type can be calibrated from publicly available data \citep{Mossong2008}.
-This in turn allows us to estimate policy-invariant infection probabilities from time series of infection and death rates using the method of simulated moments \citep{McFadden1989}.
-Since the infection probabilities are time-invariant, data collected since the beginning of the pandemic can be used for estimation.
-Moreover, since we can model the testing strategies that were in place at each point in time, we can correct the estimates for the fact that not all infections are observed.
+Another key advantage of the model is that the number of contacts an individual has of
+each contact type can be calibrated from publicly available data \citep{Mossong2008}.
+This in turn allows us to estimate policy-invariant infection probabilities from time
+series of infection and death rates using the method of simulated moments
+\citep{McFadden1989}. Since the infection probabilities are time-invariant, data
+collected since the beginning of the pandemic can be used for estimation. Moreover,
+since we can model the testing strategies that were in place at each point in time, we
+can correct the estimates for the fact that not all infections are observed.
 
-Last but not least, performing simulations whose starting point is set amidst the pandemic requires special adjustments to arrive at a realistic distribution of courses of diseases.
-We solve the initial conditions problem by matching reported infections to individuals in our data while also correcting for reporting lag and undetected cases.
+Last but not least, performing simulations whose starting point is set amidst the
+pandemic requires special adjustments to arrive at a realistic distribution of courses
+of diseases. We solve the initial conditions problem by matching reported infections to
+individuals in our data while also correcting for reporting lag and undetected cases.
 

### 2020-12-20 paper/calibration_and_estimation.tex
Update figures and pseudocode.

@@
     \includegraphics[width=\textwidth]{../figures/assortative_matching_probability_example.png}
-    \caption{Distribution of random non-work contacts by age of participants}
+    \caption{Distribution of random non-work contacts by age of participants.}
     \label{fig:assortativeness}
+    \figurenotes{
+        The figure shows the distribution of random non-work contacts by age groups. A
+        row shows the share of contacts a certain age group has with all other age
+        groups. Higher values are colored in darker red tones. The diagonal represents
+        the share of contacts with individuals from the same age group.
+    }
 \end{figure}
 
-Figure~\ref{fig:assortativeness} shows that overall contacts within age groups or age
-groups that are closer to each other are more frequent.
+Figure~\ref{fig:assortativeness} shows that assortativeness by age is especially strong
+for children and younger adults. For older people, the pattern becomes more dispersed
+around their own age group, but within-age-group contacts are still the most common
+contacts.
 

### 2020-12-20 paper/model/matching.tex
Enforced 88 character limit.
Update figures and pseudocode.

@@
 
-The empirical data described above only allows to estimate the number of contacts each person has.
-In order to simulate transmissions of Covid-19, the numbers of contacts has to be translated into actual meetings between people.
-This is achieved by matching algorithms:
+The empirical data described above only allows to estimate the number of contacts each
+person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
+to be translated into actual meetings between people. This is achieved by matching
+algorithms:
 
-As described in section \ref{sec:number_of_contacts}, some contact types are recurrent (i.e.
-the same people meet regularly), others are non-recurrent (i.e.
-it would only be by accident that two people meet twice).
-The matching process is different for recurrent and non recurrent contact models.
+As described in section \ref{sec:number_of_contacts}, some contact types are recurrent
+(i.e. the same people meet regularly), others are non-recurrent (i.e. it would only be
+by accident that two people meet twice). The matching process is different for recurrent
+and non recurrent contact models.
 
-Recurrent contacts are described by two components: 1) A variable in the background characteristics.
-An example would be a school class identifier which could come from actual data or be drawn randomly to achieve representative class sizes.
-2) A deterministic or random function that takes the value 0 (non-participating) and 1 (participating) and can depend on the weekday, date and health state.
-This can be used to model vacations, weekends or symptomatic people who stay home (see section \ref{sec:endogenous_contact_reductions} for details).
+Recurrent contacts are described by two components: 1) A variable in the background
+characteristics. An example would be a school class identifier which could come from
+actual data or be drawn randomly to achieve representative class sizes. 2) A
+deterministic or random function that takes the value 0 (non-participating) and 1
+(participating) and can depend on the weekday, date and health state. This can be used
+to model vacations, weekends or symptomatic people who stay home (see section
+\ref{sec:endogenous_contact_reductions} for details).
 
-The matching process for recurrent contacts is then extremely simple: On each simulated day, every person who does not stay home meets all other group members who do not stay home.
-The assumption that all group members have contacts with all other group members is not fully realistic, but seems like a good approximation to reality, especially in light of the suspected role of aerosol transmission for Covid-19 (\citet{Morawska2020, Anderson2020}).
+The matching process for recurrent contacts is then extremely simple: On each simulated
+day, every person who does not stay home meets all other group members who do not stay
+home. The assumption that all group members have contacts with all other group members
+is not fully realistic, but seems like a good approximation to reality, especially in
+light of the suspected role of aerosol transmission for Covid-19 (\citet{Morawska2020,
+Anderson2020}).
 
-The matching in non-recurrent contact models is more difficult and implemented in a two stage sampling procedure to allow for assortative matching.
-Currently most contact models are assortative with respect to age (it is more likely to meet people from the same age group) and county (it is more likely to meet people from the same county) but in principle any set of discrete variables can be used.
-This set of variables that influence matching probabilities introduce a discrete partition of the population into groups.
-The first stage of the two stage sampling process samples on the group level.
+The matching in non-recurrent contact models is more difficult and implemented in a two
+stage sampling procedure to allow for assortative matching. Currently most contact
+models are assortative with respect to age (it is more likely to meet people from the
+same age group) and county (it is more likely to meet people from the same county) but
+in principle any set of discrete variables can be used. This set of variables that
+influence matching probabilities introduce a discrete partition of the population into
+groups. The first stage of the two stage sampling process samples on the group level.
 The second stage on the individual level.
 
-Below, we first show pseudo code for the non-recurrent matching algorithm and then describe how the algorithm works in words.
+Below, we first show pseudo code for the non-recurrent matching algorithm and then
+describe how the algorithm works in words.
 
-%%% ToDo: Get syntax highlighting for the pseudo code. Try out different languages.
+\begin{listing}
+    \label{code:matching}
+    \begin{minted}{python}
+        while are_unmatched_contacts_left:
 
-\begin{lstlisting}
-while unmatched contacts remain:
-    model, id = draw contact model and individual
-    for contact in remaining_contacts[id, model]:
-        draw group of other person
-        draw other person from that group
-        determine if id infects other or vice versa
-        remaining_contacts[id, model] -= 1
-        remaining_contacts[other, model] -= 1
-\end{lstlisting}
+            contact_type, i = draw_contact_type_and_individual()
 
-We first randomly draw a contact type and individual.
-For each contact oth the drawn contact type that person has, we first draw the group of the other person (first stage).
-Next, we calculate the probability to be drawn for each member of the group, based on the number of remaining contacts.
-I.e.
-people who have more remaining contacts are drawn with a higher probability.
-This has to be re-calculated each time because with each matched contact, the number of remaining contacts changes.
-We then draw the other individual, determine whether an infection takes place and if so update the health states of the newly infected person.
-Finally, we reduce the number of remaining contacts of the two matched individuals by one.
+            for _ in remaining_contacts[i, contact_type]:
 
-The recalculation of matching probabilities in the second stage is computationally intensive because it requires summing up all remaining contacts in that group.
-Using a two stage sampling process where the first stage probabilities remain constant over time makes the matching computationally much more tractable because the number of computations increases quadratically in the second stage group size.
+                group_j = draw_group_of_other_person()
+                j = draw_other_person_from_that_group(group_j)
+
+                if infection_takes_places(i, j):
+                    # Update health statuses.
+
+                remaining_contacts[i, contact_type] -= 1
+                remaining_contacts[j, contact_type] -= 1
+    \end{minted}
+    \caption{Pseudo-code for the matching algorithm of non-recurrent contacts.}
+\end{listing}
+
+We first randomly draw a contact type and individual. For each contact of the drawn
+contact type that person has, we first draw the group of the other person (first stage).
+Next, we calculate the probability to be drawn for each member of the group, based on
+the number of remaining contacts. I.e. people who have more remaining contacts are drawn
+with a higher probability. This has to be re-calculated each time because with each
+matched contact, the number of remaining contacts changes. We then draw the other
+individual, determine whether an infection takes place and if so update the health
+states of the newly infected person. Finally, we reduce the number of remaining contacts
+of the two matched individuals by one.
+
+The recalculation of matching probabilities in the second stage is computationally
+intensive because it requires summing up all remaining contacts in that group. Using a
+two stage sampling process where the first stage probabilities remain constant over time
+makes the matching computationally much more tractable because the number of
+computations increases quadratically in the second stage group size.

### 2020-12-20 paper/model_validation.tex
Enforced 88 character limit.
Update figures and pseudocode.

@@
 
-We validate our model in two ways: 1) We look at the in-sample fit over the estimation period.
-2) We look at the out-of-sample fit for November.
-The last one is a challenging test for our model because there was a strong policy change between the estimation period and November.
-The model convincingly passes both tests.
+We validate our model in two ways: 1) We look at the in-sample fit over the estimation
+period. 2) We look at the out-of-sample fit for November. The last one is a challenging
+test for our model because there was a strong policy change between the estimation
+period and November. The model convincingly passes both tests.
 
@@
 
-Despite fitting only four free parameters, the in-sample fit is very good.
-The best fit is achieved in the largest age groups.
-This is so mechanically, because we weight the deviations between simulated and observed infection rates by group sizes.
-The worst fit is achieved for the 80 to 100 years old.
-The model predicts too few infections for these groups because they have very few contacts in all contact types we have included so far.
-We plan to address this issue soon by adding another contact type that captures all contacts in the data by \citet{Mossong2008} that we have not included so far.
-Moreover, we expect an improved model fit when we allow for more free parameters.
+Despite fitting only four free parameters, the in-sample fit is very good. The best fit
+is achieved in the largest age groups. This is so mechanically, because we weight the
+deviations between simulated and observed infection rates by group sizes. The worst fit
+is achieved for the 80 to 100 years old. The model predicts too few infections for these
+groups because they have very few contacts in all contact types we have included so far.
+We plan to address this issue soon by adding another contact type that captures all
+contacts in the data by \citet{Mossong2008} that we have not included so far. Moreover,
+we expect an improved model fit when we allow for more free parameters.
 
@@
     \includegraphics[width=\textwidth]{../figures/goodness_of_fit_by_age_group}
-    \caption{Actual vs. simulated infection and fatality rates}
+    \caption{Reported vs. simulated weekly incidence rates of infections.}
     \label{fig:goodness_of_fit}
+    \figurenotes{
+        The figure shows the weekly incidence rates per 100,000 people for the reported
+        (red line) versus the simulated infections rates (blue line) for age groups
+        available in the data provided by the Robert-Koch Institut.
+    }
 \end{figure}
@@
 
-We can assess the out-of-sample fit by projecting the effect of the lockdown light and comparing it to case numbers until now.
-It is important to note that this is not just a simple extrapolation of a time trend because the lockdown light only started after the estimation period.
-The out-of-sample fit can be assessed in Figure~\ref{fig:out-of-sample-fit}.
+We can assess the out-of-sample fit by projecting the effect of the lockdown light and
+comparing it to case numbers until now. It is important to note that this is not just a
+simple extrapolation of a time trend because the lockdown light only started after the
+estimation period. The out-of-sample fit can be assessed in
+Figure~\ref{fig:out-of-sample-fit}.
 
@@
     \includegraphics[width=\textwidth]{../figures/out_of_sample_validation}
-    \caption{Projected effect of the lockdown-light}
+    \caption{Predicted effect of the "Lockdown Light" on infection rates.}
     \label{fig:out-of-sample-fit}
+    \figurenotes{
+        For the time period until the beginning of November, the figure shows the weekly
+        incidence rates of infections per 100,000 people from reported (black) versus
+        simulated (blue) data. With the start of November, the projections of the three
+        scenarios, optimistic (blue), neutral (red), and pessimistic (mint green), are
+        shown until the beginning of the Christmas holidays. The actual incidence rates
+        (black) are reported until the 24th November.
+    }
 \end{figure}
@@
 The model correctly predicts the effect of the lockdown light with reasonable accuracy.
-In particular, the actual case numbers are between our neutral and pessimistic projection.
-The plot also shows that ending the lockdown light as planned on November 30 would lead to an explosive growth in case numbers in all scenarios.
+In particular, the actual case numbers are between our neutral and pessimistic
+projection. The plot also shows that ending the lockdown light as planned on November 30
+would lead to an explosive growth in case numbers in all scenarios.

### 2020-12-20 task_compile_documents.py

@@ -6,7 +6,9 @@ ROOT = Path(__file__).parent
 BLD = ROOT / "bld"
 
 
-@pytask.mark.latex
+@pytask.mark.latex(
+    ["--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd", "--shell-escape"]
+)
 @pytask.mark.depends_on(
     [ROOT / "paper" / "paper.tex"] + list(ROOT.joinpath("paper").rglob("*.tex"))
 )

### 2020-12-20 paper/abstract.tex
Enforced 88 character limit.
small adjustments.

@@
-\noindent Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs.
-Epidemiological models have a hard time predicting the effects of such fine-grained policies.
-We propose a novel simulation-based model to address this shortcoming.
-We build on state-of-the-art agent-based simulation models but replace the way contacts between susceptible and infected people take place.
-Firstly, we allow for heterogeneity in the types of contacts (e.g.
-recurrent or random) and in the infectiousness of each contact type.
-Secondly, we strictly separate the number of contacts from the probabilities that a contact leads to an infection.
-The number of contacts changes with social distancing policies, the infection probabilities remain invariant.
-This allows us to model many types of fine-grained policies that cannot easily be incorporated into other models.
-To validate our model, we show that it can accurately predict the effect of the German November lockdown even if no similar policy has been observed in the time series that were used to estimate the model parameters.
+\noindent Governments worldwide are adopting nuanced policy measures to reduce the
+number of Covid-19 cases and to balance social and economic costs. Epidemiological
+models have a hard time predicting the effects of fine-grained policies. We propose a
+novel simulation-based model to address this shortcoming. We build on state-of-the-art
+agent-based simulation models but replace the way contacts between susceptible and
+infected people take place. Firstly, we allow for heterogeneity in the types of contacts
+(e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we
+strictly separate the number of contacts from the probabilities that a contact leads to
+an infection. The number of contacts changes with social distancing policies, the
+infection probabilities remain invariant. This allows us to model many types of targeted
+policies that cannot be incorporated into other models. To validate our model, we show
+that it can predict the effect of the German November lockdown even if no similar policy
+has been observed during the time period that was used to estimate the model parameters.
 

### 2020-12-20 paper/introduction.tex
small adjustments.

@@
 The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions
-across the world. As a result, many countries were able to reduce the number of positive
-cases and could relax restrictions imposed on their societies. At the same time, the
-social and economic costs were enormous. In the second wave, countries are trying to
-implement more nuanced policies. For example, Germany imposed a ``lockdown-light'' in
-November. Some businesses are closed and certain leisure activities are prohibited but
-schools and nurseries stay open.
+across the world. Many countries were able to contain the spread of the virus and to
+relax restrictions during midyear. At the same time, the social and economic costs were
+enormous. In the second wave, countries are trying to implement more nuanced policies.
+For example, Germany imposed a ``lockdown-light'' in November. Some businesses are
+closed and certain leisure activities are prohibited but educational facilities remain
+open.
 
-Current epidemiological models have not been designed to predict the effect of such
-fine-grained policies. These models need to be extended for each new policy proposal and
+Current epidemiological models have not been designed to predict the effect of
+fine-grained policies. The models need to be extended for each new policy proposal and
 cannot be easily adapted to fast changing environments. This report describes a model
 that has been designed from the ground up to predict the effects of contact reducing
-policies in real time. At the time of this writing, it has the following features:
+policies in real time. It has the following features:
 
@@
     \item At the core of the model, people meet people based on a matching algorithm. We
-          distinguish various types of contacts. Currently, these are households,
-          leisure activities, schools, nurseries and several types of contacts at the
-          workplace. Contact types can be random or recurrent and vary in frequency.
+          distinguish various types of contacts. The contact types are households,
+          leisure activities, schools, preschools, and nurseries and several types of
+          contacts at the workplace. Contact types can be random or recurrent and vary
+          in frequency.
 
@@
 
-    \item Infection probabilities vary across contact types, but are invariant to
-          policies which reduce contacts.
+    \item Infection probabilities vary across contact types and reflect properties of
+          the contact like the location (indoor/outdoor) and the kind of interaction
+          (duration, physical contact). The probabilities are independent from the
+          number of contacts and, thus, policy-invariant.
 
-    \item The model achieves a good fit on German infection and fatality rate data even
-          if only the infection probabilities are fit to the data and the remaining
-          parameters are calibrated from the medical literature and datasets on contact
+    \item The model achieves a good fit on German data of infection rates even if only
+          the infection probabilities are fit to the data and the remaining parameters
+          are calibrated from the medical literature and datasets on contact
           frequencies.
@@
           documented and very flexible\footnotemark. We are actively looking for
-          researchers who want to use our model for their projects.
+          researchers who want to use our model for their projects and employ it in
+          different contexts.
 \end{enumerate}
@@
 }
+
+This report describes the model in an abstract way, but uses many realistic examples
+from a version that is specialized to Germany. It is important to note that this
+specialization is not baked into the model or the Python code. It is easy to adjust the
+model to other countries if data on the number of contacts and a dataset with background
+characteristics are available.
+
+More details about the German model as well as applications to currently discussed
+German policies can be found in \citet{Dorn2020a}.
 
@@
 conclude in Section~\ref{sec:conclusion}.
-
-This report describes the model in an abstract way, but uses many realistic examples
-from a version that is specialized to Germany. It is important to note that this
-specialization is not baked into the model or the Python code. It is easy to adjust the
-model to other countries if data on the number of contacts and a dataset with background
-characteristics are available.
-
-More details about the German model as well as applications to currently discussed
-German policies can be found in a companion paper
-(\citet{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new
-report.}

### 2020-12-20 paper/model/matching.tex
small adjustments.

@@
 is not fully realistic, but seems like a good approximation to reality, especially in
-light of the suspected role of aerosol transmission for Covid-19 (\citet{Morawska2020,
-Anderson2020}).
+light of the suspected role of aerosol transmission for Covid-19 \citep{Morawska2020,
+Anderson2020}.
 

### 2021-01-17 paper/introduction.tex
Release v2.0.0.

@@
           (duration, physical contact). The probabilities are independent from the
-          number of contacts and, thus, policy-invariant.
+          number of contacts and thus policy-invariant.
 
@@
           documented and very flexible\footnotemark. We are actively looking for
-          researchers who want to use our model for their projects and employ it in
-          different contexts.
+          researchers who want to use our model for their projects and apply it to other
+          contexts.
 \end{enumerate}
@@
 In Section~\ref{sec:literature_review}, we give a short overview of epidemiological
-models and how they relate to our model. We continue with a detailed model description
-in Section~\ref{sec:model} and proceed with a description of model parameters and the
-estimation in Section~\ref{sec:calibration_and_estimation}. The model is validated in
+models. We continue with a detailed description of our model in Section~\ref{sec:model}
+and proceed with a description of model parameters and the estimation in
+Section~\ref{sec:calibration_and_estimation}. The model is validated in
 Section~\ref{sec:model_validation} by assessing the in-sample fit for reported

### 2021-01-17 paper/medical_parameters.tex
Release v2.0.0.

@@
 For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8
-days of infectiousness afterwards.\footnote{Viral loads may be detected much later but 8
-days seems to be the time after which most people are culture negative, as also reported
-by \citet{Singanayagam2020}} Thus, we arrive at 0 to 13 days as the range for
-infectiousness among individuals who become symptomatic (see also figure 5). This
-duration range is very much in line with the meta-analysis’ reported evidence for
+days of infectiousness afterwards.\footnotemark Thus, we arrive at 0 to 13 days as the
+range for infectiousness among individuals who become symptomatic (see also figure 5).
+This duration range is very much in line with the meta-analysis’ reported evidence for
 asymptomatic individuals (see their figure 1). Thus, we arrive at 0 to 13 days as the
@@
 individuals.
+
+\footnotetext{
+    Viral loads may be detected much later but 8 days seems to be the time after which
+    most people are culture negative, as also reported by \citet{Singanayagam2020}.
+}
 
@@
 \href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the
-Robert Koch Institut}.
+Robert-Koch-Institut}.
 
@@
 probability mass below 0. We use it nevertheless as several other distributions (such as
-chi squared and uniform) were unable to match the variance. Discretizing this leads to
-41\% of individuals who die from Covid-19 to die after one day in intensive care. 22\%
-day after 12 days, 29\% after 20 days and 7\% after 32 days. Again, we rescale this for
-every age group among those that will not survive.
+chi squared and uniform) were unable to match the variance. Discretizing the
+distribution leads to 41\% of individuals who will die from Covid-19 after one day in
+intensive care, 22\% day after 12 days, 29\% after 20 days and 7\% after 32 days. Again,
+we rescale this for every age group among those that will not survive.
 
-They report time until recovery to have a mean of 18.8 days and a standard deviation of
+They report a mean duration of 18.8 days until recovery and a standard deviation of
 12.21 days. Approximating this with the normal distribution, we have over 5\%
-probability mass below 0. Discretizing this of those who recover in intensive care 22\%
-do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.
+probability mass below 0. Of those who recover in intensive care, 22\% do so
+after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

### 2021-01-17 paper/model/course_of_disease.tex
Release v2.0.0.

@@
     \includegraphics[width=\textwidth]{../figures/disease_progression.png}
-    \caption{Course of Disease in the model}
+    \caption{Course of Disease in the model.}
     \label{fig:course_of_disease}
+    \figurenotes{
+        The figure depicts the course of the disease from infection to either the state
+        of recovery or death.
+    }
 \end{figure}
@@
 remains in a state is drawn randomly. The parameters that govern these processes are
-taken from the literature. They can vary with the age of an individual.
+taken from the literature\footnotemark and age-dependent.
 
-Detailed information on the calibration of the disease parameters is available as part
-of our
-\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-documentation}.
+\footnotetext{
+    Detailed information on the calibration of the disease parameters is available as
+    part of our
+    \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.
+}

### 2021-01-17 paper/model/number_of_contacts.tex
Enforced 88 character limit.
Update figures and pseudocode.
Release v2.0.0.

@@
 
-Consider a hypothetical population of 1,000 individuals in which 50 were infected with a novel infectious disease.
-From this data alone, it is impossible to say whether only those 50 people had contact with an infectious person and the disease has an infection probability of 1 in each contact or whether everyone met an infectious person but the disease has an infection probability of only 5 percent per contact.
-SEIR models do not even try to distinguish contact frequency from the infectiousness of each contact and combine the two in one parameter that is not invariant to social distancing policies.
+Consider a hypothetical population of 1,000 individuals in which 50 were infected with a
+novel infectious disease. From this data alone, it is impossible to say whether only
+those 50 people had contact with an infectious person and the disease has an infection
+probability of 1 in each contact or whether everyone met an infectious person but the
+disease has an infection probability of only 5 percent per contact. SEIR models do not
+even try to distinguish contact frequency from the infectiousness of each contact and
+combine the two in one parameter that is not invariant to social distancing policies.
 
-To model social distancing policies, we need to disentangle the effects of the number of contacts of each individual and the effect of policy-invariant infection probabilities specific to each contact type.
-Since not all contacts are equally infectious, we distinguish different contact types.
+To model social distancing policies, we need to disentangle the effects of the number of
+contacts of each individual and the effect of policy-invariant infection probabilities
+specific to each contact type. Since not all contacts are equally infectious, we
+distinguish different contact types.
 
-The number and type of contacts in our model can be easily extended.
-Each type of contacts is described by a function that maps individual characteristics, health states and the date into a number of planned contacts for each individual.
-This allows to model a wide range of contact types.
+The number and type of contacts in our model can be easily extended. Each type of
+contacts is described by a function that maps individual characteristics, health states
+and the date into a number of planned contacts for each individual. This allows to model
+a wide range of contact types.
 
@@
 \begin{itemize}
-    \item Households: Each household member meets all other household members every day. The household sizes and structures are calibrated to be representative for Germany.
+    \item Households: Each household member meets all other household members every day.
+    The household sizes and structures are calibrated to be representative for Germany.
 
-    \item Random non-work contacts: Each person has contacts with randomly drawn other people which are assortative with respect to region and age group.
-This contact type reflects contacts during pure leisure activities as well as non leisure activities such as grocery shopping or medical appointments.
+    \item Random non-work contacts: Each person has contacts with randomly drawn other
+    people which are assortative with respect to region and age group. This contact type
+    reflects contacts during pure leisure activities as well as non leisure activities
+    such as grocery shopping or medical appointments.
 
-    \item Recurrent daily non-work contacts: Each person has daily recurring contacts which allows to model close relationships other than families between individuals.
+    \item Recurrent daily non-work contacts: Each person has daily recurring contacts
+    which allows to model close relationships other than families between individuals.
 
-    \item Recurrent weekly non-work contacts: Each person has weekly recurring contacts like sports groups or other weekly activities.
+    \item Recurrent weekly non-work contacts: Each person has weekly recurring contacts
+    like sports groups or other weekly activities.
 
-    \item Random work contacts: Each working adult has contact to randomly drawn other people at work.
+    \item Random work contacts: Each working adult has contact to randomly drawn other
+    people at work.
 
-    \item Recurrent daily work contacts: Each working adult meets other workers every day. This is meant to capture work colleagues.
+    \item Recurrent daily work contacts: Each working adult meets other workers every
+    day. This is meant to capture work colleagues.
 
-    \item Recurrent weekly work contacts: Each working adult meets other workers once per week.
-We randomize over the days on which the meetings take place.
-This is meant to capture meetings with clients, superiors or other colleagues which happen infrequently.
+    \item Recurrent weekly work contacts: Each working adult meets other workers once
+    per week. We randomize over the days on which the meetings take place. This is meant
+    to capture meetings with clients, superiors or other colleagues which happen
+    infrequently.
 
-    \item Schools: Each student meets all of his classmates every day.
-Class sizes are calibrated to be representative for Germany and students have the same age.
-Schools are closed on weekends and during vacations, which vary by states.
-School classes also meet three pairs of teachers every school day.
-The pairs are meant to represent interactions between teachers.
+    \item Schools: Each student meets all of his classmates every day. Class sizes are
+    calibrated to be representative for Germany and students have the same age. Schools
+    are closed on weekends and during vacations, which vary by states. School classes
+    also meet three pairs of teachers every school day. The pairs are meant to represent
+    interactions between teachers.
 
-    \item Preschools: Children who are at least three years old and younger than six may attend preschool.
-Each group of nine children interacts with the same two adults every day.
-The children in each group are of the same age.
-The remaining mechanics are similar to schools.
+    \item Preschools: Children who are at least three years old and younger than six may
+    attend preschool. Each group of nine children interacts with the same two adults
+    every day. The children in each group are of the same age. The remaining mechanics
+    are similar to schools.
 
-    \item Nurseries: Children younger than three years may attend a nursery and interact with one adult. The age of the children varies within groups. The remaining mechanics are similar to schools.
+    \item Nurseries: Children younger than three years may attend a nursery and interact
+    with one adult. The age of the children varies within groups. The remaining
+    mechanics are similar to schools.
 \end{itemize}
 
-The number of random and recurrent contacts at the workplace and at home is calibrated with data provided by \citet{Mossong2008}.
-For details see Section~\ref{sec:calibration}.
-In particular, we sample the number of contacts or group sizes from empirical distributions that sometimes depend on age.
-Instead it would also be possible to use economic or other behavioral models to predict the number of contacts.
+The number of random and recurrent contacts at the workplace and at home is calibrated
+with data provided by \citet{Mossong2008}. For details see
+Section~\ref{sec:number_of_contacts}. In particular, we sample the number of contacts or
+group sizes from empirical distributions that sometimes depend on age. It is also
+possible to use economic or other behavioral models to predict the number of contacts.
 
-Theoretically, each contact type can have its own infection probability.
-However, to reduce the number of free parameters and thus avoid a potential over-fitting we impose some constraints.
-For now, infection probabilities in schools, preschools and nurseries are equal.
-Moreover, we restrict all work contacts to have the same infection probability.\comment[id=T]{I have no idea about the current infection probabilities.
+Theoretically, each contact type can have its own infection probability. However, to
+reduce the number of free parameters and thus avoid a potential over-fitting we impose
+some constraints. For now, infection probabilities in schools, preschools and nurseries
+are equal. Moreover, we restrict all work contacts to have the same infection
+probability.\comment[id=T]{I have no idea about the current infection probabilities.
 Does it need an update?}

### 2021-01-17 task_compile_documents.py

@@ -1,0 +1,17 @@
+import pytask
+from pathlib import Path
+
+
+ROOT = Path(__file__).parent
+BLD = ROOT / "bld"
+
+
+@pytask.mark.latex(
+    ["--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd", "--shell-escape"]
+)
+@pytask.mark.depends_on(
+    [ROOT / "paper" / "paper.tex"] + list(ROOT.joinpath("paper").rglob("*.tex"))
+)
+@pytask.mark.produces(BLD / "paper.pdf")
+def task_compile_documents():
+    pass

### 2021-02-06 paper/abstract.tex
Updates and notes for more updates for the introduction.

@@
-\noindent Governments worldwide are adopting nuanced policy measures to reduce the
-number of Covid-19 cases and to balance social and economic costs. Epidemiological
-models have a hard time predicting the effects of fine-grained policies. We propose a
-novel simulation-based model to address this shortcoming. We build on state-of-the-art
-agent-based simulation models but replace the way contacts between susceptible and
-infected people take place. Firstly, we allow for heterogeneity in the types of contacts
-(e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we
-strictly separate the number of contacts from the probabilities that a contact leads to
-an infection. The number of contacts changes with social distancing policies, the
-infection probabilities remain invariant. This allows us to model many types of targeted
-policies that cannot be incorporated into other models. To validate our model, we show
-that it can predict the effect of the German November lockdown even if no similar policy
-has been observed during the time period that was used to estimate the model parameters.
+\noindent
+Governments worldwide have been adopting various and nuanced policy measures
+to reduce the number of Covid-19 cases.
+However, epidemiological models usually lack sufficient detail in meeting
+patterns to credibly predict the effects such policies, much less predict
+complementarities or substitution effects between them.
+We propose a novel simulation-based model to address these shortcomings.
+We build on state-of-the-art agent-based simulation models, greatly increasing
+the amount of detail and realism with which contacts take place.
+Firstly, we allow for heterogeneity in the types of contacts (e.g. recurrent and
+non-recurrent) and in the infectiousness of each contact type.
+Secondly, we strictly separate the number of contacts from the probabilities
+that a contact leads to an infection.
+Thus, social distancing policies can affect the number of contacts while the
+infection probabilities remain invariant.
+This allows us to model many types of targeted policies that cannot be
+incorporated into other models.
+\textcolor{red}{
+To validate our model, we show that it can predict the effect of the German
+November lockdown even if no similar policy has been observed during the time
+period that was used to estimate the model parameters.
+}
 

### 2021-02-06 paper/introduction.tex
Updates and notes for more updates for the introduction.

@@
 
+% setting
 The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions
-across the world. Many countries were able to contain the spread of the virus and to
-relax restrictions during midyear. At the same time, the social and economic costs were
-enormous. In the second wave, countries are trying to implement more nuanced policies.
-For example, Germany imposed a ``lockdown-light'' in November. Some businesses are
-closed and certain leisure activities are prohibited but educational facilities remain
-open.
+across the world.
+At large social and economic costs, these strict measures were able to suppress the
+spread of the disease and allowed governments to relax restrictions over the summer.
+Governments were more hesitant and imposed less restrictive policies when infections
+started to soar again in the fall.
+For example, Germany imposed a ``lockdown-light'' in November, closing fewer businesses
+than in spring and leaving schools open. Mobility data also showed that individuals
+reduced their mobility much less than during the first lockdown.
 
-Current epidemiological models have not been designed to predict the effect of
-fine-grained policies. The models need to be extended for each new policy proposal and
-cannot be easily adapted to fast changing environments. This report describes a model
-that has been designed from the ground up to predict the effects of contact reducing
-policies in real time. It has the following features:
+\textcolor{red}{Add image of Google Mobility reductions!}
+
+% challenge
+The less restrictive lockdown led only to a stagnation of cases and led many experts
+\textcolor{red}{(Cite No-Covid, Zero-Covid, Leopoldina, Lancet...)} to call for a
+strict lockdown to bring cases back to very low levels, such as single digit incidences.
+These calls have received additional attention since new and more contagiuos variants
+have emerged in several countries and increased the urgency to reduce transmission
+dynamics and increased awareness for the mutagenic potential of high incidences.
+\textcolor{red}{Add citations!}
+
+However, epidemiological models are not suited to advise which combination of policies
+would be able to bring infections down to sustainable levels.
+As they have not been designed to predict the effects of fine-grained policies,
+they need to be extended for each new policy proposal.
+Nor can they show the trade-offs, substitutability or complementarity of policies.
+
+% contribution
+This report describes a model that has been designed from the ground up to predict the
+effects of contact reducing policies in real time. It has the following features:
 
@@
 This report describes the model in an abstract way, but uses many realistic examples
-from a version that is specialized to Germany. It is important to note that this
-specialization is not baked into the model or the Python code. It is easy to adjust the
-model to other countries if data on the number of contacts and a dataset with background
-characteristics are available.
+from a version that is specialized to Germany. It is important to note that it would
+be easy to adjust the model to other countries if data on the number of contacts and
+a dataset with background characteristics are available.
 
@@
 German policies can be found in \citet{Dorn2020a}.
+\textcolor{red}{Update citation to home office report.}
 

### 2021-02-08 paper/introduction.tex
Update introduction and motivation.

@@
 At large social and economic costs, these strict measures were able to suppress the
-spread of the disease and allowed governments to relax restrictions over the summer.
-Governments were more hesitant and imposed less restrictive policies when infections
-started to soar again in the fall.
+spread of the disease in many countries and allowed governments to relax restrictions
+over the summer.
+When infections started to soar again in the fall, governments were more hesitant
+and imposed less restrictive policies.
 For example, Germany imposed a ``lockdown-light'' in November, closing fewer businesses
-than in spring and leaving schools open. Mobility data also showed that individuals
-reduced their mobility much less than during the first lockdown.
-
-\textcolor{red}{Add image of Google Mobility reductions!}
+than in spring and leaving schools open.
 
 % challenge
-The less restrictive lockdown led only to a stagnation of cases and led many experts
-\textcolor{red}{(Cite No-Covid, Zero-Covid, Leopoldina, Lancet...)} to call for a
-strict lockdown to bring cases back to very low levels, such as single digit incidences.
+The less restrictive lockdowns proved insufficient to stop the second wave, leading
+many experts to call for a strict lockdown to bring cases back to very low levels,
+such as single digit incidences (for example \cite{Priesemann2021}).
 These calls have received additional attention since new and more contagiuos variants
-have emerged in several countries and increased the urgency to reduce transmission
-dynamics and increased awareness for the mutagenic potential of high incidences.
-\textcolor{red}{Add citations!}
+have emerged in across the globe \citep{Duong2021} and increased the urgency to
+reduce transmission dynamics and increased awareness for the mutagenic threat posed by
+high incidences.
+Given the importance to bring down and maintain low infection numbers until
+vaccinations reach herd immunity levels, it is paramount for policy makers and the
+public to understand the effectiveness and trade-offs involved in different policies.
 
-However, epidemiological models are not suited to advise which combination of policies
-would be able to bring infections down to sustainable levels.
+However, epidemiological models are not suited for this task.
 As they have not been designed to predict the effects of fine-grained policies,
 they need to be extended for each new policy proposal.
-Nor can they show the trade-offs, substitutability or complementarity of policies.
 
@@
     \item Policies can be implemented as shutting down contact types entirely or
-          partially. The reduction of contacts can be random or systematic, e.g., to
-          allow for essential workers.
+          partially. The reduction of contacts can be random or systematic. For example,
+          it is possible to implement A/B schooling where only one half of each class
+          attends and the attending half switches on a weekly bases.
 
@@
           (duration, physical contact). The probabilities are independent from the
-          number of contacts and thus policy-invariant.
+          number of contacts and, thus, policy-invariant.
 
@@
           are calibrated from the medical literature and datasets on contact
-          frequencies.
+          frequencies and mobility reductions.
 
@@
 More details about the German model as well as applications to currently discussed
-German policies can be found in \citet{Dorn2020a}.
-\textcolor{red}{Update citation to home office report.}
+German policies can be found in \citet{Dorn2020a, Gabler2020, Gabler2021}.
 
-In Section~\ref{sec:literature_review}, we give a short overview of epidemiological
-models. We continue with a detailed description of our model in Section~\ref{sec:model}
+The report proceeds as follows: In Section~\ref{sec:literature_review}, we give a
+short overview of epidemiological models.
+We continue with a detailed description of our model in Section~\ref{sec:model}
 and proceed with a description of model parameters and the estimation in

### 2021-02-08 paper/model/initial_conditions.tex
Release v2.0.0.
Update model section.

@@
 
-Consider a situation where you want to start a simulation with the beginning set amidst the pandemic.
-It means that several thousands of individuals have already recovered from the disease, are infectious, symptomatic or in ICUs.
-Additionally, the sample of infectious people who will determine the course of the pandemic in the following periods is likely not representative of the whole population because of differences in behavior (number of contacts, assortativeness), past policies (school closures), etc..
-The distribution of courses of diseases in the population at the begin of the simulation is called initial conditions.
+Consider a situation where you want to start a simulation with the beginning set amidst
+the pandemic. It means that several thousands of individuals should already have
+recovered from the disease, be infectious, symptomatic or in intensive care at the start
+of your simulation.
+Additionally, the sample of infectious people who will determine the course of the
+pandemic in the following periods is likely not representative of the whole population
+because of differences in behavior (number of contacts, assortativeness), past policies
+(school closures), etc..
+The distribution of courses of diseases in the population at the begin of the simulation is
+called initial conditions.
 
-To come up with realistic initial conditions, we match reported infections from official data to simulated individuals by available characteristics like age and geographic information.
-The matching must be done for each day of a longer time frame like a month to have individuals with possible health states.
-Then, health statuses evolve until the begin of the simulation period without simulating infections by contacts.
-It is also possible to correct reported infections for a reporting lag and scale them up to arrive at the true number of infections.
+To come up with realistic initial conditions, we match reported infections from official
+data to simulated individuals by available characteristics like age and geographic
+information. The matching must be done for each day of a longer time frame like a month
+to have individuals with possible health states. Then, health statuses evolve until the
+begin of the simulation period without simulating infections by contacts.
+We also correct reported infections for a reporting lag and scale them up to arrive
+at the true number of infections.
 

### 2021-02-08 paper/model/matching.tex
Release v2.0.0.
Update model section.

@@
 person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
-to be translated into actual meetings between people. This is achieved by matching
-algorithms:
+to be translated into actual meetings between people. This is achieved by our matching
+algorithm:
 
@@
 characteristics. An example would be a school class identifier which could come from
-actual data or be drawn randomly to achieve representative class sizes. 2) A
-deterministic or random function that takes the value 0 (non-participating) and 1
+actual data or be drawn randomly to achieve representative class sizes.
+2) A deterministic or random function that takes the value 0 (non-participating) and 1
 (participating) and can depend on the weekday, date and health state. This can be used
@@
 
-                if infection_takes_places(i, j):
-                    # Update health statuses.
+                if infection_takes_pinfectedlaces(i, j):
+                    update_health_state_of_freshly_infected()
 
@@
     \end{minted}
-    \caption{Pseudo-code for the matching algorithm of non-recurrent contacts.}
+    \caption{Pseudo-code of the matching algorithm for non-recurrent contacts.}
 \end{listing}
@@
 Next, we calculate the probability to be drawn for each member of the group, based on
-the number of remaining contacts. I.e. people who have more remaining contacts are drawn
+the number of remaining contacts, i.e. people who have more remaining contacts are drawn
 with a higher probability. This has to be re-calculated each time because with each
@@
 individual, determine whether an infection takes place and if so update the health
-states of the newly infected person. Finally, we reduce the number of remaining contacts
+state of the newly infected person. Finally, we reduce the number of remaining contacts
 of the two matched individuals by one.

### 2021-02-08 paper/model/number_of_contacts.tex
Updates and notes for more updates for the introduction.
Update model section.

@@
 Consider a hypothetical population of 1,000 individuals in which 50 were infected with a
-novel infectious disease. From this data alone, it is impossible to say whether only
+novel infectious disease. From this alone, it is impossible to say whether only
 those 50 people had contact with an infectious person and the disease has an infection
@@
 disease has an infection probability of only 5 percent per contact. SEIR models do not
-even try to distinguish contact frequency from the infectiousness of each contact and
+distinguish contact frequency from the infectiousness of each contact and
 combine the two in one parameter that is not invariant to social distancing policies.
@@
 
-The number of random and recurrent contacts at the workplace and at home is calibrated
-with data provided by \citet{Mossong2008}. For details see
+The number of random and recurrent contacts at the workplace, during leisure activities
+and at home is calibrated with data provided by \citet{Mossong2008}. For details see
 Section~\ref{sec:number_of_contacts}. In particular, we sample the number of contacts or
@@
 some constraints. For now, infection probabilities in schools, preschools and nurseries
-are equal. Moreover, we restrict all work contacts to have the same infection
-probability.\comment[id=T]{I have no idea about the current infection probabilities.
-Does it need an update?}
+are equal. Moreover, we restrict all work contacts and all non-work contacts to have
+the same infection probability.

### 2021-02-08 paper/model/policies.tex
Enforced 88 character limit.
Update model section.

@@
 
-The main motivation of our model is to predict the effect of policies that affect the number of contacts people have.
-Examples range from school closures and lockdowns to more nuanced policies such as mandatory quarantines for symptomatic individuals or a class splitting policy where only half of the students come to school in person and the other half joins digitally with weekly rotation.
+The main motivation of our model is to predict the effect of policies that affect the
+number of contacts people have. Examples range from school closures and lockdowns to
+more nuanced policies such as mandatory quarantines for symptomatic individuals or a
+class splitting policy where only half of the students come to school in person and the
+other half joins digitally with weekly rotation.
 
-Instead of thinking of policies as completely replacing how many contacts people have, it is often more helpful to think of them as adjusting the pre-pandemic number of contacts.
+Instead of thinking of policies as completely replacing how many contacts people have,
+it is often more helpful to think of them as adjusting the pre-pandemic number of
+contacts.
 
-Therefore, we implement policies as a step that happens after the number of contacts is calculated but before individuals are matched.
+Therefore, we implement policies as a step that happens after the number of contacts is
+calculated but before individuals are matched.
 
-On an abstract level, a policy is a functions that modifies the number of contacts of one contact type.
-For example, school closures simply set all school contacts to zero.
-A lockdown where only essential workers are allowed to work means that approximately two thirds of the working population have zero work contacts and the rest has the same number of contacts as before.
+On an abstract level, a policy is a functions that modifies the number of contacts of
+one contact type. For example, school closures simply set all school contacts to zero. A
+lockdown where only essential workers are allowed to work means that approximately two
+thirds of the working population have zero work contacts and the rest has the same
+number of contacts as before.
 
-% importance of fine-grained contact types and having the entire population
-This, in conjunction with our fine-grained contact types, allows us to easily implement a wide variety of policies.
-Allowing policies to depend on the health states of the entire population means that adaptive lockdowns where, for example, schools close when a certain threshold of infections is surpassed at the county level would be as simple as determining which counties are above the threshold and then setting all school contacts in these counties to zero.
+This, in conjunction with our fine-grained contact types, allows us to easily implement
+a wide variety of policies. Allowing policies to depend on the health states of the
+entire population means that adaptive lockdowns where, for example, schools close when a
+certain threshold of infections is surpassed at the county level would be as simple as
+determining which counties are above the threshold and then setting all school contacts
+in these counties to zero.
 
-The dependency of policies on health states also makes it possible to model contact tracing.
-For example, a policy could check whether each child has a classmate who's received a positive test result and then bar all children of that class from attending school.
+The dependency of policies on health states also makes it possible to model contact
+tracing. For example, a policy could check whether each child has a classmate who's
+received a positive test result and then bar all children of that class from attending
+school.
 
-% contact tracing and how including the states means the sky is the limit
-Some policies can be easily implemented if the background characteristics are suitably extended.
-For example, a schooling policy with split classes, where each half attends school every other week can be implemented by storing whether the child would attend in even or odd weeks in the background characteristics and then using that information in the policy function.
+Some policies can be easily implemented if the background characteristics are suitably
+extended. For example, a schooling policy with A/B schooling, where each half attends
+school every other week can be implemented by storing whether the child would attend in
+even or odd weeks in the background characteristics and then using that information in
+the policy function.
 
-For some policies the exact effect on each contact type is not easy to determine.
-If this refers to a policy during the estimation period, it is possible to estimate such parameters by fitting the model to time series data of infection rates.
-This is only possible if the policy was not active during the whole estimation period and thus the infection probabilities can be identified separately.
-If instead it refers to a policy that we want to simulate, we make a scenario analysis in which the model is simulated with several assumptions about how the policy affects the number of contacts.
+For some policies the exact effect on each contact type is not easy to determine. If
+this refers to a policy during the estimation period, it is possible to estimate such
+parameters by fitting the model to time series data of infection rates. This is only
+possible if the policy was not active during the whole estimation period and thus the
+infection probabilities can be identified separately. If instead it refers to a policy
+that we want to simulate, we make a scenario analysis in which the model is simulated
+with several assumptions about how the policy affects the number of contacts.

### 2021-02-08 paper/model/summary.tex
Release v2.0.0.
Update model section.

@@
 
-To alleviate these problems we propose a different model structure. Our model inherits
-many features from agent-based simulation models but replaces the contacts between
-moving particles by contacts between individuals who work, go to school, live in a
-household and enjoy leisure activities. The structure of the model is depicted in
-Figure~\ref{fig:model_graph}.
+To predict the effects of a wide variety of fine-grained social distancing policies
+we propose a different model structure.
+Our model inherits many features from agent-based simulation models but replaces
+the contacts between moving particles by contacts between individuals who work,
+go to school, live in a household and enjoy leisure activities.
+The structure of the model is depicted in Figure~\ref{fig:model_graph}.
 
@@
     \includegraphics[width=\textwidth]{../figures/model_detailed.png}
-    \caption{Simplified graph of the model}
+    \caption{Simplified graph of the model.}
     \label{fig:model_graph}
@@
 individual. Contact models are functions that map individual characteristics into a
-predicted number of contacts. Currently we distinguish between eight types of contact
+predicted number of contacts. Currently, we distinguish between eight types of contact
 models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and
@@
 
-Once a person is infected, the disease progresses in a fairly standard way which is also
+Once a person is infected, the disease progresses in a fairly standard way which is
 depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild
@@
 
-There are two possible ways to enable testing for Covid-19 in the model. The first way
-allows to specify the ratio of known to unknown infections for each day. A random sample
-of all newly infected individuals will then receive a test result in the following days.
-The second approach consists of three steps. First, individuals demand a test because
-they, e.g., experience symptoms. Secondly, tests are allocated to individuals while
-respecting governmental access restrictions to test. Thirdly, depending of the
-capacities of laboratories, tests are processed for some time until the individual
-receives her test result.
+Our model also allows the evaluation of different testing strategies by having a detailed
+three-step model of test demand, allocation and processing:
+First, individuals demand a test because they, e.g., experience symptoms.
+Secondly, tests are allocated to individuals while respecting governmental access
+restrictions to test.
+Thirdly, depending of the capacities of laboratories, tests are processed for some time
+until the individual receives her test result.
 
@@
 collected since the beginning of the pandemic can be used for estimation. Moreover,
-since we can model the testing strategies that were in place at each point in time, we
+since we model the testing strategies that were in place at each point in time, we
 can correct the estimates for the fact that not all infections are observed.

### 2021-02-08 paper/model/testing.tex
Release v2.0.0.
Update model section.

@@
 
-The model offers two approaches to implement testing for Covid-19.
+We support to model testing as consisting of three stages.
+Firstly, we model who demands a test. Demand functions map from individual
+characteristics to a probability which is the probability for this individual to demand
+a test. There can be multiple demand functions where each function may describe a
+different channel. For example, individuals who experience symptoms or have a risk
+contact may ask for a test. Or, the ministry of education requires a negative test
+result from every teacher every second week. After the probabilities for each individual
+and every demand model are computed, individuals who demand a test as well as the
+channel is sampled.
 
-The first way can be described as an top-down approach.
-Using data on estimates for the ratio of known to all infections\footnotemark a random sample of individuals who is newly infected is sampled and assigned a test.
-The result of the test will be revealed after some duration which must take into account the time it takes for a person to develop symptoms, the availability of tests and time to process the tests.
+The second stage is the allocation phase in which demand and supply for tests are
+matched. The number of available tests can be inferred from official data and used to
+model shortages in supply. When demand exceeds supply, some individuals might be given
+preferred access to tests because of their own vulnerability or their potential to
+become a super-spreader.
 
-\footnotetext{The Dunkelzifferradar project publishes daily estimates of the dark figure of infections under \url{https://covid19.dunkelzifferradar.de/}}
+In the last and third phase, administered tests are processed.
+This step can become a bottleneck in the testing process if there are not enough
+laboratories or necessary resources available to evaluate all tests.
 
-The second way is an bottom-up approach which consists of three phases.
-In the first phase, the demand for tests is modeled.
-Demand functions map from individual characteristics to a probability which is the probability for this individual to demand a test.
-There can be multiple demand functions where each function may describe a different channel.
-For example, individuals who experience symptoms or have a risk contact may ask for a test.
-Or, the ministry of education requires a negative test result from every teacher every second week.
-After the probabilities for each individual and every demand model are computed, individuals who demand a test as well as the channel is sampled.
-
-The second stage is the allocation phase in which demand and supply for tests are matched.
-The number of available tests can be inferred from official data and used to model shortages in supply.
-When demand exceeds supply, some individuals might be given preferred access to tests because of their own vulnerability or their potential to become a super-spreader.
-
-In the last and third phase, administered tests are determined to be processed.
-This step can become a bottleneck in the testing process if there are not enough laboratories or necessary resources available to evaluate all tests.
-
-The strength of the first approach is its simplicity, but it relies on estimates for the ratio of known to unknown cases for which not much data exist in the absence of systematic and randomized testing.
-The second approach is much more detailed and in principle able to reflect individual responses and governmental restrictions on the access to tests.
-At the same time, individual behavior and restrictions are complex and the determinants of the demand, allocation and processing functions require a lot of calibration.
+In our empirical estimation we use a very simplified testing model where the number
+of tests to be distributed is calculated from estimates for the ratio of known to all
+infections.\footnotemark
+\footnotetext{The Dunkelzifferradar project publishes daily estimates of the dark figure
+of infections under \url{https://covid19.dunkelzifferradar.de/}}
+Using these estimates as well as data on the test distribution
+over age groups by the RKI\footnotemark
+\footnotetext{https://ars.rki.de/Content/COVID19/Main.aspx}
+we allocate tests firstly among the symptomatic
+and then randomly allocate tests to newly infected to fit the German test distribution.
 

### 2021-02-09 paper/calibration_and_estimation.tex
Release v2.0.0.
Update calibration and estimation section.

@@
 \subsection{Number of Contacts}
+\label{sub:number_of_contacts}
 
@@
 
+\FloatBarrier
+
 \subsection{Assortative Matching}
@@
 geographical assortativeness and just roughly calibrate it such that 60 \% of contacts
-are within the same county\footnotemark, we can calibrate the age assortativeness from
+are within the same county, we can calibrate the age assortativeness from
 the same data we use to calibrate the number of contacts.
 
-\footnotetext{We are working on improving this estimation with mobility data.}
-
-\begin{figure}[!tp]
+\begin{figure}[h]
     \centering
-    \includegraphics[width=\textwidth]{../figures/assortative_matching_probability_example.png}
+    \includegraphics[width=0.9 \textwidth]{../figures/assortative_matching_probability_example.png}
     \caption{Distribution of random non-work contacts by age of participants.}
@@
 
+\FloatBarrier
+
 \subsection{Infection Probabilities}
@@
 To calibrate infection probabilities outside of the model, it would be important to know
-the exact duration and distance of each contact type as well as virus loads. Since this
+the exact duration and distance of each contact type as well as viral loads. Since this
 is not available in any dataset, we estimate those parameters inside the model with the
@@
 
-Currently, we use data for Germany from August to October (both inclusive). We do not
-use earlier periods to save computational time. Moreover, we would be worried that the
-infection probabilities have small seasonal variation that we currently cannot model.
-However, we plan to expand the estimation period soon.
+Currently, we use data for Germany from August until November.
+We do not use earlier periods to save computational time.
+Moreover, we would be worried that the there are seasonal effects that we currently do
+not model.
 
@@
 for work contacts. 3) for households. 4) for leisure activities.
+
+\subsection{Policies}
+
+\FloatBarrier
+
+In our empirical application we distinguish four groups of contact types:
+households, education, work and other contacts.
+For households we assume that the individuals' contacts in their households do not change
+over our estimation period.
+For nurseries, preschools and schools we implement vacations as announced by the German
+federal states as well as school closures. For the moment we ignore both emergency
+childcare and that lack of childcare leads working parents to stay home.
+%
+% Schließung von Kindertagesstätten und Schulen: 37,4 Millionen ausgefallene Arbeitstage
+% http://www.iab-forum.de/schul-und-kitaschliessungen-krankheit-quarantaene-die-coronabedingten-arbeitsausfaelle-der-erwerbstaetigen-steigen-auf-592-millionen-arbeitstage/
+%
+%
+% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377:
+% In allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
+% Schüler in die Schule.
+% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430.png
+% 18% der Grundschüler in Notbetreuung in BW
+%
+For our work models\footnotemark we use the reductions in work mobility reported in the
+Google Mobility Data \citep{Google2021} to calibrate our work policies.
+Reductions in work contacts are not random but governed through a work contact priority
+where the policy changes the threshold below which workers stay home.
+Figure \ref{fig:work_multiplier} shows the share of workers that go to work in our model
+over time.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9 \textwidth]{../figures/work_multiplier.png}
+    \caption{Share of workers attending work normally.}
+    \label{fig:work_multiplier}
+    \figurenotes{
+        The figure shows the percentage of pre-pandemic work mobility taking place
+        since the start of the pandemic. In our model this is used as a proxy for the
+        share of workers that attend work as usual. Workers are ordered with the
+        importance of personal contacts for their work. Thus, if the share of
+    }
+\end{figure}
+
+\footnotetext{We distinguish non-recurrent work contacts, daily work contacts and weekly work contacts.}
+
+For the last group of contacts which cover things like leisure activities, grocery
+shopping etc. we have no reliable data by how much policies reduce them.
+In addition, they are likely to be affected by social and psychological factors such as
+pandemic fatigue and vacations. Because of this we estimate them like the infection
+probabilities to fit the time series data. We use very few change points and tie them to
+particular events such as policy announcements or particular holidays.
+
+\FloatBarrier

### 2021-02-09 paper/model/summary.tex
Mention Troeger's research.

@@
 Secondly, tests are allocated to individuals while respecting governmental access
-restrictions to test.
+restrictions to tests.
 Thirdly, depending of the capacities of laboratories, tests are processed for some time
 until the individual receives her test result.
+This framework would allow to evaluate the effect of different testing demand behavior
+and allocation policies, as are for example discussed by \cite{Troeger2020}.
 

### 2021-02-09 paper/model_validation.tex
Release v2.0.0.
Update text to model validation.

@@
 \subsection{In-Sample Fit}
+\label{sub:in_sample_fit}
 
-Despite fitting only four free parameters, the in-sample fit is very good. The best fit
-is achieved in the largest age groups. This is so mechanically, because we weight the
+Despite fitting only four infection probabilities, the in-sample fit is very good.
+The best fit is achieved in the largest age groups.
+This is so mechanically, because we weigh the
 deviations between simulated and observed infection rates by group sizes. The worst fit
-is achieved for the 80 to 100 years old. The model predicts too few infections for these
-groups because they have very few contacts in all contact types we have included so far.
-We plan to address this issue soon by adding another contact type that captures all
-contacts in the data by \citet{Mossong2008} that we have not included so far. Moreover,
-we expect an improved model fit when we allow for more free parameters.
+is achieved for the 80 to 100 years old.
+There are three reasons for this: Firstly, we only model private households at
+the moment, meaning that nursing homes are not part of the synthetic population we
+simulate. Since community housing residents are part of the German Mikrozensus
+\citep{FDSAeDBUDL2018} which forms the basis for our synthetic population, we can and
+plan to include community housing residents in the near future.
+Secondly, the elderly, especially nursing home residents, are tested more often than
+the general population leading to more cases being detected in this age group.
+We are currently working to include age-variant testing strategies,
+using \href{https://ars.rki.de/Content/COVID19/Main.aspx}{data by the RKI}
+\citep{Seifried2020}.
 
-\begin{figure}[!tp]
+\begin{figure}[h]
     \centering
-    \includegraphics[width=\textwidth]{../figures/goodness_of_fit_by_age_group}
+    \includegraphics[width=1.1\textwidth]{../figures/goodness_of_fit_by_age_group}
     \caption{Reported vs. simulated weekly incidence rates of infections.}
@@
 
+\FloatBarrier
+
 \subsection{Out-of-Sample Fit}
+\label{sub:out_of_sample_fit}
 
 We can assess the out-of-sample fit by projecting the effect of the lockdown light and
-comparing it to case numbers until now. It is important to note that this is not just a
-simple extrapolation of a time trend because the lockdown light only started after the
-estimation period. The out-of-sample fit can be assessed in
+comparing it to case numbers until mid of November. It is important to note that this
+is not just a simple extrapolation of a time trend because the lockdown light only started
+after the estimation period. The out-of-sample fit can be assessed in
 Figure~\ref{fig:out-of-sample-fit}.
@@
 In particular, the actual case numbers are between our neutral and pessimistic
-projection. The plot also shows that ending the lockdown light as planned on November 30
-would lead to an explosive growth in case numbers in all scenarios.
+projection. The plot also shows that ending the lockdown light on November 30, as was
+originally planned, would lead to an explosive growth in case numbers in all scenarios.

### 2021-02-09 task_compile_documents.py
Add references to the task dependencies.

@@ -4,14 +4,14 @@ from pathlib import Path
 
 ROOT = Path(__file__).parent
 BLD = ROOT / "bld"
+DEPENDENCIES = [ROOT / "paper" / "paper.tex", ROOT / "references.bib"]
+DEPENDENCIES += list(ROOT.joinpath("paper").rglob("*.tex"))
 
 
 @pytask.mark.latex(
     ["--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd", "--shell-escape"]
 )
-@pytask.mark.depends_on(
-    [ROOT / "paper" / "paper.tex"] + list(ROOT.joinpath("paper").rglob("*.tex"))
-)
+@pytask.mark.depends_on(DEPENDENCIES)
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass

### 2021-02-09 paper/abstract.tex
Update abstract.

--- 2020-12-20 paper/abstract.tex
@@
-\noindent Governments worldwide are adopting nuanced policy measures to reduce the
-number of Covid-19 cases and to balance social and economic costs. Epidemiological
-models have a hard time predicting the effects of fine-grained policies. We propose a
-novel simulation-based model to address this shortcoming. We build on state-of-the-art
-agent-based simulation models but replace the way contacts between susceptible and
-infected people take place. Firstly, we allow for heterogeneity in the types of contacts
-(e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we
-strictly separate the number of contacts from the probabilities that a contact leads to
-an infection. The number of contacts changes with social distancing policies, the
-infection probabilities remain invariant. This allows us to model many types of targeted
-policies that cannot be incorporated into other models. To validate our model, we show
-that it can predict the effect of the German November lockdown even if no similar policy
-has been observed during the time period that was used to estimate the model parameters.
+\noindent
+Governments worldwide have been adopting diverse and nuanced policy measures
+to contain the spread of Covid-19.
+However, epidemiological models usually lack the detailed representation of
+human meeting patterns to credibly predict the effects such policies.
+%
+We propose a novel simulation-based model to address these shortcomings.
+We build on state-of-the-art agent-based simulation models, greatly increasing
+the amount of detail and realism with which contacts take place.
+Firstly, we allow for different contact types (such as work, school, households
+or leisure), distinguish recurrent and non-recurrent contacts and allow the
+infectiousness of meetings to vary between contact types.
+% Secondly, we strictly separate the number of contacts from the probabilities
+% that a contact leads to an infection.
+Secondly, we allow agents to seek tests and react to information, such as
+experiencing symptoms, receiving a positive test or a known case among their
+contacts, by reducing their own contacts.
+% Thus, social distancing policies can affect the number of contacts while the
+% infection probabilities remain invariant.
+This allows us to model the effects of a wide array very targeted policies such
+as A/B schooling, mandatory work from home schemes or test-and-trace policies.
+To validate our model, we show that it can predict the effect of the German
+November lockdown even if no similar policy has been observed during the time
+period that was used to estimate the model parameters.
 

### 2021-02-09 task_compile_documents.py
Address some LaTeX warnings. Add force option to latex command.

@@ -9,7 +9,14 @@ DEPENDENCIES += list(ROOT.joinpath("paper").rglob("*.tex"))
 
 
 @pytask.mark.latex(
-    ["--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd", "--shell-escape"]
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
 )
 @pytask.mark.depends_on(DEPENDENCIES)
 @pytask.mark.produces(BLD / "paper.pdf")

### 2021-02-09 paper/abstract.tex
Describe A/B schooling instead of using the term.

@@
 This allows us to model the effects of a wide array very targeted policies such
-as A/B schooling, mandatory work from home schemes or test-and-trace policies.
+as split classes, mandatory work from home schemes or test-and-trace policies.
 To validate our model, we show that it can predict the effect of the German

### 2021-04-17 paper/calibration_and_estimation.tex
Mention Troeger's research.
Address some LaTeX warnings. Add force option to latex command.
Update Readme and add pdf.

@@
 probability to depend on age and county of residence. While we do not have good data on
-geographical assortativeness and just roughly calibrate it such that 60 \% of contacts
-are within the same county, we can calibrate the age assortativeness from
-the same data we use to calibrate the number of contacts.
+geographical assortativity and just roughly calibrate it such that 60 \% of contacts are
+within the same county, we can calibrate the assortative mixing by age from the same
+data we use to calibrate the number of contacts.
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
     \caption{Distribution of random non-work contacts by age of participants.}
-    \label{fig:assortativeness}
+    \label{fig:assortativity}
     \figurenotes{
@@
 
-Figure~\ref{fig:assortativeness} shows that assortativeness by age is especially strong
+Figure~\ref{fig:assortativity} shows that assortativity by age is especially strong
 for children and younger adults. For older people, the pattern becomes more dispersed
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/work_multiplier.png}
+    \includegraphics[width=1.1 \textwidth]{../figures/work_multiplier.png}
     \caption{Share of workers attending work normally.}

### 2021-04-17 paper/model/initial_conditions.tex
Update Readme and add pdf.

@@
 recovered from the disease, be infectious, symptomatic or in intensive care at the start
-of your simulation.
-Additionally, the sample of infectious people who will determine the course of the
-pandemic in the following periods is likely not representative of the whole population
-because of differences in behavior (number of contacts, assortativeness), past policies
-(school closures), etc..
-The distribution of courses of diseases in the population at the begin of the simulation is
-called initial conditions.
+of your simulation. Additionally, the sample of infectious people who will determine the
+course of the pandemic in the following periods is likely not representative of the
+whole population because of differences in behavior (number of contacts,
+assortativity), past policies (school closures), etc.. The distribution of courses of
+diseases in the population at the begin of the simulation is called initial conditions.
 
@@
 to have individuals with possible health states. Then, health statuses evolve until the
-begin of the simulation period without simulating infections by contacts.
-We also correct reported infections for a reporting lag and scale them up to arrive
-at the true number of infections.
+begin of the simulation period without simulating infections by contacts. We also
+correct reported infections for a reporting lag and scale them up to arrive at the true
+number of infections.
 

### 2021-04-17 templates/Preamble_Fonts_Times.tex
Update templates, fix errors, copy pdf to root.

% !TeX program = pdflatex
% !TeX TXS-program:compile = txs:///pdflatex/
% !TeX TS-program = pdflatex
% !BIB program = biber
% !TeX TXS-program:bibliography = txs:///biber

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  FONT SETTINGS THESIS  %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsfonts, amssymb}  % Additional symbols from AMS

\usepackage{extarrows}

% Use Times Roman for text and math
\usepackage{newtxtext}
\usepackage[smallerops, varg, upint, slantedGreek]{newtxmath}

%\usepackage[scale=1.07]{tgcursor}
%	% Use Courier as the typewriter font
%\usepackage[scale=0.93]{sourcecodepro}
%	% Use Source Code Pro as the typewriter font
\usepackage[scaled=0.835]{DejaVuSansMono}  % Scaled to match the x-height of the newtxtext Times Roman
	% Use DejaVu Sans Mono as the typewriter font, also supports Greek via the LGR encoding

\usepackage{xfrac}

\usepackage[protrusion=true, expansion=false, kerning=true]{microtype}
% This package enables so-called hanging punctuation. That is, when a punctuation sign like ":", ".", "-", etc. is found at the beginning or end of a line, it is protruded a little into the page margin. This results in "optical margin alignment," because the protrusion makes the margin alignment look straighter.
%\DisableLigatures[f]{family = {sf*, rm*}}
%	% Disable the f* ligatures for both Fira Sans and Charter because both fonts provide insufficient support
%\DisableLigatures[f]{family = {rm*}}
	% Disable the f* ligatures for Charter because it provides insufficient support
\SetExtraKerning[unit=space]
{encoding=*, family=*, series=*, size={*, normalsize, footnotesize}, font = */*/*/*/*}
{\textemdash = {325, 325},
	/ = {100, 100},
	: = { 50,   0},
	; = { 50,   0}}
\renewcommand{\textellipsis}{\mbox{.{\kern.09em}.{\kern.09em}.}}

\renewcommand{\bibfont}{%
	\normalsize%
	\addtolength{\baselineskip}{0.05\baselineskip}%
	%\setlength{\baselineskip}{\newbaselineskip}%
}  % Use same size for bibliography as for body text.

\captionsetup{footfont={rm, footnotesize}}  % Set font in \floatfoot to serif and footnotesize

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  MATH-RELATED ADJUSTMENTS  %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{mathrsfs}  % Provides the \mathscr math alphabet

% The following is taken from
% https://tex.stackexchange.com/questions/116389/automatic-upright-math-when-text-is-in-italic/116399#116399
% Filling in ``missing'' Greek glyphs for completeness
% (not really necessary, since they look identical to Latin glyphs and are thus almost never used)
% ==>
\newcommand{\omicron}{o}
\newcommand{\Digamma}{F}
\newcommand{\Alpha}  {A}
\newcommand{\Beta}   {B}
\newcommand{\Epsilon}{E}
\newcommand{\Zeta}   {Z}
\newcommand{\Eta}    {H}
\newcommand{\Iota}   {I}
\newcommand{\Kappa}  {K}
\newcommand{\Mu}     {M}
\newcommand{\Nu}     {N}
\newcommand{\Omicron}{O}
\newcommand{\Rho}    {P}
\newcommand{\Tau}    {T}
\newcommand{\Chi}    {X}
% <==

% Upright Greek letters for the \mathup command
% ==>
\makeatletter
% Save original definitions of the Greek letters
\@for\@tempa:=%
	alpha,beta,gamma,delta,epsilon,zeta,eta,theta,iota,kappa,lambda,mu,nu,xi,%
	omicron,pi,rho,sigma,varsigma,tau,upsilon,phi,chi,psi,omega,digamma,%
	Alpha,Beta,Gamma,Delta,Epsilon,Zeta,Eta,Theta,Iota,Kappa,Lambda,Mu,Nu,Xi,%
	Omicron,Pi,Rho,Sigma,Tau,Upsilon,Phi,Chi,Psi,Omega,Digamma%
	\do{%
		\expandafter\let\csname\@tempa orig\expandafter\endcsname\csname\@tempa\endcsname%
		\expandafter\let\csname\@tempa uporig\expandafter\endcsname\csname\@tempa up\endcsname%
	}%
\newcommand*{\upgreekletters}{%
	\@for\@tempa:=%
		alpha,beta,gamma,delta,epsilon,zeta,eta,theta,iota,kappa,lambda,mu,nu,xi,%
		omicron,pi,rho,sigma,varsigma,tau,upsilon,phi,chi,psi,omega,digamma,%
		Alpha,Beta,Gamma,Delta,Epsilon,Zeta,Eta,Theta,Iota,Kappa,Lambda,Mu,Nu,Xi,%
		Omicron,Pi,Rho,Sigma,Tau,Upsilon,Phi,Chi,Psi,Omega,Digamma%
		\do{%
			\expandafter\let\csname\@tempa\expandafter\endcsname\csname\@tempa up\endcsname%
		}%
}
\newcommand*{\itgreekletters}{%
	\@for\@tempa:=%
		alpha,beta,gamma,delta,epsilon,zeta,eta,theta,iota,kappa,lambda,mu,nu,xi,%
		omicron,pi,rho,sigma,varsigma,tau,upsilon,phi,chi,psi,omega,digamma,%
		Alpha,Beta,Gamma,Delta,Epsilon,Zeta,Eta,Theta,Iota,Kappa,Lambda,Mu,Nu,Xi,%
		Omicron,Pi,Rho,Sigma,Tau,Upsilon,Phi,Chi,Psi,Omega,Digamma%
		\do{%
			\expandafter\let\csname\@tempa\expandafter\endcsname\csname\@tempa orig\endcsname%
		}%
}
\makeatother
% <==

\newcommand{\mathup}[1]{\upgreekletters\mathrm{#1}\itgreekletters}

\renewcommand{\mathbf}[1]{\bm{#1}}
\newcommand{\mathbfit}[1]{\mathbf{\mathit{#1}}}
\newcommand{\mathbfup}[1]{\upgreekletters\mathbf{\mathrm{#1}}\itgreekletters}

% Since math mode uses a different font encoding, issuing \euro/\texteuro in math mode
% produces an incorrect sign. We fix this. ==>
\AtBeginDocument{%
	\newcommand*{\euro}[1]{%
		\relax\ifmmode\text{\texteuro}#1\else\texteuro #1\fi%
	}%
}
% <==

### 2021-04-17 task_compile_documents.py

@@ -1,11 +1,11 @@
-import pytask
+import shutil
 from pathlib import Path
+
+import pytask
 
 
 ROOT = Path(__file__).parent
 BLD = ROOT / "bld"
-DEPENDENCIES = [ROOT / "paper" / "paper.tex", ROOT / "references.bib"]
-DEPENDENCIES += list(ROOT.joinpath("paper").rglob("*.tex"))
 
 
 @pytask.mark.latex(
@@ -15,10 +15,15 @@ DEPENDENCIES += list(ROOT.joinpath("paper").rglob("*.tex"))
         "--synctex=1",
         "--cd",
         "--shell-escape",
-        "-f",
     ]
 )
-@pytask.mark.depends_on(DEPENDENCIES)
+@pytask.mark.depends_on(ROOT / "paper" / "paper.tex")
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass
+
+
+@pytask.mark.depends_on(BLD / "paper.pdf")
+@pytask.mark.produces(ROOT / ".." / "paper.pdf")
+def task_copy_pdf_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)

### 2021-04-17 task_compile_documents.py
pdf really to root.

@@ -24,6 +24,6 @@ def task_compile_documents():
 
 
 @pytask.mark.depends_on(BLD / "paper.pdf")
-@pytask.mark.produces(ROOT / ".." / "paper.pdf")
+@pytask.mark.produces(ROOT / "paper.pdf")
 def task_copy_pdf_to_root(depends_on, produces):
     shutil.copy(depends_on, produces)

### 2021-04-17 paper/conclusion.tex
Move to Holger's templates and some other - cannot remember - fixes.
Enforced 88 character limit.
Updates and notes for more updates for the introduction.
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
 
-We propose a simulation based model of infectious disease transmission that is designed to predict the effects of fine-grained social distancing policies.
-In particular, the model can be used to model policies such as several ways of splitting school classes or work reduction policies that affect essential and non-essential workers differently.
-Both policies would be hard to implement in standard SEIR or agent based simulation models.
+We propose a simulation based model of infectious disease transmission that is designed
+to predict the effects of fine-grained social distancing policies. In particular, the
+model can be used to model policies such as several ways of splitting school classes or
+work reduction policies that affect essential and non-essential workers differentially.
+Both policies would be hard to implement in standard SEIR or agent based simulation
+models.
 
-To predict the effects of such policies, it is not only important to have a way of expressing such flexible policies in terms of model quantities, but also to incorporate heterogeneity in disease progression as well as meeting patterns.
-We calibrate age dependent disease progression parameters from the medical literature and age dependent contact frequencies from contact diaries.
-Moreover, we distinguish ten types of contacts out of which some are only relevant for certain age groups.
+To predict the effects of such policies, it is not only important to have a way of
+expressing such flexible policies in terms of model quantities, but also to incorporate
+heterogeneity in disease progression as well as meeting patterns. We calibrate age
+dependent disease progression parameters from the medical literature and age dependent
+contact frequencies from contact diaries. Moreover, we distinguish ten types of contacts
+out of which some are only relevant for certain age groups.
 
-The model has a good fit on past German case numbers and passes an out of sample validation despite a drastic change in the policy environment between the estimation period and the validation period.
+The model has a good fit on past German case numbers and passes an out of sample
+validation despite a drastic change in the policy environment between the estimation
+period and the validation period.
 
-Despite these encouraging results we still see the model as work in progress and plan to implement more features such as a detailed model of testing and contact tracing.
-Moreover, the estimation of the infection probabilities and the model fit will improve as more data becomes available.
+Despite these encouraging results we still see the model as work in progress.
+We plan to implement more features, in particular allowing age and symptoms to affect
+the probability to seek and receive a test, opening the way to show the effect of
+different testing policies, such as those proposed by \cite{Troeger2020}.
+Moreover, the estimation of the infection probabilities and the model fit will improve
+as more data becomes available.
 
-We invite researchers from any discipline, but particularly epidemiologists to provide feedback on the model and welcome collaborations.
+We invite researchers from any discipline, but particularly epidemiologists to provide
+feedback on the model and welcome collaborations.

### 2021-04-17 paper/introduction.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
           partially. The reduction of contacts can be random or systematic. For example,
-          it is possible to implement A/B schooling where only one half of each class
-          attends and the attending half switches on a weekly bases.
+          it is possible to implement split class schooling where only one half of each
+          class attends and the attending half switches on a weekly bases.
 

### 2021-04-17 paper/literature_review.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
 The above assessment of epidemiological models is not meant as a critique. We are aware
-that those models were not designed to predict the effect of fine-grained social
+that these models were not designed to predict the effect of fine-grained social
 distancing policies in real time and are very well suited to their purpose. We invite

### 2021-04-17 paper/model/course_of_disease.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
 First, infected individuals will become infectious after one to five days. About one
-third of people stays asymptomatic. The rest develops symptoms about one to two days
+third of people remain asymptomatic. The rest develop symptoms about one to two days
 after they become infectious. Modeling asymptomatic and pre-symptomatic cases is
@@
 could model people who require hospitalization but not intensive care. So far we opted
-against that because only the ICU capacity might become a bottleneck in Germany.
+against that because only the intensive care capacities are feared to become a
+bottleneck in Germany.
 

### 2021-04-17 paper/model/policies.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
 Some policies can be easily implemented if the background characteristics are suitably
-extended. For example, a schooling policy with A/B schooling, where each half attends
-school every other week can be implemented by storing whether the child would attend in
-even or odd weeks in the background characteristics and then using that information in
-the policy function.
+extended. For example, a schooling policy of splitting and rotating classes,
+where each half attends school every other week can be implemented by storing
+whether the child would attend in even or odd weeks in the background characteristics
+and then using that information in the policy function.
 

### 2021-04-17 paper/model/testing.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
 
-We support to model testing as consisting of three stages.
-Firstly, we model who demands a test. Demand functions map from individual
-characteristics to a probability which is the probability for this individual to demand
-a test. There can be multiple demand functions where each function may describe a
-different channel. For example, individuals who experience symptoms or have a risk
-contact may ask for a test. Or, the ministry of education requires a negative test
-result from every teacher every second week. After the probabilities for each individual
-and every demand model are computed, individuals who demand a test as well as the
-channel is sampled.
+We support to model testing as consisting of three stages. Firstly, we model who demands
+a test. Demand functions map from individual characteristics to a probability which is
+the probability for this individual to demand a test. There can be multiple demand
+functions where each function may describe a different channel. For example, individuals
+who experience symptoms or have a risk contact may ask for a test. Or, the ministry of
+education requires a negative test result from every teacher every second week. After
+the probabilities for each individual and every demand model are computed, individuals
+who demand a test as well as the channel is sampled.
 
@@
 
-In the last and third phase, administered tests are processed.
-This step can become a bottleneck in the testing process if there are not enough
-laboratories or necessary resources available to evaluate all tests.
+In the last and third phase, administered tests are processed. This step can become a
+bottleneck in the testing process if there are not enough laboratories or necessary
+resources available to evaluate all tests.
 
-In our empirical estimation we use a very simplified testing model where the number
-of tests to be distributed is calculated from estimates for the ratio of known to all
-infections.\footnotemark
-\footnotetext{The Dunkelzifferradar project publishes daily estimates of the dark figure
-of infections under \url{https://covid19.dunkelzifferradar.de/}}
-Using these estimates as well as data on the test distribution
-over age groups by the RKI\footnotemark
-\footnotetext{https://ars.rki.de/Content/COVID19/Main.aspx}
-we allocate tests firstly among the symptomatic
-and then randomly allocate tests to newly infected to fit the German test distribution.
+In our empirical estimation we use a very simplified testing model where the number of
+tests to be distributed is calculated from estimates for the ratio of known to all
+infections.\footnote{The Dunkelzifferradar project publishes daily estimates of the dark
+figure of infections under \url{https://covid19.dunkelzifferradar.de/}} Using these
+estimates as well as data on the test distribution over age groups by the
+RKI\footnote{https://ars.rki.de/Content/COVID19/Main.aspx} we allocate tests firstly
+among the symptomatic and then randomly allocate tests to newly infected to fit the
+German test distribution.
 

### 2021-04-17 task_compile_documents.py

@@ -1,5 +1,7 @@
-import pytask
+import shutil
 from pathlib import Path
+
+import pytask
 
 
 ROOT = Path(__file__).parent
@@ -7,11 +9,21 @@ BLD = ROOT / "bld"
 
 
 @pytask.mark.latex(
-    ["--pdf", "--interaction=nonstopmode", "--synctex=1", "--cd", "--shell-escape"]
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+    ]
 )
-@pytask.mark.depends_on(
-    [ROOT / "paper" / "paper.tex"] + list(ROOT.joinpath("paper").rglob("*.tex"))
-)
+@pytask.mark.depends_on(ROOT / "paper" / "paper.tex")
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass
+
+
+@pytask.mark.depends_on(BLD / "paper.pdf")
+@pytask.mark.produces(ROOT / "paper.pdf")
+def task_copy_pdf_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)

### 2021-05-31 paper/introduction.tex
Update introduction.

@@
 across the world.
-At large social and economic costs, these strict measures were able to suppress the
-spread of the disease in many countries and allowed governments to relax restrictions
-over the summer.
-When infections started to soar again in the fall, governments were more hesitant
-and imposed less restrictive policies.
-For example, Germany imposed a ``lockdown-light'' in November, closing fewer businesses
-than in spring and leaving schools open.
+The reaction to second waves consisted of more targeted policy measures such as splitting school classes, closing restaurants and encouragement of home office.
+To combat the third and fourth waves, that were triggered by the spread of more infectious variants, governments relied on the same targeted measures and previously non-available tools such as large scale rapid testing and vaccinations.
 
-% challenge
-The less restrictive lockdowns proved insufficient to stop the second wave, leading
-many experts to call for a strict lockdown to bring cases back to very low levels,
-such as single digit incidences (for example \cite{Priesemann2021}).
-These calls have received additional attention since new and more contagiuos variants
-have emerged in across the globe \citep{Duong2021} and increased the urgency to
-reduce transmission dynamics and increased awareness for the mutagenic threat posed by
-high incidences.
-Given the importance to bring down and maintain low infection numbers until
-vaccinations reach herd immunity levels, it is paramount for policy makers and the
-public to understand the effectiveness and trade-offs involved in different policies.
+While this multitude of policy measures has led to declining case numbers in most countries, it becomes harder and harder to evaluate the contribution each policy had on the overall outcome.
+Moreover, the longer the pandemic lasts, the more important it becomes to accurately model heterogeneities in contact behavior and a realistic network of recurrent contacts because some sub-populations might develop herd immunity even though the full population does not.
 
-However, epidemiological models are not suited for this task.
-As they have not been designed to predict the effects of fine-grained policies,
-they need to be extended for each new policy proposal.
-
-% contribution
-This report describes a model that has been designed from the ground up to predict the
-effects of contact reducing policies in real time. It has the following features:
+The workhorse model of epidemiology, the S(E)IR model as well as many recent extensions to it are not up to this task.
+We develop an agent-based simulation model that has been designed from the ground up to predict and quantify the effects of contact reducing policies, vaccinations and testing strategies in a constantly changing policy environment.
+It has the following features:
 
@@
           contacts at the workplace. Contact types can be random or recurrent and vary
-          in frequency.
+          in frequency and infectiousness.
 
@@
           it is possible to implement split class schooling where only one half of each
-          class attends and the attending half switches on a weekly bases.
+          class attends and the attending half switches on a weekly bases. The extent to which contacts are reduced can be calibrated from observed data or estimated inside the model.
 
@@
 
-    \item The model achieves a good fit on German data of infection rates even if only
-          the infection probabilities are fit to the data and the remaining parameters
-          are calibrated from the medical literature and datasets on contact
-          frequencies and mobility reductions.
-
+    \item We distinguish detected and undetected cases. The share of detected cases varies over time and across age groups. Moreover, it can be influenced by rapid testing policies.
     \item High quality Python code for the model is freely available on Github, well
@@
 
-This report describes the model in an abstract way, but uses many realistic examples
-from a version that is specialized to Germany. It is important to note that it would
-be easy to adjust the model to other countries if data on the number of contacts and
-a dataset with background characteristics are available.
+The model achieves a good fit on German data of infection and fatality rates, even though most parameters are calibrated from the literature and observable datasets and only a few parameters are estimated inside the model.
+At accurately predicts the rise of the B117 mutation in spring 2020.
+It can also explain the surprising decline of case numbers at the end of April, without making ad hoc assumptions on behavioral changes at the time.
 
-More details about the German model as well as applications to currently discussed
-German policies can be found in \citet{Dorn2020a, Gabler2020, Gabler2021}.
+The model has previously been applied to predict the effect of schooling policies, contact tracing policies over the Christmas holidays and the effect of work from home (\citet{Dorn2020a, Gabler2020, Gabler2021}).
 
-The report proceeds as follows: In Section~\ref{sec:literature_review}, we give a
-short overview of epidemiological models.
-We continue with a detailed description of our model in Section~\ref{sec:model}
-and proceed with a description of model parameters and the estimation in
-Section~\ref{sec:calibration_and_estimation}. The model is validated in
-Section~\ref{sec:model_validation} by assessing the in-sample fit for reported
-infections from August to October and by comparing the out-of-sample fit for the period
-from November until the beginning of Christmas for different lockdown scenarios. We
-conclude in Section~\ref{sec:conclusion}.
+\textcolor{blue}{Summary of the main results}
+
+The remainder is structured as follows: In Section~\ref{sec:literature_review}, we give a short overview of epidemiological models.
+Section ~\ref{sec:background} summarizes the policy environments and dynamic of the Covid-19 pandemic in Europe, with a special focus on Germany.
+We continue with a general description of our modelling framework  as well as a specialization to the German context in section~\ref{sec:model}.
+Section~\ref{sec:calibration_and_estimation} describes our empirical datasets, sources of calibrated parameters and estimation procedure.
+\textcolor{blue}{desribe result section}.
+We conclude in section~\ref{sec:conclusion}.

### 2021-05-31 paper/model/summary.tex
Update Readme and add pdf.
Update model summary.

--- 2021-02-08 paper/model/summary.tex
@@
 
-To predict the effects of a wide variety of fine-grained social distancing policies
-we propose a different model structure.
-Our model inherits many features from agent-based simulation models but replaces
-the contacts between moving particles by contacts between individuals who work,
+To predict and quantify the effects of a wide variety of fine-grained social distancing policies, vaccinations and rapid testing, we propose a different model structure. Our
+model inherits many features from prototypical agent-based simulation models but
+replaces the contacts between moving particles by contacts between individuals who work,
 go to school, live in a household and enjoy leisure activities.
+
 The structure of the model is depicted in Figure~\ref{fig:model_graph}.
@@
 
+The infection probabilities of contacts vary with contact type, age of the susceptible
+person, and the virus strain of the infected person. Moreover, they follow a seasonal
+pattern. The strength of the seasonality effect is higher for contacts that are easy to
+be moved to an outside location in summer (such as leisure contacts) and smaller for
+contacts that take place inside even in summer (e.g. work contacts).
+
 Once a person is infected, the disease progresses in a fairly standard way which is
@@
 
-Our model also allows the evaluation of different testing strategies by having a detailed
-three-step model of test demand, allocation and processing:
-First, individuals demand a test because they, e.g., experience symptoms.
-Secondly, tests are allocated to individuals while respecting governmental access
-restrictions to test.
-Thirdly, depending of the capacities of laboratories, tests are processed for some time
-until the individual receives her test result.
+After rapid tests become available, people who work or go to school can receive rapid
+tests there. Moreover, people can decide to make a rapid test if they develop symptoms,
+have many planned contacts or observe cases in their contact network. People who have a positive rapid test demand a confirmatory PCR test with a certain
+probability. Moreover, PCR tests can be demanded because of symptoms or randomly.
 
-In addition, people who have symptoms, received a positive test, or had a risk contact
-can reduce their number of contacts across all contact types endogenously.
+This rich model of PCR and rapid tests leads to a share of detected cases that varies
+over time and across age groups. It also allows to quantify the effect of changes in
+testing policies on the dynamic of infections.
+
+People who have symptoms, received a positive test, or had a risk contact
+can reduce their number of contacts across all contact types endogenously. The extent to
+which this is done is calibrated from survey data.
 
@@
 collected since the beginning of the pandemic can be used for estimation. Moreover,
-since we model the testing strategies that were in place at each point in time, we
-can correct the estimates for the fact that not all infections are observed.
+since we model the testing strategies that were in place at each point in time, we can
+correct the estimates for the fact that not all infections are observed.
 

### 2021-05-31 paper/model/number_of_contacts.tex
Update contacts section.

@@
 
-Currently, there are the following contact types:
+In our empirical application we distinguish the following types of contacts:
 
@@
 Theoretically, each contact type can have its own infection probability. However, to
-reduce the number of free parameters and thus avoid a potential over-fitting we impose
-some constraints. For now, infection probabilities in schools, preschools and nurseries
-are equal. Moreover, we restrict all work contacts and all non-work contacts to have
-the same infection probability.
+reduce the number of free parameters and thus avoid a potential over-fitting we only
+estimate different infection probabilities for the areas work, school, preschool and
+nurseries, households and other contacts.

### 2021-06-01 figures/results/tables/student_infections_table.tex
Add results section to paper and results.

\begin{tabular}{lr}
\toprule
{} &  predicted total infections among 5-14 year olds from Easter until 2021-05-31 \\
scenario                               &                                                                               \\
\midrule
 educ open after easter  without tests &                                             843318 \\
 educ open after easter  with tests    &                                             607352 \\
 baseline                              &                                             479015 \\
 close educ after easter               &                                             475397 \\
\bottomrule
\end{tabular}

### 2021-06-01 paper/results.tex
Add results section to paper and results.

\section{Results}
\label{sec:results}

This is the results section.

\textcolor{red}{The results can be found in `figures/results`. This includes both figures
and tables for lookup of numbers and summary tables}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/one_off_and_combined/full_new_known_case}
    \caption{The Effect of Each Channel on Observed New Cases Separately}
    \label{fig:}
    \figurenotes{\textcolor{red}{\ldots}}
\end{figure}

### 2021-06-01 paper/medical_parameters.tex
Start on updating the medical parameters.

@@
 
-This section discusses the medical parameters used in the model, their sources and how
-we arrived at the distributions used in the model.\footnotemark
+This section discusses the medical parameters used in the model, their sources and how we
+arrived at the distributions used in the model.\footnotemark See
+Figure~\ref{fig:course_of_disease} for a summary of our disease progression model.
 
@@
 
-\subsubsection{Length of Presymptomatic Stage / Incubation Period}
+\subsubsection{Duration until Infectiousness and Symptoms}
 
-Estimates of the incubation period usually give a range from 2 to 12 days. A meta
-analysis by \citet{McAloon2020} comes to the conclusion that ``The incubation period
-distribution may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$
-parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55),
-respectively.'' For simplicity we discretize this distribution into four bins.
+% Latency Period
+The first medical parameter we need is the length of the period between infection and
+the start of infectiousness, the so called latent period.
+We infer it from two other measures that are more common in the medical literature:
+Firstly, the time between infection and the onset of symptoms, the incubation period.
+Secondly, the time between the start of infectiousness and the onset of symptoms.
+We assume that the latency period is the same for symptomatic and asymptomatic individuals.
 
-\subsubsection{Begin of Infectiousness}
+% Asymptomatic Share
+Once individuals become infectious a share of them goes on to develop symptoms while
+others remain asymptomatic.
+We rely on data by \cite{Davies2020} for the age-dependent probability to develop
+symptoms. It varies from 25\% for children and young adults to nearly 70\% for the
+elderly.
 
-The period between infection and onset of infectiousness is called latent or latency
-period. However, the latency period is rarely given in epidemiological reports on
-Covid-19. Instead, scientists and agencies usually report the incubation period, the
-period from infection to the onset of symptoms. A few studies used measurements of virus
-shedding to estimate infectiousness during the course of the disease. When measurements
-started before the onset of symptoms the development of the viral load before symptoms
-gives us an indication of number of days between the onset of infectiousness and
-symptoms.
+% Incubation Period
+The incubation period is usually estimated to be two to twelve days.
+A meta analysis by \citet{McAloon2020} comes to the conclusion that
+``The incubation period distribution may be modeled with a lognormal distribution with
+pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50
+(95\% CI 0.46 to 0.55), respectively.''
+For simplicity we discretize this distribution into four bins.
 
-The European Centre for Disease Prevention and Control estimates that people become
-infectious between one and two days before the symptoms set in. This is similar to
-\citet{He2020} who estimate this to take 2.3 days and is in line with \citet{Peak2020}.
+% Length of the presymptomatic stage
+The \href{European Centre for Disease Prevention and Control}{https://bit.ly/3pi18Ax}
+reports that people become infectious between one and two days before symptoms start.\footnotemark
 
-Given these numbers and the length of the incubation period we can calculate the latency
-period for symptomatic people. To our knowledge no estimates for the latency period of
-asymptomatic cases of COVID-19 exist. We assume it to be the same for symptomatic and
-asymptomatic cases.
+\footnotetext{
+This is similar to \citet{He2020} and in line with
+\citet{Peak2020}.}
 
-Thus, we arrive at the following distribution for latency periods: 40\% have one day.
-35\% have two days. 20\% have three days and 5\% have 5 days.
+Taking these estimates together, we arrive at a latent period of one to five days.
 
@@
 asymptomatic individuals as evidence suggests little differences in the transmission
-rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients
+rates between symptomatic and asymptomatic patients
 (\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic
@@
 This duration range is very much in line with the meta-analysis’ reported evidence for
-asymptomatic individuals (see their figure 1). Thus, we arrive at 0 to 13 days as the
-range for infectiousness among individuals who become symptomatic. This duration range
-is very much in line with the meta-analysis' reported evidence for asymptomatic
-individuals.
+asymptomatic individuals (see their figure 1).
 
@@
 The data on how many percent of symptomatic patients will require ICU is pretty thin. We
-rely on data by the US CDC (\citet{Stokes2020}) and
-\href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the
+rely on data by the US CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the
 OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the

### 2021-06-01 paper/abstract.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

@@
 \noindent
-Governments worldwide have been adopting diverse and nuanced policy measures
-to contain the spread of Covid-19.
-However, epidemiological models usually lack the detailed representation of
-human meeting patterns to credibly predict the effects such policies.
+In the second year of the pandemic, governments worldwide have been adopting diverse and nuanced policy measures to contain the spread of Covid-19.
+Examples are targeted contact reduction policies, mask mandates and other hygiene measures, vaccinations and large scale rapid testing.
+Standard epidemiological models are not designed to predict the effect of very fine grained policies, nor to model the effect of many interacting measures at a time.
 %
-We propose a novel simulation-based model to address these shortcomings.
-We build on state-of-the-art agent-based simulation models, greatly increasing
-the amount of detail and realism with which contacts take place.
-Firstly, we allow for different contact types (such as work, school, households
-or leisure), distinguish recurrent and non-recurrent contacts and allow the
-infectiousness of meetings to vary between contact types.
-% Secondly, we strictly separate the number of contacts from the probabilities
-% that a contact leads to an infection.
-Secondly, we allow agents to seek tests and react to information, such as
-experiencing symptoms, receiving a positive test or a known case among their
-contacts, by reducing their own contacts.
-% Thus, social distancing policies can affect the number of contacts while the
-% infection probabilities remain invariant.
-This allows us to model the effects of a wide array very targeted policies such
-as split classes, mandatory work from home schemes or test-and-trace policies.
-To validate our model, we show that it can predict the effect of the German
-November lockdown even if no similar policy has been observed during the time
-period that was used to estimate the model parameters.
+We propose a novel simulation-based model to address these shortcomings. We build on
+state-of-the-art agent-based simulation models, greatly increasing the amount of detail
+and realism with which contacts take place. Firstly, we allow for different contact
+types (such as work, school, households or leisure), distinguish recurrent and
+non-recurrent contacts and allow the infectiousness of meetings to vary between contact
+types.
+% Secondly, we strictly separate the number of contacts from the probabilities that a
+% contact leads to an infection.
+Secondly, we allow agents to seek tests and react to information, such as experiencing
+symptoms, receiving a positive test or a known case among their contacts, by reducing
+their own contacts.
+% Thus, social distancing policies can affect the number of contacts while the infection
+% probabilities remain invariant.
+The model is able to explain the surprising drop in case numbers observed in many european countries in the spring of 2021 as the combined effect of seasonality,
+vaccinations, contact reduction policies and rapid testing and allows to quantify
+the contribution each policy measure has.
 

### 2021-06-01 paper/results.tex
Add figure labels and add figure with school scenarios.
Have observed and total new cases side by side for scenarios that involve testing.

@@
 \centering
+\caption{The Effect of Different School Scenarios on Observed New Cases}
+\label{fig:school_scenarios}
+\begin{subfigure}{.6\textwidth}
   \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/one_off_and_combined/full_new_known_case}
-    \caption{The Effect of Each Channel on Observed New Cases Separately}
-    \label{fig:}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/one_off_and_combined/full_newly_infected}
+\end{subfigure}
+\caption{The Effect of Policies on Observed and Unobserved Cases}
+\label{fig:explain_decline}
 \figurenotes{\textcolor{red}{\ldots}}
 \end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/school_scenarios/full_new_known_case}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/school_scenarios/full_newly_infected}
+\end{subfigure}
+\caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
+\figurenotes{\textcolor{red}{\ldots}}
+\label{fig:school_scenarios}
+\end{figure}
+
+\FloatBarrier
+
+\input{../figures/results/tables/student_infections_table}

### 2021-06-01 task_compile_documents.py

@@ -15,6 +15,7 @@ BLD = ROOT / "bld"
         "--synctex=1",
         "--cd",
         "--shell-escape",
+        "-f",
     ]
 )
 @pytask.mark.depends_on(ROOT / "paper" / "paper.tex")

### 2021-06-01 paper/model_validation.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>
Add fitness plots.

@@
 
-We validate our model in two ways: 1) We look at the in-sample fit over the estimation
-period. 2) We look at the out-of-sample fit for November. The last one is a challenging
-test for our model because there was a strong policy change between the estimation
-period and November. The model convincingly passes both tests.
+This compares simulated data from our model with empirical data from Germany. We at
+observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
+rapid test demands. Where available we do not only look at aggregated statistics but
+also analyze the model fit for age groups.
 
-\subsection{In-Sample Fit}
+\textcolor{red}{summarize the fit}
+
+\subsection{Observed Infections}
 \label{sub:in_sample_fit}
 
-Despite fitting only four infection probabilities, the in-sample fit is very good.
-The best fit is achieved in the largest age groups.
-This is so mechanically, because we weigh the
-deviations between simulated and observed infection rates by group sizes. The worst fit
-is achieved for the 80 to 100 years old.
-There are three reasons for this: Firstly, we only model private households at
-the moment, meaning that nursing homes are not part of the synthetic population we
-simulate. Since community housing residents are part of the German Mikrozensus
-\citep{FDSAeDBUDL2018} which forms the basis for our synthetic population, we can and
-plan to include community housing residents in the near future.
-Secondly, the elderly, especially nursing home residents, are tested more often than
-the general population leading to more cases being detected in this age group.
-We are currently working to include age-variant testing strategies,
-using \href{https://ars.rki.de/Content/COVID19/Main.aspx}{data by the RKI}
-\citep{Seifried2020}.
+% Despite fitting only four infection probabilities, the in-sample fit is very good.
+% The best fit is achieved in the largest age groups.
+% This is so mechanically, because we weigh the
+% deviations between simulated and observed infection rates by group sizes. The worst fit
+% is achieved for the 80 to 100 years old.
+% There are three reasons for this: Firstly, we only model private households at
+% the moment, meaning that nursing homes are not part of the synthetic population we
+% simulate. Since community housing residents are part of the German Mikrozensus
+% \citep{FDSAeDBUDL2018} which forms the basis for our synthetic population, we can and
+% plan to include community housing residents in the near future.
+% Secondly, the elderly, especially nursing home residents, are tested more often than
+% the general population leading to more cases being detected in this age group.
+% We are currently working to include age-variant testing strategies,
+% using \href{https://ars.rki.de/Content/COVID19/Main.aspx}{data by the RKI}
+% \citep{Seifried2020}.
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
-    \includegraphics[width=1.1\textwidth]{../figures/goodness_of_fit_by_age_group}
-    \caption{Reported vs. simulated weekly incidence rates of infections.}
-    \label{fig:goodness_of_fit}
-    \figurenotes{
-        The figure shows the weekly incidence rates per 100,000 people for the reported
-        (red line) versus the simulated infections rates (blue line) for age groups
-        available in the data provided by the Robert-Koch Institut.
-    }
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/fall_fit/full_new_known_case}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/spring_fit/full_new_known_case}
+\end{subfigure}
+\caption{Simulated and Empirical Infections}
+\figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates.}
+\label{fig:aggregated_fit}
 \end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_fall_baseline_new_known_case}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_spring_baseline_new_known_case}
+\end{subfigure}
+\caption{Simulated and Empirical Infections by Age Group}
+\figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different age groups.}
+\label{fig:age_group_fit}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/state/full_fall_baseline_new_known_case}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/state/full_spring_baseline_new_known_case}
+\end{subfigure}
+\caption{Simulated and Empirical Infections by Age Group}
+\figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different federal regions.}
+\label{fig:state_fit}
+\end{figure}
+
+\FloatBarrier
+
+\section{The Spread of the B117 Mutation}
+
+\textcolor{red}{We don't have this figure in the repo yet, but the fit is perfect}
 

### 2021-06-02 paper/abstract.tex
List of bullets for report for Science, first draft.

--- 2021-02-09 paper/abstract.tex
@@
 \noindent
-Governments worldwide have been adopting diverse and nuanced policy measures
-to contain the spread of Covid-19.
-However, epidemiological models usually lack the detailed representation of
-human meeting patterns to credibly predict the effects such policies.
-%
-We propose a novel simulation-based model to address these shortcomings.
-We build on state-of-the-art agent-based simulation models, greatly increasing
-the amount of detail and realism with which contacts take place.
-Firstly, we allow for different contact types (such as work, school, households
-or leisure), distinguish recurrent and non-recurrent contacts and allow the
-infectiousness of meetings to vary between contact types.
-% Secondly, we strictly separate the number of contacts from the probabilities
-% that a contact leads to an infection.
-Secondly, we allow agents to seek tests and react to information, such as
-experiencing symptoms, receiving a positive test or a known case among their
-contacts, by reducing their own contacts.
-% Thus, social distancing policies can affect the number of contacts while the
-% infection probabilities remain invariant.
-This allows us to model the effects of a wide array very targeted policies such
-as split classes, mandatory work from home schemes or test-and-trace policies.
-To validate our model, we show that it can predict the effect of the German
-November lockdown even if no similar policy has been observed during the time
-period that was used to estimate the model parameters.
+
+% Abstracts of Research Articles and Reports should explain to the general reader why the research was done, what was found and why the results are important.
+
+% They should start with some brief BACKGROUND information: a sentence giving a broad introduction to the field comprehensible to the general reader, and then a sentence of more detailed background specific to your study.
+
+Background:
+
+Spread disease, initially NPIs, now testing and vaccines; seasonality unclear.
+
+% This should be followed by an explanation of the OBJECTIVES/METHODS and then the RESULTS.
+
+Objective: Provide a model that allows studying these things in conjunction, allowing for different virus strains.
+
+Results: Along the transition to vaccination-induced herd immunity, testing is most effective, also thanks to family structures. Seasonality contributed its share.
+
+% The final sentence should outline the main CONCLUSIONS of the study, in terms that will be comprehensible to all our readers.
+The Abstract is distinct from the main body of the text, and thus should not be the only source of background information critical to understanding the manuscript.
+Please do not include citations or abbreviations in the Abstract.
+
+Conclusions: Frequent rapid testing should remain part of strategies to contain CoViD-19.
+
+% The abstract should be 125 words or less.
 

### 2021-06-02 paper/report.tex
List of bullets for report for Science, first draft.

\section{Report}

Main contributions
\begin{itemize}
      \item Pattern: Large 2nd/3rd waves, B.1.1.7 dominant variant, steep fall in spring
      \item Many policies and developments at once: Seasonality, NPIs (private / school closures / workplace restrictions), , vaccinations, tests.
      \item Model that makes the most of many available data sources to gauge the relative effects in this transition period -- see joint distribution of infections with age, geography. But contacts with age, geography, occupations. 
      \item Intuitive -- directly work with primitives
      \item Very general approach: Limited only by size of computer and availability of data
      \item Too many policies at once for SEIR extensions
\end{itemize}

% Main Text is not divided into sub-headings for Reports.

% The manuscript should start with a brief introduction describing the paper’s significance. The introduction should provide sufficient background information to make the article intelligible to readers in other disciplines, and sufficient context that the significance of the experimental findings is clear.

\paragraph{Motivation}
\begin{itemize}
    \item CoViD-19 pandemic huge challenge. 
    \item Vaccination hold promise that will can be contained. 
    \item But most countries find themselves still many months---and often years---away from reaching herd immunity through vaccinations.
    \item Important what they do in the meantime
    \item Measures initially hard lockdown, but always crept back and not sustainable in long run
    \item Measures become increasingly diverse: Testing, Vaccination, ...
    \item Effect of seasonality unclear.
    \item Striking patterns in Europe. 
    \item Important to have models explaining the spread, which take advantage of data in a timely fashion and predict effects of interventions
\end{itemize}

\paragraph{Standard approaches }
\begin{itemize}
    \item Most common class of models: SEIR
    \item Not well suited at rapid developments, multiple changes at once, gauging effects
\end{itemize}

\paragraph{Broad description of our model}
\begin{itemize}
    \item Agent-based models promising alternative
    \item So far high level of abstraction
    \item Put interactions between heterogeneous people at center
    \item Core innovations: Recurrent contacts and assortative matching
    \item Various settings: Home, social, school, work
\end{itemize}

\paragraph{Further important features }
\begin{itemize}
    \item Vaccinations including remaining infectiousness
    \item Tests including false positives / negatives
    \item Various strains
    \item Progression of CoViD-19 and infectiousness
    \item Detailed schooling models (A/B classes, full closures, hygiene multipliers,...)
    \item Detailed work models (work from home)
    \item Model for undetected infections to make publicly disclosed data usable (only thing that requires bespoke tailoring for other countries?)
\end{itemize}

\paragraph{Advantages of our model}
\begin{itemize}
    \item Very intuitive, yet easy to extend. Curse of Dimensionality / Computing power are the only restrictions
    \item Allows to use wide array of data (contacts, census, mobility, infection rates, prevalence of strains, vaccination rates, test rates, vaccination / test efficacy, NPIs, ...)
\end{itemize}

\paragraph{Example: Germany}
\begin{itemize}
    \item Good first response
    \item Then relaxed over the summer at low incidence rates
    \item Cross-country mobility planted the seeds for fall wave
\end{itemize}

\paragraph{Story by Figures}
\begin{itemize}
    \item Fig 1: Model fit and measures
    \item Fig 2: Scenarios, Shapley
    \item Fig 3: Schooling scenarios? Including vacation effects in fall? Similar for Easter?
\end{itemize}

\paragraph{Points to mention}
\begin{itemize}
    \item Social structure / conditional block testing in families important (?)
    \item Trump-effect: More testing = more cases true for how long?
\end{itemize}

% Technical terms should be defined.

% Symbols, abbreviations, and acronyms should be defined the first time they are used.

% All tables and figures should be cited in numerical order.

% All data must be shown either in the main text or in the Supplementary Materials or must be available in an established database with accession details provided in the acknowledgements section.

% References to unpublished materials are not allowed to substantiate significant conclusions of the paper.

\section{Supplementary Material}

\begin{enumerate}
    \item Model
    \item Data
    \item Identification and Estimation
\end{enumerate}

### 2021-06-03 paper/calibration_and_estimation.tex
Add current result figures and tables.

@@
 
-The model is described by a large number of parameters that govern the number of
-contacts a person has, the likelihood of becoming infected on each contact, the
-likelihood of developing light or strong symptoms or even dying from the disease as well
-as the duration each stage of the disease takes.
+The model is described by a large number of parameters that govern the number of contacts
+a person has, the likelihood of becoming infected on each contact, the likelihood of
+developing light or strong symptoms or even dying from the disease as well as the
+duration each stage of the disease takes.
 
@@
 
-We calibrate the parameters for the predicted numbers of contacts from contact diaries
-of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
+We calibrate the parameters for the predicted numbers of contacts from contact diaries of
+over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
 \citep{Mossong2008}. Each contact diary contains all contacts an individual had
@@
 and information on the contact. Importantly, for each contact individuals entered of
-which type the contact (school, leisure, work etc.) was and how frequent the contact
-with the other person is.
+which type the contact (school, leisure, work etc.) was and how frequent the contact with
+the other person is.
 
@@
 geographical assortativity and just roughly calibrate it such that 60 \% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.
+within the same county, we can calibrate the assortative mixing by age from the same data
+we use to calibrate the number of contacts.
 
@@
         groups. Higher values are colored in darker red tones. The diagonal represents
-        the share of contacts with individuals from the same age group.
-    }
+        the share of contacts with individuals from the same age group. }
 \end{figure}
 
-Figure~\ref{fig:assortativity} shows that assortativity by age is especially strong
-for children and younger adults. For older people, the pattern becomes more dispersed
-around their own age group, but within-age-group contacts are still the most common
-contacts.
+Figure~\ref{fig:assortativity} shows that assortativity by age is especially strong for
+children and younger adults. For older people, the pattern becomes more dispersed around
+their own age group, but within-age-group contacts are still the most common contacts.
 
@@
 
-Currently, we use data for Germany from August until November.
-We do not use earlier periods to save computational time.
-Moreover, we would be worried that the there are seasonal effects that we currently do
-not model.
+Currently, we use data for Germany from August until November. We do not use earlier
+periods to save computational time. Moreover, we would be worried that the there are
+seasonal effects that we currently do not model.
 
@@
 
-In our empirical application we distinguish four groups of contact types:
-households, education, work and other contacts.
-For households we assume that the individuals' contacts in their households do not change
-over our estimation period.
-For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures. For the moment we ignore both emergency
-childcare and that lack of childcare leads working parents to stay home.
+In our empirical application we distinguish four groups of contact types: households,
+education, work and other contacts. For households we assume that the individuals'
+contacts in their households do not change over our estimation period. For nurseries,
+preschools and schools we implement vacations as announced by the German federal states
+as well as school closures. For the moment we ignore both emergency childcare and that
+lack of childcare leads working parents to stay home.
 %
@@
 \FloatBarrier
+
+\subsection{Rapid Test Demand}
+
+In our model, there are five reasons why rapid tests are done:
+\begin{enumerate}
+    \item someone plans to have work contacts
+    \item someone is an employees of an educational facility or school pupil
+    \item someone lives in a household where someone has tested positive or developed symptoms
+    \item someone has developed symptoms but has not received a PCR test
+    \item someone plans to participate in a weekly non-work meeting
+\end{enumerate}
+
+% work rapid tests
+
+For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April) that 60\% of workers who receive a test offer by their employer regularly use it.
+We assume this to be time constant.
+
+In addition, there are some surveys that allow us to trace the expansion of employers who
+offer tests to their employees. Mid march, 20\% of employers offered tests to their
+employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
+offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
+60\% until the first days of April \cite{ZDF2021}. Until mid April 70\% of workers were
+expected to receive a weekly test offer \citep{AerzteZeitung2021}. However, according to
+surveys conducted in mid April \citep{Betsch2021}, less than two thirds of individuals
+with work contacts receive a test offer. Starting on April 19th employers were required by law to provide two weekly tests to their employees \citep{Bundesanzeiger2021}.
+We assume that compliance is incomplete and only 80\% of employers actually offer tests.
+
+% educ rapid tests
+
+\textcolor{red}{Sources still missing below this}
+
+We assume that employees in educational facilities start getting tested in 2021 and that
+by March 1st 30\% of them are tested weekly.
+The share increases to 90\% for the week before Easter.
+At that time both Bavaria and Baden-Württemberg were offering tests to teachers and North-Rhine Westphalia and Lower Saxony were already testing students and tests for students and teachers were already mandatory in Saxony.
+After Easter we assume that 95\% of teachers get tested twice per week.
+
+Tests for students started later so we assume that they only start in February and only
+10\% of students get tested by March 1st. Relying on the same sources as above we
+approximate that by the week before Easter this share had increased to 40\%.
+
+After Easter we assume the share of students receiving twice weekly tests to
+based on tests already being mandatory in North-Rhine Westphalia and Bavaria while still
+being voluntary in Baden-Württemberg. There tests become mandatory on April 19th.
+
+% household member rapid tests
+
+% own symptom rapid test demand
+
+% private meeting test demand
+
+\textcolor{red}{
+    Add a section on how we calibrate rapid test demand; Mainly describe the datapoints we have and say that we usually interpolate linearly in between data points.
+(Only exception to that is private rapid test demand, which we fit to data)
+}

### 2021-06-03 paper/calibration_and_estimation.tex
Update work multiplier figure and add figure of rapid tests shares.

@@
     \centering
-    \includegraphics[width=1.1 \textwidth]{../figures/work_multiplier.png}
-    \caption{Share of workers attending work normally.}
+    \includegraphics[width=1.1 \textwidth]{../figures/results/figures/work_multiplier_since_sep}
+    \caption{Share of Workers with Work Contacts}
     \label{fig:work_multiplier}
-    \figurenotes{
-        The figure shows the percentage of pre-pandemic work mobility taking place
-        since the start of the pandemic. In our model this is used as a proxy for the
-        share of workers that attend work as usual. Workers are ordered with the
-        importance of personal contacts for their work. Thus, if the share of
-    }
+    \figurenotes{The figure shows the work mobility as reported by \cite{Google2021}.
+We take this as a proxy of the share of workers who are not in home office, i.e.
+who still have physical work contacts.}
 \end{figure}
@@
 \textcolor{red}{
-    Add a section on how we calibrate rapid test demand; Mainly describe the datapoints we have and say that we usually interpolate linearly in between data points.
-(Only exception to that is private rapid test demand, which we fit to data)
+    Add a section on how we calibrate rapid test demand; Mainly describe the
+    datapoints we have and say that we usually interpolate linearly in between data
+    points. (Only exception to that is private rapid test demand, which we fit to data)
 }
+
+\begin{figure}[ht]
+    \centering
+    \caption{Share of Individuals Having Ever Done a Rapid Test}
+    \label{fig:share_ever_rapid_test}
+    \begin{subfigure}{.6\textwidth}
+      \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/spring_fit/full_share_ever_rapid_test}
+    \end{subfigure}%
+    \begin{subfigure}{.6\textwidth}
+      \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/spring_fit/full_share_rapid_test_in_last_week}
+    \end{subfigure}
+    \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
+    \label{fig:explain_decline}
+    \figurenotes{\textcolor{red}{\ldots}}
+\end{figure}

### 2021-06-03 paper/calibration_and_estimation.tex
Update the work multiplier figure and add figure on the share of individuals with rapid tests.

@@
     \centering
-    \includegraphics[width=1.1 \textwidth]{../figures/results/figures/work_multiplier_since_sep}
+    \includegraphics[width=1.1 \textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
     \caption{Share of Workers with Work Contacts}
@@
     \centering
-    \caption{Share of Individuals Having Ever Done a Rapid Test}
+    \caption{Share of Individuals With Rapid Tests}
     \label{fig:share_ever_rapid_test}
@@
     \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
-    \label{fig:explain_decline}
-    \figurenotes{\textcolor{red}{\ldots}}
+    \label{fig:share_rapid_test_last_week}
+    \figurenotes{The figure compares the share of individuals who have ever done a rapid test or done a rapid test within the last week in our simulations to the shares reported in the \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19 Snapshot Monitoring survey}.
+The
+    left panel compares the share of individuals who have ever done a rapid test.
+The right panel compares the share of individuals who have done a rapid test within the last seven days in our simulation compared to the share reporting to have done at least weekly rapid tests in the last four weeks in the COSMO survey.
+Overall our calibration of rapid tests are slightly conservative.
+The overall share is below that in the study.
+We fit the share of weekly tests quite exactly.
+However, the study only covers adults while our share also includes children who are tested very regularly when attending school.
+    }
 \end{figure}

### 2021-06-03 paper/model_validation.tex
Add comment.
Adjust figures to use new results figures.

@@
 
-\textcolor{red}{summarize the fit}
+\textcolor{red}{summarize the fit; Some of this needs to go in the appendix. I suggest the age group and federal state fit goes there???}
 
@@
 \centering
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/fall_fit/full_new_known_case}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/spring_fit/full_new_known_case}
-\end{subfigure}
+  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
 \caption{Simulated and Empirical Infections}
@@
 \centering
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_fall_baseline_new_known_case}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_spring_baseline_new_known_case}
-\end{subfigure}
+  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
 \caption{Simulated and Empirical Infections by Age Group}
@@
 \centering
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/state/full_fall_baseline_new_known_case}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/incidences_by_group/state/full_spring_baseline_new_known_case}
-\end{subfigure}
-\caption{Simulated and Empirical Infections by Age Group}
-\figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different federal regions.}
+  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
+\caption{Simulated and Empirical Infections by Federal State}
+\figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different federal states.}
 \label{fig:state_fit}

### 2021-06-03 paper/results.tex
Adjust figures to use new results figures.

--- 2021-06-01 paper/results.tex
@@
 \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/one_off_and_combined/full_new_known_case}
-    \caption{The Effect of Each Channel on Observed New Cases Separately}
-    \label{fig:}
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+\end{subfigure}
+\caption{The Effect of Policies on Observed and Unobserved Cases}
+\label{fig:explain_decline}
 \figurenotes{\textcolor{red}{\ldots}}
 \end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/placeholder}
+  %../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/placeholder}
+  %../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected
+\end{subfigure}
+\caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
+\figurenotes{\textcolor{red}{K: One of the scenarios starts too early. Will be fixed with the next full simulations run.}}
+\label{fig:school_scenarios}
+\end{figure}
+
+\FloatBarrier
+
+\input{../figures/results/tables/student_infections_table}

### 2021-06-03 paper/introduction.tex
Line breaks.
Correct some model features and typos.

@@
 The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions
-across the world.
-The reaction to second waves consisted of more targeted policy measures such as splitting school classes, closing restaurants and encouragement of home office.
-To combat the third and fourth waves, that were triggered by the spread of more infectious variants, governments relied on the same targeted measures and previously non-available tools such as large scale rapid testing and vaccinations.
+across the world. The reaction to second waves consisted of more targeted policy
+measures such as splitting school classes, closing restaurants and encouragement of home
+office. To combat the third and fourth waves, that were triggered by the spread of more
+infectious variants, governments relied on the same targeted measures and previously
+non-available tools such as large scale rapid testing and vaccinations.
 
-While this multitude of policy measures has led to declining case numbers in most countries, it becomes harder and harder to evaluate the contribution each policy had on the overall outcome.
-Moreover, the longer the pandemic lasts, the more important it becomes to accurately model heterogeneities in contact behavior and a realistic network of recurrent contacts because some sub-populations might develop herd immunity even though the full population does not.
+While this multitude of policy measures has led to declining case numbers in most
+countries, it becomes harder and harder to evaluate the contribution each policy had on
+the overall outcome. Moreover, the longer the pandemic lasts, the more important it
+becomes to accurately model heterogeneities in contact behavior and a realistic network
+of recurrent contacts because some sub-populations might develop herd immunity even
+though the full population does not.
 
-The workhorse model of epidemiology, the S(E)IR model as well as many recent extensions to it are not up to this task.
-We develop an agent-based simulation model that has been designed from the ground up to predict and quantify the effects of contact reducing policies, vaccinations and testing strategies in a constantly changing policy environment.
-It has the following features:
+The workhorse model of epidemiology, the S(E)IR model as well as many recent extensions
+to it are not up to this task. We develop an agent-based simulation model that has been
+designed from the ground up to predict and quantify the effects of contact reducing
+policies, vaccinations and testing strategies in a constantly changing policy
+environment. It has the following features:
 
@@
           it is possible to implement split class schooling where only one half of each
-          class attends and the attending half switches on a weekly bases. The extent to which contacts are reduced can be calibrated from observed data or estimated inside the model.
+          class attends and the attending half switches on a weekly bases. The extent to
+          which contacts are reduced can be calibrated from observed data or estimated
+          inside the model.
 
@@
 
-    \item We distinguish detected and undetected cases. The share of detected cases varies over time and across age groups. Moreover, it can be influenced by rapid testing policies.
+    \item We distinguish detected and undetected cases. The share of detected cases
+          varies over time and across age groups. Moreover, it can be influenced by rapid
+          testing policies.
+
     \item High quality Python code for the model is freely available on Github, well
-          documented and very flexible\footnotemark. We are actively looking for
-          researchers who want to use our model for their projects and apply it to other
-          contexts.
+          documented and very flexible\footnotemark.
 \end{enumerate}
@@
 \footnotetext{
-    The code can be found under \url{https://github.com/covid-19-impact-lab/sid} and the
-    documentation with tutorials and background information under
+    The code can be found at \url{https://github.com/covid-19-impact-lab/sid} and the
+    documentation with tutorials and background information at
     \url{https://sid-dev.readthedocs.io/}.
@@
 
-The model achieves a good fit on German data of infection and fatality rates, even though most parameters are calibrated from the literature and observable datasets and only a few parameters are estimated inside the model.
-At accurately predicts the rise of the B117 mutation in spring 2020.
-It can also explain the surprising decline of case numbers at the end of April, without making ad hoc assumptions on behavioral changes at the time.
+The model achieves a good fit on German data of infection and fatality rates, even
+though most parameters are calibrated from the literature and observable datasets and
+only a few parameters are estimated inside the model. It accurately predicts the rise of
+the B117 mutation in spring 2020. It can also explain the surprising decline of case
+numbers at the end of April, without making ad hoc assumptions on behavioral changes at
+the time.
 
-The model has previously been applied to predict the effect of schooling policies, contact tracing policies over the Christmas holidays and the effect of work from home (\citet{Dorn2020a, Gabler2020, Gabler2021}).
+The model has previously been applied to predict the effect of schooling policies,
+contact tracing policies over the Christmas holidays and the effect of work from home
+(\citet{Dorn2020a, Gabler2020, Gabler2021}).
 
@@
 
-The remainder is structured as follows: In Section~\ref{sec:literature_review}, we give a short overview of epidemiological models.
-Section ~\ref{sec:background} summarizes the policy environments and dynamic of the Covid-19 pandemic in Europe, with a special focus on Germany.
-We continue with a general description of our modelling framework  as well as a specialization to the German context in section~\ref{sec:model}.
-Section~\ref{sec:calibration_and_estimation} describes our empirical datasets, sources of calibrated parameters and estimation procedure.
-\textcolor{blue}{desribe result section}.
-We conclude in section~\ref{sec:conclusion}.
+The remainder is structured as follows: In Section~\ref{sec:literature_review}, we give
+a short overview of epidemiological models. Section ~\ref{sec:background} summarizes the
+policy environments and dynamic of the Covid-19 pandemic in Europe, with a special focus
+on Germany. We continue with a general description of our modelling framework  as well
+as a specialization to the German context in section~\ref{sec:model}.
+Section~\ref{sec:calibration_and_estimation} describes our empirical datasets, sources
+of calibrated parameters and estimation procedure. \textcolor{blue}{desribe result
+section}. We conclude in section~\ref{sec:conclusion}.

### 2021-06-03 paper/report.tex
Correct some model features and typos.

@@
 \begin{itemize}
-    \item Vaccinations including remaining infectiousness
+    \item Vaccinations including vaccination priority according to age and systemic relevance of work
     \item Tests including false positives / negatives
     \item Various strains
-    \item Progression of CoViD-19 and infectiousness
+    \item Age dependent susceptibility and progression of CoViD-19
     \item Detailed schooling models (A/B classes, full closures, hygiene multipliers,...)
-    \item Detailed work models (work from home)
+    \item Detailed work models (work from home, work contact priority, daily, weekly and non-recurrent work contacts)
     \item Model for undetected infections to make publicly disclosed data usable (only thing that requires bespoke tailoring for other countries?)

### 2021-06-03 paper/model/testing.tex
Update testing section to include rapid tests.

@@
 
-We support to model testing as consisting of three stages. Firstly, we model who demands
-a test. Demand functions map from individual characteristics to a probability which is
-the probability for this individual to demand a test. There can be multiple demand
-functions where each function may describe a different channel. For example, individuals
-who experience symptoms or have a risk contact may ask for a test. Or, the ministry of
-education requires a negative test result from every teacher every second week. After
-the probabilities for each individual and every demand model are computed, individuals
-who demand a test as well as the channel is sampled.
+Our model includes both PCR tests which are scarce and take time until the result becomes
+available and rapid tests which are done after an individual's contacts are determined
+but before the contacts take place.
+
+PCR testing consists of three stages. Firstly, we model who demands a PCR test. Demand
+functions map from individual characteristics to a probability which is the probability
+for this individual to demand a test. There can be multiple demand functions where each
+function may describe a different channel. For example, individuals who experience
+symptoms or have a risk contact may ask for a test. Or, the ministry of education
+requires a negative test result from every teacher every second week. After the
+probabilities for each individual and every demand model are computed, individuals who
+demand a test as well as the channel is sampled.
 
@@
 
+Rapid tests are modeled much simpler. Every day before individuals have contacts they can
+decide to be tested. For example students that plan to attend school that day and have
+not done a rapid test in the last three days get a rapid test. Then they immediately
+receive the test result. After they have received their test result individuals can react
+to it by reducing their contacts. For example positively tested individuals may not go to
+work and reduce their household contacts to some degree. Who reduces their contacts to
+what degree depends on a quarantine compliance attribute.
+
+Our rapid tests include false positives and false negatives. The sensitivity of rapid tests in our model depends on when the individual has
+or will become infectious. This way we can account for the fact that rapid tests are
+likely to be false positive before infectiousness starts.
+
 % subsection testing (end)

### 2021-06-03 paper/calibration_and_estimation.tex
Adjust figures to use new results figures.
Fix geographical assortativity to 80%.

@@
 probability to depend on age and county of residence. While we do not have good data on
-geographical assortativity and just roughly calibrate it such that 60 \% of contacts are
+geographical assortativity and just roughly calibrate it such that 80\% of contacts are
 within the same county, we can calibrate the assortative mixing by age from the same data
@@
     \centering
-    \includegraphics[width=1.1 \textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
     \caption{Share of Workers with Work Contacts}
@@
     \label{fig:share_ever_rapid_test}
-    \begin{subfigure}{.6\textwidth}
+    \begin{subfigure}{.55\textwidth}
       \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/spring_fit/full_share_ever_rapid_test}
     \end{subfigure}%
-    \begin{subfigure}{.6\textwidth}
+    \begin{subfigure}{.55\textwidth}
       \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/spring_fit/full_share_rapid_test_in_last_week}

### 2021-06-03 paper/model/number_of_contacts.tex
Some smallish updates; compile on Linux.

@@
 
-In our empirical application we distinguish the following types of contacts:
+In our empirical application we distinguish the following types of contacts:\comment[id=HM]{I started updating this to understand it better myself.
+In the end, I think it will be useful to group into HH / work / schools and add an introductory sentence for each group.
+E.g., that work is for everybody who is working --- except there is a different model for teachers (?), ...}
 
@@
     \item Households: Each household member meets all other household members every day.
-    The household sizes and structures are calibrated to be representative for Germany.
+    % Household sizes and structures are calibrated to be representative for Germany.
 
-    \item Random non-work contacts: Each person has contacts with randomly drawn other
-    people which are assortative with respect to region and age group. This contact type
-    reflects contacts during pure leisure activities as well as non leisure activities
-    such as grocery shopping or medical appointments.
+    \item Recurrent daily non-work contacts representing very close family or friendship
+    relations outside the household.
 
-    \item Recurrent daily non-work contacts: Each person has daily recurring contacts
-    which allows to model close relationships other than families between individuals.
+    \item Recurrent weekly non-work contacts representing friendships, sports groups,
+    charity work, or other activities. We randomize over the days on which the meetings take place.\comment[id=HM]{True?}
 
-    \item Recurrent weekly non-work contacts: Each person has weekly recurring contacts
-    like sports groups or other weekly activities.
+    \item Random non-work contacts: Contacts with randomly drawn other
+    people, which are assortative with respect to region and age group. This contact type
+    reflects contacts during leisure activities, grocery shopping, medical appointments,
+    etc..
 
-    \item Random work contacts: Each working adult has contact to randomly drawn other
-    people at work.
+    \item Recurrent daily work contacts, capturing direct colleagues.
 
-    \item Recurrent daily work contacts: Each working adult meets other workers every
-    day. This is meant to capture work colleagues.
+    \item Recurrent weekly work contacts, representing meetings with clients, superiors, or more distant colleagues.
+    We randomize over the days on which the meetings take place.
 
-    \item Recurrent weekly work contacts: Each working adult meets other workers once
-    per week. We randomize over the days on which the meetings take place. This is meant
-    to capture meetings with clients, superiors or other colleagues which happen
-    infrequently.
+    \item Random work contacts: Working adults' contacts with randomly drawn other
+    people, which are assortative in geographical location\comment[id=HM]{True? age?}
 
@@
     also meet three pairs of teachers every school day. The pairs are meant to represent
-    interactions between teachers.
+    interactions between teachers.\comment[id=HM]{Not updated from here on. Mayb}
 

### 2021-06-03 paper/report.tex
Some smallish updates; compile on Linux.

@@
-\section{Report}
+% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance.
+Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references.
+Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions.
 
-Main contributions
+\section{Main contributions}
+
 \begin{itemize}
@@
 
-\paragraph{Motivation}
+\clearpage
+
+Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity on many dimensions.
+The development of highly effective vaccines holds the promise of containment in the medium term.
+However, most countries find themselves many months---and often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} It is thus
+
 \begin{itemize}
-    \item CoViD-19 pandemic huge challenge. 
-    \item Vaccination hold promise that will can be contained. 
-    \item But most countries find themselves still many months---and often years---away from reaching herd immunity through vaccinations.
     \item Important what they do in the meantime
-    \item Measures initially hard lockdown, but always crept back and not sustainable in long run
+    \item Measures initially hard lockdown, but always crept back and not sustainable in long run. Hardly helpful on their own for more infectious strains
     \item Measures become increasingly diverse: Testing, Vaccination, ...

### 2021-06-04 paper/medical_parameters.tex
Model description in report mostly done.

@@
 arrived at the distributions used in the model.\footnotemark See
-Figure~\ref{fig:course_of_disease} for a summary of our disease progression model.
+Figure~\ref{fig:disease_progression} for a summary of our disease progression model.
 

### 2021-06-04 paper/model/course_of_disease.tex
Start on updating the medical parameters.
Model description in report mostly done.

@@
 \subsection{Course of the Disease}
-\label{sub:course_of_disease}
+\label{sub:disease_progression}
 
-The following medical parameters describing the progression of the disease are taken
-from systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
-progresses in the way depicted in Figure~\ref{fig:course_of_disease}.
-
-\begin{figure}[!tp]
-    \includegraphics[width=\textwidth]{../figures/disease_progression.png}
-    \caption{Course of Disease in the model.}
-    \label{fig:course_of_disease}
-    \figurenotes{
-        The figure depicts the course of the disease from infection to either the state
-        of recovery or death.
-    }
-\end{figure}
+The following medical parameters describing the progression of the disease are taken from
+systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
+progresses in the way depicted in Figure~\ref{fig:disease_progression}.
 
@@
 important because those people do not reduce their contacts or demand a test and can
-potentially infect many other people \citep{Donsimoni2020}.
+potentially infect many other people \citep{Donsimoni2020}. The probability to develop
+symptoms with Covid-19 is highly age dependent with 75\% of children not developing
+symptoms \citep{Davies2020}.
 
 A small share of symptomatic people will develop strong symptoms that require intensive
-care. The exact share and time span is age-dependent. An age-dependent share of
-intensive care unit (ICU) patients will die after spending up to 32 days in intensive
-care. Moreover, if the ICU capacity was reached, all patients who require intensive care
-but do not receive it die.
+care. The exact share and time span is age-dependent. An age-dependent share of intensive
+care unit (ICU) patients will die after spending up to 32 days in intensive care.
+Moreover, if the ICU capacity was reached, all patients who require intensive care but do
+not receive it die.
 
@@
 could model people who require hospitalization but not intensive care. So far we opted
-against that because only the intensive care capacities are feared to become a
-bottleneck in Germany.
+against that because only the intensive care capacities are feared to become a bottleneck
+in Germany.
 
@@
 
-\footnotetext{
-    Detailed information on the calibration of the disease parameters is available as
-    part of our
-    \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online documentation}.
-}
+\footnotetext{ Detailed information on the calibration of the disease parameters is
+    available as part of our
+    \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
+    documentation}. }

### 2021-06-04 paper/model/testing.tex
Model description in report mostly done.

@@
 Rapid tests are modeled much simpler. Every day before individuals have contacts they can
-decide to be tested. For example students that plan to attend school that day and have
+decide to be tested.\comment[id=HM]{If we get econ readers, we should not use language like this because we do not have a choice model.} For example students that plan to attend school that day and have
 not done a rapid test in the last three days get a rapid test. Then they immediately

### 2021-06-04 paper/report.tex
Model description in report mostly done.

@@
 The development of highly effective vaccines holds the promise of containment in the medium term.
-However, most countries find themselves many months---and often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} It is thus
+However, most countries find themselves many months---and often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost importance to employ an effective mix of strategies for containing the virus.
+The most frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals.
+While this has allowed some countries to sustain equilibria with very low infection numbers\comment[id=HM]{Cite Priesemann paper or other}, most have seen large ups and downs in their infection rates.
+Containment measures have become increasingly diverse and included testing, more nuanced NPIs, and contact tracing.
+In quantitative terms, neither these policies' effect nor the influence of seasonal patterns or more infectious virus strains are well understood.
+This paper develops a model incorporating all these influences.
+The framework allows to combines a wide variety of data in a timely fashion and to predict the effects of various interventions.
+We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert something useful}
 
-\begin{itemize}
-    \item Important what they do in the meantime
-    \item Measures initially hard lockdown, but always crept back and not sustainable in long run. Hardly helpful on their own for more infectious strains
-    \item Measures become increasingly diverse: Testing, Vaccination, ...
-    \item Effect of seasonality unclear.
-    \item Striking patterns in Europe. 
-    \item Important to have models explaining the spread, which take advantage of data in a timely fashion and predict effects of interventions
-\end{itemize}
+At the core of our agent-based model are physical contacts between heterogeneous agents (See Figure~\ref{fig:model_graph}a).\footnote{We provide a detailed comparison to other approaches in \ref{sec:literature_review}, where we also review other approaches.
+The model most closely related to ours is described in \citet{Hinch2020}.} Each contact between an infectious indvidual and somebody susceptible to the disease bears the risk of transmitting the virus.
+Contacts occur in the household, at work, at school, or in other settings (leisure activities, grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of work?} Some contacts recur regularly, others occur at random.
+Random contacts are typically assortative in age and geographical location.
+Empirical applications can take the population structure from census data and the types and frequency of contacts from diary data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008}.\comment[id=HM]{cite other stuff, too?} The dimensions are chosen so that the most common NPIs can be modelled in great detail by reducing the number of contacts in a particular setting or the risk of transmitting the disease for a type of contact.
+For example, a mandate to work from home will reduce the number of work contacts to zero for a fraction of the working population.
+Schools and daycare can be closed entirely, operate at reduced capacity including an alternating schedule, or implement mitigation measures like masking requirements or air filters.\comment[id=HM]{cite \url{https://science.sciencemag.org/content/early/2021/04/28/science.abh2939}} Curfews may reduce the number of contacts in non-work/non-school settings.
+In any setting, measures like masking requirements would reduce the probability of infection associated with a contact.
 
-\paragraph{Standard approaches }
-\begin{itemize}
-    \item Most common class of models: SEIR
-    \item Not well suited at rapid developments, multiple changes at once, gauging effects
-\end{itemize}
+\begin{figure}[!tp]
+    \centering
 
-\paragraph{Broad description of our model}
-\begin{itemize}
-    \item Agent-based models promising alternative
-    \item So far high level of abstraction
-    \item Put interactions between heterogeneous people at center
-    \item Core innovations: Recurrent contacts and assortative matching
-    \item Various settings: Home, social, school, work
-\end{itemize}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \caption{{\small Model description}}
+        \label{fig:change_working_hours_by_edu}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../figures/disease_progression.png}
+        \caption{Disease progression}
+        \label{fig:disease_progression}
+    \end{subfigure}
 
-\paragraph{Further important features }
-\begin{itemize}
-    \item Vaccinations including vaccination priority according to age and systemic relevance of work
-    \item Tests including false positives / negatives
-    \item Various strains
-    \item Age dependent susceptibility and progression of CoViD-19
-    \item Detailed schooling models (A/B classes, full closures, hygiene multipliers,...)
-    \item Detailed work models (work from home, work contact priority, daily, weekly and non-recurrent work contacts)
-    \item Model for undetected infections to make publicly disclosed data usable (only thing that requires bespoke tailoring for other countries?)
-\end{itemize}
+    \vskip3ex
 
-\paragraph{Advantages of our model}
-\begin{itemize}
-    \item Very intuitive, yet easy to extend. Curse of Dimensionality / Computing power are the only restrictions
-    \item Allows to use wide array of data (contacts, census, mobility, infection rates, prevalence of strains, vaccination rates, test rates, vaccination / test efficacy, NPIs, ...)
-\end{itemize}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        PCR tests: quarantine (?)
+
+        Antigen tests: subsequent PCR tests, quarantine, work/school
+
+        Effects on other people (tracing)
+
+        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \caption{{\small PCR and antigen tests}}
+        \label{fig:pcr_antigen_tests}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        Model for detected/undetected cases
+
+        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \caption{{\small Model for translating simulated number to officially recorded cases}}
+        \label{fig:model_for_official_cases}
+    \end{subfigure}
+
+    \caption{Model description}
+\end{figure}
+
+Susceptibility to contracting CoViD-19 is dependent on age and so is the progression of the disease.
+We differentiate between an inital period of infection without being infectious or showing symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and recovery or death.
+The probabilities of transitioning between these states depend on age; their duration is random within intervals calibrated to medical literature.\comment[id=HM]{Cite}.
+Conditional on the type of contact, infectiousness independent of ages.\comment[id=HM]{Cite current Drosten paper}
+
+The model includes several other features, which are crucial to describe the evolution of the pandemic in 2020-2021.
+New virus strains with different profiles regarding infectiousness and disease progress can be introduced.
+Vaccinations induce immunity; vaccinated agents do not transmit the virus.\footnote{All of these approximations could be easily relaxed at the cost of a higher computational burden.} While vaccines are rolled out, priority may depend on age and occupation.
+Agents may demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}.
+First, polymerase chain reaction (PCR) tests directly reveal whether an individual is infected or not.
+PCR tests require some time to be processed and there are always aggregate capacity constraints.
+Second, rapid antigen tests yield immediate results.
+Specificity and sensitivity of these tests is set according to the XXX data.\comment[id=HM]{subsequent PCR tests, quarantine, work/school}.
+Demand for rapid tests will depend on policy, e.g., prices, testing requirements before work, attending school, or visiting restaurants and shops.
+
+Modelling a population of agents according to actual demographic characterists means that we can use a wide array of data to estimate the model parameters.
+Mobility data is used to inform reductions in work and social contacts.\comment[id=HM]{check} School and daycare policies can be incorporated directly from official directives.
+Infection rates by age and geographical region are matched to officially recorded numbers, so is the prevalence of virus strains.
+In order to translate simulated cases in our model to officially recorded cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
 
@@
 \begin{itemize}
-    \item Good first response
+    \item Good first response, relatively lax (no curfew, few business closures, ... )
     \item Then relaxed over the summer at low incidence rates
     \item Cross-country mobility planted the seeds for fall wave
+    \item We start our analysis at this point because modelling undetected cases would require a very different approach than in spting
 \end{itemize}
@@
 \begin{itemize}
-    \item Fig 1: Model fit and measures
-    \item Fig 2: Scenarios, Shapley
-    \item Fig 3: Schooling scenarios? Including vacation effects in fall? Similar for Easter?
+    \item Fig 2: Model fit and measures
+
+          Aggregate. Refer to appendix for fits by age group and geographical region.
+
+    \item Fig 3: Scenarios, Shapley. Why testing so effective. Form of TTI.
+
+    \item Fig 4: Schooling scenarios? Including vacation effects in fall? Similar for Easter?
+
 \end{itemize}

### 2021-06-04 paper/report.tex
Add more precise draft for figures.

@@
 New virus strains with different profiles regarding infectiousness and disease progress can be introduced.
-Vaccinations induce immunity; vaccinated agents do not transmit the virus.\footnote{All of these approximations could be easily relaxed at the cost of a higher computational burden.} While vaccines are rolled out, priority may depend on age and occupation.
+With a probability calibrated to XXX\comment[id=HM]{Klara, please add citatiion}, vaccinated agents become immune and they do not transmit the virus.
+While vaccines are rolled out, priority may depend on age and occupation.
 Agents may demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}.
@@
 
+Identification
+
 Modelling a population of agents according to actual demographic characterists means that we can use a wide array of data to estimate the model parameters.
@@
 In order to translate simulated cases in our model to officially recorded cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
+
+Simulated data have very similar structure to data that is used for regression analyses. Can use as further plausbility check.
 
@@
 
-\paragraph{Story by Figures}
+\begin{figure}[!tp]
+    \centering
+
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        Stringency index + multipliers
+
+        Contact reduction $\times$ hygiene multiplier for different types of contacts
+
+        potentially $\times$ seasonality
+
+        else put seasonality separately
         \begin{itemize}
-    \item Fig 2: Model fit and measures
+            \item Work
+            \item School
+            \item Other
+        \end{itemize}
 
-          Aggregate. Refer to appendix for fits by age group and geographical region.
+        \caption{{\small Contact reductions}}
+        \label{fig:stringency_index}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
 
-    \item Fig 3: Scenarios, Shapley. Why testing so effective. Form of TTI.
+        Number of PCR tests
 
-    \item Fig 4: Schooling scenarios? Including vacation effects in fall? Similar for Easter?
+        Fraction of pupils with antigen tests
 
-\end{itemize}
+        Fraction of workers with antigen tests
+
+        Fraction of vaccinations (effective = lagged by two weeks or date of vaccination?)
+
+        \vskip2ex
+
+        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \caption{{\small Tests and vaccinations}}
+        \label{fig:pcr_antigen_tests_vaccinations}
+    \end{subfigure}
+    \vskip3ex
+
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        % \includegraphics[width=\textwidth]{../figures/disease_progression.png}
+        Share known cases (estimated)
+
+        Share B.1.1.7
+
+        \vskip2ex
+
+        \caption{Share of known cases and fraction of B.1.1.7 strain}
+        \label{fig:share_known_cases_b117}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+        \caption{{\small Recorded cases: Empirical and simulated}}
+        \label{fig:aggregated_fit}
+    \end{subfigure}
+
+    \caption{Drivers of the pandemic and model fit, September 2020 to May 2021}
+    \label{fig:pandemic_drivers_model_fit}
+
+    Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+
+\end{figure}
+
+\begin{figure}[!tp]
+    \centering
+
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+        \caption{{\small Recorded cases: 2021 scenarios}}
+        \label{fig:full_new_known_case_cropped}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+        \caption{{\small Total cases: 2021 scenarios}}
+        \label{fig:full_newly_infected_cropped}
+    \end{subfigure}
+    \vskip3ex
+
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        % \includegraphics[width=\textwidth]{../figures/disease_progression.png}
+
+        Shapley values
+
+        \vskip2ex
+
+        \caption{Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}}
+        \label{fig:decompisition total cases}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+
+        Effects of social structure (implicit tracing) / effects of different types of testing
+
+        \vskip2ex
+
+        \caption{{\small Effects of different types of testing}}
+        \label{fig:decomposition_tests}
+    \end{subfigure}
+
+    \caption{The effect of different interventions on recorded and actual infections}
+    \label{fig:interventions_broad}
+
+    Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+
+\end{figure}
+
+\begin{figure}[!tp]
+    \centering
+
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+        % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+        \caption{{\small Effects of different schooling scenarios after Easter}}
+        \label{fig:schooling_scenarios_easter}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+        \caption{{\small TBD}}
+        \label{fig:full_newly_infected_cropped}
+    \end{subfigure}
+    \vskip3ex
+
+    \caption{Schooling / maybe home office}
+    \label{fig:interventions_school}
+
+    Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+
+\end{figure}
 
@@
 \begin{itemize}
+    \item If anything too optimistic regarding vaccinations
     \item Social structure / conditional block testing in families important (?)

### 2021-06-04 paper/report.tex
Add figures to be combined. Fix some label issues.

@@
         \caption{{\small Model description}}
-        \label{fig:change_working_hours_by_edu}
+        \label{fig:model_graph}
     \end{subfigure}
@@
         \centering
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
+
         Stringency index + multipliers
@@
 
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/share_of_educ_participants_with_rapid_test}
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/share_of_workers_with_rapid_test_offer_at_work}
+        % fraction of workers must be multiplied with 0.6 because only that many actually
+        % accept the offer
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_individuals_with_first_vaccine}
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_receiving_vaccination_per_day}
+
         Number of PCR tests
-
-        Fraction of pupils with antigen tests
-
-        Fraction of workers with antigen tests
-
-        Fraction of vaccinations (effective = lagged by two weeks or date of vaccination?)
 
@@
 
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
         \caption{{\small Tests and vaccinations}}
@@
         \centering
-        % \includegraphics[width=\textwidth]{../figures/disease_progression.png}
-        Share known cases (estimated)
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_b117_acc_to_rki}
 
-        Share B.1.1.7
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/assumed_overall_share_known_cases}
 
@@
         \caption{{\small TBD}}
-        \label{fig:full_newly_infected_cropped}
+        \label{fig:to_be_determined}
     \end{subfigure}

### 2021-06-04 paper/report.tex
Add preliminary model graphs.
Add a couple of paragraphs.

@@
 
-\clearpage
+% \clearpage
 
@@
 Containment measures have become increasingly diverse and included testing, more nuanced NPIs, and contact tracing.
-In quantitative terms, neither these policies' effect nor the influence of seasonal patterns or more infectious virus strains are well understood.
-This paper develops a model incorporating all these influences.
-The framework allows to combines a wide variety of data in a timely fashion and to predict the effects of various interventions.
-We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert something useful}
+Neither these policies' effect nor the influence of seasonal patterns or more infectious virus strains are well understood in quantitative terms.
+This paper develops a model incorporating all these factors.
+The framework allows to combine a wide variety of data in a timely fashion, making it useful to predict the effects of various interventions.
+We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert some useful numbers}
 
-At the core of our agent-based model are physical contacts between heterogeneous agents (See Figure~\ref{fig:model_graph}a).\footnote{We provide a detailed comparison to other approaches in \ref{sec:literature_review}, where we also review other approaches.
+At the core of our agent-based model are physical contacts between heterogeneous agents (Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to other approaches in \ref{sec:literature_review}.
 The model most closely related to ours is described in \citet{Hinch2020}.} Each contact between an infectious indvidual and somebody susceptible to the disease bears the risk of transmitting the virus.
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
         \caption{{\small Model description}}
-        \label{fig:model_graph}
+        \label{fig:broad_model_description}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/disease_progression.png}
+        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
         \caption{Disease progression}
@@
 
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
         \caption{{\small PCR and antigen tests}}
@@
 
-Susceptibility to contracting CoViD-19 is dependent on age and so is the progression of the disease.
+Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the progression of a possible infection.
 We differentiate between an inital period of infection without being infectious or showing symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and recovery or death.
@@
 
-Identification
+Modelling a population of agents according to actual demographic characterists means that we can use a wide array of data to identify and estimate the model's many parameters.\footnote{See section S.X of the supplementary materials for an overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the parameters from the model and then we brainstorm re notation?} Mobility data is used to model the reductions in work contacts.
+School and daycare policies are incorporated directly from official directives.
+Infection rates by age and geographical region are estimated to match to officially recorded numbers; so is the prevalence of virus strains.
+In order to translate simulated cases in our model to officially recorded cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
+\comment[id=HM]{Add a sentence.} A further advantage is that the simulated data have a structure that resembles datasets used for regression models, which allows additional plausibility checks by re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we actually do something like that.}
 
-Modelling a population of agents according to actual demographic characterists means that we can use a wide array of data to estimate the model parameters.
-Mobility data is used to inform reductions in work and social contacts.\comment[id=HM]{check} School and daycare policies can be incorporated directly from official directives.
-Infection rates by age and geographical region are matched to officially recorded numbers, so is the prevalence of virus strains.
-In order to translate simulated cases in our model to officially recorded cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
+We apply this model to Germany.
+In March and April 2020, the country broke the first wave of the pandemic fairly quickly.
+Between mid-May and mid-September, daily new infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We start our analysis in mid-September for two reasons.
+First, we do not include the first wave because the environment was very different (e.g., aggregate PCR test capacity was much lower and we would require a very different model for calculating the share of known cases) and some data was not recorded yet.\comment[id=HM]{True?}.
+Many cases during the summer of 2020 were traced to international travellers; our model is not very helpful in explaining this.
 
-Simulated data have very similar structure to data that is used for regression analyses. Can use as further plausbility check.
-
-\paragraph{Example: Germany}
-\begin{itemize}
-    \item Good first response, relatively lax (no curfew, few business closures, ... )
-    \item Then relaxed over the summer at low incidence rates
-    \item Cross-country mobility planted the seeds for fall wave
-    \item We start our analysis at this point because modelling undetected cases would require a very different approach than in spting
-\end{itemize}
+In late summer and early fall, restrictions were few.
+A distance of 1.5m had to be kept and masking requirements were enforced in public locations.\comment[id=HM]{pls double-check, add if I am missing important stuff}.
+Figure~\ref{fig:pandemic_drivers_model_fit} describes the background for our analysis.
+Figure~\ref{fig:stringency_index} Important to remember that multipliers are relative to that period
 
@@
         \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
-
         Stringency index + multipliers
@@
         \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_individuals_with_first_vaccine}
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_receiving_vaccination_per_day}
 
@@
         \centering
-
         \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
@@
     \end{subfigure}
+
     \vskip3ex

### 2021-06-05 paper/report.tex
Another paragraph, shuffled figures around a little bit.

@@
     \end{subfigure}
+    \hfill
     \begin{subfigure}[b]{0.475\textwidth}
@@
 First, we do not include the first wave because the environment was very different (e.g., aggregate PCR test capacity was much lower and we would require a very different model for calculating the share of known cases) and some data was not recorded yet.\comment[id=HM]{True?}.
-Many cases during the summer of 2020 were traced to international travellers; our model is not very helpful in explaining this.
+Second, a large fraction cases during summer of 2020 were traced to international travel,\comment[id=HM]{citation?} which is not part of this model.
 
+We describe the evolution of the pandemic and of its drivers in Figure~\ref{fig:pandemic_drivers_model_fit}.
+The black line in Figure~\ref{fig:aggregated_fit} shows the officially recorded cases; the \replace[id=HM]{color} line in
+
+The setting in our analysis 
 In late summer and early fall, restrictions were few.
 A distance of 1.5m had to be kept and masking requirements were enforced in public locations.\comment[id=HM]{pls double-check, add if I am missing important stuff}.
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the background for our analysis.
  Figure~\ref{fig:stringency_index} Important to remember that multipliers are relative to that period
@@
 
+    \begin{subfigure}[b]{0.475\textwidth}
+        \centering
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+        \caption{{\small Recorded cases: Empirical and simulated}}
+        \label{fig:aggregated_fit}
+    \end{subfigure}
+    \hfill
     \begin{subfigure}[b]{0.475\textwidth}
@@
     \end{subfigure}
-    \hfill
+
+    \vskip3ex
+
     \begin{subfigure}[b]{0.475\textwidth}
@@
     \end{subfigure}
-    \vskip3ex
-
+    \hfill
     \begin{subfigure}[b]{0.475\textwidth}
@@
     \end{subfigure}
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{\small Recorded cases: Empirical and simulated}}
-        \label{fig:aggregated_fit}
-    \end{subfigure}
-
-    \caption{Drivers of the pandemic and model fit, September 2020 to May 2021}
+    \caption{Evolution of the pandemic, its drivers, and model fit, September 2020 to May 2021}
     \label{fig:pandemic_drivers_model_fit}

### 2021-06-05 paper/report.tex
Fix mistyped command.

@@
 Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions.
-
-\section{Main contributions}
-
-\begin{itemize}
-    \item Pattern: Large 2nd/3rd waves, B.1.1.7 dominant variant, steep fall in spring
-    \item Many policies and developments at once: Seasonality, NPIs (private / school closures / workplace restrictions), , vaccinations, tests.
-    \item Model that makes the most of many available data sources to gauge the relative effects in this transition period -- see joint distribution of infections with age, geography.
-But contacts with age, geography, occupations.
-    \item Intuitive -- directly work with primitives
-    \item Very general approach: Limited only by size of computer and availability of data
-    \item Too many policies at once for SEIR extensions
-\end{itemize}
 
@@
 
+Model that makes the most of many available data sources to gauge the relative effects in this transition period -- see joint distribution of infections with age, geography.
+But contacts with age, geography, occupations.
+
 We apply this model to Germany.
@@
 First, we do not include the first wave because the environment was very different (e.g., aggregate PCR test capacity was much lower and we would require a very different model for calculating the share of known cases) and some data was not recorded yet.\comment[id=HM]{True?}.
-Second, a large fraction cases during summer of 2020 were traced to international travel,\comment[id=HM]{citation?} which is not part of this model.
+Second, a large fraction cases during summer of 2020 were traced to international travel,\comment[id=HM]{citation?}, but the precise number is difficult to model.
 
-We describe the evolution of the pandemic and of its drivers in Figure~\ref{fig:pandemic_drivers_model_fit}.
-The black line in Figure~\ref{fig:aggregated_fit} shows the officially recorded cases; the \replace[id=HM]{color} line in
+At the starting point of our analysis, the
+
+The \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020}. The Figure~\ref{fig:stringency_index}
+
+We describe the evolution of the pandemic and of its drivers in Figure~\ref{fig:pandemic_drivers_model_fit}. The black line in Figure~\ref{fig:aggregated_fit} shows the officially recorded cases;
+
+We transform the index so that 1 represents the situation in mid-September and 0 represents that all restrictions incorporated in the index are in place.
+The six weeks between  was a steep rise in infections in
 

### 2021-06-05 paper/report.tex
Add a couple of paragraphs.

@@
 The framework allows to combine a wide variety of data in a timely fashion, making it useful to predict the effects of various interventions.
-We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert some useful numbers}
+We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert some useful numbers}.
+We conclude that rapid tests should play a large role...
 
@@
 In March and April 2020, the country broke the first wave of the pandemic fairly quickly.
-Between mid-May and mid-September, daily new infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We start our analysis in mid-September for two reasons.
+Between mid-May and mid-September, daily new infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the period mid-September 2020 to the end of May 2020.
+We pick the starting date for two reasons.
 First, we do not include the first wave because the environment was very different (e.g., aggregate PCR test capacity was much lower and we would require a very different model for calculating the share of known cases) and some data was not recorded yet.\comment[id=HM]{True?}.
@@
 
-At the starting point of our analysis, the
+Figure~{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and of its drivers.
+The black line in Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
+We transform the index so that lower values represent higher levels of restrictions.
+A value of zero means all measures incorporated in the index are turned on.
+The value 1 represents the situation of mid-September, where half of all restrictions incorporated in the index are active.\comment[id=HM]{Need to find version of index with a couple of measures, probably need the API access...}.
+In the six weeks between mid September and mid October, cases increased by a factor of \replaced[id=HM]{X}{10?
+Do we have RKI numbers as table somewhere?}.
+Restrictions were somewhat tightened in mid-October and again in early November.
+New infections remained constant throughout November, before rising again in December.
+In mid December, other lockdown.
+So far, traditional measures only
 
-The \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020}. The Figure~\ref{fig:stringency_index}
+Starting in January,  becomes relevant.
+\begin{itemize}
+    \item Vaccines
+    \item Tests
+    \item Seasonality March onwards.
+    \item B.1.1.7 February onwards
+\end{itemize}
 
-We describe the evolution of the pandemic and of its drivers in Figure~\ref{fig:pandemic_drivers_model_fit}. The black line in Figure~\ref{fig:aggregated_fit} shows the officially recorded cases;
-
-We transform the index so that 1 represents the situation in mid-September and 0 represents that all restrictions incorporated in the index are in place.
-The six weeks between  was a steep rise in infections in
-
-The setting in our analysis 
-In late summer and early fall, restrictions were few.
-A distance of 1.5m had to be kept and masking requirements were enforced in public locations.\comment[id=HM]{pls double-check, add if I am missing important stuff}.
- Figure~\ref{fig:stringency_index} Important to remember that multipliers are relative to that period
+Our model:
+\begin{itemize}
+    \item Fit is great
+    \item Multipliers for three categories (how calculated precisely, refer to appendix for breakdown)
+    \item Tests fit well
+    \item Share known cases, how calculated
+\end{itemize}
 
@@
 
+Reassured by good fit and a couple of external checks (fall vacation regression?), we consider the relative effects of different measures in 2021
+\begin{itemize}
+    \item Rapid tests, 
+    \item seasonality
+    \item vaccines
+\end{itemize}
+
+lessons learned
+
 \begin{figure}[!tp]
@@
 
+Schooling very contentious. Likely damage to child development.
+
+We show
+
+Screening effect: ...
+
 \begin{figure}[!tp]

### 2021-06-05 paper/calibration_and_estimation.tex
Add explanation how we handle weekend work contacts.
Some smallish updates; compile on Linux.
Add new results figures.

@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/assortative_matching_probability_example.png}
+    \includegraphics[width=0.9 \textwidth]{../figures/assortative_matching_probability_example}
     \caption{Distribution of random non-work contacts by age of participants.}
@@
 % Schüler in die Schule.
-% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430.png
+% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430
 % 18% der Grundschüler in Notbetreuung in BW
@@
     \label{fig:work_multiplier}
-    \figurenotes{The figure shows the work mobility as reported by \cite{Google2021}.
-We take this as a proxy of the share of workers who are not in home office, i.e.
-who still have physical work contacts.}
+    \figurenotes{The figure shows the work mobility as reported by \cite{Google2021}. We
+    take this as a proxy of the share of workers who are not in home office, i.e. who
+    still have physical work contacts. The figure interpolates over weekends as we handle
+    weekend effects through information on work on weekends in the German census data we
+    use.}
 \end{figure}
@@
     \begin{subfigure}{.55\textwidth}
-      \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/spring_fit/full_share_ever_rapid_test}
+      \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
     \end{subfigure}%
     \begin{subfigure}{.55\textwidth}
-      \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/spring_fit/full_share_rapid_test_in_last_week}
+      \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
+    %   \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
     \end{subfigure}
-    \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
     \label{fig:share_rapid_test_last_week}

### 2021-06-05 paper/model/summary.tex
Model description in report mostly done.
Add new results figures.

@@
 
-To predict and quantify the effects of a wide variety of fine-grained social distancing policies, vaccinations and rapid testing, we propose a different model structure. Our
+To predict and quantify the effects of a wide variety of fine-grained social distancing
+policies, vaccinations and rapid testing, we propose a different model structure. Our
 model inherits many features from prototypical agent-based simulation models but
@@
 
-The structure of the model is depicted in Figure~\ref{fig:model_graph}.
-
-\begin{figure}[!tp]
-    \centering
-    \includegraphics[width=\textwidth]{../figures/model_detailed.png}
-    \caption{Simplified graph of the model.}
-    \label{fig:model_graph}
-\end{figure}
+The structure of the model is depicted in Figure~\ref{fig:broad_model_description}.
 
@@
 predicted number of contacts. Currently, we distinguish between eight types of contact
-models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and
-random work contacts, recurrent and random leisure contacts, and nursery, preschool, and
-school contacts.
+models which are all listed in Figure~\ref{fig:broad_model_description}: households,
+recurrent and random work contacts, recurrent and random leisure contacts, and nursery,
+preschool, and school contacts.
 
@@
 Once a person is infected, the disease progresses in a fairly standard way which is
-depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild
+depicted in Figure~\ref{fig:disease_progression}. Asymptomatic cases and cases with mild
 symptoms are infectious for some time and recover eventually. Cases with severe symptoms
@@
 tests there. Moreover, people can decide to make a rapid test if they develop symptoms,
-have many planned contacts or observe cases in their contact network. People who have a positive rapid test demand a confirmatory PCR test with a certain
-probability. Moreover, PCR tests can be demanded because of symptoms or randomly.
+have many planned contacts or observe cases in their contact network. People who have a
+positive rapid test demand a confirmatory PCR test with a certain probability. Moreover,
+PCR tests can be demanded because of symptoms or randomly.
 
@@
 
-People who have symptoms, received a positive test, or had a risk contact
-can reduce their number of contacts across all contact types endogenously. The extent to
-which this is done is calibrated from survey data.
+People who have symptoms, received a positive test, or had a risk contact can reduce
+their number of contacts across all contact types endogenously. The extent to which this
+is done is calibrated from survey data.
 

### 2021-06-05 paper/report.tex
Add new results figures.

--- 2021-06-05 paper/report.tex
@@
         \centering
-
         \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
-        Stringency index + multipliers
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency}
 
-        Contact reduction $\times$ hygiene multiplier for different types of contacts
-
-        potentially $\times$ seasonality
-
-        else put seasonality separately
-        \begin{itemize}
-            \item Work
-            \item School
-            \item Other
-        \end{itemize}
-
+        \textcolor{red}{Where to put seasonality?}
         \caption{{\small Contact reductions}}
@@
 
-        % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-
-        Effects of social structure (implicit tracing) / effects of different types of testing
-
-        \vskip2ex
-
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_rapid_tests/full_newly_infected}
         \caption{{\small Effects of different types of testing}}
@@
         \centering
-        % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
         \caption{{\small Effects of different schooling scenarios after Easter}}
@@
         \centering
-
-        % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
         \caption{{\small TBD}}

### 2021-06-05 paper/results.tex
Add very preliminary Shapley values figure.
Some smallish updates; compile on Linux.
Add new results figures.

@@
 \textcolor{red}{The results can be found in `figures/results`. This includes both figures
-and tables for lookup of numbers and summary tables}
+and tables for lookup of numbers and summary tables.}
 
@@
 \begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/placeholder}
-  %../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
 \end{subfigure}%
 \begin{subfigure}{.6\textwidth}
-  \includegraphics[width=0.9 \textwidth]{../figures/placeholder}
-  %../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
 \end{subfigure}
@@
 \input{../figures/results/tables/student_infections_table}
+
+\FloatBarrier
+
+\begin{figure}[ht]
+\centering
+% \includegraphics[width=\textwidth]{../figures/preliminary_shapley}
+\caption{Shapley Values}
+\figurenotes{\textcolor{red}{K: These are taken from Tobi's notebook as examples. Numbers
+might still change}}
+\label{fig:school_scenarios}
+
+\end{figure}

### 2021-06-05 paper/report.tex
Add some references. Address some of HM's comments. Not running through. Probably missing curly bracket somewhere.

@@
 
-Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity on many dimensions.
-The development of highly effective vaccines holds the promise of containment in the medium term.
-However, most countries find themselves many months---and often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost importance to employ an effective mix of strategies for containing the virus.
-The most frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals.
-While this has allowed some countries to sustain equilibria with very low infection numbers\comment[id=HM]{Cite Priesemann paper or other}, most have seen large ups and downs in their infection rates.
-Containment measures have become increasingly diverse and included testing, more nuanced NPIs, and contact tracing.
-Neither these policies' effect nor the influence of seasonal patterns or more infectious virus strains are well understood in quantitative terms.
-This paper develops a model incorporating all these factors.
-The framework allows to combine a wide variety of data in a timely fashion, making it useful to predict the effects of various interventions.
-We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert some useful numbers}.
-We conclude that rapid tests should play a large role...
+Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
+on many dimensions. The development of highly effective vaccines holds the promise of
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite
+some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers\comment[id=HM]{Cite Priesemann paper or other}, most have
+seen large ups and downs in their infection rates. Containment measures have become
+increasingly diverse and included testing, more nuanced NPIs, and contact tracing.
+Neither these policies' effect nor the influence of seasonal patterns or more infectious
+virus strains are well understood in quantitative terms. This paper develops a model
+incorporating all these factors. The framework allows to combine a wide variety of data
+in a timely fashion, making it useful to predict the effects of various interventions. We
+apply the model to Germany and show that rapid testing had the largest impact on the
+reduction in infections by XXX\% during the A weeks between X April and Y
+May.\comment[id=HM]{insert some useful numbers}. We conclude that rapid tests should play
+a large role...
 
-At the core of our agent-based model are physical contacts between heterogeneous agents (Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to other approaches in \ref{sec:literature_review}.
-The model most closely related to ours is described in \citet{Hinch2020}.} Each contact between an infectious indvidual and somebody susceptible to the disease bears the risk of transmitting the virus.
-Contacts occur in the household, at work, at school, or in other settings (leisure activities, grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of work?} Some contacts recur regularly, others occur at random.
-Random contacts are typically assortative in age and geographical location.
-Empirical applications can take the population structure from census data and the types and frequency of contacts from diary data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008}.\comment[id=HM]{cite other stuff, too?} The dimensions are chosen so that the most common NPIs can be modelled in great detail by reducing the number of contacts in a particular setting or the risk of transmitting the disease for a type of contact.
-For example, a mandate to work from home will reduce the number of work contacts to zero for a fraction of the working population.
-Schools and daycare can be closed entirely, operate at reduced capacity including an alternating schedule, or implement mitigation measures like masking requirements or air filters.\comment[id=HM]{cite \url{https://science.sciencemag.org/content/early/2021/04/28/science.abh2939}} Curfews may reduce the number of contacts in non-work/non-school settings.
-In any setting, measures like masking requirements would reduce the probability of infection associated with a contact.
+At the core of our agent-based model are physical contacts between heterogeneous agents
+(Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to
+other approaches in \ref{sec:literature_review}. The model most closely related to ours
+is described in \citet{Hinch2020}.} Each contact between an infectious indvidual and
+somebody susceptible to the disease bears the risk of transmitting the virus. Contacts
+occur in the household, at work, at school, or in other settings (leisure activities,
+grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of
+work?} Some contacts take place regularly, others occur at random. Random contacts are
+typically assortative in age and geographical location. Empirical applications can take
+the population structure from census data and the types and frequency of contacts from
+diary data measuring contacts before the pandemic
+\citep[e.g.][]{Mossong2008}.\comment[id=HM]{cite other stuff, too?} The dimensions are
+chosen so that the most common NPIs can be modeled in great detail by reducing the
+number of contacts in a particular setting or the risk of transmitting the disease for a
+type of contact. For example, a mandate to work from home will reduce the number of work
+contacts to zero for a fraction of the working population. Schools and daycare can be
+closed entirely, operate at reduced capacity including an alternating schedule, or
+implement mitigation measures like masking requirements or air
+filters.\comment[id=HM]{cite
+\url{https://science.sciencemag.org/content/early/2021/04/28/science.abh2939}} Curfews
+may reduce the number of contacts in non-work/non-school settings. In any setting,
+measures like masking requirements would reduce the probability of infection associated
+with a contact.
 
@@
 
-Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the progression of a possible infection.
-We differentiate between an inital period of infection without being infectious or showing symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and recovery or death.
-The probabilities of transitioning between these states depend on age; their duration is random within intervals calibrated to medical literature.\comment[id=HM]{Cite}.
-Conditional on the type of contact, infectiousness independent of ages.\comment[id=HM]{Cite current Drosten paper}
+Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the
+progression of a possible infection. We differentiate between an initial period of
+infection without being infectious or showing symptoms, being infectious (presymptomatic
+or asymptomatic), showing symptoms, requiring intensive care, and recovery or death. The
+probabilities of transitioning between these states depend on age; their duration is
+random within intervals calibrated to medical literature (for a detailed description see
+Section~\ref{sec:medical_params}. Conditional on the type of contact, infectiousness  is
+independent of ages.\comment[id=HM]{Cite current Drosten paper}
 
-The model includes several other features, which are crucial to describe the evolution of the pandemic in 2020-2021.
-New virus strains with different profiles regarding infectiousness and disease progress can be introduced.
-With a probability calibrated to XXX\comment[id=HM]{Klara, please add citatiion}, vaccinated agents become immune and they do not transmit the virus.
-While vaccines are rolled out, priority may depend on age and occupation.
-Agents may demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}.
-First, polymerase chain reaction (PCR) tests directly reveal whether an individual is infected or not.
-PCR tests require some time to be processed and there are always aggregate capacity constraints.
-Second, rapid antigen tests yield immediate results.
-Specificity and sensitivity of these tests is set according to the XXX data.\comment[id=HM]{subsequent PCR tests, quarantine, work/school}.
-Demand for rapid tests will depend on policy, e.g., prices, testing requirements before work, attending school, or visiting restaurants and shops.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different profiles regarding
+infectiousness and disease progress can be introduced. With a probability calibrated to
+\cite{Hunter2021}, vaccinated agents become immune and they do not transmit the virus
+\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.% the sources only report reduced viral shedding and reduced risk of infection.
+% Lauterbach calculated that taking the risk of infection * reduced infectiousness given
+% infection based on theses studies suggest that given exposure the risk that fully
+% vaccinated individuals pose is <10% of an unvaccinated person.
+While vaccines are rolled out, priority may depend on age and occupation. Agents may
+demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}. First,
+polymerase chain reaction (PCR) tests directly reveal whether an individual is infected
+or not. PCR tests require some time to be processed and there are always aggregate
+capacity constraints. Second, rapid antigen tests yield immediate results. Specificity
+and sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021}
+. The sensitivity of the rapid antigen tests depends on the timing of the test relative
+to the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
+work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
+requirements before work, attending school, or visiting restaurants and shops.
 
-Modelling a population of agents according to actual demographic characterists means that we can use a wide array of data to identify and estimate the model's many parameters.\footnote{See section S.X of the supplementary materials for an overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the parameters from the model and then we brainstorm re notation?} Mobility data is used to model the reductions in work contacts.
-School and daycare policies are incorporated directly from official directives.
-Infection rates by age and geographical region are estimated to match to officially recorded numbers; so is the prevalence of virus strains.
-In order to translate simulated cases in our model to officially recorded cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
-\comment[id=HM]{Add a sentence.} A further advantage is that the simulated data have a structure that resembles datasets used for regression models, which allows additional plausibility checks by re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we actually do something like that.}
+Modelling a population of agents according to actual demographic characteristics means
+that we can use a wide array of data to identify and estimate the model's many
+parameters.\footnote{See section S.X of the supplementary materials for an
+overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the
+parameters from the model and then we brainstorm re notation?} Mobility data is used to
+model the reductions in work contacts. School and daycare policies are incorporated
+directly from official directives. Infection rates by age and geographical region are
+estimated to match to officially recorded numbers; so is the prevalence of virus strains.
+In order to translate simulated cases in our model to officially recorded cases, we use
+the model depicted in Figure~\ref{fig:model_for_official_cases}. \comment[id=HM]{Add a
+sentence.} A further advantage is that the simulated data have a structure that resembles
+datasets used for regression models, which allows additional plausibility checks by
+re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we
+actually do something like that.}
 
@@
 
-We apply this model to Germany.
-In March and April 2020, the country broke the first wave of the pandemic fairly quickly.
-Between mid-May and mid-September, daily new infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the period mid-September 2020 to the end of May 2020.
-We pick the starting date for two reasons.
-First, we do not include the first wave because the environment was very different (e.g., aggregate PCR test capacity was much lower and we would require a very different model for calculating the share of known cases) and some data was not recorded yet.\comment[id=HM]{True?}.
-Second, a large fraction cases during summer of 2020 were traced to international travel,\comment[id=HM]{citation?}, but the precise number is difficult to model.
+We apply this model to Germany. In March and April 2020, the country broke the first wave
+of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
+were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
+period mid-September 2020 to the end of May 2020. We pick the starting date for two
+reasons. First, we do not include the first wave because the environment was very
+different (e.g., aggregate PCR test capacity was much lower and we would require a very
+different model for calculating the share of known cases) and some data was not recorded
+yet.\comment[id=HM]{True?}\comment[id=Klara]{ We would definitely need to model national
+and international travel which you mention below. Datawise I think everything would be
+available. The low numbers during summer in our sample size might also be a problem
+because Covid might go extinct in our model when case numbers are so low. } Second,
+a large fraction cases during summer of 2020 were traced to international
+travel,\comment[id=HM]{citation?}\comment[id=Klara]{There was a comment by minister Spahn
+that half of cases during the summer months were due to travel to the Baltic
+(\url{https://twitter.com/krstorevic/status/1396516025181556737}) but I know of no data.
+We should probably check the RKI reports of July, Aug and Sep} but the precise number is
+difficult to model.}
 
-Figure~{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and of its drivers.
-The black line in Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
-We transform the index so that lower values represent higher levels of restrictions.
-A value of zero means all measures incorporated in the index are turned on.
-The value 1 represents the situation of mid-September, where half of all restrictions incorporated in the index are active.\comment[id=HM]{Need to find version of index with a couple of measures, probably need the API access...}.
-In the six weeks between mid September and mid October, cases increased by a factor of \replaced[id=HM]{X}{10?
-Do we have RKI numbers as table somewhere?}.
-Restrictions were somewhat tightened in mid-October and again in early November.
-New infections remained constant throughout November, before rising again in December.
-In mid December, other lockdown.
-So far, traditional measures only
+Figure~{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and of
+its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially recorded
+cases; the \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the
+Oxford Response Stringency Index\cite{Hale2020}, which tracks the tightness of
+non-pharmaceutical interventions. We transform the index so that lower values represent
+higher levels of restrictions. A value of zero means all measures incorporated in the
+index are turned on. The value 1 represents the situation of mid-September, where half of
+all restrictions incorporated in the index are active.\comment[id=HM]{Need to find
+version of index with a couple of measures, probably need the API access...}. In the six
+weeks between mid September and mid October, cases increased by a factor of
+\replaced[id=HM]{X}{10? Do we have RKI numbers as table somewhere?}. Restrictions were
+somewhat tightened in mid-October and again in early November. New infections remained
+constant throughout November, before rising again in December. In mid December, other
+lockdown. So far, traditional measures only
 
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/share_of_educ_participants_with_rapid_test}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/share_of_workers_with_rapid_test_offer_at_work}
-        % fraction of workers must be multiplied with 0.6 because only that many actually
-        % accept the offer
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_individuals_with_first_vaccine}
-
-        Number of PCR tests
+        \textcolor{red}{Number of PCR tests}\comment[id=K]{I think this is not available.
+        Only a subsection of laboratories report negative PCR tests if I remember right}
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
-        \caption{{\small Effects of different schooling scenarios after Easter}}
-        \label{fig:schooling_scenarios_easter}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
+        \caption{{\small Recorded Cases: Different schooling scenarios after Easter}}
+        \label{fig:schooling_scenarios_easter_new_known_case}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
-        \caption{{\small TBD}}
-        \label{fig:to_be_determined}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \caption{{\small Total Cases: Different schooling scenarios after Easter}}
+        \label{fig:schooling_scenarios_easter_newly_infected}
     \end{subfigure}

### 2021-06-05 paper/report.tex
Restore version that runs through without some of Klaras changes.

--- 2021-06-05 paper/report.tex
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/share_of_educ_participants_with_rapid_test}
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/share_of_workers_with_rapid_test_offer_at_work}
-        % fraction of workers must be multiplied with 0.6 because only that many actually
-        % accept the offer
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
 
@@
 
-        Number of PCR tests
+        \textcolor{red}{Number of PCR tests}
 

### 2021-06-05 paper/report.tex
update figures.

@@
 At the core of our agent-based model are physical contacts between heterogeneous agents (Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to other approaches in \ref{sec:literature_review}.
-The model most closely related to ours is described in \citet{Hinch2020}.} Each contact between an infectious indvidual and somebody susceptible to the disease bears the risk of transmitting the virus.
+The model most closely related to ours is described in \citet{Hinch2020}.} Each contact between an infectious individual and somebody susceptible to the disease bears the risk of transmitting the virus.
 Contacts occur in the household, at work, at school, or in other settings (leisure activities, grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of work?} Some contacts recur regularly, others occur at random.
 Random contacts are typically assortative in age and geographical location.
-Empirical applications can take the population structure from census data and the types and frequency of contacts from diary data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008}.\comment[id=HM]{cite other stuff, too?} The dimensions are chosen so that the most common NPIs can be modelled in great detail by reducing the number of contacts in a particular setting or the risk of transmitting the disease for a type of contact.
+Empirical applications can take the population structure from census data and the types and frequency of contacts from diary data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data sets on contact types and frequencies at \url{http://www.socialcontactdata.org/} covering countries from all continents except North America and Australia.} The dimensions are chosen so that the most common NPIs can be modeled in great detail by reducing the number of contacts in a particular setting or the risk of transmitting the disease for a type of contact.
 For example, a mandate to work from home will reduce the number of work contacts to zero for a fraction of the working population.
@@
     \end{subfigure}
-
     \vskip3ex
-
     \begin{subfigure}[b]{0.475\textwidth}
@@
     \caption{Model description}
+    \label{fig:model-description}
 \end{figure}
@@
 
-Figure~{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and of its drivers.
+Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and of its drivers.
 The black line in Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
@@
         \centering
-        % \includegraphics[width=\textwidth]{../figures/disease_progression.png}
-
-        Shapley values
+        \includegraphics[width=\textwidth]{../figures/full_decomposition}
 
@@
 
-        \caption{Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}}
+        \caption{
+            Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}.
+
+            The decomposition is based on Shapley values where the individual
+            contribution of an intervention is its average contribution over different
+            sizes of coalitions (combinations with other interventions). The individual
+            contribution to a coalition is the difference between the effect size of the
+            coalition with the particular intervention and without.
+        }
         \label{fig:decompisition total cases}

### 2021-06-06 paper/model_validation.tex
Add figures to be combined. Fix some label issues.
Add missing figures/data, comment out missing figures, and clean up subcaption layout.

@@
 \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+  % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
 \caption{Simulated and Empirical Infections}
 \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates.}
-\label{fig:aggregated_fit}
+\label{fig:aggregated_fit2}
 \end{figure}
@@
 \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
+  % \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
 \caption{Simulated and Empirical Infections by Age Group}
@@
 \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
+  % \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
 \caption{Simulated and Empirical Infections by Federal State}

### 2021-06-06 paper/report.tex
Remove request for N PCR tests.
Add missing figures/data, comment out missing figures, and clean up subcaption layout.

@@
 
-        \textcolor{red}{Number of PCR tests}
-
         \vskip2ex
@@
     \end{subfigure}
-    \hfill
+    %
     \begin{subfigure}[b]{0.475\textwidth}
@@
 
-    \vskip3ex
-
     \begin{subfigure}[b]{0.475\textwidth}
@@
         \includegraphics[width=\textwidth]{../figures/full_decomposition}
-
-        \vskip2ex
-
-        \caption{
-            Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}.
-
-            The decomposition is based on Shapley values where the individual
-            contribution of an intervention is its average contribution over different
-            sizes of coalitions (combinations with other interventions). The individual
-            contribution to a coalition is the difference between the effect size of the
-            coalition with the particular intervention and without.
-        }
+        \caption{Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}.
+The decomposition is based on Shapley values where the individual contribution of an intervention is its average contribution over different sizes of coalitions (combinations with other interventions).
+The individual contribution to a coalition is the difference between the effect size of the coalition with the particular intervention and without.}
         \label{fig:decompisition total cases}
     \end{subfigure}
-    \hfill
+    %
     \begin{subfigure}[b]{0.475\textwidth}

### 2021-06-06 paper/report.tex
Small updates.

--- 2021-06-05 paper/report.tex
@@
 
-        PCR tests: quarantine (?)
-
-        Antigen tests: subsequent PCR tests, quarantine, work/school
-
-        Effects on other people (tracing)
-
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
@@
 
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        \includegraphics[width=\textwidth]{../figures/model_detailed.png}
         \caption{{\small Model for translating simulated number to officially recorded cases}}
@@
 New infections remained constant throughout November, before rising again in December.
-In mid December, other lockdown.
-So far, traditional measures only
-
-Starting in January,  becomes relevant.
-\begin{itemize}
-    \item Vaccines
-    \item Tests
-    \item Seasonality March onwards.
-    \item B.1.1.7 February onwards
-\end{itemize}
-
-Our model:
-\begin{itemize}
-    \item Fit is great
-    \item Multipliers for three categories (how calculated precisely, refer to appendix for breakdown)
-    \item Tests fit well
-    \item Share known cases, how calculated
-\end{itemize}
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2}
 
-        \textcolor{red}{Where to put seasonality?}
-        \caption{{\small Contact reductions}}
+        \caption{{\small Stringency of NPIs and changes in infectious contacts by type}}
         \label{fig:stringency_index}
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_individuals_with_first_vaccine}
-
-        \textcolor{red}{Number of PCR tests}
-
-        \vskip2ex
-
         \caption{{\small Tests and vaccinations}}
-        \label{fig:pcr_antigen_tests_vaccinations}
+        \label{fig:antigen_tests_vaccinations}
     \end{subfigure}
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/assumed_overall_share_known_cases}
-
-        \vskip2ex
-
-        \caption{Share of known cases and fraction of B.1.1.7 strain}
-        \label{fig:share_known_cases_b117}
+        \caption{Fraction of B.1.1.7 strain among measured infections}
+        \label{fig:share_b117}
     \end{subfigure}
@@
 
-    Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+    \floatfoot{
+        Note: All aggregates; See S.XXX for statistics by age group and by geographical region. Also more disaggregated data.
+
+        Sources: ... 
+    }
 
 \end{figure}
+
+Quickly rising rates in December prompted the most stringent lockdown to this date.
+Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores.
+Shortly afterwards, more factors become relevant.
+Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}.
+By the end of May, just over 40\% had received at least one dose of a vaccine.
+Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies.
+At-home tests approved by authorities were ...
+in March.\comment[id=HM]{update precise timeline}
+From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
+The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}.
+In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
+In some cases, customers were only allowed to enter with a recent negative rapid test result.
+
+The blue line in Figure~\ref{fig:aggregated_fit} shows the simulated infections based on our model.
+
+Our model:
+\begin{itemize}
+    \item Fit is great
+    \item Multipliers for three categories (how calculated precisely, refer to appendix for breakdown)
+    \item Tests fit well
+    \item Share known cases, how calculated
+\end{itemize}
 
@@
 
-        \caption{
-            Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}.
+        \caption{Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}.}
 
-            The decomposition is based on Shapley values where the individual
-            contribution of an intervention is its average contribution over different
-            sizes of coalitions (combinations with other interventions). The individual
-            contribution to a coalition is the difference between the effect size of the
-            coalition with the particular intervention and without.
-        }
-        \label{fig:decompisition total cases}
+        \label{fig:decomposition_scenarios_total_cases}
     \end{subfigure}
-    \hfill
+    %
     \begin{subfigure}[b]{0.475\textwidth}
@@
 
+    \floatfoot{
         Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
 
+        The decomposition is based on Shapley values where the individual contribution of an intervention is its average contribution over different sizes of coalitions (combinations with other interventions).
+The individual contribution to a coalition is the difference between the effect size of the coalition with the particular intervention and without.
+    }
 \end{figure}

### 2021-06-06 paper/report.tex
Add reference to low case equilibrium by Priesemann.

@@
 The most frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals.
-While this has allowed some countries to sustain equilibria with very low infection numbers\comment[id=HM]{Cite Priesemann paper or other}, most have seen large ups and downs in their infection rates.
+While this has allowed some countries to sustain equilibria with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in their infection rates.
 Containment measures have become increasingly diverse and included testing, more nuanced NPIs, and contact tracing.

### 2021-06-06 paper/report.tex
channel replaces intervention in shapley plot.

@@
 
-        The decomposition is based on Shapley values where the individual contribution of an intervention is its average contribution over different sizes of coalitions (combinations with other interventions).
-The individual contribution to a coalition is the difference between the effect size of the coalition with the particular intervention and without.
+        The decomposition is based on Shapley values where the individual contribution of a channel is its average contribution over different sizes of coalitions (combinations with other channels).
+The individual contribution to a coalition is the difference between the effect size of the coalition with the particular channel and without.
     }

### 2021-06-06 paper/report.tex
Halfway through description of results.

@@
 
-        \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
         \caption{{\small Model for translating simulated number to officially recorded cases}}
@@
     \label{fig:model-description}
+
+    \floatfoot{
+        Note: ...
+    }
+
 \end{figure}
@@
 
-The blue line in Figure~\ref{fig:aggregated_fit} shows the simulated infections based on our model.
+These developments are characteristic of many countries: The inititial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
+At broadly similar points in time, novel strains of the virus have started to pose additional challenges.
+The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well.
+This is also true for infections by age and geographical region, which are shown in the supplementary materials (S.\comment[id=HM]{reference}).
+We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic.
+Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks.
+For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%.
+This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality.
+Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual contributions.
+Two things are particularly interesting.
+First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality.
+Second, the most stringent regulations are associated with the period of strong decreases in new infection beteen late December 2020 and mid-February 2021.
+The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
+The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts.
 
-Our model:
-\begin{itemize}
-    \item Fit is great
-    \item Multipliers for three categories (how calculated precisely, refer to appendix for breakdown)
-    \item Tests fit well
-    \item Share known cases, how calculated
-\end{itemize}
-
-Reassured by good fit and a couple of external checks (fall vacation regression?), we consider the relative effects of different measures in 2021
-\begin{itemize}
-    \item Rapid tests,
-    \item seasonality
-    \item vaccines
-\end{itemize}
-
-lessons learned
+Reassured by good fit and a couple of external checks (fall vacation regression?), Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests, vaccinations, and of seasonality during 2021, assuming NPIs to be the same as in the baseline scenario.
+Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line), and three scenarios turning these factors off one by one.
 
@@
         \caption{{\small Recorded cases: 2021 scenarios}}
-        \label{fig:full_new_known_case_cropped}
+        \label{fig:2021_scenarios_recorded}
     \end{subfigure}
-    %
+    \hfill
     \begin{subfigure}[b]{0.475\textwidth}
@@
         \caption{{\small Total cases: 2021 scenarios}}
-        \label{fig:full_newly_infected_cropped}
+        \label{fig:2021_scenarios_newly_infected}
     \end{subfigure}
@@
 
-        \caption{Decomposition of effects for Figure~\ref{fig:full_newly_infected_cropped}.}
-
-        \label{fig:decomposition_scenarios_total_cases}
+        \caption{Decomposition of effects for Figure~\ref{fig:2021_scenarios_newly_infected}.}
+        \label{fig:2021_scenarios_decomposition}
     \end{subfigure}
-    %
+    \hfill
     \begin{subfigure}[b]{0.475\textwidth}
@@
         \caption{{\small Effects of different types of testing}}
-        \label{fig:decomposition_tests}
+        \label{fig:2021_scenarios_decomposition_tests}
     \end{subfigure}
@@
     \caption{The effect of different interventions on recorded and actual infections}
-    \label{fig:interventions_broad}
+    \label{fig:2021_scenarios_broad}
 
@@
 
+    \floatfoot{
         Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+    }
 

### 2021-06-07 paper/report.tex
Add reference to new Drosten paper.
Add line breaks to the report for easier diffs.

@@
 
-Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity on many dimensions.
-The development of highly effective vaccines holds the promise of containment in the medium term.
-However, most countries find themselves many months---and often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost importance to employ an effective mix of strategies for containing the virus.
-The most frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals.
-While this has allowed some countries to sustain equilibria with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in their infection rates.
-Containment measures have become increasingly diverse and included testing, more nuanced NPIs, and contact tracing.
-Neither these policies' effect nor the influence of seasonal patterns or more infectious virus strains are well understood in quantitative terms.
-This paper develops a model incorporating all these factors.
-The framework allows to combine a wide variety of data in a timely fashion, making it useful to predict the effects of various interventions.
-We apply the model to Germany and show that rapid testing had the largest impact on the reduction in infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert some useful numbers}.
-We conclude that rapid tests should play a large role...
+Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
+on many dimensions. The development of highly effective vaccines holds the promise of
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite
+some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate
+policies.}, most have seen large ups and downs in their infection rates. Containment
+measures have become increasingly diverse and included testing, more nuanced NPIs, and
+contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
+more infectious virus strains are well understood in quantitative terms. This paper
+develops a model incorporating all these factors. The framework allows to combine a wide
+variety of data in a timely fashion, making it useful to predict the effects of various
+interventions. We apply the model to Germany and show that rapid testing had the largest
+impact on the reduction in infections by XXX\% during the A weeks between X April and Y
+May.\comment[id=HM]{insert some useful numbers}. We conclude that rapid tests should play
+a large role...
 
-At the core of our agent-based model are physical contacts between heterogeneous agents (Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to other approaches in \ref{sec:literature_review}.
-The model most closely related to ours is described in \citet{Hinch2020}.} Each contact between an infectious individual and somebody susceptible to the disease bears the risk of transmitting the virus.
-Contacts occur in the household, at work, at school, or in other settings (leisure activities, grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of work?} Some contacts recur regularly, others occur at random.
-Random contacts are typically assortative in age and geographical location.
-Empirical applications can take the population structure from census data and the types and frequency of contacts from diary data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data sets on contact types and frequencies at \url{http://www.socialcontactdata.org/} covering countries from all continents except North America and Australia.} The dimensions are chosen so that the most common NPIs can be modeled in great detail by reducing the number of contacts in a particular setting or the risk of transmitting the disease for a type of contact.
-For example, a mandate to work from home will reduce the number of work contacts to zero for a fraction of the working population.
-Schools and daycare can be closed entirely, operate at reduced capacity including an alternating schedule, or implement mitigation measures like masking requirements or air filters.\comment[id=HM]{cite \url{https://science.sciencemag.org/content/early/2021/04/28/science.abh2939}} Curfews may reduce the number of contacts in non-work/non-school settings.
-In any setting, measures like masking requirements would reduce the probability of infection associated with a contact.
+At the core of our agent-based model are physical contacts between heterogeneous agents
+(Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to
+other approaches in \ref{sec:literature_review}. The model most closely related to ours
+is described in \citet{Hinch2020}.} Each contact between an infectious individual and
+somebody susceptible to the disease bears the risk of transmitting the virus. Contacts
+occur in the household, at work, at school, or in other settings (leisure activities,
+grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of
+work?} Some contacts recur regularly, others occur at random. Random contacts are
+typically assortative in age and geographical location. Empirical applications can take
+the population structure from census data and the types and frequency of contacts from
+diary data measuring contacts before the pandemic
+\citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data
+sets on contact types and frequencies at \url{http://www.socialcontactdata.org/} covering
+countries from all continents except North America and Australia.} The dimensions are
+chosen so that the most common NPIs can be modeled in great detail by reducing the number
+of contacts in a particular setting or the risk of transmitting the disease for a type of
+contact. For example, a mandate to work from home will reduce the number of work contacts
+to zero for a fraction of the working population. Schools and daycare can be closed
+entirely, operate at reduced capacity including an alternating schedule, or implement
+mitigation measures like masking requirements or air filters.\comment[id=HM]{cite
+\url{https://science.sciencemag.org/content/early/2021/04/28/science.abh2939}} Curfews
+may reduce the number of contacts in non-work/non-school settings. In any setting,
+measures like masking requirements would reduce the probability of infection associated
+with a contact.
 
@@
 
-Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the progression of a possible infection.
-We differentiate between an inital period of infection without being infectious or showing symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and recovery or death.
-The probabilities of transitioning between these states depend on age; their duration is random within intervals calibrated to medical literature.\comment[id=HM]{Cite}.
-Conditional on the type of contact, infectiousness independent of ages.\comment[id=HM]{Cite current Drosten paper}
+Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the
+progression of a possible infection. We differentiate between an inital period of
+infection without being infectious or showing symptoms, being infectious (presymptomatic
+or asymptomatic), showing symptoms, requiring intensive care, and recovery or death. The
+probabilities of transitioning between these states depend on age; their duration is
+random within intervals calibrated to medical literature.\comment[id=HM]{Cite}.
+Conditional on the type of contact, infectiousness independent of ages \citep{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution of the pandemic in 2020-2021.
-New virus strains with different profiles regarding infectiousness and disease progress can be introduced.
-With a probability calibrated to XXX\comment[id=HM]{Klara, please add citatiion}, vaccinated agents become immune and they do not transmit the virus.
-While vaccines are rolled out, priority may depend on age and occupation.
-Agents may demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}.
-First, polymerase chain reaction (PCR) tests directly reveal whether an individual is infected or not.
-PCR tests require some time to be processed and there are always aggregate capacity constraints.
-Second, rapid antigen tests yield immediate results.
-Specificity and sensitivity of these tests is set according to the XXX data.\comment[id=HM]{subsequent PCR tests, quarantine, work/school}.
-Demand for rapid tests will depend on policy, e.g., prices, testing requirements before work, attending school, or visiting restaurants and shops.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different profiles regarding
+infectiousness and disease progress can be introduced. With a probability calibrated to
+XXX\comment[id=HM]{Klara, please add citatiion}, vaccinated agents become immune and they
+do not transmit the virus. While vaccines are rolled out, priority may depend on age and
+occupation. Agents may demand two types of tests as described in
+Figure~\ref{fig:pcr_antigen_tests}. First, polymerase chain reaction (PCR) tests directly
+reveal whether an individual is infected or not. PCR tests require some time to be
+processed and there are always aggregate capacity constraints. Second, rapid antigen
+tests yield immediate results. Specificity and sensitivity of these tests is set
+according to the XXX data.\comment[id=HM]{subsequent PCR tests, quarantine, work/school}.
+Demand for rapid tests will depend on policy, e.g., prices, testing requirements before
+work, attending school, or visiting restaurants and shops.
 
-Modelling a population of agents according to actual demographic characterists means that we can use a wide array of data to identify and estimate the model's many parameters.\footnote{See section S.X of the supplementary materials for an overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the parameters from the model and then we brainstorm re notation?} Mobility data is used to model the reductions in work contacts.
-School and daycare policies are incorporated directly from official directives.
-Infection rates by age and geographical region are estimated to match to officially recorded numbers; so is the prevalence of virus strains.
-In order to translate simulated cases in our model to officially recorded cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
-\comment[id=HM]{Add a sentence.} A further advantage is that the simulated data have a structure that resembles datasets used for regression models, which allows additional plausibility checks by re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we actually do something like that.}
+Modelling a population of agents according to actual demographic characterists means that
+we can use a wide array of data to identify and estimate the model's many
+parameters.\footnote{See section S.X of the supplementary materials for an
+overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the
+parameters from the model and then we brainstorm re notation?} Mobility data is used to
+model the reductions in work contacts. School and daycare policies are incorporated
+directly from official directives. Infection rates by age and geographical region are
+estimated to match to officially recorded numbers; so is the prevalence of virus strains.
+In order to translate simulated cases in our model to officially recorded cases, we use
+the model depicted in Figure~\ref{fig:model_for_official_cases}. \comment[id=HM]{Add a
+sentence.} A further advantage is that the simulated data have a structure that resembles
+datasets used for regression models, which allows additional plausibility checks by
+re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we
+actually do something like that.}
 
-Model that makes the most of many available data sources to gauge the relative effects in this transition period -- see joint distribution of infections with age, geography.
-But contacts with age, geography, occupations.
+Model that makes the most of many available data sources to gauge the relative effects in
+this transition period -- see joint distribution of infections with age, geography. But
+contacts with age, geography, occupations.
 
-We apply this model to Germany.
-In March and April 2020, the country broke the first wave of the pandemic fairly quickly.
-Between mid-May and mid-September, daily new infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the period mid-September 2020 to the end of May 2020.
-We pick the starting date for two reasons.
-First, we do not include the first wave because the environment was very different (e.g., aggregate PCR test capacity was much lower and we would require a very different model for calculating the share of known cases) and some data was not recorded yet.\comment[id=HM]{True?}.
-Second, a large fraction cases during summer of 2020 were traced to international travel,\comment[id=HM]{citation?}, but the precise number is difficult to model.
+We apply this model to Germany. In March and April 2020, the country broke the first wave
+of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
+were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
+period mid-September 2020 to the end of May 2020. We pick the starting date for two
+reasons. First, we do not include the first wave because the environment was very
+different (e.g., aggregate PCR test capacity was much lower and we would require a very
+different model for calculating the share of known cases) and some data was not recorded
+yet.\comment[id=HM]{True?}. Second, a large fraction cases during summer of 2020 were
+traced to international travel,\comment[id=HM]{citation?}, but the precise number is
+difficult to model.
 
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and of its drivers.
-The black line in Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the \replaced[id=HM]{color}{blue?} line in Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
-We transform the index so that lower values represent higher levels of restrictions.
-A value of zero means all measures incorporated in the index are turned on.
-The value 1 represents the situation of mid-September, where half of all restrictions incorporated in the index are active.\comment[id=HM]{Need to find version of index with a couple of measures, probably need the API access...}.
-In the six weeks between mid September and mid October, cases increased by a factor of \replaced[id=HM]{X}{10?
-Do we have RKI numbers as table somewhere?}.
-Restrictions were somewhat tightened in mid-October and again in early November.
-New infections remained constant throughout November, before rising again in December.
+Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
+of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
+recorded cases; the \replaced[id=HM]{color}{blue?} line in
+Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020},
+which tracks the tightness of non-pharmaceutical interventions. We transform the index so
+that lower values represent higher levels of restrictions. A value of zero means all
+measures incorporated in the index are turned on. The value 1 represents the situation of
+mid-September, where half of all restrictions incorporated in the index are
+active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
+need the API access...}. In the six weeks between mid September and mid October, cases
+increased by a factor of \replaced[id=HM]{X}{10? Do we have RKI numbers as table
+somewhere?}. Restrictions were somewhat tightened in mid-October and again in early
+November. New infections remained constant throughout November, before rising again in
+December.
 
@@
 Quickly rising rates in December prompted the most stringent lockdown to this date.
-Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores.
-Shortly afterwards, more factors become relevant.
-Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}.
-By the end of May, just over 40\% had received at least one dose of a vaccine.
-Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies.
-At-home tests approved by authorities were ...
-in March.\comment[id=HM]{update precise timeline}
-From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
-The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}.
-In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
-In some cases, customers were only allowed to enter with a recent negative rapid test result.
+Schools and daycare centers were closed again, so were customer-facing businesses except
+for grocery and drug stores. Shortly afterwards, more factors become relevant. Just after
+Christmas, the first people were vaccinated with a focus on older age groups and medical
+personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just over 40\%
+had received at least one dose of a vaccine. Starting in January, rapid tests \dots These
+had to be administered by medical doctors or in pharmacies. At-home tests approved by
+authorities were ... in March.\comment[id=HM]{update precise timeline} From the peak of
+the second wave just before Christmas until the trough in mid-February, newly detected
+cases decreased by almost three quarters. The third wave in the spring of 2021 is
+associated with a rise in the share of the B.1.1.7 variant among all infections, depicted
+in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we
+have the table somewhere?}. In early March, some NPIs were being relaxed; e.g.,
+hairdressers and home improvement stores were allowed to open again to the public. In
+some cases, customers were only allowed to enter with a recent negative rapid test
+result.
 
-These developments are characteristic of many countries: The inititial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
-At broadly similar points in time, novel strains of the virus have started to pose additional challenges.
-The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well.
-This is also true for infections by age and geographical region, which are shown in the supplementary materials (S.\comment[id=HM]{reference}).
-We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic.
-Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks.
-For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%.
-This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality.
-Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual contributions.
-Two things are particularly interesting.
-First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality.
-Second, the most stringent regulations are associated with the period of strong decreases in new infection beteen late December 2020 and mid-February 2021.
-The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
-The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts.
+These developments are characteristic of many countries: The inititial focus on NPIs to
+slow the spread of the disease has been accompanied by vaccines and a growing acceptance
+and use of rapid tests. At broadly similar points in time, novel strains of the virus
+have started to pose additional challenges. The blue line in
+Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections
+extremely well. This is also true for infections by age and geographical region, which
+are shown in the supplementary materials (S.\comment[id=HM]{reference}). We can
+disentangle various mechanisms due to the distinct temporal variation in the drivers of
+the pandemic. Next to the stringency index, the three lines in
+Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
+regulations, and seasonality evolved since early September for each of the three broad
+contact networks. For example, a value of 0.75 for the work multiplier means that if the
+environment was the same as in September (levels of infection rates, no rapid tests or
+vaccinations, only the wildtype virus present), infections at the workplace would be
+reduced by 25\%. This reduction is the product of the effect of contact reductions,
+increased hygiene regulations, and seasonality. Along with the levels of infections,
+these measures completely determine the spread of SARS-CoV-2 in
+2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual
+contributions. Two things are particularly interesting. First, all lines broadly follow
+the stringency index and they would do so even more if we left out seasonality. Second,
+the most stringent regulations are associated with the period of strong decreases in new
+infection beteen late December 2020 and mid-February 2021. The measures were not enough,
+however, to stop the B.1.1.7 variant from spreading in the subsequent period. The steep
+drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of
+the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality
+impacting a fall in the relative infectiousness of contacts.
 
-Reassured by good fit and a couple of external checks (fall vacation regression?), Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests, vaccinations, and of seasonality during 2021, assuming NPIs to be the same as in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line), and three scenarios turning these factors off one by one.
+Reassured by good fit and a couple of external checks (fall vacation regression?),
+Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
+vaccinations, and of seasonality during 2021, assuming NPIs to be the same as in the
+baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (blue
+line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three
+factors (x line), and three scenarios turning these factors off one by one.
 

### 2021-06-07 paper/calibration_and_estimation.tex
Moved on with description of Fig. 3.

@@
         groups. Higher values are colored in darker red tones. The diagonal represents
-        the share of contacts with individuals from the same age group. }
+        the share of contacts with individuals from the same age group.
+    }
 \end{figure}
 
-Figure~\ref{fig:assortativity} shows that assortativity by age is especially strong for
+Figure~\ref{fig:assortativity}\comment[id=HM]{So no work in here?
+How about school?
+Can we add a column with marginals and another one with work / (school?), too?} shows that assortativity by age is especially strong for
 children and younger adults. For older people, the pattern becomes more dispersed around

### 2021-06-07 paper/report.tex
Moved on with description of Fig. 3.

@@
 
-    \begin{subfigure}[b]{0.475\textwidth}
+    \begin{subfigure}[b]{0.425\textwidth}
         \centering
@@
     \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
+    \begin{subfigure}[b]{0.425\textwidth}
         \centering
@@
     \vskip3ex
-    \begin{subfigure}[b]{0.475\textwidth}
+    \begin{subfigure}[b]{0.425\textwidth}
         \centering
@@
 
-    \floatfoot{
+    \floatfoot{\noindent
         Note: ...
@@
 
-    \floatfoot{
+    \floatfoot{\noindent
         Note: All aggregates; See S.XXX for statistics by age group and by geographical region. Also more disaggregated data.
@@
 Quickly rising rates in December prompted the most stringent lockdown to this date.
-Schools and daycare centers were closed again, so were customer-facing businesses except
-for grocery and drug stores. Shortly afterwards, more factors become relevant. Just after
-Christmas, the first people were vaccinated with a focus on older age groups and medical
-personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just over 40\%
-had received at least one dose of a vaccine. Starting in January, rapid tests \dots These
-had to be administered by medical doctors or in pharmacies. At-home tests approved by
-authorities were ... in March.\comment[id=HM]{update precise timeline} From the peak of
-the second wave just before Christmas until the trough in mid-February, newly detected
-cases decreased by almost three quarters. The third wave in the spring of 2021 is
-associated with a rise in the share of the B.1.1.7 variant among all infections, depicted
-in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we
-have the table somewhere?}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. In
-some cases, customers were only allowed to enter with a recent negative rapid test
-result.
+Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores.
+Shortly afterwards, more factors become relevant.
+Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}.
+By the end of May, just over 40\% had received at least one dose of a vaccine.
+Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies.
+At-home tests approved by authorities were ...
+in March.\comment[id=HM]{update precise timeline} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
+The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}.
+In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
+In some cases, customers were only allowed to enter with a recent negative rapid test result.
 
-These developments are characteristic of many countries: The inititial focus on NPIs to
-slow the spread of the disease has been accompanied by vaccines and a growing acceptance
-and use of rapid tests. At broadly similar points in time, novel strains of the virus
-have started to pose additional challenges. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections
-extremely well. This is also true for infections by age and geographical region, which
-are shown in the supplementary materials (S.\comment[id=HM]{reference}). We can
-disentangle various mechanisms due to the distinct temporal variation in the drivers of
-the pandemic. Next to the stringency index, the three lines in
-Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
-regulations, and seasonality evolved since early September for each of the three broad
-contact networks. For example, a value of 0.75 for the work multiplier means that if the
-environment was the same as in September (levels of infection rates, no rapid tests or
-vaccinations, only the wildtype virus present), infections at the workplace would be
-reduced by 25\%. This reduction is the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. Along with the levels of infections,
-these measures completely determine the spread of SARS-CoV-2 in
-2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual
-contributions. Two things are particularly interesting. First, all lines broadly follow
-the stringency index and they would do so even more if we left out seasonality. Second,
-the most stringent regulations are associated with the period of strong decreases in new
-infection beteen late December 2020 and mid-February 2021. The measures were not enough,
-however, to stop the B.1.1.7 variant from spreading in the subsequent period. The steep
-drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of
-the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality
-impacting a fall in the relative infectiousness of contacts.
+These developments are characteristic of many countries: The inititial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
+At broadly similar points in time, novel strains of the virus have started to pose additional challenges.
+The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well.
+This is also true for infections by age and geographical region, which are shown in the supplementary materials (S.\comment[id=HM]{reference}).
+We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic.
+Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks.
+For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%.
+This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality.
+Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual contributions.
+Two things are particularly interesting.
+First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality.
+Second, the most stringent regulations are associated with the period of strong decreases in new infection beteen late December 2020 and mid-February 2021.
+The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
+The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts outside of work and school\comment[id=HM]{check with corrected graph}.
 
-Reassured by good fit and a couple of external checks (fall vacation regression?),
-Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
-vaccinations, and of seasonality during 2021, assuming NPIs to be the same as in the
-baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (blue
-line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three
-factors (x line), and three scenarios turning these factors off one by one.
+Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests, vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way as in the baseline scenario.
+Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line), and three scenarios turning these factors off one by one.
+Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the model.
+Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose the difference in total infections between the scenario without any of the three factors and our main specification.
 
@@
 
-    \floatfoot{
+    \floatfoot{\noindent
     Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
@@
 \end{figure}
+
+The effect of the vaccination campaign is surprisingly small at first sight.
+The reasons for this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
+These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
+It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
+1 Mar - 31 May}.
+
+\clearpage
+
+\section*{writing barrier}
 
@@
 
-    \floatfoot{
+    \floatfoot{\noindent
         Note: All aggregates; See S.XXX for statistics by age group and by geographical region.

### 2021-06-07 paper/report.tex
Add some first comments.

--- 2021-06-07 paper/report.tex
@@
 Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the
-progression of a possible infection. We differentiate between an inital period of
+progression of a possible infection. We differentiate between an initial period of
 infection without being infectious or showing symptoms, being infectious (presymptomatic
@@
 probabilities of transitioning between these states depend on age; their duration is
-random within intervals calibrated to medical literature.\comment[id=HM]{Cite}.
+random within intervals calibrated to medical literature (for a detailed description see
+Section~\ref{sec:medical_params}).
 Conditional on the type of contact, infectiousness independent of ages \citep{Jones2021}.
@@
 infectiousness and disease progress can be introduced. With a probability calibrated to
-XXX\comment[id=HM]{Klara, please add citatiion}, vaccinated agents become immune and they
-do not transmit the virus. While vaccines are rolled out, priority may depend on age and
-occupation. Agents may demand two types of tests as described in
-Figure~\ref{fig:pcr_antigen_tests}. First, polymerase chain reaction (PCR) tests directly
-reveal whether an individual is infected or not. PCR tests require some time to be
-processed and there are always aggregate capacity constraints. Second, rapid antigen
-tests yield immediate results. Specificity and sensitivity of these tests is set
-according to the XXX data.\comment[id=HM]{subsequent PCR tests, quarantine, work/school}.
-Demand for rapid tests will depend on policy, e.g., prices, testing requirements before
-work, attending school, or visiting restaurants and shops.
+\cite{Hunter2021}, vaccinated agents become immune and they do not
+transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.
+% the sources only report reduced viral shedding and reduced risk of infection.
+% Lauterbach calculated that taking the risk of infection * reduced infectiousness given
+% infection based on theses studies suggest that given exposure the risk that fully
+% vaccinated individuals pose is <10% of an unvaccinated person.
+While vaccines are rolled out, priority may depend on age and occupation. Agents may
+demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}. First,
+polymerase chain reaction (PCR) tests directly reveal whether an individual is infected
+or not. PCR tests require some time to be processed and there are always aggregate
+capacity constraints. Second, rapid antigen tests yield immediate results. Specificity
+and sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021}
+. The sensitivity of the rapid antigen tests depends on the timing of the test relative
+to the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
+work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
+requirements before work, attending school, or visiting restaurants and shops.
 

### 2021-06-07 paper/report.tex
Update figures in report to use the new versions.

@@
 
-Modelling a population of agents according to actual demographic characterists means that
-we can use a wide array of data to identify and estimate the model's many
+Modelling a population of agents according to actual demographic characteristics means
+that we can use a wide array of data to identify and estimate the model's many
 parameters.\footnote{See section S.X of the supplementary materials for an
@@
 different model for calculating the share of known cases) and some data was not recorded
-yet.\comment[id=HM]{True?}. Second, a large fraction cases during summer of 2020 were
-traced to international travel,\comment[id=HM]{citation?}, but the precise number is
-difficult to model.
+yet.\comment[id=HM]{True?}\comment[id=Klara]{We would definitely need to model national
+and international travel which you mention below. Regarding data I think everything would
+be available. The low numbers during summer in our sample size might also be a problem
+because Covid might go extinct in our model when case numbers are so low.}. Second, a
+large fraction cases during summer of 2020 were traced to international
+travel,\comment[id=HM]{citation?}\comment[id=Klara]{There was a comment by minister Spahn
+that half of cases during the summer months were due to travel to the Baltic
+(\url{https://twitter.com/krstorevic/status/1396516025181556737}) but I know of no data.
+We should probably check the RKI reports of July, Aug and Sep}, but the precise number is difficult to model.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency_with_seasonality}
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_b117_acc_to_rki}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/assumed_overall_share_known_cases}
 

### 2021-06-07 paper/report.tex
Add Science reference by Lessler.

@@
 entirely, operate at reduced capacity including an alternating schedule, or implement
-mitigation measures like masking requirements or air filters.\comment[id=HM]{cite
-\url{https://science.sciencemag.org/content/early/2021/04/28/science.abh2939}} Curfews
+mitigation measures like masking requirements or air filters \citep{Lessler2021}. Curfews
 may reduce the number of contacts in non-work/non-school settings. In any setting,
@@
 different model for calculating the share of known cases) and some data was not recorded
-yet.\comment[id=HM]{True?}\comment[id=Klara]{We would definitely need to model national
+yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model national
 and international travel which you mention below. Regarding data I think everything would
@@
 recorded cases; the \replaced[id=HM]{color}{blue?} line in
-Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index\cite{Hale2020},
+Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
 which tracks the tightness of non-pharmaceutical interventions. We transform the index so

### 2021-06-07 paper/report.tex
Fix graph, error, add citation.

@@
 large fraction cases during summer of 2020 were traced to international
-travel,\comment[id=HM]{citation?}\comment[id=Klara]{There was a comment by minister Spahn
-that half of cases during the summer months were due to travel to the Baltic
-(\url{https://twitter.com/krstorevic/status/1396516025181556737}) but I know of no data.
-We should probably check the RKI reports of July, Aug and Sep}, but the precise number is difficult to model.
+travel,\comment[id=HM]{cite \url{https://www.rki.de/DE/Content/Infekt/EpidBull/Archiv/2021/Ausgaben/08_21.pdf}}, but the precise number is difficult to model.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/assumed_overall_share_known_cases}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_b117_acc_to_rki}
 

### 2021-06-07 paper/report.tex
Switch to plots showing the effect of single channels on the pessimistic scenario.

@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
         \caption{{\small Recorded cases: 2021 scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
         \caption{{\small Total cases: 2021 scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
         \caption{{\small Effects of different schooling scenarios after Easter}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
         \caption{{\small TBD}}

### 2021-06-07 paper/report.tex
Use scaled version of stringency index.

@@
     }
+\end{figure}
 
-\end{figure}
+EQUATIONS FOR THE FOURTH PICTURE (find names!):
+
+The probabilities for Figure~\subref{fig:model_for_official_cases} are the following:
+
+\begin{align*}
+    x &= \frac{nnewlyinfected \times shareknowncases \times P(PCR | S)}{nsymptomaticanduntested} \\
+    y &= \frac{nnewlyinfected \times shareknowncases \times P(PCR | S)}{nasymptoptomaticandinfectedanduntested} \\
+    z &= \frac{P(PCR | RT)}{survey}
+\end{align*}
+
+where untested means that individuals are not waiting on the result of a previously administered rapid test.
 
@@
 large fraction cases during summer of 2020 were traced to international
-travel,\comment[id=HM]{cite \url{https://www.rki.de/DE/Content/Infekt/EpidBull/Archiv/2021/Ausgaben/08_21.pdf}}, but the precise number is difficult to model.
+travel \citep{KochInstitut2021}, but the precise number is difficult to model.
 
@@
 need the API access...}. In the six weeks between mid September and mid October, cases
-increased by a factor of \replaced[id=HM]{X}{10? Do we have RKI numbers as table
-somewhere?}. Restrictions were somewhat tightened in mid-October and again in early
-November. New infections remained constant throughout November, before rising again in
-December.
+increased by a factor of 3.\comment[id=K]{In the next iteration all empirical data is
+in the figures/results/tables/empirical\_analogues.csv}
+% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
+Restrictions were somewhat tightened in mid-October and again in early November. New
+infections remained constant throughout November, before rising again in December.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency_with_seasonality}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2_with_seasonality}
 
@@
         \caption{Fraction of B.1.1.7 strain among measured infections}
+
+        \textcolor{red}{We'll have this compared to our baseline simulation in the next
+        iteration.}
         \label{fig:share_b117}
@@
 At-home tests approved by authorities were ...
-in March.\comment[id=HM]{update precise timeline} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
-The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}.
+in March.\comment[id=HM]{update precise timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}).
+First tests to self-administer became available starting March 6.
+However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).
+The first tests were licensed on Feb 24: \url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
+The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the empirical analogues table}.
 In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
@@
 
-These developments are characteristic of many countries: The inititial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
+These developments are characteristic of many countries: The initial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
 At broadly similar points in time, novel strains of the virus have started to pose additional challenges.
 The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well.
-This is also true for infections by age and geographical region, which are shown in the supplementary materials (S.\comment[id=HM]{reference}).
+This is also true for infections by age and geographical region, which are shown in the supplementary materials (figures~\ref{fig:age_group_fit, fig:state_fit}.
 We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic.
@@
 This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality.
-Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual contributions.
+Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} \comment[id=K]{We have not done the decomposition for 2020 yet.
+Is on my todo list} for their individual contributions.
 Two things are particularly interesting.
 First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality.
-Second, the most stringent regulations are associated with the period of strong decreases in new infection beteen late December 2020 and mid-February 2021.
+Second, the most stringent regulations are associated with the period of strong decreases in new infection between late December 2020 and mid-February 2021.
 The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
@@
 
-The effect of the vaccination campaign is surprisingly small at first sight.
-The reasons for this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
-These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
-It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
-1 Mar - 31 May}.
+The effect of the vaccination campaign is surprisingly small at first sight. The reasons
+for this are the slow start---10\% of the population had received their first vaccination
+by the end of March, % 2021-03-24
+20\% by April 19th ---and the focus on older individuals. These groups contribute less to
+the spread of the disease than others due to a lower number of contacts,
+see~\ref{fig:assortativity}. It is important to note that the initial focus of the
+campaign was to prevent deaths and severe disease; the case fatality was rate
+considerably lower during the third wave when compared to the second (4.4\% between
+October and February and 1.4\% between March and June).
 

### 2021-06-07 paper/report.tex
A couple of more paragraphs.

@@
 
-The effect of the vaccination campaign is surprisingly small at first sight. The reasons
-for this are the slow start---10\% of the population had received their first vaccination
-by the end of March, % 2021-03-24
-20\% by April 19th ---and the focus on older individuals. These groups contribute less to
-the spread of the disease than others due to a lower number of contacts,
-see~\ref{fig:assortativity}. It is important to note that the initial focus of the
-campaign was to prevent deaths and severe disease; the case fatality was rate
-considerably lower during the third wave when compared to the second (4.4\% between
-October and February and 1.4\% between March and June).
+Until mid-March, there is no visible difference between the different scenarios. Seasonality hardly changes, vaccinations  
+The effect of the vaccination campaign is surprisingly small at first sight.
+Whether considering recorded or total infections, the final level is always the highest in case the vaccination campaign had been running in isolation (red lines).
+The Shapley value decomposition shows that vaccinations contribute about 15\%\comment[id=HM]{check!} to the difference between scenarios.
+Reasons for the slow this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
+These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
+It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
+1 Mar - 31 May}.
+It is important to note that by the end of our study period, when first-dose vaccination rates reached around 40\% of the population, the numbers of new cases would have started to decline.
+
+Seasonality has a large effect in slowing the spread of SARS-CoV-2.
+By May 31, both observed and recorded cases would be reduced by a factor of four if only seasonality mattered.
+However, in this period, cases would have kept on rising throughout, just at a much lower pace.
+Nevertheless, we estimate it to be a quantitatively important factor determining the evolution of the pandemic, explaining most of the early changes and almost 40\% of the cumulative difference by the end of May.
+
+The largest effect---almost one half when considering the decompositions---comes from rapid testing.
+Here, it is crucial to differentiate between recorded cases and actual cases.
+Additional testing means that infections become known which would otherwise remain undetected.
+Figure~\ref{fig:2021_scenarios_recorded} shows that this means that until late April, recorded cases are higher than in the scenario where none of the three mechanisms is turned on.
+Compared to the scenario with vaccinations only, this point is reached only around mid May and it would be June for the comparison with the seasonality-only scenario.
+The effect on total cases, however, kicks in immediately and strongly.
+Despite the fact that only a small fraction of the population performed weekly rapid tests in March (X\%\comment[id=HM]{Put in}), the rise in new infections would be limited by XXX\% relative to the scenario without vaccinations, tests, or seasonality.
+
+Tests
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
+        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
         \caption{{\small Effects of different schooling scenarios after Easter}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
+        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
         \caption{{\small TBD}}
@@
 
-    \caption{Schooling / maybe home office}
+    \caption{Effects of different scenarios for schooling and work from home}
     \label{fig:interventions_school}

### 2021-06-07 paper/report.tex
Address some of HM's comments.

--- 2021-06-07 paper/report.tex
@@
 need the API access...}. In the six weeks between mid September and mid October, cases
-increased by a factor of \replaced[id=HM]{X}{10? Do we have RKI numbers as table
-somewhere?}. Restrictions were somewhat tightened in mid-October and again in early
-November. New infections remained constant throughout November, before rising again in
-December.
+increased by a factor of 3.\comment[id=Klara]{In the next iteration all empirical data is
+in the figures/results/tables/empirical_analogues.csv}
+% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
+Restrictions were somewhat tightened in mid-October and again in early November. New
+infections remained constant throughout November, before rising again in December.
 
@@
 At-home tests approved by authorities were ...
-in March.\comment[id=HM]{update precise timeline} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
-The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}.
+in March.\comment[id=HM]{update precise timeline}\comment[id=Klara]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}).
+First tests to self-administer became available starting March 6.
+However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
+The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=Klara]{Will be part of the empirical analogues table}.
 In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.

### 2021-06-07 paper/report.tex
Add case fatality ratio.

--- 2021-06-07 paper/report.tex
@@
 These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
-It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
-1 Mar - 31 May}.
+It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second (4.4\% between October and February and 1.4\% between March and June).
 

### 2021-06-07 paper/report.tex
Resolve some comments.

@@
 need the API access...}. In the six weeks between mid September and mid October, cases
-increased by a factor of 3.\comment[id=Klara]{In the next iteration all empirical data is
-in the figures/results/tables/empirical_analogues.csv}
+increased by a factor of 3.\comment[id=K]{In the next iteration all empirical data is
+in the figures/results/tables/empirical\_analogues.csv}
 % the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
@@
 At-home tests approved by authorities were ...
-in March.\comment[id=HM]{update precise timeline}\comment[id=Klara]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}).
+in March.\comment[id=HM]{update precise timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}).
 First tests to self-administer became available starting March 6.
-However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
-The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=Klara]{Will be part of the empirical analogues table}.
+However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).
+The first tests were licensed on Feb 24: \url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
+The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the empirical analogues table}.
 In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
@@
 
-The effect of the vaccination campaign is surprisingly small at first sight.
-The reasons for this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
-These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
-It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second (4.4\% between October and February and 1.4\% between March and June).
+The effect of the vaccination campaign is surprisingly small at first sight. The reasons
+for this are the slow start---10\% of the population had received their first vaccination
+by the end of March, % 2021-03-24
+20\% by April 19th ---and the focus on older individuals. These groups contribute less to
+the spread of the disease than others due to a lower number of contacts,
+see~\ref{fig:assortativity}. It is important to note that the initial focus of the
+campaign was to prevent deaths and severe disease; the case fatality was rate
+considerably lower during the third wave when compared to the second (4.4\% between
+October and February and 1.4\% between March and June).
 

### 2021-06-08 paper/report.tex
Outline for rest.

@@
 
-The effect of the vaccination campaign is surprisingly small at first sight. The reasons
-for this are the slow start---10\% of the population had received their first vaccination
-by the end of March, % 2021-03-24
-20\% by April 19th ---and the focus on older individuals. These groups contribute less to
-the spread of the disease than others due to a lower number of contacts,
-see~\ref{fig:assortativity}. It is important to note that the initial focus of the
-campaign was to prevent deaths and severe disease; the case fatality was rate
-considerably lower during the third wave when compared to the second (4.4\% between
-October and February and 1.4\% between March and June).
+Until mid-March, there is no visible difference between the different scenarios.
+Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
+Even thereafter, the effect of the vaccination campaign is surprisingly small at first sight.
+Whether considering recorded or total infections, the final level is always the highest in case the vaccination campaign had been running in isolation (red lines).
+The Shapley value decomposition shows that vaccinations contribute about 15\%\comment[id=HM]{check!} to the cumulative difference between scenarios.
+Reasons for the slow this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
+These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
+It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
+1 Mar - 31 May}.
+It is important to note that by the end of our study period, when first-dose vaccination rates reached around 40\% of the population, the numbers of new cases would have started to decline.
+
+Seasonality has a large effect in slowing the spread of SARS-CoV-2.
+By May 31, both observed and recorded cases would be reduced by a factor of four if only seasonality mattered.
+However, in this period, cases would have kept on rising throughout, just at a much lower pace.
+Nevertheless, we estimate it to be a quantitatively important factor determining the evolution of the pandemic, explaining most of the early changes and almost 40\% of the cumulative difference by the end of May.
+
+The largest effect---almost one half when considering the decompositions---comes from rapid testing.
+Here, it is crucial to differentiate between recorded cases and actual cases.
+Additional testing means that infections become known which would otherwise remain undetected.
+Figure~\ref{fig:2021_scenarios_recorded} shows that this means that until late April, recorded cases are higher than in the scenario where none of the three mechanisms is turned on.
+Compared to the scenario with vaccinations only, this point is reached only around mid May and it would be June for the comparison with the seasonality-only scenario.
+The effect on total cases, however, kicks in immediately and strongly.
+Despite the fact that only a small fraction of the population performed weekly rapid tests in March (X\%\comment[id=HM]{Put in}), the rise in new infections would be limited by XXX\% relative to the scenario without vaccinations, tests, or seasonality.
+
+Tests
 
@@
 
-\section*{writing barrier}
+Two particularly important areas:
+\begin{itemize}
+    \item Schooling: large costs to pupils, ... social inequalities
+    \item Work: Q2 / 2020 saw the largest drop in GDP in a long time; 
+\end{itemize}
 
-Schooling very contentious. Likely damage to child development.
-
-We show
-
-Screening effect: ...
+Results
+\begin{itemize}
+    \item Advantage of testing in both cases: Recurrent. Small cost relative to other stuff. Certain publicness.
+    \item Mandatory tests: Screening effect: ...
+\end{itemize}
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
+        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
         \caption{{\small Effects of different schooling scenarios after Easter}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
-        \caption{{\small TBD}}
+        % \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/}
+        \caption{{\small Effects of different work scenarios after Easter}}
         \label{fig:to_be_determined}
@@
 
-    \caption{Schooling / maybe home office}
+    \caption{Effects of different scenarios for schooling and working from home}
     \label{fig:interventions_school}

### 2021-06-08 paper/report.tex
Reformat.

@@
-% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance.
-Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references.
-Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions.
+% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3
+% printed pages in the journal) present important new research results of broad
+% significance. Reports should include an abstract, an introductory paragraph, up to
+% four figures or tables, and about 30 references. Materials and Methods should be
+% included in supplementary materials, which should also include information needed to
+% support the paper's conclusions.
 
@@
 
-% The manuscript should start with a brief introduction describing the paper’s significance.
-The introduction should provide sufficient background information to make the article intelligible to readers in other disciplines, and sufficient context that the significance of the experimental findings is clear.
+% The manuscript should start with a brief introduction describing the paper’s
+% significance. The introduction should provide sufficient background information to
+% make the article intelligible to readers in other disciplines, and sufficient context
+% that the significance of the experimental findings is clear.
 
@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite
-some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate
-policies.}, most have seen large ups and downs in their infection rates. Containment
-measures have become increasingly diverse and included testing, more nuanced NPIs, and
-contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
-more infectious virus strains are well understood in quantitative terms. This paper
-develops a model incorporating all these factors. The framework allows to combine a wide
-variety of data in a timely fashion, making it useful to predict the effects of various
-interventions. We apply the model to Germany and show that rapid testing had the largest
-impact on the reduction in infections by XXX\% during the A weeks between X April and Y
-May.\comment[id=HM]{insert some useful numbers}. We conclude that rapid tests should play
-a large role...
+containment in the medium term. However, most countries find themselves many
+months---and often years---away from reaching vaccination-induced herd
+immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the
+meantime, it is of utmost importance to employ an effective mix of strategies for
+containing the virus. The most frequent initial response was a set of non-pharmaceutical
+interventions (NPIs) to reduce contacts between individuals. While this has allowed some
+countries to sustain equilibria with very low infection numbers\footnote{See
+\citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
+sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in
+their infection rates. Containment measures have become increasingly diverse and
+included testing, more nuanced NPIs, and contact tracing. Neither these policies' effect
+nor the influence of seasonal patterns or more infectious virus strains are well
+understood in quantitative terms. This paper develops a model incorporating all these
+factors. The framework allows to combine a wide variety of data in a timely fashion,
+making it useful to predict the effects of various interventions. We apply the model to
+Germany and show that rapid testing had the largest impact on the reduction in
+infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert
+some useful numbers}. We conclude that rapid tests should play a large role...
 
@@
 \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data
-sets on contact types and frequencies at \url{http://www.socialcontactdata.org/} covering
-countries from all continents except North America and Australia.} The dimensions are
-chosen so that the most common NPIs can be modeled in great detail by reducing the number
-of contacts in a particular setting or the risk of transmitting the disease for a type of
-contact. For example, a mandate to work from home will reduce the number of work contacts
-to zero for a fraction of the working population. Schools and daycare can be closed
-entirely, operate at reduced capacity including an alternating schedule, or implement
-mitigation measures like masking requirements or air filters \citep{Lessler2021}. Curfews
-may reduce the number of contacts in non-work/non-school settings. In any setting,
-measures like masking requirements would reduce the probability of infection associated
-with a contact.
+sets on contact types and frequencies at \url{http://www.socialcontactdata.org/}
+covering countries from all continents except North America and Australia.} The
+dimensions are chosen so that the most common NPIs can be modeled in great detail by
+reducing the number of contacts in a particular setting or the risk of transmitting the
+disease for a type of contact. For example, a mandate to work from home will reduce the
+number of work contacts to zero for a fraction of the working population. Schools and
+daycare can be closed entirely, operate at reduced capacity including an alternating
+schedule, or implement mitigation measures like masking requirements or air filters
+\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
+settings. In any setting, measures like masking requirements would reduce the
+probability of infection associated with a contact.
 
@@
         % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
-        \caption{{\small Model for translating simulated number to officially recorded cases}}
+        \caption{{\small Model for translating simulated number to officially recorded
+        cases}}
         \label{fig:model_for_official_cases}
@@
 
-    \floatfoot{\noindent
-        Note: ...
-    }
+    \floatfoot{\noindent Note: ...}
 \end{figure}
@@
 
-where untested means that individuals are not waiting on the result of a previously administered rapid test.
+where untested means that individuals are not waiting on the result of a previously
+administered rapid test.
 
@@
 random within intervals calibrated to medical literature (for a detailed description see
-Section~\ref{sec:medical_params}).
-Conditional on the type of contact, infectiousness independent of ages \citep{Jones2021}.
+Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness
+independent of ages \citep{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different profiles regarding
+The model includes several other features, which are crucial to describe the evolution
+of the pandemic in 2020-2021. New virus strains with different profiles regarding
 infectiousness and disease progress can be introduced. With a probability calibrated to
-\cite{Hunter2021}, vaccinated agents become immune and they do not
-transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.
+\cite{Hunter2021}, vaccinated agents become immune and they do not transmit the virus
+\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.
 % the sources only report reduced viral shedding and reduced risk of infection.
@@
 capacity constraints. Second, rapid antigen tests yield immediate results. Specificity
-and sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021}
-. The sensitivity of the rapid antigen tests depends on the timing of the test relative
-to the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
-work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
-requirements before work, attending school, or visiting restaurants and shops.
+and sensitivity of these tests is set according to data by \cite{Bruemmer2021,
+Smith2021} . The sensitivity of the rapid antigen tests depends on the timing of the
+test relative to the start of infectiousness.\comment[id=HM]{subsequent PCR tests,
+quarantine, work/school}. Demand for rapid tests will depend on policy, e.g., prices,
+testing requirements before work, attending school, or visiting restaurants and shops.
 
@@
 directly from official directives. Infection rates by age and geographical region are
-estimated to match to officially recorded numbers; so is the prevalence of virus strains.
-In order to translate simulated cases in our model to officially recorded cases, we use
-the model depicted in Figure~\ref{fig:model_for_official_cases}. \comment[id=HM]{Add a
-sentence.} A further advantage is that the simulated data have a structure that resembles
-datasets used for regression models, which allows additional plausibility checks by
-re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we
-actually do something like that.}
+estimated to match to officially recorded numbers; so is the prevalence of virus
+strains. In order to translate simulated cases in our model to officially recorded
+cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
+\comment[id=HM]{Add a sentence.} A further advantage is that the simulated data have a
+structure that resembles datasets used for regression models, which allows additional
+plausibility checks by re-running the same models on the model-generated
+data.\comment[id=HM]{Only keep if we actually do something like that.}
 
-Model that makes the most of many available data sources to gauge the relative effects in
-this transition period -- see joint distribution of infections with age, geography. But
-contacts with age, geography, occupations.
+Model that makes the most of many available data sources to gauge the relative effects
+in this transition period -- see joint distribution of infections with age, geography.
+But contacts with age, geography, occupations.
 
-We apply this model to Germany. In March and April 2020, the country broke the first wave
-of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
-were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
-period mid-September 2020 to the end of May 2020. We pick the starting date for two
-reasons. First, we do not include the first wave because the environment was very
-different (e.g., aggregate PCR test capacity was much lower and we would require a very
-different model for calculating the share of known cases) and some data was not recorded
-yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model national
-and international travel which you mention below. Regarding data I think everything would
-be available. The low numbers during summer in our sample size might also be a problem
-because Covid might go extinct in our model when case numbers are so low.}. Second, a
-large fraction cases during summer of 2020 were traced to international
+We apply this model to Germany. In March and April 2020, the country broke the first
+wave of the pandemic fairly quickly. Between mid-May and mid-September, daily new
+infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We
+model the period mid-September 2020 to the end of May 2020. We pick the starting date
+for two reasons. First, we do not include the first wave because the environment was
+very different (e.g., aggregate PCR test capacity was much lower and we would require a
+very different model for calculating the share of known cases) and some data was not
+recorded yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model
+national and international travel which you mention below. Regarding data I think
+everything would be available. The low numbers during summer in our sample size might
+also be a problem because Covid might go extinct in our model when case numbers are so
+low.}. Second, a large fraction cases during summer of 2020 were traced to international
 travel \citep{KochInstitut2021}, but the precise number is difficult to model.
@@
 Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
-which tracks the tightness of non-pharmaceutical interventions. We transform the index so
-that lower values represent higher levels of restrictions. A value of zero means all
-measures incorporated in the index are turned on. The value 1 represents the situation of
-mid-September, where half of all restrictions incorporated in the index are
+which tracks the tightness of non-pharmaceutical interventions. We transform the index
+so that lower values represent higher levels of restrictions. A value of zero means all
+measures incorporated in the index are turned on. The value 1 represents the situation
+of mid-September, where half of all restrictions incorporated in the index are
 active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
-need the API access...}. In the six weeks between mid September and mid October, cases
-increased by a factor of 3.\comment[id=K]{In the next iteration all empirical data is
+need the API access...}. In the six weeks between mid September and late October, cases
+increased by a factor of XX.\comment[id=K]{In the next iteration all empirical data is
 in the figures/results/tables/empirical\_analogues.csv}
@@
 
-    \floatfoot{\noindent
-        Note: All aggregates; See S.XXX for statistics by age group and by geographical region. Also more disaggregated data.
+    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
+        by geographical region. Also more disaggregated data.
 
-        Sources: ...
-    }
+        Sources: ...}
 
@@
 Quickly rising rates in December prompted the most stringent lockdown to this date.
-Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores.
-Shortly afterwards, more factors become relevant.
-Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}.
-By the end of May, just over 40\% had received at least one dose of a vaccine.
-Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies.
-At-home tests approved by authorities were ...
-in March.\comment[id=HM]{update precise timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}).
-First tests to self-administer became available starting March 6.
-However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).
-The first tests were licensed on Feb 24: \url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
-The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the empirical analogues table}.
-In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
-In some cases, customers were only allowed to enter with a recent negative rapid test result.
+Schools and daycare centers were closed again, so were customer-facing businesses except
+for grocery and drug stores. Shortly afterwards, more factors become relevant. Just
+after Christmas, the first people were vaccinated with a focus on older age groups and
+medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just
+over 40\% had received at least one dose of a vaccine. Starting in January, rapid tests
+\dots These had to be administered by medical doctors or in pharmacies. At-home tests
+approved by authorities were ... in March.\comment[id=HM]{update precise
+timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was
+very low initially (\url{https://bit.ly/3ehmGcj}). First tests to self-administer became
+available starting March 6. However, supply was very limited in the beginning
+(\url{https://bit.ly/3xJCIn8}). The first tests were licensed on Feb 24:
+\url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}}
+From the peak of the second wave just before Christmas until the trough in mid-February,
+newly detected cases decreased by almost three quarters. The third wave in the spring of
+2021 is associated with a rise in the share of the B.1.1.7 variant among all infections,
+depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in
+April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the
+empirical analogues table}. In early March, some NPIs were being relaxed; e.g.,
+hairdressers and home improvement stores were allowed to open again to the public. In
+some cases, customers were only allowed to enter with a recent negative rapid test
+result.
 
-These developments are characteristic of many countries: The initial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
-At broadly similar points in time, novel strains of the virus have started to pose additional challenges.
-The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well.
-This is also true for infections by age and geographical region, which are shown in the supplementary materials (figures~\ref{fig:age_group_fit, fig:state_fit}.
-We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic.
-Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks.
-For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%.
-This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality.
-Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} \comment[id=K]{We have not done the decomposition for 2020 yet.
-Is on my todo list} for their individual contributions.
-Two things are particularly interesting.
-First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality.
-Second, the most stringent regulations are associated with the period of strong decreases in new infection between late December 2020 and mid-February 2021.
-The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
-The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts outside of work and school\comment[id=HM]{check with corrected graph}.
+These developments are characteristic of many countries: The initial focus on NPIs to
+slow the spread of the disease has been accompanied by vaccines and a growing acceptance
+and use of rapid tests. At broadly similar points in time, novel strains of the virus
+have started to pose additional challenges. The blue line in
+Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections
+extremely well. This is also true for infections by age and geographical region, which
+are shown in the supplementary materials (figures~\ref{fig:age_group_fit,
+fig:state_fit}. We can disentangle various mechanisms due to the distinct temporal
+variation in the drivers of the pandemic. Next to the stringency index, the three lines
+in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
+regulations, and seasonality evolved since early September for each of the three broad
+contact networks. For example, a value of 0.75 for the work multiplier means that if the
+environment was the same as in September (levels of infection rates, no rapid tests or
+vaccinations, only the wildtype virus present), infections at the workplace would be
+reduced by 25\%. This reduction is the product of the effect of contact reductions,
+increased hygiene regulations, and seasonality. Along with the levels of infections,
+these measures completely determine the spread of SARS-CoV-2 in
+2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} \comment[id=K]{We
+have not done the decomposition for 2020 yet. Is on my todo list} for their individual
+contributions. Two things are particularly interesting. First, all lines broadly follow
+the stringency index and they would do so even more if we left out seasonality. Second,
+the most stringent regulations are associated with the period of strong decreases in new
+infection between late December 2020 and mid-February 2021. The measures were not
+enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
+The steep drop in recorded cases during May 2021 is associated with rapid tests to
+X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and
+mostly seasonality impacting a fall in the relative infectiousness of contacts outside
+of work and school\comment[id=HM]{check with corrected graph}.
 
-Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests, vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way as in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line), and three scenarios turning these factors off one by one.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the model.
-Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose the difference in total infections between the scenario without any of the three factors and our main specification.
+Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
+vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way
+as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model
+fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in
+Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line),
+and three scenarios turning these factors off one by one.
+Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
+model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose
+the difference in total infections between the scenario without any of the three factors
+and our main specification.
 
@@
 
-        \caption{Decomposition of effects for Figure~\ref{fig:2021_scenarios_newly_infected}.}
+        \caption{Decomposition of effects for
+        Figure~\ref{fig:2021_scenarios_newly_infected}.}
         \label{fig:2021_scenarios_decomposition}
@@
 
-    \floatfoot{\noindent
-    Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
+    by geographical region.
 
-    The decomposition is based on Shapley values where the individual contribution of a channel is its average contribution over different sizes of coalitions (combinations with other channels).
-The individual contribution to a coalition is the difference between the effect size of the coalition with the particular channel and without.
-    }
+    The decomposition is based on Shapley values where the individual contribution of a
+    channel is its average contribution over different sizes of coalitions (combinations
+    with other channels). The individual contribution to a coalition is the difference
+    between the effect size of the coalition with the particular channel and without.}
 \end{figure}
@@
 Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
-Even thereafter, the effect of the vaccination campaign is surprisingly small at first sight.
-Whether considering recorded or total infections, the final level is always the highest in case the vaccination campaign had been running in isolation (red lines).
-The Shapley value decomposition shows that vaccinations contribute about 15\%\comment[id=HM]{check!} to the cumulative difference between scenarios.
-Reasons for the slow this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
-These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
-It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
-1 Mar - 31 May}.
-It is important to note that by the end of our study period, when first-dose vaccination rates reached around 40\% of the population, the numbers of new cases would have started to decline.
+Even thereafter, the effect of the vaccination campaign is surprisingly small at first
+sight. Whether considering recorded or total infections, the final level is always the
+highest in case the vaccination campaign had been running in isolation (red lines). The
+Shapley value decomposition shows that vaccinations contribute about
+15\%\comment[id=HM]{check!} to the cumulative difference between scenarios. Reasons for
+this are the slow start---it took until 24~March until 10\% of the population had
+received their first vaccination, the 20\% mark was reached on April 19th ---and the
+focus on older individuals. These groups contribute less to the spread of the disease
+than others due to a lower number of contacts, see~\ref{fig:assortativity}. It is
+important to note that the initial focus of the campaign was to prevent deaths and
+severe disease; the case fatality was rate considerably lower during the third wave when
+compared to the second (4.4\% between October and February and 1.4\% between March and
+June). It is important to note that by the end of our study period, when first-dose
+vaccination rates reached around 40\% of the population, the numbers of new cases would
+have started to decline.
 
-Seasonality has a large effect in slowing the spread of SARS-CoV-2.
-By May 31, both observed and recorded cases would be reduced by a factor of four if only seasonality mattered.
-However, in this period, cases would have kept on rising throughout, just at a much lower pace.
-Nevertheless, we estimate it to be a quantitatively important factor determining the evolution of the pandemic, explaining most of the early changes and almost 40\% of the cumulative difference by the end of May.
+Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
+observed and recorded cases would be reduced by a factor of four if only seasonality
+mattered. However, in this period, cases would have kept on rising throughout, just at a
+much lower pace. Nevertheless, we estimate it to be a quantitatively important factor
+determining the evolution of the pandemic, explaining most of the early changes and
+almost 40\% of the cumulative difference by the end of May.
 
-The largest effect---almost one half when considering the decompositions---comes from rapid testing.
-Here, it is crucial to differentiate between recorded cases and actual cases.
-Additional testing means that infections become known which would otherwise remain undetected.
-Figure~\ref{fig:2021_scenarios_recorded} shows that this means that until late April, recorded cases are higher than in the scenario where none of the three mechanisms is turned on.
-Compared to the scenario with vaccinations only, this point is reached only around mid May and it would be June for the comparison with the seasonality-only scenario.
-The effect on total cases, however, kicks in immediately and strongly.
-Despite the fact that only a small fraction of the population performed weekly rapid tests in March (X\%\comment[id=HM]{Put in}), the rise in new infections would be limited by XXX\% relative to the scenario without vaccinations, tests, or seasonality.
+The largest effect---almost one half when considering the decompositions---comes from
+rapid testing. Here, it is crucial to differentiate between recorded cases and actual
+cases. Additional testing means that infections become known which would otherwise
+remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
+until late April, recorded cases are higher than in the scenario where none of the three
+mechanisms is turned on. Compared to the scenario with vaccinations only, this point is
+reached only around mid May and it would be June for the comparison with the
+seasonality-only scenario. The effect on total cases, however, kicks in immediately and
+strongly. Despite the fact that only a small fraction of the population performed weekly
+rapid tests in March (X\%\comment[id=HM]{Put in}), the rise in new infections would be
+limited by XXX\% relative to the scenario without vaccinations, tests, or seasonality.
 
@@
 \begin{itemize}
-    \item Advantage of testing in both cases: Recurrent. Small cost relative to other stuff. Certain publicness.
+    \item Advantage of testing in both cases: Recurrent. Small cost relative to other
+    stuff. Certain publicness.
     \item Mandatory tests: Screening effect: ...
@@
         \centering
-        % \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/}
+        % \includegraphics[width=0.9
+        % \textwidth]{../figures/results/figures/scenario_comparisons/}
         \caption{{\small Effects of different work scenarios after Easter}}
@@
 
-    \floatfoot{\noindent
-        Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
-    }
+    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
+        by geographical region.}
 
@@
 
-% All data must be shown either in the main text or in the Supplementary Materials or must be available in an established database with accession details provided in the acknowledgements section.
+% All data must be shown either in the main text or in the Supplementary Materials or
+% must be available in an established database with accession details provided in the
+% acknowledgements section.
 
-% References to unpublished materials are not allowed to substantiate significant conclusions of the paper.
+% References to unpublished materials are not allowed to substantiate significant
+% conclusions of the paper.
 

### 2021-06-08 paper/report.tex
New results.
Improve equations.

@@
 \begin{align*}
-    x &= \frac{nnewlyinfected \times shareknowncases \times P(PCR | S)}{nsymptomaticanduntested} \\
-    y &= \frac{nnewlyinfected \times shareknowncases \times P(PCR | S)}{nasymptoptomaticandinfectedanduntested} \\
+    x &= \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_symptomatic\_and\_untested} \\
+    y &= \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
     z &= \frac{P(PCR | RT)}{survey}
@@
 directly from official directives. Infection rates by age and geographical region are
-estimated to match to officially recorded numbers; so is the prevalence of virus
-strains. In order to translate simulated cases in our model to officially recorded
-cases, we use the model depicted in Figure~\ref{fig:model_for_official_cases}.
-\comment[id=HM]{Add a sentence.} A further advantage is that the simulated data have a
-structure that resembles datasets used for regression models, which allows additional
-plausibility checks by re-running the same models on the model-generated
-data.\comment[id=HM]{Only keep if we actually do something like that.}
+estimated to match to officially recorded numbers; so is the prevalence of virus strains.
+In order to translate simulated cases in our model to officially recorded cases, we use
+the model depicted in Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a
+sentence.} A further advantage is that the simulated data have a structure that resembles
+datasets used for regression models, which allows additional plausibility checks by
+re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we
+actually do something like that.}
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_b117_acc_to_rki}
+
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
 
         \caption{Fraction of B.1.1.7 strain among measured infections}
-
-        \textcolor{red}{We'll have this compared to our baseline simulation in the next
-        iteration.}
         \label{fig:share_b117}

### 2021-06-08 paper/report.tex
Add footnote on our vaccination effectiveness decision.

@@
 
-The model includes several other features, which are crucial to describe the evolution
-of the pandemic in 2020-2021. New virus strains with different profiles regarding
-infectiousness and disease progress can be introduced. With a probability calibrated to
-\cite{Hunter2021}, vaccinated agents become immune and they do not transmit the virus
-\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.
-% the sources only report reduced viral shedding and reduced risk of infection.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different profiles regarding
+infectiousness \replaced[id=K]{and disease progress}{}\comment[id=K]{sid only supports
+differing infectiousness between variants at the moment} can be introduced. With a
+probability of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not
+transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\%
+is lower than what is usually reported for after the second dose. We choose it because in
+our model does not include booster shots and does not allow for vaccinated individuals to
+become infected and transmit the disease which is the case for Covid-19
+\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
 % Lauterbach calculated that taking the risk of infection * reduced infectiousness given
@@
 % vaccinated individuals pose is <10% of an unvaccinated person.
-While vaccines are rolled out, priority may depend on age and occupation. Agents may
+While vaccines are rolled out, priority depends on age and occupation. Agents may
 demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}. First,

### 2021-06-08 paper/report.tex
Move rapid test plot shares to the appendix. Replace plots as agreed.

@@
 the pandemic in 2020-2021. New virus strains with different profiles regarding
-infectiousness \replaced[id=K]{and disease progress}{}\comment[id=K]{sid only supports
+infectiousness \replaced[id=K]{}{and disease progress}\comment[id=K]{sid only supports
 differing infectiousness between variants at the moment} can be introduced. With a
@@
 Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
-which tracks the tightness of non-pharmaceutical interventions. We transform the index
-so that lower values represent higher levels of restrictions. A value of zero means all
-measures incorporated in the index are turned on. The value 1 represents the situation
-of mid-September, where half of all restrictions incorporated in the index are
+which tracks the tightness of non-pharmaceutical interventions. We transform the index so
+that lower values represent higher levels of restrictions. A value of zero means all
+measures incorporated in the index are turned on. The value 1 represents the situation of
+mid-September, where half of all restrictions incorporated in the index are
 active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
 need the API access...}. In the six weeks between mid September and late October, cases
-increased by a factor of XX.\comment[id=K]{In the next iteration all empirical data is
-in the figures/results/tables/empirical\_analogues.csv}
-% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
+increased by a factor of 4.% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
 Restrictions were somewhat tightened in mid-October and again in early November. New
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
 
@@
 depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in
-April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the
-empirical analogues table}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. In
-some cases, customers were only allowed to enter with a recent negative rapid test
-result.
+April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{is in the empirical
+analogues table}. In early March, some NPIs were being relaxed; e.g., hairdressers and
+home improvement stores were allowed to open again to the public. In some cases,
+customers were only allowed to enter with a recent negative rapid test result.
 

### 2021-06-08 paper/report.tex
Updated first few paragraphs.

@@
 Germany and show that rapid testing had the largest impact on the reduction in
-infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert
-some useful numbers}. We conclude that rapid tests should play a large role...
+infections by almost 80\% during the month of May 2021. We conclude that rapid tests
+have a large role to play at least as long as vaccinations have not been offered to an
+entire population.
 
@@
 occur in the household, at work, at school, or in other settings (leisure activities,
-grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of
-work?} Some contacts recur regularly, others occur at random. Random contacts are
-typically assortative in age and geographical location. Empirical applications can take
-the population structure from census data and the types and frequency of contacts from
-diary data measuring contacts before the pandemic
-\citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data
-sets on contact types and frequencies at \url{http://www.socialcontactdata.org/}
-covering countries from all continents except North America and Australia.} The
-dimensions are chosen so that the most common NPIs can be modeled in great detail by
-reducing the number of contacts in a particular setting or the risk of transmitting the
-disease for a type of contact. For example, a mandate to work from home will reduce the
-number of work contacts to zero for a fraction of the working population. Schools and
-daycare can be closed entirely, operate at reduced capacity including an alternating
-schedule, or implement mitigation measures like masking requirements or air filters
-\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
-settings. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact.
+grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
+occur at random. Random contacts are typically assortative in age and geographical
+location. Empirical applications can take the population structure from census data and
+the types and frequency of contacts from diary data measuring contacts before the
+pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to
+multiple data sets on contact types and frequencies at
+\url{http://www.socialcontactdata.org/} covering countries from all continents except
+North America and Australia.} The dimensions are chosen so that the most common NPIs can
+be modeled in great detail by reducing the number of contacts in a particular setting or
+the risk of transmitting the disease for a type of contact. For example, a mandate to
+work from home will reduce the number of work contacts to zero for a fraction of the
+working population. Schools and daycare can be closed entirely, operate at reduced
+capacity---including an alternating schedule---,or implement mitigation measures like
+masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
+of contacts in non-work/non-school settings. In any setting, measures like masking
+requirements would reduce the probability of infection associated with a contact.
 
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
+        \vskip4ex
+
         \caption{Disease progression}
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
-        \caption{{\small PCR and antigen tests}}
-        \label{fig:pcr_antigen_tests}
+        \caption{{\small Model for antigen tests}}
+        \label{fig:antigen_tests}
     \end{subfigure}
@@
         \centering
+        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
+        Model for undetected cases
+        \vskip4ex
 
-        Model for detected/undetected cases
-
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
-        \caption{{\small Model for translating simulated number to officially recorded
-        cases}}
+        \caption{{\small Model for undetected cases}}
         \label{fig:model_for_official_cases}
@@
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different profiles regarding
+The model includes several other features, which are crucial to describe the evolution
+of the pandemic in 2020-2021. New virus strains with different profiles regarding
 infectiousness \replaced[id=K]{}{and disease progress}\comment[id=K]{sid only supports
@@
 probability of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not
-transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\%
-is lower than what is usually reported for after the second dose. We choose it because in
-our model does not include booster shots and does not allow for vaccinated individuals to
-become infected and transmit the disease which is the case for Covid-19
-\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
+transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
+Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the
+second dose. We choose it because in our model does not include booster shots and does
+not allow for vaccinated individuals to become infected and transmit the disease which
+is the case for Covid-19 \citep{Petter2021, LevineTiefenbrun2021,
+Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
 % Lauterbach calculated that taking the risk of infection * reduced infectiousness given
@@
 % vaccinated individuals pose is <10% of an unvaccinated person.
-While vaccines are rolled out, priority depends on age and occupation. Agents may
-demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}. First,
-polymerase chain reaction (PCR) tests directly reveal whether an individual is infected
-or not. PCR tests require some time to be processed and there are always aggregate
-capacity constraints. Second, rapid antigen tests yield immediate results. Specificity
-and sensitivity of these tests is set according to data by \cite{Bruemmer2021,
-Smith2021} . The sensitivity of the rapid antigen tests depends on the timing of the
-test relative to the start of infectiousness.\comment[id=HM]{subsequent PCR tests,
-quarantine, work/school}. Demand for rapid tests will depend on policy, e.g., prices,
-testing requirements before work, attending school, or visiting restaurants and shops.
+While vaccines are rolled out, priority depends on age and occupation. Agents may demand
+two types of tests as described in Figure~\ref{fig:antigen_tests}. First, polymerase
+chain reaction (PCR) tests directly reveal whether an individual is infected or not. PCR
+tests require some time to be processed and there are always aggregate capacity
+constraints. Second, rapid antigen tests yield immediate results. Specificity and
+sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021} .
+The sensitivity of the rapid antigen tests depends on the timing of the test relative to
+the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
+work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
+requirements before work, attending school, or visiting restaurants and shops.
 
@@
 directly from official directives. Infection rates by age and geographical region are
-estimated to match to officially recorded numbers; so is the prevalence of virus strains.
-In order to translate simulated cases in our model to officially recorded cases, we use
-the model depicted in Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a
-sentence.} A further advantage is that the simulated data have a structure that resembles
-datasets used for regression models, which allows additional plausibility checks by
-re-running the same models on the model-generated data.\comment[id=HM]{Only keep if we
-actually do something like that.}
+estimated to match to officially recorded numbers; so is the prevalence of virus
+strains. In order to translate simulated cases in our model to officially recorded
+cases, we use the model depicted in
+Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.} A further
+advantage is that the simulated data have a structure that resembles datasets used for
+regression models, which allows additional plausibility checks by re-running the same
+models on the model-generated data.\comment[id=HM]{Only keep if we actually do something
+like that.}
 
@@
 Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
-which tracks the tightness of non-pharmaceutical interventions. We transform the index so
-that lower values represent higher levels of restrictions. A value of zero means all
-measures incorporated in the index are turned on. The value 1 represents the situation of
-mid-September, where half of all restrictions incorporated in the index are
+which tracks the tightness of non-pharmaceutical interventions. We transform the index
+so that lower values represent higher levels of restrictions. A value of zero means all
+measures incorporated in the index are turned on. The value 1 represents the situation
+of mid-September, where half of all restrictions incorporated in the index are
 active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
 need the API access...}. In the six weeks between mid September and late October, cases
-increased by a factor of 4.% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
+increased by a factor of
+4.% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
 Restrictions were somewhat tightened in mid-October and again in early November. New

### 2021-06-08 paper/report.tex
Add rough draft for model graph bottom right.

--- 2021-06-08 paper/report.tex
@@
-% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3
-% printed pages in the journal) present important new research results of broad
-% significance. Reports should include an abstract, an introductory paragraph, up to
-% four figures or tables, and about 30 references. Materials and Methods should be
-% included in supplementary materials, which should also include information needed to
-% support the paper's conclusions.
+% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance.
+Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references.
+Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions.
 
@@
 
-% The manuscript should start with a brief introduction describing the paper’s
-% significance. The introduction should provide sufficient background information to
-% make the article intelligible to readers in other disciplines, and sufficient context
-% that the significance of the experimental findings is clear.
+% The manuscript should start with a brief introduction describing the paper’s significance.
+The introduction should provide sufficient background information to make the article intelligible to readers in other disciplines, and sufficient context that the significance of the experimental findings is clear.
 
@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many
-months---and often years---away from reaching vaccination-induced herd
-immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the
-meantime, it is of utmost importance to employ an effective mix of strategies for
-containing the virus. The most frequent initial response was a set of non-pharmaceutical
-interventions (NPIs) to reduce contacts between individuals. While this has allowed some
-countries to sustain equilibria with very low infection numbers\footnote{See
-\citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
-sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in
-their infection rates. Containment measures have become increasingly diverse and
-included testing, more nuanced NPIs, and contact tracing. Neither these policies' effect
-nor the influence of seasonal patterns or more infectious virus strains are well
-understood in quantitative terms. This paper develops a model incorporating all these
-factors. The framework allows to combine a wide variety of data in a timely fashion,
-making it useful to predict the effects of various interventions. We apply the model to
-Germany and show that rapid testing had the largest impact on the reduction in
-infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert
-some useful numbers}. We conclude that rapid tests should play a large role...
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination-induced herd immunity.\comment[id=HM]{Cite
+some paper on herd immunity, maybe vaccine data} In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate
+policies.}, most have seen large ups and downs in their infection rates. Containment
+measures have become increasingly diverse and included testing, more nuanced NPIs, and
+contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
+more infectious virus strains are well understood in quantitative terms. This paper
+develops a model incorporating all these factors. The framework allows to combine a wide
+variety of data in a timely fashion, making it useful to predict the effects of various
+interventions. We apply the model to Germany and show that rapid testing had the largest
+impact on the reduction in infections by XXX\% during the A weeks between X April and Y
+May.\comment[id=HM]{insert some useful numbers}. We conclude that rapid tests should play
+a large role...
 
@@
 \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data
-sets on contact types and frequencies at \url{http://www.socialcontactdata.org/}
-covering countries from all continents except North America and Australia.} The
-dimensions are chosen so that the most common NPIs can be modeled in great detail by
-reducing the number of contacts in a particular setting or the risk of transmitting the
-disease for a type of contact. For example, a mandate to work from home will reduce the
-number of work contacts to zero for a fraction of the working population. Schools and
-daycare can be closed entirely, operate at reduced capacity including an alternating
-schedule, or implement mitigation measures like masking requirements or air filters
-\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
-settings. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact.
+sets on contact types and frequencies at \url{http://www.socialcontactdata.org/} covering
+countries from all continents except North America and Australia.} The dimensions are
+chosen so that the most common NPIs can be modeled in great detail by reducing the number
+of contacts in a particular setting or the risk of transmitting the disease for a type of
+contact. For example, a mandate to work from home will reduce the number of work contacts
+to zero for a fraction of the working population. Schools and daycare can be closed
+entirely, operate at reduced capacity including an alternating schedule, or implement
+mitigation measures like masking requirements or air filters \citep{Lessler2021}. Curfews
+may reduce the number of contacts in non-work/non-school settings. In any setting,
+measures like masking requirements would reduce the probability of infection associated
+with a contact.
 
@@
 
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
-        \caption{{\small Model for translating simulated number to officially recorded
-        cases}}
+        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
+        \caption{{\small Model for translating simulated number to officially recorded cases}}
         \label{fig:model_for_official_cases}
@@
 
-    \floatfoot{\noindent Note: ...}
+    \floatfoot{\noindent
+        Note: ...
+    }
 \end{figure}
@@
 \begin{align*}
-    x &= \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_symptomatic\_and\_untested} \\
-    y &= \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
+    x &= \frac{nnewlyinfected \times shareknowncases \times P(PCR | S)}{nsymptomaticanduntested} \\
+    y &= \frac{nnewlyinfected \times shareknowncases \times P(PCR | S)}{nasymptoptomaticandinfectedanduntested} \\
     z &= \frac{P(PCR | RT)}{survey}
@@
 
-where untested means that individuals are not waiting on the result of a previously
-administered rapid test.
+where untested means that individuals are not waiting on the result of a previously administered rapid test.
 
@@
 random within intervals calibrated to medical literature (for a detailed description see
-Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness
-independent of ages \citep{Jones2021}.
+Section~\ref{sec:medical_params}).
+Conditional on the type of contact, infectiousness independent of ages \citep{Jones2021}.
 
@@
 the pandemic in 2020-2021. New virus strains with different profiles regarding
-infectiousness \replaced[id=K]{and disease progress}{}\comment[id=K]{sid only supports
-differing infectiousness between variants at the moment} can be introduced. With a
-probability of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not
-transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\%
-is lower than what is usually reported for after the second dose. We choose it because in
-our model does not include booster shots and does not allow for vaccinated individuals to
-become infected and transmit the disease which is the case for Covid-19
-\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
+infectiousness and disease progress can be introduced. With a probability calibrated to
+\cite{Hunter2021}, vaccinated agents become immune and they do not
+transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.
+% the sources only report reduced viral shedding and reduced risk of infection.
 % Lauterbach calculated that taking the risk of infection * reduced infectiousness given
@@
 % vaccinated individuals pose is <10% of an unvaccinated person.
-While vaccines are rolled out, priority depends on age and occupation. Agents may
+While vaccines are rolled out, priority may depend on age and occupation. Agents may
 demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}. First,
@@
 capacity constraints. Second, rapid antigen tests yield immediate results. Specificity
-and sensitivity of these tests is set according to data by \cite{Bruemmer2021,
-Smith2021} . The sensitivity of the rapid antigen tests depends on the timing of the
-test relative to the start of infectiousness.\comment[id=HM]{subsequent PCR tests,
-quarantine, work/school}. Demand for rapid tests will depend on policy, e.g., prices,
-testing requirements before work, attending school, or visiting restaurants and shops.
+and sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021}
+. The sensitivity of the rapid antigen tests depends on the timing of the test relative
+to the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
+work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
+requirements before work, attending school, or visiting restaurants and shops.
 
@@
 In order to translate simulated cases in our model to officially recorded cases, we use
-the model depicted in Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a
+the model depicted in Figure~\ref{fig:model_for_official_cases}. \comment[id=HM]{Add a
 sentence.} A further advantage is that the simulated data have a structure that resembles
@@
 
-Model that makes the most of many available data sources to gauge the relative effects
-in this transition period -- see joint distribution of infections with age, geography.
-But contacts with age, geography, occupations.
+Model that makes the most of many available data sources to gauge the relative effects in
+this transition period -- see joint distribution of infections with age, geography. But
+contacts with age, geography, occupations.
 
-We apply this model to Germany. In March and April 2020, the country broke the first
-wave of the pandemic fairly quickly. Between mid-May and mid-September, daily new
-infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We
-model the period mid-September 2020 to the end of May 2020. We pick the starting date
-for two reasons. First, we do not include the first wave because the environment was
-very different (e.g., aggregate PCR test capacity was much lower and we would require a
-very different model for calculating the share of known cases) and some data was not
-recorded yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model
-national and international travel which you mention below. Regarding data I think
-everything would be available. The low numbers during summer in our sample size might
-also be a problem because Covid might go extinct in our model when case numbers are so
-low.}. Second, a large fraction cases during summer of 2020 were traced to international
-travel \citep{KochInstitut2021}, but the precise number is difficult to model.
+We apply this model to Germany. In March and April 2020, the country broke the first wave
+of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
+were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
+period mid-September 2020 to the end of May 2020. We pick the starting date for two
+reasons. First, we do not include the first wave because the environment was very
+different (e.g., aggregate PCR test capacity was much lower and we would require a very
+different model for calculating the share of known cases) and some data was not recorded
+yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model national
+and international travel which you mention below. Regarding data I think everything would
+be available. The low numbers during summer in our sample size might also be a problem
+because Covid might go extinct in our model when case numbers are so low.}. Second, a
+large fraction cases during summer of 2020 were traced to international
+travel,\comment[id=HM]{cite \url{https://www.rki.de/DE/Content/Infekt/EpidBull/Archiv/2021/Ausgaben/08_21.pdf}}, but the precise number is difficult to model.
 
@@
 Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
-which tracks the tightness of non-pharmaceutical interventions. We transform the index
-so that lower values represent higher levels of restrictions. A value of zero means all
-measures incorporated in the index are turned on. The value 1 represents the situation
-of mid-September, where half of all restrictions incorporated in the index are
+which tracks the tightness of non-pharmaceutical interventions. We transform the index so
+that lower values represent higher levels of restrictions. A value of zero means all
+measures incorporated in the index are turned on. The value 1 represents the situation of
+mid-September, where half of all restrictions incorporated in the index are
 active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
-need the API access...}. In the six weeks between mid September and late October, cases
-increased by a factor of XX.\comment[id=K]{In the next iteration all empirical data is
-in the figures/results/tables/empirical\_analogues.csv}
+need the API access...}. In the six weeks between mid September and mid October, cases
+increased by a factor of 3.\comment[id=Klara]{In the next iteration all empirical data is
+in the figures/results/tables/empirical_analogues.csv}
 % the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2_with_seasonality}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency_with_seasonality}
 
@@
         \centering
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/share_of_b117_acc_to_rki}
 
         \caption{Fraction of B.1.1.7 strain among measured infections}
+
+        \textcolor{red}{We'll have this compared to our baseline simulation in the next
+        iteration.}
         \label{fig:share_b117}
@@
 
-    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
-        by geographical region. Also more disaggregated data.
+    \floatfoot{\noindent
+        Note: All aggregates; See S.XXX for statistics by age group and by geographical region. Also more disaggregated data.
 
-        Sources: ...}
+        Sources: ...
+    }
 
@@
 Quickly rising rates in December prompted the most stringent lockdown to this date.
-Schools and daycare centers were closed again, so were customer-facing businesses except
-for grocery and drug stores. Shortly afterwards, more factors become relevant. Just
-after Christmas, the first people were vaccinated with a focus on older age groups and
-medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just
-over 40\% had received at least one dose of a vaccine. Starting in January, rapid tests
-\dots These had to be administered by medical doctors or in pharmacies. At-home tests
-approved by authorities were ... in March.\comment[id=HM]{update precise
-timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was
-very low initially (\url{https://bit.ly/3ehmGcj}). First tests to self-administer became
-available starting March 6. However, supply was very limited in the beginning
-(\url{https://bit.ly/3xJCIn8}). The first tests were licensed on Feb 24:
-\url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}}
-From the peak of the second wave just before Christmas until the trough in mid-February,
-newly detected cases decreased by almost three quarters. The third wave in the spring of
-2021 is associated with a rise in the share of the B.1.1.7 variant among all infections,
-depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in
-April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the
-empirical analogues table}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. In
-some cases, customers were only allowed to enter with a recent negative rapid test
-result.
+Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores.
+Shortly afterwards, more factors become relevant.
+Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}.
+By the end of May, just over 40\% had received at least one dose of a vaccine.
+Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies.
+At-home tests approved by authorities were ...
+in March.\comment[id=HM]{update precise timeline}\comment[id=Klara]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}).
+First tests to self-administer became available starting March 6.
+However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters.
+The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=Klara]{Will be part of the empirical analogues table}.
+In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public.
+In some cases, customers were only allowed to enter with a recent negative rapid test result.
 
-These developments are characteristic of many countries: The initial focus on NPIs to
-slow the spread of the disease has been accompanied by vaccines and a growing acceptance
-and use of rapid tests. At broadly similar points in time, novel strains of the virus
-have started to pose additional challenges. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections
-extremely well. This is also true for infections by age and geographical region, which
-are shown in the supplementary materials (figures~\ref{fig:age_group_fit,
-fig:state_fit}. We can disentangle various mechanisms due to the distinct temporal
-variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
-regulations, and seasonality evolved since early September for each of the three broad
-contact networks. For example, a value of 0.75 for the work multiplier means that if the
-environment was the same as in September (levels of infection rates, no rapid tests or
-vaccinations, only the wildtype virus present), infections at the workplace would be
-reduced by 25\%. This reduction is the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. Along with the levels of infections,
-these measures completely determine the spread of SARS-CoV-2 in
-2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} \comment[id=K]{We
-have not done the decomposition for 2020 yet. Is on my todo list} for their individual
-contributions. Two things are particularly interesting. First, all lines broadly follow
-the stringency index and they would do so even more if we left out seasonality. Second,
-the most stringent regulations are associated with the period of strong decreases in new
-infection between late December 2020 and mid-February 2021. The measures were not
-enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
-The steep drop in recorded cases during May 2021 is associated with rapid tests to
-X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and
-mostly seasonality impacting a fall in the relative infectiousness of contacts outside
-of work and school\comment[id=HM]{check with corrected graph}.
+These developments are characteristic of many countries: The initial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests.
+At broadly similar points in time, novel strains of the virus have started to pose additional challenges.
+The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well.
+This is also true for infections by age and geographical region, which are shown in the supplementary materials (figures~\ref{fig:age_group_fit, fig:state_fit}.
+We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic.
+Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks.
+For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%.
+This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality.
+Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual contributions.
+Two things are particularly interesting.
+First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality.
+Second, the most stringent regulations are associated with the period of strong decreases in new infection beteen late December 2020 and mid-February 2021.
+The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
+The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts outside of work and school\comment[id=HM]{check with corrected graph}.
 
-Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
-vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way
-as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model
-fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line),
-and three scenarios turning these factors off one by one.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose
-the difference in total infections between the scenario without any of the three factors
-and our main specification.
+Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests, vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way as in the baseline scenario.
+Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line), and three scenarios turning these factors off one by one.
+Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the model.
+Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose the difference in total infections between the scenario without any of the three factors and our main specification.
 
@@
 
-        \caption{Decomposition of effects for
-        Figure~\ref{fig:2021_scenarios_newly_infected}.}
+        \caption{Decomposition of effects for Figure~\ref{fig:2021_scenarios_newly_infected}.}
         \label{fig:2021_scenarios_decomposition}
@@
 
-    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
-    by geographical region.
+    \floatfoot{\noindent
+    Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
 
-    The decomposition is based on Shapley values where the individual contribution of a
-    channel is its average contribution over different sizes of coalitions (combinations
-    with other channels). The individual contribution to a coalition is the difference
-    between the effect size of the coalition with the particular channel and without.}
+    The decomposition is based on Shapley values where the individual contribution of a channel is its average contribution over different sizes of coalitions (combinations with other channels).
+The individual contribution to a coalition is the difference between the effect size of the coalition with the particular channel and without.
+    }
 \end{figure}
 
-Until mid-March, there is no visible difference between the different scenarios.
-Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
-Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections, the final level is always the
-highest in case the vaccination campaign had been running in isolation (red lines). The
-Shapley value decomposition shows that vaccinations contribute about
-15\%\comment[id=HM]{check!} to the cumulative difference between scenarios. Reasons for
-this are the slow start---it took until 24~March until 10\% of the population had
-received their first vaccination, the 20\% mark was reached on April 19th ---and the
-focus on older individuals. These groups contribute less to the spread of the disease
-than others due to a lower number of contacts, see~\ref{fig:assortativity}. It is
-important to note that the initial focus of the campaign was to prevent deaths and
-severe disease; the case fatality was rate considerably lower during the third wave when
-compared to the second (4.4\% between October and February and 1.4\% between March and
-June). It is important to note that by the end of our study period, when first-dose
-vaccination rates reached around 40\% of the population, the numbers of new cases would
-have started to decline.
-
-Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
-observed and recorded cases would be reduced by a factor of four if only seasonality
-mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace. Nevertheless, we estimate it to be a quantitatively important factor
-determining the evolution of the pandemic, explaining most of the early changes and
-almost 40\% of the cumulative difference by the end of May.
-
-The largest effect---almost one half when considering the decompositions---comes from
-rapid testing. Here, it is crucial to differentiate between recorded cases and actual
-cases. Additional testing means that infections become known which would otherwise
-remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
-until late April, recorded cases are higher than in the scenario where none of the three
-mechanisms is turned on. Compared to the scenario with vaccinations only, this point is
-reached only around mid May and it would be June for the comparison with the
-seasonality-only scenario. The effect on total cases, however, kicks in immediately and
-strongly. Despite the fact that only a small fraction of the population performed weekly
-rapid tests in March (X\%\comment[id=HM]{Put in}), the rise in new infections would be
-limited by XXX\% relative to the scenario without vaccinations, tests, or seasonality.
-
-Tests
+The effect of the vaccination campaign is surprisingly small at first sight.
+The reasons for this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals.
+These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}.
+It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second.\comment[id=HM]{Calculate N deaths / N cases, for RKI data 1 October - 31 Jan vs.
+1 Mar - 31 May}.
 
@@
 
-Two particularly important areas:
-\begin{itemize}
-    \item Schooling: large costs to pupils, ... social inequalities
-    \item Work: Q2 / 2020 saw the largest drop in GDP in a long time; 
-\end{itemize}
+\section*{writing barrier}
 
-Results
-\begin{itemize}
-    \item Advantage of testing in both cases: Recurrent. Small cost relative to other
-    stuff. Certain publicness.
-    \item Mandatory tests: Screening effect: ...
-\end{itemize}
+Schooling very contentious. Likely damage to child development.
+
+We show
+
+Screening effect: ...
 
@@
         \centering
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
         \caption{{\small Effects of different schooling scenarios after Easter}}
@@
         \centering
-        % \includegraphics[width=0.9
-        % \textwidth]{../figures/results/figures/scenario_comparisons/}
-        \caption{{\small Effects of different work scenarios after Easter}}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
+        \caption{{\small TBD}}
         \label{fig:to_be_determined}
@@
 
-    \caption{Effects of different scenarios for schooling and working from home}
+    \caption{Schooling / maybe home office}
     \label{fig:interventions_school}
 
-    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
-        by geographical region.}
+    \floatfoot{\noindent
+        Note: All aggregates; See S.XXX for statistics by age group and by geographical region.
+    }
 
@@
 
-% All data must be shown either in the main text or in the Supplementary Materials or
-% must be available in an established database with accession details provided in the
-% acknowledgements section.
+% All data must be shown either in the main text or in the Supplementary Materials or must be available in an established database with accession details provided in the acknowledgements section.
 
-% References to unpublished materials are not allowed to substantiate significant
-% conclusions of the paper.
+% References to unpublished materials are not allowed to substantiate significant conclusions of the paper.
 

### 2021-06-08 figures/results/tables/student_infections_table.tex
List of bullets for report for Science, first draft.
Add current result figures and tables.
Add new results figures.
New results.

@@
 \midrule
- educ open after easter  without tests &                                             843318 \\
- educ open after easter  with tests    &                                             607352 \\
- baseline                              &                                             479015 \\
- close educ after easter               &                                             475397 \\
+ educ open after easter  without tests &                                             752932 \\
+ educ open after easter  with tests    &                                             583570 \\
+ close educ after easter               &                                             443978 \\
+ baseline                              &                                             459436 \\
 \bottomrule

### 2021-06-08 paper/model_validation.tex
Address some of HM's comments.

@@
 
-% Despite fitting only four infection probabilities, the in-sample fit is very good.
-% The best fit is achieved in the largest age groups.
-% This is so mechanically, because we weigh the
-% deviations between simulated and observed infection rates by group sizes. The worst fit
-% is achieved for the 80 to 100 years old.
-% There are three reasons for this: Firstly, we only model private households at
-% the moment, meaning that nursing homes are not part of the synthetic population we
-% simulate. Since community housing residents are part of the German Mikrozensus
-% \citep{FDSAeDBUDL2018} which forms the basis for our synthetic population, we can and
-% plan to include community housing residents in the near future.
-% Secondly, the elderly, especially nursing home residents, are tested more often than
-% the general population leading to more cases being detected in this age group.
-% We are currently working to include age-variant testing strategies,
-% using \href{https://ars.rki.de/Content/COVID19/Main.aspx}{data by the RKI}
-% \citep{Seifried2020}.
+% Despite fitting only four infection probabilities, the in-sample fit is very good. The
+% best fit is achieved in the largest age groups. This is so mechanically, because we
+% weigh the deviations between simulated and observed infection rates by group sizes.
+% The worst fit is achieved for the 80 to 100 years old. There are three reasons for
+% this: Firstly, we only model private households at the moment, meaning that nursing
+% homes are not part of the synthetic population we simulate. Since community housing
+% residents are part of the German Mikrozensus \citep{FDSAeDBUDL2018} which forms the
+% basis for our synthetic population, we can and plan to include community housing
+% residents in the near future. Secondly, the elderly, especially nursing home
+% residents, are tested more often than the general population leading to more cases
+% being detected in this age group. We are currently working to include age-variant
+% testing strategies, using \href{https://ars.rki.de/Content/COVID19/Main.aspx}{data by
+% the RKI} \citep{Seifried2020}.
 
@@
 We can assess the out-of-sample fit by projecting the effect of the lockdown light and
-comparing it to case numbers until mid of November. It is important to note that this
-is not just a simple extrapolation of a time trend because the lockdown light only started
+comparing it to case numbers until mid of November. It is important to note that this is
+not just a simple extrapolation of a time trend because the lockdown light only started
 after the estimation period. The out-of-sample fit can be assessed in
@@
     \label{fig:out-of-sample-fit}
-    \figurenotes{
-        For the time period until the beginning of November, the figure shows the weekly
-        incidence rates of infections per 100,000 people from reported (black) versus
-        simulated (blue) data. With the start of November, the projections of the three
-        scenarios, optimistic (blue), neutral (red), and pessimistic (mint green), are
-        shown until the beginning of the Christmas holidays. The actual incidence rates
-        (black) are reported until the 24th November.
-    }
+    \figurenotes{For the time period until the beginning of November, the figure shows
+        the weekly incidence rates of infections per 100,000 people from reported
+        (black) versus simulated (blue) data. With the start of November, the
+        projections of the three scenarios, optimistic (blue), neutral (red), and
+        pessimistic (mint green), are shown until the beginning of the Christmas
+        holidays. The actual incidence rates (black) are reported until the 24th
+        November.}
 \end{figure}

### 2021-06-08 paper/report.tex
Re-wrote most parts.

--- 2021-06-08 paper/report.tex
@@
 \citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
-sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in
-their infection rates. Containment measures have become increasingly diverse and
-included testing, more nuanced NPIs, and contact tracing. Neither these policies' effect
-nor the influence of seasonal patterns or more infectious virus strains are well
-understood in quantitative terms. This paper develops a model incorporating all these
-factors. The framework allows to combine a wide variety of data in a timely fashion,
-making it useful to predict the effects of various interventions. We apply the model to
-Germany and show that rapid testing had the largest impact on the reduction in
-infections by almost 80\% during the month of May 2021. We conclude that rapid tests
-have a large role to play at least as long as vaccinations have not been offered to an
-entire population.
+sustained with test-trace-and-isolate policies.}, most have seen large fluctuations of
+infection rates. Containment measures have become increasingly diverse and now include
+testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
+influence of seasonal patterns or more infectious virus strains are well understood in
+quantitative terms. This paper develops a model incorporating all these factors. The
+framework allows to combine a wide variety of data in a timely fashion, making it useful
+to predict the effects of various interventions. We apply the model to Germany and show
+that rapid testing had the largest impact on the reduction in infections by almost 80\%
+during the month of May 2021. We conclude that rapid tests have a large role to play at
+least as long as vaccinations have not been offered to an entire population.
 
@@
 
-\begin{figure}[!tp]
+\begin{figure}
     \centering
@@
         \centering
-        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
-        Model for undetected cases
-        \vskip4ex
-
+        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
         \caption{{\small Model for undetected cases}}
@@
 
-EQUATIONS FOR THE FOURTH PICTURE (find names!):
-
-The probabilities for Figure~\subref{fig:model_for_official_cases} are the following:
-
-\begin{align*}
-    x &= \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_symptomatic\_and\_untested} \\
-    y &= \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
-    z &= \frac{P(PCR | RT)}{survey}
-\end{align*}
-
-where untested means that individuals are not waiting on the result of a previously
-administered rapid test.
-
-Susceptibility to contracting the SARS-CoV-2 virus is dependent on age; so is the
-progression of a possible infection. We differentiate between an initial period of
-infection without being infectious or showing symptoms, being infectious (presymptomatic
-or asymptomatic), showing symptoms, requiring intensive care, and recovery or death. The
-probabilities of transitioning between these states depend on age; their duration is
-random within intervals calibrated to medical literature (for a detailed description see
-Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness
-independent of ages \citep{Jones2021}.
+In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
+possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
+differentiate between an initial period of infection without being infectious or showing
+symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
+intensive care, and recovery or death. The probabilities of transitioning between these
+states depend on age; their duration is random within intervals calibrated to medical
+literature (for a detailed description see Section~\ref{sec:medical_params}).
+Conditional on the type of contact, infectiousness independent of ages
+\citep{Jones2021}.\comment[id=HM]{Seems broadly standard, can we cite someone?}
 
@@
 of the pandemic in 2020-2021. New virus strains with different profiles regarding
-infectiousness \replaced[id=K]{}{and disease progress}\comment[id=K]{sid only supports
-differing infectiousness between variants at the moment} can be introduced. With a
-probability of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not
-transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
-Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the
-second dose. We choose it because in our model does not include booster shots and does
-not allow for vaccinated individuals to become infected and transmit the disease which
-is the case for Covid-19 \citep{Petter2021, LevineTiefenbrun2021,
-Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
-% Lauterbach calculated that taking the risk of infection * reduced infectiousness given
-% infection based on theses studies suggest that given exposure the risk that fully
-% vaccinated individuals pose is <10% of an unvaccinated person.
-While vaccines are rolled out, priority depends on age and occupation. Agents may demand
-two types of tests as described in Figure~\ref{fig:antigen_tests}. First, polymerase
-chain reaction (PCR) tests directly reveal whether an individual is infected or not. PCR
-tests require some time to be processed and there are always aggregate capacity
-constraints. Second, rapid antigen tests yield immediate results. Specificity and
-sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021} .
-The sensitivity of the rapid antigen tests depends on the timing of the test relative to
-the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
-work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
-requirements before work, attending school, or visiting restaurants and shops.
+infectiousness can be introduced. Agents may receive a vaccination. With a probability
+of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not transmit the
+virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is lower
+than what is usually reported for after the second dose of the Biontech/Pfizer vaccine,
+which is most commonly used in Germany. We choose it because our model neither includes
+booster shots, nor does it allow vaccinated individuals who became immune to transmit
+the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.} During the vaccine
+roll-out, priority may depend on age and occupation.
+
+We include two types of tests. Polymerase chain reaction (PCR) tests directly reveal
+whether an individual is infected or not. PCR tests require some days to be processed
+and there are always aggregate capacity constraints. In contrast, rapid antigen tests
+yield immediate results. Specificity and sensitivity of these tests is set according to
+data analysed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of
+the test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows
+our model for rapid test demand. Schools may require students to be tested regularly.
+Rapid tests may be offered by employers for on-site workers. Individuals may demand
+tests for private reasons, which include having plans to meet other people\footnote{A
+positive test will make them reduce their contacts; this is why tests impact the actual
+contacts in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and
+because a household member tested positively for the virus. We endow each agent with an
+individual compliance parameter. This parameter determines whether she takes up rapid
+tests offered by employers or follows up on private reasons. Thresholds differ across
+reasons triggering rapid test demand; they are lowest for infected household members and
+highest for meeting other people.\comment[id=HM]{True?}.
 
@@
 estimated to match to officially recorded numbers; so is the prevalence of virus
-strains. In order to translate simulated cases in our model to officially recorded
-cases, we use the model depicted in
-Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.} A further
-advantage is that the simulated data have a structure that resembles datasets used for
-regression models, which allows additional plausibility checks by re-running the same
-models on the model-generated data.\comment[id=HM]{Only keep if we actually do something
-like that.}
+strains. The model yields total infections, but only a fraction of those will be
+officially recorded. We model the fraction of known cases as depicted in
+Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
+\footnote{EQUATIONS FOR THE FOURTH PICTURE (find names!):
 
-Model that makes the most of many available data sources to gauge the relative effects
-in this transition period -- see joint distribution of infections with age, geography.
-But contacts with age, geography, occupations.
+    The probabilities for Figure~\subref{fig:model_for_official_cases} are the
+    following:
+
+    \begin{align*}
+        x & = \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_symptomatic\_and\_untested}                    \\
+        y & = \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
+        z & = \frac{P(PCR | RT)}{survey}
+    \end{align*}
+    where untested means that individuals are not waiting on the result of a previously
+    administered rapid test.} A further advantage is that the simulated data have a
+    structure that resembles datasets used for regression models, which allows
+    additional plausibility checks by re-running the same models on the model-generated
+    data.\comment[id=HM]{Only keep if we actually do something like that.}
 
@@
 infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We
-model the period mid-September 2020 to the end of May 2020. We pick the starting date
+model the period mid-September 2020 to the end of May 2021. We pick the starting date
 for two reasons. First, we do not include the first wave because the environment was
 very different (e.g., aggregate PCR test capacity was much lower and we would require a
-very different model for calculating the share of known cases) and some data was not
-recorded yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model
-national and international travel which you mention below. Regarding data I think
-everything would be available. The low numbers during summer in our sample size might
-also be a problem because Covid might go extinct in our model when case numbers are so
-low.}. Second, a large fraction cases during summer of 2020 were traced to international
-travel \citep{KochInstitut2021}, but the precise number is difficult to model.
+very different model for calculating the share of known cases) Second, a large fraction
+cases during summer of 2020 were traced to international travel
+\citep{KochInstitut2021}\comment[id=HM]{Include
+\url{https://www.nature.com/articles/s41586-021-03677-y} in citation}, but the precise
+number is difficult to model.
 
@@
 measures incorporated in the index are turned on. The value 1 represents the situation
-of mid-September, where half of all restrictions incorporated in the index are
-active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
-need the API access...}. In the six weeks between mid September and late October, cases
-increased by a factor of
-4.% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
-Restrictions were somewhat tightened in mid-October and again in early November. New
-infections remained constant throughout November, before rising again in December.
+in mid-September, with restrictions on gatherings and public events, masking
+requirements, but open schools and workplaces (the raw value of the index at that point
+is 49.5). In the seven weeks between mid September and early November, cases increased
+by a factor of 10. Restrictions were somewhat tightened in mid-October and again in
+early November. New infections remained constant throughout November, before rising
+again in December, which prompted the most stringent lockdown to this date. Schools and
+daycare centers were closed again, so were customer-facing businesses except for grocery
+and drug stores. From the peak of the second wave just before Christmas until the trough
+in mid-February, newly detected cases decreased by almost three quarters. The third wave
+in the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in
+March. See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed;
+e.g., hairdressers and home improvement stores were allowed to open again to the public.
 
@@
 
-Quickly rising rates in December prompted the most stringent lockdown to this date.
-Schools and daycare centers were closed again, so were customer-facing businesses except
-for grocery and drug stores. Shortly afterwards, more factors become relevant. Just
-after Christmas, the first people were vaccinated with a focus on older age groups and
+By this time, the set of policy instruments had become much more diverse. Around the
+turn of the year, the first people were vaccinated with a focus on older age groups and
 medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just
-over 40\% had received at least one dose of a vaccine. Starting in January, rapid tests
-\dots These had to be administered by medical doctors or in pharmacies. At-home tests
-approved by authorities were ... in March.\comment[id=HM]{update precise
-timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was
-very low initially (\url{https://bit.ly/3ehmGcj}). First tests to self-administer became
-available starting March 6. However, supply was very limited in the beginning
-(\url{https://bit.ly/3xJCIn8}). The first tests were licensed on Feb 24:
-\url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}}
-From the peak of the second wave just before Christmas until the trough in mid-February,
-newly detected cases decreased by almost three quarters. The third wave in the spring of
-2021 is associated with a rise in the share of the B.1.1.7 variant among all infections,
-depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in
-April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{is in the empirical
-analogues table}. In early March, some NPIs were being relaxed; e.g., hairdressers and
-home improvement stores were allowed to open again to the public. In some cases,
-customers were only allowed to enter with a recent negative rapid test result.
+over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
+started to replace PCR tests in many medical and nursing facilities. These had to be
+administered by medical doctors or in pharmacies. At-home tests approved by authorities
+became available in mid-March, rapid test centers were opened and one test per person
+and week was made available free of charge. Depending on the state, customers were only
+allowed to enter certain stores with a recent negative rapid test result. These
+developments are characteristic of many countries: The initial focus on NPIs to slow the
+spread of the disease has been accompanied by vaccines and a growing acceptance and use
+of rapid tests. At broadly similar points in time, novel strains of the virus have
+started to pose additional challenges.
 
-These developments are characteristic of many countries: The initial focus on NPIs to
-slow the spread of the disease has been accompanied by vaccines and a growing acceptance
-and use of rapid tests. At broadly similar points in time, novel strains of the virus
-have started to pose additional challenges. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections
-extremely well. This is also true for infections by age and geographical region, which
-are shown in the supplementary materials (figures~\ref{fig:age_group_fit,
-fig:state_fit}. We can disentangle various mechanisms due to the distinct temporal
-variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
+Our model is able to track these developments very well. The blue line in
+Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
+officially recorded cases in the aggregate. This is also true for infections by age and
+geographical region, which are shown in the supplementary materials
+(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively). We can
+disentangle various mechanisms due to the distinct temporal variation in the drivers of
+the pandemic. Next to the stringency index, the three lines in
+Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
 regulations, and seasonality evolved since early September for each of the three broad
@@
 these measures completely determine the spread of SARS-CoV-2 in
-2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} \comment[id=K]{We
-have not done the decomposition for 2020 yet. Is on my todo list} for their individual
-contributions. Two things are particularly interesting. First, all lines broadly follow
-the stringency index and they would do so even more if we left out seasonality. Second,
-the most stringent regulations are associated with the period of strong decreases in new
-infection between late December 2020 and mid-February 2021. The measures were not
-enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period.
-The steep drop in recorded cases during May 2021 is associated with rapid tests to
-X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and
-mostly seasonality impacting a fall in the relative infectiousness of contacts outside
-of work and school\comment[id=HM]{check with corrected graph}.
+2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference}
+\comment[id=K]{We have not done the decomposition for 2020 yet. Is on my todo list}
+\comment[id=HM]{I guess my question was whether I was missing any factors.}
+for their individual contributions. Two things are particularly interesting. First, all
+lines broadly follow the stringency index and they would do so even more if we left out
+seasonality and school vacations (roughly the last two weeks of October, two weeks each
+around Christmas and Easter, and some days in late May). Second, the most stringent
+regulations are associated with the period of strong decreases in new infection between
+late December 2020 and mid-February 2021. The measures were not enough, however, to stop
+the B.1.1.7 variant from spreading in the subsequent period. The steep drop in recorded
+cases during May 2021 is associated with at least weekly rapid tests to around
+42~percent of the population, a vaccination rate that rose from 28\% to 43\%, and mostly
+seasonality impacting a fall in the relative infectiousness of contacts outside of work
+and school.
 
@@
 
-    The decomposition is based on Shapley values where the individual contribution of a
-    channel is its average contribution over different sizes of coalitions (combinations
-    with other channels). The individual contribution to a coalition is the difference
-    between the effect size of the coalition with the particular channel and without.}
+        The decomposition is based on Shapley values where the individual contribution
+        of a channel is its average contribution over different sizes of coalitions
+        (combinations with other channels). The individual contribution to a coalition
+        is the difference between the effect size of the coalition with the particular
+        channel and without.}
 \end{figure}
@@
     \item Trump-effect: More testing = more cases true for how long?
+    \item Towards the end, total tests go down: Preferences for tests remain the same,
+    but infection rates lower \comment[id=HM]{Janos, please fill up}
 \end{itemize}

### 2021-06-09 paper/report.tex
Formatting.
Add missing references.

@@
 reasons triggering rapid test demand; they are lowest for infected household members and
-highest for meeting other people.\comment[id=HM]{True?}.
+highest for meeting other people.\comment[id=HM]{True?}\comment[id=K]{FALSE.
+We only have one private demand threshold (over time) which is the same for own symptoms and no PCR test, household member with symptoms or positive test and for having a weekly meeting with friends.}.
 
@@
 
-We apply this model to Germany. In March and April 2020, the country broke the first
-wave of the pandemic fairly quickly. Between mid-May and mid-September, daily new
-infections were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We
-model the period mid-September 2020 to the end of May 2021. We pick the starting date
-for two reasons. First, we do not include the first wave because the environment was
-very different (e.g., aggregate PCR test capacity was much lower and we would require a
-very different model for calculating the share of known cases) Second, a large fraction
-cases during summer of 2020 were traced to international travel
-\citep{KochInstitut2021}\comment[id=HM]{Include
-\url{https://www.nature.com/articles/s41586-021-03677-y} in citation}, but the precise
-number is difficult to model.
+We apply this model to Germany. In March and April 2020, the country broke the first wave
+of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
+were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
+period mid-September 2020 to the end of May 2021. We pick the starting date for two
+reasons. First, we do not include the first wave because the environment was very
+different (e.g., aggregate PCR test capacity was much lower and we would require a very
+different model for calculating the share of known cases) Second, a large fraction cases
+during summer of 2020 were traced to international travel
+\citep{KochInstitut2021,Hodcroft2021} but the precise number is difficult to model.
 

### 2021-06-09 paper/report.tex
Updates from Zoom meeting.

@@
 
-In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
-possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
-differentiate between an initial period of infection without being infectious or showing
-symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death. The probabilities of transitioning between these
-states depend on age; their duration is random within intervals calibrated to medical
-literature (for a detailed description see Section~\ref{sec:medical_params}).
-Conditional on the type of contact, infectiousness independent of ages
-\citep{Jones2021}.\comment[id=HM]{Seems broadly standard, can we cite someone?}
+In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
+age. A possible infection progresses as shown in
+Figure~\ref{fig:disease_progression}. We differentiate between an initial period
+of infection without being infectious or showing symptoms, being infectious
+(presymptomatic or asymptomatic), showing symptoms, requiring intensive care,
+and recovery or death.\comment[id=HM]{Seems broadly standard, can we cite
+someone?} The probabilities of transitioning between these states depend on age;
+their duration is random within intervals calibrated to medical literature (for
+a detailed description see Section~\ref{sec:medical_params}). Conditional on the
+type of contact, infectiousness is independent of age \citep{Jones2021}.
 
@@
 individual compliance parameter. This parameter determines whether she takes up rapid
-tests offered by employers or follows up on private reasons. Thresholds differ across
-reasons triggering rapid test demand; they are lowest for infected household members and
-highest for meeting other people.\comment[id=HM]{True?}\comment[id=K]{FALSE.
-We only have one private demand threshold (over time) which is the same for own symptoms and no PCR test, household member with symptoms or positive test and for having a weekly meeting with friends.}.
+tests offered by employers or follows up on private reasons. The thresholds are lower for
+tests in a private setting than for tests at the workplace.\comment[id=HM]{True? @Janos}
 
@@
 that we can use a wide array of data to identify and estimate the model's many
-parameters.\footnote{See section S.X of the supplementary materials for an
-    overview.}\comment[id=HM]{Need to make a list including symbols, maybe you can fetch the
-parameters from the model and then we brainstorm re notation?} Mobility data is used to
+parameters.\footnote{See section S.XXX of the supplementary materials for an
+    overview.} Mobility data is used to
 model the reductions in work contacts. School and daycare policies are incorporated
@@
 of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
-were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
+were below 20 per Million and day.\comment[id=HM]{Cite our world in data/RKI} We model the
 period mid-September 2020 to the end of May 2021. We pick the starting date for two
@@
 
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
-of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the \replaced[id=HM]{color}{blue?} line in
-Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
-which tracks the tightness of non-pharmaceutical interventions. We transform the index
-so that lower values represent higher levels of restrictions. A value of zero means all
-measures incorporated in the index are turned on. The value 1 represents the situation
-in mid-September, with restrictions on gatherings and public events, masking
-requirements, but open schools and workplaces (the raw value of the index at that point
-is 49.5). In the seven weeks between mid September and early November, cases increased
-by a factor of 10. Restrictions were somewhat tightened in mid-October and again in
-early November. New infections remained constant throughout November, before rising
-again in December, which prompted the most stringent lockdown to this date. Schools and
-daycare centers were closed again, so were customer-facing businesses except for grocery
-and drug stores. From the peak of the second wave just before Christmas until the trough
-in mid-February, newly detected cases decreased by almost three quarters. The third wave
-in the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in
-March. See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed;
-e.g., hairdressers and home improvement stores were allowed to open again to the public.
+Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the
+pandemic and of its drivers. The black line in Figure~\ref{fig:aggregated_fit}
+shows officially recorded cases; the black line in
+Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index
+\citep{Hale2020}, which tracks the tightness of non-pharmaceutical
+interventions. We transform the index so that lower values represent higher
+levels of restrictions. A value of zero means all measures incorporated in the
+index are turned on. The value 1 represents the situation in mid-September, with
+restrictions on gatherings and public events, masking requirements, but open
+schools and workplaces (the raw value of the index at that point is 49.5). In
+the seven weeks between mid September and early November, cases increased by a
+factor of 10. Restrictions were somewhat tightened in mid-October and again in
+early November. New infections remained constant throughout November, before
+rising again in December, which prompted the most stringent lockdown to this
+date. Schools and daycare centers were closed again, so were customer-facing
+businesses except for grocery and drug stores. From the peak of the second wave
+just before Christmas until the trough in mid-February, newly detected cases
+decreased by almost three quarters. The third wave in the spring of 2021 is
+associated with the B.1.1.7 strain, which became dominant in March. See
+Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
+hairdressers and home improvement stores were allowed to open again to the
+public.
 
@@
 Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
-officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region, which are shown in the supplementary materials
+officially recorded cases in the aggregate. This is also true for infections by
+age and geographical region, which are shown in the supplementary materials
 (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively). We can
-disentangle various mechanisms due to the distinct temporal variation in the drivers of
-the pandemic. Next to the stringency index, the three lines in
-Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
-regulations, and seasonality evolved since early September for each of the three broad
-contact networks. For example, a value of 0.75 for the work multiplier means that if the
-environment was the same as in September (levels of infection rates, no rapid tests or
-vaccinations, only the wildtype virus present), infections at the workplace would be
-reduced by 25\%. This reduction is the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. Along with the levels of infections,
-these measures completely determine the spread of SARS-CoV-2 in
-2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference}
-\comment[id=K]{We have not done the decomposition for 2020 yet. Is on my todo list}
-\comment[id=HM]{I guess my question was whether I was missing any factors.}
-for their individual contributions. Two things are particularly interesting. First, all
-lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations (roughly the last two weeks of October, two weeks each
-around Christmas and Easter, and some days in late May). Second, the most stringent
-regulations are associated with the period of strong decreases in new infection between
-late December 2020 and mid-February 2021. The measures were not enough, however, to stop
-the B.1.1.7 variant from spreading in the subsequent period. The steep drop in recorded
-cases during May 2021 is associated with at least weekly rapid tests to around
-42~percent of the population, a vaccination rate that rose from 28\% to 43\%, and mostly
-seasonality impacting a fall in the relative infectiousness of contacts outside of work
-and school.
+disentangle various mechanisms due to the distinct temporal variation in the
+drivers of the pandemic. Next to the stringency index, the three lines in
+Figure~\ref{fig:stringency_index} summarize how contact reductions, increased
+hygiene regulations, and seasonality evolved since early September for each of
+the three broad contact networks. For example, a value of 0.75 for the work
+multiplier means that if the environment was the same as in September (levels of
+infection rates, no rapid tests or vaccinations, only the wildtype virus
+present), infections at the workplace would be reduced by 25\%. This reduction
+is the product of the effect of contact reductions, increased hygiene
+regulations, and seasonality. Along with the levels of infections, these
+measures determine the spread of SARS-CoV-2 in 2020. See
+Appendix~\comment[id=HM]{Reference} for separate plots of the three factors. Two
+aspects are particularly interesting. First, all lines broadly follow the
+stringency index and they would do so even more if we left out seasonality and
+school vacations (roughly the last two weeks of October, two weeks each around
+Christmas and Easter, and some days in late May). Second, the most stringent
+regulations are associated with the period of strong decreases in new infection
+between late December 2020 and mid-February 2021. The measures were not enough,
+however, to stop the B.1.1.7 variant from spreading in the subsequent period.
+The steep drop in recorded cases during May 2021 is associated with at least
+weekly rapid tests to around 42~percent of the population, a vaccination rate
+that rose from 28\% to 43\%, and mostly seasonality impacting a fall in the
+relative infectiousness of contacts outside of work and school.
 
-Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
-vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way
-as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model
-fit (the bold blue line\comment[id=HM]{Make sure this is the case}, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (x line),
-and three scenarios turning these factors off one by one.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to decompose
-the difference in total infections between the scenario without any of the three factors
-and our main specification.
+Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid
+tests, vaccinations, and of seasonality during 2021, assuming NPIs to have
+evolved the same way as in the baseline scenario.
+Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line,
+same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three
+factors (red line), and three scenarios turning these factors off one by one.
+Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs
+Shapley values to decompose the difference in total infections between the
+scenario without any of the three factors and our main specification.
 

### 2021-06-09 paper/medical_parameters.tex
Add some references. Address some of HM's comments. Not running through. Probably missing curly bracket somewhere.
Fix reference unicode error. Minor improvements to the medical parameters.

@@
 \subsection{Medical Parameters}
+\label{sec:medical_params}
 
@@
 % Latency Period
-The first medical parameter we need is the length of the period between infection and
-the start of infectiousness, the so called latent period.
-We infer it from two other measures that are more common in the medical literature:
-Firstly, the time between infection and the onset of symptoms, the incubation period.
-Secondly, the time between the start of infectiousness and the onset of symptoms.
-We assume that the latency period is the same for symptomatic and asymptomatic individuals.
+The first medical parameter we need is the length of the period between infection and the
+start of infectiousness, the so called latent period. We infer it from two other measures
+that are more common in the medical literature: Firstly, the time between infection and
+the onset of symptoms, the incubation period. Secondly, the time between the start of
+infectiousness and the onset of symptoms. We assume that the latency period is the same
+for symptomatic and asymptomatic individuals.
 
@@
 Once individuals become infectious a share of them goes on to develop symptoms while
-others remain asymptomatic.
-We rely on data by \cite{Davies2020} for the age-dependent probability to develop
-symptoms. It varies from 25\% for children and young adults to nearly 70\% for the
-elderly.
+others remain asymptomatic. We rely on data by \cite{Davies2020} for the age-dependent
+probability to develop symptoms. It varies from 25\% for children and young adults to
+nearly 70\% for the elderly.
 
 % Incubation Period
-The incubation period is usually estimated to be two to twelve days.
-A meta analysis by \citet{McAloon2020} comes to the conclusion that
-``The incubation period distribution may be modeled with a lognormal distribution with
-pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50
-(95\% CI 0.46 to 0.55), respectively.''
-For simplicity we discretize this distribution into four bins.
+The incubation period is usually estimated to be two to twelve days. A meta analysis by
+\citet{McAloon2020} comes to the conclusion that ``The incubation period distribution may
+be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\%
+CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For
+simplicity we discretize this distribution into four bins.
 
 % Length of the presymptomatic stage
-The \href{European Centre for Disease Prevention and Control}{https://bit.ly/3pi18Ax}
-reports that people become infectious between one and two days before symptoms start.\footnotemark
+The \href{https://bit.ly/3pi18Ax}{European Centre for Disease Prevention and Control}
+reports that people become infectious between one and two days before symptoms
+start.\footnotemark
 
-\footnotetext{
-This is similar to \citet{He2020} and in line with
-\citet{Peak2020}.}
+\footnotetext{This is similar to \citet{He2020} and in line with \citet{Peak2020}.}
 
@@
 
-For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8
-days of infectiousness afterwards.\footnotemark Thus, we arrive at 0 to 13 days as the
-range for infectiousness among individuals who become symptomatic (see also figure 5).
-This duration range is very much in line with the meta-analysis’ reported evidence for
-asymptomatic individuals (see their figure 1).
+For symptomatic cases they arrive at zero to five days before symptom onset (figure 2)
+and three to eight days of infectiousness afterwards.\footnotemark Thus, we arrive at 0
+to 13 days as the range for infectiousness among individuals who become symptomatic (see
+also figure 5).
 
-\footnotetext{
-    Viral loads may be detected much later but 8 days seems to be the time after which
-    most people are culture negative, as also reported by \citet{Singanayagam2020}.
-}
-
-Following this evidence we assume the following discretized distribution of the
-infectiousness period: 10\% of individuals are infectious for three days, 25\% for five
-days, another 25\% for seven days, 20\% for nine days and 20\% for eleven days.
+\footnotetext{Viral loads may be detected much later but eight days seems to be the time
+    after which most people are culture negative, as also reported by
+    \citet{Singanayagam2020}.}
 
@@
 symptomatic cases.
-
-We collapse the data to the following distribution: 10\% recover after 15 days and 30\%
-require 18, 22 or 27 days respectively.
 

### 2021-06-09 paper/model_validation.tex
Add age group and state fit back in.
Add fitness plot with single simulation lines. Fix month in reference.

@@
 \centering
-  % \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
 \caption{Simulated and Empirical Infections}
@@
 \centering
-  % \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
+  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
 \caption{Simulated and Empirical Infections by Age Group}
@@
 \centering
-  % \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
+  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
 \caption{Simulated and Empirical Infections by Federal State}

### 2021-06-09 paper/medical_parameters.tex
Simplify medical parameters section and hopefully fix references.

@@
 
-\subsubsection{Duration until Infectiousness and Symptoms}
+% Duration until Infectiousness and Symptoms ----------------------------------
 
@@
 
-\subsubsection{Duration of Infectiousness}
+% Duration of Infectiousness ----------------------------------------------------------
 
@@
 asymptomatic individuals as evidence suggests little differences in the transmission
-rates between symptomatic and asymptomatic patients
-(\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic
-individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).
+rates between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral
+load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020},
+\citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the duration of
+infectiousness is based on \citet{Byrne2020}. For symptomatic cases they arrive at zero
+to five days before symptom onset (see their figure 2) and three to eight days of
+infectiousness afterwards.\footnote{Viral loads may be detected much later but eight days
+seems to be the time after which most people are culture negative, as also reported by
+\citet{Singanayagam2020}.} Thus, we arrive at 0 to 13 days as the range for
+infectiousness among individuals who become symptomatic (see also figure 5).
 
-Our distribution of the duration of infectiousness is based on \citet{Byrne2020}.
-
-For symptomatic cases they arrive at zero to five days before symptom onset (figure 2)
-and three to eight days of infectiousness afterwards.\footnotemark Thus, we arrive at 0
-to 13 days as the range for infectiousness among individuals who become symptomatic (see
-also figure 5).
-
-\footnotetext{Viral loads may be detected much later but eight days seems to be the time
-    after which most people are culture negative, as also reported by
-    \citet{Singanayagam2020}.}
-
-\subsubsection{Duration of Symptoms}
+% Duration of Symptoms ----------------------------------------------------------
 
 We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
-Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring
-symptomatic cases.
+Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases. We
+only disaggregate by age how likely individuals are to require intensive
+care.\footnote{The length of symptoms is not very important in our model given that
+individuals mostly stop being infectious before their symptoms cease.}
 
-These numbers are only used for mild cases. We do not disaggregate by age. Note that the
-length of symptoms is not very important in our model given that individuals stop being
-infectious before their symptoms cease.
+% Time from Symptom Onset to Admission to ICU ------------------------------------------
 
-\subsubsection{Time from Symptom Onset to Admission to ICU}
-
-The data on how many percent of symptomatic patients will require ICU is pretty thin. We
-rely on data by the US CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the
-OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the
-probabilities of requiring intensive care per age group.
+For the time from symptom onset until need for intensive care we rely on data by the US
+CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
+Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of
+requiring intensive care per age group.
 
@@
 For those who will require intensive care we follow \citet{Chen2020} who estimate the
-time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.
+time from symptom onset to ICU admission as 8.5 $\pm$ 4 days. This aligns well with
+numbers reported for the time from first symptoms to hospitalization:
+\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also
+in line with the duration estimates collected by
+\href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the
+Robert-Koch-Institut}. We assume that the time between symptom onset and ICU takes 4, 6,
+8 or 10 days with equal probabilities. As we do not model nursing homes, do not focus on
+matching deaths and do not use the number of individuals in intensive care to estimate
+our parameters, these numbers are not important for our empirical results.
 
-This aligns well with numbers reported for the time from first symptoms to
-hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation
-of 4. This is also in line with the durations collected by
-\href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the
-Robert-Koch-Institut}.
-
-We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with
-equal probabilities. These times mostly matter for the ICU capacities.
-
-\subsubsection{Death and Recovery from ICU}
+% Death and Recovery from ICU ----------------------------------------------------------
 
 We take the survival probabilities and time to death and time until recovery from
-intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.
-
-They report time until death to have a mean of 11.74 days and a standard deviation of
-8.79 days. Approximating this with the normal distribution, we have nearly 10\%
-probability mass below 0. We use it nevertheless as several other distributions (such as
-chi squared and uniform) were unable to match the variance. Discretizing the
-distribution leads to 41\% of individuals who will die from Covid-19 after one day in
-intensive care, 22\% day after 12 days, 29\% after 20 days and 7\% after 32 days. Again,
-we rescale this for every age group among those that will not survive.
-
-They report a mean duration of 18.8 days until recovery and a standard deviation of
-12.21 days. Approximating this with the normal distribution, we have over 5\%
-probability mass below 0. Of those who recover in intensive care, 22\% do so
+intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
+time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
+match this approximately we discretize that 41\% of individuals who will die from
+Covid-19 do so after one day in intensive care, 22\% day after 12 days, 29\% after 20
+days and 7\% after 32 days. Again, we rescale this for every age group among those that
+will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
+reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
+days. We discretize this such that of those who recover in intensive care, 22\% do so
 after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

### 2021-06-09 paper/report.tex
Add references and solve comments.

@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many
-months---and often years---away from reaching vaccination-induced herd
-immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the
-meantime, it is of utmost importance to employ an effective mix of strategies for
-containing the virus. The most frequent initial response was a set of non-pharmaceutical
-interventions (NPIs) to reduce contacts between individuals. While this has allowed some
-countries to sustain equilibria with very low infection numbers\footnote{See
-    \citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
-    sustained with test-trace-and-isolate policies.}, most have seen large fluctuations of
-infection rates. Containment measures have become increasingly diverse and now include
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
-influence of seasonal patterns or more infectious virus strains are well understood in
-quantitative terms. This paper develops a model incorporating all these factors. The
-framework allows to combine a wide variety of data in a timely fashion, making it useful
-to predict the effects of various interventions. We apply the model to Germany and show
-that rapid testing had the largest impact on the reduction in infections by almost 80\%
-during the month of May 2021. We conclude that rapid tests have a large role to play at
-least as long as vaccinations have not been offered to an entire population.
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination-induced herd immunity
+\citep{Swaminathan2021}.\comment[id=K]{WHO Press Conference saying we won't reach herd
+immunity in 2021} In the meantime, it is of utmost importance to employ an effective mix
+of strategies for containing the virus. The most frequent initial response was a set of
+non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
+this has allowed some countries to sustain equilibria with very low infection
+numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
+numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
+fluctuations of infection rates. Containment measures have become increasingly diverse
+and now include testing, more nuanced NPIs, and contact tracing. Neither these policies'
+effect nor the influence of seasonal patterns or more infectious virus strains are well
+understood in quantitative terms. This paper develops a model incorporating all these
+factors. The framework allows to combine a wide variety of data in a timely fashion,
+making it useful to predict the effects of various interventions. We apply the model to
+Germany and show that rapid testing had the largest impact on the reduction in infections
+by almost 80\% during the month of May 2021. We conclude that rapid tests have a large
+role to play at least as long as vaccinations have not been offered to an entire
+population.
 
@@
 
-In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
-age. A possible infection progresses as shown in
-Figure~\ref{fig:disease_progression}. We differentiate between an initial period
-of infection without being infectious or showing symptoms, being infectious
-(presymptomatic or asymptomatic), showing symptoms, requiring intensive care,
-and recovery or death.\comment[id=HM]{Seems broadly standard, can we cite
-someone?} The probabilities of transitioning between these states depend on age;
-their duration is random within intervals calibrated to medical literature (for
-a detailed description see Section~\ref{sec:medical_params}). Conditional on the
-type of contact, infectiousness is independent of age \citep{Jones2021}.
+In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
+possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
+differentiate between an initial period of infection without being infectious or showing
+symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
+intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
+The probabilities of transitioning between these states depend on age; their duration is
+random within intervals calibrated to medical literature (for a detailed description see
+Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
+independent of age \citep{Jones2021}.
 
@@
 We apply this model to Germany. In March and April 2020, the country broke the first wave
-of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
-were below 20 per Million and day.\comment[id=HM]{Cite our world in data/RKI} We model the
+of the pandemic fairly quickly. Between May and mid-September, daily new infections
+were below 20 per Million and day \citep{MaxRoser2020}. We model the
 period mid-September 2020 to the end of May 2021. We pick the starting date for two

### 2021-06-09 paper/appendix_data.tex

--- 2021-06-07 paper/calibration_and_estimation.tex
@@
-\section{Calibration and Estimation}
+\section{Data and Parameters}
 \label{sec:calibration_and_estimation}
@@
 Most of these parameters can be calibrated from existing datasets or the medical
-literature. Only the infection probabilities have to be estimated inside the model by
-fitting it to time series data of case numbers and fatality rates.
+literature or calibrated from surveys and empirical datasets.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
+\caption{\textbf{Share of Individuals that do a Rapid Test.} \textcolor{red}{Talk about the interpretation of each line.}}
+\end{figure}
+
 In our model, there are five reasons why rapid tests are done:

### 2021-06-09 paper/appendix_results.tex

--- 2021-06-09 paper/model_validation.tex
@@
-\section{Model Validation}
-\label{sec:model_validation}
+\section{Additional Results}
+\label{sec:additional_results}
 
-This compares simulated data from our model with empirical data from Germany. We at
+\subsection{Simulated vs. Empirical Data}
+\label{subsec:new_cases_fit}
+
+This compares simulated data from our model with empirical data from Germany. We look at
 observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
@@
 \textcolor{red}{summarize the fit; Some of this needs to go in the appendix. I suggest the age group and federal state fit goes there???}
-
-\subsection{Observed Infections}
-\label{sub:in_sample_fit}
-
-% Despite fitting only four infection probabilities, the in-sample fit is very good. The
-% best fit is achieved in the largest age groups. This is so mechanically, because we
-% weigh the deviations between simulated and observed infection rates by group sizes.
-% The worst fit is achieved for the 80 to 100 years old. There are three reasons for
-% this: Firstly, we only model private households at the moment, meaning that nursing
-% homes are not part of the synthetic population we simulate. Since community housing
-% residents are part of the German Mikrozensus \citep{FDSAeDBUDL2018} which forms the
-% basis for our synthetic population, we can and plan to include community housing
-% residents in the near future. Secondly, the elderly, especially nursing home
-% residents, are tested more often than the general population leading to more cases
-% being detected in this age group. We are currently working to include age-variant
-% testing strategies, using \href{https://ars.rki.de/Content/COVID19/Main.aspx}{data by
-% the RKI} \citep{Seifried2020}.
 
@@
 
-\section{The Spread of the B117 Mutation}
+\subsection{Scenarios}
+\label{subsec:appendix_scenarios}
 
-\textcolor{red}{We don't have this figure in the repo yet, but the fit is perfect}
+This is the results section.
+
+\textcolor{red}{The results can be found in `figures/results`. This includes both figures
+and tables for lookup of numbers and summary tables.}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+\end{subfigure}
+\caption{The Effect of Policies on Observed and Unobserved Cases}
+\label{fig:explain_decline}
+\figurenotes{\textcolor{red}{\ldots}}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+\end{subfigure}
+\caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
+\figurenotes{\textcolor{red}{K: One of the scenarios starts too early. Will be fixed with the next full simulations run.}}
+\label{fig:school_scenarios}
+\end{figure}
 
@@
 
-\subsection{Out-of-Sample Fit}
-\label{sub:out_of_sample_fit}
+\input{../figures/results/tables/student_infections_table}
 
-We can assess the out-of-sample fit by projecting the effect of the lockdown light and
-comparing it to case numbers until mid of November. It is important to note that this is
-not just a simple extrapolation of a time trend because the lockdown light only started
-after the estimation period. The out-of-sample fit can be assessed in
-Figure~\ref{fig:out-of-sample-fit}.
-
-\begin{figure}[!tp]
-    \centering
-    \includegraphics[width=\textwidth]{../figures/out_of_sample_validation}
-    \caption{Predicted effect of the "Lockdown Light" on infection rates.}
-    \label{fig:out-of-sample-fit}
-    \figurenotes{For the time period until the beginning of November, the figure shows
-        the weekly incidence rates of infections per 100,000 people from reported
-        (black) versus simulated (blue) data. With the start of November, the
-        projections of the three scenarios, optimistic (blue), neutral (red), and
-        pessimistic (mint green), are shown until the beginning of the Christmas
-        holidays. The actual incidence rates (black) are reported until the 24th
-        November.}
-\end{figure}
-
-The model correctly predicts the effect of the lockdown light with reasonable accuracy.
-In particular, the actual case numbers are between our neutral and pessimistic
-projection. The plot also shows that ending the lockdown light on November 30, as was
-originally planned, would lead to an explosive growth in case numbers in all scenarios.
+\FloatBarrier

### 2021-06-09 paper/report.tex
Add a few points about testing.

--- 2021-06-09 paper/report.tex
@@
           but infection rates lower \comment[id=HM]{Janos, please fill up}
+    \item Actually performed tests increase until May, stay constant during May and then start to drop
+    \item This is the combination of the following things:
+    \begin{itemize}
+        \item time constant willingness to take tests
+        \item monotonically increasing availability of tests
+        \item endogenous number of situations in which a test should be taken (mainly positive tests of household members)
     \end{itemize}
+    \item This is a classic way of modeling in economics: Preferences stay constant
+          (and can thus be elicited in just one survey) but outcomes change because
+          circumstances change.
+
+\end{itemize}
+
+An important feature of the way we model rapid tests is that the number of actually
+performed test is determined endogenously. The number of weekly tests increases until
+the end of May and then starts to drop. This is because many  It is determined by the availability of tests,
+the individual compliance with testing recommendations and the occurrence of situations
+in which a test should be taken.
 
@@
 % conclusions of the paper.
-
-\section{Supplementary Material}
-
-\begin{enumerate}
-    \item Model
-    \item Data
-    \item Identification and Estimation
-\end{enumerate}

### 2021-06-09 paper/report.tex
Correct a few references.

@@
 
-We apply this model to Germany. In March and April 2020, the country broke the first wave
-of the pandemic fairly quickly. Between May and mid-September, daily new infections
-were below 20 per Million and day \citep{MaxRoser2020}. We model the
-period mid-September 2020 to the end of May 2021. We pick the starting date for two
-reasons. First, we do not include the first wave because the environment was very
-different (e.g., aggregate PCR test capacity was much lower and we would require a very
-different model for calculating the share of known cases) Second, a large fraction cases
-during summer of 2020 were traced to international travel
-\citep{KochInstitut2021,Hodcroft2021} but the precise number is difficult to model.
+We apply this model to Germany. In March and April 2020, the country broke the first
+wave of the pandemic fairly quickly. Between May and mid-September, daily new infections
+were below 20 per Million and day \citep{owidcoronavirus}. We model the period
+mid-September 2020 to the end of May 2021. We pick the starting date for two reasons.
+First, we do not include the first wave because the environment was very different
+(e.g., aggregate PCR test capacity was much lower and we would require a very different
+model for calculating the share of known cases) Second, a large fraction cases during
+summer of 2020 were traced to international travel \citep{KochInstitut2021,Hodcroft2021}
+but the precise number is difficult to model.
 

### 2021-06-09 figures/results/tables/student_infections_table.tex
New results figures and tables with both shapley decompositions.

@@
 \midrule
- educ open after easter  without tests &                                             752932 \\
- educ open after easter  with tests    &                                             583570 \\
- close educ after easter               &                                             443978 \\
- baseline                              &                                             459436 \\
+ educ open after easter  without tests &                                             775324 \\
+ educ open after easter  with tests    &                                             604078 \\
+ close educ after easter               &                                             469274 \\
+ baseline                              &                                             478080 \\
 \bottomrule

### 2021-06-09 paper/report.tex
Formatting, mostly.

@@
 
+\begin{figure}
+\centering
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
+        \caption{Decomposition of Rapid Tests}
+\end{figure}
+\begin{figure}
+\centering
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
+        \caption{Decomposition of Rapid Tests}
+\end{figure}
+
 Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination-induced herd immunity
+containment in the medium term. However, most countries find themselves many
+months---and often years---away from reaching vaccination-induced herd immunity
 \citep{Swaminathan2021}.\comment[id=K]{WHO Press Conference saying we won't reach herd
@@
 making it useful to predict the effects of various interventions. We apply the model to
-Germany and show that rapid testing had the largest impact on the reduction in infections
-by almost 80\% during the month of May 2021. We conclude that rapid tests have a large
-role to play at least as long as vaccinations have not been offered to an entire
-population.
+Germany and show that rapid testing had the largest impact on the reduction in
+infections by almost 80\% during the month of May 2021. We conclude that rapid tests
+have a large role to play at least as long as vaccinations have not been offered to an
+entire population.
 
@@
 individual compliance parameter. This parameter determines whether she takes up rapid
-tests offered by employers or follows up on private reasons. The thresholds are lower for
-tests in a private setting than for tests at the workplace.\comment[id=HM]{True? @Janos}
+tests offered by employers or follows up on private reasons. The thresholds are lower
+for tests in a private setting than for tests at the workplace.\comment[id=HM]{True?
+@Janos}
 
@@
 that we can use a wide array of data to identify and estimate the model's many
-parameters.\footnote{See section S.XXX of the supplementary materials for an
-    overview.} Mobility data is used to
-model the reductions in work contacts. School and daycare policies are incorporated
-directly from official directives. Infection rates by age and geographical region are
-estimated to match to officially recorded numbers; so is the prevalence of virus
-strains. The model yields total infections, but only a fraction of those will be
-officially recorded. We model the fraction of known cases as depicted in
+parameters.\footnote{See section S.XXX of the supplementary materials for an overview.}
+Mobility data is used to model the reductions in work contacts. School and daycare
+policies are incorporated directly from official directives. Infection rates by age and
+geographical region are estimated to match to officially recorded numbers; so is the
+prevalence of virus strains. The model yields total infections, but only a fraction of
+those will be officially recorded. We model the fraction of known cases as depicted in
 Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
-\footnote{EQUATIONS FOR THE FOURTH PICTURE (find names!):
+% \footnote{EQUATIONS FOR THE FOURTH PICTURE (find names!):
 
-    The probabilities for Figure~\subref{fig:model_for_official_cases} are the
-    following:
+%     The probabilities for Figure~\subref{fig:model_for_official_cases} are the
+%     following:
 
-    \begin{align*}
-        x & = \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_symptomatic\_and\_untested}                    \\
-        y & = \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR | S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
-        z & = \frac{P(PCR | RT)}{survey}
-    \end{align*}
-    where untested means that individuals are not waiting on the result of a previously
-    administered rapid test.} A further advantage is that the simulated data have a
-structure that resembles datasets used for regression models, which allows
-additional plausibility checks by re-running the same models on the model-generated
-data.\comment[id=HM]{Only keep if we actually do something like that.}
+%     \begin{align*} x & = \frac{n\_newly\_infected \times share\_known\_cases \times
+%         P(PCR | S)}{n\_symptomatic\_and\_untested}                    \\
+%         y & = \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR |
+%         S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
+%         z & = \frac{P(PCR | RT)}{survey} \end{align*} where untested means that
+%     individuals are not waiting on the result of a previously administered rapid
+%     test.} 
+A further advantage is that the simulated data have a structure that resembles datasets
+used for regression models, which allows additional plausibility checks by re-running
+the same models on the model-generated data.\comment[id=HM]{Only keep if we actually do
+something like that.}
 
@@
 
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the
-pandemic and of its drivers. The black line in Figure~\ref{fig:aggregated_fit}
-shows officially recorded cases; the black line in
-Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index
-\citep{Hale2020}, which tracks the tightness of non-pharmaceutical
-interventions. We transform the index so that lower values represent higher
-levels of restrictions. A value of zero means all measures incorporated in the
-index are turned on. The value 1 represents the situation in mid-September, with
-restrictions on gatherings and public events, masking requirements, but open
-schools and workplaces (the raw value of the index at that point is 49.5). In
-the seven weeks between mid September and early November, cases increased by a
-factor of 10. Restrictions were somewhat tightened in mid-October and again in
-early November. New infections remained constant throughout November, before
-rising again in December, which prompted the most stringent lockdown to this
-date. Schools and daycare centers were closed again, so were customer-facing
-businesses except for grocery and drug stores. From the peak of the second wave
-just before Christmas until the trough in mid-February, newly detected cases
-decreased by almost three quarters. The third wave in the spring of 2021 is
-associated with the B.1.1.7 strain, which became dominant in March. See
+Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
+of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
+recorded cases; the black line in Figure~\ref{fig:stringency_index} the Oxford Response
+Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
+interventions. We transform the index so that lower values represent higher levels of
+restrictions. A value of zero means all measures incorporated in the index are turned
+on. The value 1 represents the situation in mid-September, with restrictions on
+gatherings and public events, masking requirements, but open schools and workplaces (the
+raw value of the index at that point is 49.5). In the seven weeks between mid September
+and early November, cases increased by a factor of 10. Restrictions were somewhat
+tightened in mid-October and again in early November. New infections remained constant
+throughout November, before rising again in December, which prompted the most stringent
+lockdown to this date. Schools and daycare centers were closed again, so were
+customer-facing businesses except for grocery and drug stores. From the peak of the
+second wave just before Christmas until the trough in mid-February, newly detected cases
+decreased by almost three quarters. The third wave in the spring of 2021 is associated
+with the B.1.1.7 strain, which became dominant in March. See
 Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the
-public.
+hairdressers and home improvement stores were allowed to open again to the public.
 
@@
 Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
-officially recorded cases in the aggregate. This is also true for infections by
-age and geographical region, which are shown in the supplementary materials
+officially recorded cases in the aggregate. This is also true for infections by age and
+geographical region, which are shown in the supplementary materials
 (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively). We can
-disentangle various mechanisms due to the distinct temporal variation in the
-drivers of the pandemic. Next to the stringency index, the three lines in
-Figure~\ref{fig:stringency_index} summarize how contact reductions, increased
-hygiene regulations, and seasonality evolved since early September for each of
-the three broad contact networks. For example, a value of 0.75 for the work
-multiplier means that if the environment was the same as in September (levels of
-infection rates, no rapid tests or vaccinations, only the wildtype virus
-present), infections at the workplace would be reduced by 25\%. This reduction
-is the product of the effect of contact reductions, increased hygiene
-regulations, and seasonality. Along with the levels of infections, these
-measures determine the spread of SARS-CoV-2 in 2020. See
-Appendix~\comment[id=HM]{Reference} for separate plots of the three factors. Two
-aspects are particularly interesting. First, all lines broadly follow the
-stringency index and they would do so even more if we left out seasonality and
-school vacations (roughly the last two weeks of October, two weeks each around
-Christmas and Easter, and some days in late May). Second, the most stringent
-regulations are associated with the period of strong decreases in new infection
-between late December 2020 and mid-February 2021. The measures were not enough,
-however, to stop the B.1.1.7 variant from spreading in the subsequent period.
-The steep drop in recorded cases during May 2021 is associated with at least
-weekly rapid tests to around 42~percent of the population, a vaccination rate
-that rose from 28\% to 43\%, and mostly seasonality impacting a fall in the
-relative infectiousness of contacts outside of work and school.
+disentangle various mechanisms due to the distinct temporal variation in the drivers of
+the pandemic. Next to the stringency index, the three lines in
+Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
+regulations, and seasonality evolved since early September for each of the three broad
+contact networks. For example, a value of 0.75 for the work multiplier means that if the
+environment was the same as in September (levels of infection rates, no rapid tests or
+vaccinations, only the wildtype virus present), infections at the workplace would be
+reduced by 25\%. This reduction is the product of the effect of contact reductions,
+increased hygiene regulations, and seasonality. Along with the levels of infections,
+these measures determine the spread of SARS-CoV-2 in 2020. See
+Appendix~\comment[id=HM]{Reference} for separate plots of the three factors. Two aspects
+are particularly interesting. First, all lines broadly follow the stringency index and
+they would do so even more if we left out seasonality and school vacations (roughly the
+last two weeks of October, two weeks each around Christmas and Easter, and some days in
+late May). Second, the most stringent regulations are associated with the period of
+strong decreases in new infection between late December 2020 and mid-February 2021. The
+measures were not enough, however, to stop the B.1.1.7 variant from spreading in the
+subsequent period. The steep drop in recorded cases during May 2021 is associated with
+at least weekly rapid tests to around 42~percent of the population, a vaccination rate
+that rose from 28\% to 43\%, and mostly seasonality impacting a fall in the relative
+infectiousness of contacts outside of work and school.
 
-Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid
-tests, vaccinations, and of seasonality during 2021, assuming NPIs to have
-evolved the same way as in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line,
-same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the three
-factors (red line), and three scenarios turning these factors off one by one.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
-infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs
-Shapley values to decompose the difference in total infections between the
-scenario without any of the three factors and our main specification.
+Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
+vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way
+as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model
+fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any
+of the three factors (red line), and three scenarios turning these factors off one by
+one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections
+in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to
+decompose the difference in total infections between the scenario without any of the
+three factors and our main specification.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/full_decomposition}
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_area}
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_bar}
 
@@
 reached only around mid May and it would be June for the comparison with the
-seasonality-only scenario. The effect on total cases, however, kicks in immediately and
-strongly. Despite the fact that only a small fraction of the population performed weekly
-rapid tests in March (X\%\comment[id=HM]{Put in}), the rise in new infections would be
-limited by XXX\% relative to the scenario without vaccinations, tests, or seasonality.
-
-Tests
+seasonality-only scenario. The effect on total cases, however, is visible immediately.
+Despite the fact that only a small fraction of the population performed weekly rapid
+tests in March\footnote{X\%\comment[id=HM]{Put in}in the model; 27\% of respondents to
+the COSMO study reported in early March that they \textit{ever} did a rapid test)}, new
+infections on April 1 would be rise in new infections would be reduced by 53\% relative
+to the scenario without vaccinations, rapid tests, or seasonality.
 
@@
 
-Two particularly important areas:
+Two of the most contentious NPIs concern schools and mandates to work from home. In many
+countries, schools switched to remote instruction during the first wave, so did Germany. They were kept fully open until 
+ Two
+particularly important areas:
 \begin{itemize}
@@
           but infection rates lower \comment[id=HM]{Janos, please fill up}
-    \item Actually performed tests increase until May, stay constant during May and then start to drop
+    \item Actually performed tests increase until May, stay constant during May and then
+    start to drop
     \item This is the combination of the following things:
@@
               \item monotonically increasing availability of tests
-        \item endogenous number of situations in which a test should be taken (mainly positive tests of household members)
+              \item endogenous number of situations in which a test should be taken
+              (mainly positive tests of household members)
           \end{itemize}
-    \item This is a classic way of modeling in economics: Preferences stay constant
-          (and can thus be elicited in just one survey) but outcomes change because
+    \item This is a classic way of modeling in economics: Preferences stay constant (and
+          can thus be elicited in just one survey) but outcomes change because
           circumstances change.
@@
 performed test is determined endogenously. The number of weekly tests increases until
-the end of May and then starts to drop. This is because many  It is determined by the availability of tests,
-the individual compliance with testing recommendations and the occurrence of situations
-in which a test should be taken.
+the end of May and then starts to drop. This is because many  It is determined by the
+availability of tests, the individual compliance with testing recommendations and the
+occurrence of situations in which a test should be taken.
 

### 2021-06-09 paper/report.tex
Add decomposition plots.

--- 2021-06-09 paper/report.tex
@@
 % \clearpage
+
+\begin{figure}
+\centering
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
+        \caption{Decomposition of Rapid Tests}
+\end{figure}
+\begin{figure}
+\centering
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
+        \caption{Decomposition of Rapid Tests}
+\end{figure}
 

### 2021-06-09 paper/report.tex
Position note outside footnote...

--- 2021-06-09 paper/report.tex
@@
 Despite the fact that only a small fraction of the population performed weekly rapid
-tests in March\footnote{X\%\comment[id=HM]{Put in}in the model; 27\% of respondents to
-the COSMO study reported in early March that they \textit{ever} did a rapid test)}, new
-infections on April 1 would be rise in new infections would be reduced by 53\% relative
-to the scenario without vaccinations, rapid tests, or seasonality.
+tests in March\footnote{X\% in the model; 27\% of respondents to the COSMO study
+reported in early March that they \textit{ever} did a rapid test)}\comment[id=HM]{Put
+in}, new infections on April 1 would be rise in new infections would be reduced by 53\%
+relative to the scenario without vaccinations, rapid tests, or seasonality.
 

### 2021-06-09 paper/report.tex

@@
         \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
-        \caption{{\small Model description}}
+        \caption{{Model description}}
         \label{fig:broad_model_description}
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
-        \caption{{\small Model for antigen tests}}
+        \caption{{Model for antigen tests}}
         \label{fig:antigen_tests}
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
-        \caption{{\small Model for undetected cases}}
+        \caption{{Model for undetected cases}}
         \label{fig:model_for_official_cases}
@@
         \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{\small Recorded cases: Empirical and simulated}}
+        \caption{{Recorded cases: Empirical and simulated}}
         \label{fig:aggregated_fit}
@@
 
-        \caption{{\small Stringency of NPIs and changes in infectious contacts by type}}
+        \caption{{Stringency of NPIs and infectious contacts}}
         \label{fig:stringency_index}
@@
 
-        \caption{{\small Tests and vaccinations}}
+        \caption{{Tests and vaccinations}}
         \label{fig:antigen_tests_vaccinations}
@@
 
-        \caption{Fraction of B.1.1.7 strain among measured infections}
+        \caption{Fraction of B.1.1.7 strain}
         \label{fig:share_b117}
@@
         \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
-        \caption{{\small Recorded cases: 2021 scenarios}}
+        \caption{{Recorded cases: 2021 scenarios}}
         \label{fig:2021_scenarios_recorded}
@@
         \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
-        \caption{{\small Total cases: 2021 scenarios}}
+        \caption{{Total cases: 2021 scenarios}}
         \label{fig:2021_scenarios_newly_infected}
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_rapid_tests/full_newly_infected}
-        \caption{{\small Effects of different types of testing}}
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
+        \caption{{Effects of different types of testing}}
         \label{fig:2021_scenarios_decomposition_tests}
@@
 mechanisms is turned on. Compared to the scenario with vaccinations only, this point is
-reached only around mid May and it would be June for the comparison with the
-seasonality-only scenario. The effect on total cases, however, is visible immediately.
-Despite the fact that only a small fraction of the population performed weekly rapid
-tests in March\footnote{X\% in the model; 27\% of respondents to the COSMO study
-reported in early March that they \textit{ever} did a rapid test)}\comment[id=HM]{Put
-in}, new infections on April 1 would be rise in new infections would be reduced by 53\%
-relative to the scenario without vaccinations, rapid tests, or seasonality.
-
-\clearpage
+reached around mid-May and it would be June for the comparison with the seasonality-only
+scenario. The effect on total cases, however, is visible immediately. Despite the fact
+that only a small fraction of the population performed weekly rapid tests in
+March\footnote{X\% in the model; 27\% of respondents to the COSMO study reported in
+early March that they \textit{ever} did a rapid test)}\comment[id=HM]{Put in}, new
+infections on April 1 would be rise in new infections would be reduced by 53\% relative
+to the scenario without vaccinations, rapid tests, or seasonality.
 
 Two of the most contentious NPIs concern schools and mandates to work from home. In many
-countries, schools switched to remote instruction during the first wave, so did Germany. They were kept fully open until 
- Two
-particularly important areas:
+countries, schools switched to remote instruction during the first wave, so did Germany.
+They were kept fully open until Two particularly important areas:
 \begin{itemize}
@@
         \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-        \caption{{\small Effects of different schooling scenarios after Easter}}
+        \caption{{Effects of different schooling scenarios after Easter}}
         \label{fig:schooling_scenarios_easter}
@@
         % \textwidth]{../figures/results/figures/scenario_comparisons/}
-        \caption{{\small Effects of different work scenarios after Easter}}
+        \caption{{Effects of different work scenarios after Easter}}
         \label{fig:to_be_determined}

### 2021-06-09 paper/report.tex
Remove decomposition plots in the beginning.
Fixed some typos and added a comment.

@@
 % \clearpage
-
-\begin{figure}
-\centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
-        \caption{Decomposition of Rapid Tests}
-\end{figure}
-\begin{figure}
-\centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
-        \caption{Decomposition of Rapid Tests}
-\end{figure}
 
@@
 of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not transmit the
-virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is lower
-than what is usually reported for after the second dose of the Biontech/Pfizer vaccine,
-which is most commonly used in Germany. We choose it because our model neither includes
-booster shots, nor does it allow vaccinated individuals who became immune to transmit
-the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.} During the vaccine
-roll-out, priority may depend on age and occupation.
+virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\comment[id=T]{Maybe add
+that 35\% reach immunity after two weeks and 40\% after three weeks.}\footnote{75\% is
+lower than what is usually reported for after the second dose of the Biontech/Pfizer
+vaccine, which is most commonly used in Germany. We choose it because our model neither
+includes booster shots, nor does it allow vaccinated individuals who became immune to
+transmit the disease \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.} During
+the vaccine roll-out, priority may depend on age and occupation.
 
@@
 yield immediate results. Specificity and sensitivity of these tests is set according to
-data analysed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of
+data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of
 the test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows
@@
 (e.g., aggregate PCR test capacity was much lower and we would require a very different
-model for calculating the share of known cases) Second, a large fraction cases during
-summer of 2020 were traced to international travel \citep{KochInstitut2021,Hodcroft2021}
-but the precise number is difficult to model.
+model for calculating the share of known cases). Second, a large fraction of cases
+during summer of 2020 were traced to international travel
+\citep{KochInstitut2021,Hodcroft2021} but the precise number is difficult to model.
 

### 2021-06-09 paper/report.tex
Done for the night. Lacking last couple of paragraphs, abstract, description undetected cases, small stuff.

@@
 
-% \clearpage
-
 Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
@@
 \citep{Swaminathan2021}.\comment[id=K]{WHO Press Conference saying we won't reach herd
-immunity in 2021} In the meantime, it is of utmost importance to employ an effective mix
-of strategies for containing the virus. The most frequent initial response was a set of
-non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
-this has allowed some countries to sustain equilibria with very low infection
-numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
-numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
-fluctuations of infection rates. Containment measures have become increasingly diverse
-and now include testing, more nuanced NPIs, and contact tracing. Neither these policies'
-effect nor the influence of seasonal patterns or more infectious virus strains are well
-understood in quantitative terms. This paper develops a model incorporating all these
-factors. The framework allows to combine a wide variety of data in a timely fashion,
-making it useful to predict the effects of various interventions. We apply the model to
-Germany and show that rapid testing had the largest impact on the reduction in
-infections by almost 80\% during the month of May 2021. We conclude that rapid tests
-have a large role to play at least as long as vaccinations have not been offered to an
-entire population.
+immunity in 2021}\comment[id=HM]{If we find a paper (s)he cites, would be even better}
+In the meantime, it is of utmost importance to employ an effective mix of strategies for
+containing the virus. The most frequent initial response was a set of non-pharmaceutical
+interventions (NPIs) to reduce contacts between individuals. While this has allowed some
+countries to sustain equilibria with very low infection numbers\footnote{See
+    \citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
+    sustained with test-trace-and-isolate policies.}, most have seen large fluctuations of
+infection rates. Containment measures have become increasingly diverse and now include
+testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
+influence of seasonal patterns or more infectious virus strains are well understood in
+quantitative terms. This paper develops a model incorporating all these factors. The
+framework allows to combine a wide variety of data in a timely fashion, making it useful
+to predict the effects of various interventions. We apply the model to Germany and show
+that rapid testing had the largest impact on the reduction in infections by almost 80\%
+during the month of May 2021. We conclude that rapid tests have a large role to play at
+least as long as vaccinations have not been offered to an entire population.
 
@@
     includes booster shots, nor does it allow vaccinated individuals who became immune to
-transmit the disease \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.} During
-the vaccine roll-out, priority may depend on age and occupation.
+    transmit the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}. If
+    anything, these assumptions would overstate the effect of vaccines for our study period.
+    This would be different if a large fraction of vaccinated individuals had received a
+    second dose already.} During the vaccine roll-out, priority may depend on age and
+occupation.
 
@@
         \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_area}
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_bar}
-
-        \vskip2ex
-
-        \caption{Decomposition of effects for
+        % \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_bar}
+        \caption{Decomposition of the difference between the scenario without any of the
+            three factors and the main scenario in
             Figure~\ref{fig:2021_scenarios_newly_infected}.}
@@
         \centering
-
         \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
-        \caption{{Effects of different types of testing}}
+        % \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
+        \caption{Decomposition of the difference between the scenario without rapid
+            tests and the main scenario in Figure~\ref{fig:2021_scenarios_newly_infected}.}
         \label{fig:2021_scenarios_decomposition_tests}
@@
     early March that they \textit{ever} did a rapid test)}\comment[id=HM]{Put in}, new
-infections on April 1 would be rise in new infections would be reduced by 53\% relative
-to the scenario without vaccinations, rapid tests, or seasonality.
+infections on April 1 would be reduced by 53\% relative to the scenario without
+vaccinations, rapid tests, or seasonality.
+
+So why is rapid testing so effective? In order to shed more light on this question,
+Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
+scenario without rapid tests only (purple line in
+Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
+three channels for rapid tests. Tests for pupils have the smallest effect, which is
+largely explained by the relatively small number of students (XXM vs. YYM workers) and
+the fact that schools did not operate at full capacity during our period of study.
+Almost 40\% come from tests at the workplace. Despite the fact that we phase rapid tests
+for private reasons are phased in only late, they make up for more than half of the total effect.
+
+\comment[id=HM]{How many tests are actually performed along each of the three dimensions?
+Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}
+
+The reason lies in the fact that a substantial share of these tests is driven by an
+elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
+up on a positive test of a household member. The latter is essentially a form of contact
+tracing, which has been shown to be very effective.\comment[id=HM]{cite Priesemann
+paper, others?} This is also the reason for why the number of tests performed fall at
+the very end of our study period. Falling infection rates mean that there are fewer
+events triggering private test demand.
+\comment[id=HM]{Positive rates along the three channels might be nice here, if previous comment is not a quant story}
+\comment[id=HM]{I think there is a story in here about risk conditional on past exposure vs.
+prospective risk in gatherings, but it is too late to work through it.
+We might just want to calculate the probabilities.}
 
@@
 countries, schools switched to remote instruction during the first wave, so did Germany.
-They were kept fully open until Two particularly important areas:
+After the summer break, they were operating at full capacity with increased hygiene
+measures, before being closed again from mid-December on. Some states started opening
+them gradually in late February, but usual operaton was not back until the beginning of
+June. In light of the large negative effects school closures have on children and
+parents\comment[id=HM]{need to cite a couple of poor learning outcomes / mental health
+papers. There must be well-published ones}---and in particular on those with low
+socio-economic status---
+
+Effects:
 \begin{itemize}
-    \item Schooling: large costs to pupils, ... social inequalities
-    \item Work: Q2 / 2020 saw the largest drop in GDP in a long time;
+    \item Schools were essentially closed
+    \item What are the different $R_t$-values? In no-test scenario and $R_t>1$, might be
+          worth chopping off 0.05 or whatever that might be?
+    \item With tests it is really hard to make a case for that
+\end{itemize}
+
+Alternative:
+\begin{itemize}
+    \item Germany quite lenient on home office relative to other countries (employer had
+          to allow)
+    \item Testing: Rapid tests are cheap (less than 1\euro), just a nuisance.
 \end{itemize}
@@
           stuff. Certain publicness.
-    \item Mandatory tests: Screening effect: ...
+    \item Mandatory tests: Screening effect for hard-to-reach populations (do not
+          consider in this paper, also due to poor data in DE)
 \end{itemize}
@@
 \end{figure}
-
-\paragraph{Points to mention}
-\begin{itemize}
-    \item If anything too optimistic regarding vaccinations
-    \item Social structure / conditional block testing in families important (?)
-    \item Trump-effect: More testing = more cases true for how long?
-    \item Towards the end, total tests go down: Preferences for tests remain the same,
-          but infection rates lower \comment[id=HM]{Janos, please fill up}
-    \item Actually performed tests increase until May, stay constant during May and then
-    start to drop
-    \item This is the combination of the following things:
-          \begin{itemize}
-              \item time constant willingness to take tests
-              \item monotonically increasing availability of tests
-              \item endogenous number of situations in which a test should be taken
-              (mainly positive tests of household members)
-          \end{itemize}
-    \item This is a classic way of modeling in economics: Preferences stay constant (and
-          can thus be elicited in just one survey) but outcomes change because
-          circumstances change.
-
-\end{itemize}
-
-An important feature of the way we model rapid tests is that the number of actually
-performed test is determined endogenously. The number of weekly tests increases until
-the end of May and then starts to drop. This is because many  It is determined by the
-availability of tests, the individual compliance with testing recommendations and the
-occurrence of situations in which a test should be taken.
-
-% Technical terms should be defined.
-
-% Symbols, abbreviations, and acronyms should be defined the first time they are used.
-
-% All tables and figures should be cited in numerical order.
-
-% All data must be shown either in the main text or in the Supplementary Materials or
-% must be available in an established database with accession details provided in the
-% acknowledgements section.
-
-% References to unpublished materials are not allowed to substantiate significant
-% conclusions of the paper.

### 2021-06-09 paper/appendix_data.tex
Add a few points about testing.
Some cleaning up of manual comments to actual ones, formatting.

@@
 
-The model is described by a large number of parameters that govern the number of contacts
-a person has, the likelihood of becoming infected on each contact, the likelihood of
-developing light or strong symptoms or even dying from the disease as well as the
-duration each stage of the disease takes.
+The model is described by a large number of parameters that govern the number of
+contacts a person has, the likelihood of becoming infected on each contact, the
+likelihood of developing light or strong symptoms or even dying from the disease as well
+as the duration each stage of the disease takes.
 
@@
 
-We calibrate the parameters for the predicted numbers of contacts from contact diaries of
-over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
+We calibrate the parameters for the predicted numbers of contacts from contact diaries
+of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
 \citep{Mossong2008}. Each contact diary contains all contacts an individual had
@@
 and information on the contact. Importantly, for each contact individuals entered of
-which type the contact (school, leisure, work etc.) was and how frequent the contact with
-the other person is.
+which type the contact (school, leisure, work etc.) was and how frequent the contact
+with the other person is.
 
@@
 geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same data
-we use to calibrate the number of contacts.
+within the same county, we can calibrate the assortative mixing by age from the same
+data we use to calibrate the number of contacts.
 
@@
     \label{fig:assortativity}
-    \figurenotes{
-        The figure shows the distribution of random non-work contacts by age groups. A
-        row shows the share of contacts a certain age group has with all other age
-        groups. Higher values are colored in darker red tones. The diagonal represents
-        the share of contacts with individuals from the same age group.
-    }
+    \figurenotes{The figure shows the distribution of random non-work contacts by age
+        groups. A row shows the share of contacts a certain age group has with all other
+        age groups. Higher values are colored in darker red tones. The diagonal
+        represents the share of contacts with individuals from the same age group.}
 \end{figure}
 
-Figure~\ref{fig:assortativity}\comment[id=HM]{So no work in here?
-How about school?
-Can we add a column with marginals and another one with work / (school?), too?} shows that assortativity by age is especially strong for
-children and younger adults. For older people, the pattern becomes more dispersed around
-their own age group, but within-age-group contacts are still the most common contacts.
+Figure~\ref{fig:assortativity}\comment[id=HM]{So no work in here? How about school? Can
+we add a column with marginals and another one with work / (school?), too?} shows that
+assortativity by age is especially strong for children and younger adults. For older
+people, the pattern becomes more dispersed around their own age group, but
+within-age-group contacts are still the most common contacts.
 
@@
 %
-% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377:
-% In allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
+% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377: In
+% allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
 % Schüler in die Schule.
@@
 For our work models\footnotemark we use the reductions in work mobility reported in the
-Google Mobility Data \citep{Google2021} to calibrate our work policies.
-Reductions in work contacts are not random but governed through a work contact priority
-where the policy changes the threshold below which workers stay home.
-Figure \ref{fig:work_multiplier} shows the share of workers that go to work in our model
-over time.
+Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
+work contacts are not random but governed through a work contact priority where the
+policy changes the threshold below which workers stay home. Figure
+\ref{fig:work_multiplier} shows the share of workers that go to work in our model over
+time.
 
@@
     \figurenotes{The figure shows the work mobility as reported by \cite{Google2021}. We
-    take this as a proxy of the share of workers who are not in home office, i.e. who
-    still have physical work contacts. The figure interpolates over weekends as we handle
-    weekend effects through information on work on weekends in the German census data we
-    use.}
+        take this as a proxy of the share of workers who are not in home office, i.e.
+        who still have physical work contacts. The figure interpolates over weekends as
+        we handle weekend effects through information on work on weekends in the German
+        census data we use.}
 \end{figure}
@@
 For the last group of contacts which cover things like leisure activities, grocery
-shopping etc. we have no reliable data by how much policies reduce them.
-In addition, they are likely to be affected by social and psychological factors such as
-pandemic fatigue and vacations. Because of this we estimate them like the infection
-probabilities to fit the time series data. We use very few change points and tie them to
-particular events such as policy announcements or particular holidays.
+shopping etc. we have no reliable data by how much policies reduce them. In addition,
+they are likely to be affected by social and psychological factors such as pandemic
+fatigue and vacations. Because of this we estimate them like the infection probabilities
+to fit the time series data. We use very few change points and tie them to particular
+events such as policy announcements or particular holidays.
 
@@
     \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
-\caption{\textbf{Share of Individuals that do a Rapid Test.} \textcolor{red}{Talk about the interpretation of each line.}}
+    \caption{\textbf{Share of Individuals that do a Rapid Test.} \comment[id=J]{Talk
+        about the interpretation of each line.}}
+    \label{fig:rapid_test_demand}
 \end{figure}
@@
     \item someone is an employees of an educational facility or school pupil
-    \item someone lives in a household where someone has tested positive or developed symptoms
+    \item someone lives in a household where someone has tested positive or developed
+          symptoms
     \item someone has developed symptoms but has not received a PCR test
@@
 
-For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April) that 60\% of workers who receive a test offer by their employer regularly use it.
-We assume this to be time constant.
+For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April)
+that 60\% of workers who receive a test offer by their employer regularly use it. We
+assume this to be time constant.
 
-In addition, there are some surveys that allow us to trace the expansion of employers who
-offer tests to their employees. Mid march, 20\% of employers offered tests to their
-employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
-offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
-60\% until the first days of April \cite{ZDF2021}. Until mid April 70\% of workers were
-expected to receive a weekly test offer \citep{AerzteZeitung2021}. However, according to
-surveys conducted in mid April \citep{Betsch2021}, less than two thirds of individuals
-with work contacts receive a test offer. Starting on April 19th employers were required by law to provide two weekly tests to their employees \citep{Bundesanzeiger2021}.
-We assume that compliance is incomplete and only 80\% of employers actually offer tests.
+In addition, there are some surveys that allow us to trace the expansion of employers
+who offer tests to their employees. Mid march, 20\% of employers offered tests to their
+employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported
+being offered weekly rapid tests by their employer \citep{Ahlers2021}. This share
+increased to 60\% until the first days of April \cite{ZDF2021}. Until mid April 70\% of
+workers were expected to receive a weekly test offer \citep{AerzteZeitung2021}. However,
+according to surveys conducted in mid April \citep{Betsch2021}, less than two thirds of
+individuals with work contacts receive a test offer. Starting on April 19th employers
+were required by law to provide two weekly tests to their employees
+\citep{Bundesanzeiger2021}. We assume that compliance is incomplete and only 80\% of
+employers actually offer tests.
 
@@
 
-\textcolor{red}{Sources still missing below this}
+\comment[id=J]{Sources still missing below this}
 
 We assume that employees in educational facilities start getting tested in 2021 and that
-by March 1st 30\% of them are tested weekly.
-The share increases to 90\% for the week before Easter.
-At that time both Bavaria and Baden-Württemberg were offering tests to teachers and North-Rhine Westphalia and Lower Saxony were already testing students and tests for students and teachers were already mandatory in Saxony.
-After Easter we assume that 95\% of teachers get tested twice per week.
+by March 1st 30\% of them are tested weekly. The share increases to 90\% for the week
+before Easter. At that time both Bavaria and Baden-Württemberg were offering tests to
+teachers and North-Rhine Westphalia and Lower Saxony were already testing students and
+tests for students and teachers were already mandatory in Saxony. After Easter we assume
+that 95\% of teachers get tested twice per week.
 
@@
 
-After Easter we assume the share of students receiving twice weekly tests to
-based on tests already being mandatory in North-Rhine Westphalia and Bavaria while still
-being voluntary in Baden-Württemberg. There tests become mandatory on April 19th.
+After Easter we assume the share of students receiving twice weekly tests to based on
+tests already being mandatory in North-Rhine Westphalia and Bavaria while still being
+voluntary in Baden-Württemberg. There tests become mandatory on April 19th.
 
@@
 
-\textcolor{red}{
-    Add a section on how we calibrate rapid test demand; Mainly describe the
-    datapoints we have and say that we usually interpolate linearly in between data
-    points. (Only exception to that is private rapid test demand, which we fit to data)
-}
+\comment[id=J]{Add a section on how we calibrate rapid test demand; Mainly describe the
+datapoints we have and say that we usually interpolate linearly in between data points.
+(Only exception to that is private rapid test demand, which we fit to data)}
 
@@
     \label{fig:share_rapid_test_last_week}
-    \figurenotes{The figure compares the share of individuals who have ever done a rapid test or done a rapid test within the last week in our simulations to the shares reported in the \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19 Snapshot Monitoring survey}.
-The
-    left panel compares the share of individuals who have ever done a rapid test.
-The right panel compares the share of individuals who have done a rapid test within the last seven days in our simulation compared to the share reporting to have done at least weekly rapid tests in the last four weeks in the COSMO survey.
-Overall our calibration of rapid tests are slightly conservative.
-The overall share is below that in the study.
-We fit the share of weekly tests quite exactly.
-However, the study only covers adults while our share also includes children who are tested very regularly when attending school.
-    }
+    \figurenotes{The figure compares the share of individuals who have ever done a rapid
+        test or done a rapid test within the last week in our simulations to the shares
+        reported in the
+        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
+            Snapshot Monitoring survey}. The left panel compares the share of individuals
+        who have ever done a rapid test. The right panel compares the share of
+        individuals who have done a rapid test within the last seven days in our
+        simulation compared to the share reporting to have done at least weekly rapid
+        tests in the last four weeks in the COSMO survey. Overall our calibration of
+        rapid tests are slightly conservative. The overall share is below that in the
+        study. We fit the share of weekly tests quite exactly. However, the study only
+        covers adults while our share also includes children who are tested very
+        regularly when attending school.}
 \end{figure}

### 2021-06-09 paper/appendix_results.tex
Some cleaning up of manual comments to actual ones, formatting.

@@
 
-\textcolor{red}{summarize the fit; Some of this needs to go in the appendix. I suggest the age group and federal state fit goes there???}
+\comment[id=J]{summarize the fit; Some of this needs to go in the appendix. I suggest the age group and federal state fit goes there???}
 
@@
 
-\textcolor{red}{The results can be found in `figures/results`. This includes both figures
+\comment[id=K]{The results can be found in `figures/results`. This includes both figures
 and tables for lookup of numbers and summary tables.}
@@
   \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-\figurenotes{\textcolor{red}{K: One of the scenarios starts too early. Will be fixed with the next full simulations run.}}
   \label{fig:school_scenarios}

### 2021-06-10 paper/model/testing.tex
Add tasks in red on testing in appendix.

@@
 
+\textcolor{red}{
+
+To Do für Appendix an der Stelle: Institutional details.
+
+\begin{itemize}
+    \item PCR tests only at doctor's office / test centers.
+    \item Fraction PCR tests performed because of contact with infected people is small (quarantine mandate more common) -- need citation
+    \item Rapid tests available in many pop up test centers, self-administered tests in every  drug store / supermarket, prices in early June < 1 Euro
+\end{itemize}
+}
+
 Rapid tests are modeled much simpler. Every day before individuals have contacts they can

### 2021-06-10 paper/model/testing.tex
Fix latex errors.

--- 2021-06-04 paper/model/testing.tex
@@
 
+\textcolor{red}{To Do für Appendix an der Stelle: Institutional details:}
+
+\begin{itemize}
+    \item \textcolor{red}{PCR tests only at doctor's office / test centers.}
+    \item \textcolor{red}{Fraction PCR tests performed because of contact with infected
+    people is small (quarantine mandate more common) -- need citation}
+    \item \textcolor{red}{Rapid tests available in many pop up test centers,
+    self-administered tests in every  drug store / supermarket, prices in early June < 1
+    Euro}
+\end{itemize}
+
 Rapid tests are modeled much simpler. Every day before individuals have contacts they can
-decide to be tested.\comment[id=HM]{If we get econ readers, we should not use language like this because we do not have a choice model.} For example students that plan to attend school that day and have
-not done a rapid test in the last three days get a rapid test. Then they immediately
-receive the test result. After they have received their test result individuals can react
-to it by reducing their contacts. For example positively tested individuals may not go to
-work and reduce their household contacts to some degree. Who reduces their contacts to
-what degree depends on a quarantine compliance attribute.
+decide to be tested.\comment[id=HM]{If we get econ readers, we should not use language
+like this because we do not have a choice model.} For example students that plan to
+attend school that day and have not done a rapid test in the last three days get a rapid
+test. Then they immediately receive the test result. After they have received their test
+result individuals can react to it by reducing their contacts. For example positively
+tested individuals may not go to work and reduce their household contacts to some degree.
+Who reduces their contacts to what degree depends on a quarantine compliance attribute.
 
-Our rapid tests include false positives and false negatives. The sensitivity of rapid tests in our model depends on when the individual has
-or will become infectious. This way we can account for the fact that rapid tests are
-likely to be false positive before infectiousness starts.
+Our rapid tests include false positives and false negatives. The sensitivity of rapid
+tests in our model depends on when the individual has or will become infectious. This way
+we can account for the fact that rapid tests are likely to be false positive before
+infectiousness starts.
 

### 2021-06-10 paper/report.tex
Add Fig 3d.
Fix latex errors.

@@
 \citep{Swaminathan2021}.\comment[id=K]{WHO Press Conference saying we won't reach herd
-immunity in 2021}\comment[id=HM]{If we find a paper (s)he cites, would be even better}
+immunity in 2021. She cites no source.}\comment[id=HM]{If we find a paper (s)he cites, would be even better}
 In the meantime, it is of utmost importance to employ an effective mix of strategies for
@@
         \centering
-        % \includegraphics[width=0.9
-        % \textwidth]{../figures/results/figures/scenario_comparisons/}
+        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
         \caption{{Effects of different work scenarios after Easter}}

### 2021-06-10 paper/medical_parameters.tex
Remove table form appendix on medical parameters.

@@
 CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
-Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of
-requiring intensive care per age group.
-
-\begin{table}[tb]
-    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
-    \label{tab:symptomatic-to-ICU}
-    \centering
-
-    \begin{tabular}{ll}
-        \toprule
-        Age Group & Share \\
-        \midrule
-        0-9 & 0.00005 \\
-        10-19 & 0.00030 \\
-        20-29 & 0.00075 \\
-        30-39 & 0.00345 \\
-        40-49 & 0.01380 \\
-        50-59 & 0.03404 \\
-        60-69 & 0.10138 \\
-        70-79 & 0.16891 \\
-        80-100 & 0.26871 \\
-        \bottomrule
-    \end{tabular}
-
-    \tablenotes{
-        The data is taken from \citet{Stokes2020} and
-        \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the
-        OpenABM-Project}. }
-
-\end{table}
+%Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of
+%requiring intensive care per age group.
+%
+%\begin{table}[tb]
+%    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
+%    \label{tab:symptomatic-to-ICU}
+%    \centering
+%
+%    \begin{tabular}{ll}
+%        \toprule
+%        Age Group & Share \\
+%        \midrule
+%        0-9 & 0.00005 \\
+%        10-19 & 0.00030 \\
+%        20-29 & 0.00075 \\
+%        30-39 & 0.00345 \\
+%        40-49 & 0.01380 \\
+%        50-59 & 0.03404 \\
+%        60-69 & 0.10138 \\
+%        70-79 & 0.16891 \\
+%        80-100 & 0.26871 \\
+%        \bottomrule
+%    \end{tabular}
+%
+%    \tablenotes{
+%        The data is taken from \citet{Stokes2020} and \href{https://bit.ly/354pKTZ}{the
+%        OpenABM-Project}. }
+%
+%\end{table}
 
@@
 \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also
-in line with the duration estimates collected by
-\href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the
+in line with the duration estimates collected by \href{https://bit.ly/3gkGtaU}{the
 Robert-Koch-Institut}. We assume that the time between symptom onset and ICU takes 4, 6,

### 2021-06-10 paper/appendix_data.tex
Add new results.
Fix latex errors.
Update the section on the number of contacts.

@@
 
-Thus, we can use the empirical distributions from this data as pre-pandemic number of
-contacts.
+Simplifying the number of contacts, we arrive at the following distributions of the
+numbers of contacts by contact type.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
+    \caption{Number of Non Recurrent Other Contacts}
+    \label{n_contacts_other_non_recurrent}
+    \floatfoot{\noindent In the model it is sampled every day which of these numbers of
+    contacts a person is planned to have. Note that the contact diaries include such high
+    values that super spreading events are well possible in our model. The planned number
+    of contacts is reduced by policies, seasonality and individual responses to events
+    such as receiving a positive rapid test to the number of actual contacts with
+    transmission potential. Other contacts include all contacts that are not household
+    members, school contacts or work contacts, for example leisure contacts or contacts
+    during grocery shopping.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
+    \caption{Number of Daily Recurrent Other Contacts}
+    \label{n_contacts_other_daily_recurrent}
+    \floatfoot{\noindent Individuals are assigned to groups that are time constant and
+    that meet daily. The share of individuals that attend in a way that has transmission
+    potential is reduced by policies, seasonality and individual responses to events such
+    as receiving a positive rapid test. Other contacts include all contacts that are not
+    household members, school contacts or work contacts, for example leisure contacts or
+    contacts during grocery shopping.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
+    \caption{Number of Weekly Recurrent Other Contacts}
+    \label{n_contacts_other_weekly_recurrent}
+    \floatfoot{\noindent Individuals are assigned to up to four groups that are time
+    constant and that meet weekly. The share of individuals that attend in a way that has
+    transmission potential is reduced by policies, seasonality and individual responses
+    to events such as receiving a positive rapid test. Other contacts include all
+    contacts that are not household members, school contacts or work contacts, for
+    example leisure contacts or contacts during grocery shopping.}
+\end{figure}
+
+% work contacts
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
+    \caption{Number of Non Recurrent Work Contacts}
+    \label{n_contacts_work_non_recurrent}
+    \floatfoot{\noindent In the model it is sampled every day which of these numbers of
+    contacts a working person is planned to have. Note that the contact diaries include
+    such high values that super spreading events are well possible in our model. The
+    planned number of contacts is reduced by policies, seasonality and individual
+    responses to events such as receiving a positive rapid test to the number of actual
+    contacts with transmission potential. Work contacts only take place between working
+    individuals.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
+    \caption{Number of Daily Recurrent Work Contacts}
+    \label{n_contacts_work_daily_recurrent}
+    \floatfoot{\noindent Working individuals are assigned to groups that are time
+    constant and that meet daily to match the given distribution of daily work contacts.
+    You can think of these as for example colleagues with which one shares an office
+    space. The share of individuals that attend in a way that has transmission potential
+    is reduced by policies (such as a work from home mandate), seasonality and individual
+    responses to events such as receiving a positive rapid test. Work contacts only take
+    place between working individuals.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
+    \caption{Number of Weekly Recurrent Work Contacts}
+    \label{n_contacts_work_weekly_recurrent}
+    \floatfoot{\noindent Working individuals are assigned to up to 14 groups that are time
+    constant and meet weekly. Groups are scheduled to meet on separate days of the work
+    week. These contact models cover weekly team meetings etc. The share of individuals
+    that attend in a way that has transmission potential is reduced by policies,
+    seasonality and individual responses to events such as receiving a positive rapid
+    test. Work contacts only take place between working individuals.}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/assortative_matching_probability_example}
-    \caption{Distribution of random non-work contacts by age of participants.}
+    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_other_non_recurrent}
+    \caption{Distribution of Random Other Contacts by Age Group}
     \label{fig:assortativity}
@@
     \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of Individuals that do a Rapid Test.} \comment[id=J]{Talk
-        about the interpretation of each line.}}
+    \caption{\textbf{Share of Individuals that do a Rapid Test.}}
+    \figurenotes{\textcolor{red}{J: Talk about the interpretation of each line.}}
     \label{fig:rapid_test_demand}

### 2021-06-10 paper/appendix_data.tex
Complete section an assortative matching.

@@
     \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_other_non_recurrent}
-    \caption{Distribution of Random Other Contacts by Age Group}
-    \label{fig:assortativity}
-    \figurenotes{The figure shows the distribution of random non-work contacts by age
-        groups. A row shows the share of contacts a certain age group has with all other
+    \caption{Distribution of Non Recurrent Other Contacts by Age Group}
+    \label{fig:assortativity_other}
+    \floatfoot{\noindent The figure shows the distribution of non recurrent other contacts by age
+        group. A row shows the share of contacts a certain age group has with all other
         age groups. Higher values are colored in darker red tones. The diagonal
@@
 
-Figure~\ref{fig:assortativity}\comment[id=HM]{So no work in here? How about school? Can
-we add a column with marginals and another one with work / (school?), too?} shows that
-assortativity by age is especially strong for children and younger adults. For older
-people, the pattern becomes more dispersed around their own age group, but
-within-age-group contacts are still the most common contacts.
+Figure~\ref{fig:assortativity_other} shows that assortativity by age is especially strong
+for children and younger adults. For older people, the pattern becomes more dispersed
+around their own age group, but within-age-group contacts are still the most common
+contacts.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_work_non_recurrent}
+    \caption{Distribution of Random Work Contacts by Age Group}
+    \label{fig:assortativity_work}
+    \floatfoot{\noindent The figure shows the distribution of non recurrent work contacts by age
+        group. A row shows the share of contacts a certain age group has with all other
+        age groups. Higher values are colored in darker red tones. The diagonal
+        represents the share of contacts with individuals from the same age group. We
+        only show age groups that have a significant fraction of working individuals.}
+\end{figure}
+
+Figure~\ref{fig:assortativity_work} shows that assortativity by age is also important
+among work contacts.
+
+Our other two types of contacts, households and schools, get their assortativity by
+construction. Schools are groups where the same children of the mostly same age group and
+county meet with teachers every day. Household composition follows directly from the
+German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
+no work in here? How about school? Can we add a column with marginals and another one
+with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
+households addresses all questions. I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}
 
@@
     \label{fig:work_multiplier}
-    \figurenotes{The figure shows the work mobility as reported by \cite{Google2021}. We
+    \floatfoot{\noindent The figure shows the work mobility as reported by \cite{Google2021}. We
         take this as a proxy of the share of workers who are not in home office, i.e.
@@
     \caption{\textbf{Share of Individuals that do a Rapid Test.}}
-    \figurenotes{\textcolor{red}{J: Talk about the interpretation of each line.}}
+    \floatfoot{\noindent \textcolor{red}{J: Talk about the interpretation of each line.}}
     \label{fig:rapid_test_demand}
@@
     \label{fig:share_rapid_test_last_week}
-    \figurenotes{The figure compares the share of individuals who have ever done a rapid
+    \floatfoot{\noindent The figure compares the share of individuals who have ever done a rapid
         test or done a rapid test within the last week in our simulations to the shares

### 2021-06-10 paper/appendix_data.tex
Update the section on infection probabilities.

@@
 the exact duration and distance of each contact type as well as viral loads. Since this
-is not available in any dataset, we estimate those parameters inside the model with the
+is not available in any data set, we estimate those parameters inside the model with the
 method of simulated moments \citep{McFadden1989} by minimizing the distance between
@@
 
-Currently, we use data for Germany from August until November. We do not use earlier
-periods to save computational time. Moreover, we would be worried that the there are
-seasonal effects that we currently do not model.
+Currently, we use data for Germany from October 2020 until June 2021. We do not use
+earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
+limited and therefore it would be difficult to find good initial conditions for our
+simulations. In addition during the summer the case numbers were extremely low. This
+could lead to the epidemic going extinct in our simulation. Additionally, our model does
+not include international travel or other imports of cases. These would be important but
+difficult to model during the summer months.
 
-To avoid overfitting and simplify the numerical optimization problem, we only allow for
-four different probabilities: 1) for contacts in schools, preschools and nurseries. 2)
-for work contacts. 3) for households. 4) for leisure activities.
+To avoid over-fitting and simplify the numerical optimization problem, we only allow for
+five different probabilities: 1) for contacts in schools 2) contacts in preschools and
+nurseries. 3) for work contacts. 4) for households. 5) for other contacts.
+
+The infection probabilities estimated from our model are as follows: For household
+contacts (each individual meets all his or her household members every day) the infection
+probability is 10\%, for school contacts (all students and teachers meet all other
+students and teachers that attend school that day and there are three classes taking
+place each day) the probability of infection is 1.2\%, for preschools and nurseries that
+infection probability is 0.5\%. Work contacts carry a risk of 14.5\% and other activities
+(including leisure activities such as meeting friends) carry the highest risk with
+15.875\%.
 

### 2021-06-10 paper/appendix_data.tex
Incorporated first batch of comments from class.
Update policies section.

@@
     members, school contacts or work contacts, for example leisure contacts or contacts
-    during grocery shopping.}
+    during grocery shopping. We assume that individuals in households with children or
+    teachers or retired individuals have additional contacts during school vacations to
+    cover things like family visits or travel during vacations. We estimate this to be on
+    average 0.5 additional contacts per vacation day.}
 \end{figure}
@@
 preschools and schools we implement vacations as announced by the German federal states
-as well as school closures. For the moment we ignore both emergency childcare and that
-lack of childcare leads working parents to stay home.
+as well as school closures, emergency care and A / B schooling where only one half of
+students attends every other week or day. For the moment we ignore that lack of childcare
+leads working parents to stay home.
 %
@@
 %
-For our work models\footnotemark we use the reductions in work mobility reported in the
-Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
+For our work models\footnote{We distinguish non-recurrent work contacts, daily work
+contacts and weekly work contacts.} we use the reductions in work mobility reported in
+the Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
 work contacts are not random but governed through a work contact priority where the
@@
     \floatfoot{\noindent The figure shows the work mobility as reported by \cite{Google2021}. We
-        take this as a proxy of the share of workers who are not in home office, i.e.
-        who still have physical work contacts. The figure interpolates over weekends as
-        we handle weekend effects through information on work on weekends in the German
-        census data we use.}
+        take this as a proxy of the share of workers who are not in home office, i.e. who
+        still have physical work contacts. The figure interpolates over weekends as we
+        handle weekend effects through information on work on weekends in the German
+        census data we use. The figure shows the share aggregated over Germany as a
+        whole. To capture the effect that local policies, school vacations and public
+        policies have on work contacts we use the data by German state to determine which
+        workers go to work depending on the state they live in.}
 \end{figure}
-
-\footnotetext{We distinguish non-recurrent work contacts, daily work contacts and weekly work contacts.}
 

### 2021-06-10 paper/report.tex

@@
 influence of seasonal patterns or more infectious virus strains are well understood in
-quantitative terms. This paper develops a model incorporating all these factors. The
-framework allows to combine a wide variety of data in a timely fashion, making it useful
-to predict the effects of various interventions. We apply the model to Germany and show
-that rapid testing had the largest impact on the reduction in infections by almost 80\%
-during the month of May 2021. We conclude that rapid tests have a large role to play at
-least as long as vaccinations have not been offered to an entire population.
+quantitative terms.
+
+This paper develops a quantitative model incorporating all these factors. The framework
+allows to combine a wide variety of data and mechanisms in a timely fashion, making it
+useful to predict the effects of various interventions. We apply the model to Germany
+and show that rapid testing had the largest impact on the reduction in infections by
+almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
+to play at least as long as vaccinations have not been offered to an entire population.
 
@@
 pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to
-    multiple data sets on contact types and frequencies at
-    \url{http://www.socialcontactdata.org/} covering countries from all continents except
-    North America and Australia.} The dimensions are chosen so that the most common NPIs can
-be modeled in great detail by reducing the number of contacts in a particular setting or
-the risk of transmitting the disease for a type of contact. For example, a mandate to
-work from home will reduce the number of work contacts to zero for a fraction of the
-working population. Schools and daycare can be closed entirely, operate at reduced
-capacity---including an alternating schedule---,or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
-of contacts in non-work/non-school settings. In any setting, measures like masking
-requirements would reduce the probability of infection associated with a contact.
+multiple data sets on contact types and frequencies covering countries from four
+continents.}
+
+The three dimensions of contact types are chosen so that the most common NPIs can be
+modeled in great detail. NPIs affect the number of contacts or the risk of transmitting
+the disease upon having physical contact. The effect of different NPIs will generally
+vary across contact types. For example, a mandate to work from home will reduce the
+number of work contacts to zero for a fraction of the working population. Schools and
+daycare can be closed entirely, operate at reduced capacity---including an alternating
+schedule---,or implement mitigation measures like masking requirements or air filters
+\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
+settings. In any setting, measures like masking requirements would reduce the
+probability of infection associated with a contact.
 
@@
 whether an individual is infected or not. PCR tests require some days to be processed
-and there are always aggregate capacity constraints. In contrast, rapid antigen tests
-yield immediate results. Specificity and sensitivity of these tests is set according to
-data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of
-the test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows
-our model for rapid test demand. Schools may require students to be tested regularly.
-Rapid tests may be offered by employers for on-site workers. Individuals may demand
-tests for private reasons, which include having plans to meet other people\footnote{A
-    positive test will make them reduce their contacts; this is why tests impact the actual
-    contacts in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and
-because a household member tested positively for the virus. We endow each agent with an
-individual compliance parameter. This parameter determines whether she takes up rapid
-tests offered by employers or follows up on private reasons. The thresholds are lower
-for tests in a private setting than for tests at the workplace.\comment[id=HM]{True?
-@Janos}
+and there are aggregate capacity constraints throughout. In contrast, rapid antigen
+tests yield immediate results; after a phase-in period, all tests that are demanded will
+be performed. Specificity and sensitivity of these tests is set according to data
+analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the
+test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our
+model for rapid test demand. Schools may require students to be tested regularly. Rapid
+tests may be offered by employers for on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people\footnote{A positive
+test will make them reduce their contacts; this is why tests impact the actual contacts
+in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
+household member tested positively for the virus. We endow each agent with an individual
+compliance parameter. This parameter determines whether she takes up rapid tests offered
+by employers or follows up on private reasons. The thresholds are lower for tests in a
+private setting than for tests at the workplace.\comment[id=HM]{True? @Janos}
 
@@
 that we can use a wide array of data to identify and estimate the model's many
-parameters.\footnote{See section S.XXX of the supplementary materials for an overview.}
-Mobility data is used to model the reductions in work contacts. School and daycare
-policies are incorporated directly from official directives. Infection rates by age and
-geographical region are estimated to match to officially recorded numbers; so is the
-prevalence of virus strains. The model yields total infections, but only a fraction of
-those will be officially recorded. We model the fraction of known cases as depicted in
+parameters.\comment[id=HM]{
+% \footnote{
+See section S.XXX of the supplementary materials for an overview.
+% }
+} Mobility data is used to model the reductions in work contacts. School and daycare
+policies can be incorporated directly from official directives. Aggregate tests,
+vaccinations, and the prevalence of virus strains can be taken from official records and
+from survey data.\comment[id=HM]{Anything else we need to mention?} We estimate the
+infection probabilities\comment[id=HM]{need to be more precise here, cannot find in
+Appendix} that the simulated new infections predict age- and region-specific time
+series. However, the two sets of time series are not comparable directly because not all
+cases will be recorded. We model the fraction of recorded cases as depicted in
 Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
-% \footnote{EQUATIONS FOR THE FOURTH PICTURE (find names!):
 
-%     The probabilities for Figure~\subref{fig:model_for_official_cases} are the
-%     following:
-
-%     \begin{align*} x & = \frac{n\_newly\_infected \times share\_known\_cases \times
-%         P(PCR | S)}{n\_symptomatic\_and\_untested}                    \\
-%         y & = \frac{n\_newly\_infected \times share\_known\_cases \times P(PCR |
-%         S)}{n\_asymptoptomatic\_and\_infected\_and\_untested} \\
-%         z & = \frac{P(PCR | RT)}{survey} \end{align*} where untested means that
-%     individuals are not waiting on the result of a previously administered rapid
-%     test.} 
-A further advantage is that the simulated data have a structure that resembles datasets
-used for regression models, which allows additional plausibility checks by re-running
-the same models on the model-generated data.\comment[id=HM]{Only keep if we actually do
-something like that.}
-
-We apply this model to Germany. In March and April 2020, the country broke the first
-wave of the pandemic fairly quickly. Between May and mid-September, daily new infections
-were below 20 per Million and day \citep{owidcoronavirus}. We model the period
-mid-September 2020 to the end of May 2021. We pick the starting date for two reasons.
-First, we do not include the first wave because the environment was very different
-(e.g., aggregate PCR test capacity was much lower and we would require a very different
-model for calculating the share of known cases). Second, a large fraction of cases
-during summer of 2020 were traced to international travel
-\citep{KochInstitut2021,Hodcroft2021} but the precise number is difficult to model.
-
+The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
+covering the period mid-September 2020 to the end of May 2021.
 Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
@@
 on. The value 1 represents the situation in mid-September, with restrictions on
-gatherings and public events, masking requirements, but open schools and workplaces (the
-raw value of the index at that point is 49.5). In the seven weeks between mid September
-and early November, cases increased by a factor of 10. Restrictions were somewhat
-tightened in mid-October and again in early November. New infections remained constant
-throughout November, before rising again in December, which prompted the most stringent
-lockdown to this date. Schools and daycare centers were closed again, so were
-customer-facing businesses except for grocery and drug stores. From the peak of the
-second wave just before Christmas until the trough in mid-February, newly detected cases
-decreased by almost three quarters. The third wave in the spring of 2021 is associated
-with the B.1.1.7 strain, which became dominant in March. See
+gatherings and public events, masking requirements, but open schools and workplaces. In
+the seven weeks between mid September and early November, cases increased by a factor of
+10. Restrictions were somewhat tightened in mid-October and again in early November. New
+infections remained constant throughout November, before rising again in December, which
+prompted the most stringent lockdown to this date. Schools and daycare centers were
+closed again, so were customer-facing businesses except for grocery and drug stores.
+From the peak of the second wave just before Christmas until the trough in mid-February,
+newly detected cases decreased by almost three quarters. The third wave in the spring of
+2021 is associated with the B.1.1.7 strain, which became dominant in March. See
 Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public.
+hairdressers and home improvement stores were allowed to open again to the public. There
+were many changes in details of regulations afterwards, but they did not change the
+stringency index.
 
@@
 over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
-started to replace PCR tests in many medical and nursing facilities. These had to be
-administered by medical doctors or in pharmacies. At-home tests approved by authorities
-became available in mid-March, rapid test centers were opened and one test per person
-and week was made available free of charge. Depending on the state, customers were only
-allowed to enter certain stores with a recent negative rapid test result. These
-developments are characteristic of many countries: The initial focus on NPIs to slow the
-spread of the disease has been accompanied by vaccines and a growing acceptance and use
-of rapid tests. At broadly similar points in time, novel strains of the virus have
-started to pose additional challenges.
+started to replace regular PCR tests for staff in many medical and nursing facilities.
+These had to be administered by medical doctors or in pharmacies. At-home tests approved
+by authorities became available in mid-March, rapid test centers were opened and one
+test per person and week was made available free of charge. Depending on the state,
+customers were only allowed to enter certain stores with a recent negative rapid test
+result. These developments are characteristic of many countries: The initial focus on
+NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
+acceptance and use of rapid tests. At broadly similar points in time, novel strains of
+the virus have started to pose additional challenges.
 
-Our model is able to track these developments very well. The blue line in
+We draw simulated samples of agents from the distribution of recorded infections in
+September 2020 and use the model to predict recorded infection rates until the end of
+May 2021. See Supplementary Materials~\ref{sec:data_and_parameters} and \ref{sec:model}
+for detailed descriptions of the data and the model, respectively. The blue line in
 Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
@@
 geographical region, which are shown in the supplementary materials
-(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively). We can
-disentangle various mechanisms due to the distinct temporal variation in the drivers of
-the pandemic. Next to the stringency index, the three lines in
-Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
+(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).
+
+The effects of various mechanisms can be disentangled due to the distinct temporal
+variation in the drivers of the pandemic. Next to the stringency index, the three lines
+in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
 regulations, and seasonality evolved since early September for each of the three broad
@@
 vaccinations, only the wildtype virus present), infections at the workplace would be
-reduced by 25\%. This reduction is the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. Along with the levels of infections,
-these measures determine the spread of SARS-CoV-2 in 2020. See
-Appendix~\comment[id=HM]{Reference} for separate plots of the three factors. Two aspects
-are particularly interesting. First, all lines broadly follow the stringency index and
-they would do so even more if we left out seasonality and school vacations (roughly the
-last two weeks of October, two weeks each around Christmas and Easter, and some days in
-late May). Second, the most stringent regulations are associated with the period of
-strong decreases in new infection between late December 2020 and mid-February 2021. The
-measures were not enough, however, to stop the B.1.1.7 variant from spreading in the
-subsequent period. The steep drop in recorded cases during May 2021 is associated with
-at least weekly rapid tests to around 42~percent of the population, a vaccination rate
-that rose from 28\% to 43\%, and mostly seasonality impacting a fall in the relative
-infectiousness of contacts outside of work and school.
+reduced by 25\%. The lines show the product of the effect of contact reductions,
+increased hygiene regulations, and seasonality. See Appendix~\comment[id=HM]{Reference}
+for separate plots of the three factors. Two aspects are particularly interesting.
+First, all lines broadly follow the stringency index and they would do so even more if
+we left out seasonality and school vacations (roughly the last two weeks of October, two
+weeks each around Christmas and Easter, and some days in late May). Second, the most
+stringent regulations are associated with the period of strong decreases in new
+infections between late December 2020 and mid-February 2021. The reversal of the trend
+is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
+during May 2021 is associated with at least weekly rapid tests to around 42~percent of
+the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
+impacting a fall in the relative infectiousness of contacts outside of work and school.
 
-Figure~\ref{fig:2021_scenarios_broad} consider the relative effects of rapid tests,
-vaccinations, and of seasonality during 2021, assuming NPIs to have evolved the same way
-as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model
-fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any
-of the three factors (red line), and three scenarios turning these factors off one by
-one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections
-in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values to
-decompose the difference in total infections between the scenario without any of the
-three factors and our main specification.
+In order to better understand the contributios  of rapid tests, vaccinations, and of
+seasonality on the evolution of infections in 2021,
+Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
+same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
+model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
+without any of the three factors (red line), and three scenarios turning these factors
+off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
+values to decompose the difference in total infections between the scenario without any
+of the three factors and our main specification.
 

### 2021-06-10 paper/appendix_data.tex
Update a few paragraphs.

@@
 \section{Data and Parameters}
-\label{sec:calibration_and_estimation}
+\label{sec:data_and_parameters}
 
@@
 
-\subsection{Assortative Matching}
+\subsection{Contacts by age}
+\label{subsec:contacts_by_age}
 
@@
 within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.
+data we use to calibrate the number of contacts.\comment[id=HM]{Redo
+\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of
+contacts or better add a similar figure showing total number of contacts by age in all
+networks}
 
@@
 \citep{Bundesanzeiger2021}. We assume that compliance is incomplete and only 80\% of
-employers actually offer tests.
+employers actually offer tests.\comment[id=HM]{Can we use the IZA/BMAS survey? Not sur
+about citing evening news in scientific papers.}
 

### 2021-06-10 paper/appendix_data.tex
Start on educ rapid test section.

--- 2021-06-10 paper/appendix_data.tex
@@
 
-\subsection{Assortative Matching}
+\subsection{Contacts by age}
 
@@
 
+For both work and school contacts we assume that starting November with the lockdown
+light in Germany, hygiene measures (such as masks, ventilation and hand washing) became
+more strict and more conscientiously observed, leading to a reduction of 33\% in the number of contacts with the potential to transmit Covid-19.
+
 For the last group of contacts which cover things like leisure activities, grocery
@@
 to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays.
+events such as policy announcements or particular holidays.\comment[id=K]{@Janos should we already add in numbers in here?}
 
@@
 that 60\% of workers who receive a test offer by their employer regularly use it. We
-assume this to be time constant.
+assume this share to be time constant.
 
-In addition, there are some surveys that allow us to trace the expansion of employers
-who offer tests to their employees. Mid march, 20\% of employers offered tests to their
-employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported
-being offered weekly rapid tests by their employer \citep{Ahlers2021}. This share
-increased to 60\% until the first days of April \cite{ZDF2021}. Until mid April 70\% of
-workers were expected to receive a weekly test offer \citep{AerzteZeitung2021}. However,
-according to surveys conducted in mid April \citep{Betsch2021}, less than two thirds of
-individuals with work contacts receive a test offer. Starting on April 19th employers
-were required by law to provide two weekly tests to their employees
-\citep{Bundesanzeiger2021}. We assume that compliance is incomplete and only 80\% of
-employers actually offer tests.
+In addition, there are some surveys that allow us to trace the expansion of employers who
+offer tests to their employees. Mid march, 20\% of employers offered tests to their
+employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
+offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
+60\% until the first days of April \cite{ZDF2021}. Until mid April 70\% of workers were
+expected to receive a weekly test offer \citep{AerzteZeitung2021}. However, according to
+surveys conducted in mid April \citep{Betsch2021}, less than two thirds of individuals
+with work contacts receive a test offer. Starting on April 19th employers were required
+by law to provide two weekly tests to their employees \citep{Bundesanzeiger2021}. We
+assume that compliance is incomplete and only 80\% of employers actually offer
+tests.\comment[id=HM]{Can we use the IZA/BMAS survey? Not sur about citing evening news
+in scientific papers.}\comment[id=K]{I am not aware of any survey looking at compliance.
+If you know of some, that would be great. Updating this might change our results but it
+would definitely be an improvement!}
 
@@
 
+We assume that employees in educational facilities start getting tested in 2021 and that
+by March 1st 30\% of them are tested weekly.
+
+The share increases to 90\% for the week before Easter. At that time both Bavaria
+\citep{BayrischerRundfunk2021} and Baden-Württemberg \citep{MinisteriumKultus2021} were
+offering tests to teachers and North-Rhine Westphalia \cite{DPA2021} and Lower Saxony
+\citep{SueddeutscheZeitung2021} were already testing students and tests for students and
+teachers were already mandatory in Saxony \citep{SueddeutscheZeitung2021a}. After Easter
+we assume that 95\% of teachers get tested twice per week.\comment[id=K]{Mostly newspaper
+articles and one KuMi press communication. I have more sources for the smaller federal
+states but they are extremely time consuming to enter into the references.}
+
+% --------------------------------------------------------------------------------------
+%
+%def _add_educ_rapid_test_fade_in_params(params):
+%    """Add the shares how many people with educ contacts get a rapid test.
+%
+%    Sources:
+%        - 17-24 of March 2021 (Mon, 2021-03-22):
+%            - NRW had 80% tests for students before Easter (https://bit.ly/3u7z8Rx)
+%            - BY: test offers to educ_workers (https://bit.ly/3tbVX5u)
+%            - BW: only tests for educ workers (https://bit.ly/2S7251M)
+%
+%            - federal level:
+%                "In Kitas und Schulen sollen die Testmöglichkeiten "mit der
+%                steigenden Verfügbarkeit von Schnell- und Selbsttests"
+%                ausgebaut werden" (https://bit.ly/3nuCSKi)
+%            - Some KiTa workers are being tested (https://bit.ly/3nyGyus)
+%            - Self tests for students in Berlin (https://bit.ly/2ScGu8m)
+%            - Schleswig-Holstein: test offer (https://bit.ly/3eVfkuv)
+%            - mandatory tests in Saxony (https://bit.ly/3eEQGhn)
+%            - no tests yet for students in Hessia, but already ordered
+%              (https://bit.ly/3gMGJB4)
+%            - Niedersachsen had one test week before Easter (https://bit.ly/3gOOC96)
+%
+%            => assume 90% of teachers and 40% of students do rapid tests
+%
+%        - After Easter (2021-04-07):
+%            - NRW: tests are mandatory for all
+%            - Bavaria: tests are mandatory for all (https://bit.ly/3nz5fXS,
+%              https://bit.ly/2QHilX3)
+%            - BW: voluntary tests for students (https://bit.ly/3vuetaD)
+%            - Brandenburg starts with tests (https://bit.ly/3xAihZB)
+%            - Schleswig-Holstein: mandatory tests (https://bit.ly/3eVfkuv)
+%
+%            => assume 95% of teachers and 75% of students get tested
+%
+%        - BW: tests mandatory starting 2021-04-19 (https://bit.ly/3vuetaD)
+%
+%            => assume 95% of teachers and 95% of students get tested
+%
+%    """
+%    params = params.copy(deep=True)
+%
+%    loc = ("rapid_test_demand", "educ_worker_shares")
+%    params.loc[(*loc, "2020-01-01"), "value"] = 0.0
+%    params.loc[(*loc, "2021-01-01"), "value"] = 0.0
+%    # this is arbitrary to have a more convex shape
+%    params.loc[(*loc, "2021-03-01"), "value"] = 0.3
+%    params.loc[(*loc, "2021-03-22"), "value"] = 0.9
+%    params.loc[(*loc, "2021-04-07"), "value"] = 0.95
+%    params.loc[(*loc, "2021-04-19"), "value"] = 0.95
+%    params.loc[(*loc, "2021-06-01"), "value"] = 0.95
+%    params.loc[(*loc, "2025-12-31"), "value"] = 0.95
+%
+%    loc = ("rapid_test_demand", "student_shares")
+%    params.loc[(*loc, "2020-01-01"), "value"] = 0.0
+%    params.loc[(*loc, "2021-02-01"), "value"] = 0.0
+%    params.loc[(*loc, "2021-03-01"), "value"] = 0.1
+%    params.loc[(*loc, "2021-03-22"), "value"] = 0.4
+%    params.loc[(*loc, "2021-04-07"), "value"] = 0.75
+%    params.loc[(*loc, "2021-04-19"), "value"] = 0.95
+%    params.loc[(*loc, "2021-06-01"), "value"] = 1.0
+%    params.loc[(*loc, "2025-12-31"), "value"] = 1.0
+%
+%    # Assume weekly tests before Easter and twice weekly tests after Easter
+%    # We should get a fade-in through different ends of Easter vaccation
+%    params.loc[("rapid_test_demand", "educ_frequency", "before_easter"), "value"] = 7
+%    params.loc[("rapid_test_demand", "educ_frequency", "after_easter"), "value"] = 3
+%
+%    return params
+%
+% --------------------------------------------------------------------------------------
+
 \comment[id=J]{Sources still missing below this}
-
-We assume that employees in educational facilities start getting tested in 2021 and that
-by March 1st 30\% of them are tested weekly. The share increases to 90\% for the week
-before Easter. At that time both Bavaria and Baden-Württemberg were offering tests to
-teachers and North-Rhine Westphalia and Lower Saxony were already testing students and
-tests for students and teachers were already mandatory in Saxony. After Easter we assume
-that 95\% of teachers get tested twice per week.
 

### 2021-06-10 paper/appendix_data.tex

@@
 \subsection{Contacts by age}
+\label{subsec:contacts_by_age}
 

### 2021-06-10 paper/report.tex

@@
 Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections, the final level is always the
-highest in case the vaccination campaign had been running in isolation (red lines). The
-Shapley value decomposition shows that vaccinations contribute about
-15\%\comment[id=HM]{check!} to the cumulative difference between scenarios. Reasons for
-this are the slow start---it took until 24~March until 10\% of the population had
-received their first vaccination, the 20\% mark was reached on April 19th ---and the
-focus on older individuals. These groups contribute less to the spread of the disease
-than others due to a lower number of contacts, see~\ref{fig:assortativity}. It is
-important to note that the initial focus of the campaign was to prevent deaths and
-severe disease; the case fatality was rate considerably lower during the third wave when
-compared to the second (4.4\% between October and February and 1.4\% between March and
-June). It is important to note that by the end of our study period, when first-dose
-vaccination rates reached around 40\% of the population, the numbers of new cases would
-have started to decline.
+sight. Whether considering recorded or total infections with only one channel active,
+the final level is always the highest in case of the vaccination campaign (orange
+lines). The Shapley value decomposition shows that vaccinations contribute about 15\% to
+the cumulative difference between scenarios. Reasons for this are the slow start---it
+took until March~24th until 10\% of the population had received their first vaccination,
+the 20\% mark was reached on April 19th---and the focus on older individuals. These
+groups contribute less to the spread of the disease than others due to a lower number of
+contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. It is important to
+note that the initial focus of the campaign was to prevent deaths and severe disease;
+the case fatality was rate considerably lower during the third wave when compared to the
+second (4.4\% between October and February and 1.4\% between March and June). It is
+important to note that by the end of our study period, when first-dose vaccination rates
+reached around 40\% of the population, the numbers of new cases would have started to
+decline.
 
@@
 rapid testing. Here, it is crucial to differentiate between recorded cases and actual
-cases. Additional testing means that infections become known which would otherwise
+cases. Additional testing means that infections become known, which would otherwise
 remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
@@
 three channels for rapid tests. Tests for pupils have the smallest effect, which is
-largely explained by the relatively small number of students (XXM vs. YYM workers) and
-the fact that schools did not operate at full capacity during our period of study.
-Almost 40\% come from tests at the workplace. Despite the fact that we phase rapid tests
-for private reasons are phased in only late, they make up for more than half of the total effect.
+largely explained by the relatively small number of students (XXM pupils vs. YYM
+workers)\comment[id=HM]{fill in} and the fact that schools did not operate at full
+capacity during our period of study. Almost 40\% come from tests at the workplace.
+Despite the fact that we phase rapid tests for private reasons are phased in only late,
+they make up for more than half of the total effect.
 
@@
 them gradually in late February, but usual operaton was not back until the beginning of
-June. In light of the large negative effects school closures have on children and
+June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding school starting at Easter, a point where rapid tests had become widely available.
+
+In light of the large negative effects school closures have on children and
 parents\comment[id=HM]{need to cite a couple of poor learning outcomes / mental health
 papers. There must be well-published ones}---and in particular on those with low
-socio-economic status---
+socio-economic status---the results in
 
@@
         \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-        \caption{{Effects of different schooling scenarios after Easter}}
-        \label{fig:schooling_scenarios_easter}
+        \caption{{Effects of different schooling scenarios}}
+        \label{fig:school_scenarios}
     \end{subfigure}
@@
         \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
-        \caption{{Effects of different work scenarios after Easter}}
-        \label{fig:to_be_determined}
+        \caption{{Effects of different work scenarios}}
+        \label{fig:workplace_scenarios}
     \end{subfigure}
@@
 
-    \caption{Effects of different scenarios for schooling and working from home}
-    \label{fig:interventions_school}
+    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
+    \label{fig:school_workplace_scenarios}
 
-    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
-        by geographical region.}
+    \floatfoot{\noindent Note: Interventions start at Easter because there were no
+        capacity constraints for rapid tests afterwards.}
 

### 2021-06-10 paper/report.tex
Work on first batch of comments in report.tex.

@@
 of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not transmit the
-virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\comment[id=T]{Maybe add
-that 35\% reach immunity after two weeks and 40\% after three weeks.}\footnote{75\% is
+virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is
     lower than what is usually reported for after the second dose of the Biontech/Pfizer
@@
 We include two types of tests. Polymerase chain reaction (PCR) tests directly reveal
-whether an individual is infected or not. PCR tests require some days to be processed
-and there are aggregate capacity constraints throughout. In contrast, rapid antigen
-tests yield immediate results; after a phase-in period, all tests that are demanded will
-be performed. Specificity and sensitivity of these tests is set according to data
-analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the
-test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our
-model for rapid test demand. Schools may require students to be tested regularly. Rapid
-tests may be offered by employers for on-site workers. Individuals may demand tests for
-private reasons, which include having plans to meet other people\footnote{A positive
-    test will make them reduce their contacts; this is why tests impact the actual contacts
-    in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
+whether an individual is infected or not. PCR tests require some days to be processed and
+there are aggregate capacity constraints throughout. In contrast, rapid antigen tests
+yield immediate results; after a phase-in period, all tests that are demanded will be
+performed. Specificity and sensitivity of these tests is set according to data analyzed
+in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
+to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our model for rapid
+test demand. Schools may require students to be tested regularly. Rapid tests may be
+offered by employers for on-site workers. Individuals may demand tests for private
+reasons, which include having plans to meet other people\footnote{A positive test will
+make them reduce their contacts; this is why tests impact the actual contacts in
+Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
 household member tested positively for the virus. We endow each agent with an individual
 compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons. The thresholds are lower for tests in a
-private setting than for tests at the workplace.\comment[id=HM]{True? @Janos}
+by employers or follows up on private reasons and whether she reduces her contacts if she
+tests positive or develops symptoms. The thresholds are higher for tests in a private
+setting than for tests at the workplace until mid April and then compliance because of
+private reasons becomes higher than compliance at the workplace.
 
@@
 that we can use a wide array of data to identify and estimate the model's many
-parameters.\comment[id=HM]{
-% \footnote{
-See section S.XXX of the supplementary materials for an overview.
-% }
-} Mobility data is used to model the reductions in work contacts. School and daycare
+parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
+materials for an overview.} We can rely on contact diaries to get pre-pandemic
+distributions of contacts for different contact types and their assortativity by age
+group. Mobility data is used to model the reductions in work contacts. School and daycare
 policies can be incorporated directly from official directives. Aggregate tests,
-vaccinations, and the prevalence of virus strains can be taken from official records and
-from survey data.\comment[id=HM]{Anything else we need to mention?} We estimate the
-infection probabilities\comment[id=HM]{need to be more precise here, cannot find in
-Appendix} that the simulated new infections predict age- and region-specific time
-series. However, the two sets of time series are not comparable directly because not all
-cases will be recorded. We model the fraction of recorded cases as depicted in
-Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
+vaccinations and the prevalence of virus strains can be taken from official records. The
+likelihood to receive a test offer or to take up a test offer and the share of
+individuals that have ever taken a test or done weekly tests in the last month can be
+calibrated from survey data.\comment[id=HM]{Anything else we need to
+mention?}\comment[id=K]{I added some more that I think are worth mentioning.} We estimate
+the infection probabilities\comment[id=HM]{need to be more precise here, cannot find in
+Appendix}\comment[id=K]{I added \ref{subsec:infection_probs}.} that the simulated new
+infections predict age- and region-specific time series.\comment[id=K]{We never directly
+targeted the federal states if I remember it right. How large was their weight in the
+criterion function @Janos?} However, the two sets of time series are not comparable
+directly because not all cases will be recorded. We model the fraction of recorded cases
+as depicted in Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
 

### 2021-06-10 paper/report.tex
Add paragraph on schools.

@@
 reduced by 25\%. The lines show the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. See Appendix~\comment[id=HM]{Reference}
-for separate plots of the three factors. Two aspects are particularly interesting.
-First, all lines broadly follow the stringency index and they would do so even more if
-we left out seasonality and school vacations (roughly the last two weeks of October, two
-weeks each around Christmas and Easter, and some days in late May). Second, the most
-stringent regulations are associated with the period of strong decreases in new
-infections between late December 2020 and mid-February 2021. The reversal of the trend
-is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
-during May 2021 is associated with at least weekly rapid tests to around 42~percent of
-the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
-impacting a fall in the relative infectiousness of contacts outside of work and school.
+increased hygiene regulations, and seasonality. See Appendix~\ref{subsec:policies} for
+separate plots of the three factors. Two aspects are particularly interesting. First, all
+lines broadly follow the stringency index and they would do so even more if we left out
+seasonality and school vacations (roughly the last two weeks of October, two weeks each
+around Christmas and Easter, and some days in late May). Second, the most stringent
+regulations are associated with the period of strong decreases in new infections between
+late December 2020 and mid-February 2021. The reversal of the trend is associated with he
+spread of the B.1.1.7 variant. The steep drop in recorded cases during May 2021 is
+associated with at least weekly rapid tests to around 42~percent of the population, a
+vaccination rate that rose from 28\% to 43\%, and mostly seasonality impacting a fall in
+the relative infectiousness of contacts outside of work and school.
 
@@
 them gradually in late February, but usual operaton was not back until the beginning of
-June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding school starting at Easter, a point where rapid tests had become widely available.
-
-In light of the large negative effects school closures have on children and
+June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
+regarding school starting at Easter, at which point rapid tests had become widely
+available. We estimate the realized scenario to have essentially the same effect as a
+situation with closed schools. Under fully opened schools with mandatory tests, total
+infections would have been 7\% higher; this number rises to 20\% without tests. These
+effect sizes are broadly in line with emprical studies, e.g., \citet{Vlachos2021}
+\comment[id=HM]{\url{https://www.pnas.org/content/118/9/e2020834118}}
+\comment[id=HM]{Need a back-of the envelope calculation} To use another metric, the
+effective weekly reproduction number differs by $0.017$ and $0.045$, respectively. In
+light of the large negative effects school closures have on children and
 parents\comment[id=HM]{need to cite a couple of poor learning outcomes / mental health
 papers. There must be well-published ones}---and in particular on those with low
-socio-economic status---the results in
+socio-economic status---these results in conjunction with hindsight bias suggest that
+opening schools combined with a testing strategy would have been beneficial. In other
+situations, and particular when no test strategy is available, trade-offs may well be
+different.
 

### 2021-06-10 paper/appendix_data.tex
Work on first batch of comments in report.tex.
Improve on the policies section by adding multiplier plots and adding seasonality to the policy discussion.

@@
 \subsection{Infection Probabilities}
-\label{sec:estimation}
+\label{subsec:infection_probs}
 
@@
 \subsection{Policies}
+\label{subsec:policies}
 
@@
 In our empirical application we distinguish four groups of contact types: households,
-education, work and other contacts. For households we assume that the individuals'
-contacts in their households do not change over our estimation period. For nurseries,
-preschools and schools we implement vacations as announced by the German federal states
-as well as school closures, emergency care and A / B schooling where only one half of
-students attends every other week or day. For the moment we ignore that lack of childcare
-leads working parents to stay home.
+education, work and other contacts.
+% households
+For households we assume that the individuals'
+contacts in their households do not change over our estimation period.
+% educ models
+For nurseries, preschools and schools we implement vacations as announced by the German
+federal states as well as school closures, emergency care and A / B schooling where only
+one half of students attends every other week or day. For the moment we ignore that lack
+of childcare leads working parents to stay home. An approximation of the share of contacts still taking place with the different school regulations can be found in Figure~\ref{fig:school_multiplier}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/school_multiplier_comparison}
+    \caption{School Multiplier With and Without Vacations Factored In}
+    \floatfoot{\noindent}
+    \label{fig:school_multiplier}
+\end{figure}
+
 %
@@
 %
+
+% work models
 For our work models\footnote{We distinguish non-recurrent work contacts, daily work
@@
 
+% other contacts
 For the last group of contacts which cover things like leisure activities, grocery
@@
 to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays.\comment[id=K]{@Janos should we already add in numbers in here?}
+events such as policy announcements or particular holidays.
+Because of the scarce data situation we cannot distinguish between a hygiene factor (such as mask wearing) during meetings and physical distancing (such as virtual meetings with friends).
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../figures/results/figures/data/other_multiplier}
+\caption{Share of Other Contacts Still Taking Place}
+\label{fig:other_multiplier}
+\floatfoot{\noindent}
+\end{figure}
+
+% seasonality
+Another important factor in the potential for a contact to lead to an infection is the
+seasonality.\comment[id=K]{Reference missing} There are two channels through which
+seasonality affects the infectiousness of contacts. One has to do with the physical
+conditions like UV radiation and the humidity. The other has to do with where people
+meet. Especially leisure contacts are more likely to take place outdoors and individuals
+are more likely to have windows open when the weather is nicer. To capture both channels
+we allow for other contacts to have a higher seasonality than our other contact models.
+Figure~\ref{fig:seasonality} shows our seasonality factors.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
+\caption{Seasonality for Other Contacts and Other Contact Models}
+\label{fig:seasonality}
+\floatfoot{\noindent}
+\end{figure}
 

### 2021-06-10 paper/appendix_data.tex
Cleaned up comments, added sentence on model for recorded cases.

@@
 
-\subsection{Number of Contacts}
+\subsection{Number of Contacts\comment[id=HM]{Histograms seem like the way to go here. Also add household contacts, i.e., distribution of (household size minus 1).}}
 \label{sub:number_of_contacts}
@@
 
-\subsection{Infection Probabilities}
+\subsection{Infection Probabilities\comment[id=HM]{Is the data section the right place?}}
 \label{subsec:infection_probs}
@@
 
-Currently, we use data for Germany from October 2020 until June 2021. We do not use
+We use data for Germany from October 2020 until June 2021. We do not use
 earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
@@
 (including leisure activities such as meeting friends) carry the highest risk with
-15.875\%.
+15.875\%.\comment[id=HM]{Are these numbers current? In any case, a table would substitute for this and the previous paragraph. Need to mention B.1.1.7 alongside.}
 

### 2021-06-10 paper/report.tex
Improve on the policies section by adding multiplier plots and adding seasonality to the policy discussion.
Cleaned up comments, added sentence on model for recorded cases.

@@
 months---and often years---away from reaching vaccination-induced herd immunity
-\citep{Swaminathan2021}.\comment[id=K]{WHO Press Conference saying we won't reach herd
-immunity in 2021. She cites no source.}\comment[id=HM]{If we find a paper (s)he cites, would be even better}
-In the meantime, it is of utmost importance to employ an effective mix of strategies for
-containing the virus. The most frequent initial response was a set of non-pharmaceutical
-interventions (NPIs) to reduce contacts between individuals. While this has allowed some
-countries to sustain equilibria with very low infection numbers\footnote{See
-    \citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
-    sustained with test-trace-and-isolate policies.}, most have seen large fluctuations of
-infection rates. Containment measures have become increasingly diverse and now include
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
-influence of seasonal patterns or more infectious virus strains are well understood in
-quantitative terms.
+\citep{Swaminathan2021}.\comment[id=HM]{Need to find some more scientific citation
+before submission} In the meantime, it is of utmost importance to employ an effective
+mix of strategies for containing the virus. The most frequent initial response was a set
+of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
+this has allowed some countries to sustain equilibria with very low infection
+numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
+    numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
+fluctuations of infection rates. Containment measures have become increasingly diverse
+and now include testing, more nuanced NPIs, and contact tracing. Neither these policies'
+effect nor the influence of seasonal patterns or more infectious virus strains are well
+understood in quantitative terms.
 
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
-        \caption{{Model for undetected cases}}
+        \caption{{Translating all infections to recorded ones}}
         \label{fig:model_for_official_cases}
@@
 of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not transmit the
-virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is
-    lower than what is usually reported for after the second dose of the Biontech/Pfizer
-    vaccine, which is most commonly used in Germany. We choose it because our model neither
-    includes booster shots, nor does it allow vaccinated individuals who became immune to
-    transmit the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}. If
-    anything, these assumptions would overstate the effect of vaccines for our study period.
-    This would be different if a large fraction of vaccinated individuals had received a
-    second dose already.} During the vaccine roll-out, priority may depend on age and
-occupation.
+virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is lower
+    than what is usually reported for after the second dose of the Biontech/Pfizer vaccine,
+    which is most commonly used in Germany. We choose it because our model neither includes
+    booster shots, nor does it allow vaccinated individuals who became immune to transmit
+    the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}. If anything, these
+    assumptions would overstate the effect of vaccines for our study period. This would be
+    different if a large fraction of vaccinated individuals had received a second dose
+    already.} During the vaccine roll-out, priority may depend on age and occupation.
 
 We include two types of tests. Polymerase chain reaction (PCR) tests directly reveal
-whether an individual is infected or not. PCR tests require some days to be processed and
-there are aggregate capacity constraints throughout. In contrast, rapid antigen tests
-yield immediate results; after a phase-in period, all tests that are demanded will be
-performed. Specificity and sensitivity of these tests is set according to data analyzed
-in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
-to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our model for rapid
-test demand. Schools may require students to be tested regularly. Rapid tests may be
-offered by employers for on-site workers. Individuals may demand tests for private
-reasons, which include having plans to meet other people\footnote{A positive test will
-make them reduce their contacts; this is why tests impact the actual contacts in
-Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
+whether an individual is infected or not. PCR tests require some days to be processed
+and there are aggregate capacity constraints throughout. In contrast, rapid antigen
+tests yield immediate results; after a phase-in period, all tests that are demanded will
+be performed. Specificity and sensitivity of these tests is set according to data
+analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the
+test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our
+model for rapid test demand. Schools may require students to be tested regularly. Rapid
+tests may be offered by employers for on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people\footnote{A positive
+    test will make them reduce their contacts; this is why tests impact the actual contacts
+    in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
 household member tested positively for the virus. We endow each agent with an individual
 compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons and whether she reduces her contacts if she
-tests positive or develops symptoms. The thresholds are higher for tests in a private
-setting than for tests at the workplace until mid April and then compliance because of
-private reasons becomes higher than compliance at the workplace.
+by employers or follows up on private reasons and whether she reduces her contacts if
+she tests positive or develops symptoms. The thresholds are higher for tests in a
+private setting than for tests at the workplace until mid April and then compliance
+because of private reasons becomes higher than compliance at the workplace.
 
@@
 parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
-materials for an overview.} We can rely on contact diaries to get pre-pandemic
-distributions of contacts for different contact types and their assortativity by age
-group. Mobility data is used to model the reductions in work contacts. School and daycare
-policies can be incorporated directly from official directives. Aggregate tests,
-vaccinations and the prevalence of virus strains can be taken from official records. The
-likelihood to receive a test offer or to take up a test offer and the share of
-individuals that have ever taken a test or done weekly tests in the last month can be
-calibrated from survey data.\comment[id=HM]{Anything else we need to
-mention?}\comment[id=K]{I added some more that I think are worth mentioning.} We estimate
-the infection probabilities\comment[id=HM]{need to be more precise here, cannot find in
-Appendix}\comment[id=K]{I added \ref{subsec:infection_probs}.} that the simulated new
-infections predict age- and region-specific time series.\comment[id=K]{We never directly
-targeted the federal states if I remember it right. How large was their weight in the
-criterion function @Janos?} However, the two sets of time series are not comparable
-directly because not all cases will be recorded. We model the fraction of recorded cases
-as depicted in Figure~\ref{fig:model_for_official_cases}.\comment[id=HM]{Add a sentence.}
+    materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
+for different contact types and their assortativity by age group. Mobility data is used
+to model the reductions in work contacts. School and daycare policies can be
+incorporated directly from official directives. Administrative records on the number of
+tests, vaccinations by age and region, and the prevalence of virus strains are generally
+available. Surveys may ask about test offers, propensities to take them up, and past
+tests.
+
+The free parameters are the infection probabilities by contact type, see Supplementary
+Material~\ref{subsec:infection_probs}. We estimate these parameters so that  simulated
+infections by age predict the corresponding time series in the data \citep[See][for the
+    general method]{McFadden1989}. However, the two sets of time series are not comparable
+directly because not all cases will be recorded. We thus model recorded cases as
+depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of
+the share of recorded cases and use data on whether the reason for a PCR test were
+symptoms of CoViD-19 to get age-specific shares. For example,
+\comment[id=HM]{@Janos, please correct if necessary and insert reference} estimate
+unknown cases to be twice as high for children as for adults. Confirmation of a positive
+rapid test result is another reason for PCR tests, once these become available.
 
@@
         \caption{Decomposition of the difference between the scenario without rapid
-            tests and the main scenario in Figure~\ref{fig:2021_scenarios_newly_infected}.}
+            tests and the main scenario in
+            Figure~\ref{fig:2021_scenarios_newly_infected}.}
         \label{fig:2021_scenarios_decomposition_tests}
@@
 
-Effects:
-\begin{itemize}
-    \item Schools were essentially closed
-    \item What are the different $R_t$-values? In no-test scenario and $R_t>1$, might be
-          worth chopping off 0.05 or whatever that might be?
-    \item With tests it is really hard to make a case for that
-\end{itemize}
-
-Alternative:
-\begin{itemize}
-    \item Germany quite lenient on home office relative to other countries (employer had
-          to allow)
-    \item Testing: Rapid tests are cheap (less than 1\euro), just a nuisance.
-\end{itemize}
-
-Results
-\begin{itemize}
-    \item Advantage of testing in both cases: Recurrent. Small cost relative to other
-          stuff. Certain publicness.
-    \item Mandatory tests: Screening effect for hard-to-reach populations (do not
-          consider in this paper, also due to poor data in DE)
-\end{itemize}
-
 \begin{figure}[!tp]
@@
 \end{figure}
+
+Alternative:
+\begin{itemize}
+    \item Germany quite lenient on home office relative to other countries (employer had
+          to allow)
+    \item Testing: Rapid tests are cheap (less than 1\euro), just a nuisance.
+\end{itemize}
+
+Results
+\begin{itemize}
+    \item Advantage of testing in both cases: Recurrent. Small cost relative to other
+          stuff. Certain publicness.
+    \item Mandatory tests: Screening effect for hard-to-reach populations (do not
+          consider in this paper, also due to poor data in DE)
+\end{itemize}

### 2021-06-10 paper/appendix_data.tex
Fill in some numbers and references on policies, seasonality and from the simulation results.

--- 2021-06-10 paper/appendix_data.tex
@@
 Another important factor in the potential for a contact to lead to an infection is the
-seasonality.\comment[id=K]{Reference missing} There are two channels through which
+seasonality \citep{Kuehn2020} There are two channels through which
 seasonality affects the infectiousness of contacts. One has to do with the physical

### 2021-06-10 paper/report.tex
Fill in some numbers and references on policies, seasonality and from the simulation results.

--- 2021-06-10 paper/report.tex
@@
 rapid testing. Here, it is crucial to differentiate between recorded cases and actual
-cases. Additional testing means that infections become known, which would otherwise
+cases. Additional testing means that infections are detected, which would otherwise
 remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
 until late April, recorded cases are higher than in the scenario where none of the three
-mechanisms is turned on. Compared to the scenario with vaccinations only, this point is
+mechanisms are turned on. Compared to the scenario with vaccinations only, this point is
 reached around mid-May and it would be June for the comparison with the seasonality-only
@@
 that only a small fraction of the population performed weekly rapid tests in
-March\footnote{X\% in the model; 27\% of respondents to the COSMO study reported in
-    early March that they \textit{ever} did a rapid test)}\comment[id=HM]{Put in}, new
+March\footnote{8\% in the model; 27\% of respondents to the COSMO study reported in
+    early March that they \textit{ever} did a rapid test)}, new
 infections on April 1 would be reduced by 53\% relative to the scenario without
@@
 scenario without rapid tests only (purple line in
-Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
-three channels for rapid tests. Tests for pupils have the smallest effect, which is
-largely explained by the relatively small number of students (XXM pupils vs. YYM
-workers)\comment[id=HM]{fill in} and the fact that schools did not operate at full
-capacity during our period of study. Almost 40\% come from tests at the workplace.
-Despite the fact that we phase rapid tests for private reasons are phased in only late,
-they make up for more than half of the total effect.
-
-\comment[id=HM]{How many tests are actually performed along each of the three dimensions?
-Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}
+Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
+channels for rapid tests. Tests for pupils have the smallest effect, which is largely
+explained by the relatively small number of students (Only 18\% of our population are in
+the education sector (pupils, teachers etc.) versus  46\% are workers outside the
+education sector) and the fact that schools did not operate at full capacity during our
+period of study. Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for private reasons are phased in only late, they make up for more than
+half of the total effect.\comment[id=HM]{How many tests are actually performed along each of the three dimensions?
+Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}\comment[id=K]{This is only calculated very deeply inside the model.
+I'll try to save it.}
 
@@
 up on a positive test of a household member. The latter is essentially a form of contact
-tracing, which has been shown to be very effective.\comment[id=HM]{cite Priesemann
-paper, others?} This is also the reason for why the number of tests performed fall at
+tracing, which has been shown to be very effective \citep{Contreras2021,
+Kretzschmar2020}. This is also the reason for why the number of tests performed fall at
 the very end of our study period. Falling infection rates mean that there are fewer

### 2021-06-10 paper/report.tex
Add improved work scenario plot. Answer some comments.

@@
 that only a small fraction of the population performed weekly rapid tests in
-March\footnote{8\% in the model; 27\% of respondents to the COSMO study reported in
-    early March that they \textit{ever} did a rapid test)}, new
-infections on April 1 would be reduced by 53\% relative to the scenario without
-vaccinations, rapid tests, or seasonality.
+March\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
+week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
+respondents to the COSMO study reported in early March that they \textit{ever} did a
+rapid test)}, new infections on April 1 would be reduced by 53\% relative to the scenario
+without vaccinations, rapid tests, or seasonality.
 
@@
 events triggering private test demand.
-\comment[id=HM]{Positive rates along the three channels might be nice here, if previous comment is not a quant story}
-\comment[id=HM]{I think there is a story in here about risk conditional on past exposure vs.
-prospective risk in gatherings, but it is too late to work through it.
-We might just want to calculate the probabilities.}
+\comment[id=HM]{Positive rates along the three channels might be nice here, if previous
+comment is not a quant story}\comment[id=K]{Working on it}
+\comment[id=HM]{I think there is a story in here about risk conditional on past exposure
+vs. prospective risk in gatherings, but it is too late to work through it. We might just
+want to calculate the probabilities.}\comment[id=K]{Yes, we have both in the private
+reasons. Retrospective after a case in the household and prospective in anticipation of
+participation in a weekly meeting with friends. However, we cannot easily turn these
+separately on and off at the moment as they use the same multiplier. We could do,
+however, do this in the future.}
 

### 2021-06-10 paper/report.tex
Moved on with final paragraphs.

@@
 
-Alternative:
-\begin{itemize}
-    \item Germany quite lenient on home office relative to other countries (employer had
-          to allow)
-    \item Testing: Rapid tests are cheap (less than 1\euro), just a nuisance.
-\end{itemize}
+Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of employees receiving tests, testing at the workplace has larger effects than changing requirements to work from home.
+Whether the share of workers working at the usual workplace is reduced or increased by ten percent hardly has an effect on infection rates.
+Making testing mandatory twice a week---assuming independet compliance by employers and workers or 95\% each---would have reduced infections by 13\%.
+Reducing rapid tests offers by employers to the level of March would have increased them by the same amount.
+
+Despite these large effects, the results on testing likely understate the benefits.
+Disadvantaged groups of the population are less likely to be reached by testing campaigns; at the same time they have a higher risk to contract CoViD-19.
+Mandatory tests at school and at the workplace would help reach these groups, too.
+The same goes for ...
+
+The tests are
+
+ Testing: Rapid tests are cheap (less than 1\euro), just a nuisance.
 
@@
           stuff. Certain publicness.
-    \item Mandatory tests: Screening effect for hard-to-reach populations (do not
-          consider in this paper, also due to poor data in DE)
 \end{itemize}

### 2021-06-10 paper/abstract.tex
Full version, as sent to Jörg. Manually reverted 3a,b.

@@
 \noindent
-
-% Abstracts of Research Articles and Reports should explain to the general reader why the research was done, what was found and why the results are important.
-
-% They should start with some brief BACKGROUND information: a sentence giving a broad introduction to the field comprehensible to the general reader, and then a sentence of more detailed background specific to your study.
-
-Background:
-
-Spread disease, initially NPIs, now testing and vaccines; seasonality unclear.
-
-% This should be followed by an explanation of the OBJECTIVES/METHODS and then the RESULTS.
-
-Objective: Provide a model that allows studying these things in conjunction, allowing for different virus strains.
-
-Results: Along the transition to vaccination-induced herd immunity, testing is most effective, also thanks to family structures. Seasonality contributed its share.
-
-% The final sentence should outline the main CONCLUSIONS of the study, in terms that will be comprehensible to all our readers.
-The Abstract is distinct from the main body of the text, and thus should not be the only source of background information critical to understanding the manuscript.
-Please do not include citations or abbreviations in the Abstract.
-
-Conclusions: Frequent rapid testing should remain part of strategies to contain CoViD-19.
+% Abstracts of Research Articles and Reports should explain to the general reader why
+% the research was done, what was found and why the results are important. 
+% 
+% They should start with some brief BACKGROUND information: a sentence giving a broad
+% introduction to the field comprehensible to the general reader, and then a sentence of
+% more detailed background specific to your study. 
+In order to slow the spread of the CoViD-19 pandemic, governments around the world have
+enacted a wide set of policies limiting the transmission of the disease. Initially,
+these focused on non-pharmaceutical interventions; more recently, vaccinations and
+large-scale rapid testing have started to play a major role. The objective of this study
+is to explain the quantitative effects of these policies on determining the course of
+the pandemic, allowing for factors like seasonality or virus strains with different
+transmission profiles. In order to do so, the study develops an agent-based simulation
+model that is estimated on data for the second and the third wave of the CoViD-19
+pandemic in Germany. The paper finds that during a period where vaccination rates rose
+from 5\% to 40\%, large-scale rapid testing had the largest effect on reducing infection
+numbers. 
+% The final sentence should outline the main CONCLUSIONS of the study, in terms that
+% will be comprehensible to all our readers. The Abstract is distinct from the main body
+% of the text, and thus should not be the only source of background information critical
+% to understanding the manuscript. Please do not include citations or abbreviations in
+% the Abstract. 
+Frequent, large-scale rapid testing should remain part of strategies to contain
+CoViD-19; it can substitute for many non-pharmaceutical interventions that come at a
+much larger cost to individuals, society, and the economy.
 
@@
 \vspace{1cm}
-\noindent JEL Classification: C63, I18
+\noindent \textbf{JEL Classification:} C63, I18
 
-\noindent Keywords: Covid-19, agent based simulation model, public health measures
+\noindent \textbf{Keywords:} CoViD-19, agent based simulation model, rapid testing,
+non-pharmaceutical interventions

### 2021-06-10 paper/report.tex
Full version, as sent to Jörg. Manually reverted 3a,b.

@@
 settings. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact.
+probability of infection associated with a contact
+\citep{Cheng2021}\comment[id=HM]{https://science.sciencemag.org/content/early/2021/05/19/science.abg6296}
 
@@
 Despite these large effects, the results on testing likely understate the benefits.
-Disadvantaged groups of the population are less likely to be reached by testing campaigns; at the same time they have a higher risk to contract CoViD-19.
-Mandatory tests at school and at the workplace would help reach these groups, too.
-The same goes for ...
-
-The tests are
-
- Testing: Rapid tests are cheap (less than 1\euro), just a nuisance.
-
-Results
-\begin{itemize}
-    \item Advantage of testing in both cases: Recurrent. Small cost relative to other
-          stuff. Certain publicness.
-\end{itemize}
+Disadvantaged groups of the population are less likely to be reached by testing campaigns;\comment[id=HM]{seems obvious, but I did not find anything.
+COSMO data?
+We should get the raw data, btw.} at the same time they have a higher risk to contract CoViD-19.\citet{RKI-Corona-Monitoring-Bundesweit}\comment[id=HM]{\url{https://www.rki.de/DE/Content/Gesundheitsmonitoring/Studien/lid/Ergebnisse.pdf}} Mandatory tests at school and at the workplace will extend more into these groups.
+The same goes for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
+Our analysis has shown that in the transition path to high levels of vaccination and possibly thereafter, large-scale rapid testing can substitute for some NPIs.
+This comes at a fraction of the cost.
+A day of the winter lockdown in Germany is estimated to have cost XXX~Euros.
+In early June 2021, retail prices for self-administered rapid tests are frequently less than 1~Euro.

### 2021-06-11 paper/report.tex
Added two references and miniscule fixes.

@@
 probability of infection associated with a contact
-\citep{Cheng2021}\comment[id=HM]{https://science.sciencemag.org/content/early/2021/05/19/science.abg6296}
+\citep{Cheng2021}.
 
@@
 the case fatality was rate considerably lower during the third wave when compared to the
-second (4.4\% between October and February and 1.4\% between March and June). It is
-important to note that by the end of our study period, when first-dose vaccination rates
-reached around 40\% of the population, the numbers of new cases would have started to
-decline.
+second (4.4\% between October and February and 1.4\% between March and June). By the end
+of our study period, when first-dose vaccination rates reached around 40\% of the
+population, the numbers of new cases would have started to decline.
 
@@
 effect sizes are broadly in line with emprical studies, e.g., \citet{Vlachos2021}
-\comment[id=HM]{\url{https://www.pnas.org/content/118/9/e2020834118}}
 \comment[id=HM]{Need a back-of the envelope calculation} To use another metric, the

### 2021-06-11 paper/report.tex
New figures. Working on sizes.

--- 2021-06-10 paper/report.tex
@@
 half of the total effect.\comment[id=HM]{How many tests are actually performed along each of the three dimensions?
-Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}\comment[id=K]{This is only calculated very deeply inside the model.
-I'll try to save it.}
+Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}\comment[id=K]{Working on it.}
 

### 2021-06-11 paper/report.tex

--- 2021-06-11 paper/report.tex
@@
 general method]{McFadden1989}. However, the two sets of time series are not comparable
-directly because not all cases will be recorded. We thus model recorded cases as
-depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of
-the share of recorded cases and use data on whether the reason for a PCR test were
-symptoms of CoViD-19 to get age-specific shares. For example,
-\comment[id=HM]{@Janos, please correct if necessary and insert reference} estimate
-unknown cases to be twice as high for children as for adults. Confirmation of a positive
-rapid test result is another reason for PCR tests, once these become available.
+directly because not all cases will be recorded. We thus model recorded cases as depicted
+in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the share of
+recorded cases and use data on whether the reason for a PCR test were symptoms of
+CoViD-19 to get age-specific shares. For example,\comment[id=HM]{@Janos, please correct
+if necessary and insert reference} estimate unknown cases to be twice as high for
+children as for adults. Confirmation of a positive rapid test result is another reason
+for PCR tests, once these become available.
 
@@
 the very end of our study period. Falling infection rates mean that there are fewer
-events triggering private test demand.
-\comment[id=HM]{Positive rates along the three channels might be nice here, if previous
-comment is not a quant story}\comment[id=K]{Working on it}
-\comment[id=HM]{I think there is a story in here about risk conditional on past exposure
-vs. prospective risk in gatherings, but it is too late to work through it. We might just
-want to calculate the probabilities.}\comment[id=K]{Yes, we have both in the private
-reasons. Retrospective after a case in the household and prospective in anticipation of
-participation in a weekly meeting with friends. However, we cannot easily turn these
-separately on and off at the moment as they use the same multiplier. We could do,
-however, do this in the future.}
+events triggering private test demand.\comment[id=HM]{Positive rates along the three
+channels might be nice here, if previous comment is not a quant
+story}\comment[id=K]{Working on it}\comment[id=HM]{I think there is a story in here about
+risk conditional on past exposure vs. prospective risk in gatherings, but it is too late
+to work through it. We might just want to calculate the
+probabilities.}\comment[id=K]{Yes, we have both in the private reasons. Retrospective
+after a case in the household and prospective in anticipation of participation in a
+weekly meeting with friends. However, we cannot easily turn these separately on and off
+at the moment as they use the same multiplier. We could do, however, do this in the
+future.}
 
@@
 measures, before being closed again from mid-December on. Some states started opening
-them gradually in late February, but usual operaton was not back until the beginning of
-June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
-regarding school starting at Easter, at which point rapid tests had become widely
-available. We estimate the realized scenario to have essentially the same effect as a
-situation with closed schools. Under fully opened schools with mandatory tests, total
-infections would have been 7\% higher; this number rises to 20\% without tests. These
-effect sizes are broadly in line with emprical studies, e.g., \citet{Vlachos2021}
-\comment[id=HM]{Need a back-of the envelope calculation} To use another metric, the
-effective weekly reproduction number differs by $0.017$ and $0.045$, respectively. In
-light of the large negative effects school closures have on children and
-parents\comment[id=HM]{need to cite a couple of poor learning outcomes / mental health
-papers. There must be well-published ones}---and in particular on those with low
-socio-economic status---these results in conjunction with hindsight bias suggest that
-opening schools combined with a testing strategy would have been beneficial. In other
-situations, and particular when no test strategy is available, trade-offs may well be
-different.
+them gradually in late February, but usual operation was not back until the beginning of
+June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
+school starting at Easter, at which point rapid tests had become widely available. We
+estimate the realized scenario to have essentially the same effect as a situation with
+closed schools. Under fully opened schools with mandatory tests, total infections would
+have been 7\% higher; this number rises to 20\% without tests. These effect sizes are
+broadly in line with empirical studies, e.g., \citet{Vlachos2021}\comment[id=HM]{Need a
+back-of the envelope calculation} To use another metric, the effective weekly
+reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
+negative effects school closures have on children and parents\comment[id=HM]{need to cite
+a couple of poor learning outcomes / mental health papers. There must be well-published
+ones}---and in particular on those with low socio-economic status---these results in
+conjunction with hindsight bias suggest that opening schools combined with a testing
+strategy would have been beneficial. In other situations, and particular when no test
+strategy is available, trade-offs may well be different.
 

### 2021-06-11 paper/results.tex
Outline for rest.
Cleaned up: \color{red} -> \comment[id=J].

@@
 This is the results section.
-
-\textcolor{red}{The results can be found in `figures/results`. This includes both figures
-and tables for lookup of numbers and summary tables.}
 
@@
   \label{fig:explain_decline}
-\figurenotes{\textcolor{red}{\ldots}}
+  \figurenotes{\ldots}
 \end{figure}
@@
   \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-\figurenotes{\textcolor{red}{K: One of the scenarios starts too early. Will be fixed with the next full simulations run.}}
-\label{fig:school_scenarios}
+  \figurenotes{}
+  \label{fig:appdx_school_scenarios}
 \end{figure}
@@
 \FloatBarrier
-
-\begin{figure}[ht]
-\centering
-% \includegraphics[width=\textwidth]{../figures/preliminary_shapley}
-\caption{Shapley Values}
-\figurenotes{\textcolor{red}{K: These are taken from Tobi's notebook as examples. Numbers
-might still change}}
-\label{fig:school_scenarios}
-
-\end{figure}

### 2021-06-11 paper/report.tex
Add references to mental health outcomes.

@@
 weekly meeting with friends. However, we cannot easily turn these separately on and off
-at the moment as they use the same multiplier. We could do, however, do this in the
+at the moment as they use the same multiplier. We could, however, do this in the
 future.}
@@
 reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
-negative effects school closures have on children and parents\comment[id=HM]{need to cite
-a couple of poor learning outcomes / mental health papers. There must be well-published
-ones}---and in particular on those with low socio-economic status---these results in
+negative effects school closures have on children and parents \citep(Luijten2021,
+Melegari2021) \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
+health papers. There must be well-published ones}\comment[id=K]{I googled and found two
+that I think are at least not terrible but I also didn't find anything really well
+published}---and in particular on those with low socio-economic status---these results in
 conjunction with hindsight bias suggest that opening schools combined with a testing

### 2021-06-11 paper/appendix_data.tex
Cleaned up: \color{red} -> \comment[id=J].
Add seasonality reference.

@@
 % seasonality
-Another important factor in the potential for a contact to lead to an infection is the
-seasonality \citep{Kuehn2020} There are two channels through which
+Another potentially important factor for a contact to lead to an infection is the
+seasonality \citep{Kuehn2020, Carlson2020} There are two channels through which
 seasonality affects the infectiousness of contacts. One has to do with the physical
-conditions like UV radiation and the humidity. The other has to do with where people
+conditions like the temperature and the humidity. The other has to do with where people
 meet. Especially leisure contacts are more likely to take place outdoors and individuals
@@
     \caption{\textbf{Share of Individuals that do a Rapid Test.}}
-    \floatfoot{\noindent \textcolor{red}{J: Talk about the interpretation of each line.}}
+    \floatfoot{\noindent ...}
     \label{fig:rapid_test_demand}
 \end{figure}
+
+Figure~\ref{fig:rapid_test_demand}\comment[id=J]{Talk about the interpretation of each line.}
 

### 2021-06-11 paper/report.tex
Add seasonality reference.

@@
 
-In order to better understand the contributios  of rapid tests, vaccinations, and of
+In order to better understand the contributions  of rapid tests, vaccinations, and of
 seasonality on the evolution of infections in 2021,

### 2021-06-11 paper/appendix_data.tex
Update figurenotes and small stuff.

@@
 with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
-households addresses all questions. I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}
+households addresses all questions.
+I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}\comment[id=HM]{See previous comment.
+My point is that total number of contacts rather than shares are much more informative in the first place, then different group sizes are no concern anymore.}
 
@@
 
-\subsection{Policies}
-\label{subsec:policies}
+\subsection{Policies and Seasonality}
+\label{subsec:policies_seasonality}
 
@@
     \includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
-\caption{Seasonality for Other Contacts and Other Contact Models}
+    \caption{Seasonality by Type of Contact}
     \label{fig:seasonality}

### 2021-06-11 paper/report.tex
Update figurenotes and small stuff.

@@
 numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
-fluctuations of infection rates. Containment measures have become increasingly diverse
-and now include testing, more nuanced NPIs, and contact tracing. Neither these policies'
-effect nor the influence of seasonal patterns or more infectious virus strains are well
-understood in quantitative terms.
+fluctuations of infection rates over time. Containment measures have become increasingly
+diverse and now include testing, more nuanced NPIs, and contact tracing. Neither these
+policies' effect nor the influence of seasonal patterns or more infectious virus strains
+are well understood in quantitative terms.
 
@@
 almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
-to play at least as long as vaccinations have not been offered to an entire population.
+to play for at least as long as vaccinations have not been offered to an entire
+population.
 
 At the core of our agent-based model are physical contacts between heterogeneous agents
-(Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to
-    other approaches in \ref{sec:literature_review}. The model most closely related to ours
-    is described in \citet{Hinch2020}.} Each contact between an infectious individual and
-somebody susceptible to the disease bears the risk of transmitting the virus. Contacts
-occur in the household, at work, at school, or in other settings (leisure activities,
-grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
-occur at random. Random contacts are typically assortative in age and geographical
-location. Empirical applications can take the population structure from census data and
-the types and frequency of contacts from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to
-    multiple data sets on contact types and frequencies covering countries from four
-    continents.}
+(Figure~\ref{fig:broad_model_description}).\footnote{A detailed comparison with other
+approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
+model most closely related to ours is described in \citet{Hinch2020}.} Each contact
+between an infectious individual and somebody susceptible to the disease bears the risk
+of transmitting the virus. Contacts occur in the household, at work, at school, or in
+other settings (leisure activities, grocery shopping, medical appointments, etc.). Some
+contacts recur regularly, others occur at random. Random contacts are typically
+assortative in age and geographical location. Empirical applications can take the
+population structure from census data and the types and frequency of contacts from diary
+data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008,Hoang2019}.
 
@@
 settings. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact
-\citep{Cheng2021}.
+probability of infection associated with a contact \citep{Cheng2021}.
 
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
-        \vskip4ex
+        \vskip5ex
 
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
+        \vskip5ex
+
         \caption{{Translating all infections to recorded ones}}
@@
 
-    \floatfoot{\noindent Note: ...}
+    \floatfoot{\noindent \textit{Note:}
+    A % formal 
+    description of the model can be found in Supplementary Material~\ref{sec:model}.
+    Figure~\ref{fig:broad_model_description} shows the influence on an agent's contacts
+    to other agents. Demographic characteristics set the baseline number of contacts in
+    different networks. She may reduce the number of contacts due to NPIs, showing
+    symptoms, or testing positively for SARS-CoV-2. Infections may occur when a
+    susceptible agent meets an infectious agent; the probability depends on the type of
+    contact, on seasonality, and on NPIs. If infected, the infection progresses as
+    depicted in Figure~\ref{fig:disease_progression}. If rapid tests are avaible,
+    agents' demand is modeled as in Figure~\ref{fig:antigen_tests}. All pupils are
+    attending school are tested, other reasons trigger a test only for a fraction of
+    individuals. Figure~\ref{fig:model_for_official_cases} shows the model of
+    translating all infections in the simulated data to age-specific recorded
+    infections. The model uses data on the aggregate share of recorded cases, on the
+    share of positive PCR tests triggered by symptoms, and on the false positive rate of
+    rapid tests. The lower part of the graph is relevant only for periods where rapid
+    tests are available.}
 \end{figure}
 
-In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
+In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
 possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
@@
 of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_index} the Oxford Response
+recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the Oxford Response
 Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
@@
         \caption{{Stringency of NPIs and infectious contacts}}
-        \label{fig:stringency_index}
+        \label{fig:stringency_infectious_contacts}
     \end{subfigure}
@@
 
-    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
-        by geographical region. Also more disaggregated data.
-
-        Sources: ...}
-
+    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
+    Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
+    Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
+    \ref{subsec:new_cases_fit}. For legibility reasons, all lines in
+    Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
+    Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
+    value of 1 is refers to the situation at the start of our sample period and 0 means
+    that all NPIs included in the index are turned on. The other lines in
+    Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
+    contact reductions, increased hygiene regulations, and seasonality. See
+    Appendix~\ref{subsec:policies_seasonality} for separate plots of the three factors
+    by contact type.}
 \end{figure}
@@
 variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
+in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions, increased hygiene
 regulations, and seasonality evolved since early September for each of the three broad
@@
 reduced by 25\%. The lines show the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. See Appendix~\ref{subsec:policies} for
-separate plots of the three factors. Two aspects are particularly interesting. First, all
-lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations (roughly the last two weeks of October, two weeks each
-around Christmas and Easter, and some days in late May). Second, the most stringent
-regulations are associated with the period of strong decreases in new infections between
-late December 2020 and mid-February 2021. The reversal of the trend is associated with he
-spread of the B.1.1.7 variant. The steep drop in recorded cases during May 2021 is
-associated with at least weekly rapid tests to around 42~percent of the population, a
-vaccination rate that rose from 28\% to 43\%, and mostly seasonality impacting a fall in
-the relative infectiousness of contacts outside of work and school.
+increased hygiene regulations, and seasonality. Two aspects are particularly
+interesting. First, all lines broadly follow the stringency index and they would do so
+even more if we left out seasonality and school vacations (roughly the last two weeks of
+October, two weeks each around Christmas and Easter, and some days in late May). Second,
+the most stringent regulations are associated with the period of strong decreases in new
+infections between late December 2020 and mid-February 2021. The reversal of the trend
+is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
+during May 2021 is associated with at least weekly rapid tests to around 42~percent of
+the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
+impacting a fall in the relative infectiousness of contacts outside of work and school.
 
@@
 infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
-values to decompose the difference in total infections between the scenario without any
-of the three factors and our main specification.
+values\comment[id=HM]{Need reference} to decompose the difference in total infections
+between the scenario without any of the three factors and our main specification.
 
@@
 
-    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
-        by geographical region.
-
-        The decomposition is based on Shapley values where the individual contribution
-        of a channel is its average contribution over different sizes of coalitions
-        (combinations with other channels). The individual contribution to a coalition
-        is the difference between the effect size of the coalition with the particular
-        channel and without.}
+    \floatfoot{\noindent \textit{Note:} The blue line in
+        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
+        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
+        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
+        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
+        variant is introduced in the same way; vaccinations, rapid tests, and
+        seasonality remain at their January levels. The other scenarios turn these three
+        factors on one-by-one. The decompositions in
+        Figures~\ref{fig:2021_scenarios_decomposition} and
+        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
+        where the individual contribution of a channel is its average contribution over
+        different sizes of coalitions (combinations with other channels). The individual
+        contribution to a coalition is the difference between the effect size of the
+        coalition with the particular channel and without.}
 \end{figure}

### 2021-06-11 paper/report.tex
Add estimate of the costs of the winter lockdown.

@@
 infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
-values\comment[id=HM]{Need reference} to decompose the difference in total infections
+values \citep{Shapley2016} to decompose the difference in total infections
 between the scenario without any of the three factors and our main specification.
@@
 Despite these large effects, the results on testing likely understate the benefits.
-Disadvantaged groups of the population are less likely to be reached by testing campaigns;\comment[id=HM]{seems obvious, but I did not find anything.
-COSMO data?
-We should get the raw data, btw.} at the same time they have a higher risk to contract CoViD-19.\citet{RKI-Corona-Monitoring-Bundesweit}\comment[id=HM]{\url{https://www.rki.de/DE/Content/Gesundheitsmonitoring/Studien/lid/Ergebnisse.pdf}} Mandatory tests at school and at the workplace will extend more into these groups.
-The same goes for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
-Our analysis has shown that in the transition path to high levels of vaccination and possibly thereafter, large-scale rapid testing can substitute for some NPIs.
-This comes at a fraction of the cost.
-A day of the winter lockdown in Germany is estimated to have cost XXX~Euros.
-In early June 2021, retail prices for self-administered rapid tests are frequently less than 1~Euro.
+Disadvantaged groups of the population are less likely to be reached by testing
+campaigns;\comment[id=HM]{seems obvious, but I did not find anything. COSMO data? We
+should get the raw data, btw.} at the same time they have a higher risk to contract
+CoViD-19.\citet{RKI-Corona-Monitoring-Bundesweit}\comment[id=HM]{\url{https://www.rki.de/DE/Content/Gesundheitsmonitoring/Studien/lid/Ergebnisse.pdf}}
+Mandatory tests at school and at the workplace will extend more into these groups. The
+same goes for individuals who exhibit a low level of compliance with CoViD-19-related
+regulations. Our analysis has shown that in the transition path to high levels of
+vaccination and possibly thereafter, large-scale rapid testing can substitute for some
+NPIs. This comes at a fraction of the cost. A day of the winter lockdown in Germany is
+estimated to have cost 6.25~billion~Euros.\cite{Dorn2020}\comment[id=K]{This is a
+prediction by the ifo Instute back from March 2020 assuming a 2 month lockdown. I did not
+find anything citable that is earlier. I did find
+\href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
+quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
+Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
+for self-administered rapid tests are frequently less than 1.5~Euro.\footnote{Cheapest on
+\href{https://amazon.de}{Amazon} on June 11th, 2021.}

### 2021-06-11 paper/report.tex
Finished figurenotes etc.

@@
     \floatfoot{\noindent \textit{Note:}
-    A % formal 
+        A
+        % formal 
         description of the model can be found in Supplementary Material~\ref{sec:model}.
-    Figure~\ref{fig:broad_model_description} shows the influence on an agent's contacts
-    to other agents. Demographic characteristics set the baseline number of contacts in
-    different networks. She may reduce the number of contacts due to NPIs, showing
-    symptoms, or testing positively for SARS-CoV-2. Infections may occur when a
-    susceptible agent meets an infectious agent; the probability depends on the type of
-    contact, on seasonality, and on NPIs. If infected, the infection progresses as
-    depicted in Figure~\ref{fig:disease_progression}. If rapid tests are avaible,
-    agents' demand is modeled as in Figure~\ref{fig:antigen_tests}. All pupils are
-    attending school are tested, other reasons trigger a test only for a fraction of
-    individuals. Figure~\ref{fig:model_for_official_cases} shows the model of
-    translating all infections in the simulated data to age-specific recorded
-    infections. The model uses data on the aggregate share of recorded cases, on the
-    share of positive PCR tests triggered by symptoms, and on the false positive rate of
-    rapid tests. The lower part of the graph is relevant only for periods where rapid
-    tests are available.}
+        Figure~\ref{fig:broad_model_description} shows the influence on an agent's
+        contacts to other agents. Demographic characteristics set the baseline number of
+        contacts in different networks. She may reduce the number of contacts due to
+        NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
+        occur when a susceptible agent meets an infectious agent; the probability
+        depends on the type of contact, on seasonality, and on NPIs. If infected, the
+        infection progresses as depicted in Figure~\ref{fig:disease_progression}. If
+        rapid tests are avaible, agents' demand is modeled as in
+        Figure~\ref{fig:antigen_tests}. All pupils are attending school are tested,
+        other reasons trigger a test only for a fraction of individuals.
+        Figure~\ref{fig:model_for_official_cases} shows the model of translating all
+        infections in the simulated data to age-specific recorded infections. The model
+        uses data on the aggregate share of recorded cases, on the share of positive PCR
+        tests triggered by symptoms, and on the false positive rate of rapid tests. The
+        lower part of the graph is relevant only for periods where rapid tests are
+        available.}
 \end{figure}
@@
 general method]{McFadden1989}. However, the two sets of time series are not comparable
-directly because not all cases will be recorded. We thus model recorded cases as depicted
-in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the share of
-recorded cases and use data on whether the reason for a PCR test were symptoms of
-CoViD-19 to get age-specific shares. For example,\comment[id=HM]{@Janos, please correct
-if necessary and insert reference} estimate unknown cases to be twice as high for
-children as for adults. Confirmation of a positive rapid test result is another reason
-for PCR tests, once these become available.
+directly because not all cases will be recorded. We thus model recorded cases as
+depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of
+the share of recorded cases and use data on whether the reason for a PCR test were
+symptoms of CoViD-19 to get age-specific shares. For example,\comment[id=HM]{@Janos,
+please correct if necessary and insert reference} estimate unknown cases to be twice as
+high for children as for adults. Confirmation of a positive rapid test result is another
+reason for PCR tests, once these become available.
 
@@
 of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the Oxford Response
-Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
-interventions. We transform the index so that lower values represent higher levels of
-restrictions. A value of zero means all measures incorporated in the index are turned
-on. The value 1 represents the situation in mid-September, with restrictions on
-gatherings and public events, masking requirements, but open schools and workplaces. In
-the seven weeks between mid September and early November, cases increased by a factor of
-10. Restrictions were somewhat tightened in mid-October and again in early November. New
-infections remained constant throughout November, before rising again in December, which
-prompted the most stringent lockdown to this date. Schools and daycare centers were
-closed again, so were customer-facing businesses except for grocery and drug stores.
-From the peak of the second wave just before Christmas until the trough in mid-February,
-newly detected cases decreased by almost three quarters. The third wave in the spring of
-2021 is associated with the B.1.1.7 strain, which became dominant in March. See
-Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-stringency index.
+recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
+Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
+non-pharmaceutical interventions. We transform the index so that lower values represent
+higher levels of restrictions. A value of zero means all measures incorporated in the
+index are turned on. The value 1 represents the situation in mid-September, with
+restrictions on gatherings and public events, masking requirements, but open schools and
+workplaces. In the seven weeks between mid September and early November, cases increased
+by a factor of 10. Restrictions were somewhat tightened in mid-October and again in
+early November. New infections remained constant throughout November, before rising
+again in December, which prompted the most stringent lockdown to this date. Schools and
+daycare centers were closed again, so were customer-facing businesses except for grocery
+and drug stores. From the peak of the second wave just before Christmas until the trough
+in mid-February, newly detected cases decreased by almost three quarters. The third wave
+in the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in
+March. See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed;
+e.g., hairdressers and home improvement stores were allowed to open again to the public.
+There were many changes in details of regulations afterwards, but they did not change
+the stringency index.
 
@@
         Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
-    Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
-    value of 1 is refers to the situation at the start of our sample period and 0 means
-    that all NPIs included in the index are turned on. The other lines in
-    Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
-    contact reductions, increased hygiene regulations, and seasonality. See
-    Appendix~\ref{subsec:policies_seasonality} for separate plots of the three factors
-    by contact type.}
+        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
+        a value of 1 is refers to the situation at the start of our sample period and 0
+        means that all NPIs included in the index are turned on. The other lines in
+        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
+        of contact reductions, increased hygiene regulations, and seasonality. See
+        Appendix~\ref{subsec:policies_seasonality} for separate plots of the three
+        factors by contact type.}
 \end{figure}
@@
 variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions, increased hygiene
-regulations, and seasonality evolved since early September for each of the three broad
-contact networks. For example, a value of 0.75 for the work multiplier means that if the
-environment was the same as in September (levels of infection rates, no rapid tests or
-vaccinations, only the wildtype virus present), infections at the workplace would be
-reduced by 25\%. The lines show the product of the effect of contact reductions,
-increased hygiene regulations, and seasonality. Two aspects are particularly
+in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
+increased hygiene regulations, and seasonality evolved since early September for each of
+the three broad contact networks. For example, a value of 0.75 for the work multiplier
+means that if the environment was the same as in September (levels of infection rates,
+no rapid tests or vaccinations, only the wildtype virus present), infections at the
+workplace would be reduced by 25\%. The lines show the product of the effect of contact
+reductions, increased hygiene regulations, and seasonality. Two aspects are particularly
 interesting. First, all lines broadly follow the stringency index and they would do so
@@
         Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
-        where the individual contribution of a channel is its average contribution over
+        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values where
+        the individual contribution of a channel is its average contribution over
         different sizes of coalitions (combinations with other channels). The individual
@@
 respondents to the COSMO study reported in early March that they \textit{ever} did a
-rapid test)}, new infections on April 1 would be reduced by 53\% relative to the scenario
-without vaccinations, rapid tests, or seasonality.
+rapid test)}, new infections on April 1 would be reduced by 53\% relative to the
+scenario without vaccinations, rapid tests, or seasonality.
 
@@
 scenario without rapid tests only (purple line in
-Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
-channels for rapid tests. Tests for pupils have the smallest effect, which is largely
-explained by the relatively small number of students (Only 18\% of our population are in
-the education sector (pupils, teachers etc.) versus  46\% are workers outside the
-education sector) and the fact that schools did not operate at full capacity during our
-period of study. Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for private reasons are phased in only late, they make up for more than
-half of the total effect.\comment[id=HM]{How many tests are actually performed along each of the three dimensions?
-Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}\comment[id=K]{Working on it.}
+Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
+three channels for rapid tests. Tests for pupils have the smallest effect, which is
+largely explained by the relatively small number of students (Only 18\% of our
+population are in the education sector (pupils, teachers etc.) versus  46\% are workers
+outside the education sector) and the fact that schools did not operate at full capacity
+during our period of study. Almost 40\% come from tests at the workplace. Despite the
+fact that rapid tests for private reasons are phased in only late, they make up for more
+than half of the total effect.\comment[id=HM]{How many tests are actually performed
+along each of the three dimensions? Could this be a quantity story rather than the
+directed testing story I am making up here as wishful thinking?}\comment[id=K]{Working
+on it.}
 
@@
 tracing, which has been shown to be very effective \citep{Contreras2021,
-Kretzschmar2020}. This is also the reason for why the number of tests performed fall at
-the very end of our study period. Falling infection rates mean that there are fewer
-events triggering private test demand.\comment[id=HM]{Positive rates along the three
-channels might be nice here, if previous comment is not a quant
-story}\comment[id=K]{Working on it}\comment[id=HM]{I think there is a story in here about
-risk conditional on past exposure vs. prospective risk in gatherings, but it is too late
-to work through it. We might just want to calculate the
+Kretzschmar2020,FetzerGraeber2021}. This is also the reason for why the number of tests
+performed fall at the very end of our study period. Falling infection rates mean that
+there are fewer events triggering private test demand.\comment[id=HM]{Positive rates
+along the three channels might be nice here, if previous comment is not a quant
+story}\comment[id=K]{Working on it}\comment[id=HM]{I think there is a story in here
+about risk conditional on past exposure vs. prospective risk in gatherings, but it is
+too late to work through it. We might just want to calculate the
 probabilities.}\comment[id=K]{Yes, we have both in the private reasons. Retrospective
@@
 them gradually in late February, but usual operation was not back until the beginning of
-June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
-school starting at Easter, at which point rapid tests had become widely available. We
-estimate the realized scenario to have essentially the same effect as a situation with
-closed schools. Under fully opened schools with mandatory tests, total infections would
-have been 7\% higher; this number rises to 20\% without tests. These effect sizes are
-broadly in line with empirical studies, e.g., \citet{Vlachos2021}\comment[id=HM]{Need a
-back-of the envelope calculation} To use another metric, the effective weekly
-reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
-negative effects school closures have on children and parents \citep(Luijten2021,
-Melegari2021) \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
-health papers. There must be well-published ones}\comment[id=K]{I googled and found two
-that I think are at least not terrible but I also didn't find anything really well
-published}---and in particular on those with low socio-economic status---these results in
-conjunction with hindsight bias suggest that opening schools combined with a testing
-strategy would have been beneficial. In other situations, and particular when no test
-strategy is available, trade-offs may well be different.
+June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
+regarding school starting at Easter, at which point rapid tests had become widely
+available. We estimate the realized scenario to have essentially the same effect as a
+situation with closed schools. Under fully opened schools with mandatory tests, total
+infections would have been 7\% higher; this number rises to 20\% without tests. These
+effect sizes are broadly in line with empirical studies, e.g.,
+\citet{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
+another metric, the effective weekly reproduction number differs by $0.017$ and $0.045$,
+respectively. In light of the large negative effects school closures have on children
+and parents \citep(Luijten2021, Melegari2021) \comment[id=HM]{need to cite a couple of
+poor learning outcomes / mental health papers. There must be well-published
+ones}\comment[id=K]{I googled and found two that I think are at least not terrible but I
+also didn't find anything really well published}---and in particular on those with low
+socio-economic status---these results in conjunction with hindsight bias suggest that
+opening schools combined with a testing strategy would have been beneficial. In other
+situations, and particular when no test strategy is available, trade-offs may well be
+different.
 
@@
 
-    \floatfoot{\noindent Note: Interventions start at Easter because there were no
-        capacity constraints for rapid tests afterwards.}
+    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
+        scenario; they are the same as in
+        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
+        because there were no capacity constraints for rapid tests afterwards.}
 
@@
 
-Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of employees receiving tests, testing at the workplace has larger effects than changing requirements to work from home.
-Whether the share of workers working at the usual workplace is reduced or increased by ten percent hardly has an effect on infection rates.
-Making testing mandatory twice a week---assuming independet compliance by employers and workers or 95\% each---would have reduced infections by 13\%.
-Reducing rapid tests offers by employers to the level of March would have increased them by the same amount.
+Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of employees
+receiving tests, testing at the workplace has larger effects than changing requirements
+to work from home. Whether the share of workers working at the usual workplace is
+reduced or increased by ten percent hardly has an effect on infection rates. Making
+testing mandatory twice a week---assuming independent compliance by employers and workers
+or 95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
+employers to the level of March would have increased them by the same amount.
 
@@
 Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
-for self-administered rapid tests are frequently less than 1.5~Euro.\footnote{Cheapest on
-\href{https://amazon.de}{Amazon} on June 11th, 2021.}
+for self-administered rapid tests are frequently less than 1~Euro.\comment[id=HM]{Cheapest on
+\href{https://amazon.de}{Amazon} on June 11th, 2021. --- DM 95ct}

### 2021-06-11 paper/report.tex
Add RKI source on disadvantaged groups.

--- 2021-06-11 paper/report.tex
@@
 should get the raw data, btw.} at the same time they have a higher risk to contract
-CoViD-19.\citet{RKI-Corona-Monitoring-Bundesweit}\comment[id=HM]{\url{https://www.rki.de/DE/Content/Gesundheitsmonitoring/Studien/lid/Ergebnisse.pdf}}
-Mandatory tests at school and at the workplace will extend more into these groups. The
-same goes for individuals who exhibit a low level of compliance with CoViD-19-related
-regulations. Our analysis has shown that in the transition path to high levels of
-vaccination and possibly thereafter, large-scale rapid testing can substitute for some
-NPIs. This comes at a fraction of the cost. A day of the winter lockdown in Germany is
-estimated to have cost 6.25~billion~Euros.\cite{Dorn2020}\comment[id=K]{This is a
-prediction by the ifo Instute back from March 2020 assuming a 2 month lockdown. I did not
-find anything citable that is earlier. I did find
+CoViD-19.\citet{KochInstitute2021a} Mandatory tests at school and at the workplace will
+extend more into these groups. The same goes for individuals who exhibit a low level of
+compliance with CoViD-19-related regulations. Our analysis has shown that in the
+transition path to high levels of vaccination and possibly thereafter, large-scale rapid
+testing can substitute for some NPIs. This comes at a fraction of the cost. A day of the
+winter lockdown in Germany is estimated to have cost
+6.25~billion~Euros.\cite{Dorn2020}\comment[id=K]{This is a prediction by the ifo Instute
+back from March 2020 assuming a 2 month lockdown. I did not find anything citable that is
+earlier. I did find
 \href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A

### 2021-06-11 paper/appendix_results.tex
Cleaned up: \color{red} -> \comment[id=J].
Finished figurenotes etc.
Add rapid test statistics figures. Add share known cases and rapid test statistics to the appendix.

@@
 observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
-rapid test demands. Where available we do not only look at aggregated statistics but
-also analyze the model fit for age groups.
-
-\comment[id=J]{summarize the fit; Some of this needs to go in the appendix. I suggest the age group and federal state fit goes there???}
+rapid test demands. Where available we do not only look at aggregated statistics but also
+analyze the model fit for age groups.\comment[id=J]{summarize the fit; Some of this needs
+to go in the appendix. I suggest the age group and federal state fit goes there???}
 
@@
 
+\subsection{Share of Cases that are Detected}
+\label{subsec:appendix_share_known_cases}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
+  \caption{Share of Detected Cases by Age Group in the Main Prediction}
+  \label{fig:share_known_cases_by_age_group}
+  \floatfoot{\noindent \textcolor{red}{\ldots}}
+\end{figure}
+
+It's noteworthy that the share of detected cases increases rapidly in May for the five to fourteen year olds.
+This is a direct result of the mandatory tests in school.\comment[id=K]{We don't do this at the moment but we could also create plots showing the share of detected cases between scenarios - either for the full population or for a age group of interest.}
+
+\subsection{Rapid Tests}
+\label{subsec:appendix_rapid_tests}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/demand_shares}
+  \caption{Share of the Population Demanding a Rapid Test for a Particular Reason}
+  \label{fig:rapid_tests_by_reason}
+  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without PCR demand and other contact demand together form the private demand category.
+Also note that these do not add up to the total share of demanded rapid tests as individuals may have more than one reason to demand a rapid test on any given day.
+\textcolor{red}{\ldots}}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_demand}
+  \caption{Share of Individuals that Demand a Rapid Test for a Particular Reason that are Actually Infected}
+  \label{fig:share_of_rapid_tests_by_reason_by_infected}
+  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without PCR demand and other contact demand together form the private demand category.
+Also note that this is not the same as individuals getting a positive rapid test.
+The sensitivity is quite low before individuals become infectious.
+Therefore, if individuals are still in the latent period of their infection they are likely to get a false positive rapid test.
+\textcolor{red}{\ldots}}
+\end{figure}
+
+\FloatBarrier
+
 \subsection{Scenarios}
@@
 
-This is the results section.
-
-\comment[id=K]{The results can be found in `figures/results`. This includes both figures
-and tables for lookup of numbers and summary tables.}
+This is the results section.\comment[id=K]{The results can be found in `figures/results`.
+This includes both figures and tables for lookup of numbers and summary tables.}
 
@@
   \label{fig:explain_decline}
-  \figurenotes{\textcolor{red}{\ldots}}
+  \figurenotes{\ldots}
 \end{figure}
@@
   \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-  \label{fig:school_scenarios}
+  \label{fig:school_scenarios_detailed}
 \end{figure}

### 2021-06-11 paper/report.tex
Add rapid test statistics figures. Add share known cases and rapid test statistics to the appendix.

@@
 reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
-negative effects school closures have on children and parents \citep(Luijten2021,
-Melegari2021) \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
+negative effects school closures have on children and parents \citep{Luijten2021,
+Melegari2021} \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
 health papers. There must be well-published ones}\comment[id=K]{I googled and found two
@@
 should get the raw data, btw.} at the same time they have a higher risk to contract
-CoViD-19.\citet{KochInstitute2021a} Mandatory tests at school and at the workplace will
+CoViD-19.\citet{KochInstitut2021a} Mandatory tests at school and at the workplace will
 extend more into these groups. The same goes for individuals who exhibit a low level of
@@
 testing can substitute for some NPIs. This comes at a fraction of the cost. A day of the
-winter lockdown in Germany is estimated to have cost
-6.25~billion~Euros.\cite{Dorn2020}\comment[id=K]{This is a prediction by the ifo Instute
-back from March 2020 assuming a 2 month lockdown. I did not find anything citable that is
-earlier. I did find
+winter lockdown in Germany is estimated to have cost 6.25~billion~Euros
+\citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo Instute back from March
+2020 assuming a 2 month lockdown. I did not find anything citable that is more recent. I did
+find
 \href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
 quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
-Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
+Milliarden Euro pro Woche“} for the winter lockdown. That would be much less than calculated in March.} In early June 2021, retail prices
 for self-administered rapid tests are frequently less than 1.5~Euro.\footnote{Cheapest on

### 2021-06-11 paper/report.tex
Add references to the new rapid test statistics in the comments.

@@
 education sector) and the fact that schools did not operate at full capacity during our
-period of study. Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for private reasons are phased in only late, they make up for more than
-half of the total effect.\comment[id=HM]{How many tests are actually performed along each of the three dimensions?
-Could this be a quantity story rather than the directed testing story I am making up here as wishful thinking?}\comment[id=K]{Working on it.}
+period of study. Almost 40\% come from tests at the workplace. Despite the fact that
+rapid tests for private reasons are phased in only late, they make up for more than half
+of the total effect.\comment[id=HM]{How many tests are actually performed along each of
+the three dimensions? Could this be a quantity story rather than the directed testing
+story I am making up here as wishful thinking?}\comment[id=K]{See subsection
+\ref{subsec:appendix_rapid_tests}.}
 
@@
 channels might be nice here, if previous comment is not a quant
-story}\comment[id=K]{Working on it}\comment[id=HM]{I think there is a story in here about
+story}\comment[id=K]{See figure~\ref{fig:share_of_rapid_tests_by_reason_by_infected}}\comment[id=HM]{I think there is a story in here about
 risk conditional on past exposure vs. prospective risk in gatherings, but it is too late
 to work through it. We might just want to calculate the
-probabilities.}\comment[id=K]{Yes, we have both in the private reasons. Retrospective
-after a case in the household and prospective in anticipation of participation in a
-weekly meeting with friends. However, we cannot easily turn these separately on and off
-at the moment as they use the same multiplier. We could, however, do this in the
-future.}
+probabilities.}\comment[id=K]{I don't understand which probabilities you mean.}
 

### 2021-06-14 paper/report.tex
Add Janos' comments on the report.

@@
 
-Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
+Since early 2020,\comment[id=J]{\textbf{Features that are not highlighted enough:}
+
+\begin{itemize}
+    \item Low sensitivity of rapid tests before and on day of infectiousness onset
+    \item Different seasonalities for different contact types (and quite strong for other contacts)
+    \item Low number of breakpoints in parameters / policies compared to breakpoints in
+    cases / reproduction rate
+\end{itemize}
+
+} the CoViD-19 pandemic has presented an enormous challenge to humanity
 on many dimensions. The development of highly effective vaccines holds the promise of
@@
 
-We include two types of tests. Polymerase chain reaction (PCR) tests directly reveal
-whether an individual is infected or not. PCR tests require some days to be processed
-and there are aggregate capacity constraints throughout. In contrast, rapid antigen
-tests yield immediate results; after a phase-in period, all tests that are demanded will
-be performed. Specificity and sensitivity of these tests is set according to data
-analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the
-test relative to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our
-model for rapid test demand. Schools may require students to be tested regularly. Rapid
-tests may be offered by employers for on-site workers. Individuals may demand tests for
-private reasons, which include having plans to meet other people\footnote{A positive
-    test will make them reduce their contacts; this is why tests impact the actual contacts
-    in Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
+We include two types of tests. Polymerase chain reaction (PCR) tests
+directly\comment[id=J]{It is not clear what "directly" means} reveal whether an
+individual is infected or not. PCR tests require some days to be processed and there are
+aggregate capacity constraints throughout. In contrast, rapid antigen tests yield
+immediate results; after a phase-in period, all tests that are demanded will be
+performed. Specificity and sensitivity of these tests is set according to data analyzed
+in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
+to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our model for rapid
+test demand. Schools may require students to be tested regularly. Rapid tests may be
+offered by employers for on-site workers. Individuals may demand tests for private
+reasons, which include having plans to meet other people\footnote{A positive test will
+make them reduce their contacts; this is why tests impact the actual contacts in
+Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
 household member tested positively for the virus. We endow each agent with an individual
 compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons and whether she reduces her contacts if
-she tests positive or develops symptoms. The thresholds are higher for tests in a
-private setting than for tests at the workplace until mid April and then compliance
-because of private reasons becomes higher than compliance at the workplace.
+by employers or follows up on private reasons and whether she reduces her contacts if she
+tests positive or develops symptoms. The thresholds are higher for tests in a private
+setting than for tests at the workplace until mid April and then compliance because of
+private reasons becomes higher than compliance at the workplace.\comment[id=J]{Shouldn't
+we mention that this is because of availability / high costs?}
 
@@
 for different contact types and their assortativity by age group. Mobility data is used
-to model the reductions in work contacts. School and daycare policies can be
-incorporated directly from official directives. Administrative records on the number of
-tests, vaccinations by age and region, and the prevalence of virus strains are generally
+to model the reductions in work contacts. School and daycare policies can be incorporated
+directly from official directives. Administrative records on the number of tests,
+vaccinations by age and region, and the prevalence of virus strains are generally
 available. Surveys may ask about test offers, propensities to take them up, and past
@@
 children as for adults. Confirmation of a positive rapid test result is another reason
-for PCR tests, once these become available.
+for PCR tests, once these become available.\comment[id=J]{Free parameters also include:
+
+- One hygiene multiplier for school an work contacts after November 2
+- A few "other" multipliers including two lockdown fatigue effects
+- Number of extra contacts during holidays
+
+Strictly speaking also:
+- The fade in speed of the rapid tests
+- Share known cases around holidays }
 
@@
 We draw simulated samples of agents from the distribution of recorded infections in
-September 2020 and use the model to predict recorded infection rates until the end of
-May 2021. See Supplementary Materials~\ref{sec:data_and_parameters} and \ref{sec:model}
-for detailed descriptions of the data and the model, respectively. The blue line in
-Figure~\ref{fig:aggregated_fit} shows our model's predictions are very close to
-officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region, which are shown in the supplementary materials
-(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).
+September 2020 and use the model to predict recorded infection rates until the end of May
+2021.\comment[id=J]{I do not understand} See Supplementary
+Materials~\ref{sec:data_and_parameters} and \ref{sec:model} for detailed descriptions of
+the data and the model, respectively. The blue line in Figure~\ref{fig:aggregated_fit}
+shows our model's predictions are very close to officially recorded cases in the
+aggregate. This is also true for infections by age and geographical region, which are
+shown in the supplementary materials (Figures~\ref{fig:age_group_fit} and
+\ref{fig:state_fit}, respectively).
 
@@
 In order to better understand the contributions  of rapid tests, vaccinations, and of
-seasonality on the evolution of infections in 2021,
-Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
-same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
-model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
-without any of the three factors (red line), and three scenarios turning these factors
-off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
-infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
-values \citep{Shapley2016} to decompose the difference in total infections
-between the scenario without any of the three factors and our main specification.
+seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
+considers various scenarios. NPIs are always the same as in the baseline scenario.
+Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
+Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
+and three scenarios turning these factors off one by one.
+Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
+model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
+\citep{Shapley2016} to decompose the difference in total infections between the scenario
+without any of the three factors and our main specification.\comment[id=J]{I would not
+mention the word "coalition" in the notes to the figure on the Shapley values.
+Combination with other channels is better. I think I would prefer to have a short section
+on shapley values in the appendix (why use them, how do they work) and just link to that
+in the figure note}
 
@@
         to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and
-        seasonality remain at their January levels. The other scenarios turn these three
-        factors on one-by-one. The decompositions in
-        Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
+        variant is introduced in the same way; vaccinations, rapid tests, and seasonality
+        remain at their January levels. The other scenarios turn these three factors on
+        one-by-one. The decompositions in Figures~\ref{fig:2021_scenarios_decomposition}
+        and \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
         where the individual contribution of a channel is its average contribution over
@@
 extend more into these groups. The same goes for individuals who exhibit a low level of
-compliance with CoViD-19-related regulations. Our analysis has shown that in the
-transition path to high levels of vaccination and possibly thereafter, large-scale rapid
-testing can substitute for some NPIs. This comes at a fraction of the cost. A day of the
-winter lockdown in Germany is estimated to have cost 6.25~billion~Euros
-\citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo Instute back from March
-2020 assuming a 2 month lockdown. I did not find anything citable that is more recent. I did
-find
+compliance with CoViD-19-related regulations.
+
+Our analysis\comment[id=J]{The conclusion should be slightly longer and more separated from the rest} has shown that in the transition path to high levels of vaccination and
+possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
+at a fraction of the cost. A day of the winter lockdown in Germany is estimated to have
+cost 6.25~billion~Euros \citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo
+Instute back from March 2020 assuming a 2 month lockdown. I did not find anything citable
+that is more recent. I did find
 \href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
 quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
-Milliarden Euro pro Woche“} for the winter lockdown. That would be much less than calculated in March.} In early June 2021, retail prices
-for self-administered rapid tests are frequently less than 1.5~Euro.\footnote{Cheapest on
-\href{https://amazon.de}{Amazon} on June 11th, 2021.}
+Milliarden Euro pro Woche“} for the winter lockdown. That would be much less than
+calculated in March.} In early June 2021, retail prices for self-administered rapid tests
+are frequently less than 1.5~Euro.\footnote{Cheapest on \href{https://amazon.de}{Amazon}
+on June 11th, 2021.}

### 2021-06-14 figures/results/tables/student_infections_table.tex
New results from clean run with full 30 seeds.

@@
 \midrule
- educ open after easter  without tests &                                             775324 \\
- educ open after easter  with tests    &                                             604078 \\
- close educ after easter               &                                             469274 \\
- baseline                              &                                             478080 \\
+ educ open after easter  without tests &                                             705858 \\
+ educ open after easter  with tests    &                                             539545 \\
+ close educ after easter               &                                             414920 \\
+ baseline                              &                                             425200 \\
 \bottomrule

### 2021-06-15 paper/appendix_data.tex
New results from clean run with full 30 seeds.

@@
 
+\FloatBarrier
+
+\subsection{The Synthetic Population}
+
+\textcolor{red}{\ldots}
+
+\FloatBarrier
+
 \subsection{Number of Contacts\comment[id=HM]{Histograms seem like the way to go here. Also add household contacts, i.e., distribution of (household size minus 1).}}
@@
 Simplifying the number of contacts, we arrive at the following distributions of the
-numbers of contacts by contact type.
+numbers of contacts by contact type.\comment[id=K]{Put plots into one figure.}
 
@@
     \centering
+
+    \begin{subfigure}[b]{0.25\textwidth}
+        \centering
+
         \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
@@
         \label{n_contacts_other_non_recurrent}
-    \floatfoot{\noindent In the model it is sampled every day which of these numbers of
-        contacts a person is planned to have. Note that the contact diaries include such high
-        values that super spreading events are well possible in our model. The planned number
-        of contacts is reduced by policies, seasonality and individual responses to events
-        such as receiving a positive rapid test to the number of actual contacts with
-        transmission potential. Other contacts include all contacts that are not household
-        members, school contacts or work contacts, for example leisure contacts or contacts
-        during grocery shopping. We assume that individuals in households with children or
-        teachers or retired individuals have additional contacts during school vacations to
-        cover things like family visits or travel during vacations. We estimate this to be on
-        average 0.5 additional contacts per vacation day.}
-\end{figure}
+    \end{subfigure}
 
-\begin{figure}
+    \hfill
+
+    \begin{subfigure}[b]{0.25\textwidth}
         \centering
@@
         \label{n_contacts_other_daily_recurrent}
-    \floatfoot{\noindent Individuals are assigned to groups that are time constant and
-        that meet daily. The share of individuals that attend in a way that has transmission
-        potential is reduced by policies, seasonality and individual responses to events such
-        as receiving a positive rapid test. Other contacts include all contacts that are not
-        household members, school contacts or work contacts, for example leisure contacts or
-        contacts during grocery shopping.}
-\end{figure}
+    \end{subfigure}
 
-\begin{figure}
+    \hfill
+
+    \begin{subfigure}[b]{0.25\textwidth}
         \centering
@@
         \label{n_contacts_other_weekly_recurrent}
-    \floatfoot{\noindent Individuals are assigned to up to four groups that are time
-        constant and that meet weekly. The share of individuals that attend in a way that has
-        transmission potential is reduced by policies, seasonality and individual responses
-        to events such as receiving a positive rapid test. Other contacts include all
-        contacts that are not household members, school contacts or work contacts, for
-        example leisure contacts or contacts during grocery shopping.}
+    \end{subfigure}
+
+    \caption{Number of Contacts of the Other Contact Type}
+    \label{fig:n_contacts_other}
+
+    \floatfoot{\noindent
+        \textit{Note:} Other contacts include all contacts that are not household
+        members, school contacts or work contacts, for example leisure contacts or
+        contacts during grocery shopping. The planned number of contacts is reduced by
+        policies, seasonality and individual responses to events such as receiving a
+        positive rapid test to the number of actual contacts with transmission potential.        % non recurrent
+        In the model it is sampled every day which of the numbers of non recurrent
+        contacts a person is planned to have. Note that the contact diaries include such
+        high values that super spreading events are well possible in our model through
+        non recurrent other models. We assume that individuals in households with
+        children or teachers or retired individuals have additional non recurrent
+        contacts during school vacations to cover things like family visits or travel
+        during vacations. We estimate this to be on average 0.5 additional contacts per
+        vacation day.
+        % daily and weekly
+        For the recurrent other contacts, individuals are assigned to groups that are
+        time constant and that meet daily or weekly. The share of individuals that attend
+        in a way that has transmission potential is reduced by policies, seasonality and
+        individual responses to events such as receiving a positive rapid test. For
+        weekly contacts, individuals are assigned to up to four groups that are time
+        constant and that meet weekly. The day on which meetings take place varies
+        between groups but stays the same for each group.
+    }
 \end{figure}
+
+\textcolor{red}{Arbeitsplots auch in eine gemeinsame Graphik packen}
 
@@
         test. Work contacts only take place between working individuals.}
+\end{figure}
+
+An exception where we do not rely on the data by \cite{Mossong2008} are the household
+contacts. Since household are included in the the German microcensus\comment[id=K]{Cite
+microcensus} on which we build our synthetic population we simply assume for the
+household contact model that individuals meet all other household members every day.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/household}
+    \caption{Number of Household Contacts}
+    \label{n_contacts_hh}
+    \floatfoot{\noindent
+        \textit{Note:} Every individual meets all other household members every day. The
+        German microcensus sampled full households such that our synthetic population
+        automatically fits population characteristics such as size and age distribution.}
 \end{figure}

### 2021-06-15 paper/report.tex
Add number of household contacts.

@@
 
-\begin{figure}
+\begin{figure}   % Figure 1
     \centering
@@
 of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the Oxford Response
-Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
-interventions. We transform the index so that lower values represent higher levels of
-restrictions. A value of zero means all measures incorporated in the index are turned
-on. The value 1 represents the situation in mid-September, with restrictions on
-gatherings and public events, masking requirements, but open schools and workplaces. In
-the seven weeks between mid September and early November, cases increased by a factor of
-10. Restrictions were somewhat tightened in mid-October and again in early November. New
-infections remained constant throughout November, before rising again in December, which
-prompted the most stringent lockdown to this date. Schools and daycare centers were
-closed again, so were customer-facing businesses except for grocery and drug stores.
-From the peak of the second wave just before Christmas until the trough in mid-February,
-newly detected cases decreased by almost three quarters. The third wave in the spring of
-2021 is associated with the B.1.1.7 strain, which became dominant in March. See
-Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-stringency index.
+recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
+Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
+non-pharmaceutical interventions. We transform the index so that lower values represent
+higher levels of restrictions. A value of zero means all measures incorporated in the
+index are turned on. The value 1 represents the situation in mid-September, with
+restrictions on gatherings and public events, masking requirements, but open schools and
+workplaces. In the seven weeks between mid September and early November, cases increased
+by a factor of 10. Restrictions were somewhat tightened in mid-October and again in
+early November. New infections remained constant throughout November, before rising
+again in December, which prompted the most stringent lockdown to this date. Schools and
+daycare centers were closed again, so were customer-facing businesses except for grocery
+and drug stores. From the peak of the second wave just before Christmas until the trough
+in mid-February, newly detected cases decreased by almost three quarters. The third wave
+in the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in
+March. See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed;
+e.g., hairdressers and home improvement stores were allowed to open again to the public.
+There were many changes in details of regulations afterwards, but they did not change
+the stringency index.
 
-\begin{figure}[!tp]
+\begin{figure}[!tp]   % Figure 2
     \centering
@@
 events triggering private test demand.\comment[id=HM]{Positive rates along the three
-channels might be nice here, if previous comment is not a quant
-story}\comment[id=K]{See figure~\ref{fig:share_of_rapid_tests_by_reason_by_infected}}\comment[id=HM]{I think there is a story in here about
-risk conditional on past exposure vs. prospective risk in gatherings, but it is too late
-to work through it. We might just want to calculate the
+channels might be nice here, if previous comment is not a quant story}\comment[id=K]{See
+figure~\ref{fig:share_of_rapid_tests_by_reason_by_infected}} \comment[id=HM]{I think
+there is a story in here about risk conditional on past exposure vs. prospective risk in
+gatherings, but it is too late to work through it. We might just want to calculate the
 probabilities.}\comment[id=K]{I don't understand which probabilities you mean.}
@@
 quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
-Milliarden Euro pro Woche“} for the winter lockdown. That would be much less than
-calculated in March.} In early June 2021, retail prices for self-administered rapid tests
-are frequently less than 1.5~Euro.\footnote{Cheapest on \href{https://amazon.de}{Amazon}
-on June 11th, 2021.}
+Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
+for self-administered rapid tests are frequently less than 1~Euro.\comment[id=HM]{Cheapest on
+\href{https://amazon.de}{Amazon} on June 11th, 2021. --- DM 95ct}

### 2021-06-15 paper/appendix_data.tex
Change K's comment color. Put rapid test params section in separate tex file and add text with URLs.

@@
 
+\input{rapid_test_params}
+
+All shares of individuals that would take a rapid test if the conditions were met can be
+seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
+of each line.}
+
 \begin{figure}
@@
 \end{figure}
-
-Figure~\ref{fig:rapid_test_demand}\comment[id=J]{Talk about the interpretation of each line.}
-
-In our model, there are five reasons why rapid tests are done:
-\begin{enumerate}
-    \item someone plans to have work contacts
-    \item someone is an employees of an educational facility or school pupil
-    \item someone lives in a household where someone has tested positive or developed
-          symptoms
-    \item someone has developed symptoms but has not received a PCR test
-    \item someone plans to participate in a weekly non-work meeting
-\end{enumerate}
-
-% work rapid tests
-
-For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April)
-that 60\% of workers who receive a test offer by their employer regularly use it. We
-assume this share to be time constant.
-
-In addition, there are some surveys that allow us to trace the expansion of employers who
-offer tests to their employees. Mid march, 20\% of employers offered tests to their
-employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
-offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
-60\% until the first days of April \cite{ZDF2021}. Until mid April 70\% of workers were
-expected to receive a weekly test offer \citep{AerzteZeitung2021}. However, according to
-surveys conducted in mid April \citep{Betsch2021}, less than two thirds of individuals
-with work contacts receive a test offer. Starting on April 19th employers were required
-by law to provide two weekly tests to their employees \citep{Bundesanzeiger2021}. We
-assume that compliance is incomplete and only 80\% of employers actually offer
-tests.\comment[id=HM]{Can we use the IZA/BMAS survey? Not sur about citing evening news
-in scientific papers.}\comment[id=K]{I am not aware of any survey looking at compliance.
-If you know of some, that would be great. Updating this might change our results but it
-would definitely be an improvement!}
-
-% educ rapid tests
-
-We assume that employees in educational facilities start getting tested in 2021 and that
-by March 1st 30\% of them are tested weekly.
-
-The share increases to 90\% for the week before Easter. At that time both Bavaria
-\citep{BayrischerRundfunk2021} and Baden-Württemberg \citep{MinisteriumKultus2021} were
-offering tests to teachers and North-Rhine Westphalia \cite{DPA2021} and Lower Saxony
-\citep{SueddeutscheZeitung2021} were already testing students and tests for students and
-teachers were already mandatory in Saxony \citep{SueddeutscheZeitung2021a}. After Easter
-we assume that 95\% of teachers get tested twice per week.\comment[id=K]{Mostly newspaper
-articles and one KuMi press communication. I have more sources for the smaller federal
-states but they are extremely time consuming to enter into the references.}
-
-% --------------------------------------------------------------------------------------
-%
-%def _add_educ_rapid_test_fade_in_params(params):
-%    """Add the shares how many people with educ contacts get a rapid test.
-%
-%    Sources:
-%        - 17-24 of March 2021 (Mon, 2021-03-22):
-%            - NRW had 80% tests for students before Easter (https://bit.ly/3u7z8Rx)
-%            - BY: test offers to educ_workers (https://bit.ly/3tbVX5u)
-%            - BW: only tests for educ workers (https://bit.ly/2S7251M)
-%
-%            - federal level:
-%                "In Kitas und Schulen sollen die Testmöglichkeiten "mit der
-%                steigenden Verfügbarkeit von Schnell- und Selbsttests"
-%                ausgebaut werden" (https://bit.ly/3nuCSKi)
-%            - Some KiTa workers are being tested (https://bit.ly/3nyGyus)
-%            - Self tests for students in Berlin (https://bit.ly/2ScGu8m)
-%            - Schleswig-Holstein: test offer (https://bit.ly/3eVfkuv)
-%            - mandatory tests in Saxony (https://bit.ly/3eEQGhn)
-%            - no tests yet for students in Hessia, but already ordered
-%              (https://bit.ly/3gMGJB4)
-%            - Niedersachsen had one test week before Easter (https://bit.ly/3gOOC96)
-%
-%            => assume 90% of teachers and 40% of students do rapid tests
-%
-%        - After Easter (2021-04-07):
-%            - NRW: tests are mandatory for all
-%            - Bavaria: tests are mandatory for all (https://bit.ly/3nz5fXS,
-%              https://bit.ly/2QHilX3)
-%            - BW: voluntary tests for students (https://bit.ly/3vuetaD)
-%            - Brandenburg starts with tests (https://bit.ly/3xAihZB)
-%            - Schleswig-Holstein: mandatory tests (https://bit.ly/3eVfkuv)
-%
-%            => assume 95% of teachers and 75% of students get tested
-%
-%        - BW: tests mandatory starting 2021-04-19 (https://bit.ly/3vuetaD)
-%
-%            => assume 95% of teachers and 95% of students get tested
-%
-%    """
-%    params = params.copy(deep=True)
-%
-%    loc = ("rapid_test_demand", "educ_worker_shares")
-%    params.loc[(*loc, "2020-01-01"), "value"] = 0.0
-%    params.loc[(*loc, "2021-01-01"), "value"] = 0.0
-%    # this is arbitrary to have a more convex shape
-%    params.loc[(*loc, "2021-03-01"), "value"] = 0.3
-%    params.loc[(*loc, "2021-03-22"), "value"] = 0.9
-%    params.loc[(*loc, "2021-04-07"), "value"] = 0.95
-%    params.loc[(*loc, "2021-04-19"), "value"] = 0.95
-%    params.loc[(*loc, "2021-06-01"), "value"] = 0.95
-%    params.loc[(*loc, "2025-12-31"), "value"] = 0.95
-%
-%    loc = ("rapid_test_demand", "student_shares")
-%    params.loc[(*loc, "2020-01-01"), "value"] = 0.0
-%    params.loc[(*loc, "2021-02-01"), "value"] = 0.0
-%    params.loc[(*loc, "2021-03-01"), "value"] = 0.1
-%    params.loc[(*loc, "2021-03-22"), "value"] = 0.4
-%    params.loc[(*loc, "2021-04-07"), "value"] = 0.75
-%    params.loc[(*loc, "2021-04-19"), "value"] = 0.95
-%    params.loc[(*loc, "2021-06-01"), "value"] = 1.0
-%    params.loc[(*loc, "2025-12-31"), "value"] = 1.0
-%
-%    # Assume weekly tests before Easter and twice weekly tests after Easter
-%    # We should get a fade-in through different ends of Easter vaccation
-%    params.loc[("rapid_test_demand", "educ_frequency", "before_easter"), "value"] = 7
-%    params.loc[("rapid_test_demand", "educ_frequency", "after_easter"), "value"] = 3
-%
-%    return params
-%
-% --------------------------------------------------------------------------------------
-
-\comment[id=J]{Sources still missing below this}
-
-Tests for students started later so we assume that they only start in February and only
-10\% of students get tested by March 1st. Relying on the same sources as above we
-approximate that by the week before Easter this share had increased to 40\%.
-
-After Easter we assume the share of students receiving twice weekly tests to based on
-tests already being mandatory in North-Rhine Westphalia and Bavaria while still being
-voluntary in Baden-Württemberg. There tests become mandatory on April 19th.
-
-% household member rapid tests
-
-% own symptom rapid test demand
-
-% private meeting test demand
-
-\comment[id=J]{Add a section on how we calibrate rapid test demand; Mainly describe the
-datapoints we have and say that we usually interpolate linearly in between data points.
-(Only exception to that is private rapid test demand, which we fit to data)}
 

### 2021-06-15 paper/rapid_test_params.tex
Change K's comment color. Put rapid test params section in separate tex file and add text with URLs.

In our model, there are five reasons why rapid tests are done:\comment[id=J]{Add a
section on how we calibrate rapid test demand; Mainly describe the datapoints we have and
say that we usually interpolate linearly in between data points. (Only exception to that
is private rapid test demand, which we fit to data)}

\begin{enumerate}
    \item someone plans to have work contacts
    \item someone is an employee of an educational facility or a school pupil
    \item a household member has tested positive or developed symptoms
    \item someone has developed symptoms but has not received a PCR test
    \item someone plans to participate in a weekly non-work meeting
\end{enumerate}

\subsubsection{work rapid tests}

For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April)
that 60\% of workers who receive a test offer by their employer regularly use it. We
assume this share to be time constant.

In addition, there are some surveys that allow us to trace the expansion of employers who
offer tests to their employees. Mid march, 20\% of employers offered tests to their
employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
60\% until the first days of April \cite{ZDF2021}.

\textcolor{red}{ToDo: Find the survey that the ZDF is citing here}

Until mid April 70\% of workers were expected to receive a
weekly test offer \citep{AerzteZeitung2021}. However, according to surveys conducted in
mid April \citep{Betsch2021}, less than two thirds of individuals with work contacts
receive a test offer. Starting on April 19th employers were required by law to provide
two weekly tests to their employees \citep{Bundesanzeiger2021}. We assume that compliance
is incomplete and only 80\% of employers actually offer tests.

\subsubsection{educ rapid tests}

We assume that employees in educational facilities start getting tested in 2021 and that
by March 1st 30\% of them are tested weekly. The share increases to 90\% for the week
before Easter. At that time both Bavaria \citep{BayrischerRundfunk2021} and
Baden-Württemberg \citep{MinisteriumKultus2021} were offering tests to teachers and
North-Rhine
Westphalia\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen}}
\cite{DPA2021} and Lower Saxony \citep{SueddeutscheZeitung2021} were already testing
students and tests for students and teachers were already mandatory in Saxony
\citep{SueddeutscheZeitung2021a}. After Easter we assume that 95\% of teachers get tested
twice per week.

Tests for students started
later\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen}}
\citep{MinisteriumKultus2021} so we assume that they only start in February and only 10\%
of students get tested by March 1st. Relying on the same sources as above we approximate
that by the week before Easter this share had increased to 40\%.\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen}, }

After Easter the share of students receiving twice weekly tests is set to 75\%. This as
based on tests becoming mandatory becoming mandatory in Bavaria after Easter
break\footnote{Bavaria\footnote{\url{https://bit.ly/3nz5fXS}}, in North-Rhine Westphalia
on April
12th\footnote{https://www.schulministerium.nrw/ministerium/schulverwaltung/schulmail-archiv/14042021-schulbetrieb-im-wechselunterricht-ab-montag},
\url{https://bit.ly/2QHilX3}} and on the 19th in
Baden-Württemberg\footnote{https://bit.ly/3vuetaD, https://bit.ly/3vuetaD}.

To limit our degrees of freedom, we only have one parameter that governs how many
individuals do a rapid test because of any of the private demand reasons (own symptoms
but no PCR test, planned weekly leisure meeting or a symptomatic or positively tested
household member).

We assume that there is no private rapid test demand until March when both the citizens'
tests and rapid tests for lay people started to become available
\footnote{\url{https://bit.ly/3ehmGcj}, \url{https://bit.ly/3xJCIn8}} and other access to
rapid tests was very limited.

According to the COSMO study\footnote{\url{https://bit.ly/2QSFAgR}} 63\% would have been
willing to take a test in the round of 23rd of February 2021 when an acquaintance would
have tested positive. Since this is only asking for willingness not actual behavior and
the demand when meeting with friends is very likely lower, we take this as the upper
bound of private rapid test demand which is reached on May 4th. To cover that many people
are likely to have sought and done their first rapid test before the Easter holidays to
meet friends or family, we let the share of individuals doing rapid tests in that time
increase more rapidly than before and after. By end of March 25\% of individuals would do
a rapid test due to a private reason.

### 2021-06-15 paper/report.tex
Fix a couple of to-dos.

@@
 
-We include two types of tests. Polymerase chain reaction (PCR) tests
-directly\comment[id=J]{It is not clear what "directly" means} reveal whether an
-individual is infected or not. PCR tests require some days to be processed and there are
-aggregate capacity constraints throughout. In contrast, rapid antigen tests yield
-immediate results; after a phase-in period, all tests that are demanded will be
-performed. Specificity and sensitivity of these tests is set according to data analyzed
-in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
-to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our model for rapid
-test demand. Schools may require students to be tested regularly. Rapid tests may be
-offered by employers for on-site workers. Individuals may demand tests for private
-reasons, which include having plans to meet other people\footnote{A positive test will
-make them reduce their contacts; this is why tests impact the actual contacts in
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
+individual is infected or not; there is no uncertainty to the result. PCR tests require
+some days to be processed and there are aggregate capacity constraints throughout. In
+contrast, rapid antigen tests yield immediate results; after a phase-in period, all
+tests that are demanded will be performed. Specificity and sensitivity of these tests is
+set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
+the timing of the test relative to the start of infectiousness.
+Figure~\ref{fig:antigen_tests} shows our model for rapid test demand. Schools may
+require students to be tested regularly. Rapid tests may be offered by employers for
+on-site workers. Individuals may demand tests for private reasons, which include having
+plans to meet other people\footnote{A positive test will make them reduce their
+contacts; this is why tests impact the actual contacts in
 Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
@@
 compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons and whether she reduces her contacts if she
-tests positive or develops symptoms. The thresholds are higher for tests in a private
-setting than for tests at the workplace until mid April and then compliance because of
-private reasons becomes higher than compliance at the workplace.\comment[id=J]{Shouldn't
-we mention that this is because of availability / high costs?}
+by employers or follows up on private reasons and whether she reduces her contacts if
+she tests positive or develops symptoms. The thresholds are higher for tests in a
+private setting than for tests at the workplace until mid April and then compliance
+because of private reasons becomes higher than compliance at the
+workplace.\comment[id=J]{Shouldn't we mention that this is because of availability /
+high costs?}
 
@@
 The free parameters are the infection probabilities by contact type, see Supplementary
-Material~\ref{subsec:infection_probs}. We estimate these parameters so that simulated
-infections by age predict the corresponding time series in the data \citep[See][for the
-general method]{McFadden1989}. However, the two sets of time series are not comparable
-directly because not all cases will be recorded. We thus model recorded cases as depicted
-in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the share of
-recorded cases and use data on whether the reason for a PCR test were symptoms of
-CoViD-19 to get age-specific shares. For example,\comment[id=HM]{@Janos, please correct
+Material~\ref{subsec:infection_probs}. We estimate these parameters so that age-specific
+simulated infections predict the corresponding time series in the data \citep[see][for
+the general method]{McFadden1989}. However, the two sets of time series are not
+comparable directly because not all cases will be recorded. We thus model recorded cases
+as depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of
+the share of recorded cases and use data on whether CoViD-19 symptoms led to a PCR test
+in order to get age-specific shares. For example,\comment[id=HM]{@Janos, please correct
 if necessary and insert reference} estimate unknown cases to be twice as high for
@@
 
-Our analysis\comment[id=J]{The conclusion should be slightly longer and more separated from the rest} has shown that in the transition path to high levels of vaccination and
+Our analysis\comment[id=J]{The conclusion should be slightly longer and more separated
+from the rest} has shown that in the transition path to high levels of vaccination and
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
@@
 cost 6.25~billion~Euros \citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo
-Instute back from March 2020 assuming a 2 month lockdown. I did not find anything citable
-that is more recent. I did find
+Instute back from March 2020 assuming a 2 month lockdown. I did not find anything
+citable that is more recent. I did find
 \href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
@@
 Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
-for self-administered rapid tests are frequently less than 1~Euro.\comment[id=HM]{Cheapest on
-\href{https://amazon.de}{Amazon} on June 11th, 2021. --- DM 95ct}
+for self-administered rapid tests are frequently less than
+1~Euro.\comment[id=HM]{Cheapest on \href{https://amazon.de}{Amazon} on June 11th, 2021.
+--- DM 95ct}

### 2021-06-16 figures/results/tables/student_infections_table.tex
Add new figures with fixed bug in the mandatory work test scenario and fixed vaccination data (i.e. Fig 2c).

@@
 \midrule
- educ open after easter  without tests &                                             705858 \\
- educ open after easter  with tests    &                                             539545 \\
- close educ after easter               &                                             414920 \\
- baseline                              &                                             425200 \\
+ educ open after easter  without tests &                                             808831 \\
+ educ open after easter  with tests    &                                             627182 \\
+ close educ after easter               &                                             478064 \\
+ baseline                              &                                             490880 \\
 \bottomrule

### 2021-06-16 paper/appendix_data.tex
Red texts -> comments, except for some ldots in figurenotes.
Move estimated parameters to model section. Address some comments.

@@
 
-\subsection{The Synthetic Population}
-
-\textcolor{red}{\ldots}
+\subsection{The Synthetic Population\comment[id=K]{To be written}}
 
@@
 \end{figure}
-
-\textcolor{red}{Arbeitsplots auch in eine gemeinsame Graphik packen}
 
@@
 
-\FloatBarrier
-
-\subsection{Infection Probabilities\comment[id=HM]{Is the data section the right place?}}
-\label{subsec:infection_probs}
-
-To calibrate infection probabilities outside of the model, it would be important to know
-the exact duration and distance of each contact type as well as viral loads. Since this
-is not available in any data set, we estimate those parameters inside the model with the
-method of simulated moments \citep{McFadden1989} by minimizing the distance between
-simulated and observed infection rates. Since our model includes a lot of randomness, we
-average simulated infection rates over several model runs.
-
-We use data for Germany from October 2020 until June 2021. We do not use
-earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
-limited and therefore it would be difficult to find good initial conditions for our
-simulations. In addition during the summer the case numbers were extremely low. This
-could lead to the epidemic going extinct in our simulation. Additionally, our model does
-not include international travel or other imports of cases. These would be important but
-difficult to model during the summer months.
-
-To avoid over-fitting and simplify the numerical optimization problem, we only allow for
-five different probabilities: 1) for contacts in schools 2) contacts in preschools and
-nurseries. 3) for work contacts. 4) for households. 5) for other contacts.
-
-The infection probabilities estimated from our model are as follows: For household
-contacts (each individual meets all his or her household members every day) the infection
-probability is 10\%, for school contacts (all students and teachers meet all other
-students and teachers that attend school that day and there are three classes taking
-place each day) the probability of infection is 1.2\%, for preschools and nurseries that
-infection probability is 0.5\%. Work contacts carry a risk of 14.5\% and other activities
-(including leisure activities such as meeting friends) carry the highest risk with
-15.875\%.\comment[id=HM]{Are these numbers current? In any case, a table would substitute for this and the previous paragraph. Need to mention B.1.1.7 alongside.}
-
 \subsection{Policies and Seasonality}
@@
 light in Germany, hygiene measures (such as masks, ventilation and hand washing) became
-more strict and more conscientiously observed, leading to a reduction of 33\% in the number of contacts with the potential to transmit Covid-19.
+more strict and more conscientiously observed, leading to a reduction of 33\% in the
+number of contacts with the potential to transmit Covid-19.
 
@@
 to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays.
-Because of the scarce data situation we cannot distinguish between a hygiene factor (such as mask wearing) during meetings and physical distancing (such as virtual meetings with friends).
+events such as policy announcements or particular holidays. Because of the scarce data
+situation we cannot distinguish between a hygiene factor (such as mask wearing) during
+meetings and physical distancing (such as virtual meetings with friends).
 
@@
 \FloatBarrier
-
-\subsection{Rapid Test Demand}
 

### 2021-06-16 paper/appendix_results.tex
Move estimated parameters to model section. Address some comments.

@@
 rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups.\comment[id=J]{summarize the fit; Some of this needs
-to go in the appendix. I suggest the age group and federal state fit goes there???}
+analyze the model fit for age groups.\comment[id=J]{summarize the
+fit}\comment[id=HM]{Also need to show simulated total infections somewhere. So far only
+ever shown for 2021.}
 
@@
   \caption{Simulated and Empirical Infections}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates.}
+  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
+  reported versus the simulated infections rates.}
   \label{fig:aggregated_fit2}
@@
   \caption{Simulated and Empirical Infections by Age Group}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different age groups.}
+  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
+  reported versus the simulated infections rates for different age groups.}
   \label{fig:age_group_fit}
@@
   \caption{Simulated and Empirical Infections by Federal State}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different federal states.}
+  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
+  reported versus the simulated infections rates for different federal states.}
   \label{fig:state_fit}
@@
 
-\subsection{Share of Cases that are Detected}
+\subsection{Share of Cases that are Detected  \comment[id=K]{Figure notes missing}}
 \label{subsec:appendix_share_known_cases}
@@
   \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent \textcolor{red}{\ldots}}
+  \floatfoot{\noindent}
 \end{figure}
 
-It's noteworthy that the share of detected cases increases rapidly in May for the five to fourteen year olds.
-This is a direct result of the mandatory tests in school.\comment[id=K]{We don't do this at the moment but we could also create plots showing the share of detected cases between scenarios - either for the full population or for a age group of interest.}
+It's noteworthy that the share of detected cases increases rapidly in May for the five to
+fourteen year olds. This is a direct result of the mandatory tests in
+school.
 
-\subsection{Rapid Tests}
+\subsection{Rapid Tests \comment[id=K]{The figures will likely change layout later}}
 \label{subsec:appendix_rapid_tests}
@@
   \label{fig:rapid_tests_by_reason}
-  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without PCR demand and other contact demand together form the private demand category.
-Also note that these do not add up to the total share of demanded rapid tests as individuals may have more than one reason to demand a rapid test on any given day.
-\textcolor{red}{\ldots}}
+  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without
+  PCR demand and other contact demand together form the private demand category. Also
+  note that these do not add up to the total share of demanded rapid tests as individuals
+  may have more than one reason to demand a rapid test on any given day. }
 \end{figure}
@@
   \label{fig:share_of_rapid_tests_by_reason_by_infected}
-  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without PCR demand and other contact demand together form the private demand category.
-Also note that this is not the same as individuals getting a positive rapid test.
-The sensitivity is quite low before individuals become infectious.
-Therefore, if individuals are still in the latent period of their infection they are likely to get a false positive rapid test.
-\textcolor{red}{\ldots}}
+  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without
+  PCR demand and other contact demand together form the private demand category. Also
+  note that this is not the same as individuals getting a positive rapid test. The
+  sensitivity is quite low before individuals become infectious. Therefore, if
+  individuals are still in the latent period of their infection they are likely to get a
+  false positive rapid test.}
 \end{figure}

### 2021-06-16 paper/model/estimated_parameters.tex
Move estimated parameters to model section. Address some comments.

\subsection{Estimated Parameters}
\label{subsec:estimated_params}

%%% The free parameters are the infection probabilities by contact category, one hygiene
%%% multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
%%% that govern the reduction of other contacts\comment[id=K]{Must be updated when the
%%% estimation is finished.}, the number of extra contacts during holidays, the share of
%%% detected cases around holidays and the fade in speed of the rapid tests as well as one
%%% parameter that governs the import of B.1.1.7 cases in January 2021.

\FloatBarrier

% infection probabilities

To calibrate infection probabilities outside of the model, it would be important to know
the exact duration and distance of each contact type as well as viral loads. Since this
is not available in any data set, we estimate those parameters inside the model with the
method of simulated moments \citep{McFadden1989} by minimizing the distance between
simulated and observed infection rates. Since our model includes a lot of randomness, we
average simulated infection rates over several model runs.

We use data for Germany from October 2020 until June 2021. We do not use
earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
limited and therefore it would be difficult to find good initial conditions for our
simulations. In addition during the summer the case numbers were extremely low. This
could lead to the epidemic going extinct in our simulation. Additionally, our model does
not include international travel or other imports of cases. These would be important but
difficult to model during the summer months.

To avoid over-fitting and simplify the numerical optimization problem, we only allow for
five different probabilities: 1) for contacts in schools 2) contacts in preschools and
nurseries. 3) for work contacts. 4) for households. 5) for other contacts.

The infection probabilities estimated from our model are as follows: For household
contacts (each individual meets all his or her household members every day) the infection
probability is 10\%, for school contacts (all students and teachers meet all other
students and teachers that attend school that day and there are three classes taking
place each day) the probability of infection is 1.2\%, for preschools and nurseries that
infection probability is 0.5\%. Work contacts carry a risk of 14.5\% and other activities
(including leisure activities such as meeting friends) carry the highest risk with
15.875\%.\comment[id=HM]{Are these numbers current? In any case, a table would substitute for this and the previous paragraph. Need to mention B.1.1.7 alongside.}\comment[id=K]{Missing to add data for the other estimated parameters (import of b117 etc)}

### 2021-06-16 paper/report.tex
Move estimated parameters to model section. Address some comments.

@@
 
-Since early 2020,\comment[id=J]{\textbf{Features that are not highlighted enough:}
-
-\begin{itemize}
-    \item Low sensitivity of rapid tests before and on day of infectiousness onset
-    \item Different seasonalities for different contact types (and quite strong for other contacts)
-    \item Low number of breakpoints in parameters / policies compared to breakpoints in
-    cases / reproduction rate
-\end{itemize}
-
-} the CoViD-19 pandemic has presented an enormous challenge to humanity
-on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many
-months---and often years---away from reaching vaccination-induced herd immunity
-\citep{Swaminathan2021}.\comment[id=HM]{Need to find some more scientific citation
-before submission} In the meantime, it is of utmost importance to employ an effective
-mix of strategies for containing the virus. The most frequent initial response was a set
-of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
-this has allowed some countries to sustain equilibria with very low infection
-numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
-numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
-fluctuations of infection rates over time. Containment measures have become increasingly
-diverse and now include testing, more nuanced NPIs, and contact tracing. Neither these
-policies' effect nor the influence of seasonal patterns or more infectious virus strains
-are well understood in quantitative terms.
+Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
+sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
+presented an enormous challenge to humanity on many dimensions. The development of highly
+effective vaccines holds the promise of containment in the medium term. However, most
+countries find themselves many months---and often years---away from reaching
+vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
+some more scientific citation before submission} In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate
+policies.}, most have seen large fluctuations of infection rates over time. Containment
+measures have become increasingly diverse and now include testing, more nuanced NPIs, and
+contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
+more infectious virus strains are well understood in quantitative terms.
 
@@
 some days to be processed and there are aggregate capacity constraints throughout. In
-contrast, rapid antigen tests yield immediate results; after a phase-in period, all
-tests that are demanded will be performed. Specificity and sensitivity of these tests is
-set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
-the timing of the test relative to the start of infectiousness.
-Figure~\ref{fig:antigen_tests} shows our model for rapid test demand. Schools may
-require students to be tested regularly. Rapid tests may be offered by employers for
-on-site workers. Individuals may demand tests for private reasons, which include having
-plans to meet other people\footnote{A positive test will make them reduce their
-contacts; this is why tests impact the actual contacts in
-Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
-household member tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons and whether she reduces her contacts if
-she tests positive or develops symptoms. The thresholds are higher for tests in a
-private setting than for tests at the workplace until mid April and then compliance
-because of private reasons becomes higher than compliance at the
-workplace.\comment[id=J]{Shouldn't we mention that this is because of availability /
-high costs?}
+contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
+that are demanded will be performed. Specificity and sensitivity of these tests is set
+according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
+timing of the test relative to the start of infectiousness.
+Figure~\ref{fig:antigen_tests} shows our model for rapid test demand. Schools may require
+students to be tested regularly. Rapid tests may be offered by employers for on-site
+workers. Individuals may demand tests for private reasons, which include having plans to
+meet other people\footnote{A positive test will make them reduce their contacts; this is
+why tests impact the actual contacts in Figure~\ref{fig:model-description}.}, showing
+symptoms of CoViD-19, and because a household member tested positively for the virus. We
+endow each agent with an individual compliance parameter. This parameter determines
+whether she takes up rapid tests offered by employers or follows up on private reasons
+and whether she reduces her contacts if she tests positive or develops symptoms. The
+thresholds fall over time and are calibrated to survey data where possible (see
+Supplementary Materials~\ref{subsec:rapid_test_demand}).
+% The thresholds are higher for tests in a private setting than for tests at the
+% workplace until mid April and then compliance because of private reasons becomes higher
+% than compliance at the workplace.
 
@@
 
-The free parameters are the infection probabilities by contact type, see Supplementary
-Material~\ref{subsec:infection_probs}. We estimate these parameters so that age-specific
-simulated infections predict the corresponding time series in the data \citep[see][for
-the general method]{McFadden1989}. However, the two sets of time series are not
-comparable directly because not all cases will be recorded. We thus model recorded cases
-as depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of
-the share of recorded cases and use data on whether CoViD-19 symptoms led to a PCR test
-in order to get age-specific shares. For example,\comment[id=HM]{@Janos, please correct
-if necessary and insert reference} estimate unknown cases to be twice as high for
-children as for adults. Confirmation of a positive rapid test result is another reason
-for PCR tests, once these become available.\comment[id=J]{Free parameters also include:
+The free parameters are the infection probabilities by contact category, one hygiene
+multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
+that govern the reduction of other contacts\comment[id=K]{Must be updated when the
+estimation is finished.}, the number of extra contacts during holidays, the share of
+detected cases around holidays and the fade in speed of the rapid tests as well as one
+parameter that governs the import of B.1.1.7 cases in January 2021. For details, see
+Supplementary Material~\ref{subsec:estimated_params},~\ref{subsec:rapid_test_demand}.
 
-- One hygiene multiplier for school an work contacts after November 2
-- A few "other" multipliers including two lockdown fatigue effects
-- Number of extra contacts during holidays
-
-Strictly speaking also:
-- The fade in speed of the rapid tests
-- Share known cases around holidays }
+We estimate these parameters such that the model matches the following empirical
+quantities: 1. official case numbers for each age group and region, 2. deaths, 3. share
+of B.1.1.7. \citep[See][for the general method]{McFadden1989}. However, only a fraction
+of infections actually enter the official case numbers. We thus model official cases as
+depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the
+share of detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As
+the share of asymptomatic individuals varies by age group, this gives us age-specific
+shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases
+by age group over time in our model). Our estimates suggest that only 18 to 28\% of cases
+are detected for school age children while the rate is 30 to 55\% for those above age
+80.\comment[id=K]{Must be verified once results are final.} Confirmation of a positive
+rapid test result is another reason for PCR tests, once these become available.
 
@@
 workplaces. In the seven weeks between mid September and early November, cases increased
-by a factor of 10. Restrictions were somewhat tightened in mid-October and again in
-early November. New infections remained constant throughout November, before rising
-again in December, which prompted the most stringent lockdown to this date. Schools and
-daycare centers were closed again, so were customer-facing businesses except for grocery
-and drug stores. From the peak of the second wave just before Christmas until the trough
-in mid-February, newly detected cases decreased by almost three quarters. The third wave
-in the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in
-March. See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed;
-e.g., hairdressers and home improvement stores were allowed to open again to the public.
-There were many changes in details of regulations afterwards, but they did not change
-the stringency index.
+by a factor of 10. Restrictions were somewhat tightened in mid-October and again in early
+November. New infections remained constant throughout November, before rising again in
+December, which prompted the most stringent lockdown to this date. Schools and daycare
+centers were closed again, so were customer-facing businesses except for grocery and drug
+stores. From the peak of the second wave just before Christmas until the trough in
+mid-February, newly detected cases decreased by almost three quarters. The third wave in
+the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in March.
+See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
+hairdressers and home improvement stores were allowed to open again to the public. There
+were many changes in details of regulations afterwards, but they did not change the
+stringency index.
 
@@
 September 2020 and use the model to predict recorded infection rates until the end of May
-2021.\comment[id=J]{I do not understand} See Supplementary
+2021.\comment[id=J]{This needs more explanation. I will rewrite this.} See Supplementary
 Materials~\ref{sec:data_and_parameters} and \ref{sec:model} for detailed descriptions of

### 2021-06-17 paper/model/summary.tex
Got started re-writing with Radost's comments on the side.

@@
 
-The structure of the model is depicted in Figure~\ref{fig:broad_model_description}.
+The structure of the model is depicted in Figure~\ref{fig:model_contacts_infections}.
+\comment[id=HM]{Mentioned by Tim in the course, just main paper: Modular structure may deserve a sentence}
 
@@
 predicted number of contacts. Currently, we distinguish between eight types of contact
-models which are all listed in Figure~\ref{fig:broad_model_description}: households,
+models which are all listed in Figure~\ref{fig:model_contacts_infections}: households,
 recurrent and random work contacts, recurrent and random leisure contacts, and nursery,
@@
 Once a person is infected, the disease progresses in a fairly standard way which is
-depicted in Figure~\ref{fig:disease_progression}. Asymptomatic cases and cases with mild
+depicted in Figure~\ref{fig:model_disease_progression}. Asymptomatic cases and cases with mild
 symptoms are infectious for some time and recover eventually. Cases with severe symptoms

### 2021-06-17 paper/report.tex
Got started re-writing with Radost's comments on the side.

@@
 contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical
+with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
 equilibrium at low case numbers which is sustained with test-trace-and-isolate
-policies.}, most have seen large fluctuations of infection rates over time. Containment
+policies.} most have seen large fluctuations of infection rates over time. Containment
 measures have become increasingly diverse and now include testing, more nuanced NPIs, and
@@
 
-This paper develops a quantitative model incorporating all these factors. The framework
-allows to combine a wide variety of data and mechanisms in a timely fashion, making it
-useful to predict the effects of various interventions. We apply the model to Germany
-and show that rapid testing had the largest impact on the reduction in infections by
-almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
-to play for at least as long as vaccinations have not been offered to an entire
-population.
+This paper develops a quantitative model incorporating these factors simultaneously. The
+framework allows to combine a wide variety of data and mechanisms in a timely fashion,
+making it useful to predict the effects of various interventions. We apply the model to
+Germany and show that rapid testing had the largest impact on the reduction in
+infections by almost 80\% during the month of May 2021. We conclude that rapid tests
+have a large role to play for at least as long as vaccinations have not been offered to
+an entire population.
 
 At the core of our agent-based model are physical contacts between heterogeneous agents
-(Figure~\ref{fig:broad_model_description}).\footnote{A detailed comparison with other
+(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
 approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
@@
 between an infectious individual and somebody susceptible to the disease bears the risk
-of transmitting the virus. Contacts occur in the household, at work, at school, or in
-other settings (leisure activities, grocery shopping, medical appointments, etc.). Some
-contacts recur regularly, others occur at random. Random contacts are typically
-assortative in age and geographical location. Empirical applications can take the
-population structure from census data and the types and frequency of contacts from diary
-data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008,Hoang2019}.
+of transmitting the virus. Contacts occur in up to four networks: Within the household,
+at work, at school, or in other settings (leisure activities, grocery shopping, medical
+appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
+applications can take the population and household structure from census data and the
+network-specific frequencies of contacts from diary data measuring contacts before the
+pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
+differentiated by age and geographical location see Section~\ref{sub:number_of_contacts}
+of the Supplementary Material.
 
-The three dimensions of contact types are chosen so that the most common NPIs can be
+The four contact networks are chosen so that the most common NPIs can be
 modeled in great detail. NPIs affect the number of contacts or the risk of transmitting
@@
 daycare can be closed entirely, operate at reduced capacity---including an alternating
-schedule---,or implement mitigation measures like masking requirements or air filters
+schedule---, or implement mitigation measures like masking requirements or air filters
 \citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
-        \caption{{Model description}}
-        \label{fig:broad_model_description}
+        \caption{{Model for contacts and infections}}
+        \label{fig:model_contacts_infections}
     \end{subfigure}
@@
         \caption{Disease progression}
-        \label{fig:disease_progression}
+        \label{fig:model_disease_progression}
     \end{subfigure}
@@
         \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
-        \caption{{Model for antigen tests}}
-        \label{fig:antigen_tests}
+        \caption{{Model for rapid tests}}
+        \label{fig:model_rapid_tests}
     \end{subfigure}
@@
         \caption{{Translating all infections to recorded ones}}
-        \label{fig:model_for_official_cases}
+        \label{fig:model_official_cases}
     \end{subfigure}
@@
     \caption{Model description}
-    \label{fig:model-description}
+    \label{fig:model_description}
 
@@
         description of the model can be found in Supplementary Material~\ref{sec:model}.
-        Figure~\ref{fig:broad_model_description} shows the influence on an agent's
+        Figure~\ref{fig:model_contacts_infections} shows the influence on an agent's
         contacts to other agents. Demographic characteristics set the baseline number of
@@
         depends on the type of contact, on seasonality, and on NPIs. If infected, the
-        infection progresses as depicted in Figure~\ref{fig:disease_progression}. If
+        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}. If
         rapid tests are avaible, agents' demand is modeled as in
-        Figure~\ref{fig:antigen_tests}. All pupils are attending school are tested,
+        Figure~\ref{fig:model_rapid_tests}. All pupils are attending school are tested,
         other reasons trigger a test only for a fraction of individuals.
-        Figure~\ref{fig:model_for_official_cases} shows the model of translating all
+        Figure~\ref{fig:model_official_cases} shows the model of translating all
         infections in the simulated data to age-specific recorded infections. The model
@@
 In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
-possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
+possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
 differentiate between an initial period of infection without being infectious or showing
@@
 some days to be processed and there are aggregate capacity constraints throughout. In
-contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
-that are demanded will be performed. Specificity and sensitivity of these tests is set
-according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the start of infectiousness.
-Figure~\ref{fig:antigen_tests} shows our model for rapid test demand. Schools may require
-students to be tested regularly. Rapid tests may be offered by employers for on-site
-workers. Individuals may demand tests for private reasons, which include having plans to
-meet other people\footnote{A positive test will make them reduce their contacts; this is
-why tests impact the actual contacts in Figure~\ref{fig:model-description}.}, showing
-symptoms of CoViD-19, and because a household member tested positively for the virus. We
-endow each agent with an individual compliance parameter. This parameter determines
-whether she takes up rapid tests offered by employers or follows up on private reasons
-and whether she reduces her contacts if she tests positive or develops symptoms. The
-thresholds fall over time and are calibrated to survey data where possible (see
-Supplementary Materials~\ref{subsec:rapid_test_demand}).
+contrast, rapid antigen tests yield immediate results; after a phase-in period, all
+tests that are demanded will be performed. Specificity and sensitivity of these tests is
+set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
+the timing of the test relative to the onset of infectiousness.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require students to be tested regularly. Rapid tests may be offered by employers for
+on-site workers. Individuals may demand tests for private reasons, which include having
+plans to meet other people,\footnote{A positive test will make them reduce their
+contacts; this is why tests impact the actual contacts in
+Figure~\ref{fig:model_description}.} showing symptoms of CoViD-19, and because a
+household member tested positively for the virus. We endow each agent with an individual
+compliance parameter. This parameter determines whether she takes up rapid tests offered
+by employers or follows up on private reasons.\comment[id=HM]{Removed ``and whether she
+reduces her contacts if she tests positive or develops symptoms'' -- does not fit here.
+If anything, add footnote, but it seems like this is a detail that is just distracting
+right here. Make sure to describe clearly and at length in Appendix.}
+
+% The thresholds fall over time and are calibrated to survey data where possible (see
+% Supplementary Materials~\ref{subsec:rapid_test_demand}).
+
 % The thresholds are higher for tests in a private setting than for tests at the
-% workplace until mid April and then compliance because of private reasons becomes higher
-% than compliance at the workplace.
+% workplace until mid April and then compliance because of private reasons becomes
+% higher than compliance at the workplace.
 
@@
 for different contact types and their assortativity by age group. Mobility data is used
-to model the reductions in work contacts. School and daycare policies can be incorporated
+to model the evolution of work contacts. School and daycare policies can be incorporated
 directly from official directives. Administrative records on the number of tests,
@@
 of infections actually enter the official case numbers. We thus model official cases as
-depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the
+depicted in Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the
 share of detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As
@@
 Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. We transform the index so that lower values represent
-higher levels of restrictions. A value of zero means all measures incorporated in the
-index are turned on. The value 1 represents the situation in mid-September, with
-restrictions on gatherings and public events, masking requirements, but open schools and
-workplaces. In the seven weeks between mid September and early November, cases increased
-by a factor of 10. Restrictions were somewhat tightened in mid-October and again in early
-November. New infections remained constant throughout November, before rising again in
-December, which prompted the most stringent lockdown to this date. Schools and daycare
-centers were closed again, so were customer-facing businesses except for grocery and drug
-stores. From the peak of the second wave just before Christmas until the trough in
-mid-February, newly detected cases decreased by almost three quarters. The third wave in
-the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in March.
-See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-stringency index.
+non-pharmaceutical interventions. For legibility, we transform the index so that lower
+values represent higher levels of restrictions. A value of zero means all measures
+incorporated in the index are turned on. The value 1 represents the situation in
+mid-September, with restrictions on gatherings and public events, masking requirements,
+but open schools and workplaces. In the seven weeks between mid September and early
+November, cases increased by a factor of 10. Restrictions were somewhat tightened in
+mid-October and again in early November. New infections remained constant throughout
+November, before rising again in December, which prompted the most stringent lockdown to
+this date. Schools and daycare centers were closed again, so were customer-facing
+businesses except for grocery and drug stores. From the peak of the second wave just
+before Christmas until the trough in mid-February, newly detected cases decreased by
+almost three quarters. The third wave in the spring of 2021 is associated with the
+B.1.1.7 strain, which became dominant in March. See Figure~\ref{fig:share_b117}. In
+early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
+stores were allowed to open again to the public. There were many changes in details of
+regulations afterwards, but they did not change the stringency index.
 
@@
 that only a small fraction of the population performed weekly rapid tests in
-March\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
+March,\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
 week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
 respondents to the COSMO study reported in early March that they \textit{ever} did a
-rapid test)}, new infections on April 1 would be reduced by 53\% relative to the
+rapid test)} new infections on April 1 would be reduced by 53\% relative to the
 scenario without vaccinations, rapid tests, or seasonality.

### 2021-06-17 paper/model/estimated_parameters.tex
Put estimated infection probabilities into a table.

@@
 
-% infection probabilities
-
-To calibrate infection probabilities outside of the model, it would be important to know
-the exact duration and distance of each contact type as well as viral loads. Since this
-is not available in any data set, we estimate those parameters inside the model with the
+We estimate parameters that could not be calibrated inside the model with the
 method of simulated moments \citep{McFadden1989} by minimizing the distance between
@@
 
+% infection probabilities
+
 To avoid over-fitting and simplify the numerical optimization problem, we only allow for
@@
 
-The infection probabilities estimated from our model are as follows: For household
-contacts (each individual meets all his or her household members every day) the infection
-probability is 10\%, for school contacts (all students and teachers meet all other
-students and teachers that attend school that day and there are three classes taking
-place each day) the probability of infection is 1.2\%, for preschools and nurseries that
-infection probability is 0.5\%. Work contacts carry a risk of 14.5\% and other activities
-(including leisure activities such as meeting friends) carry the highest risk with
-15.875\%.\comment[id=HM]{Are these numbers current?
-In any case, a table would substitute for this and the previous paragraph.
-Need to mention B.1.1.7 alongside.}\comment[id=K]{Missing to add data for the other estimated parameters (import of b117 etc)}
+The infection probabilities estimated from our model are as shown in
+Table~\ref{tab:infection_probs}.\comment[id=K]{Missing to add data for the other
+estimated parameters (import of b117 etc)}\comment[id=K]{Update table after the new
+estimation is finished}
+
+\begin{table}[tb]
+    \caption{Estimated Infection Probabilities}
+    \label{tab:infection_probs}
+    \centering
+    \begin{tabular}{ll}
+        \toprule
+        Contact Type & Infection Probability \\
+        \midrule
+        preschool and nursery & 00.500\% \\
+        school & 01.200\% \\
+        household & 10.000\% \\
+        work & 14.500\% \\
+        other & 15.875\% \\
+        \bottomrule
+    \end{tabular}
+    \tablenotes{School contacts take place Monday through Friday. All students and
+teachers meet all other students and teachers that attend school that day and there are
+three classes taking place each day. Work contacts can be daily, weekly or non recurrent.
+Other contacts can be daily, weekly or non recurrent. We do not differ the infection
+probability depending on how often a contact takes place.}
+\end{table}

### 2021-06-17 paper/appendix_data.tex
Polish model summary.

@@
         For the recurrent other contacts, individuals are assigned to groups that are
-        time constant and that meet daily or weekly. The share of individuals that attend
+        time constant and that meet daily or weekly. The share of individuals who attend
         in a way that has transmission potential is reduced by policies, seasonality and
@@
         You can think of these as for example colleagues with which one shares an office
-        space. The share of individuals that attend in a way that has transmission potential
+        space. The share of individuals who attend in a way that has transmission potential
         is reduced by policies (such as a work from home mandate), seasonality and individual
@@
 
-All shares of individuals that would take a rapid test if the conditions were met can be
+All shares of individuals who would take a rapid test if the conditions were met can be
 seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
@@
     \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of Individuals that do a Rapid Test.}}
+    \caption{\textbf{Share of individuals who do a Rapid Test.}}
     \floatfoot{\noindent ...}

### 2021-06-17 paper/abstract.tex
No obvious stuff left, at least nothing obvious at this time of the night.

@@
 the pandemic, allowing for factors like seasonality or virus strains with different
-transmission profiles. In order to do so, the study develops an agent-based simulation
-model that is estimated on data for the second and the third wave of the CoViD-19
+transmission profiles. To do so, the study develops an agent-based simulation
+model, which is estimated using data for the second and the third wave of the CoViD-19
 pandemic in Germany. The paper finds that during a period where vaccination rates rose
@@
 % the Abstract. 
-Frequent, large-scale rapid testing should remain part of strategies to contain
+Frequent large-scale rapid testing should remain part of strategies to contain
 CoViD-19; it can substitute for many non-pharmaceutical interventions that come at a

### 2021-06-17 paper/model/course_of_disease.tex
Got started re-writing with Radost's comments on the side.
No obvious stuff left, at least nothing obvious at this time of the night.

@@
 systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
-progresses in the way depicted in Figure~\ref{fig:disease_progression}.
+progresses in the way depicted in Figure~\ref{fig:model_disease_progression}.
 
@@
 remains in a state is drawn randomly. The parameters that govern these processes are
-taken from the literature\footnotemark and age-dependent.
-
-\footnotetext{ Detailed information on the calibration of the disease parameters is
-    available as part of our
+taken from the literature \footnote{ Detailed information on the calibration of the
+disease parameters is available as part of our
 \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-    documentation}. }
+documentation}.} and age-dependent.

### 2021-06-17 paper/model/policies.tex
No obvious stuff left, at least nothing obvious at this time of the night.

@@
-\subsection{Reducing Numbers of Contacts Through Policies}
+\subsection{Reducing Numbers of Contacts via NPIs}
 \label{sec:policies}
@@
 thirds of the working population have zero work contacts and the rest has the same
-number of contacts as before.
+number of contacts as before.\comment[id=HM]{Mention here or in results that school
+closures $\neq$ vacations because of additional contacts}
 

### 2021-06-17 paper/report.tex
No obvious stuff left, at least nothing obvious at this time of the night.

@@
 sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
-presented an enormous challenge to humanity on many dimensions. The development of highly
-effective vaccines holds the promise of containment in the medium term. However, most
-countries find themselves many months---and often years---away from reaching
+presented an enormous challenge to humanity on many dimensions. The development of
+highly effective vaccines holds the promise of containment in the medium term. However,
+most countries find themselves many months---and often years---away from reaching
 vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
@@
 frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate
-policies.} most have seen large fluctuations of infection rates over time. Containment
-measures have become increasingly diverse and now include testing, more nuanced NPIs, and
-contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
-more infectious virus strains are well understood in quantitative terms.
+contacts between individuals. While this has allowed some countries to sustain
+equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
+theoretical equilibrium at low case numbers which is sustained with
+test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
+over time. Containment measures have become increasingly diverse and now include
+testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
+influence of seasonal patterns or more infectious virus strains are well understood in
+quantitative terms.
 
@@
 making it useful to predict the effects of various interventions. We apply the model to
-Germany and show that rapid testing had the largest impact on the reduction in
-infections by almost 80\% during the month of May 2021. We conclude that rapid tests
-have a large role to play for at least as long as vaccinations have not been offered to
+Germany, where new infections fell by almost 80\% during the month of May 2021. Our
+analysis shows that frequent and large-scale rapid testing caused the bulk of this
+decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that it
+should have a large role for at least as long as vaccinations have not been offered to
 an entire population.
@@
 
-The four contact networks are chosen so that the most common NPIs can be
-modeled in great detail. NPIs affect the number of contacts or the risk of transmitting
-the disease upon having physical contact. The effect of different NPIs will generally
-vary across contact types. For example, a mandate to work from home will reduce the
-number of work contacts to zero for a fraction of the working population. Schools and
-daycare can be closed entirely, operate at reduced capacity---including an alternating
-schedule---, or implement mitigation measures like masking requirements or air filters
+The four contact networks are chosen so that the most common NPIs can be modeled in
+great detail. NPIs affect the number of contacts or the risk of transmitting the disease
+upon having physical contact. The effect of different NPIs will generally vary across
+contact types. For example, a mandate to work from home will reduce the number of work
+contacts to zero for a fraction of the working population. Schools and daycare can be
+closed entirely, operate at reduced capacity---including an alternating schedule---, or
+implement mitigation measures like masking requirements or air filters
 \citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
@@
         depends on the type of contact, on seasonality, and on NPIs. If infected, the
-        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}. If
-        rapid tests are avaible, agents' demand is modeled as in
+        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}.
+        If rapid tests are avaible, agents' demand is modeled as in
         Figure~\ref{fig:model_rapid_tests}. All pupils are attending school are tested,
@@
 
-% The thresholds fall over time and are calibrated to survey data where possible (see
-% Supplementary Materials~\ref{subsec:rapid_test_demand}).
-
-% The thresholds are higher for tests in a private setting than for tests at the
-% workplace until mid April and then compliance because of private reasons becomes
-% higher than compliance at the workplace.
-
 Modelling a population of agents according to actual demographic characteristics means
@@
 available. Surveys may ask about test offers, propensities to take them up, and past
-tests.
+tests. Other studies' estimates of the seasonality of infections can be incorporated
+directly. The remaining parameters---most notably, these include infection probabilities
+by contact network and the effects of NPIs, see Supplementary
+Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
+matches features of the data \citep[see][for the general method, also described in
+Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
+more specific once Appendix has new structure}. We keep the number of free parameters
+low in order to avoid overfitting. In our application, these features include  official
+case numbers for each age group and region, deaths, and the share of the B.1.1.7 strain.
 
-The free parameters are the infection probabilities by contact category, one hygiene
-multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
-that govern the reduction of other contacts\comment[id=K]{Must be updated when the
-estimation is finished.}, the number of extra contacts during holidays, the share of
-detected cases around holidays and the fade in speed of the rapid tests as well as one
-parameter that governs the import of B.1.1.7 cases in January 2021. For details, see
-Supplementary Material~\ref{subsec:estimated_params},~\ref{subsec:rapid_test_demand}.
-
-We estimate these parameters such that the model matches the following empirical
-quantities: 1. official case numbers for each age group and region, 2. deaths, 3. share
-of B.1.1.7. \citep[See][for the general method]{McFadden1989}. However, only a fraction
-of infections actually enter the official case numbers. We thus model official cases as
-depicted in Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the
-share of detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As
-the share of asymptomatic individuals varies by age group, this gives us age-specific
-shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases
-by age group over time in our model). Our estimates suggest that only 18 to 28\% of cases
-are detected for school age children while the rate is 30 to 55\% for those above age
-80.\comment[id=K]{Must be verified once results are final.} Confirmation of a positive
-rapid test result is another reason for PCR tests, once these become available.
+The main issue with official case numbers is that they will contain only a fraction of
+all infections. We thus model official cases as depicted in
+Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
+detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As the share
+of asymptomatic individuals varies by age group, this gives us age-specific shares (see
+Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age
+group over time in our model). Our estimates suggest that---in the absence of rapid
+testing---the detection rate is two to three times higher for individuals above age 80
+compared to school age children.\comment[id=K]{Must be verified once results are final.}
+Once rapid test become available, confirmation of a positive result is another reason
+leading to positive PCR tests.
 
@@
 Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. For legibility, we transform the index so that lower
+non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
+never use it directly. For legibility reasons, we transform the index so that lower
 values represent higher levels of restrictions. A value of zero means all measures
@@
 mid-October and again in early November. New infections remained constant throughout
-November, before rising again in December, which prompted the most stringent lockdown to
-this date. Schools and daycare centers were closed again, so were customer-facing
-businesses except for grocery and drug stores. From the peak of the second wave just
-before Christmas until the trough in mid-February, newly detected cases decreased by
-almost three quarters. The third wave in the spring of 2021 is associated with the
-B.1.1.7 strain, which became dominant in March. See Figure~\ref{fig:share_b117}. In
-early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
-stores were allowed to open again to the public. There were many changes in details of
-regulations afterwards, but they did not change the stringency index.
+November before rising again in December, prompting the most stringent lockdown to this
+date. Schools and daycare centers were closed again, so were customer-facing businesses
+except for grocery and drug stores. From the peak of the second wave just before
+Christmas until the trough in mid-February, newly detected cases decreased by almost
+three quarters. The third wave in the spring of 2021 is associated with the B.1.1.7
+strain, which became dominant in March (Figure~\ref{fig:share_b117}). In early March,
+some NPIs were being relaxed; e.g., hairdressers and home improvement stores were
+allowed to open again to the public. There were many changes in details of regulations
+afterwards, but they did not change the stringency index.
 
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
 
-        \caption{{Tests and vaccinations}}
-        \label{fig:antigen_tests_vaccinations}
+        \caption{Fraction of B.1.1.7 strain}
+        \label{fig:share_b117}
     \end{subfigure}
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
 
-        \caption{Fraction of B.1.1.7 strain}
-        \label{fig:share_b117}
+        \caption{{Tests and vaccinations}}
+        \label{fig:antigen_tests_vaccinations}
     \end{subfigure}
@@
 turn of the year, the first people were vaccinated with a focus on older age groups and
-medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just
+medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). By the end of May, just
 over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
@@
 
-We draw simulated samples of agents from the distribution of recorded infections in
-September 2020 and use the model to predict recorded infection rates until the end of May
-2021.\comment[id=J]{This needs more explanation. I will rewrite this.} See Supplementary
-Materials~\ref{sec:data_and_parameters} and \ref{sec:model} for detailed descriptions of
-the data and the model, respectively. The blue line in Figure~\ref{fig:aggregated_fit}
-shows our model's predictions are very close to officially recorded cases in the
-aggregate. This is also true for infections by age and geographical region, which are
-shown in the supplementary materials (Figures~\ref{fig:age_group_fit} and
-\ref{fig:state_fit}, respectively).
+We draw simulated samples of agents from the population structure in September 2020 and
+use the model to predict recorded infection rates until the end of May 2021. See
+Supplementary Material~\ref{sub:initial_conditions} for a detailed description of this
+procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows our model's
+predictions are very close to officially recorded cases in the aggregate. This is also
+true for infections by age and geographical region, which are shown in the supplementary
+materials (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).
 
@@
 infections between late December 2020 and mid-February 2021. The reversal of the trend
-is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
-during May 2021 is associated with at least weekly rapid tests to around 42~percent of
-the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
+is associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
+during May 2021 is associated with at least weekly rapid tests to 42\% of the
+population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
 impacting a fall in the relative infectiousness of contacts outside of work and school.
 
-In order to better understand the contributions  of rapid tests, vaccinations, and of
-seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
-considers various scenarios. NPIs are always the same as in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
-and three scenarios turning these factors off one by one.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
-\citep{Shapley2016} to decompose the difference in total infections between the scenario
-without any of the three factors and our main specification.\comment[id=J]{I would not
-mention the word "coalition" in the notes to the figure on the Shapley values.
-Combination with other channels is better. I think I would prefer to have a short section
-on shapley values in the appendix (why use them, how do they work) and just link to that
-in the figure note}
+In order to better understand the contributions  of rapid tests, vaccinations, and
+seasonality on the evolution of infections in 2021,
+Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
+same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
+model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
+without any of the three factors (red line), and three scenarios turning these factors
+off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
+values \citep{Shapley2016} to decompose the difference in total infections between the
+scenario without any of the three factors and our main specification.
 
@@
         to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and seasonality
-        remain at their January levels. The other scenarios turn these three factors on
-        one-by-one. The decompositions in Figures~\ref{fig:2021_scenarios_decomposition}
-        and \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
-        where the individual contribution of a channel is its average contribution over
+        variant is introduced in the same way; vaccinations, rapid tests, and
+        seasonality remain at their January levels. The other scenarios turn these three
+        factors on one-by-one. The decompositions in
+        Figures~\ref{fig:2021_scenarios_decomposition} and
+        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values where
+        the individual contribution of a channel is its average contribution over
         different sizes of coalitions (combinations with other channels). The individual
@@
 groups contribute less to the spread of the disease than others due to a lower number of
-contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. It is important to
-note that the initial focus of the campaign was to prevent deaths and severe disease;
-the case fatality was rate considerably lower during the third wave when compared to the
-second (4.4\% between October and February and 1.4\% between March and June). By the end
-of our study period, when first-dose vaccination rates reached around 40\% of the
-population, the numbers of new cases would have started to decline.
+contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
+study period, when first-dose vaccination rates reached around 40\% of the population,
+the numbers of new cases would have started to decline. It is important to note that the
+initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
+fatality was rate considerably lower during the third wave when compared to the second
+(4.4\% between October and February and 1.4\% between March and the end of May).
 
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace. Nevertheless, we estimate it to be a quantitatively important factor
-determining the evolution of the pandemic, explaining most of the early changes and
-almost 40\% of the cumulative difference by the end of May.
+much lower pace \citep[this is in line with results in][]{Gavenciak2021}. Nevertheless,
+we estimate seasonality to be a quantitatively important factor determining the
+evolution of the pandemic, explaining most of the early changes and close to 40\% of the
+cumulative difference by the end of May.
 
@@
 rapid testing. Here, it is crucial to differentiate between recorded cases and actual
-cases. Additional testing means that infections are detected, which would otherwise
-remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
-until late April, recorded cases are higher than in the scenario where none of the three
-mechanisms are turned on. Compared to the scenario with vaccinations only, this point is
-reached around mid-May and it would be June for the comparison with the seasonality-only
-scenario. The effect on total cases, however, is visible immediately. Despite the fact
-that only a small fraction of the population performed weekly rapid tests in
-March,\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
-week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
-respondents to the COSMO study reported in early March that they \textit{ever} did a
-rapid test)} new infections on April 1 would be reduced by 53\% relative to the
-scenario without vaccinations, rapid tests, or seasonality.
+cases. Additional testing means that infections are recorded, which would otherwise
+remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is
+large. Until late April, recorded cases are higher in the scenario with rapid testing
+alone when compared to the setting where none of the three mechanisms are turned on.
+Compared to the scenario with vaccinations only, this point is reached around mid-May
+and it would be June for the comparison with the seasonality-only scenario. The effect
+on total cases, however, is visible immediately in
+Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only a small
+fraction of the population performed weekly rapid tests in March,\footnote{4\% to 6\% of
+respondents to the COSMO study report having done a rapid test during the first week of
+March; 8\% in the model have ever done a rapid test at that point.}\comment[id=HM]{would
+be great to have total tests performed in March, model vs. data is not important here.
+Just first week of March looks too small} new infections on April 1 would be reduced by
+53\% relative to the scenario without vaccinations, rapid tests, or seasonality.
 
@@
 scenario without rapid tests only (purple line in
-Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
-channels for rapid tests. Tests for pupils have the smallest effect, which is largely
-explained by the relatively small number of students (Only 18\% of our population are in
-the education sector (pupils, teachers etc.) versus  46\% are workers outside the
-education sector) and the fact that schools did not operate at full capacity during our
-period of study. Almost 40\% come from tests at the workplace. Despite the fact that
-rapid tests for private reasons are phased in only late, they make up for more than half
-of the total effect.\comment[id=HM]{How many tests are actually performed along each of
-the three dimensions? Could this be a quantity story rather than the directed testing
-story I am making up here as wishful thinking?}\comment[id=K]{See subsection
-\ref{subsec:appendix_rapid_tests}.}
-
-The reason lies in the fact that a substantial share of these tests is driven by an
-elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. The latter is essentially a form of contact
-tracing, which has been shown to be very effective \citep{Contreras2021,
-Kretzschmar2020}. This is also the reason for why the number of tests performed fall at
-the very end of our study period. Falling infection rates mean that there are fewer
-events triggering private test demand.\comment[id=HM]{Positive rates along the three
-channels might be nice here, if previous comment is not a quant story}\comment[id=K]{See
-figure~\ref{fig:share_of_rapid_tests_by_reason_by_infected}} \comment[id=HM]{I think
-there is a story in here about risk conditional on past exposure vs. prospective risk in
-gatherings, but it is too late to work through it. We might just want to calculate the
-probabilities.}\comment[id=K]{I don't understand which probabilities you mean.}
+Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
+three channels for rapid tests. Tests at schools have the smallest effect, which is
+largely explained by schools not operating at full capacity during our period of study
+and the relatively small number of students.\footnote{18\% of our population are in the
+education sector (pupils, teachers etc.); 46\% are workers outside the education sector}
+Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for
+private reasons are phased in only late, they make up for more than half of the total
+effect. The reason lies in the fact that a substantial share of these tests is driven by
+an elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or
+following up on a positive test of a household member. The latter is essentially a form
+of contact tracing, which has been shown to be very effective \citep{Contreras2021,
+Kretzschmar2020,FetzerGraeber2021}.
 
@@
 After the summer break, they were operating at full capacity with increased hygiene
-measures, before being closed again from mid-December on. Some states started opening
-them gradually in late February, but usual operation was not back until the beginning of
-June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
-school starting at Easter, at which point rapid tests had become widely available. We
-estimate the realized scenario to have essentially the same effect as a situation with
-closed schools. Under fully opened schools with mandatory tests, total infections would
-have been 7\% higher; this number rises to 20\% without tests. These effect sizes are
-broadly in line with empirical studies, e.g., \citet{Vlachos2021}\comment[id=HM]{Need a
-back-of the envelope calculation} To use another metric, the effective weekly
-reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
-negative effects school closures have on children and parents \citep{Luijten2021,
-Melegari2021} \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
-health papers. There must be well-published ones}\comment[id=K]{I googled and found two
-that I think are at least not terrible but I also didn't find anything really well
-published}---and in particular on those with low socio-economic status---these results in
-conjunction with hindsight bias suggest that opening schools combined with a testing
-strategy would have been beneficial. In other situations, and particular when no test
-strategy is available, trade-offs may well be different.
+measures, before being closed again from mid-December onwards. Some states started
+opening them gradually in late February, but operation at normal capacity did not resume
+until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
+different policies regarding school starting at Easter, at which point rapid tests had
+become widely available. We estimate the realized scenario to have essentially the same
+effect as a situation with closed schools. Under fully opened schools with mandatory
+tests, total infections would have been 7\% higher; this number rises to 20\% without
+tests. These effect sizes are broadly in line with empirical studies, e.g.,
+\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
+another metric, the effective weekly reproduction number differs by $0.017$ and $0.045$,
+respectively. In light of the large negative effects school closures have on children
+and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
+socio-economic status---these results in conjunction with hindsight bias suggest that
+opening schools combined with a testing strategy would have been beneficial. In other
+situations, and particular when rapid test are not available at scale, trade-offs may
+well be different.
 
@@
 
-Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of employees
-receiving tests, testing at the workplace has larger effects than changing requirements
-to work from home. Whether the share of workers working at the usual workplace is
-reduced or increased by ten percent hardly has an effect on infection rates. Making
-testing mandatory twice a week---assuming independent compliance by employers and workers
-or 95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
-employers to the level of March would have increased them by the same amount.
+Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
+receiving tests, testing at the workplace has larger effects than mandating employees to
+work from home. Whether the share of workers working at the usual workplace is reduced
+or increased by ten percent hardly has an effect on infection rates. Making testing
+mandatory twice a week---assuming independent compliance by employers and workers or
+95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
+employers to the level of March would have increased infections by the same amount.
 
-Despite these large effects, the results on testing likely understate the benefits.
-Disadvantaged groups of the population are less likely to be reached by testing
-campaigns;\comment[id=HM]{seems obvious, but I did not find anything. COSMO data? We
-should get the raw data, btw.} at the same time they have a higher risk to contract
-CoViD-19.\citet{KochInstitut2021a} Mandatory tests at school and at the workplace will
+Our analysis has shown that in the transition path to high levels of vaccination and
+possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
+at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
+estimated to have cost around 20~Euros per capita \citet{Wollmershauser2021}; retail
+prices for rapid tests were below one Euro in early June 2021. Despite these large
+effects, the results on testing likely understate the benefits. Disadvantaged groups are
+less likely to be reached by testing campaigns relying on voluntary participation
+\citep[e.g.][]{StillmanTonin2021}; at the same time they have a higher risk to contract
+CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the workplace will
 extend more into these groups. The same goes for individuals who exhibit a low level of
-compliance with CoViD-19-related regulations.
-
-Our analysis\comment[id=J]{The conclusion should be slightly longer and more separated
-from the rest} has shown that in the transition path to high levels of vaccination and
-possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A day of the winter lockdown in Germany is estimated to have
-cost 6.25~billion~Euros \citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo
-Instute back from March 2020 assuming a 2 month lockdown. I did not find anything
-citable that is more recent. I did find
-\href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
-quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
-Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
-for self-administered rapid tests are frequently less than
-1~Euro.\comment[id=HM]{Cheapest on \href{https://amazon.de}{Amazon} on June 11th, 2021.
---- DM 95ct}
+compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
+programmes allow a much quicker roll-out, making it an  short-term alternative.

### 2021-06-18 paper/appendix_results.tex
Polish model summary.
Add explanation on shapley values.

@@
   \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_demand}
-  \caption{Share of Individuals that Demand a Rapid Test for a Particular Reason that are Actually Infected}
+  \caption{Share of individuals who Demand a Rapid Test for a Particular Reason that are Actually Infected}
   \label{fig:share_of_rapid_tests_by_reason_by_infected}
@@
 \FloatBarrier
+
+\subsection{Shapley Value} % (fold)
+\label{sub:shapley_value}
+
+We decompose the effects of different NPIs and seasonality on the infection rates with
+Shapley values. Shapley values \citep{Shapley2016} are a concept in game theory to
+divide payoffs between a coalition of players. It allows to assign a single value to the
+contribution of an NPI or seasonality which takes into account substitutional and
+complementary effects with other factors.
+
+More formally, define a coalitional game with $N$ players and a super-additive function
+$\nu$ which maps subsets of $N$ to the real numbers. The function $\nu$ is also called
+the characteristic function and assigns a value to a coalition. Then, the Shapley value
+$\phi$ for player $i$ is
+
+\begin{align*}
+    \phi_i(\nu) = \frac{1}{|N| !} \sum_{S \subseteq N \setminus \{i\}} |S| ! (|N| - |S| - 1)! (\nu(S \cup \{i\}) - \nu(S))
+\end{align*}
+
+The last term $(\nu(S \cup \{i\}) - \nu(S))$ is the marginal contribution of player $i$
+minus the coalition without player $i$. Then, compute the sum of marginal contributions
+over all subsets $S$ of $N$ which do not include player $i$. Each marginal contribution
+has to be multiplied by all combinations of other players in $S$ which precede $i$ and
+all possible combinations of remaining players which follow player $i$ in the coalition.
+To arrive at the Shapley value for player $i$, divide the sum by the total number of
+combinations.
+
+The Shapley value has some properties.
+
+\begin{description}
+  \item[Efficiency] The sum of Shapley values is equal to the value of a coalition
+  formed by all players.
+  \item[Symmetry] The Shapley does not depend on the label of a player but only on its
+  position in the characteristic function.
+  \item[Linearity] The Shapley value depends linearly on the values from the
+  characteristic function $\nu$.
+  \item[Dummy Axiom] The Shapley value of a player who contributes nothing to any
+  coalition is 0.
+\end{description}
+
+To produce Figure~\ref{fig:2021_scenarios_decomposition} and
+Figure~\ref{fig:2021_scenarios_decomposition_tests}, we calculate the Shapley values of
+each factor in the comparison on the cumulative number of saved infections between the
+main scenario and the scenario without any of the factors for every day. Then, we divide
+up the saved infections on a particular day according to the Shapley values for the same
+day which yields the daily saved infections for each factor.
+
+% subsection shapley_value (end)

### 2021-06-18 paper/report.tex
Add explanation on shapley values.

@@
         Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values where
-        the individual contribution of a channel is its average contribution over
-        different sizes of coalitions (combinations with other channels). The individual
-        contribution to a coalition is the difference between the effect size of the
-        coalition with the particular channel and without.}
+        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values which
+        are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
 \end{figure}

### 2021-06-18 paper/appendix_data.tex
Add figure descriptions to the data appendix.

@@
     \label{n_contacts_work_non_recurrent}
-    \floatfoot{\noindent In the model it is sampled every day which of these numbers of
-        contacts a working person is planned to have. Note that the contact diaries include
-        such high values that super spreading events are well possible in our model. The
-        planned number of contacts is reduced by policies, seasonality and individual
-        responses to events such as receiving a positive rapid test to the number of actual
-        contacts with transmission potential. Work contacts only take place between working
-        individuals.}
+    \floatfoot{\noindent \textit{Note:} In the model it is sampled every day which of
+        these numbers of contacts a working person is planned to have. Note that the
+        contact diaries include such high values that super spreading events are well
+        possible in our model. The planned number of contacts is reduced by policies,
+        seasonality and individual responses to events such as receiving a positive rapid
+        test to the number of actual contacts with transmission potential. Work contacts
+        only take place between working individuals.}
 \end{figure}
@@
     \label{n_contacts_work_daily_recurrent}
-    \floatfoot{\noindent Working individuals are assigned to groups that are time
-        constant and that meet daily to match the given distribution of daily work contacts.
-        You can think of these as for example colleagues with which one shares an office
-        space. The share of individuals who attend in a way that has transmission potential
-        is reduced by policies (such as a work from home mandate), seasonality and individual
-        responses to events such as receiving a positive rapid test. Work contacts only take
-        place between working individuals.}
+    \floatfoot{\noindent \textit{Note:} Working individuals are assigned to groups that
+        are time constant and that meet daily to match the given distribution of daily
+        work contacts. You can think of these as for example colleagues with which one
+        shares an office space. The share of individuals who attend in a way that has
+        transmission potential is reduced by policies (such as a work from home mandate),
+        seasonality and individual responses to events such as receiving a positive rapid
+        test. Work contacts only take place between working individuals.}
 \end{figure}
@@
     \label{n_contacts_work_weekly_recurrent}
-    \floatfoot{\noindent  Working individuals are assigned to up to 14 groups that are time
-        constant and meet weekly. Groups are scheduled to meet on separate days of the work
-        week. These contact models cover weekly team meetings etc. The share of individuals
-        that attend in a way that has transmission potential is reduced by policies,
-        seasonality and individual responses to events such as receiving a positive rapid
-        test. Work contacts only take place between working individuals.}
+    \floatfoot{\noindent  \textit{Note:} Working individuals are assigned to up to 14
+        groups that are time constant and meet weekly. Groups are scheduled to meet on
+        separate days of the work week. These contact models cover weekly team meetings
+        etc. The share of individuals that attend in a way that has transmission
+        potential is reduced by policies, seasonality and individual responses to events
+        such as receiving a positive rapid test. Work contacts only take place between
+        working individuals.}
 \end{figure}
@@
     \label{fig:assortativity_other}
-    \floatfoot{\noindent The figure shows the distribution of non recurrent other contacts by age
-        group. A row shows the share of contacts a certain age group has with all other
-        age groups. Higher values are colored in darker red tones. The diagonal
-        represents the share of contacts with individuals from the same age group.}
+    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
+        recurrent other contacts by age group. A row shows the share of contacts a
+        certain age group has with all other age groups. Higher values are colored in
+        darker red tones. The diagonal represents the share of contacts with individuals
+        from the same age group.}
 \end{figure}
@@
     \label{fig:assortativity_work}
-    \floatfoot{\noindent The figure shows the distribution of non recurrent work contacts by age
-        group. A row shows the share of contacts a certain age group has with all other
-        age groups. Higher values are colored in darker red tones. The diagonal
-        represents the share of contacts with individuals from the same age group. We
-        only show age groups that have a significant fraction of working individuals.}
+    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
+        recurrent work contacts by age group. A row shows the share of contacts a certain
+        age group has with all other age groups. Higher values are colored in darker red
+        tones. The diagonal represents the share of contacts with individuals from the
+        same age group. We only show age groups that have a significant fraction of
+        working individuals.}
 \end{figure}
@@
     \caption{School Multiplier With and Without Vacations Factored In}
-    \floatfoot{\noindent}
+    \floatfoot{\noindent \textit{Note:} The dates on which schools have vacation are
+    decided at the federal level. Vacations are directly implemented in our model with no
+    school contacts taking place on weekends and during vacations (by federal state) just
+    like the schooling mode (full operation, emergency care, rotating schemes with half
+    class sizes etc.). The figure is thus only an illustration that roughly shows the
+    share of contacts taking place compared to pre-pandemic level with and without
+    vacations. The difference between the lines show when vacations take place and to
+    what degree. For example all states have fall vacations but the timing varies
+    strongly between states.}
     \label{fig:school_multiplier}
@@
     \label{fig:work_multiplier}
-    \floatfoot{\noindent The figure shows the work mobility as reported by \cite{Google2021}. We
-        take this as a proxy of the share of workers who are not in home office, i.e. who
-        still have physical work contacts. The figure interpolates over weekends as we
-        handle weekend effects through information on work on weekends in the German
-        census data we use. The figure shows the share aggregated over Germany as a
-        whole. To capture the effect that local policies, school vacations and public
-        policies have on work contacts we use the data by German state to determine which
-        workers go to work depending on the state they live in.}
+    \floatfoot{\noindent \textit{Note:} The figure shows the work mobility as reported by
+        \cite{Google2021}. We take this as a proxy of the share of workers who are not in
+        home office, i.e. who still have physical work contacts. The figure interpolates
+        over weekends as we handle weekend effects through information on work on
+        weekends in the German census data we use. The figure shows the share aggregated
+        over Germany as a whole. To capture the effect that local policies, school
+        vacations and public policies have on work contacts we use the data on the level
+        of the federal states to determine which workers go to work depending on the
+        state they live in.}
 \end{figure}
@@
 situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with friends).
+meetings and physical distancing (such as virtual meetings with
+friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
+which phases we have and why the switching points are where they are.}
 
@@
     \includegraphics[width=\textwidth]{../figures/results/figures/data/other_multiplier}
-    \caption{Share of Other Contacts Still Taking Place}
+    \caption{Share of Pre-Pandemic Other Contacts Taking Place with Infection Potential}
     \label{fig:other_multiplier}
-    \floatfoot{\noindent}
+    \floatfoot{\noindent \textit{Note:} All values are estimated. We try to use as little
+    switching points as possible and tie them to political events (such as lockdown
+    announcements) unless changes are used to capture anticipation or pandemic fatigue
+    (for example we model an anticipation of the November lockdown and model lockdown
+    fatigue in early March).}
 \end{figure}
@@
     \label{fig:seasonality}
-    \floatfoot{\noindent}
+    \floatfoot{\noindent \textit{Note:} We model seasonality as a factor that reduces the
+    probability of infection of all encounters. The factor depends on the day and is
+    calculated from a sinus shaped function with its maximum on January 1st. Since
+    seasonality can affect the transmission both through physical conditions such as
+    temperature and humidity as well as through the numbers of contacts that take place
+    outside we assume two seasonality factors. One for other contacts which we expect to
+    be strongly affected by fairer weather with a maximum reduction of 30\% in the
+    infection probability. The other seasonality only makes contacts up to 20\% less
+    infectious and is applied to household, work and school contacts.}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of individuals who do a Rapid Test.}}
-    \floatfoot{\noindent ...}
+    \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
+    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
+    planning to have education contacts, work contacts, developing symptoms without
+    access to a PCR test, having a household member with a positive test or symptoms. In
+    each case whether a rapid test is done depends on how long it has been since the
+    individual's last rapid test and her individual compliance parameters. As an example,
+    take a worker in May. In that time workers are encouraged to test themselves twice
+    weekly but there is no general requirement to test themselves. If the worker has not
+    done a test within the last four days in our model she will demand a test if her
+    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
     \label{fig:rapid_test_demand}
@@
     \label{fig:share_rapid_test_last_week}
-    \floatfoot{\noindent The figure compares the share of individuals who have ever done a rapid
-        test or done a rapid test within the last week in our simulations to the shares
-        reported in the
+    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
+        have ever done a rapid test or done a rapid test within the last week in our
+        simulations to the shares reported in the
         \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
-            Snapshot Monitoring survey}. The left panel compares the share of individuals
-        who have ever done a rapid test. The right panel compares the share of
-        individuals who have done a rapid test within the last seven days in our
-        simulation compared to the share reporting to have done at least weekly rapid
-        tests in the last four weeks in the COSMO survey. Overall our calibration of
-        rapid tests are slightly conservative. The overall share is below that in the
-        study. We fit the share of weekly tests quite exactly. However, the study only
-        covers adults while our share also includes children who are tested very
-        regularly when attending school.}
+        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
+        have ever done a rapid test. The right panel compares the share of individuals
+        who have done a rapid test within the last seven days in our simulation compared
+        to the share reporting to have done at least weekly rapid tests in the last four
+        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
+        conservative. The overall share is below that in the study. We fit the share of
+        weekly tests quite exactly. However, the study only covers adults while our share
+        also includes children who are tested very regularly when attending school.}
 \end{figure}

### 2021-06-18 paper/rapid_test_params.tex
Red texts -> comments, except for some ldots in figurenotes.
Move estimated parameters to model section. Address some comments.
Add figure descriptions to the data appendix.

@@
+\subsection{Rapid Test Demand}
+\label{subsec:rapid_test_demand}
+
 In our model, there are five reasons why rapid tests are done:\comment[id=J]{Add a
@@
 
-\subsubsection{work rapid tests}
+%%% \subsubsection{work rapid tests}
 
@@
 offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
-60\% until the first days of April \cite{ZDF2021}.
+60\% until the first days of April \cite{ZDF2021}.\comment[id=K]{ToDo: Find the survey
+that the ZDF is citing here}
 
-\textcolor{red}{ToDo: Find the survey that the ZDF is citing here}
+Until mid April 70\% of workers were expected to receive a weekly test offer
+\citep{AerzteZeitung2021}. However, according to surveys conducted in mid April
+\citep{Betsch2021}, less than two thirds of individuals with work contacts receive a test
+offer. Starting on April 19th employers were required by law to provide two weekly tests
+to their employees \citep{Bundesanzeiger2021}. We assume that compliance is incomplete
+and only 80\% of employers actually offer tests.
 
-Until mid April 70\% of workers were expected to receive a
-weekly test offer \citep{AerzteZeitung2021}. However, according to surveys conducted in
-mid April \citep{Betsch2021}, less than two thirds of individuals with work contacts
-receive a test offer. Starting on April 19th employers were required by law to provide
-two weekly tests to their employees \citep{Bundesanzeiger2021}. We assume that compliance
-is incomplete and only 80\% of employers actually offer tests.
-
-\subsubsection{educ rapid tests}
+%%\subsubsection{educ rapid tests}
 
@@
 of students get tested by March 1st. Relying on the same sources as above we approximate
-that by the week before Easter this share had increased to 40\%.\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen}, }
+that by the week before Easter this share had increased to
+40\%.\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen},
+}
 
@@
 Baden-Württemberg\footnote{https://bit.ly/3vuetaD, https://bit.ly/3vuetaD}.
+
+%%\subsubsection{private rapid tests}
 

### 2021-06-18 paper/model/policies.tex
Rewrite policies section in model appendix.

--- 2021-02-08 paper/model/policies.tex
@@
-\subsection{Reducing Numbers of Contacts Through Policies}
+\subsection{Reducing Numbers of Contacts via NPIs}
 \label{sec:policies}
 
-The main motivation of our model is to predict the effect of policies that affect the
-number of contacts people have. Examples range from school closures and lockdowns to
-more nuanced policies such as mandatory quarantines for symptomatic individuals or a
-class splitting policy where only half of the students come to school in person and the
-other half joins digitally with weekly rotation.
+Our model makes it very easy to model a wide range of NPIs, either in isolation or
+simultaneously. This is important for two reasons: Firstly, it allows to predict and
+quantify the effect of novel NPIs. Secondly, it allows to model the actually implemented
+policy environment in great detail, which is necessary to use use the full time series
+of infections and fatality rates to estimate the model parameters.\footnote{
+See \citet{Avery2020} for an explanation why it can be harmful to use too long time
+series to estimate simple SEIR type models.}
 
@@
 it is often more helpful to think of them as adjusting the pre-pandemic number of
-contacts.
-
-Therefore, we implement policies as a step that happens after the number of contacts is
+contacts. Therefore, we implement policies as a step that happens after the number of contacts is
 calculated but before individuals are matched.
@@
 On an abstract level, a policy is a functions that modifies the number of contacts of
-one contact type. For example, school closures simply set all school contacts to zero. A
-lockdown where only essential workers are allowed to work means that approximately two
-thirds of the working population have zero work contacts and the rest has the same
-number of contacts as before.
+one contact type. This function can be random or deterministic. For example, school
+closures simply set all school contacts to zero. A work from home mandate leads to a
+share of workers staying home every day whereas those who cannot work from home are
+unaffected. Hygiene measures at work randomly reduce the number of infectious contacts
+for all workers who still go to work.
 
-This, in conjunction with our fine-grained contact types, allows us to easily implement
-a wide variety of policies. Allowing policies to depend on the health states of the
-entire population means that adaptive lockdowns where, for example, schools close when a
-certain threshold of infections is surpassed at the county level would be as simple as
-determining which counties are above the threshold and then setting all school contacts
-in these counties to zero.
+Policies can also interact. For example, school vacations are temporally reducing school
+contacts to zero while at the same time increasing other contacts to account for
+increased leisure activities and family visits during this time. This is important to
+reproduce the finding that school vacations do not reduce infection numbers even though
+schools lead to infections when open \citep{Isphording2021}.
 
-The dependency of policies on health states also makes it possible to model contact
-tracing. For example, a policy could check whether each child has a classmate who's
-received a positive test result and then bar all children of that class from attending
-school.
+The most complex policies are typically found in the education sector. Since the
+beginning of 2021 schools have switched back and fourth between full closures,
+split class approaches with alternating schedules for some or all age groups and
+reopening while maintaining hygiene measures. On top of that there are different
+policies for allowing young students whose parents work full time to attend school
+even on days where they normally would not. For details on how we calibrate these
+policies see Section~\ref{subsec:policies_seasonality}.
 
-Some policies can be easily implemented if the background characteristics are suitably
-extended. For example, a schooling policy with A/B schooling, where each half attends
-school every other week can be implemented by storing whether the child would attend in
-even or odd weeks in the background characteristics and then using that information in
-the policy function.
+Importantly, policies can depend on the health states of participating individuals.
+This allows to quarantine entire school classes if one student tested positive or
+to implement official or private contact tracing.
 
 For some policies the exact effect on each contact type is not easy to determine. If
-this refers to a policy during the estimation period, it is possible to estimate such
-parameters by fitting the model to time series data of infection rates. This is only
-possible if the policy was not active during the whole estimation period and thus the
-infection probabilities can be identified separately. If instead it refers to a policy
-that we want to simulate, we make a scenario analysis in which the model is simulated
-with several assumptions about how the policy affects the number of contacts.
+this refers to a policy has been active during the estimation period, it is possible to
+estimate such parameters by fitting the model to time series data of infection rates.
+This is only possible if the policy was not active during the whole estimation period
+and thus the infection probabilities can be identified separately. We do this to account
+for hygiene measures at school and in the workplace that have been in effect since
+November 2020.

### 2021-06-18 paper/appendix_results.tex
Add rapid test statistics with new layout.

@@
 
-\subsection{Rapid Tests \comment[id=K]{The figures will likely change layout later}}
+\subsection{Rapid Tests}
 \label{subsec:appendix_rapid_tests}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/demand_shares}
-  \caption{Share of the Population Demanding a Rapid Test for a Particular Reason}
-  \label{fig:rapid_tests_by_reason}
-  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without
-  PCR demand and other contact demand together form the private demand category. Also
-  note that these do not add up to the total share of demanded rapid tests as individuals
-  may have more than one reason to demand a rapid test on any given day. }
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/educ_demand_share}
+    \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
+    \label{fig:educ_rapid_test_demand}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/work_demand_share}
+    \caption{Share of the Population Demanding a Rapid Test due to Work}
+    \label{fig:work_rapid_test_demand}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/private_demand_share}
+    \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
+  \label{fig:private_rapid_test_demand}
+  \end{subfigure}
+  \caption{Rapid Test Demand by Reason}
+  \label{fig:rapid_test_demand_by_reason}
+  \floatfoot{\noindent \textit{Note:}
+    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
+    school) as well as school pupils. After Easter the required frequency of tests is
+    increased from once per week to twice per week. Work rapid tests are demanded by
+    individuals that still have work contacts, i.e. do not work from home. The share of
+    employers offering rapid tests increases over the time frame and the frequency of
+    testing is also increased. Tests are demanded by individuals for one of three private
+    reasons: having developed symptoms without access to a PCR test, having a household
+    member that has tested positive or developed symptoms or having planned weekly
+    meeting with friends.
+  }
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_demand}
-  \caption{Share of individuals who Demand a Rapid Test for a Particular Reason that are Actually Infected}
-  \label{fig:share_of_rapid_tests_by_reason_by_infected}
-  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without
-  PCR demand and other contact demand together form the private demand category. Also
-  note that this is not the same as individuals getting a positive rapid test. The
-  sensitivity is quite low before individuals become infectious. Therefore, if
-  individuals are still in the latent period of their infection they are likely to get a
-  false positive rapid test.}
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
+    \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
+    Individuals}
+    \label{fig:share_infected_among_educ_rapid_test_demand}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
+    \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
+    \label{fig:share_infected_among_work_rapid_test_demand}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
+    \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
+  \label{fig:share_infected_among_private_rapid_test_demand}
+  \end{subfigure}
+  \caption{Share of Rapid Tests Demanded by Infected Individuals by Reason}
+  \label{fig:share_infected_among_rapid_test_demand_by_reason}
+  \floatfoot{\noindent \textit{Note:}
+    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
+    school) as well as school pupils. After Easter the required frequency of tests is
+    increased from once per week to twice per week. Work rapid tests are demanded by
+    individuals that still have work contacts, i.e. do not work from home. The share of
+    employers offering rapid tests increases over the time frame and the frequency of
+    testing is also increased. Tests are demanded by individuals for one of three private
+    reasons: having developed symptoms without access to a PCR test, having a household
+    member that has tested positive or developed symptoms or having planned weekly
+    meeting with friends. Private rapid tests have a very high share of infected
+    individuals because they are mostly triggered by events that make infection likely.
+    Remember however, that one reason is that a household member has a positive rapid
+    test. This means that work and education rapid tests which have a low rate of
+    infected individuals trigger more targeted rapid tests. They also have a much higher
+    volume of tests. As can be seen in the decomposition (see
+    Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
+    important for the overall effect.}
 \end{figure}

### 2021-06-18 paper/model/policies.tex
Merge policies and endogenous contact reductions.

@@
 November 2020.
+
+Not all things that reduce contacts compared to the pre-pandemic level are driven
+by NPIs. Therefore, we also model endogenous contact reductions that can depend on
+the health state of individuals, known risk contacts or the local incidence of
+infections. Examples are strong contact reductions for symptomatic individuals or those
+who have a positive PCR or rapid tests or contact reductions when a houshold member
+tested positive. The extent to which contacts are reduced can be calibrated from
+surveys. For an application of our model showcasing private contact tracing in the
+context of the Christmas holidays see \cite{Gabler2020}.

### 2021-06-18 paper/model/matching.tex
Apply suggestions from code review
Update Readme and add pdf.
Update templates, fix errors, copy pdf to root.
No obvious stuff left, at least nothing obvious at this time of the night.
Update matching, remove pseudocode.

@@
 person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
-to be translated into actual meetings between people. This is achieved by our matching
+to be translated into actual meetings between people. This is achieved by a matching
 algorithm:
@@
 
-Recurrent contacts are described by two components: 1) A variable in the background
-characteristics. An example would be a school class identifier which could come from
-actual data or be drawn randomly to achieve representative class sizes.
-2) A deterministic or random function that takes the value 0 (non-participating) and 1
+Recurrent contacts are described by two components: 1) A set of time invariant groups, such as school classes. The groups can be sampled from empirical data or created by
+randomly matching simulated individuals into groups. 2) A deterministic or random function that takes the value 0 (non-participating) and 1
 (participating) and can depend on the weekday, date and health state. This can be used
@@
 home. The assumption that all group members have contacts with all other group members
-is not fully realistic, but seems like a good approximation to reality, especially in
+is not fully realistic, but a good approximation to reality, especially in
 light of the suspected role of aerosol transmission for Covid-19 \citep{Morawska2020,
-Anderson2020}.
+Anderson2020}. Alternatively, the infection probability of recurrent contact types can
+be interpreted as being the product of a true infection probability and the probability
+that an actual contact takes place.
 
@@
 
-Below, we first show pseudo code for the non-recurrent matching algorithm and then
-describe how the algorithm works in words.
-
-\begin{listing}
-    \label{code:matching}
-    \begin{minted}{python}
-        while are_unmatched_contacts_left:
-
-            contact_type, i = draw_contact_type_and_individual()
-
-            for _ in remaining_contacts[i, contact_type]:
-
-                group_j = draw_group_of_other_person()
-                j = draw_other_person_from_that_group(group_j)
-
-                if infection_takes_pinfectedlaces(i, j):
-                    update_health_state_of_freshly_infected()
-
-                remaining_contacts[i, contact_type] -= 1
-                remaining_contacts[j, contact_type] -= 1
-    \end{minted}
-    \caption{Pseudo-code of the matching algorithm for non-recurrent contacts.}
-\end{listing}
-
-We first randomly draw a contact type and individual. For each contact of the drawn
-contact type that person has, we first draw the group of the other person (first stage).
+The algorithm works as follows: First we randomly draw a contact type and individual.
+For each contact of the drawn contact type that person has, we first draw the group of
+the other person (first stage).
 Next, we calculate the probability to be drawn for each member of the group, based on
 the number of remaining contacts, i.e. people who have more remaining contacts are drawn
-with a higher probability. This has to be re-calculated each time because with each
+with a higher probability. The probabilities have to be re-calculated each time because
+with each
 matched contact, the number of remaining contacts changes. We then draw the other
-individual, determine whether an infection takes place and if so update the health
-state of the newly infected person. Finally, we reduce the number of remaining contacts
-of the two matched individuals by one.
+individual, determine whether an infection takes place and if so update the health state
+of the newly infected person. Finally, we reduce the number of remaining contacts of the
+two matched individuals by one.\comment[id=J]{Add formulae here.}
 
@@
 makes the matching computationally much more tractable because the number of
-computations increases quadratically in the second stage group size.
+computations only increases quadratically in the size of the second group and not
+quadratically in the size of the entire simulated population.

### 2021-06-18 paper/appendix_results.tex
Add many figure descriptions and start adding some more plots.

@@
 \subsection{Simulated vs. Empirical Data}
-\label{subsec:new_cases_fit}
+\label{subsec:fit_results}
 
@@
 rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups.\comment[id=J]{summarize the
-fit}\comment[id=HM]{Also need to show simulated total infections somewhere. So far only
-ever shown for 2021.}
+analyze the model fit for age groups and federal states.\comment[id=J]{summarize the
+fit}
 
@@
   \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
-  \caption{Simulated and Empirical Infections}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates.}
+  \caption{Fit Over the Full Simulation Time Frame with Single Simulation Runs}
+  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
+  100,000 people for the reported simulated infections rates. The mean infection rate is
+  the thick blue line. Single simulation runs are plotted in lighter and thinner lines.
+  The official case numbers as reported by the Robert-Koch-Institut are plotted in black.
+  The fit is overall very good. The higher the mean incidence and the stronger the growth
+  the more variance there is between simulation runs. We averaged over 30 simulation
+  runs.}
   \label{fig:aggregated_fit2}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_newly_infected_with_single_runs}
+  \caption{Development of the Total Infections Over the Full Simulation Time Frame with
+  Single Simulation Runs}
+  \floatfoot{\noindent \textit{Note:} The figure shows the true weekly incidence rates
+  per 100,000 people, including undetected cases. The mean infection rate is the thick
+  blue line. Single simulation runs are plotted in lighter and thinner lines. The higher
+  the mean incidence and the stronger the growth the more variance there is between
+  simulation runs. We averaged over 30 simulation runs.}
+  \label{fig:newly_infected_in_baseline}
 \end{figure}
@@
   \caption{Simulated and Empirical Infections by Age Group}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates for different age groups.}
+  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
+  100,000 people for the reported versus the simulated infections rates for different age
+  groups. The age group of individuals above 80 needs to be interpreted with caution
+  because our synthetic population only includes private households, i.e. nursing homes
+  are not represented in our model. They accounted for many cases and deaths in the
+  winter of 2020 and many 80 to 100 year olds live in these facilities. However, the
+  official data does not contain information on whether cases were nursing home
+  inhabitants or not. We averaged over 30 simulation runs.}
   \label{fig:age_group_fit}
@@
   \caption{Simulated and Empirical Infections by Federal State}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates for different federal states.}
+  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
+  100,000 people for the reported versus the simulated infections rates for different
+  federal states. We averaged over 30 simulation runs.}
   \label{fig:state_fit}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
+  \caption{Share of Vaccinated Individuals}
+  \floatfoot{\noindent \textit{Note:} The figure shows the rate of individuals that are
+  vaccinated in our synthetic population versus in the general German population. Note
+  that we excluded the vaccinations that were given to nursing homes, approximately the
+  first percent of the German population that were vaccinated. Overall, our model covers
+  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
+  the population are vaccinated. Our vaccinations work imperfectly but we do not model
+  different vaccines nor do we distinguish between first and second shot.}
+  \label{fig:newly_infected_in_baseline}
 \end{figure}
@@
 
-\subsection{Share of Cases that are Detected  \comment[id=K]{Figure notes missing}}
+\subsection{Share of Cases that are Detected}
 \label{subsec:appendix_share_known_cases}
@@
   \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group in the Main Prediction}
+  \caption{Share of Detected Cases by Age Group}
   \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent}
+  \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
+  reported as an official case via PCR confirmation. We use the overall share of known
+  cases that was estimated through the case fatality ratio by the
+  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
+  then assume it to be constant as vaccinations of the elderly strongly affect the case
+  fatality rate which the project does not account for. To get from an overall share of
+  detected cases to the share of cases that is detected in each age group we use that
+  asymptomatic cases are much less likely to be detected. As our model covers age
+  specific asymptomatic rates this endogenously leads to group specific share known cases
+  that verify that infections in younger age groups are under-detected. Starting in 2021
+  in addition to the overall numbers of detected cases through symptoms and the share
+  known cases, cases are also detected through confirmation of positive rapid tests. This
+  leads to an increase in the share of known cases for all age groups but in particular
+  for the younger age groups that are covered extensively with rapid tests through the
+  rapid test requirement for participating in school.}
 \end{figure}
@@
     member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. Private rapid tests have a very high share of infected
-    individuals because they are mostly triggered by events that make infection likely.
-    Remember however, that one reason is that a household member has a positive rapid
-    test. This means that work and education rapid tests which have a low rate of
-    infected individuals trigger more targeted rapid tests. They also have a much higher
-    volume of tests. As can be seen in the decomposition (see
-    Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
-    important for the overall effect.}
+    meeting with friends. Private rapid tests have a much higher share of infected
+    individuals because they are mostly triggered by events that make an infection
+    likely. Remember however, that one reason is that a household member has a positive
+    rapid test. This means that work and education rapid tests which have a low rate of
+    infected individuals trigger more targeted rapid tests through the household demand.
+    They also have a much higher volume of tests. As can be seen in the decomposition
+    (Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
+    important for the overall effect. It appears that the combination of wide testing to
+    find infection chains plus targeted tests to break those chains taking together
+    is important.}
 \end{figure}
@@
 \label{subsec:appendix_scenarios}
-
-This is the results section.\comment[id=K]{The results can be found in `figures/results`.
-This includes both figures and tables for lookup of numbers and summary tables.}
 
@@
   \label{fig:explain_decline}
-  \figurenotes{\ldots}
+  \floatfoot{\noindent \textit{Note:} \ldots}
 \end{figure}

### 2021-06-18 paper/report.tex
Went through draft once more.

@@
 test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
-over time. Containment measures have become increasingly diverse and now include
+over time. Containment measures have become increasingly diverse and now include rapid
 testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
-influence of seasonal patterns or more infectious virus strains are well understood in
-quantitative terms.
+influence of seasonal patterns or of more infectious virus strains are well understood
+in quantitative terms.
 
@@
 pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
-differentiated by age and geographical location see Section~\ref{sub:number_of_contacts}
-of the Supplementary Material.
+differentiated by age and geographical location (see Section~\ref{sub:number_of_contacts}
+of the Supplementary Material).
 
@@
 symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
-The probabilities of transitioning between these states depend on age; their duration is
-random within intervals calibrated to medical literature (for a detailed description see
+intensive care, and recovery or death \citep[similar
+to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
+transitioning between these states depend on age; their duration is random within
+intervals calibrated to medical literature (for a detailed description see
 Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
@@
 Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
-more specific once Appendix has new structure}. We keep the number of free parameters
-low in order to avoid overfitting. In our application, these features include  official
-case numbers for each age group and region, deaths, and the share of the B.1.1.7 strain.
+more specific once Appendix has new structure} In our application, we keep the number of
+free parameters low in order to avoid overfitting. The data features to be matched
+include official case numbers for each age group and region, deaths, and the share of
+the B.1.1.7 strain.
 
@@
 These had to be administered by medical doctors or in pharmacies. At-home tests approved
-by authorities became available in mid-March, rapid test centers were opened and one
-test per person and week was made available free of charge. Depending on the state,
-customers were only allowed to enter certain stores with a recent negative rapid test
-result. These developments are characteristic of many countries: The initial focus on
-NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
-acceptance and use of rapid tests. At broadly similar points in time, novel strains of
-the virus have started to pose additional challenges.
+by authorities became available in mid-March, rapid test centers were opened, and one
+test per person and week was made available free of charge. In several states, customers
+were only allowed to enter certain stores with a recent negative rapid test result.
+These developments are characteristic of many countries: The initial focus on NPIs to
+slow the spread of the disease has been accompanied by vaccines and a growing acceptance
+and use of rapid tests. At broadly similar points in time, novel strains of the virus
+have started to pose additional challenges.
 
@@
 Supplementary Material~\ref{sub:initial_conditions} for a detailed description of this
-procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows our model's
+procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model's
 predictions are very close to officially recorded cases in the aggregate. This is also
@@
         Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values which
+        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
         are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
@@
 rapid testing. Here, it is crucial to differentiate between recorded cases and actual
-cases. Additional testing means that infections are recorded, which would otherwise
-remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is
-large. Until late April, recorded cases are higher in the scenario with rapid testing
-alone when compared to the setting where none of the three mechanisms are turned on.
-Compared to the scenario with vaccinations only, this point is reached around mid-May
-and it would be June for the comparison with the seasonality-only scenario. The effect
-on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only a small
-fraction of the population performed weekly rapid tests in March,\footnote{4\% to 6\% of
-respondents to the COSMO study report having done a rapid test during the first week of
-March; 8\% in the model have ever done a rapid test at that point.}\comment[id=HM]{would
-be great to have total tests performed in March, model vs. data is not important here.
-Just first week of March looks too small} new infections on April 1 would be reduced by
-53\% relative to the scenario without vaccinations, rapid tests, or seasonality.
+cases. Additional testing means that additional infections will be recorded which would
+otherwise remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this
+effect is large. Until late April, recorded cases are higher in the scenario with rapid
+testing alone when compared to the setting where none of the three mechanisms are turned
+on. Compared to the scenario with vaccinations only, this point is reached around
+mid-May and it would be June for the comparison with the seasonality-only scenario. The
+effect on total cases, however, is visible immediately in
+Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\%  of the
+population performed weekly rapid tests in March on average, new infections on April 1
+would be reduced by 53\% relative to the scenario without vaccinations, rapid tests, or
+seasonality.
 

### 2021-06-18 paper/model/testing.tex
Cleaned up: \color{red} -> \comment[id=J].
First draft for section on testing.

--- 2021-06-04 paper/model/testing.tex
@@
 
-Our model includes both PCR tests which are scarce and take time until the result becomes
-available and rapid tests which are done after an individual's contacts are determined
-but before the contacts take place.
+Having a realistic model of PCR and rapid tests is crucial for two reasons: Firstly,
+only via a testing model the simulated infections from the model can be made comparable
+to official case numbers. Secondly, individuals with undetected or not yet detected
+infections are an important driver of the pandemic.
 
-PCR testing consists of three stages. Firstly, we model who demands a PCR test. Demand
-functions map from individual characteristics to a probability which is the probability
-for this individual to demand a test. There can be multiple demand functions where each
-function may describe a different channel. For example, individuals who experience
-symptoms or have a risk contact may ask for a test. Or, the ministry of education
-requires a negative test result from every teacher every second week. After the
-probabilities for each individual and every demand model are computed, individuals who
-demand a test as well as the channel is sampled.
+In principle, our modeling approach is flexible enough to incorporate mechanistic
+test demand, allocation and processing models. However, there is not enough data
+available to calibrate such a mechanistic model.
 
-The second stage is the allocation phase in which demand and supply for tests are
-matched. The number of available tests can be inferred from official data and used to
-model shortages in supply. When demand exceeds supply, some individuals might be given
-preferred access to tests because of their own vulnerability or their potential to
-become a super-spreader.
+Therefore we aim for a simpler model of PCR and rapid tests that can be calibrated with
+available data on test demand and availability and and, nevertheless, can produce
+a share of undetected cases that varies over time and across age groups and agrees with
+other estimates over the time periods where they are available.
 
-In the last and third phase, administered tests are processed. This step can become a
-bottleneck in the testing process if there are not enough laboratories or necessary
-resources available to evaluate all tests.
+PCR tests are modeled since the beginning of the simulation and determine whether a
+infections is officially recorded. Rapid tests are only added at the beginning of
+2021. Positive rapid tests do not enter official case numbers directly, but most people
+with a positive rapid tests demand a confirmatory PCR test. However, positive rapid test
+can have a strong effect on the infection dynamics because they trigger contact
+reductions.
 
-In our empirical estimation we use a very simplified testing model where the number of
-tests to be distributed is calculated from estimates for the ratio of known to all
-infections.\footnote{The Dunkelzifferradar project publishes daily estimates of the dark
-figure of infections under \url{https://covid19.dunkelzifferradar.de/}} Using these
-estimates as well as data on the test distribution over age groups by the
-RKI\footnote{https://ars.rki.de/Content/COVID19/Main.aspx} we allocate tests firstly
-among the symptomatic and then randomly allocate tests to newly infected to fit the
-German test distribution.
+During 2020 people can demand PCR tests either because they have symptoms or randomly.
+The probability that a PCR test is performed in each of the two situations depends on
+the number of new infections and the number of available tests. It thus varies strongly
+over time and is unknown.
 
-Rapid tests are modeled much simpler. Every day before individuals have contacts they can
-decide to be tested.\comment[id=HM]{If we get econ readers, we should not use language like this because we do not have a choice model.} For example students that plan to attend school that day and have
-not done a rapid test in the last three days get a rapid test. Then they immediately
-receive the test result. After they have received their test result individuals can react
-to it by reducing their contacts. For example positively tested individuals may not go to
-work and reduce their household contacts to some degree. Who reduces their contacts to
-what degree depends on a quarantine compliance attribute.
+To distribute the correct number of PCR tests among symptomatic and asymptomatic
+infections without knowing explicit test demand probabilities, we use the following
+approach: First we calculate the total number of positive PCR tests by multiplying the
+number of newly infected individuals with an estimate of the share of detected cases
+from the Dunkelzifferradar project.
+\footnote{The Dunkelzifferradar project publishes daily estimates of the dark figure of infections under \url{https://covid19.dunkelzifferradar.de/}}.
+Next we determine how many of these tests
+should go to symptomatic and asymptomatic individuals from data by the Robert Koch
+Institut.\comment[id=J]{add source} Then we sample the individuals to which those test
+are allocated from the pools of symptomatic and asyptomatic infected but not yet tested
+individuals.
 
-Our rapid tests include false positives and false negatives. The sensitivity of rapid tests in our model depends on when the individual has
-or will become infectious. This way we can account for the fact that rapid tests are
-likely to be false positive before infectiousness starts.
+Sampling uniformly from the pool of symptomatic individuals ensures that age groups
+who are more likely to develop symptoms are also more likely to receive tests. Thus the
+share of detected cases is much higher for adults than for children in time
+periods where many tests are done because of symptoms which is in line with the
+estimates from the literature.\comment[id=J]{need source}
 
-% subsection testing (end)
+At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
+of detected cases from the Dunkelzifferradar can no longer be used because it is based
+on case fatality rates which drastically change due to vaccinations. Secondly, rapid
+tests become available at a large scale.
+
+We solve the first challenge by assuming that the share of detected cases would have
+remained on the level it reached before Christmas if rapid tests had not become
+available. While this is only an approximation to reality, changes in the share of
+detected cases that would have happened without rapid tests are very likely to be small
+compared to the changes caused by rapid tests.
+
+The second challenge is solved by mechanistic rapid test demand models for the
+workplace, schools and by private individuals. The calibration of these models is
+described in Section~\ref{rapid_test_demand}.
+Figure~\ref{fig:antigen_tests_vaccinations} that the number of performed rapid tests in
+the model fits the empirical data well (where empirical data is available).
+
+In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
+falsely negative. While the specificity of rapid tests is constant over time, the
+their sensitivity strongly depends on whether the tested individual is already
+infectious and if so for how long he has been infectious. Before the onset of
+infectiousness the sensitivity is very low (35 \%). On the first day of infectiousness
+it is much higher (88 \%) but still lower than during the remaining infectious
+period (92 \%). After infectiousness stops, the sensitivity drops to 50 \%.
+
+Modeling the diagnostic gap before and at the beginning of infectiousness is very
+important to address concerns that rapid tests are too unreliable to serve as screening
+devices.
+
+We do not distinguish between self administered and rapid tests and those that are
+administered by medical personal. While there were concerns that self administered tests
+are less reliable, a recent study has found no basis for this
+concern.\comment[id=J]{cite charite study}
+
+While rapid tests do not directly enter official case numbers, many positively tested
+individuals confirm their rapid test result with a PCR test.\comment[id=J]{Need to
+link to data appendix once the number is there and maybe mention the number}.
+Importantly, those PCR tests are made in addition to the tests that would have been
+done otherwise. Section~\ref{subsec:appendix_share_known_cases} discusses the effect
+of rapid tests on the share of detected cases.

### 2021-06-18 paper/appendix_results.tex
AddR_t and Share B117 to appendix results.

@@
   different vaccines nor do we distinguish between first and second shot.}
-  \label{fig:newly_infected_in_baseline}
+  \label{fig:fit_vaccinations}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
+  \caption{Effective Replication Number $R_t$ in the Model and as Reported by the
+  Robert-Koch-Institute}
+  \floatfoot{\noindent \textit{Note:} The figure shows the effective replication number
+  ($R_t$) as reported by the RKI and as calculated in our model. The $R_t$ gives the
+  average number of new infections caused by one infected individual. The $R_t$ in our
+  model broadly follows the $R_t$ reported by the RKI. Two trends stand out. Firstly, the
+  RKI's $R_t$ drops faster in November. This could be due to a change in the testing
+  policy that focused tests on the elderly when the second wave hit Germany and led to a
+  decline in the overall share of detected cases. The second difference is from mid
+  February to mid March where the RKI's reported $R_t$ increased more rapidly than that
+  in our model. Here the opposite effect can be expected. During this time rapid tests
+  increased strongly leading to more cases being detected. In the short term this leads
+  an $R_t$ estimation that is based on detected cases to overestimate the replication
+  number.}
+  \label{fig:fit_r_effective}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117_with_single_runs}
+  \caption{Share of B.1.1.7 in the Model and as Reported by the Robert-Koch-Institute}
+  \floatfoot{\noindent \textit{Note:} The figure shows the share of B.1.1.7 as
+  reported by the RKI and as calculated in our model. We only introduce a few cases over
+  the cause of January. From then B.1.1.7 takes over endogenously through its increased
+  infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases per 100,000 inhabitants.}
+  \label{fig:fit_r_effective}
 \end{figure}

### 2021-06-18 paper/report.tex

@@
 
-Our analysis has shown that in the transition path to high levels of vaccination and
+Our analysis has shown that during the transition to high levels of vaccination and
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
 at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 20~Euros per capita \citet{Wollmershauser2021}; retail
+estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
 prices for rapid tests were below one Euro in early June 2021. Despite these large
@@
 less likely to be reached by testing campaigns relying on voluntary participation
-\citep[e.g.][]{StillmanTonin2021}; at the same time they have a higher risk to contract
-CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the workplace will
-extend more into these groups. The same goes for individuals who exhibit a low level of
-compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
-programmes allow a much quicker roll-out, making it an  short-term alternative.
+\citep[e.g.][]{StillmanTonin2021}; at the same time, these groups have a higher risk to
+contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the
+workplace will extend more into these groups. The same goes for individuals who exhibit
+a low level of compliance with CoViD-19-related regulations. Compared to vaccinations,
+rapid testing programmes allow a much quicker roll-out, making it arguably the most
+effective tool to contain the pandemic in the short run.

### 2021-06-18 paper/model/estimated_parameters.tex
Improve section on estimated parameters.

@@
 
-We estimate parameters that could not be calibrated inside the model with the
+We estimate parameters that cannot be calibrated outside of the model with the
 method of simulated moments \citep{McFadden1989} by minimizing the distance between
@@
 
-We use data for Germany from October 2020 until June 2021. We do not use
+We fit our model to data for Germany from October 2020 until June 2021. We do not use
 earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
@@
 To avoid over-fitting and simplify the numerical optimization problem, we only allow for
-five different probabilities: 1) for contacts in schools 2) contacts in preschools and
-nurseries. 3) for work contacts. 4) for households. 5) for other contacts.
+five different infection probabilities: 1) for contacts in schools 2) contacts in
+preschools and nurseries. 3) for work contacts. 4) for households. 5) for other
+contacts.\comment[id=J]{Write that the exact numbers are hard to interpret and show
+the infections by channel heatmap more prominently than the infection probabilities}
 
-The infection probabilities estimated from our model are as shown in
-Table~\ref{tab:infection_probs}.\comment[id=K]{Missing to add data for the other
-estimated parameters (import of b117 etc)}\comment[id=K]{Update table after the new
-estimation is finished}
+% policies
+We also estimate a parameter that reflects the effect of hygiene measures
+after November 2020 at work and in educational facilities. This parameter reduces
+infectiousness of contacts by one third. In total those are 10 parameters. The
+breakpoints the contact reduction changes are not determined from data but from
+announcements of policy changes. Moreover, we constrain the estimated contact reduction
+to follow the shape of the stringency index. The resulting contact reduction can be
+seen in Figure~\ref{fig:other_multiplier}
 
-\begin{table}[tb]
-    \caption{Estimated Infection Probabilities}
-    \label{tab:infection_probs}
-    \centering
-    \begin{tabular}{ll}
-        \toprule
-        Contact Type & Infection Probability \\
-        \midrule
-        preschool and nursery & 00.500\% \\
-        school & 01.200\% \\
-        household & 10.000\% \\
-        work & 14.500\% \\
-        other & 15.875\% \\
-        \bottomrule
-    \end{tabular}
-    \tablenotes{School contacts take place Monday through Friday. All students and
-teachers meet all other students and teachers that attend school that day and there are
-three classes taking place each day. Work contacts can be daily, weekly or non recurrent.
-Other contacts can be daily, weekly or non recurrent. We do not differ the infection
-probability depending on how often a contact takes place.}
-\end{table}
+% b117 rate
+Finally we estimate one parameter that governs the introduction of the B.1.1.7 virus
+variant in January 2021. This parameter implies that at the end of January roughly one
+case per 100 000 individuals per day is imported. After January we do not model imported
+cases of B.1.1.7 anymore because they are negligible compared to the endogenous growth
+of that virus variant.
+
+% The infection probabilities estimated from our model are as shown in
+% Table~\ref{tab:infection_probs}.\comment[id=K]{Missing to add data for the other
+% estimated parameters (import of b117 etc)}\comment[id=K]{Update table after the new
+% estimation is finished}
+
+% \begin{table}[tb]
+%     \caption{Estimated Infection Probabilities}
+%     \label{tab:infection_probs}
+%     \centering
+%     \begin{tabular}{ll}
+%         \toprule
+%         Contact Type & Infection Probability \\
+%         \midrule
+%         preschool and nursery & 00.500\% \\
+%         school & 01.200\% \\
+%         household & 10.000\% \\
+%         work & 14.500\% \\
+%         other & 15.875\% \\
+%         \bottomrule
+%     \end{tabular}
+%     \tablenotes{}
+% \end{table}

### 2021-06-21 src/figures/results/tables/student_infections_table.tex
Restructure for arxiv submission.

--- 2021-06-16 figures/results/tables/student_infections_table.tex
@@
 \midrule
- educ open after easter  without tests &                                             808831 \\
- educ open after easter  with tests    &                                             627182 \\
- close educ after easter               &                                             478064 \\
- baseline                              &                                             490880 \\
+ educ open after easter  without tests &                                             751930 \\
+ educ open after easter  with tests    &                                             567537 \\
+ close educ after easter               &                                             436232 \\
+ baseline                              &                                             452727 \\
 \bottomrule

### 2021-06-21 src/paper/appendix_data.tex
Restructure for arxiv submission.

--- 2021-06-18 paper/appendix_data.tex
@@
 
-\input{medical_parameters}
+\input{paper/medical_parameters}
 
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
         \caption{Number of Non Recurrent Other Contacts}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
         \caption{Number of Daily Recurrent Other Contacts}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
         \caption{Number of Weekly Recurrent Other Contacts}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
     \caption{Number of Non Recurrent Work Contacts}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
     \caption{Number of Daily Recurrent Work Contacts}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
     \caption{Number of Weekly Recurrent Work Contacts}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/household}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
     \caption{Number of Household Contacts}
@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_other_non_recurrent}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_other_non_recurrent}
     \caption{Distribution of Non Recurrent Other Contacts by Age Group}
@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_work_non_recurrent}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_work_non_recurrent}
     \caption{Distribution of Random Work Contacts by Age Group}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/school_multiplier_comparison}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/school_multiplier_comparison}
     \caption{School Multiplier With and Without Vacations Factored In}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/work_multiplier_since_sep}
     \caption{Share of Workers with Work Contacts}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/other_multiplier}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/other_multiplier}
     \caption{Share of Pre-Pandemic Other Contacts Taking Place with Infection Potential}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/seasonality}
     \caption{Seasonality by Type of Contact}
@@
 
-\input{rapid_test_params}
+\input{paper/rapid_test_params}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
     \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
@@
     \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
     \end{subfigure}%
     \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
         %   \caption{Share of Individuals Having Done a Rapid Test in the Last Week}

### 2021-06-21 src/paper/appendix_results.tex
Restructure for arxiv submission.

--- 2021-06-18 paper/appendix_results.tex
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
   \caption{Fit Over the Full Simulation Time Frame with Single Simulation Runs}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_newly_infected_with_single_runs}
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_newly_infected_with_single_runs}
   \caption{Development of the Total Infections Over the Full Simulation Time Frame with
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
+  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
   \caption{Simulated and Empirical Infections by Age Group}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
+  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
   \caption{Simulated and Empirical Infections by Federal State}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
   \caption{Share of Vaccinated Individuals}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
   \caption{Effective Replication Number $R_t$ in the Model and as Reported by the
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117_with_single_runs}
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117_with_single_runs}
   \caption{Share of B.1.1.7 in the Model and as Reported by the Robert-Koch-Institute}
@@
   infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases per 100,000 inhabitants.}
-  \label{fig:fit_r_effective}
+  \label{fig:fit_share_b117}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
+  \includegraphics[width=\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
   \caption{Share of Detected Cases by Age Group}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/educ_demand_share}
+    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/educ_demand_share}
     \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/work_demand_share}
+    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/work_demand_share}
     \caption{Share of the Population Demanding a Rapid Test due to Work}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/private_demand_share}
+    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/private_demand_share}
     \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
+    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
     \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
+    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
     \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
+    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
     \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
@@
   \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
   \end{subfigure}%
   \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
   \end{subfigure}
@@
   \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
   \end{subfigure}%
   \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
   \end{subfigure}
@@
 
-\input{../figures/results/tables/student_infections_table}
+\input{figures/results/tables/student_infections_table}
 

### 2021-06-21 src/paper/model/course_of_disease.tex
Restructure for arxiv submission.

--- 2021-06-17 paper/model/course_of_disease.tex
@@
 
-The following medical parameters describing the progression of the disease are taken from
-systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
-progresses in the way depicted in Figure~\ref{fig:model_disease_progression}.
+The disease progression in the model is fairly standard. It is depicted in
+Figure~\ref{fig:model_disease_progression} and the values and source of the relevant
+parameters are describes in Section~\ref{sec:medical_params}.
 
-First, infected individuals will become infectious after one to five days. About one
-third of people remain asymptomatic. The rest develop symptoms about one to two days
-after they become infectious. Modeling asymptomatic and pre-symptomatic cases is
-important because those people do not reduce their contacts or demand a test and can
-potentially infect many other people \citep{Donsimoni2020}. The probability to develop
+First, infected individuals will become infectious after one to five days. Overall,
+about one third of people remain asymptomatic. The rest develop symptoms about one to
+two days after they become infectious. Modeling asymptomatic and pre-symptomatic cases
+is important because those people do not reduce their contacts nor do they have an
+elevated probability to demand a test. Thus they can potentially infect many other
+people \citep{Donsimoni2020}. The probability to develop
 symptoms with Covid-19 is highly age dependent with 75\% of children not developing
-symptoms \citep{Davies2020}.
+clinical symptoms \citep{Davies2020}.
 
@@
 \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-documentation}.} and age-dependent.
+documentation}.}.\comment[id=J]{Replace this by a link to the relevant table with
+parameters in the data appendix, once we have that table.}

### 2021-06-21 src/paper/model/estimated_parameters.tex
Restructure for arxiv submission.

--- 2021-06-18 paper/model/estimated_parameters.tex
@@
 method of simulated moments \citep{McFadden1989} by minimizing the distance between
-simulated and observed infection rates. Since our model includes a lot of randomness, we
+simulated and observed infection rates and fatality rates (disaggregated by region
+and age groups). Since our model includes a lot of randomness, we
 average simulated infection rates over several model runs.

### 2021-06-21 src/paper/model/number_of_contacts.tex
Restructure for arxiv submission.

--- 2021-06-03 paper/model/number_of_contacts.tex
@@
 those 50 people had contact with an infectious person and the disease has an infection
-probability of 1 in each contact or whether everyone met an infectious person but the
+probability of 1 per contact or whether everyone met an infectious person but the
 disease has an infection probability of only 5 percent per contact. SEIR models do not
@@
 
-In our empirical application we distinguish the following types of contacts:\comment[id=HM]{I started updating this to understand it better myself.
+In our empirical application we distinguish the following  contact types that
+are depicted in Figure~\ref{fig:model_contacts_infections} and can be further grouped
+in the categories household, work, education and others.
+
+types of contacts:\comment[id=HM]{I started updating this to understand it better myself.
 In the end, I think it will be useful to group into HH / work / schools and add an introductory sentence for each group.
@@
 
-    \item Recurrent daily non-work contacts representing very close family or friendship
-    relations outside the household.
+    \item Recurrent work contacts, capturing contacts with coworkers, repeating
+          clients and superiors. Some of these recurrent contacts take place on
+          every workday, others just once per week.
 
-    \item Recurrent weekly non-work contacts representing friendships, sports groups,
-    charity work, or other activities. We randomize over the days on which the meetings take place.\comment[id=HM]{True?}
-
-    \item Random non-work contacts: Contacts with randomly drawn other
-    people, which are assortative with respect to region and age group. This contact type
-    reflects contacts during leisure activities, grocery shopping, medical appointments,
-    etc..
-
-    \item Recurrent daily work contacts, capturing direct colleagues.
-
-    \item Recurrent weekly work contacts, representing meetings with clients, superiors, or more distant colleagues.
-    We randomize over the days on which the meetings take place.
-
-    \item Random work contacts: Working adults' contacts with randomly drawn other
-    people, which are assortative in geographical location\comment[id=HM]{True? age?}
+    \item Random work contacts: Working adults have contacts with randomly drawn other
+          people, which are assortative in geographical location and age.
 
@@
     are closed on weekends and during vacations, which vary by states. School classes
-    also meet three pairs of teachers every school day. The pairs are meant to represent
-    interactions between teachers.\comment[id=HM]{Not updated from here on. Mayb}
+    also meet six teachers everyday and some of the teachers meet each other.
 
@@
     mechanics are similar to schools.
+
+    \item Random other contacts: Contacts with randomly drawn other
+    people, which are assortative with respect to geographic location and and age group.
+    This contact type reflects contacts during leisure activities, grocery shopping,
+    medical appointments, etc..
+
+    \item Recurrent other contacts representing contacts with friends neighbors or
+        family members who do not live in the same household. Some of these contacts
+        happen daily, others only once per week.
+
 \end{itemize}
@@
 Section~\ref{sec:number_of_contacts}. In particular, we sample the number of contacts or
-group sizes from empirical distributions that sometimes depend on age. It is also
+group sizes from empirical distributions that sometimes depend on age. It would also be
 possible to use economic or other behavioral models to predict the number of contacts.
 
-Theoretically, each contact type can have its own infection probability. However, to
+% Theoretically, each contact type can have its own infection probability. However, to
 reduce the number of free parameters and thus avoid a potential over-fitting we only

### 2021-06-21 src/paper/model/summary.tex
Restructure for arxiv submission.

--- 2021-06-17 paper/model/summary.tex
@@
 The structure of the model is depicted in Figure~\ref{fig:model_contacts_infections}.
-\comment[id=HM]{Mentioned by Tim in the course, just main paper: Modular structure may deserve a sentence}
 
-The background characteristics include age, county and occupation of each simulated
-individual. Contact models are functions that map individual characteristics into a
-predicted number of contacts. Currently, we distinguish between eight types of contact
+We distinguish between eight types of contact
 models which are all listed in Figure~\ref{fig:model_contacts_infections}: households,
@@
 
-The predicted number of contacts is translated into infections by a matching algorithm.
+The number of contacts is translated into infections by a matching algorithm.
 There are different matching algorithms for recurrent contacts (e.g. classmates, family
-members) and non-recurrent contacts (e.g. clients, contacts in supermarkets). The
-infection probability can differ for each contact type. All types of contacts can be
-assortative with respect to geographic and demographic characteristics.
+members) and non-recurrent contacts (e.g. clients, contacts in supermarkets).
+All types of contacts can be assortative with respect to geographic and demographic
+characteristics.
 
@@
 
-Another key advantage of the model is that the number of contacts an individual has of
+An important feature of the model is that the number of contacts an individual has of
 each contact type can be calibrated from publicly available data \citep{Mossong2008}.
@@
 
-Last but not least, performing simulations whose starting point is set amidst the
-pandemic requires special adjustments to arrive at a realistic distribution of courses
-of diseases. We solve the initial conditions problem by matching reported infections to
-individuals in our data while also correcting for reporting lag and undetected cases.
-
-In the following sections we describe each of the model components in more detail.
+The model has a very modular structure and can easily be extended to distinguish more
+contact types, add more stages to the disease progression, implement new policies
+or test demand models. The main bottleneck is not complexity or computational cost
+but the availability of data to calibrate the additional model features.

### 2021-06-21 src/paper/report.tex
Restructure for arxiv submission.

--- 2021-06-18 paper/report.tex
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
+        \includegraphics[width=\textwidth]{figures/model-graph-top-left}
         \caption{{Model for contacts and infections}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
+        \includegraphics[width=\textwidth]{figures/model-graph-top-right}
         \vskip5ex
@@
 
-        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
+        \includegraphics[width=\textwidth]{figures/model-graph-bottom-left}
         \caption{{Model for rapid tests}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
+        \includegraphics[width=\textwidth]{figures/model-graph-bottom-right}
         \vskip5ex
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
         \caption{{Recorded cases: Empirical and simulated}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2_with_seasonality}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/stringency2_with_seasonality}
 
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
 
@@
 
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
 
@@
         Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
-        \ref{subsec:new_cases_fit}. For legibility reasons, all lines in
+        \ref{subsec:fit_results}. For legibility reasons, all lines in
         Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
         \caption{{Recorded cases: 2021 scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
         \caption{{Total cases: 2021 scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_area}
-        % \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_bar}
+        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_area}
+        % \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_bar}
         \caption{Decomposition of the difference between the scenario without any of the
@@
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
-        % \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
+        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_area}
+        % \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_bar}
         \caption{Decomposition of the difference between the scenario without rapid
@@
         \centering
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
         \caption{{Effects of different schooling scenarios}}
@@
         \centering
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
         \caption{{Effects of different work scenarios}}

### 2021-06-21 task_compile_documents.py

@@ -18,7 +18,7 @@ BLD = ROOT / "bld"
         "-f",
     ]
 )
-@pytask.mark.depends_on(ROOT / "paper" / "paper.tex")
+@pytask.mark.depends_on(ROOT / "src" / "paper.tex")
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass

### 2021-06-21 src/paper.tex
More adjustments for arxiv.

\pdfoutput=1

\input{paper/preamble}

\begin{document}

\selectlanguage{USenglish}
\frenchspacing

\input{paper/titlepage}

\maketitle
\pagenumbering{arabic}
\setcounter{page}{1}

\newpage

\input{paper/report}

\phantomsection % Needed for the bibliography bookmark to point to the correct page

\begin{refcontext}[sorting=nyt]  % Sort BIBLIOGRAPHY by alphabet (while CITATIONS are sorted by year)
\printbibliography[heading=bibintoc]
\end{refcontext}

\clearpage

\begin{center}
    {
        \Large 
        Supplementary Material for\\[2ex]

        \textbf{\papertitle}
    }
    \vskip4ex
    {
        \large
        Janoś Gabler,

        Tobias Raabe,

        Klara Röhrl,

        Hans-Martin von Gaudecker

    }
\end{center}

\clearpage

\begin{appendices}
    \begin{refsection}

    \input{paper/appendix_data}

    \FloatBarrier

    \input{paper/appendix_model}

    \input{paper/appendix_results}

    \printbibliography[heading=bibintoc]

    \end{refsection}
\end{appendices}

\end{document}

### 2021-06-21 src/paper/appendix_data.tex
Clean up comments. Put shapley values into separate file. Put comments back in.

@@
 
-\subsection{Number of Contacts\comment[id=HM]{Histograms seem like the way to go here. Also add household contacts, i.e., distribution of (household size minus 1).}}
+\subsection{Number of Contacts}
 \label{sub:number_of_contacts}
@@
 Simplifying the number of contacts, we arrive at the following distributions of the
-numbers of contacts by contact type.\comment[id=K]{Put plots into one figure.}
+numbers of contacts by contact type.
 
@@
 
-Our other two types of contacts, households and schools, get their assortativity by
-construction. Schools are groups where the same children of the mostly same age group and
+Our two other types of contacts, households and schools, get their assortativity by
+construction. Schools are groups where the same children of the mostly same age and
 county meet with teachers every day. Household composition follows directly from the
-German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
-no work in here? How about school? Can we add a column with marginals and another one
-with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
-households addresses all questions.
-I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}\comment[id=HM]{See previous comment.
-My point is that total number of contacts rather than shares are much more informative in the first place, then different group sizes are no concern anymore.}
+German microcensus data we use to construct our synthetic population.
 

### 2021-06-21 src/paper/appendix_results.tex
Clean up comments. Put shapley values into separate file. Put comments back in.

@@
 \FloatBarrier
-
-\input{figures/results/tables/student_infections_table}
-
-\FloatBarrier
-
-\subsection{Shapley Value} % (fold)
-\label{sub:shapley_value}
-
-We decompose the effects of different NPIs and seasonality on the infection rates with
-Shapley values. Shapley values \citep{Shapley2016} are a concept in game theory to
-divide payoffs between a coalition of players. It allows to assign a single value to the
-contribution of an NPI or seasonality which takes into account substitutional and
-complementary effects with other factors.
-
-More formally, define a coalitional game with $N$ players and a super-additive function
-$\nu$ which maps subsets of $N$ to the real numbers. The function $\nu$ is also called
-the characteristic function and assigns a value to a coalition. Then, the Shapley value
-$\phi$ for player $i$ is
-
-\begin{align*}
-    \phi_i(\nu) = \frac{1}{|N| !} \sum_{S \subseteq N \setminus \{i\}} |S| ! (|N| - |S| - 1)! (\nu(S \cup \{i\}) - \nu(S))
-\end{align*}
-
-The last term $(\nu(S \cup \{i\}) - \nu(S))$ is the marginal contribution of player $i$
-minus the coalition without player $i$. Then, compute the sum of marginal contributions
-over all subsets $S$ of $N$ which do not include player $i$. Each marginal contribution
-has to be multiplied by all combinations of other players in $S$ which precede $i$ and
-all possible combinations of remaining players which follow player $i$ in the coalition.
-To arrive at the Shapley value for player $i$, divide the sum by the total number of
-combinations.
-
-The Shapley value has some properties.
-
-\begin{description}
-  \item[Efficiency] The sum of Shapley values is equal to the value of a coalition
-  formed by all players.
-  \item[Symmetry] The Shapley does not depend on the label of a player but only on its
-  position in the characteristic function.
-  \item[Linearity] The Shapley value depends linearly on the values from the
-  characteristic function $\nu$.
-  \item[Dummy Axiom] The Shapley value of a player who contributes nothing to any
-  coalition is 0.
-\end{description}
-
-To produce Figure~\ref{fig:2021_scenarios_decomposition} and
-Figure~\ref{fig:2021_scenarios_decomposition_tests}, we calculate the Shapley values of
-each factor in the comparison on the cumulative number of saved infections between the
-main scenario and the scenario without any of the factors for every day. Then, we divide
-up the saved infections on a particular day according to the Shapley values for the same
-day which yields the daily saved infections for each factor.
-
-% subsection shapley_value (end)

### 2021-06-21 src/paper/model/shapley_values.tex
Clean up comments. Put shapley values into separate file. Put comments back in.

--- 2021-06-18 paper/appendix_results.tex
@@
-\section{Additional Results}
-\label{sec:additional_results}
-
-\subsection{Simulated vs. Empirical Data}
-\label{subsec:new_cases_fit}
-
-This compares simulated data from our model with empirical data from Germany. We look at
-observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
-rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups.\comment[id=J]{summarize the
-fit}\comment[id=HM]{Also need to show simulated total infections somewhere. So far only
-ever shown for 2021.}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
-  \caption{Simulated and Empirical Infections}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates.}
-  \label{fig:aggregated_fit2}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Age Group}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates for different age groups.}
-  \label{fig:age_group_fit}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Federal State}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates for different federal states.}
-  \label{fig:state_fit}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Share of Cases that are Detected  \comment[id=K]{Figure notes missing}}
-\label{subsec:appendix_share_known_cases}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group in the Main Prediction}
-  \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent}
-\end{figure}
-
-It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
-
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/educ_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
-    \label{fig:educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/work_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test due to Work}
-    \label{fig:work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/private_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
-  \label{fig:private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Rapid Test Demand by Reason}
-  \label{fig:rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends.
-  }
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
-    \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
-    Individuals}
-    \label{fig:share_infected_among_educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
-    \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
-    \label{fig:share_infected_among_work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
-    \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
-  \label{fig:share_infected_among_private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Share of Rapid Tests Demanded by Infected Individuals by Reason}
-  \label{fig:share_infected_among_rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. Private rapid tests have a very high share of infected
-    individuals because they are mostly triggered by events that make infection likely.
-    Remember however, that one reason is that a household member has a positive rapid
-    test. This means that work and education rapid tests which have a low rate of
-    infected individuals trigger more targeted rapid tests. They also have a much higher
-    volume of tests. As can be seen in the decomposition (see
-    Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
-    important for the overall effect.}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Scenarios}
-\label{subsec:appendix_scenarios}
-
-This is the results section.\comment[id=K]{The results can be found in `figures/results`.
-This includes both figures and tables for lookup of numbers and summary tables.}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
-  \end{subfigure}
-  \caption{The Effect of Policies on Observed and Unobserved Cases}
-  \label{fig:explain_decline}
-  \figurenotes{\ldots}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-  \end{subfigure}
-  \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-  \label{fig:school_scenarios_detailed}
-\end{figure}
-
-\FloatBarrier
-
-\input{../figures/results/tables/student_infections_table}
-
-\FloatBarrier
-
-\subsection{Shapley Value} % (fold)
+\subsection{Shapley Values} % (fold)
 \label{sub:shapley_value}

### 2021-06-21 src/paper/appendix_data.tex
Fix typos in testing add two sentences on how we create our synthetic population.

@@
 \subsection{The Synthetic Population\comment[id=K]{To be written}}
+
+We build a synthetic population based on the German microcensus \citep{FDSAeDBUDL2018}.
+We only use private households, i.e. exclude living arrangements such as nursing homes as
+non-private households vary widely in size and it is very difficult to know which
+contacts take place in such households.
+
+We sample households to build our synthetic population of over one million households
+keeping for each individual their age, gender, occupation and whether they work on
+Saturdays and Sundays.
 
@@
 An exception where we do not rely on the data by \cite{Mossong2008} are the household
-contacts. Since household are included in the the German microcensus\comment[id=K]{Cite
-microcensus} on which we build our synthetic population we simply assume for the
-household contact model that individuals meet all other household members every day.
+contacts. Since household are included in the the German microcensus
+\citep{FDSAeDBUDL2018} on which we build our synthetic population we simply assume for
+the household contact model that individuals meet all other household members every day.
 

### 2021-06-21 src/paper/model/testing.tex
Fix typos in testing add two sentences on how we create our synthetic population.

--- 2021-06-18 paper/model/testing.tex
@@
 
-In principle, our modeling approach is flexible enough to incorporate mechanistic
-test demand, allocation and processing models. However, there is not enough data
-available to calibrate such a mechanistic model.
+In principle, our modeling approach is flexible enough to incorporate mechanistic test
+demand, allocation and processing models. However, there is not enough data available to
+calibrate such a mechanistic model.
 
-Therefore we aim for a simpler model of PCR and rapid tests that can be calibrated with
-available data on test demand and availability and and, nevertheless, can produce
-a share of undetected cases that varies over time and across age groups and agrees with
-other estimates over the time periods where they are available.
+Therefore, we aim for a simpler model of PCR and rapid tests that can be calibrated with
+available data on test demand and availability and -- nevertheless -- can produce a share
+of undetected cases that varies over time and across age groups and agrees with other
+estimates over the time periods where they are available.
 
 PCR tests are modeled since the beginning of the simulation and determine whether a
-infections is officially recorded. Rapid tests are only added at the beginning of
-2021. Positive rapid tests do not enter official case numbers directly, but most people
-with a positive rapid tests demand a confirmatory PCR test. However, positive rapid test
-can have a strong effect on the infection dynamics because they trigger contact
-reductions.
+infections is officially recorded. Rapid tests are only added at the beginning of 2021.
+Positive rapid tests do not enter official case numbers directly, but most people with a
+positive rapid tests demand a confirmatory PCR test. However, positive rapid tests can
+have a strong effect on the infection dynamics because they trigger contact reductions
+and additional rapid tests.
 
 During 2020 people can demand PCR tests either because they have symptoms or randomly.
-The probability that a PCR test is performed in each of the two situations depends on
-the number of new infections and the number of available tests. It thus varies strongly
-over time and is unknown.
+The probability that a PCR test is performed in each of the two situations depends on the
+number of new infections and the number of available tests. Thus, it varies strongly over
+time and is unknown.
 
@@
 infections without knowing explicit test demand probabilities, we use the following
-approach: First we calculate the total number of positive PCR tests by multiplying the
-number of newly infected individuals with an estimate of the share of detected cases
-from the Dunkelzifferradar project.
-\footnote{The Dunkelzifferradar project publishes daily estimates of the dark figure of infections under \url{https://covid19.dunkelzifferradar.de/}}.
-Next we determine how many of these tests
+approach: First, we calculate the total number of positive PCR tests by multiplying the
+number of newly infected individuals with an estimate of the share of detected cases from
+the Dunkelzifferradar project. \footnote{The Dunkelzifferradar project publishes daily
+estimates of the dark figure of infections under
+\url{https://covid19.dunkelzifferradar.de/}} Next, we determine how many of these tests
 should go to symptomatic and asymptomatic individuals from data by the Robert Koch
-Institut.\comment[id=J]{add source} Then we sample the individuals to which those test
+Institut.\comment[id=J]{add source} Then, we sample the individuals to which those tests
 are allocated from the pools of symptomatic and asyptomatic infected but not yet tested
@@
 
-Sampling uniformly from the pool of symptomatic individuals ensures that age groups
-who are more likely to develop symptoms are also more likely to receive tests. Thus the
-share of detected cases is much higher for adults than for children in time
-periods where many tests are done because of symptoms which is in line with the
-estimates from the literature.\comment[id=J]{need source}
+Sampling uniformly from the pool of symptomatic individuals ensures that age groups who
+are more likely to develop symptoms are also more likely to receive tests. Thus, the
+share of detected cases is much higher for the elderly than for children in time periods
+where many tests are done because of symptoms which is in line with the estimates from
+the literature.\comment[id=J]{need source}
 
 At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
-of detected cases from the Dunkelzifferradar can no longer be used because it is based
-on case fatality rates which drastically change due to vaccinations. Secondly, rapid
-tests become available at a large scale.
+of detected cases from the Dunkelzifferradar can no longer be used because it is based on
+case fatality rates which drastically change due to vaccinations. Secondly, rapid tests
+become available at a large scale.
 
 We solve the first challenge by assuming that the share of detected cases would have
-remained on the level it reached before Christmas if rapid tests had not become
+remained at the level it reached before Christmas if rapid tests had not become
 available. While this is only an approximation to reality, changes in the share of
@@
 
-The second challenge is solved by mechanistic rapid test demand models for the
-workplace, schools and by private individuals. The calibration of these models is
-described in Section~\ref{rapid_test_demand}.
-Figure~\ref{fig:antigen_tests_vaccinations} that the number of performed rapid tests in
-the model fits the empirical data well (where empirical data is available).
+The second challenge is solved by mechanistic rapid test demand models for the workplace,
+schools and by private individuals. The calibration of these models is described in
+Section~\ref{subsec:rapid_test_demand}. Figure~\ref{fig:antigen_tests_vaccinations} shows
+that the number of performed rapid tests in the model fits the empirical data well (where
+empirical data is available).
 
 In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
-falsely negative. While the specificity of rapid tests is constant over time, the
-their sensitivity strongly depends on whether the tested individual is already
-infectious and if so for how long he has been infectious. Before the onset of
-infectiousness the sensitivity is very low (35 \%). On the first day of infectiousness
-it is much higher (88 \%) but still lower than during the remaining infectious
-period (92 \%). After infectiousness stops, the sensitivity drops to 50 \%.
+falsely negative. While the specificity of rapid tests is constant over time, their
+sensitivity strongly depends on whether the tested individual is already infectious and
+if so for how long he has been infectious. Before the onset of infectiousness the
+sensitivity is very low (35\%). On the first day of infectiousness it is much higher
+(88\%) but still lower than during the remaining infectious period (92\%). After
+infectiousness stops, the sensitivity drops to 50 \%.
 
@@
 
-We do not distinguish between self administered and rapid tests and those that are
-administered by medical personal. While there were concerns that self administered tests
-are less reliable, a recent study has found no basis for this
-concern.\comment[id=J]{cite charite study}
+We do not distinguish between self administered rapid tests and those that are
+administered by medical personnel. While there were concerns that self administered tests
+are less reliable, a recent study has found no basis for this concern.\comment[id=J]{cite
+charite study on self administered tests}
 

### 2021-06-21 src/paper/report.tex
Fix numbers for relative effects (still old seasonality).

@@
-% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3
-% printed pages in the journal) present important new research results of broad
-% significance. Reports should include an abstract, an introductory paragraph, up to
-% four figures or tables, and about 30 references. Materials and Methods should be
-% included in supplementary materials, which should also include information needed to
-% support the paper's conclusions.
-
-% Main Text is not divided into sub-headings for Reports.
-
-% The manuscript should start with a brief introduction describing the paper’s
-% significance. The introduction should provide sufficient background information to
-% make the article intelligible to readers in other disciplines, and sufficient context
-% that the significance of the experimental findings is clear.
-
 Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
@@
 pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
-differentiated by age and geographical location (see Section~\ref{sub:number_of_contacts}
-of the Supplementary Material).
+differentiated by age and geographical location (see
+Section~\ref{sub:number_of_contacts} of the Supplementary Material).
 
@@
         \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_area}
-        % \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_bar}
         \caption{Decomposition of the difference between the scenario without any of the
@@
         \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_area}
-        % \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_bar}
         \caption{Decomposition of the difference between the scenario without rapid
@@
 the final level is always the highest in case of the vaccination campaign (orange
-lines). The Shapley value decomposition shows that vaccinations contribute about 15\% to
-the cumulative difference between scenarios. Reasons for this are the slow start---it
-took until March~24th until 10\% of the population had received their first vaccination,
-the 20\% mark was reached on April 19th---and the focus on older individuals. These
-groups contribute less to the spread of the disease than others due to a lower number of
+lines). The Shapley value decomposition shows that vaccinations contribute 17\% to the
+cumulative difference between scenarios. Reasons for this are the slow start---it took
+until March~24th until 10\% of the population had received their first vaccination, the
+20\% mark was reached on April 19th---and the focus on older individuals. These groups
+contribute less to the spread of the disease than others due to a lower number of
 contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
@@
 we estimate seasonality to be a quantitatively important factor determining the
-evolution of the pandemic, explaining most of the early changes and close to 40\% of the
+evolution of the pandemic, explaining most of the early changes and 38\% of the
 cumulative difference by the end of May.
 
-The largest effect---almost one half when considering the decompositions---comes from
-rapid testing. Here, it is crucial to differentiate between recorded cases and actual
-cases. Additional testing means that additional infections will be recorded which would
-otherwise remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this
-effect is large. Until late April, recorded cases are higher in the scenario with rapid
-testing alone when compared to the setting where none of the three mechanisms are turned
-on. Compared to the scenario with vaccinations only, this point is reached around
-mid-May and it would be June for the comparison with the seasonality-only scenario. The
-effect on total cases, however, is visible immediately in
+The largest effect---45\% in the decompositions---comes from rapid testing. Here, it is
+crucial to differentiate between recorded cases and actual cases. Additional testing
+means that additional infections will be recorded which would otherwise remain
+undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large.
+Until late April, recorded cases are higher in the scenario with rapid testing alone
+when compared to the setting where none of the three mechanisms are turned on. Compared
+to the scenario with vaccinations only, this point is reached around mid-May and it
+would be June for the comparison with the seasonality-only scenario. The effect on total
+cases, however, is visible immediately in
 Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\%  of the

### 2021-06-21 src/paper/appendix_data.tex
Remove all comments by hand for arxiv.

@@
 
-\subsection{The Synthetic Population\comment[id=K]{To be written}}
+\subsection{The Synthetic Population} %&%&%&\comment[id=K]{To be written}}
 
@@
 within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.\comment[id=HM]{Redo
-\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of
-contacts or better add a similar figure showing total number of contacts by age in all
-networks}
+data we use to calibrate the number of contacts.
+%&%&%&\comment[id=HM]{Redo \ref{fig:assortativity_other} /\ref{fig:assortativity_work}
+%&%&%&with total number of contacts or better add a similar figure showing total number of
+%&%&%&contacts by age in all networks}
 
@@
 
-% other contacts
-For the last group of contacts which cover things like leisure activities, grocery
-shopping etc. we have no reliable data by how much policies reduce them. In addition,
-they are likely to be affected by social and psychological factors such as pandemic
-fatigue and vacations. Because of this we estimate them like the infection probabilities
-to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays. Because of the scarce data
-situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with
-friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
-which phases we have and why the switching points are where they are.}
+% other contacts For the last group of contacts which cover things like leisure
+activities, grocery shopping etc. we have no reliable data by how much policies reduce
+them. In addition, they are likely to be affected by social and psychological factors
+such as pandemic fatigue and vacations. Because of this we estimate them like the
+infection probabilities to fit the time series data. We use very few change points and
+tie them to particular events such as policy announcements or particular holidays.
+Because of the scarce data situation we cannot distinguish between a hygiene factor (such
+as mask wearing) during meetings and physical distancing (such as virtual meetings with
+friends).%&%&%&\comment[id=K]{@Janos: Maybe make more concrete when the estimation is
+%&%&%& finished which phases we have and why the switching points are where they are.}
 
@@
 All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
-of each line.}
+seen in Figure~\ref{fig:rapid_test_demand}.%&%&%&\comment[id=J]{Talk about the
+%&%&%& interpretation of each line.}
 

### 2021-06-21 src/paper/appendix_results.tex
Remove all comments by hand for arxiv.

@@
 rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups and federal states.\comment[id=J]{summarize the
-fit}
+analyze the model fit for age groups and federal states.%&%&%&\comment[id=J]{summarize the fit}
 

### 2021-06-21 src/paper/model/course_of_disease.tex
Remove all comments by hand for arxiv.

@@
 \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-documentation}.}.\comment[id=J]{Replace this by a link to the relevant table with
-parameters in the data appendix, once we have that table.}
+documentation}.}.
+%&%&%& \comment[id=J]{Replace this by a link to the relevant table with parameters in the
+%data appendix, once we have that table.}

### 2021-06-21 src/paper/model/number_of_contacts.tex
Clean up comments. Put shapley values into separate file. Put comments back in.
Remove all comments by hand for arxiv.

@@
 
-types of contacts:\comment[id=HM]{I started updating this to understand it better myself.
-In the end, I think it will be useful to group into HH / work / schools and add an introductory sentence for each group.
-E.g., that work is for everybody who is working --- except there is a different model for teachers (?), ...}
+types of contacts:
+%&%&%& \comment[id=HM]{I started updating this to understand it better myself.
+%&%&%& In the end, I think it will be useful to group into HH / work / schools and add an
+%&%&%& introductory sentence for each group. E.g., that work is for everybody who is working ---
+%&%&%& except there is a different model for teachers (?), ...}
 
@@
 % Theoretically, each contact type can have its own infection probability. However, to
-reduce the number of free parameters and thus avoid a potential over-fitting we only
-estimate different infection probabilities for the areas work, school, preschool and
-nurseries, households and other contacts.
+% reduce the number of free parameters and thus avoid a potential over-fitting we only
+% estimate different infection probabilities for the areas work, school, preschool and
+% nurseries, households and other contacts.

### 2021-06-21 src/paper/model/testing.tex
Remove all comments by hand for arxiv.

@@
 should go to symptomatic and asymptomatic individuals from data by the Robert Koch
-Institut.\comment[id=J]{add source} Then, we sample the individuals to which those tests
-are allocated from the pools of symptomatic and asyptomatic infected but not yet tested
-individuals.
+Institut. %&%&%& \comment[id=J]{add source}
+Then, we sample the individuals to which those tests are allocated from the pools of
+symptomatic and asyptomatic infected but not yet tested individuals.
 
@@
 where many tests are done because of symptoms which is in line with the estimates from
-the literature.\comment[id=J]{need source}
+the literature. %&%&%& \comment[id=J]{need source}
 
@@
 administered by medical personnel. While there were concerns that self administered tests
-are less reliable, a recent study has found no basis for this concern.\comment[id=J]{cite
-charite study on self administered tests}
+are less reliable, a recent study has found no basis for this concern.
+%&%&%& \comment[id=J]{cite charite study on self administered tests}
 
 While rapid tests do not directly enter official case numbers, many positively tested
-individuals confirm their rapid test result with a PCR test.\comment[id=J]{Need to
-link to data appendix once the number is there and maybe mention the number}.
+individuals confirm their rapid test result with a PCR test.
+%&%&%& \comment[id=J]{Need to link to data appendix once the number is there and maybe
+%&%&%& mention the number}.
 Importantly, those PCR tests are made in addition to the tests that would have been

### 2021-06-21 src/paper/rapid_test_params.tex
Remove all comments by hand for arxiv.

--- 2021-06-18 paper/rapid_test_params.tex
@@
 
-In our model, there are five reasons why rapid tests are done:\comment[id=J]{Add a
-section on how we calibrate rapid test demand; Mainly describe the datapoints we have and
-say that we usually interpolate linearly in between data points. (Only exception to that
-is private rapid test demand, which we fit to data)}
+In our model, there are five reasons why rapid tests are done:
+%&%&%& \comment[id=J]{Add a
+%&%&%& section on how we calibrate rapid test demand; Mainly describe the datapoints we have and
+%&%&%& say that we usually interpolate linearly in between data points. (Only exception to that
+%&%&%& is private rapid test demand, which we fit to data)}
 
@@
 offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
-60\% until the first days of April \cite{ZDF2021}.\comment[id=K]{ToDo: Find the survey
-that the ZDF is citing here}
+60\% until the first days of April \cite{ZDF2021}.
+%&%&%& \comment[id=K]{ToDo: Find the survey that the ZDF is citing here}
 

### 2021-06-21 src/paper/report.tex
Remove all comments by hand for arxiv.

--- 2021-06-21 src/paper/report.tex
@@
 
-Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
-sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
+Since early 2020,
+%&%&%& \comment[id=J]{Put stronger focus on time dependent rapid test
+%&%&%& sensitivity (30\% before day of onset of infectiousness)}
+the CoViD-19 pandemic has
 presented an enormous challenge to humanity on many dimensions. The development of
@@
 most countries find themselves many months---and often years---away from reaching
-vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
-some more scientific citation before submission} In the meantime, it is of utmost
+vaccination-induced herd immunity \citep{Swaminathan2021}.
+%&%&%& \comment[id=HM]{Need to find some more scientific citation before submission}
+In the meantime, it is of utmost
 importance to employ an effective mix of strategies for containing the virus. The most
@@
 intensive care, and recovery or death \citep[similar
-to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
-transitioning between these states depend on age; their duration is random within
-intervals calibrated to medical literature (for a detailed description see
+to][]{Grimm2021}.
+%&%&%& \comment[id=HM]{Need to look for non-economists}
+The probabilities of transitioning between these states depend on age; their duration is
+random within intervals calibrated to medical literature (for a detailed description see
 Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
@@
 compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons.\comment[id=HM]{Removed ``and whether she
-reduces her contacts if she tests positive or develops symptoms'' -- does not fit here.
-If anything, add footnote, but it seems like this is a detail that is just distracting
-right here. Make sure to describe clearly and at length in Appendix.}
+by employers or follows up on private reasons.
+%&%&%& \comment[id=HM]{Removed ``and whether she reduces her contacts if she tests positive or
+%&%&%& develops symptoms'' -- does not fit here. If anything, add footnote, but it seems like
+%&%&%& this is a detail that is just distracting right here. Make sure to describe clearly and
+%&%&%& at length in Appendix.}
 
@@
 matches features of the data \citep[see][for the general method, also described in
-Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
-more specific once Appendix has new structure} In our application, we keep the number of
-free parameters low in order to avoid overfitting. The data features to be matched
-include official case numbers for each age group and region, deaths, and the share of
-the B.1.1.7 strain.
+Supplementary Material \ref{sec:model}]{McFadden1989}.
+%&%&%& \comment[id=HM]{Make referencemore specific once Appendix has new structure}
+In our application, we keep the number of free parameters low in order to avoid
+overfitting. The data features to be matched include official case numbers for each age
+group and region, deaths, and the share of the B.1.1.7 strain.
 
@@
 testing---the detection rate is two to three times higher for individuals above age 80
-compared to school age children.\comment[id=K]{Must be verified once results are final.}
+compared to school age children.
+%&%&%& \comment[id=K]{Must be verified once results are final.}
 Once rapid test become available, confirmation of a positive result is another reason
@@
 tests. These effect sizes are broadly in line with empirical studies, e.g.,
-\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
-another metric, the effective weekly reproduction number differs by $0.017$ and $0.045$,
-respectively. In light of the large negative effects school closures have on children
-and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
-socio-economic status---these results in conjunction with hindsight bias suggest that
+\citep{Vlachos2021}
+%&%&%& \comment[id=HM]{Need a back-of the envelope calculation}
+To use another metric, the effective weekly reproduction number differs by $0.017$ and
+$0.045$, respectively. In light of the large negative effects school closures have on
+children and parents \citep{Luijten2021, Melegari2021}---and in particular on those with
+low socio-economic status---these results in conjunction with hindsight bias suggest that
 opening schools combined with a testing strategy would have been beneficial. In other

### 2021-06-21 src/paper/report.tex
Put comments back in.

--- 2021-06-21 src/paper/report.tex
@@
 sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
-presented an enormous challenge to humanity on many dimensions. The development of
-highly effective vaccines holds the promise of containment in the medium term. However,
-most countries find themselves many months---and often years---away from reaching
+presented an enormous challenge to humanity on many dimensions. The development of highly
+effective vaccines holds the promise of containment in the medium term. However, most
+countries find themselves many months---and often years---away from reaching
 vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
@@
 frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain
-equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
-theoretical equilibrium at low case numbers which is sustained with
-test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
-over time. Containment measures have become increasingly diverse and now include rapid
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
-influence of seasonal patterns or of more infectious virus strains are well understood
-in quantitative terms.
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
+most have seen large fluctuations of infection rates over time. Containment measures have
+become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
+tracing. Neither these policies' effect nor the influence of seasonal patterns or of more
+infectious virus strains are well understood in quantitative terms.
 
@@
 intensive care, and recovery or death \citep[similar
-to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
-transitioning between these states depend on age; their duration is random within
-intervals calibrated to medical literature (for a detailed description see
+to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists}
+The probabilities of transitioning between these states depend on age; their duration is
+random within intervals calibrated to medical literature (for a detailed description see
 Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
@@
 some days to be processed and there are aggregate capacity constraints throughout. In
-contrast, rapid antigen tests yield immediate results; after a phase-in period, all
-tests that are demanded will be performed. Specificity and sensitivity of these tests is
-set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
-the timing of the test relative to the onset of infectiousness.
+contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
+that are demanded will be performed. Specificity and sensitivity of these tests is set
+according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
+timing of the test relative to the onset of infectiousness.
 Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
@@
 matches features of the data \citep[see][for the general method, also described in
-Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
-more specific once Appendix has new structure} In our application, we keep the number of
-free parameters low in order to avoid overfitting. The data features to be matched
-include official case numbers for each age group and region, deaths, and the share of
-the B.1.1.7 strain.
+Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make referencemore
+specific once Appendix has new structure} In our application, we keep the number of free
+parameters low in order to avoid overfitting. The data features to be matched include
+official case numbers for each age group and region, deaths, and the share of the B.1.1.7
+strain.
 
@@
 tests. These effect sizes are broadly in line with empirical studies, e.g.,
-\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
-another metric, the effective weekly reproduction number differs by $0.017$ and $0.045$,
-respectively. In light of the large negative effects school closures have on children
-and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
-socio-economic status---these results in conjunction with hindsight bias suggest that
+\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation}
+To use another metric, the effective weekly reproduction number differs by $0.017$ and
+$0.045$, respectively. In light of the large negative effects school closures have on
+children and parents \citep{Luijten2021, Melegari2021}---and in particular on those with
+low socio-economic status---these results in conjunction with hindsight bias suggest that
 opening schools combined with a testing strategy would have been beneficial. In other

### 2021-06-21 src/paper/model/testing.tex

--- 2021-06-21 src/paper/model/testing.tex
@@
 should go to symptomatic and asymptomatic individuals from data by the Robert Koch
-Institut.\comment[id=J]{add source} Then, we sample the individuals to which those tests
-are allocated from the pools of symptomatic and asyptomatic infected but not yet tested
-individuals.
+Institut. \comment[id=J]{add source}
+Then, we sample the individuals to which those tests are allocated from the pools of
+symptomatic and asyptomatic infected but not yet tested individuals.
 
@@
 where many tests are done because of symptoms which is in line with the estimates from
-the literature.\comment[id=J]{need source}
+the literature. \comment[id=J]{need source}
 
@@
 While rapid tests do not directly enter official case numbers, many positively tested
-individuals confirm their rapid test result with a PCR test.\comment[id=J]{Need to
-link to data appendix once the number is there and maybe mention the number}.
-Importantly, those PCR tests are made in addition to the tests that would have been
-done otherwise. Section~\ref{subsec:appendix_share_known_cases} discusses the effect
-of rapid tests on the share of detected cases.
+individuals confirm their rapid test result with a PCR test.\comment[id=J]{Need to link
+to data appendix once the number is there and maybe mention the number}. Importantly,
+those PCR tests are made in addition to the tests that would have been done otherwise.
+Section~\ref{subsec:appendix_share_known_cases} discusses the effect of rapid tests on
+the share of detected cases.

### 2021-06-22 src/paper/report.tex
Add some comments. Add all tex files as pytask dependencies. Fix typo in README on how to propose replacements.

@@
 (Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
-approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
-model most closely related to ours is described in \citet{Hinch2020}.} Each contact
-between an infectious individual and somebody susceptible to the disease bears the risk
-of transmitting the virus. Contacts occur in up to four networks: Within the household,
-at work, at school, or in other settings (leisure activities, grocery shopping, medical
+approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The model
+most closely related to ours is described in \citet{Hinch2020}.} Each contact between an
+infectious individual and somebody susceptible to the disease bears the risk of
+transmitting the virus. Contacts occur in up to four networks: Within the household, at
+work, at school, or in other settings (leisure activities, grocery shopping, medical
 appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
@@
 network-specific frequencies of contacts from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
-differentiated by age and geographical location (see
-Section~\ref{sub:number_of_contacts} of the Supplementary Material).
+pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network,
+\replaced[id=K]{meeting probabilities depend on}{contacts are differentiated by}
+age and
+geographical location (see Section~\ref{sub:number_of_contacts} of the Supplementary
+Material).
 
@@
 
-    \floatfoot{\noindent \textit{Note:}
-        A
-        % formal
+    \floatfoot{\noindent \textit{Note:} A        % formal
         description of the model can be found in Supplementary Material~\ref{sec:model}.
@@
         NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
-        occur when a susceptible agent meets an infectious agent; the probability
-        depends on the type of contact, on seasonality, and on NPIs. If infected, the
-        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}.
-        If rapid tests are avaible, agents' demand is modeled as in
-        Figure~\ref{fig:model_rapid_tests}. All pupils are attending school are tested,
-        other reasons trigger a test only for a fraction of individuals.
-        Figure~\ref{fig:model_official_cases} shows the model of translating all
-        infections in the simulated data to age-specific recorded infections. The model
-        uses data on the aggregate share of recorded cases, on the share of positive PCR
-        tests triggered by symptoms, and on the false positive rate of rapid tests. The
-        lower part of the graph is relevant only for periods where rapid tests are
-        available.}
+        occur when a susceptible agent meets an infectious agent; the probability depends
+        on the type of contact, on seasonality, and on NPIs. If infected, the infection
+        progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
+        tests are available, agents' demand is modeled as in
+        Figure~\ref{fig:model_rapid_tests}. \textcolor{blue}{For each reason only a share
+        of eligible individuals receive a test. The share of eligible individuals that
+        get tested depends on the reason and changes over time (see
+        Figure~\ref{fig:rapid_test_demand}).} \textcolor{red}{All pupils are attending
+        school are tested, other reasons trigger a test only for a fraction of
+        individuals.} Figure~\ref{fig:model_official_cases} shows the model of
+        translating all infections in the simulated data to age-specific recorded
+        infections. The model uses data on the aggregate share of recorded cases, on the
+        share of positive PCR tests triggered by symptoms, \textcolor{red}{and on the
+        false positive rate of rapid tests}. The lower part of the graph is relevant only
+        for periods where rapid tests are available.}
 \end{figure}
@@
 all infections. We thus model official cases as depicted in
-Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
-detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As the share
-of asymptomatic individuals varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age
-group over time in our model). Our estimates suggest that---in the absence of rapid
-testing---the detection rate is two to three times higher for individuals above age 80
-compared to school age children.\comment[id=K]{Must be verified once results are final.}
-Once rapid test become available, confirmation of a positive result is another reason
-leading to positive PCR tests.
+Figure~\ref{fig:model_official_cases}.\comment[id=K]{I don't understand why we mention
+the false positive rates in the floatfoot of this. I would either mention rapid test
+sensitivity and specificity in general or not mention them at all.} We take aggregate
+estimates of the share of detected cases and use data on whether CoViD-19 symptoms led to
+a PCR test. As the share of asymptomatic individuals varies by age group, this gives us
+age-specific shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of
+known cases by age group over time in our model). Our estimates suggest that---in the
+absence of rapid testing---the detection rate is two to three times higher for
+individuals above age 80 compared to school age children.\comment[id=K]{Must be verified
+once results are final.} Once rapid test become available, confirmation of a positive
+result is another reason leading to positive PCR tests.
 

### 2021-06-22 task_compile_documents.py

@@ -6,6 +6,8 @@ import pytask
 
 ROOT = Path(__file__).parent
 BLD = ROOT / "bld"
+SRC = ROOT / "src"
+_DEPENDENCIES = SRC.rglob("*.tex")
 
 
 @pytask.mark.latex(
@@ -18,7 +20,7 @@ BLD = ROOT / "bld"
         "-f",
     ]
 )
-@pytask.mark.depends_on(ROOT / "src" / "paper.tex")
+@pytask.mark.depends_on([SRC / "paper.tex", *_DEPENDENCIES])
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass

### 2021-06-22 src/paper/report.tex
More comments.

@@
 
-The four contact networks are chosen so that the most common NPIs can be modeled in
-great detail. NPIs affect the number of contacts or the risk of transmitting the disease
-upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from home will reduce the number of work
-contacts to zero for a fraction of the working population. Schools and daycare can be
-closed entirely, operate at reduced capacity---including an alternating schedule---, or
+The four contact networks are chosen so that the most common NPIs can be modeled in great
+detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
+having physical contact. The effect of different NPIs will generally vary across contact
+types. For example, a mandate to work from home will reduce the number of work contacts
+to zero for a fraction of the working population. Schools and daycare can be closed
+entirely, operate at reduced capacity---including an alternating schedule---, or
 implement mitigation measures like masking requirements or air filters
-\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
-settings. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact \citep{Cheng2021}.
+\citep{Lessler2021}. Curfews may reduce the number of contacts in
+non-work/non-school/non-work settings. In any setting, measures like masking requirements
+would reduce the probability of infection associated with a contact \citep{Cheng2021}.
 
@@
 intensive care, and recovery or death \citep[similar
-to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists}
-The probabilities of transitioning between these states depend on age; their duration is
-random within intervals calibrated to medical literature (for a detailed description see
+to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
+transitioning between these states depend on age; their duration is random within
+intervals calibrated to medical literature (for a detailed description see
 Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
-independent of age \citep{Jones2021}.
+independent of age \citep{Jones2021}.\comment[id=K]{This might be misleading because the
+susceptibility to catch the virus \textbf{is} age-dependent.}
 
@@
 contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
-that are demanded will be performed. Specificity and sensitivity of these tests is set
-according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the onset of infectiousness.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require students to be tested regularly. Rapid tests may be offered by employers for
-on-site workers. Individuals may demand tests for private reasons, which include having
-plans to meet other people,\footnote{A positive test will make them reduce their
-contacts; this is why tests impact the actual contacts in
-Figure~\ref{fig:model_description}.} showing symptoms of CoViD-19, and because a
-household member tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid tests offered
-by employers or follows up on private reasons.\comment[id=HM]{Removed ``and whether she
-reduces her contacts if she tests positive or develops symptoms'' -- does not fit here.
-If anything, add footnote, but it seems like this is a detail that is just distracting
-right here. Make sure to describe clearly and at length in Appendix.}
+that are demanded will be performed.\comment[id=K]{Might be problematic to talk about
+supply and demand here. We don't have supply and demand (except for maybe employers and
+employees) and I would not interpret the increases in April as an easing of supply side
+constraints.} Specificity and sensitivity of these tests is set according to data
+analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test
+relative to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our
+model for rapid test demand. Schools may require staff and students to be tested
+regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
+demand tests for private reasons, which include having plans to meet other people,
+showing symptoms of CoViD-19, and because a household member tested positively for the
+virus. We endow each agent with an individual compliance parameter. This parameter
+determines whether she takes up rapid tests\deleted[id=K]{offered by employers or
+follows up on private reasons} % compliance also determines who is tested in education settings
+.\footnote{A positive test result (of either kind), as well as symptoms leads most
+individuals to reduce their contacts; this is why tests impact the actual contacts in
+Figure~\ref{fig:model_description}.}
+
+%%% ===================================================================================
 

### 2021-06-22 src/paper/report.tex
Update group specific share known cases number.

@@
 matches features of the data \citep[see][for the general method, also described in
-Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make referencemore
+Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference more
 specific once Appendix has new structure} In our application, we keep the number of free
@@
 known cases by age group over time in our model). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is two to three times higher for
-individuals above age 80 compared to school age children.\comment[id=K]{Must be verified
-once results are final.} Once rapid test become available, confirmation of a positive
-result is another reason leading to positive PCR tests.
+absence of rapid testing---the detection rate is on average 80\% higher for individuals
+above age 80 compared to school age children. Once rapid test become available,
+confirmation of a positive result is another reason leading to positive PCR tests.
 

### 2021-06-22 src/paper/appendix_data.tex
Update paper with new results.

--- 2021-06-21 src/paper/appendix_data.tex
@@
 
-\subsection{The Synthetic Population\comment[id=K]{To be written}}
+\subsection{The Synthetic Population \comment[id=K]{To be written}}
 
@@
 geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.\comment[id=HM]{Redo
-\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of
-contacts or better add a similar figure showing total number of contacts by age in all
-networks}
+within the same county, we can calibrate the assortative mixing by age from the same data
+we use to calibrate the number of contacts.\comment[id=HM]{Redo
+\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of contacts
+or better add a similar figure showing total number of contacts by age in all networks}
 
@@
 
-% other contacts
-For the last group of contacts which cover things like leisure activities, grocery
-shopping etc. we have no reliable data by how much policies reduce them. In addition,
+For the last group of contacts, which cover things like leisure activities, grocery
+shopping, etc., we have no reliable data by how much policies reduce them. In addition,
 they are likely to be affected by social and psychological factors such as pandemic

### 2021-06-22 src/paper/report.tex
Update paper with new results.

@@
 frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
-most have seen large fluctuations of infection rates over time. Containment measures have
-become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
-tracing. Neither these policies' effect nor the influence of seasonal patterns or of more
-infectious virus strains are well understood in quantitative terms.
+contacts between individuals. While this has allowed some countries to sustain
+equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
+theoretical equilibrium at low case numbers which is sustained with
+test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
+over time. Containment measures have become increasingly diverse and now include rapid
+testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
+influence of seasonal patterns or of more infectious virus strains are well understood
+in quantitative terms.
 
@@
 Germany, where new infections fell by almost 80\% during the month of May 2021. Our
-analysis shows that frequent and large-scale rapid testing caused the bulk of this
-decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that it
-should have a large role for at least as long as vaccinations have not been offered to
-an entire population.
+analysis shows that, aside from seasonality, frequent and large-scale rapid testing
+caused the bulk of this decrease, which is in line with prior predictions
+\citep{Mina2021}. We conclude that it should have a large role for at least as long as
+vaccinations have not been offered to an entire population.
 
@@
 (Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
-approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The model
-most closely related to ours is described in \citet{Hinch2020}.} Each contact between an
-infectious individual and somebody susceptible to the disease bears the risk of
-transmitting the virus. Contacts occur in up to four networks: Within the household, at
-work, at school, or in other settings (leisure activities, grocery shopping, medical
+approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
+model most closely related to ours is described in \citet{Hinch2020}.} Each contact
+between an infectious individual and somebody susceptible to the disease bears the risk
+of transmitting the virus. Contacts occur in up to four networks: Within the household,
+at work, at school, or in other settings (leisure activities, grocery shopping, medical
 appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
@@
 network-specific frequencies of contacts from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network,
-\replaced[id=K]{meeting probabilities depend on}{contacts are differentiated by}
-age and
-geographical location (see Section~\ref{sub:number_of_contacts} of the Supplementary
-Material).
+pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
+depend on age and geographical location (see Supplementary Material
+Section~\ref{sub:number_of_contacts}).
 
-The four contact networks are chosen so that the most common NPIs can be modeled in great
-detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
-having physical contact. The effect of different NPIs will generally vary across contact
-types. For example, a mandate to work from home will reduce the number of work contacts
-to zero for a fraction of the working population. Schools and daycare can be closed
-entirely, operate at reduced capacity---including an alternating schedule---, or
+The four contact networks are chosen so that the most common NPIs can be modeled in
+great detail. NPIs affect the number of contacts or the risk of transmitting the disease
+upon having physical contact. The effect of different NPIs will generally vary across
+contact types. For example, a mandate to work from home will reduce the number of work
+contacts to zero for a fraction of the working population. Schools and daycare can be
+closed entirely, operate at reduced capacity---including an alternating schedule---, or
 implement mitigation measures like masking requirements or air filters
 \citep{Lessler2021}. Curfews may reduce the number of contacts in
-non-work/non-school/non-work settings. In any setting, measures like masking requirements
-would reduce the probability of infection associated with a contact \citep{Cheng2021}.
+non-work/non-school/non-work settings. In any setting, measures like masking
+requirements would reduce the probability of infection associated with a contact
+\citep{Cheng2021}.
 
@@
 
-    \floatfoot{\noindent \textit{Note:} A        % formal
+    \floatfoot{\noindent \textit{Note:} A% formal
         description of the model can be found in Supplementary Material~\ref{sec:model}.
@@
         NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
-        occur when a susceptible agent meets an infectious agent; the probability depends
-        on the type of contact, on seasonality, and on NPIs. If infected, the infection
-        progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
-        tests are available, agents' demand is modeled as in
-        Figure~\ref{fig:model_rapid_tests}. \textcolor{blue}{For each reason only a share
-        of eligible individuals receive a test. The share of eligible individuals that
-        get tested depends on the reason and changes over time (see
-        Figure~\ref{fig:rapid_test_demand}).} \textcolor{red}{All pupils are attending
-        school are tested, other reasons trigger a test only for a fraction of
-        individuals.} Figure~\ref{fig:model_official_cases} shows the model of
+        occur when a susceptible agent meets an infectious agent; the probability
+        depends on the type of contact, on seasonality, and on NPIs. If infected, the
+        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}.
+        If rapid tests are available, agents' demand is modeled as in
+        Figure~\ref{fig:model_rapid_tests}. All reasons trigger a test only for a
+        fraction of individuals depending on an individual compliance parameter; the
+        thresholds for triggering test demand differ across reasons and they may depend
+        on calendar time. Figure~\ref{fig:model_official_cases} shows the model of
         translating all infections in the simulated data to age-specific recorded
         infections. The model uses data on the aggregate share of recorded cases, on the
-        share of positive PCR tests triggered by symptoms, \textcolor{red}{and on the
-        false positive rate of rapid tests}. The lower part of the graph is relevant only
-        for periods where rapid tests are available.}
+        share of positive PCR tests triggered by symptoms, and on the false positive
+        rate of rapid tests. The lower part of the graph is relevant only for periods
+        where rapid tests are available.}
 \end{figure}
@@
 symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar
-to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
-transitioning between these states depend on age; their duration is random within
-intervals calibrated to medical literature (for a detailed description see
-Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
-independent of age \citep{Jones2021}.\comment[id=K]{This might be misleading because the
-susceptibility to catch the virus \textbf{is} age-dependent.}
+intensive care, and recovery or death \citep[similar to][]{Grimm2021}.
+\comment[id=HM]{Need to look for non-economists} The probabilities of transitioning
+between these states depend on age; their duration is random within intervals calibrated
+to medical literature (for a detailed description see Section~\ref{sec:medical_params}).
+Conditional on the type of contact, infectiousness is independent of age
+\citep{Jones2021}.
 
@@
 some days to be processed and there are aggregate capacity constraints throughout. In
-contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
-that are demanded will be performed.\comment[id=K]{Might be problematic to talk about
-supply and demand here. We don't have supply and demand (except for maybe employers and
-employees) and I would not interpret the increases in April as an easing of supply side
-constraints.} Specificity and sensitivity of these tests is set according to data
-analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test
-relative to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
-demand tests for private reasons, which include having plans to meet other people,
-showing symptoms of CoViD-19, and because a household member tested positively for the
-virus. We endow each agent with an individual compliance parameter. This parameter
-determines whether she takes up rapid tests\deleted[id=K]{offered by employers or
-follows up on private reasons} % compliance also determines who is tested in education settings
-.\footnote{A positive test result (of either kind), as well as symptoms leads most
+contrast, rapid antigen tests yield immediate results; after a phase-in period, all
+tests that are demanded will be performed. Specificity and sensitivity of these tests is
+set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
+the timing of the test relative to the onset of infectiousness.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require staff and students to be tested regularly. Rapid tests may be offered by
+employers to on-site workers. Individuals may demand tests for private reasons, which
+include having plans to meet other people, showing symptoms of CoViD-19, and because a
+household member tested positively for the virus. We endow each agent with an individual
+compliance parameter. This parameter determines whether she takes up rapid
+tests.\footnote{A positive test result (of either kind), as well as symptoms leads most
 individuals to reduce their contacts; this is why tests impact the actual contacts in
 Figure~\ref{fig:model_description}.}
-
-%%% ===================================================================================
 
@@
 matches features of the data \citep[see][for the general method, also described in
-Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference more
-specific once Appendix has new structure} In our application, we keep the number of free
-parameters low in order to avoid overfitting. The data features to be matched include
-official case numbers for each age group and region, deaths, and the share of the B.1.1.7
-strain.
+Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
+more specific once Appendix has new structure} In our application, we keep the number of
+free parameters low in order to avoid overfitting. The data features to be matched
+include official case numbers for each age group and region, deaths, and the share of
+the B.1.1.7 strain.
 
@@
 all infections. We thus model official cases as depicted in
-Figure~\ref{fig:model_official_cases}.\comment[id=K]{I don't understand why we mention
-the false positive rates in the floatfoot of this. I would either mention rapid test
-sensitivity and specificity in general or not mention them at all.} We take aggregate
-estimates of the share of detected cases and use data on whether CoViD-19 symptoms led to
-a PCR test. As the share of asymptomatic individuals varies by age group, this gives us
-age-specific shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of
-known cases by age group over time in our model). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is on average 80\% higher for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
+Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
+detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As the share
+of asymptomatic individuals varies by age group, this gives us age-specific shares (see
+Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age
+group over time in our model). Our estimates suggest that---in the absence of rapid
+testing---the detection rate is 80\% higher on average for individuals above age 80
+compared to school age children. Once rapid test become available, confirmation of a
+positive result is another reason leading to positive PCR tests.
 
@@
         Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
-        a value of 1 is refers to the situation at the start of our sample period and 0
+        a value of 1 refers to the situation at the start of our sample period and 0
         means that all NPIs included in the index are turned on. The other lines in
@@
 the final level is always the highest in case of the vaccination campaign (orange
-lines). The Shapley value decomposition shows that vaccinations contribute 17\% to the
+lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
 cumulative difference between scenarios. Reasons for this are the slow start---it took
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace \citep[this is in line with results in][]{Gavenciak2021}. Nevertheless,
-we estimate seasonality to be a quantitatively important factor determining the
-evolution of the pandemic, explaining most of the early changes and 38\% of the
-cumulative difference by the end of May.
+much lower pace \citep[this is in line with results in][, which our seasonality measure
+is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
+quantitatively important factor determining the evolution of the pandemic, explaining
+most of the early changes and 43\% of the cumulative difference by the end of May.
 
-The largest effect---45\% in the decompositions---comes from rapid testing. Here, it is
-crucial to differentiate between recorded cases and actual cases. Additional testing
+A similar-sized effect---41\% in the decompositions---comes from rapid testing. Here, it
+is crucial to differentiate between recorded cases and actual cases. Additional testing
 means that additional infections will be recorded which would otherwise remain
@@
 Until late April, recorded cases are higher in the scenario with rapid testing alone
-when compared to the setting where none of the three mechanisms are turned on. Compared
-to the scenario with vaccinations only, this point is reached around mid-May and it
-would be June for the comparison with the seasonality-only scenario. The effect on total
-cases, however, is visible immediately in
+when compared to the setting where none of the three mechanisms are turned on. The
+effect on total cases, however, is visible immediately in
 Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\%  of the
@@
 and the relatively small number of students.\footnote{18\% of our population are in the
-education sector (pupils, teachers etc.); 46\% are workers outside the education sector}
+education sector (pupils, teachers, etc.); 46\% are workers outside the education sector}
 Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for
@@
 effect as a situation with closed schools. Under fully opened schools with mandatory
-tests, total infections would have been 7\% higher; this number rises to 20\% without
+tests, total infections would have been 6\% higher; this number rises to 20\% without
 tests. These effect sizes are broadly in line with empirical studies, e.g.,
-\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation}
-To use another metric, the effective weekly reproduction number differs by $0.017$ and
-$0.045$, respectively. In light of the large negative effects school closures have on
-children and parents \citep{Luijten2021, Melegari2021}---and in particular on those with
-low socio-economic status---these results in conjunction with hindsight bias suggest that
+\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
+another metric, the effective weekly reproduction number differs by $0.018$ and $0.052$,
+respectively. In light of the large negative effects school closures have on children
+and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
+socio-economic status---these results in conjunction with hindsight bias suggest that
 opening schools combined with a testing strategy would have been beneficial. In other
@@
 work from home. Whether the share of workers working at the usual workplace is reduced
-or increased by ten percent hardly has an effect on infection rates. Making testing
-mandatory twice a week---assuming independent compliance by employers and workers or
-95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
-employers to the level of March would have increased infections by the same amount.
+or increased by ten percent changes infection rates by 2.5\% or less in either
+direction. Making testing mandatory twice a week---assuming independent compliance by
+employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
+rapid tests offers by employers to the level of March would have increased infections by
+13\%.
 

### 2021-06-22 src/paper/appendix_data.tex
Add seasonality description.

@@
 
-For the last group of contacts, which cover things like leisure activities, grocery
-shopping, etc., we have no reliable data by how much policies reduce them. In addition,
-they are likely to be affected by social and psychological factors such as pandemic
-fatigue and vacations. Because of this we estimate them like the infection probabilities
-to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays. Because of the scarce data
-situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with
-friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
-which phases we have and why the switching points are where they are.}
+% other contacts For the last group of contacts which cover things like leisure
+activities, grocery shopping etc. we have no reliable data by how much policies reduce
+them. In addition, they are likely to be affected by social and psychological factors
+such as pandemic fatigue and vacations. Because of this we estimate them like the
+infection probabilities to fit the time series data. We use very few change points and
+tie them to particular events such as policy announcements or particular holidays.
+Because of the scarce data situation we cannot distinguish between a hygiene factor (such
+as mask wearing) during meetings and physical distancing (such as virtual meetings with
+friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is
+finished which phases we have and why the switching points are where they are.}
 
@@
     outside we assume two seasonality factors. One for other contacts which we expect to
-    be strongly affected by fairer weather with a maximum reduction of 30\% in the
-    infection probability. The other seasonality only makes contacts up to 20\% less
+    be strongly affected by fairer weather with a maximum reduction of 42\% in the
+    infection probability. The other seasonality only makes contacts up to 21\% less
     infectious and is applied to household, work and school contacts.}

### 2021-06-22 src/paper/report.tex

@@
 tests, total infections would have been 6\% higher; this number rises to 20\% without
-tests. These effect sizes are broadly in line with empirical studies, e.g.,
-\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
+tests. These effect sizes are broadly in line with empirical studies
+\citep[e.g.,][]{Vlachos2021}. \comment[id=HM]{Need a back-of the envelope calculation} To use
 another metric, the effective weekly reproduction number differs by $0.018$ and $0.052$,

### 2021-06-22 src/paper/report.tex
Temp.

--- 2021-06-22 src/paper/report.tex
@@
         tests are available, agents' demand is modeled as in
-        Figure~\ref{fig:model_rapid_tests}. \textcolor{blue}{For each reason only a share
-        of eligible individuals receive a test. The share of eligible individuals that
-        get tested depends on the reason and changes over time (see
-        Figure~\ref{fig:rapid_test_demand}).} \textcolor{red}{All pupils are attending
-        school are tested, other reasons trigger a test only for a fraction of
-        individuals.} Figure~\ref{fig:model_official_cases} shows the model of
-        translating all infections in the simulated data to age-specific recorded
-        infections. The model uses data on the aggregate share of recorded cases, on the
-        share of positive PCR tests triggered by symptoms, \textcolor{red}{and on the
-        false positive rate of rapid tests}. The lower part of the graph is relevant only
-        for periods where rapid tests are available.}
+        Figure~\ref{fig:model_rapid_tests}. For each reason only a share of eligible
+        individuals receive a test depending on an individual compliance parameter.
+        Figure~\ref{fig:model_official_cases} shows the model of translating all
+        infections in the simulated data to age-specific recorded infections. The model
+        uses data on the aggregate share of recorded cases, on the share of positive PCR
+        tests triggered by symptoms, and on the false positive rate of rapid tests. The
+        lower part of the graph is relevant only for periods where rapid tests are
+        available.}
 \end{figure}
@@
 contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
-that are demanded will be performed.\comment[id=K]{Might be problematic to talk about
-supply and demand here. We don't have supply and demand (except for maybe employers and
-employees) and I would not interpret the increases in April as an easing of supply side
-constraints.} Specificity and sensitivity of these tests is set according to data
-analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test
-relative to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
-demand tests for private reasons, which include having plans to meet other people,
-showing symptoms of CoViD-19, and because a household member tested positively for the
-virus. We endow each agent with an individual compliance parameter. This parameter
-determines whether she takes up rapid tests\deleted[id=K]{offered by employers or
-follows up on private reasons} % compliance also determines who is tested in education settings
+that are demanded will be performed. Specificity and sensitivity of these tests is set
+according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
+timing of the test relative to the onset of infectiousness.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require staff and students to be tested regularly. Rapid tests may be offered by
+employers to on-site workers. Individuals may demand tests for private reasons, which
+include having plans to meet other people, showing symptoms of CoViD-19, and because a
+household member tested positively for the virus. We endow each agent with an individual
+compliance parameter. This parameter determines whether she takes up rapid
+tests\deleted[id=K]{offered by employers or follows up on private reasons}% compliance also determines who is tested in education settings
 .\footnote{A positive test result (of either kind), as well as symptoms leads most
@@
 Figure~\ref{fig:model_description}.}
-
-%%% ===================================================================================
 
@@
 all infections. We thus model official cases as depicted in
-Figure~\ref{fig:model_official_cases}.\comment[id=K]{I don't understand why we mention
-the false positive rates in the floatfoot of this. I would either mention rapid test
-sensitivity and specificity in general or not mention them at all.} We take aggregate
-estimates of the share of detected cases and use data on whether CoViD-19 symptoms led to
-a PCR test. As the share of asymptomatic individuals varies by age group, this gives us
-age-specific shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of
-known cases by age group over time in our model). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is on average 80\% higher for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
+Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
+detected cases and use data on \replaced[id=K]{the share of PCR tests administered to
+people with CoViD-19 symptoms}{whether CoViD-19 symptoms led to a PCR test}. As the share
+of asymptomatic individuals varies by age group, this gives us age-specific shares (see
+Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age group
+over time in our model). Our estimates suggest that---in the absence of rapid
+testing---the detection rate is on average 80\% higher for individuals above age 80
+compared to school age children. Once rapid test become available, confirmation of a
+positive result is another reason leading to positive PCR tests.
 
@@
 the three broad contact networks. For example, a value of 0.75 for the work multiplier
-means that if the environment was the same as in September (levels of infection rates,
-no rapid tests or vaccinations, only the wildtype virus present), infections at the
+means that if the environment was the same as in September (levels of infection rates, no
+rapid tests or vaccinations, only the wildtype virus present), infections at the
 workplace would be reduced by 25\%. The lines show the product of the effect of contact
@@
 the most stringent regulations are associated with the period of strong decreases in new
-infections between late December 2020 and mid-February 2021. The reversal of the trend
-is associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
-during May 2021 is associated with at least weekly rapid tests to 42\% of the
-population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
-impacting a fall in the relative infectiousness of contacts outside of work and school.
+infections between late December 2020 and mid-February 2021. The reversal of the trend is
+associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
+during May 2021 is associated with at least weekly rapid tests for 42\% of the
+population, a vaccination rate that rose from 28\% to 43\%, and \deleted[id=K]{mostly}
+seasonality \replaced[id=K]{further lowering}{impacting a fall in} the relative
+infectiousness of contacts \deleted[id=K]{outside of work and school.}
 
 In order to better understand the contributions  of rapid tests, vaccinations, and
-seasonality on the evolution of infections in 2021,
-Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
-same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
-model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
-without any of the three factors (red line), and three scenarios turning these factors
-off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
+considers various scenarios. NPIs are always the same as in the baseline scenario.
+Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
+Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
+and three scenarios turning \replaced[id=K]{each factor on individually}{these factors
+off one by one}. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
 infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
@@
 Until mid-March, there is no visible difference between the different scenarios.
-Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
+Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
 Even thereafter, the effect of the vaccination campaign is surprisingly small at first
@@
 contribute less to the spread of the disease than others due to a lower number of
-contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
-study period, when first-dose vaccination rates reached around 40\% of the population,
-the numbers of new cases would have started to decline. It is important to note that the
-initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
-fatality was rate considerably lower during the third wave when compared to the second
-(4.4\% between October and February and 1.4\% between March and the end of May).
+contacts\replaced[id=K]{as they do not work, do not go to school and tend to live in small households}{, see Supplementary Material~\ref{subsec:contacts_by_age}}.
+% We don't draw age specific number of contacts because the sample sizes would be to
+% small.
+By the end of our study period, when first-dose vaccination rates reached around 40\% of
+the population, the numbers of new cases would have started to decline. It is important
+to note that the initial focus of the campaign was to prevent deaths and severe disease.
+Indeed, the case fatality was rate considerably lower during the third wave when compared
+to the second (4.4\% between October and February and 1.4\% between March and the end of
+May).\comment[id=K]{This needs a source.}
 

### 2021-06-22 src/figures/results/tables/student_infections_table.tex
Create arXiv preprint v2. (#3)

@@
 \midrule
- educ open after easter  without tests &                                             751930 \\
- educ open after easter  with tests    &                                             567537 \\
- close educ after easter               &                                             436232 \\
- baseline                              &                                             452727 \\
+ educ open after easter  without tests &                                             725914 \\
+ educ open after easter  with tests    &                                             552894 \\
+ close educ after easter               &                                             430594 \\
+ baseline                              &                                             439077 \\
 \bottomrule

### 2021-06-22 src/paper/abstract.tex
Create arXiv preprint v2. (#3)

--- 2021-06-17 paper/abstract.tex
@@
 \noindent
-% Abstracts of Research Articles and Reports should explain to the general reader why
-% the research was done, what was found and why the results are important. 
-% 
-% They should start with some brief BACKGROUND information: a sentence giving a broad
-% introduction to the field comprehensible to the general reader, and then a sentence of
-% more detailed background specific to your study. 
 In order to slow the spread of the CoViD-19 pandemic, governments around the world have
@@
 the pandemic, allowing for factors like seasonality or virus strains with different
-transmission profiles. To do so, the study develops an agent-based simulation
-model, which is estimated using data for the second and the third wave of the CoViD-19
-pandemic in Germany. The paper finds that during a period where vaccination rates rose
-from 5\% to 40\%, large-scale rapid testing had the largest effect on reducing infection
-numbers. 
-% The final sentence should outline the main CONCLUSIONS of the study, in terms that
-% will be comprehensible to all our readers. The Abstract is distinct from the main body
-% of the text, and thus should not be the only source of background information critical
-% to understanding the manuscript. Please do not include citations or abbreviations in
-% the Abstract. 
-Frequent large-scale rapid testing should remain part of strategies to contain
-CoViD-19; it can substitute for many non-pharmaceutical interventions that come at a
-much larger cost to individuals, society, and the economy.
-
-% The abstract should be 125 words or less.
+transmission profiles. To do so, the study develops an agent-based simulation model,
+which is estimated using data for the second and the third wave of the CoViD-19 pandemic
+in Germany. The paper finds that during a period where vaccination rates rose from 5\%
+to 40\%, rapid testing had the largest effect on reducing infection numbers. Frequent
+large-scale rapid testing should remain part of strategies to contain CoViD-19; it can
+substitute for many non-pharmaceutical interventions that come at a much larger cost to
+individuals, society, and the economy.
 

### 2021-06-22 src/paper/medical_parameters.tex
Create arXiv preprint v2. (#3)

--- 2021-06-10 paper/medical_parameters.tex
@@
 arrived at the distributions used in the model.\footnotemark See
-Figure~\ref{fig:disease_progression} for a summary of our disease progression model.
+Figure~\ref{fig:model_disease_progression} for a summary of our disease progression model.
 

### 2021-06-22 src/paper/model/seasonality.tex
Create arXiv preprint v2. (#3)

\subsection{Seasonality}
\label{subsec:seasonality}

It is widely acknowledged that the transmission of SARS-CoV-2 is subject to seasonal
influences. Infectiousness is increased in winter when most contacts take place inside
and the immune system is weakened by low levels of vitamin D, dry air and large
temperature swings. For a detailed overview of possible drivers see
\cite{KronfeldSchor2021}.

We follow \cite{Kuehn2020} and \cite{Gavenciak2021} in modeling seasonality in the
transmission of SARS-CoV-2 as a multiplicative factor on infection probabilities. The
factor follows a sine curve that reaches its maximum at January first and its minimum
on June 30.

For simplicity we normalize the factor to reach one at its maximum. Thus the formula of
the seasonality factor is given by:

\begin{equation}
    s_k(t) = 1 + 0.5 \kappa_k  sin \left ( \pi  \left (\frac{1}{2} + \frac{t}{182.5}\right ) \right ) - 0.5 \kappa_k
\end{equation}

Where $\kappa_k$ is difference in the seasonality factor between peak infectiousness
and lowest infectiousness.

The subscript $k$ is needed because the strength of the seasonality effect differs
across contact types: Work, household and school contacts are likely to take place
inside even in summer. Thus they are only subject to seasonality due to factors that
influence the immune system. Other contacts (for example meeting friends and while doing
leisure activities) are mostly happening outside in the summer. Therefore, transmission
via those contacts should have a stronger seasonal pattern.

We calibrate $kappa_{strong}$ to 0.42 and $kappa_{weak}$ to 0.21. This is in line with
\cite{Gavenciak2021} and \cite{Kuehn2020}.

### 2021-06-22 src/paper/model/testing.tex
Create arXiv preprint v2. (#3)

@@
 Section~\ref{subsec:appendix_share_known_cases} discusses the effect of rapid tests on
-the share of detected cases.
+the share of detected cases.\comment[id=K]{Talk here about reactions to tests and
+symptoms with contact reductions. (If not here, find other place where to talk about this
+in detail.)}

### 2021-06-22 src/paper/rapid_test_params.tex
Create arXiv preprint v2. (#3)

--- 2021-06-18 paper/rapid_test_params.tex
@@
 
-In our model, there are five reasons why rapid tests are done:\comment[id=J]{Add a
-section on how we calibrate rapid test demand; Mainly describe the datapoints we have and
-say that we usually interpolate linearly in between data points. (Only exception to that
-is private rapid test demand, which we fit to data)}
+In our model, there are five reasons why rapid tests are done:\comment[id=J]{ Add a
+    section on how we calibrate rapid test demand; Mainly describe the datapoints we have
+    and say that we usually interpolate linearly in between data points. (Only exception
+    to that is private rapid test demand, which we fit to data)
+}
 

### 2021-06-22 src/paper/report.tex
Create arXiv preprint v2. (#3)

--- 2021-06-22 src/paper/report.tex
@@
         NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
-        occur when a susceptible agent meets an infectious agent; the probability
-        depends on the type of contact, on seasonality, and on NPIs. If infected, the
-        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}.
-        If rapid tests are available, agents' demand is modeled as in
+        occur when a susceptible agent meets an infectious agent; the probability depends
+        on the type of contact, on seasonality, and on NPIs. If infected, the infection
+        progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
+        tests are available, agents' demand is modeled as in
         Figure~\ref{fig:model_rapid_tests}. All reasons trigger a test only for a
@@
         infections. The model uses data on the aggregate share of recorded cases, on the
-        share of positive PCR tests triggered by symptoms, and on the false positive
-        rate of rapid tests. The lower part of the graph is relevant only for periods
-        where rapid tests are available.}
+        share of positive PCR tests triggered by symptoms, and on the false positive rate
+        of rapid tests. The lower part of the graph is relevant only for periods where
+        rapid tests are available.}
     \end{figure}
@@
 symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Grimm2021}.
-\comment[id=HM]{Need to look for non-economists} The probabilities of transitioning
-between these states depend on age; their duration is random within intervals calibrated
-to medical literature (for a detailed description see Section~\ref{sec:medical_params}).
-Conditional on the type of contact, infectiousness is independent of age
-\citep{Jones2021}.
+intensive care, and recovery or death \citep[similar
+to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
+transitioning between these states depend on age; their duration is random within
+intervals calibrated to medical literature (for a detailed description see
+Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
+independent of age \citep{Jones2021}.
 
@@
 Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
-detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As the share
+detected cases and use data on \replaced[id=K]{the share of PCR tests administered to
+people with CoViD-19 symptoms}{whether CoViD-19 symptoms led to a PCR test}. As the share
 of asymptomatic individuals varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age
-group over time in our model). Our estimates suggest that---in the absence of rapid
+Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age group
+over time in our model). Our estimates suggest that---in the absence of rapid
 testing---the detection rate is 80\% higher on average for individuals above age 80

### 2021-06-22 src/paper/results.tex
Create arXiv preprint v2. (#3)

--- 2021-06-21 src/paper/appendix_results.tex
@@
-\section{Additional Results}
-\label{sec:additional_results}
+\section{Results}
+\label{sec:results}
 
-\subsection{Simulated vs. Empirical Data}
-\label{subsec:fit_results}
-
-This compares simulated data from our model with empirical data from Germany. We look at
-observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
-rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups and federal states.\comment[id=J]{summarize the
-fit}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
-  \caption{Fit Over the Full Simulation Time Frame with Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported simulated infections rates. The mean infection rate is
-  the thick blue line. Single simulation runs are plotted in lighter and thinner lines.
-  The official case numbers as reported by the Robert-Koch-Institut are plotted in black.
-  The fit is overall very good. The higher the mean incidence and the stronger the growth
-  the more variance there is between simulation runs. We averaged over 30 simulation
-  runs.}
-  \label{fig:aggregated_fit2}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_newly_infected_with_single_runs}
-  \caption{Development of the Total Infections Over the Full Simulation Time Frame with
-  Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the true weekly incidence rates
-  per 100,000 people, including undetected cases. The mean infection rate is the thick
-  blue line. Single simulation runs are plotted in lighter and thinner lines. The higher
-  the mean incidence and the stronger the growth the more variance there is between
-  simulation runs. We averaged over 30 simulation runs.}
-  \label{fig:newly_infected_in_baseline}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Age Group}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different age
-  groups. The age group of individuals above 80 needs to be interpreted with caution
-  because our synthetic population only includes private households, i.e. nursing homes
-  are not represented in our model. They accounted for many cases and deaths in the
-  winter of 2020 and many 80 to 100 year olds live in these facilities. However, the
-  official data does not contain information on whether cases were nursing home
-  inhabitants or not. We averaged over 30 simulation runs.}
-  \label{fig:age_group_fit}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Federal State}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different
-  federal states. We averaged over 30 simulation runs.}
-  \label{fig:state_fit}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
-  \caption{Share of Vaccinated Individuals}
-  \floatfoot{\noindent \textit{Note:} The figure shows the rate of individuals that are
-  vaccinated in our synthetic population versus in the general German population. Note
-  that we excluded the vaccinations that were given to nursing homes, approximately the
-  first percent of the German population that were vaccinated. Overall, our model covers
-  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
-  the population are vaccinated. Our vaccinations work imperfectly but we do not model
-  different vaccines nor do we distinguish between first and second shot.}
-  \label{fig:fit_vaccinations}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
-  \caption{Effective Replication Number $R_t$ in the Model and as Reported by the
-  Robert-Koch-Institute}
-  \floatfoot{\noindent \textit{Note:} The figure shows the effective replication number
-  ($R_t$) as reported by the RKI and as calculated in our model. The $R_t$ gives the
-  average number of new infections caused by one infected individual. The $R_t$ in our
-  model broadly follows the $R_t$ reported by the RKI. Two trends stand out. Firstly, the
-  RKI's $R_t$ drops faster in November. This could be due to a change in the testing
-  policy that focused tests on the elderly when the second wave hit Germany and led to a
-  decline in the overall share of detected cases. The second difference is from mid
-  February to mid March where the RKI's reported $R_t$ increased more rapidly than that
-  in our model. Here the opposite effect can be expected. During this time rapid tests
-  increased strongly leading to more cases being detected. In the short term this leads
-  an $R_t$ estimation that is based on detected cases to overestimate the replication
-  number.}
-  \label{fig:fit_r_effective}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117_with_single_runs}
-  \caption{Share of B.1.1.7 in the Model and as Reported by the Robert-Koch-Institute}
-  \floatfoot{\noindent \textit{Note:} The figure shows the share of B.1.1.7 as
-  reported by the RKI and as calculated in our model. We only introduce a few cases over
-  the cause of January. From then B.1.1.7 takes over endogenously through its increased
-  infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases per 100,000 inhabitants.}
-  \label{fig:fit_share_b117}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Share of Cases that are Detected}
-\label{subsec:appendix_share_known_cases}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group}
-  \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
-  reported as an official case via PCR confirmation. We use the overall share of known
-  cases that was estimated through the case fatality ratio by the
-  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
-  then assume it to be constant as vaccinations of the elderly strongly affect the case
-  fatality rate which the project does not account for. To get from an overall share of
-  detected cases to the share of cases that is detected in each age group we use that
-  asymptomatic cases are much less likely to be detected. As our model covers age
-  specific asymptomatic rates this endogenously leads to group specific share known cases
-  that verify that infections in younger age groups are under-detected. Starting in 2021
-  in addition to the overall numbers of detected cases through symptoms and the share
-  known cases, cases are also detected through confirmation of positive rapid tests. This
-  leads to an increase in the share of known cases for all age groups but in particular
-  for the younger age groups that are covered extensively with rapid tests through the
-  rapid test requirement for participating in school.}
-\end{figure}
-
-It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
-
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/educ_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
-    \label{fig:educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/work_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test due to Work}
-    \label{fig:work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/private_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
-  \label{fig:private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Rapid Test Demand by Reason}
-  \label{fig:rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends.
-  }
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
-    \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
-    Individuals}
-    \label{fig:share_infected_among_educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
-    \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
-    \label{fig:share_infected_among_work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
-    \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
-  \label{fig:share_infected_among_private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Share of Rapid Tests Demanded by Infected Individuals by Reason}
-  \label{fig:share_infected_among_rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. Private rapid tests have a much higher share of infected
-    individuals because they are mostly triggered by events that make an infection
-    likely. Remember however, that one reason is that a household member has a positive
-    rapid test. This means that work and education rapid tests which have a low rate of
-    infected individuals trigger more targeted rapid tests through the household demand.
-    They also have a much higher volume of tests. As can be seen in the decomposition
-    (Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
-    important for the overall effect. It appears that the combination of wide testing to
-    find infection chains plus targeted tests to break those chains taking together
-    is important.}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Scenarios}
-\label{subsec:appendix_scenarios}
+This is the results section.
 
@@
   \label{fig:explain_decline}
-  \floatfoot{\noindent \textit{Note:} \ldots}
+  \figurenotes{\ldots}
 \end{figure}
@@
   \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-  \label{fig:school_scenarios_detailed}
+  \figurenotes{}
+  \label{fig:appdx_school_scenarios}
 \end{figure}
@@
 \FloatBarrier
+
+\input{figures/results/tables/student_infections_table}
+
+\FloatBarrier

### 2021-06-22 task_compile_documents.py

@@ -6,6 +6,8 @@ import pytask
 
 ROOT = Path(__file__).parent
 BLD = ROOT / "bld"
+SRC = ROOT / "src"
+_DEPENDENCIES = SRC.rglob("*.tex")
 
 
 @pytask.mark.latex(
@@ -15,9 +17,10 @@ BLD = ROOT / "bld"
         "--synctex=1",
         "--cd",
         "--shell-escape",
+        "-f",
     ]
 )
-@pytask.mark.depends_on(ROOT / "paper" / "paper.tex")
+@pytask.mark.depends_on([SRC / "paper.tex", *_DEPENDENCIES])
 @pytask.mark.produces(BLD / "paper.pdf")
 def task_compile_documents():
     pass

### 2021-06-23 src/paper.tex
Remove suggestion by Annika.
Add some comments. Add all tex files as pytask dependencies. Fix typo in README on how to propose replacements.
Commenting out, new pdf, and updated readme.
Put comments back in. Minor corrections in the README. Add references to our code in the first paragraph of the appendix.

@@
 
-\input{paper/preamble}
+% To ``accept'' all changes in the document:
+% \PassOptionsToPackage{final}{changes}
+
+\documentclass[a4paper, 10pt]{article}
+% fleqn: flush-left aligned equations
+
+\newcommand{\papertitle}{
+The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs
+}
+
+\input{templates/Preamble_Fundamental_packages}
+\hypersetup{
+    pdftitle={\papertitle},
+    pdfauthor={Janoś Gabler, Tobias Raabe, Klara Röhrl, Hans-Martin von Gaudecker}
+}
+
+% packages ======================================================================
+
+\input{templates/Preamble_Page_layout_general}
+\input{templates/Preamble_Page_layout_paper}
+\subimport{templates/}{Preamble_Fonts_Charter_FiraSans}
+\input{templates/Preamble_Sansserif_math}
+\input{templates/Preamble_BibLaTeX_AER}
+%\input{templates/Preamble_BibLaTeX_APA}
+\input{templates/Preamble_Commenting_and_highlighting}
+\input{templates/Preamble_Miscellaneous}
+
+% Add global hyphenation rules to avoid unwelcome linebreaks:
+% \hyphenation{Bu-shong Della-Vigna}
+
+\bibliography{references.bib}
+
+% \usepackage[newfloat, outputdir=../bld, cache=false]{minted}
+
+% Define environments to comment on the sidebar of the PDF.
+\definechangesauthor[name={Tobias}, color=alizarin]{T}
+\definechangesauthor[name={Janos}, color=internationalorange]{J}
+\definechangesauthor[name={Klara}, color=dartmouthgreen]{K}
+\definechangesauthor[name={HM}, color=blue]{HM}
+% Use \replaced[id=T]{<to-be-replaced>}{<new>} or the same with \deleted to propose
+% changes.
 
@@
 
+    This paper aims to be completely reproducible. The code for the general model can be
+    found at
+    \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
+    and its documentation at
+    \href{https://sid-dev.readthedocs.io/en/latest/}{https://sid-dev.readthedocs.io/en/latest/}.
+    The code for the application to Germany can be found at
+    \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}
+    with a very rudimentary documentation at
+    \href{https://sid-germany.readthedocs.io/en/latest/}{https://sid-germany.readthedocs.io/en/latest/}.
+
     \input{paper/appendix_data}

### 2021-06-23 src/paper/report.tex
Put comments back in. Minor corrections in the README. Add references to our code in the first paragraph of the appendix.

@@
 scenario without rapid tests only (purple line in
-Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
-three channels for rapid tests. Tests at schools have the smallest effect, which is
-largely explained by schools not operating at full capacity during our period of study
-and the relatively small number of students.\footnote{18\% of our population are in the
-education sector (pupils, teachers, etc.); 46\% are workers outside the education sector}
-Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for
-private reasons are phased in only late, they make up for more than half of the total
-effect. The reason lies in the fact that a substantial share of these tests is driven by
-an elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or
-following up on a positive test of a household member. The latter is essentially a form
-of contact tracing, which has been shown to be very effective \citep{Contreras2021,
+Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
+channels for rapid tests. Tests at schools have the smallest effect, which is largely
+explained by schools not operating at full capacity during our period of study and the
+relatively small number of students.\footnote{18\% of our population are in the education
+sector (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost
+40\% come from tests at the workplace. Despite the fact that rapid tests for private
+reasons are phased in only late, they make up for more than half of the total effect. The
+reason lies in the fact that a substantial share of these tests is driven by an elevated
+probability to carry the virus, i.e., showing symptoms of CoViD-19 or following up on a
+positive test of a household member. The latter is essentially a form of contact tracing,
+which has been shown to be very effective \citep{Contreras2021,
 Kretzschmar2020,FetzerGraeber2021}.

### 2021-06-23 src/paper/appendix_results.tex
Add figures on rapid test shares in the results appendix.

--- 2021-06-21 src/paper/appendix_results.tex
@@
 rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups and federal states.\comment[id=J]{summarize the
-fit}
+analyze the model fit for age groups and federal states.\comment[id=J]{summarize the fit}
 
@@
 
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_doing_rapid_test_today}
+  \caption{Share of Individuals Doing a Rapid Test per Day}
+  \label{fig:fit_share_doing_rapid_test}
+  \floatfoot{\noindent \textit{Note:} The plateau in the end is the result of falling
+  case numbers that lead to less private rapid tests being demanded due to contact
+  tracing. The two biggest jumps are due to public holidays, i.e. the end of Easter
+  vacations in the first half of April and May 1st. These lead many individuals to not
+  test themselves (even though the threshold to be tested falls during the time) due to
+  vacations from school and work. As a result more individuals than usual are due to be
+  tested because of frequency requirements after public holidays. }
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+  \caption{Share of Individuals That has Ever Done a Rapid Test}
+  \label{fig:fit_share_ever_rapid_test}
+  \floatfoot{\noindent \textit{Note:} This figure clearly shows that overall our
+  assumptions on rapid tests are conservative. In our model, the share of individuals who
+  have ever done a rapid test lies  consistently 10 percentage points below the share
+  reported in the COSMO study \citep{Betsch2021}. The main reason for this is that we
+  introduce rapid tests only at the start of 2021. However, using the available data to
+  calibrate our rapid test parameters and estimating the remaining parameters to fit the
+  official case numbers we arrive at a shape that is very similar to that implied by the
+  survey results. This together with fitting the share of Germans with weekly rapid tests
+  well (see Figure~\ref{fig:fit_share_rapid_test_within_last_week} makes us confident
+  that our rapid test model is a good -- and especially not over-confident --
+  representation of rapid testing in Germany.}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
+  \caption{Share of Individuals Who Have Done a Rapid Test in the Last Week}
+  \label{fig:fit_share_rapid_test_within_last_week}
+  \floatfoot{\noindent \textit{Note:} The two lines are not perfectly compatible. The
+  COSMO study \citep{Betsch2021} only asked if individuals had done at least one
+  self-administered rapid test per week in the last four weeks. This means there are two
+  differences between the simulated line and the empirical data points. Firstly, the
+  simulated line encompasses all rapid tests since we do not differentiate between
+  self-administered and medically administered rapid tests. Secondly, in our model we
+  cannot verify the rhythm over the last four weeks but just observe the last time of a
+  rapid test. However, both work and school rapid tests which make up the largest share
+  of rapid tests (see Figure~\ref{fig:rapid_test_demand_by_reason}) are both required on
+  a weekly or twice weekly basis while the private rapid test demand which are partly
+  just triggered by events make up only a small fraction of tests. }
+\end{figure}
+
 \FloatBarrier

### 2021-06-23 src/paper/appendix_results.tex
Add figure description to last figures.

@@
   \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
+  \begin{subfigure}[b]{.49\textwidth}
+    \centering
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_new_known_case}
+    \caption{Reported Cases}
+    \label{fig:work_scenarios_new_known_case}
   \end{subfigure}
-  \caption{The Effect of Policies on Observed and Unobserved Cases}
-  \label{fig:explain_decline}
-  \floatfoot{\noindent \textit{Note:} \ldots}
+  \hfill
+  \begin{subfigure}[b]{.49\textwidth}
+    \centering
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+    \caption{Total Cases}
+    \label{fig:work_scenarios_newly_infected}
+  \end{subfigure}
+  \caption{The Effect of Different Work Scenarios on Reported and Total Cases}
+  \label{fig:work_scenarios_detailed}
+  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
+  policy changes took place at Easter until the end of our simulation period (end of
+  May). We vary the share of workers that work from home and how many tests are performed
+  at work relative to our baseline scenario. Making it mandatory to test all employees
+  that do not work from home markedly reduces cases -- even when only assuming 95\%
+  compliance on both the employer and the employee side. As before, the observed cases
+  can be misleading because more testing leads to more detected cases. It takes two to
+  three weeks for the reduction in new infections effect to dominate the increased
+  detection effect. Furthermore, the two opposing effects lead to a smaller effect size
+  than is actually the case.}
 \end{figure}
+
+ \comment[id=K]{Regarding figure \ref{fig:work_scenarios_detailed}: Look up numbers and
+ add them to the description.}
 
@@
   \centering
-  \begin{subfigure}{.6\textwidth}
+  \begin{subfigure}[b]{.49\textwidth}
+    \centering
     \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
+    \caption{Reported Cases}
+    \label{fig:school_scenarios_new_known_case}
   \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
+  \hfill
+  \begin{subfigure}[b]{.49\textwidth}
+    \centering
     \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+    \caption{Total Cases}
+    \label{fig:school_scenarios_newly_infected}
   \end{subfigure}
-  \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
+  \caption{The Effect of Different School Scenarios on Reported and Total Cases}
   \label{fig:school_scenarios_detailed}
+  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
+   policy changes took place at Easter until the end of our simulation period (end of
+   May). Apart from the enacted school policies as our baseline we simulate how cases
+   would have developed if schools had been closed completely as the strictest possible
+   counterfactual scenario and two opening models: One where schools open normally (with
+   hygiene measures) without any testing in the education sector and one where schools
+   open normally but testing shares develop as in the baseline scenario. Our simulations
+   suggest that the enacted policies were as effective as keeping schools closed. Opening
+   schools with the testing schemes that were in place after Easter would have had a
+   small effect on the overall incidence. However, this is mainly due to the stringent
+   testing that was in place in schools by that time. Had schools opened without testing
+   requirements the total incidence would have been up to 50 points higher, though this
+   would have been less visible in the reported cases.}
 \end{figure}
 
+ \comment[id=K]{Regarding figure \ref{fig:school_scenarios_detailed}: Look up numbers and
+ add them to the description.}
+
 \FloatBarrier

### 2021-06-23 src/paper/model/summary.tex
Add figure description to last figures.

@@
 tests there. Moreover, people can decide to make a rapid test if they develop symptoms,
-have many planned contacts or observe cases in their contact network. People who have a
-positive rapid test demand a confirmatory PCR test with a certain probability. Moreover,
-PCR tests can be demanded because of symptoms or randomly.
+have many planned contacts\comment[id=K]{at the moment it's implemented as planing to
+participate in a weekly leisure meeting. Should we be more explicit here?} or observe
+cases in their contact network. People who have a positive rapid test demand a
+confirmatory PCR test with a certain probability. Moreover, PCR tests can be demanded
+because of symptoms or randomly.
 

### 2021-06-25 src/paper/report.tex
Add comments by Philipp Eisenhauer.

@@
 influence of seasonal patterns or of more infectious virus strains are well understood
-in quantitative terms.
+in quantitative terms.\comment[id=K]{feedback from Philipp:
+\begin{itemize}
+    \item he would suggest to lay more focus on explaining the mechanisms behind our
+    results
+    \item he suggests we should try to get some form of confidence bands or standard errors
+    \item he got the wrong impression that we only have three mechanisms and that the
+    rapid tests are the residual.
+\end{itemize}
+}
 

### 2021-06-28 src/paper/literature_review.tex
Create arXiv preprint v2. (#3)
Move all comments in the references to comment fields so they are not printed as part of the bibliography.

--- 2021-04-17 paper/literature_review.tex
@@
-\section{Literature Review}
+\subsection{Literature Review}
 \label{sec:literature_review}
@@
 We build on two strands of literature: Recent extensions of the epidemiological SEIR
-model and agent-based simulation models.
+model and agent-based simulation models.\comment[id=K]{Write a sentence about the
+\href{https://www.itwm.fraunhofer.de/de/presse-publikationen/presseinformationen/2021/2021-06-22_Dritte_Welle_Starker-Effekt-von-Schnelltests-an-Schulen.html}{Frauenhofer
+paper} coming to similar conclusions: }
 

### 2021-06-28 src/paper/appendix_data.tex
Add bullet points for synthetic population.

@@
 Saturdays and Sundays.
+
+\begin{itemize}
+    \item \textcolor{red}{Describe microcensus campus file and which data we use and for what}
+    \item \textcolor{red}{Describe size and sampling of households}
+    \item \textcolor{red}{counties are randomly drawn}\comment[id=K]{Find out why!}
+    \item \textcolor{red}{how we create the occupation column}
+    \item \textcolor{red}{describe how we create the contact model group ids for school,
+    work and other}
+    \item \textcolor{red}{Describe work contact priority, educ contact priority and
+    rapid test compliance parameter and quarantine compliance.}
+    \item \textcolor{red}{vaccination rank creation}
+\end{itemize}
 
@@
 
-% other contacts For the last group of contacts which cover things like leisure
-activities, grocery shopping etc. we have no reliable data by how much policies reduce
-them. In addition, they are likely to be affected by social and psychological factors
-such as pandemic fatigue and vacations. Because of this we estimate them like the
-infection probabilities to fit the time series data. We use very few change points and
-tie them to particular events such as policy announcements or particular holidays.
-Because of the scarce data situation we cannot distinguish between a hygiene factor (such
-as mask wearing) during meetings and physical distancing (such as virtual meetings with
-friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is
-finished which phases we have and why the switching points are where they are.}
+For the last group of contacts, which cover things like leisure activities, grocery
+shopping, etc., we have no reliable data by how much policies reduce them. In addition,
+they are likely to be affected by social and psychological factors such as pandemic
+fatigue and vacations. Because of this we estimate them like the infection probabilities
+to fit the time series data. We use very few change points and tie them to particular
+events such as policy announcements or particular holidays. Because of the scarce data
+situation we cannot distinguish between a hygiene factor (such as mask wearing) during
+meetings and physical distancing (such as virtual meetings with
+friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
+which phases we have and why the switching points are where they are.}
 

### 2021-06-28 src/paper/model/param_tables.tex
Very first version of adding parameter tables.

\subsection{Overview Model Parameters} % (fold)
\label{sub:param_tables}

\begin{table}
\centering
\caption{contacts_and_matching}
\label{tab:contacts_and_matching}
\input{../bld/param_tables/contacts_and_matching}
\end{table}

\begin{table}
\centering
\caption{policies}
\label{tab:policies}
\input{../bld/param_tables/policies}
\end{table}

\begin{table}
\centering
\caption{infection_probs}
\label{tab:infection_probs}
\input{../bld/param_tables/infection_probs}
\end{table}

\begin{table}
\centering
\caption{disease_model}
\label{tab:disease_model}
\input{../bld/param_tables/disease_model}
\end{table}

\begin{table}
\centering
\caption{pcr_testing_and_case_detection}
\label{tab:pcr_testing_and_case_detection}
\input{../bld/param_tables/pcr_testing_and_case_detection}
\end{table}

\begin{table}
\centering
\caption{rapid_testing}
\label{tab:rapid_testing}
\input{../bld/param_tables/rapid_testing}
\end{table}

\begin{table}
\centering
\caption{behavioral_response}
\label{tab:behavioral_response}
\input{../bld/param_tables/behavioral_response}
\end{table}

\begin{table}
\centering
\caption{vaccinations}
\label{tab:vaccinations}
\input{../bld/param_tables/vaccinations}
\end{table}

### 2021-06-28 src/parameter_tables/task_convert_csv_to_latex.py

@@ -1,0 +1,16 @@
+import pytask
+import pandas as pd
+from pathlib import Path
+
+PARAM_TABLE_PATH = Path(__file__).parent
+BLD = PARAM_TABLE_PATH.parent.parent / "bld"
+SRC = PARAM_TABLE_PATH / "csv"
+
+PARAMETRIZATION = [(p, BLD / "param_tables" / f"{p.stem}.tex") for p in SRC.iterdir()]
+
+
+@pytask.mark.parametrize("depends_on, produces", PARAMETRIZATION)
+def task_convert_csv_to_tex(depends_on, produces):
+    df = pd.read_csv(depends_on)
+    with open(produces, "w") as f:
+        f.write(df.to_latex(na_rep=""))

### 2021-06-28 task_compile_documents.py

@@ -7,7 +7,7 @@ import pytask
 ROOT = Path(__file__).parent
 BLD = ROOT / "bld"
 SRC = ROOT / "src"
-_DEPENDENCIES = SRC.rglob("*.tex")
+_DEPENDENCIES = [*SRC.rglob("*.tex"), *(BLD / "param_tables").rglob("*.tex")]
 
 
 @pytask.mark.latex(

### 2021-06-29 src/parameter_tables/task_convert_csv_to_latex.py
Nicely format tables.

@@ -12,5 +12,16 @@ PARAMETRIZATION = [(p, BLD / "param_tables" / f"{p.stem}.tex") for p in SRC.iter
 @pytask.mark.parametrize("depends_on, produces", PARAMETRIZATION)
 def task_convert_csv_to_tex(depends_on, produces):
     df = pd.read_csv(depends_on)
+    df = df.dropna(axis=0)  # drop empty rows
     with open(produces, "w") as f:
-        f.write(df.to_latex(na_rep=""))
+        with pd.option_context("max_colwidth", 10000):
+            table = df.to_latex(
+                na_rep="",
+                escape=False,
+                longtable=False,
+                index=False,
+                column_format=r"p{4cm}p{2cm}p{2cm}p{2cm}p{9cm}",
+            )
+        table = table.replace(r"\begin{tabular}", r"\begin{tabular*}{\textwidth}")
+        table = table.replace(r"\end{tabular}", r"\end{tabular*}")
+        f.write(table)

### 2021-06-29 src/parameter_tables/task_convert_csv_to_latex.py
Fill the parameter description tables.

@@ -12,7 +12,7 @@ PARAMETRIZATION = [(p, BLD / "param_tables" / f"{p.stem}.tex") for p in SRC.iter
 @pytask.mark.parametrize("depends_on, produces", PARAMETRIZATION)
 def task_convert_csv_to_tex(depends_on, produces):
     df = pd.read_csv(depends_on)
-    df = df.dropna(axis=0)  # drop empty rows
+    df = df.dropna(axis=0, how="all")  # drop empty rows
     with open(produces, "w") as f:
         with pd.option_context("max_colwidth", 10000):
             table = df.to_latex(

### 2021-06-29 src/paper/appendix_data.tex
Minor improvements.
Start on describing the synthetic population in more detail.

@@
 non-private households vary widely in size and it is very difficult to know which
-contacts take place in such households.
+contacts take place in such living arrangements.
 
@@
 keeping for each individual their age, gender, occupation and whether they work on
-Saturdays and Sundays.
+Saturdays and Sundays. For each household we draw its county and set the corresponding
+federal state.% county is not part of the microcensus campus file
+We draw for young children whether they attend a nursery or not \citep{Destatis2020}.
 
 \begin{itemize}
-    \item \textcolor{red}{Describe microcensus campus file and which data we use and for what}
-    \item \textcolor{red}{Describe size and sampling of households}
-    \item \textcolor{red}{counties are randomly drawn}\comment[id=K]{Find out why!}
-    \item \textcolor{red}{how we create the occupation column}
     \item \textcolor{red}{describe how we create the contact model group ids for school,
@@
         \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+        \caption{Share of Individuals That Have Ever Done a Rapid Test}
     \end{subfigure}%
@@
         \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        %   \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
+        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
     \end{subfigure}

### 2021-06-30 src/parameter_tables/task_convert_csv_to_latex.py
2021 as integer in the overview table.

@@ -21,6 +21,7 @@ def task_convert_csv_to_tex(depends_on, produces):
                 longtable=False,
                 index=False,
                 column_format=r"p{4cm}p{2cm}p{2cm}p{2cm}p{9cm}",
+                float_format="%.0f",
             )
         table = table.replace(r"\begin{tabular}", r"\begin{tabular*}{\textwidth}")
         table = table.replace(r"\end{tabular}", r"\end{tabular*}")

### 2021-06-30 src/paper/model/matching.tex
Restructure for arxiv submission.
Remove all comments by hand for arxiv.
Nicely format names. Add rapid test identifier to rapid test parameters.

--- 2021-06-18 paper/model/matching.tex
@@
 to be translated into actual meetings between people. This is achieved by a matching
-algorithm:
+algorithm:\comment[id=K]{comment from HM: Describe what happens if matching is not
+perfect (i.e. if some people remain)}
 
@@
 to model vacations, weekends or symptomatic people who stay home (see section
-\ref{sec:endogenous_contact_reductions} for details).
+\ref{sec:policies} for details).
 
@@
 of the newly infected person. Finally, we reduce the number of remaining contacts of the
-two matched individuals by one.\comment[id=J]{Add formulae here.}
+two matched individuals by one. \comment[id=J]{Add formulae here.}
 

### 2021-06-30 src/paper/appendix_data.tex
Move synthetic population todos to comment.

@@
 
-\subsection{The Synthetic Population \comment[id=K]{To be written}}
+\subsection{The Synthetic Population}
 
@@
 federal state.% county is not part of the microcensus campus file
-We draw for young children whether they attend a nursery or not \citep{Destatis2020}.
-
+We draw for young children whether they attend a nursery or not
+\citep{Destatis2020}.\comment[id=K]{Missing still:
     \begin{itemize}
-    \item \textcolor{red}{describe how we create the contact model group ids for school,
-    work and other}
-    \item \textcolor{red}{Describe work contact priority, educ contact priority and
-    rapid test compliance parameter and quarantine compliance.}
-    \item \textcolor{red}{vaccination rank creation}
+        \item describe how we create the contact model group ids for school,
+        work and other
+        \item describe work contact priority, educ contact priority and
+        rapid test compliance parameter and quarantine compliance.
+        \item vaccination rank creation
     \end{itemize}
+}
 

### 2021-07-01 src/paper/literature_review.tex
Add comment on Twitter feedback on other agent based models.

@@
 
-Another commonly used model class in epidemiology are agent-based simulation models. In
-these models individuals are simulated as moving particles. Infections take place when
-two particles come closer than a certain contact radius (e.g. \citet{Silva2020} and
-\citet{Cuevas2020}). While the simulation approach makes it easy to incorporate
-heterogeneity in disease progression, it is hard to incorporate heterogeneity in meeting
-patterns. Moreover, policies are modeled as changes in the contact radius or momentum
-equation of the particles. The translation from real policies to corresponding model
-parameters is a hard task.
+Another commonly used model class in epidemiology are agent-based simulation
+models.\comment[id=K]{See
+\href{https://twitter.com/Patrick_M_Econ/status/1410543606939721730?s=20}{this tweet} for
+a reference to some models that are similar to ours} In these models individuals are
+simulated as moving particles. Infections take place when two particles come closer than
+a certain contact radius (e.g. \citet{Silva2020} and \citet{Cuevas2020}). While the
+simulation approach makes it easy to incorporate heterogeneity in disease progression, it
+is hard to incorporate heterogeneity in meeting patterns. Moreover, policies are modeled
+as changes in the contact radius or momentum equation of the particles. The translation
+from real policies to corresponding model parameters is a hard task.
 

### 2021-07-02 src/paper/rapid_test_params.tex
Transfer Roman's nice references.

@@
 offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
-60\% until the first days of April \cite{ZDF2021}.\comment[id=K]{ToDo: Find the survey
-that the ZDF is citing here}
+61\% until the first days of April \citep{BMWI2021, IZA2021}.
 
-Until mid April 70\% of workers were expected to receive a weekly test offer
-\citep{AerzteZeitung2021}. However, according to surveys conducted in mid April
+Until mid April 72\% of workers were expected to receive a weekly test offer
+\citep{BMWI2021, IZA2021}. However, according to surveys conducted in mid April
 \citep{Betsch2021}, less than two thirds of individuals with work contacts receive a test
@@
 by March 1st 30\% of them are tested weekly. The share increases to 90\% for the week
-before Easter. At that time both Bavaria \citep{BayrischerRundfunk2021} and
-Baden-Württemberg \citep{MinisteriumKultus2021} were offering tests to teachers and
-North-Rhine
-Westphalia\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen}}
-\cite{DPA2021} and Lower Saxony \citep{SueddeutscheZeitung2021} were already testing
+before Easter. At that time both Bavaria \citep{STMGP2021} and Baden-Württemberg
+\citep{MinisteriumKultus2021} were offering tests to teachers and North-Rhine Westphalia
+\citep{SchulministeriumNRW2021} and Lower Saxony \citep{NSMK2021} were already testing
 students and tests for students and teachers were already mandatory in Saxony
-\citep{SueddeutscheZeitung2021a}. After Easter we assume that 95\% of teachers get tested
-twice per week.
+\citep{SMK2021}. After Easter we assume that 95\% of teachers get tested twice per week.
 
-Tests for students started
-later\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen}}
-\citep{MinisteriumKultus2021} so we assume that they only start in February and only 10\%
-of students get tested by March 1st. Relying on the same sources as above we approximate
-that by the week before Easter this share had increased to
-40\%.\footnote{\url{https://www.land.nrw/de/pressemitteilung/umfassende-informationen-fuer-die-schulen-zu-corona-selbsttests-fuer-schuelerinnen},
-}
+Tests for students started later \citep{MinisteriumKultus2021, SchulministeriumNRW2021}
+so we assume that they only start in February and only 10\% of students get tested by
+March 1st. Relying on the same sources as above we approximate that by the week before
+Easter this share had increased to 40\% \citep{SchulministeriumNRW2021}.
 
-After Easter the share of students receiving twice weekly tests is set to 75\%. This as
-based on tests becoming mandatory becoming mandatory in Bavaria after Easter
-break\footnote{Bavaria\footnote{\url{https://bit.ly/3nz5fXS}}, in North-Rhine Westphalia
-on April
-12th\footnote{https://www.schulministerium.nrw/ministerium/schulverwaltung/schulmail-archiv/14042021-schulbetrieb-im-wechselunterricht-ab-montag},
-\url{https://bit.ly/2QHilX3}} and on the 19th in
-Baden-Württemberg\footnote{https://bit.ly/3vuetaD, https://bit.ly/3vuetaD}.
+After Easter the share of students receiving twice weekly tests is set to 75\%. This is
+based on tests becoming mandatory in Bavaria \citep{BayerischeStaatskanzlei2021} and
+North Rhine-Westphalia \citep{SchulministeriumNRW2021b} after their Easter breaks and
+on the 19th in Baden-Württemberg \citep{KMBaWue2021}.
 
@@
 tests and rapid tests for lay people started to become available
-\footnote{\url{https://bit.ly/3ehmGcj}, \url{https://bit.ly/3xJCIn8}} and other access to
-rapid tests was very limited.
+\citep{Bundesanzeiger2021a, Bundesregierung2021} and other access to rapid tests was very
+limited.
 
-According to the COSMO study\footnote{\url{https://bit.ly/2QSFAgR}} 63\% would have been
-willing to take a test in the round of 23rd of February 2021 when an acquaintance would
-have tested positive. Since this is only asking for willingness not actual behavior and
-the demand when meeting with friends is very likely lower, we take this as the upper
-bound of private rapid test demand which is reached on May 4th. To cover that many people
-are likely to have sought and done their first rapid test before the Easter holidays to
-meet friends or family, we let the share of individuals doing rapid tests in that time
+According to the COSMO study \citep{Betsch2021a} 63\% would have been willing to take a
+test in the round of 23rd of February 2021 when an acquaintance would have tested
+positive. Since this is only asking for willingness not actual behavior and the demand
+when meeting with friends is very likely lower, we take this as the upper bound of
+private rapid test demand which is reached on May 4th. To cover that many people are
+likely to have sought and done their first rapid test before the Easter holidays to meet
+friends or family, we let the share of individuals doing rapid tests in that time
 increase more rapidly than before and after. By end of March 25\% of individuals would do

### 2021-07-05 src/paper/appendix_data.tex
Replace Smith pre-print with published version.

@@
 federal state.% county is not part of the microcensus campus file
-We draw for young children whether they attend a nursery or not
-\citep{Destatis2020}.\comment[id=K]{Missing still:
+We draw for young children whether they attend a nursery or
+not.\citep{Destatis2020}.\comment[id=K]{Missing still:
     \begin{itemize}

### 2021-07-05 src/paper/model/estimated_parameters.tex
Remove all comments by hand for arxiv.
Put comments back in.
Add comments by Philipp Eisenhauer.
Replace Smith pre-print with published version.

@@
-\subsection{Estimated Parameters}
+\subsection{Estimated Parameters}\comment[id=K]{Philipp suggests to write more about
+identification. Both in general and to take seasonality as an example of how we identify
+both calibrated and estimated parameters. }
 \label{subsec:estimated_params}
@@
 preschools and nurseries. 3) for work contacts. 4) for households. 5) for other
-contacts.\comment[id=J]{Write that the exact numbers are hard to interpret and show
-the infections by channel heatmap more prominently than the infection probabilities}
+contacts.\comment[id=J]{Write that the exact numbers are hard to interpret and show the
+infections by channel heatmap more prominently than the infection
+probabilities}\comment[id=K]{Compare to
+\href{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3867682}{these estimates}}
 

### 2021-07-06 src/figures/results/tables/student_infections_table.tex
New results uploaded without adjusting version.txt and no deleting previous results.

@@
 \midrule
- educ open after easter  without tests &                                             725914 \\
- educ open after easter  with tests    &                                             552894 \\
- close educ after easter               &                                             430594 \\
- baseline                              &                                             439077 \\
+ educ open after easter  without tests &                                             696963 \\
+ educ open after easter  with tests    &                                             545057 \\
+ close educ after easter               &                                             422360 \\
+ baseline                              &                                             432320 \\
 \bottomrule

### 2021-07-07 src/paper/report.tex
Smallish updates.

--- 2021-06-23 src/paper/report.tex
@@
-Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
-sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
-presented an enormous challenge to humanity on many dimensions. The development of
-highly effective vaccines holds the promise of containment in the medium term. However,
-most countries find themselves many months---and often years---away from reaching
-vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
-some more scientific citation before submission} In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain
+Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
+on many dimensions. The development of highly effective vaccines holds the promise of
+containment in the medium term. However, most countries find themselves many
+months---and often years---away from reaching vaccination levels that would end the
+pandemic or even protect the most vulnerable\citep{Swaminathan2021}.\comment[id=HM]{Use
+\url{https://www.nature.com/articles/s41562-021-01160-2} instead} In the meantime, it is
+of utmost importance to employ an effective mix of strategies for containing the virus.
+The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
+to reduce contacts between individuals. While this has allowed some countries to sustain
 equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
@@
 Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
-detected cases and use data on \replaced[id=K]{the share of PCR tests administered to
-people with CoViD-19 symptoms}{whether CoViD-19 symptoms led to a PCR test}. As the share
-of asymptomatic individuals varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age group
-over time in our model). Our estimates suggest that---in the absence of rapid
-testing---the detection rate is 80\% higher on average for individuals above age 80
-compared to school age children. Once rapid test become available, confirmation of a
-positive result is another reason leading to positive PCR tests.
+detected cases and use data on the share of PCR tests administered because of CoViD-19
+symptoms. As the share of asymptomatic individuals varies by age group, this gives us
+age-specific shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share
+of known cases by age group over time in our model). Our estimates suggest that---in the
+absence of rapid testing---the detection rate is 80\% higher on average for individuals
+above age 80 compared to school age children. Once rapid test become available,
+confirmation of a positive result is another reason leading to positive PCR tests.
 
-The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
+The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
 covering the period mid-September 2020 to the end of May 2021.
@@
 November before rising again in December, prompting the most stringent lockdown to this
-date. Schools and daycare centers were closed again, so were customer-facing businesses
-except for grocery and drug stores. From the peak of the second wave just before
-Christmas until the trough in mid-February, newly detected cases decreased by almost
-three quarters. The third wave in the spring of 2021 is associated with the B.1.1.7
-strain, which became dominant in March (Figure~\ref{fig:share_b117}). In early March,
-some NPIs were being relaxed; e.g., hairdressers and home improvement stores were
-allowed to open again to the public. There were many changes in details of regulations
-afterwards, but they did not change the stringency index.
+date. Schools and daycare centers were closed, so were customer-facing businesses except
+for grocery and drug stores. From the peak of the second wave just before Christmas
+until the trough in mid-February, newly detected cases decreased by almost three
+quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 strain,
+which became dominant in March (Figure~\ref{fig:share_b117}). In early March, some NPIs
+were being relaxed; e.g., hairdressers and home improvement stores were allowed to open
+again to the public. There were many changes in details of regulations afterwards, but
+they did not change the overall stringency index.
 
@@
 
-By this time, the set of policy instruments had become much more diverse. Around the
+By March 2021, the set of policy instruments had become much more diverse. Around the
 turn of the year, the first people were vaccinated with a focus on older age groups and
-medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). By the end of May, just
-over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
-started to replace regular PCR tests for staff in many medical and nursing facilities.
-These had to be administered by medical doctors or in pharmacies. At-home tests approved
-by authorities became available in mid-March, rapid test centers were opened, and one
-test per person and week was made available free of charge. In several states, customers
-were only allowed to enter certain stores with a recent negative rapid test result.
-These developments are characteristic of many countries: The initial focus on NPIs to
-slow the spread of the disease has been accompanied by vaccines and a growing acceptance
-and use of rapid tests. At broadly similar points in time, novel strains of the virus
-have started to pose additional challenges.
+medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May,
+43\% had received at least one dose of a vaccine. In late 2020, rapid tests started to
+replace regular PCR tests for staff in many medical and nursing facilities. These had to
+be administered by medical doctors or in pharmacies. At-home tests approved by
+authorities became available in mid-March, rapid test centers were opened, and one test
+per person and week was made available free of charge. In several states, customers were
+only allowed to enter certain stores with a recent negative rapid test result. These
+developments are characteristic of many countries: The initial focus on NPIs to slow the
+spread of the disease has been accompanied by vaccines and a growing acceptance and use
+of rapid tests. At broadly similar points in time, novel strains of the virus have
+started to pose additional challenges.
 
@@
 the three broad contact networks. For example, a value of 0.75 for the work multiplier
-means that if the environment was the same as in September (levels of infection rates, no
-rapid tests or vaccinations, only the wildtype virus present), infections at the
+means that if the environment was the same as in September (levels of infection rates,
+no rapid tests or vaccinations, only the wildtype virus present), infections at the
 workplace would be reduced by 25\%. The lines show the product of the effect of contact
@@
 the most stringent regulations are associated with the period of strong decreases in new
-infections between late December 2020 and mid-February 2021. The reversal of the trend is
-associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
+infections between late December 2020 and mid-February 2021. The reversal of the trend
+is associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
 during May 2021 is associated with at least weekly rapid tests for 42\% of the
-population, a vaccination rate that rose from 28\% to 43\%, and \deleted[id=K]{mostly}
-seasonality \replaced[id=K]{further lowering}{impacting a fall in} the relative
-infectiousness of contacts \deleted[id=K]{outside of work and school.}
+population, a vaccination rate that rose from 28\% to 43\%, and seasonality lowering the
+relative infectiousness of contacts, most strongly outside of work and school.
 
@@
 contribute less to the spread of the disease than others due to a lower number of
-contacts\replaced[id=K]{as they do not work, do not go to school and tend to live in small households}{, see Supplementary Material~\ref{subsec:contacts_by_age}}.
-% We don't draw age specific number of contacts because the sample sizes would be to
-% small.
-By the end of our study period, when first-dose vaccination rates reached around 40\% of
-the population, the numbers of new cases would have started to decline. It is important
-to note that the initial focus of the campaign was to prevent deaths and severe disease.
-Indeed, the case fatality was rate considerably lower during the third wave when compared
-to the second (4.4\% between October and February and 1.4\% between March and the end of
+contacts see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
+study period, when first-dose vaccination rates reached around 40\% of the population,
+the numbers of new cases would have started to decline. It is important to note that the
+initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
+fatality was rate considerably lower during the third wave when compared to the second
+(4.4\% between October and February and 1.4\% between March and the end of
 May).\comment[id=K]{This needs a source.}

### 2021-07-08 src/paper/appendix_data.tex
Add finished section on the synthetic population.

--- 2021-06-30 src/paper/appendix_data.tex
@@
 Saturdays and Sundays. For each household we draw its county and set the corresponding
-federal state.% county is not part of the microcensus campus file
-We draw for young children whether they attend a nursery or not
-\citep{Destatis2020}.\comment[id=K]{Missing still:
-    \begin{itemize}
-        \item describe how we create the contact model group ids for school,
-        work and other
-        \item describe work contact priority, educ contact priority and
-        rapid test compliance parameter and quarantine compliance.
-        \item vaccination rank creation
-    \end{itemize}
-}
+federal state.% we draw the county because county is not part of the campus file
+
+We randomly assign 35\% of children below three to attend a nursery \citep{Destatis2020}.
+For children between three and six years old, we assume all go to preschool (officially
+92.5\% according to \cite{Destatis2020}). Nursery groups consist
+
+Children that attend a nursery meet in groups of four plus one adult care taker every
+weekday when there are no school vacations. Preschool children meet in groups of nine
+with two adult care takers. These groups are mixed with respect to age but all belong to
+the same state and mostly to the same county. For nurseries the group size is four, for
+preschools it's nine \citep{BertelsmannStiftung2019}.
+
+Every child that goes to school is part of three different classes that meet every day on
+weekdays when there are no vacations and no school policies are in place.\footnote{We
+implement vacations on the federal state level.} Each class consists of approximately 23
+students \citep{OECD2013} and two teachers. All students in a class are of the same age
+and live in the same state and if mostly also in the same county. In addition, each
+child gets assigned a value that captures his or her need to attend nursery, preschool or
+school. This allows us to capture various degrees of emergency care that can be granted
+while educational facilities are closed or function in some kind of rotating schedule.
+
+Workers are assigned to a daily meeting work group. The group sizes vary to match the
+number of daily repeating work contacts reported by working individuals in
+\cite{Mossong2008}. These groups only consist of workers that work in the same county.
+For a distribution of the number of daily recurring work contacts see
+Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
+we match each worker with up to 14 other workers to match the number of reported weekly
+work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}. Each pair is
+assigned a weekend on which they always meet in the absence of work policies. 80\% of
+these contacts are individuals from the same county.
+% work contact priority
+In the same way children have a educational priority determining if they are entitled to
+emergency care workers are assigned a work contact priority that captures how necessary
+their work is and to which degree they can work from home. This means that it's always
+the same individuals that continue to have work contacts when work from home mandates of
+a certain strictness are in place. Since our work contact priority is uniformly normal
+from zero to one, we can use the reduction in work mobility as reported by
+\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.
+
+In addition to creating groups for educational facilities and work we also have other
+recurring contacts to represent things like groups of friends or sports teams that
+practice regularly together. Both daily and weekly groups are created analogously to the
+work groups but matching the numbers in Figure~\ref{n_contacts_other_daily_recurrent} and
+Figure~\ref{n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
+highly assortative by age all individuals that have a daily leisure contact are matched
+with a person not only from the same county but also from the same age group.
+
+The individuals in our population can react to events such as developing symptoms that
+are typical of CoViD-19, a positive PCR test or a positive rapid test by reducing their
+contacts. To determine who would reduce their contacts in such a situation or demand a
+rapid test to avoid quarantine we introduce a quarantine compliance parameter. Similarly
+we introduce a rapid test compliance parameter that determines in which order individuals
+start demanding rapid tests when rapid tests become increasingly available. This makes
+sure that when for example only 10\% of workers get tested, it's the same workers that
+have access to tests every week.
+
+Lastly, for the distribution of vaccinations every individual is assigned a vaccination
+group and a vaccination rank from that group that creates a complete vaccination queue
+over the population including 15\% that refuse to be vaccinated \citep{RKI2021}. The
+vaccination groups are created to match the recommendations by the Ständige
+Impfkommission \citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized
+more than recommended by the commission.}% To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
+% put adolescents and children into two groups that follow after the general population.
+% These groups do not become eligible within our simulation frame until June.
 

### 2021-07-09 src/paper/report.tex
Fix abstract and add reference for global vaccination process.

@@
 months---and often years---away from reaching vaccination levels that would end the
-pandemic or even protect the most vulnerable\citep{Swaminathan2021}.\comment[id=HM]{Use
-\url{https://www.nature.com/articles/s41562-021-01160-2} instead} In the meantime, it is
+pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
 of utmost importance to employ an effective mix of strategies for containing the virus.

### 2021-07-09 src/figures/results/tables/student_infections_table.tex
Add new figures to repo. Mostly not in paper yet.

@@
 \midrule
- educ open after easter  without tests &                                             696963 \\
- educ open after easter  with tests    &                                             545057 \\
- close educ after easter               &                                             422360 \\
- baseline                              &                                             432320 \\
+ educ open after easter  without tests &                                            1037016 \\
+ educ open after easter  with tests    &                                             797195 \\
+ close educ after easter               &                                             614035 \\
+ baseline                              &                                             627864 \\
 \bottomrule

### 2021-07-09 src/paper/appendix_results.tex
Add comments by Philipp Eisenhauer.
Add new figures to repo. Mostly not in paper yet.

@@
-\section{Additional Results}
+\section{Additional Results}\comment[id=K]{Philipp suggests to do a sensitivity analysis
+for the seasonality}
 \label{sec:additional_results}
@@
 
+\FloatBarrier
+
 \subsection{Rapid Tests}
 \label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/educ_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
-    \label{fig:educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/work_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test due to Work}
-    \label{fig:work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/private_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
-  \label{fig:private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Rapid Test Demand by Reason}
-  \label{fig:rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends.
-  }
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
-    \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
-    Individuals}
-    \label{fig:share_infected_among_educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
-    \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
-    \label{fig:share_infected_among_work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
-    \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
-  \label{fig:share_infected_among_private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Share of Rapid Tests Demanded by Infected Individuals by Reason}
-  \label{fig:share_infected_among_rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. Private rapid tests have a much higher share of infected
-    individuals because they are mostly triggered by events that make an infection
-    likely. Remember however, that one reason is that a household member has a positive
-    rapid test. This means that work and education rapid tests which have a low rate of
-    infected individuals trigger more targeted rapid tests through the household demand.
-    They also have a much higher volume of tests. As can be seen in the decomposition
-    (Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
-    important for the overall effect. It appears that the combination of wide testing to
-    find infection chains plus targeted tests to break those chains taking together
-    is important.}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_doing_rapid_test_today}
-  \caption{Share of Individuals Doing a Rapid Test per Day}
-  \label{fig:fit_share_doing_rapid_test}
-  \floatfoot{\noindent \textit{Note:} The plateau in the end is the result of falling
-  case numbers that lead to less private rapid tests being demanded due to contact
-  tracing. The two biggest jumps are due to public holidays, i.e. the end of Easter
-  vacations in the first half of April and May 1st. These lead many individuals to not
-  test themselves (even though the threshold to be tested falls during the time) due to
-  vacations from school and work. As a result more individuals than usual are due to be
-  tested because of frequency requirements after public holidays. }
-\end{figure}
 
@@
   just triggered by events make up only a small fraction of tests. }
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
+  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels}
+  \label{fig:rapid_test_demand}
+  \floatfoot{\noindent \textit{Note:}
+    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
+    school) as well as school pupils. After Easter the required frequency of tests is
+    increased from once per week to twice per week. Work rapid tests are demanded by
+    individuals that still have work contacts, i.e. do not work from home. The share of
+    employers offering rapid tests increases over the time frame and the frequency of
+    testing is also increased. Tests are demanded by individuals for one of three private
+    reasons: having developed symptoms without access to a PCR test, having a household
+    member that has tested positive or developed symptoms or having planned weekly
+    meeting with friends. }
 \end{figure}

### 2021-07-09 src/paper/appendix_results.tex
Fix last plots.

@@
   \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-  \caption{Share of Individuals That has Ever Done a Rapid Test}
+  \caption{Share of Individuals Who Have Ever Done a Rapid Test}
   \label{fig:fit_share_ever_rapid_test}
@@
   \label{fig:fit_share_rapid_test_within_last_week}
-  \floatfoot{\noindent \textit{Note:} The two lines are not perfectly compatible. The
-  COSMO study \citep{Betsch2021} only asked if individuals had done at least one
-  self-administered rapid test per week in the last four weeks. This means there are two
-  differences between the simulated line and the empirical data points. Firstly, the
-  simulated line encompasses all rapid tests since we do not differentiate between
-  self-administered and medically administered rapid tests. Secondly, in our model we
-  cannot verify the rhythm over the last four weeks but just observe the last time of a
-  rapid test. However, both work and school rapid tests which make up the largest share
-  of rapid tests (see Figure~\ref{fig:rapid_test_demand_by_reason}) are both required on
-  a weekly or twice weekly basis while the private rapid test demand which are partly
-  just triggered by events make up only a small fraction of tests. }
+  \floatfoot{\noindent \textit{Note:} Our modelling of rapid tests leads to very similar rates of individuals that are tested at least weekly as in the COSMO study \citep{Betsch2021}.
+  In that study, individuals were asked if they performed rapaid tests at least weekly in
+  the last four weeks. In our model we cannot verify the rhythm over the last four
+  weeks but just observe the last time of a rapid test. However, both work and school
+  rapid tests which make up the largest share of rapid tests (see
+  Figure~\ref{fig:rapid_test_demand_by_channel}) are both required on a weekly or twice
+  weekly basis while the private rapid test demand which are partly just triggered by
+  events make up only a small fraction of tests.}
 \end{figure}
@@
   \caption{Share of the Population Demanding a Rapid Test Because of Different Channels}
-  \label{fig:rapid_test_demand}
+  \label{fig:rapid_test_demand_by_channel}
   \floatfoot{\noindent \textit{Note:}
@@
     meeting with friends. }
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_true_positive}
+        \caption{Number of Discovered Cases Due to Rapid Tests by Channel}
+        \label{fig:rapid_tests_number_true_positive}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_false_positive}
+        \caption{Number of False Positive Rapid Tests by Channel}
+        \label{fig:rapid_tests_number_false_positive}
+    \end{subfigure}
+    \vskip3ex
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_true_negative}
+        \caption{Number of True Negative Rapid Tests by Channel}
+        \label{fig:rapid_tests_number_true_negative}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_false_negative}
+        \caption{Number of False Negative Rapid Tests by Channel}
+        \label{fig:rapid_tests_number_false_negative}
+    \end{subfigure}
+    \vskip3ex
+    \caption{Rapid Test Results}
+    \label{fig:rapid_test_results}
+
+    \floatfoot{\noindent \textit{Note:}
+      The number of rapid tests of each category are upscaled to the full German
+      population. \textcolor{red}{To be written.}
+    }
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
+        \caption{Rate of True Positive Rapid Tests by Channel}
+        \label{fig:rapid_tests_true_positive_rate}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_positive_rate}
+        \caption{Rate of False Positive Rapid Tests by Channel}
+        \label{fig:rapid_tests_false_positive_rate}
+    \end{subfigure}
+    \vskip3ex
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
+        \caption{Rate of True Negative Rapid Tests by Channel}
+        \label{fig:rapid_tests_true_negative_rate}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_negative_rate}
+        \caption{Rate of False Negative Rapid Tests by Channel}
+        \label{fig:rapid_tests_false_negative_rate}
+    \end{subfigure}
+    \vskip3ex
+    \caption{Rapid Test Results}
+    \label{fig:rapid_test_results}
+
+    \floatfoot{\noindent \textit{Note:}
+      . \textcolor{red}{To be written.}
+    }
 \end{figure}

### 2021-07-09 src/paper/appendix_results.tex
Refactor the results section. Add random rapid test scenarios and robustness check scenarios (no descriptions yet).

--- 2021-07-09 src/paper/appendix_results.tex
@@
 
-\subsection{Simulated vs. Empirical Data}
-\label{subsec:fit_results}
+\input{paper/results/fit}
 
-This compares simulated data from our model with empirical data from Germany. We look at
-observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
-rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups and federal states.\comment[id=J]{summarize the fit}
+\input{paper/results/share_known_cases}
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
-  \caption{Fit Over the Full Simulation Time Frame with Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported simulated infections rates. The mean infection rate is
-  the thick blue line. Single simulation runs are plotted in lighter and thinner lines.
-  The official case numbers as reported by the Robert-Koch-Institut are plotted in black.
-  The fit is overall very good. The higher the mean incidence and the stronger the growth
-  the more variance there is between simulation runs. We averaged over 30 simulation
-  runs.}
-  \label{fig:aggregated_fit2}
-\end{figure}
+\input{paper/results/rapid_tests_statistics}
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_newly_infected_with_single_runs}
-  \caption{Development of the Total Infections Over the Full Simulation Time Frame with
-  Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the true weekly incidence rates
-  per 100,000 people, including undetected cases. The mean infection rate is the thick
-  blue line. Single simulation runs are plotted in lighter and thinner lines. The higher
-  the mean incidence and the stronger the growth the more variance there is between
-  simulation runs. We averaged over 30 simulation runs.}
-  \label{fig:newly_infected_in_baseline}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Age Group}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different age
-  groups. The age group of individuals above 80 needs to be interpreted with caution
-  because our synthetic population only includes private households, i.e. nursing homes
-  are not represented in our model. They accounted for many cases and deaths in the
-  winter of 2020 and many 80 to 100 year olds live in these facilities. However, the
-  official data does not contain information on whether cases were nursing home
-  inhabitants or not. We averaged over 30 simulation runs.}
-  \label{fig:age_group_fit}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Federal State}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different
-  federal states. We averaged over 30 simulation runs.}
-  \label{fig:state_fit}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
-  \caption{Share of Vaccinated Individuals}
-  \floatfoot{\noindent \textit{Note:} The figure shows the rate of individuals that are
-  vaccinated in our synthetic population versus in the general German population. Note
-  that we excluded the vaccinations that were given to nursing homes, approximately the
-  first percent of the German population that were vaccinated. Overall, our model covers
-  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
-  the population are vaccinated. Our vaccinations work imperfectly but we do not model
-  different vaccines nor do we distinguish between first and second shot.}
-  \label{fig:fit_vaccinations}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
-  \caption{Effective Replication Number $R_t$ in the Model and as Reported by the
-  Robert-Koch-Institute}
-  \floatfoot{\noindent \textit{Note:} The figure shows the effective replication number
-  ($R_t$) as reported by the RKI and as calculated in our model. The $R_t$ gives the
-  average number of new infections caused by one infected individual. The $R_t$ in our
-  model broadly follows the $R_t$ reported by the RKI. Two trends stand out. Firstly, the
-  RKI's $R_t$ drops faster in November. This could be due to a change in the testing
-  policy that focused tests on the elderly when the second wave hit Germany and led to a
-  decline in the overall share of detected cases. The second difference is from mid
-  February to mid March where the RKI's reported $R_t$ increased more rapidly than that
-  in our model. Here the opposite effect can be expected. During this time rapid tests
-  increased strongly leading to more cases being detected. In the short term this leads
-  an $R_t$ estimation that is based on detected cases to overestimate the replication
-  number.}
-  \label{fig:fit_r_effective}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117_with_single_runs}
-  \caption{Share of B.1.1.7 in the Model and as Reported by the Robert-Koch-Institute}
-  \floatfoot{\noindent \textit{Note:} The figure shows the share of B.1.1.7 as
-  reported by the RKI and as calculated in our model. We only introduce a few cases over
-  the cause of January. From then B.1.1.7 takes over endogenously through its increased
-  infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases per 100,000 inhabitants.}
-  \label{fig:fit_share_b117}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Share of Cases that are Detected}
-\label{subsec:appendix_share_known_cases}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group}
-  \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
-  reported as an official case via PCR confirmation. We use the overall share of known
-  cases that was estimated through the case fatality ratio by the
-  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
-  then assume it to be constant as vaccinations of the elderly strongly affect the case
-  fatality rate which the project does not account for. To get from an overall share of
-  detected cases to the share of cases that is detected in each age group we use that
-  asymptomatic cases are much less likely to be detected. As our model covers age
-  specific asymptomatic rates this endogenously leads to group specific share known cases
-  that verify that infections in younger age groups are under-detected. Starting in 2021
-  in addition to the overall numbers of detected cases through symptoms and the share
-  known cases, cases are also detected through confirmation of positive rapid tests. This
-  leads to an increase in the share of known cases for all age groups but in particular
-  for the younger age groups that are covered extensively with rapid tests through the
-  rapid test requirement for participating in school.}
-\end{figure}
-
-It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
-
-\FloatBarrier
-
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-  \caption{Share of Individuals That has Ever Done a Rapid Test}
-  \label{fig:fit_share_ever_rapid_test}
-  \floatfoot{\noindent \textit{Note:} This figure clearly shows that overall our
-  assumptions on rapid tests are conservative. In our model, the share of individuals who
-  have ever done a rapid test lies  consistently 10 percentage points below the share
-  reported in the COSMO study \citep{Betsch2021}. The main reason for this is that we
-  introduce rapid tests only at the start of 2021. However, using the available data to
-  calibrate our rapid test parameters and estimating the remaining parameters to fit the
-  official case numbers we arrive at a shape that is very similar to that implied by the
-  survey results. This together with fitting the share of Germans with weekly rapid tests
-  well (see Figure~\ref{fig:fit_share_rapid_test_within_last_week} makes us confident
-  that our rapid test model is a good -- and especially not over-confident --
-  representation of rapid testing in Germany.}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-  \caption{Share of Individuals Who Have Done a Rapid Test in the Last Week}
-  \label{fig:fit_share_rapid_test_within_last_week}
-  \floatfoot{\noindent \textit{Note:} The two lines are not perfectly compatible. The
-  COSMO study \citep{Betsch2021} only asked if individuals had done at least one
-  self-administered rapid test per week in the last four weeks. This means there are two
-  differences between the simulated line and the empirical data points. Firstly, the
-  simulated line encompasses all rapid tests since we do not differentiate between
-  self-administered and medically administered rapid tests. Secondly, in our model we
-  cannot verify the rhythm over the last four weeks but just observe the last time of a
-  rapid test. However, both work and school rapid tests which make up the largest share
-  of rapid tests (see Figure~\ref{fig:rapid_test_demand_by_reason}) are both required on
-  a weekly or twice weekly basis while the private rapid test demand which are partly
-  just triggered by events make up only a small fraction of tests. }
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
-  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels}
-  \label{fig:rapid_test_demand}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. }
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Scenarios}
-\label{subsec:appendix_scenarios}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_new_known_case}
-    \caption{Reported Cases}
-    \label{fig:work_scenarios_new_known_case}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
-    \caption{Total Cases}
-    \label{fig:work_scenarios_newly_infected}
-  \end{subfigure}
-  \caption{The Effect of Different Work Scenarios on Reported and Total Cases}
-  \label{fig:work_scenarios_detailed}
-  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
-  policy changes took place at Easter until the end of our simulation period (end of
-  May). We vary the share of workers that work from home and how many tests are performed
-  at work relative to our baseline scenario. Making it mandatory to test all employees
-  that do not work from home markedly reduces cases -- even when only assuming 95\%
-  compliance on both the employer and the employee side. As before, the observed cases
-  can be misleading because more testing leads to more detected cases. It takes two to
-  three weeks for the reduction in new infections effect to dominate the increased
-  detection effect. Furthermore, the two opposing effects lead to a smaller effect size
-  than is actually the case.}
-\end{figure}
-
- \comment[id=K]{Regarding figure \ref{fig:work_scenarios_detailed}: Look up numbers and
- add them to the description.}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
-    \caption{Reported Cases}
-    \label{fig:school_scenarios_new_known_case}
-  \end{subfigure}%
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-    \caption{Total Cases}
-    \label{fig:school_scenarios_newly_infected}
-  \end{subfigure}
-  \caption{The Effect of Different School Scenarios on Reported and Total Cases}
-  \label{fig:school_scenarios_detailed}
-  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
-   policy changes took place at Easter until the end of our simulation period (end of
-   May). Apart from the enacted school policies as our baseline we simulate how cases
-   would have developed if schools had been closed completely as the strictest possible
-   counterfactual scenario and two opening models: One where schools open normally (with
-   hygiene measures) without any testing in the education sector and one where schools
-   open normally but testing shares develop as in the baseline scenario. Our simulations
-   suggest that the enacted policies were as effective as keeping schools closed. Opening
-   schools with the testing schemes that were in place after Easter would have had a
-   small effect on the overall incidence. However, this is mainly due to the stringent
-   testing that was in place in schools by that time. Had schools opened without testing
-   requirements the total incidence would have been up to 50 points higher, though this
-   would have been less visible in the reported cases.}
-\end{figure}
-
- \comment[id=K]{Regarding figure \ref{fig:school_scenarios_detailed}: Look up numbers and
- add them to the description.}
-
-\FloatBarrier
+\input{paper/results/scenarios}

### 2021-07-09 src/paper/results/fit.tex
Refactor the results section. Add random rapid test scenarios and robustness check scenarios (no descriptions yet).

--- 2021-06-18 paper/appendix_results.tex
@@
-\section{Additional Results}
-\label{sec:additional_results}
-
 \subsection{Simulated vs. Empirical Data}
@@
 \FloatBarrier
-
-\subsection{Share of Cases that are Detected}
-\label{subsec:appendix_share_known_cases}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group}
-  \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
-  reported as an official case via PCR confirmation. We use the overall share of known
-  cases that was estimated through the case fatality ratio by the
-  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
-  then assume it to be constant as vaccinations of the elderly strongly affect the case
-  fatality rate which the project does not account for. To get from an overall share of
-  detected cases to the share of cases that is detected in each age group we use that
-  asymptomatic cases are much less likely to be detected. As our model covers age
-  specific asymptomatic rates this endogenously leads to group specific share known cases
-  that verify that infections in younger age groups are under-detected. Starting in 2021
-  in addition to the overall numbers of detected cases through symptoms and the share
-  known cases, cases are also detected through confirmation of positive rapid tests. This
-  leads to an increase in the share of known cases for all age groups but in particular
-  for the younger age groups that are covered extensively with rapid tests through the
-  rapid test requirement for participating in school.}
-\end{figure}
-
-It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
-
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/educ_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
-    \label{fig:educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/work_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test due to Work}
-    \label{fig:work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/private_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
-  \label{fig:private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Rapid Test Demand by Reason}
-  \label{fig:rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends.
-  }
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
-    \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
-    Individuals}
-    \label{fig:share_infected_among_educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
-    \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
-    \label{fig:share_infected_among_work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
-    \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
-  \label{fig:share_infected_among_private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Share of Rapid Tests Demanded by Infected Individuals by Reason}
-  \label{fig:share_infected_among_rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. Private rapid tests have a much higher share of infected
-    individuals because they are mostly triggered by events that make an infection
-    likely. Remember however, that one reason is that a household member has a positive
-    rapid test. This means that work and education rapid tests which have a low rate of
-    infected individuals trigger more targeted rapid tests through the household demand.
-    They also have a much higher volume of tests. As can be seen in the decomposition
-    (Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
-    important for the overall effect. It appears that the combination of wide testing to
-    find infection chains plus targeted tests to break those chains taking together
-    is important.}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Scenarios}
-\label{subsec:appendix_scenarios}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
-  \end{subfigure}
-  \caption{The Effect of Policies on Observed and Unobserved Cases}
-  \label{fig:explain_decline}
-  \floatfoot{\noindent \textit{Note:} \ldots}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-  \end{subfigure}
-  \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-  \label{fig:school_scenarios_detailed}
-\end{figure}
-
-\FloatBarrier
-
-\input{../figures/results/tables/student_infections_table}
-
-\FloatBarrier
-
-\subsection{Shapley Value} % (fold)
-\label{sub:shapley_value}
-
-We decompose the effects of different NPIs and seasonality on the infection rates with
-Shapley values. Shapley values \citep{Shapley2016} are a concept in game theory to
-divide payoffs between a coalition of players. It allows to assign a single value to the
-contribution of an NPI or seasonality which takes into account substitutional and
-complementary effects with other factors.
-
-More formally, define a coalitional game with $N$ players and a super-additive function
-$\nu$ which maps subsets of $N$ to the real numbers. The function $\nu$ is also called
-the characteristic function and assigns a value to a coalition. Then, the Shapley value
-$\phi$ for player $i$ is
-
-\begin{align*}
-    \phi_i(\nu) = \frac{1}{|N| !} \sum_{S \subseteq N \setminus \{i\}} |S| ! (|N| - |S| - 1)! (\nu(S \cup \{i\}) - \nu(S))
-\end{align*}
-
-The last term $(\nu(S \cup \{i\}) - \nu(S))$ is the marginal contribution of player $i$
-minus the coalition without player $i$. Then, compute the sum of marginal contributions
-over all subsets $S$ of $N$ which do not include player $i$. Each marginal contribution
-has to be multiplied by all combinations of other players in $S$ which precede $i$ and
-all possible combinations of remaining players which follow player $i$ in the coalition.
-To arrive at the Shapley value for player $i$, divide the sum by the total number of
-combinations.
-
-The Shapley value has some properties.
-
-\begin{description}
-  \item[Efficiency] The sum of Shapley values is equal to the value of a coalition
-  formed by all players.
-  \item[Symmetry] The Shapley does not depend on the label of a player but only on its
-  position in the characteristic function.
-  \item[Linearity] The Shapley value depends linearly on the values from the
-  characteristic function $\nu$.
-  \item[Dummy Axiom] The Shapley value of a player who contributes nothing to any
-  coalition is 0.
-\end{description}
-
-To produce Figure~\ref{fig:2021_scenarios_decomposition} and
-Figure~\ref{fig:2021_scenarios_decomposition_tests}, we calculate the Shapley values of
-each factor in the comparison on the cumulative number of saved infections between the
-main scenario and the scenario without any of the factors for every day. Then, we divide
-up the saved infections on a particular day according to the Shapley values for the same
-day which yields the daily saved infections for each factor.
-
-% subsection shapley_value (end)

### 2021-07-09 src/paper/results/rapid_tests_statistics.tex
Refactor the results section. Add random rapid test scenarios and robustness check scenarios (no descriptions yet).

\subsection{Rapid Tests}
\label{subsec:appendix_rapid_tests}

\begin{figure}[ht] % Ever Rapid Test
  \centering
  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
  \caption{Share of Individuals Who Have Ever Done a Rapid Test}
  \label{fig:fit_share_ever_rapid_test}
  \floatfoot{\noindent \textit{Note:} This figure clearly shows that overall our
  assumptions on rapid tests are conservative. In our model, the share of individuals who
  have ever done a rapid test lies  consistently 10 percentage points below the share
  reported in the COSMO study \citep{Betsch2021}. The main reason for this is that we
  introduce rapid tests only at the start of 2021. However, using the available data to
  calibrate our rapid test parameters and estimating the remaining parameters to fit the
  official case numbers we arrive at a shape that is very similar to that implied by the
  survey results. This together with fitting the share of Germans with weekly rapid tests
  well (see Figure~\ref{fig:fit_share_rapid_test_within_last_week} makes us confident
  that our rapid test model is a good -- and especially not over-confident --
  representation of rapid testing in Germany.}
\end{figure}

\begin{figure}[ht] % Tested in the last week
  \centering
  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
  \caption{Share of Individuals Who Have Done a Rapid Test in the Last Week}
  \label{fig:fit_share_rapid_test_within_last_week}
  \floatfoot{\noindent \textit{Note:} Our modelling of rapid tests leads to very similar rates of individuals that are tested at least weekly as in the COSMO study \citep{Betsch2021}.
  In that study, individuals were asked if they performed rapaid tests at least weekly in
  the last four weeks. In our model we cannot verify the rhythm over the last four
  weeks but just observe the last time of a rapid test. However, both work and school
  rapid tests which make up the largest share of rapid tests (see
  Figure~\ref{fig:rapid_test_demand_by_channel}) are both required on a weekly or twice
  weekly basis while the private rapid test demand which are partly just triggered by
  events make up only a small fraction of tests.}
\end{figure}

\begin{figure}[ht] % Share tested per day
  \centering
  \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels}
  \label{fig:rapid_test_demand_by_channel}
  \floatfoot{\noindent \textit{Note:}
    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
    school) as well as school pupils. After Easter the required frequency of tests is
    increased from once per week to twice per week. Work rapid tests are demanded by
    individuals that still have work contacts, i.e. do not work from home. The share of
    employers offering rapid tests increases over the time frame and the frequency of
    testing is also increased. Tests are demanded by individuals for one of three private
    reasons: having developed symptoms without access to a PCR test, having a household
    member that has tested positive or developed symptoms or having planned weekly
    meeting with friends. }
\end{figure}

\begin{figure}   % Number of True Positive / False Positive / True Negative / False Negative
    \centering
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_true_positive}
        \caption{Number of Discovered Cases Due to Rapid Tests by Channel}
        \label{fig:rapid_tests_number_true_positive}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_false_positive}
        \caption{Number of False Positive Rapid Tests by Channel}
        \label{fig:rapid_tests_number_false_positive}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_true_negative}
        \caption{Number of True Negative Rapid Tests by Channel}
        \label{fig:rapid_tests_number_true_negative}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/number_false_negative}
        \caption{Number of False Negative Rapid Tests by Channel}
        \label{fig:rapid_tests_number_false_negative}
    \end{subfigure}
    \vskip3ex
    \caption{Rapid Test Results}
    \label{fig:rapid_test_results_numbers}

    \floatfoot{\noindent \textit{Note:}
      The number of rapid tests of each category are upscaled to the full German
      population. \textcolor{red}{To be written.}
    }
\end{figure}

\begin{figure} % True Positive / False Positive / True Negative / False Negative Rate
    \centering
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
        \caption{Rate of True Positive Rapid Tests by Channel}
        \label{fig:rapid_tests_true_positive_rate}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_positive_rate}
        \caption{Rate of False Positive Rapid Tests by Channel}
        \label{fig:rapid_tests_false_positive_rate}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
        \caption{Rate of True Negative Rapid Tests by Channel}
        \label{fig:rapid_tests_true_negative_rate}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_negative_rate}
        \caption{Rate of False Negative Rapid Tests by Channel}
        \label{fig:rapid_tests_false_negative_rate}
    \end{subfigure}
    \vskip3ex
    \caption{Rapid Test Rates by Channel}
    \label{fig:rapid_test_results_rates}

    \floatfoot{\noindent \textit{Note:}
      \textcolor{red}{To be written.}
    }
\end{figure}

\begin{figure} % Share of Tests That Are True Positive / False Positive / True Negative / False Negative
    \centering
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
        \caption{Share of Tests That Are True Positive by Channel}
        \label{fig:rapid_tests_true_positive_rate}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_positive_rate}
        \caption{Share of Tests That Are False Positive by Channel}
        \label{fig:rapid_tests_false_positive_rate}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
        \caption{Share of Tests That Are True Negative by Channel}
        \label{fig:rapid_tests_true_negative_rate}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_negative_rate}
        \caption{Share of Tests That Are False Negative by Channel}
        \label{fig:rapid_tests_false_negative_rate}
    \end{subfigure}
    \vskip3ex
    \caption{Share of Rapid Tests by Outcome and True Infection Status}
    \label{fig:rapid_test_results_test_shares}

    \floatfoot{\noindent \textit{Note:}
      \textcolor{red}{To be written.}
    }
\end{figure}

\FloatBarrier

### 2021-07-09 src/paper/results/scenarios.tex
Refactor the results section. Add random rapid test scenarios and robustness check scenarios (no descriptions yet).

\subsection{Scenarios}
\label{subsec:appendix_scenarios}

\begin{figure}[ht] % Work Scenarios
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_new_known_case}
    \caption{Reported Cases}
    \label{fig:work_scenarios_new_known_case}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
    \caption{Total Cases}
    \label{fig:work_scenarios_newly_infected}
  \end{subfigure}
  \caption{The Effect of Different Work Scenarios on Reported and Total Cases}
  \label{fig:work_scenarios_detailed}
  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
  policy changes took place at Easter until the end of our simulation period (end of
  May). We vary the share of workers that work from home and how many tests are performed
  at work relative to our baseline scenario. Making it mandatory to test all employees
  that do not work from home markedly reduces cases -- even when only assuming 95\%
  compliance on both the employer and the employee side. As before, the observed cases
  can be misleading because more testing leads to more detected cases. It takes two to
  three weeks for the reduction in new infections effect to dominate the increased
  detection effect. Furthermore, the two opposing effects lead to a smaller effect size
  than is actually the case.}
\end{figure}

 \comment[id=K]{Regarding figure \ref{fig:work_scenarios_detailed}: Look up numbers and
 add them to the description.}

\begin{figure}[ht] % School Scenarios
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
    \caption{Reported Cases}
    \label{fig:school_scenarios_new_known_case}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
    \caption{Total Cases}
    \label{fig:school_scenarios_newly_infected}
  \end{subfigure}
  \caption{The Effect of Different School Scenarios on Reported and Total Cases}
  \label{fig:school_scenarios_detailed}
  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
   policy changes took place at Easter until the end of our simulation period (end of
   May). Apart from the enacted school policies as our baseline we simulate how cases
   would have developed if schools had been closed completely as the strictest possible
   counterfactual scenario and two opening models: One where schools open normally (with
   hygiene measures) without any testing in the education sector and one where schools
   open normally but testing shares develop as in the baseline scenario. Our simulations
   suggest that the enacted policies were as effective as keeping schools closed. Opening
   schools with the testing schemes that were in place after Easter would have had a
   small effect on the overall incidence. However, this is mainly due to the stringent
   testing that was in place in schools by that time. Had schools opened without testing
   requirements the total incidence would have been up to 50 points higher, though this
   would have been less visible in the reported cases.}
\end{figure}

 \comment[id=K]{Regarding figure \ref{fig:school_scenarios_detailed}: Look up numbers and
 add them to the description.}

\begin{figure}[ht] % Robustness Check
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_new_known_case}
    \caption{Reported Cases}
    \label{fig:robustness_check_new_known_case}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_newly_infected}
    \caption{Total Cases}
    \label{fig:robustness_check_newly_infected}
  \end{subfigure}
  \caption{Robustness Check}
  \label{fig:robustness_check_detailed}
  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
\end{figure}

 \comment[id=K]{@Janos: Please add caption and description for the robustness check }

\begin{figure}[ht] % Random Rapid Tests
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/random_rapid_tests_vs_baseline/full_new_known_case}
    \caption{Reported Cases}
    \label{fig:random_rapid_tests_new_known_case}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/random_rapid_tests_vs_baseline/full_newly_infected}
    \caption{Total Cases}
    \label{fig:random_rapid_tests_newly_infected}
  \end{subfigure}
  \caption{The Role of Targeted and Compliance Driven Rapid Test Demand}
  \label{fig:random_rapid_tests_detailed}
  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
\end{figure}

\FloatBarrier

### 2021-07-09 src/paper/results/share_known_cases.tex
Refactor the results section. Add random rapid test scenarios and robustness check scenarios (no descriptions yet).

--- 2021-06-18 paper/appendix_results.tex
@@
-\section{Additional Results}
-\label{sec:additional_results}
-
-\subsection{Simulated vs. Empirical Data}
-\label{subsec:new_cases_fit}
-
-This compares simulated data from our model with empirical data from Germany. We look at
-observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
-rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups.\comment[id=J]{summarize the
-fit}\comment[id=HM]{Also need to show simulated total infections somewhere. So far only
-ever shown for 2021.}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
-  \caption{Simulated and Empirical Infections}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates.}
-  \label{fig:aggregated_fit2}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Age Group}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates for different age groups.}
-  \label{fig:age_group_fit}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
-  \caption{Simulated and Empirical Infections by Federal State}
-  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the
-  reported versus the simulated infections rates for different federal states.}
-  \label{fig:state_fit}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Share of Cases that are Detected  \comment[id=K]{Figure notes missing}}
+\subsection{Share of Cases that are Detected}
 \label{subsec:appendix_share_known_cases}
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group in the Main Prediction}
+  \includegraphics[width=\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
+  \caption{Share of Detected Cases by Age Group}
   \label{fig:share_known_cases_by_age_group}
-  \floatfoot{\noindent}
+  \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
+  reported as an official case via PCR confirmation. We use the overall share of known
+  cases that was estimated through the case fatality ratio by the
+  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
+  then assume it to be constant as vaccinations of the elderly strongly affect the case
+  fatality rate which the project does not account for. To get from an overall share of
+  detected cases to the share of cases that is detected in each age group we use that
+  asymptomatic cases are much less likely to be detected. As our model covers age
+  specific asymptomatic rates this endogenously leads to group specific share known cases
+  that verify that infections in younger age groups are under-detected. Starting in 2021
+  in addition to the overall numbers of detected cases through symptoms and the share
+  known cases, cases are also detected through confirmation of positive rapid tests. This
+  leads to an increase in the share of known cases for all age groups but in particular
+  for the younger age groups that are covered extensively with rapid tests through the
+  rapid test requirement for participating in school.}
 \end{figure}
@@
 
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/educ_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test in an Education Setting}
-    \label{fig:educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/work_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test due to Work}
-    \label{fig:work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/private_demand_share}
-    \caption{Share of the Population Demanding a Rapid Test for Private Reasons}
-  \label{fig:private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Rapid Test Demand by Reason}
-  \label{fig:rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends.
-  }
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_educ_demand}
-    \caption{Share of Rapid Tests in the Educational Setting Demanded by Infected
-    Individuals}
-    \label{fig:share_infected_among_educ_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_work_demand}
-    \caption{Share of Work Rapid Tests Demanded by Infected Individuals}
-    \label{fig:share_infected_among_work_rapid_test_demand}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.425\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_private_demand}
-    \caption{Share of Private Rapid Tests Demanded by Infected Individuals}
-  \label{fig:share_infected_among_private_rapid_test_demand}
-  \end{subfigure}
-  \caption{Share of Rapid Tests Demanded by Infected Individuals by Reason}
-  \label{fig:share_infected_among_rapid_test_demand_by_reason}
-  \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. Private rapid tests have a very high share of infected
-    individuals because they are mostly triggered by events that make infection likely.
-    Remember however, that one reason is that a household member has a positive rapid
-    test. This means that work and education rapid tests which have a low rate of
-    infected individuals trigger more targeted rapid tests. They also have a much higher
-    volume of tests. As can be seen in the decomposition (see
-    Figure~\ref{fig:2021_scenarios_decomposition_tests} every category of tests is
-    important for the overall effect.}
-\end{figure}
-
 \FloatBarrier
-
-\subsection{Scenarios}
-\label{subsec:appendix_scenarios}
-
-This is the results section.\comment[id=K]{The results can be found in `figures/results`.
-This includes both figures and tables for lookup of numbers and summary tables.}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
-  \end{subfigure}
-  \caption{The Effect of Policies on Observed and Unobserved Cases}
-  \label{fig:explain_decline}
-  \figurenotes{\ldots}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
-  \end{subfigure}%
-  \begin{subfigure}{.6\textwidth}
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-  \end{subfigure}
-  \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
-  \label{fig:school_scenarios_detailed}
-\end{figure}
-
-\FloatBarrier
-
-\input{../figures/results/tables/student_infections_table}
-
-\FloatBarrier
-
-\subsection{Shapley Value} % (fold)
-\label{sub:shapley_value}
-
-We decompose the effects of different NPIs and seasonality on the infection rates with
-Shapley values. Shapley values \citep{Shapley2016} are a concept in game theory to
-divide payoffs between a coalition of players. It allows to assign a single value to the
-contribution of an NPI or seasonality which takes into account substitutional and
-complementary effects with other factors.
-
-More formally, define a coalitional game with $N$ players and a super-additive function
-$\nu$ which maps subsets of $N$ to the real numbers. The function $\nu$ is also called
-the characteristic function and assigns a value to a coalition. Then, the Shapley value
-$\phi$ for player $i$ is
-
-\begin{align*}
-    \phi_i(\nu) = \frac{1}{|N| !} \sum_{S \subseteq N \setminus \{i\}} |S| ! (|N| - |S| - 1)! (\nu(S \cup \{i\}) - \nu(S))
-\end{align*}
-
-The last term $(\nu(S \cup \{i\}) - \nu(S))$ is the marginal contribution of player $i$
-minus the coalition without player $i$. Then, compute the sum of marginal contributions
-over all subsets $S$ of $N$ which do not include player $i$. Each marginal contribution
-has to be multiplied by all combinations of other players in $S$ which precede $i$ and
-all possible combinations of remaining players which follow player $i$ in the coalition.
-To arrive at the Shapley value for player $i$, divide the sum by the total number of
-combinations.
-
-The Shapley value has some properties.
-
-\begin{description}
-  \item[Efficiency] The sum of Shapley values is equal to the value of a coalition
-  formed by all players.
-  \item[Symmetry] The Shapley does not depend on the label of a player but only on its
-  position in the characteristic function.
-  \item[Linearity] The Shapley value depends linearly on the values from the
-  characteristic function $\nu$.
-  \item[Dummy Axiom] The Shapley value of a player who contributes nothing to any
-  coalition is 0.
-\end{description}
-
-To produce Figure~\ref{fig:2021_scenarios_decomposition} and
-Figure~\ref{fig:2021_scenarios_decomposition_tests}, we calculate the Shapley values of
-each factor in the comparison on the cumulative number of saved infections between the
-main scenario and the scenario without any of the factors for every day. Then, we divide
-up the saved infections on a particular day according to the Shapley values for the same
-day which yields the daily saved infections for each factor.
-
-% subsection shapley_value (end)

### 2021-07-09 src/paper/results/fit.tex
Add plot of vaccinations by age group.

@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]   % new known case with single runs
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]  % newly infected with single runs
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]  % new known case by age group
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]   % new known case by federal state
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]   % fit of vaccinations
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]   % vaccinations by age group
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
+  \caption{Simulated and Empirical Infections by Federal State}
+  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
+  100,000 people for the reported versus the simulated infections rates for different
+  federal states. We averaged over 30 simulation runs.}
+  \label{fig:state_fit}
+\end{figure}
+
+\begin{figure}[ht]   % R effective
   \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht]   % Share B.1.1.7
   \centering

### 2021-07-09 src/paper/results/rapid_tests_statistics.tex
Fix copy paste errors.

@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_true_positive}
         \caption{Share of Tests That Are True Positive by Channel}
-        \label{fig:rapid_tests_true_positive_rate}
+        \label{fig:rapid_tests_testshare_true_positive}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_positive_rate}
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_false_positive}
         \caption{Share of Tests That Are False Positive by Channel}
-        \label{fig:rapid_tests_false_positive_rate}
+        \label{fig:rapid_tests_testshare_false_positive}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_true_negative}
         \caption{Share of Tests That Are True Negative by Channel}
-        \label{fig:rapid_tests_true_negative_rate}
+        \label{fig:rapid_tests_testshare_true_negative}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/false_negative_rate}
+        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_false_negative}
         \caption{Share of Tests That Are False Negative by Channel}
-        \label{fig:rapid_tests_false_negative_rate}
+        \label{fig:rapid_tests_testshare_false_negative}
     \end{subfigure}

### 2021-07-09 src/paper/appendix_data.tex
Add comments from HM and implement some fast suggestions.

@@
 
+    \vskip3ex
     \caption{Number of Contacts of the Other Contact Type}
@@
 within the same county, we can calibrate the assortative mixing by age from the same data
-we use to calibrate the number of contacts.\comment[id=HM]{Redo
-\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of contacts
-or better add a similar figure showing total number of contacts by age in all networks}
+we use to calibrate the number of contacts.
 
@@
 
+\FloatBarrier
+
 \subsection{Policies and Seasonality}
 \label{subsec:policies_seasonality}
-
-\FloatBarrier
 

### 2021-07-09 src/paper/results/fit.tex
Fix copy paste errors.
Add comments from HM and implement some fast suggestions.

@@
 
+ \comment[id=K]{More text in this subsection. For \ref{fig:age_group_fit}: Explain why we
+ don't fit the 80 to 100 year olds}
+
 \begin{figure}[ht]  % new known case by age group
@@
 
+ \comment[id=K]{Explain differences between simulated and actual cases in figure
+ \ref{fig:state_fit}. (Our states are uniform with respect to their age distribution.}
+
 \begin{figure}[ht]   % new known case by federal state
   \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
+  \includegraphics[height=0.95\textheight]{figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
   \caption{Simulated and Empirical Infections by Federal State}
@@
   \caption{Simulated and Empirical Infections by Federal State}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different
-  federal states. We averaged over 30 simulation runs.}
+  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
   \label{fig:state_fit}

### 2021-07-09 src/paper/results/rapid_tests_statistics.tex
Add comments from HM and implement some fast suggestions.

@@
   \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
-  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels}
+  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels
+  on a Given Day}
   \label{fig:rapid_test_demand_by_channel}
@@
     \centering
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
-        \caption{Rate of True Positive Rapid Tests by Channel}
-        \label{fig:rapid_tests_true_positive_rate}
-    \end{subfigure}
-    \hfill
+    % \begin{subfigure}[b]{0.425\textwidth}
+    %     \centering
+    %     \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
+    %     \caption{Rate of True Positive Rapid Tests by Channel}
+    %     \label{fig:rapid_tests_true_positive_rate}
+    % \end{subfigure}
+    % \hfill
     \begin{subfigure}[b]{0.425\textwidth}
@@
     \end{subfigure}
-    \vskip3ex
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
-        \caption{Rate of True Negative Rapid Tests by Channel}
-        \label{fig:rapid_tests_true_negative_rate}
-    \end{subfigure}
+    % \vskip3ex
+    % \begin{subfigure}[b]{0.425\textwidth}
+    %     \centering
+    %     \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
+    %     \caption{Rate of True Negative Rapid Tests by Channel}
+    %     \label{fig:rapid_tests_true_negative_rate}
+    % \end{subfigure}
     \hfill

### 2021-07-09 src/paper/results/scenarios.tex
Add comments from HM and implement some fast suggestions.

@@
 
- \comment[id=K]{@Janos: Please add caption and description for the robustness check }
+ \comment[id=K]{@Janos: Please add caption and description for the robustness check.}
+
+ \comment[id=K]{HM: Why are the baseline cases above the three other scenarios?
+
+  \begin{itemize}
+    \item there are \textbf{less} vaccinations in the robustness check
+    \item the "true" work multiplier is on average approx. equal to the value that we set
+  \end{itemize}
+
+  How do the numbers of tests compare?
+
+ }
 
@@
 
+ \commend[id=K]{Describe that random tests are more effective at detecting cases and that
+ explains the bump in the random test scenario detection in the beginning. However, cases
+ fall immediately more in the non-random scenario because the targeted demand for tests
+ by household members leads to a much more efficient interruption of infection chains.}
+
 \FloatBarrier

### 2021-07-09 src/paper/appendix_data.tex
Add model valiadation section. Restructure results.

@@
 Impfkommission \citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized
-more than recommended by the commission.}% To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
+more than recommended by the commission.}\comment[id=K]{Write how that plays out in our simulations for the different age groups.}
+
+% To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
 % put adolescents and children into two groups that follow after the general population.
 % These groups do not become eligible within our simulation frame until June.
+
+% \begin{figure}[ht]   % vaccinations by age group
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
+%   \caption{Simulated and Empirical Infections by Federal State}
+%   \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
+%   \label{fig:vaccinations_by_age_group}
+% \end{figure}
 
@@
 \subsection{Number of Contacts}
-\label{sub:number_of_contacts}
+\label{subsec:data_number_of_contacts}
 
@@
 
-In our empirical application we distinguish four groups of contact types: households,
-education, work and other contacts.
-% households
+In our empirical application\comment[id=K]{Improve language} we distinguish four groups
+of contact types: households, education, work and other contacts.% households
 For households we assume that the individuals'
@@
 
-% work models
-For our work models\footnote{We distinguish non-recurrent work contacts, daily work
-    contacts and weekly work contacts.} we use the reductions in work mobility reported in
-the Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
-work contacts are not random but governed through a work contact priority where the
-policy changes the threshold below which workers stay home. Figure
-\ref{fig:work_multiplier} shows the share of workers that go to work in our model over
-time.
+% work
+For our work contacts we use the reductions in work mobility reported in the Google
+Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in work
+contacts are not random but governed through a work contact priority where the policy
+changes the threshold below which workers stay home. Figure \ref{fig:work_multiplier}
+shows the share of workers that go to work in our model over time.
 
@@
 All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
-of each line.}
+seen in Figure~\ref{fig:rapid_test_demand}.
 
@@
 \end{figure}
+
+ \comment[id=K]{J: Explain that we don't fit the share ever tested that well because our rapid test compliance is completely fixed. }

### 2021-07-09 src/paper/literature_review.tex
Add model valiadation section. Restructure results.

@@
 We build on two strands of literature: Recent extensions of the epidemiological SEIR
-model and agent-based simulation models.\comment[id=K]{Write a sentence about the
-\href{https://www.itwm.fraunhofer.de/de/presse-publikationen/presseinformationen/2021/2021-06-22_Dritte_Welle_Starker-Effekt-von-Schnelltests-an-Schulen.html}{Frauenhofer
-paper} coming to similar conclusions: }
+model and agent-based simulation models.
 

### 2021-07-09 src/paper/model/number_of_contacts.tex
Add model valiadation section. Restructure results.

--- 2021-06-21 src/paper/model/number_of_contacts.tex
@@
 types of contacts:\comment[id=HM]{I started updating this to understand it better myself.
-In the end, I think it will be useful to group into HH / work / schools and add an introductory sentence for each group.
-E.g., that work is for everybody who is working --- except there is a different model for teachers (?), ...}
+In the end, I think it will be useful to group into HH / work / schools and add an
+introductory sentence for each group. E.g., that work is for everybody who is working ---
+except there is a different model for teachers (?), ...}
 
@@
 and at home is calibrated with data provided by \citet{Mossong2008}. For details see
-Section~\ref{sec:number_of_contacts}. In particular, we sample the number of contacts or
-group sizes from empirical distributions that sometimes depend on age. It would also be
-possible to use economic or other behavioral models to predict the number of contacts.
+Section~\ref{subsec:data_number_of_contacts}. In particular, we sample the number of
+contacts or group sizes from empirical distributions that sometimes depend on age. It
+would also be possible to use economic or other behavioral models to predict the number
+of contacts.
 
 % Theoretically, each contact type can have its own infection probability. However, to
-reduce the number of free parameters and thus avoid a potential over-fitting we only
-estimate different infection probabilities for the areas work, school, preschool and
-nurseries, households and other contacts.
+% reduce the number of free parameters and thus avoid a potential over-fitting we only
+% estimate different infection probabilities for the areas work, school, preschool and
+% nurseries, households and other contacts.

### 2021-07-09 src/paper/rapid_test_params.tex
Add model valiadation section. Restructure results.

@@
 when meeting with friends is very likely lower, we take this as the upper bound of
-private rapid test demand which is reached on May 4th. To cover that many people are
-likely to have sought and done their first rapid test before the Easter holidays to meet
-friends or family, we let the share of individuals doing rapid tests in that time
-increase more rapidly than before and after. By end of March 25\% of individuals would do
-a rapid test due to a private reason.
+private rapid test demand which is reached in the beginning of May. To cover that many
+people are likely to have sought and done their first rapid test before the Easter
+holidays to meet friends or family, we let the share of individuals doing rapid tests in
+that time increase more rapidly than before and after. By end of March 25\% of
+individuals would do a rapid test due to a private reason.

### 2021-07-09 src/paper/report.tex
Add model valiadation section. Restructure results.

@@
 depend on age and geographical location (see Supplementary Material
-Section~\ref{sub:number_of_contacts}).
+Section~\ref{subsec:data_number_of_contacts}).
 

### 2021-07-09 src/paper/results/fit.tex
Add model valiadation section. Restructure results.

@@
-\subsection{Simulated vs. Empirical Data}
+\subsection{Model Fit}
 \label{subsec:fit_results}
@@
   \label{fig:aggregated_fit2}
-\end{figure}
-
-\begin{figure}[ht]  % newly infected with single runs
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_newly_infected_with_single_runs}
-  \caption{Development of the Total Infections Over the Full Simulation Time Frame with
-  Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the true weekly incidence rates
-  per 100,000 people, including undetected cases. The mean infection rate is the thick
-  blue line. Single simulation runs are plotted in lighter and thinner lines. The higher
-  the mean incidence and the stronger the growth the more variance there is between
-  simulation runs. We averaged over 30 simulation runs.}
-  \label{fig:newly_infected_in_baseline}
 \end{figure}
@@
 
-\begin{figure}[ht]   % vaccinations by age group
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
-  \caption{Simulated and Empirical Infections by Federal State}
-  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
-  \label{fig:state_fit}
-\end{figure}
-
 \begin{figure}[ht]   % R effective

### 2021-07-09 src/paper/results/rapid_tests_statistics.tex
Add model valiadation section. Restructure results.

@@
 \label{subsec:appendix_rapid_tests}
-
-\begin{figure}[ht] % Ever Rapid Test
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-  \caption{Share of Individuals Who Have Ever Done a Rapid Test}
-  \label{fig:fit_share_ever_rapid_test}
-  \floatfoot{\noindent \textit{Note:} This figure clearly shows that overall our
-  assumptions on rapid tests are conservative. In our model, the share of individuals who
-  have ever done a rapid test lies  consistently 10 percentage points below the share
-  reported in the COSMO study \citep{Betsch2021}. The main reason for this is that we
-  introduce rapid tests only at the start of 2021. However, using the available data to
-  calibrate our rapid test parameters and estimating the remaining parameters to fit the
-  official case numbers we arrive at a shape that is very similar to that implied by the
-  survey results. This together with fitting the share of Germans with weekly rapid tests
-  well (see Figure~\ref{fig:fit_share_rapid_test_within_last_week} makes us confident
-  that our rapid test model is a good -- and especially not over-confident --
-  representation of rapid testing in Germany.}
-\end{figure}
-
-\begin{figure}[ht] % Tested in the last week
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-  \caption{Share of Individuals Who Have Done a Rapid Test in the Last Week}
-  \label{fig:fit_share_rapid_test_within_last_week}
-  \floatfoot{\noindent \textit{Note:} Our modelling of rapid tests leads to very similar rates of individuals that are tested at least weekly as in the COSMO study \citep{Betsch2021}.
-  In that study, individuals were asked if they performed rapaid tests at least weekly in
-  the last four weeks. In our model we cannot verify the rhythm over the last four
-  weeks but just observe the last time of a rapid test. However, both work and school
-  rapid tests which make up the largest share of rapid tests (see
-  Figure~\ref{fig:rapid_test_demand_by_channel}) are both required on a weekly or twice
-  weekly basis while the private rapid test demand which are partly just triggered by
-  events make up only a small fraction of tests.}
-\end{figure}
 

### 2021-07-09 src/paper/results/scenarios.tex
Add model valiadation section. Restructure results.

--- 2021-07-09 src/paper/results/scenarios.tex
@@
 
-\begin{figure}[ht] % Robustness Check
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_new_known_case}
-    \caption{Reported Cases}
-    \label{fig:robustness_check_new_known_case}
-  \end{subfigure}%
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_newly_infected}
-    \caption{Total Cases}
-    \label{fig:robustness_check_newly_infected}
-  \end{subfigure}
-  \caption{Robustness Check}
-  \label{fig:robustness_check_detailed}
-  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
-\end{figure}
-
- \comment[id=K]{@Janos: Please add caption and description for the robustness check }
-
 \begin{figure}[ht] % Random Rapid Tests
@@
 
+ \comment[id=K]{Describe that random tests are more effective at detecting cases and that
+ explains the bump in the random test scenario detection in the beginning. However, cases
+ fall immediately more in the non-random scenario because the targeted demand for tests
+ by household members leads to a much more efficient interruption of infection chains
+ because we catch the household members very early in their infectious period.}
+
 \FloatBarrier

### 2021-07-09 src/paper/results/validation.tex
Add model valiadation section. Restructure results.

\subsection{Model Validation}
\label{subsec:model_validation}

\begin{figure}[ht] % Robustness Check
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_new_known_case}
    \caption{Reported Cases}
    \label{fig:robustness_check_new_known_case}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_newly_infected}
    \caption{Total Cases}
    \label{fig:robustness_check_newly_infected}
  \end{subfigure}
  \caption{Robustness Check}
  \label{fig:robustness_check_detailed}
  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
\end{figure}

 \comment[id=K]{@Janos: Please add caption and description for the robustness check.}

### 2021-07-10 src/paper/appendix_data.tex
Move data text to different text files. Arange number of contacts figure.

--- 2021-07-09 src/paper/appendix_data.tex
@@
 
-\input{paper/medical_parameters}
+\input{paper/data/medical_parameters}
 
-\FloatBarrier
+\input{paper/data/synthetic_population}
 
-\subsection{The Synthetic Population}
+\input{paper/data/number_of_contacts}
 
-We build a synthetic population based on the German microcensus \citep{FDSAeDBUDL2018}.
-We only use private households, i.e. exclude living arrangements such as nursing homes as
-non-private households vary widely in size and it is very difficult to know which
-contacts take place in such living arrangements.
+\input{paper/data/assortativity}
 
-We sample households to build our synthetic population of over one million households
-keeping for each individual their age, gender, occupation and whether they work on
-Saturdays and Sundays. For each household we draw its county and set the corresponding
-federal state.% we draw the county because county is not part of the campus file
-
-We randomly assign 35\% of children below three to attend a nursery \citep{Destatis2020}.
-For children between three and six years old, we assume all go to preschool (officially
-92.5\% according to \cite{Destatis2020}). Nursery groups consist
-
-Children that attend a nursery meet in groups of four plus one adult care taker every
-weekday when there are no school vacations. Preschool children meet in groups of nine
-with two adult care takers. These groups are mixed with respect to age but all belong to
-the same state and mostly to the same county. For nurseries the group size is four, for
-preschools it's nine \citep{BertelsmannStiftung2019}.
-
-Every child that goes to school is part of three different classes that meet every day on
-weekdays when there are no vacations and no school policies are in place.\footnote{We
-implement vacations on the federal state level.} Each class consists of approximately 23
-students \citep{OECD2013} and two teachers. All students in a class are of the same age
-and live in the same state and if mostly also in the same county. In addition, each
-child gets assigned a value that captures his or her need to attend nursery, preschool or
-school. This allows us to capture various degrees of emergency care that can be granted
-while educational facilities are closed or function in some kind of rotating schedule.
-
-Workers are assigned to a daily meeting work group. The group sizes vary to match the
-number of daily repeating work contacts reported by working individuals in
-\cite{Mossong2008}. These groups only consist of workers that work in the same county.
-For a distribution of the number of daily recurring work contacts see
-Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
-we match each worker with up to 14 other workers to match the number of reported weekly
-work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}. Each pair is
-assigned a weekend on which they always meet in the absence of work policies. 80\% of
-these contacts are individuals from the same county.
-% work contact priority
-In the same way children have a educational priority determining if they are entitled to
-emergency care workers are assigned a work contact priority that captures how necessary
-their work is and to which degree they can work from home. This means that it's always
-the same individuals that continue to have work contacts when work from home mandates of
-a certain strictness are in place. Since our work contact priority is uniformly normal
-from zero to one, we can use the reduction in work mobility as reported by
-\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.
-
-In addition to creating groups for educational facilities and work we also have other
-recurring contacts to represent things like groups of friends or sports teams that
-practice regularly together. Both daily and weekly groups are created analogously to the
-work groups but matching the numbers in Figure~\ref{n_contacts_other_daily_recurrent} and
-Figure~\ref{n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
-highly assortative by age all individuals that have a daily leisure contact are matched
-with a person not only from the same county but also from the same age group.
-
-The individuals in our population can react to events such as developing symptoms that
-are typical of CoViD-19, a positive PCR test or a positive rapid test by reducing their
-contacts. To determine who would reduce their contacts in such a situation or demand a
-rapid test to avoid quarantine we introduce a quarantine compliance parameter. Similarly
-we introduce a rapid test compliance parameter that determines in which order individuals
-start demanding rapid tests when rapid tests become increasingly available. This makes
-sure that when for example only 10\% of workers get tested, it's the same workers that
-have access to tests every week.
-
-Lastly, for the distribution of vaccinations every individual is assigned a vaccination
-group and a vaccination rank from that group that creates a complete vaccination queue
-over the population including 15\% that refuse to be vaccinated \citep{RKI2021}. The
-vaccination groups are created to match the recommendations by the Ständige
-Impfkommission \citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized
-more than recommended by the commission.}% To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
-% put adolescents and children into two groups that follow after the general population.
-% These groups do not become eligible within our simulation frame until June.
-
-\FloatBarrier
-
-\subsection{Number of Contacts}
-\label{sub:number_of_contacts}
-
-We calibrate the parameters for the predicted numbers of contacts from contact diaries
-of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
-\citep{Mossong2008}. Each contact diary contains all contacts an individual had
-throughout one day, including information on the other person (such as age and gender)
-and information on the contact. Importantly, for each contact individuals entered of
-which type the contact (school, leisure, work etc.) was and how frequent the contact
-with the other person is.
-
-Simplifying the number of contacts, we arrive at the following distributions of the
-numbers of contacts by contact type.
-
-\begin{figure}
-    \centering
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
-        \caption{Number of Non Recurrent Other Contacts}
-        \label{n_contacts_other_non_recurrent}
-    \end{subfigure}
-
-    \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
-        \caption{Number of Daily Recurrent Other Contacts}
-        \label{n_contacts_other_daily_recurrent}
-    \end{subfigure}
-
-    \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
-        \caption{Number of Weekly Recurrent Other Contacts}
-        \label{n_contacts_other_weekly_recurrent}
-    \end{subfigure}
-
-    \vskip3ex
-    \caption{Number of Contacts of the Other Contact Type}
-    \label{fig:n_contacts_other}
-
-    \floatfoot{\noindent
-        \textit{Note:} Other contacts include all contacts that are not household
-        members, school contacts or work contacts, for example leisure contacts or
-        contacts during grocery shopping. The planned number of contacts is reduced by
-        policies, seasonality and individual responses to events such as receiving a
-        positive rapid test to the number of actual contacts with transmission potential.        % non recurrent
-        In the model it is sampled every day which of the numbers of non recurrent
-        contacts a person is planned to have. Note that the contact diaries include such
-        high values that super spreading events are well possible in our model through
-        non recurrent other models. We assume that individuals in households with
-        children or teachers or retired individuals have additional non recurrent
-        contacts during school vacations to cover things like family visits or travel
-        during vacations. We estimate this to be on average 0.5 additional contacts per
-        vacation day.
-        % daily and weekly
-        For the recurrent other contacts, individuals are assigned to groups that are
-        time constant and that meet daily or weekly. The share of individuals who attend
-        in a way that has transmission potential is reduced by policies, seasonality and
-        individual responses to events such as receiving a positive rapid test. For
-        weekly contacts, individuals are assigned to up to four groups that are time
-        constant and that meet weekly. The day on which meetings take place varies
-        between groups but stays the same for each group.
-    }
-\end{figure}
-
-% work contacts
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
-    \caption{Number of Non Recurrent Work Contacts}
-    \label{n_contacts_work_non_recurrent}
-    \floatfoot{\noindent \textit{Note:} In the model it is sampled every day which of
-        these numbers of contacts a working person is planned to have. Note that the
-        contact diaries include such high values that super spreading events are well
-        possible in our model. The planned number of contacts is reduced by policies,
-        seasonality and individual responses to events such as receiving a positive rapid
-        test to the number of actual contacts with transmission potential. Work contacts
-        only take place between working individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
-    \caption{Number of Daily Recurrent Work Contacts}
-    \label{n_contacts_work_daily_recurrent}
-    \floatfoot{\noindent \textit{Note:} Working individuals are assigned to groups that
-        are time constant and that meet daily to match the given distribution of daily
-        work contacts. You can think of these as for example colleagues with which one
-        shares an office space. The share of individuals who attend in a way that has
-        transmission potential is reduced by policies (such as a work from home mandate),
-        seasonality and individual responses to events such as receiving a positive rapid
-        test. Work contacts only take place between working individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
-    \caption{Number of Weekly Recurrent Work Contacts}
-    \label{n_contacts_work_weekly_recurrent}
-    \floatfoot{\noindent  \textit{Note:} Working individuals are assigned to up to 14
-        groups that are time constant and meet weekly. Groups are scheduled to meet on
-        separate days of the work week. These contact models cover weekly team meetings
-        etc. The share of individuals that attend in a way that has transmission
-        potential is reduced by policies, seasonality and individual responses to events
-        such as receiving a positive rapid test. Work contacts only take place between
-        working individuals.}
-\end{figure}
-
-An exception where we do not rely on the data by \cite{Mossong2008} are the household
-contacts. Since household are included in the the German microcensus
-\citep{FDSAeDBUDL2018} on which we build our synthetic population we simply assume for
-the household contact model that individuals meet all other household members every day.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
-    \caption{Number of Household Contacts}
-    \label{n_contacts_hh}
-    \floatfoot{\noindent
-        \textit{Note:} Every individual meets all other household members every day. The
-        German microcensus sampled full households such that our synthetic population
-        automatically fits population characteristics such as size and age distribution.}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Contacts by age}
-\label{subsec:contacts_by_age}
-
-As mentioned in section \ref{sec:matching}, the probability that two individuals are
-matched can depend on background characteristics. In particular, we allow this
-probability to depend on age and county of residence. While we do not have good data on
-geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same data
-we use to calibrate the number of contacts.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_other_non_recurrent}
-    \caption{Distribution of Non Recurrent Other Contacts by Age Group}
-    \label{fig:assortativity_other}
-    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
-        recurrent other contacts by age group. A row shows the share of contacts a
-        certain age group has with all other age groups. Higher values are colored in
-        darker red tones. The diagonal represents the share of contacts with individuals
-        from the same age group.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_other} shows that assortativity by age is especially strong
-for children and younger adults. For older people, the pattern becomes more dispersed
-around their own age group, but within-age-group contacts are still the most common
-contacts.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_work_non_recurrent}
-    \caption{Distribution of Random Work Contacts by Age Group}
-    \label{fig:assortativity_work}
-    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
-        recurrent work contacts by age group. A row shows the share of contacts a certain
-        age group has with all other age groups. Higher values are colored in darker red
-        tones. The diagonal represents the share of contacts with individuals from the
-        same age group. We only show age groups that have a significant fraction of
-        working individuals.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_work} shows that assortativity by age is also important
-among work contacts.
-
-Our two other types of contacts, households and schools, get their assortativity by
-construction. Schools are groups where the same children of the mostly same age and
-county meet with teachers every day. Household composition follows directly from the
-German microcensus data we use to construct our synthetic population.
-
-\FloatBarrier
-
-\subsection{Policies and Seasonality}
-\label{subsec:policies_seasonality}
-
-In our empirical application we distinguish four groups of contact types: households,
-education, work and other contacts.
-% households
-For households we assume that the individuals'
-contacts in their households do not change over our estimation period.
-% educ models
-For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures, emergency care and A / B schooling where only
-one half of students attends every other week or day. For the moment we ignore that lack
-of childcare leads working parents to stay home. An approximation of the share of contacts still taking place with the different school regulations can be found in Figure~\ref{fig:school_multiplier}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/school_multiplier_comparison}
-    \caption{School Multiplier With and Without Vacations Factored In}
-    \floatfoot{\noindent \textit{Note:} The dates on which schools have vacation are
-    decided at the federal level. Vacations are directly implemented in our model with no
-    school contacts taking place on weekends and during vacations (by federal state) just
-    like the schooling mode (full operation, emergency care, rotating schemes with half
-    class sizes etc.). The figure is thus only an illustration that roughly shows the
-    share of contacts taking place compared to pre-pandemic level with and without
-    vacations. The difference between the lines show when vacations take place and to
-    what degree. For example all states have fall vacations but the timing varies
-    strongly between states.}
-    \label{fig:school_multiplier}
-\end{figure}
-
-%
-% Schließung von Kindertagesstätten und Schulen: 37,4 Millionen ausgefallene Arbeitstage
-% http://www.iab-forum.de/schul-und-kitaschliessungen-krankheit-quarantaene-die-coronabedingten-arbeitsausfaelle-der-erwerbstaetigen-steigen-auf-592-millionen-arbeitstage/
-%
-%
-% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377: In
-% allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
-% Schüler in die Schule.
-% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430
-% 18% der Grundschüler in Notbetreuung in BW
-%
-
-% work models
-For our work models\footnote{We distinguish non-recurrent work contacts, daily work
-    contacts and weekly work contacts.} we use the reductions in work mobility reported in
-the Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
-work contacts are not random but governed through a work contact priority where the
-policy changes the threshold below which workers stay home. Figure
-\ref{fig:work_multiplier} shows the share of workers that go to work in our model over
-time.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/work_multiplier_since_sep}
-    \caption{Share of Workers with Work Contacts}
-    \label{fig:work_multiplier}
-    \floatfoot{\noindent \textit{Note:} The figure shows the work mobility as reported by
-        \cite{Google2021}. We take this as a proxy of the share of workers who are not in
-        home office, i.e. who still have physical work contacts. The figure interpolates
-        over weekends as we handle weekend effects through information on work on
-        weekends in the German census data we use. The figure shows the share aggregated
-        over Germany as a whole. To capture the effect that local policies, school
-        vacations and public policies have on work contacts we use the data on the level
-        of the federal states to determine which workers go to work depending on the
-        state they live in.}
-\end{figure}
-
-For both work and school contacts we assume that starting November with the lockdown
-light in Germany, hygiene measures (such as masks, ventilation and hand washing) became
-more strict and more conscientiously observed, leading to a reduction of 33\% in the
-number of contacts with the potential to transmit Covid-19.
-
-For the last group of contacts, which cover things like leisure activities, grocery
-shopping, etc., we have no reliable data by how much policies reduce them. In addition,
-they are likely to be affected by social and psychological factors such as pandemic
-fatigue and vacations. Because of this we estimate them like the infection probabilities
-to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays. Because of the scarce data
-situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with
-friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
-which phases we have and why the switching points are where they are.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/other_multiplier}
-    \caption{Share of Pre-Pandemic Other Contacts Taking Place with Infection Potential}
-    \label{fig:other_multiplier}
-    \floatfoot{\noindent \textit{Note:} All values are estimated. We try to use as little
-    switching points as possible and tie them to political events (such as lockdown
-    announcements) unless changes are used to capture anticipation or pandemic fatigue
-    (for example we model an anticipation of the November lockdown and model lockdown
-    fatigue in early March).}
-\end{figure}
-
-% seasonality
-Another potentially important factor for a contact to lead to an infection is the
-seasonality \citep{Kuehn2020, Carlson2020} There are two channels through which
-seasonality affects the infectiousness of contacts. One has to do with the physical
-conditions like the temperature and the humidity. The other has to do with where people
-meet. Especially leisure contacts are more likely to take place outdoors and individuals
-are more likely to have windows open when the weather is nicer. To capture both channels
-we allow for other contacts to have a higher seasonality than our other contact models.
-Figure~\ref{fig:seasonality} shows our seasonality factors.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/seasonality}
-    \caption{Seasonality by Type of Contact}
-    \label{fig:seasonality}
-    \floatfoot{\noindent \textit{Note:} We model seasonality as a factor that reduces the
-    probability of infection of all encounters. The factor depends on the day and is
-    calculated from a sinus shaped function with its maximum on January 1st. Since
-    seasonality can affect the transmission both through physical conditions such as
-    temperature and humidity as well as through the numbers of contacts that take place
-    outside we assume two seasonality factors. One for other contacts which we expect to
-    be strongly affected by fairer weather with a maximum reduction of 42\% in the
-    infection probability. The other seasonality only makes contacts up to 21\% less
-    infectious and is applied to household, work and school contacts.}
-\end{figure}
-
-\FloatBarrier
-
-\input{paper/rapid_test_params}
-
-All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
-of each line.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
-    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
-    planning to have education contacts, work contacts, developing symptoms without
-    access to a PCR test, having a household member with a positive test or symptoms. In
-    each case whether a rapid test is done depends on how long it has been since the
-    individual's last rapid test and her individual compliance parameters. As an example,
-    take a worker in May. In that time workers are encouraged to test themselves twice
-    weekly but there is no general requirement to test themselves. If the worker has not
-    done a test within the last four days in our model she will demand a test if her
-    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
-    \label{fig:rapid_test_demand}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \caption{Share of Individuals With Rapid Tests}
-    \label{fig:share_ever_rapid_test}
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-        \caption{Share of Individuals That Have Ever Done a Rapid Test}
-    \end{subfigure}%
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
-    \end{subfigure}
-    \label{fig:share_rapid_test_last_week}
-    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
-        have ever done a rapid test or done a rapid test within the last week in our
-        simulations to the shares reported in the
-        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
-        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
-        have ever done a rapid test. The right panel compares the share of individuals
-        who have done a rapid test within the last seven days in our simulation compared
-        to the share reporting to have done at least weekly rapid tests in the last four
-        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
-        conservative. The overall share is below that in the study. We fit the share of
-        weekly tests quite exactly. However, the study only covers adults while our share
-        also includes children who are tested very regularly when attending school.}
-\end{figure}
+\input{paper/data/policies}

### 2021-07-10 src/paper/data/assortativity.tex
Move data text to different text files. Arange number of contacts figure.

--- 2021-06-16 paper/appendix_data.tex
@@
-\section{Data and Parameters}
-\label{sec:data_and_parameters}
-
-The model is described by a large number of parameters that govern the number of
-contacts a person has, the likelihood of becoming infected on each contact, the
-likelihood of developing light or strong symptoms or even dying from the disease as well
-as the duration each stage of the disease takes.
-
-Most of these parameters can be calibrated from existing datasets or the medical
-literature or calibrated from surveys and empirical datasets.
-
-\input{medical_parameters}
-
-\FloatBarrier
-
-\subsection{The Synthetic Population\comment[id=K]{To be written}}
-
-\FloatBarrier
-
-\subsection{Number of Contacts\comment[id=HM]{Histograms seem like the way to go here. Also add household contacts, i.e., distribution of (household size minus 1).}}
-\label{sub:number_of_contacts}
-
-We calibrate the parameters for the predicted numbers of contacts from contact diaries
-of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
-\citep{Mossong2008}. Each contact diary contains all contacts an individual had
-throughout one day, including information on the other person (such as age and gender)
-and information on the contact. Importantly, for each contact individuals entered of
-which type the contact (school, leisure, work etc.) was and how frequent the contact
-with the other person is.
-
-Simplifying the number of contacts, we arrive at the following distributions of the
-numbers of contacts by contact type.\comment[id=K]{Put plots into one figure.}
-
-\begin{figure}
-    \centering
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
-        \caption{Number of Non Recurrent Other Contacts}
-        \label{n_contacts_other_non_recurrent}
-    \end{subfigure}
-
-    \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
-        \caption{Number of Daily Recurrent Other Contacts}
-        \label{n_contacts_other_daily_recurrent}
-    \end{subfigure}
-
-    \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
-        \caption{Number of Weekly Recurrent Other Contacts}
-        \label{n_contacts_other_weekly_recurrent}
-    \end{subfigure}
-
-    \caption{Number of Contacts of the Other Contact Type}
-    \label{fig:n_contacts_other}
-
-    \floatfoot{\noindent
-        \textit{Note:} Other contacts include all contacts that are not household
-        members, school contacts or work contacts, for example leisure contacts or
-        contacts during grocery shopping. The planned number of contacts is reduced by
-        policies, seasonality and individual responses to events such as receiving a
-        positive rapid test to the number of actual contacts with transmission potential.        % non recurrent
-        In the model it is sampled every day which of the numbers of non recurrent
-        contacts a person is planned to have. Note that the contact diaries include such
-        high values that super spreading events are well possible in our model through
-        non recurrent other models. We assume that individuals in households with
-        children or teachers or retired individuals have additional non recurrent
-        contacts during school vacations to cover things like family visits or travel
-        during vacations. We estimate this to be on average 0.5 additional contacts per
-        vacation day.
-        % daily and weekly
-        For the recurrent other contacts, individuals are assigned to groups that are
-        time constant and that meet daily or weekly. The share of individuals that attend
-        in a way that has transmission potential is reduced by policies, seasonality and
-        individual responses to events such as receiving a positive rapid test. For
-        weekly contacts, individuals are assigned to up to four groups that are time
-        constant and that meet weekly. The day on which meetings take place varies
-        between groups but stays the same for each group.
-    }
-\end{figure}
-
-% work contacts
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
-    \caption{Number of Non Recurrent Work Contacts}
-    \label{n_contacts_work_non_recurrent}
-    \floatfoot{\noindent In the model it is sampled every day which of these numbers of
-        contacts a working person is planned to have. Note that the contact diaries include
-        such high values that super spreading events are well possible in our model. The
-        planned number of contacts is reduced by policies, seasonality and individual
-        responses to events such as receiving a positive rapid test to the number of actual
-        contacts with transmission potential. Work contacts only take place between working
-        individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
-    \caption{Number of Daily Recurrent Work Contacts}
-    \label{n_contacts_work_daily_recurrent}
-    \floatfoot{\noindent Working individuals are assigned to groups that are time
-        constant and that meet daily to match the given distribution of daily work contacts.
-        You can think of these as for example colleagues with which one shares an office
-        space. The share of individuals that attend in a way that has transmission potential
-        is reduced by policies (such as a work from home mandate), seasonality and individual
-        responses to events such as receiving a positive rapid test. Work contacts only take
-        place between working individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
-    \caption{Number of Weekly Recurrent Work Contacts}
-    \label{n_contacts_work_weekly_recurrent}
-    \floatfoot{\noindent  Working individuals are assigned to up to 14 groups that are time
-        constant and meet weekly. Groups are scheduled to meet on separate days of the work
-        week. These contact models cover weekly team meetings etc. The share of individuals
-        that attend in a way that has transmission potential is reduced by policies,
-        seasonality and individual responses to events such as receiving a positive rapid
-        test. Work contacts only take place between working individuals.}
-\end{figure}
-
-An exception where we do not rely on the data by \cite{Mossong2008} are the household
-contacts. Since household are included in the the German microcensus\comment[id=K]{Cite
-microcensus} on which we build our synthetic population we simply assume for the
-household contact model that individuals meet all other household members every day.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/household}
-    \caption{Number of Household Contacts}
-    \label{n_contacts_hh}
-    \floatfoot{\noindent
-        \textit{Note:} Every individual meets all other household members every day. The
-        German microcensus sampled full households such that our synthetic population
-        automatically fits population characteristics such as size and age distribution.}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Contacts by age}
-\label{subsec:contacts_by_age}
+\subsection{Assortativity}
+\label{subsec:assortativity}
 
@@
 geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.\comment[id=HM]{Redo
-\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of
-contacts or better add a similar figure showing total number of contacts by age in all
-networks}
+within the same county, we can calibrate the assortative mixing by age from the same data
+we use to calibrate the number of contacts.
 
@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_other_non_recurrent}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_other_non_recurrent}
     \caption{Distribution of Non Recurrent Other Contacts by Age Group}
     \label{fig:assortativity_other}
-    \floatfoot{\noindent The figure shows the distribution of non recurrent other contacts by age
-        group. A row shows the share of contacts a certain age group has with all other
-        age groups. Higher values are colored in darker red tones. The diagonal
-        represents the share of contacts with individuals from the same age group.}
+    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
+        recurrent other contacts by age group. A row shows the share of contacts a
+        certain age group has with all other age groups. Higher values are colored in
+        darker red tones. The diagonal represents the share of contacts with individuals
+        from the same age group.}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_work_non_recurrent}
+    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_work_non_recurrent}
     \caption{Distribution of Random Work Contacts by Age Group}
     \label{fig:assortativity_work}
-    \floatfoot{\noindent The figure shows the distribution of non recurrent work contacts by age
-        group. A row shows the share of contacts a certain age group has with all other
-        age groups. Higher values are colored in darker red tones. The diagonal
-        represents the share of contacts with individuals from the same age group. We
-        only show age groups that have a significant fraction of working individuals.}
+    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
+        recurrent work contacts by age group. A row shows the share of contacts a certain
+        age group has with all other age groups. Higher values are colored in darker red
+        tones. The diagonal represents the share of contacts with individuals from the
+        same age group. We only show age groups that have a significant fraction of
+        working individuals.}
 \end{figure}
@@
 
-Our other two types of contacts, households and schools, get their assortativity by
-construction. Schools are groups where the same children of the mostly same age group and
+Our two other types of contacts, households and schools, get their assortativity by
+construction. Schools are groups where the same children of the mostly same age and
 county meet with teachers every day. Household composition follows directly from the
-German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
-no work in here? How about school? Can we add a column with marginals and another one
-with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
-households addresses all questions.
-I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}\comment[id=HM]{See previous comment.
-My point is that total number of contacts rather than shares are much more informative in the first place, then different group sizes are no concern anymore.}
-
-\subsection{Policies and Seasonality}
-\label{subsec:policies_seasonality}
+German microcensus data we use to construct our synthetic population.
 
 \FloatBarrier
-
-In our empirical application we distinguish four groups of contact types: households,
-education, work and other contacts.
-% households
-For households we assume that the individuals'
-contacts in their households do not change over our estimation period.
-% educ models
-For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures, emergency care and A / B schooling where only
-one half of students attends every other week or day. For the moment we ignore that lack
-of childcare leads working parents to stay home. An approximation of the share of contacts still taking place with the different school regulations can be found in Figure~\ref{fig:school_multiplier}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/school_multiplier_comparison}
-    \caption{School Multiplier With and Without Vacations Factored In}
-    \floatfoot{\noindent}
-    \label{fig:school_multiplier}
-\end{figure}
-
-%
-% Schließung von Kindertagesstätten und Schulen: 37,4 Millionen ausgefallene Arbeitstage
-% http://www.iab-forum.de/schul-und-kitaschliessungen-krankheit-quarantaene-die-coronabedingten-arbeitsausfaelle-der-erwerbstaetigen-steigen-auf-592-millionen-arbeitstage/
-%
-%
-% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377: In
-% allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
-% Schüler in die Schule.
-% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430
-% 18% der Grundschüler in Notbetreuung in BW
-%
-
-% work models
-For our work models\footnote{We distinguish non-recurrent work contacts, daily work
-    contacts and weekly work contacts.} we use the reductions in work mobility reported in
-the Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
-work contacts are not random but governed through a work contact priority where the
-policy changes the threshold below which workers stay home. Figure
-\ref{fig:work_multiplier} shows the share of workers that go to work in our model over
-time.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
-    \caption{Share of Workers with Work Contacts}
-    \label{fig:work_multiplier}
-    \floatfoot{\noindent The figure shows the work mobility as reported by \cite{Google2021}. We
-        take this as a proxy of the share of workers who are not in home office, i.e. who
-        still have physical work contacts. The figure interpolates over weekends as we
-        handle weekend effects through information on work on weekends in the German
-        census data we use. The figure shows the share aggregated over Germany as a
-        whole. To capture the effect that local policies, school vacations and public
-        policies have on work contacts we use the data by German state to determine which
-        workers go to work depending on the state they live in.}
-\end{figure}
-
-For both work and school contacts we assume that starting November with the lockdown
-light in Germany, hygiene measures (such as masks, ventilation and hand washing) became
-more strict and more conscientiously observed, leading to a reduction of 33\% in the
-number of contacts with the potential to transmit Covid-19.
-
-% other contacts
-For the last group of contacts which cover things like leisure activities, grocery
-shopping etc. we have no reliable data by how much policies reduce them. In addition,
-they are likely to be affected by social and psychological factors such as pandemic
-fatigue and vacations. Because of this we estimate them like the infection probabilities
-to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays. Because of the scarce data
-situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with friends).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/other_multiplier}
-    \caption{Share of Other Contacts Still Taking Place}
-    \label{fig:other_multiplier}
-    \floatfoot{\noindent}
-\end{figure}
-
-% seasonality
-Another potentially important factor for a contact to lead to an infection is the
-seasonality \citep{Kuehn2020, Carlson2020} There are two channels through which
-seasonality affects the infectiousness of contacts. One has to do with the physical
-conditions like the temperature and the humidity. The other has to do with where people
-meet. Especially leisure contacts are more likely to take place outdoors and individuals
-are more likely to have windows open when the weather is nicer. To capture both channels
-we allow for other contacts to have a higher seasonality than our other contact models.
-Figure~\ref{fig:seasonality} shows our seasonality factors.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
-    \caption{Seasonality by Type of Contact}
-    \label{fig:seasonality}
-    \floatfoot{\noindent}
-\end{figure}
-
-\FloatBarrier
-
-\input{rapid_test_params}
-
-All shares of individuals that would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
-of each line.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of Individuals that do a Rapid Test.}}
-    \floatfoot{\noindent ...}
-    \label{fig:rapid_test_demand}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \caption{Share of Individuals With Rapid Tests}
-    \label{fig:share_ever_rapid_test}
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-    \end{subfigure}%
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        %   \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
-    \end{subfigure}
-    \label{fig:share_rapid_test_last_week}
-    \floatfoot{\noindent The figure compares the share of individuals who have ever done a rapid
-        test or done a rapid test within the last week in our simulations to the shares
-        reported in the
-        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
-            Snapshot Monitoring survey}. The left panel compares the share of individuals
-        who have ever done a rapid test. The right panel compares the share of
-        individuals who have done a rapid test within the last seven days in our
-        simulation compared to the share reporting to have done at least weekly rapid
-        tests in the last four weeks in the COSMO survey. Overall our calibration of
-        rapid tests are slightly conservative. The overall share is below that in the
-        study. We fit the share of weekly tests quite exactly. However, the study only
-        covers adults while our share also includes children who are tested very
-        regularly when attending school.}
-\end{figure}

### 2021-07-10 src/paper/data/number_of_contacts.tex
Move data text to different text files. Arange number of contacts figure.

--- 2021-06-17 paper/appendix_data.tex
@@
-\section{Data and Parameters}
-\label{sec:data_and_parameters}
-
-The model is described by a large number of parameters that govern the number of
-contacts a person has, the likelihood of becoming infected on each contact, the
-likelihood of developing light or strong symptoms or even dying from the disease as well
-as the duration each stage of the disease takes.
-
-Most of these parameters can be calibrated from existing datasets or the medical
-literature or calibrated from surveys and empirical datasets.
-
-\input{medical_parameters}
-
-\FloatBarrier
-
-\subsection{The Synthetic Population\comment[id=K]{To be written}}
-
-\FloatBarrier
-
-\subsection{Number of Contacts\comment[id=HM]{Histograms seem like the way to go here. Also add household contacts, i.e., distribution of (household size minus 1).}}
-\label{sub:number_of_contacts}
+\subsection{Number of Contacts}
+\label{subsec:data_number_of_contacts}
 
@@
     \centering
-
-    \begin{subfigure}[b]{0.25\textwidth}
+    % other contacts
+    \begin{subfigure}[b]{0.3\textwidth}
         \centering
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
         \caption{Number of Non Recurrent Other Contacts}
@@
     \end{subfigure}
-
     \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
+    \begin{subfigure}[b]{0.3\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
         \caption{Number of Daily Recurrent Other Contacts}
@@
     \end{subfigure}
-
     \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
+    \begin{subfigure}[b]{0.3\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
         \caption{Number of Weekly Recurrent Other Contacts}
@@
 
-    \caption{Number of Contacts of the Other Contact Type}
+    \vskip3ex
+
+    % work contacts
+    \begin{subfigure}[b]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
+        \caption{Number of Non Recurrent Work Contacts}
+        \label{n_contacts_work_non_recurrent}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
+        \caption{Number of Daily Recurrent Work Contacts}
+        \label{n_contacts_work_daily_recurrent}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
+        \caption{Number of Weekly Recurrent Work Contacts}
+        \label{n_contacts_work_weekly_recurrent}
+    \end{subfigure}
+    \vskip3ex
+
+    \caption{Number of Contacts of the Different Contact Types}
     \label{fig:n_contacts_other}
-
     \floatfoot{\noindent
-        \textit{Note:} Other contacts include all contacts that are not household
-        members, school contacts or work contacts, for example leisure contacts or
-        contacts during grocery shopping. The planned number of contacts is reduced by
-        policies, seasonality and individual responses to events such as receiving a
-        positive rapid test to the number of actual contacts with transmission potential.        % non recurrent
+        \textit{Note:} The upper row shows the distribution of the number of other
+        contacts individuals report. Other contacts include all contacts that are not
+        household members, school contacts or work contacts, for example leisure
+        contacts. The planned number of contacts is reduced by policies, seasonality and
+        individual responses to events such as receiving a positive rapid test to the
+        number of actual contacts with transmission potential.
+        % non recurrent
         In the model it is sampled every day which of the numbers of non recurrent
@@
         between groups but stays the same for each group.
-    }
+        % work contacts
+        The lower row shows the distribution of the different types of work contacts.
+        Work contacts only take place between working individuals.
+        % non recurrent
+        In the model it is sampled every day which number of non recurrent work contacts
+        a working person is planned to have. Note that the contact diaries include such
+        high values that super spreading events are well possible in our model. The
+        planned number of contacts is reduced by policies, seasonality and individual
+        responses to events such as receiving a positive rapid test to the number of
+        actual contacts with transmission potential.
+        % daily
+        Working individuals are assigned to groups that are time constant and that meet
+        daily to match the given distribution of daily work contacts. This contact type
+        covers contacts such as colleagues with which one shares an office space. The
+        share of individuals who attend in a way that has transmission potential is
+        reduced by policies (such as a work from home mandate), seasonality and
+        individual responses to events such as receiving a positive rapid test.
+        % weekly
+        Working individuals are assigned to up to 14 groups that are time constant and
+        meet weekly. Groups are scheduled to meet on separate days of the work week.
+        These contact models cover weekly team meetings etc. The share of individuals
+        that attend in a way that has transmission potential is reduced by policies,
+        seasonality and individual responses to events such as receiving a positive rapid
+        test.}
 \end{figure}
 
-% work contacts
+An exception where we do not rely on the data by \cite{Mossong2008} are the household
+contacts. Since household are included in the the German microcensus
+\citep{FDSAeDBUDL2018} on which we build our synthetic population we simply assume for
+the household contact model that individuals meet all other household members every day.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
-    \caption{Number of Non Recurrent Work Contacts}
-    \label{n_contacts_work_non_recurrent}
-    \floatfoot{\noindent In the model it is sampled every day which of these numbers of
-        contacts a working person is planned to have. Note that the contact diaries include
-        such high values that super spreading events are well possible in our model. The
-        planned number of contacts is reduced by policies, seasonality and individual
-        responses to events such as receiving a positive rapid test to the number of actual
-        contacts with transmission potential. Work contacts only take place between working
-        individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
-    \caption{Number of Daily Recurrent Work Contacts}
-    \label{n_contacts_work_daily_recurrent}
-    \floatfoot{\noindent Working individuals are assigned to groups that are time
-        constant and that meet daily to match the given distribution of daily work contacts.
-        You can think of these as for example colleagues with which one shares an office
-        space. The share of individuals who attend in a way that has transmission potential
-        is reduced by policies (such as a work from home mandate), seasonality and individual
-        responses to events such as receiving a positive rapid test. Work contacts only take
-        place between working individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
-    \caption{Number of Weekly Recurrent Work Contacts}
-    \label{n_contacts_work_weekly_recurrent}
-    \floatfoot{\noindent  Working individuals are assigned to up to 14 groups that are time
-        constant and meet weekly. Groups are scheduled to meet on separate days of the work
-        week. These contact models cover weekly team meetings etc. The share of individuals
-        that attend in a way that has transmission potential is reduced by policies,
-        seasonality and individual responses to events such as receiving a positive rapid
-        test. Work contacts only take place between working individuals.}
-\end{figure}
-
-An exception where we do not rely on the data by \cite{Mossong2008} are the household
-contacts. Since household are included in the the German microcensus\comment[id=K]{Cite
-microcensus} on which we build our synthetic population we simply assume for the
-household contact model that individuals meet all other household members every day.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/household}
+    \includegraphics[width=0.4\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
     \caption{Number of Household Contacts}
@@
 \FloatBarrier
-
-\subsection{Contacts by age}
-\label{subsec:contacts_by_age}
-
-As mentioned in section \ref{sec:matching}, the probability that two individuals are
-matched can depend on background characteristics. In particular, we allow this
-probability to depend on age and county of residence. While we do not have good data on
-geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.\comment[id=HM]{Redo
-\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of
-contacts or better add a similar figure showing total number of contacts by age in all
-networks}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_other_non_recurrent}
-    \caption{Distribution of Non Recurrent Other Contacts by Age Group}
-    \label{fig:assortativity_other}
-    \floatfoot{\noindent The figure shows the distribution of non recurrent other contacts by age
-        group. A row shows the share of contacts a certain age group has with all other
-        age groups. Higher values are colored in darker red tones. The diagonal
-        represents the share of contacts with individuals from the same age group.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_other} shows that assortativity by age is especially strong
-for children and younger adults. For older people, the pattern becomes more dispersed
-around their own age group, but within-age-group contacts are still the most common
-contacts.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_work_non_recurrent}
-    \caption{Distribution of Random Work Contacts by Age Group}
-    \label{fig:assortativity_work}
-    \floatfoot{\noindent The figure shows the distribution of non recurrent work contacts by age
-        group. A row shows the share of contacts a certain age group has with all other
-        age groups. Higher values are colored in darker red tones. The diagonal
-        represents the share of contacts with individuals from the same age group. We
-        only show age groups that have a significant fraction of working individuals.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_work} shows that assortativity by age is also important
-among work contacts.
-
-Our other two types of contacts, households and schools, get their assortativity by
-construction. Schools are groups where the same children of the mostly same age group and
-county meet with teachers every day. Household composition follows directly from the
-German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
-no work in here? How about school? Can we add a column with marginals and another one
-with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
-households addresses all questions.
-I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}\comment[id=HM]{See previous comment.
-My point is that total number of contacts rather than shares are much more informative in the first place, then different group sizes are no concern anymore.}
-
-\subsection{Policies and Seasonality}
-\label{subsec:policies_seasonality}
-
-\FloatBarrier
-
-In our empirical application we distinguish four groups of contact types: households,
-education, work and other contacts.
-% households
-For households we assume that the individuals'
-contacts in their households do not change over our estimation period.
-% educ models
-For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures, emergency care and A / B schooling where only
-one half of students attends every other week or day. For the moment we ignore that lack
-of childcare leads working parents to stay home. An approximation of the share of contacts still taking place with the different school regulations can be found in Figure~\ref{fig:school_multiplier}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/school_multiplier_comparison}
-    \caption{School Multiplier With and Without Vacations Factored In}
-    \floatfoot{\noindent}
-    \label{fig:school_multiplier}
-\end{figure}
-
-%
-% Schließung von Kindertagesstätten und Schulen: 37,4 Millionen ausgefallene Arbeitstage
-% http://www.iab-forum.de/schul-und-kitaschliessungen-krankheit-quarantaene-die-coronabedingten-arbeitsausfaelle-der-erwerbstaetigen-steigen-auf-592-millionen-arbeitstage/
-%
-%
-% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377: In
-% allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
-% Schüler in die Schule.
-% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430
-% 18% der Grundschüler in Notbetreuung in BW
-%
-
-% work models
-For our work models\footnote{We distinguish non-recurrent work contacts, daily work
-    contacts and weekly work contacts.} we use the reductions in work mobility reported in
-the Google Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in
-work contacts are not random but governed through a work contact priority where the
-policy changes the threshold below which workers stay home. Figure
-\ref{fig:work_multiplier} shows the share of workers that go to work in our model over
-time.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/work_multiplier_since_sep}
-    \caption{Share of Workers with Work Contacts}
-    \label{fig:work_multiplier}
-    \floatfoot{\noindent The figure shows the work mobility as reported by \cite{Google2021}. We
-        take this as a proxy of the share of workers who are not in home office, i.e. who
-        still have physical work contacts. The figure interpolates over weekends as we
-        handle weekend effects through information on work on weekends in the German
-        census data we use. The figure shows the share aggregated over Germany as a
-        whole. To capture the effect that local policies, school vacations and public
-        policies have on work contacts we use the data by German state to determine which
-        workers go to work depending on the state they live in.}
-\end{figure}
-
-For both work and school contacts we assume that starting November with the lockdown
-light in Germany, hygiene measures (such as masks, ventilation and hand washing) became
-more strict and more conscientiously observed, leading to a reduction of 33\% in the
-number of contacts with the potential to transmit Covid-19.
-
-% other contacts
-For the last group of contacts which cover things like leisure activities, grocery
-shopping etc. we have no reliable data by how much policies reduce them. In addition,
-they are likely to be affected by social and psychological factors such as pandemic
-fatigue and vacations. Because of this we estimate them like the infection probabilities
-to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays. Because of the scarce data
-situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with friends).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/other_multiplier}
-    \caption{Share of Other Contacts Still Taking Place}
-    \label{fig:other_multiplier}
-    \floatfoot{\noindent}
-\end{figure}
-
-% seasonality
-Another potentially important factor for a contact to lead to an infection is the
-seasonality \citep{Kuehn2020, Carlson2020} There are two channels through which
-seasonality affects the infectiousness of contacts. One has to do with the physical
-conditions like the temperature and the humidity. The other has to do with where people
-meet. Especially leisure contacts are more likely to take place outdoors and individuals
-are more likely to have windows open when the weather is nicer. To capture both channels
-we allow for other contacts to have a higher seasonality than our other contact models.
-Figure~\ref{fig:seasonality} shows our seasonality factors.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
-    \caption{Seasonality by Type of Contact}
-    \label{fig:seasonality}
-    \floatfoot{\noindent}
-\end{figure}
-
-\FloatBarrier
-
-\input{rapid_test_params}
-
-All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
-of each line.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of individuals who do a Rapid Test.}}
-    \floatfoot{\noindent ...}
-    \label{fig:rapid_test_demand}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \caption{Share of Individuals With Rapid Tests}
-    \label{fig:share_ever_rapid_test}
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-    \end{subfigure}%
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        %   \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
-    \end{subfigure}
-    \label{fig:share_rapid_test_last_week}
-    \floatfoot{\noindent The figure compares the share of individuals who have ever done a rapid
-        test or done a rapid test within the last week in our simulations to the shares
-        reported in the
-        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
-            Snapshot Monitoring survey}. The left panel compares the share of individuals
-        who have ever done a rapid test. The right panel compares the share of
-        individuals who have done a rapid test within the last seven days in our
-        simulation compared to the share reporting to have done at least weekly rapid
-        tests in the last four weeks in the COSMO survey. Overall our calibration of
-        rapid tests are slightly conservative. The overall share is below that in the
-        study. We fit the share of weekly tests quite exactly. However, the study only
-        covers adults while our share also includes children who are tested very
-        regularly when attending school.}
-\end{figure}

### 2021-07-10 src/paper/data/policies.tex
Move data text to different text files. Arange number of contacts figure.

--- 2021-06-18 paper/appendix_data.tex
@@
-\section{Data and Parameters}
-\label{sec:data_and_parameters}
+\subsection{Policies}
+\label{subsec:policies}
 
-The model is described by a large number of parameters that govern the number of
-contacts a person has, the likelihood of becoming infected on each contact, the
-likelihood of developing light or strong symptoms or even dying from the disease as well
-as the duration each stage of the disease takes.
-
-Most of these parameters can be calibrated from existing datasets or the medical
-literature or calibrated from surveys and empirical datasets.
-
-\input{medical_parameters}
-
-\FloatBarrier
-
-\subsection{The Synthetic Population\comment[id=K]{To be written}}
-
-\FloatBarrier
-
-\subsection{Number of Contacts\comment[id=HM]{Histograms seem like the way to go here. Also add household contacts, i.e., distribution of (household size minus 1).}}
-\label{sub:number_of_contacts}
-
-We calibrate the parameters for the predicted numbers of contacts from contact diaries
-of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
-\citep{Mossong2008}. Each contact diary contains all contacts an individual had
-throughout one day, including information on the other person (such as age and gender)
-and information on the contact. Importantly, for each contact individuals entered of
-which type the contact (school, leisure, work etc.) was and how frequent the contact
-with the other person is.
-
-Simplifying the number of contacts, we arrive at the following distributions of the
-numbers of contacts by contact type.\comment[id=K]{Put plots into one figure.}
-
-\begin{figure}
-    \centering
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
-        \caption{Number of Non Recurrent Other Contacts}
-        \label{n_contacts_other_non_recurrent}
-    \end{subfigure}
-
-    \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
-        \caption{Number of Daily Recurrent Other Contacts}
-        \label{n_contacts_other_daily_recurrent}
-    \end{subfigure}
-
-    \hfill
-
-    \begin{subfigure}[b]{0.25\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
-        \caption{Number of Weekly Recurrent Other Contacts}
-        \label{n_contacts_other_weekly_recurrent}
-    \end{subfigure}
-
-    \caption{Number of Contacts of the Other Contact Type}
-    \label{fig:n_contacts_other}
-
-    \floatfoot{\noindent
-        \textit{Note:} Other contacts include all contacts that are not household
-        members, school contacts or work contacts, for example leisure contacts or
-        contacts during grocery shopping. The planned number of contacts is reduced by
-        policies, seasonality and individual responses to events such as receiving a
-        positive rapid test to the number of actual contacts with transmission potential.        % non recurrent
-        In the model it is sampled every day which of the numbers of non recurrent
-        contacts a person is planned to have. Note that the contact diaries include such
-        high values that super spreading events are well possible in our model through
-        non recurrent other models. We assume that individuals in households with
-        children or teachers or retired individuals have additional non recurrent
-        contacts during school vacations to cover things like family visits or travel
-        during vacations. We estimate this to be on average 0.5 additional contacts per
-        vacation day.
-        % daily and weekly
-        For the recurrent other contacts, individuals are assigned to groups that are
-        time constant and that meet daily or weekly. The share of individuals who attend
-        in a way that has transmission potential is reduced by policies, seasonality and
-        individual responses to events such as receiving a positive rapid test. For
-        weekly contacts, individuals are assigned to up to four groups that are time
-        constant and that meet weekly. The day on which meetings take place varies
-        between groups but stays the same for each group.
-    }
-\end{figure}
-
-% work contacts
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
-    \caption{Number of Non Recurrent Work Contacts}
-    \label{n_contacts_work_non_recurrent}
-    \floatfoot{\noindent \textit{Note:} In the model it is sampled every day which of
-        these numbers of contacts a working person is planned to have. Note that the
-        contact diaries include such high values that super spreading events are well
-        possible in our model. The planned number of contacts is reduced by policies,
-        seasonality and individual responses to events such as receiving a positive rapid
-        test to the number of actual contacts with transmission potential. Work contacts
-        only take place between working individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
-    \caption{Number of Daily Recurrent Work Contacts}
-    \label{n_contacts_work_daily_recurrent}
-    \floatfoot{\noindent \textit{Note:} Working individuals are assigned to groups that
-        are time constant and that meet daily to match the given distribution of daily
-        work contacts. You can think of these as for example colleagues with which one
-        shares an office space. The share of individuals who attend in a way that has
-        transmission potential is reduced by policies (such as a work from home mandate),
-        seasonality and individual responses to events such as receiving a positive rapid
-        test. Work contacts only take place between working individuals.}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
-    \caption{Number of Weekly Recurrent Work Contacts}
-    \label{n_contacts_work_weekly_recurrent}
-    \floatfoot{\noindent  \textit{Note:} Working individuals are assigned to up to 14
-        groups that are time constant and meet weekly. Groups are scheduled to meet on
-        separate days of the work week. These contact models cover weekly team meetings
-        etc. The share of individuals that attend in a way that has transmission
-        potential is reduced by policies, seasonality and individual responses to events
-        such as receiving a positive rapid test. Work contacts only take place between
-        working individuals.}
-\end{figure}
-
-An exception where we do not rely on the data by \cite{Mossong2008} are the household
-contacts. Since household are included in the the German microcensus\comment[id=K]{Cite
-microcensus} on which we build our synthetic population we simply assume for the
-household contact model that individuals meet all other household members every day.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/distributions_of_the_number_of_contacts/household}
-    \caption{Number of Household Contacts}
-    \label{n_contacts_hh}
-    \floatfoot{\noindent
-        \textit{Note:} Every individual meets all other household members every day. The
-        German microcensus sampled full households such that our synthetic population
-        automatically fits population characteristics such as size and age distribution.}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Contacts by age}
-\label{subsec:contacts_by_age}
-
-As mentioned in section \ref{sec:matching}, the probability that two individuals are
-matched can depend on background characteristics. In particular, we allow this
-probability to depend on age and county of residence. While we do not have good data on
-geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same
-data we use to calibrate the number of contacts.\comment[id=HM]{Redo
-\ref{fig:assortativity_other} /\ref{fig:assortativity_work} with total number of
-contacts or better add a similar figure showing total number of contacts by age in all
-networks}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_other_non_recurrent}
-    \caption{Distribution of Non Recurrent Other Contacts by Age Group}
-    \label{fig:assortativity_other}
-    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
-        recurrent other contacts by age group. A row shows the share of contacts a
-        certain age group has with all other age groups. Higher values are colored in
-        darker red tones. The diagonal represents the share of contacts with individuals
-        from the same age group.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_other} shows that assortativity by age is especially strong
-for children and younger adults. For older people, the pattern becomes more dispersed
-around their own age group, but within-age-group contacts are still the most common
-contacts.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/data/assortativity_work_non_recurrent}
-    \caption{Distribution of Random Work Contacts by Age Group}
-    \label{fig:assortativity_work}
-    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
-        recurrent work contacts by age group. A row shows the share of contacts a certain
-        age group has with all other age groups. Higher values are colored in darker red
-        tones. The diagonal represents the share of contacts with individuals from the
-        same age group. We only show age groups that have a significant fraction of
-        working individuals.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_work} shows that assortativity by age is also important
-among work contacts.
-
-Our other two types of contacts, households and schools, get their assortativity by
-construction. Schools are groups where the same children of the mostly same age group and
-county meet with teachers every day. Household composition follows directly from the
-German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
-no work in here? How about school? Can we add a column with marginals and another one
-with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
-households addresses all questions.
-I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}\comment[id=HM]{See previous comment.
-My point is that total number of contacts rather than shares are much more informative in the first place, then different group sizes are no concern anymore.}
-
-\subsection{Policies and Seasonality}
-\label{subsec:policies_seasonality}
-
-\FloatBarrier
-
-In our empirical application we distinguish four groups of contact types: households,
-education, work and other contacts.
-% households
+In our empirical application\comment[id=K]{Improve language} we distinguish four groups
+of contact types: households, education, work and other contacts.% households
 For households we assume that the individuals'
@@
 
-% seasonality
-Another potentially important factor for a contact to lead to an infection is the
-seasonality \citep{Kuehn2020, Carlson2020} There are two channels through which
-seasonality affects the infectiousness of contacts. One has to do with the physical
-conditions like the temperature and the humidity. The other has to do with where people
-meet. Especially leisure contacts are more likely to take place outdoors and individuals
-are more likely to have windows open when the weather is nicer. To capture both channels
-we allow for other contacts to have a higher seasonality than our other contact models.
-Figure~\ref{fig:seasonality} shows our seasonality factors.
+\FloatBarrier
+
+\input{paper/rapid_test_params}
+
+All shares of individuals who would take a rapid test if the conditions were met can be
+seen in Figure~\ref{fig:rapid_test_demand}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
-    \caption{Seasonality by Type of Contact}
-    \label{fig:seasonality}
-    \floatfoot{\noindent \textit{Note:} We model seasonality as a factor that reduces the
-    probability of infection of all encounters. The factor depends on the day and is
-    calculated from a sinus shaped function with its maximum on January 1st. Since
-    seasonality can affect the transmission both through physical conditions such as
-    temperature and humidity as well as through the numbers of contacts that take place
-    outside we assume two seasonality factors. One for other contacts which we expect to
-    be strongly affected by fairer weather with a maximum reduction of 30\% in the
-    infection probability. The other seasonality only makes contacts up to 20\% less
-    infectious and is applied to household, work and school contacts.}
-\end{figure}
-
-\FloatBarrier
-
-\input{rapid_test_params}
-
-All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.\comment[id=J]{Talk about the interpretation
-of each line.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/results/figures/data/testing/rapid_test_demand_shares}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
     \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
@@
     \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+        \caption{Share of Individuals That Have Ever Done a Rapid Test}
     \end{subfigure}%
     \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        %   \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
+        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
     \end{subfigure}
@@
 \end{figure}
+
+ \comment[id=K]{J: Explain that we don't fit the share ever tested that well because our rapid test compliance is completely fixed. }
+
+\FloatBarrier

### 2021-07-10 src/paper/data/synthetic_population.tex
Move data text to different text files. Arange number of contacts figure.

\subsection{The Synthetic Population}
\label{subsec:synthetic_population}

We build a synthetic population based on the German microcensus \citep{FDSAeDBUDL2018}.
We only use private households, i.e. exclude living arrangements such as nursing homes as
non-private households vary widely in size and it is very difficult to know which
contacts take place in such living arrangements.

We sample households to build our synthetic population of over one million households
keeping for each individual their age, gender, occupation and whether they work on
Saturdays and Sundays. For each household we draw its county and set the corresponding
federal state.% we draw the county because county is not part of the campus file

We randomly assign 35\% of children below three to attend a nursery \citep{Destatis2020}.
For children between three and six years old, we assume all go to preschool (officially
92.5\% according to \cite{Destatis2020}). Nursery groups consist

Children that attend a nursery meet in groups of four plus one adult care taker every
weekday when there are no school vacations. Preschool children meet in groups of nine
with two adult care takers. These groups are mixed with respect to age but all belong to
the same state and mostly to the same county. For nurseries the group size is four, for
preschools it's nine \citep{BertelsmannStiftung2019}.

Every child that goes to school is part of three different classes that meet every day on
weekdays when there are no vacations and no school policies are in place.\footnote{We
implement vacations on the federal state level.} Each class consists of approximately 23
students \citep{OECD2013} and two teachers. All students in a class are of the same age
and live in the same state and if mostly also in the same county. In addition, each
child gets assigned a value that captures his or her need to attend nursery, preschool or
school. This allows us to capture various degrees of emergency care that can be granted
while educational facilities are closed or function in some kind of rotating schedule.

Workers are assigned to a daily meeting work group. The group sizes vary to match the
number of daily repeating work contacts reported by working individuals in
\cite{Mossong2008}. These groups only consist of workers that work in the same county.
For a distribution of the number of daily recurring work contacts see
Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
we match each worker with up to 14 other workers to match the number of reported weekly
work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}. Each pair is
assigned a weekend on which they always meet in the absence of work policies. 80\% of
these contacts are individuals from the same county.
% work contact priority
In the same way children have a educational priority determining if they are entitled to
emergency care workers are assigned a work contact priority that captures how necessary
their work is and to which degree they can work from home. This means that it's always
the same individuals that continue to have work contacts when work from home mandates of
a certain strictness are in place. Since our work contact priority is uniformly normal
from zero to one, we can use the reduction in work mobility as reported by
\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.

In addition to creating groups for educational facilities and work we also have other
recurring contacts to represent things like groups of friends or sports teams that
practice regularly together. Both daily and weekly groups are created analogously to the
work groups but matching the numbers in Figure~\ref{n_contacts_other_daily_recurrent} and
Figure~\ref{n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
highly assortative by age all individuals that have a daily leisure contact are matched
with a person not only from the same county but also from the same age group.

The individuals in our population can react to events such as developing symptoms that
are typical of CoViD-19, a positive PCR test or a positive rapid test by reducing their
contacts. To determine who would reduce their contacts in such a situation or demand a
rapid test to avoid quarantine we introduce a quarantine compliance parameter. Similarly
we introduce a rapid test compliance parameter that determines in which order individuals
start demanding rapid tests when rapid tests become increasingly available. This makes
sure that when for example only 10\% of workers get tested, it's the same workers that
have access to tests every week.

Lastly, for the distribution of vaccinations every individual is assigned a vaccination
group and a vaccination rank from that group that creates a complete vaccination queue
over the population including 15\% that refuse to be vaccinated \citep{RKI2021}. The
vaccination groups are created to match the recommendations by the Ständige
Impfkommission \citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized
more than recommended by the commission.}\comment[id=K]{Write how that plays out in our simulations for the different age groups.}

% To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
% put adolescents and children into two groups that follow after the general population.
% These groups do not become eligible within our simulation frame until June.

% \begin{figure}[ht]   % vaccinations by age group
%   \centering
%   \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
%   \caption{Simulated and Empirical Infections by Federal State}
%   \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
%   \label{fig:vaccinations_by_age_group}
% \end{figure}

\FloatBarrier
os

### 2021-07-10 src/paper/model/seasonality.tex
Add comments by Philipp Eisenhauer.
Make clear which parameters have distributions. Nicer math formatting. Clearer parameter names. Mention normalizations.
Add model valiadation section. Restructure results.
Move data text to different text files. Arange number of contacts figure.

@@
-\subsection{Seasonality}
+\subsection{Seasonality}\comment[id=K]{Philipp suggests to write that we choose upper
+bounds}
 \label{subsec:seasonality}
@@
 
-For simplicity we normalize the factor to reach one at its maximum. Thus the formula of
+For simplicity we normalize the factor to reach one at its maximum. Thus, the formula of
 the seasonality factor is given by:
@@
 \begin{equation}
-    s_k(t) = 1 + 0.5 \kappa_k  sin \left ( \pi  \left (\frac{1}{2} + \frac{t}{182.5}\right ) \right ) - 0.5 \kappa_k
+\label{eq:seasonality}
+    s_{c, t} = 1 + 0.5 \kappa_c  sin \left ( \pi  \left (\frac{1}{2} + \frac{t}{182.5}\right ) \right ) - 0.5 \kappa_c
 \end{equation}
@@
 
-We calibrate $kappa_{strong}$ to 0.42 and $kappa_{weak}$ to 0.21. This is in line with
+We calibrate $\kappa_{strong}$ to 0.42 and $\kappa_{weak}$ to 0.21. This is in line with
 \cite{Gavenciak2021} and \cite{Kuehn2020}.
+
+The two seasonality curves are shown in Figure~\ref{fig:seasonality}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/data/seasonality}
+    \caption{Seasonality by Type of Contact}
+    \label{fig:seasonality}
+    \floatfoot{\noindent \textit{Note:} We model seasonality as a factor that reduces the
+    probability of infection of all encounters. The factor depends on the day and is
+    calculated from a sinus shaped function with its maximum on January 1st. Since
+    seasonality can affect the transmission both through physical conditions such as
+    temperature and humidity as well as through the numbers of contacts that take place
+    outside we assume two seasonality factors. One for other contacts which we expect to
+    be strongly affected by fairer weather with a maximum reduction of 42\% in the
+    infection probability. The other seasonality only makes contacts up to 21\% less
+    infectious and is applied to household, work and school contacts.}
+\end{figure}

### 2021-07-10 src/paper/model/estimated_parameters.tex
Add first version of estimated parameters table.

@@
 % policies
-We also estimate a parameter that reflects the effect of hygiene measures
-after November 2020 at work and in educational facilities. This parameter reduces
-infectiousness of contacts by one third. In total those are 10 parameters. The
-breakpoints the contact reduction changes are not determined from data but from
-announcements of policy changes. Moreover, we constrain the estimated contact reduction
-to follow the shape of the stringency index. The resulting contact reduction can be
-seen in Figure~\ref{fig:other_multiplier}
+We also estimate a parameter that reflects the effect of hygiene measures after November
+2020 at work and in educational facilities. This parameter reduces infectiousness of
+contacts by one third. In total those are ten parameters.\comment[id=K]{I think here you
+refer to the other multipliers but you don't mention them. Also, since there is no change on the 12th of April it's only nine changes} The breakpoints the contact
+reduction changes are not determined from data but from announcements of policy changes.
+Moreover, we constrain the estimated contact reduction to follow the shape of the
+stringency index. The resulting contact reduction can be seen in
+Figure~\ref{fig:other_multiplier}
 
@@
 
-% The infection probabilities estimated from our model are as shown in
-% Table~\ref{tab:infection_probs}.\comment[id=K]{Missing to add data for the other
-% estimated parameters (import of b117 etc)}\comment[id=K]{Update table after the new
-% estimation is finished}
-
-% \begin{table}[tb]
-%     \caption{Estimated Infection Probabilities}
-%     \label{tab:infection_probs}
-%     \centering
-%     \begin{tabular}{ll}
-%         \toprule
-%         Contact Type & Infection Probability \\
-%         \midrule
-%         preschool and nursery & 00.500\% \\
-%         school & 01.200\% \\
-%         household & 10.000\% \\
-%         work & 14.500\% \\
-%         other & 15.875\% \\
-%         \bottomrule
-%     \end{tabular}
-%     \tablenotes{}
-% \end{table}
+\begin{table}[htb]
+    \centering
+    \caption{Estimated Parameters}
+    \label{tab:estimated_params}
+    \input{../bld/param_tables/estimates}
+\end{table}

### 2021-07-10 src/paper/report.tex
Add first version of estimated parameters table.

@@
         of contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~\ref{subsec:policies_seasonality} for separate plots of the three
+        Appendix~\ref{subsec:policies} for separate plots of the three
         factors by contact type.}
@@
 contribute less to the spread of the disease than others due to a lower number of
-contacts see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
+contacts see Supplementary Material~\ref{subsec:assortativity}. By the end of our
 study period, when first-dose vaccination rates reached around 40\% of the population,

### 2021-07-10 src/parameter_tables/task_convert_csv_to_latex.py

@@ -13,6 +13,14 @@ PARAMETRIZATION = [(p, BLD / "param_tables" / f"{p.stem}.tex") for p in SRC.iter
 def task_convert_csv_to_tex(depends_on, produces):
     df = pd.read_csv(depends_on)
     df = df.dropna(axis=0, how="all")  # drop empty rows
+    if len(df.columns) == 5:
+        col_format = r"p{4cm}p{2cm}p{2cm}p{2cm}p{9cm}"
+        float_format = "%.0f"
+    elif len(df.columns) == 3:
+        col_format = r"p{3cm}p{1.5cm}p{9cm}"
+        float_format = "%.3f"
+    else:
+        raise ValueError(f"Unexpected number of columns: {len(df.columns)}")
     with open(produces, "w") as f:
         with pd.option_context("max_colwidth", 10000):
             table = df.to_latex(
@@ -20,8 +28,8 @@ def task_convert_csv_to_tex(depends_on, produces):
                 escape=False,
                 longtable=False,
                 index=False,
-                column_format=r"p{4cm}p{2cm}p{2cm}p{2cm}p{9cm}",
-                float_format="%.0f",
+                column_format=col_format,
+                float_format=float_format,
             )
         table = table.replace(r"\begin{tabular}", r"\begin{tabular*}{\textwidth}")
         table = table.replace(r"\end{tabular}", r"\end{tabular*}")

### 2021-07-10 src/paper/data/policies.tex
Start on extending the model fit section.

@@
 \FloatBarrier
-
-\input{paper/rapid_test_params}
-
-All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
-    \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
-    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
-    planning to have education contacts, work contacts, developing symptoms without
-    access to a PCR test, having a household member with a positive test or symptoms. In
-    each case whether a rapid test is done depends on how long it has been since the
-    individual's last rapid test and her individual compliance parameters. As an example,
-    take a worker in May. In that time workers are encouraged to test themselves twice
-    weekly but there is no general requirement to test themselves. If the worker has not
-    done a test within the last four days in our model she will demand a test if her
-    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
-    \label{fig:rapid_test_demand}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \caption{Share of Individuals With Rapid Tests}
-    \label{fig:share_ever_rapid_test}
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-        \caption{Share of Individuals That Have Ever Done a Rapid Test}
-    \end{subfigure}%
-    \begin{subfigure}{.55\textwidth}
-        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
-    \end{subfigure}
-    \label{fig:share_rapid_test_last_week}
-    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
-        have ever done a rapid test or done a rapid test within the last week in our
-        simulations to the shares reported in the
-        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
-        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
-        have ever done a rapid test. The right panel compares the share of individuals
-        who have done a rapid test within the last seven days in our simulation compared
-        to the share reporting to have done at least weekly rapid tests in the last four
-        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
-        conservative. The overall share is below that in the study. We fit the share of
-        weekly tests quite exactly. However, the study only covers adults while our share
-        also includes children who are tested very regularly when attending school.}
-\end{figure}
-
- \comment[id=K]{J: Explain that we don't fit the share ever tested that well because our rapid test compliance is completely fixed. }
-
-\FloatBarrier

### 2021-07-10 src/paper/results/fit.tex
Start on extending the model fit section.

@@
 
-This compares simulated data from our model with empirical data from Germany. We look at
-observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
-rapid test demands. Where available we do not only look at aggregated statistics but also
-analyze the model fit for age groups and federal states.\comment[id=J]{summarize the fit}
+This section compares simulated data from our model with empirical data from Germany. We
+look at observed infections, fatality rates,\comment[id=K]{should we include them?} the
+spread of the B.1.1.7 mutation, vaccinations\comment[id=K]{We fit vaccinations in the
+population by construction. Maybe this should go to a different section?} and rapid test
+demand. Where available we do not only look at aggregated statistics but also analyze the
+model fit for age groups and federal states.
 
-\begin{figure}[ht]   % new known case with single runs
+% summary of the fit
+Overall, our model achieves an excellent fit of the two waves of infections with few free
+parameters (Figure~\ref{fig:aggregated_fit2}. As a result the effective replication
+number is similar to that reported by the Robert Koch-Institut (see
+Figure~\ref{fig:fit_r_effective}). This excellent fit is also achieved for most age
+groups in Germany. The fit is also good for many German federal states even though our
+synthetic population does not represent the differences in the distribution of age groups
+and they only had a small weight in the optimization. Despite, the fact that the number
+of performed rapid tests and their distribution in the population results endogenously
+inside our model we fit the share of the population with at least a weekly rapid test
+very well and err on the side of too few individuals that have ever done a rapid test.
+
+% observed infections
+
+\begin{figure}[ht]   % observed infections with single runs
   \centering
@@
 
- \comment[id=K]{More text in this subsection. For \ref{fig:age_group_fit}: Explain why we
- don't fit the 80 to 100 year olds}
+\textcolor{red}{...} High levels of infections often coincide with increased statistical
+uncertainty, as for example in April and May.
 
-\begin{figure}[ht]  % new known case by age group
+\begin{figure}[ht]  % observed infections by age group
   \centering
@@
 
- \comment[id=K]{Explain differences between simulated and actual cases in figure
- \ref{fig:state_fit}. (Our states are uniform with respect to their age distribution.}
+\textcolor{red}{...} One important exception are the 80 to 100 year olds where we do not
+fit the large number of cases in the winter of 2020/2021. This is because we do not model
+nursing homes where a large fraction of this age group lives and a very large fraction of
+cases in that age group occurred.
 
-\begin{figure}[ht]   % new known case by federal state
+\begin{figure}[ht]   % observed infections by federal state
   \centering
@@
 
-\begin{figure}[ht]   % fit of vaccinations
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
-  \caption{Share of Vaccinated Individuals}
-  \floatfoot{\noindent \textit{Note:} The figure shows the rate of individuals that are
-  vaccinated in our synthetic population versus in the general German population. Note
-  that we excluded the vaccinations that were given to nursing homes, approximately the
-  first percent of the German population that were vaccinated. Overall, our model covers
-  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
-  the population are vaccinated. Our vaccinations work imperfectly but we do not model
-  different vaccines nor do we distinguish between first and second shot.}
-  \label{fig:fit_vaccinations}
-\end{figure}
+Our states are uniform with respect to their age distribution. \textcolor{red}{...}
+
+Our fit of the effective replication number $R_t$ closely follows the values reported by
+the RKI. \textcolor{red}{ldots}
 
@@
 
+% fatality rates: MISSING
+
 \begin{figure}[ht]   % Share B.1.1.7
@@
 
+\begin{figure}[ht]   % fit of vaccinations
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
+  \caption{Share of Vaccinated Individuals}
+  \floatfoot{\noindent \textit{Note:} The figure shows the rate of individuals that are
+  vaccinated in our synthetic population versus in the general German population. Note
+  that we excluded the vaccinations that were given to nursing homes, approximately the
+  first percent of the German population that were vaccinated. Overall, our model covers
+  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
+  the population are vaccinated. Our vaccinations work imperfectly but we do not model
+  different vaccines nor do we distinguish between first and second shot.}
+  \label{fig:fit_vaccinations}
+\end{figure}
+
+\begin{figure}[ht]     % rapid test demand
+    \centering
+    \caption{Share of Individuals With Rapid Tests}
+    \label{fig:share_ever_rapid_test}
+    \begin{subfigure}{.55\textwidth}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+        \caption{Share of Individuals That Have Ever Done a Rapid Test}
+    \end{subfigure}%
+    \begin{subfigure}{.55\textwidth}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
+        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
+    \end{subfigure}
+    \label{fig:share_rapid_test_last_week}
+    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
+        have ever done a rapid test or done a rapid test within the last week in our
+        simulations to the shares reported in the
+        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
+        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
+        have ever done a rapid test. The right panel compares the share of individuals
+        who have done a rapid test within the last seven days in our simulation compared
+        to the share reporting to have done at least weekly rapid tests in the last four
+        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
+        conservative. The overall share is below that in the study. We fit the share of
+        weekly tests quite exactly. However, the study only covers adults while our share
+        also includes children who are tested very regularly when attending school.}
+\end{figure}
+
 \FloatBarrier

### 2021-07-10 src/paper/data/rapid_test_demand.tex
Continue extending the fit section.

--- 2021-07-09 src/paper/rapid_test_params.tex
@@
 individuals would do a rapid test due to a private reason.
+
+All shares of individuals who would take a rapid test if the conditions were met can be
+seen in Figure~\ref{fig:rapid_test_demand}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
+    \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
+    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
+    planning to have education contacts, work contacts, developing symptoms without
+    access to a PCR test, having a household member with a positive test or symptoms. In
+    each case whether a rapid test is done depends on how long it has been since the
+    individual's last rapid test and her individual compliance parameters. As an example,
+    take a worker in May. In that time workers are encouraged to test themselves twice
+    weekly but there is no general requirement to test themselves. If the worker has not
+    done a test within the last four days in our model she will demand a test if her
+    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
+    \label{fig:rapid_test_demand}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \caption{Share of Individuals With Rapid Tests}
+    \label{fig:share_ever_rapid_test}
+    \begin{subfigure}{.55\textwidth}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
+        \caption{Share of Individuals That Have Ever Done a Rapid Test}
+    \end{subfigure}%
+    \begin{subfigure}{.55\textwidth}
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
+        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
+    \end{subfigure}
+    \label{fig:share_rapid_test_last_week}
+    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
+        have ever done a rapid test or done a rapid test within the last week in our
+        simulations to the shares reported in the
+        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
+        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
+        have ever done a rapid test. The right panel compares the share of individuals
+        who have done a rapid test within the last seven days in our simulation compared
+        to the share reporting to have done at least weekly rapid tests in the last four
+        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
+        conservative. The overall share is below that in the study. We fit the share of
+        weekly tests quite exactly. However, the study only covers adults while our share
+        also includes children who are tested very regularly when attending school.}
+\end{figure}
+
+ \comment[id=K]{J: Explain that we don't fit the share ever tested that well because our rapid test compliance is completely fixed. }
+
+\FloatBarrier

### 2021-07-10 src/paper/results/fit.tex
Continue extending the fit section.

@@
 % observed infections
+Our fit of the infection rates in Germany between October 2020 and June 2021 is very
+good. The incidence in our model follows the shape of the reported incidence almost
+perfectly. The increase in December is somewhat less steep than it was. This can be
+attributed to the large number of cases in nursing homes during that time that our model
+cannot capture because it does not include nursing homes. It is very noticeable that high
+levels of infections often coincide with increased statistical uncertainty, as for
+example in April and May.
 
@@
 
-\textcolor{red}{...} High levels of infections often coincide with increased statistical
-uncertainty, as for example in April and May.
+% observed infections by age group
+Zooming into the different age groups in Figure~\ref{fit:age_group_fit}, we can see that
+our model is also able to reproduce the infection rates on this level. Differences
+between the age groups allow us to identify the infection probabilities because different
+age groups have contacts of different types (e.g. Mostly children go to school, the share
+of workers is highest in the 35 to 59 age bracket etc.). There are three exceptions to
+our fit. Firstly, the age group of 80 to 100 year olds due to the absence of nursing
+homes in our synthetic population. Secondly, the increase in adolescents in April. It is
+surprising that we do not match this increase despite explicitly modeling the
+introduction of rapid tests for students. However, it can be expected that every rapid
+test in a school setting is followed up by a PCR test while only 82\% of our population
+\citep{Betsch2021} demand a confirming PCR test when receiving a positive rapid test. The
+last age group where our fit does not fit the empirical curve very well is the group of
+15 to 34 year olds. This group has been touted to have a very active social life and it
+has often been suspected that this age group also stayed more socially active during the
+pandemic. Since our model does not include either effect it's unsurprising that the
+actual case numbers in this group are consistently higher than our simulated case
+numbers.\comment[id=K]{Does anyone have an idea why we do not match the shape from Oct to
+Dec?}
 
@@
 
-\textcolor{red}{...} One important exception are the 80 to 100 year olds where we do not
-fit the large number of cases in the winter of 2020/2021. This is because we do not model
-nursing homes where a large fraction of this age group lives and a very large fraction of
-cases in that age group occurred.
+% observed infections by federal state
+Looking at another disaggr
+Our states are uniform with respect to their age distribution.
 
@@
 \end{figure}
-
-Our states are uniform with respect to their age distribution. \textcolor{red}{...}
 

### 2021-07-10 src/paper/results/fit.tex
Temp.

@@
 
+%%% WRITE MORE POSITIVELY
+
 This section compares simulated data from our model with empirical data from Germany. We
-look at observed infections, fatality rates,\comment[id=K]{should we include them?} the
-spread of the B.1.1.7 mutation, vaccinations\comment[id=K]{We fit vaccinations in the
-population by construction. Maybe this should go to a different section?} and rapid test
-demand. Where available we do not only look at aggregated statistics but also analyze the
-model fit for age groups and federal states.
+look at observed infections, the spread of the B.1.1.7 mutation,
+vaccinations\comment[id=K]{We fit vaccinations in the population by construction. Maybe
+this should go to a different section?} and rapid test demand. Where available we do not
+only look at aggregated statistics but also analyze the model fit for age groups and
+federal states.
 
@@
 parameters (Figure~\ref{fig:aggregated_fit2}. As a result the effective replication
-number is similar to that reported by the Robert Koch-Institut (see
-Figure~\ref{fig:fit_r_effective}). This excellent fit is also achieved for most age
-groups in Germany. The fit is also good for many German federal states even though our
-synthetic population does not represent the differences in the distribution of age groups
-and they only had a small weight in the optimization. Despite, the fact that the number
-of performed rapid tests and their distribution in the population results endogenously
-inside our model we fit the share of the population with at least a weekly rapid test
-very well and err on the side of too few individuals that have ever done a rapid test.
+number is similar to that reported by the RKI (see Figure~\ref{fig:fit_r_effective}).
+This excellent fit is also achieved for most age groups in Germany. The fit is also good
+for many German federal states. Despite, the fact that the number of performed rapid
+tests and their distribution in the population results endogenously inside our model we
+fit the share of the population with at least a weekly rapid test very well and err on
+the side of too few individuals that have ever done a rapid test.
 
@@
 good. The incidence in our model follows the shape of the reported incidence almost
-perfectly. The increase in December is somewhat less steep than it was. This can be
-attributed to the large number of cases in nursing homes during that time that our model
-cannot capture because it does not include nursing homes. It is very noticeable that high
-levels of infections often coincide with increased statistical uncertainty, as for
-example in April and May.
+perfectly. % The increase in December is somewhat less steep than it was. This can be
+% attributed to the large number of cases in nursing homes during that time that our
+% model cannot capture because it does not include nursing homes. \comment[id=K]{Write
+% about uncertainty and that we take enough simulations to have a stable mean}
 
@@
 % observed infections by age group
-Zooming into the different age groups in Figure~\ref{fit:age_group_fit}, we can see that
+Zooming into the different age groups in Figure~\ref{fig:age_group_fit}, we can see that
 our model is also able to reproduce the infection rates on this level. Differences
@@
 age groups have contacts of different types (e.g. Mostly children go to school, the share
-of workers is highest in the 35 to 59 age bracket etc.). There are three exceptions to
-our fit. Firstly, the age group of 80 to 100 year olds due to the absence of nursing
-homes in our synthetic population. Secondly, the increase in adolescents in April. It is
-surprising that we do not match this increase despite explicitly modeling the
-introduction of rapid tests for students. However, it can be expected that every rapid
-test in a school setting is followed up by a PCR test while only 82\% of our population
-\citep{Betsch2021} demand a confirming PCR test when receiving a positive rapid test. The
-last age group where our fit does not fit the empirical curve very well is the group of
-15 to 34 year olds. This group has been touted to have a very active social life and it
-has often been suspected that this age group also stayed more socially active during the
-pandemic. Since our model does not include either effect it's unsurprising that the
-actual case numbers in this group are consistently higher than our simulated case
-numbers.\comment[id=K]{Does anyone have an idea why we do not match the shape from Oct to
-Dec?}
+of workers is highest in the 35 to 59 age bracket etc.). There are only three
+\textcolor{red}{\ldots}. Firstly, the age group of 80 to 100 year olds due to the absence
+of nursing homes in our synthetic population. Secondly, the increase in adolescents in
+April. It is surprising that we do not match this increase despite explicitly modeling
+the introduction of rapid tests for students. However, it can be expected that every
+rapid test in a school setting is followed up by a PCR test while only 82\% of our
+population \citep{Betsch2021} demand a confirming PCR test when receiving a positive
+rapid test. The last age group where our fit does not fit the empirical curve very well
+is the group of 15 to 34 year olds. This group has been touted to have a very active
+social life and it has often been suspected that this age group also stayed more socially
+active during the pandemic. Since our model does not include either effect it's
+unsurprising that the actual case numbers in this group are consistently higher than our
+simulated case numbers.\comment[id=K]{Does anyone have an idea why we do not match the
+shape from Oct to Dec?}
 
@@
     \caption{Share of Individuals With Rapid Tests}
-    \label{fig:share_ever_rapid_test}
+    \label{fig:share_ever_rapid_test2}
     \begin{subfigure}{.55\textwidth}
@@
     \end{subfigure}
-    \label{fig:share_rapid_test_last_week}
+    \label{fig:share_rapid_test_last_week2}
     \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who

### 2021-07-10 src/paper/data/share_known_cases.tex
Improve the model fit text and add paragraphs on R_t and vaccinations.

--- 2021-07-09 src/paper/results/share_known_cases.tex
@@
-\subsection{Share of Cases that are Detected}
-\label{subsec:appendix_share_known_cases}
+\subsection{Share of Detected Cases}
+\label{subsec:data_share_known_cases}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
-  \caption{Share of Detected Cases by Age Group}
-  \label{fig:share_known_cases_by_age_group}
+  \includegraphics[width=\textwidth]{figures/results/figures/data/testing/assumed_overall_share_known_cases}
+  \caption{Share of Detected Cases}
+  \label{fig:share_known_cases_data}
   \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
@@
   then assume it to be constant as vaccinations of the elderly strongly affect the case
-  fatality rate which the project does not account for. To get from an overall share of
-  detected cases to the share of cases that is detected in each age group we use that
-  asymptomatic cases are much less likely to be detected. As our model covers age
-  specific asymptomatic rates this endogenously leads to group specific share known cases
-  that verify that infections in younger age groups are under-detected. Starting in 2021
-  in addition to the overall numbers of detected cases through symptoms and the share
-  known cases, cases are also detected through confirmation of positive rapid tests. This
-  leads to an increase in the share of known cases for all age groups but in particular
-  for the younger age groups that are covered extensively with rapid tests through the
-  rapid test requirement for participating in school.}
+  fatality rate which the project does not account for. Starting in 2021 in addition to
+  the overall numbers of detected cases through symptoms and the share known cases, cases
+  are also detected through confirmation of positive rapid tests which happens
+  endogenously inside the model. For the public holidays of Christmas and Easter we lower
+  the share of detected cases as fewer PCR tests are available during public holidays.}
 \end{figure}
 
-It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
-
 \FloatBarrier

### 2021-07-10 src/paper/results/fit.tex
Improve the model fit text and add paragraphs on R_t and vaccinations.

@@
 
-%%% WRITE MORE POSITIVELY
-
 This section compares simulated data from our model with empirical data from Germany. We
-look at observed infections, the spread of the B.1.1.7 mutation,
-vaccinations\comment[id=K]{We fit vaccinations in the population by construction. Maybe
-this should go to a different section?} and rapid test demand. Where available we do not
-only look at aggregated statistics but also analyze the model fit for age groups and
-federal states.
+look at observed infections (overall as well as by age group and federal state), the
+spread of the B.1.1.7 mutation, vaccinations\comment[id=K]{We fit vaccinations in the
+population by construction. Maybe this should go to a different section?} and rapid test
+demand.
 
@@
 for many German federal states. Despite, the fact that the number of performed rapid
-tests and their distribution in the population results endogenously inside our model we
+tests and their distribution in the population result endogenously inside our model we
 fit the share of the population with at least a weekly rapid test very well and err on
@@
 % observed infections
-Our fit of the infection rates in Germany between October 2020 and June 2021 is very
-good. The incidence in our model follows the shape of the reported incidence almost
-perfectly. % The increase in December is somewhat less steep than it was. This can be
-% attributed to the large number of cases in nursing homes during that time that our
-% model cannot capture because it does not include nursing homes. \comment[id=K]{Write
-% about uncertainty and that we take enough simulations to have a stable mean}
+Our fit of the infection rates in Germany between October 2020 and June 2021 is
+excellent. The incidence in our model matches both the levels and the shape of the
+reported incidence almost perfectly. When the prevalence of the virus is high and during
+phases of exponential growth, the effect of random events on the incidence is large. That
+is why it's important that we rely on 30 simulation runs with different seed sequences to
+average over these random differences and have a stable mean.
 
@@
   \caption{Fit Over the Full Simulation Time Frame with Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported simulated infections rates. The mean infection rate is
-  the thick blue line. Single simulation runs are plotted in lighter and thinner lines.
-  The official case numbers as reported by the Robert-Koch-Institut are plotted in black.
-  The fit is overall very good. The higher the mean incidence and the stronger the growth
-  the more variance there is between simulation runs. We averaged over 30 simulation
-  runs.}
+  \floatfoot{\noindent \textit{Note:} The figure shows the daily incidence rate per
+  million for the reported simulated infections rates. The mean infection rate is the
+  thick blue line. Single simulation runs are plotted in lighter and thinner lines. The
+  official case numbers as reported by the RKI are plotted in black. The fit is overall
+  very good. The higher the mean incidence and the stronger the growth the more variance
+  there is between simulation runs. We averaged over 30 simulation runs.}
   \label{fig:aggregated_fit2}
@@
 age groups have contacts of different types (e.g. Mostly children go to school, the share
-of workers is highest in the 35 to 59 age bracket etc.). There are only three
-\textcolor{red}{\ldots}. Firstly, the age group of 80 to 100 year olds due to the absence
-of nursing homes in our synthetic population. Secondly, the increase in adolescents in
-April. It is surprising that we do not match this increase despite explicitly modeling
-the introduction of rapid tests for students. However, it can be expected that every
-rapid test in a school setting is followed up by a PCR test while only 82\% of our
-population \citep{Betsch2021} demand a confirming PCR test when receiving a positive
-rapid test. The last age group where our fit does not fit the empirical curve very well
-is the group of 15 to 34 year olds. This group has been touted to have a very active
-social life and it has often been suspected that this age group also stayed more socially
-active during the pandemic. Since our model does not include either effect it's
+of workers is highest in the 35 to 59 age bracket etc.). There are very few points where
+our simulated incidences and the empirical incidences do not
+coincide:\comment[id=K]{@Janos: I think you can probably present this better...} Firstly,
+the age group of 80 to 100 year olds due to the absence of nursing homes in our synthetic
+population. Secondly, the increase in adolescents in April. We model the introduction of
+rapid tests for students and do see an increase both in the observed infections and the
+share of detected cases among students (see
+Figure~\ref{fig:share_known_cases_by_age_group}). However, we follow \cite{Betsch2021}
+that 82\% of all rapid tests are followed up with a PCR test. It is likely that that
+share is higher for rapid tests that were done in schools. The last age group where our
+fit does not fit the empirical curve very well is the group of 15 to 34 year olds. This
+group has been touted to have a very active social life and it has often been suspected
+that this age group also stayed more socially active during the pandemic. However, our
+model does not include either effect because of very little data on both and hence it's
 unsurprising that the actual case numbers in this group are consistently higher than our
@@
 
+\FloatBarrier
+
 % observed infections by federal state
-Looking at another disaggr
-Our states are uniform with respect to their age distribution.
+Our model fit is also very good for the different German federal states. This holds not
+only for the large states such as North Rhine-Westphalia or Bavaria but also for many
+smaller states such as Hessen or Rhineland-Palatinate. This shows that the inclusion of
+school vacations and relying on work reductions by \cite{Google2021} combined with the
+assortativity of contacts on the county level is sufficient to represent many local
+differences.  %
+However, admittedly, there are several patterns that our model does not reproduce - for
+example because we do not have cross-border travel nor population density. This explains
+why our model produces too many infections for federal states such as Schleswig-Holstein
+and why it predicts too few infections for states, such as Saxony, bordering high
+incidence countries.\comment[id=K]{The Czech Republic had very high numbers in October
+(max > 800) but they fell during Nov to 220 and then climbed again to 800 by Jan 12th so
+that might be only part of the answer why Saxony had such high numbers.}
+However, these missing features are irrelevant for our main conclusions.
 
@@
 
+\FloatBarrier
+
 Our fit of the effective replication number $R_t$ closely follows the values reported by
-the RKI. \textcolor{red}{ldots}
+the RKI even though we calculate $R_t$ on all infected individuals not just the detected
+cases. This explains why the $R_t$ in our simulations is higher during phases where the
+share of detected cases falls such as in the fall of 2020 (see
+Figure~\ref{fig:share_known_cases_data}) where the RKI underestimated the effective
+replication number due to observing a lower share of cases and is lower than the $R_t$
+reported by the RKI in spring where the share of known cases increased due to increased
+rapid testing.
 
@@
 
-% fatality rates: MISSING
+% B.1.1.7
+We fit the proliferation of the B.1.1.7 variant quite exactly despite only introducing a
+few cases in January as can be seen in Figure~\ref{fig:fit_share_b117}. Since we only
+model B.1.1.7 and do not include other variants B.1.1.7 reaches a share of nearly 100\%
+by May while the true rate plateaued at 90\% and B.1.1.7 was then replaced by B.1.617.2
+as the dominant variant in June. However, B.1.1.7 still made up over 90\% of all cases in
+Germany at the end of our simulation period.
 
@@
 \end{figure}
+
+% vaccinations
+
+The fit of the share of vaccinated individuals can be seen in
+Figure~\ref{fig:fit_vaccinations}. In Germany, vaccines were rolled according to four
+priority groups. The first vaccines were mostly reserved for nursing homes and some
+selected professions such as first responders. Since we do not have nursing home
+inhabitants in our model, we subtract the first percent of vaccinations which is
+equivalent to the share of Germans living in nursing homes. Afterwards, the share of
+vaccinated individuals in the population follows the German increase quite exactly. We
+took great care to model the prioritization of older individuals and professions that
+cannot reduce physical contact easily such as teachers or medical staff (see
+Section~\ref{subsec:synthetic_population}. How the share of vaccinated individuals
+develops by age group can be seen in Figure~\ref{fig:vaccinations_by_age_group}.
 
@@
 \end{figure}
+
+% rapid test demand
 

### 2021-07-10 src/paper/results/fit.tex
Finish first version of the model fit section.

@@
 % rapid test demand
+The most difficult moment to match in our model is the rapid test demand. Since we have
+five different channels through which individuals demand rapid tests and many of the
+demand curves are at least partially calibrated through survey data. It is therefore very
+reassuring that we fit the share of individuals that do weekly rapid tests almost
+perfectly. For the share of individuals that have ever done a rapid test our model is
+conservative. There are two reasons for this: Firstly, we only introduce rapid tests in
+2021. This is due to the small number of rapid tests in 2020 as well as a lack of data on
+the number of random tests and who received them. Secondly, there is no randomness in an
+individual's rapid test demand. This reduces the randomness in our model at the cost of
+having less individuals that only occasionally test themselves. \comment[id=K]{Write that
+this makes us underestimate the effect of random tests?} However,
+Section~\ref{subsec:model_validation} shows that our results are very robust to changes
+in the exact shares of individuals demanding rapid tests.
 

### 2021-07-10 src/paper/data/share_known_cases.tex
Improve the fit text. Add text to the data section of the share known cases and improve its figure.

@@
 \label{subsec:data_share_known_cases}
+
+ \comment[id=K]{Stoye called this the ascertainment rate if I remember right. Maybe we should replace this everywhere}
+
+One important feature of our model is that we distinguish between undetected and detected
+cases and that we model which cases are detected and which are not (see \ref{sub:testing}
+for a detailed description for how we model both rapid and PCR tests). For our model it
+is important to have an estimate for the share of cases that is detected in the absence
+of rapid tests. For this we rely on the
+\href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} which relies on estimates
+of the case fatality rate to estimate the number of total cases given the number of
+deaths which are suspected to be perfectly observable. For 2020, we follow the reported
+share of detected cases quite closely only interpolating the phase of November 2020 where
+the testing policy was changed to give more priority to vulnerable groups. This lead to
+more tests being done among older age groups which are also at the highest risk of dying
+from CoViD-19 and an overestimation of the share of detected cases.
+% After Christmas 2020
+Similarly, the start of vaccinations after Christmas 2020 that were predominantly given
+to nursing homes in the beginning and other vulnerable groups in spring, led to a
+decrease in the number of deadly cases, leading to an increasing upward bias in the share
+of detected cases. This is why we stop using the Dunkelzifferradar after Christmas.
+Instead, we assume that that the share of detected cases would have stayed the same in
+the absence of rapid tests. Thus, we also achieve in our model an increase in the share
+of detected cases but this is driven from inside our model through increased testing and
+follow-ups on positive rapid tests (see Section~\ref{subsec:results_share_known_cases}).
+% Christmas and Easter
+Lastly, we include reductions in the share known cases due to the two major holidays in
+our simulation period, Christmas and Easter. During both holidays where many laboratories
+do not process tests and most physicians' offices are closed, less PCR tests were done
+leading to short and large drops in the share of known cases.
 

### 2021-07-10 src/paper/data/synthetic_population.tex
Improve the fit text. Add text to the data section of the share known cases and improve its figure.

@@
 
-% To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
-% put adolescents and children into two groups that follow after the general population.
-% These groups do not become eligible within our simulation frame until June.
+To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
+put adolescents and children into two groups that follow after the general population.
+These groups do not become eligible within our simulation frame until June.
 
-% \begin{figure}[ht]   % vaccinations by age group
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
-%   \caption{Simulated and Empirical Infections by Federal State}
-%   \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
-%   \label{fig:vaccinations_by_age_group}
-% \end{figure}
+\begin{figure}[ht]   % vaccinations by age group
+  \centering
+  \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
+  \caption{Simulated and Empirical Infections by Federal State}
+  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
+  \label{fig:vaccinations_by_age_group}
+\end{figure}
 
 \FloatBarrier
-os

### 2021-07-10 src/paper/results/fit.tex
Improve the fit text. Add text to the data section of the share known cases and improve its figure.

@@
 only for the large states such as North Rhine-Westphalia or Bavaria but also for many
-smaller states such as Hessen or Rhineland-Palatinate. This shows that the inclusion of
-school vacations and relying on work reductions by \cite{Google2021} combined with the
-assortativity of contacts on the county level is sufficient to represent many local
-differences.  %
-However, admittedly, there are several patterns that our model does not reproduce - for
-example because we do not have cross-border travel nor population density. This explains
-why our model produces too many infections for federal states such as Schleswig-Holstein
-and why it predicts too few infections for states, such as Saxony, bordering high
-incidence countries.\comment[id=K]{The Czech Republic had very high numbers in October
-(max > 800) but they fell during Nov to 220 and then climbed again to 800 by Jan 12th so
-that might be only part of the answer why Saxony had such high numbers.}
-However, these missing features are irrelevant for our main conclusions.
+smaller states such as Hessen or Rhineland-Palatinate. This shows that using school
+vacations dates and work mobility reductions by \cite{Google2021} at the state level
+combined with county and age group specific initial conditions (see
+Section~\ref{sub:initial_conditions}) and county level assortativity of contacts is
+sufficient to represent many local differences.  %
+The fit is especially good given that our model does not aim to have a high local
+resolution. For example we abstract from population density and cross-border travel. It
+is, thus, unsurprising that there are states that we do not match well, such as very
+thinly populated Mecklenburg-Vorpommern and Schleswig-Holstein or Saxony with its large
+border to the Czech Republic that had a much higher incidence than
+Germany.\comment[id=K]{The Czech Republic had very high numbers in October (max > 800)
+but they fell during Nov to 220 and then climbed again to 800 by Jan 12th so that might
+be only part of the answer why Saxony had such high numbers.}
 

### 2021-07-10 src/paper/results/share_known_cases.tex
Fix copy paste errors.
Add model valiadation section. Restructure results.
Improve the model fit text and add paragraphs on R_t and vaccinations.
Add section to the group specific share known cases results.

@@
-\subsection{Share of Cases that are Detected}
-\label{subsec:appendix_share_known_cases}
+\subsection{Share of Detected Cases}
+\label{subsec:results_share_known_cases}
+
+We explain how we model the detection of cases in Section~\ref{sub:testing} and show the
+overall share of known cases that we use as input for our testing model in
+Section~\ref{subsec:data_share_known_cases}.
+% 2020
+Given the differential demand by symptomatic and asymptomatic individuals our model
+endogenously results in group specific share known cases that are shown in
+Figure~\ref{fig:share_known_cases_by_age_group}. Since older individuals are much more
+likely to develop symptoms many more cases among them are detected.
+% 2021
+Starting in 2021 the share known cases not only depends on the overall share known cases
+we feed into our model but also on the number of performed rapid tests. As more and more
+(often asymptomatic and presymptomatic) individuals receive positive rapid tests and
+receive confirmation, the share of detected cases starts to increase, especially after
+March when rapid tests became much more widely available in Germany. This effect is very
+visible for school pupils, (5-14, green line). As twice weekly rapid tests become
+mandatory in schools and students start going to school after the Easter break, the share
+of detected cases for that group increases drastically from 25\% to over 60\% between
+April and June.
 
@@
 It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
+fourteen year olds. This is a direct result of the mandatory tests in school.
 

### 2021-07-10 src/paper/model/estimated_parameters.tex
Temp.
Section on estimated parameters is ready for review.

@@
-\subsection{Estimated Parameters}\comment[id=K]{Philipp suggests to write more about
-identification. Both in general and to take seasonality as an example of how we identify
-both calibrated and estimated parameters. }
+\subsection{Estimated Parameters}
 \label{subsec:estimated_params}
-
-%%% The free parameters are the infection probabilities by contact category, one hygiene
-%%% multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
-%%% that govern the reduction of other contacts\comment[id=K]{Must be updated when the
-%%% estimation is finished.}, the number of extra contacts during holidays, the share of
-%%% detected cases around holidays and the fade in speed of the rapid tests as well as one
-%%% parameter that governs the import of B.1.1.7 cases in January 2021.
 
@@
 
-We fit our model to data for Germany from October 2020 until June 2021. We do not use
-earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
-limited and therefore it would be difficult to find good initial conditions for our
-simulations. In addition during the summer the case numbers were extremely low. This
-could lead to the epidemic going extinct in our simulation. Additionally, our model does
-not include international travel or other imports of cases. These would be important but
-difficult to model during the summer months.
-
-% infection probabilities
-
-To avoid over-fitting and simplify the numerical optimization problem, we only allow for
-five different infection probabilities: 1) for contacts in schools 2) contacts in
-preschools and nurseries. 3) for work contacts. 4) for households. 5) for other
-contacts.\comment[id=J]{Write that the exact numbers are hard to interpret and show the
-infections by channel heatmap more prominently than the infection
-probabilities}\comment[id=K]{Compare to
-\href{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3867682}{these estimates}}
-
-% policies
-We also estimate a parameter that reflects the effect of hygiene measures after November
-2020 at work and in educational facilities. This parameter reduces infectiousness of
-contacts by one third. In total those are ten parameters.\comment[id=K]{I think here you
-refer to the other multipliers but you don't mention them. Also, since there is no change on the 12th of April it's only nine changes} The breakpoints the contact
-reduction changes are not determined from data but from announcements of policy changes.
-Moreover, we constrain the estimated contact reduction to follow the shape of the
-stringency index. The resulting contact reduction can be seen in
-Figure~\ref{fig:other_multiplier}
-
-% b117 rate
-Finally we estimate one parameter that governs the introduction of the B.1.1.7 virus
-variant in January 2021. This parameter implies that at the end of January roughly one
-case per 100 000 individuals per day is imported. After January we do not model imported
-cases of B.1.1.7 anymore because they are negligible compared to the endogenous growth
-of that virus variant.
+All estimated parameters are described in Table~\ref{tab:estimated_params}.
 
@@
 \end{table}
+
+We fit our model to data for Germany from October 2020 until June 2021. We do not use
+earlier periods for three reasons. Firstly, in the beginning PCR tests were very scarce
+and the reported case numbers unreliable. Secondly, during the summer the case
+numbers were extremely low. This could lead to the epidemic going extinct in our
+simulation. Thirdly, over the summer, imported cases from touristic travel were likely
+important for the infection dynamic but there is not enough data to include them into
+our model.
+
+% infection probabilities
+To avoid over-fitting and simplify the numerical optimization problem, we only allow for
+five different infection probabilities: 1) for contacts in schools 2) contacts in
+preschools and nurseries. 3) for work contacts. 4) for households. 5) for other
+contacts.
+
+Since the infectiousness of a contact between an infectious and a susceptible
+person depends on many things, the numerical values of the infection probabilities in
+Table~\ref{tab:estimated_params} only reflect a base probability. This
+base probability is modified by a seasonality factor, an age specific
+susceptiblity factor and an infectiousness factor that depends on the virus strand
+of the infected person. The base infection probability is only equal to the actual
+infection probability when all of those factors are 1. This would be the
+case for a contact between an 80+ year old susceptible person with a person who is
+infected with the B.1.1.7 strand of the virus on January first.
+
+It is not possible to rank different types of contacts according to their
+infectiousness just from the numerical values of the infection probabilities.
+There are two reasons for this: Firstly, for computational reasons the
+seasonality factor is normalized such that it reaches 1 at its peak. It has thus
+a lower average for contact types with strong seasonality (e.g. other contacts) than for
+contact types with weak seasonality (e.g. work contacts). Secondly, for household and
+school contacts we do not have data on whether people actually have physical contact.
+Thus the infection probabilities for those contact types are actually the product of the
+probability to actually have physical contact on a given day and the infection
+probability of that contact.
+
+In order to get a feeling for the infectiousness of each contact type it
+is more intuitive to look at how many infections were actually caused by each contact
+type. This is depicted in Figure~\ref{fig:infection_heatmap}. We can see that work and
+other contacts are the main driver of the pandemic, followed by infections in
+households. Schools and preschools contribute fewer infections which is to be expected
+given that there are much fewer students than working adults in the German population.
+Nevertheless, Figure~\ref{fig:school_scenarios_detailed} shows that schools do have a
+notable effect on the infection dynamic in the long run.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/infection_channels.png}
+    \caption{\textbf{Daily share of infections by contact type}}
+    \floatfoot{\noindent \textit{Note:} Daily hare of infections that were contributed
+    by each contact type. Darker colors mean that a larger share of infections were
+    contributed by that contact type. The majority of infections take place in the
+    workplace, in households and via other contacts. Schools and preschools contribute
+    less infections, especially after hygiene measures have been introduced.}
+    \label{fig:infection_heatmap}
+\end{figure}
+
+We also estimate a parameter that reflects the effect of hygiene measures
+at work and in educational facilities. This parameter becomes active in November 2020
+when stricter mask mandates and distancing rules were introduced. It is estimated to
+reduce infectiousness of contacts by one third.
+
+Moreover, we estimate nine different multipliers that reflect how strongly other
+contacts are reduced over time. The dates at which we switch between the multipliers
+usually coincides with policy changes and is not determined from the data. The only
+exception to this are slight adjustments to parameters to incorporate lockdown fatigue
+(towards the end of a lockdown period) or precautionary contact reductions (in times
+of high incidences right before a lockdown is enacted).
+
+While we estimate 9 different values for the other contact multiplier, they are not
+estimated completely freely. In particular we ensure that the ordering of the parameter
+values is consistent with the stringency of policies. For example, the strongest
+contact reduction was estimated for January 2021 during where very strict measures
+and curfews were in place, whereas the weakest contact reduction was in October 2020
+where policies were very lenient.
+
+Since we do not have good data on the reduction of other contacts, it is not possible
+to separately estimate parameters for contact reduction and the effect of hygiene
+measures. The reported other multipliers in Table~\ref{tab:estimated_params} are thus
+a combination of contact reduction and hygiene measures.
+
+Finally we estimate one parameter that governs the introduction of the B.1.1.7 virus
+variant in January 2021. This parameter implies that at the end of January roughly one
+case per 100 000 individuals per day is imported. After January we do not model imported
+cases of B.1.1.7 anymore because they are negligible compared to the endogenous growth
+of that virus variant.
+
+While a formal identification argument is beyond the scope of this paper, below we give
+a rough intuition which features of the data help us to estimate each parameter.
+
+The different infection probabilities can be separately identified because the degree to
+which each contact type is active varies over time (e.g. school closures, vacations and
+different work from home policies) and they affect different subgroups of the population
+differently (e.g. $\beta_{school}$ most strongly affects kids whereas
+$\beta_{work}$ has the strongest effect on adults in working age and $\beta_{other}$
+affects all age groups equally).
+The hygiene and other multipliers can be identified because they are only active in
+certain time periods. However, it is necessary to normalize one other multiplier to
+1 because there is no period without any contact reduction in our data.
+The introduction parameter for the B.1.1.7 mutation can be identified from the share of
+that virus strand in the population. The rapid test
+demand parameters are identified because rapid tests first lead to a very steep increase
+in observed cases and then to a sudden decrease -- in a time where almost all other
+things in the model would not cause a change in trend.

### 2021-07-10 src/paper/data/rapid_test_demand.tex

--- 2021-07-02 src/paper/rapid_test_params.tex
@@
 test in the round of 23rd of February 2021 when an acquaintance would have tested
-positive. Since this is only asking for willingness not actual behavior and the demand
-when meeting with friends is very likely lower, we take this as the upper bound of
-private rapid test demand which is reached on May 4th. To cover that many people are
-likely to have sought and done their first rapid test before the Easter holidays to meet
-friends or family, we let the share of individuals doing rapid tests in that time
-increase more rapidly than before and after. By end of March 25\% of individuals would do
-a rapid test due to a private reason.
+positive. Since this is only asking for willingness not actual behavior, we take this as
+the upper bound of private rapid test demand which is reached in the beginning of May. To
+cover that many people are likely to have sought and done their first rapid test before
+the Easter holidays to meet friends or family, we let the share of individuals doing
+rapid tests in that time increase more rapidly than before and after. By end of March
+25\% of individuals would do a rapid test due to a private reason.
+
+All shares of individuals who would take a rapid test if the conditions were met can be
+seen in Figure~\ref{fig:rapid_test_demand}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
+    \caption{Share of Individuals Doing a Rapid Test.}
+    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
+    planning to have education contacts, work contacts, developing symptoms without
+    access to a PCR test, having a household member with a positive test or symptoms. In
+    each case whether a rapid test is done depends on how long it has been since the
+    individual's last rapid test and her individual compliance parameters. As an example,
+    take a worker in May. In that time workers are encouraged to test themselves twice
+    weekly but there is no general requirement to test themselves. If the worker has not
+    done a test within the last four days in our model she will demand a test if her
+    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
+    \label{fig:rapid_test_demand}
+\end{figure}
+
+\FloatBarrier

### 2021-07-10 src/paper/data/share_known_cases.tex

@@
 One important feature of our model is that we distinguish between undetected and detected
-cases and that we model which cases are detected and which are not (see \ref{sub:testing}
-for a detailed description for how we model both rapid and PCR tests). For our model it
-is important to have an estimate for the share of cases that is detected in the absence
-of rapid tests. For this we rely on the
+cases and that we model which cases are detected and which are not (see
+Section~\ref{sub:testing} for a detailed description for how we model both rapid and PCR
+tests). For our model it is important to have an estimate for the share of cases that is
+detected in the absence of rapid tests. For this we rely on the
 \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} which relies on estimates
@@
 more tests being done among older age groups which are also at the highest risk of dying
-from CoViD-19 and an overestimation of the share of detected cases.
-% After Christmas 2020
+from CoViD-19 and an overestimation of the share of detected cases.% After Christmas 2020
 Similarly, the start of vaccinations after Christmas 2020 that were predominantly given

### 2021-07-10 src/paper/model/testing.tex

--- 2021-06-21 src/paper/model/testing.tex
@@
 those PCR tests are made in addition to the tests that would have been done otherwise.
-Section~\ref{subsec:appendix_share_known_cases} discusses the effect of rapid tests on
-the share of detected cases.
+Section~\ref{subsec:results_share_known_cases} discusses the effect of rapid tests on
+the share of detected cases.\comment[id=K]{Talk here about reactions to tests and
+symptoms with contact reductions. (If not here, find other place where to talk about this
+in detail.)}

### 2021-07-10 src/paper/report.tex

@@
 date. Schools and daycare centers were closed, so were customer-facing businesses except
-for grocery and drug stores. From the peak of the second wave just before Christmas
-until the trough in mid-February, newly detected cases decreased by almost three
-quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 strain,
-which became dominant in March (Figure~\ref{fig:share_b117}). In early March, some NPIs
-were being relaxed; e.g., hairdressers and home improvement stores were allowed to open
-again to the public. There were many changes in details of regulations afterwards, but
-they did not change the overall stringency index.
+for grocery and drug stores. From the peak of the second wave just before Christmas until
+the trough in mid-February, newly detected cases decreased by almost three quarters. The
+third wave in the spring of 2021 is associated with the B.1.1.7 strain, which became
+dominant in March (Figure~\ref{fig:share_b117}) and stayed dominant for our entire
+simulation period.\footnote{B.1.617.2 (Delta) reached Germany in April but still
+accounted for less than 5\% of cases at the end of our simulation period.}
+In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
+stores were allowed to open again to the public. There were many changes in details of
+regulations afterwards, but they did not change the overall stringency index.
 

### 2021-07-10 src/paper/data/policies.tex
Move plot of other multiplier to estimated parameter section.

@@
 
-For the last group of contacts, which cover things like leisure activities, grocery
-shopping, etc., we have no reliable data by how much policies reduce them. In addition,
-they are likely to be affected by social and psychological factors such as pandemic
-fatigue and vacations. Because of this we estimate them like the infection probabilities
-to fit the time series data. We use very few change points and tie them to particular
-events such as policy announcements or particular holidays. Because of the scarce data
-situation we cannot distinguish between a hygiene factor (such as mask wearing) during
-meetings and physical distancing (such as virtual meetings with
-friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
-which phases we have and why the switching points are where they are.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/other_multiplier}
-    \caption{Share of Pre-Pandemic Other Contacts Taking Place with Infection Potential}
-    \label{fig:other_multiplier}
-    \floatfoot{\noindent \textit{Note:} All values are estimated. We try to use as little
-    switching points as possible and tie them to political events (such as lockdown
-    announcements) unless changes are used to capture anticipation or pandemic fatigue
-    (for example we model an anticipation of the November lockdown and model lockdown
-    fatigue in early March).}
-\end{figure}
-
 \FloatBarrier

### 2021-07-10 src/paper/model/estimated_parameters.tex
Move plot of other multiplier to estimated parameter section.

@@
 (towards the end of a lockdown period) or precautionary contact reductions (in times
-of high incidences right before a lockdown is enacted).
+of high incidences right before a lockdown is enacted). The estimated other multipliers
+are also depicted in Figure~\ref{fig:other_multiplier}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/data/other_multiplier}
+    \caption{Share of Pre-Pandemic Other Contacts Taking Place with Infection Potential}
+    \label{fig:other_multiplier}
+    \floatfoot{\noindent \textit{Note:} Values of the other multiplier. All values are
+    estimated via the method of simulated moments. The rationale behind each switching
+    point is described in Table~\ref{tab:estimated_params}}
+\end{figure}
 

### 2021-07-11 src/paper/data/synthetic_population.tex
Add figure description in the synthetic population section.

@@
 the same individuals that continue to have work contacts when work from home mandates of
-a certain strictness are in place. Since our work contact priority is uniformly normal
-from zero to one, we can use the reduction in work mobility as reported by
+a certain strictness are in place. We use the reduction in work mobility reported by
 \cite{Google2021} as a proxy for the share of individuals that do not have work contacts.
@@
   \includegraphics[width=\textwidth]{figures/results/figures/vaccinations/spring_baseline}
-  \caption{Simulated and Empirical Infections by Federal State}
-  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written.}}
+  \caption{Vaccination Rates by Age Group}
+  \floatfoot{\noindent \textit{Note:} An individual's vaccination priority depends on her
+  work contact priority, her age group and a random component to capture medical
+  preconditions like diabetes. 15\% of the population refuse to be vaccinated.
+  Adolescents would be vaccinated after the general population and children last. The
+  figure clearly shows that the first vaccinations go to some workers with very high work
+  contact priority and to the 80 to 100 age group followed by the 60 to 79 year olds.
+  Both groups are saturated with vaccinations by mid March and start of May respectively.
+  By June a third of the younger adults have received the vaccination but these groups
+  still remain far from herd immunity thresholds.}
   \label{fig:vaccinations_by_age_group}

### 2021-07-11 src/paper/data/assortativity.tex
Join single plots and make single plots smaller plus various small improvements.

@@
 
-As mentioned in section \ref{sec:matching}, the probability that two individuals are
+As explained in section \ref{sec:matching}, the probability that two individuals are
 matched can depend on background characteristics. In particular, we allow this
@@
 geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from the same data
-we use to calibrate the number of contacts.
+within the same county, we can calibrate the assortative mixing by age from
+\cite{Mossong2008}.
 
@@
     \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_other_non_recurrent}
-    \caption{Distribution of Non Recurrent Other Contacts by Age Group}
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/data/assortativity_other_non_recurrent}
+        \caption{{Distribution of Non Recurrent Other Contacts by Age Group}}
         \label{fig:assortativity_other}
+    \end{subfigure}
+    \hfill
+    % work assortativity
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_work_non_recurrent}
+        \caption{{Distribution of Non Recurrent Work Contacts by Age Group}}
+        \label{fig:assortativity_work}
+    \end{subfigure}
+
+    \vskip3ex
+
+    \caption{Assortativity by Age Group for Non Recurrent Other and Work Contacts}
     \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
-        recurrent other contacts by age group. A row shows the share of contacts a
-        certain age group has with all other age groups. Higher values are colored in
-        darker red tones. The diagonal represents the share of contacts with individuals
-        from the same age group.}
+        recurrent contacts by age group for other contacts on the left and work contacts
+        on the right. A row shows the share of contacts a certain age group has with all
+        other age groups. Higher values are colored in darker red tones. The diagonal
+        represents the share of contacts with individuals from the same age group. The
+        80-100 age group for other contacts was so small that we assumed for them to have
+        the same contact distribution as the 70-79 year olds. For work contacts, we only
+        show age groups that have a significant fraction of working individuals. The
+        matrix must not be symmetric because the groups have very different sizes.}
+
+    \label{fig:assortativity}
+
 \end{figure}
 
-Figure~\ref{fig:assortativity_other} shows that assortativity by age is especially strong
-for children and younger adults. For older people, the pattern becomes more dispersed
-around their own age group, but within-age-group contacts are still the most common
-contacts.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/data/assortativity_work_non_recurrent}
-    \caption{Distribution of Random Work Contacts by Age Group}
-    \label{fig:assortativity_work}
-    \floatfoot{\noindent \textit{Note:} The figure shows the distribution of non
-        recurrent work contacts by age group. A row shows the share of contacts a certain
-        age group has with all other age groups. Higher values are colored in darker red
-        tones. The diagonal represents the share of contacts with individuals from the
-        same age group. We only show age groups that have a significant fraction of
-        working individuals.}
-\end{figure}
-
-Figure~\ref{fig:assortativity_work} shows that assortativity by age is also important
-among work contacts.
+Figure~\ref{fig:assortativity_other} shows that assortativity of the other contacts by
+age is especially strong for children and adolescents. For older people, the pattern
+becomes more dispersed around their own age group, but within-age-group contacts are
+still the most common contacts. Figure~\ref{fig:assortativity_work} shows that
+assortativity by age is also important among work contacts.
 

### 2021-07-11 src/paper/data/policies.tex
Join single plots and make single plots smaller plus various small improvements.

@@
 
-In our empirical application\comment[id=K]{Improve language} we distinguish four groups
-of contact types: households, education, work and other contacts.% households
-For households we assume that the individuals'
-contacts in their households do not change over our estimation period.
+Our policies usually affect one of our four contact types: households, education, work
+and other contacts.
+% households
+For households we assume that the individuals' contacts in their households do not change
+over our estimation period.
 % educ models
@@
 federal states as well as school closures, emergency care and A / B schooling where only
-one half of students attends every other week or day. For the moment we ignore that lack
-of childcare leads working parents to stay home. An approximation of the share of contacts still taking place with the different school regulations can be found in Figure~\ref{fig:school_multiplier}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/school_multiplier_comparison}
-    \caption{School Multiplier With and Without Vacations Factored In}
-    \floatfoot{\noindent \textit{Note:} The dates on which schools have vacation are
-    decided at the federal level. Vacations are directly implemented in our model with no
-    school contacts taking place on weekends and during vacations (by federal state) just
-    like the schooling mode (full operation, emergency care, rotating schemes with half
-    class sizes etc.). The figure is thus only an illustration that roughly shows the
-    share of contacts taking place compared to pre-pandemic level with and without
-    vacations. The difference between the lines show when vacations take place and to
-    what degree. For example all states have fall vacations but the timing varies
-    strongly between states.}
-    \label{fig:school_multiplier}
-\end{figure}
-
-%
-% Schließung von Kindertagesstätten und Schulen: 37,4 Millionen ausgefallene Arbeitstage
-% http://www.iab-forum.de/schul-und-kitaschliessungen-krankheit-quarantaene-die-coronabedingten-arbeitsausfaelle-der-erwerbstaetigen-steigen-auf-592-millionen-arbeitstage/
-%
-%
-% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377: In
-% allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
-% Schüler in die Schule.
-% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430
-% 18% der Grundschüler in Notbetreuung in BW
-%
-
+one half of students attends every other week or day.
+% For the moment we ignore that lack of childcare leads working parents to stay home.
+An approximation of the share of contacts still taking place with the different school
+regulations can be found in Figure~\ref{fig:school_multiplier}.
 % work
-For our work contacts we use the reductions in work mobility reported in the Google
+For work contacts we use the reductions in work mobility reported in the Google
 Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in work
@@
 changes the threshold below which workers stay home. Figure \ref{fig:work_multiplier}
-shows the share of workers that go to work in our model over time.
+shows the share of workers that go to work over time at the federal German level. We use
+the data on the state level.
+
+In addition, for both work and school contacts we assume that hygiene measures (such as
+masks, ventilation and hand washing) became more strict and more conscientiously observed
+in November, leading to a reduction of 33\% in the number of contacts with the potential
+to transmit Covid-19.
 
@@
     \centering
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/data/school_multiplier_comparison}
+        \caption{{School Attendance}}
+        \label{fig:school_multiplier}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
         \includegraphics[width=\textwidth]{figures/results/figures/data/work_multiplier_since_sep}
-    \caption{Share of Workers with Work Contacts}
+        \caption{Work Attendance}
         \label{fig:work_multiplier}
-    \floatfoot{\noindent \textit{Note:} The figure shows the work mobility as reported by
-        \cite{Google2021}. We take this as a proxy of the share of workers who are not in
-        home office, i.e. who still have physical work contacts. The figure interpolates
-        over weekends as we handle weekend effects through information on work on
-        weekends in the German census data we use. The figure shows the share aggregated
-        over Germany as a whole. To capture the effect that local policies, school
-        vacations and public policies have on work contacts we use the data on the level
-        of the federal states to determine which workers go to work depending on the
-        state they live in.}
+    \end{subfigure}
+
+    \vskip3ex
+
+    \caption{The Effect of School and Work Policies on the Number Contacts Taking Place}
+
+    \floatfoot{\noindent \textit{Note:} The left figure shows the approximate share of
+    school contacts taking place with and without vacations factored in. The dates on
+    which schools have vacation are decided at the state level. Vacations are directly
+    implemented in our model with no school contacts taking place on weekends and during
+    vacations (by federal state) just like the schooling mode (full operation, emergency
+    care, rotating schemes with half class sizes etc.). The figure is thus only an
+    illustration that shows the approximate share of contacts taking place compared to
+    the pre-pandemic level with and without vacations. The difference between the lines
+    show when vacations take place and to what degree. For example all states have fall
+    vacations but the timing varies strongly between states while all states close
+    schools during the Christmas holidays. The right figure shows the work mobility as
+    reported by \cite{Google2021}. We take this as a proxy of the share of workers who
+    still have physical work contacts. The figure interpolates over weekends as we handle
+    weekend effects through information on work on weekends in the German census data we
+    use. The figure shows the share for Germany as a whole. To capture the effect that
+    local policies, school vacations, etc. have on work contacts we use the data on the
+    state level to determine which workers go to work depending on the state they live
+    in.}
+    \label{fig:multipliers}
 \end{figure}
 
-For both work and school contacts we assume that starting November with the lockdown
-light in Germany, hygiene measures (such as masks, ventilation and hand washing) became
-more strict and more conscientiously observed, leading to a reduction of 33\% in the
-number of contacts with the potential to transmit Covid-19.
+Lastly, for the other contacts category we did not calibrate them from data but estimated
+the policy effects on them. Refer to Section~\ref{subsec:estimated_params} and Figure~\ref{fig:other_multiplier} in particular.
 

### 2021-07-11 src/paper/results/fit.tex
Join single plots and make single plots smaller plus various small improvements.

@@
 look at observed infections (overall as well as by age group and federal state), the
-spread of the B.1.1.7 mutation, vaccinations\comment[id=K]{We fit vaccinations in the
-population by construction. Maybe this should go to a different section?} and rapid test
-demand.
+effective replication number,  the spread of the B.1.1.7 mutation,
+vaccinations\comment[id=K]{We fit vaccinations in the population by construction. Maybe
+this should go to a different section?} and rapid test demand.
 
@@
   \centering
+  \begin{subfigure}[b]{0.425\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
-  \caption{Fit Over the Full Simulation Time Frame with Single Simulation Runs}
-  \floatfoot{\noindent \textit{Note:} The figure shows the daily incidence rate per
-  million for the reported simulated infections rates. The mean infection rate is the
+    \caption{Observed Incidence in the Model and as Reported by the RKI}
+    \label{fig:aggregated_fit2}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.425\textwidth}
+    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
+    \caption{Effective Replication Number $R_t$ in the Model and as Reported by the RKI}
+    \label{fig:fit_r_effective}
+  \end{subfigure}
+
+  \caption{Model Fit of the Reported Cases and the Effective Replication Number}
+  \label{fig:incidence_and_r_effective}
+
+  \floatfoot{\noindent \textit{Note:} Both figures show averages and single runs. The
+  average is the thick line. Single runs are shown as lighter and thinner lines. We
+  averaged and show 30 simulation runs. The left figure shows the daily incidence rate
+  per million for the reported simulated infections rates. The mean infection rate is the
   thick blue line. Single simulation runs are plotted in lighter and thinner lines. The
@@
   very good. The higher the mean incidence and the stronger the growth the more variance
-  there is between simulation runs. We averaged over 30 simulation runs.}
-  \label{fig:aggregated_fit2}
+  there is between simulation runs. The right figure shows the effective replication
+  number ($R_t$) as reported by the RKI and as calculated in our model. The $R_t$ gives
+  the average number of new infections caused by one infected individual. The $R_t$ in
+  our model broadly follows the $R_t$ reported by the RKI. Two trends stand out. Firstly,
+  the RKI's $R_t$ drops faster in November. This could be due to a change in the testing
+  policy that focused tests on the elderly when the second wave hit Germany and led to a
+  decline in the overall share of detected cases. The second difference is from mid
+  February to mid March where the RKI's reported $R_t$ increased more rapidly than that
+  in our model. Here the opposite effect can be expected. During this time rapid tests
+  increased strongly leading to more cases being detected. In the short term this leads
+  an $R_t$ estimation that is based on detected cases to overestimate the replication
+  number.}
 \end{figure}
+
+% R effective
+Our fit of the effective replication number $R_t$ closely follows the values reported by
+the RKI (see Figure~\ref{fig:fit_r_effective} even though we calculate $R_t$ on all
+infected individuals not just the detected cases. This explains why the $R_t$ in our
+simulations is higher during phases where the share of detected cases falls such as in
+the fall of 2020 (see Figure~\ref{fig:share_known_cases_data}) where the RKI
+underestimated the effective replication number due to observing a lower share of cases
+and is lower than the $R_t$ reported by the RKI in spring where the share of known cases
+increased due to increased rapid testing.
+
+\FloatBarrier
 
@@
 
-Our fit of the effective replication number $R_t$ closely follows the values reported by
-the RKI even though we calculate $R_t$ on all infected individuals not just the detected
-cases. This explains why the $R_t$ in our simulations is higher during phases where the
-share of detected cases falls such as in the fall of 2020 (see
-Figure~\ref{fig:share_known_cases_data}) where the RKI underestimated the effective
-replication number due to observing a lower share of cases and is lower than the $R_t$
-reported by the RKI in spring where the share of known cases increased due to increased
-rapid testing.
-
-\begin{figure}[ht]   % R effective
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_r_effective_with_single_runs}
-  \caption{Effective Replication Number $R_t$ in the Model and as Reported by the
-  Robert-Koch-Institute}
-  \floatfoot{\noindent \textit{Note:} The figure shows the effective replication number
-  ($R_t$) as reported by the RKI and as calculated in our model. The $R_t$ gives the
-  average number of new infections caused by one infected individual. The $R_t$ in our
-  model broadly follows the $R_t$ reported by the RKI. Two trends stand out. Firstly, the
-  RKI's $R_t$ drops faster in November. This could be due to a change in the testing
-  policy that focused tests on the elderly when the second wave hit Germany and led to a
-  decline in the overall share of detected cases. The second difference is from mid
-  February to mid March where the RKI's reported $R_t$ increased more rapidly than that
-  in our model. Here the opposite effect can be expected. During this time rapid tests
-  increased strongly leading to more cases being detected. In the short term this leads
-  an $R_t$ estimation that is based on detected cases to overestimate the replication
-  number.}
-  \label{fig:fit_r_effective}
-\end{figure}
-
 % B.1.1.7
@@
 
-\begin{figure}[ht]   % Share B.1.1.7
+\begin{figure}[ht]
+  \centering
+
+  \begin{subfigure}[b]{0.425\textwidth}   % Share B.1.1.7
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117_with_single_runs}
-  \caption{Share of B.1.1.7 in the Model and as Reported by the Robert-Koch-Institute}
-  \floatfoot{\noindent \textit{Note:} The figure shows the share of B.1.1.7 as
+    \caption{Share of B.1.1.7 in the Model and as Reported by the RKI}
+    \label{fig:fit_share_b117}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.425\textwidth}    % Vaccinations
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
+    \caption{Share of Vaccinated Individuals in the Model and the German Population}
+    \label{fig:fit_vaccinations}
+  \end{subfigure}
+  \caption{Model Fit of the Share of B.1.1.7 and Vaccinations}
+  \floatfoot{\noindent \textit{Note:} The left figure shows the share of B.1.1.7 as
   reported by the RKI and as calculated in our model. We only introduce a few cases over
   the cause of January. From then B.1.1.7 takes over endogenously through its increased
-  infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases per 100,000 inhabitants.}
-  \label{fig:fit_share_b117}
+  infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases
+  per 100,000 inhabitants. The right figure shows the rate of individuals that are
+  vaccinated in our synthetic population versus in the general German population. Note
+  that we excluded the vaccinations that were given to nursing homes, approximately the
+  first percent of the German population that were vaccinated. Overall, our model covers
+  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
+  the population are vaccinated. Our vaccinations work imperfectly but we do not model
+  different vaccines nor do we distinguish between first and second shot.}
 \end{figure}
@@
 % vaccinations
-
 The fit of the share of vaccinated individuals can be seen in
@@
 
-\begin{figure}[ht]   % fit of vaccinations
-  \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_ever_vaccinated}
-  \caption{Share of Vaccinated Individuals}
-  \floatfoot{\noindent \textit{Note:} The figure shows the rate of individuals that are
-  vaccinated in our synthetic population versus in the general German population. Note
-  that we excluded the vaccinations that were given to nursing homes, approximately the
-  first percent of the German population that were vaccinated. Overall, our model covers
-  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
-  the population are vaccinated. Our vaccinations work imperfectly but we do not model
-  different vaccines nor do we distinguish between first and second shot.}
-  \label{fig:fit_vaccinations}
-\end{figure}
+\FloatBarrier
 
@@
     \label{fig:share_ever_rapid_test2}
-    \begin{subfigure}{.55\textwidth}
+    \begin{subfigure}{.425\textwidth}
         \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
-        \caption{Share of Individuals That Have Ever Done a Rapid Test}
+        \caption{Share of the Population That Has Ever Done a Rapid Test}
     \end{subfigure}%
-    \begin{subfigure}{.55\textwidth}
+    \hfill
+    \begin{subfigure}{.425\textwidth}
         \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
-        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
+        \caption{Share of the Population That Did a Rapid Test in the Last Week}
     \end{subfigure}

### 2021-07-11 src/paper/data/synthetic_population.tex
Abbreviate Robert Koch Institute and introduce abbreviation at the first mention.

@@
 group and a vaccination rank from that group that creates a complete vaccination queue
-over the population including 15\% that refuse to be vaccinated \citep{RKI2021}. The
-vaccination groups are created to match the recommendations by the Ständige
-Impfkommission \citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized
-more than recommended by the commission.}\comment[id=K]{Write how that plays out in our simulations for the different age groups.}
+over the population including 15\% that refuse to be vaccinated as reported by the Robert
+Koch Institute (RKI, \cite{RKI2021}). The vaccination groups are created to match the
+recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
+that teachers were prioritized more than recommended by the
+commission.}\comment[id=K]{Write how that plays out in our simulations for the different
+age groups.}
 

### 2021-07-11 src/paper/model/testing.tex
Abbreviate Robert Koch Institute and introduce abbreviation at the first mention.

@@
 \url{https://covid19.dunkelzifferradar.de/}} Next, we determine how many of these tests
-should go to symptomatic and asymptomatic individuals from data by the Robert Koch
-Institut. \comment[id=J]{add source}
-Then, we sample the individuals to which those tests are allocated from the pools of
-symptomatic and asyptomatic infected but not yet tested individuals.
+should go to symptomatic and asymptomatic individuals from data by the
+RKI.\comment[id=J]{add source} Then, we sample the individuals to which those tests are
+allocated from the pools of symptomatic and asyptomatic infected but not yet tested
+individuals.
 

### 2021-07-11 src/paper/data/medical_parameters.tex
Move data text to different text files. Arange number of contacts figure.
Abbreviate Robert Koch Institute and introduce abbreviation at the first mention.
urls to references.

--- 2021-06-22 src/paper/medical_parameters.tex
@@
 % Length of the presymptomatic stage
-The \href{https://bit.ly/3pi18Ax}{European Centre for Disease Prevention and Control}
-reports that people become infectious between one and two days before symptoms
-start.\footnotemark
+The European Centre for Disease Prevention and Control \citep{ECDC2021} reports that
+people become infectious between one and two days before symptoms start.\footnotemark
 
@@
 For the time from symptom onset until need for intensive care we rely on data by the US
-CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
+CDC \citep{Stokes2020} and the OpenABM Project \citep{Hinch2021}.
+
 %Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of
@@
 %
-%    \tablenotes{
-%        The data is taken from \citet{Stokes2020} and \href{https://bit.ly/354pKTZ}{the
-%        OpenABM-Project}. }
+%    \tablenotes{The data is taken from \citet{Stokes2020} and citet{Hinch2021}.}
 %
@@
 \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also
-in line with the duration estimates collected by \href{https://bit.ly/3gkGtaU}{the
-Robert-Koch-Institut}. We assume that the time between symptom onset and ICU takes 4, 6,
-8 or 10 days with equal probabilities. As we do not model nursing homes, do not focus on
+in line with the duration estimates collected by the Robert Koch Institute
+\citep{RKI2021b}. We assume that the time between symptom onset and ICU takes 4, 6, 8 or
+10 days with equal probabilities. As we do not model nursing homes, do not focus on
 matching deaths and do not use the number of individuals in intensive care to estimate
@@
 We take the survival probabilities and time to death and time until recovery from
-intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
-time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
-match this approximately we discretize that 41\% of individuals who will die from
-Covid-19 do so after one day in intensive care, 22\% day after 12 days, 29\% after 20
-days and 7\% after 32 days. Again, we rescale this for every age group among those that
-will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
-reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
-days. We discretize this such that of those who recover in intensive care, 22\% do so
-after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.
+intensive care from \citet{Hinch2021}. They report time until death to have a mean of
+11.74 days and a standard deviation of 8.79 days. To match this approximately we
+discretize that 41\% of individuals who will die from Covid-19 do so after one day in
+intensive care, 22\% day after 12 days, 29\% after 20 days and 7\% after 32 days. Again,
+we rescale this for every age group among those that will not survive. For survivors the
+OpenABM Project reports a mean duration of 18.8 days until recovery and a standard
+deviation of 12.21 days. We discretize this such that of those who recover in intensive
+care, 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45
+days.
+
+\FloatBarrier

### 2021-07-11 src/paper/data/share_known_cases.tex
Join single plots and make single plots smaller plus various small improvements.
urls to references.

@@
 detected in the absence of rapid tests. For this we rely on the
-\href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} which relies on estimates
+\citet{Dunkelzifferradar2020} which relies on estimates
 of the case fatality rate to estimate the number of total cases given the number of
@@
 decrease in the number of deadly cases, leading to an increasing upward bias in the share
-of detected cases. This is why we stop using the Dunkelzifferradar after Christmas.
+of detected cases. This is why we stop using the \citet{Dunkelzifferradar2020} after Christmas.
 Instead, we assume that that the share of detected cases would have stayed the same in
@@
   cases that was estimated through the case fatality ratio by the
-  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
+  \citet{Dunkelzifferradar2020} for all of 2020 and
   then assume it to be constant as vaccinations of the elderly strongly affect the case

### 2021-07-11 src/paper/data/synthetic_population.tex
urls to references.

--- 2021-07-10 src/paper/data/synthetic_population.tex
@@
 group and a vaccination rank from that group that creates a complete vaccination queue
-over the population including 15\% that refuse to be vaccinated \citep{RKI2021}. The
+over the population including 15\% that refuse to be vaccinated \citep{RKI2021c}. The
 vaccination groups are created to match the recommendations by the Ständige

### 2021-07-11 src/paper/model/testing.tex
urls to references.

@@
 number of newly infected individuals with an estimate of the share of detected cases from
-the Dunkelzifferradar project. \footnote{The Dunkelzifferradar project publishes daily
-estimates of the dark figure of infections under
-\url{https://covid19.dunkelzifferradar.de/}} Next, we determine how many of these tests
-should go to symptomatic and asymptomatic individuals from data by the
-RKI.\comment[id=J]{add source} Then, we sample the individuals to which those tests are
-allocated from the pools of symptomatic and asyptomatic infected but not yet tested
-individuals.
+the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how many of these tests
+should go to symptomatic and asymptomatic individuals from data by the Robert Koch
+Institute. \comment[id=J]{add source}
+Then, we sample the individuals to which those tests are allocated from the pools of
+symptomatic and asymptomatic infected but not yet tested individuals.
 
@@
 At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
-of detected cases from the Dunkelzifferradar can no longer be used because it is based on
+of detected cases from \citet{Dunkelzifferradar2020} can no longer be used because it is based on
 case fatality rates which drastically change due to vaccinations. Secondly, rapid tests

### 2021-07-11 src/paper/results/fit.tex
urls to references.

--- 2021-07-10 src/paper/results/fit.tex
@@
         have ever done a rapid test or done a rapid test within the last week in our
-        simulations to the shares reported in the
-        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
-        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
+        simulations to the shares reported in \citet{Betsch2021b}. The left panel compares the share of individuals who
         have ever done a rapid test. The right panel compares the share of individuals

### 2021-07-11 src/paper/results/share_known_cases.tex
Join single plots and make single plots smaller plus various small improvements.
urls to references.

@@
   reported as an official case via PCR confirmation. We use the overall share of known
-  cases that was estimated through the case fatality ratio by the
-  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
+  cases that was estimated through the case fatality ratio by \citet{Dunkelzifferradar2020} for all of 2020 and
   then assume it to be constant as vaccinations of the elderly strongly affect the case

### 2021-07-11 src/paper/data/assortativity.tex
Make sections A.1 to A.4 ready for review.

@@
 probability to depend on age and county of residence. While we do not have good data on
-geographical assortativity and just roughly calibrate it such that 80\% of contacts are
-within the same county, we can calibrate the assortative mixing by age from
-\cite{Mossong2008}.
+geographical assortativity and set it such that 80\% of contacts are within the same
+county, we can calibrate the assortativity by age from \cite{Mossong2008}.
 
@@
         the same contact distribution as the 70-79 year olds. For work contacts, we only
-        show age groups that have a significant fraction of working individuals. The
-        matrix must not be symmetric because the groups have very different sizes.}
+        show age groups that have a significant fraction of working individuals.}
 
@@
 
-Our two other types of contacts, households and schools, get their assortativity by
-construction. Schools are groups where the same children of the mostly same age and
-county meet with teachers every day. Household composition follows directly from the
-German microcensus data we use to construct our synthetic population.
+For recurrent contacts, we constructed groups to have the following features: Recurrent
+work contacts are not assortative by age. Daily work groups are always of the same county
+and weekly work contacts are to 80\% with workers from the same county. Other recurrent
+contacts are constructed the same way but we impose for daily contacts that they are
+always with individuals from the same age group. School classes are groups where the same
+children of the mostly same age and county meet with teachers every day. Nurseries and
+preschools mix children by age but match them to come mostly from the same county.
+Household age composition follows directly from the German microcensus data we use to
+construct our synthetic population.
 

### 2021-07-11 src/paper/data/medical_parameters.tex
Make sections A.1 to A.4 ready for review.

--- 2021-06-10 paper/medical_parameters.tex
@@
 This section discusses the medical parameters used in the model, their sources and how we
-arrived at the distributions used in the model.\footnotemark See
-Figure~\ref{fig:disease_progression} for a summary of our disease progression model.
-
-\footnotetext{Additional information can be found in the
+arrived at the distributions used in the model.\footnote{Additional information can be
+found in the
 \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-documentation}.}
+documentation}.} See Figure~\ref{fig:model_disease_progression} for a summary of our
+disease progression model.
 
@@
 The first medical parameter we need is the length of the period between infection and the
-start of infectiousness, the so called latent period. We infer it from two other measures
-that are more common in the medical literature: Firstly, the time between infection and
-the onset of symptoms, the incubation period. Secondly, the time between the start of
-infectiousness and the onset of symptoms. We assume that the latency period is the same
-for symptomatic and asymptomatic individuals.
+start of infectiousness, the so called latent period. \cite{Zhao2021} estimate the latent
+period to last 3.3 days (95\% CI: 0.2, 7.9). In line with this estimate our latent period
+lasts one to five days with a slightly lower average of 2 days.
 
@@
 % Length of the presymptomatic stage
-The \href{https://bit.ly/3pi18Ax}{European Centre for Disease Prevention and Control}
-reports that people become infectious between one and two days before symptoms
-start.\footnotemark
-
-\footnotetext{This is similar to \citet{He2020} and in line with \citet{Peak2020}.}
-
-Taking these estimates together, we arrive at a latent period of one to five days.
+Similar to \citet{Peak2020} and in line with \citet{He2020} we set the length of the
+presymptomatic stage to be one to two days.
 
@@
 seems to be the time after which most people are culture negative, as also reported by
-\citet{Singanayagam2020}.} Thus, we arrive at 0 to 13 days as the range for
+\citet{Singanayagam2020}.} Thus, we arrive at zero to 13 days as the range for
 infectiousness among individuals who become symptomatic (see also figure 5).
@@
 
-For the time from symptom onset until need for intensive care we rely on data by the US
-CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
-%Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of
-%requiring intensive care per age group.
-%
-%\begin{table}[tb]
-%    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
-%    \label{tab:symptomatic-to-ICU}
-%    \centering
-%
-%    \begin{tabular}{ll}
-%        \toprule
-%        Age Group & Share \\
-%        \midrule
-%        0-9 & 0.00005 \\
-%        10-19 & 0.00030 \\
-%        20-29 & 0.00075 \\
-%        30-39 & 0.00345 \\
-%        40-49 & 0.01380 \\
-%        50-59 & 0.03404 \\
-%        60-69 & 0.10138 \\
-%        70-79 & 0.16891 \\
-%        80-100 & 0.26871 \\
-%        \bottomrule
-%    \end{tabular}
-%
-%    \tablenotes{
-%        The data is taken from \citet{Stokes2020} and \href{https://bit.ly/354pKTZ}{the
-%        OpenABM-Project}. }
-%
-%\end{table}
+For the time from symptom onset until need for intensive care we rely on data by
+\cite{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
 
 For those who will require intensive care we follow \citet{Chen2020} who estimate the
-time from symptom onset to ICU admission as 8.5 $\pm$ 4 days. This aligns well with
+time from symptom onset to ICU admission as $8.5 \pm 4$ days. This aligns well with
 numbers reported for the time from first symptoms to hospitalization:
-\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of 4. This is also
-in line with the duration estimates collected by \href{https://bit.ly/3gkGtaU}{the
-Robert-Koch-Institut}. We assume that the time between symptom onset and ICU takes 4, 6,
-8 or 10 days with equal probabilities. As we do not model nursing homes, do not focus on
-matching deaths and do not use the number of individuals in intensive care to estimate
-our parameters, these numbers are not important for our empirical results.
+\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
+that the time between symptom onset and ICU takes four, six, eight or ten days with equal
+probabilities.
 
@@
 match this approximately we discretize that 41\% of individuals who will die from
-Covid-19 do so after one day in intensive care, 22\% day after 12 days, 29\% after 20
+Covid-19 do so after one day in intensive care, 22\% day after twelve days, 29\% after 20
 days and 7\% after 32 days. Again, we rescale this for every age group among those that
@@
 after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.
+
+\FloatBarrier

### 2021-07-11 src/paper/data/synthetic_population.tex
Make sections A.1 to A.4 ready for review.

--- 2021-07-11 src/paper/data/synthetic_population.tex
@@
 We sample households to build our synthetic population of over one million households
-keeping for each individual their age, gender, occupation and whether they work on
-Saturdays and Sundays. For each household we draw its county and set the corresponding
-federal state.% we draw the county because county is not part of the campus file
+keeping for each of the 2.3 million individuals their age, gender, occupation and whether
+they work on Saturdays and Sundays. For each household we draw its county and set the
+corresponding federal state.% we draw the county because county is not part of the campus file
 
@@
 For children between three and six years old, we assume all go to preschool (officially
-92.5\% according to \cite{Destatis2020}). Nursery groups consist
+92.5\% according to \cite{Destatis2020}).
+% nurseries and preschools
+Children that attend a nursery meet in groups of four \citep{BertelsmannStiftung2019}
+plus one adult care taker every weekday when there are no school vacations. Preschool
+children meet in groups of nine \citep{BertelsmannStiftung2019} with two adult care
+takers. These groups are mixed with respect to age but all belong to the same state and
+mostly to the same county.
 
-Children that attend a nursery meet in groups of four plus one adult care taker every
-weekday when there are no school vacations. Preschool children meet in groups of nine
-with two adult care takers. These groups are mixed with respect to age but all belong to
-the same state and mostly to the same county. For nurseries the group size is four, for
-preschools it's nine \citep{BertelsmannStiftung2019}.
-
+% school
 Every child that goes to school is part of three different classes that meet every day on
@@
 students \citep{OECD2013} and two teachers. All students in a class are of the same age
-and live in the same state and if mostly also in the same county. In addition, each
+and live in the same state and mostly also in the same county. In addition, each
 child gets assigned a value that captures his or her need to attend nursery, preschool or
 school. This allows us to capture various degrees of emergency care that can be granted
-while educational facilities are closed or function in some kind of rotating schedule.
+while educational facilities are closed or are on some kind of rotating schedule.
 
+% workers
 Workers are assigned to a daily meeting work group. The group sizes vary to match the
@@
 Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
-we match each worker with up to 14 other workers to match the number of reported weekly
-work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}. Each pair is
-assigned a weekend on which they always meet in the absence of work policies. 80\% of
-these contacts are individuals from the same county.
-% work contact priority
-In the same way children have a educational priority determining if they are entitled to
+we match each worker with up to 14 other workers into pairs to match the number of
+reported weekly work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}.
+Each pair is assigned a weekday on which they always meet in the absence of work
+policies. 80\% of these contacts are individuals from the same county.% work contact priority
+In the same way children have an educational priority determining if they are entitled to
 emergency care workers are assigned a work contact priority that captures how necessary
@@
 
+% other recurring contacts
 In addition to creating groups for educational facilities and work we also have other
@@
 
+% individual responses
 The individuals in our population can react to events such as developing symptoms that
@@
 contacts. To determine who would reduce their contacts in such a situation or demand a
-rapid test to avoid quarantine we introduce a quarantine compliance parameter. Similarly
-we introduce a rapid test compliance parameter that determines in which order individuals
-start demanding rapid tests when rapid tests become increasingly available. This makes
-sure that when for example only 10\% of workers get tested, it's the same workers that
-have access to tests every week.
+rapid test we introduce a quarantine compliance parameter. Similarly we introduce a rapid
+test compliance parameter that determines in which order individuals start demanding
+rapid tests when rapid tests become increasingly available. This makes sure that when for
+example only 10\% of workers get tested, it's the same workers that have access to tests
+every week.
 
@@
 recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
-that teachers were prioritized more than recommended by the
-commission.}\comment[id=K]{Write how that plays out in our simulations for the different
-age groups.}
-
-To cover that the Pfizer-BioNTech vaccine was later approved for younger age groups we
-put adolescents and children into two groups that follow after the general population.
-These groups do not become eligible within our simulation frame until June.
+that teachers were prioritized more than recommended by the commission.} To cover that
+the Pfizer-BioNTech vaccine was later approved for younger age groups we put adolescents
+and children into two groups that follow after the general population. These groups do
+not become eligible within our simulation frame until June. The way vaccinations are rolled out in our model is shown in Figure~\ref{fig:vaccinations_by_age_group}.
 

### 2021-07-11 src/paper/appendix_data.tex
Start on extending the model fit section.

@@
 
-The model is described by a large number of parameters that govern the number of
-contacts a person has, the likelihood of becoming infected on each contact, the
-likelihood of developing light or strong symptoms or even dying from the disease as well
-as the duration each stage of the disease takes.
+The model is described by a large number of parameters that govern the number of contacts
+a person has, the reduction in contacts due to NPIs, the demand for rapid tests and PCR
+tests, the likelihood of becoming infected on each contact, the likelihood of developing
+light or strong symptoms or even dying from the disease as well as the duration each
+stage of the disease takes.
 
-Most of these parameters can be calibrated from existing datasets or the medical
-literature or calibrated from surveys and empirical datasets.
+Many of these parameters can be calibrated from the medical literature or calibrated from
+surveys and other empirical datasets.
 
@@
 \input{paper/data/policies}
+
+\input{paper/data/rapid_test_demand}
+
+\input{paper/data/share_known_cases}

### 2021-07-11 src/paper/data/medical_parameters.tex

@@
 
-This section discusses the medical parameters used in the model, their sources and how we
-arrived at the distributions used in the model.\footnote{Additional information can be
-found in the
+This section discusses the medical parameters used in the model, their sources and how
+we arrived at the distributions used in the model.\footnote{Additional information can
+be found in the
 \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
@@
 % Latency Period
-The first medical parameter we need is the length of the period between infection and the
-start of infectiousness, the so called latent period. \cite{Zhao2021} estimate the latent
-period to last 3.3 days (95\% CI: 0.2, 7.9). In line with this estimate our latent period
-lasts one to five days with a slightly lower average of 2 days.
+The first medical parameter we need is the length of the period between infection and
+the start of infectiousness, the so called latent period. \cite{Zhao2021} estimate the
+latent period to last 3.3 days (95\% CI: 0.2, 7.9). In line with this estimate our
+latent period lasts one to five days with a slightly lower average of 2 days.
 
@@
 The incubation period is usually estimated to be two to twelve days. A meta analysis by
-\citet{McAloon2020} comes to the conclusion that ``The incubation period distribution may
-be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\%
-CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For
-simplicity we discretize this distribution into four bins.
+\citet{McAloon2020} comes to the conclusion that ``The incubation period distribution
+may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters
+(95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55),
+respectively.'' For simplicity we discretize this distribution into four bins.
 
@@
 to five days before symptom onset (see their figure 2) and three to eight days of
-infectiousness afterwards.\footnote{Viral loads may be detected much later but eight days
-seems to be the time after which most people are culture negative, as also reported by
-\citet{Singanayagam2020}.} Thus, we arrive at zero to 13 days as the range for
+infectiousness afterwards.\footnote{Viral loads may be detected much later but eight
+days seems to be the time after which most people are culture negative, as also reported
+by \citet{Singanayagam2020}.} Thus, we arrive at zero to 13 days as the range for
 infectiousness among individuals who become symptomatic (see also figure 5).
@@
 We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
-Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases. We
-only disaggregate by age how likely individuals are to require intensive
+Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases.
+We only disaggregate by age how likely individuals are to require intensive
 care.\footnote{The length of symptoms is not very important in our model given that
@@
 For the time from symptom onset until need for intensive care we rely on data by
-\cite{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.
+\cite{Stokes2020}) and \citep{Hinch2021}.
 
@@
 numbers reported for the time from first symptoms to hospitalization:
+
 \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
-that the time between symptom onset and ICU takes four, six, eight or ten days with equal
-probabilities.
+that the time between symptom onset and ICU takes four, six, eight or ten days with
+equal probabilities.
 
@@
 We take the survival probabilities and time to death and time until recovery from
-intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
-time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
-match this approximately we discretize that 41\% of individuals who will die from
-Covid-19 do so after one day in intensive care, 22\% day after twelve days, 29\% after 20
-days and 7\% after 32 days. Again, we rescale this for every age group among those that
-will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
-reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
-days. We discretize this such that of those who recover in intensive care, 22\% do so
-after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.
+intensive care from the \citet{Hinch2021}. They report time until death to have a mean
+of 11.74 days and a standard deviation of 8.79 days. To match this approximately we
+discretize that 41\% of individuals who will die from Covid-19 do so after one day in
+intensive care, 22\% day after twelve days, 29\% after 20 days and 7\% after 32 days.
+Again, we rescale this for every age group among those that will not survive. For
+survivors the OpenABM Project reports a mean duration of 18.8 days until recovery and a
+standard deviation of 12.21 days. We discretize this such that of those who recover in
+intensive care, 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and
+18\% after 45 days.
 

### 2021-07-11 src/paper/data/number_of_contacts.tex
Join single plots and make single plots smaller plus various small improvements.

@@
 and information on the contact. Importantly, for each contact individuals entered of
-which type the contact (school, leisure, work etc.) was and how frequent the contact
-with the other person is.
+which type the contact (school, leisure, work etc.) was and how frequent the contact with
+the other person is. Binning the number of contacts for very high numbers, we arrive at
+the distributions of the numbers of contacts by contact type as shown in Figure~\ref{fig:n_contacts}.
 
-Simplifying the number of contacts, we arrive at the following distributions of the
-numbers of contacts by contact type.
-
-\begin{figure}
+\begin{figure}[ht]
     \centering
@@
         \caption{Number of Non Recurrent Other Contacts}
-        \label{n_contacts_other_non_recurrent}
+        \label{fig:n_contacts_other_non_recurrent}
     \end{subfigure}
@@
         \caption{Number of Daily Recurrent Other Contacts}
-        \label{n_contacts_other_daily_recurrent}
+        \label{fig:n_contacts_other_daily_recurrent}
     \end{subfigure}
@@
         \caption{Number of Weekly Recurrent Other Contacts}
-        \label{n_contacts_other_weekly_recurrent}
+        \label{fig:n_contacts_other_weekly_recurrent}
     \end{subfigure}
@@
         \caption{Number of Non Recurrent Work Contacts}
-        \label{n_contacts_work_non_recurrent}
+        \label{fig:n_contacts_work_non_recurrent}
     \end{subfigure}
@@
         \caption{Number of Daily Recurrent Work Contacts}
-        \label{n_contacts_work_daily_recurrent}
+        \label{fig:n_contacts_work_daily_recurrent}
     \end{subfigure}
@@
         \caption{Number of Weekly Recurrent Work Contacts}
-        \label{n_contacts_work_weekly_recurrent}
+        \label{fig:n_contacts_work_weekly_recurrent}
     \end{subfigure}
@@
     \caption{Number of Contacts of the Different Contact Types}
-    \label{fig:n_contacts_other}
+    \label{fig:n_contacts}
     \floatfoot{\noindent
-        \textit{Note:} The upper row shows the distribution of the number of other
-        contacts individuals report. Other contacts include all contacts that are not
-        household members, school contacts or work contacts, for example leisure
-        contacts. The planned number of contacts is reduced by policies, seasonality and
-        individual responses to events such as receiving a positive rapid test to the
-        number of actual contacts with transmission potential.
-        % non recurrent
+        \textit{Note:} This figure shows the pre-pandemic number of contacts individuals
+        report of different contact types.
+        % non recurrent contacts
         In the model it is sampled every day which of the numbers of non recurrent
@@
         high values that super spreading events are well possible in our model through
-        non recurrent other models. We assume that individuals in households with
-        children or teachers or retired individuals have additional non recurrent
-        contacts during school vacations to cover things like family visits or travel
-        during vacations. We estimate this to be on average 0.5 additional contacts per
-        vacation day.
-        % daily and weekly
-        For the recurrent other contacts, individuals are assigned to groups that are
-        time constant and that meet daily or weekly. The share of individuals who attend
-        in a way that has transmission potential is reduced by policies, seasonality and
-        individual responses to events such as receiving a positive rapid test. For
-        weekly contacts, individuals are assigned to up to four groups that are time
-        constant and that meet weekly. The day on which meetings take place varies
-        between groups but stays the same for each group.
+        non recurrent models.
+        % recurrent contacts
+        For recurrent contacts individuals are put into groups that meet either every day
+        or on a particular week day every day. For work contacts, meetings can only take
+        place on work days.
+        % reductions that can happen
+        The pre-pandemic number of contacts with transmission potential is reduced by
+        policies, seasonality and individual responses to events such as
+        receiving a positive rapid test to the number of actual contacts with
+        transmission potential.
+        % other
+        The upper row shows the distribution of the number of other contacts individuals
+        report. Other contacts include all contacts that are not household members,
+        school contacts or work contacts, for example leisure contacts. We assume that
+        individuals in households with children or teachers or retired individuals have
+        additional non recurrent other contacts during school vacations to cover things
+        like family visits or travel during vacations.
         % work contacts
         The lower row shows the distribution of the different types of work contacts.
-        Work contacts only take place between working individuals.
-        % non recurrent
-        In the model it is sampled every day which number of non recurrent work contacts
-        a working person is planned to have. Note that the contact diaries include such
-        high values that super spreading events are well possible in our model. The
-        planned number of contacts is reduced by policies, seasonality and individual
-        responses to events such as receiving a positive rapid test to the number of
-        actual contacts with transmission potential.
-        % daily
-        Working individuals are assigned to groups that are time constant and that meet
-        daily to match the given distribution of daily work contacts. This contact type
-        covers contacts such as colleagues with which one shares an office space. The
-        share of individuals who attend in a way that has transmission potential is
-        reduced by policies (such as a work from home mandate), seasonality and
-        individual responses to events such as receiving a positive rapid test.
-        % weekly
-        Working individuals are assigned to up to 14 groups that are time constant and
-        meet weekly. Groups are scheduled to meet on separate days of the work week.
-        These contact models cover weekly team meetings etc. The share of individuals
-        that attend in a way that has transmission potential is reduced by policies,
-        seasonality and individual responses to events such as receiving a positive rapid
-        test.}
+        Work contacts only take place between working individuals.}
 \end{figure}
 
+\FloatBarrier
+
 An exception where we do not rely on the data by \cite{Mossong2008} are the household
-contacts. Since household are included in the the German microcensus
+contacts. Since households are included in the the German microcensus
 \citep{FDSAeDBUDL2018} on which we build our synthetic population we simply assume for
-the household contact model that individuals meet all other household members every day.
+the household contacts that individuals meet all other household members every day. The
+number of household contacts that happen every day is shown in
+Figure~\ref{fig:n_contacts_hh}.
 
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
+    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
     \caption{Number of Household Contacts}
-    \label{n_contacts_hh}
+    \label{fig:n_contacts_hh}
     \floatfoot{\noindent

### 2021-07-11 src/paper/data/synthetic_population.tex

@@
 Each pair is assigned a weekday on which they always meet in the absence of work
-policies. 80\% of these contacts are individuals from the same county.% work contact priority
+policies. 80\% of these contacts are individuals from the same county.
+% work contact priority
 In the same way children have an educational priority determining if they are entitled to
@@
 group and a vaccination rank from that group that creates a complete vaccination queue
-over the population including 15\% that refuse to be vaccinated as reported by the Robert
-Koch Institute (RKI, \cite{RKI2021}). The vaccination groups are created to match the
-recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
-that teachers were prioritized more than recommended by the commission.} To cover that
-the Pfizer-BioNTech vaccine was later approved for younger age groups we put adolescents
-and children into two groups that follow after the general population. These groups do
-not become eligible within our simulation frame until June. The way vaccinations are rolled out in our model is shown in Figure~\ref{fig:vaccinations_by_age_group}.
+over the population including 15\% that refuse to be vaccinated as reported by the
+Robert Koch Institute (RKI, \cite{RKI2021c}). The vaccination groups are created to
+match the recommendations by the Ständige Impfkommission
+\citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized more than
+recommended by the commission.} To cover that the Pfizer-BioNTech vaccine was later
+approved for younger age groups we put adolescents and children into two groups that
+follow after the general population. These groups do not become eligible within our
+simulation frame until June. The way vaccinations are rolled out in our model is shown
+in Figure~\ref{fig:vaccinations_by_age_group}.
 

### 2021-07-11 src/paper/model/testing.tex

@@
 approach: First, we calculate the total number of positive PCR tests by multiplying the
-number of newly infected individuals with an estimate of the share of detected cases from
-the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how many of these tests
-should go to symptomatic and asymptomatic individuals from data by the Robert Koch
-Institute. \comment[id=J]{add source}
-Then, we sample the individuals to which those tests are allocated from the pools of
-symptomatic and asymptomatic infected but not yet tested individuals.
+number of newly infected individuals with an estimate of the share of detected cases
+from the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how
+many of these tests should go to symptomatic and asymptomatic individuals from data by
+the RKI.\comment[id=J]{add source} Then, we sample the individuals to which those tests
+are allocated from the pools of symptomatic and asymptomatic infected but not yet tested
+individuals.
 
@@
 At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
-of detected cases from \citet{Dunkelzifferradar2020} can no longer be used because it is based on
-case fatality rates which drastically change due to vaccinations. Secondly, rapid tests
-become available at a large scale.
+of detected cases from \citet{Dunkelzifferradar2020} can no longer be used because it is
+based on case fatality rates which drastically change due to vaccinations. Secondly,
+rapid tests become available at a large scale.
 
@@
 
-The second challenge is solved by mechanistic rapid test demand models for the workplace,
-schools and by private individuals. The calibration of these models is described in
-Section~\ref{subsec:rapid_test_demand}. Figure~\ref{fig:antigen_tests_vaccinations} shows
-that the number of performed rapid tests in the model fits the empirical data well (where
-empirical data is available).
+The second challenge is solved by mechanistic rapid test demand models for the
+workplace, schools and by private individuals. The calibration of these models is
+described in Section~\ref{subsec:rapid_test_demand}.
+Figure~\ref{fig:antigen_tests_vaccinations} shows that the number of performed rapid
+tests in the model fits the empirical data well (where empirical data is available).
 
@@
 We do not distinguish between self administered rapid tests and those that are
-administered by medical personnel. While there were concerns that self administered tests
-are less reliable, a recent study has found no basis for this concern.\comment[id=J]{cite
-charite study on self administered tests}
+administered by medical personnel. While there were concerns that self administered
+tests are less reliable, a recent study has found no basis for this
+concern.\comment[id=J]{cite charite study on self administered tests}
 
@@
 the share of detected cases.\comment[id=K]{Talk here about reactions to tests and
-symptoms with contact reductions. (If not here, find other place where to talk about this
-in detail.)}
+symptoms with contact reductions. (If not here, find other place where to talk about
+this in detail.)}

### 2021-07-11 src/paper/results/rapid_tests_statistics.tex

@@
   \centering
+  \begin{subfigure}[b]{0.3\textwidth}
       \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
-  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels
+      \caption{Share of the Population Doing a Rapid Test Because of Different Channels
       on a Given Day}
       \label{fig:rapid_test_demand_by_channel}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.3\textwidth}
+      \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_positive}
+      \caption{Share of the Population Testing Positive Because of Different Channels
+      on a Given Day}
+      \label{fig:pos_rapid_tests_by_channel}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.3\textwidth}
+      \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_negative}
+      \caption{Share of the Population Testing Negative Because of Different Channels
+      on a Given Day}
+      \label{fig:neg_rapid_tests_by_channel}
+  \end{subfigure}
+  \caption{Rapid Test Shares in the Population by Channel}
   \floatfoot{\noindent \textit{Note:}
@@
     member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. }
+    meeting with friends. The left figure shows the share of the population demanding a
+    rapid test on a given day. The middle figure shows the share of the population
+    testing positive on a given day (true and false positives). The right figure shows
+    the share of the population testing negative on a given day (true and false
+    negatives).}
 \end{figure}

### 2021-07-11 src/paper/results/fit.tex
Fix some references and remove weekly incidences from plot descriptions.

@@
   \caption{Simulated and Empirical Infections by Age Group}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different age
-  groups. The age group of individuals above 80 needs to be interpreted with caution
-  because our synthetic population only includes private households, i.e. nursing homes
-  are not represented in our model. They accounted for many cases and deaths in the
-  winter of 2020 and many 80 to 100 year olds live in these facilities. However, the
-  official data does not contain information on whether cases were nursing home
-  inhabitants or not. We averaged over 30 simulation runs.}
+  \floatfoot{\noindent \textit{Note:} The figure shows the number of reported versus
+  simulated cases per one million people per day for different age groups. The age group
+  of individuals above 80 needs to be interpreted with caution because our synthetic
+  population only includes private households, i.e. nursing homes are not represented in
+  our model. They accounted for many cases and deaths in the winter of 2020 and many 80
+  to 100 year olds live in these facilities. However, the official data does not contain
+  information on whether cases were nursing home inhabitants or not. We averaged over 30
+  simulation runs.}
   \label{fig:age_group_fit}
@@
   \caption{Simulated and Empirical Infections by Federal State}
-  \floatfoot{\noindent \textit{Note:} The figure shows the weekly incidence rates per
-  100,000 people for the reported versus the simulated infections rates for different
-  federal states. We averaged over 30 simulation runs.}
+  \floatfoot{\noindent \textit{Note:} The figure shows the number of reported versus
+  simulated cases per one million people per day for different federal states. We
+  averaged over 30 simulation runs.}
   \label{fig:state_fit}

### 2021-07-11 src/paper/data/policies.tex
Make policy data section ready for review.

@@
 For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures, emergency care and A / B schooling where only
-one half of students attends every other week or day.
-% For the moment we ignore that lack of childcare leads working parents to stay home.
+federal states as well as school closures, emergency care and rotating schooling
+schedules where only one half of students attends every other week or day.% For the moment we ignore that lack of childcare leads working parents to stay home.
 An approximation of the share of contacts still taking place with the different school
@@
 % work
-For work contacts we use the reductions in work mobility reported in the Google
-Mobility Data \citep{Google2021} to calibrate our work policies. Reductions in work
-contacts are not random but governed through a work contact priority where the policy
-changes the threshold below which workers stay home. Figure \ref{fig:work_multiplier}
-shows the share of workers that go to work over time at the federal German level. We use
-the data on the state level.
-
+For work contacts we use the reductions in work mobility reported by the Google Mobility
+Data \citep{Google2021} to calibrate our work policies. Reductions in work contacts are
+not random but governed through a work contact priority where the policy changes the
+threshold below which workers stay home. Figure \ref{fig:work_multiplier} shows the share
+of workers that go to work over time at the federal German level. We use the data on the
+state level to account for local holidays and differences in state regulations.
+% hygiene umltiplier
 In addition, for both work and school contacts we assume that hygiene measures (such as
@@
 
-    \caption{The Effect of School and Work Policies on the Number Contacts Taking Place}
+    \caption{The Contact Reduction Effects of School and Work Attendance Policies}
 
@@
 
-Lastly, for the other contacts category we did not calibrate them from data but estimated
-the policy effects on them. Refer to Section~\ref{subsec:estimated_params} and Figure~\ref{fig:other_multiplier} in particular.
+Lastly, for the other contacts category we could not calibrate the policies from data but
+estimated the policy effects. The estimation and values are detailed in
+Section~\ref{subsec:estimated_params} and Figure~\ref{fig:other_multiplier}.
 

### 2021-07-11 src/paper/data/rapid_test_demand.tex
Typos in the estimated param table and get rapid test demand data section ready for review.

@@
 to their employees \citep{Bundesanzeiger2021}. We assume that compliance is incomplete
-and only 80\% of employers actually offer tests.
+and only 80\% of employers actually offer tests. We interpolate between these points
+linearly, arriving at the blue line in Figure~\ref{fig:rapid_test_demand}.
 
@@
 based on tests becoming mandatory in Bavaria \citep{BayerischeStaatskanzlei2021} and
-North Rhine-Westphalia \citep{SchulministeriumNRW2021b} after their Easter breaks and
-on the 19th in Baden-Württemberg \citep{KMBaWue2021}.
+North Rhine-Westphalia \citep{SchulministeriumNRW2021b} after their Easter breaks and on
+the 19th in Baden-Württemberg \citep{KMBaWue2021}. Again, we interpolate linearly between
+these points and arrive at the purple line for teachers and the red line for school
+students in Figure~\ref{fig:rapid_test_demand}.
 
@@
 positive. Since this is only asking for willingness not actual behavior, we take this as
-the upper bound of private rapid test demand which is reached in the beginning of May. To
-cover that many people are likely to have sought and done their first rapid test before
-the Easter holidays to meet friends or family, we let the share of individuals doing
-rapid tests in that time increase more rapidly than before and after. By end of March
-25\% of individuals would do a rapid test due to a private reason.
-
-All shares of individuals who would take a rapid test if the conditions were met can be
-seen in Figure~\ref{fig:rapid_test_demand}.
+the upper bound of private rapid test demand which we estimate in our model to be reached
+in the beginning of May. To cover that many people are likely to have sought and done
+their first rapid test before the Easter holidays we add another point that we estimate
+for the rapid test demand around Easter. Similarly, we estimate one point in mid March
+when tests started to become available in grocery stores and pharmacies which we estimate
+in our model. The resulting share of private rapid test demand is shown as the green line
+in Figure~\ref{fig:rapid_test_demand}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
+    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
     \caption{Share of Individuals Doing a Rapid Test.}
+    \label{fig:rapid_test_demand}
     \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
@@
     each case whether a rapid test is done depends on how long it has been since the
-    individual's last rapid test and her individual compliance parameters. As an example,
+    individual's last rapid test and her individual compliance parameter. As an example,
     take a worker in May. In that time workers are encouraged to test themselves twice

### 2021-07-11 src/paper/results/validation.tex
Add intro for validation.

@@
 \label{subsec:model_validation}
+
+Achieving a good in-sample fit does not necessarily guarantee that our model will also
+be able to make out of sample predictions. For example, it could be that the results
+are very sensitive to the exact number of vaccinations, the work mobility multiplier
+or the number of rapid tests that are performed -- all of which are things that cannot
+be exactly known ex-ante.
+
+In this section we compare simulated infections that use all available data with
+out of sample predictions that only use data that was available at March 1 2021.
+
+For the out of sample predictions we predict the number of vaccinations between March
+and June with a simple linear regression model that was fitted on vaccine data from
+February. The work mobility multiplier is predicted to be constant at a value of
+0.75, which is an approximate average of the second half of February.
+
+The area that is frought with the most uncerainty are the introduction of rapid test,
+because it comprises both, supply and demand factors. Moreover, accurately predicting
+the number of rapid tests ist expected to be important because rapid tests play a large
+role for the transmission dynamic.
 
@@
   \end{subfigure}
-  \caption{Robustness Check}
+  \caption{Out of sample prediction for reported and total cases from March to June
+    2021. The ex-post scenario is an in-sample prediction that uses all available
+    information and is very close to actual case numbers. For the other scenarios data
+    on vaccinations, work mobility and rapid tests that became available after March 1
+    have been replaced by prediction models that are calibrated with data from February.
+    Moreover, they do not model a lower number of detected cases over the easter
+    holidays. The different scenarios make different assumptions on the date at which
+    full availability of rapid tests is reached. While the out of sample predictions
+    differ substantially for the exact case numbers at the beginning of June (between
+    20 and 70 cases per million), they can all reproduce the decline in case numbers
+    that is jointly driven by seasonality, large scale rapid tests and vaccinations.
+  }
   \label{fig:robustness_check_detailed}

### 2021-07-11 src/paper/data/share_known_cases.tex
Make data section on share known cases ready for review. Add figures showing the parameter demand shares in different robustness scenarios.

@@
 tests). For our model it is important to have an estimate for the share of cases that is
-detected in the absence of rapid tests. For this we rely on the
-\citet{Dunkelzifferradar2020} which relies on estimates
-of the case fatality rate to estimate the number of total cases given the number of
-deaths which are suspected to be perfectly observable. For 2020, we follow the reported
-share of detected cases quite closely only interpolating the phase of November 2020 where
-the testing policy was changed to give more priority to vulnerable groups. This lead to
-more tests being done among older age groups which are also at the highest risk of dying
-from CoViD-19 and an overestimation of the share of detected cases.% After Christmas 2020
-Similarly, the start of vaccinations after Christmas 2020 that were predominantly given
-to nursing homes in the beginning and other vulnerable groups in spring, led to a
-decrease in the number of deadly cases, leading to an increasing upward bias in the share
-of detected cases. This is why we stop using the \citet{Dunkelzifferradar2020} after Christmas.
-Instead, we assume that that the share of detected cases would have stayed the same in
-the absence of rapid tests. Thus, we also achieve in our model an increase in the share
-of detected cases but this is driven from inside our model through increased testing and
-follow-ups on positive rapid tests (see Section~\ref{subsec:results_share_known_cases}).
+detected in the absence of rapid tests. For this we rely on the Dunkelzifferradar Project
+\citep{Dunkelzifferradar2020} which relies on estimates of the case fatality rate to
+estimate the number of total cases given the number of CoViD-19 deaths which are assumed
+to be perfectly observable. For 2020, we follow the reported share of detected cases
+quite closely only interpolating the phase of November 2020 where the testing policy was
+changed to give more priority to vulnerable groups. This lead to more tests being done
+among older age groups which are also at the highest risk of dying from CoViD-19 and an
+overestimation of the share of detected cases.
+% After Christmas 2020
+Since vaccinations started after Christmas 2020 and these were predominantly given to
+nursing homes in the beginning and other vulnerable groups in spring, we expect the
+relationship between deaths and the number of total infections to change rapidly in 2021.
+This is why we stop using the share of detected cases estimated by the Dunkelzifferradar
+after Christmas. Instead, we assume that the share of detected cases would have
+stayed the same in the absence of rapid tests. Thus, we also achieve in our model an
+increase in the share of detected cases but this is driven from inside our model through
+increased rapid testing and follow-up on positive rapid tests (see
+Section~\ref{subsec:results_share_known_cases} and \ref{sub:testing}).
 % Christmas and Easter
-Lastly, we include reductions in the share known cases due to the two major holidays in
-our simulation period, Christmas and Easter. During both holidays where many laboratories
-do not process tests and most physicians' offices are closed, less PCR tests were done
-leading to short and large drops in the share of known cases.
+Lastly, we model reductions in the share known cases due to the two major holidays in our
+simulation period, Christmas and Easter. During both holidays many laboratories did not
+process tests and most physicians' offices were closed, leading to less PCR tests and
+short and large drops in the share of known cases. The resulting share of detected cases
+in the absence of rapid tests is shown in Figure~\ref{fig:share_known_cases_data}.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/data/testing/assumed_overall_share_known_cases}
-  \caption{Share of Detected Cases}
+  \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/assumed_overall_share_known_cases}
+  \caption{Share of Detected Cases in the Absence of Rapid Tests}
   \label{fig:share_known_cases_data}

### 2021-07-11 src/paper/report.tex
urls to references.
Add figure showing rapid test fade in parameters for the robustness check scenarios and format the results figure.

@@
 
-    \begin{subfigure}[b]{0.475\textwidth}
+    \begin{subfigure}[b]{0.425\textwidth}
         \centering
-        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
         \caption{{Effects of different schooling scenarios}}
@@
     \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
+    \begin{subfigure}[b]{0.425\textwidth}
         \centering
-        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
         \caption{{Effects of different work scenarios}}
@@
 \citep[e.g.][]{StillmanTonin2021}; at the same time, these groups have a higher risk to
-contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the
+contract CoViD-19 \citep{RKI2021a}. Mandatory tests at school and at the
 workplace will extend more into these groups. The same goes for individuals who exhibit

### 2021-07-11 src/paper/results/validation.tex
Add figure showing rapid test fade in parameters for the robustness check scenarios and format the results figure.

@@
 
-The area that is frought with the most uncerainty are the introduction of rapid test,
-because it comprises both, supply and demand factors. Moreover, accurately predicting
-the number of rapid tests ist expected to be important because rapid tests play a large
+The area that is fraught with the most uncertainty are the introduction of rapid test,
+because it comprises both supply and demand factors. Moreover, accurately predicting
+the number of rapid tests is expected to be important because rapid tests play a large
 role for the transmission dynamic.
+
+\begin{figure}[ht] % Robustness Check Rapid Test Fade In Params
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand/robustness_check_params_early_shares}
+    \caption{Rapid Test Parameters: Early Scenario}
+    \label{fig:robustness_early_params}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand/robustness_check_params_medium_shares}
+    \caption{Rapid Test Parameters: Medium Scenario}
+    \label{fig:robustness_medium_params}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand/robustness_check_params_late_shares}
+    \caption{Rapid Test Parameters: Late Scenario}
+    \label{fig:robustness_late_params}
+  \end{subfigure}
+
+  \caption{Rapid Test Introduction Parameters in the Three Scenarios}
+  \label{fig:robustness_check_rapid_test_params}
+
+  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
+\end{figure}
 
@@
   \centering
-  \begin{subfigure}[b]{.49\textwidth}
+  \begin{subfigure}[b]{.425\textwidth}
     \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_new_known_case}
+    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_new_known_case}
     \caption{Reported Cases}
@@
   \hfill
-  \begin{subfigure}[b]{.49\textwidth}
+  \begin{subfigure}[b]{.425\textwidth}
     \centering
-    \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_newly_infected}
+    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_newly_infected}
     \caption{Total Cases}
@@
   \end{subfigure}
-  \caption{Out of sample prediction for reported and total cases from March to June
-    2021. The ex-post scenario is an in-sample prediction that uses all available
-    information and is very close to actual case numbers. For the other scenarios data
-    on vaccinations, work mobility and rapid tests that became available after March 1
-    have been replaced by prediction models that are calibrated with data from February.
-    Moreover, they do not model a lower number of detected cases over the easter
-    holidays. The different scenarios make different assumptions on the date at which
-    full availability of rapid tests is reached. While the out of sample predictions
-    differ substantially for the exact case numbers at the beginning of June (between
-    20 and 70 cases per million), they can all reproduce the decline in case numbers
-    that is jointly driven by seasonality, large scale rapid tests and vaccinations.
-  }
+  \caption{Out of Sample Prediction for Reported and Total Cases from March to June
+  2021.}
   \label{fig:robustness_check_detailed}
-  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
+  \floatfoot{\noindent \textit{Note:} The ex-post scenario is an in-sample prediction
+  that uses all available information and is very close to actual case numbers. For the
+  other scenarios data on vaccinations, work mobility and rapid tests that became
+  available after March 1 have been replaced by prediction models that are calibrated
+  with data from February. Moreover, they do not model a lower number of detected cases
+  over the easter holidays. The different scenarios make different assumptions on the
+  date at which full availability of rapid tests is reached. While the out of sample
+  predictions differ substantially for the exact case numbers at the beginning of June
+  (between 20 and 70 cases per million), they can all reproduce the decline in case
+  numbers that is jointly driven by seasonality, large scale rapid tests and
+  vaccinations. \textcolor{red}{To be written}}
 \end{figure}

### 2021-07-11 src/paper/results/share_known_cases.tex
Make share of detected cases results section ready for review.

--- 2021-07-09 src/paper/results/share_known_cases.tex
@@
-\subsection{Share of Cases that are Detected}
-\label{subsec:appendix_share_known_cases}
+\subsection{Share of Detected Cases}
+\label{subsec:results_share_known_cases}
+
+We show the age group specific share of detected cases in
+Figure~\ref{fig:share_known_cases_by_age_group}.\footnote{We explain how we model the
+detection of cases in Section~\ref{sub:testing} and show the overall share of detected
+cases that we use as input for our testing model in
+Section~\ref{subsec:data_share_known_cases}.}
+
+In 2020 the share of detected cases is much higher in older age groups as they are much
+more likely to develop symptoms and symptomatic individuals are much more likely to be
+tested. Starting in 2021 the share of detected cases not only depends on the overall
+share of detected cases we feed into our model but also on the number of performed rapid
+tests. As more and more (often asymptomatic and presymptomatic) individuals receive
+positive rapid tests and receive confirmation, the share of detected cases starts to
+increase, especially after March when rapid tests became much more widely available. This
+effect is very visible for school pupils, (5-14, green line). As twice weekly rapid tests
+become mandatory in schools and students start going to school after the Easter break,
+the share of detected cases for that group increases drastically from 25\% to over 60\%
+between April and June.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
+  \includegraphics[width=0.5\textwidth]{figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
   \caption{Share of Detected Cases by Age Group}
@@
   \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
-  reported as an official case via PCR confirmation. We use the overall share of known
-  cases that was estimated through the case fatality ratio by the
-  \href{https://covid19.dunkelzifferradar.de/}{Dunkelzifferradar} for all of 2020 and
-  then assume it to be constant as vaccinations of the elderly strongly affect the case
-  fatality rate which the project does not account for. To get from an overall share of
-  detected cases to the share of cases that is detected in each age group we use that
-  asymptomatic cases are much less likely to be detected. As our model covers age
-  specific asymptomatic rates this endogenously leads to group specific share known cases
-  that verify that infections in younger age groups are under-detected. Starting in 2021
-  in addition to the overall numbers of detected cases through symptoms and the share
-  known cases, cases are also detected through confirmation of positive rapid tests. This
-  leads to an increase in the share of known cases for all age groups but in particular
-  for the younger age groups that are covered extensively with rapid tests through the
-  rapid test requirement for participating in school.}
+  reported as an official case via PCR confirmation for each age group in our population.
+  We use the overall share of detected cases that was estimated through the case fatality
+  ratio by \citet{Dunkelzifferradar2020} for all of 2020 and then assume it to be
+  constant as vaccinations of the elderly strongly affect the case fatality rate which
+  the project does not account for. To get from an overall share of detected cases to the
+  share of cases that is detected in each age group we use that asymptomatic cases are
+  much less likely to be detected. As our model covers age specific asymptomatic rates
+  this endogenously leads to group specific share of detected cases that suggest that
+  infections in younger age groups are under-detected. Starting in 2021 in addition to
+  the overall numbers of detected cases through symptoms and the share of detected cases,
+  cases are also detected through confirmation of positive rapid tests. This leads to an
+  increase in the share of detected cases for all age groups but in particular for the
+  younger age groups that are covered extensively with rapid tests through the rapid test
+  requirement for participating in school. Age groups that have a high share of workers
+  also disproportionately increase their detection rate.}
 \end{figure}
 
-It's noteworthy that the share of detected cases increases rapidly in May for the five to
-fourteen year olds. This is a direct result of the mandatory tests in
-school.
-
 \FloatBarrier

### 2021-07-11 src/paper/results/validation.tex
Out of sample validation ready for review.

@@
 and June with a simple linear regression model that was fitted on vaccine data from
-February. The work mobility multiplier is predicted to be constant at a value of
-0.75, which is an approximate average of the second half of February.
+February. This prediction model is pessimistic compared to the actual number of
+vaccinations. The work mobility multiplier is predicted to be constant at a value of
+0.75, which is an approximate average of the second half of February. This turned out
+to be optimistic.
 
-The area that is fraught with the most uncertainty are the introduction of rapid test,
+The area that is fraught with the most uncertainty is the introduction of rapid test,
 because it comprises both supply and demand factors. Moreover, accurately predicting
@@
 role for the transmission dynamic.
+
+We therefore make a scenario analysis with different assumptions on the availability
+of rapid tests. The number of rapid tests performed in each scenario can be seen in
+Figure~\ref{fig:robustness_check_rapid_test_params}. All scenarios are the same
+unti March 1 and have the same level of rapid tests when all supply constraints are
+resolved. They differ in the date at which the full number of tests is reached. For
+students and teachers the full number of rapid tests is reached after the easter
+holidays. For rapid tests in the workplace and private rapid tests it is reached between
+May 1 and June 10, depending on the scenario.
 
@@
 
-  \caption{Rapid Test Introduction Parameters in the Three Scenarios}
+  \caption{Rapid Test Introduction in the Three Scenarios}
   \label{fig:robustness_check_rapid_test_params}
 
-  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
+  \floatfoot{\noindent \textit{Note:} Number of rapid tests performed in the different
+  prediction scenarios. All scenarios are the same
+  unti March 1 and have the same level of rapid tests when all supply constraints are
+  resolved. They differ in the date at which the full number of tests is reached. For
+  students and teachers the full number of rapid tests is reached after the easter
+  holidays. For rapid tests in the workplace and private rapid tests it is reached
+  between May 1 and June 10, depending on the scenario.
+  }
 \end{figure}
+
+Moreover, the out of sample predictions assume that the share of detected cases that
+would have obtained without rapid tests is not affected by the easter holidays because
+the extent to which this was the case was estimated from case numbers in April.
+
+The results of the out of sample prediction are displayed in
+Figure~\ref{robustness_check_detailed}. While all scenarios considerably deviate from
+the ex-post scenario, they all reproduce the steep increase of cases until the end
+of April, followed by a decline until June. We can therefore conclude that our main
+results are not sensitive to measurement errors in the number of rapid tests,
+vaccinations or mobility data.
 
@@
   numbers that is jointly driven by seasonality, large scale rapid tests and
-  vaccinations. \textcolor{red}{To be written}}
+  vaccinations.}
 \end{figure}
-
- \comment[id=K]{@Janos: Please add caption and description for the robustness check.}

### 2021-07-11 src/paper/data/synthetic_population.tex
Fix figure references.

@@
 For a distribution of the number of daily recurring work contacts see
-Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
+Figure~\ref{fig:n_contacts_work_daily_recurrent}. To match the number of weekly work groups
 we match each worker with up to 14 other workers into pairs to match the number of
-reported weekly work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}.
+reported weekly work contacts shown in Figure~\ref{fig:n_contacts_work_weekly_recurrent}.
 Each pair is assigned a weekday on which they always meet in the absence of work
@@
 practice regularly together. Both daily and weekly groups are created analogously to the
-work groups but matching the numbers in Figure~\ref{n_contacts_other_daily_recurrent} and
-Figure~\ref{n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
+work groups but matching the numbers in Figure~\ref{fig:n_contacts_other_daily_recurrent} and
+Figure~\ref{fig:n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
 highly assortative by age all individuals that have a daily leisure contact are matched

### 2021-07-11 src/paper/results/rapid_tests_statistics.tex
Fix figure references.

@@
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
+\subsection{Rapid Test Statistics}
+\label{subsec:results_rapid_test_statistics}
+
+An important feature in our model is that our rapid tests are imperfectly specific and
+imperfectly sensitive. Especially when people have not become infectious yet they are
+likely to return a false positive result. Here we look at how rapid tests expanded over
+time and to what degree they are useful as a screening device despite missing positive
+cases and returning false positive results.
+
+We start with the share of the population doing a rapid test, receiving a positive rapid
+test and receiving a negative rapid test over time by the channel through which the test
+was demanded in Figures~\ref{fig:rapid_test_demand_by_channel,
+fig:pos_rapid_tests_by_channel} and \ref{fig:neg_rapid_tests_by_channel} respectively.
+Overall, the share of the population getting a rapid test on a given day increases from
+2\% in Mid March to over 10\% by May. The work rapid tests are a little ragged because of
+public holidays and weekend effects that lead to jumps around Mondays. For education
+rapid tests both vacations (first half of April) as well as the opening of schools in May
+are very visible in the rapid test demand. Overall, work tests make up the largest
+fraction of rapid tests. The image is very similar for the share of positive tests,
+except that the overall number of positive tests starts decreasing in May as rapid test
+expansion comes to a halt and cases fall, especially the positive share of private rapid
+tests falls as less and less individuals seek a rapid test because of a positive test in
+their household.
 

### 2021-07-11 src/paper/results/rapid_tests_statistics.tex
Add figure descriptions and main text for the rapid test statistics results.

@@
 An important feature in our model is that our rapid tests are imperfectly specific and
-imperfectly sensitive. Especially when people have not become infectious yet they are
-likely to return a false positive result. Here we look at how rapid tests expanded over
-time and to what degree they are useful as a screening device despite missing positive
-cases and returning false positive results.
+imperfectly sensitive. Especially when people have not become infectious yet rapid tests
+are likely to return a false positive result. Here we look at how rapid tests expanded
+over time and to what degree they are useful as a screening device despite missing
+positive cases and returning false positive results.
 
@@
 test and receiving a negative rapid test over time by the channel through which the test
-was demanded in Figures~\ref{fig:rapid_test_demand_by_channel,
-fig:pos_rapid_tests_by_channel} and \ref{fig:neg_rapid_tests_by_channel} respectively.
-Overall, the share of the population getting a rapid test on a given day increases from
-2\% in Mid March to over 10\% by May. The work rapid tests are a little ragged because of
-public holidays and weekend effects that lead to jumps around Mondays. For education
-rapid tests both vacations (first half of April) as well as the opening of schools in May
-are very visible in the rapid test demand. Overall, work tests make up the largest
-fraction of rapid tests. The image is very similar for the share of positive tests,
-except that the overall number of positive tests starts decreasing in May as rapid test
-expansion comes to a halt and cases fall, especially the positive share of private rapid
-tests falls as less and less individuals seek a rapid test because of a positive test in
-their household.
+was demanded in Figures~\ref{fig:rapid_test_demand_by_channel},
+\ref{fig:pos_rapid_tests_by_channel} and \ref{fig:neg_rapid_tests_by_channel},
+respectively. Overall, the share of the population getting a rapid test on a given day
+increases from 2\% in mid March to over 10\% by May. The work rapid tests are a little
+ragged because of public holidays. For education rapid tests both vacations (first half
+of April) as well as the opening of schools in May are very visible in the rapid test
+demand. Overall, work tests make up the largest fraction of rapid tests. The image is
+very similar for the share of positive tests, except that the overall number of positive
+tests starts decreasing in May as rapid test expansion comes to a halt and cases fall,
+especially the positive share of private rapid tests falls as less and less individuals
+seek a rapid test because of a positive test in their household.
 
@@
     member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. The left figure shows the share of the population demanding a
+    meeting with friends. The left figure shows the share of the population doing a
     rapid test on a given day. The middle figure shows the share of the population
@@
 \end{figure}
+
+\FloatBarrier
+
+Next, we show the tests split by whether they are true positive, false positive, true
+negative or false negative (see Figure~\ref{fig:rapid_test_results_numbers}) in numbers
+per million individuals to have a more legible metric.
+% true positive
+The number of true positives (Figure~\ref{fig:rapid_tests_number_true_positive}) rapidly
+increases and peaks at the end of April with over 200 cases per million detected through
+rapid tests per day. This means that our model suggests that Germany was able to detect
+up to 16,600 cases per day that would have likely gone undetected otherwise. The most
+powerful tool for detecting cases are the private rapid tests. This is because a large
+share of them are targeted, i.e. triggered by events in the household. Thus, we have to
+be careful to interpret them, as many private tests are likely to be triggered by cases
+detected through work rapid tests. Thus, it's unsurprising that their Shapley value
+(Figure~\ref{fig:2021_scenarios_decomposition_tests}) implies that their contribution is
+only 50\% of the effect of rapid tests with work rapid tests accounting for over 40\% of
+the effect of rapid tests and schools only accounting for 7\%.
+% false positive and true negative
+Turning to the number of false positive tests
+(Figure~\ref{fig:rapid_tests_number_false_positive}) we see that the number of false
+positives increases until May during which the number of actual cases declines and
+testing expands, both leading to more false positive test results. The same pattern
+emerges for the true negative tests (Figure~\ref{fig:rapid_tests_number_true_negative})
+but on a much larger scale. Overall the number of false positives is negligible. Less
+than 0.07\% of the population test false positive.
+% false negative
+Lastly, concerning worries about the accuracy of rapid tests, our model has very
+encouraging results. Despite modeling in detail the low sensitivity of rapid tests before
+the onset of infectiousness, there are very few (less than 80 cases per million per day)
+false negative results as can be seen in
+Figure~\ref{fig:rapid_tests_false_negative_rate}. The false negative rate is highest for
+private demand because many household members perform the rapid test in response to an
+infected household member because they themselves become infectious.
 
@@
     \floatfoot{\noindent \textit{Note:}
-      The number of rapid tests of each category are upscaled to the full German
-      population. \textcolor{red}{To be written.}
-    }
+    Each panel shows the number of rapid tests per million inhabitants that fall into the
+    respective category. Private rapid tests are especially good at detecting cases but
+    since they are often triggered by rapid tests from other channels, the other groups
+    of tests, especially rapid tests at the workplace, also play an important role for
+    containing the pandemic. Since they are often triggered by rapid tests in the
+    household private rapid tests are especially able to prevent infections because they
+    are done soon after infection has taken place but are also more prone to return false
+    negative results (see bottom right panel) because they are more likely to be
+    performed before the onset of infectiousness.}
 \end{figure}
+
+\FloatBarrier
+
+We now turn to the false positive and false negative rates, i.e. the share of positive
+tests that go to people who are not truly infected and the share of negative tests that
+go to individuals that are actually infected. The lower these numbers the more accurate
+the test results. As can be seen in Figure~\ref{fig:rapid_tests_false_positive_rate} the
+share of false positive rates is very high. On average 60\% to 93\% of positive tests are
+received by individuals that are not infected. This is due to the low base rate, which
+falls over time leading to an increase in the false positive rate. Again, private rapid
+tests are an exception with a much lower false positive rate. Here the base rate of
+individuals doing a rapid test through this channel is much higher because these tests
+are often triggered by an event that makes it likely that the person doing the test is
+actually infected. On the other hand, For the most part negative rapid tests are very
+accurate with less than 0.2\% going to infected individuals at any point in time. One
+exception are again the private rapid tests. Here, because they are triggered and are
+thus performed very early, the chance of a false negative result is very high in the
+beginning. As the incidences drop and less private rapid tests are triggered by positive
+rapid tests in the household and a larger percentage of private rapid tests are done for
+leisure activities, the false negative rate steeply decreases.
 
@@
     \floatfoot{\noindent \textit{Note:}
-      \textcolor{red}{To be written.}
+    The left panel shows the share of positive tests that are given to people who are not
+    infected. This share is large as can be expected with a very low baseline rate of
+    positive individuals. As the incidence in the population drops, the false positive
+    rate increases. An exception are the private rapid tests because they are --
+    especially when the incidence is high -- often triggered by events that make it
+    likely that the test taker is infected and therefore their false positive rate is
+    much lower. The right panel shows the false negative rate in the population, i.e. the
+    share of negative test results that are mistakenly given to infected individuals. Our
+    simulations show that rapid tests are very good at detecting infected individuals
+    overall. The only exception are private rapid tests because triggered rapid tests are
+    likely to be performed early during an infection when the sensitivity of rapid tests
+    is still rather low.
     }
 \end{figure}
+
+\FloatBarrier
+
+ \comment[id=K]{@Janos: Please write the section on the testshares.}
 
@@
     \floatfoot{\noindent \textit{Note:}
-      \textcolor{red}{To be written.}
-    }
+    This figure shows the share of all tests that are true positive, false positive, true
+    negative and false negative. The share of true positives is very low, except for the
+    targeted private rapid tests. The share of false positives is also very low, except
+    for the private rapid tests that are more likely to be performed early during the
+    infection when the sensitivity of rapid tests is still very low. Private tests also
+    are the exception among the true negative results. Negative positive test results
+    must be interpreted with caution when an individual is suspected to have been
+    exposed, especially when that exposure has been recently as they are likely to return
+    false negative results in the first few days of an infection. }
 \end{figure}

### 2021-07-11 src/paper/results/rapid_tests_statistics.tex
Rewrite section on rapid test statistics.

--- 2021-07-11 src/paper/results/rapid_tests_statistics.tex
@@
-\subsection{Rapid Tests}
-\label{subsec:appendix_rapid_tests}
+\subsection{Simulated Rapid Test Statistics}
+\label{subsec:results_rapid_test_statistics}
+
+In order to make most use out of limited data sources on rapid test usage, we model the
+number of performed rapid tests as a result of time invariant willingness to do rapid
+tests and time varying supply side factors and events that trigger rapid tests. Thus,
+the parameters described in Section~\ref{subsec:rapid_test_demand} are only indirectly
+related to the number of rapid tests that are actually performed in the model. When it
+comes to positive and negative rapid tests, there is even an additional layer because
+rapid tests imperfectly sensitive and specific.
+
+In this section we look at how rapid tests expanded in our simulations over time and to
+what degree they are useful as a screening device despite their imperfections.
+
+We start with the share of the population doing a rapid test, receiving a positive rapid
+test and receiving a negative rapid test over time by the channel through which the test
+was demanded in Figures~\ref{fig:rapid_test_demand_by_channel},
+\ref{fig:pos_rapid_tests_by_channel} and \ref{fig:neg_rapid_tests_by_channel},
+respectively. Overall, the share of the population getting a rapid test on a given day
+increases from 2\% in mid March to over 10\% by May. The work rapid tests are a little
+ragged because of public holidays. For education rapid tests both vacations (first half
+of April) as well as the opening of schools in May are very visible in the rapid test
+demand. Overall, work tests make up the largest fraction of rapid tests. The image is
+very similar for the share of positive tests, except that the overall number of positive
+tests starts decreasing in May as rapid test expansion comes to a halt and cases fall,
+especially the positive share of private rapid tests falls as less and less individuals
+are triggered to seek a rapid test because of a risk contact in their household.
 
@@
     Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as school pupils. After Easter the required frequency of tests is
+    school) as well as pupils. After Easter the required frequency of tests is
     increased from once per week to twice per week. Work rapid tests are demanded by
@@
     employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Tests are demanded by individuals for one of three private
+    testing is also increased. Private tests are demanded by individuals for one of three
     reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned weekly
-    meeting with friends. The left figure shows the share of the population demanding a
+    member that has tested positive or developed symptoms or having planned a weekly
+    meeting with friends. The left figure shows the share of the population doing a
     rapid test on a given day. The middle figure shows the share of the population
@@
 \end{figure}
+
+\FloatBarrier
+
+Next, we show the tests split by whether they are true positive, false positive, true
+negative or false negative (see Figure~\ref{fig:rapid_test_results_numbers}) in numbers
+per million individuals to make the metric comparable to incidences.
+% true positive
+The number of true positives (Figure~\ref{fig:rapid_tests_number_true_positive}) rapidly
+increases and peaks at the end of April with over 200 cases per million detected through
+rapid tests per day. This means that our model suggests that Germany was able to detect
+up to 16,600 cases per day that would have likely gone undetected otherwise. The most
+powerful tool for detecting cases are the private rapid tests. This is because a large
+share of them are targeted, i.e. triggered by events in the household. However, this
+does not mean that rapid tests in the workplace or at school are useless. It is rather
+the combination of large scale screening at work and in schools and very efficient
+follow up tests whenever those screening tests detected a case. Shapley values
+(Figure~\ref{fig:2021_scenarios_decomposition_tests}) take this into acount and
+assign about 50\% of the overall reduction of case numbers via rapid tests to private
+rapid tests with work and school rapid tests accounting for 40\% and 7\%,
+respectively.
+
+Such a large effect of rapid tests seems to be at odds with the general perception that
+they are not very reliable. However, this is not the case, which can be seen from the
+number of
+false negative tests. At its peak at the beginning of May, there are more than 80 false
+negative rapid tests per million inhabitants per day. While this number might sound low
+in isolation, it becomes quite large when set in context: We estimate that at May first
+there are about 200 detected new infections per million inhabitants and day and roughly
+the same number of undetected cases. Thus there are only 400 per million inhabitants
+who could test falsely negative to begin with and not all of them will make a rapid
+test. This shows clearly that the large effect of rapid tests on the infection dynamic
+is not driven by unrealistic assumptions about their sensitivity but rather by the
+fact that there was a very large number of infected individuals who did not know they
+are positive. Detecting and isolating some of them is enough to slow down the
+overall infection dynamic.
 
@@
     \label{fig:rapid_test_results_numbers}
+    \floatfoot{\noindent \textit{Note:}
+    Each panel shows the number of rapid tests per million inhabitants that fall into the
+    respective category. Private rapid tests are especially good at detecting cases but
+    since they are often triggered by rapid tests from other channels, the other groups
+    of tests, especially rapid tests at the workplace, also play an important role for
+    containing the pandemic.}
+\end{figure}
 
-    \floatfoot{\noindent \textit{Note:}
-      The number of rapid tests of each category are upscaled to the full German
-      population. \textcolor{red}{To be written.}
-    }
-\end{figure}
+\FloatBarrier
+
+A similar picture arises, when looking at the false positive rate,
+i.e. the share of positive tests that go to people who are not infected.
+Figure~\ref{fig:rapid_tests_false_positive_rate} shws that the false
+positive rate is very high. On average 60\% to 93\% of positive tests are
+received by individuals that are not infected. The false positive rate increases over
+time. This is due to the low prevalence of infections in the population, which falls
+over time. Again, private rapid tests are an exception with a much lower false
+positive rate because those tests are primarily demanded when there is a high likelihood
+of beeing infected. The low false negative rate of 0.2\% looks very low. As discussed
+above this is deceiving and just a mechanical consequence of a very low prevalence
+of the disease and the many rapid tests done by non-infected people.
 
@@
     \floatfoot{\noindent \textit{Note:}
-      \textcolor{red}{To be written.}
-    }
-\end{figure}
-
-\begin{figure} % Share of Tests That Are True Positive / False Positive / True Negative / False Negative
-    \centering
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_true_positive}
-        \caption{Share of Tests That Are True Positive by Channel}
-        \label{fig:rapid_tests_testshare_true_positive}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_false_positive}
-        \caption{Share of Tests That Are False Positive by Channel}
-        \label{fig:rapid_tests_testshare_false_positive}
-    \end{subfigure}
-    \vskip3ex
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_true_negative}
-        \caption{Share of Tests That Are True Negative by Channel}
-        \label{fig:rapid_tests_testshare_true_negative}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/testshare_false_negative}
-        \caption{Share of Tests That Are False Negative by Channel}
-        \label{fig:rapid_tests_testshare_false_negative}
-    \end{subfigure}
-    \vskip3ex
-    \caption{Share of Rapid Tests by Outcome and True Infection Status}
-    \label{fig:rapid_test_results_test_shares}
-
-    \floatfoot{\noindent \textit{Note:}
-      \textcolor{red}{To be written.}
-    }
+    The left panel shows the share of positive tests that are given to people who are not
+    infected. This share is large as can be expected with a very low baseline rate of
+    positive individuals. As the incidence in the population drops, the false positive
+    rate increases. An exception are the private rapid tests because they are --
+    especially when the incidence is high -- often triggered by events that make it
+    likely that the test taker is infected and therefore their false positive rate is
+    much lower. The right panel shows the false negative rate in the population, i.e. the
+    share of negative test results that are mistakenly given to infected individuals.
+    This is very low because there are many truly negative tests in times of low
+    incidences and large scale screening tests.
 \end{figure}

### 2021-07-11 src/paper/report.tex
Fixed document, some typos, removed duplicated parameter.

@@
 which has been shown to be very effective \citep{Contreras2021,
-Kretzschmar2020,FetzerGraeber2021}.
+Kretzschmar2020,Fetzer2020}.
 
@@
 \citep[e.g.][]{StillmanTonin2021}; at the same time, these groups have a higher risk to
-contract CoViD-19 \citep{RKI2021a}. Mandatory tests at school and at the
+contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the
 workplace will extend more into these groups. The same goes for individuals who exhibit

### 2021-07-11 src/paper/results/rapid_tests_statistics.tex
Fixed document, some typos, removed duplicated parameter.

@@
 follow up tests whenever those screening tests detected a case. Shapley values
-(Figure~\ref{fig:2021_scenarios_decomposition_tests}) take this into acount and
+(Figure~\ref{fig:2021_scenarios_decomposition_tests}) take this into account and
 assign about 50\% of the overall reduction of case numbers via rapid tests to private
@@
 i.e. the share of positive tests that go to people who are not infected.
-Figure~\ref{fig:rapid_tests_false_positive_rate} shws that the false
+Figure~\ref{fig:rapid_tests_false_positive_rate} shows that the false
 positive rate is very high. On average 60\% to 93\% of positive tests are
@@
 positive rate because those tests are primarily demanded when there is a high likelihood
-of beeing infected. The low false negative rate of 0.2\% looks very low. As discussed
+of being infected. The low false negative rate of 0.2\% looks very low. As discussed
 above this is deceiving and just a mechanical consequence of a very low prevalence
@@
     This is very low because there are many truly negative tests in times of low
-    incidences and large scale screening tests.
+    incidences and large scale screening tests.}
 \end{figure}

### 2021-07-12 src/paper/results/rapid_tests_statistics.tex
Polish rapid test statistics section.

--- 2021-07-11 src/paper/results/rapid_tests_statistics.tex
@@
-\subsection{Simulated Rapid Test Statistics}
+\subsection{Simulated Rapid Tests}
 \label{subsec:results_rapid_test_statistics}
@@
 negative rapid tests per million inhabitants per day. While this number might sound low
-in isolation, it becomes quite large when set in context: We estimate that at May first
-there are about 200 detected new infections per million inhabitants and day and roughly
-the same number of undetected cases. Thus there are only 400 per million inhabitants
-who could test falsely negative to begin with and not all of them will make a rapid
-test. This shows clearly that the large effect of rapid tests on the infection dynamic
+in isolation, it becomes quite large when set in context: We estimate that at at the
+the same time only 200 cases per million are detected by rapid tests. This means that
+even at the time where discovery of cases via rapid tests is at its peak, almost
+30\% of of people who are infected and make a test are not discovered by the test.
+
+This shows clearly that the large effect of rapid tests on the infection dynamic
 is not driven by unrealistic assumptions about their sensitivity but rather by the

### 2021-07-12 src/paper/model/estimated_parameters.tex
Add notation from the parameter tables to the report main text (mostly as comments).

@@
 
-We fit our model to data for Germany from October 2020 until June 2021. We do not use
-earlier periods for three reasons. Firstly, in the beginning PCR tests were very scarce
-and the reported case numbers unreliable. Secondly, during the summer the case
+We fit our model to data for Germany from mid September 2020 until June 2021. We do not
+use earlier periods for three reasons. Firstly, in the beginning PCR tests were very
+scarce and the reported case numbers unreliable. Secondly, during the summer the case
 numbers were extremely low. This could lead to the epidemic going extinct in our
 simulation. Thirdly, over the summer, imported cases from touristic travel were likely
-important for the infection dynamic but there is not enough data to include them into
-our model.
+important for the infection dynamic but there is not enough data to include them into our
+model.
 
@@
     \includegraphics[width=\textwidth]{figures/infection_channels.png}
-    \caption{\textbf{Daily share of infections by contact type}}
+    \caption{Daily share of infections by contact type}
     \floatfoot{\noindent \textit{Note:} Daily hare of infections that were contributed
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/other_multiplier}
+    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/other_multiplier}
     \caption{Share of Pre-Pandemic Other Contacts Taking Place with Infection Potential}

### 2021-07-12 src/paper/report.tex
Add notation from the parameter tables to the report main text (mostly as comments).

@@
 making it useful to predict the effects of various interventions. We apply the model to
-Germany, where new infections fell by almost 80\% during the month of May 2021. Our
-analysis shows that, aside from seasonality, frequent and large-scale rapid testing
-caused the bulk of this decrease, which is in line with prior predictions
-\citep{Mina2021}. We conclude that it should have a large role for at least as long as
-vaccinations have not been offered to an entire population.
+Germany, where new infections fell by almost 80\% during the May 2021. Our analysis shows
+that, aside from seasonality, frequent and large-scale rapid testing caused the bulk of
+this decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that
+it should have a large role for at least as long as vaccinations have not been offered to
+an entire population.
 
@@
 (Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
-approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
-model most closely related to ours is described in \citet{Hinch2020}.} Each contact
-between an infectious individual and somebody susceptible to the disease bears the risk
-of transmitting the virus. Contacts occur in up to four networks: Within the household,
-at work, at school, or in other settings (leisure activities, grocery shopping, medical
+approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The model
+most closely related to ours is described in \citet{Hinch2020}.} Each contact between an
+infectious individual and somebody susceptible to the disease bears the risk of
+transmitting the virus\comment[id=K]{infection probability per contact for contact type
+$c$: $\beta_{c}$}. Contacts occur in up to four networks: Within the household, at work,
+at school, or in other settings (leisure activities, grocery shopping, medical
 appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
 applications can take the population and household structure from census data and the
-network-specific frequencies of contacts from diary data measuring contacts before the
+network-specific frequencies of contacts\comment[id=K]{probability to have $n$ contacts
+of contact type $c$: $\eta_{c,\:n}$} from diary data measuring contacts before the
 pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
 depend on age and geographical location (see Supplementary Material
-Section~\ref{subsec:data_number_of_contacts}).
+Section~\ref{subsec:assortativity}).\comment[id=K]{Share of contacts of age group and
+county of i to meet someone of j's age group and county: $\alpha_{work,\:age_i,
+age_j,\:county_i,\:county_j}$}
 
-The four contact networks are chosen so that the most common NPIs can be modeled in
-great detail. NPIs affect the number of contacts or the risk of transmitting the disease
-upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from home will reduce the number of work
-contacts to zero for a fraction of the working population. Schools and daycare can be
-closed entirely, operate at reduced capacity---including an alternating schedule---, or
-implement mitigation measures like masking requirements or air filters
-\citep{Lessler2021}. Curfews may reduce the number of contacts in
-non-work/non-school/non-work settings. In any setting, measures like masking
-requirements would reduce the probability of infection associated with a contact
-\citep{Cheng2021}.
+The four contact networks are chosen so that the most common NPIs can be modeled in great
+detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
+having physical contact. The effect of different NPIs will generally vary across contact
+types. For example, a mandate to work from\comment[id=K]{$\rho_{attend,\,t}$} home will
+reduce the number of work contacts to zero for a fraction of the working population.
+Schools and daycare can be closed entirely, operate at reduced capacity---including an
+alternating schedule---, or implement mitigation measures like masking requirements or
+air filters \citep{Lessler2021}. Curfews\comment[id=K]{would enter into $\rho_{other\:,t}$} may reduce the number of contacts in
+non-work/non-school/non-work settings. In any setting, measures like masking requirements
+would reduce the probability of infection associated with a contact \citep{Cheng2021}.\comment[id=K]{$\rho_{hygiene}$}
 
@@
 
-    \floatfoot{\noindent \textit{Note:} A% formal
-        description of the model can be found in Supplementary Material~\ref{sec:model}.
-        Figure~\ref{fig:model_contacts_infections} shows the influence on an agent's
-        contacts to other agents. Demographic characteristics set the baseline number of
-        contacts in different networks. She may reduce the number of contacts due to
-        NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
-        occur when a susceptible agent meets an infectious agent; the probability depends
-        on the type of contact, on seasonality, and on NPIs. If infected, the infection
+    \floatfoot{\noindent \textit{Note:}
+        A~description of the model can be found in Supplementary
+        Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
+        influence on an agent's contacts to other agents. Demographic characteristics set
+        the baseline number of contacts in different networks ($\eta$). She may reduce
+        the number of contacts due to NPIs, showing symptoms, or testing positively for
+        SARS-CoV-2 ($\tau$). Infections may occur when a susceptible agent meets an
+        infectious agent; the probability depends on the type of contact ($\beta_c$), on
+        seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If infected, the infection
         progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
@@
         thresholds for triggering test demand differ across reasons and they may depend
-        on calendar time. Figure~\ref{fig:model_official_cases} shows the model of
-        translating all infections in the simulated data to age-specific recorded
-        infections. The model uses data on the aggregate share of recorded cases, on the
-        share of positive PCR tests triggered by symptoms, and on the false positive rate
-        of rapid tests. The lower part of the graph is relevant only for periods where
-        rapid tests are available.}
+        on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
+        Figure~\ref{fig:model_official_cases} shows the model of translating all
+        infections in the simulated data to age-specific recorded infections. The model
+        uses data on the aggregate share of recorded cases ($\psi$), on the share of
+        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and on the false
+        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part of
+        the graph is relevant only for periods where rapid tests are available.}
     \end{figure}
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
-possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
-differentiate between an initial period of infection without being infectious or showing
-symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar
-to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
-transitioning between these states depend on age; their duration is random within
-intervals calibrated to medical literature (for a detailed description see
-Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
-independent of age \citep{Jones2021}.
+In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
+age\comment[id=K]{$\zeta_{a}$}. A possible infection progresses as shown in
+Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period of
+infection without being infectious or showing symptoms, being infectious (presymptomatic
+or asymptomatic), showing symptoms, requiring intensive care, and recovery or death
+\citep[similar to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The
+probabilities of transitioning between these states depend on age;\comment[id=K]{$\gamma$
+for all transition probabilities} their duration is random within intervals calibrated to
+the medical literature (for a detailed description see Section~\ref{sec:medical_params}).
+Conditional on the type of contact, infectiousness is independent of age
+\citep{Jones2021}.\comment[id=K]{would be $\beta_{c}$}
 
-The model includes several other features, which are crucial to describe the evolution
-of the pandemic in 2020-2021. New virus strains with different profiles regarding
-infectiousness can be introduced. Agents may receive a vaccination. With a probability
-of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not transmit the
-virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is lower
-than what is usually reported for after the second dose of the Biontech/Pfizer vaccine,
-which is most commonly used in Germany. We choose it because our model neither includes
-booster shots, nor does it allow vaccinated individuals who became immune to transmit
-the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}. If anything, these
-assumptions would overstate the effect of vaccines for our study period. This would be
-different if a large fraction of vaccinated individuals had received a second dose
-already.} During the vaccine roll-out, priority may depend on age and occupation.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different profiles regarding
+infectiousness\comment[id=K]{$\sigma_v$} can be introduced. Agents may receive a
+vaccination. With a probability of 75\% \citep{Hunter2021}, vaccinated agents become
+immune and they do not transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
+Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the second
+dose of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We choose it
+because our model neither includes booster shots, nor does it allow vaccinated
+individuals who became immune to transmit the disease\citep{Petter2021,
+LevineTiefenbrun2021, Pritchard2021}. If anything, these assumptions would overstate the
+effect of vaccines for our study period. This would be different if a large fraction of
+vaccinated individuals had received a second dose already.} During the vaccine roll-out,
+priority may depend on age and occupation.\comment[id=K]{Maybe mention that we have
+refusers?}
 
@@
 individual is infected or not; there is no uncertainty to the result. PCR tests require
-some days to be processed and there are aggregate capacity constraints throughout. In
-contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
-that are demanded will be performed. Specificity and sensitivity of these tests is set
+some days to be processed\comment[id=K]{$\gamma_{PCR,\:d}$ where $d=$ number of days} and
+there are aggregate capacity constraints throughout. In contrast, rapid antigen tests
+yield immediate results; after a phase-in period, all tests that are demanded will be
+performed. Specificity\comment[id=K]{$p_{negative|not\:infected}$} and
+sensitivity\comment[id=K]{"$p_{positive|infected,\;i,\:t}$", where $i$ is the individual
+and $t$ is the date to determine time until/since infectiousness} of these tests is set
 according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
@@
 Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and because a
-household member tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{A positive test result (of either kind), as well as symptoms leads most
-individuals to reduce their contacts; this is why tests impact the actual contacts in
-Figure~\ref{fig:model_description}.}
+require staff and students to be tested regularly.\comment[id=K]{"$\pi_{teacher,\:t}$"
+etc. } Rapid tests may be offered by employers to on-site workers. Individuals may demand
+tests for private reasons, which include having plans to meet other people, showing
+symptoms of CoViD-19, and because a household member tested positively for the virus. We
+endow each agent with an individual compliance parameter. This parameter determines
+whether she takes up rapid tests.\footnote{A positive test result (of either kind), as
+well as symptoms leads most individuals to reduce their contacts; this is why tests
+impact the actual contacts in Figure~\ref{fig:model_description}.}
 
@@
 materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
-for different contact types and their assortativity by age group. Mobility data is used
-to model the evolution of work contacts. School and daycare policies can be incorporated
-directly from official directives. Administrative records on the number of tests,
-vaccinations by age and region, and the prevalence of virus strains are generally
-available. Surveys may ask about test offers, propensities to take them up, and past
-tests. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, these include infection probabilities
-by contact network and the effects of NPIs, see Supplementary
-Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method, also described in
-Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
-more specific once Appendix has new structure} In our application, we keep the number of
+for different contact types and their assortativity by age group\comment[id=K]{Should we
+add all our sources here?}. Mobility data is used to model the evolution of work
+contacts. School and daycare policies can be incorporated directly from official
+directives. Administrative records on the number of tests, vaccinations by age and
+region, and the prevalence of virus strains are generally available. Surveys may ask
+about test offers, propensities to take them up, and past tests. Other studies' estimates
+of the seasonality of infections can be incorporated directly. The remaining
+parameters---most notably, these include infection probabilities by contact network and
+the effects of some NPIs, see Supplementary Material~\ref{subsec:estimated_params}---will
+be chosen numerically so that the model matches features of the data \citep[see][for the
+general method, also described in Supplementary Material
+\ref{subsec:estimated_param}]{McFadden1989}. In our application, we keep the number of
 free parameters low in order to avoid overfitting. The data features to be matched
-include official case numbers for each age group and region, deaths, and the share of
-the B.1.1.7 strain.
+include official case numbers for each age group and region, deaths, and the share of the
+B.1.1.7 strain.
 
@@
 Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
-detected cases and use data on the share of PCR tests administered because of CoViD-19
-symptoms. As the share of asymptomatic individuals varies by age group, this gives us
-age-specific shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share
-of known cases by age group over time in our model). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is 80\% higher on average for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
+detected cases\comment[id=K]{$\psi$} and use data on the share of PCR tests administered
+because of CoViD-19 symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of
+asymptomatic individuals varies by age group, this gives us age-specific shares (see
+Figure~\ref{fig:share_known_cases_by_age_group} for the share of detected cases by age
+group over time in our model). Our estimates suggest that---in the absence of rapid
+testing---the detection rate is 80\% higher on average for individuals above age 80
+compared to school age children. Once rapid test become available, confirmation of a
+positive result\comment[id=K]{prob of demanding confirmation is $\chi_{confirmation}$} is
+another reason leading to positive PCR tests.
 
@@
 values represent higher levels of restrictions. A value of zero means all measures
-incorporated in the index are turned on. The value 1 represents the situation in
+incorporated in the index are turned on. The value one represents the situation in
 mid-September, with restrictions on gatherings and public events, masking requirements,
 but open schools and workplaces. In the seven weeks between mid September and early
-November, cases increased by a factor of 10. Restrictions were somewhat tightened in
+November, cases increased by a factor of ten. Restrictions were somewhat tightened in
 mid-October and again in early November. New infections remained constant throughout
@@
         Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
-        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
-        a value of 1 refers to the situation at the start of our sample period and 0
+        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
+        value of one refers to the situation at the start of our sample period and zero
         means that all NPIs included in the index are turned on. The other lines in
-        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
-        of contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~\ref{subsec:policies} for separate plots of the three
-        factors by contact type.}
+        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
+        contact reductions, increased hygiene regulations, and seasonality. See
+        Appendix~\ref{subsec:policies} for separate plots of the three factors by contact
+        type.}
 \end{figure}
@@
 use the model to predict recorded infection rates until the end of May 2021. See
-Supplementary Material~\ref{sub:initial_conditions} for a detailed description of this
-procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model's
-predictions are very close to officially recorded cases in the aggregate. This is also
-true for infections by age and geographical region, which are shown in the supplementary
-materials (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).
+Supplementary Material sections~\ref{subsec:synthetic_population} and
+\ref{sub:initial_conditions} for details. The blue line in
+Figure~\ref{fig:aggregated_fit} shows that our model's predictions are very close to
+officially recorded cases in the aggregate. This is also true for infections by age and
+geographical region, which are shown in the supplementary materials
+(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).
 
@@
 population, a vaccination rate that rose from 28\% to 43\%, and seasonality lowering the
-relative infectiousness of contacts, most strongly outside of work and school.
+relative infectiousness of contacts, most strongly for other contacts.
 
@@
 Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
-and three scenarios turning \replaced[id=K]{each factor on individually}{these factors
-off one by one}. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
-infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
-values \citep{Shapley2016} to decompose the difference in total infections between the
-scenario without any of the three factors and our main specification.
+and three scenarios turning each of these factors on individually.
+Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
+model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
+\citep{Shapley2016} to decompose the difference in total infections between the scenario
+without any of the three factors and our main specification.
 
@@
         to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and
-        seasonality remain at their January levels. The other scenarios turn these three
-        factors on one-by-one. The decompositions in
+        variant is introduced in the same way; vaccinations, rapid tests, and seasonality
+        remain at their January levels. The other scenarios turn each of these three
+        factors on individually. The decompositions in
         Figures~\ref{fig:2021_scenarios_decomposition} and
@@
 Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections with only one channel active,
-the final level is always the highest in case of the vaccination campaign (orange
-lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
-cumulative difference between scenarios. Reasons for this are the slow start---it took
-until March~24th until 10\% of the population had received their first vaccination, the
-20\% mark was reached on April 19th---and the focus on older individuals. These groups
+sight. Whether considering recorded or total infections with only one channel active, the
+final level is always the highest in case of the vaccination campaign (orange lines). The
+Shapley value decomposition shows that vaccinations contribute 16\% to the cumulative
+difference between scenarios. Reasons for this are the slow start---it took until
+March~24th until 10\% of the population had received their first vaccination, the 20\%
+mark was reached on April 19th---and the focus on older individuals. These groups
 contribute less to the spread of the disease than others due to a lower number of
-contacts see Supplementary Material~\ref{subsec:assortativity}. By the end of our
-study period, when first-dose vaccination rates reached around 40\% of the population,
-the numbers of new cases would have started to decline. It is important to note that the
+contacts (see Supplementary
+Material~\ref{subsec:assortativity}).\comment[id=K]{Assortativity is the wrong section
+here. They have few contacts in our model because they tend to live in small households
+and don't have work or educ contacts. They are also more likely to meet other old people
+but that's a different channel from what is written in the text.} By the end of our study
+period, when first-dose vaccination rates reached around 40\% of the population, the
+numbers of new cases would have started to decline. It is important to note that the
 initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
@@
 
-A similar-sized effect---41\% in the decompositions---comes from rapid testing. Here, it
+A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
 is crucial to differentiate between recorded cases and actual cases. Additional testing
@@
 undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large.
-Until late April, recorded cases are higher in the scenario with rapid testing alone
-when compared to the setting where none of the three mechanisms are turned on. The
-effect on total cases, however, is visible immediately in
+Until late April, recorded cases are higher in the scenario with rapid testing alone when
+compared to the setting where none of the three mechanisms are turned on. The effect on
+total cases, however, is visible immediately in
 Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
 population performed weekly rapid tests in March on average, new infections on April 1
-would be reduced by 53\% relative to the scenario without vaccinations, rapid tests, or
-seasonality.
+would have been reduced by 53\% relative to the scenario without vaccinations, rapid
+tests, or seasonality.
 
@@
 Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
-scenario without rapid tests only (purple line in
-Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
-channels for rapid tests. Tests at schools have the smallest effect, which is largely
-explained by schools not operating at full capacity during our period of study and the
-relatively small number of students.\footnote{18\% of our population are in the education
-sector (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost
-40\% come from tests at the workplace. Despite the fact that rapid tests for private
-reasons are phased in only late, they make up for more than half of the total effect. The
-reason lies in the fact that a substantial share of these tests is driven by an elevated
+scenario without rapid tests only and the main specification into the three channels for
+rapid tests. Tests at schools have the smallest effect, which is largely explained by
+schools not operating at full capacity during our period of study and the relatively
+small number of students.\footnote{18\% of our population are in the education sector
+(pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
+come from tests at the workplace. Despite the fact that rapid tests for private reasons
+are phased in quite late, they make up for more than half of the total effect. The reason
+lies in the fact that a substantial share of these tests is driven by an elevated
 probability to carry the virus, i.e., showing symptoms of CoViD-19 or following up on a
@@
 
-Two of the most contentious NPIs concern schools and mandates to work from home. In many
-countries, schools switched to remote instruction during the first wave, so did Germany.
-After the summer break, they were operating at full capacity with increased hygiene
-measures, before being closed again from mid-December onwards. Some states started
-opening them gradually in late February, but operation at normal capacity did not resume
-until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
-different policies regarding school starting at Easter, at which point rapid tests had
-become widely available. We estimate the realized scenario to have essentially the same
-effect as a situation with closed schools. Under fully opened schools with mandatory
-tests, total infections would have been 6\% higher; this number rises to 20\% without
-tests. These effect sizes are broadly in line with empirical studies
-\citep[e.g.,][]{Vlachos2021}. \comment[id=HM]{Need a back-of the envelope calculation} To use
-another metric, the effective weekly reproduction number differs by $0.018$ and $0.052$,
-respectively. In light of the large negative effects school closures have on children
-and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
-socio-economic status---these results in conjunction with hindsight bias suggest that
-opening schools combined with a testing strategy would have been beneficial. In other
-situations, and particular when rapid test are not available at scale, trade-offs may
-well be different.
+Two of the most contentious NPIs concern schools and mandates to work from
+home.\comment[id=K]{We should decide if we want to keep this section or prefer to show
+some rapid test statistics.} In many countries, schools switched to remote instruction
+during the first wave, so did Germany. After the summer break, they were operating at
+full capacity with increased hygiene measures, before being closed again from
+mid-December onward. Some states started opening them gradually in late February, but
+operation at normal capacity did not resume until the beginning of June.
+Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
+schools starting at Easter, at which point rapid tests had become widely available. We
+estimate the realized scenario to have essentially the same effect as a situation with
+closed schools. Under fully opened schools with mandatory tests, total infections would
+have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
+broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}. \comment[id=HM]{Need
+a back-of the envelope calculation} To use another metric, the effective weekly
+reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
+negative effects school closures have on children and parents \citep{Luijten2021,
+Melegari2021}---and in particular on those with low socio-economic status---these results
+in conjunction with hindsight bias suggest that opening schools combined with a testing
+strategy would have been beneficial. In other situations, and particular when rapid test
+are not available at scale, trade-offs may well be different.
 

### 2021-07-12 src/paper/data/medical_parameters.tex
Use greek letters in medical parameters section.

@@
 % Latency Period
-The first medical parameter we need is the length of the period between infection and
-the start of infectiousness, the so called latent period. \cite{Zhao2021} estimate the
-latent period to last 3.3 days (95\% CI: 0.2, 7.9). In line with this estimate our
-latent period lasts one to five days with a slightly lower average of 2 days.
+The first medical parameter we need is the distribution of the length of the period
+between infection and the start of infectiousness, the so called latent period,
+$\gamma_{infectious}$. \cite{Zhao2021} estimate the latent period to last 3.3 days (95\%
+CI: 0.2, 7.9). In line with this estimate our latent period lasts one to five days with a
+slightly lower average of two days.
 
@@
 nearly 70\% for the elderly.
-
-% Incubation Period
-The incubation period is usually estimated to be two to twelve days. A meta analysis by
-\citet{McAloon2020} comes to the conclusion that ``The incubation period distribution
-may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters
-(95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55),
-respectively.'' For simplicity we discretize this distribution into four bins.
-
 % Length of the presymptomatic stage
 Similar to \citet{Peak2020} and in line with \citet{He2020} we set the length of the
-presymptomatic stage to be one to two days.
+presymptomatic stage of age group $a$, $\gamma_{symptoms,\:a}$ to be one or two days. The
+probability to become symptomatic for age group $a$ is split equally between one and two
+days. This combined with our latency period leads to an incubation period that is in line
+with the meta analysis by \citet{McAloon2020}.
 
@@
 
-We assume that the duration of infectiousness is the same for both symptomatic and
-asymptomatic individuals as evidence suggests little differences in the transmission
-rates between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral
-load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020},
-\citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the duration of
-infectiousness is based on \citet{Byrne2020}. For symptomatic cases they arrive at zero
-to five days before symptom onset (see their figure 2) and three to eight days of
-infectiousness afterwards.\footnote{Viral loads may be detected much later but eight
-days seems to be the time after which most people are culture negative, as also reported
-by \citet{Singanayagam2020}.} Thus, we arrive at zero to 13 days as the range for
-infectiousness among individuals who become symptomatic (see also figure 5).
+We assume that the duration of infectiousness ($\gamma_{stop\:infectious}$) is the same
+for both symptomatic and asymptomatic individuals as evidence suggests little differences
+in the transmission rates between symptomatic and asymptomatic patients (\citet{Yin2020})
+and that the viral load between symptomatic and asymptomatic individuals are similar
+(\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the
+duration of infectiousness is based on \citet{Byrne2020}. For symptomatic cases they
+arrive at zero to five days before symptom onset (see their figure 2) and three to eight
+days of infectiousness afterwards.\footnote{Viral loads may be detected much later but
+eight days seems to be the time after which most people are culture negative, as also
+reported by \citet{Singanayagam2020}.} Excluding the most extreme combinations, we arrive
+at 3 to 11 days as the duration of infectiousness.
 
@@
 We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
-Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases.
-We only disaggregate by age how likely individuals are to require intensive
-care.\footnote{The length of symptoms is not very important in our model given that
-individuals mostly stop being infectious before their symptoms cease.}
-
-% Time from Symptom Onset to Admission to ICU ------------------------------------------
+Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases
+($\gamma_{stop\:symptoms}$). We only disaggregate by age how likely individuals
+are to require intensive care. %
 
 For the time from symptom onset until need for intensive care we rely on data by
-\cite{Stokes2020}) and \citep{Hinch2021}.
-
-For those who will require intensive care we follow \citet{Chen2020} who estimate the
-time from symptom onset to ICU admission as $8.5 \pm 4$ days. This aligns well with
-numbers reported for the time from first symptoms to hospitalization:
-
-\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
-that the time between symptom onset and ICU takes four, six, eight or ten days with
-equal probabilities.
+\cite{Stokes2020}) and \cite{Hinch2021} ($\gamma_{icu,\:a}$). For those who will require
+intensive care we follow \citet{Chen2020} who estimate the time from symptom onset to ICU
+admission as $8.5 \pm 4$ days. This aligns well with numbers reported for the time from
+first symptoms to hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a
+standard deviation of four. We assume that the time between symptom onset and ICU takes
+four, six, eight or ten days with equal probabilities.
 
 % Death and Recovery from ICU ----------------------------------------------------------
-
-We take the survival probabilities and time to death and time until recovery from
-intensive care from the \citet{Hinch2021}. They report time until death to have a mean
-of 11.74 days and a standard deviation of 8.79 days. To match this approximately we
-discretize that 41\% of individuals who will die from Covid-19 do so after one day in
-intensive care, 22\% day after twelve days, 29\% after 20 days and 7\% after 32 days.
-Again, we rescale this for every age group among those that will not survive. For
-survivors the OpenABM Project reports a mean duration of 18.8 days until recovery and a
-standard deviation of 12.21 days. We discretize this such that of those who recover in
-intensive care, 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and
-18\% after 45 days.
+We take the survival probabilities and time to death and time until recovery
+($\gamma_{stop\:icu\:a}$ and $\gamma_{dead,\:a}$) from intensive care from
+\citet{Hinch2021}. They report time until death to have a mean of 11.74 days and a
+standard deviation of 8.79 days. To match this we discretize that 41\% of individuals who
+will die from Covid-19 do so after one day in intensive care, 22\% day after twelve days,
+29\% after 20 days and 7\% after 32 days. Again, we rescale this for every age group
+among those that will not survive. For survivors \cite{Hinch2020} reports a mean duration
+of 18.8 days until recovery and a standard deviation of 12.21 days. We discretize this
+such that of those who recover in intensive care, 22\% do so after one day, 30\% after 15
+days, 28\% after 25 days and 18\% after 45 days.
 

### 2021-07-12 src/paper/results/fit.tex
Use greek letters in medical parameters section.

@@
 Overall, our model achieves an excellent fit of the two waves of infections with few free
-parameters (Figure~\ref{fig:aggregated_fit2}. As a result the effective replication
-number is similar to that reported by the RKI (see Figure~\ref{fig:fit_r_effective}).
-This excellent fit is also achieved for most age groups in Germany. The fit is also good
-for many German federal states. Despite, the fact that the number of performed rapid
-tests and their distribution in the population result endogenously inside our model we
-fit the share of the population with at least a weekly rapid test very well and err on
-the side of too few individuals that have ever done a rapid test.
+parameters (Figure~\ref{fig:aggregated_fit2}). As a result the effective replication
+number also closely follows that reported by the RKI (see
+Figure~\ref{fig:fit_r_effective}). We also achieve an excellent fit for most age groups
+in Germany. The fit is also good for many German federal states. Despite, the fact that
+the number of performed rapid tests and their distribution in the population result
+endogenously inside our model we fit the share of the population with at least a weekly
+rapid test very well and err on the side of too few individuals that have ever done a
+rapid test.
 
@@
 excellent. The incidence in our model matches both the levels and the shape of the
-reported incidence almost perfectly. When the prevalence of the virus is high and during
-phases of exponential growth, the effect of random events on the incidence is large. That
-is why it's important that we rely on 30 simulation runs with different seed sequences to
-average over these random differences and have a stable mean.
+reported incidence almost perfectly. When the prevalence of the virus is high
+\textcolor{red}{and during exponential growth phases} the effect of random events on the
+incidence is large. That is why it's important that we rely on 30 simulation runs with
+different seed sequences and average over these random differences to have a mean that's
+robust to the randomness in our model.
 
@@
   the average number of new infections caused by one infected individual. The $R_t$ in
-  our model broadly follows the $R_t$ reported by the RKI. Two trends stand out. Firstly,
-  the RKI's $R_t$ drops faster in November. This could be due to a change in the testing
-  policy that focused tests on the elderly when the second wave hit Germany and led to a
-  decline in the overall share of detected cases. The second difference is from mid
-  February to mid March where the RKI's reported $R_t$ increased more rapidly than that
-  in our model. Here the opposite effect can be expected. During this time rapid tests
-  increased strongly leading to more cases being detected. In the short term this leads
-  an $R_t$ estimation that is based on detected cases to overestimate the replication
-  number.}
+  our model broadly follows the $R_t$ reported by the RKI. Two differences stand out.
+  Firstly, the RKI's $R_t$ drops faster in November. \textcolor{red}{This is likely due
+  \st{to a change in the testing policy that focused tests on the elderly when the second
+  wave hit Germany and led } to a decline in the estimated overall share of detected
+  cases}. %when the second wave hit Germany.
+  The second difference is from mid February to mid March where the RKI's reported $R_t$
+  increased more rapidly than that in our model. Here the opposite effect can be
+  expected. During this time rapid tests increased strongly leading to more cases being
+  detected. In the short term this leads an $R_t$ estimation that is based on detected
+  cases to overestimate the replication number.}
 \end{figure}
@@
 between the age groups allow us to identify the infection probabilities because different
-age groups have contacts of different types (e.g. Mostly children go to school, the share
+age groups have contacts of different types (e.g. mostly children go to school, the share
 of workers is highest in the 35 to 59 age bracket etc.). There are very few points where
@@
 Figure~\ref{fig:share_known_cases_by_age_group}). However, we follow \cite{Betsch2021}
-that 82\% of all rapid tests are followed up with a PCR test. It is likely that that
-share is higher for rapid tests that were done in schools. The last age group where our
-fit does not fit the empirical curve very well is the group of 15 to 34 year olds. This
-group has been touted to have a very active social life and it has often been suspected
-that this age group also stayed more socially active during the pandemic. However, our
-model does not include either effect because of very little data on both and hence it's
-unsurprising that the actual case numbers in this group are consistently higher than our
-simulated case numbers.\comment[id=K]{Does anyone have an idea why we do not match the
-shape from Oct to Dec?}
+and assume that 82\% of all rapid tests are followed up with a PCR test---irrespective of
+why they took place. It is likely that that share is higher for rapid tests that were
+done in schools. The last age group where our fit does not fit the empirical curve very
+well is the group of 15 to 34 year olds. This group has been touted to have a very active
+social life and it has often been suspected that this age group also stayed more socially
+active during the pandemic. However, our model does not include either effect because of
+very little data on both and hence it's unsurprising that the actual case numbers in this
+group are consistently higher than our simulated case numbers.\comment[id=K]{Does anyone
+have an idea why we do not match their shape from Oct to Dec?}
 
@@
   to 100 year olds live in these facilities. However, the official data does not contain
-  information on whether cases were nursing home inhabitants or not. We averaged over 30
+  information on whether cases were nursing home inhabitants or not. We average over 30
   simulation runs.}
@@
 but they fell during Nov to 220 and then climbed again to 800 by Jan 12th so that might
-be only part of the answer why Saxony had such high numbers.}
+be only part of the story why Saxony had such high numbers.}
 
@@
 few cases in January as can be seen in Figure~\ref{fig:fit_share_b117}. Since we only
-model B.1.1.7 and do not include other variants B.1.1.7 reaches a share of nearly 100\%
-by May while the true rate plateaued at 90\% and B.1.1.7 was then replaced by B.1.617.2
-as the dominant variant in June. However, B.1.1.7 still made up over 90\% of all cases in
-Germany at the end of our simulation period.
+model B.1.1.7 and do not include other variants, B.1.1.7 reaches a share of nearly 100\%
+by May while the true rate plateaued at 90\% as B.1.617.2 started gained traction in
+Germany. Given that B.1.617.2 made up less than 5\% even at the end of our simulation
+period, we did not include it in our model.
 
@@
   the cause of January. From then B.1.1.7 takes over endogenously through its increased
-  infectiousness. We model no other features of B.1.1.7. At most we introduce 0.75 cases
+  infectiousness. We model no other features of B.1.1.7. We introduce at most 0.986 cases
   per 100,000 inhabitants. The right figure shows the rate of individuals that are
   vaccinated in our synthetic population versus in the general German population. Note
-  that we excluded the vaccinations that were given to nursing homes, approximately the
-  first percent of the German population that were vaccinated. Overall, our model covers
-  a time frame that goes from zero vaccinated individuals to a state where over 40\% of
-  the population are vaccinated. Our vaccinations work imperfectly but we do not model
-  different vaccines nor do we distinguish between first and second shot.}
+  that we excluded the vaccinations that were given to nursing homes (approx. one percent
+  of the German population) that were vaccinated. Overall, our model covers a time frame
+  that goes from zero vaccinated individuals to a state where over 40\% of the population
+  are vaccinated. Our vaccinations work imperfectly but we do not model different
+  vaccines nor do we distinguish between first and second shot.}
 \end{figure}
@@
 equivalent to the share of Germans living in nursing homes. Afterwards, the share of
-vaccinated individuals in the population follows the German increase quite exactly. We
-took great care to model the prioritization of older individuals and professions that
-cannot reduce physical contact easily such as teachers or medical staff (see
+vaccinated individuals in the population follows the German increase exactly. We took
+great care to model the prioritization of older individuals and professions that cannot
+reduce physical contact easily such as teachers or medical staff (see
 Section~\ref{subsec:synthetic_population}. How the share of vaccinated individuals
@@
 the number of random tests and who received them. Secondly, there is no randomness in an
-individual's rapid test demand. This reduces the randomness in our model at the cost of
-having less individuals that only occasionally test themselves. \comment[id=K]{Write that
-this makes us underestimate the effect of random tests?} However,
-Section~\ref{subsec:model_validation} shows that our results are very robust to changes
-in the exact shares of individuals demanding rapid tests.
+individual's rapid test demand in our model. This reduces the randomness in our model at
+the cost of having less individuals that only occasionally test themselves.
+ \comment[id=K]{Write that this makes us underestimate the effect of random tests?}
+However, Section~\ref{subsec:model_validation} shows that our results are very robust to
+changes in the exact shares of individuals demanding rapid tests.
 
@@
         have ever done a rapid test or done a rapid test within the last week in our
-        simulations to the shares reported in \citet{Betsch2021b}. The left panel compares the share of individuals who
-        have ever done a rapid test. The right panel compares the share of individuals
-        who have done a rapid test within the last seven days in our simulation compared
-        to the share reporting to have done at least weekly rapid tests in the last four
-        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
-        conservative. The overall share is below that in the study. We fit the share of
-        weekly tests quite exactly. However, the study only covers adults while our share
-        also includes children who are tested very regularly when attending school.}
+        simulations to the shares reported in \citet{Betsch2021b}. The left panel
+        compares the share of individuals who have ever done a rapid test. The right
+        panel compares the share of individuals who have done a rapid test within the
+        last seven days in our simulation compared to the share reporting to have done at
+        least weekly rapid tests in the last four weeks in the COSMO survey. Overall our
+        calibration of rapid tests are slightly conservative. The overall share is below
+        that in the study. We fit the share of weekly tests quite exactly. % However, the
+        % study only covers adults while our share also includes children who are tested
+        % very regularly when attending school.
+      }
 \end{figure}

### 2021-07-12 src/paper/data/number_of_contacts.tex
Use parameter notation everywhere in the data section.

@@
 the other person is. Binning the number of contacts for very high numbers, we arrive at
-the distributions of the numbers of contacts by contact type as shown in Figure~\ref{fig:n_contacts}.
+the distributions of the numbers of contacts by type of contact ($\eta_{c}$) as shown in
+Figure~\ref{fig:n_contacts}.
 
@@
         \textit{Note:} This figure shows the pre-pandemic number of contacts individuals
-        report of different contact types.
+        report of different contact types ($\eta_c$).
         % non recurrent contacts
@@
         The pre-pandemic number of contacts with transmission potential is reduced by
-        policies, seasonality and individual responses to events such as
-        receiving a positive rapid test to the number of actual contacts with
-        transmission potential.
+        policies ($\rho$), seasonality ($\kappa$) and individual responses ($\tau$) to
+        events such as receiving a positive rapid test to the number of actual contacts
+        with transmission potential.
         % other
         The upper row shows the distribution of the number of other contacts individuals
-        report. Other contacts include all contacts that are not household members,
-        school contacts or work contacts, for example leisure contacts. We assume that
-        individuals in households with children or teachers or retired individuals have
-        additional non recurrent other contacts during school vacations to cover things
-        like family visits or travel during vacations.
+        report ($\eta_{other}$). Other contacts include all contacts that are not
+        household members, school contacts or work contacts, for example leisure
+        contacts. We assume that individuals in households with children or teachers or
+        retired individuals have additional non recurrent other contacts during school
+        vacations to cover things like family visits or travel during vacations.
         % work contacts
-        The lower row shows the distribution of the different types of work contacts.
-        Work contacts only take place between working individuals.}
+        The lower row shows the distribution of the different types of work contacts
+        ($\eta_{work}$). Work contacts only take place between working individuals.}
 \end{figure}

### 2021-07-12 src/paper/data/policies.tex
Use parameter notation everywhere in the data section.

@@
 
-Our policies usually affect one of our four contact types: households, education, work
-and other contacts.
+Our policies (denoted by $\rho$) usually affect one of three contact types: education,
+work and other contacts.
 % households
-For households we assume that the individuals' contacts in their households do not change
-over our estimation period.
+Germany had no policies limiting contacts within households so there are no policies on
+them in our model.\footnote{Household contacts can, however, be reduced when individuals
+quarantine themselves after developing symptoms, for example. This happens to a lesser
+degree than other contacts to capture difficulties in isolation within the home.}
 % educ models
@@
 federal states as well as school closures, emergency care and rotating schooling
-schedules where only one half of students attends every other week or day.% For the moment we ignore that lack of childcare leads working parents to stay home.
+schedules where only one half of students attends every other week or day.
+% For the moment we ignore that lack of childcare leads working parents to stay home.
 An approximation of the share of contacts still taking place with the different school
@@
 For work contacts we use the reductions in work mobility reported by the Google Mobility
-Data \citep{Google2021} to calibrate our work policies. Reductions in work contacts are
-not random but governed through a work contact priority where the policy changes the
-threshold below which workers stay home. Figure \ref{fig:work_multiplier} shows the share
-of workers that go to work over time at the federal German level. We use the data on the
-state level to account for local holidays and differences in state regulations.
+Data \citep{Google2021} to calibrate the reduction in physical work contacts
+($\rho_{w,\:attend,\:t}$). Reductions in work contacts are not random but governed
+through a work contact priority where the policy changes the threshold below which
+workers stay home. Figure \ref{fig:work_multiplier} shows the share of workers that go to
+work over time at the federal German level. We use the data on the state level to account
+for local holidays and differences in state regulations.
 % hygiene umltiplier
@@
 in November, leading to a reduction of 33\% in the number of contacts with the potential
-to transmit Covid-19.
+to transmit Covid-19 ($\rho_{hygiene}$).
 
@@
     vacations (by federal state) just like the schooling mode (full operation, emergency
-    care, rotating schemes with half class sizes etc.). The figure is thus only an
+    care, rotating schemes with half class sizes etc.). The figure is, thus, only an
     illustration that shows the approximate share of contacts taking place compared to
@@
     reported by \cite{Google2021}. We take this as a proxy of the share of workers who
-    still have physical work contacts. The figure interpolates over weekends as we handle
-    weekend effects through information on work on weekends in the German census data we
-    use. The figure shows the share for Germany as a whole. To capture the effect that
-    local policies, school vacations, etc. have on work contacts we use the data on the
-    state level to determine which workers go to work depending on the state they live
-    in.}
+    still have physical work contacts ($\rho_{w,\:attend,\:t}$). The figure interpolates
+    over weekends as we handle weekend effects through information on work on weekends in
+    the German census data we use. The figure shows the share for Germany as a whole. To
+    capture the effect that local policies, school vacations, etc. have on work contacts
+    we use the data on the state level to determine which workers go to work depending on
+    the state they live in.}
     \label{fig:multipliers}
@@
 
-Lastly, for the other contacts category we could not calibrate the policies from data but
-estimated the policy effects. The estimation and values are detailed in
-Section~\ref{subsec:estimated_params} and Figure~\ref{fig:other_multiplier}.
+Lastly, for the other contacts category ($\rho_{other,\:t}$) we could not calibrate the
+policies from data but estimated the policy effects. The estimation and values are
+detailed in Section~\ref{subsec:estimated_params} and Figure~\ref{fig:other_multiplier}.
 

### 2021-07-12 src/paper/data/rapid_test_demand.tex
Make data section on share known cases ready for review. Add figures showing the parameter demand shares in different robustness scenarios.
Fixed document, some typos, removed duplicated parameter.
Use parameter notation everywhere in the data section.

@@
 
-In our model, there are five reasons why rapid tests are done:\comment[id=J]{ Add a
-    section on how we calibrate rapid test demand; Mainly describe the datapoints we have
-    and say that we usually interpolate linearly in between data points. (Only exception
-    to that is private rapid test demand, which we fit to data)
-}
+In our model, there are five reasons why rapid tests are done:
 
@@
 For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April)
-that 60\% of workers who receive a test offer by their employer regularly use it. We
-assume this share to be time constant.
+that 60\% of workers who receive a test offer by their employer regularly use it
+($\pi_{w,\:d}$). We assume this share to be time constant.
 
 In addition, there are some surveys that allow us to trace the expansion of employers who
-offer tests to their employees. Mid march, 20\% of employers offered tests to their
-employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
-offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
-61\% until the first days of April \citep{BMWI2021, IZA2021}.
-
-Until mid April 72\% of workers were expected to receive a weekly test offer
-\citep{BMWI2021, IZA2021}. However, according to surveys conducted in mid April
-\citep{Betsch2021}, less than two thirds of individuals with work contacts receive a test
-offer. Starting on April 19th employers were required by law to provide two weekly tests
-to their employees \citep{Bundesanzeiger2021}. We assume that compliance is incomplete
-and only 80\% of employers actually offer tests. We interpolate between these points
-linearly, arriving at the blue line in Figure~\ref{fig:rapid_test_demand}.
+offer tests to their employees ($\pi_{w,\:s,\:t}$). Mid march, 20\% of employers offered
+tests to their employees \citep{DIHK2021}. In the second half of March, 23\% of employees
+reported being offered weekly rapid tests by their employer \citep{Ahlers2021}. This
+share increased to 61\% until the first days of April \citep{BMWI2021, IZA2021}. Until
+mid April 72\% of workers were expected to receive a weekly test offer \citep{BMWI2021,
+IZA2021}. However, according to surveys conducted in mid April \citep{Betsch2021}, less
+than two thirds of individuals with work contacts receive a test offer. Starting on April
+19th employers were required by law to provide two weekly tests to their employees
+\citep{Bundesanzeiger2021}. We assume that compliance is incomplete and only 80\% of
+employers actually offer tests. We interpolate between these points linearly, arriving at
+the blue line in Figure~\ref{fig:rapid_test_demand}. In addition, we increase the
+frequency of testing ($\theta_{t,\:work}$) from weekly to twice weekly during April.
 
@@
 We assume that employees in educational facilities start getting tested in 2021 and that
-by March 1st 30\% of them are tested weekly. The share increases to 90\% for the week
-before Easter. At that time both Bavaria \citep{STMGP2021} and Baden-Württemberg
+by March 1st 30\% of them ($\pi_{teacher,\:t}$) are tested weekly
+($\theta_{before\:Easter,\:educ} = 7$). The share increases to 90\% for the week before
+Easter. At that time both Bavaria \citep{STMGP2021} and Baden-Württemberg
 \citep{MinisteriumKultus2021} were offering tests to teachers and North-Rhine Westphalia
@@
 students and tests for students and teachers were already mandatory in Saxony
-\citep{SMK2021}. After Easter we assume that 95\% of teachers get tested twice per week.
+\citep{SMK2021}. After Easter we assume that 95\% of teachers get tested twice per week
+($\theta_{after\:Easter,\:educ} = 3$).
 
@@
 so we assume that they only start in February and only 10\% of students get tested by
-March 1st. Relying on the same sources as above we approximate that by the week before
-Easter this share had increased to 40\% \citep{SchulministeriumNRW2021}.
-
-After Easter the share of students receiving twice weekly tests is set to 75\%. This is
-based on tests becoming mandatory in Bavaria \citep{BayerischeStaatskanzlei2021} and
-North Rhine-Westphalia \citep{SchulministeriumNRW2021b} after their Easter breaks and on
-the 19th in Baden-Württemberg \citep{KMBaWue2021}. Again, we interpolate linearly between
-these points and arrive at the purple line for teachers and the red line for school
-students in Figure~\ref{fig:rapid_test_demand}.
+March 1st ($\pi_{students,\:t}$). Relying on the same sources as above we approximate
+that by the week before Easter this share had increased to 40\%
+\citep{SchulministeriumNRW2021}. After Easter the share of students receiving twice
+weekly tests is set to 75\%. This is based on tests becoming mandatory in Bavaria
+\citep{BayerischeStaatskanzlei2021} and North Rhine-Westphalia
+\citep{SchulministeriumNRW2021b} after their Easter breaks and on the 19th in
+Baden-Württemberg \citep{KMBaWue2021}, after which we assume twice weekly rapid tests to
+be mandatory for all students in Germany. Again, we interpolate linearly between these
+points and arrive at the purple line for teachers and the red line for school students in
+Figure~\ref{fig:rapid_test_demand}.
 
@@
 To limit our degrees of freedom, we only have one parameter that governs how many
-individuals do a rapid test because of any of the private demand reasons (own symptoms
-but no PCR test, planned weekly leisure meeting or a symptomatic or positively tested
-household member).
-
+individuals do a rapid test because of any of the private demand reasons
+($\pi_{private,\:t}$).\footnote{The reasons that can lead to an individual doing a rapid
+test for private reasons are own symptoms but no PCR test, planned weekly leisure meeting
+or a symptomatic or positively tested household member}
 We assume that there is no private rapid test demand until March when both the citizens'
@@
 in our model. The resulting share of private rapid test demand is shown as the green line
-in Figure~\ref{fig:rapid_test_demand}.
+in Figure~\ref{fig:rapid_test_demand} (also see Section~\ref{subsec:estimated_params} for
+details on the estimation).
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
+    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/rapid_test_demand/baseline_shares}
     \caption{Share of Individuals Doing a Rapid Test.}
-    \label{fig:rapid_test_demand}
     \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
-    planning to have education contacts, work contacts, developing symptoms without
-    access to a PCR test, having a household member with a positive test or symptoms. In
-    each case whether a rapid test is done depends on how long it has been since the
-    individual's last rapid test and her individual compliance parameter. As an example,
-    take a worker in May. In that time workers are encouraged to test themselves twice
-    weekly but there is no general requirement to test themselves. If the worker has not
-    done a test within the last four days in our model she will demand a test if her
+    planning to have education contacts ($\pi_{teacher,\:t}$ or $\pi_{students,\:t}$),
+    work contacts ($\pi_{w,\:d}$ and $\pi_{w,\:s,\:t}$), developing symptoms without
+    access to a PCR test, having a household member with a positive test or symptoms
+    ($\pi_{private,\:t}$). In each case whether a rapid test is done depends on how long
+    it has been since the individual's last rapid test and her individual compliance
+    parameter. As an example, take a worker in May. In that time workers are encouraged
+    to test themselves twice weekly (or every three days, i.e. $\theta_{May,\:work} = 3$)
+    but there is no general requirement to test themselves. If the worker has not done a
+    test within the last four days in our model she will demand a test if her
     (time-constant) compliance parameter belongs to the upper 60\% in the population.}

### 2021-07-12 src/paper/data/share_known_cases.tex
Use parameter notation everywhere in the data section.

@@
 tests). For our model it is important to have an estimate for the share of cases that is
-detected in the absence of rapid tests. For this we rely on the Dunkelzifferradar Project
-\citep{Dunkelzifferradar2020} which relies on estimates of the case fatality rate to
-estimate the number of total cases given the number of CoViD-19 deaths which are assumed
-to be perfectly observable. For 2020, we follow the reported share of detected cases
-quite closely only interpolating the phase of November 2020 where the testing policy was
-changed to give more priority to vulnerable groups. This lead to more tests being done
-among older age groups which are also at the highest risk of dying from CoViD-19 and an
-overestimation of the share of detected cases.
+detected in the absence of rapid tests ($\psi_t$). For this we rely on the
+\cite[Dunkelzifferradar Project][]{Dunkelzifferradar2020} which uses estimates of the case
+fatality rate to estimate the number of total cases given the number of CoViD-19 deaths
+which are assumed to be perfectly observable. For 2020, we follow the reported share of
+detected cases quite closely \textcolor{red}{only interpolating the phase of November
+2020 where the testing policy was changed to give more priority to vulnerable groups
+\citep{RKI2020a}. This lead to more tests being done among older age groups which are
+also at the highest risk of dying from CoViD-19. We suspect that this led the estimated
+share of detected cases to increase without a corresponding increase in the share of
+detected cases in the overall population as PCR tests, a rare resource at the time, were
+only moved to groups with a high case fatality ratio.}
+
 % After Christmas 2020
@@
 This is why we stop using the share of detected cases estimated by the Dunkelzifferradar
-after Christmas. Instead, we assume that the share of detected cases would have
-stayed the same in the absence of rapid tests. Thus, we also achieve in our model an
-increase in the share of detected cases but this is driven from inside our model through
-increased rapid testing and follow-up on positive rapid tests (see
+after Christmas. Instead, we assume that the share of detected cases would have stayed
+the same in the absence of rapid tests. Thus, we also achieve in our model an increase in
+the share of detected cases but this is driven from inside our model through increased
+rapid testing which lead follow-up PCR tests when they are positive (see
 Section~\ref{subsec:results_share_known_cases} and \ref{sub:testing}).
+
 % Christmas and Easter
@@
   reported as an official case via PCR confirmation. We use the overall share of known
-  cases that was estimated through the case fatality ratio by the
-  \citet{Dunkelzifferradar2020} for all of 2020 and
-  then assume it to be constant as vaccinations of the elderly strongly affect the case
-  fatality rate which the project does not account for. Starting in 2021 in addition to
-  the overall numbers of detected cases through symptoms and the share known cases, cases
-  are also detected through confirmation of positive rapid tests which happens
-  endogenously inside the model. For the public holidays of Christmas and Easter we lower
-  the share of detected cases as fewer PCR tests are available during public holidays.}
+  cases that was estimated through the case fatality ratio by the \cite[Dunkelzifferradar
+  Project][]{Dunkelzifferradar2020} for all of 2020 and then assume it to be constant as
+  vaccinations of the elderly strongly affect the case fatality rate which the project
+  does not account for. Starting in 2021 in addition to the overall numbers of detected
+  cases through symptoms and a random component, cases are also detected through
+  confirmation of positive rapid tests which happens endogenously inside the model. For
+  the public holidays of Christmas and Easter we lower the share of detected cases as
+  fewer PCR tests are available during public holidays. See
+  Figure~\ref{fig:share_known_cases_by_age_group} for how the share of detected cases
+  develops in our model for each age group}.
 \end{figure}

### 2021-07-12 src/paper/data/synthetic_population.tex
Use parameter notation everywhere in the data section.

@@
 the same individuals that continue to have work contacts when work from home mandates of
-a certain strictness are in place. We use the reduction in work mobility reported by
-\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.
+a certain strictness are in place.
 
@@
 contacts. To determine who would reduce their contacts in such a situation or demand a
-rapid test we introduce a quarantine compliance parameter. Similarly we introduce a rapid
-test compliance parameter that determines in which order individuals start demanding
-rapid tests when rapid tests become increasingly available. This makes sure that when for
-example only 10\% of workers get tested, it's the same workers that have access to tests
-every week.
+rapid test we introduce a quarantine compliance parameter. Similarly, we introduce a
+rapid test compliance parameter that determines in which order individuals start
+demanding rapid tests when rapid tests become increasingly available. This makes sure
+that when for example only 10\% of workers get tested, it's the same workers that have
+access to tests every week.
 
@@
 group and a vaccination rank from that group that creates a complete vaccination queue
-over the population including 15\% that refuse to be vaccinated as reported by the
-Robert Koch Institute (RKI, \cite{RKI2021c}). The vaccination groups are created to
-match the recommendations by the Ständige Impfkommission
-\citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized more than
-recommended by the commission.} To cover that the Pfizer-BioNTech vaccine was later
-approved for younger age groups we put adolescents and children into two groups that
-follow after the general population. These groups do not become eligible within our
-simulation frame until June. The way vaccinations are rolled out in our model is shown
-in Figure~\ref{fig:vaccinations_by_age_group}.
+over the population including a share that refuses to be vaccinated ($\xi$) which we
+calibrate to  15\% \citep{RKI2021c}. The vaccination groups are created to match the
+recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
+that teachers were prioritized more than recommended by the commission.} To cover that
+the Pfizer-BioNTech vaccine was later approved for younger age groups we put adolescents
+and children into two groups that follow after the general population. These groups do
+not become eligible within our simulation frame until June. The way vaccinations are
+rolled out in our model is shown in Figure~\ref{fig:vaccinations_by_age_group}.
 
@@
   work contact priority, her age group and a random component to capture medical
-  preconditions like diabetes. 15\% of the population refuse to be vaccinated.
+  preconditions like diabetes. 15\% of the population refuse to be vaccinated ($\xi$).
   Adolescents would be vaccinated after the general population and children last. The

### 2021-07-12 src/paper/literature_review.tex

@@
 
-We build on two strands of literature: Recent extensions of the epidemiological SEIR
-model and agent-based simulation models.
-
+The workhorse model in epidemiology is the SEIR model. However,
 The traditional SEIR model is not fine-grained enough to model nuanced policies. This
@@
 Another commonly used model class in epidemiology are agent-based simulation
-models.\comment[id=K]{See
-\href{https://twitter.com/Patrick_M_Econ/status/1410543606939721730?s=20}{this tweet} for
-a reference to some models that are similar to ours} In these models individuals are
+models. In a prototypical agent-based simulation model, individuals are
 simulated as moving particles. Infections take place when two particles come closer than
@@
 
-\citet{Hinch2020} is a recent extension of the prototypical agent-based simulation model
-that replaces moving particles by contact networks for households, work and random
-contacts. This model is similar in spirit to ours but focuses on contact tracing rather
-than social distancing policies.
+These shortcomings have motivated variations of agent-based simulation models where
+moving particles have been replaced by contact networks for households, work and random
+contacts.
 
-The above assessment of epidemiological models is not meant as a critique. We are aware
-that these models were not designed to predict the effect of fine-grained social
-distancing policies in real time and are very well suited to their purpose. We invite
-epidemiologists to provide feedback and collaborate to improve our model.
+The OpenABM-Covid-19 model by \citet{Hinch2020} is the closest in spirit to ours. They
+model detailed contact networks for workplaces, schools and households and can evaluate
+the effect of several NPIs. The main focus of their application are contact tracing
+policies.
+
+\citet{Aleta2020} develop an agent-based simulation model with a very high geographical
+resolution by estimating contact networks from fine grained mobility data for the
+Boston metropolitan area. They use this model to show how NPIs, contact tracing and PCR
+testing can influence the infection dynamics. However, they do not calibrate their model
+match actual infection numbers which makes it more suitable to explore the general
+mechanics of different disease mitigation measures than for their quantitative
+evaluation.
+
+\citet{Bicher2021} implement a model that simulates the complete Austrian population
+with almost nine million people. It uses the same data provided by \citet{Mossong2008}
+as we use to calibrate contact networks for households, workplaces and schools. The
+model focuses and analyzing the effect of different contact tracing strategies. It is
+calibrated with data from the first wave (February 21 to April 9).
+
+Moreover, there are several forthcoming applications of agent-based simulation models
+with contact networks. Examples are \citet{Basurto2020}, \citet{DelliGatti2020} and
+\citet{Mellacher2020}.
+
+Our model combines elements from the above models and adds several new features. To
+the best of our knowledge, ours is the only model that can explain observed case
+numbers with an excellent fit over more than 9 Months. We also have the most fine
+grained model of schools and NPIs that affect schools. Finally, we have an extremely
+detailed model of PCR and rapid tests with a share of detected cases that varies over
+time and across age groups.

### 2021-07-12 src/paper/model/number_of_contacts.tex
Remove comment.

@@
 
-types of contacts:\comment[id=HM]{I started updating this to understand it better myself.
-In the end, I think it will be useful to group into HH / work / schools and add an
-introductory sentence for each group. E.g., that work is for everybody who is working ---
-except there is a different model for teachers (?), ...}
+types of contacts:
 

### 2021-07-12 src/paper/results/fit.tex
Remove red from some sections.

@@
 reported incidence almost perfectly. When the prevalence of the virus is high
-\textcolor{red}{and during exponential growth phases} the effect of random events on the
+and during exponential growth phases the effect of random events on the
 incidence is large. That is why it's important that we rely on 30 simulation runs with
@@
   our model broadly follows the $R_t$ reported by the RKI. Two differences stand out.
-  Firstly, the RKI's $R_t$ drops faster in November. \textcolor{red}{This is likely due
-  \st{to a change in the testing policy that focused tests on the elderly when the second
-  wave hit Germany and led } to a decline in the estimated overall share of detected
-  cases}. %when the second wave hit Germany.
+  Firstly, the RKI's $R_t$ drops faster in November. This is likely due to a decline in
+  the estimated overall share of detected cases when the second wave hit Germany.
   The second difference is from mid February to mid March where the RKI's reported $R_t$

### 2021-07-12 src/paper/results/fit.tex
Polish goodness of fit section.

--- 2021-07-12 src/paper/results/fit.tex
@@
 effective replication number,  the spread of the B.1.1.7 mutation,
-vaccinations\comment[id=K]{We fit vaccinations in the population by construction. Maybe
-this should go to a different section?} and rapid test demand.
+vaccinations.
 
@@
 Figure~\ref{fig:fit_r_effective}). We also achieve an excellent fit for most age groups
-in Germany. The fit is also good for many German federal states. Despite, the fact that
-the number of performed rapid tests and their distribution in the population result
-endogenously inside our model we fit the share of the population with at least a weekly
-rapid test very well and err on the side of too few individuals that have ever done a
-rapid test.
+in Germany. The fit is also good for many German federal states. Despite the fact that
+the number of performed rapid tests and their distribution in the population are
+determined endogenously in our model, we fit the share of the population with at least a
+weekly rapid test very well. For the share of individuals who have ever done a rapid
+test we err on the side of too few test.
 
@@
 excellent. The incidence in our model matches both the levels and the shape of the
-reported incidence almost perfectly. When the prevalence of the virus is high
-\textcolor{red}{and during exponential growth phases} the effect of random events on the
-incidence is large. That is why it's important that we rely on 30 simulation runs with
-different seed sequences and average over these random differences to have a mean that's
-robust to the randomness in our model.
+reported incidence almost perfectly. When the prevalence of the virus is high and
+especially after explosive growth phases, the effect of random events on the
+incidence is large. Therefore all reported simulations average over at least 30
+simulation runs which is enough to reduce the sampling uncertainty to a negligible
+level.
 
@@
   averaged and show 30 simulation runs. The left figure shows the daily incidence rate
-  per million for the reported simulated infections rates. The mean infection rate is the
+  per million for the simulated reported infection rates. The mean infection rate is the
   thick blue line. Single simulation runs are plotted in lighter and thinner lines. The
@@
 Zooming into the different age groups in Figure~\ref{fig:age_group_fit}, we can see that
-our model is also able to reproduce the infection rates on this level. Differences
-between the age groups allow us to identify the infection probabilities because different
-age groups have contacts of different types (e.g. mostly children go to school, the share
-of workers is highest in the 35 to 59 age bracket etc.). There are very few points where
-our simulated incidences and the empirical incidences do not
-coincide:\comment[id=K]{@Janos: I think you can probably present this better...} Firstly,
-the age group of 80 to 100 year olds due to the absence of nursing homes in our synthetic
-population. Secondly, the increase in adolescents in April. We model the introduction of
-rapid tests for students and do see an increase both in the observed infections and the
-share of detected cases among students (see
-Figure~\ref{fig:share_known_cases_by_age_group}). However, we follow \cite{Betsch2021}
-and assume that 82\% of all rapid tests are followed up with a PCR test---irrespective of
-why they took place. It is likely that that share is higher for rapid tests that were
-done in schools. The last age group where our fit does not fit the empirical curve very
-well is the group of 15 to 34 year olds. This group has been touted to have a very active
-social life and it has often been suspected that this age group also stayed more socially
-active during the pandemic. However, our model does not include either effect because of
-very little data on both and hence it's unsurprising that the actual case numbers in this
-group are consistently higher than our simulated case numbers.\comment[id=K]{Does anyone
-have an idea why we do not match their shape from Oct to Dec?}
+our model is also able to reproduce the infection rates on this level.
+The only major deviation from this pattern is that
+our model predicts too few infections for the 80 to 100 year olds.
+This was to be expected because our synthetic population does not include inhabitants of
+nursing homes. Outbreaks in nursing homes led to a large number of infections among
+the oldest during the second wave of the pandemic in Germany.
+Moreover, the model predicts too few observed infections for the 15 to 34 years old at
+the end of 2020 and the 5 to 14 years old in April and May 2021. The former is likely
+due to the fact that this age group has a very active social life which is not fully
+captured by our contact networks. The latter probably comes from too optimistic
+assumptions on school reopenings.
 
@@
   our model. They accounted for many cases and deaths in the winter of 2020 and many 80
-  to 100 year olds live in these facilities. However, the official data does not contain
-  information on whether cases were nursing home inhabitants or not. We average over 30
+  to 100 year olds live in these facilities. We average over 30
   simulation runs.}
@@
 border to the Czech Republic that had a much higher incidence than
-Germany.\comment[id=K]{The Czech Republic had very high numbers in October (max > 800)
-but they fell during Nov to 220 and then climbed again to 800 by Jan 12th so that might
-be only part of the story why Saxony had such high numbers.}
+Germany.
 
@@
 The fit of the share of vaccinated individuals can be seen in
-Figure~\ref{fig:fit_vaccinations}. In Germany, vaccines were rolled according to four
+Figure~\ref{fig:fit_vaccinations}. In Germany, vaccines were rolled out according to four
 priority groups. The first vaccines were mostly reserved for nursing homes and some
@@
 perfectly. For the share of individuals that have ever done a rapid test our model is
-conservative. There are two reasons for this: Firstly, we only introduce rapid tests in
-2021. This is due to the small number of rapid tests in 2020 as well as a lack of data on
-the number of random tests and who received them. Secondly, there is no randomness in an
-individual's rapid test demand in our model. This reduces the randomness in our model at
-the cost of having less individuals that only occasionally test themselves.
- \comment[id=K]{Write that this makes us underestimate the effect of random tests?}
-However, Section~\ref{subsec:model_validation} shows that our results are very robust to
-changes in the exact shares of individuals demanding rapid tests.
+conservative. There are two reasons for this: Firstly, we do not model people who have
+made rapid tests out of curiosity once they became available. Secondly, in the model,
+the decision to take a rapid test is based on a time invariant individual specific
+compliance factor without any additional random components. While this captures
+important features of rapid test demand it abstracts to people who turn down rapid tests
+most of the time but accept them sometimes. Fortunately,
+Section~\ref{subsec:model_validation} shows
+that our main results are robust to changes in the exact shares of individuals demanding
+rapid tests.
 

### 2021-07-12 src/paper/data/share_known_cases.tex

@@
 detected in the absence of rapid tests ($\psi_t$). For this we rely on the
-\cite[Dunkelzifferradar Project][]{Dunkelzifferradar2020} which uses estimates of the case
-fatality rate to estimate the number of total cases given the number of CoViD-19 deaths
-which are assumed to be perfectly observable. For 2020, we follow the reported share of
-detected cases quite closely \textcolor{red}{only interpolating the phase of November
-2020 where the testing policy was changed to give more priority to vulnerable groups
-\citep{RKI2020a}. This lead to more tests being done among older age groups which are
-also at the highest risk of dying from CoViD-19. We suspect that this led the estimated
-share of detected cases to increase without a corresponding increase in the share of
-detected cases in the overall population as PCR tests, a rare resource at the time, were
-only moved to groups with a high case fatality ratio.}
+\cite[Dunkelzifferradar Project][]{Dunkelzifferradar2020} which uses estimates of the
+case fatality rate to estimate the number of total cases given the number of CoViD-19
+deaths which are assumed to be perfectly observable. For 2020, we follow the reported
+share of detected cases quite closely. One exception is the phase of November 2020 where
+we interpolate to maintain monotonicity during the fall as there was no reason why the
+share of detected cases should have risen in that time\footnote{The testing policy
+changed in November \citep{RKI2020a}. However, this only moved the rare PCR tests more
+towards vulnerable groups.}
 

### 2021-07-12 src/paper/model/number_of_contacts.tex

@@
 Consider a hypothetical population of 1,000 individuals in which 50 were infected with a
-novel infectious disease. From this alone, it is impossible to say whether only
-those 50 people had contact with an infectious person and the disease has an infection
-probability of 1 per contact or whether everyone met an infectious person but the
-disease has an infection probability of only 5 percent per contact. SEIR models do not
-distinguish contact frequency from the infectiousness of each contact and
-combine the two in one parameter that is not invariant to social distancing policies.
+novel infectious disease. From this alone, it is impossible to say whether only those 50
+people had contact with an infectious person and the disease has an infection probability
+of one or whether everyone met one infectious person but the disease has an infection
+probability of only 5 percent per contact. SEIR models do not distinguish between the
+number of contacts and the infectiousness of each contact. Instead, they combine the two
+into one parameter that is not invariant to social distancing policies.
 
 To model social distancing policies, we need to disentangle the effects of the number of
-contacts of each individual and the effect of policy-invariant infection probabilities
-specific to each contact type. Since not all contacts are equally infectious, we
-distinguish different contact types.
+contacts of each individual and the effect of mostly policy-invariant infection
+probabilities specific to each contact type. Since not all contacts are equally
+infectious, we distinguish different contact types.
 

### 2021-07-12 src/paper/results/fit.tex
Add parameter notation, shorten sentences and fix some wordings in the fit section.

@@
 parameters (Figure~\ref{fig:aggregated_fit2}). As a result the effective replication
-number also closely follows that reported by the RKI (see
+number $R_t$ also closely follows that reported by the RKI (see
 Figure~\ref{fig:fit_r_effective}). We also achieve an excellent fit for most age groups
@@
 determined endogenously in our model, we fit the share of the population with at least a
-weekly rapid test very well. For the share of individuals who have ever done a rapid
-test we err on the side of too few test.
+weekly rapid test very well. For the share of individuals who have ever done a rapid test
+we err on the side of too few test.
 
@@
   our model broadly follows the $R_t$ reported by the RKI. Two differences stand out.
-  Firstly, the RKI's $R_t$ drops faster in November. \textcolor{red}{This is likely due
-  \st{to a change in the testing policy that focused tests on the elderly when the second
-  wave hit Germany and led } to a decline in the estimated overall share of detected
-  cases}. %when the second wave hit Germany.
-  The second difference is from mid February to mid March where the RKI's reported $R_t$
-  increased more rapidly than that in our model. Here the opposite effect can be
-  expected. During this time rapid tests increased strongly leading to more cases being
-  detected. In the short term this leads an $R_t$ estimation that is based on detected
-  cases to overestimate the replication number.}
+  Firstly, the RKI's $R_t$ drops faster in November. This is likely due to a decline in
+  the estimated overall share of detected cases ($\psi_t$) when the second wave hit
+  Germany. The second difference is from mid February to mid March where the RKI's
+  reported $R_t$ increased more rapidly than that in our model. Here the opposite effect
+  can be expected. During this time rapid tests increased strongly leading to more cases
+  being detected. In the short term this leads an $R_t$ estimation that is based on
+  detected cases to overestimate the replication number.}
 \end{figure}
@@
 Our fit of the effective replication number $R_t$ closely follows the values reported by
-the RKI (see Figure~\ref{fig:fit_r_effective} even though we calculate $R_t$ on all
+the RKI (see Figure~\ref{fig:fit_r_effective}) even though we calculate $R_t$ on all
 infected individuals not just the detected cases. This explains why the $R_t$ in our
-simulations is higher during phases where the share of detected cases falls such as in
-the fall of 2020 (see Figure~\ref{fig:share_known_cases_data}) where the RKI
-underestimated the effective replication number due to observing a lower share of cases
-and is lower than the $R_t$ reported by the RKI in spring where the share of known cases
-increased due to increased rapid testing.
+simulations is higher during phases where the share of detected cases ($\psi_t$) falls.
+This is the case in the fall of 2020 (see Figure~\ref{fig:share_known_cases_data}) where
+the RKI underestimated the effective replication number due to observing a falling share
+of cases. Analogously, the $R_t$ in our simulations is lower than the $R_t$ reported by
+the RKI in spring where the share of known cases increased due to increased rapid
+testing.
 
@@
 Zooming into the different age groups in Figure~\ref{fig:age_group_fit}, we can see that
-our model is also able to reproduce the infection rates on this level.
-The only major deviation from this pattern is that
-our model predicts too few infections for the 80 to 100 year olds.
-This was to be expected because our synthetic population does not include inhabitants of
-nursing homes. Outbreaks in nursing homes led to a large number of infections among
-the oldest during the second wave of the pandemic in Germany.
-Moreover, the model predicts too few observed infections for the 15 to 34 years old at
-the end of 2020 and the 5 to 14 years old in April and May 2021. The former is likely
-due to the fact that this age group has a very active social life which is not fully
-captured by our contact networks. The latter probably comes from too optimistic
-assumptions on school reopenings.
+our model is also able to reproduce the infection rates on this level. The only major
+deviation from this pattern is that our model predicts too few infections for the 80 to
+100 year olds. This was to be expected because our synthetic population does not include
+inhabitants of nursing homes. Outbreaks in nursing homes led to a large number of
+infections among the oldest during the second wave of the pandemic in Germany. Moreover,
+the model predicts too few observed infections for the 15 to 34 years old at the end of
+2020 and the 5 to 14 years old in April and May 2021. The former is likely due to the
+fact that this age group has a very active social life which is not fully captured by our
+contact networks. The latter probably comes from too optimistic\comment[id=K]{cautious or
+slow? I think optimistic might be ambiguous here.} assumptions on school reopenings.
 
@@
 We fit the proliferation of the B.1.1.7 variant quite exactly despite only introducing a
-few cases in January as can be seen in Figure~\ref{fig:fit_share_b117}. Since we only
-model B.1.1.7 and do not include other variants, B.1.1.7 reaches a share of nearly 100\%
-by May while the true rate plateaued at 90\% as B.1.617.2 started gained traction in
-Germany. Given that B.1.617.2 made up less than 5\% even at the end of our simulation
-period, we did not include it in our model.
+few cases in January ($\omega_{B.1.1.7,t}$) as can be seen in
+Figure~\ref{fig:fit_share_b117}. Since we only model B.1.1.7 and do not include other
+variants, B.1.1.7 reaches a share of nearly 100\% by May while the true rate plateaued at
+90\%. By the end of May B.1.617.2 gained traction in Germany. However, given that
+B.1.617.2 made up less than 5\% even at the end of our simulation period, we did not
+include it in our model.
 
@@
   reported by the RKI and as calculated in our model. We only introduce a few cases over
-  the cause of January. From then B.1.1.7 takes over endogenously through its increased
-  infectiousness. We model no other features of B.1.1.7. We introduce at most 0.986 cases
-  per 100,000 inhabitants. The right figure shows the rate of individuals that are
-  vaccinated in our synthetic population versus in the general German population. Note
-  that we excluded the vaccinations that were given to nursing homes (approx. one percent
-  of the German population) that were vaccinated. Overall, our model covers a time frame
-  that goes from zero vaccinated individuals to a state where over 40\% of the population
-  are vaccinated. Our vaccinations work imperfectly but we do not model different
-  vaccines nor do we distinguish between first and second shot.}
+  the course of January. From then B.1.1.7 takes over endogenously through its increased
+  infectiousness ($\sigma_{B.1.1.7}$). We model no other features of B.1.1.7. We
+  introduce at most 0.986 cases per 100,000 inhabitants. The right figure shows the rate
+  of individuals that are vaccinated in our synthetic population versus in the general
+  German population. Note that we excluded the vaccinations that were given to nursing
+  homes (approx. one percent of the German population). Overall, our model covers a time
+  frame that goes from zero vaccinated individuals to a state where over 40\% of the
+  population are vaccinated. Our vaccinations work imperfectly but we do not model
+  different different vaccines nor do we distinguish between first and second shot.}
 \end{figure}
@@
 reduce physical contact easily such as teachers or medical staff (see
-Section~\ref{subsec:synthetic_population}. How the share of vaccinated individuals
-develops by age group can be seen in Figure~\ref{fig:vaccinations_by_age_group}.
+Section~\ref{subsec:synthetic_population} and Figure~\ref{fig:vaccinations_by_age_group}
+for the vaccination rates in our model by age group).
 
@@
 % rapid test demand
-The most difficult moment to match in our model is the rapid test demand. Since we have
-five different channels through which individuals demand rapid tests and many of the
-demand curves are at least partially calibrated through survey data. It is therefore very
-reassuring that we fit the share of individuals that do weekly rapid tests almost
+The most difficult moment to match in our model is the rapid test demand. This is because
+we have five different channels through which individuals demand rapid tests and many of
+the demand curves are at least partially calibrated through survey data. It is therefore
+very reassuring that we fit the share of individuals that do weekly rapid tests almost
 perfectly. For the share of individuals that have ever done a rapid test our model is
 conservative. There are two reasons for this: Firstly, we do not model people who have
-made rapid tests out of curiosity once they became available. Secondly, in the model,
-the decision to take a rapid test is based on a time invariant individual specific
-compliance factor without any additional random components. While this captures
-important features of rapid test demand it abstracts to people who turn down rapid tests
-most of the time but accept them sometimes. Fortunately,
-Section~\ref{subsec:model_validation} shows
-that our main results are robust to changes in the exact shares of individuals demanding
-rapid tests.
+made rapid tests out of curiosity once they became available. Secondly, in the model, the
+decision to take a rapid test is based on a time invariant individual specific compliance
+factor without any additional random components. While this captures important features
+of rapid test demand it abstracts from people who turn down rapid tests most of the time
+but accept them sometimes. Fortunately, Section~\ref{subsec:model_validation} shows that
+our main results are robust to changes in the exact shares of individuals demanding rapid
+tests.
 

### 2021-07-12 src/paper/results/rapid_tests_statistics.tex
Add parameter notation where possible and fix typos.

@@
 number of performed rapid tests as a result of time invariant willingness to do rapid
-tests and time varying supply side factors and events that trigger rapid tests. Thus,
-the parameters described in Section~\ref{subsec:rapid_test_demand} are only indirectly
-related to the number of rapid tests that are actually performed in the model. When it
-comes to positive and negative rapid tests, there is even an additional layer because
-rapid tests imperfectly sensitive and specific.
+tests and time varying supply side factors and events that trigger rapid tests. Thus, the
+$\pi$ parameters governing when individuals do rapid tests described in
+Section~\ref{subsec:rapid_test_demand} are only indirectly related to the number of rapid
+tests that are actually performed in the model. When it comes to positive and negative
+rapid tests, there is even an additional layer because rapid tests are imperfectly
+sensitive and specific.
 
@@
 even at the time where discovery of cases via rapid tests is at its peak, almost
-30\% of of people who are infected and make a test are not discovered by the test.
+30\% of people who are infected and make a test are not discovered by the test.
 

### 2021-07-12 src/paper/results/share_known_cases.tex
Add parameter notation where possible and fix typos.

@@
 tested. Starting in 2021 the share of detected cases not only depends on the overall
-share of detected cases we feed into our model but also on the number of performed rapid
-tests. As more and more (often asymptomatic and presymptomatic) individuals receive
-positive rapid tests and receive confirmation, the share of detected cases starts to
-increase, especially after March when rapid tests became much more widely available. This
-effect is very visible for school pupils, (5-14, green line). As twice weekly rapid tests
-become mandatory in schools and students start going to school after the Easter break,
-the share of detected cases for that group increases drastically from 25\% to over 60\%
-between April and June.
+share of detected cases ($\psi_t$) we feed into our model but also on the number of
+performed rapid tests (governed by the $\pi$ parameters). As more and more (often
+asymptomatic and presymptomatic) individuals receive positive rapid tests and seek and
+get confirming PCR tests, the share of detected cases starts to increase, especially
+after March when rapid tests became much more widely available. This effect is very
+visible for school pupils (5-14, green line). As twice weekly rapid tests become
+mandatory in schools and students start going to school after the Easter break, the share
+of detected cases for that group increases drastically from 25\% to over 60\% between
+April and June.
 
@@
   reported as an official case via PCR confirmation for each age group in our population.
-  We use the overall share of detected cases that was estimated through the case fatality
-  ratio by \citet{Dunkelzifferradar2020} for all of 2020 and then assume it to be
-  constant as vaccinations of the elderly strongly affect the case fatality rate which
-  the project does not account for. To get from an overall share of detected cases to the
-  share of cases that is detected in each age group we use that asymptomatic cases are
-  much less likely to be detected. As our model covers age specific asymptomatic rates
-  this endogenously leads to group specific share of detected cases that suggest that
-  infections in younger age groups are under-detected. Starting in 2021 in addition to
-  the overall numbers of detected cases through symptoms and the share of detected cases,
-  cases are also detected through confirmation of positive rapid tests. This leads to an
-  increase in the share of detected cases for all age groups but in particular for the
-  younger age groups that are covered extensively with rapid tests through the rapid test
-  requirement for participating in school. Age groups that have a high share of workers
-  also disproportionately increase their detection rate.}
+  We use the overall share of detected cases ($\psi_t$) that was estimated through the
+  case fatality ratio by \citet{Dunkelzifferradar2020} for all of 2020 and then assume it
+  to be constant as vaccinations of the elderly strongly affect the case fatality rate
+  which the project does not account for. To get from an overall share of detected cases
+  to the share of cases that is detected in each age group we use that asymptomatic cases
+  are much less likely to be detected. As our model covers age specific asymptomatic
+  rates this endogenously leads to group specific share of detected cases that suggest
+  that infections in younger age groups are under-detected. Starting in 2021 in addition
+  to the overall numbers of detected cases through $\psi_t$, cases are also detected
+  through confirmation of positive rapid tests. This leads to an increase in the share of
+  detected cases for all age groups but in particular for the younger age groups that are
+  covered extensively with rapid tests through the rapid test requirement for
+  participating in school. Age groups that have high shares of workers also have
+  disproportionately increases in their detection rates.}
 \end{figure}

### 2021-07-12 src/paper/results/validation.tex
Fix figure references.
Rewrite section on rapid test statistics.
Add parameter notation where possible and fix typos.

@@
 
-Achieving a good in-sample fit does not necessarily guarantee that our model will also
-be able to make out of sample predictions. For example, it could be that the results
-are very sensitive to the exact number of vaccinations, the work mobility multiplier
-or the number of rapid tests that are performed -- all of which are things that cannot
-be exactly known ex-ante.
+Achieving a good in-sample fit does not necessarily guarantee that our model will also be
+able to make out of sample predictions. For example, it could be that the results are
+very sensitive to the exact number of vaccinations, the work mobility multiplier
+($\rho_{w,\:attend,\:t}$) or the number of rapid tests (governed by the $\pi$ parameters)
+that are performed -- all of which are things that cannot be exactly known ex-ante.
 
@@
 
-For the out of sample predictions we predict the number of vaccinations between March
-and June with a simple linear regression model that was fitted on vaccine data from
-February. This prediction model is pessimistic compared to the actual number of
-vaccinations. The work mobility multiplier is predicted to be constant at a value of
-0.75, which is an approximate average of the second half of February. This turned out
+For the out of sample predictions we predict the number of vaccinations between March and
+June with a simple linear regression model that was fitted on vaccine data from February.
+This prediction model is pessimistic compared to the actual number of vaccinations. The
+work mobility multiplier ($\rho_{w,\:attend,\:t}$) is predicted to be constant at a value
+of 0.75, which is an approximate average of the second half of February. This turned out
 to be optimistic.
 
-The area that is fraught with the most uncertainty is the introduction of rapid test,
-because it comprises both supply and demand factors. Moreover, accurately predicting
-the number of rapid tests is expected to be important because rapid tests play a large
-role for the transmission dynamic.
+The area that is fraught with the most uncertainty is the introduction of rapid tests,
+because it comprises both supply and demand factors. Moreover, accurately predicting the
+number of rapid tests is expected to be important because rapid tests play a large role
+for the transmission dynamic.
 
-We therefore make a scenario analysis with different assumptions on the availability
-of rapid tests. The number of rapid tests performed in each scenario can be seen in
-Figure~\ref{fig:robustness_check_rapid_test_params}. All scenarios are the same
-unti March 1 and have the same level of rapid tests when all supply constraints are
-resolved. They differ in the date at which the full number of tests is reached. For
-students and teachers the full number of rapid tests is reached after the easter
-holidays. For rapid tests in the workplace and private rapid tests it is reached between
+We therefore make a scenario analysis with different assumptions on the availability of
+rapid tests. The number of rapid tests performed in each scenario can be seen in
+Figure~\ref{fig:robustness_check_rapid_test_params}. All scenarios are the same until
+March 1 and have the same level of rapid tests when all supply constraints are resolved.
+They differ in the date at which the full number of tests is reached. For students
+($\pi_{students,\:t}$) and teachers ($\pi_{teacher,\:t}$) the full number of rapid tests
+is reached after the Easter holidays in all scenarios. For rapid tests in the workplace
+($\pi_{w,\:s,\:t}$) and private rapid tests ($\pi_{private,\:t}$) it is reached between
 May 1 and June 10, depending on the scenario.
@@
   \floatfoot{\noindent \textit{Note:} Number of rapid tests performed in the different
-  prediction scenarios. All scenarios are the same
-  unti March 1 and have the same level of rapid tests when all supply constraints are
-  resolved. They differ in the date at which the full number of tests is reached. For
-  students and teachers the full number of rapid tests is reached after the easter
-  holidays. For rapid tests in the workplace and private rapid tests it is reached
-  between May 1 and June 10, depending on the scenario.
-  }
+  prediction scenarios. All scenarios are the same until March 1 and have the same level
+  of rapid tests when all supply constraints are resolved. They differ in the date at
+  which the full number of tests is reached. For students ($\pi_{students,\:t}$) and
+  teachers ($\pi_{teacher,\:t}$) the full number of rapid tests is reached after the
+  Easter holidays in all scenarios. For rapid tests in the workplace ($\pi_{w,\:s,\:t}$)
+  and private rapid tests ($\pi_{private,\:t}$) it is reached between May 1 and June 10,
+  depending on the scenario. }
 \end{figure}
 
-Moreover, the out of sample predictions assume that the share of detected cases that
-would have obtained without rapid tests is not affected by the easter holidays because
-the extent to which this was the case was estimated from case numbers in April.
+Moreover, the out of sample predictions assume that the share of detected cases
+($\psi_t$) that would have been obtained without rapid tests is not affected by the
+Easter holidays because the extent to which this was the case was estimated from case
+numbers in April.
 
 The results of the out of sample prediction are displayed in
-Figure~\ref{robustness_check_detailed}. While all scenarios considerably deviate from
-the ex-post scenario, they all reproduce the steep increase of cases until the end
-of April, followed by a decline until June. We can therefore conclude that our main
-results are not sensitive to measurement errors in the number of rapid tests,
-vaccinations or mobility data.
+Figure~\ref{fig:robustness_check_detailed}. While all scenarios considerably deviate from
+the ex-post scenario, they all reproduce the steep increase of cases until the end of
+April, followed by a decline until June. We can therefore conclude that our main results
+are not sensitive to measurement errors in the number of rapid tests, vaccinations or
+mobility data.
 
@@
   with data from February. Moreover, they do not model a lower number of detected cases
-  over the easter holidays. The different scenarios make different assumptions on the
+  over the Easter holidays. The different scenarios make different assumptions on the
   date at which full availability of rapid tests is reached. While the out of sample
@@
 \end{figure}
+
+\FloatBarrier

### 2021-07-13 src/paper.tex
Add science style.
Went through paper once.

@@
 \input{templates/Preamble_Sansserif_math}
-\input{templates/Preamble_BibLaTeX_AER}
-%\input{templates/Preamble_BibLaTeX_APA}
+% \input{templates/Preamble_BibLaTeX_AER}
+% \input{templates/Preamble_BibLaTeX_APA}
+\input{templates/Preamble_BibLaTeX_Science}
 \input{templates/Preamble_Commenting_and_highlighting}
@@
 
+\tableofcontents
+
 \clearpage
@@
     \begin{refsection}
-
-    This paper aims to be completely reproducible. The code for the general model can be
-    found at
-    \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
-    and its documentation at
-    \href{https://sid-dev.readthedocs.io/en/latest/}{https://sid-dev.readthedocs.io/en/latest/}.
-    The code for the application to Germany can be found at
-    \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}
-    with a very rudimentary documentation at
-    \href{https://sid-germany.readthedocs.io/en/latest/}{https://sid-germany.readthedocs.io/en/latest/}.
 

### 2021-07-13 src/paper/data/course_of_disease.tex
Went through paper once.

--- 2021-07-12 src/paper/data/medical_parameters.tex
@@
-\subsection{Medical Parameters}
-\label{sec:medical_params}
+\subsection{Course of Disease}
+\label{sub:data_course_of_disease}
 
-This section discusses the medical parameters used in the model, their sources and how
-we arrived at the distributions used in the model.\footnote{Additional information can
-be found in the
+This section discusses the parameters governing the course of disease, their sources and
+how we arrived at the distributions used in the paper.\footnote{Additional information
+can be found in the
 \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
@@
 % Latency Period
-The first medical parameter we need is the distribution of the length of the period
-between infection and the start of infectiousness, the so called latent period,
-$\gamma_{infectious}$. \cite{Zhao2021} estimate the latent period to last 3.3 days (95\%
-CI: 0.2, 7.9). In line with this estimate our latent period lasts one to five days with a
-slightly lower average of two days.
+We denote the latent period---i.e., the time span between infection and the start of
+infectiousness---by $\gamma_{infectious}$. \cite{Zhao2021} estimate the latent period to
+last 3.3 days (95\% CI: 0.2, 7.9) on average. In line with this estimate our latent
+period lasts one to five days.
 
 % Asymptomatic Share
-Once individuals become infectious a share of them goes on to develop symptoms while
+Once individuals become infectious, a share of them goes on to develop symptoms while
 others remain asymptomatic. We rely on data by \cite{Davies2020} for the age-dependent

### 2021-07-13 src/paper/data/synthetic_population.tex
Improve wording of number of contacts in the model section.
Add parameter notation, shorten sentences and fix some wordings in the fit section.
Went through paper once.

@@
 practice regularly together. Both daily and weekly groups are created analogously to the
-work groups but matching the numbers in Figure~\ref{fig:n_contacts_other_daily_recurrent} and
-Figure~\ref{fig:n_contacts_other_weekly_recurrent}. In addition, since leisure contacts are
-highly assortative by age all individuals that have a daily leisure contact are matched
-with a person not only from the same county but also from the same age group.
+work groups but matching the numbers in Figure~\ref{fig:n_contacts_other_daily_recurrent}
+and Figure~\ref{fig:n_contacts_other_weekly_recurrent}. In addition, since leisure
+contacts are highly assortative by age all individuals that have a daily leisure contact
+are matched with a person not only from the same county but also from the same age group.
 
@@
 
+% vaccination rank
 Lastly, for the distribution of vaccinations every individual is assigned a vaccination
@@
   \floatfoot{\noindent \textit{Note:} An individual's vaccination priority depends on her
-  work contact priority, her age group and a random component to capture medical
+  work contact priority, her age group and a random component to capture
   preconditions like diabetes. 15\% of the population refuse to be vaccinated ($\xi$).

### 2021-07-13 src/paper/literature_review.tex
Went through paper once.

@@
 
-The workhorse model in epidemiology is the SEIR model. However,
-The traditional SEIR model is not fine-grained enough to model nuanced policies. This
-has motivated a large number of researchers to extend the standard model to allow for
-more heterogeneity and flexibility. Examples are \citet{Grimm2020},
-\citet{Donsimoni2020} and \citet{Acemoglu2020} who develop multi group SEIR models to
-analyze the effects of targeted lockdowns and \citet{Berger2020} who extend the SEIR
-model to analyze testing and conditional quarantines. For a more comprehensive review
-see \citet{Avery2020}. Others have used the results of a standard SEIR model as input
-for economic models that estimate the cost of policies (e.g. \citet{Dorn2020}).
+The workhorse model in epidemiology is the SEIR model. However, The traditional SEIR
+model is not fine-grained enough to model nuanced policies.\comment[id=HM]{Still old
+motivation. Hard to get the timing of tests and results right?} This has motivated a
+large number of researchers to extend the standard model to allow for more heterogeneity
+and flexibility. Examples are \citet{Grimm2020}, \citet{Donsimoni2020} and
+\citet{Acemoglu2020}\comment[id=HM]{We need to cite more strategically. Cut Grimm,
+Donsinomi. More state of the art SEIR in other fields. Or drop SEIR altogether? If I
+read that as a Michael Meyer-Hermann type, I would think, ``ah, these economists
+again...''} who develop multi group SEIR models to analyze the effects of targeted
+lockdowns and \citet{Berger2020} who extend the SEIR model to analyze testing and
+conditional quarantines. For a more comprehensive review see \citet{Avery2020}. Others
+have used the results of a standard SEIR model as input for economic models that
+estimate the cost of policies (e.g. \citet{Dorn2020}).
 
@@
 
-The OpenABM-Covid-19 model by \citet{Hinch2020} is the closest in spirit to ours. They
-model detailed contact networks for workplaces, schools and households and can evaluate
-the effect of several NPIs. The main focus of their application are contact tracing
-policies.
+The OpenABM-Covid-19 model by \citet{Hinch2020} is the closest in spirit to
+ours\comment[id=HM]{I'd rather be close to Aleta et al. ;-)}. They model detailed
+contact networks for workplaces, schools and households and can evaluate the effect of
+several NPIs. The main focus of their application are contact tracing policies.
 
@@
 
-\citet{Bicher2021} implement a model that simulates the complete Austrian population
-with almost nine million people. It uses the same data provided by \citet{Mossong2008}
-as we use to calibrate contact networks for households, workplaces and schools. The
-model focuses and analyzing the effect of different contact tracing strategies. It is
-calibrated with data from the first wave (February 21 to April 9).
+\citet{Bicher2021} simulate the Austrian population during the first wave (February 21
+to April 9, 2020). It uses the same data provided by \citet{Mossong2008} as we use to
+calibrate contact networks for households, workplaces and schools. The model focuses on
+analyzing the effect of different contact tracing strategies.
 
-Moreover, there are several forthcoming applications of agent-based simulation models
-with contact networks. Examples are \citet{Basurto2020}, \citet{DelliGatti2020} and
-\citet{Mellacher2020}.
+Moreover, there are several forthcoming applications\comment[id=HM]{What is that
+supposed to mean?} of agent-based simulation models with contact networks. Examples are
+\citet{Basurto2020}, \citet{DelliGatti2020} and \citet{Mellacher2020}.
 

### 2021-07-13 src/paper/model/code.tex
Went through paper once.

The code used for this paper is open source and available under the MIT License. It is
split into two parts
\begin{itemize}
    \item The code for the model can be found at
   \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
   and its documentation at
   \href{https://sid-dev.readthedocs.io}{https://sid-dev.readthedocs.io}.
   \item The code for the application to Germany can be found at
   \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}
   with a shorter documentation at
   \href{https://sid-germany.readthedocs.io}{https://sid-germany.readthedocs.io}.
\end{itemize}

### 2021-07-13 src/paper/model/course_of_disease.tex
Went through paper once.

--- 2021-06-21 src/paper/model/course_of_disease.tex
@@
-\subsection{Course of the Disease}
-\label{sub:disease_progression}
+\subsection{Course of Disease}
+\label{sub:model_course_of_disease}
 
@@
 Figure~\ref{fig:model_disease_progression} and the values and source of the relevant
-parameters are describes in Section~\ref{sec:medical_params}.
+parameters are describes in Section~\ref{sub:data_course_of_disease}.
 
@@
 
-It would be easy to make the course of disease even more fine-grained. For example, we
-could model people who require hospitalization but not intensive care. So far we opted
-against that because only the intensive care capacities are feared to become a bottleneck
-in Germany.
+% It would be easy to make the course of disease even more fine-grained. For example, we
+% could model people who require hospitalization but not intensive care. So far we opted
+% against that because only the intensive care capacities are feared to become a bottleneck
+% in Germany.
 

### 2021-07-13 src/paper/report.tex
Use parameter notation everywhere in the data section.
Went through paper once.

@@
 
-The four contact networks are chosen so that the most common NPIs can be modeled in great
-detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
-having physical contact. The effect of different NPIs will generally vary across contact
-types. For example, a mandate to work from\comment[id=K]{$\rho_{attend,\,t}$} home will
-reduce the number of work contacts to zero for a fraction of the working population.
-Schools and daycare can be closed entirely, operate at reduced capacity---including an
-alternating schedule---, or implement mitigation measures like masking requirements or
-air filters \citep{Lessler2021}. Curfews\comment[id=K]{would enter into $\rho_{other\:,t}$} may reduce the number of contacts in
-non-work/non-school/non-work settings. In any setting, measures like masking requirements
-would reduce the probability of infection associated with a contact \citep{Cheng2021}.\comment[id=K]{$\rho_{hygiene}$}
+The four contact networks are chosen so that the most common NPIs can be modeled in
+great detail. NPIs affect the number of contacts or the risk of transmitting the disease
+upon having physical contact. The effect of different NPIs will generally vary across
+contact types. For example, a mandate to work from\comment[id=K]{$\rho_{attend,\,t}$}
+home will reduce the number of work contacts to zero for a fraction of the working
+population. Schools and daycare can be closed entirely, operate at reduced
+capacity---including an alternating schedule---, or implement mitigation measures like
+masking requirements or air filters \citep{Lessler2021}. Curfews\comment[id=K]{would
+enter into $\rho_{other\:,t}$} may reduce the number of contacts in settings outside of
+work and school. In any setting, measures like masking requirements would reduce the
+probability of infection associated with a contact
+\citep{Cheng2021}.\comment[id=K]{$\rho_{hygiene}$}
 
@@
 age\comment[id=K]{$\zeta_{a}$}. A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period of
-infection without being infectious or showing symptoms, being infectious (presymptomatic
-or asymptomatic), showing symptoms, requiring intensive care, and recovery or death
-\citep[similar to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The
-probabilities of transitioning between these states depend on age;\comment[id=K]{$\gamma$
-for all transition probabilities} their duration is random within intervals calibrated to
-the medical literature (for a detailed description see Section~\ref{sec:medical_params}).
+Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has
+lost the timing. Need to explain d-subscript, maybe even better to add stuff like 1-3
+days back in? Needs to be comprehensible without recurring to appendix} We differentiate
+between an initial period of infection without being infectious or showing symptoms,
+being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
+care, and recovery or death \citep[similar to][]{Grimm2021}.\comment[id=HM]{Need to look
+for non-economists -- this must be fairly standard?} The probabilities of transitioning
+between these states depend on age; their duration is random and calibrated to medical
+literature (for a detailed description see Section~\ref{sub:data_course_of_disease}).
 Conditional on the type of contact, infectiousness is independent of age
@@
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different profiles regarding
+The model includes several other features, which are crucial to describe the evolution
+of the pandemic in 2020-2021. New virus strains with different profiles regarding
 infectiousness\comment[id=K]{$\sigma_v$} can be introduced. Agents may receive a
@@
 immune and they do not transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
-Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the second
-dose of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We choose it
-because our model neither includes booster shots, nor does it allow vaccinated
+Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the
+second dose of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We
+choose it because our model neither includes booster shots, nor does it allow vaccinated
 individuals who became immune to transmit the disease\citep{Petter2021,
@@
 vaccinated individuals had received a second dose already.} During the vaccine roll-out,
-priority may depend on age and occupation.\comment[id=K]{Maybe mention that we have
-refusers?}
+priority may depend on age and occupation; some individuals will refuse vaccination
+offers.
 
@@
 individual is infected or not; there is no uncertainty to the result. PCR tests require
-some days to be processed\comment[id=K]{$\gamma_{PCR,\:d}$ where $d=$ number of days} and
-there are aggregate capacity constraints throughout. In contrast, rapid antigen tests
+at least one day to be processed\comment[id=K]{$\gamma_{PCR,\:d}$ where $d=$ number of
+days} and there are aggregate capacity constraints. In contrast, rapid antigen tests
 yield immediate results; after a phase-in period, all tests that are demanded will be
@@
 require staff and students to be tested regularly.\comment[id=K]{"$\pi_{teacher,\:t}$"
-etc. } Rapid tests may be offered by employers to on-site workers. Individuals may demand
-tests for private reasons, which include having plans to meet other people, showing
-symptoms of CoViD-19, and because a household member tested positively for the virus. We
-endow each agent with an individual compliance parameter. This parameter determines
-whether she takes up rapid tests.\footnote{A positive test result (of either kind), as
-well as symptoms leads most individuals to reduce their contacts; this is why tests
-impact the actual contacts in Figure~\ref{fig:model_description}.}
+etc. } Rapid tests may be offered by employers to on-site workers. Individuals may
+demand tests for private reasons, which include having plans to meet other people,
+showing symptoms of CoViD-19, and because a household member tested positively for the
+virus. We endow each agent with an individual compliance parameter. This parameter
+determines whether she takes up rapid tests.\footnote{A positive test result (of either
+kind), as well as symptoms leads most individuals to reduce their contacts; this is why
+tests impact the actual contacts in Figure~\ref{fig:model_description}.}
 
@@
 parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
-materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
-for different contact types and their assortativity by age group\comment[id=K]{Should we
-add all our sources here?}. Mobility data is used to model the evolution of work
-contacts. School and daycare policies can be incorporated directly from official
-directives. Administrative records on the number of tests, vaccinations by age and
-region, and the prevalence of virus strains are generally available. Surveys may ask
-about test offers, propensities to take them up, and past tests. Other studies' estimates
-of the seasonality of infections can be incorporated directly. The remaining
-parameters---most notably, these include infection probabilities by contact network and
-the effects of some NPIs, see Supplementary Material~\ref{subsec:estimated_params}---will
-be chosen numerically so that the model matches features of the data \citep[see][for the
-general method, also described in Supplementary Material
-\ref{subsec:estimated_param}]{McFadden1989}. In our application, we keep the number of
-free parameters low in order to avoid overfitting. The data features to be matched
-include official case numbers for each age group and region, deaths, and the share of the
-B.1.1.7 strain.
+materials for the precise sources.} Contact diaries yield pre-pandemic distributions of
+contacts for different contact types and their assortativity by age group. Mobility data
+is used to model the evolution of work contacts. School and daycare policies can be
+incorporated directly from official directives. Administrative records on the number of
+tests, vaccinations by age and region, and the prevalence of virus strains are generally
+available. Surveys may ask about test offers, propensities to take them up, and past
+tests. Other studies' estimates of the seasonality of infections can be incorporated
+directly. The remaining parameters---most notably, these include infection probabilities
+by contact network and the effects of some NPIs, see Supplementary
+Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
+matches features of the data \citep[see][for the general method, also described in
+Supplementary Material \ref{subsec:estimated_params}]{McFadden1989}. In our application,
+we keep the number of free parameters low in order to avoid overfitting. The data
+features to be matched include official case numbers for each age group and region,
+deaths, and the share of the B.1.1.7 strain.
 
@@
 all infections. We thus model official cases as depicted in
-Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
-detected cases\comment[id=K]{$\psi$} and use data on the share of PCR tests administered
-because of CoViD-19 symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of
-asymptomatic individuals varies by age group, this gives us age-specific shares (see
+Figure~\ref{fig:model_official_cases}\comment[id=HM]{Replace P2, P(compliance) by
+symbols in figure}. We take aggregate estimates of the share of detected
+cases\comment[id=K]{$\psi$} and use data on the share of PCR tests administered because
+of CoViD-19 symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of asymptomatic
+individuals varies by age group, this gives us age-specific shares (see
 Figure~\ref{fig:share_known_cases_by_age_group} for the share of detected cases by age
@@
 compared to school age children. Once rapid test become available, confirmation of a
-positive result\comment[id=K]{prob of demanding confirmation is $\chi_{confirmation}$} is
-another reason leading to positive PCR tests.
+positive result\comment[id=K]{prob of demanding confirmation is $\chi_{confirmation}$}
+is another reason leading to positive PCR tests.
 
@@
 turn of the year, the first people were vaccinated with a focus on older age groups and
-medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May,
-43\% had received at least one dose of a vaccine. In late 2020, rapid tests started to
+medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May, 43\%
+had received at least one dose of a vaccine. In late 2020, rapid tests started to
 replace regular PCR tests for staff in many medical and nursing facilities. These had to
@@
 Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections with only one channel active, the
-final level is always the highest in case of the vaccination campaign (orange lines). The
-Shapley value decomposition shows that vaccinations contribute 16\% to the cumulative
-difference between scenarios. Reasons for this are the slow start---it took until
-March~24th until 10\% of the population had received their first vaccination, the 20\%
-mark was reached on April 19th---and the focus on older individuals. These groups
+sight. Whether considering recorded or total infections with only one channel active,
+the final level is always the highest in case of the vaccination campaign (orange
+lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
+cumulative difference between scenarios. Reasons for this are the slow start---it took
+until March~24th until 10\% of the population had received their first vaccination, the
+20\% mark was reached on April 19th---and the focus on older individuals. These groups
 contribute less to the spread of the disease than others due to a lower number of
-contacts (see Supplementary
-Material~\ref{subsec:assortativity}).\comment[id=K]{Assortativity is the wrong section
-here. They have few contacts in our model because they tend to live in small households
-and don't have work or educ contacts. They are also more likely to meet other old people
-but that's a different channel from what is written in the text.} By the end of our study
-period, when first-dose vaccination rates reached around 40\% of the population, the
-numbers of new cases would have started to decline. It is important to note that the
-initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
-fatality was rate considerably lower during the third wave when compared to the second
-(4.4\% between October and February and 1.4\% between March and the end of
-May).\comment[id=K]{This needs a source.}
+contacts. By the end of our study period, when first-dose vaccination rates reached
+around 40\% of the population, the numbers of new cases would have started to decline.
+It is important to note that the initial focus of the campaign was to prevent deaths and
+severe disease. Indeed, the case fatality was rate considerably lower during the third
+wave when compared to the second (4.4\% between October and February and 1.4\% between
+March and the end of May).\comment[id=K]{This needs a source.}\comment[id=HM]{Didn't we
+calculate that ourselves at some point?}
 

### 2021-07-13 src/paper/results/rapid_tests_statistics.tex
Went through paper once.

@@
         \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
-      \caption{Share of the Population Doing a Rapid Test Because of Different Channels
-      on a Given Day}
+        \caption{Share of the Population Doing a Rapid Test on a Given Day, by Channels}
         \label{fig:rapid_test_demand_by_channel}
@@
         \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_positive}
-      \caption{Share of the Population Testing Positive Because of Different Channels
-      on a Given Day}
+        \caption{Share of the Population Testing Positive on a Given Day, by Channels}
         \label{fig:pos_rapid_tests_by_channel}
@@
         \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_negative}
-      \caption{Share of the Population Testing Negative Because of Different Channels
-      on a Given Day}
+        \caption{Share of the Population Testing Negative on a Given Day, by Channels}
         \label{fig:neg_rapid_tests_by_channel}
@@
     \floatfoot{\noindent \textit{Note:}
-    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
-    school) as well as pupils. After Easter the required frequency of tests is
-    increased from once per week to twice per week. Work rapid tests are demanded by
-    individuals that still have work contacts, i.e. do not work from home. The share of
-    employers offering rapid tests increases over the time frame and the frequency of
-    testing is also increased. Private tests are demanded by individuals for one of three
-    reasons: having developed symptoms without access to a PCR test, having a household
-    member that has tested positive or developed symptoms or having planned a weekly
-    meeting with friends. The left figure shows the share of the population doing a
-    rapid test on a given day. The middle figure shows the share of the population
-    testing positive on a given day (true and false positives). The right figure shows
-    the share of the population testing negative on a given day (true and false
-    negatives).}
+        Rapid tests in the education setting are demanded by teachers (nursery,
+        preschool and school) as well as pupils. After Easter the required frequency of
+        tests is increased from once per week to twice per week. Work rapid tests are
+        demanded by individuals that still have work contacts, i.e. do not work from
+        home. The share of employers offering rapid tests increases over the time frame
+        and the frequency of testing is also increased. Private tests are demanded by
+        individuals for one of three reasons: having developed symptoms without access
+        to a PCR test, having a household member that has tested positive or developed
+        symptoms or having planned a weekly meeting with friends. The left figure shows
+        the share of the population doing a rapid test on a given day. The middle figure
+        shows the share of the population testing positive on a given day (true and
+        false positives). The right figure shows the share of the population testing
+        negative on a given day (true and false negatives).}
 \end{figure}

### 2021-07-13 src/paper/model/matching.tex
Add formal description of matching.

--- 2021-06-18 paper/model/matching.tex
@@
 
-Recurrent contacts are described by two components: 1) A set of time invariant groups, such as school classes. The groups can be sampled from empirical data or created by
-randomly matching simulated individuals into groups. 2) A deterministic or random function that takes the value 0 (non-participating) and 1
-(participating) and can depend on the weekday, date and health state. This can be used
-to model vacations, weekends or symptomatic people who stay home (see section
-\ref{sec:endogenous_contact_reductions} for details).
+%% Recurrent contacts ============================================================
+Recurrent contacts are described by two components: 1) A set of time invariant groups,
+such as school classes or groups of co-workers. Those groups are generated once at the
+beginning of the simulation. The groups can be sampled from empirical data or created by
+randomly matching simulated individuals into groups. 2) A deterministic or random
+function that takes the value 0 (non-participating) and 1 (participating) and can depend
+on the weekday, date and health state. This can be used to model vacations, weekends or
+symptomatic people who stay home (see section \ref{sec:policies} for details).
 
-The matching process for recurrent contacts is then extremely simple: On each simulated
-day, every person who does not stay home meets all other group members who do not stay
-home. The assumption that all group members have contacts with all other group members
+Given those two components, the disease transmission for recurrent contacts is extremely
+simple: On each simulated day, every person who does not stay home meets all other group
+members who do not stay home. If there is a contact between individual $i$ who is
+infected with virus variant $v$ and infectious and individual $j$ who is in age group
+$a$ and susceptible, then $j$ becomes infected with the following probability
+
+\begin{equation}
+    \label{eq:infection_probability_formula}
+    P(infection) = \beta_c \cdot s_{c, t} \cdot \sigma_v \cdot \zeta_a
+\end{equation}
+
+where $beta_c$ denotes the base infection probability of contact type $c$,
+$s_{c, t}$ is a seasonality factor between zero and one that depends on the contact
+type $c$ and time $t$ (see Equation~\ref{eq:seasonality}), $\sigma_v$ is the
+infectiousness factor of virus variant $v$ and $\zeta_a$ is an age dependent
+susceptibility factor.
+
+The assumption that all group members have contacts with all other group members
 is not fully realistic, but a good approximation to reality, especially in
@@
 
-The matching in non-recurrent contact models is more difficult and implemented in a two
-stage sampling procedure to allow for assortative matching. Currently most contact
-models are assortative with respect to age (it is more likely to meet people from the
-same age group) and county (it is more likely to meet people from the same county) but
-in principle any set of discrete variables can be used. This set of variables that
-influence matching probabilities introduce a discrete partition of the population into
-groups. The first stage of the two stage sampling process samples on the group level.
-The second stage on the individual level.
+%% Non recurrent contacts
 
-The algorithm works as follows: First we randomly draw a contact type and individual.
-For each contact of the drawn contact type that person has, we first draw the group of
-the other person (first stage).
-Next, we calculate the probability to be drawn for each member of the group, based on
-the number of remaining contacts, i.e. people who have more remaining contacts are drawn
-with a higher probability. The probabilities have to be re-calculated each time because
-with each
-matched contact, the number of remaining contacts changes. We then draw the other
-individual, determine whether an infection takes place and if so update the health state
-of the newly infected person. Finally, we reduce the number of remaining contacts of the
-two matched individuals by one.\comment[id=J]{Add formulae here.}
+The matching of non-recurrent contact types is more difficult because the contact
+network is resampled randomly every day. Moreover, it needs to allow for
+assortative matching. In our application, all
+random contacts are assortative with respect to age group $a$ (it is usually more likely
+to meet
+people from the same age group) and county (it is more likely to meet people from the
+same county) but in principle any set of discrete variables can be used. This set of
+variables that influence matching probabilities introduce a discrete partition of the
+population into groups.
 
-The recalculation of matching probabilities in the second stage is computationally
-intensive because it requires summing up all remaining contacts in that group. Using a
-two stage sampling process where the first stage probabilities remain constant over time
-makes the matching computationally much more tractable because the number of
-computations only increases quadratically in the size of the second group and not
-quadratically in the size of the entire simulated population.
+Below we first describe one iteration of a simplified matching algorithm that
+illustrates what we want to achieve. In practice, we approximate the result of this
+matching algorithm by a two stage sampling procedure that is computationally more
+efficient. The matching is done for each non-recurrent contact type $c$. The following
+step is repeated until no individual has unmatched contacts left. Let $z$ be an iteration
+counter for the matching algorithm and $i$ denote the individual whose unmatched
+contacts we are trying to match.
+
+Let $K_{z,i,c}$ denote the number of unmatched contacts of individual $i$ of contact type
+$c$ before iteration $z$ is completed. Note that $K_{z,i,c} \leq n_{ic}$ which is the
+total number of contacts individual $i$ has of type $c$.
+
+Let $a_i$ denote $i$'s age group and $county_i$ her county of residence.
+
+We first draw individual $j$ from the distribution defined by probability mass function
+$F_{z}$ over individuals $j \neq i$ in the synthetic population where the probability
+$f_{zj}$ is calculated as follows:
+
+\begin{equation}
+    \label{eq:combined_matching_probabilities}
+    f_{zj} = \underbrace{\alpha_{c,a_i,a_j, county_i, county_j}\vphantom{\frac{bla}{\sum_i^k}}}_{\text{Group Probability}} \cdot \underbrace{\frac{K_{z,j,c}}{\sum_{l=1}^N K_{z,l,c} \cdot \mathbb{I}_{county_l=county_j \land a_l = a_j}}}_{\text{Individual probability}}
+\end{equation}
+
+We then draw an individual $j$. If one of the two participants is susceptible and the
+other one is infectious, we sample whether an infection takes place. The success
+probability for this event is calculated as in
+Equation~\ref{eq:infection_probability_formula}. Finally we update the remaining
+numbers of unmatched contacts by setting:
+
+\begin{align}
+    K_{z+1,i,c} = K_{z, i, c} - 1
+    K_{z+1, j, c} = K_{z, j, c} - 1
+\end{align}
+
+The runtime of this algorithm scales roughly cubic in the number $N$ of simulated
+individuals. This is because the number of iterations is proportional to $N$, in each
+iteration we have evaluate Equation~\ref{eq:combined_matching_probabilities} $N$ times
+and each evaluation of that equation entails a sum over $N$ individuals.
+
+This makes it prohibitively expensive. We
+therefore replace the above algorithm by a two stage sampling procedure, where we
+first sample the group from which individual $j$ will be drawn according to the
+group probabilities defined in Equation~\ref{eq:combined_matching_probabilities}.
+Next we sample an individual from this group with the Individual probabilities defined
+in Equation~\ref{eq:combined_matching_probabilities}.
+
+Thus, while the calculation of any given second stage probability entails exactly the
+same number of calculations as before we do not have to calculate a second stage
+probability for all simulated individuals but only for those who are members of
+the group that was sampled in the first stage.
+
+It is easy to see that ex-ante the probability of being sampled are identical between
+the two stage sampling and the one stage sampling. The only drawback is
+that towards the end of the matching process it becomes possible to sample a group in
+which no unmatched contacts are left. In our empirical application this happens
+extremely rarely. This is so for two reasons: Firstly, the first stage sampling
+probabilities have been estimated from the same dataset as the numbers of contacts so
+there cannot be any mismatches such as for example a group that has a low probability
+of being sampled in the first stage but where all members have many contacts.
+Secondly, the group sizes are relatively large and we go over individuals in random
+order. Therefore, groups where no unmatched contacts remain only occur very late in the
+matching process.\footnote{If unmatched contacts were a concern one could simply
+use the fast two stage sampling process for a first pass over contacts and then
+match all remaining contacts with the slow algorithm.}

### 2021-07-13 src/parameter_tables/task_convert_csv_to_latex.py
Estimated parameter table in landscape format.

@@ -17,7 +17,7 @@ def task_convert_csv_to_tex(depends_on, produces):
         col_format = r"p{4cm}p{2cm}p{2cm}p{2cm}p{9cm}"
         float_format = "%.0f"
     elif len(df.columns) == 3:
-        col_format = r"p{3cm}p{1.5cm}p{6.5cm}"
+        col_format = r"p{3cm}p{2cm}p{15cm}"
         float_format = "%.3f"
     else:
         raise ValueError(f"Unexpected number of columns: {len(df.columns)}")

### 2021-07-13 src/paper/literature_review.tex
work on literature review.

@@
 
-The workhorse model in epidemiology is the SEIR model. However, The traditional SEIR
-model is not fine-grained enough to model nuanced policies.\comment[id=HM]{Still old
-motivation. Hard to get the timing of tests and results right?} This has motivated a
-large number of researchers to extend the standard model to allow for more heterogeneity
-and flexibility. Examples are \citet{Grimm2020}, \citet{Donsimoni2020} and
-\citet{Acemoglu2020}\comment[id=HM]{We need to cite more strategically. Cut Grimm,
-Donsinomi. More state of the art SEIR in other fields. Or drop SEIR altogether? If I
-read that as a Michael Meyer-Hermann type, I would think, ``ah, these economists
-again...''} who develop multi group SEIR models to analyze the effects of targeted
-lockdowns and \citet{Berger2020} who extend the SEIR model to analyze testing and
-conditional quarantines. For a more comprehensive review see \citet{Avery2020}. Others
-have used the results of a standard SEIR model as input for economic models that
-estimate the cost of policies (e.g. \citet{Dorn2020}).
-
-While the popularity of the SEIR model is mainly due to its simplicity, the extensions
-are quite complex. It is unlikely that there will be a SEIR model that combines all
-proposed extensions. Moreover, the extensions do not address other key issues: The main
-parameter of the SEIR model, the basic reproduction number ($R_0$), is not
-policy-invariant. It is a composite of the number of contacts each person has and the
-infection probability of the contacts. In fact, policy simulations are done by setting
-$R_0$ to a different value but it is hard to translate a real policy into the value of
-$R_0$ it will induce. In other words, SEIR models are not suited for evaluating the
-effect of policies which have never been experienced before.
-
-Another commonly used model class in epidemiology are agent-based simulation
+A commonly used model class in epidemiology are agent-based simulation
 models. In a prototypical agent-based simulation model, individuals are
@@
 testing can influence the infection dynamics. However, they do not calibrate their model
-match actual infection numbers which makes it more suitable to explore the general
+to match actual infection numbers which makes it more suitable to explore the general
 mechanics of different disease mitigation measures than for their quantitative
@@
 
-\citet{Bicher2021} simulate the Austrian population during the first wave (February 21
-to April 9, 2020). It uses the same data provided by \citet{Mossong2008} as we use to
-calibrate contact networks for households, workplaces and schools. The model focuses on
-analyzing the effect of different contact tracing strategies.
+\citet{Bicher2021} simulate the entire Austrian population. The use data from the first
+wave (February 21 to April 9, 2020) to calibrate ther model and predict the effect of
+different NPIs and contact tracing policies until November 2020. They the same data
+provided by \citet{Mossong2008} as we use to calibrate contact networks for households,
+workplaces and schools. The model focuses on analyzing the effect of different contact
+tracing strategies and not on modelling enacted Austrian policies over a long period
+of time.
 
 Moreover, there are several forthcoming applications\comment[id=HM]{What is that
-supposed to mean?} of agent-based simulation models with contact networks. Examples are
+supposed to mean?}\comment[id=J]{Does that mean we should not cite the working
+papers here?} of agent-based simulation models with contact networks. Examples are
 \citet{Basurto2020}, \citet{DelliGatti2020} and \citet{Mellacher2020}.
 
-Our model combines elements from the above models and adds several new features. To
-the best of our knowledge, ours is the only model that can explain observed case
-numbers with an excellent fit over more than 9 Months. We also have the most fine
-grained model of schools and NPIs that affect schools. Finally, we have an extremely
-detailed model of PCR and rapid tests with a share of detected cases that varies over
-time and across age groups.
+Our model combines elements from the above models and adds other ones. To the best of
+our knowledge, our model is the only one with the following features:
+
+1. The free model parameters have been estimated with the method of simulated moments
+\citep{McFadden1989}. Despite having few free parameters our model does en excellent
+job in explaining observed case numbers and the spread of the B.1.1.7 mutation over
+more than nine months of data.
+
+2. We have an extremely fine grained of schools and preschools.
+We can thus easily model all NPIs that have been implemented in Germany in the past 9
+months. This includes complete school closures, phases where only those students whose
+parents could not find any private childcare arrangement could attend, split class
+approaches for some or all age groups and combinations thereof. Moreover, we can
+account for additional hygiene measures whose effect is estimated inside the model.
+
+3. we model the evolution of the pandemic and all enacted policies over since the
+start of the second wave. Since the vast majority of cases has occurred in that time
+period and we also model unobserved infections our simulations take into account that
+many people are already immune because they are recovered from an infection and that
+this immunity is not spread randomly across the population.
+
+4. We have an extremely detailed model of PCR and rapid tests with a share of detected
+cases that varies over time and across age groups. We fit estimates of the share of
+known cases from other studies for time periods where those are available but also
+estimate a plausible share of known cases after vaccinations become widespread and
+methodologies that rely on a case fatality ratio to identify the share of known
+cases are not applicable anymore.

### 2021-07-13 src/paper/model/param_tables.tex
Nicely format tables.
Minor improvements.
Merge tables.
Shortened some parts of main text, added discussion of rapid tsts.

@@
+\clearpage
+
+\begin{landscape}
+
 \subsection{Overview Model Parameters} % (fold)
@@
 
-\begin{table}
+\comment[id=HM]{Copied from main text. Do we describe the parameter for fraction of
+vaccination-induced immunity/sterility  somewhere? I could not find it. Nothing special
+about the 75\%...
+
+Also, start out more positively. Something like ``We do not distinguish between first
+and second doses for computational reasons. We thus use a weighted average ...''
+
+75\% is lower than what is usually reported for after the second dose
+of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We choose it
+because our model neither includes booster shots, nor does it allow vaccinated
+individuals who became immune to transmit the disease\citep{Petter2021,
+LevineTiefenbrun2021, Pritchard2021}. If anything, these assumptions would overstate the
+effect of vaccines for our study period. This would be different if a large fraction of
+vaccinated individuals had received a second dose already.
+
+}
+
+\FloatBarrier
+
+\begin{table}[htb]
 \centering
-\caption{contacts_and_matching}
+\caption{Contacts, Matching and Policies}
 \label{tab:contacts_and_matching}
-\input{../bld/param_tables/contacts_and_matching}
+\input{../bld/param_tables/contacts_matching_policies}
 \end{table}
 
-\begin{table}
+\begin{table}[htb]
 \centering
-\caption{policies}
-\label{tab:policies}
-\input{../bld/param_tables/policies}
-\end{table}
-
-\begin{table}
-\centering
-\caption{infection_probs}
+\caption{Infection Probabilities and Virus Variants}
 \label{tab:infection_probs}
@@
 
-\begin{table}
+\begin{table}[htb]
 \centering
-\caption{disease_model}
-\label{tab:disease_model}
+\caption{Disease and Vaccination Model}
+\label{tab:disease_and_vacc_model}
 \input{../bld/param_tables/disease_model}
@@
 
-\begin{table}
+\begin{table}[htb]
 \centering
-\caption{pcr_testing_and_case_detection}
+\caption{Rapid Testing}
+\label{tab:rapid_testing}
+\input{../bld/param_tables/rapid_testing}
+\end{table}
+
+\begin{table}[htb]
+\centering
+\caption{PCR Testing, Case Detection and Behavioral Response}
 \label{tab:pcr_testing_and_case_detection}
@@
 
-\begin{table}
-\centering
-\caption{rapid_testing}
-\label{tab:rapid_testing}
-\input{../bld/param_tables/rapid_testing}
-\end{table}
+\FloatBarrier
 
-\begin{table}
-\centering
-\caption{behavioral_response}
-\label{tab:behavioral_response}
-\input{../bld/param_tables/behavioral_response}
-\end{table}
+\end{landscape}
 
-\begin{table}
-\centering
-\caption{vaccinations}
-\label{tab:vaccinations}
-\input{../bld/param_tables/vaccinations}
-\end{table}
+\FloatBarrier

### 2021-07-13 src/paper/report.tex
Shortened some parts of main text, added discussion of rapid tsts.

@@
 making it useful to predict the effects of various interventions. We apply the model to
-Germany, where new infections fell by almost 80\% during the May 2021. Our analysis shows
-that, aside from seasonality, frequent and large-scale rapid testing caused the bulk of
-this decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that
-it should have a large role for at least as long as vaccinations have not been offered to
-an entire population.
+Germany, where new infections fell by almost 80\% during the May 2021. Our analysis
+shows that, aside from seasonality, frequent and large-scale rapid testing caused the
+bulk of this decrease, which is in line with prior predictions \citep{Mina2021}. We
+conclude that it should have a large role for at least as long as vaccinations have not
+been offered to an entire population.
 
-At the core of our agent-based model are physical contacts between heterogeneous agents
-(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
-approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The model
-most closely related to ours is described in \citet{Hinch2020}.} Each contact between an
-infectious individual and somebody susceptible to the disease bears the risk of
-transmitting the virus\comment[id=K]{infection probability per contact for contact type
-$c$: $\beta_{c}$}. Contacts occur in up to four networks: Within the household, at work,
-at school, or in other settings (leisure activities, grocery shopping, medical
-appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
-applications can take the population and household structure from census data and the
-network-specific frequencies of contacts\comment[id=K]{probability to have $n$ contacts
-of contact type $c$: $\eta_{c,\:n}$} from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
-depend on age and geographical location (see Supplementary Material
-Section~\ref{subsec:assortativity}).\comment[id=K]{Share of contacts of age group and
+At the core of our agent-based model \citep[][we review more literature in Supplementary
+    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts
+between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
+the disease bears the risk of transmitting the virus\comment[id=K]{infection probability
+per contact for contact type $c$: $\beta_{c}$}. Contacts occur in up to four networks:
+Within the household, at work, at school, or in other settings (leisure activities,
+grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
+occur at random. Empirical applications can take the population and household structure
+from census data and the network-specific frequencies of
+contacts\comment[id=K]{probability to have $n$ contacts of contact type $c$:
+$\eta_{c,\:n}$} from diary data measuring contacts before the pandemic
+\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
+on age and geographical location (see Supplementary
+Material~\ref{subsec:assortativity}). \comment[id=K]{Share of contacts of age group and
 county of i to meet someone of j's age group and county: $\alpha_{work,\:age_i,
@@
         Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
-        influence on an agent's contacts to other agents. Demographic characteristics set
-        the baseline number of contacts in different networks ($\eta$). She may reduce
-        the number of contacts due to NPIs, showing symptoms, or testing positively for
-        SARS-CoV-2 ($\tau$). Infections may occur when a susceptible agent meets an
-        infectious agent; the probability depends on the type of contact ($\beta_c$), on
-        seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If infected, the infection
-        progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
-        tests are available, agents' demand is modeled as in
-        Figure~\ref{fig:model_rapid_tests}. All reasons trigger a test only for a
-        fraction of individuals depending on an individual compliance parameter; the
-        thresholds for triggering test demand differ across reasons and they may depend
-        on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
+        influence on an agent's contacts to other agents. Demographic characteristics
+        set the baseline number of contacts in different networks ($\eta$). She may
+        reduce the number of contacts due to NPIs, showing symptoms, or testing
+        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
+        agent meets an infectious agent; the probability depends on the type of contact
+        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
+        infected, the infection progresses as depicted in
+        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
+        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
+        trigger a test only for a fraction of individuals depending on an individual
+        compliance parameter; the thresholds for triggering test demand differ across
+        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
         Figure~\ref{fig:model_official_cases} shows the model of translating all
@@
         positive PCR tests triggered by symptoms ($\chi_{symptom}$), and on the false
-        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part of
-        the graph is relevant only for periods where rapid tests are available.}
+        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
+        of the graph is relevant only for periods where rapid tests are available.}
 \end{figure}
@@
 being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
-care, and recovery or death \citep[similar to][]{Grimm2021}.\comment[id=HM]{Need to look
-for non-economists -- this must be fairly standard?} The probabilities of transitioning
-between these states depend on age; their duration is random and calibrated to medical
-literature (for a detailed description see Section~\ref{sub:data_course_of_disease}).
-Conditional on the type of contact, infectiousness is independent of age
+care, and recovery or death \citep[similar to][]{Grimm2021}.\comment[id=HM]{Would the
+model in \citet{Aleta2020} work?} The probabilities of transitioning between these
+states depend on age; their duration is random and calibrated to medical literature (for
+a detailed description see Supplementary Material~\ref{sub:data_course_of_disease}). Conditional on the
+type of contact, infectiousness is independent of age
 \citep{Jones2021}.\comment[id=K]{would be $\beta_{c}$}
@@
 The model includes several other features, which are crucial to describe the evolution
-of the pandemic in 2020-2021. New virus strains with different profiles regarding
-infectiousness\comment[id=K]{$\sigma_v$} can be introduced. Agents may receive a
-vaccination. With a probability of 75\% \citep{Hunter2021}, vaccinated agents become
-immune and they do not transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
-Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the
-second dose of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We
-choose it because our model neither includes booster shots, nor does it allow vaccinated
-individuals who became immune to transmit the disease\citep{Petter2021,
-LevineTiefenbrun2021, Pritchard2021}. If anything, these assumptions would overstate the
-effect of vaccines for our study period. This would be different if a large fraction of
-vaccinated individuals had received a second dose already.} During the vaccine roll-out,
-priority may depend on age and occupation; some individuals will refuse vaccination
-offers.
+of the pandemic in 2020-2021. New virus strains with different infectiousness
+profiles\comment[id=K]{$\sigma_v$} may appear. Vaccines may become available. During the
+vaccine roll-out, priority may depend on age and occupation; vaccine hesitancy is
+modelled by some individuals refusing vaccination offers. With some probability,
+vaccinated agents become immune and do not transmit the virus \citep{Hunter2021,
+    LevineTiefenbrun2021, Petter2021, Pritchard2021}.
 
@@
 days} and there are aggregate capacity constraints. In contrast, rapid antigen tests
-yield immediate results; after a phase-in period, all tests that are demanded will be
-performed. Specificity\comment[id=K]{$p_{negative|not\:infected}$} and
+yield immediate results. Specificity\comment[id=K]{$p_{negative|not\:infected}$} and
 sensitivity\comment[id=K]{"$p_{positive|infected,\;i,\:t}$", where $i$ is the individual
@@
 according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the onset of infectiousness.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly.\comment[id=K]{"$\pi_{teacher,\:t}$"
-etc. } Rapid tests may be offered by employers to on-site workers. Individuals may
-demand tests for private reasons, which include having plans to meet other people,
-showing symptoms of CoViD-19, and because a household member tested positively for the
-virus. We endow each agent with an individual compliance parameter. This parameter
-determines whether she takes up rapid tests.\footnote{A positive test result (of either
-kind), as well as symptoms leads most individuals to reduce their contacts; this is why
-tests impact the actual contacts in Figure~\ref{fig:model_description}.}
+timing of the test relative to the onset of infectiousness. After a phase-in period, all
+tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
+model for rapid test demand. Schools may require staff and students to be tested
+regularly.\comment[id=K]{"$\pi_{teacher,\:t}$" etc.} Rapid tests may be offered by
+employers to on-site workers. Individuals may demand tests for private reasons, which
+include having plans to meet other people, showing symptoms of CoViD-19, and a household
+member having tested positively for the virus. We endow each agent with an individual
+compliance parameter. This parameter determines whether she takes up rapid
+tests.\footnote{Positive test results or symptoms leads most individuals to reduce their
+    contacts; this is why tests impact the actual contacts in
+    Figure~\ref{fig:model_description}.}
 
@@
 that we can use a wide array of data to identify and estimate the model's many
-parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
-materials for the precise sources.} Contact diaries yield pre-pandemic distributions of
-contacts for different contact types and their assortativity by age group. Mobility data
-is used to model the evolution of work contacts. School and daycare policies can be
-incorporated directly from official directives. Administrative records on the number of
-tests, vaccinations by age and region, and the prevalence of virus strains are generally
+parameters.\footnote{See Supplementary Material~\ref{sec:data_and_parameters} for a
+    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
+different contact types and their assortativity by age group. Mobility data is used to
+model the evolution of work contacts. School and daycare policies can be incorporated
+directly from official directives. Administrative records on the number of tests,
+vaccinations by age and region, and the prevalence of virus strains are generally
 available. Surveys may ask about test offers, propensities to take them up, and past
@@
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method, also described in
-Supplementary Material \ref{subsec:estimated_params}]{McFadden1989}. In our application,
-we keep the number of free parameters low in order to avoid overfitting. The data
-features to be matched include official case numbers for each age group and region,
-deaths, and the share of the B.1.1.7 strain.
+matches features of the data \citep[see][for the general method]{McFadden1989}. In our
+application, we keep the number of free parameters low in order to avoid overfitting.
+The data features to be matched include official case numbers for each age group and
+region, deaths, and the share of the B.1.1.7 strain.
 
 The main issue with official case numbers is that they will contain only a fraction of
-all infections. We thus model official cases as depicted in
-Figure~\ref{fig:model_official_cases}\comment[id=HM]{Replace P2, P(compliance) by
-symbols in figure}. We take aggregate estimates of the share of detected
-cases\comment[id=K]{$\psi$} and use data on the share of PCR tests administered because
-of CoViD-19 symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of asymptomatic
-individuals varies by age group, this gives us age-specific shares (see
+all infections. In the German case, this specifically amounts to positive PCR tests. We
+thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
+mortality-based\comment[id=HM]{added ``mortality-based'' -- correct?} aggregate
+estimates of the share of detected cases \comment[id=K]{$\psi$} and use data on the
+share of PCR tests administered because of CoViD-19
+symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of asymptomatic individuals
+varies by age group, this gives us age-specific shares (see
 Figure~\ref{fig:share_known_cases_by_age_group} for the share of detected cases by age
@@
 date. Schools and daycare centers were closed, so were customer-facing businesses except
-for grocery and drug stores. From the peak of the second wave just before Christmas until
-the trough in mid-February, newly detected cases decreased by almost three quarters. The
-third wave in the spring of 2021 is associated with the B.1.1.7 strain, which became
-dominant in March (Figure~\ref{fig:share_b117}) and stayed dominant for our entire
-simulation period.\footnote{B.1.617.2 (Delta) reached Germany in April but still
-accounted for less than 5\% of cases at the end of our simulation period.}
-In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
-stores were allowed to open again to the public. There were many changes in details of
-regulations afterwards, but they did not change the overall stringency index.
+for grocery and drug stores. From the peak of the second wave just before Christmas
+until the trough in mid-February, newly detected cases decreased by almost three
+quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
+strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
+    (Delta) reached Germany in April but still accounted for less than 5\% of cases at the
+    end of our simulation period.} In early March, some NPIs were relaxed; e.g.,
+hairdressers and home improvement stores were allowed to open again to the public. There
+were many changes in details of regulations afterwards, but they did not change the
+overall stringency index.
 
@@
         Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
-        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
-        value of one refers to the situation at the start of our sample period and zero
-        means that all NPIs included in the index are turned on. The other lines in
-        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
-        contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~\ref{subsec:policies} for separate plots of the three factors by contact
-        type.}
+        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
+        a value of one refers to the situation at the start of our sample period and
+        zero means that all NPIs included in the index are turned on. The other lines in
+        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
+        of contact reductions, increased hygiene regulations, and seasonality. See
+        Appendix~\ref{subsec:policies} for separate plots of the three factors by
+        contact type.}
 \end{figure}
@@
 be administered by medical doctors or in pharmacies. At-home tests approved by
-authorities became available in mid-March, rapid test centers were opened, and one test
+authorities became available in mid-March. Rapid test centers were opened, and one test
 per person and week was made available free of charge. In several states, customers were
@@
 use the model to predict recorded infection rates until the end of May 2021. See
-Supplementary Material sections~\ref{subsec:synthetic_population} and
+Supplementary Materials~\ref{subsec:synthetic_population} and
 \ref{sub:initial_conditions} for details. The blue line in
@@
 officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region, which are shown in the supplementary materials
-(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).
+geographical region (see Supplementary Material~\ref{sec:subsec:fit_results}).
 
@@
 no rapid tests or vaccinations, only the wildtype virus present), infections at the
-workplace would be reduced by 25\%. The lines show the product of the effect of contact
-reductions, increased hygiene regulations, and seasonality. Two aspects are particularly
-interesting. First, all lines broadly follow the stringency index and they would do so
-even more if we left out seasonality and school vacations (roughly the last two weeks of
-October, two weeks each around Christmas and Easter, and some days in late May). Second,
-the most stringent regulations are associated with the period of strong decreases in new
-infections between late December 2020 and mid-February 2021. The reversal of the trend
-is associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
-during May 2021 is associated with at least weekly rapid tests for 42\% of the
-population, a vaccination rate that rose from 28\% to 43\%, and seasonality lowering the
-relative infectiousness of contacts, most strongly for other contacts.
+workplace would be reduced by 25\%. Two aspects are particularly interesting. First, all
+lines broadly follow the stringency index and they would do so even more if we left out
+seasonality and school vacations (roughly the last two weeks of October, two weeks each
+around Christmas and Easter, and some days in late May). Second, the most stringent
+regulations coincide with the period of decreasing infection rates between late December
+2020 and mid-February 2021. The subsequent reversal of the trend is associated with the
+spread of the B.1.1.7 variant. During the steep drop in recorded cases during May 2021,
+for 42\% of the population took at least one rapid tests per week, the first-dose vaccination
+rate rose from 28\% to 43\%, and seasonality lowered the relative infectiousness
+of contacts.
 
 In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
-considers various scenarios. NPIs are always the same as in the baseline scenario.
+seasonality on the evolution of infections in 2021,
+Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always held
+constant at their values in the baseline scenario.
 Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
-and three scenarios turning each of these factors on individually.
+Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red
+line), and three scenarios turning each of these factors on individually.
 Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
@@
         to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and seasonality
-        remain at their January levels. The other scenarios turn each of these three
-        factors on individually. The decompositions in
+        variant is introduced in the same way; vaccinations, rapid tests, and
+        seasonality remain at their January levels. The other scenarios turn each of
+        these three factors on individually. The decompositions in
         Figures~\ref{fig:2021_scenarios_decomposition} and
@@
 lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
-cumulative difference between scenarios. Reasons for this are the slow start---it took
-until March~24th until 10\% of the population had received their first vaccination, the
-20\% mark was reached on April 19th---and the focus on older individuals. These groups
-contribute less to the spread of the disease than others due to a lower number of
-contacts. By the end of our study period, when first-dose vaccination rates reached
-around 40\% of the population, the numbers of new cases would have started to decline.
-It is important to note that the initial focus of the campaign was to prevent deaths and
-severe disease. Indeed, the case fatality was rate considerably lower during the third
-wave when compared to the second (4.4\% between October and February and 1.4\% between
-March and the end of May).\comment[id=K]{This needs a source.}\comment[id=HM]{Didn't we
-calculate that ourselves at some point?}
+cumulative difference between scenarios. Reasons for the low share are the slow
+start---it took until March~24th until 10\% of the population had received their first
+vaccination, the 20\% mark was reached on April 19th---and the focus on older
+individuals. These groups contribute less to the spread of the disease than others due
+to a lower number of contacts. By the end of our study period, when first-dose
+vaccination rates reached 43\% of the population, the numbers of new cases would have
+started to decline. It is important to note that the initial focus of the campaign was
+to prevent deaths and severe disease. Indeed, the case fatality was rate considerably
+lower during the third wave when compared to the second (4.4\% between October and
+February and 1.4\% between March and the end of May).\comment[id=K]{This needs a
+source.}\comment[id=HM]{Didn't we calculate that ourselves at some point?}
 
@@
 means that additional infections will be recorded which would otherwise remain
-undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large.
-Until late April, recorded cases are higher in the scenario with rapid testing alone when
-compared to the setting where none of the three mechanisms are turned on. The effect on
-total cases, however, is visible immediately in
+undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
+may persist for some time. Until late April, recorded cases are higher in the scenario
+with rapid testing alone when compared to the setting where none of the three mechanisms
+are turned on. The effect on total cases, however, is visible immediately in
 Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
-population performed weekly rapid tests in March on average, new infections on April 1
+population performed weekly rapid tests in March on average, new infections on April~1
 would have been reduced by 53\% relative to the scenario without vaccinations, rapid
@@
 come from tests at the workplace. Despite the fact that rapid tests for private reasons
-are phased in quite late, they make up for more than half of the total effect. The reason
-lies in the fact that a substantial share of these tests is driven by an elevated
-probability to carry the virus, i.e., showing symptoms of CoViD-19 or following up on a
-positive test of a household member. The latter is essentially a form of contact tracing,
-which has been shown to be very effective \citep{Contreras2021,
-Kretzschmar2020,Fetzer2020}.
+are phased in only in mid-March, they make up for more than half of the total effect.
+The reason lies in the fact that a substantial share of these tests is driven by an
+elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
+up on a positive test of a household member. The latter is essentially a form of contact
+tracing, which has been shown to be very effective \citep{Contreras2021,
+    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
+Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
+administered randomly in the population would not nearly have been as effective.
 
@@
 have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
-broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}. \comment[id=HM]{Need
-a back-of the envelope calculation} To use another metric, the effective weekly
-reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
-negative effects school closures have on children and parents \citep{Luijten2021,
-Melegari2021}---and in particular on those with low socio-economic status---these results
-in conjunction with hindsight bias suggest that opening schools combined with a testing
-strategy would have been beneficial. In other situations, and particular when rapid test
-are not available at scale, trade-offs may well be different.
+broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.
+\comment[id=HM]{Need a back-of the envelope calculation} To use another metric, the
+effective weekly reproduction number differs by $0.018$ and $0.052$, respectively. In
+light of the large negative effects school closures have on children and parents
+\citep{Luijten2021, Melegari2021}---and in particular on those with low socio-economic
+status---these results in conjunction with hindsight bias suggest that opening schools
+combined with a testing strategy would have been beneficial. In other situations, and
+particular when rapid test are not available at scale, trade-offs may well be different.
 
@@
 estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
-prices for rapid tests were below one Euro in early June 2021. Despite these large
-effects, the results on testing likely understate the benefits. Disadvantaged groups are
-less likely to be reached by testing campaigns relying on voluntary participation
-\citep[e.g.][]{StillmanTonin2021}; at the same time, these groups have a higher risk to
-contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the
-workplace will extend more into these groups. The same goes for individuals who exhibit
-a low level of compliance with CoViD-19-related regulations. Compared to vaccinations,
-rapid testing programmes allow a much quicker roll-out, making it arguably the most
-effective tool to contain the pandemic in the short run.
+prices for rapid tests were below one Euro in early June 2021. While we do not
+distinguish between self-administered rapid tests and point of care rapid tests, the
+former are likely to play a larger role for indication-driven testing. Widespread
+availability at low prices seems important. However, they rely on purely voluntary
+participation in a non-public setting. The benefit of point-of-care rapid tests as a
+precondition to participate in leisure activities as well as mandatory tests at the
+workplace or at school come from screening the entire population. This is important
+because disadvantaged groups are less likely to be reached by testing campaigns relying
+on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
+groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
+tests at school and at the workplace will extend more into these groups. The same goes
+for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
+Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
+it arguably the most effective tool to contain the pandemic in the short run.

### 2021-07-13 src/paper/results/rapid_tests_statistics.tex
Shortened some parts of main text, added discussion of rapid tsts.

@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.49\textwidth}
         \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
@@
     \hfill
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.49\textwidth}
         \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_positive}
@@
         \label{fig:pos_rapid_tests_by_channel}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_negative}
-        \caption{Share of the Population Testing Negative on a Given Day, by Channels}
-        \label{fig:neg_rapid_tests_by_channel}
     \end{subfigure}
@@
         to a PCR test, having a household member that has tested positive or developed
-        symptoms or having planned a weekly meeting with friends. The left figure shows
-        the share of the population doing a rapid test on a given day. The middle figure
-        shows the share of the population testing positive on a given day (true and
-        false positives). The right figure shows the share of the population testing
-        negative on a given day (true and false negatives).}
+        symptoms or having planned a weekly meeting with friends.
+        Panel~\subref{fig:rapid_test_demand_by_channel} shows the share of the
+        population doing a rapid test on a given day.
+        Panel~\subref{fig:pos_rapid_tests_by_channel} shows the share of the population
+        testing positive on a given day (true and false positives).}
 \end{figure}
@@
 Such a large effect of rapid tests seems to be at odds with the general perception that
-they are not very reliable. However, this is not the case, which can be seen from the
-number of
-false negative tests. At its peak at the beginning of May, there are more than 80 false
-negative rapid tests per million inhabitants per day. While this number might sound low
-in isolation, it becomes quite large when set in context: We estimate that at at the
-the same time only 200 cases per million are detected by rapid tests. This means that
-even at the time where discovery of cases via rapid tests is at its peak, almost
-30\% of people who are infected and make a test are not discovered by the test.
+they are not very reliable. However, this is not the case,\comment[id=HM]{I do not
+understand. 80 fals neg for 200 true pos does not seem too reliable? I guess the
+important bit is the reference cat. -- catching 200 pos vs. nothing} which can be seen
+from the number of false negative tests. At its peak at the beginning of May, there are
+more than 80 false negative rapid tests per million inhabitants per day. While this
+number might sound low in isolation, it becomes quite large when set in context: We
+estimate that at at the the same time only 200 cases per million are detected by rapid
+tests. This means that even at the time where discovery of cases via rapid tests is at
+its peak, almost 30\% of people who are infected and make a test are not discovered by
+the test.
 

### 2021-07-14 src/paper/report.tex
Fix broken references.

@@
 officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region (see Supplementary Material~\ref{sec:subsec:fit_results}).
+geographical region (see Supplementary Material~\ref{subsec:fit_results}).
 
@@
 have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
-broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.
-\comment[id=HM]{Need a back-of the envelope calculation} To use another metric, the
-effective weekly reproduction number differs by $0.018$ and $0.052$, respectively. In
-light of the large negative effects school closures have on children and parents
-\citep{Luijten2021, Melegari2021}---and in particular on those with low socio-economic
-status---these results in conjunction with hindsight bias suggest that opening schools
-combined with a testing strategy would have been beneficial. In other situations, and
-particular when rapid test are not available at scale, trade-offs may well be different.
+broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.\comment[id=HM]{Need
+a back-of the envelope calculation} To use another metric, the effective weekly
+reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
+negative effects school closures have on children and parents \citep{Luijten2021,
+Melegari2021}---and in particular on those with low socio-economic status---these results
+in conjunction with hindsight bias suggest that opening schools combined with a testing
+strategy would have been beneficial. In other situations, and particular when rapid test
+are not available at scale, trade-offs may well be different.
 

### 2021-07-14 src/paper/report.tex
Add paragraphs on work and school scenarios.

@@
 age\comment[id=K]{$\zeta_{a}$}. A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has
-lost the timing. Need to explain d-subscript, maybe even better to add stuff like 1-3
-days back in? Needs to be comprehensible without recurring to appendix} We differentiate
+Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has lost
+the timing. Need to explain d-subscript, maybe even better to add stuff like 1-3 days
+back in? Needs to be comprehensible without recurring to appendix} We differentiate
 between an initial period of infection without being infectious or showing symptoms,
 being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
-care, and recovery or death \citep[similar to][]{Grimm2021}.\comment[id=HM]{Would the
-model in \citet{Aleta2020} work?} The probabilities of transitioning between these
-states depend on age; their duration is random and calibrated to medical literature (for
-a detailed description see Supplementary Material~\ref{sub:data_course_of_disease}). Conditional on the
-type of contact, infectiousness is independent of age
-\citep{Jones2021}.\comment[id=K]{would be $\beta_{c}$}
+care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities of
+transitioning between these states depend on age; their duration is random and calibrated
+to medical literature (for a detailed description see Supplementary
+Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
+infectiousness is independent of age \citep{Jones2021}.\comment[id=K]{would be
+$\beta_{c}$}
 

### 2021-07-14 src/paper/results/scenarios.tex
Show URLs in the References.
Fix broken references.
Add paragraphs on work and school scenarios.

@@
-\subsection{Scenarios}
+\subsection{Scenarios} \comment[id=K]{Add parameter notation when writing this.}
 \label{subsec:appendix_scenarios}
+
+When a country is facing a surge in infections, policy makers face the difficult decision
+to weigh which instrument to apply and the extremity of each instrument. Here we
+complement our analysis of the effectiveness of vaccinations and rapid tests by showing
+the effects of rapid test policies vis-à-vis the more traditional NPIs, work from home
+mandates and school closures. All scenarios start after Easter (April 6). Our analyses
+show that many socially costly NPIs can be avoided through strong rapid testing policies.
+
+Figure~\ref{fig:work_scenarios_detailed} shows the effects of different work policies on
+the infections in the general population. We compare four scenarios with our baseline
+scenario: Keeping the share of workers having physical work contacts the same as in our
+baseline scenario the orange line shows what would have happened with rapid testing in
+firms at the level of mid March (orange line) where only 14\% of workers regularly did
+rapid tests. %
+We also include a scenario what would have happened if rapid tests had become truly
+mandatory after Easter\footnote{Starting on April 19th employers were required by law to
+provide two weekly tests to their employees \citep{Bundesanzeiger2021}. However,
+voluntarily only 60\% of workers regularly test themselves when offered tests
+(\cite{Betsch2021}, 20th/21st of April).}, assuming a 95\% compliance rate on both the
+employer and the employee side. On the work from home dimension we compare our baseline
+scenario with 10\% more or less work from home compared to the baseline scenario. For the
+total cases, the picture is very clear. Given the testing policy Germany had in place
+during that time (twice weekly tests done by 35\% to 50\% of workers over that time
+frame) whether 70\% (10\% below the actual mobility) or 85\% (10\% above the actual
+mobility) of workers attend work physically makes little difference for the incidence. On
+the other hand, the effect of a laxer or more ambitious testing policy for firms make an
+important difference: As can be seen in Figure~\ref{fig:work_scenarios_newly_infected}
+the gap between the two scenarios becomes quite sizable with a difference of over 80
+incidence points around May 1.\comment[id=K]{Talk about detected cases?}
 
@@
   \label{fig:work_scenarios_detailed}
-  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
-  policy changes took place at Easter until the end of our simulation period (end of
-  May). We vary the share of workers that work from home and how many tests are performed
-  at work relative to our baseline scenario. Making it mandatory to test all employees
-  that do not work from home markedly reduces cases -- even when only assuming 95\%
+  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after
+  different hypothetical policy changes take place at Easter until the end of our
+  simulation period. We vary the share of workers that work from home and how many tests
+  are performed at work relative to our baseline scenario. Making it mandatory to test
+  all employees that do not work from home markedly reduces cases -- assuming 95\%
   compliance on both the employer and the employee side. As before, the observed cases
@@
 
- \comment[id=K]{Regarding figure \ref{fig:work_scenarios_detailed}: Look up numbers and
- add them to the description.}
+\FloatBarrier
+
+The second commonly employed and also very contentious NPI we look at are school
+closures. Due to the very high incidence we approximate the German schooling policy as
+generous emergency care with rotating on-site schooling for graduating classes for April.
+In May where cases fall and schools gradually opened, we approximate the policy as
+rotating on-site schooling for most students (except for children eligible for emergency
+care and graduating classes who attend in full). We compare this baseline scenario to
+simply keeping schools completely closed (the brown line) and opening schools normally
+(but maintaining our hygiene multiplier to account for mask wearing, ventilation etc.)
+with and without tests.
+
+As can be seen, the transmission potential in schools is very low both in the generous
+emergency setting as well as the rotating operation. The difference to keeping schools
+completely closed for the general population is truly negligible. Also, consistent
+testing makes schools very safe to open. Would have schools opened directly after Easter
+given the testing rates Germany managed at schools during that time, the total incidence
+would have only been at most 14 points higher. Tests, however, are crucial here. Had
+schools opened completely without any testing of students and staff, schools would have
+become a driver of the pandemic with up to 50 incidence points more compared to the
+baseline scenario.
 
@@
   \label{fig:school_scenarios_detailed}
-  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after the
-   policy changes took place at Easter until the end of our simulation period (end of
-   May). Apart from the enacted school policies as our baseline we simulate how cases
-   would have developed if schools had been closed completely as the strictest possible
-   counterfactual scenario and two opening models: One where schools open normally (with
-   hygiene measures) without any testing in the education sector and one where schools
-   open normally but testing shares develop as in the baseline scenario. Our simulations
-   suggest that the enacted policies were as effective as keeping schools closed. Opening
-   schools with the testing schemes that were in place after Easter would have had a
-   small effect on the overall incidence. However, this is mainly due to the stringent
-   testing that was in place in schools by that time. Had schools opened without testing
-   requirements the total incidence would have been up to 50 points higher, though this
-   would have been less visible in the reported cases.}
+  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after
+   different hypothetical policy changes take place at Easter until the end of our
+   simulation period. Apart from the enacted school policies as our baseline
+   we simulate how cases would have developed if schools had been closed completely as
+   the strictest possible counterfactual scenario and two opening models: One where
+   schools open normally (with hygiene measures) without any testing in the education
+   sector and one where schools open normally but testing shares develop as in the
+   baseline scenario.
+   %Our simulations suggest that the enacted policies were as effective as keeping schools
+   %closed. Opening schools with the testing schemes that were in place after Easter would
+   %have had a small effect on the overall incidence. However, this is mainly due to the
+   %stringent testing that was in place in schools by that time. Had schools opened
+   %without testing requirements the total incidence would have been up to 50 points
+   %higher, though this would have been less visible in the reported cases.
+   }
 \end{figure}
-
- \comment[id=K]{Regarding figure \ref{fig:school_scenarios_detailed}: Look up numbers and
- add them to the description.}
 
@@
 
+ \comment[id=K]{Write that at the end of our simulation period nearly everyone in the
+ educ sector did rapid tests because they are mandatory but still 40-50\% refusers in the
+ work and private sector}
+
 \FloatBarrier

### 2021-07-14 src/paper/results/scenarios.tex
Add scenairo section.

@@
 
-When a country is facing a surge in infections, policy makers face the difficult decision
-to weigh which instrument to apply and the extremity of each instrument. Here we
-complement our analysis of the effectiveness of vaccinations and rapid tests by showing
-the effects of rapid test policies vis-à-vis the more traditional NPIs, work from home
-mandates and school closures. All scenarios start after Easter (April 6). Our analyses
-show that many socially costly NPIs can be avoided through strong rapid testing policies.
+Here we complement our analysis of the effectiveness of vaccinations and
+rapid tests by showing the effects of rapid test policies vis-à-vis the more traditional
+NPIs, work from home mandates and school closures. All scenarios start after Easter
+(April 6). Our analyses show that many socially costly NPIs can be avoided through strong
+rapid testing policies.
 
@@
 mobility) of workers attend work physically makes little difference for the incidence. On
-the other hand, the effect of a laxer or more ambitious testing policy for firms make an
-important difference: As can be seen in Figure~\ref{fig:work_scenarios_newly_infected}
-the gap between the two scenarios becomes quite sizable with a difference of over 80
-incidence points around May 1.\comment[id=K]{Talk about detected cases?}
+the other hand, the effect of a laxer or more ambitious testing policy for firms is
+sizable: As can be seen in Figure~\ref{fig:work_scenarios_newly_infected} the gap between
+the two scenarios grows to over 80 incidence points around May 1.%
+As in other scenarios, the observed cases can be misleading because more testing leads to
+more detected cases. It takes two to three weeks for the reduction in new infections to
+dominate the increased detection. Furthermore, the two opposing effects lead to a
+smaller effect size than is actually the case.
 
@@
   \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after
-  different hypothetical policy changes take place at Easter until the end of our
-  simulation period. We vary the share of workers that work from home and how many tests
-  are performed at work relative to our baseline scenario. Making it mandatory to test
-  all employees that do not work from home markedly reduces cases -- assuming 95\%
-  compliance on both the employer and the employee side. As before, the observed cases
-  can be misleading because more testing leads to more detected cases. It takes two to
-  three weeks for the reduction in new infections effect to dominate the increased
-  detection effect. Furthermore, the two opposing effects lead to a smaller effect size
-  than is actually the case.}
+  different hypothetical work policy changes take place at Easter until the end of our
+  simulation period. We vary the share of workers that have physical work contacts (10\%
+  more or less compared to the share in the baseline scenario, 85\% or 70\% of workers,
+  respectively) and how many tests are performed at work relative to our baseline
+  scenario. As an ambitious scenario we implement mandatory tests for all employees that
+  do not work from home, assuming 95\% compliance on both the employer and the employee
+  side. On the other hand, we show what would have happened if the test offers had fallen
+  back to the level of mid March (only 14\% of workers are tested regularly). The
+  observed cases can be misleading because more testing leads to more detected cases. It
+  takes two to three weeks for the reduction in new infections to dominate the increased
+  detection. Furthermore, the two opposing effects lead to a smaller effect size than is
+  actually the case.}
 \end{figure}
@@
 The second commonly employed and also very contentious NPI we look at are school
-closures. Due to the very high incidence we approximate the German schooling policy as
+closures. Due to the very high incidence we model the German schooling policy as
 generous emergency care with rotating on-site schooling for graduating classes for April.
-In May where cases fall and schools gradually opened, we approximate the policy as
+In May where cases fall and schools gradually opened, we model the policy as
 rotating on-site schooling for most students (except for children eligible for emergency
@@
 emergency setting as well as the rotating operation. The difference to keeping schools
-completely closed for the general population is truly negligible. Also, consistent
-testing makes schools very safe to open. Would have schools opened directly after Easter
-given the testing rates Germany managed at schools during that time, the total incidence
-would have only been at most 14 points higher. Tests, however, are crucial here. Had
+completely closed is very small. Also, consistent testing reduces the transmission
+potential at schools strongly. Had schools opened directly after Easter given the testing
+rates Germany managed at schools during that time, the total incidence would have been
+only been 9 incidence points higher on average. Tests, however, are crucial here. Had
 schools opened completely without any testing of students and staff, schools would have
-become a driver of the pandemic with up to 50 incidence points more compared to the
-baseline scenario.
+added up to 50 incidence points.
 
@@
   \floatfoot{\noindent \textit{Note:} The figure shows the development of cases after
-   different hypothetical policy changes take place at Easter until the end of our
-   simulation period. Apart from the enacted school policies as our baseline
-   we simulate how cases would have developed if schools had been closed completely as
-   the strictest possible counterfactual scenario and two opening models: One where
-   schools open normally (with hygiene measures) without any testing in the education
-   sector and one where schools open normally but testing shares develop as in the
-   baseline scenario.
-   %Our simulations suggest that the enacted policies were as effective as keeping schools
+   different hypothetical school policy changes take place at Easter until the end of our
+   simulation period. Apart from the enacted school policies as our baseline we simulate
+   how cases would have developed if schools had been closed completely as the strictest
+   possible counterfactual scenario and two opening models: One where schools open
+   normally (with hygiene measures) without any testing in the education sector and one
+   where schools open normally but testing develops as in the baseline scenario.   %Our simulations suggest that the enacted policies were as effective as keeping schools
    %closed. Opening schools with the testing schemes that were in place after Easter would
@@
 \end{figure}
+
+\FloatBarrier
+
+Lastly, we shed some light on the role our rapid test demand channels play
+for the effect of rapid tests on case numbers. To do so we ran two scenarios where we
+allocated rapid tests either completely randomly in the entire population or among 70\%
+of the population to account for the fact that a share of the population might refuse or
+be very hard to reach with rapid tests.\footnote{We calculate the number of rapid tests
+as in our baseline model. This leads to similar numbers of rapid tests. However given the
+higher incidence in our random scenarios these scenarios have a slightly higher number of
+rapid tests.}
+
+Figure \ref{fig:random_rapid_tests_detailed} shows how the incidence of detected and
+total cases develops in the two random scenarios (red and purple line) relative to our
+baseline scenario (blue line). Two things stand out: Firstly, the total number of cases
+falls much faster in our baseline scenario compared to the two random scenarios.
+Secondly, this is not because the share of detected cases is higher in the baseline
+scenario; in fact, it is even slightly lower until end of April.
+
+There are two mechanisms that can explain these surprising facts: Firstly, tests at the
+workplace predominantly target a group that has many contacts. Thus, catching infections
+in this group prevents more infections than in the general population. Secondly, rapid
+tests that are done because of private contact tracing are more effective at interrupting
+infection chains because they catch many infections in an early stage. Isolating infected
+individuals early on means that there are fewer days on which they can infect others.
+%
+The difference between the two random scenarios are small. This is likely due to only a
+small fraction of the population being tested on any given day.
 
@@
   \end{subfigure}
+
   \caption{The Role of Targeted and Compliance Driven Rapid Test Demand}
   \label{fig:random_rapid_tests_detailed}
-  \floatfoot{\noindent \textit{Note:} \textcolor{red}{To be written}}
+
+  \floatfoot{\noindent \textit{Note:} The figure shows the development of cases in two
+  scenarios where rapid tests are distributed randomly in the population compared to our
+  baseline scenario after Easter. In the baseline scenario rapid tests are targeted to
+  workers, students, teachers and individuals at high risk of being infected including a
+  weekly or twice weekly spacing between rapid tests. In the scenario with 30\% refusers
+  tests are randomly distributed among 70\% of the population who are identified as
+  compliers.
+  }
 \end{figure}
 
- \comment[id=K]{Describe that random tests are more effective at detecting cases and that
- explains the bump in the random test scenario detection in the beginning. However, cases
- fall immediately more in the non-random scenario because the targeted demand for tests
- by household members leads to a much more efficient interruption of infection chains
- because we catch the household members very early in their infectious period.}
-
- \comment[id=K]{Write that at the end of our simulation period nearly everyone in the
- educ sector did rapid tests because they are mandatory but still 40-50\% refusers in the
- work and private sector}
-
 \FloatBarrier

### 2021-07-14 src/paper/report.tex
Fix typos and remove notation comments.

@@
 over time. Containment measures have become increasingly diverse and now include rapid
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
+testing, more nuanced NPIs, and contact tracing. Neither these policies' effects nor the
 influence of seasonal patterns or of more infectious virus strains are well understood
@@
 making it useful to predict the effects of various interventions. We apply the model to
-Germany, where new infections fell by almost 80\% during the May 2021. Our analysis
+Germany, where new infections fell by almost 80\% during May 2021. Our analysis
 shows that, aside from seasonality, frequent and large-scale rapid testing caused the
@@
 between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
-the disease bears the risk of transmitting the virus\comment[id=K]{infection probability
-per contact for contact type $c$: $\beta_{c}$}. Contacts occur in up to four networks:
+the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
 Within the household, at work, at school, or in other settings (leisure activities,
@@
 from census data and the network-specific frequencies of
-contacts\comment[id=K]{probability to have $n$ contacts of contact type $c$:
-$\eta_{c,\:n}$} from diary data measuring contacts before the pandemic
+contacts from diary data measuring contacts before the pandemic
 \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
 on age and geographical location (see Supplementary
-Material~\ref{subsec:assortativity}). \comment[id=K]{Share of contacts of age group and
-county of i to meet someone of j's age group and county: $\alpha_{work,\:age_i,
-        age_j,\:county_i,\:county_j}$}
+Material~\ref{subsec:assortativity}).
 
@@
 upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from\comment[id=K]{$\rho_{attend,\,t}$}
+contact types. For example, a mandate to work from
 home will reduce the number of work contacts to zero for a fraction of the working
@@
 capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. Curfews\comment[id=K]{would
-enter into $\rho_{other\:,t}$} may reduce the number of contacts in settings outside of
+masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
+of contacts in settings outside of
 work and school. In any setting, measures like masking requirements would reduce the
 probability of infection associated with a contact
-\citep{Cheng2021}.\comment[id=K]{$\rho_{hygiene}$}
+\citep{Cheng2021}.
 
@@
 In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
-age\comment[id=K]{$\zeta_{a}$}. A possible infection progresses as shown in
+age. A possible infection progresses as shown in
 Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has lost
@@
 Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \citep{Jones2021}.\comment[id=K]{would be
-$\beta_{c}$}
+infectiousness is independent of age \citep{Jones2021}.
 
@@
 of the pandemic in 2020-2021. New virus strains with different infectiousness
-profiles\comment[id=K]{$\sigma_v$} may appear. Vaccines may become available. During the
+profiles may appear. Vaccines may become available. During the
 vaccine roll-out, priority may depend on age and occupation; vaccine hesitancy is
@@
 individual is infected or not; there is no uncertainty to the result. PCR tests require
-at least one day to be processed\comment[id=K]{$\gamma_{PCR,\:d}$ where $d=$ number of
-days} and there are aggregate capacity constraints. In contrast, rapid antigen tests
-yield immediate results. Specificity\comment[id=K]{$p_{negative|not\:infected}$} and
-sensitivity\comment[id=K]{"$p_{positive|infected,\;i,\:t}$", where $i$ is the individual
-and $t$ is the date to determine time until/since infectiousness} of these tests is set
+at least one day to be processed and there are aggregate capacity constraints. In
+contrast, rapid antigen tests
+yield immediate results. Specificity and
+sensitivity of these tests is set
 according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
@@
 model for rapid test demand. Schools may require staff and students to be tested
-regularly.\comment[id=K]{"$\pi_{teacher,\:t}$" etc.} Rapid tests may be offered by
+regularly. Rapid tests may be offered by
 employers to on-site workers. Individuals may demand tests for private reasons, which
@@
 mortality-based\comment[id=HM]{added ``mortality-based'' -- correct?} aggregate
-estimates of the share of detected cases \comment[id=K]{$\psi$} and use data on the
+estimates of the share of detected cases and use data on the
 share of PCR tests administered because of CoViD-19
-symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of asymptomatic individuals
+symptoms. As the share of asymptomatic individuals
 varies by age group, this gives us age-specific shares (see
@@
 compared to school age children. Once rapid test become available, confirmation of a
-positive result\comment[id=K]{prob of demanding confirmation is $\chi_{confirmation}$}
+positive result
 is another reason leading to positive PCR tests.

### 2021-07-14 src/paper/report.tex
Minor polishing of report.

@@
 the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
+ \comment[id=J]{I would replace networks by areas because we have more than four networks}
 Within the household, at work, at school, or in other settings (leisure activities,
@@
         Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
-        influence on an agent's contacts to other agents. Demographic characteristics
-        set the baseline number of contacts in different networks ($\eta$). She may
+        influence of an agent's contacts to other agents on infections. Demographic characteristics
+        set the baseline number of contacts in different networks ($\eta$). The agents may
         reduce the number of contacts due to NPIs, showing symptoms, or testing
@@
         infections in the simulated data to age-specific recorded infections. The model
-        uses data on the aggregate share of recorded cases ($\psi$), on the share of
-        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and on the false
+        uses data on the aggregate share of recorded cases ($\psi$), the share of
+        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and the false
         positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
-        of the graph is relevant only for periods where rapid tests are available.}
+        of the graph is relevant only for periods where rapid tests are available. All
+        parameters are explained in Appendix\ref{sub:param_tables}.}
 \end{figure}
@@
 compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms leads most individuals to reduce their
+tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
     contacts; this is why tests impact the actual contacts in
@@
 Modelling a population of agents according to actual demographic characteristics means
-that we can use a wide array of data to identify and estimate the model's many
+that we can use a wide array of data to identify and calibrate the model's many
 parameters.\footnote{See Supplementary Material~\ref{sec:data_and_parameters} for a
@@
 thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
-mortality-based\comment[id=HM]{added ``mortality-based'' -- correct?} aggregate
-estimates of the share of detected cases and use data on the
+mortality-based aggregate estimates of the share of detected cases and use data on the
 share of PCR tests administered because of CoViD-19
@@
 varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group} for the share of detected cases by age
-group over time in our model). Our estimates suggest that---in the absence of rapid
-testing---the detection rate is 80\% higher on average for individuals above age 80
-compared to school age children. Once rapid test become available, confirmation of a
-positive result
-is another reason leading to positive PCR tests.
+Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
+absence of rapid testing---the detection rate is 80\% higher on average for individuals
+above age 80 compared to school age children. Once rapid test become available,
+confirmation of a positive result is another reason leading to positive PCR tests.
 
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
+        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
+        For legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}
@@
 started to decline. It is important to note that the initial focus of the campaign was
-to prevent deaths and severe disease. Indeed, the case fatality was rate considerably
+to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
 lower during the third wave when compared to the second (4.4\% between October and
-February and 1.4\% between March and the end of May).\comment[id=K]{This needs a
-source.}\comment[id=HM]{Didn't we calculate that ourselves at some point?}
+February and 1.4\% between March and the end of May).
 
 Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
-observed and recorded cases would be reduced by a factor of four if only seasonality
+observed and total cases would be reduced by a factor of four if only seasonality
 mattered. However, in this period, cases would have kept on rising throughout, just at a
@@
 Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
-scenario without rapid tests only and the main specification into the three channels for
+scenario without rapid tests and the main specification into the three channels for
 rapid tests. Tests at schools have the smallest effect, which is largely explained by
@@
 Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not nearly have been as effective.
+administered randomly in the population would not have been nearly as effective.
 
 Two of the most contentious NPIs concern schools and mandates to work from
-home.\comment[id=K]{We should decide if we want to keep this section or prefer to show
-some rapid test statistics.} In many countries, schools switched to remote instruction
+home. In many countries, schools switched to remote instruction
 during the first wave, so did Germany. After the summer break, they were operating at
@@
 Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
-schools starting at Easter, at which point rapid tests had become widely available. We
+schools starting after Easter, at which point rapid tests had become widely available. We
 estimate the realized scenario to have essentially the same effect as a situation with
@@
         Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
-        because there were no capacity constraints for rapid tests afterwards.}
+        because there were no capacity constraints for rapid tests afterwards.
+        For legibility reasons, all lines are rolling 7-day averages.}
 
@@
 estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
+ \comment[id=J]{Check this number and add to which time period it refers}
 prices for rapid tests were below one Euro in early June 2021. While we do not

### 2021-07-14 src/paper/data/course_of_disease.tex
More polishing.

@@
 This section discusses the parameters governing the course of disease, their sources and
-how we arrived at the distributions used in the paper.\footnote{Additional information
-can be found in the
-\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-documentation}.} See Figure~\ref{fig:model_disease_progression} for a summary of our
+how we arrived at the distributions used in the paper.
+See Figure~\ref{fig:model_disease_progression} for a summary of our
 disease progression model.

### 2021-07-14 src/paper/results/fit.tex
More polishing.

@@
 look at observed infections (overall as well as by age group and federal state), the
-effective replication number,  the spread of the B.1.1.7 mutation,
-vaccinations.
+effective replication number,  the spread of B.1.1.7 and vaccinations.
 
@@
   averaged and show 30 simulation runs. The left figure shows the daily incidence rate
-  per million for the simulated reported infection rates. The mean infection rate is the
-  thick blue line. Single simulation runs are plotted in lighter and thinner lines. The
+  per million for the simulated reported infection rates. The
   official case numbers as reported by the RKI are plotted in black. The fit is overall
-  very good. The higher the mean incidence and the stronger the growth the more variance
-  there is between simulation runs. The right figure shows the effective replication
+  very good. The right figure shows the effective replication
   number ($R_t$) as reported by the RKI and as calculated in our model. The $R_t$ gives
@@
   being detected. In the short term this leads an $R_t$ estimation that is based on
-  detected cases to overestimate the replication number.}
+  detected cases to overestimate the replication number.
+  For legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}
@@
 fact that this age group has a very active social life which is not fully captured by our
-contact networks. The latter probably comes from too optimistic\comment[id=K]{cautious or
-slow? I think optimistic might be ambiguous here.} assumptions on school reopenings.
+contact networks. The latter probably comes from a too conservative model of school
+reopenings.
 
@@
   to 100 year olds live in these facilities. We average over 30
-  simulation runs.}
+  simulation runs. For legibility reasons, all lines are rolling 7-day averages.}
   \label{fig:age_group_fit}
@@
   simulated cases per one million people per day for different federal states. We
-  averaged over 30 simulation runs.}
+  averaged over 30 simulation runs. For legibility reasons, all lines are rolling 7-day averages.}
   \label{fig:state_fit}
@@
   the course of January. From then B.1.1.7 takes over endogenously through its increased
-  infectiousness ($\sigma_{B.1.1.7}$). We model no other features of B.1.1.7. We
-  introduce at most 0.986 cases per 100,000 inhabitants. The right figure shows the rate
+  infectiousness ($\sigma_{B.1.1.7}$). The right figure shows the rate
   of individuals that are vaccinated in our synthetic population versus in the general
-  German population. Note that we excluded the vaccinations that were given to nursing
-  homes (approx. one percent of the German population). Overall, our model covers a time
-  frame that goes from zero vaccinated individuals to a state where over 40\% of the
-  population are vaccinated. Our vaccinations work imperfectly but we do not model
-  different different vaccines nor do we distinguish between first and second shot.}
+  German population. For legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}
@@
 conservative. There are two reasons for this: Firstly, we do not model people who have
-made rapid tests out of curiosity once they became available. Secondly, in the model, the
+done rapid tests out of curiosity once they became available. Secondly, in the model, the
 decision to take a rapid test is based on a time invariant individual specific compliance
@@
     \label{fig:share_rapid_test_last_week2}
-    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
-        have ever done a rapid test or done a rapid test within the last week in our
-        simulations to the shares reported in \citet{Betsch2021b}. The left panel
-        compares the share of individuals who have ever done a rapid test. The right
+    \floatfoot{\noindent \textit{Note:} Both panels compare empirical and simulated
+        rapid test demands. The empirical data comes from \citet{Betsch2021b}.
+        The left panel compares the share of individuals who have ever done a rapid test.
+        The right
         panel compares the share of individuals who have done a rapid test within the
@@
         calibration of rapid tests are slightly conservative. The overall share is below
-        that in the study. We fit the share of weekly tests quite exactly. % However, the
+        that in the study. We fit the share of weekly tests quite exactly.
+        For legibility reasons, all simulated lines are rolling 7-day averages. % However, the
         % study only covers adults while our share also includes children who are tested

### 2021-07-14 src/paper/results/rapid_tests_statistics.tex
Fix broken references.
More polishing.

@@
 
-In order to make most use out of limited data sources on rapid test usage, we model the
+In order to make the most use out of limited data sources on rapid test usage, we model the
 number of performed rapid tests as a result of time invariant willingness to do rapid
@@
 
-We start with the share of the population doing a rapid test, receiving a positive rapid
-test and receiving a negative rapid test over time by the channel through which the test
+We start with the share of the population doing a rapid test and receiving a positive rapid
+test over time by the channel through which the test
 was demanded in Figures~\ref{fig:rapid_test_demand_by_channel},
-\ref{fig:pos_rapid_tests_by_channel} and \ref{fig:neg_rapid_tests_by_channel},
-respectively. Overall, the share of the population getting a rapid test on a given day
-increases from 2\% in mid March to over 10\% by May. The work rapid tests are a little
-ragged because of public holidays. For education rapid tests both vacations (first half
-of April) as well as the opening of schools in May are very visible in the rapid test
-demand. Overall, work tests make up the largest fraction of rapid tests. The image is
-very similar for the share of positive tests, except that the overall number of positive
-tests starts decreasing in May as rapid test expansion comes to a halt and cases fall,
-especially the positive share of private rapid tests falls as less and less individuals
-are triggered to seek a rapid test because of a risk contact in their household.
+\ref{fig:pos_rapid_tests_by_channel}, respectively. Overall, the share of the population
+getting a rapid test on a given day increases from 2\% in mid March to over 10\% by May.
+The work rapid tests are a little ragged because of public holidays. For education rapid
+tests both vacations (first half of April) as well as the opening of schools in May are
+very visible in the rapid test demand. Overall, work tests make up the largest fraction
+of rapid tests. The image is very similar for the share of positive tests, except that
+the overall number of positive tests starts decreasing in May as rapid test expansion
+comes to a halt and cases fall, especially the positive share of private rapid tests
+falls as less and less individuals are triggered to seek a rapid test because of a risk
+contact in their household.
 
@@
 per million individuals to make the metric comparable to incidences.
-% true positive
+
 The number of true positives (Figure~\ref{fig:rapid_tests_number_true_positive}) rapidly
@@
 share of them are targeted, i.e. triggered by events in the household. However, this
-does not mean that rapid tests in the workplace or at school are useless. It is rather
+does not mean that rapid tests in the workplace or at school are less important. It is rather
 the combination of large scale screening at work and in schools and very efficient
@@
 fact that there was a very large number of infected individuals who did not know they
-are positive. Detecting and isolating some of them is enough to slow down the
+are infected. Detecting and isolating some of them is enough to slow down the
 overall infection dynamic.
@@
     \vskip3ex
-    \caption{Rapid Test Results}
+    \caption{Simulated Rapid Test Statistics}
     \label{fig:rapid_test_results_numbers}
@@
         of tests, especially rapid tests at the workplace, also play an important role for
-        containing the pandemic.}
+        containing the pandemic. All results are averaged over 30 simulation runs.
+        For legibility reasons, all lines are rolling 7-day averages.
+        }
 \end{figure}
@@
 positive rate because those tests are primarily demanded when there is a high likelihood
-of being infected. The low false negative rate of 0.2\% looks very low. As discussed
+of being infected. The false negative rate of 0.2\% looks very low. As discussed
 above this is deceiving and just a mechanical consequence of a very low prevalence
@@
     \vskip3ex
-    \caption{Rapid Test Rates by Channel}
+    \caption{False Positive and False Negative Rates by Channel}
     \label{fig:rapid_test_results_rates}
@@
         much lower. The right panel shows the false negative rate in the population, i.e. the
-        share of negative test results that are mistakenly given to infected individuals.
+        share of negative tests done by infected individuals.
         This is very low because there are many truly negative tests in times of low

### 2021-07-14 src/paper/results/share_known_cases.tex
More polishing.

--- 2021-07-11 src/paper/results/share_known_cases.tex
@@
 
-We show the age group specific share of detected cases in
-Figure~\ref{fig:share_known_cases_by_age_group}.\footnote{We explain how we model the
-detection of cases in Section~\ref{sub:testing} and show the overall share of detected
-cases that we use as input for our testing model in
-Section~\ref{subsec:data_share_known_cases}.}
+This section shows the share of detected cases for different age groups. See
+Section~\ref{sub:testing} for an explanation of how we model the detection of cases
+and Section~\ref{subsec:data_share_known_cases} for the calibration of the relevant
+parameters.
 
-In 2020 the share of detected cases is much higher in older age groups as they are much
-more likely to develop symptoms and symptomatic individuals are much more likely to be
-tested. Starting in 2021 the share of detected cases not only depends on the overall
-share of detected cases we feed into our model but also on the number of performed rapid
-tests. As more and more (often asymptomatic and presymptomatic) individuals receive
-positive rapid tests and receive confirmation, the share of detected cases starts to
-increase, especially after March when rapid tests became much more widely available. This
-effect is very visible for school pupils, (5-14, green line). As twice weekly rapid tests
-become mandatory in schools and students start going to school after the Easter break,
-the share of detected cases for that group increases drastically from 25\% to over 60\%
-between April and June.
+The share of detected cases fall drastically from October to December when the
+incidence of CoViD-19 skyrocketed, PCR tests were still scarce and official contact
+tracing became impossible due to the sheer amount of cases.
+
+As rapid tests become available and more and more individuals receive positive rapid
+tests and seek PCR tests, the share of detected cases starts to increase. While first
+rapid tests are available since the beginning of 2021 the effect only becomes substantial
+after March when access to rapid tests was greatly expanded.
+
+Overall, the share of detected cases is much higher in older age groups. This is because
+the likelihood to develop symptoms increases with age and symptomatic cases are
+more likely to be detected.
+
+A notable exception is that school age children (5-14, green line) overtake the next age
+group in May 2021. This comes from a particularly strong increase in their share of
+detected cases after Easter, when weekly rapid tests become mandatory in schools.
 
@@
   \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
-  reported as an official case via PCR confirmation for each age group in our population.
-  We use the overall share of detected cases that was estimated through the case fatality
-  ratio by \citet{Dunkelzifferradar2020} for all of 2020 and then assume it to be
-  constant as vaccinations of the elderly strongly affect the case fatality rate which
-  the project does not account for. To get from an overall share of detected cases to the
-  share of cases that is detected in each age group we use that asymptomatic cases are
-  much less likely to be detected. As our model covers age specific asymptomatic rates
-  this endogenously leads to group specific share of detected cases that suggest that
-  infections in younger age groups are under-detected. Starting in 2021 in addition to
-  the overall numbers of detected cases through symptoms and the share of detected cases,
-  cases are also detected through confirmation of positive rapid tests. This leads to an
-  increase in the share of detected cases for all age groups but in particular for the
-  younger age groups that are covered extensively with rapid tests through the rapid test
-  requirement for participating in school. Age groups that have a high share of workers
-  also disproportionately increase their detection rate.}
+  reported as an official case for each age group in our simulated population.
+  For legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}

### 2021-07-14 src/paper/results/validation.tex
More polishing.

@@
 ($\rho_{w,\:attend,\:t}$) or the number of rapid tests (governed by the $\pi$ parameters)
-that are performed -- all of which are things that cannot be exactly known ex-ante.
+that are performed -- all of which are things that cannot be known exactly ex-ante.
 
@@
   and private rapid tests ($\pi_{private,\:t}$) it is reached between May 1 and June 10,
-  depending on the scenario. }
+  depending on the scenario.}
 \end{figure}
@@
   numbers that is jointly driven by seasonality, large scale rapid tests and
-  vaccinations.}
+  vaccinations. For legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}

### 2021-07-14 src/paper/report.tex
Go over report references. Add comments for unpublished articles.

@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many
-months---and often years---away from reaching vaccination levels that would end the
-pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
-of utmost importance to employ an effective mix of strategies for containing the virus.
-The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
-to reduce contacts between individuals. While this has allowed some countries to sustain
-equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
-    theoretical equilibrium at low case numbers which is sustained with
-    test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
-over time. Containment measures have become increasingly diverse and now include rapid
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effects nor the
-influence of seasonal patterns or of more infectious virus strains are well understood
-in quantitative terms.
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination levels that would end the pandemic or even
+protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate
+policies.}\comment[id=K]{Contreras2021 is not published but only on medrxiv} most have
+seen large fluctuations of infection rates over time. Containment measures have become
+increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
+tracing. Neither these policies' effects nor the influence of seasonal patterns or of
+more infectious virus strains are well understood in quantitative terms.
 
@@
 
-The model includes several other features, which are crucial to describe the evolution
-of the pandemic in 2020-2021. New virus strains with different infectiousness
-profiles may appear. Vaccines may become available. During the
-vaccine roll-out, priority may depend on age and occupation; vaccine hesitancy is
-modelled by some individuals refusing vaccination offers. With some probability,
-vaccinated agents become immune and do not transmit the virus \citep{Hunter2021,
-    LevineTiefenbrun2021, Petter2021, Pritchard2021}.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
+appear. Vaccines may become available. During the vaccine roll-out, priority may depend
+on age and occupation; vaccine hesitancy is modelled by some individuals refusing
+vaccination offers. With some probability, vaccinated agents become immune and do not
+transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021,
+Pritchard2021}.\comment[id=K]{Hunter2021, Petter2021 and Pritchard not published
+(medrxiv)}
 
@@
 much lower pace \citep[this is in line with results in][, which our seasonality measure
-    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
-quantitatively important factor determining the evolution of the pandemic, explaining
-most of the early changes and 43\% of the cumulative difference by the end of May.
+is based on]{Gavenciak2021}.\comment[id=K]{Gavenciak2021 is not published} Nevertheless,
+we estimate seasonality to be a quantitatively important factor determining the evolution
+of the pandemic, explaining most of the early changes and 43\% of the cumulative
+difference by the end of May.
 
@@
 because disadvantaged groups are less likely to be reached by testing campaigns relying
-on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
-groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
-tests at school and at the workplace will extend more into these groups. The same goes
-for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
-Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
-it arguably the most effective tool to contain the pandemic in the short run.
+on voluntary participation \citep[e.g.][]{StillmanTonin2021};\comment[id=K]{IZA DP} at
+the same time, these groups have a higher risk to contract CoViD-19
+\citep{KochInstitut2021a}. Mandatory tests at school and at the workplace will extend
+more into these groups. The same goes for individuals who exhibit a low level of
+compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
+programmes allow a much quicker roll-out, making it arguably the most effective tool to
+contain the pandemic in the short run.

### 2021-07-14 src/paper/literature_review.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

@@
 
-\citet{Bicher2021} simulate the entire Austrian population. The use data from the first
-wave (February 21 to April 9, 2020) to calibrate ther model and predict the effect of
-different NPIs and contact tracing policies until November 2020. They the same data
-provided by \citet{Mossong2008} as we use to calibrate contact networks for households,
+\citet{Bicher2021} simulate the entire Austrian population. They use data from the first
+wave (February 21 to April 9, 2020) to calibrate their model and predict the effect of
+different NPIs and contact tracing policies until November 2020. They use the same data
+provided by \citet{Mossong2008} as we to calibrate contact networks for households,
 workplaces and schools. The model focuses on analyzing the effect of different contact
@@
 
-Our model combines elements from the above models and adds other ones. To the best of
+Our model combines elements from the above models and adds several others. To the best of
 our knowledge, our model is the only one with the following features:
@@
 1. The free model parameters have been estimated with the method of simulated moments
-\citep{McFadden1989}. Despite having few free parameters our model does en excellent
+   \citep{McFadden1989}. Despite having few free parameters our model does an excellent
    job in explaining observed case numbers and the spread of the B.1.1.7 mutation over
@@
 
-2. We have an extremely fine grained of schools and preschools.
-We can thus easily model all NPIs that have been implemented in Germany in the past 9
-months. This includes complete school closures, phases where only those students whose
-parents could not find any private childcare arrangement could attend, split class
-approaches for some or all age groups and combinations thereof. Moreover, we can
-account for additional hygiene measures whose effect is estimated inside the model.
+2. We have an extremely fine grained \replaced[id=K]{representation}{} of schools and
+   preschools. We can thus easily model all \replaced[id=K]{schooling policies}{NPIs}
+   that have been implemented in Germany in the past nine months. This includes complete
+   school closures, phases where only those students whose parents could not find any
+   private childcare arrangement could attend, split class approaches for some or all age
+   groups and combinations thereof. Moreover, we can account for additional hygiene
+   measures whose effect is estimated inside the model.
 
-3. we model the evolution of the pandemic and all enacted policies over since the
-start of the second wave. Since the vast majority of cases has occurred in that time
-period and we also model unobserved infections our simulations take into account that
-many people are already immune because they are recovered from an infection and that
-this immunity is not spread randomly across the population.
+3. We model the evolution of the pandemic and all enacted policies since the start of the
+   second wave. Since the vast majority of cases has occurred in that time period and we
+   also model unobserved infections our simulations take into account that many people
+   are already immune because they have recovered from an infection and that this
+   immunity is not spread randomly across the population.
 
@@
    cases that varies over time and across age groups. We fit estimates of the share of
-known cases from other studies for time periods where those are available but also
-estimate a plausible share of known cases after vaccinations become widespread and
-methodologies that rely on a case fatality ratio to identify the share of known
-cases are not applicable anymore.
+   known cases from other studies for time periods where those are
+   available\comment[id=K]{Citation here?} but also estimate a plausible share of known
+   cases after vaccinations become widespread and methodologies that rely on the case
+   fatality ratio to identify the share of known cases are not applicable anymore.
+
+ \comment[id=K]{Four is of course plenty but what about weekly recurring contacts or using the google mobility data? -> maybe add these to point 2?}

### 2021-07-14 src/paper/model/course_of_disease.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

@@
 Figure~\ref{fig:model_disease_progression} and the values and source of the relevant
-parameters are describes in Section~\ref{sub:data_course_of_disease}.
+parameters are described in Section~\ref{sub:data_course_of_disease}.
 
@@
 remains in a state is drawn randomly. The parameters that govern these processes are
-taken from the literature \footnote{ Detailed information on the calibration of the
-disease parameters is available as part of our
-\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
-documentation}.}.\comment[id=J]{Replace this by a link to the relevant table with
-parameters in the data appendix, once we have that table.}
+taken from the literature and detailed in Section~\ref{sub:data_course_of_disease}. For
+an overview of our disease progression parameters see
+Table~\ref{tab:disease_and_vacc_model}.\comment[id=K]{@Janos: Is this what you wanted
+("Replace this by a link to the relevant table with parameters in the data appendix, once
+we have that table")}

### 2021-07-14 src/paper/model/matching.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

@@
 The empirical data described above only allows to estimate the number of contacts each
-person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
+person has. In order to simulate transmissions of Covid-19, the numbers of contacts have
 to be translated into actual meetings between people. This is achieved by a matching
@@
 function that takes the value 0 (non-participating) and 1 (participating) and can depend
-on the weekday, date and health state. This can be used to model vacations, weekends or
-symptomatic people who stay home (see section \ref{sec:policies} for details).
+on the weekday, date and health state.\comment[id=K]{Actually the full states and params
+enter into the contact functions} This can be used to model things like vacations,
+weekends or symptomatic people who stay home (see section \ref{sec:policies} for
+details).
 
@@
 
-where $beta_c$ denotes the base infection probability of contact type $c$,
+where $\beta_c$ denotes the base infection probability of contact type $c$,
 $s_{c, t}$ is a seasonality factor between zero and one that depends on the contact
@@
 
-The matching of non-recurrent contact types is more difficult because the contact
-network is resampled randomly every day. Moreover, it needs to allow for
-assortative matching. In our application, all
-random contacts are assortative with respect to age group $a$ (it is usually more likely
-to meet
-people from the same age group) and county (it is more likely to meet people from the
-same county) but in principle any set of discrete variables can be used. This set of
-variables that influence matching probabilities introduce a discrete partition of the
-population into groups.
+The matching of non-recurrent contact types is more difficult because the contact network
+is resampled randomly every day. Moreover, it needs to allow for assortative matching. In
+our application, all random contacts are assortative with respect to age group $a$ (it is
+usually more likely to meet people from the same age group) and county (it is more likely
+to meet people from the same county) but in principle any set of discrete variables can
+be used. This set of variables that influence matching probabilities
+\replaced[id=K]{induce}{introduce} a discrete partition of the population into groups.
 
@@
     \label{eq:combined_matching_probabilities}
-    f_{zj} = \underbrace{\alpha_{c,a_i,a_j, county_i, county_j}\vphantom{\frac{bla}{\sum_i^k}}}_{\text{Group Probability}} \cdot \underbrace{\frac{K_{z,j,c}}{\sum_{l=1}^N K_{z,l,c} \cdot \mathbb{I}_{county_l=county_j \land a_l = a_j}}}_{\text{Individual probability}}
+    f_{zj} = \underbrace{\alpha_{c,a_i,a_j, county_i, county_j}\vphantom{\frac{bla}{\sum_i^k}}}_{\text{Group Probability}} \cdot \underbrace{\frac{K_{z,j,c}}{\sum_{l=1}^N K_{z,l,c} \cdot \mathbb{I}_{county_l=county_j \land a_l = a_j}}}_{\text{Individual Probability}}
 \end{equation}
@@
 \begin{align}
-    K_{z+1,i,c} = K_{z, i, c} - 1
-    K_{z+1, j, c} = K_{z, j, c} - 1
+    K_{z+1,i,c} &= K_{z, i, c} - 1 \\
+    K_{z+1, j, c} &= K_{z, j, c} - 1
 \end{align}

### 2021-07-14 src/paper/model/policies.tex
Add first version of estimated parameters table.
Fix typos, make suggestions and add comments to sections B.1 to B.7.

--- 2021-06-18 paper/model/policies.tex
@@
 \label{sec:policies}
+
+ \comment[id=K]{I think we need to make sure we distinguish very clearly what is possible
+ and what's implemented.}
 
@@
 quantify the effect of novel NPIs. Secondly, it allows to model the actually implemented
-policy environment in great detail, which is necessary to use use the full time series
+policy environment in great detail, which is necessary to use the full time series
 of infections and fatality rates to estimate the model parameters.\footnote{
@@
 The most complex policies are typically found in the education sector. Since the
-beginning of 2021 schools have switched back and fourth between full closures,
-split class approaches with alternating schedules for some or all age groups and
-reopening while maintaining hygiene measures. On top of that there are different
-policies for allowing young students whose parents work full time to attend school
-even on days where they normally would not. For details on how we calibrate these
-policies see Section~\ref{subsec:policies_seasonality}.
+beginning of \replaced[id=K]{2020}{2021} schools have switched back and fourth between
+full closures, split class approaches with alternating schedules for some or all age
+groups and reopening while maintaining hygiene measures. On top of that there are
+different policies for allowing young students whose parents work full time to attend
+school even on days where they normally would not. For details on how we calibrate these
+policies see Section~\ref{subsec:policies}.\comment[id=K]{Did I write enough on the school policies? (No sources or details so far...)}
 
 Importantly, policies can depend on the health states of participating individuals.
-This allows to quarantine entire school classes if one student tested positive or
-to implement official or private contact tracing.
+\replaced[id=K]{Thus, it would be possible}{This allows} to quarantine entire school
+classes if one student tested positive \replaced[id=K]{and many other forms of contact
+tracing}{or to implement official or private contact tracing}.
 
-For some policies the exact effect on each contact type is not easy to determine. If
-this refers to a policy has been active during the estimation period, it is possible to
+For some policies the exact effect on each contact type is not easy to determine. If this
+refers to a policy that has been active during the estimation period, it is possible to
 estimate such parameters by fitting the model to time series data of infection rates.
-This is only possible if the policy was not active during the whole estimation period
-and thus the infection probabilities can be identified separately. We do this to account
-for hygiene measures at school and in the workplace that have been in effect since
-November 2020.
+This is only possible if the policy was not active during the whole estimation period and
+thus the infection probabilities can be identified separately. We do this to account for
+hygiene measures at school and in the workplace that have been in effect since November
+2020.\comment[id=K]{I don't understand this paragraph.}
+
+Not all things that reduce contacts compared to the pre-pandemic level are driven by
+NPIs. Therefore, we also model endogenous contact reductions that can depend on the
+health state of individuals \replaced[id=K]{}{, known risk contacts} or the local
+incidence. Examples in our model are contact reductions for symptomatic individuals or
+those who have a positive PCR or rapid tests \replaced[id=K]{or contact reductions when a
+household member tested positive}. The extent to which contacts are reduced can be
+calibrated from surveys. For an application of our model showcasing private contact
+tracing in the context of the Christmas holidays see \cite{Gabler2020}.

### 2021-07-14 src/paper/model/summary.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

--- 2021-06-21 src/paper/model/summary.tex
@@
 To predict and quantify the effects of a wide variety of fine-grained social distancing
-policies, vaccinations and rapid testing, we propose a different model structure. Our
-model inherits many features from prototypical agent-based simulation models but
-replaces the contacts between moving particles by contacts between individuals who work,
-go to school, live in a household and enjoy leisure activities.
+policies, vaccinations and rapid testing, we \replaced[id=K]{use the following}{propose a
+different} model structure. \replaced[id=K]{}{It inherits many features from prototypical
+agent-based simulation models but replaces the contacts between moving particles by
+contacts between individuals who work, go to school, live in a household and enjoy
+leisure activities.}\comment[id=K]{This has been moved to the Literature Review, I
+think.}
 
@@
 Once a person is infected, the disease progresses in a fairly standard way which is
-depicted in Figure~\ref{fig:model_disease_progression}. Asymptomatic cases and cases with mild
-symptoms are infectious for some time and recover eventually. Cases with severe symptoms
-additionally require hospitalization and lead to either recovery or death.
+depicted in Figure~\ref{fig:model_disease_progression}. Asymptomatic cases and cases with
+mild symptoms are infectious for some time and recover eventually. Cases with severe
+symptoms additionally require hospitalization and lead to either recovery or death.
 
 After rapid tests become available, people who work or go to school can receive rapid
-tests there. Moreover, people can decide to make a rapid test if they develop symptoms,
-have many planned contacts or observe cases in their contact network. People who have a
-positive rapid test demand a confirmatory PCR test with a certain probability. Moreover,
-PCR tests can be demanded because of symptoms or randomly.
+tests there. Moreover, people can decide to do a rapid test if they develop symptoms,
+have many planned contacts\comment[id=K]{at the moment it's implemented as planing to
+participate in a weekly leisure meeting. Should we be more explicit here?} or
+\replaced[id=K]{have a sick or positively tested household member}{observe cases in their
+contact network}. People who have a positive rapid test demand a confirmatory PCR test
+with a certain probability. Moreover, PCR tests can be demanded because of symptoms or
+randomly.
 
@@
 
-People who have symptoms, received a positive test, or had a risk contact can reduce
-their number of contacts across all contact types endogenously. The extent to which this
-is done is calibrated from survey data.
+People who have symptoms\replaced[id=K]{ or}{,} received a positive test,
+\replaced[id=K]{}{or had a risk contact} can reduce their number of contacts across all
+contact types endogenously. The extent to which this is done is calibrated from survey
+data.
 
@@
 or test demand models. The main bottleneck is not complexity or computational cost
-but the availability of data to calibrate the additional model features.
+but the availability of data to calibrate additional model features.

### 2021-07-14 src/paper/model/testing.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

@@
 
-Having a realistic model of PCR and rapid tests is crucial for two reasons: Firstly,
-only via a testing model the simulated infections from the model can be made comparable
-to official case numbers. Secondly, individuals with undetected or not yet detected
-infections are an important driver of the pandemic.
+ \comment[id=K]{Talk here about reactions to tests
+and symptoms with contact reductions. (If not here, find other place where to talk about
+this in detail.)}
+
+Having a realistic model of PCR and rapid tests is crucial for two reasons: Firstly, only
+via a testing model can the simulated infections from the model be made comparable to
+official case numbers. Secondly, individuals with undetected or not yet detected
+infections are an important driver of the pandemic: \textcolor{blue}{One of the features
+that makes SARS-CoV-2 so hard to control is that it can be transmitted both by pre- and
+asymptomatic individuals which makes it difficult for individuals to isolate when they
+pose a risk to others. Our model captures this because developing symptoms or receiving a
+positive test result leads many people to reduce their contacts, especially contacts
+outside the own household.}\comment[id=K]{I think these multipliers $\tau$ are not
+mentioned in the data section.}
 
@@
 
-Therefore, we aim for a simpler model of PCR and rapid tests that can be calibrated with
-available data on test demand and availability and -- nevertheless -- can produce a share
-of undetected cases that varies over time and across age groups and agrees with other
-estimates over the time periods where they are available.
+Therefore, we \replaced[id=K]{build}{aim for} a simpler model of PCR and rapid tests that
+can be calibrated with available data on test demand and availability and -- nevertheless
+-- can produce a share of undetected cases that varies over time and across age groups
+and agrees with other estimates over the time periods where they are available.
 
 PCR tests are modeled since the beginning of the simulation and determine whether a
-infections is officially recorded. Rapid tests are only added at the beginning of 2021.
+infection is officially recorded. Rapid tests are only added at the beginning of 2021.
 Positive rapid tests do not enter official case numbers directly, but most people with a
@@
 At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
-of detected cases from \citet{Dunkelzifferradar2020} can no longer be used because it is
-based on case fatality rates which drastically change due to vaccinations. Secondly,
-rapid tests become available at a large scale.
+of detected cases from Dunkelzifferradar project \citep{Dunkelzifferradar2020} can no
+longer be used because it is based on the case fatality rate which changes drastically
+due to vaccinations. Secondly, rapid tests become available at a large scale.
 
@@
 tests in the model fits the empirical data well (where empirical data is available).
+ \comment[id=K]{Maybe refer to the more detailed \ref{fig:share_rapid_test_last_week2}}
 
 In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
-falsely negative. While the specificity of rapid tests is constant over time, their
-sensitivity strongly depends on whether the tested individual is already infectious and
-if so for how long he has been infectious. Before the onset of infectiousness the
-sensitivity is very low (35\%). On the first day of infectiousness it is much higher
-(88\%) but still lower than during the remaining infectious period (92\%). After
-infectiousness stops, the sensitivity drops to 50 \%.
+falsely negative. While the specificity of rapid tests is constant, their sensitivity
+strongly depends \replaced[id=K]{on the timing of the rapid test relative to the start of
+infectiousness}{on whether the tested individual is already infectious and if so for how
+long he has been infectious}. Before the onset of infectiousness the sensitivity is very
+low (35\%). On the first day of infectiousness it is much higher (88\%) but still lower
+than during the remaining infectious period (92\%). After infectiousness stops, the
+sensitivity drops to 50 \%.
 
@@
 
-While rapid tests do not directly enter official case numbers, many positively tested
-individuals confirm their rapid test result with a PCR test.\comment[id=J]{Need to link
-to data appendix once the number is there and maybe mention the number}. Importantly,
-those PCR tests are made in addition to the tests that would have been done otherwise.
-Section~\ref{subsec:results_share_known_cases} discusses the effect of rapid tests on
-the share of detected cases.\comment[id=K]{Talk here about reactions to tests and
-symptoms with contact reductions. (If not here, find other place where to talk about
-this in detail.)}
+While rapid tests do not directly enter official case numbers, 82\%
+($\chi_{confirmation}$) of positively tested individuals seek a PCR test
+\citep{Betsch2021}.\comment[id=J]{Need to link to data appendix once the number is there
+and maybe mention the number}\comment[id=K]{Not mentioned in the data section.}
+Importantly, those PCR tests are made in addition to the tests that would have been done
+otherwise. Section~\ref{subsec:results_share_known_cases} discusses the effect of rapid
+tests on the share of detected cases.

### 2021-07-14 src/paper/model/estimated_parameters.tex
Estimated parameter table in landscape format.
Finish going over appendix B. More type corrections, suggestions and comments.

@@
 
-We estimate parameters that cannot be calibrated outside of the model with the
-method of simulated moments \citep{McFadden1989} by minimizing the distance between
-simulated and observed infection rates and fatality rates (disaggregated by region
-and age groups). Since our model includes a lot of randomness, we
-average simulated infection rates over several model runs.
+We estimate parameters that cannot be calibrated outside of the model with the method of
+simulated moments \citep{McFadden1989} by minimizing the distance between simulated and
+observed infection rates (disaggregated by region and age groups) and fatality rates.
+Since our model includes a lot of randomness, we average simulated infection rates over
+several model runs.
 
 All estimated parameters are described in Table~\ref{tab:estimated_params}.
+
+\begin{landscape}
 
@@
 \end{table}
+
+\end{landscape}
 
@@
 To avoid over-fitting and simplify the numerical optimization problem, we only allow for
-five different infection probabilities: 1) for contacts in schools 2) contacts in
+five different infection probabilities: 1) for contacts in schools 2) for contacts in
 preschools and nurseries. 3) for work contacts. 4) for households. 5) for other
@@
 type. This is depicted in Figure~\ref{fig:infection_heatmap}. We can see that work and
-other contacts are the main driver of the pandemic, followed by infections in
+other contacts are the main drivers of the pandemic, followed by infections in
 households. Schools and preschools contribute fewer infections which is to be expected
@@
     \caption{Daily share of infections by contact type}
-    \floatfoot{\noindent \textit{Note:} Daily hare of infections that were contributed
+    \floatfoot{\noindent \textit{Note:} Daily share of infections that were contributed
     by each contact type. Darker colors mean that a larger share of infections were
@@
 
-Moreover, we estimate nine different multipliers that reflect how strongly other
-contacts are reduced over time. The dates at which we switch between the multipliers
-usually coincides with policy changes and is not determined from the data. The only
-exception to this are slight adjustments to parameters to incorporate lockdown fatigue
-(towards the end of a lockdown period) or precautionary contact reductions (in times
-of high incidences right before a lockdown is enacted). The estimated other multipliers
-are also depicted in Figure~\ref{fig:other_multiplier}.
+Moreover, we estimate nine different multipliers that reflect how strongly other contacts
+are reduced over time. The dates at which we switch between the multipliers usually
+coincide with policy changes and is not determined from the \replaced[id=K]{case
+numbers}{data}. The only exception to this are slight adjustments to parameters to
+incorporate lockdown fatigue (towards the end of a lockdown period) or precautionary
+contact reductions (in times of high incidences right before a lockdown is enacted). The
+estimated other multipliers are also depicted in Figure~\ref{fig:other_multiplier}.
 

### 2021-07-14 src/paper/model/initial_conditions.tex
Finish going over appendix B. More type corrections, suggestions and comments.

--- 2021-04-17 paper/model/initial_conditions.tex
@@
 of your simulation. Additionally, the sample of infectious people who will determine the
-course of the pandemic in the following periods is likely not representative of the
-whole population because of differences in behavior (number of contacts,
-assortativity), past policies (school closures), etc.. The distribution of courses of
-diseases in the population at the begin of the simulation is called initial conditions.
+course of the pandemic in the following periods is likely not representative of the whole
+population because of differences in behavior (number of contacts, assortativity), past
+policies (school closures), etc.. The distribution of \replaced[id=K]{health
+states}{courses of diseases} in the population at the beginning of the simulation is
+called initial conditions.
 
@@
 information. The matching must be done for each day of a longer time frame like a month
-to have individuals with possible health states. Then, health statuses evolve until the
-begin of the simulation period without simulating infections by contacts. We also
-correct reported infections for a reporting lag and scale them up to arrive at the true
-number of infections.
+to have individuals with all possible health states. Then, health statuses evolve until
+the beginning of the simulation period without simulating infections by contacts. We also
+correct reported infections for a reporting lag and scale them up \replaced[id=K]{by the
+share of detected cases}{} to arrive at the true number of infections.
 

### 2021-07-14 src/paper/model/seasonality.tex
Join single plots and make single plots smaller plus various small improvements.
Fix typos, make suggestions and add comments to sections B.1 to B.7.
Finish going over appendix B. More type corrections, suggestions and comments.

@@
-\subsection{Seasonality}\comment[id=K]{Philipp suggests to write that we choose upper
-bounds}
+\subsection{Seasonality}
 \label{subsec:seasonality}
@@
 transmission of SARS-CoV-2 as a multiplicative factor on infection probabilities. The
-factor follows a sine curve that reaches its maximum at January first and its minimum
+factor follows a sine curve that reaches its maximum at January 1 and its minimum
 on June 30.
@@
 
-Where $\kappa_k$ is difference in the seasonality factor between peak infectiousness
+Where $\kappa_c$ is difference in the seasonality factor between peak infectiousness
 and lowest infectiousness.
 
-The subscript $k$ is needed because the strength of the seasonality effect differs
+The subscript $c$ is needed because the strength of the seasonality effect differs
 across contact types: Work, household and school contacts are likely to take place
@@
 
-The two seasonality curves are shown in Figure~\ref{fig:seasonality}
+The two seasonality curves are shown in Figure~\ref{fig:seasonality}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/results/figures/data/seasonality}
+    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/seasonality}
     \caption{Seasonality by Type of Contact}
@@
     probability of infection of all encounters. The factor depends on the day and is
-    calculated from a sinus shaped function with its maximum on January 1st. Since
+    calculated from a sinus shaped function with its maximum on January 1. Since
     seasonality can affect the transmission both through physical conditions such as

### 2021-07-14 src/paper/model/code.tex
Finish going over appendix B. More type corrections, suggestions and comments.
Finish most things in section B.

@@
-The code used for this paper is open source and available under the MIT License. It is
-split into two parts
+\subsection{Reproducibility}
+\label{subsec:code}
+
+The source code used for this paper is open source and available under the MIT License.
+It is split into two parts
 \begin{itemize}
-    \item The code for the model can be found at
+    \item The source code for the model can be found at
    \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
@@
    \href{https://sid-dev.readthedocs.io}{https://sid-dev.readthedocs.io}.
-   \item The code for the application to Germany can be found at
+   \item The source code for the application to Germany can be found at
    \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}

### 2021-07-14 src/paper/model/course_of_disease.tex
Finish most things in section B.

@@
 an overview of our disease progression parameters see
-Table~\ref{tab:disease_and_vacc_model}.\comment[id=K]{@Janos: Is this what you wanted
-("Replace this by a link to the relevant table with parameters in the data appendix, once
-we have that table")}
+Table~\ref{tab:disease_and_vacc_model}.

### 2021-07-14 src/paper/model/matching.tex
Finish most things in section B.

@@
 function that takes the value 0 (non-participating) and 1 (participating) and can depend
-on the weekday, date and health state.\comment[id=K]{Actually the full states and params
-enter into the contact functions} This can be used to model things like vacations,
-weekends or symptomatic people who stay home (see section \ref{sec:policies} for
-details).
+on the weekday, date and health states of the entire population. This can be used to
+model things like vacations, weekends or symptomatic people who stay home (see section
+\ref{sec:policies} for details).
 
@@
 to meet people from the same county) but in principle any set of discrete variables can
-be used. This set of variables that influence matching probabilities
-\replaced[id=K]{induce}{introduce} a discrete partition of the population into groups.
+be used. This set of variables that influence matching probabilities induce a discrete
+partition of the population into groups.
 

### 2021-07-14 src/paper/model/policies.tex
Finish most things in section B.

@@
 The most complex policies are typically found in the education sector. Since the
-beginning of \replaced[id=K]{2020}{2021} schools have switched back and fourth between
-full closures, split class approaches with alternating schedules for some or all age
-groups and reopening while maintaining hygiene measures. On top of that there are
-different policies for allowing young students whose parents work full time to attend
-school even on days where they normally would not. For details on how we calibrate these
-policies see Section~\ref{subsec:policies}.\comment[id=K]{Did I write enough on the school policies? (No sources or details so far...)}
+beginning of 2021 schools have switched back and fourth between full closures, split
+class approaches with alternating schedules for some or all age groups and reopening
+while maintaining hygiene measures. On top of that there are different policies for
+allowing young students whose parents work full time to attend school even on days where
+they normally would not. For details on how we calibrate these policies see
+Section~\ref{subsec:policies}.
 
-Importantly, policies can depend on the health states of participating individuals.
-\replaced[id=K]{Thus, it would be possible}{This allows} to quarantine entire school
-classes if one student tested positive \replaced[id=K]{and many other forms of contact
-tracing}{or to implement official or private contact tracing}.
-
-For some policies the exact effect on each contact type is not easy to determine. If this
-refers to a policy that has been active during the estimation period, it is possible to
-estimate such parameters by fitting the model to time series data of infection rates.
-This is only possible if the policy was not active during the whole estimation period and
-thus the infection probabilities can be identified separately. We do this to account for
-hygiene measures at school and in the workplace that have been in effect since November
-2020.\comment[id=K]{I don't understand this paragraph.}
+Importantly, policies can depend on the health states of participating individuals. For
+example children rarely go to school when they have symptoms. It would even be possible
+to quarantine entire school classes if one student tested positive and many other forms
+of contact tracing. For an application of our model showcasing private contact tracing in
+the context of the Christmas holidays see \cite{Gabler2020}.
 
 Not all things that reduce contacts compared to the pre-pandemic level are driven by
-NPIs. Therefore, we also model endogenous contact reductions that can depend on the
-health state of individuals \replaced[id=K]{}{, known risk contacts} or the local
-incidence. Examples in our model are contact reductions for symptomatic individuals or
-those who have a positive PCR or rapid tests \replaced[id=K]{or contact reductions when a
-household member tested positive}. The extent to which contacts are reduced can be
-calibrated from surveys. For an application of our model showcasing private contact
-tracing in the context of the Christmas holidays see \cite{Gabler2020}.
+NPIs. Therefore, we also model endogenous contact reductions that depend on the health
+state of individuals. Other possible factors could include things such as the local
+incidence. The extent to which contacts are reduced can be calibrated from surveys.

### 2021-07-14 src/paper/model/summary.tex
Finish most things in section B.

--- 2021-06-21 src/paper/model/summary.tex
@@
 
-To predict and quantify the effects of a wide variety of fine-grained social distancing
-policies, vaccinations and rapid testing, we propose a different model structure. Our
-model inherits many features from prototypical agent-based simulation models but
-replaces the contacts between moving particles by contacts between individuals who work,
-go to school, live in a household and enjoy leisure activities.
+We use an agent-based simulation model with detailed contact networks. The model
+structure is depicted in Figure~\ref{fig:model_contacts_infections}.
 
-The structure of the model is depicted in Figure~\ref{fig:model_contacts_infections}.
-
-We distinguish between eight types of contact
-models which are all listed in Figure~\ref{fig:model_contacts_infections}: households,
-recurrent and random work contacts, recurrent and random leisure contacts, and nursery,
-preschool, and school contacts.
+We distinguish between eight types of contacts which are all listed in
+Figure~\ref{fig:model_contacts_infections}: households, recurrent and random work
+contacts, recurrent and random leisure contacts, as well as nursery, preschool, and
+school contacts.
 
@@
 Once a person is infected, the disease progresses in a fairly standard way which is
-depicted in Figure~\ref{fig:model_disease_progression}. Asymptomatic cases and cases with mild
-symptoms are infectious for some time and recover eventually. Cases with severe symptoms
-additionally require hospitalization and lead to either recovery or death.
+depicted in Figure~\ref{fig:model_disease_progression}. Asymptomatic cases and cases with
+mild symptoms are infectious for some time and recover eventually. Cases with severe
+symptoms additionally require hospitalization and lead to either recovery or death.
 
 After rapid tests become available, people who work or go to school can receive rapid
-tests there. Moreover, people can decide to make a rapid test if they develop symptoms,
-have many planned contacts or observe cases in their contact network. People who have a
-positive rapid test demand a confirmatory PCR test with a certain probability. Moreover,
-PCR tests can be demanded because of symptoms or randomly.
+tests there. Moreover, people can decide to do a rapid test if they develop symptoms,
+have many planned contacts or have a sick or positively tested household member. People
+who have a positive rapid test demand a confirmatory PCR test with a certain probability.
+Moreover, PCR tests can be demanded because of symptoms or randomly.
 
@@
 
-People who have symptoms, received a positive test, or had a risk contact can reduce
-their number of contacts across all contact types endogenously. The extent to which this
-is done is calibrated from survey data.
+People who have symptoms or received a positive test can reduce their number of contacts
+across all contact types endogenously. The extent to which this is done is calibrated
+from survey data.
 

### 2021-07-14 src/paper/model/testing.tex
Finish most things in section B.

@@
 \label{sub:testing}
-
- \comment[id=K]{Talk here about reactions to tests
-and symptoms with contact reductions. (If not here, find other place where to talk about
-this in detail.)}
 
@@
 official case numbers. Secondly, individuals with undetected or not yet detected
-infections are an important driver of the pandemic: \textcolor{blue}{One of the features
-that makes SARS-CoV-2 so hard to control is that it can be transmitted both by pre- and
-asymptomatic individuals which makes it difficult for individuals to isolate when they
-pose a risk to others. Our model captures this because developing symptoms or receiving a
-positive test result leads many people to reduce their contacts, especially contacts
-outside the own household.}\comment[id=K]{I think these multipliers $\tau$ are not
-mentioned in the data section.}
+infections are an important driver of the pandemic.
 
@@
 
-Therefore, we \replaced[id=K]{build}{aim for} a simpler model of PCR and rapid tests that
-can be calibrated with available data on test demand and availability and -- nevertheless
+Therefore, we build a simpler model of PCR and rapid tests that can be calibrated with
+available data on test demand and availability and -- nevertheless
 -- can produce a share of undetected cases that varies over time and across age groups
@@
 share of detected cases is much higher for the elderly than for children in time periods
-where many tests are done because of symptoms which is in line with the estimates from
-the literature. \comment[id=J]{need source}
+where many tests are done because of symptoms.
 
@@
 tests in the model fits the empirical data well (where empirical data is available).
- \comment[id=K]{Maybe refer to the more detailed \ref{fig:share_rapid_test_last_week2}}
 
@@
 falsely negative. While the specificity of rapid tests is constant, their sensitivity
-strongly depends \replaced[id=K]{on the timing of the rapid test relative to the start of
-infectiousness}{on whether the tested individual is already infectious and if so for how
-long he has been infectious}. Before the onset of infectiousness the sensitivity is very
-low (35\%). On the first day of infectiousness it is much higher (88\%) but still lower
-than during the remaining infectious period (92\%). After infectiousness stops, the
-sensitivity drops to 50 \%.
+strongly depends on the timing of the rapid test relative to the start of infectiousness.
+Before the onset of infectiousness the sensitivity is very low (35\%). On the first day
+of infectiousness it is much higher (88\%) but still lower than during the remaining
+infectious period (92\%). After infectiousness stops, the sensitivity drops to 50 \%.
 
@@
 tests are less reliable, a recent study has found no basis for this
-concern.\comment[id=J]{cite charite study on self administered tests}
+concern \citep{Lindner2020}.
 
@@
 ($\chi_{confirmation}$) of positively tested individuals seek a PCR test
-\citep{Betsch2021}.\comment[id=J]{Need to link to data appendix once the number is there
-and maybe mention the number}\comment[id=K]{Not mentioned in the data section.}
-Importantly, those PCR tests are made in addition to the tests that would have been done
-otherwise. Section~\ref{subsec:results_share_known_cases} discusses the effect of rapid
-tests on the share of detected cases.
+\citep{Betsch2021}. Importantly, those PCR tests are made in addition to the tests that
+would have been done otherwise. Section~\ref{subsec:results_share_known_cases} discusses
+the effect of rapid tests on the share of detected cases.

### 2021-07-14 src/paper/data/course_of_disease.tex
Polishing.

@@
 standard deviation of 8.79 days. To match this we discretize that 41\% of individuals who
-will die from Covid-19 do so after one day in intensive care, 22\% day after twelve days,
+will die from Covid-19 do so after one day in intensive care, 22\% die after twelve days,
 29\% after 20 days and 7\% after 32 days. Again, we rescale this for every age group

### 2021-07-14 src/paper/data/policies.tex

@@
 % educ models
-For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures, emergency care and rotating schooling
-schedules where only one half of students attends every other week or day.
-% For the moment we ignore that lack of childcare leads working parents to stay home.
-An approximation of the share of contacts still taking place with the different school
-regulations can be found in Figure~\ref{fig:school_multiplier}.
+\input{paper/data/school_policies}
+
 % work

### 2021-07-14 src/paper/data/school_policies.tex

For nurseries, preschools and schools we implement vacations as announced by the German
federal states as well as school closures, emergency care and rotating schooling
schedules where only one half of students attends every other week or day.
% For the moment we ignore that lack of childcare leads working parents to stay home.
An approximation of the share of contacts still taking place with the different school
regulations can be found in Figure~\ref{fig:school_multiplier}.

### 2021-07-14 src/paper/data/policies.tex
Polishing.

@@
 masks, ventilation and hand washing) became more strict and more conscientiously observed
-in November, leading to a reduction of 33\% in the number of contacts with the potential
+in November 2020, leading to a reduction of 33\% in the number of contacts with the potential
 to transmit Covid-19 ($\rho_{hygiene}$).
@@
     \floatfoot{\noindent \textit{Note:} The left figure shows the approximate share of
-    school contacts taking place with and without vacations factored in. The dates on
-    which schools have vacation are decided at the state level. Vacations are directly
-    implemented in our model with no school contacts taking place on weekends and during
-    vacations (by federal state) just like the schooling mode (full operation, emergency
-    care, rotating schemes with half class sizes etc.). The figure is, thus, only an
+    school contacts taking place with and without vacations factored in. In contrast
+    to other policies, school policies are not implemented via multipliers but as
+    mechanistic models (e.g. split class approaches in with emergency care). For the
+    above plot we assigned approximate multipliers to those policies. The figure is,
+    thus, only an
     illustration that shows the approximate share of contacts taking place compared to
-    the pre-pandemic level with and without vacations. The difference between the lines
-    show when vacations take place and to what degree. For example all states have fall
-    vacations but the timing varies strongly between states while all states close
-    schools during the Christmas holidays. The right figure shows the work mobility as
+    the pre-pandemic level with and without vacations.
+    The right figure shows the work mobility as
     reported by \cite{Google2021}. We take this as a proxy of the share of workers who

### 2021-07-14 src/paper/data/school_policies.tex
Add young educ policies with sources.

@@
+% educ policies
+
+% summary
 For nurseries, preschools and schools we implement vacations as announced by the German
 federal states as well as school closures, emergency care and rotating schooling
-schedules where only one half of students attends every other week or day.
-% For the moment we ignore that lack of childcare leads working parents to stay home.
-An approximation of the share of contacts still taking place with the different school
-regulations can be found in Figure~\ref{fig:school_multiplier}.
+schedules where only one half of students attends every other week or day. An
+approximation of the share of contacts still taking place with the different school
+regulations can be found in Figure~\ref{fig:school_multiplier}. Note that schooling
+policies are decided on the state level and usually involved rules that depend on local
+incidences. We simplify these rules to one federal policy from the federal incidence and
+the policies of the three most populous federal states (North Rhine-Westphalia, Bavaria
+and Baden-Württemberg).
+
+% young educ
+The policies for preschools and nurseries are similar to the school policies but simpler.
+Until November children attend completely normally, starting in November with increased
+hygiene measures. Nurseries and preschools stay open until mid December. From mid
+December until January 10, nurseries and preschools were nearly completely closed. If
+parents could credibly demonstrate that both parents work in systemically relevant
+professions and no other child care arrangement was possible, nurseries and preschools
+offered so called ``emergency care''. We assume 10\% of children qualified and used
+emergency care during this time. After January 10 when more and more parents returned to
+work the rules for emergency care were made less strict and we assume a third children
+attended to nursery and preschool. This policy stayed in place until February 20.
+Afterwards, preschools and nurseries opened normally until mid March. Then during the
+third wave the restrictions February were put back into place until end of April when
+nurseries and preschools opened again and stayed open \citep{KiTa_BY, KiTa_NRW, KiTa_BW,
+KiTa_BWa, KiTa_BYa}.
+
+% school

### 2021-07-14 src/paper/data/school_policies.tex

@@
 the policies of the three most populous federal states (North Rhine-Westphalia, Bavaria
-and Baden-Württemberg).
+and Baden-Württemberg). The testing policies for schools are described in
+Section~\ref{subsec:rapid_test_demand}
+
+% school
+Until November schools were open normally. Starting in November, we assume that increased
+hygiene measures were taken or that hygiene measures were more carefully observed as the
+incidence increased. Schools stayed open until mid December.
+% mid Dec until mid Jan
+From mid December until January 10 schools closed only offered so called ``emergency
+care'' for young children whose parents could credibly demonstrate that both had to work
+and had no other child care arrangement. Approximately 25\% of primary students and 5\%
+of secondary students attended school as a result.
+% Jan and Feb
+After January 10 when parents had returned to work the rules for emergency care were
+relaxed and we assume a third of primary children and 10\% of secondary students attended
+school as a result. In addition, graduating classes (most adolescents between 16 and 18)
+were allowed to return to school in a rotating scheme where each class was split in two
+groups. Relying on anecdotal evidence we assume that the groups rotate on a daily basis.
+% mid Feb to mid March
+Starting on February 22 primary students were also allowed to return to school on a
+rotating basis until mid March.
+% mid March to Easter
+We summarize the school policy from mid March until Easter as all students being on a
+rotating school schedule. In addition, children that qualify for emergency care also
+attend on days where there group is scheduled to not attend school physically
+% after Easter
+After the Easter break schools were mostly closed again. Part of this was a federal law,
+the so called ``Bundesnotbremse'' \citep{Notbremse2021} that set some binding rules for
+schools based on local incidences that were binding at the time. As a result, most states
+adjusted there schooling policies and during April most schools were closed with
+emergency care arrangements as in the time from January 10 to February 21. As cases fell
+schools were allowed to gradually open. We summarize this as students being on the same
+rotating schedule as from mid March to Easter starting on May 1 \citep{school_BW,
+school_BWa, school_BWb , school_BY, school_BYa, school_NRW, school_NRWa, school_NRWb,
+school_NRWc}.
 
@@
 The policies for preschools and nurseries are similar to the school policies but simpler.
-Until November children attend completely normally, starting in November with increased
-hygiene measures. Nurseries and preschools stay open until mid December. From mid
+Until November children attended completely normally, starting in November with increased
+hygiene measures. Nurseries and preschools stayed open until mid December. From mid
 December until January 10, nurseries and preschools were nearly completely closed. If
@@
 offered so called ``emergency care''. We assume 10\% of children qualified and used
-emergency care during this time. After January 10 when more and more parents returned to
-work the rules for emergency care were made less strict and we assume a third children
-attended to nursery and preschool. This policy stayed in place until February 20.
-Afterwards, preschools and nurseries opened normally until mid March. Then during the
-third wave the restrictions February were put back into place until end of April when
-nurseries and preschools opened again and stayed open \citep{KiTa_BY, KiTa_NRW, KiTa_BW,
-KiTa_BWa, KiTa_BYa}.
-
-% school
+emergency care during this time. After January 10 when parents had returned to work the
+rules for emergency care were relaxed and we assume a third of children attended nursery
+and preschool. This policy stayed in place until February 20. Afterwards, preschools and
+nurseries were open normally (still with increased hygiene measures) until mid March.
+Then during the third wave the restrictions February were put back into place until end
+of April when nurseries and preschools opened again and stayed open for the rest of our
+simulation period with increased hygiene measures in place. \citep{KiTa_BY, KiTa_NRW,
+KiTa_BW, KiTa_BWa, KiTa_BYa, school_BWb}.

### 2021-07-14 src/paper/data/share_known_cases.tex

@@
 \label{subsec:data_share_known_cases}
-
- \comment[id=K]{Stoye called this the ascertainment rate if I remember right. Maybe we should replace this everywhere}
 
@@
 % Christmas and Easter
-Lastly, we model reductions in the share known cases due to the two major holidays in our
+Lastly, we model reductions in the share of detected cases due to the two major holidays in our
 simulation period, Christmas and Easter. During both holidays many laboratories did not
@@
 short and large drops in the share of known cases. The resulting share of detected cases
-in the absence of rapid tests is shown in Figure~\ref{fig:share_known_cases_data}.
+in the absence of rapid tests is shown in Figure~\ref{fig:share_known_cases_data} and
+was estimated to fit the data.
 

### 2021-07-15 src/paper/data/course_of_disease.tex
Add values and refer to sections in tables B.2 until B.4. Add paragraph on immunity and infection probability in the course of disease section.

@@
 disease progression model.
+
+% Susceptibility and Infectiousness
+ \comment[id=K]{@Janos: read this.}
+The first stage of any disease is the infection. As detailed in
+Equation~\ref{eq:infection_probability_formula} the infection
+probability depends on the contact type, the calendar date to determine the seasonality,
+the age group of the susceptible person and the variant the infectious person is
+carrying.
+%
+The base infection probability of each contact type ($\beta_c$) is estimated inside our
+model (Section~\ref{subsec:estimated_params}). How we model and set the seasonality is
+detailed in Section~\ref{subsec:seasonality}. For the susceptibility of each age group
+($\zeta_a$) we take the estimates of \citet{Davies2020} (Extended Data Fig. 4). Lastly,
+we calibrate the infectiousness of the B.1.1.7 variant ($\sigma_{B.1.1.7}$) from
+\citet{Davies2021} to 1.67.
 
@@
 
+% Immunity and Vaccination -------------------------------------------------------------
+
+ \comment[id=K]{@Janos: Read this!}
+
+Individuals can become immune either through infection ($\gamma_{immune}$) or vaccination
+($\gamma_{vacc,\:d}$). As reinfections are estimated to be extremely rare
+\citep{AbuRaddad2020}, we simply set the immunity period to one year with probability
+one, i.e. everyone that has been infected enjoys immunity for the rest of the simulation
+period.
+
+% vaccinations
+The second route to immunity is through vaccination. Germany has mostly relied on the
+Pfizer-BioNTech BNT162b2 and Oxford-AstraZeneca ChAdOx1-S vaccines with smaller shares of
+the Moderna and Johnson\&Johnson vaccines \citep{Impfdashboard}. As \citet{Pritchard2021}
+and \citet{Harris2021} find no difference in the effectiveness between the two most
+common vaccines, we do not distinguish between vaccine types.
+
+Immunity is binary in our model, i.e. individuals become either sterile immune or remain
+susceptible. Thus, we cannot simply use the reported effectiveness but must also include
+the risk of asymptomatic and sub-clinical reinfection among the vaccinated in our
+probability to become immune upon vaccination. This is important as there is by now ample
+evidence that vaccinated individuals can still get infected with SARS-CoV-2 and transmit
+the disease \citep{Petter2021, LevineTiefenbrun2021, Harris2021}.
+
+The reported effectiveness for BNT162b2 is estimated to be 90\% 21 days after the first
+shot \citep{Hunter2021}. The effectiveness does not increase much through the booster
+shot as \citet{Thompson2021} report 90\% (95\% CI = 68\%–97\%) effectiveness against
+PCR-confirmed infections after two doses for mRNA vaccines in general. We therefore do
+not distinguish between the first and the booster shot.
+
+On the other hand, \citet{Lipsitch2021} report a lower bound on transmission for
+the very similar Moderna vaccine of 61\%.
+
+Taking this evidence together we simplify to 35\% of individuals developing immunity 14
+days after their vaccination and 40\% after 21 days. For the remaining 25\% we assume
+that the vaccine does not protect them. %
+This may seem slightly optimistic for how early vaccinations may already have an effect
+when looking at infection rates like \citet{Hunter2021}. However, it can be expected that
+individuals underestimate the time it takes for the vaccine to work and behave less
+carefully after receiving a vaccination. We err on the side of being too optimistic for
+the effectiveness of vaccine and assume a significant protection after 14 days as
+\citet{LevineTiefenbrun2021} report strongly reduced viral loads already 12 to 28 days
+after the first vaccine.
+
 \FloatBarrier

### 2021-07-15 src/paper/data/testing_and_behavioral_response.tex
Add section for pcr testing, case detection and behavioral response. Finishes adding a paragraph or sentence for every parameter in the tables of B.12.

\subsection{PCR Testing and Behavioral Response}
\label{subsec:pcr_testing_and_behavioral_response}

 \comment[id=K]{@Janos: Read this}

After showing how we calibrate rapid test demand and the share of detected cases, this
section describes the remaining parameters for our testing model. Refer to
Section~\ref{sub:testing} for a description of the full testing model.

% share of tests for symptomatics
From the share of detected cases and the number of infections we arrive at the number of
positive PCR tests in our model. A share of these positive tests is given to symptomatic
individuals ($\chi_{symptom,\:t}$). This share is calibrated from German data on case
characteristics \citep{RKICaseCharacteristics} and shown in
Figure~\ref{fig:share_pcr_to_symptomatic}. We keep $\chi_{symptom,\:t}$ constant after
Christmas because the RKI data does not include if a PCR test was done to confirm a
positive rapid test and this share is used for PCR test demand without prior rapid test
indication

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/used_share_of_pcr_tests_going_to_symptomatics}
    \caption{Share of Positive PCR Tests Administered to Symptomatic Individuals}
    \label{fig:share_pcr_to_symptomatic}
    \floatfoot{\noindent \textit{Note:} The share of positive PCR tests that are
    administered to symptomatic individuals ($\chi_{symptom,\:t}$). Since it was not
    recorded for every case if the person was symptomatic or not we take the midpoint
    between the upper and lower bound. We keep the share constant after Christmas because
    the RKI data does not include if a PCR test was done to confirm a positive rapid test
    and this share is used for PCR test demand without prior rapid test indication.}
\end{figure}

% time until test result
PCR tests take one to four days until their result is revealed to the individual
($\gamma_{PCR,\:d}$). Relying on the ARS data \citep{ARS2020} we calculate that 33\% of
individuals receive the test result after one day, 50\% after two days, 10\% after three
days and 7\% after four days.

% share with positive rapid test requesting pcr
To model the demand for PCR tests through rapid tests, we only need the share of
individuals that seek a PCR test to confirm a positive rapid test result
($\chi_{confirmation}$). We calibrate this from \citet{Betsch2021} who asked this
hypothetical question in March of 2021. There 82\% of Germans reported that they would
follow up on a positive rapid test with a PCR test.

% Behavioral Response ------------------------------------------------------------------

Lastly, we need to set the parameters that decide how individuals reduce their contacts
after certain events, $\tau$. We distinguish between the reduction in household contacts
(which are harder to avoid) and non household contacts. There are three events which
trigger potential contact reductions: showing symptoms of CoViD-19, having received a
positive rapid test and having received a positive PCR test. The only survey data we are
aware of on this is \citet{Betsch2021} where 85\% of individuals claimed they
would isolate and restrict their contatcs after a positive rapid test. We assume this
reduction for non household contacts. As household contacts are much more difficult to
avoid, we assume that they are only reduced by 30\%. We assume the same behavior for
individuals that develop symptoms. Lastly, we assume the response to a positive PCR test
to be stronger than in the other two cases and set the reduction of non household
contacts to 95\% and the reduction of household contacts to 50\%.

\FloatBarrier

### 2021-07-15 src/paper/results/validation.tex

@@
 
+Another form of validating our model is to see how well our main results align with
+other studies that evaluate the effect of large scale rapid testing. Of course, this
+has to be taken with a grain of salt as the effect of any rapid testing policy depends
+on the incidence of the disease in the population, how well other testing policies
+such as PCR tests are working, the effect of seasonality and NPIs that are in place.
+
+Nevertheless, it is reassuring that other studies find effect sizes in the same order
+of magnitude.
+
+\citet{Pavelka2021} estimates that a mass testing campaign in Slovakia in October and
+November 2020 where approximately 65 \% of the population took a rapid test within a two
+week period lead to a reduction in case numbers of 70 \% three weeks after the start
+of the intervention. Moreover, they find that this strong reduction in
+cases cannot be explained by isolation of people who tested positive alone but only when
+they took into account that household members of people who tested positive reduced
+their contacts.
+
+While we do not model the exact scenario of \citet{Pavelka2021}, we can roughly compare
+their estimates with our predictions for the difference between the baseline scenario
+and and a scenario without rapid tests. In May about 45\% of people do
+at least one rapid test in every week. Taking into account that there are many repeated
+testers the number of people who do a test within a two week period is probably
+slightly less than the 65 \% from the intervention in Slovakia. On the other hand, we
+have many people who do more than one rapid test in that time which also leads to the
+detection of cases. Our model predicts that the observed incidence with tests is
+approximately 65 \% lower than without tests after three weeks.
+Thus we have an effect size in the same order of magnitude but are slightly
+less optimistic regarding the efficiency of rapid test.
+
+\citet{Berger2021} analyse the effect of twice weekly rapid testing in schools. They
+have two main findings: Firstly, rapid tests reduced the share of undetected cases
+among students by a factor between 2 and 4. Secondly, open schools with mandatory
+testing might leads to the same or even lower numbers of infections than closed
+schools. The estimates are based on infection numbers after the Easter holiday.
+
+Again, we do not directly simulated their scenarios but can roughly compare our results
+to theirs. We estimate a share of undetected cases of approximately 75 \% among school
+age children (5 to 14 years) at the beginning of April, see
+Figure~\ref{fig:share_known_cases_by_age_group}. This drops to slightly less than 40 \%
+at the end of our simulation period. Thus in the long run, mandatory tests at schools
+led to a reduction of the share of undetected cases by a factor of more than 1.8 which
+is just slightly below the factor of 2 to 4 predicted by \citet{Berger2021}.
+
+Similarly we are slightly less optimistic for the effect opening schools with testing
+compared to closing schools. While they predict that opening schools could even be
+beneficial we estimate that it would lead to a slight increase in case numbers
+see Figure~\ref{fig:school_scenarios_detailed}).
+
+% In May 45 % per week are tested
+% Incidence pessimistic = 1750
+% Incidence with mass testing = 750
+
 \FloatBarrier

### 2021-07-15 src/science/paper.tex
Add preliminary science draft.

--- 2021-07-14 src/paper/report.tex
@@
+% Use only LaTeX2e, calling the article.cls class and 12-point type.
+
+\documentclass[12pt]{article}
+
+% Users of the {thebibliography} environment or BibTeX should use the
+% scicite.sty package, downloadable from *Science* at
+% http://www.sciencemag.org/authors/preparing-manuscripts-using-latex
+% This package should properly format in-text
+% reference calls and reference-list numbers.
+
+\usepackage{scicite}
+
+\usepackage{times}
+
+\usepackage{subcaption}
+\usepackage{graphicx}
+\usepackage{floatrow}
+
+% The preamble here sets up a lot of new/revised commands and
+% environments.  It's annoying, but please do *not* try to strip these
+% out into a separate .sty file (which could lead to the loss of some
+% information when we convert the file to other formats).  Instead, keep
+% them in the preamble of your main LaTeX source file.
+
+% The following parameters seem to provide a reasonable page setup.
+
+\topmargin 0.0cm
+\oddsidemargin 0.2cm
+\textwidth 16cm
+\textheight 21cm
+\footskip 1.0cm
+
+%The next command sets up an environment for the abstract to your paper.
+
+\newenvironment{sciabstract}{%
+\begin{quote} \bf}
+{\end{quote}}
+
+% Include your paper's title here
+
+\title{The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs}
+
+% Place the author information here.  Please hand-code the contact
+% information and notecalls; do *not* use \footnote commands.  Let the
+% author contact information appear immediately below the author names
+% as shown.  We would also prefer that you don't change the type-size
+% settings shown here.
+
+\author
+{Janoś Gabler,$^{1, 2}$ Tobias Raabe,$^{3}$ Klara Röhrl,$^{1}$ Hans-Martin von Gaudecker,$^{2,4\ast}$\\
+\\
+\normalsize{$^{1}$Bonn Graduate School of Economics}\\
+\normalsize{$^{2}$IZA Institute of Labor Economics}\\
+\normalsize{$^{3}$Private sector}\\
+\normalsize{$^{4}$Rheinische Friedrich-Wilhelms-Universität Bonn}\\
+\\
+\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  hmgaudecker@uni-bonn.de.}
+}
+
+% Include the date command, but leave its argument blank.
+
+\date{}
+
+%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
+
+\begin{document}
+
+% Double-space the manuscript.
+
+\baselineskip24pt
+
+% Make the title.
+
+\maketitle
+
+% Place your abstract within the special {sciabstract} environment.
+
+\begin{sciabstract}
+    In order to slow the spread of the CoViD-19 pandemic, governments around the world
+    have enacted a wide set of policies limiting the transmission of the disease.
+    Initially, these focused on non-pharmaceutical interventions; more recently,
+    vaccinations and large-scale rapid testing have started to play a major role. The
+    objective of this study is to explain the quantitative effects of these policies on
+    determining the course of the pandemic, allowing for factors like seasonality or
+    virus strains with different transmission profiles. To do so, the study develops an
+    agent-based simulation model, which is estimated using data for the second and the
+    third wave of the CoViD-19 pandemic in Germany. The paper finds that during a period
+    where vaccination rates rose from 5\% to 40\%, seasonality and rapid testing had the
+    largest effect on reducing infection numbers. Frequent large-scale rapid testing
+    should remain part of strategies to contain CoViD-19; it can substitute for many
+    non-pharmaceutical interventions that come at a much larger cost to individuals,
+    society, and the economy.
+\end{sciabstract}
+
 Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many
-months---and often years---away from reaching vaccination levels that would end the
-pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
-of utmost importance to employ an effective mix of strategies for containing the virus.
-The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
-to reduce contacts between individuals. While this has allowed some countries to sustain
-equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
-    theoretical equilibrium at low case numbers which is sustained with
-    test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
-over time. Containment measures have become increasingly diverse and now include rapid
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effects nor the
-influence of seasonal patterns or of more infectious virus strains are well understood
-in quantitative terms.
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination levels that would end the pandemic or even
+protect the most vulnerable \cite{Mathieu2021}. In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers,\footnote{See \cite{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate
+policies.} most have
+seen large fluctuations of infection rates over time. Containment measures have become
+increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
+tracing. Neither these policies' effects nor the influence of seasonal patterns or of
+more infectious virus strains are well understood in quantitative terms.
 
@@
 shows that, aside from seasonality, frequent and large-scale rapid testing caused the
-bulk of this decrease, which is in line with prior predictions \citep{Mina2021}. We
+bulk of this decrease, which is in line with prior predictions \cite{Mina2021}. We
 conclude that it should have a large role for at least as long as vaccinations have not
@@
 
-At the core of our agent-based model \citep[][we review more literature in Supplementary
-    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts
-between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
-the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
- \comment[id=J]{I would replace networks by areas because we have more than four networks}
-Within the household, at work, at school, or in other settings (leisure activities,
-grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
-occur at random. Empirical applications can take the population and household structure
-from census data and the network-specific frequencies of
-contacts from diary data measuring contacts before the pandemic
-\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
+At the core of our agent-based model \cite{Aleta2020,Hinch2020} (we review more
+literature in Supplementary Material~\ref{sec:literature_review}) are physical contacts
+between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact
+between an infectious individual and somebody susceptible to the disease bears the risk
+of transmitting the virus. Contacts occur in up to four networks: Within the household,
+at work, at school, or in other settings (leisure activities, grocery shopping, medical
+appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
+applications can take the population and household structure from census data and the
+network-specific frequencies of contacts from diary data measuring contacts before the
+pandemic \cite{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
 on age and geographical location (see Supplementary
@@
 capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
+masking requirements or air filters \cite{Lessler2021}. Curfews may reduce the number
 of contacts in settings outside of
@@
 probability of infection associated with a contact
-\citep{Cheng2021}.
+\cite{Cheng2021}.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-top-left}
+        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
         \caption{{Model for contacts and infections}}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-top-right}
+        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
         \vskip5ex
@@
 
-        \includegraphics[width=\textwidth]{figures/model-graph-bottom-left}
+        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
         \caption{{Model for rapid tests}}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-bottom-right}
+        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
         \vskip5ex
@@
         Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
-        influence of an agent's contacts to other agents on infections. Demographic characteristics
-        set the baseline number of contacts in different networks ($\eta$). The agents may
-        reduce the number of contacts due to NPIs, showing symptoms, or testing
-        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
-        agent meets an infectious agent; the probability depends on the type of contact
-        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
-        infected, the infection progresses as depicted in
+        influence of an agent's contacts to other agents on infections. Demographic
+        characteristics set the baseline number of contacts in different networks
+        ($\eta$). The agents may reduce the number of contacts due to NPIs, showing
+        symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may occur
+        when a susceptible agent meets an infectious agent; the probability depends on
+        the type of contact ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs
+        ($\rho_{c,\:t}$). If infected, the infection progresses as depicted in
         Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
@@
         of the graph is relevant only for periods where rapid tests are available. All
-        parameters are explained in Appendix\ref{sub:param_tables}.}
+        parameters are explained in Appendix~\ref{sub:param_tables}.}
 \end{figure}
@@
 age. A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has lost
-the timing. Need to explain d-subscript, maybe even better to add stuff like 1-3 days
-back in? Needs to be comprehensible without recurring to appendix} We differentiate
+Figure~\ref{fig:model_disease_progression}. We differentiate
 between an initial period of infection without being infectious or showing symptoms,
 being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
-care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities of
+care, and recovery or death (similar to \cite{Aleta2020}). The probabilities of
 transitioning between these states depend on age; their duration is random and calibrated
@@
 Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \citep{Jones2021}.
+infectiousness is independent of age \cite{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution
-of the pandemic in 2020-2021. New virus strains with different infectiousness
-profiles may appear. Vaccines may become available. During the
-vaccine roll-out, priority may depend on age and occupation; vaccine hesitancy is
-modelled by some individuals refusing vaccination offers. With some probability,
-vaccinated agents become immune and do not transmit the virus \citep{Hunter2021,
-    LevineTiefenbrun2021, Petter2021, Pritchard2021}.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
+appear. Vaccines may become available. During the vaccine roll-out, priority may depend
+on age and occupation; vaccine hesitancy is modelled by some individuals refusing
+vaccination offers. With some probability, vaccinated agents become immune and do not
+transmit the virus \cite{Hunter2021, LevineTiefenbrun2021, Petter2021,
+Pritchard2021}.
 
@@
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}. In our
+matches features of the data (see \cite{McFadden1989} for the general method). In our
 application, we keep the number of free parameters low in order to avoid overfitting.
@@
 recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
-Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
+Oxford Response Stringency Index \cite{Hale2020}, which tracks the tightness of
 non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
         \caption{{Recorded cases: Empirical and simulated}}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/data/stringency2_with_seasonality}
+        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2_with_seasonality}
 
@@
 
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
 
@@
 
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
 
@@
 model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
-\citep{Shapley2016} to decompose the difference in total infections between the scenario
+\cite{Shapley2016} to decompose the difference in total infections between the scenario
 without any of the three factors and our main specification.
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
         \caption{{Recorded cases: 2021 scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
         \caption{{Total cases: 2021 scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_area}
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_area}
         \caption{Decomposition of the difference between the scenario without any of the
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_area}
+        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
         \caption{Decomposition of the difference between the scenario without rapid
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace \citep[this is in line with results in][, which our seasonality measure
-    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
+much lower pace (This is in line with results in \cite{Gavenciak2021} which our
+seasonality measure is based on). Nevertheless, we estimate seasonality to be a
 quantitatively important factor determining the evolution of the pandemic, explaining
@@
 up on a positive test of a household member. The latter is essentially a form of contact
-tracing, which has been shown to be very effective \citep{Contreras2021,
+tracing, which has been shown to be very effective \cite{Contreras2021,
     Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
@@
 
-Two of the most contentious NPIs concern schools and mandates to work from
-home. In many countries, schools switched to remote instruction
-during the first wave, so did Germany. After the summer break, they were operating at
-full capacity with increased hygiene measures, before being closed again from
-mid-December onward. Some states started opening them gradually in late February, but
-operation at normal capacity did not resume until the beginning of June.
-Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
-schools starting after Easter, at which point rapid tests had become widely available. We
-estimate the realized scenario to have essentially the same effect as a situation with
-closed schools. Under fully opened schools with mandatory tests, total infections would
-have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
-broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.\comment[id=HM]{Need
-a back-of the envelope calculation} To use another metric, the effective weekly
-reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
-negative effects school closures have on children and parents \citep{Luijten2021,
-Melegari2021}---and in particular on those with low socio-economic status---these results
-in conjunction with hindsight bias suggest that opening schools combined with a testing
-strategy would have been beneficial. In other situations, and particular when rapid test
-are not available at scale, trade-offs may well be different.
+Two of the most contentious NPIs concern schools and mandates to work from home. In many
+countries, schools switched to remote instruction during the first wave, so did Germany.
+After the summer break, they were operating at full capacity with increased hygiene
+measures, before being closed again from mid-December onward. Some states started
+opening them gradually in late February, but operation at normal capacity did not resume
+until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
+different policies regarding schools starting after Easter, at which point rapid tests
+had become widely available. We estimate the realized scenario to have essentially the
+same effect as a situation with closed schools. Under fully opened schools with
+mandatory tests, total infections would have been 6\% higher; this number rises to 20\%
+without tests. These effect sizes are broadly in line with empirical studies (e.g.,
+\cite{Vlachos2021}). To use another metric, the effective weekly reproduction number
+differs by $0.018$ and $0.052$, respectively. In light of the large negative effects
+school closures have on children and parents \cite{Luijten2021, Melegari2021}---and in
+particular on those with low socio-economic status---these results in conjunction with
+hindsight bias suggest that opening schools combined with a testing strategy would have
+been beneficial. In other situations, and particular when rapid test are not available
+at scale, trade-offs may well be different.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
         \caption{{Effects of different schooling scenarios}}
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
         \caption{{Effects of different work scenarios}}
@@
 at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
- \comment[id=J]{Check this number and add to which time period it refers}
+estimated to have cost around 20~Euros per capita \cite{Wollmershauser2021}; retail
 prices for rapid tests were below one Euro in early June 2021. While we do not
@@
 because disadvantaged groups are less likely to be reached by testing campaigns relying
-on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
-groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
-tests at school and at the workplace will extend more into these groups. The same goes
-for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
-Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
-it arguably the most effective tool to contain the pandemic in the short run.
+on voluntary participation (e.g., \cite{StillmanTonin2021}); at
+the same time, these groups have a higher risk to contract CoViD-19
+\cite{KochInstitut2021a}. Mandatory tests at school and at the workplace will extend
+more into these groups. The same goes for individuals who exhibit a low level of
+compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
+programmes allow a much quicker roll-out, making it arguably the most effective tool to
+contain the pandemic in the short run.
+
+\bibliography{bibliography}
+
+\bibliographystyle{Science}
+
+\section*{Acknowledgments}
+Include acknowledgments of funding, any patents pending, where raw data for the paper are deposited, etc.
+
+%Here you should list the contents of your Supplementary Materials -- below is an example.
+%You should include a list of Supplementary figures, Tables, and any references that appear only in the SM.
+%Note that the reference numbering continues from the main text to the SM.
+% In the example below, Refs. 4-10 were cited only in the SM.
+\section*{Supplementary materials}
+Materials and Methods\\
+Supplementary Text\\
+Figs. S1 to S3\\
+Tables S1 to S4\\
+References \textit{(4-10)}
+
+\end{document}

### 2021-07-15 task_compile_documents.py

@@ -26,6 +26,22 @@ def task_compile_documents():
     pass
 
 
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "science" / "paper.tex", *_DEPENDENCIES])
+@pytask.mark.produces(BLD / "science_report.pdf")
+def task_compile_science_report():
+    pass
+
+
 @pytask.mark.depends_on(BLD / "paper.pdf")
 @pytask.mark.produces(ROOT / "paper.pdf")
 def task_copy_pdf_to_root(depends_on, produces):

### 2021-07-15 src/paper/data/school_policies.tex
Typos etc in validation exercise and some improvements to the references trying to resolve error.

@@
 % educ policies
+
+ \comment[id=K]{@Janos: Read this!}
 
@@
 schools were allowed to gradually open. We summarize this as students being on the same
-rotating schedule as from mid March to Easter starting on May 1 \citep{school_BW,
-school_BWa, school_BWb , school_BY, school_BYa, school_NRW, school_NRWa, school_NRWb,
-school_NRWc}.
+rotating schedule as from mid March to Easter starting on May 1 \citep{schoolBW,
+schoolBWa, schoolBWb, schoolBY, schoolBYa, schoolNRW, schoolNRWa, schoolNRWb,
+schoolNRWc}.
 
@@
 simulation period with increased hygiene measures in place. \citep{KiTa_BY, KiTa_NRW,
-KiTa_BW, KiTa_BWa, KiTa_BYa, school_BWb}.
+KiTa_BW, KiTa_BWa, KiTa_BYa, schoolBWb}.

### 2021-07-15 src/paper/results/validation.tex
Typos etc in validation exercise and some improvements to the references trying to resolve error.

--- 2021-07-14 src/paper/results/validation.tex
@@
 very sensitive to the exact number of vaccinations, the work mobility multiplier
-($\rho_{w,\:attend,\:t}$) or the number of rapid tests (governed by the $\pi$ parameters)
-that are performed -- all of which are things that cannot be known exactly ex-ante.
+($\rho_{w,\:attend,\:t}$) or the number of performed rapid tests (governed by the $\pi$
+parameters) -- all of which are things that cannot be known exactly ex-ante.
 
@@
   numbers that is jointly driven by seasonality, large scale rapid tests and
-  vaccinations. For legibility reasons, all lines are rolling 7-day averages.}
+  vaccinations. For legibility reasons, all lines are rolling 7-day averages. Each line
+  is the average over 30 simulation runs.}
 \end{figure}
 
+Another form of validating our model is to see how well our main results align with
+other studies that evaluate the effect of large scale rapid testing. Of course, this
+has to be taken with a grain of salt as the effect of any rapid testing policy depends
+on the incidence of the disease in the population, how well other testing policies
+such as PCR tests are working, the effect of seasonality and NPIs that are in place.
+
+Nevertheless, it is reassuring that other studies find effect sizes in the same order
+of magnitude.
+
+\citet{Pavelka2021} estimate that a mass testing campaign in Slovakia in October and
+November 2020 where approximately 65 \% of the population took a rapid test within a two
+week period lead to a reduction in case numbers of 70 \% three weeks after the start
+of the intervention. Moreover, they find that this strong reduction in
+cases cannot be explained by isolation of people who tested positive alone but only when
+they took into account that household members of people who tested positive reduced
+their contacts.
+
+While we do not model the exact scenario of \citet{Pavelka2021}, we can roughly compare
+their estimates with our predictions for the difference between the baseline scenario
+and and a scenario without rapid tests. In May about 45\% of people do
+at least one rapid test in every week. Taking into account that there are many repeated
+testers the number of people who do a test within a two week period is probably
+slightly less than the 65\% from the intervention in Slovakia. On the other hand, we
+have many people who do more than one rapid test in that time which also leads to the
+detection of cases. Our model predicts that the observed incidence with tests is
+approximately 65\% lower than without tests after three weeks.
+Thus we have an effect size in the same order of magnitude but are slightly
+less optimistic regarding the efficiency of rapid tests.
+
+\citet{Berger2021} analyse the effect of twice weekly rapid testing in schools. They
+have two main findings: Firstly, rapid tests reduced the share of undetected cases
+among students by a factor between two and four. Secondly, open schools with mandatory
+testing might lead to the same or even lower numbers of infections than closed
+schools. The estimates are based on infection numbers after the Easter holiday.
+
+Again, we do not directly simulate their scenarios but can roughly compare our results
+to theirs. We estimate a share of undetected cases of approximately 75\% among school
+age children (five to 14 years) at the beginning of April, see
+Figure~\ref{fig:share_known_cases_by_age_group}. This drops to slightly less than 40\%
+at the end of our simulation period. Thus in the long run, mandatory tests at schools
+led to a reduction of the share of undetected cases by a factor of more than 1.8 which
+is just slightly below the factor of two to four predicted by \citet{Berger2021}.
+
+Similarly we are slightly less optimistic for the effect of opening schools with testing
+compared to closing schools. While they predict that opening schools could even be
+beneficial we estimate that it would lead to a slight increase in case numbers see
+Figure~\ref{fig:school_scenarios_detailed}).
+
+% In May 45 % per week are tested
+% Incidence pessimistic = 1750
+% Incidence with mass testing = 750
+
 \FloatBarrier

### 2021-07-15 src/paper/report.tex
Replace some working papers with published versions. 50 EUR costs per day and capita and remove no longer needed comments.

--- 2021-07-14 src/paper/report.tex
@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many
-months---and often years---away from reaching vaccination levels that would end the
-pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
-of utmost importance to employ an effective mix of strategies for containing the virus.
-The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
-to reduce contacts between individuals. While this has allowed some countries to sustain
-equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
-    theoretical equilibrium at low case numbers which is sustained with
-    test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
-over time. Containment measures have become increasingly diverse and now include rapid
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effects nor the
-influence of seasonal patterns or of more infectious virus strains are well understood
-in quantitative terms.
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination levels that would end the pandemic or even
+protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
+most have seen large fluctuations of infection rates over time. Containment measures have
+become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
+tracing. Neither these policies' effects nor the influence of seasonal patterns or of
+more infectious virus strains are well understood in quantitative terms.
 
@@
 
-The model includes several other features, which are crucial to describe the evolution
-of the pandemic in 2020-2021. New virus strains with different infectiousness
-profiles may appear. Vaccines may become available. During the
-vaccine roll-out, priority may depend on age and occupation; vaccine hesitancy is
-modelled by some individuals refusing vaccination offers. With some probability,
-vaccinated agents become immune and do not transmit the virus \citep{Hunter2021,
-    LevineTiefenbrun2021, Petter2021, Pritchard2021}.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
+appear. Vaccines may become available. During the vaccine roll-out, priority may depend
+on age and occupation; vaccine hesitancy is modelled by some individuals refusing
+vaccination offers. With some probability, vaccinated agents become immune and do not
+transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.
 
@@
 much lower pace \citep[this is in line with results in][, which our seasonality measure
-    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
-quantitatively important factor determining the evolution of the pandemic, explaining
-most of the early changes and 43\% of the cumulative difference by the end of May.
+is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a quantitatively
+important factor determining the evolution of the pandemic, explaining most of the early
+changes and 43\% of the cumulative difference by the end of May.
 
@@
 at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
+estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
  \comment[id=J]{Check this number and add to which time period it refers}
@@
 on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
-groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
-tests at school and at the workplace will extend more into these groups. The same goes
-for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
+groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests
+at school and at the workplace will extend more into these groups. The same goes for
+individuals who exhibit a low level of compliance with CoViD-19-related regulations.
 Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making

### 2021-07-15 src/paper/model/code.tex

@@
 \end{itemize}
+
+We are grateful to the authors and contributors of the following software packages upon
+which our software is built: conda \cite{Anaconda2016}, conda-forge
+\cite{CondaForge2015} dask \cite{Rocklin2015}, estimagic \cite{Gabler2020a}, holoviews
+\cite{Stevens2015}, matplotlib \cite{Hunter2007}, numba \cite{Lam2015}, numpy
+\cite{Harris2020}, pandas \cite{Reback2020,McKinney2010}, pytask \cite{Raabe2020},
+Python \cite{VanRossum1995}, scipy \cite{Virtanen2020}.

### 2021-07-15 src/paper/data/course_of_disease.tex
Hopefully final polishing.

@@
 % Susceptibility and Infectiousness
- \comment[id=K]{@Janos: read this.}
 The first stage of any disease is the infection. As detailed in
@@
 The base infection probability of each contact type ($\beta_c$) is estimated inside our
-model (Section~\ref{subsec:estimated_params}). How we model and set the seasonality is
-detailed in Section~\ref{subsec:seasonality}. For the susceptibility of each age group
-($\zeta_a$) we take the estimates of \citet{Davies2020} (Extended Data Fig. 4). Lastly,
-we calibrate the infectiousness of the B.1.1.7 variant ($\sigma_{B.1.1.7}$) from
+model (Section~\ref{subsec:estimated_params}). How we model and calibrate the seasonality
+effect is detailed in Section~\ref{subsec:seasonality}. For the susceptibility of each
+age group ($\zeta_a$) we take the estimates of \citet{Davies2020} (Extended Data Fig. 4).
+Lastly, we calibrate the infectiousness of the B.1.1.7 variant ($\sigma_{B.1.1.7}$) from
 \citet{Davies2021} to 1.67.
@@
 
- \comment[id=K]{@Janos: Read this!}
-
 Individuals can become immune either through infection ($\gamma_{immune}$) or vaccination
-($\gamma_{vacc,\:d}$). As reinfections are estimated to be extremely rare
-\citep{AbuRaddad2020}, we simply set the immunity period to one year with probability
-one, i.e. everyone that has been infected enjoys immunity for the rest of the simulation
-period.
+($\gamma_{vacc,\:d}$). As reinfections are very rare \citep{AbuRaddad2020}, we
+set the immunity period to one year with probability one, i.e. everyone that has
+been infected enjoys immunity for the rest of the simulation period.
 
@@
 
-Immunity is binary in our model, i.e. individuals become either sterile immune or remain
-susceptible. Thus, we cannot simply use the reported effectiveness but must also include
-the risk of asymptomatic and sub-clinical reinfection among the vaccinated in our
-probability to become immune upon vaccination. This is important as there is by now ample
-evidence that vaccinated individuals can still get infected with SARS-CoV-2 and transmit
-the disease \citep{Petter2021, LevineTiefenbrun2021, Harris2021}.
+Immunity is binary in our model, i.e. individuals achieve either sterile immunity or
+remain susceptible. Thus, we cannot simply use the reported effectiveness but must also
+include the risk of asymptomatic and sub-clinical reinfection among the vaccinated in our
+probability to become immune upon vaccination. This is important as there is ample
+evidence by now that vaccinated individuals can still get infected with SARS-CoV-2 and
+transmit the disease \citep{Petter2021, LevineTiefenbrun2021, Harris2021}.
 
@@
 
-Taking this evidence together we simplify to 35\% of individuals developing immunity 14
-days after their vaccination and 40\% after 21 days. For the remaining 25\% we assume
-that the vaccine does not protect them. %
-This may seem slightly optimistic for how early vaccinations may already have an effect
-when looking at infection rates like \citet{Hunter2021}. However, it can be expected that
-individuals underestimate the time it takes for the vaccine to work and behave less
-carefully after receiving a vaccination. We err on the side of being too optimistic for
-the effectiveness of vaccine and assume a significant protection after 14 days as
-\citet{LevineTiefenbrun2021} report strongly reduced viral loads already 12 to 28 days
-after the first vaccine.
+To strike a middle ground we assume that 75\% of individuals achieve sterile immunity
+after vaccination. This is split into 35\% reaching immunity after 14 days after the
+first shot and 40\% reaching immunity after 21 days.
 

### 2021-07-15 src/paper/data/school_policies.tex
Hopefully final polishing.

--- 2021-07-14 src/paper/data/school_policies.tex
@@
 For nurseries, preschools and schools we implement vacations as announced by the German
-federal states as well as school closures, emergency care and rotating schooling
-schedules where only one half of students attends every other week or day. An
-approximation of the share of contacts still taking place with the different school
-regulations can be found in Figure~\ref{fig:school_multiplier}. Note that schooling
-policies are decided on the state level and usually involved rules that depend on local
-incidences. We simplify these rules to one federal policy from the federal incidence and
-the policies of the three most populous federal states (North Rhine-Westphalia, Bavaria
-and Baden-Württemberg).
+federal states as well as school closures, emergency care and rotating schedules where
+only one half of students attends every other week or day. An approximation of the share
+of contacts still taking place with the different school regulations can be found in
+Figure~\ref{fig:school_multiplier}. Note that schooling policies differ between states
+and usually involve rules that depend on local incidences. We simplify these rules to one
+federal policy from the federal incidence and the policies of the three most populous
+federal states (North Rhine-Westphalia, Bavaria and Baden-Württemberg). The testing
+policies for schools are described in Section~\ref{subsec:rapid_test_demand}.
+
+% school
+Until November schools were open normally. Starting in November, we assume that increased
+hygiene measures were taken. Schools stayed open until mid December.
+% mid Dec until mid Jan
+From mid December until January 10 schools closed and only offered so called ``emergency
+care'' for young children whose parents could credibly demonstrate that both had to work
+and had no other child care arrangement. Approximately 25\% of primary school children and 5\%
+of secondary students attended school as a result.
+% Jan and Feb
+After January 10 when parents had returned to work the rules for emergency care were
+relaxed and approximately a third of primary school children and 10\% of secondary students
+attended school as a result. In addition, graduating classes (most adolescents between 16
+and 18) were allowed to return to school in a rotating scheme where each class was split
+in two groups. Relying on anecdotal evidence we assume that the groups rotate on a daily
+basis.% mid Feb to mid March
+Starting on February 22 primary school children were also allowed to return to school on
+a rotating basis until mid March.
+% mid March to Easter
+We summarize the school policy from mid March until Easter as all students being on a
+rotating school schedule. In addition, children that qualify for emergency care also
+attend on days where their group is scheduled to not attend school physically.
+% after Easter
+After the Easter break schools were mostly closed again. Part of this was a federal law,
+the so called ``Bundesnotbremse'' \citep{Notbremse2021} that set rules for schools based
+on local incidences that were binding at the time. As a result, most states adjusted
+their schooling policies and during April most schools were closed with emergency care
+arrangements as in the time from January 10 to February 21. As cases fell schools were
+allowed to gradually open. We summarize this as students being on the same rotating
+schedule as from mid March to Easter starting on May 1 \citep{schoolBW, schoolBWa,
+schoolBWb, schoolBY, schoolBYa, schoolNRW, schoolNRWa, schoolNRWb, schoolNRWc}.
 
@@
 offered so called ``emergency care''. We assume 10\% of children qualified and used
-emergency care during this time. After January 10 when more and more parents returned to
-work the rules for emergency care were made less strict and we assume a third children
-attended to nursery and preschool. This policy stayed in place until February 20.
-Afterwards, preschools and nurseries opened normally until mid March. Then during the
-third wave the restrictions February were put back into place until end of April when
-nurseries and preschools opened again and stayed open \citep{KiTa_BY, KiTa_NRW, KiTa_BW,
-KiTa_BWa, KiTa_BYa}.
-
-% school
+emergency care during this time. After January 10 when parents had returned to work the
+rules for emergency care were relaxed and we assume a third of children attended nursery
+and preschool. This policy stayed in place until February 20. Afterwards, preschools and
+nurseries were open normally (maintaining increased hygiene measures) until mid March.
+Then during the third wave the restrictions February were put back into place until end
+of April when nurseries and preschools opened again and stayed open for the rest of our
+simulation period - maintaining increased hygiene measures. \citep{KiTa_BY, KiTa_NRW,
+KiTa_BW, KiTa_BWa, KiTa_BYa, schoolBWb}.

### 2021-07-15 src/paper/report.tex
Hopefully final polishing.

@@
 At the core of our agent-based model \citep[][we review more literature in Supplementary
-    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts
-between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
-the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
- \comment[id=J]{I would replace networks by areas because we have more than four networks}
-Within the household, at work, at school, or in other settings (leisure activities,
-grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
-occur at random. Empirical applications can take the population and household structure
-from census data and the network-specific frequencies of
-contacts from diary data measuring contacts before the pandemic
-\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
-on age and geographical location (see Supplementary
+Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
+heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
+an infectious individual and somebody susceptible to the disease bears the risk of
+transmitting the virus. Contacts occur in up to four networks: Within the household, at
+work, at school, or in other settings (leisure activities, grocery shopping, medical
+appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
+applications can take the population and household structure from census data and the
+network-specific frequencies of contacts from diary data measuring contacts before the
+pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
+depend on age and geographical location (see Supplementary
 Material~\ref{subsec:assortativity}).
@@
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
-age. A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has lost
-the timing. Need to explain d-subscript, maybe even better to add stuff like 1-3 days
-back in? Needs to be comprehensible without recurring to appendix} We differentiate
-between an initial period of infection without being infectious or showing symptoms,
-being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
-care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities of
-transitioning between these states depend on age; their duration is random and calibrated
-to medical literature (for a detailed description see Supplementary
+In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
+possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
+differentiate between an initial period of infection without being infectious or showing
+symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
+intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
+of transitioning between these states depend on age; their duration is random and
+calibrated to medical literature (for a detailed description see Supplementary
 Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
@@
 
-Two of the most contentious NPIs concern schools and mandates to work from
-home. In many countries, schools switched to remote instruction
-during the first wave, so did Germany. After the summer break, they were operating at
-full capacity with increased hygiene measures, before being closed again from
-mid-December onward. Some states started opening them gradually in late February, but
-operation at normal capacity did not resume until the beginning of June.
-Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
-schools starting after Easter, at which point rapid tests had become widely available. We
-estimate the realized scenario to have essentially the same effect as a situation with
-closed schools. Under fully opened schools with mandatory tests, total infections would
-have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
-broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.\comment[id=HM]{Need
-a back-of the envelope calculation} To use another metric, the effective weekly
-reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
-negative effects school closures have on children and parents \citep{Luijten2021,
+Two of the most contentious NPIs concern schools and mandates to work from home. In many
+countries, schools switched to remote instruction during the first wave, so did Germany.
+After the summer break, they were operating at full capacity with increased hygiene
+measures, before being closed again from mid-December onward. Some states started opening
+them gradually in late February, but operation at normal capacity did not resume until
+the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
+policies regarding schools starting after Easter, at which point rapid tests had become
+widely available. We estimate the realized scenario to have essentially the same effect
+as a situation with closed schools. Under fully opened schools with mandatory tests,
+total infections would have been 6\% higher; this number rises to 20\% without tests.
+These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
+Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
+large negative effects school closures have on children and parents \citep{Luijten2021,
 Melegari2021}---and in particular on those with low socio-economic status---these results
@@
 estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
- \comment[id=J]{Check this number and add to which time period it refers}
-prices for rapid tests were below one Euro in early June 2021. While we do not
-distinguish between self-administered rapid tests and point of care rapid tests, the
-former are likely to play a larger role for indication-driven testing. Widespread
-availability at low prices seems important. However, they rely on purely voluntary
-participation in a non-public setting. The benefit of point-of-care rapid tests as a
-precondition to participate in leisure activities as well as mandatory tests at the
-workplace or at school come from screening the entire population. This is important
-because disadvantaged groups are less likely to be reached by testing campaigns relying
-on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
-groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests
-at school and at the workplace will extend more into these groups. The same goes for
-individuals who exhibit a low level of compliance with CoViD-19-related regulations.
+prices for rapid tests were below one Euro in early June 2021 and below five Euros for
+firms. While we do not distinguish between self-administered rapid tests and point of
+care rapid tests, the former are likely to play a larger role for indication-driven
+testing. Widespread availability at low prices seems important. However, they rely on
+purely voluntary participation in a non-public setting. The benefit of point-of-care
+rapid tests as a precondition to participate in leisure activities as well as mandatory
+tests at the workplace or at school come from screening the entire population. This is
+important because disadvantaged groups are less likely to be reached by testing campaigns
+relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time,
+these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
+tests at school and at the workplace will extend more into these groups. The same goes
+for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
 Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making

### 2021-07-15 src/paper/report.tex
2930

@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination levels that would end the pandemic or even
-protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
+containment in the medium term. However, most countries find themselves months or years
+away from reaching vaccination levels that would end the pandemic \citep{Mathieu2021}.
+In the meantime, it is of utmost importance to employ an effective mix of strategies for
+containing the virus. The most frequent initial response was a set of non-pharmaceutical
+interventions (NPIs) to reduce contacts between individuals. While this has allowed some
+countries to sustain equilibria with very low infection numbers \citep{Contreras2021}
 most have seen large fluctuations of infection rates over time. Containment measures have
@@
 capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
-of contacts in settings outside of
-work and school. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact
+masking requirements or air filters \citep{Lessler2021}. In any setting, measures like
+masking requirements would reduce the probability of infection
 \citep{Cheng2021}.
@@
         A~description of the model can be found in Supplementary
-        Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
-        influence of an agent's contacts to other agents on infections. Demographic
-        characteristics set the baseline number of contacts in different networks
-        ($\eta$). The agents may reduce the number of contacts due to NPIs, showing
-        symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may occur
-        when a susceptible agent meets an infectious agent; the probability depends on
-        the type of contact ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs
+        Material~\ref{sec:supplementary_text}. Figure~\ref{fig:model_contacts_infections}
+        shows the influence of an agent's contacts to other agents on infections.
+        Demographic characteristics set the baseline number of contacts in different
+        networks ($\eta$). The agents may reduce the number of contacts due to NPIs,
+        showing symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may
+        occur when a susceptible agent meets an infectious agent; the probability depends
+        on the type of contact ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs
         ($\rho_{c,\:t}$). If infected, the infection progresses as depicted in
-        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
-        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
-        trigger a test only for a fraction of individuals depending on an individual
-        compliance parameter; the thresholds for triggering test demand differ across
-        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
+        Figure~\ref{fig:model_disease_progression}. If rapid tests are available, agents'
+        demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons trigger a
+        test only for a fraction of individuals depending on an individual compliance
+        parameter; the thresholds for triggering test demand differ across reasons and
+        they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
         Figure~\ref{fig:model_official_cases} shows the model of translating all
@@
         positive PCR tests triggered by symptoms ($\chi_{symptom}$), and the false
-        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
-        of the graph is relevant only for periods where rapid tests are available. All
+        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part of
+        the graph is relevant only for periods where rapid tests are available. All
         parameters are explained in Appendix~\ref{sub:param_tables}.}
@@
 of transitioning between these states depend on age; their duration is random and
-calibrated to medical literature (for a detailed description see Supplementary
+calibrated to medical literature (see Supplementary
 Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
@@
 at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests
-yield immediate results. Specificity and
-sensitivity of these tests is set
-according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the onset of infectiousness. After a phase-in period, all
-tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by
+contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
+these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
+sensitivity depends on the timing of the test relative to the onset of infectiousness.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require staff and students to be tested regularly. Rapid tests may be offered by
 employers to on-site workers. Individuals may demand tests for private reasons, which
@@
 compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-    contacts; this is why tests impact the actual contacts in
-    Figure~\ref{fig:model_description}.}
+tests. Positive tests and symptoms lead most individuals to reduce their contacts.
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~\ref{sec:data_and_parameters} for a
+parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
     complete description.} Contact diaries yield pre-pandemic distributions of contacts for
@@
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}. In our
-application, we keep the number of free parameters low in order to avoid overfitting.
-The data features to be matched include official case numbers for each age group and
-region, deaths, and the share of the B.1.1.7 strain.
+matches features of the data \citep[see][for the general method]{McFadden1989}.
 
@@
 varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is 80\% higher on average for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
+Figure~\ref{fig:share_known_cases_by_age_group}).
+% Our estimates suggest that---in the
+% absence of rapid testing---the detection rate is 80\% higher on average for individuals
+% above age 80 compared to school age children. Once rapid test become available,
+% confirmation of a positive result is another reason leading to positive PCR tests.
 
-The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
-covering the period mid-September 2020 to the end of May 2021.
+The model is applied to the second and third waves of infections in Germany,
+covering the period September 2020 to May 2021.
 Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
@@
     \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
-        Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
+        Material~\ref{sec:materials_and_methods}. Age- and region-specific analogues to
         Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
@@
 had received at least one dose of a vaccine. In late 2020, rapid tests started to
-replace regular PCR tests for staff in many medical and nursing facilities. These had to
-be administered by medical doctors or in pharmacies. At-home tests approved by
+replace regular PCR tests for staff in many medical and nursing facilities.
+At-home tests approved by
 authorities became available in mid-March. Rapid test centers were opened, and one test
@@
 We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until the end of May 2021. See
+use the model to predict recorded infection rates until May 2021. See
 Supplementary Materials~\ref{subsec:synthetic_population} and
@@
 lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations (roughly the last two weeks of October, two weeks each
-around Christmas and Easter, and some days in late May). Second, the most stringent
+seasonality and school vacations. Second, the most stringent
 regulations coincide with the period of decreasing infection rates between late December
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
-        For legibility reasons, all lines are rolling 7-day averages.}
+        are explained in Appendix~\ref{sub:shapley_value}.
+        All lines are rolling 7-day averages.}
 \end{figure}
@@
 Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
-Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections with only one channel active,
-the final level is always the highest in case of the vaccination campaign (orange
-lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
+Even thereafter, the effect of the vaccination campaign is surprisingly small.
+The Shapley value decomposition shows that vaccinations contribute 16\% to the
 cumulative difference between scenarios. Reasons for the low share are the slow
@@
 vaccination, the 20\% mark was reached on April 19th---and the focus on older
-individuals. These groups contribute less to the spread of the disease than others due
-to a lower number of contacts. By the end of our study period, when first-dose
+individuals. By the end of our study period, when first-dose
 vaccination rates reached 43\% of the population, the numbers of new cases would have
-started to decline. It is important to note that the initial focus of the campaign was
-to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
-lower during the third wave when compared to the second (4.4\% between October and
-February and 1.4\% between March and the end of May).
+started to decline. However, the vaccination campaign reached its goal of protecting
+risk groups.
+
+%It is important to note that the initial focus of the campaign was
+%to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
+%lower during the third wave when compared to the second (4.4\% between October and
+%February and 1.4\% between March and the end of May).
 
@@
 much lower pace \citep[this is in line with results in][, which our seasonality measure
-is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a quantitatively
-important factor determining the evolution of the pandemic, explaining most of the early
-changes and 43\% of the cumulative difference by the end of May.
+is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be an
+important factor, explaining most of the early changes and 43\% of the cumulative
+difference by the end of May.
 
@@
 tracing, which has been shown to be very effective \citep{Contreras2021,
-    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
+    Fetzer2021,Kretzschmar2020}. A deeper analysis in Supplementary
 Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not have been nearly as effective.
-
-Two of the most contentious NPIs concern schools and mandates to work from home. In many
-countries, schools switched to remote instruction during the first wave, so did Germany.
-After the summer break, they were operating at full capacity with increased hygiene
-measures, before being closed again from mid-December onward. Some states started opening
-them gradually in late February, but operation at normal capacity did not resume until
-the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
-policies regarding schools starting after Easter, at which point rapid tests had become
-widely available. We estimate the realized scenario to have essentially the same effect
-as a situation with closed schools. Under fully opened schools with mandatory tests,
-total infections would have been 6\% higher; this number rises to 20\% without tests.
-These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
-Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
-large negative effects school closures have on children and parents \citep{Luijten2021,
-Melegari2021}---and in particular on those with low socio-economic status---these results
-in conjunction with hindsight bias suggest that opening schools combined with a testing
-strategy would have been beneficial. In other situations, and particular when rapid test
-are not available at scale, trade-offs may well be different.
-
-\begin{figure}[!tp]
-    \centering
-
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-        \caption{{Effects of different schooling scenarios}}
-        \label{fig:school_scenarios}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
-        \caption{{Effects of different work scenarios}}
-        \label{fig:workplace_scenarios}
-    \end{subfigure}
-    \vskip3ex
-
-    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
-    \label{fig:school_workplace_scenarios}
-
-    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
-        scenario; they are the same as in
-        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
-        because there were no capacity constraints for rapid tests afterwards.
-        For legibility reasons, all lines are rolling 7-day averages.}
-
-\end{figure}
-
-Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
-receiving tests, testing at the workplace has larger effects than mandating employees to
-work from home. Whether the share of workers working at the usual workplace is reduced
-or increased by ten percent changes infection rates by 2.5\% or less in either
-direction. Making testing mandatory twice a week---assuming independent compliance by
-employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
-rapid tests offers by employers to the level of March would have increased infections by
-13\%.
+administered randomly in the population would not have been nearly as effective and that
+mandatory tests at schools almost offset the increased infection risk of opening
+schools (see Figure~\ref{fig:school_scenarios}).
 
@@
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
+at a fraction of the cost. A day of the fairly strict lockdown in early 2021 is
 estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail

### 2021-07-15 task_compile_documents.py

@@ -59,7 +59,7 @@ def task_extract_supplementary_material(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "15-78",
+            "15-67",
             "--",
             produces.as_posix(),
         ]

### 2021-07-15 src/paper/report.tex
2762

@@
 heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
-an infectious individual and somebody susceptible to the disease bears the risk of
-transmitting the virus. Contacts occur in up to four networks: Within the household, at
+infectious and susceptible individuals bears the risk of transmitting the virus.
+Contacts occur in up to four networks: Within the household, at
 work, at school, or in other settings (leisure activities, grocery shopping, medical
@@
 upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from
-home will reduce the number of work contacts to zero for a fraction of the working
-population. Schools and daycare can be closed entirely, operate at reduced
-capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. In any setting, measures like
-masking requirements would reduce the probability of infection
-\citep{Cheng2021}.
+contact types. See Section~\ref{sec:policies} for details on how we model NPIs.
+ In any setting, mask mandates or air filtes would reduce the probability
+of infection \citep{Lessler2021, Cheng2021}.
 
@@
 symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
-of transitioning between these states depend on age; their duration is random and
-calibrated to medical literature (see Supplementary
-Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \citep{Jones2021}.
+intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The calibration
+of the age dependent probabilities of transitioning between these states  is described
+in Supplementary Material~\ref{sub:data_course_of_disease}). Conditional on the type of
+contact, infectiousness is independent of age \citep{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
-appear. Vaccines may become available. During the vaccine roll-out, priority may depend
-on age and occupation; vaccine hesitancy is modelled by some individuals refusing
-vaccination offers. With some probability, vaccinated agents become immune and do not
-transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.
+The model includes several other crucial features. New virus strains with different
+infectiousness profiles may appear. Vaccines may become available. During the vaccine
+roll-out, priority may depend on age and occupation; vaccine hesitancy is modelled by
+some individuals refusing vaccination offers. With some probability, vaccinated agents
+become immune and do not transmit the virus \citep{Hunter2021, LevineTiefenbrun2021,
+Petter2021, Pritchard2021}.
 
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
-individual is infected or not; there is no uncertainty to the result. PCR tests require
-at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
+certainty whether an individual is infected or not; They require at least one day to be
+processed and are scarce. In contrast, rapid antigen tests yield immediate results.
+Specificity and sensitivity of
 these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
@@
 parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
-different contact types and their assortativity by age group. Mobility data is used to
-model the evolution of work contacts. School and daycare policies can be incorporated
-directly from official directives. Administrative records on the number of tests,
-vaccinations by age and region, and the prevalence of virus strains are generally
-available. Surveys may ask about test offers, propensities to take them up, and past
-tests. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, these include infection probabilities
+    complete description.}
+Examples are contact diaries, administrative records on case numbers, virus straints,
+tests and vaccines as well as surveys on private rapid test demand.
+Other studies' estimates of the seasonality of infections can be incorporated
+directly. The remaining parameters---most notably, the infection probabilities
 by contact network and the effects of some NPIs, see Supplementary
@@
 vaccination rates reached 43\% of the population, the numbers of new cases would have
-started to decline. However, the vaccination campaign reached its goal of protecting
-risk groups.
+started to decline.
 
@@
 are turned on. The effect on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
-population performed weekly rapid tests in March on average, new infections on April~1
-would have been reduced by 53\% relative to the scenario without vaccinations, rapid
-tests, or seasonality.
+Figure~\ref{fig:2021_scenarios_newly_infected}.
+
+% Despite the fact that only 10\% of the
+% population performed weekly rapid tests in March on average, new infections on April~1
+% would have been reduced by 53\% relative to the scenario without vaccinations, rapid
+% tests, or seasonality.
 
@@
 elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. The latter is essentially a form of contact
+up on a positive test of a household member. This is essentially a form of contact
 tracing, which has been shown to be very effective \citep{Contreras2021,
-    Fetzer2021,Kretzschmar2020}. A deeper analysis in Supplementary
+    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
 Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests

### 2021-07-15 src/paper/model/code.tex

@@
 \cite{Harris2020}, pandas \cite{Reback2020,McKinney2010}, pytask \cite{Raabe2020},
-Python \cite{VanRossum1995}, scipy \cite{Virtanen2020}.
+Python \cite{VanRossum1995}, scipy \cite{Virtanen2020}, and seaborn \cite{Waskom2021}.

### 2021-07-15 src/paper/report.tex
shorten.

@@
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
-possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
-differentiate between an initial period of infection without being infectious or showing
-symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The calibration
-of the age dependent probabilities of transitioning between these states  is described
-in Supplementary Material~\ref{sub:data_course_of_disease}). Conditional on the type of
-contact, infectiousness is independent of age \citep{Jones2021}.
+In the model, a possible infection progresses as shown in
+Figure~\ref{fig:model_disease_progression} (similar to \cite{Aleta2020}, more in
+Supplementary Material~\ref{sub:data_course_of_disease}). Whereas susceptibility to the
+viruses differs by age, the infection probabilities are independent of age conditioned
+on the type of contact \cite{Jones2021}.
 

### 2021-07-15 src/paper/report.tex
Done.

@@
-Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
-on many dimensions. The development of highly effective vaccines holds the promise of
+The development of highly effective vaccines holds the promise of
 containment in the medium term. However, most countries find themselves months or years
 away from reaching vaccination levels that would end the pandemic \citep{Mathieu2021}.
-In the meantime, it is of utmost importance to employ an effective mix of strategies for
-containing the virus. The most frequent initial response was a set of non-pharmaceutical
-interventions (NPIs) to reduce contacts between individuals. While this has allowed some
+While non-pharmaceutical interventions (NPIs) have allowed some
 countries to sustain equilibria with very low infection numbers \citep{Contreras2021}
-most have seen large fluctuations of infection rates over time. Containment measures have
-become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
-tracing. Neither these policies' effects nor the influence of seasonal patterns or of
-more infectious virus strains are well understood in quantitative terms.
+most have seen large fluctuations of infection rates over time. It is therefore of utmost
+importance to employ an effective mix of strategies for containing the virus.
+However, the effects of diverse policies and their interplay with seasonal patterns
+are not well understood.
 
 This paper develops a quantitative model incorporating these factors simultaneously. The
-framework allows to combine a wide variety of data and mechanisms in a timely fashion,
+framework allows to combine a wide variety of data and mechanisms,
 making it useful to predict the effects of various interventions. We apply the model to
@@
 shows that, aside from seasonality, frequent and large-scale rapid testing caused the
-bulk of this decrease, which is in line with prior predictions \citep{Mina2021}. We
-conclude that it should have a large role for at least as long as vaccinations have not
-been offered to an entire population.
+bulk of this decrease, which is in line with prior predictions \citep{Mina2021}.
 
@@
 Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
-heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
-infectious and susceptible individuals bears the risk of transmitting the virus.
-Contacts occur in up to four networks: Within the household, at
-work, at school, or in other settings (leisure activities, grocery shopping, medical
-appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
-applications can take the population and household structure from census data and the
+heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that potentially cause
+infections. Contacts occur in up to four networks: Within the household, at
+work, at school, or in other settings. Some contacts recur regularly, others occur at random.
+Empirical applications can take the population and household structure from census data and the
 network-specific frequencies of contacts from diary data measuring contacts before the
@@
 Supplementary Material~\ref{sub:data_course_of_disease}). Whereas susceptibility to the
-viruses differs by age, the infection probabilities are independent of age conditioned
+viruses differs by age, the infectiousness is independent of age conditioned
 on the type of contact \cite{Jones2021}.
@@
     complete description.}
-Examples are contact diaries, administrative records on case numbers, virus straints,
+Examples are contact diaries, administrative records on case numbers, virus strains,
 tests and vaccines as well as surveys on private rapid test demand.
@@
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}.
-
-The main issue with official case numbers is that they will contain only a fraction of
-all infections. In the German case, this specifically amounts to positive PCR tests. We
-thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
-mortality-based aggregate estimates of the share of detected cases and use data on the
-share of PCR tests administered because of CoViD-19
-symptoms. As the share of asymptomatic individuals
-varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group}).
-% Our estimates suggest that---in the
-% absence of rapid testing---the detection rate is 80\% higher on average for individuals
-% above age 80 compared to school age children. Once rapid test become available,
-% confirmation of a positive result is another reason leading to positive PCR tests.
+matches features of the data \citep[see][for the general method]{McFadden1989}. By
+modelling PCR and rapid tests in detail, we can translate infection numbers in the model
+into observed infections before matching them to the data. The mechanism is
+depicted in Figure~\ref{fig:model_official_cases} and described in Supplementary
+Material~\ref{sub:testing}. Figure~\ref{fig:share_known_cases_by_age_group}
+shows the resulting share of detected cases.
 
@@
 of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
-Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
-never use it directly. For legibility reasons, we transform the index so that lower
-values represent higher levels of restrictions. A value of zero means all measures
-incorporated in the index are turned on. The value one represents the situation in
-mid-September, with restrictions on gatherings and public events, masking requirements,
-but open schools and workplaces. In the seven weeks between mid September and early
-November, cases increased by a factor of ten. Restrictions were somewhat tightened in
+recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} a
+rescaled Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
+non-pharmaceutical interventions. Between mid September and early
+November, cases increased tenfold. Restrictions were somewhat tightened in
 mid-October and again in early November. New infections remained constant throughout
@@
 quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
-strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
-    (Delta) reached Germany in April but still accounted for less than 5\% of cases at the
-    end of our simulation period.} In early March, some NPIs were relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-overall stringency index.
+strain, which became dominant in March (Figure~\ref{fig:share_b117}).
+In early March, some NPIs were relaxed. There were many changes in details of
+regulations afterwards, but they did not change the overall stringency index.
 
@@
 
-By March 2021, the set of policy instruments had become much more diverse. Around the
+Around the
 turn of the year, the first people were vaccinated with a focus on older age groups and
@@
 spread of the disease has been accompanied by vaccines and a growing acceptance and use
-of rapid tests. At broadly similar points in time, novel strains of the virus have
-started to pose additional challenges.
+of rapid tests.
 
@@
 In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections in 2021,
+seasonality on the evolution of infections,
 Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always held
@@
 Until mid-March, there is no visible difference between the different scenarios.
-Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
 Even thereafter, the effect of the vaccination campaign is surprisingly small.
@@
 cumulative difference between scenarios. Reasons for the low share are the slow
-start---it took until March~24th until 10\% of the population had received their first
-vaccination, the 20\% mark was reached on April 19th---and the focus on older
-individuals. By the end of our study period, when first-dose
+start and the focus on older individuals who typically have fewer contacts.
+By the end of our study period, when first-dose
 vaccination rates reached 43\% of the population, the numbers of new cases would have
 started to decline.
-
-%It is important to note that the initial focus of the campaign was
-%to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
-%lower during the third wave when compared to the second (4.4\% between October and
-%February and 1.4\% between March and the end of May).
 
@@
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A day of the fairly strict lockdown in early 2021 is
-estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
-prices for rapid tests were below one Euro in early June 2021 and below five Euros for
-firms. While we do not distinguish between self-administered rapid tests and point of
-care rapid tests, the former are likely to play a larger role for indication-driven
-testing. Widespread availability at low prices seems important. However, they rely on
-purely voluntary participation in a non-public setting. The benefit of point-of-care
-rapid tests as a precondition to participate in leisure activities as well as mandatory
-tests at the workplace or at school come from screening the entire population. This is
-important because disadvantaged groups are less likely to be reached by testing campaigns
-relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time,
-these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
-tests at school and at the workplace will extend more into these groups. The same goes
-for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
+at a fraction of the cost. A day of strict lockdown is
+estimated to have cost around 50~Euros per capita \citep{Dorn2020b}; retail
+prices for rapid tests were below one Euro in early June 2021. Widespread availability
+of self-administered tests at low prices are likely to play a role for indication-driven
+testing. Mandatory tests in schools and at the workplace are important to screen the
+entire population. Especially, because disadvantaged groups are less likely to be
+reached by testing campaigns relying on voluntary participation
+\citep[e.g.][]{StillmanTonin2021}; at the same time,
+these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}.
 Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making

### 2021-07-15 src/figures/results/tables/student_infections_table.tex

--- 2021-07-06 src/figures/results/tables/student_infections_table.tex
@@
 \midrule
- educ open after easter  without tests &                                             696963 \\
- educ open after easter  with tests    &                                             545057 \\
- close educ after easter               &                                             422360 \\
- baseline                              &                                             432320 \\
+ educ open after easter  without tests &                                             951185 \\
+ educ open after easter  with tests    &                                             722502 \\
+ close educ after easter               &                                             559526 \\
+ baseline                              &                                             582466 \\
 \bottomrule

### 2021-07-15 src/paper/abstract.tex

@@
 In order to slow the spread of the CoViD-19 pandemic, governments around the world have
-enacted a wide set of policies limiting the transmission of the disease. Initially,
-these focused on non-pharmaceutical interventions; more recently, vaccinations and
-large-scale rapid testing have started to play a major role. The objective of this study
-is to explain the quantitative effects of these policies on determining the course of
-the pandemic, allowing for factors like seasonality or virus strains with different
+enacted a wide set of policies limiting the transmission of the disease. Initially, these
+focused on non-pharmaceutical interventions; more recently, vaccinations and large-scale
+rapid testing have started to play a major role. The objective of this study is to
+explain the quantitative effects of these policies on determining the course of the
+pandemic, allowing for factors like seasonality or virus strains with different
 transmission profiles. To do so, the study develops an agent-based simulation model,
 which is estimated using data for the second and the third wave of the CoViD-19 pandemic
-in Germany. The paper finds that during a period where vaccination rates rose from 5\%
-to 40\%, rapid testing had the largest effect on reducing infection numbers. Frequent
-large-scale rapid testing should remain part of strategies to contain CoViD-19; it can
-substitute for many non-pharmaceutical interventions that come at a much larger cost to
-individuals, society, and the economy.
+in Germany. The paper finds that during a period where vaccination rates rose from 5\% to
+40\%, seasonality and rapid testing had the largest effect on reducing infection numbers.
+Frequent large-scale rapid testing should remain part of strategies to contain CoViD-19;
+it can substitute for many non-pharmaceutical interventions that come at a much larger
+cost to individuals, society, and the economy.
 

### 2021-07-15 src/paper/data/number_of_contacts.tex

@@
         For recurrent contacts individuals are put into groups that meet either every day
-        or on a particular week day every day. For work contacts, meetings can only take
-        place on work days.
-        % reductions that can happen
-        The pre-pandemic number of contacts with transmission potential is reduced by
-        policies ($\rho$), seasonality ($\kappa$) and individual responses ($\tau$) to
-        events such as receiving a positive rapid test to the number of actual contacts
-        with transmission potential.
+        or on a particular week day every day.
+        % For work contacts, meetings can only take place on work days.
+        % % reductions that can happen
+        % The pre-pandemic number of contacts with transmission potential is reduced by
+        % policies ($\rho$), seasonality ($\kappa$) and individual responses ($\tau$) to
+        % events such as receiving a positive rapid test to the number of actual contacts
+        % with transmission potential.
         % other

### 2021-07-15 src/paper/data/testing_and_behavioral_response.tex

@@
 
- \comment[id=K]{@Janos: Read this}
-
-After showing how we calibrate rapid test demand and the share of detected cases, this
-section describes the remaining parameters for our testing model. Refer to
+This section describes the remaining parameters for our testing model. Refer to
 Section~\ref{sub:testing} for a description of the full testing model.
@@
 From the share of detected cases and the number of infections we arrive at the number of
-positive PCR tests in our model. A share of these positive tests is given to symptomatic
+positive PCR tests in our model. A share of these positive tests goes to symptomatic
 individuals ($\chi_{symptom,\:t}$). This share is calibrated from German data on case
@@
 positive rapid test and this share is used for PCR test demand without prior rapid test
-indication
+indication.
 
@@
 individuals that seek a PCR test to confirm a positive rapid test result
-($\chi_{confirmation}$). We calibrate this from \citet{Betsch2021} who asked this
+($\chi_{confirmation}$). We calibrate this from \citet{Betsch2021} who asked this as a
 hypothetical question in March of 2021. There 82\% of Germans reported that they would

### 2021-07-15 src/paper/literature_review.tex

@@
 moving particles have been replaced by contact networks for households, work and random
-contacts.
+contacts. The OpenABM-Covid-19 model by \citet{Hinch2020} and the model by
+\citet{Aleta2020} are the closest in spirit to ours.
 
-The OpenABM-Covid-19 model by \citet{Hinch2020} is the closest in spirit to
-ours\comment[id=HM]{I'd rather be close to Aleta et al. ;-)}. They model detailed
-contact networks for workplaces, schools and households and can evaluate the effect of
-several NPIs. The main focus of their application are contact tracing policies.
+The OpenABM-Covid-19 model by \citet{Hinch2020} also uses detailed contact networks for
+workplaces, schools and households and can evaluate the effect of several NPIs. The main
+focus of their application are contact tracing policies. Recently they have also added
+support for multiple virus strains and vaccinations.
 
@@
 
-Moreover, there are several forthcoming applications\comment[id=HM]{What is that
-supposed to mean?}\comment[id=J]{Does that mean we should not cite the working
-papers here?} of agent-based simulation models with contact networks. Examples are
+Moreover, there are several working papers that develop agent-based simulation models
+with contact networks in conjunction with economic models. Examples are
 \citet{Basurto2020}, \citet{DelliGatti2020} and \citet{Mellacher2020}.
@@
 
-2. We have an extremely fine grained \replaced[id=K]{representation}{} of schools and
-   preschools. We can thus easily model all \replaced[id=K]{schooling policies}{NPIs}
-   that have been implemented in Germany in the past nine months. This includes complete
-   school closures, phases where only those students whose parents could not find any
-   private childcare arrangement could attend, split class approaches for some or all age
-   groups and combinations thereof. Moreover, we can account for additional hygiene
-   measures whose effect is estimated inside the model.
+2. We have an extremely fine grained representation of schools and preschools. We can
+   thus easily model all schooling policies that have been implemented in Germany in the
+   past nine months. This includes complete school closures, phases where only those
+   students whose parents could not find any private childcare arrangement could attend,
+   split class approaches for some or all age groups and combinations thereof. Moreover,
+   we can account for additional hygiene measures whose effect is estimated inside the
+   model.
 
@@
 4. We have an extremely detailed model of PCR and rapid tests with a share of detected
-   cases that varies over time and across age groups. We fit estimates of the share of
-   known cases from other studies for time periods where those are
-   available\comment[id=K]{Citation here?} but also estimate a plausible share of known
-   cases after vaccinations become widespread and methodologies that rely on the case
-   fatality ratio to identify the share of known cases are not applicable anymore.
+   cases that varies over time and across age groups.
 
- \comment[id=K]{Four is of course plenty but what about weekly recurring contacts or using the google mobility data? -> maybe add these to point 2?}
+5. Our model is designed to combine information from many different data sources.
+   Examples are surveys on rapid test demand \citep{Betsch2021}, reaction to test results
+   \citep{Betsch2021}, contact diaries \citep{Mossong2008}, share of detected cases
+   \citep{Dunkelzifferradar2020} and many more.

### 2021-07-15 src/paper/model/estimated_parameters.tex
Finish most things in section B.

@@
 are reduced over time. The dates at which we switch between the multipliers usually
-coincide with policy changes and is not determined from the \replaced[id=K]{case
-numbers}{data}. The only exception to this are slight adjustments to parameters to
-incorporate lockdown fatigue (towards the end of a lockdown period) or precautionary
-contact reductions (in times of high incidences right before a lockdown is enacted). The
-estimated other multipliers are also depicted in Figure~\ref{fig:other_multiplier}.
+coincide with policy changes and is not determined from the case numbers. The only
+exception to this are slight adjustments to parameters to incorporate lockdown fatigue
+(towards the end of a lockdown period) or precautionary contact reductions (in times of
+high incidences right before a lockdown is enacted). The estimated other multipliers are
+also depicted in Figure~\ref{fig:other_multiplier}.
 
@@
 
-While we estimate 9 different values for the other contact multiplier, they are not
+While we estimate nine different values for the other contact multiplier, they are not
 estimated completely freely. In particular we ensure that the ordering of the parameter

### 2021-07-15 src/paper/model/initial_conditions.tex

--- 2021-04-17 paper/model/initial_conditions.tex
@@
 of your simulation. Additionally, the sample of infectious people who will determine the
-course of the pandemic in the following periods is likely not representative of the
-whole population because of differences in behavior (number of contacts,
-assortativity), past policies (school closures), etc.. The distribution of courses of
-diseases in the population at the begin of the simulation is called initial conditions.
+course of the pandemic in the following periods is likely not representative of the whole
+population because of differences in behavior (number of contacts, assortativity), past
+policies (school closures), etc.. The distribution of health states in the population at
+the beginning of the simulation is called initial conditions.
 
 To come up with realistic initial conditions, we match reported infections from official
-data to simulated individuals by available characteristics like age and geographic
-information. The matching must be done for each day of a longer time frame like a month
-to have individuals with possible health states. Then, health statuses evolve until the
-begin of the simulation period without simulating infections by contacts. We also
-correct reported infections for a reporting lag and scale them up to arrive at the true
-number of infections.
+data to simulated individuals by age group and county. We use one month of data to
+generate initial conditions with in all possible health states. Meanwhile health states
+evolve until the beginning of the simulation period without simulating infections by
+contacts. We also correct reported infections for a reporting lag and scale them up by
+the share of detected cases to arrive at the true number of infections.
 

### 2021-07-15 src/paper/model/testing.tex

@@
 many of these tests should go to symptomatic and asymptomatic individuals from data by
-the RKI.\comment[id=J]{add source} Then, we sample the individuals to which those tests
+the RKI \citep{ARS2020}. Then, we sample the individuals to which those tests
 are allocated from the pools of symptomatic and asymptomatic infected but not yet tested
@@
 In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
-falsely negative. While the specificity of rapid tests is constant, their sensitivity
-strongly depends on the timing of the rapid test relative to the start of infectiousness.
-Before the onset of infectiousness the sensitivity is very low (35\%). On the first day
-of infectiousness it is much higher (88\%) but still lower than during the remaining
-infectious period (92\%). After infectiousness stops, the sensitivity drops to 50 \%.
+falsely negative. While the specificity of rapid tests is calibrated at 99.4\%
+\citep{Bruemmer2021}, their sensitivity strongly depends on the timing of the rapid test
+relative to the start of infectiousness \citep{Smith2021}. Before the onset of
+infectiousness the sensitivity is very low (35\%). On the first day of infectiousness it
+is much higher (88\%) but still lower than during the remaining infectious period (92\%).
+After infectiousness stops, the sensitivity drops to 50 \%.
 

### 2021-07-15 src/paper/report.tex

--- 2021-07-15 src/paper/report.tex
@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves months or years
-away from reaching vaccination levels that would end the pandemic \citep{Mathieu2021}.
-In the meantime, it is of utmost importance to employ an effective mix of strategies for
-containing the virus. The most frequent initial response was a set of non-pharmaceutical
-interventions (NPIs) to reduce contacts between individuals. While this has allowed some
-countries to sustain equilibria with very low infection numbers \citep{Contreras2021}
+containment in the medium term. However, most countries find themselves many months---and
+often years---away from reaching vaccination levels that would end the pandemic or even
+protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
+importance to employ an effective mix of strategies for containing the virus. The most
+frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
+contacts between individuals. While this has allowed some countries to sustain equilibria
+with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
 most have seen large fluctuations of infection rates over time. Containment measures have
@@
 heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
-infectious and susceptible individuals bears the risk of transmitting the virus.
-Contacts occur in up to four networks: Within the household, at
+an infectious individual and somebody susceptible to the disease bears the risk of
+transmitting the virus. Contacts occur in up to four networks: Within the household, at
 work, at school, or in other settings (leisure activities, grocery shopping, medical
@@
 upon having physical contact. The effect of different NPIs will generally vary across
-contact types. See Section~\ref{sec:policies} for details on how we model NPIs.
- In any setting, mask mandates or air filtes would reduce the probability
-of infection \citep{Lessler2021, Cheng2021}.
+contact types. For example, a mandate to work from
+home will reduce the number of work contacts to zero for a fraction of the working
+population. Schools and daycare can be closed entirely, operate at reduced
+capacity---including an alternating schedule---, or implement mitigation measures like
+masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
+of contacts in settings outside of
+work and school. In any setting, measures like masking requirements would reduce the
+probability of infection associated with a contact
+\citep{Cheng2021}.
 
@@
 symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The calibration
-of the age dependent probabilities of transitioning between these states  is described
-in Supplementary Material~\ref{sub:data_course_of_disease}). Conditional on the type of
-contact, infectiousness is independent of age \citep{Jones2021}.
+intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
+of transitioning between these states depend on age; their duration is random and
+calibrated to medical literature (for a detailed description see Supplementary
+Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
+infectiousness is independent of age \citep{Jones2021}.
 
-The model includes several other crucial features. New virus strains with different
-infectiousness profiles may appear. Vaccines may become available. During the vaccine
-roll-out, priority may depend on age and occupation; vaccine hesitancy is modelled by
-some individuals refusing vaccination offers. With some probability, vaccinated agents
-become immune and do not transmit the virus \citep{Hunter2021, LevineTiefenbrun2021,
-Petter2021, Pritchard2021}.
+The model includes several other features, which are crucial to describe the evolution of
+the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
+appear. Vaccines may become available. During the vaccine roll-out, priority may depend
+on age and occupation; vaccine hesitancy is modelled by some individuals refusing
+vaccination offers. With some probability, vaccinated agents become immune and do not
+transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.
 
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
-certainty whether an individual is infected or not; They require at least one day to be
-processed and are scarce. In contrast, rapid antigen tests yield immediate results.
-Specificity and sensitivity of
-these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
-sensitivity depends on the timing of the test relative to the onset of infectiousness.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
+individual is infected or not; there is no uncertainty to the result. PCR tests require
+at least one day to be processed and there are aggregate capacity constraints. In
+contrast, rapid antigen tests
+yield immediate results. Specificity and
+sensitivity of these tests is set
+according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
+timing of the test relative to the onset of infectiousness. After a phase-in period, all
+tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
+model for rapid test demand. Schools may require staff and students to be tested
+regularly. Rapid tests may be offered by
 employers to on-site workers. Individuals may demand tests for private reasons, which
@@
 compliance parameter. This parameter determines whether she takes up rapid
-tests. Positive tests and symptoms lead most individuals to reduce their contacts.
+tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
+    contacts; this is why tests impact the actual contacts in
+    Figure~\ref{fig:model_description}.}
 
@@
 parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-    complete description.}
-Examples are contact diaries, administrative records on case numbers, virus straints,
-tests and vaccines as well as surveys on private rapid test demand.
-Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, the infection probabilities
+    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
+different contact types and their assortativity by age group. Mobility data is used to
+model the evolution of work contacts. School and daycare policies can be incorporated
+directly from official directives. Administrative records on the number of tests,
+vaccinations by age and region, and the prevalence of virus strains are generally
+available. Surveys may ask about test offers, propensities to take them up, and past
+tests. Other studies' estimates of the seasonality of infections can be incorporated
+directly. The remaining parameters---most notably, these include infection probabilities
 by contact network and the effects of some NPIs, see Supplementary
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}.
+matches features of the data \citep[see][for the general method]{McFadden1989}. In our
+application, we keep the number of free parameters low in order to avoid overfitting.
+The data features to be matched include official case numbers for each age group and
+region, deaths, and the share of the B.1.1.7 strain.
 
@@
 varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group}).
-% Our estimates suggest that---in the
-% absence of rapid testing---the detection rate is 80\% higher on average for individuals
-% above age 80 compared to school age children. Once rapid test become available,
-% confirmation of a positive result is another reason leading to positive PCR tests.
+Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
+absence of rapid testing---the detection rate is 80\% higher on average for individuals
+above age 80 compared to school age children. Once rapid test become available,
+confirmation of a positive result is another reason leading to positive PCR tests.
 
-The model is applied to the second and third waves of infections in Germany,
-covering the period September 2020 to May 2021.
+The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
+covering the period mid-September 2020 to the end of May 2021.
 Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
@@
 had received at least one dose of a vaccine. In late 2020, rapid tests started to
-replace regular PCR tests for staff in many medical and nursing facilities.
-At-home tests approved by
+replace regular PCR tests for staff in many medical and nursing facilities. These had to
+be administered by medical doctors or in pharmacies. At-home tests approved by
 authorities became available in mid-March. Rapid test centers were opened, and one test
@@
 We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until May 2021. See
+use the model to predict recorded infection rates until the end of May 2021. See
 Supplementary Materials~\ref{subsec:synthetic_population} and
@@
 lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations. Second, the most stringent
+seasonality and school vacations (roughly the last two weeks of October, two weeks each
+around Christmas and Easter, and some days in late May). Second, the most stringent
 regulations coincide with the period of decreasing infection rates between late December
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained in Appendix~\ref{sub:shapley_value}.
-        All lines are rolling 7-day averages.}
+        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
+        For legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}
@@
 Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
-Even thereafter, the effect of the vaccination campaign is surprisingly small.
-The Shapley value decomposition shows that vaccinations contribute 16\% to the
+Even thereafter, the effect of the vaccination campaign is surprisingly small at first
+sight. Whether considering recorded or total infections with only one channel active,
+the final level is always the highest in case of the vaccination campaign (orange
+lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
 cumulative difference between scenarios. Reasons for the low share are the slow
@@
 vaccination, the 20\% mark was reached on April 19th---and the focus on older
-individuals. By the end of our study period, when first-dose
+individuals. These groups contribute less to the spread of the disease than others due
+to a lower number of contacts. By the end of our study period, when first-dose
 vaccination rates reached 43\% of the population, the numbers of new cases would have
-started to decline.
-
-%It is important to note that the initial focus of the campaign was
-%to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
-%lower during the third wave when compared to the second (4.4\% between October and
-%February and 1.4\% between March and the end of May).
+started to decline. It is important to note that the initial focus of the campaign was
+to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
+lower during the third wave when compared to the second (4.4\% between October and
+February and 1.4\% between March and the end of May).
 
@@
 much lower pace \citep[this is in line with results in][, which our seasonality measure
-is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be an
-important factor, explaining most of the early changes and 43\% of the cumulative
-difference by the end of May.
+is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a quantitatively
+important factor determining the evolution of the pandemic, explaining most of the early
+changes and 43\% of the cumulative difference by the end of May.
 
@@
 are turned on. The effect on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}.
-
-% Despite the fact that only 10\% of the
-% population performed weekly rapid tests in March on average, new infections on April~1
-% would have been reduced by 53\% relative to the scenario without vaccinations, rapid
-% tests, or seasonality.
+Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
+population performed weekly rapid tests in March on average, new infections on April~1
+would have been reduced by 53\% relative to the scenario without vaccinations, rapid
+tests, or seasonality.
 
@@
 elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. This is essentially a form of contact
+up on a positive test of a household member. The latter is essentially a form of contact
 tracing, which has been shown to be very effective \citep{Contreras2021,
@@
 Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not have been nearly as effective and that
-mandatory tests at schools almost offset the increased infection risk of opening
-schools (see Figure~\ref{fig:school_scenarios}).
+administered randomly in the population would not have been nearly as effective.
+
+Two of the most contentious NPIs concern schools and mandates to work from home. In many
+countries, schools switched to remote instruction during the first wave, so did Germany.
+After the summer break, they were operating at full capacity with increased hygiene
+measures, before being closed again from mid-December onward. Some states started opening
+them gradually in late February, but operation at normal capacity did not resume until
+the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
+policies regarding schools starting after Easter, at which point rapid tests had become
+widely available. We estimate the realized scenario to have essentially the same effect
+as a situation with closed schools. Under fully opened schools with mandatory tests,
+total infections would have been 6\% higher; this number rises to 20\% without tests.
+These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
+Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
+large negative effects school closures have on children and parents \citep{Luijten2021,
+Melegari2021}---and in particular on those with low socio-economic status---these results
+in conjunction with hindsight bias suggest that opening schools combined with a testing
+strategy would have been beneficial. In other situations, and particular when rapid test
+are not available at scale, trade-offs may well be different.
+
+\begin{figure}[!tp]
+    \centering
+
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \caption{{Effects of different schooling scenarios}}
+        \label{fig:school_scenarios}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+        \caption{{Effects of different work scenarios}}
+        \label{fig:workplace_scenarios}
+    \end{subfigure}
+    \vskip3ex
+
+    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
+    \label{fig:school_workplace_scenarios}
+
+    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
+        scenario; they are the same as in
+        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
+        because there were no capacity constraints for rapid tests afterwards.
+        For legibility reasons, all lines are rolling 7-day averages.}
+
+\end{figure}
+
+Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
+receiving tests, testing at the workplace has larger effects than mandating employees to
+work from home. Whether the share of workers working at the usual workplace is reduced
+or increased by ten percent changes infection rates by 2.5\% or less in either
+direction. Making testing mandatory twice a week---assuming independent compliance by
+employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
+rapid tests offers by employers to the level of March would have increased infections by
+13\%.
 
@@
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A day of the fairly strict lockdown in early 2021 is
+at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
 estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail

### 2021-07-15 src/paper/report.tex
Finish formatting of report and suppmat.

--- 2021-07-15 src/paper/report.tex
@@
-The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves months or years
-away from reaching vaccination levels that would end the pandemic \citep{Mathieu2021}.
-While non-pharmaceutical interventions (NPIs) have allowed some
-countries to sustain equilibria with very low infection numbers \citep{Contreras2021}
-most have seen large fluctuations of infection rates over time. It is therefore of utmost
-importance to employ an effective mix of strategies for containing the virus.
-However, the effects of diverse policies and their interplay with seasonal patterns
-are not well understood.
+The development of highly effective vaccines holds the promise of containment in the
+medium term. However, most countries find themselves months or years away from reaching
+vaccination levels that would end the pandemic \citep{Mathieu2021}. While
+non-pharmaceutical interventions (NPIs) have allowed some countries to sustain
+equilibria with very low infection numbers \citep{Contreras2021} most have seen large
+fluctuations of infection rates over time. It is therefore of utmost importance to
+employ an effective mix of strategies for containing the virus. However, the effects of
+diverse policies and their interplay with seasonal patterns are not well understood.
 
 This paper develops a quantitative model incorporating these factors simultaneously. The
-framework allows to combine a wide variety of data and mechanisms,
-making it useful to predict the effects of various interventions. We apply the model to
-Germany, where new infections fell by almost 80\% during May 2021. Our analysis
-shows that, aside from seasonality, frequent and large-scale rapid testing caused the
-bulk of this decrease, which is in line with prior predictions \citep{Mina2021}.
+framework allows to combine a wide variety of data and mechanisms, making it useful to
+predict the effects of various interventions. We apply the model to Germany, where new
+infections fell by almost 80\% during May 2021. Our analysis shows that, aside from
+seasonality, frequent and large-scale rapid testing caused the bulk of this decrease,
+which is in line with prior predictions \citep{Mina2021}.
 
-At the core of our agent-based model \citep[][we review more literature in Supplementary
-Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
-heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that potentially cause
-infections. Contacts occur in up to four networks: Within the household, at
-work, at school, or in other settings. Some contacts recur regularly, others occur at random.
-Empirical applications can take the population and household structure from census data and the
-network-specific frequencies of contacts from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
-depend on age and geographical location (see Supplementary
-Material~\ref{subsec:assortativity}).
+At the core of our agent-based model \cite{Aleta2020,Hinch2020} (we review more
+literature in Supplementary Material~\ref{sec:literature_review}) are physical contacts
+between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that
+potentially cause infections. Contacts occur in up to four networks: Within the
+household, at work, at school, or in other settings. Some contacts recur regularly,
+others occur at random. Empirical applications can take the population and household
+structure from census data and the network-specific frequencies of contacts from diary
+data measuring contacts before the pandemic (e.g. \cite{Mossong2008,Hoang2019}). Within
+each network, meeting frequencies depend on age and geographical location (see
+Supplementary Material~\ref{subsec:assortativity}).
 
@@
 upon having physical contact. The effect of different NPIs will generally vary across
-contact types. See Section~\ref{sec:policies} for details on how we model NPIs.
-In any setting, mask mandates or air filtes would reduce the probability
-of infection \citep{Lessler2021, Cheng2021}.
+contact types. See Section~\ref{sec:policies} for details on how we model NPIs. In any
+setting, mask mandates or air filters would reduce the probability of infection
+\citep{Lessler2021, Cheng2021}.
 
@@
 Supplementary Material~\ref{sub:data_course_of_disease}). Whereas susceptibility to the
-viruses differs by age, the infectiousness is independent of age conditioned
-on the type of contact \cite{Jones2021}.
+viruses differs by age, the infectiousness is independent of age conditioned on the type
+of contact \cite{Jones2021}.
 
@@
 infectiousness profiles may appear. Vaccines may become available. During the vaccine
-roll-out, priority may depend on age and occupation; vaccine hesitancy is modelled by
+roll-out, priority may depend on age and occupation; vaccine hesitancy is modeled by
 some individuals refusing vaccination offers. With some probability, vaccinated agents
-become immune and do not transmit the virus \citep{Hunter2021, LevineTiefenbrun2021,
+become immune and do not transmit the virus \cite{Hunter2021, LevineTiefenbrun2021,
 Petter2021, Pritchard2021}.
@@
 processed and are scarce. In contrast, rapid antigen tests yield immediate results.
-Specificity and sensitivity of
-these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
-sensitivity depends on the timing of the test relative to the onset of infectiousness.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and a household
-member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests. Positive tests and symptoms lead most individuals to reduce their contacts.
+Specificity and sensitivity of these tests is set according to data analyzed in
+\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
+to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our model for
+rapid test demand. Schools may require staff and students to be tested regularly. Rapid
+tests may be offered by employers to on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people, showing symptoms of
+CoViD-19, and a household member having tested positively for the virus. We endow each
+agent with an individual compliance parameter. This parameter determines whether she
+takes up rapid tests. Positive tests and symptoms lead most individuals to reduce their
+contacts.
 
@@
 parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-    complete description.}
-Examples are contact diaries, administrative records on case numbers, virus strains,
-tests and vaccines as well as surveys on private rapid test demand.
-Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, the infection probabilities
-by contact network and the effects of some NPIs, see Supplementary
+complete description.} Examples are contact diaries, administrative records on case
+numbers, virus strains, tests and vaccines as well as surveys on private rapid test
+demand. Other studies' estimates of the seasonality of infections can be incorporated
+directly. The remaining parameters---most notably, the infection probabilities by
+contact network and the effects of some NPIs, see Supplementary
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}. By
+matches features of the data (see \cite{McFadden1989} for the general method). By
 modelling PCR and rapid tests in detail, we can translate infection numbers in the model
-into observed infections before matching them to the data. The mechanism is
-depicted in Figure~\ref{fig:model_official_cases} and described in Supplementary
-Material~\ref{sub:testing}. Figure~\ref{fig:share_known_cases_by_age_group}
-shows the resulting share of detected cases.
+into observed infections before matching them to the data. The mechanism is depicted in
+Figure~\ref{fig:model_official_cases} and described in Supplementary
+Material~\ref{sub:testing}. Figure~\ref{fig:share_known_cases_by_age_group} shows the
+resulting share of detected cases.
 
-The model is applied to the second and third waves of infections in Germany,
-covering the period September 2020 to May 2021.
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
-of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} a
-rescaled Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. Between mid September and early
-November, cases increased tenfold. Restrictions were somewhat tightened in
-mid-October and again in early November. New infections remained constant throughout
-November before rising again in December, prompting the most stringent lockdown to this
-date. Schools and daycare centers were closed, so were customer-facing businesses except
-for grocery and drug stores. From the peak of the second wave just before Christmas
-until the trough in mid-February, newly detected cases decreased by almost three
-quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
-strain, which became dominant in March (Figure~\ref{fig:share_b117}).
-In early March, some NPIs were relaxed. There were many changes in details of
-regulations afterwards, but they did not change the overall stringency index.
+The model is applied to the second and third waves of infections in Germany, covering
+the period September 2020 to May 2021. Figure~\ref{fig:pandemic_drivers_model_fit}
+describes the evolution of the pandemic and of its drivers. The black line in
+Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the black line in
+Figure~\ref{fig:stringency_infectious_contacts} a rescaled Oxford Response Stringency
+Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
+Between mid September and early November, cases increased tenfold. Restrictions were
+somewhat tightened in mid-October and again in early November. New infections remained
+constant throughout November before rising again in December, prompting the most
+stringent lockdown to this date. Schools and daycare centers were closed, so were
+customer-facing businesses except for grocery and drug stores. From the peak of the
+second wave just before Christmas until the trough in mid-February, newly detected cases
+decreased by almost three quarters. The third wave in the spring of 2021 is associated
+with the B.1.1.7 (Alpha) strain, which became dominant in March
+(Figure~\ref{fig:share_b117}). In early March, some NPIs were relaxed. There were many
+changes in details of regulations afterwards, but they did not change the overall
+stringency index.
 
@@
 
-Around the
-turn of the year, the first people were vaccinated with a focus on older age groups and
-medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May, 43\%
-had received at least one dose of a vaccine. In late 2020, rapid tests started to
-replace regular PCR tests for staff in many medical and nursing facilities.
-At-home tests approved by
-authorities became available in mid-March. Rapid test centers were opened, and one test
-per person and week was made available free of charge. In several states, customers were
-only allowed to enter certain stores with a recent negative rapid test result. These
-developments are characteristic of many countries: The initial focus on NPIs to slow the
-spread of the disease has been accompanied by vaccines and a growing acceptance and use
-of rapid tests.
+Around the turn of the year, the first people were vaccinated with a focus on older age
+groups and medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of
+May, 43\% had received at least one dose of a vaccine. In late 2020, rapid tests started
+to replace regular PCR tests for staff in many medical and nursing facilities. At-home
+tests approved by authorities became available in mid-March. Rapid test centers were
+opened, and one test per person and week was made available free of charge. In several
+states, customers were only allowed to enter certain stores with a recent negative rapid
+test result. These developments are characteristic of many countries: The initial focus
+on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
+acceptance and use of rapid tests.
 
 We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until May 2021. See
-Supplementary Materials~\ref{subsec:synthetic_population} and
-\ref{sub:initial_conditions} for details. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model's predictions are very close to
-officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region (see Supplementary Material~\ref{subsec:fit_results}).
+use the model to predict recorded infection rates until May 2021. See Supplementary
+Materials~\ref{subsec:synthetic_population} and \ref{sub:initial_conditions} for
+details. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model's
+predictions are very close to officially recorded cases in the aggregate. This is also
+true for infections by age and geographical region (see Supplementary
+Material~\ref{subsec:fit_results}).
 
@@
 lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations. Second, the most stringent
-regulations coincide with the period of decreasing infection rates between late December
-2020 and mid-February 2021. The subsequent reversal of the trend is associated with the
-spread of the B.1.1.7 variant. During the steep drop in recorded cases during May 2021,
-for 42\% of the population took at least one rapid tests per week, the first-dose vaccination
-rate rose from 28\% to 43\%, and seasonality lowered the relative infectiousness
-of contacts.
+seasonality and school vacations. Second, the most stringent regulations coincide with
+the period of decreasing infection rates between late December 2020 and mid-February
+2021. The subsequent reversal of the trend is associated with the spread of the B.1.1.7
+variant. During the steep drop in recorded cases during May 2021, for 42\% of the
+population took at least one rapid tests per week, the first-dose vaccination rate rose
+from 28\% to 43\%, and seasonality lowered the relative infectiousness of contacts.
 
 In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections,
-Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always held
-constant at their values in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red
-line), and three scenarios turning each of these factors on individually.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
-\citep{Shapley2016} to decompose the difference in total infections between the scenario
-without any of the three factors and our main specification.
+seasonality on the evolution of infections, Figure~\ref{fig:2021_scenarios_broad}
+considers various scenarios. NPIs are always held constant at their values in the
+baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the
+blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the
+three factors (red line), and three scenarios turning each of these factors on
+individually. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
+values \citep{Shapley2016} to decompose the difference in total infections between the
+scenario without any of the three factors and our main specification.
 
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained in Appendix~\ref{sub:shapley_value}.
-        All lines are rolling 7-day averages.}
+        are explained in Appendix~\ref{sub:shapley_value}. All lines are rolling 7-day
+        averages.
+    }
 \end{figure}
 
-Until mid-March, there is no visible difference between the different scenarios.
-Even thereafter, the effect of the vaccination campaign is surprisingly small.
-The Shapley value decomposition shows that vaccinations contribute 16\% to the
-cumulative difference between scenarios. Reasons for the low share are the slow
-start and the focus on older individuals who typically have fewer contacts.
-By the end of our study period, when first-dose
-vaccination rates reached 43\% of the population, the numbers of new cases would have
-started to decline.
+Until mid-March, there is no visible difference between the different scenarios. Even
+thereafter, the effect of the vaccination campaign is surprisingly small. The Shapley
+value decomposition shows that vaccinations contribute 16\% to the cumulative difference
+between scenarios. Reasons for the low share are the slow start and the focus on older
+individuals who typically have fewer contacts. By the end of our study period, when
+first-dose vaccination rates reached 43\% of the population, the numbers of new cases
+would have started to decline.
 
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace \citep[this is in line with results in][, which our seasonality measure
-is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be an
+much lower pace (this is in line with results in \cite{Gavenciak2021}, which our
+seasonality measure is based on). Nevertheless, we estimate seasonality to be an
 important factor, explaining most of the early changes and 43\% of the cumulative
@@
 Figure~\ref{fig:2021_scenarios_newly_infected}.
-
-% Despite the fact that only 10\% of the
-% population performed weekly rapid tests in March on average, new infections on April~1
-% would have been reduced by 53\% relative to the scenario without vaccinations, rapid
-% tests, or seasonality.
 
@@
 up on a positive test of a household member. This is essentially a form of contact
-tracing, which has been shown to be very effective \citep{Contreras2021,
+tracing, which has been shown to be very effective \cite{Contreras2021,
 Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
@@
 administered randomly in the population would not have been nearly as effective and that
-mandatory tests at schools almost offset the increased infection risk of opening
-schools (see Figure~\ref{fig:school_scenarios}).
+mandatory tests at schools almost offset the increased infection risk of opening schools
+(see Figure~\ref{fig:school_scenarios}).
 
@@
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A day of strict lockdown is
-estimated to have cost around 50~Euros per capita \citep{Dorn2020b}; retail
-prices for rapid tests were below one Euro in early June 2021. Widespread availability
-of self-administered tests at low prices are likely to play a role for indication-driven
-testing. Mandatory tests in schools and at the workplace are important to screen the
-entire population. Especially, because disadvantaged groups are less likely to be
-reached by testing campaigns relying on voluntary participation
-\citep[e.g.][]{StillmanTonin2021}; at the same time,
-these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}.
-Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
-it arguably the most effective tool to contain the pandemic in the short run.
+at a fraction of the cost. A day of strict lockdown is estimated to have cost around
+50~Euros per capita \cite{Dorn2020b}; retail prices for rapid tests were below one Euro
+in early June 2021. Widespread availability of self-administered tests at low prices are
+likely to play a role for indication-driven testing. Mandatory tests in schools and at
+the workplace are important to screen the entire population. Especially, because
+disadvantaged groups are less likely to be reached by testing campaigns relying on
+voluntary participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups
+have a higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to
+vaccinations, rapid testing programmes allow a much quicker roll-out, making it arguably
+the most effective tool to contain the pandemic in the short run.

### 2021-07-15 src/science/paper.tex
Finish formatting of report and suppmat.

@@
 
-Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
-on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination levels that would end the pandemic or even
-protect the most vulnerable \cite{Mathieu2021}. In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \cite{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate
-policies.} most have
-seen large fluctuations of infection rates over time. Containment measures have become
-increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
-tracing. Neither these policies' effects nor the influence of seasonal patterns or of
-more infectious virus strains are well understood in quantitative terms.
+The development of highly effective vaccines holds the promise of containment in the
+medium term. However, most countries find themselves months or years away from reaching
+vaccination levels that would end the pandemic \cite{Mathieu2021}. While
+non-pharmaceutical interventions (NPIs) have allowed some countries to sustain
+equilibria with very low infection numbers \cite{Contreras2021} most have seen large
+fluctuations of infection rates over time. It is therefore of utmost importance to
+employ an effective mix of strategies for containing the virus. However, the effects of
+diverse policies and their interplay with seasonal patterns are not well understood.
 
 This paper develops a quantitative model incorporating these factors simultaneously. The
-framework allows to combine a wide variety of data and mechanisms in a timely fashion,
-making it useful to predict the effects of various interventions. We apply the model to
-Germany, where new infections fell by almost 80\% during May 2021. Our analysis
-shows that, aside from seasonality, frequent and large-scale rapid testing caused the
-bulk of this decrease, which is in line with prior predictions \cite{Mina2021}. We
-conclude that it should have a large role for at least as long as vaccinations have not
-been offered to an entire population.
+framework allows to combine a wide variety of data and mechanisms, making it useful to
+predict the effects of various interventions. We apply the model to Germany, where new
+infections fell by almost 80\% during May 2021. Our analysis shows that, aside from
+seasonality, frequent and large-scale rapid testing caused the bulk of this decrease,
+which is in line with prior predictions \cite{Mina2021}.
 
 At the core of our agent-based model \cite{Aleta2020,Hinch2020} (we review more
-literature in Supplementary Material~\ref{sec:literature_review}) are physical contacts
-between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact
-between an infectious individual and somebody susceptible to the disease bears the risk
-of transmitting the virus. Contacts occur in up to four networks: Within the household,
-at work, at school, or in other settings (leisure activities, grocery shopping, medical
-appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
-applications can take the population and household structure from census data and the
-network-specific frequencies of contacts from diary data measuring contacts before the
-pandemic \cite{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
-on age and geographical location (see Supplementary
-Material~\ref{subsec:assortativity}).
+literature in Supplementary Material~Fig.~B.1) are physical contacts
+between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that
+potentially cause infections. Contacts occur in up to four networks: Within the
+household, at work, at school, or in other settings. Some contacts recur regularly,
+others occur at random. Empirical applications can take the population and household
+structure from census data and the network-specific frequencies of contacts from diary
+data measuring contacts before the pandemic (e.g. \cite{Mossong2008,Hoang2019}). Within
+each network, meeting frequencies depend on age and geographical location (see
+Supplementary Material~A.4).
 
@@
 upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from
-home will reduce the number of work contacts to zero for a fraction of the working
-population. Schools and daycare can be closed entirely, operate at reduced
-capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \cite{Lessler2021}. Curfews may reduce the number
-of contacts in settings outside of
-work and school. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact
-\cite{Cheng2021}.
+contact types. See Section~B.4 for details on how we model NPIs. In any
+setting, mask mandates or air filters would reduce the probability of infection
+\cite{Lessler2021, Cheng2021}.
 
@@
     \floatfoot{\noindent \textit{Note:}
-        A~description of the model can be found in Supplementary
-        Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
-        influence of an agent's contacts to other agents on infections. Demographic
-        characteristics set the baseline number of contacts in different networks
-        ($\eta$). The agents may reduce the number of contacts due to NPIs, showing
-        symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may occur
-        when a susceptible agent meets an infectious agent; the probability depends on
-        the type of contact ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs
-        ($\rho_{c,\:t}$). If infected, the infection progresses as depicted in
+        A~description of the model can be found in Supplementary Material~B.
+        Figure~\ref{fig:model_contacts_infections} shows the influence of an agent's
+        contacts to other agents on infections. Demographic characteristics set the
+        baseline number of contacts in different networks ($\eta$). The agents may
+        reduce the number of contacts due to NPIs, showing symptoms, or testing
+        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
+        agent meets an infectious agent; the probability depends on the type of contact
+        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
+        infected, the infection progresses as depicted in
         Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
@@
         of the graph is relevant only for periods where rapid tests are available. All
-        parameters are explained in Appendix~\ref{sub:param_tables}.}
+        parameters are explained in Appendix~A.11.}
 \end{figure}
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
-age. A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}. We differentiate
-between an initial period of infection without being infectious or showing symptoms,
-being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
-care, and recovery or death (similar to \cite{Aleta2020}). The probabilities of
-transitioning between these states depend on age; their duration is random and calibrated
-to medical literature (for a detailed description see Supplementary
-Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \cite{Jones2021}.
+In the model, a possible infection progresses as shown in
+Figure~\ref{fig:model_disease_progression} (similar to \cite{Aleta2020}, more in
+Supplementary Material~A.1). Whereas susceptibility to the viruses differs by age, the
+infectiousness is independent of age conditioned on the type of contact
+\cite{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
-appear. Vaccines may become available. During the vaccine roll-out, priority may depend
-on age and occupation; vaccine hesitancy is modelled by some individuals refusing
-vaccination offers. With some probability, vaccinated agents become immune and do not
-transmit the virus \cite{Hunter2021, LevineTiefenbrun2021, Petter2021,
-Pritchard2021}.
+The model includes several other crucial features. New virus strains with different
+infectiousness profiles may appear. Vaccines may become available. During the vaccine
+roll-out, priority may depend on age and occupation; vaccine hesitancy is modeled by
+some individuals refusing vaccination offers. With some probability, vaccinated agents
+become immune and do not transmit the virus \cite{Hunter2021, LevineTiefenbrun2021,
+Petter2021, Pritchard2021}.
 
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
-individual is infected or not; there is no uncertainty to the result. PCR tests require
-at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests
-yield immediate results. Specificity and
-sensitivity of these tests is set
-according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the onset of infectiousness. After a phase-in period, all
-tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and a household
-member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-    contacts; this is why tests impact the actual contacts in
-    Figure~\ref{fig:model_description}.}
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
+certainty whether an individual is infected or not; They require at least one day to be
+processed and are scarce. In contrast, rapid antigen tests yield immediate results.
+Specificity and sensitivity of these tests is set according to data analyzed in
+\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
+to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our model for
+rapid test demand. Schools may require staff and students to be tested regularly. Rapid
+tests may be offered by employers to on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people, showing symptoms of
+CoViD-19, and a household member having tested positively for the virus. We endow each
+agent with an individual compliance parameter. This parameter determines whether she
+takes up rapid tests. Positive tests and symptoms lead most individuals to reduce their
+contacts.
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~\ref{sec:data_and_parameters} for a
-    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
-different contact types and their assortativity by age group. Mobility data is used to
-model the evolution of work contacts. School and daycare policies can be incorporated
-directly from official directives. Administrative records on the number of tests,
-vaccinations by age and region, and the prevalence of virus strains are generally
-available. Surveys may ask about test offers, propensities to take them up, and past
-tests. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, these include infection probabilities
-by contact network and the effects of some NPIs, see Supplementary
-Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data (see \cite{McFadden1989} for the general method). In our
-application, we keep the number of free parameters low in order to avoid overfitting.
-The data features to be matched include official case numbers for each age group and
-region, deaths, and the share of the B.1.1.7 strain.
+parameters.\footnote{See Supplementary Material~A for a complete description.} Examples
+are contact diaries, administrative records on case numbers, virus strains, tests and
+vaccines as well as surveys on private rapid test demand. Other studies' estimates of
+the seasonality of infections can be incorporated directly. The remaining
+parameters---most notably, the infection probabilities by contact network and the
+effects of some NPIs, see Supplementary Material~A.9---will be chosen numerically so
+that the model matches features of the data (see \cite{McFadden1989} for the general
+method). By modelling PCR and rapid tests in detail, we can translate infection numbers
+in the model into observed infections before matching them to the data. The mechanism is
+depicted in Figure~\ref{fig:model_official_cases} and described in Supplementary
+Material~B.7. Figure~B.9 shows the resulting share of detected cases.
 
-The main issue with official case numbers is that they will contain only a fraction of
-all infections. In the German case, this specifically amounts to positive PCR tests. We
-thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
-mortality-based aggregate estimates of the share of detected cases and use data on the
-share of PCR tests administered because of CoViD-19
-symptoms. As the share of asymptomatic individuals
-varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is 80\% higher on average for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
-
-The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
-covering the period mid-September 2020 to the end of May 2021.
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
-of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
-Oxford Response Stringency Index \cite{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
-never use it directly. For legibility reasons, we transform the index so that lower
-values represent higher levels of restrictions. A value of zero means all measures
-incorporated in the index are turned on. The value one represents the situation in
-mid-September, with restrictions on gatherings and public events, masking requirements,
-but open schools and workplaces. In the seven weeks between mid September and early
-November, cases increased by a factor of ten. Restrictions were somewhat tightened in
-mid-October and again in early November. New infections remained constant throughout
-November before rising again in December, prompting the most stringent lockdown to this
-date. Schools and daycare centers were closed, so were customer-facing businesses except
-for grocery and drug stores. From the peak of the second wave just before Christmas
-until the trough in mid-February, newly detected cases decreased by almost three
-quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
-strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
-    (Delta) reached Germany in April but still accounted for less than 5\% of cases at the
-    end of our simulation period.} In early March, some NPIs were relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-overall stringency index.
+The model is applied to the second and third waves of infections in Germany, covering
+the period September 2020 to May 2021. Figure~\ref{fig:pandemic_drivers_model_fit}
+describes the evolution of the pandemic and of its drivers. The black line in
+Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the black line in
+Figure~\ref{fig:stringency_infectious_contacts} a rescaled Oxford Response Stringency
+Index \cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
+Between mid September and early November, cases increased tenfold. Restrictions were
+somewhat tightened in mid-October and again in early November. New infections remained
+constant throughout November before rising again in December, prompting the most
+stringent lockdown to this date. Schools and daycare centers were closed, so were
+customer-facing businesses except for grocery and drug stores. From the peak of the
+second wave just before Christmas until the trough in mid-February, newly detected cases
+decreased by almost three quarters. The third wave in the spring of 2021 is associated
+with the B.1.1.7 (Alpha) strain, which became dominant in March
+(Figure~\ref{fig:share_b117}). In early March, some NPIs were relaxed. There were many
+changes in details of regulations afterwards, but they did not change the overall
+stringency index.
 
@@
     \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
-        Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
-        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
-        \ref{subsec:fit_results}. For legibility reasons, all lines in
-        Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
-        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
-        a value of one refers to the situation at the start of our sample period and
-        zero means that all NPIs included in the index are turned on. The other lines in
+        Material~A. Age- and region-specific analogues to
+        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material B.10. For
+        legibility reasons, all lines in Figure~\ref{fig:stringency_infectious_contacts}
+        are rolling 7-day averages. The Oxford Response Stringency Index is scaled as $2
+        \cdot (1 -  x / 100)$, so that a value of one refers to the situation at the
+        start of our sample period and zero means that all NPIs included in the index
+        are turned on. The other lines in
         Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
         of contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~\ref{subsec:policies} for separate plots of the three factors by
-        contact type.}
+        Appendix~A.5 for separate plots of the three factors by contact type.}
 \end{figure}
 
-By March 2021, the set of policy instruments had become much more diverse. Around the
-turn of the year, the first people were vaccinated with a focus on older age groups and
-medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May, 43\%
-had received at least one dose of a vaccine. In late 2020, rapid tests started to
-replace regular PCR tests for staff in many medical and nursing facilities. These had to
-be administered by medical doctors or in pharmacies. At-home tests approved by
-authorities became available in mid-March. Rapid test centers were opened, and one test
-per person and week was made available free of charge. In several states, customers were
-only allowed to enter certain stores with a recent negative rapid test result. These
-developments are characteristic of many countries: The initial focus on NPIs to slow the
-spread of the disease has been accompanied by vaccines and a growing acceptance and use
-of rapid tests. At broadly similar points in time, novel strains of the virus have
-started to pose additional challenges.
+Around the turn of the year, the first people were vaccinated with a focus on older age
+groups and medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of
+May, 43\% had received at least one dose of a vaccine. In late 2020, rapid tests started
+to replace regular PCR tests for staff in many medical and nursing facilities. At-home
+tests approved by authorities became available in mid-March. Rapid test centers were
+opened, and one test per person and week was made available free of charge. In several
+states, customers were only allowed to enter certain stores with a recent negative rapid
+test result. These developments are characteristic of many countries: The initial focus
+on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
+acceptance and use of rapid tests.
 
 We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until the end of May 2021. See
-Supplementary Materials~\ref{subsec:synthetic_population} and
-\ref{sub:initial_conditions} for details. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model's predictions are very close to
-officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region (see Supplementary Material~\ref{subsec:fit_results}).
+use the model to predict recorded infection rates until May 2021. See Supplementary
+Materials~A.2 and B.9 for details. The blue line in Figure~\ref{fig:aggregated_fit}
+shows that our model's predictions are very close to officially recorded cases in the
+aggregate. This is also true for infections by age and geographical region (see
+Supplementary Material~B.10).
 
@@
 lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations (roughly the last two weeks of October, two weeks each
-around Christmas and Easter, and some days in late May). Second, the most stringent
-regulations coincide with the period of decreasing infection rates between late December
-2020 and mid-February 2021. The subsequent reversal of the trend is associated with the
-spread of the B.1.1.7 variant. During the steep drop in recorded cases during May 2021,
-for 42\% of the population took at least one rapid tests per week, the first-dose vaccination
-rate rose from 28\% to 43\%, and seasonality lowered the relative infectiousness
-of contacts.
+seasonality and school vacations. Second, the most stringent regulations coincide with
+the period of decreasing infection rates between late December 2020 and mid-February
+2021. The subsequent reversal of the trend is associated with the spread of the B.1.1.7
+variant. During the steep drop in recorded cases during May 2021, for 42\% of the
+population took at least one rapid tests per week, the first-dose vaccination rate rose
+from 28\% to 43\%, and seasonality lowered the relative infectiousness of contacts.
 
 In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections in 2021,
-Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always held
-constant at their values in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red
-line), and three scenarios turning each of these factors on individually.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
-\cite{Shapley2016} to decompose the difference in total infections between the scenario
-without any of the three factors and our main specification.
+seasonality on the evolution of infections, Figure~\ref{fig:2021_scenarios_broad}
+considers various scenarios. NPIs are always held constant at their values in the
+baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the
+blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the
+three factors (red line), and three scenarios turning each of these factors on
+individually. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
+values \cite{Shapley2016} to decompose the difference in total infections between the
+scenario without any of the three factors and our main specification.
 
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
-        For legibility reasons, all lines are rolling 7-day averages.}
+        are explained in Appendix~A.10. All lines are rolling 7-day averages.
+    }
 \end{figure}
 
-Until mid-March, there is no visible difference between the different scenarios.
-Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
-Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections with only one channel active,
-the final level is always the highest in case of the vaccination campaign (orange
-lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
-cumulative difference between scenarios. Reasons for the low share are the slow
-start---it took until March~24th until 10\% of the population had received their first
-vaccination, the 20\% mark was reached on April 19th---and the focus on older
-individuals. These groups contribute less to the spread of the disease than others due
-to a lower number of contacts. By the end of our study period, when first-dose
-vaccination rates reached 43\% of the population, the numbers of new cases would have
-started to decline. It is important to note that the initial focus of the campaign was
-to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
-lower during the third wave when compared to the second (4.4\% between October and
-February and 1.4\% between March and the end of May).
+Until mid-March, there is no visible difference between the different scenarios. Even
+thereafter, the effect of the vaccination campaign is surprisingly small. The Shapley
+value decomposition shows that vaccinations contribute 16\% to the cumulative difference
+between scenarios. Reasons for the low share are the slow start and the focus on older
+individuals who typically have fewer contacts. By the end of our study period, when
+first-dose vaccination rates reached 43\% of the population, the numbers of new cases
+would have started to decline.
 
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace (This is in line with results in \cite{Gavenciak2021} which our
-seasonality measure is based on). Nevertheless, we estimate seasonality to be a
-quantitatively important factor determining the evolution of the pandemic, explaining
-most of the early changes and 43\% of the cumulative difference by the end of May.
+much lower pace (this is in line with results in \cite{Gavenciak2021}, which our
+seasonality measure is based on). Nevertheless, we estimate seasonality to be an
+important factor, explaining most of the early changes and 43\% of the cumulative
+difference by the end of May.
 
@@
 are turned on. The effect on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
-population performed weekly rapid tests in March on average, new infections on April~1
-would have been reduced by 53\% relative to the scenario without vaccinations, rapid
-tests, or seasonality.
+Figure~\ref{fig:2021_scenarios_newly_infected}.
 
@@
 elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. The latter is essentially a form of contact
+up on a positive test of a household member. This is essentially a form of contact
 tracing, which has been shown to be very effective \cite{Contreras2021,
-    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
-Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not have been nearly as effective.
-
-Two of the most contentious NPIs concern schools and mandates to work from home. In many
-countries, schools switched to remote instruction during the first wave, so did Germany.
-After the summer break, they were operating at full capacity with increased hygiene
-measures, before being closed again from mid-December onward. Some states started
-opening them gradually in late February, but operation at normal capacity did not resume
-until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
-different policies regarding schools starting after Easter, at which point rapid tests
-had become widely available. We estimate the realized scenario to have essentially the
-same effect as a situation with closed schools. Under fully opened schools with
-mandatory tests, total infections would have been 6\% higher; this number rises to 20\%
-without tests. These effect sizes are broadly in line with empirical studies (e.g.,
-\cite{Vlachos2021}). To use another metric, the effective weekly reproduction number
-differs by $0.018$ and $0.052$, respectively. In light of the large negative effects
-school closures have on children and parents \cite{Luijten2021, Melegari2021}---and in
-particular on those with low socio-economic status---these results in conjunction with
-hindsight bias suggest that opening schools combined with a testing strategy would have
-been beneficial. In other situations, and particular when rapid test are not available
-at scale, trade-offs may well be different.
-
-\begin{figure}[!tp]
-    \centering
-
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-        \caption{{Effects of different schooling scenarios}}
-        \label{fig:school_scenarios}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
-        \caption{{Effects of different work scenarios}}
-        \label{fig:workplace_scenarios}
-    \end{subfigure}
-    \vskip3ex
-
-    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
-    \label{fig:school_workplace_scenarios}
-
-    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
-        scenario; they are the same as in
-        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
-        because there were no capacity constraints for rapid tests afterwards.
-        For legibility reasons, all lines are rolling 7-day averages.}
-
-\end{figure}
-
-Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
-receiving tests, testing at the workplace has larger effects than mandating employees to
-work from home. Whether the share of workers working at the usual workplace is reduced
-or increased by ten percent changes infection rates by 2.5\% or less in either
-direction. Making testing mandatory twice a week---assuming independent compliance by
-employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
-rapid tests offers by employers to the level of March would have increased infections by
-13\%.
+Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary Material~B.14
+shows that the same amount of rapid tests administered randomly in the population would
+not have been nearly as effective and that mandatory tests at schools almost offset the
+increased infection risk of opening schools (see Figure~B.13a).
 
@@
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 20~Euros per capita \cite{Wollmershauser2021}; retail
-prices for rapid tests were below one Euro in early June 2021. While we do not
-distinguish between self-administered rapid tests and point of care rapid tests, the
-former are likely to play a larger role for indication-driven testing. Widespread
-availability at low prices seems important. However, they rely on purely voluntary
-participation in a non-public setting. The benefit of point-of-care rapid tests as a
-precondition to participate in leisure activities as well as mandatory tests at the
-workplace or at school come from screening the entire population. This is important
-because disadvantaged groups are less likely to be reached by testing campaigns relying
-on voluntary participation (e.g., \cite{StillmanTonin2021}); at
-the same time, these groups have a higher risk to contract CoViD-19
-\cite{KochInstitut2021a}. Mandatory tests at school and at the workplace will extend
-more into these groups. The same goes for individuals who exhibit a low level of
-compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
-programmes allow a much quicker roll-out, making it arguably the most effective tool to
-contain the pandemic in the short run.
+at a fraction of the cost. A day of strict lockdown is estimated to have cost around
+50~Euros per capita \cite{Dorn2020b}; retail prices for rapid tests were below one Euro
+in early June 2021. Widespread availability of self-administered tests at low prices are
+likely to play a role for indication-driven testing. Mandatory tests in schools and at
+the workplace are important to screen the entire population. Especially, because
+disadvantaged groups are less likely to be reached by testing campaigns relying on
+voluntary participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups
+have a higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to
+vaccinations, rapid testing programmes allow a much quicker roll-out, making it arguably
+the most effective tool to contain the pandemic in the short run.
 
@@
 \section*{Acknowledgments}
-Include acknowledgments of funding, any patents pending, where raw data for the paper are deposited, etc.
 
-%Here you should list the contents of your Supplementary Materials -- below is an example.
-%You should include a list of Supplementary figures, Tables, and any references that appear only in the SM.
-%Note that the reference numbering continues from the main text to the SM.
-% In the example below, Refs. 4-10 were cited only in the SM.
+The authors are grateful for support by the Deutsche Forschungsgemeinschaft (DFG, German
+Research Foundation) under Germany´s Excellence Strategy – EXC 2126/1– 390838866 – and
+through CRC-TR 224 (Projects A02 and C01), by the IZA Institute of Labor Economics, and
+by the Google Cloud CoViD-19 research credits program.
+
+%Here you should list the contents of your Supplementary Materials -- below is an
+%example. You should include a list of Supplementary figures, Tables, and any references
+%that appear only in the SM. Note that the reference numbering continues from the main
+%text to the SM. In the example below, Refs. 4-10 were cited only in the SM.
+
 \section*{Supplementary materials}
-Materials and Methods\\
-Supplementary Text\\
-Figs. S1 to S3\\
-Tables S1 to S4\\
-References \textit{(4-10)}
+
+Appendix A - Materials and Methods\\
+Appendix B - Supplementary Text\\
+Figs. A.1-A.10, B.1-B.16\\
+Tables A.1-A.6\\
+References \textit{(26-111)}
 

### 2021-07-15 task_compile_documents.py

@@ -59,7 +59,7 @@ def task_extract_supplementary_material(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "15-67",
+            "12-78",
             "--",
             produces.as_posix(),
         ]

### 2021-07-15 src/science/cover_letter.tex
Add draft for cover letter.

--- 2021-07-15 src/science/paper.tex
@@
-% Use only LaTeX2e, calling the article.cls class and 12-point type.
-
-\documentclass[12pt]{article}
-
-% Users of the {thebibliography} environment or BibTeX should use the
-% scicite.sty package, downloadable from *Science* at
-% http://www.sciencemag.org/authors/preparing-manuscripts-using-latex
-% This package should properly format in-text
-% reference calls and reference-list numbers.
-
-\usepackage{scicite}
-
-\usepackage{times}
-
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{floatrow}
-
-% The preamble here sets up a lot of new/revised commands and
-% environments.  It's annoying, but please do *not* try to strip these
-% out into a separate .sty file (which could lead to the loss of some
-% information when we convert the file to other formats).  Instead, keep
-% them in the preamble of your main LaTeX source file.
-
-% The following parameters seem to provide a reasonable page setup.
-
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm
-\textheight 21cm
-\footskip 1.0cm
-
-%The next command sets up an environment for the abstract to your paper.
-
-\newenvironment{sciabstract}{%
-\begin{quote} \bf}
-{\end{quote}}
-
-% Include your paper's title here
+\documentclass[a4paper, 10pt]{article}
 
 \title{The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs}
+
+\usepackage{url}
 
@@
 
-% Include the date command, but leave its argument blank.
-
-\date{}
-
-%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
-
 \begin{document}
-
-% Double-space the manuscript.
-
-\baselineskip24pt
-
-% Make the title.
 
@@
 
-% Place your abstract within the special {sciabstract} environment.
+Main point: Our main methodological contribution is to calibrate a very detailed and
+flexible agent based model to a long time series of data during the pandemic. Our core
+substantive result is the importance of rapid testing as a supplement to NPIs. Due to
+the slow vaccine roll out world wide, this will remain topical for some time.
 
-\begin{sciabstract}
-    In order to slow the spread of the CoViD-19 pandemic, governments around the world
-    have enacted a wide set of policies limiting the transmission of the disease.
-    Initially, these focused on non-pharmaceutical interventions; more recently,
-    vaccinations and large-scale rapid testing have started to play a major role. The
-    objective of this study is to explain the quantitative effects of these policies on
-    determining the course of the pandemic, allowing for factors like seasonality or
-    virus strains with different transmission profiles. To do so, the study develops an
-    agent-based simulation model, which is estimated using data for the second and the
-    third wave of the CoViD-19 pandemic in Germany. The paper finds that during a period
-    where vaccination rates rose from 5\% to 40\%, seasonality and rapid testing had the
-    largest effect on reducing infection numbers. Frequent large-scale rapid testing
-    should remain part of strategies to contain CoViD-19; it can substitute for many
-    non-pharmaceutical interventions that come at a much larger cost to individuals,
-    society, and the economy.
-\end{sciabstract}
+There were no submissions to other journals, nor discussions with editors prior to
+submitting the report.
 
-The development of highly effective vaccines holds the promise of containment in the
-medium term. However, most countries find themselves months or years away from reaching
-vaccination levels that would end the pandemic \cite{Mathieu2021}. While
-non-pharmaceutical interventions (NPIs) have allowed some countries to sustain
-equilibria with very low infection numbers \cite{Contreras2021} most have seen large
-fluctuations of infection rates over time. It is therefore of utmost importance to
-employ an effective mix of strategies for containing the virus. However, the effects of
-diverse policies and their interplay with seasonal patterns are not well understood.
+A draft of this report was reviewed by Prof. Jörg Stoye and Prof. Philipp Eisenhauer.
 
-This paper develops a quantitative model incorporating these factors simultaneously. The
-framework allows to combine a wide variety of data and mechanisms, making it useful to
-predict the effects of various interventions. We apply the model to Germany, where new
-infections fell by almost 80\% during May 2021. Our analysis shows that, aside from
-seasonality, frequent and large-scale rapid testing caused the bulk of this decrease,
-which is in line with prior predictions \cite{Mina2021}.
+A draft of this report has been published as a pre-print on arXiv
+(\url{https://arxiv.org/abs/2106.11129}). It is not under consideration for publication
+elsewhere.
 
-At the core of our agent-based model \cite{Aleta2020,Hinch2020} (we review more
-literature in Supplementary Material~Fig.~B.1) are physical contacts
-between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that
-potentially cause infections. Contacts occur in up to four networks: Within the
-household, at work, at school, or in other settings. Some contacts recur regularly,
-others occur at random. Empirical applications can take the population and household
-structure from census data and the network-specific frequencies of contacts from diary
-data measuring contacts before the pandemic (e.g. \cite{Mossong2008,Hoang2019}). Within
-each network, meeting frequencies depend on age and geographical location (see
-Supplementary Material~A.4).
+All data is freely available in linked to in the supplementary material. Only one data
+set, the microcensus of 2010, provided by the FDZ (Forschungsdatenzentren) requires
+registration prior to access
+(\url{http://www.forschungsdatenzentrum.de/de/campus-files})
 
-The four contact networks are chosen so that the most common NPIs can be modeled in
-great detail. NPIs affect the number of contacts or the risk of transmitting the disease
-upon having physical contact. The effect of different NPIs will generally vary across
-contact types. See Section~B.4 for details on how we model NPIs. In any
-setting, mask mandates or air filters would reduce the probability of infection
-\cite{Lessler2021, Cheng2021}.
-
-\begin{figure}   % Figure 1
-    \centering
-
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
-        \caption{{Model for contacts and infections}}
-        \label{fig:model_contacts_infections}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
-        \vskip5ex
-
-        \caption{Disease progression}
-        \label{fig:model_disease_progression}
-    \end{subfigure}
-    \vskip3ex
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
-        \caption{{Model for rapid tests}}
-        \label{fig:model_rapid_tests}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
-        \vskip5ex
-
-        \caption{{Translating all infections to recorded ones}}
-        \label{fig:model_official_cases}
-    \end{subfigure}
-
-    \caption{Model description}
-    \label{fig:model_description}
-
-    \floatfoot{\noindent \textit{Note:}
-        A~description of the model can be found in Supplementary Material~B.
-        Figure~\ref{fig:model_contacts_infections} shows the influence of an agent's
-        contacts to other agents on infections. Demographic characteristics set the
-        baseline number of contacts in different networks ($\eta$). The agents may
-        reduce the number of contacts due to NPIs, showing symptoms, or testing
-        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
-        agent meets an infectious agent; the probability depends on the type of contact
-        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
-        infected, the infection progresses as depicted in
-        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
-        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
-        trigger a test only for a fraction of individuals depending on an individual
-        compliance parameter; the thresholds for triggering test demand differ across
-        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
-        Figure~\ref{fig:model_official_cases} shows the model of translating all
-        infections in the simulated data to age-specific recorded infections. The model
-        uses data on the aggregate share of recorded cases ($\psi$), the share of
-        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and the false
-        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
-        of the graph is relevant only for periods where rapid tests are available. All
-        parameters are explained in Appendix~A.11.}
-\end{figure}
-
-In the model, a possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression} (similar to \cite{Aleta2020}, more in
-Supplementary Material~A.1). Whereas susceptibility to the viruses differs by age, the
-infectiousness is independent of age conditioned on the type of contact
-\cite{Jones2021}.
-
-The model includes several other crucial features. New virus strains with different
-infectiousness profiles may appear. Vaccines may become available. During the vaccine
-roll-out, priority may depend on age and occupation; vaccine hesitancy is modeled by
-some individuals refusing vaccination offers. With some probability, vaccinated agents
-become immune and do not transmit the virus \cite{Hunter2021, LevineTiefenbrun2021,
-Petter2021, Pritchard2021}.
-
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
-certainty whether an individual is infected or not; They require at least one day to be
-processed and are scarce. In contrast, rapid antigen tests yield immediate results.
-Specificity and sensitivity of these tests is set according to data analyzed in
-\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
-to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our model for
-rapid test demand. Schools may require staff and students to be tested regularly. Rapid
-tests may be offered by employers to on-site workers. Individuals may demand tests for
-private reasons, which include having plans to meet other people, showing symptoms of
-CoViD-19, and a household member having tested positively for the virus. We endow each
-agent with an individual compliance parameter. This parameter determines whether she
-takes up rapid tests. Positive tests and symptoms lead most individuals to reduce their
-contacts.
-
-Modelling a population of agents according to actual demographic characteristics means
-that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~A for a complete description.} Examples
-are contact diaries, administrative records on case numbers, virus strains, tests and
-vaccines as well as surveys on private rapid test demand. Other studies' estimates of
-the seasonality of infections can be incorporated directly. The remaining
-parameters---most notably, the infection probabilities by contact network and the
-effects of some NPIs, see Supplementary Material~A.9---will be chosen numerically so
-that the model matches features of the data (see \cite{McFadden1989} for the general
-method). By modelling PCR and rapid tests in detail, we can translate infection numbers
-in the model into observed infections before matching them to the data. The mechanism is
-depicted in Figure~\ref{fig:model_official_cases} and described in Supplementary
-Material~B.7. Figure~B.9 shows the resulting share of detected cases.
-
-The model is applied to the second and third waves of infections in Germany, covering
-the period September 2020 to May 2021. Figure~\ref{fig:pandemic_drivers_model_fit}
-describes the evolution of the pandemic and of its drivers. The black line in
-Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the black line in
-Figure~\ref{fig:stringency_infectious_contacts} a rescaled Oxford Response Stringency
-Index \cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
-Between mid September and early November, cases increased tenfold. Restrictions were
-somewhat tightened in mid-October and again in early November. New infections remained
-constant throughout November before rising again in December, prompting the most
-stringent lockdown to this date. Schools and daycare centers were closed, so were
-customer-facing businesses except for grocery and drug stores. From the peak of the
-second wave just before Christmas until the trough in mid-February, newly detected cases
-decreased by almost three quarters. The third wave in the spring of 2021 is associated
-with the B.1.1.7 (Alpha) strain, which became dominant in March
-(Figure~\ref{fig:share_b117}). In early March, some NPIs were relaxed. There were many
-changes in details of regulations afterwards, but they did not change the overall
-stringency index.
-
-\begin{figure}[!tp]   % Figure 2
-    \centering
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{Recorded cases: Empirical and simulated}}
-        \label{fig:aggregated_fit}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2_with_seasonality}
-
-        \caption{{Stringency of NPIs and infectious contacts}}
-        \label{fig:stringency_infectious_contacts}
-    \end{subfigure}
-
-    \vskip3ex
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
-
-        \caption{Fraction of B.1.1.7 strain}
-        \label{fig:share_b117}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
-
-        \caption{{Tests and vaccinations}}
-        \label{fig:antigen_tests_vaccinations}
-    \end{subfigure}
-
-    \caption{Evolution of the pandemic, its drivers, and model fit, September 2020 to May 2021}
-    \label{fig:pandemic_drivers_model_fit}
-
-    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
-        Material~A. Age- and region-specific analogues to
-        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material B.10. For
-        legibility reasons, all lines in Figure~\ref{fig:stringency_infectious_contacts}
-        are rolling 7-day averages. The Oxford Response Stringency Index is scaled as $2
-        \cdot (1 -  x / 100)$, so that a value of one refers to the situation at the
-        start of our sample period and zero means that all NPIs included in the index
-        are turned on. The other lines in
-        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
-        of contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~A.5 for separate plots of the three factors by contact type.}
-\end{figure}
-
-Around the turn of the year, the first people were vaccinated with a focus on older age
-groups and medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of
-May, 43\% had received at least one dose of a vaccine. In late 2020, rapid tests started
-to replace regular PCR tests for staff in many medical and nursing facilities. At-home
-tests approved by authorities became available in mid-March. Rapid test centers were
-opened, and one test per person and week was made available free of charge. In several
-states, customers were only allowed to enter certain stores with a recent negative rapid
-test result. These developments are characteristic of many countries: The initial focus
-on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
-acceptance and use of rapid tests.
-
-We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until May 2021. See Supplementary
-Materials~A.2 and B.9 for details. The blue line in Figure~\ref{fig:aggregated_fit}
-shows that our model's predictions are very close to officially recorded cases in the
-aggregate. This is also true for infections by age and geographical region (see
-Supplementary Material~B.10).
-
-The effects of various mechanisms can be disentangled due to the distinct temporal
-variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
-increased hygiene regulations, and seasonality evolved since early September for each of
-the three broad contact networks. For example, a value of 0.75 for the work multiplier
-means that if the environment was the same as in September (levels of infection rates,
-no rapid tests or vaccinations, only the wildtype virus present), infections at the
-workplace would be reduced by 25\%. Two aspects are particularly interesting. First, all
-lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations. Second, the most stringent regulations coincide with
-the period of decreasing infection rates between late December 2020 and mid-February
-2021. The subsequent reversal of the trend is associated with the spread of the B.1.1.7
-variant. During the steep drop in recorded cases during May 2021, for 42\% of the
-population took at least one rapid tests per week, the first-dose vaccination rate rose
-from 28\% to 43\%, and seasonality lowered the relative infectiousness of contacts.
-
-In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections, Figure~\ref{fig:2021_scenarios_broad}
-considers various scenarios. NPIs are always held constant at their values in the
-baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the
-blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the
-three factors (red line), and three scenarios turning each of these factors on
-individually. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
-infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
-values \cite{Shapley2016} to decompose the difference in total infections between the
-scenario without any of the three factors and our main specification.
-
-\begin{figure}[!tp]
-    \centering
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
-        \caption{{Recorded cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_recorded}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
-        \caption{{Total cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_newly_infected}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_area}
-        \caption{Decomposition of the difference between the scenario without any of the
-            three factors and the main scenario in
-            Figure~\ref{fig:2021_scenarios_newly_infected}.}
-        \label{fig:2021_scenarios_decomposition}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
-        \caption{Decomposition of the difference between the scenario without rapid
-            tests and the main scenario in
-            Figure~\ref{fig:2021_scenarios_newly_infected}.}
-        \label{fig:2021_scenarios_decomposition_tests}
-    \end{subfigure}
-
-    \caption{The effect of different interventions on recorded and actual infections}
-    \label{fig:2021_scenarios_broad}
-
-    \floatfoot{\noindent \textit{Note:} The blue line in
-        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
-        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
-        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
-        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and
-        seasonality remain at their January levels. The other scenarios turn each of
-        these three factors on individually. The decompositions in
-        Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained in Appendix~A.10. All lines are rolling 7-day averages.
-    }
-\end{figure}
-
-Until mid-March, there is no visible difference between the different scenarios. Even
-thereafter, the effect of the vaccination campaign is surprisingly small. The Shapley
-value decomposition shows that vaccinations contribute 16\% to the cumulative difference
-between scenarios. Reasons for the low share are the slow start and the focus on older
-individuals who typically have fewer contacts. By the end of our study period, when
-first-dose vaccination rates reached 43\% of the population, the numbers of new cases
-would have started to decline.
-
-Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
-observed and total cases would be reduced by a factor of four if only seasonality
-mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace (this is in line with results in \cite{Gavenciak2021}, which our
-seasonality measure is based on). Nevertheless, we estimate seasonality to be an
-important factor, explaining most of the early changes and 43\% of the cumulative
-difference by the end of May.
-
-A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
-is crucial to differentiate between recorded cases and actual cases. Additional testing
-means that additional infections will be recorded which would otherwise remain
-undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
-may persist for some time. Until late April, recorded cases are higher in the scenario
-with rapid testing alone when compared to the setting where none of the three mechanisms
-are turned on. The effect on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}.
-
-So why is rapid testing so effective? In order to shed more light on this question,
-Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
-scenario without rapid tests and the main specification into the three channels for
-rapid tests. Tests at schools have the smallest effect, which is largely explained by
-schools not operating at full capacity during our period of study and the relatively
-small number of students.\footnote{18\% of our population are in the education sector
-(pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
-come from tests at the workplace. Despite the fact that rapid tests for private reasons
-are phased in only in mid-March, they make up for more than half of the total effect.
-The reason lies in the fact that a substantial share of these tests is driven by an
-elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. This is essentially a form of contact
-tracing, which has been shown to be very effective \cite{Contreras2021,
-Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary Material~B.14
-shows that the same amount of rapid tests administered randomly in the population would
-not have been nearly as effective and that mandatory tests at schools almost offset the
-increased infection risk of opening schools (see Figure~B.13a).
-
-Our analysis has shown that during the transition to high levels of vaccination and
-possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A day of strict lockdown is estimated to have cost around
-50~Euros per capita \cite{Dorn2020b}; retail prices for rapid tests were below one Euro
-in early June 2021. Widespread availability of self-administered tests at low prices are
-likely to play a role for indication-driven testing. Mandatory tests in schools and at
-the workplace are important to screen the entire population. Especially, because
-disadvantaged groups are less likely to be reached by testing campaigns relying on
-voluntary participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups
-have a higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to
-vaccinations, rapid testing programmes allow a much quicker roll-out, making it arguably
-the most effective tool to contain the pandemic in the short run.
-
-\bibliography{bibliography}
-
-\bibliographystyle{Science}
-
-\section*{Acknowledgments}
-
-The authors are grateful for support by the Deutsche Forschungsgemeinschaft (DFG, German
-Research Foundation) under Germany´s Excellence Strategy – EXC 2126/1– 390838866 – and
-through CRC-TR 224 (Projects A02 and C01), by the IZA Institute of Labor Economics, and
-by the Google Cloud CoViD-19 research credits program.
-
-%Here you should list the contents of your Supplementary Materials -- below is an
-%example. You should include a list of Supplementary figures, Tables, and any references
-%that appear only in the SM. Note that the reference numbering continues from the main
-%text to the SM. In the example below, Refs. 4-10 were cited only in the SM.
-
-\section*{Supplementary materials}
-
-Appendix A - Materials and Methods\\
-Appendix B - Supplementary Text\\
-Figs. A.1-A.10, B.1-B.16\\
-Tables A.1-A.6\\
-References \textit{(26-111)}
+The submitted materials include sections on materials and methods and supplementary
+text.
 

### 2021-07-15 task_compile_documents.py

@@ -43,6 +43,22 @@ def task_compile_science_report():
     pass
 
 
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "science" / "cover_letter.tex", *_DEPENDENCIES])
+@pytask.mark.produces(BLD / "cover_letter.pdf")
+def task_compile_cover_letter():
+    pass
+
+
 @pytask.mark.depends_on(BLD / "paper.pdf")
 @pytask.mark.produces(ROOT / "paper.pdf")
 def task_copy_pdf_to_root(depends_on, produces):

### 2021-07-15 src/science/paper.tex
Hardcode bib to paper.

@@
 
-\bibliography{bibliography}
+\begin{thebibliography}{10}
 
-\bibliographystyle{Science}
+\bibitem{Mathieu2021}
+E.~Mathieu, {\it et~al.\/}, {\it Nature Human Behaviour\/}  (2021).
+
+\bibitem{Contreras2021}
+S.~Contreras, {\it et~al.\/}, {\it arXiv\/}  (2021).
+
+\bibitem{Mina2021}
+M.~J. Mina, K.~G. Andersen, {\it Science\/} {\bf 371}, 126 (2021).
+
+\bibitem{Aleta2020}
+A.~Aleta, {\it et~al.\/}, {\it Nature Human Behaviour\/} {\bf 4}, 964 (2020).
+
+\bibitem{Hinch2020}
+R.~Hinch, {\it et~al.\/}, {\it medRxiv\/}  (2020).
+
+\bibitem{Mossong2008}
+J.~Mossong, {\it et~al.\/}, {\it PLoS medicine\/} {\bf 5} (2008).
+
+\bibitem{Hoang2019}
+T.~Hoang, {\it et~al.\/}, {\it Epidemiology\/} {\bf 30}, 723 (2019).
+
+\bibitem{Lessler2021}
+J.~Lessler, {\it et~al.\/}, {\it Science\/} {\bf 372}, 1092 (2021).
+
+\bibitem{Cheng2021}
+Y.~Cheng, {\it et~al.\/}, {\it Science\/}  (2021).
+
+\bibitem{Jones2021}
+T.~C. Jones, {\it et~al.\/}, {\it Science\/}  (2021).
+
+\bibitem{Hunter2021}
+P.~R. Hunter, J.~Brainard, {\it medRxiv\/}  (2021).
+
+\bibitem{LevineTiefenbrun2021}
+M.~Levine-Tiefenbrun, {\it et~al.\/}, {\it Nature medicine\/} {\bf 27}, 790
+  (2021).
+
+\bibitem{Petter2021}
+E.~Petter, {\it et~al.\/}, {\it medRxiv\/}  (2021).
+
+\bibitem{Pritchard2021}
+E.~Pritchard, {\it et~al.\/} .
+
+\bibitem{Bruemmer2021}
+L.~E. Br\"{u}mmer, {\it et~al.\/}, {\it medRxiv\/}  (2021).
+
+\bibitem{Smith2021}
+R.~L. Smith, {\it et~al.\/}, {\it The Journal of Infectious Diseases\/}
+  (2021). Jiab337.
+
+\bibitem{McFadden1989}
+D.~McFadden, {\it Econometrica: Journal of the Econometric Society\/} pp.
+  995--1026 (1989).
+
+\bibitem{Hale2020}
+T.~Hale, {\it et~al.\/}, {\it Blavatnik School of Government\/}  (2020).
+
+\bibitem{Shapley2016}
+L.~S. Shapley, {\it 17. A value for n-person games\/} (Princeton University
+  Press, 2016).
+
+\bibitem{Gavenciak2021}
+T.~Gaven{\v c}iak, {\it et~al.\/}, {\it medRxiv\/}  (2021).
+
+\bibitem{Fetzer2021}
+T.~Fetzer, T.~Graeber, {\it Forthcoming, Proceedings of the National Academy of
+  Sciences\/}  (2021).
+
+\bibitem{Kretzschmar2020}
+M.~E. Kretzschmar, {\it et~al.\/}, {\it The Lancet Public Health\/} {\bf 5},
+  e452 (2020).
+
+\bibitem{Dorn2020b}
+F.~Dorn, {\it et~al.\/}, {\it ifo Schnelldienst\/} {\bf 73}, 29 (2020).
+
+\bibitem{StillmanTonin2021}
+S.~Stillman, M.~Tonin, Communities and testing for {COVID-19} (2021). IZA
+  Discussion Paper No. 14012.
+
+\bibitem{KochInstitut2021a}
+{Robert Koch-Institut}, Corona-monitoring bundesweit (rki-soep-studie), {\it
+  Tech. rep.\/}, {Robert Koch-Institut} (2021).
+
+\end{thebibliography}
 

### 2021-07-15 task_compile_documents.py
Create task to put all submission material for the report in one folder.

@@ -38,7 +38,7 @@ def task_compile_documents():
     ]
 )
 @pytask.mark.depends_on([SRC / "science" / "paper.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "science_report.pdf")
+@pytask.mark.produces(BLD / "science" / "paper.pdf")
 def task_compile_science_report():
     pass
 
@@ -80,3 +80,153 @@ def task_extract_supplementary_material(depends_on, produces):
             produces.as_posix(),
         ]
     )
+
+
+@pytask.mark.depends_on(
+    {
+        "paper": SRC / "science" / "paper.tex",
+        "scicite": SRC / "science" / "scicite.sty",
+        "model_graph_top_left": SRC / "figures" / "model-graph-top-left.pdf",
+        "model_graph_top_right": SRC / "figures" / "model-graph-top-right.pdf",
+        "model_graph_bottom_left": SRC / "figures" / "model-graph-bottom-left.pdf",
+        "model_graph_bottom_right": SRC / "figures" / "model-graph-bottom-right.pdf",
+        "full_new_known_case": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "combined_fit"
+        / "full_new_known_case.pdf",
+        "stringency2_with_seasonality": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "data"
+        / "stringency2_with_seasonality.pdf",
+        "full_share_b117": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "combined_fit"
+        / "full_share_b117.pdf",
+        "full_share_rapid_test_in_last_week_and_vaccinated": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "combined_fit"
+        / "full_share_rapid_test_in_last_week_and_vaccinated.pdf",
+        "full_new_known_case_": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "effect_of_channels_on_pessimistic_scenario"
+        / "full_new_known_case.pdf",
+        "full_newly_infected_": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "effect_of_channels_on_pessimistic_scenario"
+        / "full_newly_infected.pdf",
+        "full_decomposition_channels_area": SRC
+        / "figures"
+        / "results"
+        / "figures"
[…]

### 2021-07-16 src/science/cover_letter.tex
Tiny updates to cover letter.

@@
 \author
-{Janoś Gabler,$^{1, 2}$ Tobias Raabe,$^{3}$ Klara Röhrl,$^{1}$ Hans-Martin von Gaudecker,$^{2,4\ast}$\\
+{Janoś Gabler,$^{1, 2}$\\Tobias Raabe,$^{3}$\\Klara Röhrl,$^{1}$\\Hans-Martin von Gaudecker,$^{2,4\ast}$\\
 \\
@@
 
-Main point: Our main methodological contribution is to calibrate a very detailed and
-flexible agent based model to a long time series of data during the pandemic. Our core
-substantive result is the importance of rapid testing as a supplement to NPIs. Due to
-the slow vaccine roll out world wide, this will remain topical for some time.
+\noindent
+\textbf{Main point}
+\vskip1ex
 
+\noindent
+Our main methodological contribution is to calibrate a very detailed and flexible agent
+based model to a long time series of data during the CoVid-19 pandemic. Our core
+substantive result is the quantitative importance of rapid testing as a supplement to
+NPIs. Due to the slow vaccine roll-out in most countries, this will remain topical for
+some time.
+
+\vskip2ex
+
+\noindent
 There were no submissions to other journals, nor discussions with editors prior to
@@
 
+\vskip2ex
+
+\noindent
 A draft of this report was reviewed by Prof. Jörg Stoye and Prof. Philipp Eisenhauer.
 
+\vskip2ex
+
+\noindent
 A draft of this report has been published as a pre-print on arXiv
@@
 
+\vskip2ex
+
+\noindent
 All data is freely available in linked to in the supplementary material. Only one data
@@
 
+\vskip2ex
+
+\noindent
 The submitted materials include sections on materials and methods and supplementary

### 2021-07-18 task_compile_documents.py
push last changes.

@@ -230,3 +230,21 @@ def task_prepare_submission_material(depends_on, produces):
             produces[name].write_text(text)
         else:
             shutil.copy(depends_on[name], produces[name])
+
+
+
+@pytask.mark.skipif(shutil.which("qpdf") is None, reason="Combination needs qpdf.")
+@pytask.mark.depends_on([BLD / "science" / "paper.pdf", BLD / "supplementary_material.pdf"])
+@pytask.mark.produces(BLD / "science" / "full_paper.pdf")
+def task_create_combined_report(depends_on, produces):
+    subprocess.run(
+        [
+            "qpdf",
+            "--empty",
+            "--pages",
+            depends_on[0].as_posix(),
+            depends_on[1].as_posix(),
+            "--",
+            produces.as_posix(),
+        ]
+    )

### 2021-07-21 src/paper/data/synthetic_population.tex
Version 5: Final long version  (#4)

@@
 % school
-Every child that goes to school is part of three different classes that meet every day on
-weekdays when there are no vacations and no school policies are in place.\footnote{We
+Every child that goes to school is part of a school class. Each school class meets
+three times per weekday, each time with a different set of two teachers, unless
+there are vacations or policies that suspend schools.\footnote{We
 implement vacations on the federal state level.} Each class consists of approximately 23
-students \citep{OECD2013} and two teachers. All students in a class are of the same age
+students \citep{OECD2013}. All students in a class are of the same age
 and live in the same state and mostly also in the same county. In addition, each

### 2021-07-21 src/paper/model/number_of_contacts.tex
Version 5: Final long version  (#4)

@@
 people had contact with an infectious person and the disease has an infection probability
-of one or whether everyone met one infectious person but the disease has an infection
-probability of only 5 percent per contact. SEIR models do not distinguish between the
-number of contacts and the infectiousness of each contact. Instead, they combine the two
-into one parameter that is not invariant to social distancing policies.
+per contact ($\beta$) of one or whether everyone met one infectious person but the
+disease has an infection probability of only 5 percent per contact. SEIR models do not
+distinguish between the number of contacts ($\eta$) and the infectiousness of each
+contact ($\beta$). Instead, they combine the two into one parameter that is not invariant
+to social distancing policies.
 
@@
 contacts of each individual and the effect of mostly policy-invariant infection
-probabilities specific to each contact type. Since not all contacts are equally
-infectious, we distinguish different contact types.
+probabilities specific to each contact type.
 
@@
 are depicted in Figure~\ref{fig:model_contacts_infections} and can be further grouped
-in the categories household, work, education and others.
-
-types of contacts:
+in the categories household, work, education and others:
 
@@
 
-    \item Recurrent work contacts, capturing contacts with coworkers, repeating
-          clients and superiors. Some of these recurrent contacts take place on
-          every workday, others just once per week.
+    \item Recurrent work contacts: These capture contacts with coworkers, repeating
+          clients and superiors. Some of these recurrent contacts take place on every
+          workday, others just once per week. The contacts are assortative in
+          geographical location and age.
 
-    \item Random work contacts: Working adults have contacts with randomly drawn other
-          people, which are assortative in geographical location and age.
+    \item Non recurrent work contacts: Working adults have contacts with randomly drawn
+          other people, which are assortative in geographical location and age.
 
     \item Schools: Each student meets all of his classmates every day. Class sizes are
-    calibrated to be representative for Germany and students have the same age. Schools
-    are closed on weekends and during vacations, which vary by states. School classes
-    also meet six teachers everyday and some of the teachers meet each other.
+    calibrated to be representative for Germany and students have the same age and mostly
+    live in the same county. Schools are closed on weekends and during vacations, which
+    vary by states. School classes also meet six teachers every day and some of the
+    teachers meet each other.
 
-    \item Preschools: Children who are at least three years old and younger than six may
-    attend preschool. Each group of nine children interacts with the same two adults
-    every day. The children in each group are of the same age. The remaining mechanics
-    are similar to schools.
+    \item Preschools: Children who are between three and six years old attend preschool.
+    Each group consists of nine children of mixed ages and two adults who live mostly in
+    the same county. They all meet each other every work day when there are no vacations.
 
     \item Nurseries: Children younger than three years may attend a nursery and interact
-    with one adult. The age of the children varies within groups. The remaining
-    mechanics are similar to schools.
+    with one adult. The age of the children varies within groups but all live in the same
+    county. They all meet each other every work day when there are no vacations.
 
-    \item Random other contacts: Contacts with randomly drawn other
+    \item Non recurrent other contacts: Contacts with randomly drawn other
     people, which are assortative with respect to geographic location and and age group.
@@
 
-    \item Recurrent other contacts representing contacts with friends neighbors or
-        family members who do not live in the same household. Some of these contacts
-        happen daily, others only once per week.
+    \item Recurrent other contacts representing contacts with friends neighbours or
+    family members who do not live in the same household. Some of these contacts happen
+    daily, others only once per week. They are assortative in geographic location and
+    age.
 
@@
 Section~\ref{subsec:data_number_of_contacts}. In particular, we sample the number of
-contacts or group sizes from empirical distributions that sometimes depend on age. It
-would also be possible to use economic or other behavioral models to predict the number
-of contacts.
+contacts or group sizes from empirical distributions. It would also be possible to use
+economic or other behavioral models to predict the number of contacts.
 

### 2021-07-21 src/paper/results/rapid_tests_statistics.tex
Version 5: Final long version  (#4)

@@
 Such a large effect of rapid tests seems to be at odds with the general perception that
-they are not very reliable. However, this is not the case,\comment[id=HM]{I do not
-understand. 80 fals neg for 200 true pos does not seem too reliable? I guess the
-important bit is the reference cat. -- catching 200 pos vs. nothing} which can be seen
-from the number of false negative tests. At its peak at the beginning of May, there are
-more than 80 false negative rapid tests per million inhabitants per day. While this
-number might sound low in isolation, it becomes quite large when set in context: We
-estimate that at at the the same time only 200 cases per million are detected by rapid
-tests. This means that even at the time where discovery of cases via rapid tests is at
-its peak, almost 30\% of people who are infected and make a test are not discovered by
-the test.
+they are not very reliable. However, one has to differentiate between the reliability of
+one test in isolation and the effect imperfect tests can have when employed at a large
+scale. On average our tests have a sensitivity of slightly more than 70\%. This means
+they miss almost 30\% of infections among the tested. Of course perfect tests would have
+an even larger effect but the relevant number to compare is that up to 200 cases per
+million are detected by rapid tests every day which would have otherwise gone undetected.
 
-This shows clearly that the large effect of rapid tests on the infection dynamic
+This clearly shows that the large effect of rapid tests on the infection dynamic
 is not driven by unrealistic assumptions about their sensitivity but rather by the

### 2021-07-21 src/paper/report.tex
Shorten the paper to 2500 words (#5)

@@
 between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that
-potentially cause infections. Contacts occur in up to four networks: Within the
+potentially cause infections. Contacts occur in up to four areas: Within the
 household, at work, at school, or in other settings. Some contacts recur regularly,
@@
 
-The four contact networks are chosen so that the most common NPIs can be modeled in
-great detail. NPIs affect the number of contacts or the risk of transmitting the disease
-upon having physical contact. The effect of different NPIs will generally vary across
-contact types. See Section~\ref{sec:policies} for details on how we model NPIs. In any
-setting, mask mandates or air filters would reduce the probability of infection
-\citep{Lessler2021, Cheng2021}.
+The contact networks are chosen so that the most common NPIs can be modeled in great
+detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
+having physical contact. NPIs' effects on meeting frequencies will vary across contact
+types. Mask mandates or air filters would reduce the probability of infection
+\citep{Lessler2021, Cheng2021}. See Supplementary Material~\ref{sec:policies} for
+details on how we model NPIs.
 
@@
 We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
-certainty whether an individual is infected or not; They require at least one day to be
-processed and are scarce. In contrast, rapid antigen tests yield immediate results.
+certainty whether an individual is infected or not; they are scarce and require at least
+one day to be processed. In contrast, rapid antigen tests yield immediate results.
 Specificity and sensitivity of these tests is set according to data analyzed in
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-complete description.} Examples are contact diaries, administrative records on case
+parameters (see Supplementary Material~\ref{sec:materials_and_methods} for a
+complete description). Examples are contact diaries, administrative records on case
 numbers, virus strains, tests and vaccines as well as surveys on private rapid test
@@
 Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not have been nearly as effective and that
-mandatory tests at schools almost offset the increased infection risk of opening schools
-(see Figure~\ref{fig:school_scenarios}).
+administered randomly in the population would not have been nearly as effective.
+Furthermore, mandating tests at schools almost offsets the increased infection risk of
+opening schools (see Figure~\ref{fig:school_scenarios}).
 
@@
 likely to play a role for indication-driven testing. Mandatory tests in schools and at
-the workplace are important to screen the entire population. Especially, because
-disadvantaged groups are less likely to be reached by testing campaigns relying on
-voluntary participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups
-have a higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to
-vaccinations, rapid testing programmes allow a much quicker roll-out, making it arguably
-the most effective tool to contain the pandemic in the short run.
+the workplace are important to screen the entire population, also because disadvantaged
+groups are less likely to be reached by testing campaigns relying on voluntary
+participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups have a
+higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to vaccinations,
+rapid testing programmes allow a much quicker roll-out, making it arguably the most
+effective tool to contain the pandemic in the short run.

### 2021-07-21 task_compile_documents.py

@@ -38,8 +38,24 @@ def task_compile_documents():
     ]
 )
 @pytask.mark.depends_on([SRC / "science" / "paper.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "science_report.pdf")
+@pytask.mark.produces(BLD / "science" / "paper.pdf")
 def task_compile_science_report():
+    pass
+
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "science" / "cover_letter.tex", *_DEPENDENCIES])
+@pytask.mark.produces(BLD / "cover_letter.pdf")
+def task_compile_cover_letter():
     pass
 
 
@@ -59,7 +75,175 @@ def task_extract_supplementary_material(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "15-78",
+            "12-78",
+            "--",
+            produces.as_posix(),
+        ]
+    )
+
+
+@pytask.mark.depends_on(
+    {
+        "paper": SRC / "science" / "paper.tex",
+        "scicite": SRC / "science" / "scicite.sty",
+        "model_graph_top_left": SRC / "figures" / "model-graph-top-left.pdf",
+        "model_graph_top_right": SRC / "figures" / "model-graph-top-right.pdf",
+        "model_graph_bottom_left": SRC / "figures" / "model-graph-bottom-left.pdf",
+        "model_graph_bottom_right": SRC / "figures" / "model-graph-bottom-right.pdf",
+        "full_new_known_case": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "combined_fit"
+        / "full_new_known_case.pdf",
+        "stringency2_with_seasonality": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "data"
+        / "stringency2_with_seasonality.pdf",
+        "full_share_b117": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "combined_fit"
+        / "full_share_b117.pdf",
+        "full_share_rapid_test_in_last_week_and_vaccinated": SRC
+        / "figures"
+        / "results"
+        / "figures"
+        / "scenario_comparisons"
+        / "combined_fit"
[…]

### 2021-08-05 task_compile_documents.py
Go back to longer version.

@@ -38,24 +38,8 @@ def task_compile_documents():
     ]
 )
 @pytask.mark.depends_on([SRC / "science" / "paper.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "science" / "paper.pdf")
+@pytask.mark.produces(BLD / "science_report.pdf")
 def task_compile_science_report():
-    pass
-
-
-@pytask.mark.latex(
-    [
-        "--pdf",
-        "--interaction=nonstopmode",
-        "--synctex=1",
-        "--cd",
-        "--shell-escape",
-        "-f",
-    ]
-)
-@pytask.mark.depends_on([SRC / "science" / "cover_letter.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "cover_letter.pdf")
-def task_compile_cover_letter():
     pass
 
 
@@ -75,175 +59,7 @@ def task_extract_supplementary_material(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "12-78",
-            "--",
-            produces.as_posix(),
-        ]
-    )
-
-
-@pytask.mark.depends_on(
-    {
-        "paper": SRC / "science" / "paper.tex",
-        "scicite": SRC / "science" / "scicite.sty",
-        "model_graph_top_left": SRC / "figures" / "model-graph-top-left.pdf",
-        "model_graph_top_right": SRC / "figures" / "model-graph-top-right.pdf",
-        "model_graph_bottom_left": SRC / "figures" / "model-graph-bottom-left.pdf",
-        "model_graph_bottom_right": SRC / "figures" / "model-graph-bottom-right.pdf",
-        "full_new_known_case": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "scenario_comparisons"
-        / "combined_fit"
-        / "full_new_known_case.pdf",
-        "stringency2_with_seasonality": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "data"
-        / "stringency2_with_seasonality.pdf",
-        "full_share_b117": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "scenario_comparisons"
-        / "combined_fit"
-        / "full_share_b117.pdf",
-        "full_share_rapid_test_in_last_week_and_vaccinated": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "scenario_comparisons"
-        / "combined_fit"
[…]

### 2021-08-05 src/paper/abstract.tex
Headline, first batch of changes.

@@
 In order to slow the spread of the CoViD-19 pandemic, governments around the world have
-enacted a wide set of policies limiting the transmission of the disease. Initially, these
-focused on non-pharmaceutical interventions; more recently, vaccinations and large-scale
-rapid testing have started to play a major role. The objective of this study is to
-explain the quantitative effects of these policies on determining the course of the
-pandemic, allowing for factors like seasonality or virus strains with different
+enacted a wide set of policies limiting the transmission of the disease. Initially,
+these focused on non-pharmaceutical interventions; more recently, vaccinations and
+large-scale rapid testing have started to play a major role. The objective of this study
+is to explain the quantitative effects of these policies on determining the course of
+the pandemic, allowing for factors like seasonality or virus strains with different
 transmission profiles. To do so, the study develops an agent-based simulation model,
-which is estimated using data for the second and the third wave of the CoViD-19 pandemic
-in Germany. The paper finds that during a period where vaccination rates rose from 5\% to
-40\%, seasonality and rapid testing had the largest effect on reducing infection numbers.
-Frequent large-scale rapid testing should remain part of strategies to contain CoViD-19;
-it can substitute for many non-pharmaceutical interventions that come at a much larger
-cost to individuals, society, and the economy.
+which explicitly takes into account test demand and behavioral changes following
+positive tests. is estimated using data for the second and the third wave of the
+CoViD-19 pandemic in Germany. The paper finds that during a period where vaccination
+rates rose from 5\% to 40\%, seasonality and rapid testing had the largest effect on
+reducing infection numbers. Frequent large-scale rapid testing should remain part of
+strategies to contain CoViD-19; it can substitute for many non-pharmaceutical
+interventions that come at a much larger cost to individuals, society, and the economy.
 

### 2021-08-05 src/paper/report.tex
Headline, first batch of changes.

--- 2021-07-15 src/paper/report.tex
@@
 
+\section{Model description}
+
 At the core of our agent-based model \citep[][we review more literature in Supplementary
-Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
+Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021} are physical contacts between
 heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
@@
 
+\section{Second and third waves of the CoViD-19 pandemic in Germany}
+
 The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
@@
 strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
-    (Delta) reached Germany in April but still accounted for less than 5\% of cases at the
-    end of our simulation period.} In early March, some NPIs were relaxed; e.g.,
+(Delta) was first detected in Germany in April; at the end of our simulation period it
+accounted for less than 5\% of cases.} In early March, some NPIs were relaxed; e.g.,
 hairdressers and home improvement stores were allowed to open again to the public. There
@@
 started to pose additional challenges.
+
+\section{Results}
 
@@
 
-Two of the most contentious NPIs concern schools and mandates to work from home. In many
-countries, schools switched to remote instruction during the first wave, so did Germany.
-After the summer break, they were operating at full capacity with increased hygiene
-measures, before being closed again from mid-December onward. Some states started opening
-them gradually in late February, but operation at normal capacity did not resume until
-the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
-policies regarding schools starting after Easter, at which point rapid tests had become
-widely available. We estimate the realized scenario to have essentially the same effect
-as a situation with closed schools. Under fully opened schools with mandatory tests,
-total infections would have been 6\% higher; this number rises to 20\% without tests.
-These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
-Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
-large negative effects school closures have on children and parents \citep{Luijten2021,
-Melegari2021}---and in particular on those with low socio-economic status---these results
-in conjunction with hindsight bias suggest that opening schools combined with a testing
-strategy would have been beneficial. In other situations, and particular when rapid test
-are not available at scale, trade-offs may well be different.
+\section{Discussion and conclusions}
+
+Having quantified the effects of various mechanisms, we now simulate hypothetical
+scenarios comparing changes in NPIs and testing regimes. Two of the most contentious
+NPIs concern schools and mandates to work from home. In many countries, schools switched
+to remote instruction during the first wave, so did Germany. After the summer break,
+they were operating at full capacity with increased hygiene measures, before being
+closed again from mid-December onward. Some states started opening them gradually in
+late February, but operation at normal capacity did not resume until the beginning of
+June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
+regarding schools starting after Easter, at which point rapid tests had become widely
+available. We estimate the realized scenario to have essentially the same effect as a
+situation with closed schools. Under fully opened schools with mandatory tests, total
+infections would have been 6\% higher; this number rises to 20\% without tests. These
+effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
+Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of
+the large negative effects school closures have on children and parents
+\citep{Luijten2021, Melegari2021}---and in particular on those with low socio-economic
+status---these results in conjunction with hindsight bias suggest that opening schools
+combined with a testing strategy would have been beneficial. In other situations, and
+particular when rapid test are not available at scale, trade-offs may well be different.
 

### 2021-08-05 task_compile_documents.py

@@ -22,36 +22,30 @@ _DEPENDENCIES = [*SRC.rglob("*.tex"), *(BLD / "param_tables").rglob("*.tex")]
     ]
 )
 @pytask.mark.depends_on([SRC / "paper.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "paper.pdf")
+@pytask.mark.produces(BLD / "paper-with-appendix.pdf")
 def task_compile_documents():
     pass
 
-
-@pytask.mark.latex(
-    [
-        "--pdf",
-        "--interaction=nonstopmode",
-        "--synctex=1",
-        "--cd",
-        "--shell-escape",
-        "-f",
-    ]
-)
-@pytask.mark.depends_on([SRC / "science" / "paper.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "science_report.pdf")
-def task_compile_science_report():
-    pass
-
-
-@pytask.mark.depends_on(BLD / "paper.pdf")
+@pytask.mark.skipif(shutil.which("qpdf") is None, reason="Extraction needs qpdf.")
+@pytask.mark.depends_on(BLD / "paper-with-appendix.pdf")
 @pytask.mark.produces(ROOT / "paper.pdf")
-def task_copy_pdf_to_root(depends_on, produces):
-    shutil.copy(depends_on, produces)
+def task_extract_paper(depends_on, produces):
+    subprocess.run(
+        [
+            "qpdf",
+            "--empty",
+            "--pages",
+            depends_on.as_posix(),
+            "1-14",
+            "--",
+            produces.as_posix(),
+        ]
+    )
 
 
 @pytask.mark.skipif(shutil.which("qpdf") is None, reason="Extraction needs qpdf.")
-@pytask.mark.depends_on(BLD / "paper.pdf")
-@pytask.mark.produces(BLD / "supplementary_material.pdf")
+@pytask.mark.depends_on(BLD / "paper-with-appendix.pdf")
+@pytask.mark.produces("supplementary-material.pdf")
 def task_extract_supplementary_material(depends_on, produces):
     subprocess.run(
         [
@@ -59,7 +53,7 @@ def task_extract_supplementary_material(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "15-78",
+            "15-77",
             "--",
             produces.as_posix(),
         ]

### 2021-08-05 src/cover-letter.tex
Add cover letter.

% !TEX program = pdflatex
% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  FUNDAMENTAL SETTINGS  %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[USenglish, 10pt, UBonn_letterhead]{scrlttr2}
% To change the language, set the respective class option to ngerman or UKenglish
% To increase the size of the body font, set the respective class option to 11pt or 12pt

\newkomavar{suppressanniversary}
\newkomavar{suppressbankaccount}
\newkomavar{suppresssigspacing}
\newkomavar{suppressbackaddress}

%%%%%%%%%%%%%%%%%%%%%
%%  SENDER'S DATA  %%
%%%%%%%%%%%%%%%%%%%%%

\IfLanguageName{ngerman}{%
    % German:
    \newcommand{\shortuniversity}{Universität Bonn}
    \newcommand{\higherlevelunit}{Fachbereich Wirtschafts\-wissenschaften}
    \setkomavar{higherlevelunitcaps}{\caps{\higherlevelunit}}
    \setkomavar{institute}{Institut für Angewandte Mikroökonomik}%
    \setkomavar{fromcountry}{Deutschland}
    \setkomavar{place}{Bonn}
    \setkomavar{fromname}{Prof.~Dr.~Hans-Martin v. Gaudecker}
    \setkomavar{signature}{\\[-3ex]\includegraphics[width=8cm]{\string~/admin/unterschrift_lang}\\[0ex]Hans-Martin v. Gaudecker}
}{
    % US or UK English:
    \newcommand{\shortuniversity}{University of Bonn}
    \newcommand{\higherlevelunit}{Department \newline of \newline Economics}
    \setkomavar{higherlevelunitcaps}{\caps{Department} \\ \caps{of} \\ \caps{Economics}}
    \setkomavar{institute}{Institute for Applied~Microeconomics}%
    \setkomavar{fromcountry}{Germany}
    \setkomavar{place}{}
    \setkomavar{fromname}{Prof.~Dr.~Hans-Martin von Gaudecker}
    \setkomavar{signature}{\\[-3ex]\includegraphics[width=8cm]{\string~/admin/unterschrift_lang}\\[0ex]Hans-Martin von Gaudecker}
}
\setkomavar{shortuniversity}{\caps{\shortuniversity}}
\setkomavar{higherlevelunit}{\higherlevelunit}
\setkomavar{shortinstitute}{\caps{IAME}}
\setkomavar{fromaddress}{Lennéstraße~43 \\ 53113~Bonn}
\setkomavar{fromphone}{+49~228~73-9357}
\setkomavar{secretary}{Simone Jost}
\setkomavar{secretaryphone}{+49~228~73-9238}
\setkomavar{fromemail}{\email{hmgaudecker@uni-bonn.de}}
\setkomavar{fromurl}{\urlhttps{www.wiwi.uni-bonn.de/gaudecker}}
\setkomavar{refnum}{}

%%%%%%%%%%%%%%%%%%%
%%  LETTER BODY  %%
%%%%%%%%%%%%%%%%%%%

\begin{document}

\setkomavar{date}{\today}

\setkomavar{subject}{%
    The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs
}

\begin{letter}{%
        To the Editors of\\
        Nature Human Behaviour \\
    }

    \opening{Dear colleague,}

    Many thanks for considering our manuscript for possible publication in
    \textit{Nature Human Behavior}. Our hope is that modelling the evolution of the
    CoViD-19 pandemic and the relative effectiveness of diverse interventions are of
    interest to a broad readership.

    We incorporate detailed NPIs, different tests, and vaccinations in an agent-based
    model. Importantly, we do so in a way that allows to calibrate its parameters based
    on routinely available data. We apply the model to the second and third waves of the
    CoViD-19 pandemic in Germany over a period of 9 months, achieving a good fit to
    actual infection rates.

    Our core substantive result is that rapid testing has been very effective relative
    to many NPIs in a low-medium vaccination environment.

    \closing{Kind regards,}

\end{letter}

\end{document}

### 2021-08-05 task_compile_documents.py

@@ -58,3 +58,18 @@ def task_extract_supplementary_material(depends_on, produces):
             produces.as_posix(),
         ]
     )
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "cover-letter.tex"])
+@pytask.mark.produces(BLD / "cover-letter.pdf")
+def task_compile_documents():
+    pass

### 2021-08-06 task_compile_documents.py
Skip if basics for Uni Bonn letter are not around.

@@ -1,3 +1,4 @@
+import getpass
 import subprocess
 import shutil
 from pathlib import Path
@@ -59,6 +60,7 @@ def task_extract_supplementary_material(depends_on, produces):
         ]
     )
 
+@pytask.mark.skipif(getpass.getuser() != "hmg", reason="Need UniBN letter basics.")
 @pytask.mark.latex(
     [
         "--pdf",
@@ -71,5 +73,5 @@ def task_extract_supplementary_material(depends_on, produces):
 )
 @pytask.mark.depends_on([SRC / "cover-letter.tex"])
 @pytask.mark.produces(BLD / "cover-letter.pdf")
-def task_compile_documents():
+def task_compile_cover_letter():
     pass

### 2021-08-06 src/cover-letter.tex
Version as submitted.

@@
     CoViD-19 pandemic and the relative effectiveness of diverse interventions are of
-    interest to a broad readership.
+    interest to a broad readership. On way to view the manuscript is as taking the same
+    basic approach as the work by Aleta et al.---published a year ago in this
+    journal---and demonstrating its usefulness for modelling and understanding the
+    spread of the virus under different policies.
 
     We incorporate detailed NPIs, different tests, and vaccinations in an agent-based
-    model. Importantly, we do so in a way that allows to calibrate its parameters based
-    on routinely available data. We apply the model to the second and third waves of the
-    CoViD-19 pandemic in Germany over a period of 9 months, achieving a good fit to
-    actual infection rates.
+    model, taking into account behavioural reactions to symptoms or positive test
+    results. Importantly, we do so in a way that allows to calibrate its parameters
+    based on routinely available data. We apply the model to the second and third waves
+    of the CoViD-19 pandemic in Germany over a period of 9 months, achieving a good fit
+    to actual infection rates.
 
     Our core substantive result is that rapid testing has been very effective relative
-    to many NPIs in a low-medium vaccination environment.
+    to many NPIs in a low-medium vaccination environment. This is the relevant scenario
+    for most countries across the world at present; it seems also applicable to countries
+    with high vaccination rates under the conditions of the delta variant.
 

### 2021-08-17 src/paper/literature_review.tex

@@
 moving particles have been replaced by contact networks for households, work and random
-contacts. The OpenABM-Covid-19 model by \citet{Hinch2020} and the model by
+contacts. The OpenABM-Covid-19 model by \citet{Hinch2021} and the model by
 \citet{Aleta2020} are the closest in spirit to ours.
 
-The OpenABM-Covid-19 model by \citet{Hinch2020} also uses detailed contact networks for
+The OpenABM-Covid-19 model by \citet{Hinch2021} also uses detailed contact networks for
 workplaces, schools and households and can evaluate the effect of several NPIs. The main

### 2021-08-22 src/sars-cov-2-spread-policies-pres.tex
Presentation with graphs.

--- 2021-07-21 src/paper/report.tex
@@
-The development of highly effective vaccines holds the promise of containment in the
-medium term. However, most countries find themselves months or years away from reaching
-vaccination levels that would end the pandemic \citep{Mathieu2021}. While
-non-pharmaceutical interventions (NPIs) have allowed some countries to sustain
-equilibria with very low infection numbers \citep{Contreras2021} most have seen large
-fluctuations of infection rates over time. It is therefore of utmost importance to
-employ an effective mix of strategies for containing the virus. However, the effects of
-diverse policies and their interplay with seasonal patterns are not well understood.
+\documentclass[11pt]{beamer}
+% \documentclass[11pt,handout]{beamer}
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage{float, afterpage, rotating, graphicx}
+\usepackage{epstopdf}
+\usepackage{longtable, booktabs, tabularx}
+\usepackage{fancyvrb, moreverb, relsize}
+\usepackage{eurosym, calc}
+\usepackage{amsmath, amssymb, amsfonts, amsthm, bm}
 
-This paper develops a quantitative model incorporating these factors simultaneously. The
-framework allows to combine a wide variety of data and mechanisms, making it useful to
-predict the effects of various interventions. We apply the model to Germany, where new
-infections fell by almost 80\% during May 2021. Our analysis shows that, aside from
-seasonality, frequent and large-scale rapid testing caused the bulk of this decrease,
-which is in line with prior predictions \citep{Mina2021}.
+\usepackage{natbib}
+\bibliographystyle{rusnat}
 
-At the core of our agent-based model \cite{Aleta2020,Hinch2020} (we review more
-literature in Supplementary Material~\ref{sec:literature_review}) are physical contacts
-between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that
-potentially cause infections. Contacts occur in up to four areas: Within the
-household, at work, at school, or in other settings. Some contacts recur regularly,
-others occur at random. Empirical applications can take the population and household
-structure from census data and the network-specific frequencies of contacts from diary
-data measuring contacts before the pandemic (e.g. \cite{Mossong2008,Hoang2019}). Within
-each network, meeting frequencies depend on age and geographical location (see
-Supplementary Material~\ref{subsec:assortativity}).
+\hypersetup{colorlinks=true, linkcolor=black, anchorcolor=black, citecolor=black, filecolor=black, menucolor=black, runcolor=black, urlcolor=black}
 
-The contact networks are chosen so that the most common NPIs can be modeled in great
-detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
-having physical contact. NPIs' effects on meeting frequencies will vary across contact
-types. Mask mandates or air filters would reduce the probability of infection
-\citep{Lessler2021, Cheng2021}. See Supplementary Material~\ref{sec:policies} for
-details on how we model NPIs.
+\setbeamertemplate{footline}[frame number]
+\setbeamertemplate{navigation symbols}{}
+\setbeamertemplate{frametitle}{\centering\vspace{1ex}\insertframetitle\par}
 
-\begin{figure}   % Figure 1
+\begin{document}
+
+\title{The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs}
+
+\author[Janoś Gabler, Tobias Raabe, Klara Röhrl, Hans-Martin von Gaudecker]
+{
+{\bf Janoś Gabler, Tobias Raabe, Klara Röhrl, Hans-Martin von Gaudecker}\\
+{\small Universität Bonn and IZA}\\[1ex]
+}
+
+\begin{frame}
+    \titlepage
+    \note{~}
+\end{frame}
+
+\begin{frame}
     \centering
+    \includegraphics[height=.9\textheight]{figures/model-graph-top-left}
+\end{frame}
 
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-top-left}
-        \caption{{Model for contacts and infections}}
-        \label{fig:model_contacts_infections}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/model-graph-top-right}
-        \vskip5ex
+\end{frame}
 
-        \caption{Disease progression}
-        \label{fig:model_disease_progression}
-    \end{subfigure}
-    \vskip3ex
-    \begin{subfigure}[b]{0.425\textwidth}
+\begin{frame}
     \centering
+    \includegraphics[width=\textwidth]{figures/model-graph-bottom-left}
+\end{frame}
 
-        \includegraphics[width=\textwidth]{figures/model-graph-bottom-left}
-        \caption{{Model for rapid tests}}
-        \label{fig:model_rapid_tests}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/model-graph-bottom-right}
-        \vskip5ex
+\end{frame}
 
-        \caption{{Translating all infections to recorded ones}}
-        \label{fig:model_official_cases}
-    \end{subfigure}
-
-    \caption{Model description}
-    \label{fig:model_description}
-
-    \floatfoot{\noindent \textit{Note:}
-        A~description of the model can be found in Supplementary
-        Material~\ref{sec:supplementary_text}. Figure~\ref{fig:model_contacts_infections}
-        shows the influence of an agent's contacts to other agents on infections.
-        Demographic characteristics set the baseline number of contacts in different
-        networks ($\eta$). The agents may reduce the number of contacts due to NPIs,
-        showing symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may
-        occur when a susceptible agent meets an infectious agent; the probability depends
-        on the type of contact ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs
-        ($\rho_{c,\:t}$). If infected, the infection progresses as depicted in
-        Figure~\ref{fig:model_disease_progression}. If rapid tests are available, agents'
-        demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons trigger a
-        test only for a fraction of individuals depending on an individual compliance
-        parameter; the thresholds for triggering test demand differ across reasons and
-        they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
-        Figure~\ref{fig:model_official_cases} shows the model of translating all
-        infections in the simulated data to age-specific recorded infections. The model
-        uses data on the aggregate share of recorded cases ($\psi$), the share of
-        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and the false
-        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part of
-        the graph is relevant only for periods where rapid tests are available. All
-        parameters are explained in Appendix~\ref{sub:param_tables}.}
-\end{figure}
-
-In the model, a possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression} (similar to \cite{Aleta2020}, more in
-Supplementary Material~\ref{sub:data_course_of_disease}). Whereas susceptibility to the
-viruses differs by age, the infectiousness is independent of age conditioned on the type
-of contact \cite{Jones2021}.
-
-The model includes several other crucial features. New virus strains with different
-infectiousness profiles may appear. Vaccines may become available. During the vaccine
-roll-out, priority may depend on age and occupation; vaccine hesitancy is modeled by
-some individuals refusing vaccination offers. With some probability, vaccinated agents
-become immune and do not transmit the virus \cite{Hunter2021, LevineTiefenbrun2021,
-Petter2021, Pritchard2021}.
-
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
-certainty whether an individual is infected or not; they are scarce and require at least
-one day to be processed. In contrast, rapid antigen tests yield immediate results.
-Specificity and sensitivity of these tests is set according to data analyzed in
-\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
-to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our model for
-rapid test demand. Schools may require staff and students to be tested regularly. Rapid
-tests may be offered by employers to on-site workers. Individuals may demand tests for
-private reasons, which include having plans to meet other people, showing symptoms of
-CoViD-19, and a household member having tested positively for the virus. We endow each
-agent with an individual compliance parameter. This parameter determines whether she
-takes up rapid tests. Positive tests and symptoms lead most individuals to reduce their
-contacts.
-
-Modelling a population of agents according to actual demographic characteristics means
-that we can use a wide array of data to identify and calibrate the model's many
-parameters (see Supplementary Material~\ref{sec:materials_and_methods} for a
-complete description). Examples are contact diaries, administrative records on case
-numbers, virus strains, tests and vaccines as well as surveys on private rapid test
-demand. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, the infection probabilities by
-contact network and the effects of some NPIs, see Supplementary
-Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data (see \cite{McFadden1989} for the general method). By
-modelling PCR and rapid tests in detail, we can translate infection numbers in the model
-into observed infections before matching them to the data. The mechanism is depicted in
-Figure~\ref{fig:model_official_cases} and described in Supplementary
-Material~\ref{sub:testing}. Figure~\ref{fig:share_known_cases_by_age_group} shows the
-resulting share of detected cases.
-
-The model is applied to the second and third waves of infections in Germany, covering
-the period September 2020 to May 2021. Figure~\ref{fig:pandemic_drivers_model_fit}
-describes the evolution of the pandemic and of its drivers. The black line in
-Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the black line in
-Figure~\ref{fig:stringency_infectious_contacts} a rescaled Oxford Response Stringency
-Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
-Between mid September and early November, cases increased tenfold. Restrictions were
-somewhat tightened in mid-October and again in early November. New infections remained
-constant throughout November before rising again in December, prompting the most
-stringent lockdown to this date. Schools and daycare centers were closed, so were
-customer-facing businesses except for grocery and drug stores. From the peak of the
-second wave just before Christmas until the trough in mid-February, newly detected cases
-decreased by almost three quarters. The third wave in the spring of 2021 is associated
-with the B.1.1.7 (Alpha) strain, which became dominant in March
-(Figure~\ref{fig:share_b117}). In early March, some NPIs were relaxed. There were many
-changes in details of regulations afterwards, but they did not change the overall
-stringency index.
-
-\begin{figure}[!tp]   % Figure 2
-    \centering
-
-    \begin{subfigure}[b]{0.475\textwidth}
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{Recorded cases: Empirical and simulated}}
-        \label{fig:aggregated_fit}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
+\end{frame}
+
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/data/stringency2_with_seasonality}
+\end{frame}
 
-        \caption{{Stringency of NPIs and infectious contacts}}
-        \label{fig:stringency_infectious_contacts}
-    \end{subfigure}
+\begin{frame}
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
+\end{frame}
 
-    \vskip3ex
+\begin{frame}
+    \centering
+    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
+\end{frame}
 
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
-
-        \caption{Fraction of B.1.1.7 strain}
-        \label{fig:share_b117}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
-
-        \caption{{Tests and vaccinations}}
-        \label{fig:antigen_tests_vaccinations}
-    \end{subfigure}
-
-    \caption{Evolution of the pandemic, its drivers, and model fit, September 2020 to May 2021}
-    \label{fig:pandemic_drivers_model_fit}
-
-    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
-        Material~\ref{sec:materials_and_methods}. Age- and region-specific analogues to
-        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
-        \ref{subsec:fit_results}. For legibility reasons, all lines in
-        Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
-        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
-        a value of one refers to the situation at the start of our sample period and
-        zero means that all NPIs included in the index are turned on. The other lines in
-        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
-        of contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~\ref{subsec:policies} for separate plots of the three factors by
-        contact type.}
-\end{figure}
-
-Around the turn of the year, the first people were vaccinated with a focus on older age
-groups and medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of
-May, 43\% had received at least one dose of a vaccine. In late 2020, rapid tests started
-to replace regular PCR tests for staff in many medical and nursing facilities. At-home
-tests approved by authorities became available in mid-March. Rapid test centers were
-opened, and one test per person and week was made available free of charge. In several
-states, customers were only allowed to enter certain stores with a recent negative rapid
-test result. These developments are characteristic of many countries: The initial focus
-on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
-acceptance and use of rapid tests.
-
-We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until May 2021. See Supplementary
-Materials~\ref{subsec:synthetic_population} and \ref{sub:initial_conditions} for
-details. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model's
-predictions are very close to officially recorded cases in the aggregate. This is also
-true for infections by age and geographical region (see Supplementary
-Material~\ref{subsec:fit_results}).
-
-The effects of various mechanisms can be disentangled due to the distinct temporal
-variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
-increased hygiene regulations, and seasonality evolved since early September for each of
-the three broad contact networks. For example, a value of 0.75 for the work multiplier
-means that if the environment was the same as in September (levels of infection rates,
-no rapid tests or vaccinations, only the wildtype virus present), infections at the
-workplace would be reduced by 25\%. Two aspects are particularly interesting. First, all
-lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations. Second, the most stringent regulations coincide with
-the period of decreasing infection rates between late December 2020 and mid-February
-2021. The subsequent reversal of the trend is associated with the spread of the B.1.1.7
-variant. During the steep drop in recorded cases during May 2021, for 42\% of the
-population took at least one rapid tests per week, the first-dose vaccination rate rose
-from 28\% to 43\%, and seasonality lowered the relative infectiousness of contacts.
-
-In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections, Figure~\ref{fig:2021_scenarios_broad}
-considers various scenarios. NPIs are always held constant at their values in the
-baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the
-blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the
-three factors (red line), and three scenarios turning each of these factors on
-individually. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
-infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
-values \citep{Shapley2016} to decompose the difference in total infections between the
-scenario without any of the three factors and our main specification.
-
-\begin{figure}[!tp]
-    \centering
-
-    \begin{subfigure}[b]{0.475\textwidth}
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
-        \caption{{Recorded cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_recorded}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
+\end{frame}
+
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
-        \caption{{Total cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_newly_infected}
-    \end{subfigure}
+\end{frame}
 
-    \begin{subfigure}[b]{0.475\textwidth}
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_area}
-        \caption{Decomposition of the difference between the scenario without any of the
-            three factors and the main scenario in
-            Figure~\ref{fig:2021_scenarios_newly_infected}.}
-        \label{fig:2021_scenarios_decomposition}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
+\end{frame}
+
+\begin{frame}
     \centering
     \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_area}
-        \caption{Decomposition of the difference between the scenario without rapid
-            tests and the main scenario in
-            Figure~\ref{fig:2021_scenarios_newly_infected}.}
-        \label{fig:2021_scenarios_decomposition_tests}
-    \end{subfigure}
+\end{frame}
 
-    \caption{The effect of different interventions on recorded and actual infections}
-    \label{fig:2021_scenarios_broad}
+% Print black screen only in presentation mode for finishing up.
+\mode<beamer> {
+    \beamersetaveragebackground{black}
+    \begin{frame}
+        \frametitle{}
+    \end{frame}
 
-    \floatfoot{\noindent \textit{Note:} The blue line in
-        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
-        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
-        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
-        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and
-        seasonality remain at their January levels. The other scenarios turn each of
-        these three factors on individually. The decompositions in
-        Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained in Appendix~\ref{sub:shapley_value}. All lines are rolling 7-day
-        averages.
+    \beamersetaveragebackground{white}
 }
-\end{figure}
 
-Until mid-March, there is no visible difference between the different scenarios. Even
-thereafter, the effect of the vaccination campaign is surprisingly small. The Shapley
-value decomposition shows that vaccinations contribute 16\% to the cumulative difference
-between scenarios. Reasons for the low share are the slow start and the focus on older
-individuals who typically have fewer contacts. By the end of our study period, when
-first-dose vaccination rates reached 43\% of the population, the numbers of new cases
-would have started to decline.
+% \begin{frame}[allowframebreaks]
+%     \frametitle{References}
 
-Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
-observed and total cases would be reduced by a factor of four if only seasonality
-mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace (this is in line with results in \cite{Gavenciak2021}, which our
-seasonality measure is based on). Nevertheless, we estimate seasonality to be an
-important factor, explaining most of the early changes and 43\% of the cumulative
-difference by the end of May.
+%     \bibliography{refs}
+% \end{frame}
 
-A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
-is crucial to differentiate between recorded cases and actual cases. Additional testing
-means that additional infections will be recorded which would otherwise remain
-undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
-may persist for some time. Until late April, recorded cases are higher in the scenario
-with rapid testing alone when compared to the setting where none of the three mechanisms
-are turned on. The effect on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}.
-
-So why is rapid testing so effective? In order to shed more light on this question,
-Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
-scenario without rapid tests and the main specification into the three channels for
-rapid tests. Tests at schools have the smallest effect, which is largely explained by
-schools not operating at full capacity during our period of study and the relatively
-small number of students.\footnote{18\% of our population are in the education sector
-(pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
-come from tests at the workplace. Despite the fact that rapid tests for private reasons
-are phased in only in mid-March, they make up for more than half of the total effect.
-The reason lies in the fact that a substantial share of these tests is driven by an
-elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. This is essentially a form of contact
-tracing, which has been shown to be very effective \cite{Contreras2021,
-Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
-Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not have been nearly as effective.
-Furthermore, mandating tests at schools almost offsets the increased infection risk of
-opening schools (see Figure~\ref{fig:school_scenarios}).
-
-Our analysis has shown that during the transition to high levels of vaccination and
-possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A day of strict lockdown is estimated to have cost around
-50~Euros per capita \cite{Dorn2020b}; retail prices for rapid tests were below one Euro
-in early June 2021. Widespread availability of self-administered tests at low prices are
-likely to play a role for indication-driven testing. Mandatory tests in schools and at
-the workplace are important to screen the entire population, also because disadvantaged
-groups are less likely to be reached by testing campaigns relying on voluntary
-participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups have a
-higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to vaccinations,
-rapid testing programmes allow a much quicker roll-out, making it arguably the most
-effective tool to contain the pandemic in the short run.
+\end{document}

### 2021-08-22 task_compile_documents.py

@@ -23,7 +23,23 @@ _DEPENDENCIES = [*SRC.rglob("*.tex"), *(BLD / "param_tables").rglob("*.tex")]
 )
 @pytask.mark.depends_on([SRC / "paper.tex", *_DEPENDENCIES])
 @pytask.mark.produces(BLD / "paper.pdf")
-def task_compile_documents():
+def task_compile_paper():
+    pass
+
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "sars-cov-2-spread-policies-pres.tex", *_DEPENDENCIES])
+@pytask.mark.produces(BLD / "sars-cov-2-spread-policies-pres.pdf")
+def task_compile_pres():
     pass
 
 
@@ -163,20 +179,23 @@ def task_extract_supplementary_material(depends_on, produces):
         / "science"
         / "figures"
         / "model-graph-bottom-right.pdf",
-        "full_new_known_case": BLD / "science"
+        "full_new_known_case": BLD
+        / "science"
         / "figures"
         / "results"
         / "figures"
         / "scenario_comparisons"
         / "combined_fit"
         / "full_new_known_case.pdf",
-        "stringency2_with_seasonality": BLD / "science"
+        "stringency2_with_seasonality": BLD
+        / "science"
         / "figures"
         / "results"
         / "figures"
         / "data"
         / "stringency2_with_seasonality.pdf",
-        "full_share_b117": BLD / "science"
+        "full_share_b117": BLD
+        / "science"
         / "figures"
         / "results"
         / "figures"
@@ -232,9 +251,10 @@ def task_prepare_submission_material(depends_on, produces):
             shutil.copy(depends_on[name], produces[name])
 
 
-
 @pytask.mark.skipif(shutil.which("qpdf") is None, reason="Combination needs qpdf.")
-@pytask.mark.depends_on([BLD / "science" / "paper.pdf", BLD / "supplementary_material.pdf"])
+@pytask.mark.depends_on(
+    [BLD / "science" / "paper.pdf", BLD / "supplementary_material.pdf"]
+)
 @pytask.mark.produces(BLD / "science" / "full_paper.pdf")
 def task_create_combined_report(depends_on, produces):
     subprocess.run(

### 2021-08-23 src/cover-letter.tex
Version for Scientific Reports.

@@
         To the Editors of\\
-        Nature Human Behaviour \\
+        Scientific Reports \\
     }
@@
     Many thanks for considering our manuscript for possible publication in
-    \textit{Nature Human Behavior}. Our hope is that modelling the evolution of the
+    \textit{scientific reports}. Our hope is that modelling the evolution of the
     CoViD-19 pandemic and the relative effectiveness of diverse interventions are of
-    interest to a broad readership. On way to view the manuscript is as taking the same
-    basic approach as the work by Aleta et al.---published a year ago in this
-    journal---and demonstrating its usefulness for modelling and understanding the
-    spread of the virus under different policies.
+    interest to a broad readership. We make both methodological and substantive
+    contributions.
 

### 2021-08-27 src/paper/report.tex
Fix missed Hinch reference.

@@
 framework allows to combine a wide variety of data and mechanisms in a timely fashion,
-making it useful to predict the effects of various interventions. We apply the model to
-Germany, where new infections fell by almost 80\% during May 2021. Our analysis
-shows that, aside from seasonality, frequent and large-scale rapid testing caused the
-bulk of this decrease, which is in line with prior predictions \citep{Mina2021}. We
-conclude that it should have a large role for at least as long as vaccinations have not
-been offered to an entire population.
+making it useful to predict the effects of various interventions. Behavioral reactions
+to symptoms or positive tests are explicitly taken into account. We apply the model to
+Germany, where new infections fell by almost 80\% during May 2021. Our analysis shows
+that, aside from seasonality, frequent and large-scale rapid testing caused the bulk of
+this decrease, which is in line with prior predictions \citep{Mina2021}. We conclude
+that it should have a large role for at least as long as vaccinations have not been
+offered to an entire population.
 
@@
 At the core of our agent-based model \citep[][we review more literature in Supplementary
-Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021} are physical contacts between
+Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021a} are physical contacts between
 heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between

### 2021-08-27 src/paper/literature_review.tex
Go back to longer version, nature submission, update references. (#6)

@@
 
-A commonly used model class in epidemiology are agent-based simulation
-models. In a prototypical agent-based simulation model, individuals are
-simulated as moving particles. Infections take place when two particles come closer than
-a certain contact radius (e.g. \citet{Silva2020} and \citet{Cuevas2020}). While the
-simulation approach makes it easy to incorporate heterogeneity in disease progression, it
-is hard to incorporate heterogeneity in meeting patterns. Moreover, policies are modeled
-as changes in the contact radius or momentum equation of the particles. The translation
-from real policies to corresponding model parameters is a hard task.
+A commonly used model class in epidemiology are agent-based simulation models.
+In a prototypical agent-based simulation model, individuals are simulated as
+moving particles. Infections take place when two particles come closer than a
+certain contact radius (e.g. \citet{Silva2020} and \citet{Cuevas2020}). While
+the simulation approach makes it easy to incorporate heterogeneity in disease
+progression, it is hard to incorporate heterogeneity in meeting patterns.
+Moreover, policies are modeled as changes in the contact radius or momentum
+equation of the particles. The translation from real policies to corresponding
+model parameters is a hard task.
 
-These shortcomings have motivated variations of agent-based simulation models where
-moving particles have been replaced by contact networks for households, work and random
-contacts. The OpenABM-Covid-19 model by \citet{Hinch2021} and the model by
-\citet{Aleta2020} are the closest in spirit to ours.
+These shortcomings have motivated variations of agent-based simulation models
+where moving particles have been replaced by contact networks for households,
+work and random contacts. The OpenABM-Covid-19 model by \citet{Hinch2021a} and
+the model by \citet{Aleta2020} are the closest in spirit to ours.
 
-The OpenABM-Covid-19 model by \citet{Hinch2021} also uses detailed contact networks for
-workplaces, schools and households and can evaluate the effect of several NPIs. The main
-focus of their application are contact tracing policies. Recently they have also added
-support for multiple virus strains and vaccinations.
+The OpenABM-Covid-19 model by \citet{Hinch2021a} also uses detailed contact
+networks for workplaces, schools and households and can evaluate the effect of
+several NPIs. The main focus of their application are contact tracing policies
+\citep{Abueg2021}. Recently they have also added support for multiple virus
+strains and vaccinations.
 
-\citet{Aleta2020} develop an agent-based simulation model with a very high geographical
-resolution by estimating contact networks from fine grained mobility data for the
-Boston metropolitan area. They use this model to show how NPIs, contact tracing and PCR
-testing can influence the infection dynamics. However, they do not calibrate their model
-to match actual infection numbers which makes it more suitable to explore the general
-mechanics of different disease mitigation measures than for their quantitative
-evaluation.
+\citet{Aleta2020} develop an agent-based simulation model with a very high
+geographical resolution by estimating contact networks from fine grained
+mobility data for the Boston metropolitan area. They use this model to show how
+NPIs, contact tracing and PCR testing can influence the infection dynamics.
+However, they do not calibrate their model to match actual infection numbers
+which makes it more suitable to explore the general mechanics of different
+disease mitigation measures than for their quantitative evaluation.
 

### 2021-08-27 task_compile_documents.py

@@ -1,3 +1,4 @@
+import getpass
 import subprocess
 import shutil
 from pathlib import Path
@@ -268,3 +269,37 @@ def task_create_combined_report(depends_on, produces):
             produces.as_posix(),
         ]
     )
+
+
+@pytask.mark.skipif(shutil.which("qpdf") is None, reason="Extraction needs qpdf.")
+@pytask.mark.depends_on(BLD / "paper.pdf")
+@pytask.mark.produces(ROOT / "paper-no-appendix.pdf")
+def task_extract_paper(depends_on, produces):
+    subprocess.run(
+        [
+            "qpdf",
+            "--empty",
+            "--pages",
+            depends_on.as_posix(),
+            "1-14",
+            "--",
+            produces.as_posix(),
+        ]
+    )
+
+
+@pytask.mark.skipif(getpass.getuser() != "hmg", reason="Need UniBN letter basics.")
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "cover-letter.tex"])
+@pytask.mark.produces(BLD / "cover-letter.pdf")
+def task_compile_cover_letter():
+    pass

### 2021-08-27 src/paper/results/scenarios.tex
Various fixes that got lost in the previous PR.

@@
-\subsection{Scenarios} \comment[id=K]{Add parameter notation when writing this.}
+\subsection{Scenarios}
 \label{subsec:appendix_scenarios}
@@
 rapid testing policies.
+
+Two of the most contentious NPIs concern schools and mandates to work from home. In many
+countries, schools switched to remote instruction during the first wave, so did Germany.
+After the summer break, they were operating at full capacity with increased hygiene
+measures, before being closed again from mid-December onward. Some states started opening
+them gradually in late February, but operation at normal capacity did not resume until
+the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
+policies regarding schools starting after Easter, at which point rapid tests had become
+widely available. We estimate the realized scenario to have essentially the same effect
+as a situation with closed schools. Under fully opened schools with mandatory tests,
+total infections would have been 6\% higher; this number rises to 20\% without tests.
+These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
+Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
+large negative effects school closures have on children and parents \citep{Luijten2021,
+Melegari2021}---and in particular on those with low socio-economic status---these results
+in conjunction with hindsight bias suggest that opening schools combined with a testing
+strategy would have been beneficial. In other situations, and particular when rapid test
+are not available at scale, trade-offs may well be different.
+
+\begin{figure}[!tp]
+    \centering
+
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \caption{{Effects of different schooling scenarios}}
+        \label{fig:school_scenarios}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.425\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+        \caption{{Effects of different work scenarios}}
+        \label{fig:workplace_scenarios}
+    \end{subfigure}
+    \vskip3ex
+
+    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
+    \label{fig:school_workplace_scenarios}
+
+    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
+        scenario; they are the same as in
+        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
+        because there were no capacity constraints for rapid tests afterwards.
+        For legibility reasons, all lines are rolling 7-day averages.}
+
+\end{figure}
+
+Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
+receiving tests, testing at the workplace has larger effects than mandating employees to
+work from home. Whether the share of workers working at the usual workplace is reduced
+or increased by ten percent changes infection rates by 2.5\% or less in either
+direction. Making testing mandatory twice a week---assuming independent compliance by
+employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
+rapid tests offers by employers to the level of March would have increased infections by
+13\%.
 

### 2021-09-03 src/science/paper.tex

--- 2021-07-15 src/science/paper.tex
@@
 
-The four contact networks are chosen so that the most common NPIs can be modeled in
-great detail. NPIs affect the number of contacts or the risk of transmitting the disease
-upon having physical contact. The effect of different NPIs will generally vary across
-contact types. See Section~B.4 for details on how we model NPIs. In any
-setting, mask mandates or air filters would reduce the probability of infection
-\cite{Lessler2021, Cheng2021}.
+The contact networks are chosen so that the most common NPIs can be modeled in great
+detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
+having physical contact. NPIs' effects on meeting frequencies will vary across contact
+types. Mask mandates or air filters would reduce the probability of infection
+\cite{Lessler2021, Cheng2021}. See Supplementary Material~B.4 for
+details on how we model NPIs.
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~A for a complete description.} Examples
-are contact diaries, administrative records on case numbers, virus strains, tests and
+parameters (see Supplementary Material~A for a complete description). Examples are
+contact diaries, administrative records on case numbers, virus strains, tests and
 vaccines as well as surveys on private rapid test demand. Other studies' estimates of
@@
 shows that the same amount of rapid tests administered randomly in the population would
-not have been nearly as effective and that mandatory tests at schools almost offset the
-increased infection risk of opening schools (see Figure~B.13a).
+not have been nearly as effective. Furthermore, mandating tests at schools almost
+offsets the increased infection risk of opening schools (see Figure~B.13a).
 
@@
 likely to play a role for indication-driven testing. Mandatory tests in schools and at
-the workplace are important to screen the entire population. Especially, because
-disadvantaged groups are less likely to be reached by testing campaigns relying on
-voluntary participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups
-have a higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to
-vaccinations, rapid testing programmes allow a much quicker roll-out, making it arguably
-the most effective tool to contain the pandemic in the short run.
+the workplace are important to screen the entire population, also because disadvantaged
+groups are less likely to be reached by testing campaigns relying on voluntary
+participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups have a
+higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to vaccinations,
+rapid testing programmes allow a much quicker roll-out, making it arguably the most
+effective tool to contain the pandemic in the short run.
+
+% \bibliography{bibliography}
+% \bibliographystyle{Science}
 

### 2021-09-03 src/cover-letter.tex
Re-submitted version.

--- 2021-08-05 src/cover-letter.tex
@@
     Many thanks for considering our manuscript for possible publication in
-    \textit{Nature Human Behavior}. Our hope is that modelling the evolution of the
+    \textit{scientific reports}. Our hope is that modelling the evolution of the
     CoViD-19 pandemic and the relative effectiveness of diverse interventions are of
-    interest to a broad readership.
+    interest to a broad readership. We make both methodological and substantive
+    contributions.\\[0.5ex]
+    We incorporate detailed NPIs, different tests, and vaccinations in an agent-based
+    model, taking into account behavioural reactions to symptoms or positive test
+    results. Importantly, we do so in a way that allows to calibrate its parameters
+    based on routinely available data. We apply the model to the second and third waves
+    of the CoViD-19 pandemic in Germany over a period of 9 months, achieving a good fit
+    to actual infection rates.\\[0.5ex]
+    Our core substantive result is that rapid testing has been very effective relative
+    to many NPIs in a low-medium vaccination environment. This is the relevant scenario
+    for most countries across the world at present; it seems also applicable to countries
+    with high vaccination rates under the conditions of the delta variant.\\[0.5ex]
+    A prior submission required clarification regarding the pre-print servers we had
+    uploaded the paper to; see the e-mail conversation with Dr.{} Sabrya Carim. These are the links to the 
+    pre-prints (all exactly the same paper):\\[-2ex]
+    \begin{itemize}
+        \item \href{https://arxiv.org/abs/2106.11129}{arXiv.org}\\[-2ex]
+        \item \href{https://www.iza.org/publications/dp/14510/the-effectiveness-of-strategies-to-contain-sars-cov-2-testing-vaccinations-and-npis}{IZA}\\[-2ex]
+        \item \href{https://www.econtribute.de/RePEc/ajk/ajkdps/ECONtribute_100_2021.pdf}{ECONtribute}\\[-2ex]
+        \item \href{https://www.crctr224.de/en/research-output/discussion-papers/archive/2021/DP302}{Collaborative Research Center TR-224 Bonn/Mannheim}\\[2ex]
+    \end{itemize}
 
-    We incorporate detailed NPIs, different tests, and vaccinations in an agent-based
-    model. Importantly, we do so in a way that allows to calibrate its parameters based
-    on routinely available data. We apply the model to the second and third waves of the
-    CoViD-19 pandemic in Germany over a period of 9 months, achieving a good fit to
-    actual infection rates.
-
-    Our core substantive result is that rapid testing has been very effective relative
-    to many NPIs in a low-medium vaccination environment.
-
-    \closing{Kind regards,}
+    \closing{Many thanks for your consideration and please let me know should any further questions arise. Kind regards,}
 

### 2021-09-03 src/paper/report.tex
Changes requested by manuscript checker.

@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination levels that would end the pandemic or even
-protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
-most have seen large fluctuations of infection rates over time. Containment measures have
-become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
-tracing. Neither these policies' effects nor the influence of seasonal patterns or of
-more infectious virus strains are well understood in quantitative terms.
+containment in the medium term. However, most countries find themselves many
+months---and often years---away from reaching vaccination levels that would end the
+pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
+of utmost importance to employ an effective mix of strategies for containing the virus.
+The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
+to reduce contacts between individuals. While this has allowed some countries to sustain
+equilibria with very low infection numbers (see \citet{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate policies)
+most have seen large fluctuations of infection rates over time. Containment measures
+have become increasingly diverse and now include rapid testing, more nuanced NPIs, and
+contact tracing. Neither these policies' effects nor the influence of seasonal patterns
+or of more infectious virus strains are well understood in quantitative terms.
 
@@
 at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests
-yield immediate results. Specificity and
-sensitivity of these tests is set
-according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the onset of infectiousness. After a phase-in period, all
-tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by
+contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
+these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
+sensitivity depends on the timing of the test relative to the onset of infectiousness.
+After a phase-in period, all tests that are demanded will be performed.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require staff and students to be tested regularly. Rapid tests may be offered by
 employers to on-site workers. Individuals may demand tests for private reasons, which
@@
 member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-    contacts; this is why tests impact the actual contacts in
-    Figure~\ref{fig:model_description}.}
+compliance parameter. This parameter determines whether she takes up rapid tests
+(positive test results or symptoms lead most individuals to reduce their contacts; this
+is why tests impact the actual contacts in Figure~\ref{fig:model_description}).
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
-different contact types and their assortativity by age group. Mobility data is used to
-model the evolution of work contacts. School and daycare policies can be incorporated
-directly from official directives. Administrative records on the number of tests,
-vaccinations by age and region, and the prevalence of virus strains are generally
-available. Surveys may ask about test offers, propensities to take them up, and past
-tests. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, these include infection probabilities
-by contact network and the effects of some NPIs, see Supplementary
+parameters (see Supplementary Material~\ref{sec:materials_and_methods} for a complete
+description). Contact diaries yield pre-pandemic distributions of contacts for different
+contact types and their assortativity by age group. Mobility data is used to model the
+evolution of work contacts. School and daycare policies can be incorporated directly
+from official directives. Administrative records on the number of tests, vaccinations by
+age and region, and the prevalence of virus strains are generally available. Surveys may
+ask about test offers, propensities to take them up, and past tests. Other studies'
+estimates of the seasonality of infections can be incorporated directly. The remaining
+parameters---most notably, these include infection probabilities by contact network and
+the effects of some NPIs, see Supplementary
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
@@
 quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
-strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
+strain, which became dominant in March (Figure~\ref{fig:share_b117}, note that B.1.617.2
 (Delta) was first detected in Germany in April; at the end of our simulation period it
-accounted for less than 5\% of cases.} In early March, some NPIs were relaxed; e.g.,
+accounted for less than 5\% of cases.). In early March, some NPIs were relaxed; e.g.,
 hairdressers and home improvement stores were allowed to open again to the public. There
@@
 schools not operating at full capacity during our period of study and the relatively
-small number of students.\footnote{18\% of our population are in the education sector
-    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
-come from tests at the workplace. Despite the fact that rapid tests for private reasons
-are phased in only in mid-March, they make up for more than half of the total effect.
-The reason lies in the fact that a substantial share of these tests is driven by an
-elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. The latter is essentially a form of contact
+small number of students (18\% of our population are in the education sector (pupils,
+teachers, etc.); 46\% are workers outside the education sector). Almost 40\% come from
+tests at the workplace. Despite the fact that rapid tests for private reasons are phased
+in only in mid-March, they make up for more than half of the total effect. The reason
+lies in the fact that a substantial share of these tests is driven by an elevated
+probability to carry the virus, i.e., showing symptoms of CoViD-19 or following up on a
+positive test of a household member. The latter is essentially a form of contact
 tracing, which has been shown to be very effective \citep{Contreras2021,

### 2021-09-29 src/paper/data/rapid_test_demand.tex
Polishing.
Add additional references by HiWis to the rapid test demand section.

@@
     \item someone plans to have work contacts
-    \item someone is an employee of an educational facility or a school pupil
+    \item someone is an employee or student of an educational facility
     \item a household member has tested positive or developed symptoms
@@
 
-In addition, there are some surveys that allow us to trace the expansion of employers who
-offer tests to their employees ($\pi_{w,\:s,\:t}$). Mid march, 20\% of employers offered
-tests to their employees \citep{DIHK2021}. In the second half of March, 23\% of employees
-reported being offered weekly rapid tests by their employer \citep{Ahlers2021}. This
-share increased to 61\% until the first days of April \citep{BMWI2021, IZA2021}. Until
-mid April 72\% of workers were expected to receive a weekly test offer \citep{BMWI2021,
-IZA2021}. However, according to surveys conducted in mid April \citep{Betsch2021}, less
-than two thirds of individuals with work contacts receive a test offer. Starting on April
-19th employers were required by law to provide two weekly tests to their employees
-\citep{Bundesanzeiger2021}. We assume that compliance is incomplete and only 80\% of
-employers actually offer tests. We interpolate between these points linearly, arriving at
-the blue line in Figure~\ref{fig:rapid_test_demand}. In addition, we increase the
-frequency of testing ($\theta_{t,\:work}$) from weekly to twice weekly during April.
+In addition, there are some surveys that allow us to trace the expansion of employers
+who offer tests to their employees ($\pi_{w,\:s,\:t}$). Mid march, 20\% of employers
+offered tests to their employees \citep{DIHK2021}. In the second half of March, 23\% of
+employees reported being offered weekly rapid tests by their employer
+\citep{Ahlers2021}\footnote{\cite{IZA2021b} reports a somewhat higher share of 35\%
+around March 20}. This share increased to 61\% until the first days of April
+\citep{IZA2021c, IZA2021}. Until mid April 72\% of workers were expected to receive a
+weekly test offer \citep{IZA2021c, IZA2021}. However, according to surveys conducted in
+mid April \citep{Betsch2021}, less than two thirds of individuals with work contacts
+receive a test offer.\footnote{\cite{IZA2021d} report 80\% in mid April which is about
+12\% points less than what we arrive at.} Starting on April 19th employers were required
+by law to provide two weekly tests to their employees \citep{Bundesanzeiger2021}.
+However, compliance was very incomplete and only reached 80\% \citep{IAB2021b,
+Ahlers2021b}. We interpolate between these points linearly, arriving at
+Figure~\ref{fig:work_rapid_test_offers}. In addition, we increase the frequency of
+testing ($\theta_{t,\:work}$) from weekly to twice weekly during April.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/share_of_workers_with_rapid_test_offer_at_work}
+    \caption{Share of Individuals Being Offered a Rapid Test at Work.}
+    \floatfoot{\noindent \textit{Note:} Employees were never subject to mandatory tests.
+    According to surveys only 60\% of workers regularly used rapid test offers by their
+    employers (\cite{Betsch2021}, 20th/21st of April).}
+\end{figure}
+\FloatBarrier
 

### 2021-09-29 src/paper/data/rapid_test_demand.tex
Add educ rapid test references from HiWis.

@@
 by March 1st 30\% of them ($\pi_{teacher,\:t}$) are tested weekly
-($\theta_{before\:Easter,\:educ} = 7$). The share increases to 90\% for the week before
+($\theta_{before\:Easter,\:educ} = 7$) \citep{schoolNRWd, schoolBWc,
+schoolBYb,SueddeutscheZeitung2021b,Zeit2021}. The share increases to 90\% for the week before
 Easter. At that time both Bavaria \citep{STMGP2021} and Baden-Württemberg

### 2021-09-29 src/paper/report.tex
Add educ rapid test references from HiWis.

--- 2021-08-27 src/paper/report.tex
@@
 at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests
-yield immediate results. Specificity and
-sensitivity of these tests is set
-according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
-timing of the test relative to the onset of infectiousness. After a phase-in period, all
-tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by
+contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
+these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021,
+Toptan2020}; sensitivity depends on the timing of the test relative to the onset of
+infectiousness. After a phase-in period, all tests that are demanded will be performed.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require staff and students to be tested regularly. Rapid tests may be offered by
 employers to on-site workers. Individuals may demand tests for private reasons, which

### 2021-09-29 src/paper/results/rapid_tests_statistics.tex
Add NRW data on number of performed citizens' rapid tests.

@@
 
-We start with the share of the population doing a rapid test and receiving a positive rapid
-test over time by the channel through which the test
-was demanded in Figures~\ref{fig:rapid_test_demand_by_channel},
-\ref{fig:pos_rapid_tests_by_channel}, respectively. Overall, the share of the population
-getting a rapid test on a given day increases from 2\% in mid March to over 10\% by May.
-The work rapid tests are a little ragged because of public holidays. For education rapid
-tests both vacations (first half of April) as well as the opening of schools in May are
-very visible in the rapid test demand. Overall, work tests make up the largest fraction
-of rapid tests. The image is very similar for the share of positive tests, except that
-the overall number of positive tests starts decreasing in May as rapid test expansion
-comes to a halt and cases fall, especially the positive share of private rapid tests
-falls as less and less individuals are triggered to seek a rapid test because of a risk
-contact in their household.
+We start with the share of the population doing a rapid test and receiving a positive rapid test
+over time by the channel through which the test was demanded in
+Figures~\ref{fig:rapid_test_demand_by_channel}, \ref{fig:pos_rapid_tests_by_channel}, respectively.
+Overall, the share of the population getting a rapid test on a given day increases from 2\% in mid
+March to over 10\% by May. The work rapid tests are a little ragged because of public holidays. For
+education rapid tests both vacations (first half of April) as well as the opening of schools in May
+are very visible in the rapid test demand. Overall, work tests make up the largest fraction of rapid
+tests. The image is very similar for the share of positive tests, except that the overall number of
+positive tests starts decreasing in May as rapid test expansion comes to a halt and cases fall,
+especially the positive share of private rapid tests falls as less and less individuals are
+triggered to seek a rapid test because of a risk contact in their household. The share of private
+rapid tests aligns with numbers reported by states on the demand for the free citizen tests
+\citep{NRW_rapid_tests}.
 
@@
 
-A similar picture arises, when looking at the false positive rate,
-i.e. the share of positive tests that go to people who are not infected.
-Figure~\ref{fig:rapid_tests_false_positive_rate} shows that the false
-positive rate is very high. On average 60\% to 93\% of positive tests are
-received by individuals that are not infected. The false positive rate increases over
-time. This is due to the low prevalence of infections in the population, which falls
-over time. Again, private rapid tests are an exception with a much lower false
-positive rate because those tests are primarily demanded when there is a high likelihood
-of being infected. The false negative rate of 0.2\% looks very low. As discussed
-above this is deceiving and just a mechanical consequence of a very low prevalence
-of the disease and the many rapid tests done by non-infected people.
+A similar picture arises, when looking at the false positive rate, i.e. the share of
+positive tests that go to people who are not infected.
+Figure~\ref{fig:rapid_tests_false_positive_rate} shows that the false positive rate is
+very high. On average 60\% to 93\% of positive tests are received by individuals that
+are not infected which is broadly in line with rates later reported in the German media
+\citep{SueddeutscheZeitung2021c}. The false positive rate increases over time. This is
+due to the low prevalence of infections in the population, which falls over time. Again,
+private rapid tests are an exception with a much lower false positive rate because those
+tests are primarily demanded when there is a high likelihood of being infected. The
+false negative rate of 0.2\% looks very low. As discussed above this is deceiving and
+just a mechanical consequence of a very low prevalence of the disease and the many rapid
+tests done by non-infected people.
 

### 2021-09-29 src/paper/model/testing.tex
Add some last references.

@@
 detected cases that would have happened without rapid tests are very likely to be small 
-compared to the changes caused by rapid tests.
+compared to the changes caused by rapid tests. In addition, the number of PCR tests per 
+1000 remains in the same 1.5 to 2.5 window without a visible trend over our estimation 
+period \citep{owid_n_pcr_tests} which is in line with less than 0.08\% of the 
+population receiving a positive rapid test on any given day in our model.
 
@@
 is much higher (88\%) but still lower than during the remaining infectious period (92\%).
-After infectiousness stops, the sensitivity drops to 50 \%.
+After infectiousness stops, the sensitivity drops to 50\%.
 
@@
 administered by medical personnel. While there were concerns that self administered
-tests are less reliable, a recent study has found no basis for this
-concern \citep{Lindner2020}.
+tests are less reliable, a recent study has found no basis for this concern
+\citep{Lindner2020}.
 

### 2021-09-29 src/paper/report.tex
Adjust lockdown costs.

@@
 at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
+estimated to have cost around 20-30~Euros per capita \citep{Wollmershauser2021}; retail
 prices for rapid tests were below one Euro in early June 2021 and below five Euros for

### 2022-03-04 src/paper/report.tex
Add reference for susceptibility by age to main text.

@@
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
-possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
-differentiate between an initial period of infection without being infectious or showing
-symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
-of transitioning between these states depend on age; their duration is random and
-calibrated to medical literature (for a detailed description see Supplementary
-Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \citep{Jones2021}.
+In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
+age \citep{Davies2020}. A possible infection progresses as shown in
+Figure~\ref{fig:model_disease_progression}. We differentiate between an initial
+period of infection without being infectious or showing symptoms, being
+infectious (presymptomatic or asymptomatic), showing symptoms, requiring
+intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The
+probabilities of transitioning between these states depend on age; their
+duration is random and calibrated to medical literature (for a detailed
+description see Supplementary Material~\ref{sub:data_course_of_disease}).
+Conditional on the type of contact, infectiousness is independent of age
+\citep{Jones2021}.
 

### 2022-03-08 src/paper/report.tex
Add reference how susceptibility varies with age.

--- 2021-09-29 src/paper/report.tex
@@
-Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
-on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination levels that would end the pandemic or even
-protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
-most have seen large fluctuations of infection rates over time. Containment measures have
-become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
-tracing. Neither these policies' effects nor the influence of seasonal patterns or of
-more infectious virus strains are well understood in quantitative terms.
+The development of highly effective vaccines holds the promise of containment in the
+medium term. However, most countries find themselves months or years away from reaching
+vaccination levels that would end the pandemic \citep{Mathieu2021}. While
+non-pharmaceutical interventions (NPIs) have allowed some countries to sustain
+equilibria with very low infection numbers \citep{Contreras2021} most have seen large
+fluctuations of infection rates over time. It is therefore of utmost importance to
+employ an effective mix of strategies for containing the virus. However, the effects of
+diverse policies and their interplay with seasonal patterns are not well understood.
 
 This paper develops a quantitative model incorporating these factors simultaneously. The
-framework allows to combine a wide variety of data and mechanisms in a timely fashion,
-making it useful to predict the effects of various interventions. Behavioral reactions
-to symptoms or positive tests are explicitly taken into account. We apply the model to
-Germany, where new infections fell by almost 80\% during May 2021. Our analysis shows
-that, aside from seasonality, frequent and large-scale rapid testing caused the bulk of
-this decrease, which is in line with prior predictions \citep{Mina2021}. We conclude
-that it should have a large role for at least as long as vaccinations have not been
-offered to an entire population.
+framework allows to combine a wide variety of data and mechanisms, making it useful to
+predict the effects of various interventions. We apply the model to Germany, where new
+infections fell by almost 80\% during May 2021. Our analysis shows that, aside from
+seasonality, frequent and large-scale rapid testing caused the bulk of this decrease,
+which is in line with prior predictions \citep{Mina2021}.
 
-\section{Model description}
+At the core of our agent-based model \cite{Aleta2020,Hinch2020} (we review more
+literature in Supplementary Material~\ref{sec:literature_review}) are physical contacts
+between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}) that
+potentially cause infections. Contacts occur in up to four areas: Within the
+household, at work, at school, or in other settings. Some contacts recur regularly,
+others occur at random. Empirical applications can take the population and household
+structure from census data and the network-specific frequencies of contacts from diary
+data measuring contacts before the pandemic (e.g. \cite{Mossong2008,Hoang2019}). Within
+each network, meeting frequencies depend on age and geographical location (see
+Supplementary Material~\ref{subsec:assortativity}).
 
-At the core of our agent-based model \citep[][we review more literature in Supplementary
-Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021a} are physical contacts between
-heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
-an infectious individual and somebody susceptible to the disease bears the risk of
-transmitting the virus. Contacts occur in up to four networks: Within the household, at
-work, at school, or in other settings (leisure activities, grocery shopping, medical
-appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
-applications can take the population and household structure from census data and the
-network-specific frequencies of contacts from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
-depend on age and geographical location (see Supplementary
-Material~\ref{subsec:assortativity}).
-
-The four contact networks are chosen so that the most common NPIs can be modeled in
-great detail. NPIs affect the number of contacts or the risk of transmitting the disease
-upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from
-home will reduce the number of work contacts to zero for a fraction of the working
-population. Schools and daycare can be closed entirely, operate at reduced
-capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
-of contacts in settings outside of
-work and school. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact
-\citep{Cheng2021}.
+The contact networks are chosen so that the most common NPIs can be modeled in great
+detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
+having physical contact. NPIs' effects on meeting frequencies will vary across contact
+types. Mask mandates or air filters would reduce the probability of infection
+\citep{Lessler2021, Cheng2021}. See Supplementary Material~\ref{sec:policies} for
+details on how we model NPIs.
 
@@
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
-possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
-differentiate between an initial period of infection without being infectious or showing
-symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
-of transitioning between these states depend on age; their duration is random and
-calibrated to medical literature (for a detailed description see Supplementary
-Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \citep{Jones2021}.
+In the model, a possible infection progresses as shown in
+Figure~\ref{fig:model_disease_progression} (similar to \cite{Aleta2020}, more in
+Supplementary Material~\ref{sub:data_course_of_disease}). Whereas susceptibility
+to the viruses differs by age \citep{Davies2020}, the infectiousness is
+independent of age conditioned on the type of contact \cite{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
-appear. Vaccines may become available. During the vaccine roll-out, priority may depend
-on age and occupation; vaccine hesitancy is modelled by some individuals refusing
-vaccination offers. With some probability, vaccinated agents become immune and do not
-transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.
+The model includes several other crucial features. New virus strains with different
+infectiousness profiles may appear. Vaccines may become available. During the vaccine
+roll-out, priority may depend on age and occupation; vaccine hesitancy is modeled by
+some individuals refusing vaccination offers. With some probability, vaccinated agents
+become immune and do not transmit the virus \cite{Hunter2021, LevineTiefenbrun2021,
+Petter2021, Pritchard2021}.
 
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
-individual is infected or not; there is no uncertainty to the result. PCR tests require
-at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
-these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021,
-Toptan2020}; sensitivity depends on the timing of the test relative to the onset of
-infectiousness. After a phase-in period, all tests that are demanded will be performed.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and a household
-member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-contacts; this is why tests impact the actual contacts in
-Figure~\ref{fig:model_description}.}
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
+certainty whether an individual is infected or not; they are scarce and require at least
+one day to be processed. In contrast, rapid antigen tests yield immediate results.
+Specificity and sensitivity of these tests is set according to data analyzed in
+\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
+to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our model for
+rapid test demand. Schools may require staff and students to be tested regularly. Rapid
+tests may be offered by employers to on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people, showing symptoms of
+CoViD-19, and a household member having tested positively for the virus. We endow each
+agent with an individual compliance parameter. This parameter determines whether she
+takes up rapid tests. Positive tests and symptoms lead most individuals to reduce their
+contacts.
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
-different contact types and their assortativity by age group. Mobility data is used to
-model the evolution of work contacts. School and daycare policies can be incorporated
-directly from official directives. Administrative records on the number of tests,
-vaccinations by age and region, and the prevalence of virus strains are generally
-available. Surveys may ask about test offers, propensities to take them up, and past
-tests. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, these include infection probabilities
-by contact network and the effects of some NPIs, see Supplementary
+parameters (see Supplementary Material~\ref{sec:materials_and_methods} for a
+complete description). Examples are contact diaries, administrative records on case
+numbers, virus strains, tests and vaccines as well as surveys on private rapid test
+demand. Other studies' estimates of the seasonality of infections can be incorporated
+directly. The remaining parameters---most notably, the infection probabilities by
+contact network and the effects of some NPIs, see Supplementary
 Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}. In our
-application, we keep the number of free parameters low in order to avoid overfitting.
-The data features to be matched include official case numbers for each age group and
-region, deaths, and the share of the B.1.1.7 strain.
+matches features of the data (see \cite{McFadden1989} for the general method). By
+modelling PCR and rapid tests in detail, we can translate infection numbers in the model
+into observed infections before matching them to the data. The mechanism is depicted in
+Figure~\ref{fig:model_official_cases} and described in Supplementary
+Material~\ref{sub:testing}. Figure~\ref{fig:share_known_cases_by_age_group} shows the
+resulting share of detected cases.
 
-The main issue with official case numbers is that they will contain only a fraction of
-all infections. In the German case, this specifically amounts to positive PCR tests. We
-thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
-mortality-based aggregate estimates of the share of detected cases and use data on the
-share of PCR tests administered because of CoViD-19
-symptoms. As the share of asymptomatic individuals
-varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is 80\% higher on average for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
-
-\section{Second and third waves of the CoViD-19 pandemic in Germany}
-
-The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
-covering the period mid-September 2020 to the end of May 2021.
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
-of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
-Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
-never use it directly. For legibility reasons, we transform the index so that lower
-values represent higher levels of restrictions. A value of zero means all measures
-incorporated in the index are turned on. The value one represents the situation in
-mid-September, with restrictions on gatherings and public events, masking requirements,
-but open schools and workplaces. In the seven weeks between mid September and early
-November, cases increased by a factor of ten. Restrictions were somewhat tightened in
-mid-October and again in early November. New infections remained constant throughout
-November before rising again in December, prompting the most stringent lockdown to this
-date. Schools and daycare centers were closed, so were customer-facing businesses except
-for grocery and drug stores. From the peak of the second wave just before Christmas
-until the trough in mid-February, newly detected cases decreased by almost three
-quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
-strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
-(Delta) was first detected in Germany in April; at the end of our simulation period it
-accounted for less than 5\% of cases.} In early March, some NPIs were relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-overall stringency index.
+The model is applied to the second and third waves of infections in Germany, covering
+the period September 2020 to May 2021. Figure~\ref{fig:pandemic_drivers_model_fit}
+describes the evolution of the pandemic and of its drivers. The black line in
+Figure~\ref{fig:aggregated_fit} shows officially recorded cases; the black line in
+Figure~\ref{fig:stringency_infectious_contacts} a rescaled Oxford Response Stringency
+Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
+Between mid September and early November, cases increased tenfold. Restrictions were
+somewhat tightened in mid-October and again in early November. New infections remained
+constant throughout November before rising again in December, prompting the most
+stringent lockdown to this date. Schools and daycare centers were closed, so were
+customer-facing businesses except for grocery and drug stores. From the peak of the
+second wave just before Christmas until the trough in mid-February, newly detected cases
+decreased by almost three quarters. The third wave in the spring of 2021 is associated
+with the B.1.1.7 (Alpha) strain, which became dominant in March
+(Figure~\ref{fig:share_b117}). In early March, some NPIs were relaxed. There were many
+changes in details of regulations afterwards, but they did not change the overall
+stringency index.
 
@@
 
-By March 2021, the set of policy instruments had become much more diverse. Around the
-turn of the year, the first people were vaccinated with a focus on older age groups and
-medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May, 43\%
-had received at least one dose of a vaccine. In late 2020, rapid tests started to
-replace regular PCR tests for staff in many medical and nursing facilities. These had to
-be administered by medical doctors or in pharmacies. At-home tests approved by
-authorities became available in mid-March. Rapid test centers were opened, and one test
-per person and week was made available free of charge. In several states, customers were
-only allowed to enter certain stores with a recent negative rapid test result. These
-developments are characteristic of many countries: The initial focus on NPIs to slow the
-spread of the disease has been accompanied by vaccines and a growing acceptance and use
-of rapid tests. At broadly similar points in time, novel strains of the virus have
-started to pose additional challenges.
-
-\section{Results}
+Around the turn of the year, the first people were vaccinated with a focus on older age
+groups and medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of
+May, 43\% had received at least one dose of a vaccine. In late 2020, rapid tests started
+to replace regular PCR tests for staff in many medical and nursing facilities. At-home
+tests approved by authorities became available in mid-March. Rapid test centers were
+opened, and one test per person and week was made available free of charge. In several
+states, customers were only allowed to enter certain stores with a recent negative rapid
+test result. These developments are characteristic of many countries: The initial focus
+on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
+acceptance and use of rapid tests.
 
 We draw simulated samples of agents from the population structure in September 2020 and
-use the model to predict recorded infection rates until the end of May 2021. See
-Supplementary Materials~\ref{subsec:synthetic_population} and
-\ref{sub:initial_conditions} for details. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model's predictions are very close to
-officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region (see Supplementary Material~\ref{subsec:fit_results}).
+use the model to predict recorded infection rates until May 2021. See Supplementary
+Materials~\ref{subsec:synthetic_population} and \ref{sub:initial_conditions} for
+details. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model's
+predictions are very close to officially recorded cases in the aggregate. This is also
+true for infections by age and geographical region (see Supplementary
+Material~\ref{subsec:fit_results}).
 
@@
 lines broadly follow the stringency index and they would do so even more if we left out
-seasonality and school vacations (roughly the last two weeks of October, two weeks each
-around Christmas and Easter, and some days in late May). Second, the most stringent
-regulations coincide with the period of decreasing infection rates between late December
-2020 and mid-February 2021. The subsequent reversal of the trend is associated with the
-spread of the B.1.1.7 variant. During the steep drop in recorded cases during May 2021,
-for 42\% of the population took at least one rapid tests per week, the first-dose vaccination
-rate rose from 28\% to 43\%, and seasonality lowered the relative infectiousness
-of contacts.
+seasonality and school vacations. Second, the most stringent regulations coincide with
+the period of decreasing infection rates between late December 2020 and mid-February
+2021. The subsequent reversal of the trend is associated with the spread of the B.1.1.7
+variant. During the steep drop in recorded cases during May 2021, for 42\% of the
+population took at least one rapid tests per week, the first-dose vaccination rate rose
+from 28\% to 43\%, and seasonality lowered the relative infectiousness of contacts.
 
 In order to better understand the contributions of rapid tests, vaccinations, and
-seasonality on the evolution of infections in 2021,
-Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always held
-constant at their values in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red
-line), and three scenarios turning each of these factors on individually.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
-\citep{Shapley2016} to decompose the difference in total infections between the scenario
-without any of the three factors and our main specification.
+seasonality on the evolution of infections, Figure~\ref{fig:2021_scenarios_broad}
+considers various scenarios. NPIs are always held constant at their values in the
+baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the
+blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario without any of the
+three factors (red line), and three scenarios turning each of these factors on
+individually. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
+infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
+values \citep{Shapley2016} to decompose the difference in total infections between the
+scenario without any of the three factors and our main specification.
 
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
-        For legibility reasons, all lines are rolling 7-day averages.}
+        are explained in Appendix~\ref{sub:shapley_value}. All lines are rolling 7-day
+        averages.
+    }
 \end{figure}
 
-Until mid-March, there is no visible difference between the different scenarios.
-Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
-Even thereafter, the effect of the vaccination campaign is surprisingly small at first
-sight. Whether considering recorded or total infections with only one channel active,
-the final level is always the highest in case of the vaccination campaign (orange
-lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
-cumulative difference between scenarios. Reasons for the low share are the slow
-start---it took until March~24th until 10\% of the population had received their first
-vaccination, the 20\% mark was reached on April 19th---and the focus on older
-individuals. These groups contribute less to the spread of the disease than others due
-to a lower number of contacts. By the end of our study period, when first-dose
-vaccination rates reached 43\% of the population, the numbers of new cases would have
-started to decline. It is important to note that the initial focus of the campaign was
-to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
-lower during the third wave when compared to the second (4.4\% between October and
-February and 1.4\% between March and the end of May).
+Until mid-March, there is no visible difference between the different scenarios. Even
+thereafter, the effect of the vaccination campaign is surprisingly small. The Shapley
+value decomposition shows that vaccinations contribute 16\% to the cumulative difference
+between scenarios. Reasons for the low share are the slow start and the focus on older
+individuals who typically have fewer contacts. By the end of our study period, when
+first-dose vaccination rates reached 43\% of the population, the numbers of new cases
+would have started to decline.
 
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace \citep[this is in line with results in][, which our seasonality measure
-is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a quantitatively
-important factor determining the evolution of the pandemic, explaining most of the early
-changes and 43\% of the cumulative difference by the end of May.
+much lower pace (this is in line with results in \cite{Gavenciak2021}, which our
+seasonality measure is based on). Nevertheless, we estimate seasonality to be an
+important factor, explaining most of the early changes and 43\% of the cumulative
+difference by the end of May.
 
@@
 elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. The latter is essentially a form of contact
-tracing, which has been shown to be very effective \citep{Contreras2021,
+up on a positive test of a household member. This is essentially a form of contact
+tracing, which has been shown to be very effective \cite{Contreras2021,
 Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
@@
 administered randomly in the population would not have been nearly as effective.
-
-\section{Discussion and conclusions}
-
-Having quantified the effects of various mechanisms, we now simulate hypothetical
-scenarios comparing changes in NPIs and testing regimes. Two of the most contentious
-NPIs concern schools and mandates to work from home. In many countries, schools switched
-to remote instruction during the first wave, so did Germany. After the summer break,
-they were operating at full capacity with increased hygiene measures, before being
-closed again from mid-December onward. Some states started opening them gradually in
-late February, but operation at normal capacity did not resume until the beginning of
-June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
-regarding schools starting after Easter, at which point rapid tests had become widely
-available. We estimate the realized scenario to have essentially the same effect as a
-situation with closed schools. Under fully opened schools with mandatory tests, total
-infections would have been 6\% higher; this number rises to 20\% without tests. These
-effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
-Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of
-the large negative effects school closures have on children and parents
-\citep{Luijten2021, Melegari2021}---and in particular on those with low socio-economic
-status---these results in conjunction with hindsight bias suggest that opening schools
-combined with a testing strategy would have been beneficial. In other situations, and
-particular when rapid test are not available at scale, trade-offs may well be different.
-
-\begin{figure}[!tp]
-    \centering
-
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-        \caption{{Effects of different schooling scenarios}}
-        \label{fig:school_scenarios}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
-        \caption{{Effects of different work scenarios}}
-        \label{fig:workplace_scenarios}
-    \end{subfigure}
-    \vskip3ex
-
-    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
-    \label{fig:school_workplace_scenarios}
-
-    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
-        scenario; they are the same as in
-        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
-        because there were no capacity constraints for rapid tests afterwards.
-        For legibility reasons, all lines are rolling 7-day averages.}
-
-\end{figure}
-
-Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
-receiving tests, testing at the workplace has larger effects than mandating employees to
-work from home. Whether the share of workers working at the usual workplace is reduced
-or increased by ten percent changes infection rates by 2.5\% or less in either
-direction. Making testing mandatory twice a week---assuming independent compliance by
-employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
-rapid tests offers by employers to the level of March would have increased infections by
-13\%.
+Furthermore, mandating tests at schools almost offsets the increased infection risk of
+opening schools (see Figure~\ref{fig:school_scenarios}).
 
@@
 possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
-at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
-prices for rapid tests were below one Euro in early June 2021 and below five Euros for
-firms. While we do not distinguish between self-administered rapid tests and point of
-care rapid tests, the former are likely to play a larger role for indication-driven
-testing. Widespread availability at low prices seems important. However, they rely on
-purely voluntary participation in a non-public setting. The benefit of point-of-care
-rapid tests as a precondition to participate in leisure activities as well as mandatory
-tests at the workplace or at school come from screening the entire population. This is
-important because disadvantaged groups are less likely to be reached by testing campaigns
-relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time,
-these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
-tests at school and at the workplace will extend more into these groups. The same goes
-for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
-Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
-it arguably the most effective tool to contain the pandemic in the short run.
+at a fraction of the cost. A day of strict lockdown is estimated to have cost around
+50~Euros per capita \cite{Dorn2020b}; retail prices for rapid tests were below one Euro
+in early June 2021. Widespread availability of self-administered tests at low prices are
+likely to play a role for indication-driven testing. Mandatory tests in schools and at
+the workplace are important to screen the entire population, also because disadvantaged
+groups are less likely to be reached by testing campaigns relying on voluntary
+participation (e.g. \cite{StillmanTonin2021}); at the same time, these groups have a
+higher risk to contract CoViD-19 \cite{KochInstitut2021a}. Compared to vaccinations,
+rapid testing programmes allow a much quicker roll-out, making it arguably the most
+effective tool to contain the pandemic in the short run.

### 2022-03-08 src/paper/results/robustness_to_sensitivity_assumptions.tex
Add results and section on robustness checks with less sensitive rapid tests.

\subsection{Robustness to assumptions about rapid test sensitivity}
\label{subsec:robustness_rapid_test_sensitivity}

Our result that frequent and large scale rapid testing was responsible for a
large decrease in cases in the spring of 2021 in Germany seems to be at odds
with the general perception that rapid tests are unreliable. They also seem to
contradict recent studies showing that the actual sensitivity of rapid tests
might be lower than advertised (e.g.\cite{Scheiblauer2021}).

This is, however, not the case. It is important to distinguish studies that
analyze the sensitivity and specificity of rapid tests in isolation and studies
that analyze the aggregate effect of testing strategies within an
epidemiological model.

This section shows that our results are robust to making less favorable
assumptions on rapid test sensitivity. We proceed by describing several possible
ways of calibrating rapid test sensitivity profiles based on recent studies.
Since none of these methods is inherently better than the others, we make
simulations with two sensitivity profiles: The average over all methods and the
lower envelope over all methods.

Both profiles imply much lower sensitivities of rapid tests than used in our
main results, especially in the later stage of an infection. However, the main
results stay very robust. The original result was that rapid tests, seasonality
and vaccinations are responsible for 42\%, 43\% and 16\%, respectively. With the
average profile, the effect of rapid tests goes down to 41 \%. With the lower
envelope profile, which is an extremely unfavorable assumption, it becomes 38\%.

% what do we need
The effect of rapid tests on infection dynamics strongly depends on \emph{when}
an infection is detected. Earlier detection means that it is more likely that
the infection has not yet been discovered for a different reason (e.g. due to
the onset of symptoms) and that the infected person can be isolated before
spreading the disease to others.

The sensitivity of rapid tests depends on the viral load in the repiratory
tract. It is low at the beginning of an infection (especially before the onset
of infectiousness), high in the first few days of infectiousness and then
gradually decreasing towards the end of infectiousness.

We thus need to calibrate a profile of rapid test sensitivities based on the
number of days until or since the onset of infectiousness.

% what is available -> two step mapping
Unfortunately, such sensitivity profiles are not usually reported in studies. We
thus need to create them by combining two types of studies: 1. Studies that
report the viral load in terms of threshold cycle ($C_t$) values determined by
PCR test 
(e.g. \cite{Cosentino2022}
, \cite{Ong2021}, 
\cite{Bonnet2022},
\cite{Jang2021} and \cite{Zuin2021}
). 2. Studies that report the ensitivity of
rapid tests for different $C_t$ values 
(e.g. \cite{Scheiblauer2021},
\cite{Bruemmer2021}).

% what options do we have to configure the mappings
It is natural to assume that the evolution of $C_t$ values over time as well as
the effect of $C_t$ values on rapid test sensitivity are continuous functions.
However, the results of the available studies are usually reported in a
discretized way. This leads to multiple ways of calculating the sensitivity
profiles. Some try to recover the underlying continuous functions using
interpolation or regression, others simply use the discretized values.

% Describe CT value formula vs. bin averages
For the calibration of $C_t$ values over time we can either use discretized
values for several time bins from \cite{Ong2021} and \cite{Jang2021}.
Alternatively, we can use linearized formulas for calculating sensitivities over
time \cite{Cosentino2022}.

% Describe sensitivity lookupd vs. interpolation
For the mapping of $C_t$ values to rapid test sensitivities, again we have two
options. First, we can simply look up the discretized values for the three $C_t$
bins provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35).
Secondly, we can use linear regression to estimate a continous mapping for the
relationship by assuming that the $C_t$ values of each bin are achieved exactly
at the bin midpoints and that the relationship is linear.

% Properties of the different methods
In general, using discretized values can lead to an underestimation of peak
sensitivities and an overestimation of very low sensitivities. This is because
discretization is essentially a smoothing device. On the other hand, it hat the
advantage of simply working with published results, without introducing any
tuning parameters or other assumptions.

% resulting profiles.
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/results/figures/data/testing/sensitivity_params_with_different_methods}
    \caption{Rapid test sensitivity profiles}
    \floatfoot{\noindent \textit{Note:} The figure shows estimated sensitivities
    of rapid tests over the course of an infection. The x-axis shows days
    relative to the onset of infectiousness. The y-axis shows the estimated
    rapid test sensitivities. The grey lines are the raw sensitivity estimates
    obtained with different methods of dealing with discretized data. The blue
    line shows their average and the yellow line their lower envelope. The
    turquoise line are the sensitivities used for the main results of the paper.
    It is much higher than the new estimates in the later stage of an infection.
    Luckily this has only a very small effect on the overall effectiveness of
    rapid tests on the infection dynamics.}
    \label{fig:sensitivity_assumptions}
\end{figure}

\FloatBarrier

Figure \ref{fig:sensitivity_assumptions} shows that the updated sensitivity
estimates are lower than the ones used for our original results, especially
towards the end of an infection. However, the main results barely change.

\begin{figure}
    \centering

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/results/figures/data/testing/full_decomposition_channels_area_average}
        \caption{Shapley decomposition with average of sensitivity estimates}
    \end{subfigure}
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/results/figures/data/testing/full_decomposition_channels_area_lower_envelope}
        \caption{Shapley decomposition with lower envelope of sensitivity estimates}
    \end{subfigure}

    \caption{Updated shapley decompositions}
    \floatfoot{\noindent \textit{Note:} The figure shows updated versions of the
    shapley decomposition in figure \ref{fig:2021_scenarios_decomposition}. The
    decomposition is based on 20 model runs with different random seeds. The
    share attributed to each channel is rounded to the next full percentage
    point to acknowledge the remaining sampling uncertainty.}
    \label{fig:robustness_to_sensitivity_assumptions}
\end{figure}

\FloatBarrier

%

### 2022-03-08 src/paper/results/robustness_to_sensitivity_assumptions.tex
Fix problems with references.

@@
 with the general perception that rapid tests are unreliable. They also seem to
-contradict recent studies showing that the actual sensitivity of rapid tests
-might be lower than advertised (e.g.\cite{Scheiblauer2021}).
+contradict recent studies showing that the actual sensitivity of rapid tests t
+be lower than advertised (e.g.\cite{Scheiblauer2021}).
 
@@
 report the viral load in terms of threshold cycle ($C_t$) values determined by
-PCR test 
-(e.g. \cite{Cosentino2022}
-, \cite{Ong2021}, 
-\cite{Bonnet2022},
-\cite{Jang2021} and \cite{Zuin2021}
-). 2. Studies that report the ensitivity of
-rapid tests for different $C_t$ values 
-(e.g. \cite{Scheiblauer2021},
+PCR test (e.g. \cite{Cosentino2022}, \cite{Ong2021}, \cite{Bonnet2022},
+\cite{Jang2021} and \cite{Zuin2021}). 2. Studies that report the ensitivity of
+rapid tests for different $C_t$ values (e.g. \cite{Scheiblauer2021},
 \cite{Bruemmer2021}).
@@
 options. First, we can simply look up the discretized values for the three $C_t$
-bins provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35).
-Secondly, we can use linear regression to estimate a continous mapping for the
+provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35). Secondly,
+we can use linear regression to estimate a continous mapping for the
 relationship by assuming that the $C_t$ values of each bin are achieved exactly

### 2022-03-08 src/paper/model/testing.tex
Refer to sensitivity robustness check when citing Smith2021.

--- 2021-07-15 src/paper/model/testing.tex
@@
 
-In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
-falsely negative. While the specificity of rapid tests is calibrated at 99.4\%
-\citep{Bruemmer2021}, their sensitivity strongly depends on the timing of the rapid test
-relative to the start of infectiousness \citep{Smith2021}. Before the onset of
-infectiousness the sensitivity is very low (35\%). On the first day of infectiousness it
-is much higher (88\%) but still lower than during the remaining infectious period (92\%).
-After infectiousness stops, the sensitivity drops to 50 \%.
+In contrast to PCR tests, rapid tests are not perfect and can be falsely
+positive or falsely negative. While the specificity of rapid tests is calibrated
+at 99.4\% \citep{Bruemmer2021}, their sensitivity strongly depends on the timing
+of the rapid test relative to the start of infectiousness. We follow
+\cite{Smith2021} for our main results: Before the onset of infectiousness the
+sensitivity is very low (35\%). On the first day of infectiousness it is much
+higher (88\%) but still lower than during the remaining infectious period
+(92\%). After infectiousness stops, the sensitivity drops to 50\%. We show that
+our results are robust to more conservative assumptions in
+Section~\ref{subsec:robustness_rapid_test_sensitivity}.
 

### 2022-03-08 src/paper/report.tex
Refer to sensitivity robustness check when citing Smith2021.

@@
 We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
-certainty whether an individual is infected or not; they are scarce and require at least
-one day to be processed. In contrast, rapid antigen tests yield immediate results.
-Specificity and sensitivity of these tests is set according to data analyzed in
-\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
-to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our model for
-rapid test demand. Schools may require staff and students to be tested regularly. Rapid
-tests may be offered by employers to on-site workers. Individuals may demand tests for
-private reasons, which include having plans to meet other people, showing symptoms of
-CoViD-19, and a household member having tested positively for the virus. We endow each
-agent with an individual compliance parameter. This parameter determines whether she
-takes up rapid tests. Positive tests and symptoms lead most individuals to reduce their
-contacts.
+certainty whether an individual is infected or not; they are scarce and require
+at least one day to be processed. In contrast, rapid antigen tests yield
+immediate results. Specificity and sensitivity of these tests is set according
+to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
+timing of the test relative to the onset of infectiousness (we show our results
+are robust to more conservative assumptions in
+Section~\ref{subsec:robustness_rapid_test_sensitivity}).
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand.
+Schools may require staff and students to be tested regularly. Rapid tests may
+be offered by employers to on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people, showing
+symptoms of CoViD-19, and a household member having tested positively for the
+virus. We endow each agent with an individual compliance parameter. This
+parameter determines whether she takes up rapid tests. Positive tests and
+symptoms lead most individuals to reduce their contacts.
 

### 2022-03-08 src/science/paper.tex
Add declaration of interest.
Refer to sensitivity robustness check when citing Smith2021.

@@
 
+\section*{Ethical declarations}
+
+\subsection*{Competing interests}
+
+The authors declare no competing interests.
+
 %Here you should list the contents of your Supplementary Materials -- below is an

### 2022-03-08 src/paper/results/robustness_to_sensitivity_assumptions.tex
Talk about pre and asymptomatics' Ct values.

@@
 thus need to create them by combining two types of studies: 1. Studies that
-report the viral load in terms of threshold cycle ($C_t$) values determined by
+report the viral load in terms of threshold cycle ($Ct$) values determined by
 PCR test (e.g. \cite{Cosentino2022}, \cite{Ong2021}, \cite{Bonnet2022},
 \cite{Jang2021} and \cite{Zuin2021}). 2. Studies that report the ensitivity of
-rapid tests for different $C_t$ values (e.g. \cite{Scheiblauer2021},
+rapid tests for different $Ct$ values (e.g. \cite{Scheiblauer2021},
 \cite{Bruemmer2021}).
@@
 % what options do we have to configure the mappings
-It is natural to assume that the evolution of $C_t$ values over time as well as
-the effect of $C_t$ values on rapid test sensitivity are continuous functions.
+It is natural to assume that the evolution of $Ct$ values over time as well as
+the effect of $Ct$ values on rapid test sensitivity are continuous functions.
 However, the results of the available studies are usually reported in a
@@
 % Describe CT value formula vs. bin averages
-For the calibration of $C_t$ values over time we can either use discretized
+For the calibration of $Ct$ values over time we can either use discretized
 values for several time bins from \cite{Ong2021} and \cite{Jang2021}.
 Alternatively, we can use linearized formulas for calculating sensitivities over
-time \cite{Cosentino2022}.
+time \cite{Cosentino2022} and complement it with interpolations of data points
+from \cite{Jang2021} in the pre-infectious stage. Throughout we assume that the
+$Ct$ values of individuals who eventually develop symptoms and those who do not
+follow the same trajectory. This is in line with results by \cite{Zuin2021} and
+recent evidence that rapid tests excel at discovering asymptomatic cases
+\cite{Rosella2022}.
 
 % Describe sensitivity lookupd vs. interpolation
-For the mapping of $C_t$ values to rapid test sensitivities, again we have two
-options. First, we can simply look up the discretized values for the three $C_t$
+For the mapping of $Ct$ values to rapid test sensitivities, again we have two
+options. First, we can simply look up the discretized values for the three $Ct$
 provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35). Secondly,
 we can use linear regression to estimate a continous mapping for the
-relationship by assuming that the $C_t$ values of each bin are achieved exactly
+relationship by assuming that the $Ct$ values of each bin are achieved exactly
 at the bin midpoints and that the relationship is linear.

### 2022-03-09 src/rev1_reply_comments.tex
Remove outdated folder, add reply to comments.

\documentclass[a4paper,11pt]{article}
\usepackage{a4wide}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{afterpage,rotating,natbib,graphicx,setspace,xcolor}
\usepackage{longtable,booktabs,tabularx}
\usepackage{chngcntr}
\usepackage{eurosym,calc}
\usepackage{amsmath,amssymb,amsfonts,amsthm,delarray}
\usepackage{bm}
\usepackage{bbm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xfrac}
\usepackage{listings}
\usepackage{adjustbox}
\usepackage{xr}
\usepackage{xr-hyper}
\usepackage[unicode=true]{hyperref}
\usepackage{afterpage}
\usepackage{clipboard}

\definecolor{darkblue}{rgb}{0, 0, 0.8}
\hypersetup{
    colorlinks=true,
    linkcolor=darkblue,
    anchorcolor=darkblue,
    citecolor=darkblue,
    filecolor=darkblue,
    menucolor=darkblue,
    runcolor=darkblue,
    urlcolor=darkblue
}
\usepackage{enumerate}
\usepackage{caption}
\usepackage{booktabs,caption,rotating}
\usepackage[capposition=top]{floatrow}
% \usepackage{bbm}
\usepackage{epstopdf}
\usepackage{longtable, tabularx, array}
\usepackage{colortbl}

% \usepackage[british]{babel}
% \usepackage[useregional]{datetime2}
% \DTMlangsetup[en-GB]{showdayofmonth=false}

\usepackage{subcaption}
\usepackage{eurosym}

% \setlength{\marginparwidth}{2cm}
\usepackage[
    colorinlistoftodos,
    backgroundcolor=white,
    prependcaption,
    textsize=tiny,
    disable,
]{todonotes}

%%%%PROPOSITION AND PROOF ENVIRONMENTS%%%%
\theoremstyle{plain}
\newtheorem{result}{Result}
\newtheorem{assumption}{Assumption}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

% don't let footnotes split across pages
\interfootnotelinepenalty=10000
\widowpenalty=10000
\clubpenalty=10000

\hyphenation{bet-ween}
\hyphenation{with-out}

\frenchspacing

\newlength{\parindentaux}
\setlength{\parindentaux}{\parindent}

\usepackage{enumitem}
\setlist[itemize]{%
    leftmargin=\parindentaux,
    listparindent=\parindentaux
}

% \usepackage{calc}  % Needed to be able to calculate, e.g., sum of dimensions
%\usepackage{ifthen}

\makeatletter
\newcommand*\@myenumerate{enumerate}
\newcommand*\@myitemize{itemize}
\newcommand{\comment}[1]{{\color{black!60}#1}}
\newcommand{\response}[1]{{%
		\color{black}%
		\par\frenchspacing%
		\ifx\@currenvir\@myenumerate%
		\addtolength{\leftskip}{-\parindentaux}%
		\rm\noindent #1\par%
		\addtolength{\leftskip}{\parindentaux}%
		\else%
		\ifx\@currenvir\@myitemize%
		\addtolength{\leftskip}{-\parindentaux}%
		\rm\noindent #1\par%
		\addtolength{\leftskip}{\parindentaux}%
		\else%
		\rm\noindent #1\par%
		\fi%
		\fi%
}}
\makeatother

\renewcommand{\theequation}{\roman{equation}}
\renewcommand{\thetable}{\Roman{table}}
\renewcommand{\thefigure}{\Roman{figure}}
\renewcommand{\thesection}{\Roman{section}}

\setlength{\parindent}{0ex}

\begin{document}

\title{\large Reply to the Editor's and Reviewers' comments on: \\[2ex]
    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations and Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
    \large Submitted to Scientific Reports (Submission ID db917d3c-b56a-4129-8eb1-a62a716aaede)\\[-6ex]
}

\author{}
\date{}

\maketitle

\section*{Editors}

\comment{
The title should have a clear, precise scientific meaning and should not contain a colon. Where possible, the title should be read as one concise sentence.
}
\vskip1ex
\response{%
    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations and Contact Restrictions for Containing the Covid-19 Pandemic''
}

\section*{Reviewer 1}
\comment{
The authors of the manuscript presented here have used mathematical models to examine the impact of different measures to contain the SARS-CoV-2 pandemic. Using epidemiological data from Germany, the contribution of rapid antigen testing is considered in particular.

The topic studied here is of great importance for the control of the coronavirus pandemic. The presentation of the scientific methodology is sufficiently detailed, especially with the supplementary materials. It allows the reader to follow the research process in detail. However, I noticed a significant deficiency when reviewing the submitted work. Therefore, the manuscript should definitely be revised again before publication: A crucial point of the authors is to point out the special benefit of a rapid test strategy. However, for the calculation of the corresponding influencing factors in their mathematical model, the authors simplify, specifically, the sensitivity of rapid antigen tests in an unacceptable way. In the development of the mathematical model, apparently only the knowledge available a few months ago on the performance of rapid antigen tests was taken into account. This is evident from literature references mentioned above. In part, incorrect references are also used, especially for the assumption of the high sensitivity of the rapid tests. For example, reference No. 74 explicitly does not refer to rapid antigen tests, but to an immunological laboratory test method (Quidel SARS Sofia antigen FIA). Especially in light of the benefit of rapid antigen tests described by the authors, it should be considered that many of the commercially marketed tests show only extremely poor performance. This has recently been demonstrated by several publications:
\begin{itemize}
    \item     Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic tests for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer, Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and Corman, Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange, Constanze and Emmerich, Petra and Müller, Michael and Knauer, Olivia and Nübling, C Micha, Eurosurveillance, 26, 2100441 (2021), \url{https://doi.org/10.2807/1560-7917.ES.2021.26.44.2100441}
    \item Assessment of SARS-CoV-2 rapid antigen tests, Özcürümez, Mustafa, Katsounas, Antonios, Holdenrieder, Stefan, von Meyer, Alexander, Renz, Harald and Wölfel, Roman, Journal of Laboratory Medicine, vol. 45, no. 3, 2021, pp. 143-148. \url{https://doi.org/10.1515/labmed-2021-0036}
\end{itemize}

Especially in the case of a mass application of antigen tests, it is therefore to be expected that products with lowest sensitivity will also be used in the general population. The authors should at least model this with an alternative calculation and discuss it accordingly. Such an improvement would increase the importance of this manuscript enormously.

Furthermore, I recommend including a declaration of potential conflicts of interest in the manuscript. This may be particularly important in light of the significance of the findings from the study for the continued operation of industrial companies.

\vskip1ex

\response{Many thanks for these insightful comments. We respond in three parts.}

\begin{enumerate}
    \item In part, incorrect references are also used, especially for the assumption of the high sensitivity of the rapid tests. For example, reference No. 74 explicitly does not refer to rapid antigen tests, but to an immunological laboratory test method (Quidel SARS Sofia antigen FIA).

          \response{%
              We thank the reviewer for spotting this mistake on our part. We removed the reference and only use current information as detailed in the next item.
          }

    \item In the development of the mathematical model, apparently only the knowledge available a few months ago on the performance of rapid antigen tests was taken into account.

          The authors should at least model this with an alternative calculation and discuss it accordingly. Such an improvement would increase the importance of this manuscript enormously.

          \response{%
              We agree that the references we used have become outdated in the meantime (we submitted the paper in  August 2021). Indeed, the numbers we used turned out to be too optimistic in light of newer studies.

              We therefore ran additional simulations with updated and more conservative assumptions on the sensitivity of rapid tests (Appendix B.12) and found that our main results are remarkably robust to this change. The percentage of the decrease in spring 2021 attributed to rapid tests decreases by only 4 percentage points even in the least favourable scenario. We describe our approach for calculating the adjusted sensitivities in the next answer.
          }
    \item I recommend including a declaration of potential conflicts of interest in the manuscript.
          \response{We have added such a declaration. There are no competing interests.}

\end{enumerate}
}

\section{Reviewer 2}

\comment{
This is a comprehensive and important study that highlights the relevance of testing to stop the spread of the pandemic.

I cannot comment on the validity of the underlaying mathematical analysis, but I would like to comment on some assumptions made, which I think require further clarification or discussion.

\begin{enumerate}
    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful if behavior of people is already changed by the fact there is a pandemic?
          \response{%
              We account for individual responses in behaviour to the pandemic through a time-variant multiplier that reduces the number of other (i.e. non-work, non-household, non-educational) contacts. The values at change points that mark policy announcements or fatigue effects are estimated. They are detailed in Section A.9.

              Since we estimate both the infection probability of this contact type and the multiplier, one of them has to be normalised. We decided to normalise the multiplier to 1 in October 2020. This may appear as if we assumed that pre-pandemic behaviour had resumed during that time in Germany. However, this is not the case. In fact, we would achieve the exact same fit if we doubled the infection probability and halved the multiplier. This is because halving the multiplier means only half the number of contacts take place. However, because of the doubling of the infection probability, in expectation we would arrive at the same number of infections in that contact category.
          }
    \item They state that susceptibility is dependent on age. This requires references and needs to be further explained. Also, younger populations are highly susceptible, they may just not contribute transmission as much. On that same note, infectiousness may not be independent of age

          Edward Goldstein, Marc Lipsitch, Muge Cevik, On the Effect of Age on the Transmission of SARS-CoV-2 in Households, Schools, and the Community, The Journal of Infectious Diseases, Volume 223, Issue 3, 1 February 2021, Pages 362–369, \url{https://doi.org/10.1093/infdis/jiaa691}

          \response{%
              We now cite our main reference for the susceptibilities for each age group also in the main text. We directly take the susceptibilities given by Davies et al. 2020, in their Extended Data Fig. 4.
              We base our assumption on infectiousness being independent of age on Jones et al. 2021. They report only slight differences in viral loads between German adults and children: ``The least infectious youngest children have 78\% (61, 94) of the peak culture probability of adults aged 45 to 55''. Furthermore, they caution that these lower numbers may be due to different sample taking - such as the usage of smaller pediatric swab devices - rather than actual lower viral loads and conclude that the differences, if existent, are likely not clinically relevant.
          }
    \item It is not accurate that the vaccines stop transmission. This has to be adapted accordingly. Also, the authors should state which vaccines they based this analysis on, as in which ones were in use in Germany at that time and how does that affect the relative transmissibility as it cannot be assumed that all vaccines will have the same effect.
          \response{%
              We talk in detail about the vaccines that were used in Germany and on which we base our analysis on in the last third of appendix A.1.
              We agree that vaccines do not stop transmission completely. We simplify the way vaccines reduce transmission risk considerably by assuming that 75\% of individuals achieve sterile immunity through a vaccination while the remaining 25\% remain fully susceptible. [Continue here]
          }
\end{enumerate}
}

% \setstretch{1}
% \bibliographystyle{ecta}
% \bibliography{refs}

\end{document}

### 2022-03-09 task_compile_documents.py

@@ -37,25 +37,9 @@ def task_compile_documents():
         "-f",
     ]
 )
-@pytask.mark.depends_on([SRC / "science" / "paper.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "science" / "paper.pdf")
-def task_compile_science_report():
-    pass
-
-
-@pytask.mark.latex(
-    [
-        "--pdf",
-        "--interaction=nonstopmode",
-        "--synctex=1",
-        "--cd",
-        "--shell-escape",
-        "-f",
-    ]
-)
-@pytask.mark.depends_on([SRC / "science" / "cover_letter.tex", *_DEPENDENCIES])
-@pytask.mark.produces(BLD / "cover_letter.pdf")
-def task_compile_cover_letter():
+@pytask.mark.depends_on([SRC / "rev1_reply_comments.tex"])
+@pytask.mark.produces(BLD / "rev1_reply_comments.pdf")
+def task_compile_reply():
     pass
 
 
@@ -81,170 +65,3 @@ def task_extract_supplementary_material(depends_on, produces):
         ]
     )
 
-
-@pytask.mark.depends_on(
-    {
-        "paper": SRC / "science" / "paper.tex",
-        "scicite": SRC / "science" / "scicite.sty",
-        "model_graph_top_left": SRC / "figures" / "model-graph-top-left.pdf",
-        "model_graph_top_right": SRC / "figures" / "model-graph-top-right.pdf",
-        "model_graph_bottom_left": SRC / "figures" / "model-graph-bottom-left.pdf",
-        "model_graph_bottom_right": SRC / "figures" / "model-graph-bottom-right.pdf",
-        "full_new_known_case": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "scenario_comparisons"
-        / "combined_fit"
-        / "full_new_known_case.pdf",
-        "stringency2_with_seasonality": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "data"
-        / "stringency2_with_seasonality.pdf",
-        "full_share_b117": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "scenario_comparisons"
-        / "combined_fit"
-        / "full_share_b117.pdf",
-        "full_share_rapid_test_in_last_week_and_vaccinated": SRC
-        / "figures"
-        / "results"
-        / "figures"
-        / "scenario_comparisons"
-        / "combined_fit"
[…]

### 2022-03-09 src/paper.tex
Add preliminary science draft.
Final version.
Headline, first batch of changes.
Version as submitted.
Various fixes that got lost in the previous PR.
Add declaration of competing interests.
Add declaration of interest.
Title consistent.

@@
 \newcommand{\papertitle}{
-The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs
+    The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic
 }
@@
 
-\begin{refcontext}[sorting=nyt]  % Sort BIBLIOGRAPHY by alphabet (while CITATIONS are sorted by year)
+\begin{refcontext}[sorting=none]  % Sort BIBLIOGRAPHY by appearance.
     \printbibliography[heading=bibintoc]
 \end{refcontext}
+
+\section*{Ethical declarations}
+
+\subsection*{Competing interests}
+
+The authors declare no competing interests.
 
@@
 
-    \input{paper/appendix_data}
+        \input{paper/appendix_materials_and_methods}
 
@@
 
-    \input{paper/appendix_model}
+        \input{paper/appendix_supplementary_text}
 
-    \input{paper/appendix_results}
-
+        \begin{refcontext}[sorting=none]
             \printbibliography[heading=bibintoc]
+        \end{refcontext}
 

### 2022-03-09 src/rev1_reply_comments.tex
Title consistent.

@@
 \title{\large Reply to the Editor's and Reviewers' comments on: \\[2ex]
-    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations and Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
+    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
     \large Submitted to Scientific Reports (Submission ID db917d3c-b56a-4129-8eb1-a62a716aaede)\\[-6ex]
@@
 \response{%
-    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations and Contact Restrictions for Containing the Covid-19 Pandemic''
+    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic''
 }

### 2022-03-09 src/rev1_reply_comments.tex
Transfer from Google Doc new answers regarding vaccinations and pre-pandemic contact frequencies.

@@
           \response{%
+
+          Other models also use the Mossong data: OpenABM-Covid-19 model (Hinch2020), Bicher2021.
+
+          We only use the pre-pandemic diary data as the starting point.
+Our policies, which include non-mandated individual behavior adjustments, then reduce contacts taking the type of contact into account and for the most part not randomly but in very specific ways.
+This allows us to account for behavior reactions and policy effects that target different types of contacts differently.
+For example, household contacts continue to take place no matter the policy environment.
+Work from home does not keep random individuals at home but every individual that works has a time-constant value that determines when he or she will stay home.
+Which share of the population stays home is taken from Google mobility data.
+In the education sector we do not use the Mossong data but use completely mechanistic models.
+This gives us confidence that we can sufficiently accurately represent the contact patterns during the pandemic and account for changes induced by the pandemic.
+
+          Also, our excellent fit for most age groups gives us confidence that we manage to match actual meeting patterns.
+If we were far off for example for the number of work contacts, we would not be able to match either the infections of individuals in the working age (35-59) or of those that mostly do not work anymore (60-79).
+
+          \textbf{Answer in case there is confusion because the other multiplier is 1 in fall 2020:}
+
           We account for individual responses in behaviour to the pandemic through a time-variant multiplier that reduces the number of other (i.e.
@@
           \response{%
-              We talk in detail about the vaccines that were used in Germany and on which we base our analysis on in the last third of appendix A.1.
           We agree that vaccines do not stop transmission completely.
 We simplify the way vaccines reduce transmission risk considerably by assuming that 75\% of individuals achieve sterile immunity through a vaccination while the remaining 25\% remain fully susceptible.
-[Continue here]
+We are comfortable with this assumption for several reasons:
+          Firstly, the existing evidence supports the assumption that all four vaccines approved in the EU (Biontech/Pfizer, Moderna, AstraZeneca and Johnson&Johnson) provide similar levels of protection in the first few months after vaccination and against the Alpha variant (for details see the last third of appendix A.1), which is the only time period and variant our simulations cover.
+          Secondly, given the surprisingly good performance of rapid tests we tried to err on the side of being too optimistic with our modelling of vaccine performance.
+This is  why we chose the perfect protection from vaccination for a large part of the population.
+          Thirdly, reinfections are not a big issue for the time period that we look at.
+The incidences were small and the chance of more than one exposure to Covid are negligibly small.
+Thus, the value added by a probabilistic model of vaccination protection is very small for the period that we cover.
           }

### 2022-03-09 src/rev1_reply_comments.tex
Allow compilation.

@@
 We are comfortable with this assumption for several reasons:
-          Firstly, the existing evidence supports the assumption that all four vaccines approved in the EU (Biontech/Pfizer, Moderna, AstraZeneca and Johnson&Johnson) provide similar levels of protection in the first few months after vaccination and against the Alpha variant (for details see the last third of appendix A.1), which is the only time period and variant our simulations cover.
+              Firstly, the existing evidence supports the assumption that all four vaccines approved in the EU (Biontech/Pfizer, Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of protection in the first few months after vaccination and against the Alpha variant (for details see the last third of appendix A.1), which is the only time period and variant our simulations cover.
               Secondly, given the surprisingly good performance of rapid tests we tried to err on the side of being too optimistic with our modelling of vaccine performance.

### 2022-03-09 src/paper/report.tex

--- 2022-03-04 src/paper/report.tex
@@
 
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
-individual is infected or not; there is no uncertainty to the result. PCR tests require
-at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
-these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021,
-Toptan2020}; sensitivity depends on the timing of the test relative to the onset of
-infectiousness. After a phase-in period, all tests that are demanded will be performed.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and a household
-member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-contacts; this is why tests impact the actual contacts in
-Figure~\ref{fig:model_description}.}
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal
+whether an individual is infected or not; there is no uncertainty to the result.
+PCR tests require at least one day to be processed and there are aggregate
+capacity constraints. In contrast, rapid antigen tests yield immediate results.
+Specificity and sensitivity of these tests is set according to data analyzed in
+\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test
+relative to the onset of infectiousness  (we show our results are robust to more
+conservative assumptions in
+Section~\ref{subsec:robustness_rapid_test_sensitivity}). After a phase-in
+period, all tests that are demanded will be performed.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand.
+Schools may require staff and students to be tested regularly. Rapid tests may
+be offered by employers to on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people, showing
+symptoms of CoViD-19, and a household member having tested positively for the
+virus. We endow each agent with an individual compliance parameter. This
+parameter determines whether she takes up rapid tests.\footnote{Positive test
+    results or symptoms lead most individuals to reduce their contacts; this is why
+    tests impact the actual contacts in Figure~\ref{fig:model_description}.}
 
@@
 at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 20-30~Euros per capita \citep{Wollmershauser2021}; retail
+estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
 prices for rapid tests were below one Euro in early June 2021 and below five Euros for

### 2022-03-09 src/paper/literature_review.tex
Consistent wrapping.

@@
 
-A commonly used model class in epidemiology are agent-based simulation models.
-In a prototypical agent-based simulation model, individuals are simulated as
-moving particles. Infections take place when two particles come closer than a
-certain contact radius (e.g. \citet{Silva2020} and \citet{Cuevas2020}). While
-the simulation approach makes it easy to incorporate heterogeneity in disease
-progression, it is hard to incorporate heterogeneity in meeting patterns.
-Moreover, policies are modeled as changes in the contact radius or momentum
-equation of the particles. The translation from real policies to corresponding
-model parameters is a hard task.
+A commonly used model class in epidemiology are agent-based simulation models. In a
+prototypical agent-based simulation model, individuals are simulated as moving
+particles. Infections take place when two particles come closer than a certain contact
+radius (e.g. \citet{Silva2020} and \citet{Cuevas2020}). While the simulation approach
+makes it easy to incorporate heterogeneity in disease progression, it is hard to
+incorporate heterogeneity in meeting patterns. Moreover, policies are modeled as changes
+in the contact radius or momentum equation of the particles. The translation from real
+policies to corresponding model parameters is a hard task.
 
-These shortcomings have motivated variations of agent-based simulation models
-where moving particles have been replaced by contact networks for households,
-work and random contacts. The OpenABM-Covid-19 model by \citet{Hinch2021a} and
-the model by \citet{Aleta2020} are the closest in spirit to ours.
+These shortcomings have motivated variations of agent-based simulation models where
+moving particles have been replaced by contact networks for households, work and random
+contacts. The OpenABM-Covid-19 model by \citet{Hinch2021a} and the model by
+\citet{Aleta2020} are the closest in spirit to ours.
 
-The OpenABM-Covid-19 model by \citet{Hinch2021a} also uses detailed contact
-networks for workplaces, schools and households and can evaluate the effect of
-several NPIs. The main focus of their application are contact tracing policies
-\citep{Abueg2021}. Recently they have also added support for multiple virus
-strains and vaccinations.
+The OpenABM-Covid-19 model by \citet{Hinch2021a} also uses detailed contact networks for
+workplaces, schools and households and can evaluate the effect of several NPIs. The main
+focus of their application are contact tracing policies \citep{Abueg2021}. Recently they
+have also added support for multiple virus strains and vaccinations.
 
-\citet{Aleta2020} develop an agent-based simulation model with a very high
-geographical resolution by estimating contact networks from fine grained
-mobility data for the Boston metropolitan area. They use this model to show how
-NPIs, contact tracing and PCR testing can influence the infection dynamics.
-However, they do not calibrate their model to match actual infection numbers
-which makes it more suitable to explore the general mechanics of different
-disease mitigation measures than for their quantitative evaluation.
+\citet{Aleta2020} develop an agent-based simulation model with a very high geographical
+resolution by estimating contact networks from fine grained mobility data for the Boston
+metropolitan area. They use this model to show how NPIs, contact tracing and PCR testing
+can influence the infection dynamics. However, they do not calibrate their model to
+match actual infection numbers which makes it more suitable to explore the general
+mechanics of different disease mitigation measures than for their quantitative
+evaluation.
 
@@
 workplaces and schools. The model focuses on analyzing the effect of different contact
-tracing strategies and not on modelling enacted Austrian policies over a long period
-of time.
+tracing strategies and not on modelling enacted Austrian policies over a long period of
+time.
 
@@
 
-Our model combines elements from the above models and adds several others. To the best of
-our knowledge, our model is the only one with the following features:
+Our model combines elements from the above models and adds several others. To the best
+of our knowledge, our model is the only one with the following features:
 
@@
 
-3. We model the evolution of the pandemic and all enacted policies since the start of the
-   second wave. Since the vast majority of cases has occurred in that time period and we
-   also model unobserved infections our simulations take into account that many people
-   are already immune because they have recovered from an infection and that this
+3. We model the evolution of the pandemic and all enacted policies since the start of
+the second wave. Since the vast majority of cases has occurred in that time period
+and we also model unobserved infections our simulations take into account that many
+people are already immune because they have recovered from an infection and that this
 immunity is not spread randomly across the population.
@@
 5. Our model is designed to combine information from many different data sources.
-   Examples are surveys on rapid test demand \citep{Betsch2021}, reaction to test results
-   \citep{Betsch2021}, contact diaries \citep{Mossong2008}, share of detected cases
-   \citep{Dunkelzifferradar2020} and many more.
+Examples are surveys on rapid test demand \citep{Betsch2021}, reaction to test
+results \citep{Betsch2021}, contact diaries \citep{Mossong2008}, share of detected
+cases \citep{Dunkelzifferradar2020} and many more.

### 2022-03-09 src/paper/report.tex
Consistent wrapping.

@@
 on many dimensions. The development of highly effective vaccines holds the promise of
-containment in the medium term. However, most countries find themselves many months---and
-often years---away from reaching vaccination levels that would end the pandemic or even
-protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
-importance to employ an effective mix of strategies for containing the virus. The most
-frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
-contacts between individuals. While this has allowed some countries to sustain equilibria
-with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
-    equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
-most have seen large fluctuations of infection rates over time. Containment measures have
-become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
-tracing. Neither these policies' effects nor the influence of seasonal patterns or of
-more infectious virus strains are well understood in quantitative terms.
+containment in the medium term. However, most countries find themselves many
+months---and often years---away from reaching vaccination levels that would end the
+pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
+of utmost importance to employ an effective mix of strategies for containing the virus.
+The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
+to reduce contacts between individuals. While this has allowed some countries to sustain
+equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
+    theoretical equilibrium at low case numbers which is sustained with
+    test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
+over time. Containment measures have become increasingly diverse and now include rapid
+testing, more nuanced NPIs, and contact tracing. Neither these policies' effects nor the
+influence of seasonal patterns or of more infectious virus strains are well understood
+in quantitative terms.
 
@@
 At the core of our agent-based model \citep[][we review more literature in Supplementary
-    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021a} are physical contacts between
-heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
-an infectious individual and somebody susceptible to the disease bears the risk of
-transmitting the virus. Contacts occur in up to four networks: Within the household, at
-work, at school, or in other settings (leisure activities, grocery shopping, medical
-appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
-applications can take the population and household structure from census data and the
-network-specific frequencies of contacts from diary data measuring contacts before the
-pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
+    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021a} are physical contacts
+between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each
+contact between an infectious individual and somebody susceptible to the disease
+bears the risk of transmitting the virus. Contacts occur in up to four networks:
+Within the household, at work, at school, or in other settings (leisure activities,
+grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
+occur at random. Empirical applications can take the population and household
+structure from census data and the network-specific frequencies of contacts from
+diary data measuring contacts before the pandemic
+\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
 depend on age and geographical location (see Supplementary
@@
 upon having physical contact. The effect of different NPIs will generally vary across
-contact types. For example, a mandate to work from
-home will reduce the number of work contacts to zero for a fraction of the working
-population. Schools and daycare can be closed entirely, operate at reduced
-capacity---including an alternating schedule---, or implement mitigation measures like
-masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
-of contacts in settings outside of
+contact types. For example, a mandate to work from home will reduce the number of work
+contacts to zero for a fraction of the working population. Schools and daycare can be
+closed entirely, operate at reduced capacity---including an alternating schedule---, or
+implement mitigation measures like masking requirements or air filters
+\citep{Lessler2021}. Curfews may reduce the number of contacts in settings outside of
 work and school. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact
-\citep{Cheng2021}.
+probability of infection associated with a contact \citep{Cheng2021}.
 
@@
         A~description of the model can be found in Supplementary
-        Material~\ref{sec:supplementary_text}. Figure~\ref{fig:model_contacts_infections}
-        shows the influence of an agent's contacts to other agents on infections.
-        Demographic characteristics set the baseline number of contacts in different
-        networks ($\eta$). The agents may reduce the number of contacts due to NPIs,
-        showing symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may
-        occur when a susceptible agent meets an infectious agent; the probability depends
-        on the type of contact ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs
-        ($\rho_{c,\:t}$). If infected, the infection progresses as depicted in
-        Figure~\ref{fig:model_disease_progression}. If rapid tests are available, agents'
-        demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons trigger a
-        test only for a fraction of individuals depending on an individual compliance
-        parameter; the thresholds for triggering test demand differ across reasons and
-        they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
+        Material~\ref{sec:supplementary_text}.
+        Figure~\ref{fig:model_contacts_infections} shows the influence of an agent's
+        contacts to other agents on infections. Demographic characteristics set the
+        baseline number of contacts in different networks ($\eta$). The agents may
+        reduce the number of contacts due to NPIs, showing symptoms, or testing
+        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
+        agent meets an infectious agent; the probability depends on the type of contact
+        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
+        infected, the infection progresses as depicted in
+        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
+        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
+        trigger a test only for a fraction of individuals depending on an individual
+        compliance parameter; the thresholds for triggering test demand differ across
+        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
         Figure~\ref{fig:model_official_cases} shows the model of translating all
@@
         positive PCR tests triggered by symptoms ($\chi_{symptom}$), and the false
-        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part of
-        the graph is relevant only for periods where rapid tests are available. All
+        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
+        of the graph is relevant only for periods where rapid tests are available. All
         parameters are explained in Appendix~\ref{sub:param_tables}.}
@@
 
-In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
-age \citep{Davies2020}. A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}. We differentiate between an initial
-period of infection without being infectious or showing symptoms, being
-infectious (presymptomatic or asymptomatic), showing symptoms, requiring
-intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The
-probabilities of transitioning between these states depend on age; their
-duration is random and calibrated to medical literature (for a detailed
-description see Supplementary Material~\ref{sub:data_course_of_disease}).
-Conditional on the type of contact, infectiousness is independent of age
-\citep{Jones2021}.
+In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age
+\citep{Davies2020}. A possible infection progresses as shown in
+Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period
+of infection without being infectious or showing symptoms, being infectious
+(presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and
+recovery or death \citep[similar to][]{Aleta2020}. The probabilities of transitioning
+between these states depend on age; their duration is random and calibrated to medical
+literature (for a detailed description see Supplementary
+Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
+infectiousness is independent of age \citep{Jones2021}.
 
-The model includes several other features, which are crucial to describe the evolution of
-the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
-appear. Vaccines may become available. During the vaccine roll-out, priority may depend
-on age and occupation; vaccine hesitancy is modelled by some individuals refusing
+The model includes several other features, which are crucial to describe the evolution
+of the pandemic in 2020-2021. New virus strains with different infectiousness profiles
+may appear. Vaccines may become available. During the vaccine roll-out, priority may
+depend on age and occupation; vaccine hesitancy is modelled by some individuals refusing
 vaccination offers. With some probability, vaccinated agents become immune and do not
@@
 
-We include two types of tests. Polymerase chain reaction (PCR) tests reveal
-whether an individual is infected or not; there is no uncertainty to the result.
-PCR tests require at least one day to be processed and there are aggregate
-capacity constraints. In contrast, rapid antigen tests yield immediate results.
-Specificity and sensitivity of these tests is set according to data analyzed in
-\cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test
-relative to the onset of infectiousness  (we show our results are robust to more
-conservative assumptions in
-Section~\ref{subsec:robustness_rapid_test_sensitivity}). After a phase-in
-period, all tests that are demanded will be performed.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand.
-Schools may require staff and students to be tested regularly. Rapid tests may
-be offered by employers to on-site workers. Individuals may demand tests for
-private reasons, which include having plans to meet other people, showing
-symptoms of CoViD-19, and a household member having tested positively for the
-virus. We endow each agent with an individual compliance parameter. This
-parameter determines whether she takes up rapid tests.\footnote{Positive test
-    results or symptoms lead most individuals to reduce their contacts; this is why
-    tests impact the actual contacts in Figure~\ref{fig:model_description}.}
+We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
+individual is infected or not; there is no uncertainty to the result. PCR tests require
+at least one day to be processed and there are aggregate capacity constraints. In
+contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
+these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
+sensitivity depends on the timing of the test relative to the onset of infectiousness
+(we show our results are robust to more conservative assumptions in
+Section~\ref{subsec:robustness_rapid_test_sensitivity}). After a phase-in period, all
+tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
+model for rapid test demand. Schools may require staff and students to be tested
+regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
+demand tests for private reasons, which include having plans to meet other people,
+showing symptoms of CoViD-19, and a household member having tested positively for the
+virus. We endow each agent with an individual compliance parameter. This parameter
+determines whether she takes up rapid tests.\footnote{Positive test results or symptoms
+    lead most individuals to reduce their contacts; this is why tests impact the actual
+    contacts in Figure~\ref{fig:model_description}.}
 
@@
 mortality-based aggregate estimates of the share of detected cases and use data on the
-share of PCR tests administered because of CoViD-19
-symptoms. As the share of asymptomatic individuals
-varies by age group, this gives us age-specific shares (see
+share of PCR tests administered because of CoViD-19 symptoms. As the share of
+asymptomatic individuals varies by age group, this gives us age-specific shares (see
 Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
@@
 spread of the B.1.1.7 variant. During the steep drop in recorded cases during May 2021,
-for 42\% of the population took at least one rapid tests per week, the first-dose vaccination
-rate rose from 28\% to 43\%, and seasonality lowered the relative infectiousness
-of contacts.
+for 42\% of the population took at least one rapid tests per week, the first-dose
+vaccination rate rose from 28\% to 43\%, and seasonality lowered the relative
+infectiousness of contacts.
 
@@
         \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
-        For legibility reasons, all lines are rolling 7-day averages.}
+        are explained more thoroughly in Appendix~\ref{sub:shapley_value}. For
+        legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}
@@
 much lower pace \citep[this is in line with results in][, which our seasonality measure
-    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a quantitatively
-important factor determining the evolution of the pandemic, explaining most of the early
-changes and 43\% of the cumulative difference by the end of May.
+    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
+quantitatively important factor determining the evolution of the pandemic, explaining
+most of the early changes and 43\% of the cumulative difference by the end of May.
 
@@
         Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
-        because there were no capacity constraints for rapid tests afterwards.
-        For legibility reasons, all lines are rolling 7-day averages.}
+        because there were no capacity constraints for rapid tests afterwards. For
+        legibility reasons, all lines are rolling 7-day averages.}
 
@@
 tests at the workplace or at school come from screening the entire population. This is
-important because disadvantaged groups are less likely to be reached by testing campaigns
-relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time,
-these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
-tests at school and at the workplace will extend more into these groups. The same goes
-for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
-Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
-it arguably the most effective tool to contain the pandemic in the short run.
+important because disadvantaged groups are less likely to be reached by testing
+campaigns relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the
+same time, these groups have a higher risk to contract CoViD-19
+\citep{KochInstitut2021a}. Mandatory tests at school and at the workplace will extend
+more into these groups. The same goes for individuals who exhibit a low level of
+compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
+programmes allow a much quicker roll-out, making it arguably the most effective tool to
+contain the pandemic in the short run.

### 2022-03-09 src/paper/results/rapid_tests_statistics.tex
Consistent wrapping.

@@
 
-In order to make the most use out of limited data sources on rapid test usage, we model the
-number of performed rapid tests as a result of time invariant willingness to do rapid
-tests and time varying supply side factors and events that trigger rapid tests. Thus, the
-$\pi$ parameters governing when individuals do rapid tests described in
-Section~\ref{subsec:rapid_test_demand} are only indirectly related to the number of rapid
-tests that are actually performed in the model. When it comes to positive and negative
-rapid tests, there is even an additional layer because rapid tests are imperfectly
-sensitive and specific.
+In order to make the most use out of limited data sources on rapid test usage, we model
+the number of performed rapid tests as a result of time invariant willingness to do
+rapid tests and time varying supply side factors and events that trigger rapid tests.
+Thus, the $\pi$ parameters governing when individuals do rapid tests described in
+Section~\ref{subsec:rapid_test_demand} are only indirectly related to the number of
+rapid tests that are actually performed in the model. When it comes to positive and
+negative rapid tests, there is even an additional layer because rapid tests are
+imperfectly sensitive and specific.
 
@@
 
-We start with the share of the population doing a rapid test and receiving a positive rapid test
-over time by the channel through which the test was demanded in
-Figures~\ref{fig:rapid_test_demand_by_channel}, \ref{fig:pos_rapid_tests_by_channel}, respectively.
-Overall, the share of the population getting a rapid test on a given day increases from 2\% in mid
-March to over 10\% by May. The work rapid tests are a little ragged because of public holidays. For
-education rapid tests both vacations (first half of April) as well as the opening of schools in May
-are very visible in the rapid test demand. Overall, work tests make up the largest fraction of rapid
-tests. The image is very similar for the share of positive tests, except that the overall number of
-positive tests starts decreasing in May as rapid test expansion comes to a halt and cases fall,
-especially the positive share of private rapid tests falls as less and less individuals are
-triggered to seek a rapid test because of a risk contact in their household. The share of private
-rapid tests aligns with numbers reported by states on the demand for the free citizen tests
-\citep{NRW_rapid_tests}.
+We start with the share of the population doing a rapid test and receiving a positive
+rapid test over time by the channel through which the test was demanded in
+Figures~\ref{fig:rapid_test_demand_by_channel}, \ref{fig:pos_rapid_tests_by_channel},
+respectively. Overall, the share of the population getting a rapid test on a given day
+increases from 2\% in mid March to over 10\% by May. The work rapid tests are a little
+ragged because of public holidays. For education rapid tests both vacations (first half
+of April) as well as the opening of schools in May are very visible in the rapid test
+demand. Overall, work tests make up the largest fraction of rapid tests. The image is
+very similar for the share of positive tests, except that the overall number of positive
+tests starts decreasing in May as rapid test expansion comes to a halt and cases fall,
+especially the positive share of private rapid tests falls as less and less individuals
+are triggered to seek a rapid test because of a risk contact in their household.
 
@@
 share of them are targeted, i.e. triggered by events in the household. However, this
-does not mean that rapid tests in the workplace or at school are less important. It is rather
-the combination of large scale screening at work and in schools and very efficient
-follow up tests whenever those screening tests detected a case. Shapley values
-(Figure~\ref{fig:2021_scenarios_decomposition_tests}) take this into account and
-assign about 50\% of the overall reduction of case numbers via rapid tests to private
-rapid tests with work and school rapid tests accounting for 40\% and 7\%,
-respectively.
+does not mean that rapid tests in the workplace or at school are less important. It is
+rather the combination of large scale screening at work and in schools and very
+efficient follow up tests whenever those screening tests detected a case. Shapley values
+(Figure~\ref{fig:2021_scenarios_decomposition_tests}) take this into account and assign
+about 50\% of the overall reduction of case numbers via rapid tests to private rapid
+tests with work and school rapid tests accounting for 40\% and 7\%, respectively.
 
@@
 an even larger effect but the relevant number to compare is that up to 200 cases per
-million are detected by rapid tests every day which would have otherwise gone undetected.
+million are detected by rapid tests every day which would have otherwise gone
+undetected.
 
-This clearly shows that the large effect of rapid tests on the infection dynamic
-is not driven by unrealistic assumptions about their sensitivity but rather by the
-fact that there was a very large number of infected individuals who did not know they
-are infected. Detecting and isolating some of them is enough to slow down the
-overall infection dynamic.
+This clearly shows that the large effect of rapid tests on the infection dynamic is not
+driven by unrealistic assumptions about their sensitivity but rather by the fact that
+there was a very large number of infected individuals who did not know they are
+infected. Detecting and isolating some of them is enough to slow down the overall
+infection dynamic.
 
-\begin{figure}   % Number of True Positive / False Positive / True Negative / False Negative
+\begin{figure}   % Number of True Positive / False Positive / True Negative / False
+   Negative
    \centering
@@
    \floatfoot{\noindent \textit{Note:}
-        Each panel shows the number of rapid tests per million inhabitants that fall into the
-        respective category. Private rapid tests are especially good at detecting cases but
-        since they are often triggered by rapid tests from other channels, the other groups
-        of tests, especially rapid tests at the workplace, also play an important role for
-        containing the pandemic. All results are averaged over 30 simulation runs.
-        For legibility reasons, all lines are rolling 7-day averages.
-        }
+      Each panel shows the number of rapid tests per million inhabitants that fall
+      into the respective category. Private rapid tests are especially good at
+      detecting cases but since they are often triggered by rapid tests from other
+      channels, the other groups of tests, especially rapid tests at the workplace,
+      also play an important role for containing the pandemic. All results are
+      averaged over 30 simulation runs. For legibility reasons, all lines are rolling
+      7-day averages. }
 \end{figure}
@@
 very high. On average 60\% to 93\% of positive tests are received by individuals that
-are not infected which is broadly in line with rates later reported in the German media
-\citep{SueddeutscheZeitung2021c}. The false positive rate increases over time. This is
-due to the low prevalence of infections in the population, which falls over time. Again,
-private rapid tests are an exception with a much lower false positive rate because those
-tests are primarily demanded when there is a high likelihood of being infected. The
-false negative rate of 0.2\% looks very low. As discussed above this is deceiving and
-just a mechanical consequence of a very low prevalence of the disease and the many rapid
-tests done by non-infected people.
+are not infected. The false positive rate increases over time. This is due to the low
+prevalence of infections in the population, which falls over time. Again, private rapid
+tests are an exception with a much lower false positive rate because those tests are
+primarily demanded when there is a high likelihood of being infected. The false negative
+rate of 0.2\% looks very low. As discussed above this is deceiving and just a mechanical
+consequence of a very low prevalence of the disease and the many rapid tests done by
+non-infected people.
 
@@
    \centering
-    % \begin{subfigure}[b]{0.425\textwidth}
-    %     \centering
+   % \begin{subfigure}[b]{0.425\textwidth} \centering
    %     \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_positive_rate}
    %     \caption{Rate of True Positive Rapid Tests by Channel}
-    %     \label{fig:rapid_tests_true_positive_rate}
-    % \end{subfigure}
-    % \hfill
+   %     \label{fig:rapid_tests_true_positive_rate} \end{subfigure} \hfill
    \begin{subfigure}[b]{0.425\textwidth}
@@
    \end{subfigure}
-    % \vskip3ex
-    % \begin{subfigure}[b]{0.425\textwidth}
-    %     \centering
+   % \vskip3ex \begin{subfigure}[b]{0.425\textwidth} \centering
    % \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/true_negative_rate}
    % \caption{Rate of True Negative Rapid Tests by Channel}
-    %     \label{fig:rapid_tests_true_negative_rate}
-    % \end{subfigure}
+   % \label{fig:rapid_tests_true_negative_rate} \end{subfigure}
    \hfill
@@
    \floatfoot{\noindent \textit{Note:}
-        The left panel shows the share of positive tests that are given to people who are not
-        infected. This share is large as can be expected with a very low baseline rate of
-        positive individuals. As the incidence in the population drops, the false positive
-        rate increases. An exception are the private rapid tests because they are --
-        especially when the incidence is high -- often triggered by events that make it
-        likely that the test taker is infected and therefore their false positive rate is
-        much lower. The right panel shows the false negative rate in the population, i.e. the
-        share of negative tests done by infected individuals.
-        This is very low because there are many truly negative tests in times of low
-        incidences and large scale screening tests.}
+      The left panel shows the share of positive tests that are given to people who
+      are not infected. This share is large as can be expected with a very low
+      baseline rate of positive individuals. As the incidence in the population drops,
+      the false positive rate increases. An exception are the private rapid tests
+      because they are -- especially when the incidence is high -- often triggered by
+      events that make it likely that the test taker is infected and therefore their
+      false positive rate is much lower. The right panel shows the false negative rate
+      in the population, i.e. the share of negative tests done by infected
+      individuals. This is very low because there are many truly negative tests in
+      times of low incidences and large scale screening tests.}
 \end{figure}

### 2022-03-09 src/rev1_reply_comments.tex
Consistent wrapping.

@@
 \definecolor{darkblue}{rgb}{0, 0, 0.8}
-\hypersetup{
-    colorlinks=true,
-    linkcolor=darkblue,
-    anchorcolor=darkblue,
-    citecolor=darkblue,
-    filecolor=darkblue,
-    menucolor=darkblue,
-    runcolor=darkblue,
-    urlcolor=darkblue
-}
+\hypersetup{ colorlinks=true, linkcolor=darkblue, anchorcolor=darkblue,
+    citecolor=darkblue, filecolor=darkblue, menucolor=darkblue, runcolor=darkblue,
+    urlcolor=darkblue }
 \usepackage{enumerate}
@@
 
-% \usepackage[british]{babel}
-% \usepackage[useregional]{datetime2}
+% \usepackage[british]{babel} \usepackage[useregional]{datetime2}
 % \DTMlangsetup[en-GB]{showdayofmonth=false}
@@
 % \setlength{\marginparwidth}{2cm}
-\usepackage[
-    colorinlistoftodos,
-    backgroundcolor=white,
-    prependcaption,
-    textsize=tiny,
-    disable,
-]{todonotes}
+\usepackage[ colorinlistoftodos, backgroundcolor=white, prependcaption, textsize=tiny,
+    disable, ]{todonotes}
 
@@
 
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max}
 
 % don't let footnotes split across pages
-\interfootnotelinepenalty=10000
-\widowpenalty=10000
-\clubpenalty=10000
+\interfootnotelinepenalty=10000 \widowpenalty=10000 \clubpenalty=10000
 
@@
 \setlist[itemize]{%
-    leftmargin=\parindentaux,
-    listparindent=\parindentaux
-}
+    leftmargin=\parindentaux, listparindent=\parindentaux }
 
@@
 \makeatletter
-\newcommand*\@myenumerate{enumerate}
-\newcommand*\@myitemize{itemize}
-\newcommand{\comment}[1]{{\color{black!60}#1}}
-\newcommand{\response}[1]{{%
+\newcommand*\@myenumerate{enumerate} \newcommand*\@myitemize{itemize}
+\newcommand{\comment}[1]{{\color{black!60}#1}} \newcommand{\response}[1]{{%
 		\color{black}%
@@
 
-\renewcommand{\theequation}{\roman{equation}}
-\renewcommand{\thetable}{\Roman{table}}
-\renewcommand{\thefigure}{\Roman{figure}}
-\renewcommand{\thesection}{\Roman{section}}
+\renewcommand{\theequation}{\roman{equation}} \renewcommand{\thetable}{\Roman{table}}
+\renewcommand{\thefigure}{\Roman{figure}} \renewcommand{\thesection}{\Roman{section}}
 
@@
 \title{\large Reply to the Editor's and Reviewers' comments on: \\[2ex]
-    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
-    \large Submitted to Scientific Reports (Submission ID db917d3c-b56a-4129-8eb1-a62a716aaede)\\[-6ex]
+    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and
+    Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
+    \large Submitted to Scientific Reports (Submission ID
+    db917d3c-b56a-4129-8eb1-a62a716aaede)\\[-6ex]
 }
@@
 
-\comment{
-The title should have a clear, precise scientific meaning and should not contain a colon. Where possible, the title should be read as one concise sentence.
-}
-\vskip1ex
-\response{%
-    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic''
-}
+\comment{ The title should have a clear, precise scientific meaning and should not
+contain a colon. Where possible, the title should be read as one concise sentence. }
+\vskip1ex \response{%
+    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2:
+    Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic
+    Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic'' }
 
 \section*{Reviewer 1}
-\comment{
-The authors of the manuscript presented here have used mathematical models to examine the impact of different measures to contain the SARS-CoV-2 pandemic.
-Using epidemiological data from Germany, the contribution of rapid antigen testing is considered in particular.
+\comment{ The authors of the manuscript presented here have used mathematical models to
+examine the impact of different measures to contain the SARS-CoV-2 pandemic. Using
+epidemiological data from Germany, the contribution of rapid antigen testing is
+considered in particular.
 
-The topic studied here is of great importance for the control of the coronavirus pandemic.
-The presentation of the scientific methodology is sufficiently detailed, especially with the supplementary materials.
-It allows the reader to follow the research process in detail.
-However, I noticed a significant deficiency when reviewing the submitted work.
-Therefore, the manuscript should definitely be revised again before publication: A crucial point of the authors is to point out the special benefit of a rapid test strategy.
-However, for the calculation of the corresponding influencing factors in their mathematical model, the authors simplify, specifically, the sensitivity of rapid antigen tests in an unacceptable way.
-In the development of the mathematical model, apparently only the knowledge available a few months ago on the performance of rapid antigen tests was taken into account.
-This is evident from literature references mentioned above.
-In part, incorrect references are also used, especially for the assumption of the high sensitivity of the rapid tests.
-For example, reference No.
-74 explicitly does not refer to rapid antigen tests, but to an immunological laboratory test method (Quidel SARS Sofia antigen FIA).
-Especially in light of the benefit of rapid antigen tests described by the authors, it should be considered that many of the commercially marketed tests show only extremely poor performance.
-This has recently been demonstrated by several publications:
+The topic studied here is of great importance for the control of the coronavirus
+pandemic. The presentation of the scientific methodology is sufficiently detailed,
+especially with the supplementary materials. It allows the reader to follow the research
+process in detail. However, I noticed a significant deficiency when reviewing the
+submitted work. Therefore, the manuscript should definitely be revised again before
+publication: A crucial point of the authors is to point out the special benefit of a
+rapid test strategy. However, for the calculation of the corresponding influencing
+factors in their mathematical model, the authors simplify, specifically, the sensitivity
+of rapid antigen tests in an unacceptable way. In the development of the mathematical
+model, apparently only the knowledge available a few months ago on the performance of
+rapid antigen tests was taken into account. This is evident from literature references
+mentioned above. In part, incorrect references are also used, especially for the
+assumption of the high sensitivity of the rapid tests. For example, reference No. 74
+explicitly does not refer to rapid antigen tests, but to an immunological laboratory
+test method (Quidel SARS Sofia antigen FIA). Especially in light of the benefit of rapid
+antigen tests described by the authors, it should be considered that many of the
+commercially marketed tests show only extremely poor performance. This has recently been
+demonstrated by several publications:
 \begin{itemize}
-    \item     Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic tests for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer, Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and Corman, Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange, Constanze and Emmerich, Petra and Müller, Michael and Knauer, Olivia and Nübling, C Micha, Eurosurveillance, 26, 2100441 (2021), \url{https://doi.org/10.2807/1560-7917.ES.2021.26.44.2100441}
-    \item Assessment of SARS-CoV-2 rapid antigen tests, Özcürümez, Mustafa, Katsounas, Antonios, Holdenrieder, Stefan, von Meyer, Alexander, Renz, Harald and Wölfel, Roman, Journal of Laboratory Medicine, vol.
-45, no.
-3, 2021, pp.
-143-148.
+    \item     Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic
+          tests for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer,
+          Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and Corman,
+          Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange, Constanze and
+          Emmerich, Petra and Müller, Michael and Knauer, Olivia and Nübling, C Micha,
+          Eurosurveillance, 26, 2100441 (2021),
+          \url{https://doi.org/10.2807/1560-7917.ES.2021.26.44.2100441}
+    \item Assessment of SARS-CoV-2 rapid antigen tests, Özcürümez, Mustafa, Katsounas,
+          Antonios, Holdenrieder, Stefan, von Meyer, Alexander, Renz, Harald and Wölfel,
+          Roman, Journal of Laboratory Medicine, vol. 45, no. 3, 2021, pp. 143-148.
           \url{https://doi.org/10.1515/labmed-2021-0036}
@@
 
-Especially in the case of a mass application of antigen tests, it is therefore to be expected that products with lowest sensitivity will also be used in the general population.
-The authors should at least model this with an alternative calculation and discuss it accordingly.
-Such an improvement would increase the importance of this manuscript enormously.
+Especially in the case of a mass application of antigen tests, it is therefore to be
+expected that products with lowest sensitivity will also be used in the general
+population. The authors should at least model this with an alternative calculation and
+discuss it accordingly. Such an improvement would increase the importance of this
+manuscript enormously.
 
-Furthermore, I recommend including a declaration of potential conflicts of interest in the manuscript.
-This may be particularly important in light of the significance of the findings from the study for the continued operation of industrial companies.
+Furthermore, I recommend including a declaration of potential conflicts of interest in
+the manuscript. This may be particularly important in light of the significance of the
+findings from the study for the continued operation of industrial companies.
 
@@
 \begin{enumerate}
-    \item In part, incorrect references are also used, especially for the assumption of the high sensitivity of the rapid tests.
-For example, reference No.
-74 explicitly does not refer to rapid antigen tests, but to an immunological laboratory test method (Quidel SARS Sofia antigen FIA).
+    \item In part, incorrect references are also used, especially for the assumption of
+          the high sensitivity of the rapid tests. For example, reference No. 74 explicitly
+          does not refer to rapid antigen tests, but to an immunological laboratory test
+          method (Quidel SARS Sofia antigen FIA).
 
           \response{%
-              We thank the reviewer for spotting this mistake on our part. We removed the reference and only use current information as detailed in the next item.
-          }
+              We thank the reviewer for spotting this mistake on our part. We removed
+              the reference and only use current information as detailed in the next item. }
 
-    \item In the development of the mathematical model, apparently only the knowledge available a few months ago on the performance of rapid antigen tests was taken into account.
+    \item In the development of the mathematical model, apparently only the knowledge
+          available a few months ago on the performance of rapid antigen tests was taken into
+          account.
 
-          The authors should at least model this with an alternative calculation and discuss it accordingly. Such an improvement would increase the importance of this manuscript enormously.
+          The authors should at least model this with an alternative calculation and
+          discuss it accordingly. Such an improvement would increase the importance of
+          this manuscript enormously.
 
           \response{%
-              We agree that the references we used have become outdated in the meantime (we submitted the paper in  August 2021).
-Indeed, the numbers we used turned out to be too optimistic in light of newer studies.
+              We agree that the references we used have become outdated in the meantime
+              (we submitted the paper in  August 2021). Indeed, the numbers we used
+              turned out to be too optimistic in light of newer studies.
 
-              We therefore ran additional simulations with updated and more conservative assumptions on the sensitivity of rapid tests (Appendix B.12) and found that our main results are remarkably robust to this change.
-The percentage of the decrease in spring 2021 attributed to rapid tests decreases by only 4 percentage points even in the least favourable scenario.
-We describe our approach for calculating the adjusted sensitivities in the next answer.
-          }
-    \item I recommend including a declaration of potential conflicts of interest in the manuscript.
+              We therefore ran additional simulations with updated and more conservative
+              assumptions on the sensitivity of rapid tests (Appendix B.12) and found that
+              our main results are remarkably robust to this change. The percentage of the
+              decrease in spring 2021 attributed to rapid tests decreases by only 4
+              percentage points even in the least favourable scenario. We describe our
+              approach for calculating the adjusted sensitivities in the next answer. }
+    \item I recommend including a declaration of potential conflicts of interest in the
+          manuscript.
           \response{We have added such a declaration. There are no competing interests.}
@@
 
-\comment{
-This is a comprehensive and important study that highlights the relevance of testing to stop the spread of the pandemic.
+\comment{ This is a comprehensive and important study that highlights the relevance of
+testing to stop the spread of the pandemic.
 
-I cannot comment on the validity of the underlaying mathematical analysis, but I would like to comment on some assumptions made, which I think require further clarification or discussion.
+I cannot comment on the validity of the underlaying mathematical analysis, but I would
+like to comment on some assumptions made, which I think require further clarification or
+discussion.
 
 \begin{enumerate}
-    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful if behavior of people is already changed by the fact there is a pandemic?
+    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful
+          if behavior of people is already changed by the fact there is a pandemic?
           \response{%
 
-              Other models also use the Mossong data: OpenABM-Covid-19 model (Hinch2020), Bicher2021.
+              Other models also use the Mossong data: OpenABM-Covid-19 model
+              (Hinch2020), Bicher2021.
 
-              We only use the pre-pandemic diary data as the starting point.
-Our policies, which include non-mandated individual behavior adjustments, then reduce contacts taking the type of contact into account and for the most part not randomly but in very specific ways.
-This allows us to account for behavior reactions and policy effects that target different types of contacts differently.
-For example, household contacts continue to take place no matter the policy environment.
-Work from home does not keep random individuals at home but every individual that works has a time-constant value that determines when he or she will stay home.
-Which share of the population stays home is taken from Google mobility data.
-In the education sector we do not use the Mossong data but use completely mechanistic models.
-This gives us confidence that we can sufficiently accurately represent the contact patterns during the pandemic and account for changes induced by the pandemic.
+              We only use the pre-pandemic diary data as the starting point. Our
+              policies, which include non-mandated individual behavior adjustments, then
+              reduce contacts taking the type of contact into account and for the most
+              part not randomly but in very specific ways. This allows us to account for
+              behavior reactions and policy effects that target different types of
+              contacts differently. For example, household contacts continue to take
+              place no matter the policy environment. Work from home does not keep
+              random individuals at home but every individual that works has a
+              time-constant value that determines when he or she will stay home. Which
+              share of the population stays home is taken from Google mobility data. In
+              the education sector we do not use the Mossong data but use completely
+              mechanistic models. This gives us confidence that we can sufficiently
+              accurately represent the contact patterns during the pandemic and account
+              for changes induced by the pandemic.
 
-              Also, our excellent fit for most age groups gives us confidence that we manage to match actual meeting patterns.
-If we were far off for example for the number of work contacts, we would not be able to match either the infections of individuals in the working age (35-59) or of those that mostly do not work anymore (60-79).
+              Also, our excellent fit for most age groups gives us confidence that we
+              manage to match actual meeting patterns. If we were far off for example
+              for the number of work contacts, we would not be able to match either the
+              infections of individuals in the working age (35-59) or of those that
+              mostly do not work anymore (60-79).
 
@@
 
-              We account for individual responses in behaviour to the pandemic through a time-variant multiplier that reduces the number of other (i.e.
-non-work, non-household, non-educational) contacts.
-The values at change points that mark policy announcements or fatigue effects are estimated.
-They are detailed in Section A.9.
+              We account for individual responses in behaviour to the pandemic through a
+              time-variant multiplier that reduces the number of other (i.e. non-work,
+              non-household, non-educational) contacts. The values at change points that
+              mark policy announcements or fatigue effects are estimated. They are
+              detailed in Section A.9.
 
-              Since we estimate both the infection probability of this contact type and the multiplier, one of them has to be normalised.
-We decided to normalise the multiplier to 1 in October 2020.
-This may appear as if we assumed that pre-pandemic behaviour had resumed during that time in Germany.
-However, this is not the case.
-In fact, we would achieve the exact same fit if we doubled the infection probability and halved the multiplier.
-This is because halving the multiplier means only half the number of contacts take place.
-However, because of the doubling of the infection probability, in expectation we would arrive at the same number of infections in that contact category.
-          }
-    \item They state that susceptibility is dependent on age.
-This requires references and needs to be further explained.
-Also, younger populations are highly susceptible, they may just not contribute transmission as much.
-On that same note, infectiousness may not be independent of age
+              Since we estimate both the infection probability of this contact type and
+              the multiplier, one of them has to be normalised. We decided to normalise the
+              multiplier to 1 in October 2020. This may appear as if we assumed that
+              pre-pandemic behaviour had resumed during that time in Germany. However, this
+              is not the case. In fact, we would achieve the exact same fit if we doubled
+              the infection probability and halved the multiplier. This is because halving
+              the multiplier means only half the number of contacts take place. However,
+              because of the doubling of the infection probability, in expectation we would
+              arrive at the same number of infections in that contact category. }
+    \item They state that susceptibility is dependent on age. This requires references
+          and needs to be further explained. Also, younger populations are highly susceptible,
+          they may just not contribute transmission as much. On that same note, infectiousness
+          may not be independent of age
 
-          Edward Goldstein, Marc Lipsitch, Muge Cevik, On the Effect of Age on the Transmission of SARS-CoV-2 in Households, Schools, and the Community, The Journal of Infectious Diseases, Volume 223, Issue 3, 1 February 2021, Pages 362–369, \url{https://doi.org/10.1093/infdis/jiaa691}
+          Edward Goldstein, Marc Lipsitch, Muge Cevik, On the Effect of Age on the
+          Transmission of SARS-CoV-2 in Households, Schools, and the Community, The
+          Journal of Infectious Diseases, Volume 223, Issue 3, 1 February 2021, Pages
+          362–369, \url{https://doi.org/10.1093/infdis/jiaa691}
 
           \response{%
-              We now cite our main reference for the susceptibilities for each age group also in the main text.
-We directly take the susceptibilities given by Davies et al.
-2020, in their Extended Data Fig.
-4.
-              We base our assumption on infectiousness being independent of age on Jones et al.
-2021.
-They report only slight differences in viral loads between German adults and children: ``The least infectious youngest children have 78\% (61, 94) of the peak culture probability of adults aged 45 to 55''.
-Furthermore, they caution that these lower numbers may be due to different sample taking - such as the usage of smaller pediatric swab devices - rather than actual lower viral loads and conclude that the differences, if existent, are likely not clinically relevant.
-          }
-    \item It is not accurate that the vaccines stop transmission.
-This has to be adapted accordingly.
-Also, the authors should state which vaccines they based this analysis on, as in which ones were in use in Germany at that time and how does that affect the relative transmissibility as it cannot be assumed that all vaccines will have the same effect.
-          \response{%
-              We agree that vaccines do not stop transmission completely.
-We simplify the way vaccines reduce transmission risk considerably by assuming that 75\% of individuals achieve sterile immunity through a vaccination while the remaining 25\% remain fully susceptible.
-We are comfortable with this assumption for several reasons:
-              Firstly, the existing evidence supports the assumption that all four vaccines approved in the EU (Biontech/Pfizer, Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of protection in the first few months after vaccination and against the Alpha variant (for details see the last third of appendix A.1), which is the only time period and variant our simulations cover.
-              Secondly, given the surprisingly good performance of rapid tests we tried to err on the side of being too optimistic with our modelling of vaccine performance.
-This is  why we chose the perfect protection from vaccination for a large part of the population.
-              Thirdly, reinfections are not a big issue for the time period that we look at.
-The incidences were small and the chance of more than one exposure to Covid are negligibly small.
-Thus, the value added by a probabilistic model of vaccination protection is very small for the period that we cover.
-          }
+              We now cite our main reference for the susceptibilities for each age group
+              also in the main text. We directly take the susceptibilities given by
+              Davies et al. 2020, in their Extended Data Fig. 4. We base our assumption
+              on infectiousness being independent of age on Jones et al. 2021. They
+              report only slight differences in viral loads between German adults and
+              children: ``The least infectious youngest children have 78\% (61, 94) of
+              the peak culture probability of adults aged 45 to 55''. Furthermore, they
+              caution that these lower numbers may be due to different sample taking -
+              such as the usage of smaller pediatric swab devices - rather than actual
+              lower viral loads and conclude that the differences, if existent, are
+              likely not clinically relevant. }
+    \item It is not accurate that the vaccines stop transmission. This has to be adapted
+          accordingly. Also, the authors should state which vaccines they based this
+          analysis on, as in which ones were in use in Germany at that time and how does
+          that affect the relative transmissibility as it cannot be assumed that all
+          vaccines will have the same effect. \response{%
+              We agree that vaccines do not stop transmission completely. We simplify
+              the way vaccines reduce transmission risk considerably by assuming that
+              75\% of individuals achieve sterile immunity through a vaccination while
+              the remaining 25\% remain fully susceptible. We are comfortable with this
+              assumption for several reasons: Firstly, the existing evidence supports
+              the assumption that all four vaccines approved in the EU (Biontech/Pfizer,
+              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
+              protection in the first few months after vaccination and against the Alpha
+              variant (for details see the last third of appendix A.1), which is the
+              only time period and variant our simulations cover. Secondly, given the
+              surprisingly good performance of rapid tests we tried to err on the side
+              of being too optimistic with our modelling of vaccine performance. This is
+              why we chose the perfect protection from vaccination for a large part of
+              the population. Thirdly, reinfections are not a big issue for the time
+              period that we look at. The incidences were small and the chance of more
+              than one exposure to Covid are negligibly small. Thus, the value added by
+              a probabilistic model of vaccination protection is very small for the
+              period that we cover. }
 \end{enumerate}
@@
 
-% \setstretch{1}
-% \bibliographystyle{ecta}
-% \bibliography{refs}
+% \setstretch{1} \bibliographystyle{ecta} \bibliography{refs}
 

### 2022-03-09 src/rev1_reply_comments.tex
Shorten title a little bit, smaller formatting changes, include external doc.

@@
 % \setlength{\marginparwidth}{2cm}
-\usepackage[ colorinlistoftodos, backgroundcolor=white, prependcaption, textsize=tiny,
-    disable, ]{todonotes}
+\usepackage[colorinlistoftodos, backgroundcolor=white, prependcaption, textsize=tiny,
+    disable]{todonotes}
 
@@
 
+\externaldocument{paper}
+
 \renewcommand{\theequation}{\roman{equation}} \renewcommand{\thetable}{\Roman{table}}
@@
 \title{\large Reply to the Editor's and Reviewers' comments on: \\[2ex]
-    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and
-    Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
-    \large Submitted to Scientific Reports (Submission ID
+    \LARGE The Effectiveness of Testing, Vaccinations, and Contact Restrictions for
+    Containing the Covid-19 Pandemic\\[2ex]
+    \large Submitted to Scientific Reports \\(Submission ID
     db917d3c-b56a-4129-8eb1-a62a716aaede)\\[-6ex]
@@
     We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2:
-    Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic
-    Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic'' }
+    Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Testing, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic''
+}
 
 \section*{Reviewer 1}
-\comment{ The authors of the manuscript presented here have used mathematical models to
+\comment{The authors of the manuscript presented here have used mathematical models to
 examine the impact of different measures to contain the SARS-CoV-2 pandemic. Using
@@
 \begin{itemize}
-    \item     Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic
-          tests for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer,
-          Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and Corman,
-          Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange, Constanze and
-          Emmerich, Petra and Müller, Michael and Knauer, Olivia and Nübling, C Micha,
-          Eurosurveillance, 26, 2100441 (2021),
+    \item Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic tests
+          for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer,
+          Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and
+          Corman, Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange,
+          Constanze and Emmerich, Petra and Müller, Michael and Knauer, Olivia and
+          Nübling, C Micha, Eurosurveillance, 26, 2100441 (2021),
           \url{https://doi.org/10.2807/1560-7917.ES.2021.26.44.2100441}
@@
           \response{%
-              We thank the reviewer for spotting this mistake on our part. We removed
-              the reference and only use current information as detailed in the next item. }
+              We would like to thank the reviewer for spotting this mistake of ours. We
+              removed the reference.}
 
@@
           \response{%
-              We agree that the references we used have become outdated in the meantime
-              (we submitted the paper in  August 2021). Indeed, the numbers we used
-              turned out to be too optimistic in light of newer studies.
+              We agree that the references we used have become outdated in the meantime.
+              Note that we submitted the original manuscript in early September 2021;
+              the calculations had been done some time before that.
 
               We therefore ran additional simulations with updated and more conservative
-              assumptions on the sensitivity of rapid tests (Appendix B.12) and found that
+              assumptions on the sensitivity of rapid tests (Appendix~\ref{subsec:robustness_rapid_test_sensitivity}) and found that
               our main results are remarkably robust to this change. The percentage of the

### 2022-03-09 task_compile_documents.py

@@ -27,22 +27,6 @@ def task_compile_documents():
     pass
 
 
-@pytask.mark.latex(
-    [
-        "--pdf",
-        "--interaction=nonstopmode",
-        "--synctex=1",
-        "--cd",
-        "--shell-escape",
-        "-f",
-    ]
-)
-@pytask.mark.depends_on([SRC / "rev1_reply_comments.tex"])
-@pytask.mark.produces(BLD / "rev1_reply_comments.pdf")
-def task_compile_reply():
-    pass
-
-
 @pytask.mark.depends_on(BLD / "paper.pdf")
 @pytask.mark.produces(ROOT / "paper.pdf")
 def task_copy_pdf_to_root(depends_on, produces):
@@ -64,4 +48,26 @@ def task_extract_supplementary_material(depends_on, produces):
             produces.as_posix(),
         ]
     )
+
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on([SRC / "rev1_reply_comments.tex", BLD / "paper.pdf"])
+@pytask.mark.produces(BLD / "rev1_reply_comments.pdf")
+def task_compile_reply():
+    pass
+
+
+@pytask.mark.depends_on(BLD / "rev1_reply_comments.pdf")
+@pytask.mark.produces(ROOT / "rev1_reply_comments.pdf")
+def task_copy_pdf_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)
 

### 2022-03-09 src/paper/report.tex
Remove Smith2021 almost completely, use clipboard package.

@@
 contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
-these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
-sensitivity depends on the timing of the test relative to the onset of infectiousness
-(we show our results are robust to more conservative assumptions in
-Section~\ref{subsec:robustness_rapid_test_sensitivity}). After a phase-in period, all
-tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
-model for rapid test demand. Schools may require staff and students to be tested
-regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
-demand tests for private reasons, which include having plans to meet other people,
-showing symptoms of CoViD-19, and a household member having tested positively for the
-virus. We endow each agent with an individual compliance parameter. This parameter
-determines whether she takes up rapid tests.\footnote{Positive test results or symptoms
-    lead most individuals to reduce their contacts; this is why tests impact the actual
-    contacts in Figure~\ref{fig:model_description}.}
+these tests is set according to data analyzed in
+\label{r.1.a}\Copy{r.1.a}{\cite{Bruemmer2021,Scheiblauer2021,Özcürümez2021};
+    sensitivity depends on the timing of the test relative to the onset of infectiousness.}
+After a phase-in period, all tests that are demanded will be performed.
+Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
+require staff and students to be tested regularly. Rapid tests may be offered by
+employers to on-site workers. Individuals may demand tests for private reasons, which
+include having plans to meet other people, showing symptoms of CoViD-19, and a household
+member having tested positively for the virus. We endow each agent with an individual
+compliance parameter. This parameter determines whether she takes up rapid
+tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
+    contacts; this is why tests impact the actual contacts in
+    Figure~\ref{fig:model_description}.}
 

### 2022-03-09 src/rev1_reply_comments.tex
Remove Smith2021 almost completely, use clipboard package.

@@
 \usepackage[T1]{fontenc}
-\usepackage{afterpage,rotating,natbib,graphicx,setspace,xcolor}
+\usepackage{afterpage,rotating,graphicx,setspace,xcolor}
 \usepackage{longtable,booktabs,tabularx}
@@
 \usepackage{clipboard}
+\input{templates/Preamble_BibLaTeX_Science}
+\openclipboard{clipboard-rev1}
 
@@
 
+\bibliography{references.bib}
+
 \begin{document}
@@
 
+\vskip1ex
+
+\response{%
+    We would like to thank the reviewer for these insightful comments. First of all, we
+    agree that the references we used have become outdated in the meantime. Note that we
+    submitted the original manuscript in early September 2021; the analysis had been
+    finalised in June already.
+
+    Many thanks for pointing out that we misclassified the Quidel SARS Sofia antigen
+    FIA test; we removed the reference.
+
+    Indeed, in the line of the additional evidence, the numbers we use for test
+    sensitivity seem to be in the upper part of plausible values. On
+    page~\pageref{r.1.a} of the paper, we cite the two references you pointed out (numbering of references as in this document):
+
+    \textit{\Paste{r.1.a}}
+
+    We therefore followed
+    your suggestion and ran additional analyses with more conservative assumptions on
+    the sensitivity of rapid tests. We report on these in
+    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}. Our results are
+    remarkably robust to this change. We write:
+
+    The percentage of the decrease in spring 2021
+    attributed to rapid tests decreases by only 4 percentage points even in the least
+    favourable scenario. We describe our approach for calculating the adjusted
+    sensitivities in the next answer. }
+
 Furthermore, I recommend including a declaration of potential conflicts of interest in
@@
 
-\vskip1ex
-
-\response{Many thanks for these insightful comments. We respond in three parts.}
-
-\begin{enumerate}
-    \item In part, incorrect references are also used, especially for the assumption of
-          the high sensitivity of the rapid tests. For example, reference No. 74 explicitly
-          does not refer to rapid antigen tests, but to an immunological laboratory test
-          method (Quidel SARS Sofia antigen FIA).
-
-          \response{%
-              We would like to thank the reviewer for spotting this mistake of ours. We
-              removed the reference.}
-
-    \item In the development of the mathematical model, apparently only the knowledge
-          available a few months ago on the performance of rapid antigen tests was taken into
-          account.
-
-          The authors should at least model this with an alternative calculation and
-          discuss it accordingly. Such an improvement would increase the importance of
-          this manuscript enormously.
-
-          \response{%
-              We agree that the references we used have become outdated in the meantime.
-              Note that we submitted the original manuscript in early September 2021;
-              the calculations had been done some time before that.
-
-              We therefore ran additional simulations with updated and more conservative
-              assumptions on the sensitivity of rapid tests (Appendix~\ref{subsec:robustness_rapid_test_sensitivity}) and found that
-              our main results are remarkably robust to this change. The percentage of the
-              decrease in spring 2021 attributed to rapid tests decreases by only 4
-              percentage points even in the least favourable scenario. We describe our
-              approach for calculating the adjusted sensitivities in the next answer. }
-    \item I recommend including a declaration of potential conflicts of interest in the
-          manuscript.
 \response{We have added such a declaration. There are no competing interests.}
 
-\end{enumerate}
 }
 
-\section{Reviewer 2}
+\clearpage
+\section*{Reviewer 2}
 
@@
 
-% \setstretch{1} \bibliographystyle{ecta} \bibliography{refs}
+\begin{refcontext}[sorting=none]  % Sort BIBLIOGRAPHY by appearance.
+    \printbibliography[heading=bibintoc]
+\end{refcontext}
 

### 2022-03-09 task_compile_documents.py

@@ -27,10 +27,21 @@ def task_compile_documents():
     pass
 
 
+@pytask.mark.skipif(shutil.which("qpdf") is None, reason="Extraction needs qpdf.")
 @pytask.mark.depends_on(BLD / "paper.pdf")
-@pytask.mark.produces(ROOT / "paper.pdf")
-def task_copy_pdf_to_root(depends_on, produces):
-    shutil.copy(depends_on, produces)
+@pytask.mark.produces(BLD / "paper-no-appendix.pdf")
+def task_extract_paper_body(depends_on, produces):
+    subprocess.run(
+        [
+            "qpdf",
+            "--empty",
+            "--pages",
+            depends_on.as_posix(),
+            "1-11",
+            "--",
+            produces.as_posix(),
+        ]
+    )
 
 
 @pytask.mark.skipif(shutil.which("qpdf") is None, reason="Extraction needs qpdf.")
@@ -48,6 +59,24 @@ def task_extract_supplementary_material(depends_on, produces):
             produces.as_posix(),
         ]
     )
+
+
+@pytask.mark.depends_on(BLD / "paper.pdf")
+@pytask.mark.produces(ROOT / "paper.pdf")
+def task_copy_paper_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)
+
+
+@pytask.mark.depends_on(BLD / "paper-no-appendix.pdf")
+@pytask.mark.produces(ROOT / "paper-no-appendix.pdf")
+def task_copy_paper_no_appendix_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)
+
+
+@pytask.mark.depends_on(BLD / "supplementary_material.pdf")
+@pytask.mark.produces(ROOT / "supplementary_material.pdf")
+def task_copy_supplementary_material_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)
 
 
 @pytask.mark.latex(

### 2022-03-10 src/paper/data/course_of_disease.tex
Go back to longer version, nature submission, update references. (#6)
Mostly done except for Smith reference, switching machines.

@@
 29\% after 20 days and 7\% after 32 days. Again, we rescale this for every age group
-among those that will not survive. For survivors \cite{Hinch2020} reports a mean duration
+among those that will not survive. For survivors \cite{Hinch2021} reports a mean duration
 of 18.8 days until recovery and a standard deviation of 12.21 days. We discretize this
@@
 % vaccinations
+\label{r.2.c}
 The second route to immunity is through vaccination. Germany has mostly relied on the

### 2022-03-10 src/paper/report.tex
Mostly done except for Smith reference, switching machines.

@@
 In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age
-\citep{Davies2020}. A possible infection progresses as shown in
+\citep{Davies2020}.\label{r.2.b} A possible infection progresses as shown in
 Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period
@@
 would have been reduced by 53\% relative to the scenario without vaccinations, rapid
-tests, or seasonality.
+tests, or seasonality. \label{r.1.b}\Copy{r.1.b}{In
+    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}, we provide a detailed analysis
+    of whether our results are robust regarding the sensitivity parameters we assume for
+    rapid tests. Even if we take a very pessimistic stance, the effect is only reduced from
+    42\% to 38\%.}
 
@@
 scenarios comparing changes in NPIs and testing regimes. Two of the most contentious
-NPIs concern schools and mandates to work from home. In many countries, schools switched
-to remote instruction during the first wave, so did Germany. After the summer break,
-they were operating at full capacity with increased hygiene measures, before being
-closed again from mid-December onward. Some states started opening them gradually in
-late February, but operation at normal capacity did not resume until the beginning of
-June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
-regarding schools starting after Easter, at which point rapid tests had become widely
-available. We estimate the realized scenario to have essentially the same effect as a
-situation with closed schools. Under fully opened schools with mandatory tests, total
-infections would have been 6\% higher; this number rises to 20\% without tests. These
-effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
+NPIs concern schools and mandates to work from home.\label{ref.2.a} In many countries,
+schools switched to remote instruction during the first wave, so did Germany. After the
+summer break, they were operating at full capacity with increased hygiene measures,
+before being closed again from mid-December onward. Some states started opening them
+gradually in late February, but operation at normal capacity did not resume until the
+beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
+policies regarding schools starting after Easter, at which point rapid tests had become
+widely available. We estimate the realized scenario to have essentially the same effect
+as a situation with closed schools. Under fully opened schools with mandatory tests,
+total infections would have been 6\% higher; this number rises to 20\% without tests.
+These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
     Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of

### 2022-03-10 src/paper/results/robustness_to_sensitivity_assumptions.tex
Mostly done except for Smith reference, switching machines.

@@
 
-Our result that frequent and large scale rapid testing was responsible for a
-large decrease in cases in the spring of 2021 in Germany seems to be at odds
-with the general perception that rapid tests are unreliable. They also seem to
-contradict recent studies showing that the actual sensitivity of rapid tests t
-be lower than advertised (e.g.\cite{Scheiblauer2021}).
+\Copy{r.1.b-appendix}{
+    Our main results are based on rapid test sensitivities read from clinical trials.
+Recent studies showing that the actual sensitivity of rapid tests may be lower than that \citep[e.g.,][]{Scheiblauer2021}.
 
-This is, however, not the case. It is important to distinguish studies that
-analyze the sensitivity and specificity of rapid tests in isolation and studies
-that analyze the aggregate effect of testing strategies within an
-epidemiological model.
+    This section shows that our results are robust to making less favorable assumptions
+    on rapid test sensitivity. We proceed by describing several possible ways of
+    calibrating rapid test sensitivity profiles based on recent studies. Since none of
+    these methods is inherently better than the others, we make simulations with two
+    sensitivity profiles: The average over all methods and the lower envelope over all
+    methods.
 
-This section shows that our results are robust to making less favorable
-assumptions on rapid test sensitivity. We proceed by describing several possible
-ways of calibrating rapid test sensitivity profiles based on recent studies.
-Since none of these methods is inherently better than the others, we make
-simulations with two sensitivity profiles: The average over all methods and the
-lower envelope over all methods.
-
-Both profiles imply much lower sensitivities of rapid tests than used in our
-main results, especially in the later stage of an infection. However, the main
-results stay very robust. The original result was that rapid tests, seasonality
-and vaccinations are responsible for 42\%, 43\% and 16\%, respectively. With the
-average profile, the effect of rapid tests goes down to 41 \%. With the lower
-envelope profile, which is an extremely unfavorable assumption, it becomes 38\%.
+    Both profiles imply lower sensitivities of rapid tests than used in our main
+    results. This is especially true during the later stage of an infection. However,
+    the main results stay very robust. The original result was that rapid tests,
+    seasonality and vaccinations are responsible for 42\%, 43\% and 16\%, respectively.
+    With the average profile, the effect of rapid tests decreases to 41 \%. With the
+    lower envelope profile, which is an extremely unfavorable assumption, it becomes
+    38\%.
 
     % what do we need
-The effect of rapid tests on infection dynamics strongly depends on \emph{when}
-an infection is detected. Earlier detection means that it is more likely that
-the infection has not yet been discovered for a different reason (e.g. due to
-the onset of symptoms) and that the infected person can be isolated before
-spreading the disease to others.
+    The effect of rapid tests on infection dynamics strongly depends on \emph{when} an
+    infection is detected. Earlier detection means that it is more likely that the
+    infection has not yet been discovered for a different reason (e.g. due to the onset
+    of symptoms) and that the infected person can be isolated before spreading the
+    disease to others.
 
-The sensitivity of rapid tests depends on the viral load in the repiratory
-tract. It is low at the beginning of an infection (especially before the onset
-of infectiousness), high in the first few days of infectiousness and then
-gradually decreasing towards the end of infectiousness.
+    The sensitivity of rapid tests depends on the viral load in the respiratory tract.
+    It is low at the beginning of an infection (especially before the onset of
+    infectiousness), high in the first few days of infectiousness and then gradually
+    decreasing towards the end of infectiousness.
 
-We thus need to calibrate a profile of rapid test sensitivities based on the
-number of days until or since the onset of infectiousness.
+    We thus need to calibrate a profile of rapid test sensitivities based on the number
+    of days until or since the onset of infectiousness.
 
@@
     Unfortunately, such sensitivity profiles are not usually reported in studies. We
-thus need to create them by combining two types of studies: 1. Studies that
-report the viral load in terms of threshold cycle ($Ct$) values determined by
-PCR test (e.g. \cite{Cosentino2022}, \cite{Ong2021}, \cite{Bonnet2022},
-\cite{Jang2021} and \cite{Zuin2021}). 2. Studies that report the ensitivity of
-rapid tests for different $Ct$ values (e.g. \cite{Scheiblauer2021},
-\cite{Bruemmer2021}).
+    thus need to create them by combining two types of studies: 1. Studies that report
+    the viral load in terms of threshold cycle ($Ct$) values determined by PCR test
+    \citep[e.g.][]{Cosentino2022,Ong2021,Bonnet2022,Jang2021,Zuin2021}. 2. Studies that
+    report the sensitivity of rapid tests for different $Ct$ values
+    \citep[e.g.][]{Scheiblauer2021,Bruemmer2021}.
 
     % what options do we have to configure the mappings
-It is natural to assume that the evolution of $Ct$ values over time as well as
-the effect of $Ct$ values on rapid test sensitivity are continuous functions.
-However, the results of the available studies are usually reported in a
-discretized way. This leads to multiple ways of calculating the sensitivity
-profiles. Some try to recover the underlying continuous functions using
-interpolation or regression, others simply use the discretized values.
+    It is natural to assume that the evolution of $Ct$ values over time as well as the
+    effect of $Ct$ values on rapid test sensitivity are continuous functions. However,
+    the results of the available studies are usually reported in a discretized way. This
+    leads to multiple ways of calculating the sensitivity profiles. Some try to recover
+    the underlying continuous functions using interpolation or regression, others simply
+    use the discretized values.
 
     % Describe CT value formula vs. bin averages
-For the calibration of $Ct$ values over time we can either use discretized
-values for several time bins from \cite{Ong2021} and \cite{Jang2021}.
-Alternatively, we can use linearized formulas for calculating sensitivities over
-time \cite{Cosentino2022} and complement it with interpolations of data points
-from \cite{Jang2021} in the pre-infectious stage. Throughout we assume that the
-$Ct$ values of individuals who eventually develop symptoms and those who do not
-follow the same trajectory. This is in line with results by \cite{Zuin2021} and
-recent evidence that rapid tests excel at discovering asymptomatic cases
-\cite{Rosella2022}.
+    For the calibration of $Ct$ values over time we can either use discretized values
+    for several time bins from \cite{Ong2021} and \cite{Jang2021}. Alternatively, we can
+    use linearized formulas for calculating sensitivities over time \cite{Cosentino2022}
+    and complement it with interpolations of data points from \cite{Jang2021} in the
+    pre-infectious stage. Throughout we assume that the $Ct$ values of individuals who
+    eventually develop symptoms and those who do not follow the same trajectory. This is
+    in line with results by \cite{Zuin2021} and recent evidence that rapid tests excel
+    at discovering asymptomatic cases \cite{Rosella2022}.
 
@@
     options. First, we can simply look up the discretized values for the three $Ct$
-provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35). Secondly,
-we can use linear regression to estimate a continous mapping for the
-relationship by assuming that the $Ct$ values of each bin are achieved exactly
-at the bin midpoints and that the relationship is linear.
+    provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35). Secondly, we
+    can use linear regression to estimate a continous mapping for the relationship by
+    assuming that the $Ct$ values of each bin are achieved exactly at the bin midpoints
+    and that the relationship is linear.
 
@@
     sensitivities and an overestimation of very low sensitivities. This is because
-discretization is essentially a smoothing device. On the other hand, it hat the
-advantage of simply working with published results, without introducing any
-tuning parameters or other assumptions.
+    discretization is essentially a smoothing device. On the other hand, it has the
+    advantage of simply working with published results, without introducing any tuning
+    parameters or other assumptions.
 
@@
         \caption{Rapid test sensitivity profiles}
-    \floatfoot{\noindent \textit{Note:} The figure shows estimated sensitivities
-    of rapid tests over the course of an infection. The x-axis shows days
-    relative to the onset of infectiousness. The y-axis shows the estimated
-    rapid test sensitivities. The grey lines are the raw sensitivity estimates
-    obtained with different methods of dealing with discretized data. The blue
-    line shows their average and the yellow line their lower envelope. The
-    turquoise line are the sensitivities used for the main results of the paper.
-    It is much higher than the new estimates in the later stage of an infection.
-    Luckily this has only a very small effect on the overall effectiveness of
-    rapid tests on the infection dynamics.}
+        \floatfoot{\noindent \textit{Note:} The figure shows estimated sensitivities of
+            rapid tests over the course of an infection. The x-axis shows days relative
+            to the onset of infectiousness. The y-axis shows the estimated rapid test
+            sensitivities. The grey lines are the raw sensitivity estimates obtained
+            with different methods of dealing with discretized data. The blue line shows
+            their average and the yellow line their lower envelope. The turquoise line
+            are the test sensitivities used for the main results of the paper.}
         \label{fig:sensitivity_assumptions}
@@
 
-\FloatBarrier
+    % \FloatBarrier
 
     Figure \ref{fig:sensitivity_assumptions} shows that the updated sensitivity
-estimates are lower than the ones used for our original results, especially
-towards the end of an infection. However, the main results barely change.
+    estimates are lower than the ones used for our original results, especially towards
+    the end of an infection. However, the main results barely change. This is due to the
+    fact that the differences are most important towards the later stage of an
+    infection. Uncovering an infection that was previously undetected at that stage does
+    not have a large effect on infection dynamics.
 
@@
 
-    \caption{Updated shapley decompositions}
+        \caption{Shapley decompositions for different values of rapid test sensitivity}
         \floatfoot{\noindent \textit{Note:} The figure shows updated versions of the
-    shapley decomposition in figure \ref{fig:2021_scenarios_decomposition}. The
-    decomposition is based on 20 model runs with different random seeds. The
+            Shapley decomposition in figure \ref{fig:2021_scenarios_decomposition}. The
+            decompositions are based on 20 model runs with different random seeds. The
             share attributed to each channel is rounded to the next full percentage
@@
 
+}
 \FloatBarrier
-
-%

### 2022-03-10 src/rev1_reply_comments.tex
Mostly done except for Smith reference, switching machines.

@@
     Many thanks for pointing out that we misclassified the Quidel SARS Sofia antigen
-    FIA test; we removed the reference.
+    FIA test; we removed the reference.{\bf \color{red}DOUBLE-CHECK -- this is Smith2021}
 
@@
     sensitivity seem to be in the upper part of plausible values. On
-    page~\pageref{r.1.a} of the paper, we cite the two references you pointed out (numbering of references as in this document):
+    page~\pageref{r.1.a} of the paper, we cite the two references you pointed out
+    (numbering of references relates to this document):
 
@@
 
-    We therefore followed
-    your suggestion and ran additional analyses with more conservative assumptions on
-    the sensitivity of rapid tests. We report on these in
-    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}. Our results are
-    remarkably robust to this change. We write:
+    We followed your suggestion and ran additional analyses with more conservative
+    assumptions on the sensitivity of rapid tests. The results are very robust. We write
+    in the main text on page~\pageref{r.1.b}
 
-    The percentage of the decrease in spring 2021
-    attributed to rapid tests decreases by only 4 percentage points even in the least
-    favourable scenario. We describe our approach for calculating the adjusted
-    sensitivities in the next answer. }
+    \textit{\Paste{r.1.b}}
+
+    The robustness is due to the fact that over the course of an infection, the largest differences between the parameters we use in our main analysis and those based on other studies are found during the later stages of an infection.
+Uncovering an infection that was previously undetected at that stage does
+    not have a large effect on subsequent infection dynamics. Our whole analysis in
+    Appendix~\ref{subsec:robustness_rapid_test_sensitivity} reads as follows:
+
+    \textit{\Paste{r.1.b-appendix}}
+
+}
 
@@
 
-              Other models also use the Mossong data: OpenABM-Covid-19 model
-              (Hinch2020), Bicher2021.
+              Many thanks for this comment.
+Ideally, we would of course use real-time data.
+For example, several studies have used smartphone tracking data for this purpose.
+However, they are not detailed enough for our requirements, because they do not provide any information about \textit{who} is meeting with each other.
+These dimensions make up our contact networks, which are absolutely crucial.
 
-              We only use the pre-pandemic diary data as the starting point. Our
-              policies, which include non-mandated individual behavior adjustments, then
-              reduce contacts taking the type of contact into account and for the most
-              part not randomly but in very specific ways. This allows us to account for
-              behavior reactions and policy effects that target different types of
-              contacts differently. For example, household contacts continue to take
-              place no matter the policy environment. Work from home does not keep
-              random individuals at home but every individual that works has a
-              time-constant value that determines when he or she will stay home. Which
-              share of the population stays home is taken from Google mobility data. In
-              the education sector we do not use the Mossong data but use completely
-              mechanistic models. This gives us confidence that we can sufficiently
-              accurately represent the contact patterns during the pandemic and account
-              for changes induced by the pandemic.
+              Furthermore, note that we only use the pre-pandemic diary data as the
+              starting point. Contact restrictions and individual behavior adjustments
+              reduce contacts. These are active throughout the period of analysis and
+              they vary with the strictness of measure. In order to do so, it is crucial
+              to take the type of contact into account; see for example our analysis of
+              potential school closures on Page~\pageref{ref.2.a}. Closed schools mean
+              that children have far fewer and parents have somewhat fewer contacts than
+              at  and for the most part not randomly but in very specific ways. This
+              allows us to account for behavior reactions and policy effects that target
+              different types of contacts differently. For example, household contacts
+              continue to take place no matter the policy environment. Work from home
+              does not keep random individuals at home but every individual that works
+              has a time-constant value that determines when he or she will stay home.
+              Which share of the population stays home is taken from Google mobility
+              data. In the education sector we do not use the Mossong data but use
+              completely mechanistic models. This gives us confidence that we can
+              sufficiently accurately represent the contact patterns during the pandemic
+              and account for changes induced by the pandemic.
 

### 2022-03-10 src/rev1_reply_comments.tex
Answers to ref 2.

@@
           if behavior of people is already changed by the fact there is a pandemic?
+
           \response{%
@@
               that children have far fewer and parents have somewhat fewer contacts than
-              at  and for the most part not randomly but in very specific ways. This
-              allows us to account for behavior reactions and policy effects that target
-              different types of contacts differently. For example, household contacts
-              continue to take place no matter the policy environment. Work from home
-              does not keep random individuals at home but every individual that works
-              has a time-constant value that determines when he or she will stay home.
-              Which share of the population stays home is taken from Google mobility
-              data. In the education sector we do not use the Mossong data but use
-              completely mechanistic models. This gives us confidence that we can
-              sufficiently accurately represent the contact patterns during the pandemic
-              and account for changes induced by the pandemic.
+              without school closures.
 
-              Also, our excellent fit for most age groups gives us confidence that we
-              manage to match actual meeting patterns. If we were far off for example
-              for the number of work contacts, we would not be able to match either the
-              infections of individuals in the working age (35-59) or of those that
-              mostly do not work anymore (60-79).
-
-              \textbf{Answer in case there is confusion because the other multiplier is 1 in fall 2020:}
-
-              We account for individual responses in behaviour to the pandemic through a
-              time-variant multiplier that reduces the number of other (i.e. non-work,
-              non-household, non-educational) contacts. The values at change points that
-              mark policy announcements or fatigue effects are estimated. They are
-              detailed in Section A.9.
-
-              Since we estimate both the infection probability of this contact type and
-              the multiplier, one of them has to be normalised. We decided to normalise the
-              multiplier to 1 in October 2020. This may appear as if we assumed that
-              pre-pandemic behaviour had resumed during that time in Germany. However, this
-              is not the case. In fact, we would achieve the exact same fit if we doubled
-              the infection probability and halved the multiplier. This is because halving
-              the multiplier means only half the number of contacts take place. However,
-              because of the doubling of the infection probability, in expectation we would
-              arrive at the same number of infections in that contact category. }
+              Finally, the very good fit of our model to official infection rates across
+              age groups increases our confidence that we manage to match actual meeting
+              patterns. If we were far off for, for example, the number of work contacts,
+              we would not be able to match either the infections of individuals in the
+              working age (35-59) or of those that mostly do not work anymore (60-79).
+          }
     \item They state that susceptibility is dependent on age. This requires references
@@
           \response{%
-              We now cite our main reference for the susceptibilities for each age group
-              also in the main text. We directly take the susceptibilities given by
-              Davies et al. 2020, in their Extended Data Fig. 4. We base our assumption
-              on infectiousness being independent of age on Jones et al. 2021. They
-              report only slight differences in viral loads between German adults and
-              children: ``The least infectious youngest children have 78\% (61, 94) of
-              the peak culture probability of adults aged 45 to 55''. Furthermore, they
-              caution that these lower numbers may be due to different sample taking -
-              such as the usage of smaller pediatric swab devices - rather than actual
-              lower viral loads and conclude that the differences, if existent, are
-              likely not clinically relevant. }
+              Many thanks for pointing this out, indeed we had not cited our key reference for susceptibilities by age group \citep{Davies2020} in the main text. We do so now on Page~\ref{r.2.b}.
+
+              We base our assumption on infectiousness being independent of age on
+              \citep{Jones2021}. That paper reports only slight differences in viral
+              loads between German adults and children: ``The least infectious youngest
+              children have 78\% (61, 94) of the peak culture probability of adults aged
+              45 to 55''. Furthermore, they caution that these lower numbers may be due
+              to different sample taking---such as the usage of smaller pediatric swab
+              devices---rather than actual lower viral loads and conclude that the
+              differences, if existent, are likely not clinically relevant.
+
+                  {\bf ToDo: Check paper suggestion
+                      \url{https://doi.org/10.1093/infdis/jiaa691} -- are they actually checking
+                      infectiousness conditional on contact type? Or could we even reproduce
+                      what they are doing (not suggesting we should, certainly no need to spend
+                      time on it)}
+              %   In any case, while the jury is certainly still out on that particular issue, we are confident that different assumptions would not change our results much. 
+          }
     \item It is not accurate that the vaccines stop transmission. This has to be adapted
@@
           that affect the relative transmissibility as it cannot be assumed that all
-          vaccines will have the same effect. \response{%
-              We agree that vaccines do not stop transmission completely. We simplify
-              the way vaccines reduce transmission risk considerably by assuming that
-              75\% of individuals achieve sterile immunity through a vaccination while
-              the remaining 25\% remain fully susceptible. We are comfortable with this
-              assumption for several reasons: Firstly, the existing evidence supports
-              the assumption that all four vaccines approved in the EU (Biontech/Pfizer,
+          vaccines will have the same effect.
+
+          \response{%
+              Many thanks, it is of course true that vaccines do not stop transmission
+              completely. We do not assume that, either. However, we have not been able
+              to find data on the way in which this affects different population groups
+              for the wild type and the $\alpha$-variant.
+
+              In our model, we thus simplify the way vaccines reduce transmission risk.
+              In particular, we assume that that 75\% of individuals achieve sterile
+              immunity through a vaccination while the remaining 25\% remain fully
+              susceptible. SFirstly, the existing evidence supports the
+              assumption that all four vaccines approved in the EU (Biontech/Pfizer,
               Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
-              protection in the first few months after vaccination and against the Alpha
-              variant (for details see the last third of appendix A.1), which is the
-              only time period and variant our simulations cover. Secondly, given the
-              surprisingly good performance of rapid tests we tried to err on the side
-              of being too optimistic with our modelling of vaccine performance. This is
-              why we chose the perfect protection from vaccination for a large part of
-              the population. Thirdly, reinfections are not a big issue for the time
-              period that we look at. The incidences were small and the chance of more
-              than one exposure to Covid are negligibly small. Thus, the value added by
-              a probabilistic model of vaccination protection is very small for the
-              period that we cover. }
+              protection in the first few months after vaccination against the wild type
+              and the $\alpha$ variant (for details see Appendix~\ref{r.2.c}, starting
+              on Page~\pageref{r.2.c}), which is the only time period and variant our
+              simulations cover. Secondly, given the surprisingly good performance of
+              rapid tests we tried to err on the side of being too optimistic with our
+              modelling of vaccine performance. This is why we chose the perfect
+              protection from vaccination for a large part of the population. Thirdly,
+              reinfections are not a big issue for the time period that we look at. The
+              incidences were small and the chance of more than one exposure to Covid
+              are negligibly small. Thus, the value added by a probabilistic model of
+              vaccination protection is very small for the period that we cover. }
 \end{enumerate}

### 2022-03-10 src/paper/report.tex
Re-wrote answers re ref 1.

@@
 at least one day to be processed and there are aggregate capacity constraints. In
-contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
-these tests is set according to data analyzed in
-\label{r.1.a}\Copy{r.1.a}{\cite{Bruemmer2021,Scheiblauer2021,Özcürümez2021};
-    sensitivity depends on the timing of the test relative to the onset of infectiousness.}
-After a phase-in period, all tests that are demanded will be performed.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and a household
-member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-    contacts; this is why tests impact the actual contacts in
-    Figure~\ref{fig:model_description}.}
+contrast, rapid antigen tests yield immediate results.
+\label{r.1.a}\Copy{r.1.a}{Specificity and sensitivity of these tests is set according to
+data analyzed in \citet{Bruemmer2021,Scheiblauer2021,Özcürümez2021}; sensitivity depends
+on the timing of the test relative to the onset of infectiousness as in
+\citet{Smith2021}. We provide a robustness to different assumptions in
+Appendix~\ref{subsec:robustness_rapid_test_sensitivity}.} After a phase-in period, all
+tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
+model for rapid test demand. Schools may require staff and students to be tested
+regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
+demand tests for private reasons, which include having plans to meet other people,
+showing symptoms of CoViD-19, and a household member having tested positively for the
+virus. We endow each agent with an individual compliance parameter. This parameter
+determines whether she takes up rapid tests.\footnote{Positive test results or symptoms
+lead most individuals to reduce their contacts; this is why tests impact the actual
+contacts in Figure~\ref{fig:model_description}.}
 

### 2022-03-10 src/rev1_reply_comments.tex
Re-wrote answers re ref 1.

@@
     submitted the original manuscript in early September 2021; the analysis had been
-    finalised in June already.
+    finalised in June of last year. Indeed, in the line of the additional evidence, the
+    numbers we use for test sensitivity appear to be in the upper part of plausible
+    values. We have followed your suggestion and report detailed robustness analyses in
+    the paper and its Appendix.
 
-    Many thanks for pointing out that we misclassified the Quidel SARS Sofia antigen
-    FIA test; we removed the reference.{\bf \color{red}DOUBLE-CHECK -- this is Smith2021}
+    Before describing our changes, we would like to thank you for pointing out that the
+    test analysed in \citet{Smith2021} is not a rapid test. While this is true, the
+    crucial feature in \citet{Smith2021} is that the paper reports the sensitivity over
+    the course of an infection. We have not found a comparable study. Getting this
+    profile right is important particularly in the beginning of an infection, as we
+    detail shortly. At that point, the differences between different methods to assess
+    test sensitivity by timing within an infection are relatively small.
 
-    Indeed, in the line of the additional evidence, the numbers we use for test
-    sensitivity seem to be in the upper part of plausible values. On
-    page~\pageref{r.1.a} of the paper, we cite the two references you pointed out
-    (numbering of references relates to this document):
+    On page~\pageref{r.1.a} of the paper, we cite the references you pointed out
+    (numbering of references relates to this document) and note that we conduct a
+    robustness analysis:\\[1ex]
 
@@
 
+    \vskip1ex
+
     We followed your suggestion and ran additional analyses with more conservative
     assumptions on the sensitivity of rapid tests. The results are very robust. We write
-    in the main text on page~\pageref{r.1.b}
+    in the main text on page~\pageref{r.1.b}:\\[1ex]
 
@@
 
-    The robustness is due to the fact that over the course of an infection, the largest differences between the parameters we use in our main analysis and those based on other studies are found during the later stages of an infection.
-Uncovering an infection that was previously undetected at that stage does
+    \vskip1ex
+
+    The robustness is due to the fact that over the course of an infection, the largest
+    differences between the parameters we use in our main analysis and those based on
+    imputing values from other studies are found during the later stages of an
+    infection. Uncovering an infection that was previously undetected at that stage does
     not have a large effect on subsequent infection dynamics. Our whole analysis in
-    Appendix~\ref{subsec:robustness_rapid_test_sensitivity} reads as follows:
+    Appendix~\ref{subsec:robustness_rapid_test_sensitivity} reads as follows:\\[1ex]
 
@@
 
+~\\[4ex]
+
 Furthermore, I recommend including a declaration of potential conflicts of interest in
 the manuscript. This may be particularly important in light of the significance of the
-findings from the study for the continued operation of industrial companies.
+findings from the study for the continued operation of industrial companies.\\[2ex]
 
-\response{We have added such a declaration. There are no competing interests.}
+\response{We have added such a declaration at the end of the paper:\\[2ex]
+
+\textit{\Paste{r.1.c}}
+}
 
@@
 
-              Many thanks for this comment.
-Ideally, we would of course use real-time data.
-For example, several studies have used smartphone tracking data for this purpose.
-However, they are not detailed enough for our requirements, because they do not provide any information about \textit{who} is meeting with each other.
-These dimensions make up our contact networks, which are absolutely crucial.
+              Many thanks for this comment. Ideally, we would of course use real-time
+              data. For example, several studies have used smartphone tracking data for
+              this purpose. However, any such data are not detailed enough for our
+              requirements because they do not provide any information about
+              \textit{who} is meeting with each other. These dimensions make up our
+              contact networks, which are absolutely crucial.
 

### 2022-03-10 src/rev1_reply_comments.tex
Tiny updates.

@@
               age groups increases our confidence that we manage to match actual meeting
-              patterns. If we were far off for, for example, the number of work contacts,
-              we would not be able to match either the infections of individuals in the
-              working age (35-59) or of those that mostly do not work anymore (60-79).
+              patterns. If we were far off for, for example, the number of work
+              contacts, we would not be able to match either the infections of
+              individuals during working age (35-59) or of those that mostly do not work
+              anymore (60-79). 
             }
@@
           \response{%
-              Many thanks for pointing this out, indeed we had not cited our key reference for susceptibilities by age group \citep{Davies2020} in the main text. We do so now on Page~\ref{r.2.b}.
+              Many thanks for pointing this out, indeed we had not cited our key
+              reference for susceptibilities by age group \citep{Davies2020} in the main
+              text. We do so now on Page~\ref{r.2.b}.
 
               We base our assumption on infectiousness being independent of age on
-              \citep{Jones2021}. That paper reports only slight differences in viral
+              \citet{Jones2021}. That paper reports only slight differences in viral
               loads between German adults and children: ``The least infectious youngest
@@
 
-                  {\bf ToDo: Check paper suggestion
+                  {\bf \color{red} ToDo: Check paper suggestion
                       \url{https://doi.org/10.1093/infdis/jiaa691} -- are they actually checking
@@
                       time on it)}
-              %   In any case, while the jury is certainly still out on that particular issue, we are confident that different assumptions would not change our results much. 
+
+              In any case, while the jury is certainly still out on that particular
+              issue, we are confident that different assumptions would not be quantitatively
+              meaningful to change our results in a significant way.
+
           }

### 2022-03-10 src/rev1_reply_comments.tex
Incorporate Janos' comments.

@@
               reference for susceptibilities by age group \citep{Davies2020} in the main
-              text. We do so now on Page~\ref{r.2.b}.
+              text. In fact, the paper you suggested \citep{Goldstein2021} comes to
+              similar conclusions: \texit{``We found evidence that compared to
+                  younger/middle-aged adults, children younger than 10 years have
+                  significantly lower estimated susceptibility to SARS-CoV-2 infection,
+                  while adults older than 60 years have elevated susceptibility to
+                  infection.''} The current version cites both papers on Page~\ref{r.2.b}.
 
               We base our assumption on infectiousness being independent of age on
-              \citet{Jones2021}. That paper reports only slight differences in viral
+              \citet{Jones2021}. That study was the only one we could find that was able
+              differentiated between age differences in susceptibility (given a risk
+              contact) vs. age differences in contact patterns. Moreover, they had fine
+              grained age bins. That paper reports only slight differences in viral
               loads between German adults and children: ``The least infectious youngest
@@
               devices---rather than actual lower viral loads and conclude that the
-              differences, if existent, are likely not clinically relevant.
-
-                  {\bf \color{red} ToDo: Check paper suggestion
-                      \url{https://doi.org/10.1093/infdis/jiaa691} -- are they actually checking
-                      infectiousness conditional on contact type? Or could we even reproduce
-                      what they are doing (not suggesting we should, certainly no need to spend
-                      time on it)}
-
-              In any case, while the jury is certainly still out on that particular
-              issue, we are confident that different assumptions would not be quantitatively
-              meaningful to change our results in a significant way.
+              differences, if existent, are likely not clinically relevant. In fact,
+              \citet{Goldstein2021}  reach similar conclusions, writing: \textit{``There
+                  is limited evidence in the literature regarding age-related differences in
+                  infectivity, although point estimates in several studies suggest that
+                  infectivity may increase somewhat with age.''} We are thus confident that
+              different assumptions in the range of plausible values would not be
+              to change our results in a quantitatively meaningful way.
 

### 2022-03-10 src/rev1_reply_comments.tex
Clean up references, fix typo.

@@
               text. In fact, the paper you suggested \citep{Goldstein2021} comes to
-              similar conclusions: \texit{``We found evidence that compared to
+              similar conclusions: \textit{``We found evidence that compared to
                   younger/middle-aged adults, children younger than 10 years have

### 2022-03-11 src/paper/results/robustness_to_sensitivity_assumptions.tex
Add missing word.
Co-authored-by: Janos Gabler <janos.gabler@gmail.com>
Implmeent Klara's suggestions from zulip.

@@
     sensitivity profiles: The average over all methods and the lower envelope over all
-    methods.
+    methods using recent studies.
 
@@
     For the mapping of $Ct$ values to rapid test sensitivities, again we have two
-    options. First, we can simply look up the discretized values for the three $Ct$
+    options. First, we can simply look up the discretized values for the three $Ct$ bins
     provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35). Secondly, we
@@
     the end of an infection. However, the main results barely change. This is due to the
-    fact that the differences are most important towards the later stage of an
+    fact that the differences are largest towards the later stage of an
     infection. Uncovering an infection that was previously undetected at that stage does

### 2022-03-11 src/rev1_reply_comments.tex
Implmeent Klara's suggestions from zulip.

@@
 
+\large{\textbf{\textcolor{red}{SUMMARY}}}
+
 \section*{Editors}
@@
 
+\FloatBarrier
+
 Furthermore, I recommend including a declaration of potential conflicts of interest in
@@
               reduce contacts. These are active throughout the period of analysis and
-              they vary with the strictness of measure. In order to do so, it is crucial
+              they vary with the strictness of measures. In order to do so, it is crucial
               to take the type of contact into account; see for example our analysis of
@@
               \citet{Jones2021}. That study was the only one we could find that was able
-              differentiated between age differences in susceptibility (given a risk
+              to differentiate between age differences in susceptibility (given a risk
               contact) vs. age differences in contact patterns. Moreover, they had fine
@@
                   infectivity may increase somewhat with age.''} We are thus confident that
-              different assumptions in the range of plausible values would not be
-              to change our results in a quantitatively meaningful way.
+              different assumptions in the range of plausible values would not
+              change our results in a quantitatively meaningful way.
 
@@
               immunity through a vaccination while the remaining 25\% remain fully
-              susceptible. SFirstly, the existing evidence supports the
+              susceptible. Firstly, the existing evidence supports the
               assumption that all four vaccines approved in the EU (Biontech/Pfizer,

### 2022-03-11 src/rev1_reply_comments.tex
Update references.
Add summary.

@@
 
-\large{\textbf{\textcolor{red}{SUMMARY}}}
+\section*{Summary}
+
+We thank the editors and referees for their excellent comments. Revising the
+paper to address them has allowed us to substantially improve our paper.
+
+% strongest change came through sensitivity comment
+The largest adjustment was warrented with respect to the sensitivity of rapid
+tests. We fully agree that our calibration appears optimistic in light of the
+studies that have come out since our submission. We therefore verified that 
+our results and conclusions are not affected when repeating the simulations 
+with rapid tests with a much lower sensitivity. For that, we added a robustness
+check with sensitivities that are derived from more recent
+sources, relying on the progression of viral load over the course of an
+infection and rapid test sensitivity with respect to different viral loads. 
+These new sensitivities are on average significantly lower than those in our 
+main specification. We are happy to report that even the most conservative 
+estimates only lead to a small decrease in the estimated role of rapid tests.
+
+The remaining comments have allowed us to improve our text significantly to be
+clearer and better referenced. We also updated our citations for papers that we
+had cited as pre-prints and that have been published in the meantime.
+In our view the text has become clearer, more legible and better referenced 
+as a result.
 
@@
               We base our assumption on infectiousness being independent of age on
-              \citet{Jones2021}. That study was the only one we could find that was able
-              to differentiate between age differences in susceptibility (given a risk
-              contact) vs. age differences in contact patterns. Moreover, they had fine
-              grained age bins. That paper reports only slight differences in viral
+              \citet{Jones2021}. That paper reports only slight differences in viral
               loads between German adults and children: ``The least infectious youngest
@@
               devices---rather than actual lower viral loads and conclude that the
-              differences, if existent, are likely not clinically relevant. In fact,
+              differences, if existent, are likely not clinically relevant.
+
+              In fact,
               \citet{Goldstein2021}  reach similar conclusions, writing: \textit{``There

### 2022-03-11 src/paper/report.tex
Incorporate Janos' comments.
Final version for resubmission.

@@
 In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age
-\citep{Davies2020}.\label{r.2.b} A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period
-of infection without being infectious or showing symptoms, being infectious
+\citep{Davies2020,Goldstein2021}.\label{r.2.b} A possible infection progresses as shown
+in Figure~\ref{fig:model_disease_progression}. We differentiate between an initial
+period of infection without being infectious or showing symptoms, being infectious
 (presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and
@@
     on the timing of the test relative to the onset of infectiousness as in
-\citet{Smith2021}. We provide a robustness to different assumptions in
+    \citet{Smith2021}. We analyse robustness to different assumptions in
     Appendix~\ref{subsec:robustness_rapid_test_sensitivity}.} After a phase-in period, all
@@
     of whether our results are robust regarding the sensitivity parameters we assume for
-    rapid tests. Even if we take a very pessimistic stance, the effect is only reduced from
-    42\% to 38\%.}
+    rapid tests. Even if we take a pessimistic stance, the effect is only reduced from 42\%
+    to 38\%.}
 

### 2022-03-11 src/rev1_reply_comments.tex
Final version for resubmission.

@@
 \usepackage{colortbl}
+\usepackage{placeins}
 
@@
 We thank the editors and referees for their excellent comments. Revising the
-paper to address them has allowed us to substantially improve our paper.
+paper to address them has allowed us to substantially improve our paper. \\[1ex]
 
-% strongest change came through sensitivity comment
-The largest adjustment was warrented with respect to the sensitivity of rapid
-tests. We fully agree that our calibration appears optimistic in light of the
-studies that have come out since our submission. We therefore verified that 
-our results and conclusions are not affected when repeating the simulations 
-with rapid tests with a much lower sensitivity. For that, we added a robustness
-check with sensitivities that are derived from more recent
-sources, relying on the progression of viral load over the course of an
-infection and rapid test sensitivity with respect to different viral loads. 
-These new sensitivities are on average significantly lower than those in our 
-main specification. We are happy to report that even the most conservative 
-estimates only lead to a small decrease in the estimated role of rapid tests.
+The largest adjustment was warranted with respect to the sensitivity of rapid tests. We
+fully agree that our calibration appears optimistic in light of the studies that have
+come out since our submission. We therefore verified that our results and conclusions
+are not affected when repeating the simulations under a much lower sensitivity of rapid
+tests. For that, we added a robustness check with sensitivities that are derived from
+more recent sources, relying on the progression of viral load over the course of an
+infection and rapid test sensitivity with respect to different viral loads. These new
+sensitivities are on average significantly lower than those in our main specification.
+We are happy to report that even the most conservative estimates lead to only a minor
+decrease in the estimated role of rapid tests. \\[1ex]
 
-The remaining comments have allowed us to improve our text significantly to be
-clearer and better referenced. We also updated our citations for papers that we
-had cited as pre-prints and that have been published in the meantime.
-In our view the text has become clearer, more legible and better referenced 
-as a result.
+The remaining comments have allowed us to improve our text significantly to be clearer
+and better referenced. We also updated our citations for papers that we cited as
+pre-prints and that have been published in the meantime. In our view the text has become
+clearer, more legible and better referenced as a result.\\[1ex]
+
+We now address the comments by the Editors and Reviewers in turn.
 
@@
 }
+
+\clearpage
 
@@
     submitted the original manuscript in early September 2021; the analysis had been
-    finalised in June of last year. Indeed, in the line of the additional evidence, the
+    finalised in June of last year. Indeed, in the light of the additional evidence, the
     numbers we use for test sensitivity appear to be in the upper part of plausible
     values. We have followed your suggestion and report detailed robustness analyses in
-    the paper and its Appendix.
+    the paper and its appendix.
 
@@
 
-    \textit{\Paste{r.1.a}}
-
-    \vskip1ex
+    \textit{\Paste{r.1.a}}\\[1ex]
 
@@
 
-    \textit{\Paste{r.1.b}}
-
-    \vskip1ex
+    \textit{\Paste{r.1.b}}\\[1ex]
 
@@
 
-\response{We have added such a declaration at the end of the paper:\\[2ex]
+\response{We have added such a declaration at the end of the paper:
 
@@
           \response{%
-              Many thanks, it is of course true that vaccines do not stop transmission
-              completely. We do not assume that, either. However, we have not been able
-              to find data on the way in which this affects different population groups
-              for the wild type and the $\alpha$-variant.
+              Many thanks, it is of course true that vaccines reduce transmissions, but
+              do not stop them completely. We do not assume that, either. However, we
+              have not been able to find data on the way in which this affects different
+              population groups for the wild type and the $\alpha$-variant.
 
@@
               immunity through a vaccination while the remaining 25\% remain fully
-              susceptible. Firstly, the existing evidence supports the
-              assumption that all four vaccines approved in the EU (Biontech/Pfizer,
-              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
+              susceptible. Firstly, the existing evidence supports the assumption that
+              all four vaccines approved in the EU (Biontech/Pfizer, Moderna,
+              Oxford/AstraZeneca and Johnson \& Johnson) provide similar levels of
               protection in the first few months after vaccination against the wild type
@@
               are negligibly small. Thus, the value added by a probabilistic model of
-              vaccination protection is very small for the period that we cover. }
+              vaccination protection is very small for the period that we cover. At the
+              same time, it comes at prohibitive computational costs.}
 \end{enumerate}

### 2022-03-11 src/cover-letter.tex
Remove commenting, add cover letter.

--- 2021-08-23 src/cover-letter.tex
@@
 \setkomavar{subject}{%
-    The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs
+    The Effectiveness of Testing, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic
 }
@@
 
-    Many thanks for considering our manuscript for possible publication in
-    \textit{scientific reports}. Our hope is that modelling the evolution of the
-    CoViD-19 pandemic and the relative effectiveness of diverse interventions are of
-    interest to a broad readership. We make both methodological and substantive
-    contributions.
+    Many thanks for inviting a revision of manuscript for possible publication in
+    \textit{Scientific Reports}. We are confident that we were able  to address all
+    points that have been raised, please see our detailed response.
 
-    We incorporate detailed NPIs, different tests, and vaccinations in an agent-based
-    model, taking into account behavioural reactions to symptoms or positive test
-    results. Importantly, we do so in a way that allows to calibrate its parameters
-    based on routinely available data. We apply the model to the second and third waves
-    of the CoViD-19 pandemic in Germany over a period of 9 months, achieving a good fit
-    to actual infection rates.
-
-    Our core substantive result is that rapid testing has been very effective relative
-    to many NPIs in a low-medium vaccination environment. This is the relevant scenario
-    for most countries across the world at present; it seems also applicable to countries
-    with high vaccination rates under the conditions of the delta variant.
+    We are looking forward to hearing from you.
 

### 2022-03-11 task_compile_documents.py

@@ -37,7 +37,7 @@ def task_extract_paper_body(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "1-11",
+            "1-15",
             "--",
             produces.as_posix(),
         ]
@@ -54,7 +54,7 @@ def task_extract_supplementary_material(depends_on, produces):
             "--empty",
             "--pages",
             depends_on.as_posix(),
-            "12-78",
+            "16-83",
             "--",
             produces.as_posix(),
         ]
@@ -97,6 +97,28 @@ def task_compile_reply():
 
 @pytask.mark.depends_on(BLD / "rev1_reply_comments.pdf")
 @pytask.mark.produces(ROOT / "rev1_reply_comments.pdf")
-def task_copy_pdf_to_root(depends_on, produces):
+def task_copy_reply_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)
+
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.depends_on(SRC / "cover-letter.tex")
+@pytask.mark.produces(BLD / "cover-letter.pdf")
+def task_compile_reply():
+    pass
+
+
+@pytask.mark.depends_on(BLD / "cover-letter.pdf")
+@pytask.mark.produces(ROOT / "cover-letter.pdf")
+def task_copy_cover_letter_to_root(depends_on, produces):
     shutil.copy(depends_on, produces)
 

### 2022-03-20 src/panel_figures/2021_scenarios.tex
Add article for submission and multi-figures to single figures.

% \documentclass[varwidth,10pt, border=10pt]{wlscirep}
\documentclass[varwidth, border=10pt, class=wlscirep, 10pt]{standalone}

\RequirePackage{helvet}
\RequirePackage{courier}

% border={10pt 5pt} % left/right bottom/top
% border={10pt 10pt 0pt 10pt} % left right bottom top

\usepackage{subcaption}
\usepackage[
    font=footnotesize,
    labelfont={bf,sf,footnotesize},
    labelsep=period,
    justification=raggedright
]{caption}
\usepackage{graphicx}

\begin{document}

\begin{figure}
    \centering

    \begin{subfigure}[b]{6cm}
        \centering
        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
        \caption{{Recorded cases: 2021 scenarios}}
        \label{fig:2021_scenarios_recorded}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{6cm}
        \centering
        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
        \caption{{Total cases: 2021 scenarios}}
        \label{fig:2021_scenarios_newly_infected}
    \end{subfigure}

    \begin{subfigure}[b]{6cm}
        \centering
        \includegraphics[width=5cm]{../figures/results/figures/full_decomposition_channels_area}
        \caption{Decomposition of the difference between the scenario without any of the
            three factors and the main scenario in
            Figure~3b.}
        \label{fig:2021_scenarios_decomposition}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{6cm}
        \centering
        \includegraphics[width=5cm]{../figures/results/figures/full_decomposition_rapid_tests_area}
        \caption{Decomposition of the difference between the scenario without rapid
            tests and the main scenario in
            Figure~3b.}
        \label{fig:2021_scenarios_decomposition_tests}
    \end{subfigure}
\end{figure}

\end{document}

### 2022-03-20 src/panel_figures/evolution_pandemic.tex
Add article for submission and multi-figures to single figures.

--- 2022-03-20 src/panel_figures/2021_scenarios.tex
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
-        \caption{{Recorded cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_recorded}
+        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
+        \caption{{Recorded cases: Empirical and simulated}}
+        \label{fig:aggregated_fit}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
-        \caption{{Total cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_newly_infected}
+        \includegraphics[width=5cm]{../figures/results/figures/data/stringency2_with_seasonality}
+
+        \caption{{Stringency of NPIs and infectious contacts}}
+        \label{fig:stringency_infectious_contacts}
     \end{subfigure}
-
+    \vskip3ex
     \begin{subfigure}[b]{6cm}
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/full_decomposition_channels_area}
-        \caption{Decomposition of the difference between the scenario without any of the
-            three factors and the main scenario in
-            Figure~3b.}
-        \label{fig:2021_scenarios_decomposition}
+
+        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
+
+        \caption{Fraction of B.1.1.7 strain}
+        \label{fig:share_b117}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/full_decomposition_rapid_tests_area}
-        \caption{Decomposition of the difference between the scenario without rapid
-            tests and the main scenario in
-            Figure~3b.}
-        \label{fig:2021_scenarios_decomposition_tests}
+
+        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
+
+        \caption{{Tests and vaccinations}}
+        \label{fig:antigen_tests_vaccinations}
     \end{subfigure}

### 2022-03-20 src/panel_figures/model_description.tex
Add article for submission and multi-figures to single figures.

--- 2022-03-20 src/panel_figures/evolution_pandemic.tex
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{Recorded cases: Empirical and simulated}}
-        \label{fig:aggregated_fit}
+        \includegraphics[width=5cm]{../figures/model-graph-top-left}
+        \caption{Model for contacts and infections}
+        \label{fig:model_contacts_infections}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/data/stringency2_with_seasonality}
+        \includegraphics[width=5cm]{../figures/model-graph-top-right}
 
-        \caption{{Stringency of NPIs and infectious contacts}}
-        \label{fig:stringency_infectious_contacts}
+        \caption{Disease progression}
+        \label{fig:model_disease_progression}
     \end{subfigure}
@@
 
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
-
-        \caption{Fraction of B.1.1.7 strain}
-        \label{fig:share_b117}
+        \includegraphics[width=5cm]{../figures/model-graph-bottom-left}
+        \caption{Model for rapid tests}
+        \label{fig:model_rapid_tests}
     \end{subfigure}
@@
         \centering
+        \includegraphics[width=5cm]{../figures/model-graph-bottom-right}
 
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
+        \caption{Translating all infections to recorded ones}
+        \label{fig:model_official_cases}
+    \end{subfigure}
 
-        \caption{{Tests and vaccinations}}
-        \label{fig:antigen_tests_vaccinations}
-    \end{subfigure}
 \end{figure}

### 2022-03-20 src/panel_figures/school_workplace_scenarios.tex
Add article for submission and multi-figures to single figures.

--- 2022-03-20 src/panel_figures/evolution_pandemic.tex
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{Recorded cases: Empirical and simulated}}
-        \label{fig:aggregated_fit}
+        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
+        \caption{{Effects of different schooling scenarios}}
+        \label{fig:school_scenarios}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=5cm]{../figures/results/figures/data/stringency2_with_seasonality}
-
-        \caption{{Stringency of NPIs and infectious contacts}}
-        \label{fig:stringency_infectious_contacts}
-    \end{subfigure}
-    \vskip3ex
-    \begin{subfigure}[b]{6cm}
-        \centering
-
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
-
-        \caption{Fraction of B.1.1.7 strain}
-        \label{fig:share_b117}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{6cm}
-        \centering
-
-        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
-
-        \caption{{Tests and vaccinations}}
-        \label{fig:antigen_tests_vaccinations}
+        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
+        \caption{{Effects of different work scenarios}}
+        \label{fig:workplace_scenarios}
     \end{subfigure}

### 2022-03-20 src/scientific_reports_submission/article.tex
Add article for submission and multi-figures to single figures.

--- 2022-03-10 src/paper/report.tex
@@
+\documentclass[fleqn,10pt]{wlscirep}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\title{
+    The Effectiveness of Testing, Vaccinations and Contact Restrictions for Containing
+    the CoViD-19 Pandemic
+}
+
+\author[1, 2, +]{Janoś Gabler}
+\author[3, +]{Tobias Raabe}
+\author[1, +]{Klara Röhrl}
+\author[2, 4, *, +]{Hans-Martin von Gaudecker}
+
+\affil[1]{Bonn Graduate School of Economics, Bonn, 53113, Germany}
+\affil[2]{IZA Institute of Labor Economics, Bonn, 53113, Germany}
+\affil[3]{Private sector}
+\affil[4]{Rheinische Friedrich-Wilhelms-Universität Bonn, Bonn, 53113, Germany}
+
+\affil[*]{hmgaudecker@uni-bonn.de}
+
+\affil[+]{These authors contributed equally to this work}
+
+\keywords{
+    CoViD-19, agent based simulation model, rapid testing, non-pharmaceutical
+    interventions
+}
+
+\begin{abstract}
+    In order to slow the spread of the CoViD-19 pandemic, governments around the world
+    have enacted a wide set of policies limiting the transmission of the disease.
+    Initially, these focused on non-pharmaceutical interventions; more recently,
+    vaccinations and large-scale rapid testing have started to play a major role. The
+    objective of this study is to explain the quantitative effects of these policies on
+    determining the course of the pandemic, allowing for factors like seasonality or
+    virus strains with different transmission profiles. To do so, the study develops an
+    agent-based simulation model, which explicitly takes into account test demand and
+    behavioral changes following positive tests. The model is estimated using data for
+    the second and the third wave of the CoViD-19 pandemic in Germany. The paper finds
+    that during a period where vaccination rates rose from 5\% to 40\%, seasonality and
+    rapid testing had the largest effect on reducing infection numbers. Frequent
+    large-scale rapid testing should remain part of strategies to contain CoViD-19; it
+    can substitute for many non-pharmaceutical interventions that come at a much larger
+    cost to individuals, society, and the economy.
+\end{abstract}
+
+\begin{document}
+
+\flushbottom
+\maketitle
+\thispagestyle{empty}
+
+\section*{Introduction}
+
 Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
@@
 months---and often years---away from reaching vaccination levels that would end the
-pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
+pandemic or even protect the most vulnerable \cite{Mathieu2021}. In the meantime, it is
 of utmost importance to employ an effective mix of strategies for containing the virus.
@@
 to reduce contacts between individuals. While this has allowed some countries to sustain
-equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
-    theoretical equilibrium at low case numbers which is sustained with
-    test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
-over time. Containment measures have become increasingly diverse and now include rapid
-testing, more nuanced NPIs, and contact tracing. Neither these policies' effects nor the
-influence of seasonal patterns or of more infectious virus strains are well understood
-in quantitative terms.
+equilibria with very low infection numbers (see \cite{Contreras2021} for a theoretical
+equilibrium at low case numbers which is sustained with test-trace-and-isolate
+policies), most have seen large fluctuations of infection rates over time. Containment
+measures have become increasingly diverse and now include rapid testing, more nuanced
+NPIs, and contact tracing. Neither these policies' effects nor the influence of seasonal
+patterns or of more infectious virus strains are well understood in quantitative terms.
 
@@
 that, aside from seasonality, frequent and large-scale rapid testing caused the bulk of
-this decrease, which is in line with prior predictions \citep{Mina2021}. We conclude
+this decrease, which is in line with prior predictions \cite{Mina2021}. We conclude
 that it should have a large role for at least as long as vaccinations have not been
@@
 
-At the core of our agent-based model \citep[][we review more literature in Supplementary
-    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021a} are physical contacts
-between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each
-contact between an infectious individual and somebody susceptible to the disease
-bears the risk of transmitting the virus. Contacts occur in up to four networks:
-Within the household, at work, at school, or in other settings (leisure activities,
-grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
-occur at random. Empirical applications can take the population and household
-structure from census data and the network-specific frequencies of contacts from
-diary data measuring contacts before the pandemic
-\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
-depend on age and geographical location (see Supplementary
-Material~\ref{subsec:assortativity}).
+At the core of our agent-based model---we review more literature in Supplementary
+Material~B.1 \cite{Aleta2020,Hinch2021a}---are physical contacts between heterogeneous
+agents (Figure~1a). Each contact between an infectious individual and somebody
+susceptible to the disease bears the risk of transmitting the virus. Contacts occur in
+up to four networks: Within the household, at work, at school, or in other settings
+(leisure activities, grocery shopping, medical appointments, etc.). Some contacts recur
+regularly, others occur at random. Empirical applications can take the population and
+household structure from census data and the network-specific frequencies of contacts
+from diary data measuring contacts before the pandemic \cite{Mossong2008,Hoang2019}.
+Within each network, meeting frequencies depend on age and geographical location (see
+Supplementary Material~A.4).
 
@@
 implement mitigation measures like masking requirements or air filters
-\citep{Lessler2021}. Curfews may reduce the number of contacts in settings outside of
+\cite{Lessler2021}. Curfews may reduce the number of contacts in settings outside of
 work and school. In any setting, measures like masking requirements would reduce the
-probability of infection associated with a contact \citep{Cheng2021}.
+probability of infection associated with a contact \cite{Cheng2021}.
 
-\begin{figure}   % Figure 1
+\begin{figure}[tb]
     \centering
-
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-top-left}
-        \caption{{Model for contacts and infections}}
-        \label{fig:model_contacts_infections}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-top-right}
-        \vskip5ex
-
-        \caption{Disease progression}
-        \label{fig:model_disease_progression}
-    \end{subfigure}
-    \vskip3ex
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{figures/model-graph-bottom-left}
-        \caption{{Model for rapid tests}}
-        \label{fig:model_rapid_tests}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/model-graph-bottom-right}
-        \vskip5ex
-
-        \caption{{Translating all infections to recorded ones}}
-        \label{fig:model_official_cases}
-    \end{subfigure}
-
-    \caption{Model description}
-    \label{fig:model_description}
-
-    \floatfoot{\noindent \textit{Note:}
-        A~description of the model can be found in Supplementary
-        Material~\ref{sec:supplementary_text}.
-        Figure~\ref{fig:model_contacts_infections} shows the influence of an agent's
-        contacts to other agents on infections. Demographic characteristics set the
-        baseline number of contacts in different networks ($\eta$). The agents may
-        reduce the number of contacts due to NPIs, showing symptoms, or testing
-        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
-        agent meets an infectious agent; the probability depends on the type of contact
-        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
-        infected, the infection progresses as depicted in
-        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
-        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
-        trigger a test only for a fraction of individuals depending on an individual
-        compliance parameter; the thresholds for triggering test demand differ across
-        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
-        Figure~\ref{fig:model_official_cases} shows the model of translating all
+    \includegraphics[width=\textwidth]{../../bld/panel_figures/fig-model-description.pdf}
+    \caption{
+        A~description of the model can be found in Supplementary Material~B. Figure~1a
+        shows the influence of an agent's contacts to other agents on infections.
+        Demographic characteristics set the baseline number of contacts in different
+        networks ($\eta$). The agents may reduce the number of contacts due to NPIs,
+        showing symptoms, or testing positively for SARS-CoV-2 ($\tau$). Infections may
+        occur when a susceptible agent meets an infectious agent; the probability
+        depends on the type of contact ($\beta_c$), on seasonality ($\kappa_c$), and on
+        NPIs ($\rho_{c,\:t}$). If infected, the infection progresses as depicted in
+        Figure~1b. If rapid tests are available, agents' demand is modeled as in
+        Figure~1c. All reasons trigger a test only for a fraction of individuals
+        depending on an individual compliance parameter; the thresholds for triggering
+        test demand differ across reasons and they may depend on calendar time
+        ($\pi_{c,\:t}$ and $\tau_{c,\:t}$). Figure~1d shows the model of translating all
         infections in the simulated data to age-specific recorded infections. The model
@@
         of the graph is relevant only for periods where rapid tests are available. All
-        parameters are explained in Appendix~\ref{sub:param_tables}.}
+        parameters are explained in Supplementary Material~A.11.
+    }
+    \label{fig:model_description}
 \end{figure}
@@
 In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age
-\citep{Davies2020}.\label{r.2.b} A possible infection progresses as shown in
-Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period
-of infection without being infectious or showing symptoms, being infectious
-(presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and
-recovery or death \citep[similar to][]{Aleta2020}. The probabilities of transitioning
-between these states depend on age; their duration is random and calibrated to medical
-literature (for a detailed description see Supplementary
-Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
-infectiousness is independent of age \citep{Jones2021}.
+\cite{Davies2020,Goldstein2021}. A possible infection progresses as shown in Figure~1b.
+We differentiate between an initial period of infection without being infectious or
+showing symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms,
+requiring intensive care, and recovery or death similar to \cite{Aleta2020}. The
+probabilities of transitioning between these states depend on age; their duration is
+random and calibrated to medical literature (for a detailed description see
+Supplementary Material~A.1). Conditional on the type of contact, infectiousness is
+independent of age \cite{Jones2021}.
 
@@
 vaccination offers. With some probability, vaccinated agents become immune and do not
-transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.
+transmit the virus \cite{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.
 
@@
 these tests is set according to data analyzed in
-\label{r.1.a}\Copy{r.1.a}{\cite{Bruemmer2021,Scheiblauer2021,Özcürümez2021};
-    sensitivity depends on the timing of the test relative to the onset of infectiousness.}
-After a phase-in period, all tests that are demanded will be performed.
-Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
-require staff and students to be tested regularly. Rapid tests may be offered by
-employers to on-site workers. Individuals may demand tests for private reasons, which
-include having plans to meet other people, showing symptoms of CoViD-19, and a household
-member having tested positively for the virus. We endow each agent with an individual
-compliance parameter. This parameter determines whether she takes up rapid
-tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
-    contacts; this is why tests impact the actual contacts in
-    Figure~\ref{fig:model_description}.}
+\cite{Bruemmer2021,Scheiblauer2021,Ozcurumez2021}; sensitivity depends on the timing of
+the test relative to the onset of infectiousness as in \cite{Smith2021}. We analyse
+robustness to different assumptions in Supplementary Material~A.12. After a phase-in
+period, all tests that are demanded will be performed. Figure~1c shows our model for
+rapid test demand. Schools may require staff and students to be tested regularly. Rapid
+tests may be offered by employers to on-site workers. Individuals may demand tests for
+private reasons, which include having plans to meet other people, showing symptoms of
+CoViD-19, and a household member having tested positively for the virus. We endow each
+agent with an individual compliance parameter. This parameter determines whether she
+takes up rapid tests. (Positive test results or symptoms lead most individuals to reduce
+their contacts; this is why tests impact the actual contacts in Figure~1.)
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
-    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
-different contact types and their assortativity by age group. Mobility data is used to
-model the evolution of work contacts. School and daycare policies can be incorporated
-directly from official directives. Administrative records on the number of tests,
-vaccinations by age and region, and the prevalence of virus strains are generally
-available. Surveys may ask about test offers, propensities to take them up, and past
-tests. Other studies' estimates of the seasonality of infections can be incorporated
-directly. The remaining parameters---most notably, these include infection probabilities
-by contact network and the effects of some NPIs, see Supplementary
-Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
-matches features of the data \citep[see][for the general method]{McFadden1989}. In our
-application, we keep the number of free parameters low in order to avoid overfitting.
-The data features to be matched include official case numbers for each age group and
-region, deaths, and the share of the B.1.1.7 strain.
+parameters (see Supplementary Material~A for a complete description). Contact diaries
+yield pre-pandemic distributions of contacts for different contact types and their
+assortativity by age group. Mobility data is used to model the evolution of work
+contacts. School and daycare policies can be incorporated directly from official
+directives. Administrative records on the number of tests, vaccinations by age and
+region, and the prevalence of virus strains are generally available. Surveys may ask
+about test offers, propensities to take them up, and past tests. Other studies'
+estimates of the seasonality of infections can be incorporated directly. The remaining
+parameters---most notably, these include infection probabilities by contact network and
+the effects of some NPIs, see Supplementary Material~A.9---will be chosen numerically so
+that the model matches features of the data (see \cite{McFadden1989} for the general
+method). In our application, we keep the number of free parameters low in order to avoid
+overfitting. The data features to be matched include official case numbers for each age
+group and region, deaths, and the share of the B.1.1.7 strain.
 
@@
 all infections. In the German case, this specifically amounts to positive PCR tests. We
-thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
-mortality-based aggregate estimates of the share of detected cases and use data on the
-share of PCR tests administered because of CoViD-19 symptoms. As the share of
-asymptomatic individuals varies by age group, this gives us age-specific shares (see
-Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
-absence of rapid testing---the detection rate is 80\% higher on average for individuals
-above age 80 compared to school age children. Once rapid test become available,
-confirmation of a positive result is another reason leading to positive PCR tests.
+thus model recorded cases as depicted in Figure~1d. We take mortality-based aggregate
+estimates of the share of detected cases and use data on the share of PCR tests
+administered because of CoViD-19 symptoms. As the share of asymptomatic individuals
+varies by age group, this gives us age-specific shares (see Figure~B.11). Our estimates
+suggest that---in the absence of rapid testing---the detection rate is 80\% higher on
+average for individuals above age 80 compared to school age children. Once rapid test
+become available, confirmation of a positive result is another reason leading to
+positive PCR tests.
 
-\section{Second and third waves of the CoViD-19 pandemic in Germany}
+\section*{Second and third waves of the CoViD-19 pandemic in Germany}
 
 The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
-covering the period mid-September 2020 to the end of May 2021.
-Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
-of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
-recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
-Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
-non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
-never use it directly. For legibility reasons, we transform the index so that lower
-values represent higher levels of restrictions. A value of zero means all measures
-incorporated in the index are turned on. The value one represents the situation in
-mid-September, with restrictions on gatherings and public events, masking requirements,
-but open schools and workplaces. In the seven weeks between mid September and early
-November, cases increased by a factor of ten. Restrictions were somewhat tightened in
-mid-October and again in early November. New infections remained constant throughout
-November before rising again in December, prompting the most stringent lockdown to this
-date. Schools and daycare centers were closed, so were customer-facing businesses except
-for grocery and drug stores. From the peak of the second wave just before Christmas
-until the trough in mid-February, newly detected cases decreased by almost three
-quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
-strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
-    (Delta) was first detected in Germany in April; at the end of our simulation period it
-    accounted for less than 5\% of cases.} In early March, some NPIs were relaxed; e.g.,
-hairdressers and home improvement stores were allowed to open again to the public. There
-were many changes in details of regulations afterwards, but they did not change the
-overall stringency index.
+covering the period mid-September 2020 to the end of May 2021. Figure~2a describes the
+evolution of the pandemic and of its drivers. The black line in Figure~2a shows
+officially recorded cases; the black line in Figure~2b the Oxford Response Stringency
+Index \cite{Hale2020}, which tracks the tightness of non-pharmaceutical interventions.
+The index is shown for illustration of the NPIs, we never use it directly. For
+legibility reasons, we transform the index so that lower values represent higher levels
+of restrictions. A value of zero means all measures incorporated in the index are turned
+on. The value one represents the situation in mid-September, with restrictions on
+gatherings and public events, masking requirements, but open schools and workplaces. In
+the seven weeks between mid September and early November, cases increased by a factor of
+ten. Restrictions were somewhat tightened in mid-October and again in early November.
+New infections remained constant throughout November before rising again in December,
+prompting the most stringent lockdown to this date. Schools and daycare centers were
+closed, so were customer-facing businesses except for grocery and drug stores. From the
+peak of the second wave just before Christmas until the trough in mid-February, newly
+detected cases decreased by almost three quarters. The third wave in the spring of 2021
+is associated with the B.1.1.7 (Alpha) strain, which became dominant in March
+(Figure~2c)---B.1.617.2 (Delta) was first detected in Germany in April; at the end of
+our simulation period it accounted for less than 5\% of cases. In early March, some NPIs
+were relaxed; e.g., hairdressers and home improvement stores were allowed to open again
+to the public. There were many changes in details of regulations afterwards, but they
+did not change the overall stringency index.
 
-\begin{figure}[!tp]   % Figure 2
+\begin{figure}[tb]
     \centering
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
-        \caption{{Recorded cases: Empirical and simulated}}
-        \label{fig:aggregated_fit}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/data/stringency2_with_seasonality}
-
-        \caption{{Stringency of NPIs and infectious contacts}}
-        \label{fig:stringency_infectious_contacts}
-    \end{subfigure}
-
-    \vskip3ex
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}
-
-        \caption{Fraction of B.1.1.7 strain}
-        \label{fig:share_b117}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}
-
-        \caption{{Tests and vaccinations}}
-        \label{fig:antigen_tests_vaccinations}
-    \end{subfigure}
-
-    \caption{Evolution of the pandemic, its drivers, and model fit, September 2020 to May 2021}
+    \includegraphics[width=\textwidth]{../../bld/panel_figures/fig-evolution-pandemic.pdf}
+    \caption{
+        Evolution of the pandemic, its drivers, and model fit, September 2020 to May
+        2021: Data sources are described in Supplementary Material~A. Age- and
+        region-specific analogues to Figure~2a can be found in Supplementary
+        Material~B.10. For legibility reasons, all lines in Figure~2b are rolling 7-day
+        averages. The Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x /
+        100)$, so that a value of one refers to the situation at the start of our sample
+        period and zero means that all NPIs included in the index are turned on. The
+        other lines in Figure~2b show the product of the effect of contact reductions,
+        increased hygiene regulations, and seasonality. See Appendix~A.5 for separate
+        plots of the three factors by contact type.
+    }
     \label{fig:pandemic_drivers_model_fit}
-
-    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
-        Material~\ref{sec:materials_and_methods}. Age- and region-specific analogues to
-        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
-        \ref{subsec:fit_results}. For legibility reasons, all lines in
-        Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
-        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
-        a value of one refers to the situation at the start of our sample period and
-        zero means that all NPIs included in the index are turned on. The other lines in
-        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
-        of contact reductions, increased hygiene regulations, and seasonality. See
-        Appendix~\ref{subsec:policies} for separate plots of the three factors by
-        contact type.}
 \end{figure}
@@
 turn of the year, the first people were vaccinated with a focus on older age groups and
-medical staff (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May, 43\%
-had received at least one dose of a vaccine. In late 2020, rapid tests started to
-replace regular PCR tests for staff in many medical and nursing facilities. These had to
-be administered by medical doctors or in pharmacies. At-home tests approved by
-authorities became available in mid-March. Rapid test centers were opened, and one test
-per person and week was made available free of charge. In several states, customers were
-only allowed to enter certain stores with a recent negative rapid test result. These
-developments are characteristic of many countries: The initial focus on NPIs to slow the
-spread of the disease has been accompanied by vaccines and a growing acceptance and use
-of rapid tests. At broadly similar points in time, novel strains of the virus have
-started to pose additional challenges.
+medical staff (Figure~2d). Until the end of May, 43\% had received at least one dose of
+a vaccine. In late 2020, rapid tests started to replace regular PCR tests for staff in
+many medical and nursing facilities. These had to be administered by medical doctors or
+in pharmacies. At-home tests approved by authorities became available in mid-March.
+Rapid test centers were opened, and one test per person and week was made available free
+of charge. In several states, customers were only allowed to enter certain stores with a
+recent negative rapid test result. These developments are characteristic of many
+countries: The initial focus on NPIs to slow the spread of the disease has been
+accompanied by vaccines and a growing acceptance and use of rapid tests. At broadly
+similar points in time, novel strains of the virus have started to pose additional
+challenges.
 
-\section{Results}
+\section*{Results}
 
@@
 use the model to predict recorded infection rates until the end of May 2021. See
-Supplementary Materials~\ref{subsec:synthetic_population} and
-\ref{sub:initial_conditions} for details. The blue line in
-Figure~\ref{fig:aggregated_fit} shows that our model's predictions are very close to
+Supplementary Materials~A.2 and B.9 for details. The blue line in
+Figure~2a shows that our model's predictions are very close to
 officially recorded cases in the aggregate. This is also true for infections by age and
-geographical region (see Supplementary Material~\ref{subsec:fit_results}).
+geographical region (see Supplementary Material~B.10).
 
@@
 variation in the drivers of the pandemic. Next to the stringency index, the three lines
-in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
+in Figure~2b summarize how contact reductions,
 increased hygiene regulations, and seasonality evolved since early September for each of
@@
 Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always held
-constant at their values in the baseline scenario.
-Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
-Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red
-line), and three scenarios turning each of these factors on individually.
-Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
-model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
-\citep{Shapley2016} to decompose the difference in total infections between the scenario
+constant at their values in the baseline scenario. Figure~3a shows the model fit (the
+blue line, same as in Figure~2a), a scenario without any of the three factors (red
+line), and three scenarios turning each of these factors on individually. Figure~3b does
+the same for total infections in the model. Figure~3c employs Shapley values
+\cite{Shapley2016} to decompose the difference in total infections between the scenario
 without any of the three factors and our main specification.
 
-\begin{figure}[!tp]
+\begin{figure}[tb]
     \centering
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
-        \caption{{Recorded cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_recorded}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
-        \caption{{Total cases: 2021 scenarios}}
-        \label{fig:2021_scenarios_newly_infected}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_area}
-        \caption{Decomposition of the difference between the scenario without any of the
-            three factors and the main scenario in
-            Figure~\ref{fig:2021_scenarios_newly_infected}.}
-        \label{fig:2021_scenarios_decomposition}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.475\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_area}
-        \caption{Decomposition of the difference between the scenario without rapid
-            tests and the main scenario in
-            Figure~\ref{fig:2021_scenarios_newly_infected}.}
-        \label{fig:2021_scenarios_decomposition_tests}
-    \end{subfigure}
-
-    \caption{The effect of different interventions on recorded and actual infections}
+    \includegraphics[width=\textwidth]{../../bld/panel_figures/fig-2021-scenarios.pdf}
+    \caption{
+        The effect of different interventions on recorded and actual infections. The
+        blue line in Figure~3a is the same as in Figure~2a and refers to our baseline
+        scenario, so does the blue line in Figure~3b. The red lines refer to a situation
+        where NPIs evolve as in the baseline scenario and the B.1.1.7 variant is
+        introduced in the same way; vaccinations, rapid tests, and seasonality remain at
+        their January levels. The other scenarios turn each of these three factors on
+        individually. The decompositions in Figures~3c and 3d are based on Shapley
+        values, which are explained more thoroughly in Appendix~A.10. For legibility
+        reasons, all lines are rolling 7-day averages.
+    }
     \label{fig:2021_scenarios_broad}
-
-    \floatfoot{\noindent \textit{Note:} The blue line in
-        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
-        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
-        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
-        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
-        variant is introduced in the same way; vaccinations, rapid tests, and
-        seasonality remain at their January levels. The other scenarios turn each of
-        these three factors on individually. The decompositions in
-        Figures~\ref{fig:2021_scenarios_decomposition} and
-        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
-        are explained more thoroughly in Appendix~\ref{sub:shapley_value}. For
-        legibility reasons, all lines are rolling 7-day averages.}
 \end{figure}
@@
 mattered. However, in this period, cases would have kept on rising throughout, just at a
-much lower pace \citep[this is in line with results in][, which our seasonality measure
-    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
+much lower pace this is in line with results in \cite{Gavenciak2021}, which our
+seasonality measure is based on. Nevertheless, we estimate seasonality to be a
 quantitatively important factor determining the evolution of the pandemic, explaining
@@
 means that additional infections will be recorded which would otherwise remain
-undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
-may persist for some time. Until late April, recorded cases are higher in the scenario
-with rapid testing alone when compared to the setting where none of the three mechanisms
-are turned on. The effect on total cases, however, is visible immediately in
-Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
-population performed weekly rapid tests in March on average, new infections on April~1
-would have been reduced by 53\% relative to the scenario without vaccinations, rapid
-tests, or seasonality. \label{r.1.b}\Copy{r.1.b}{In
-    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}, we provide a detailed analysis
-    of whether our results are robust regarding the sensitivity parameters we assume for
-    rapid tests. Even if we take a very pessimistic stance, the effect is only reduced from
-    42\% to 38\%.}
+undetected. Figure~3a shows that this effect is large and may persist for some time.
+Until late April, recorded cases are higher in the scenario with rapid testing alone
+when compared to the setting where none of the three mechanisms are turned on. The
+effect on total cases, however, is visible immediately in Figure~3b. Despite the fact
+that only 10\% of the population performed weekly rapid tests in March on average, new
+infections on April~1 would have been reduced by 53\% relative to the scenario without
+vaccinations, rapid tests, or seasonality. In Supplementary Material~B.12, we provide a
+detailed analysis of whether our results are robust regarding the sensitivity parameters
+we assume for rapid tests. Even if we take a pessimistic stance, the effect is only
+reduced from 42\% to 38\%.
 
 So why is rapid testing so effective? In order to shed more light on this question,
-Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
-scenario without rapid tests and the main specification into the three channels for
-rapid tests. Tests at schools have the smallest effect, which is largely explained by
-schools not operating at full capacity during our period of study and the relatively
-small number of students.\footnote{18\% of our population are in the education sector
-    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
-come from tests at the workplace. Despite the fact that rapid tests for private reasons
-are phased in only in mid-March, they make up for more than half of the total effect.
-The reason lies in the fact that a substantial share of these tests is driven by an
-elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
-up on a positive test of a household member. The latter is essentially a form of contact
-tracing, which has been shown to be very effective \citep{Contreras2021,
-    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
-Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
-administered randomly in the population would not have been nearly as effective.
+Figure~3d decomposes the difference in the scenario without rapid tests and the main
+specification into the three channels for rapid tests. Tests at schools have the
+smallest effect, which is largely explained by schools not operating at full capacity
+during our period of study and the relatively small number of students. (18\% of our
+population are in the education sector, e.g., pupils, teachers; 46\% are workers outside
+the education sector.) Almost 40\% come from tests at the workplace. Despite the fact
+that rapid tests for private reasons are phased in only in mid-March, they make up for
+more than half of the total effect. The reason lies in the fact that a substantial share
+of these tests is driven by an elevated probability to carry the virus, i.e., showing
+symptoms of CoViD-19 or following up on a positive test of a household member. The
+latter is essentially a form of contact tracing, which has been shown to be very
+effective \cite{Contreras2021, Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in
+Supplementary Material~B.15 shows that the same amount of rapid tests administered
+randomly in the population would not have been nearly as effective.
 
-\section{Discussion and conclusions}
+\section*{Discussion and conclusions}
 
@@
 scenarios comparing changes in NPIs and testing regimes. Two of the most contentious
-NPIs concern schools and mandates to work from home.\label{ref.2.a} In many countries,
-schools switched to remote instruction during the first wave, so did Germany. After the
-summer break, they were operating at full capacity with increased hygiene measures,
-before being closed again from mid-December onward. Some states started opening them
-gradually in late February, but operation at normal capacity did not resume until the
-beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
-policies regarding schools starting after Easter, at which point rapid tests had become
-widely available. We estimate the realized scenario to have essentially the same effect
-as a situation with closed schools. Under fully opened schools with mandatory tests,
-total infections would have been 6\% higher; this number rises to 20\% without tests.
-These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
-    Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of
-the large negative effects school closures have on children and parents
-\citep{Luijten2021, Melegari2021}---and in particular on those with low socio-economic
-status---these results in conjunction with hindsight bias suggest that opening schools
-combined with a testing strategy would have been beneficial. In other situations, and
-particular when rapid test are not available at scale, trade-offs may well be different.
+NPIs concern schools and mandates to work from home. In many countries, schools switched
+to remote instruction during the first wave, so did Germany. After the summer break,
+they were operating at full capacity with increased hygiene measures, before being
+closed again from mid-December onward. Some states started opening them gradually in
+late February, but operation at normal capacity did not resume until the beginning of
+June. Figure~4a shows the effects of different policies regarding schools starting after
+Easter, at which point rapid tests had become widely available. We estimate the realized
+scenario to have essentially the same effect as a situation with closed schools. Under
+fully opened schools with mandatory tests, total infections would have been 6\% higher;
+this number rises to 20\% without tests. These effect sizes are broadly in line with
+empirical studies (e.g. \cite{Vlachos2021, Berger2021}, see Supplementary Material~B.11
+for a comparison). In light of the large negative effects school closures have on
+children and parents \cite{Luijten2021, Melegari2021}---and in particular on those with
+low socio-economic status---these results in conjunction with hindsight bias suggest
+that opening schools combined with a testing strategy would have been beneficial. In
+other situations, and particular when rapid test are not available at scale, trade-offs
+may well be different.
 
-\begin{figure}[!tp]
+\begin{figure}[tb]
     \centering
-
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
-        \caption{{Effects of different schooling scenarios}}
-        \label{fig:school_scenarios}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.425\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
-        \caption{{Effects of different work scenarios}}
-        \label{fig:workplace_scenarios}
-    \end{subfigure}
-    \vskip3ex
-
-    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
+    \includegraphics[width=\textwidth]{../../bld/panel_figures/fig-school-workplace-scenarios.pdf}
+    \caption{
+        Effects of different scenarios for policies regarding schools and workplaces.
+        Blue lines in both figures refer to our baseline scenario; they are the same as
+        in Figure~3b. Interventions start at Easter because there were no capacity
+        constraints for rapid tests afterwards. For legibility reasons, all lines are
+        rolling 7-day averages.
+    }
     \label{fig:school_workplace_scenarios}
-
-    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
-        scenario; they are the same as in
-        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
-        because there were no capacity constraints for rapid tests afterwards. For
-        legibility reasons, all lines are rolling 7-day averages.}
-
 \end{figure}
 
-Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
+Figure~4b shows that with a large fraction of workers
 receiving tests, testing at the workplace has larger effects than mandating employees to
@@
 at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
-estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
+estimated to have cost around 50~Euros per capita \cite{Wollmershauser2021}; retail
 prices for rapid tests were below one Euro in early June 2021 and below five Euros for
@@
 important because disadvantaged groups are less likely to be reached by testing
-campaigns relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the
+campaigns relying on voluntary participation (e.g. \cite{StillmanTonin2021}); at the
 same time, these groups have a higher risk to contract CoViD-19
-\citep{KochInstitut2021a}. Mandatory tests at school and at the workplace will extend
+\cite{KochInstitut2021a}. Mandatory tests at school and at the workplace will extend
 more into these groups. The same goes for individuals who exhibit a low level of
@@
 contain the pandemic in the short run.
+
+\section*{Data Availability}
+
+The source code for the epidemiological model is publicly available in this Github
+repository: https://github.com/covid-19-impact-lab/sid. The source code of the research
+project and the data is available in this repository:
+https://github.com/covid-19-impact-lab/sid-germany.
+
+\bibliography{references}
+
+\section*{Acknowledgements (not compulsory)}
+
+The authors are grateful for support by the Deutsche Forschungsgemeinschaft (DFG, German
+Research Foundation) under Germany´s Excellence Strategy – EXC 2126/1– 390838866 – and
+through CRC-TR 224 (Projects A02 and C01), by the IZA Institute of Labor Economics, and
+by the Google Cloud CoViD-19 research credits program.
+
+\section*{Author contributions statement}
+
+Must include all authors, identified by initials, for example: A.A. conceived the
+experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the
+results.  All authors reviewed the manuscript.
+
+\section*{Additional information}
+
+To include, in this order: \textbf{Accession codes} (where applicable);
+
+\textbf{Competing interests}: The authors declare no competing interests.
+
+\end{document}

### 2022-03-20 src/panel_figures/task_create_panel_figures.py

@@ -1,0 +1,50 @@
+"""Create single figures from multi-panel figures.
+
+For publications multi-panel figures are usually not allowed since necessary packages
+are not supported, etc..
+
+https://texblog.org/2015/10/07/combining-sub-figures-to-a-single-figure-for-submission-to-journal/
+
+"""
+
+import pytask
+from pathlib import Path
+
+
+BLD = Path(__file__).parents[2].joinpath("bld").resolve()
+SRC = Path(__file__).parents[1].resolve()
+
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
+@pytask.mark.parametrize(
+    "depends_on, produces",
+    [
+        (
+            SRC / "panel_figures" / "model_description.tex",
+            BLD / "panel_figures" / "fig-model-description.pdf",
+        ),
+        (
+            SRC / "panel_figures" / "evolution_pandemic.tex",
+            BLD / "panel_figures" / "fig-evolution-pandemic.pdf",
+        ),
+        (
+            SRC / "panel_figures" / "2021_scenarios.tex",
+            BLD / "panel_figures" / "fig-2021-scenarios.pdf",
+        ),
+        (
+            SRC / "panel_figures" / "school_workplace_scenarios.tex",
+            BLD / "panel_figures" / "fig-school-workplace-scenarios.pdf",
+        ),
+    ],
+)
+def task_create_panel_figures():
+    pass

### 2022-03-20 task_compile_documents.py

@@ -113,7 +113,7 @@ def task_copy_reply_to_root(depends_on, produces):
 )
 @pytask.mark.depends_on(SRC / "cover-letter.tex")
 @pytask.mark.produces(BLD / "cover-letter.pdf")
-def task_compile_reply():
+def task_compile_cover_letter():
     pass
 
 
@@ -122,3 +122,19 @@ def task_compile_reply():
 def task_copy_cover_letter_to_root(depends_on, produces):
     shutil.copy(depends_on, produces)
 
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+        "-e", "$bibtex_fudge=1"
+    ]
+)
+@pytask.mark.depends_on(SRC / "scientific_reports_submission" / "article.tex")
+@pytask.mark.produces(BLD / "scientific_reports_submission" / "article.pdf")
+def task_compile_scientific_reports_submission():
+    pass

### 2022-03-21 src/scientific_reports_submission/article.tex
Remove paper toc from supplmat toc, fix references in article.

@@
 
-\author[1, 2, +]{Janoś Gabler}
-\author[3, +]{Tobias Raabe}
-\author[1, +]{Klara Röhrl}
-\author[2, 4, *, +]{Hans-Martin von Gaudecker}
+\author[1, 2]{Janoś Gabler}
+\author[3]{Tobias Raabe}
+\author[1]{Klara Röhrl}
+\author[2, 4, *]{Hans-Martin von Gaudecker}
 
@@
 \affil[*]{hmgaudecker@uni-bonn.de}
-
-\affil[+]{These authors contributed equally to this work}
 
@@
 to reduce contacts between individuals. While this has allowed some countries to sustain
-equilibria with very low infection numbers (see \cite{Contreras2021} for a theoretical
-equilibrium at low case numbers which is sustained with test-trace-and-isolate
-policies), most have seen large fluctuations of infection rates over time. Containment
-measures have become increasingly diverse and now include rapid testing, more nuanced
-NPIs, and contact tracing. Neither these policies' effects nor the influence of seasonal
-patterns or of more infectious virus strains are well understood in quantitative terms.
+equilibria with very low infection numbers\textcolor{red}{---see \cite{Contreras2021}
+for a theoretical equilibrium at low case numbers which is sustained with
+test-trace-and-isolate policies---}, most have seen large fluctuations of infection
+rates over time. Containment measures have become increasingly diverse and now include
+rapid testing, more nuanced NPIs, and contact tracing. Neither these policies' effects
+nor the influence of seasonal patterns or of more infectious virus strains are well
+understood in quantitative terms.
 
@@
 
-At the core of our agent-based model---we review more literature in Supplementary
-Material~B.1 \cite{Aleta2020,Hinch2021a}---are physical contacts between heterogeneous
+At the core of our agent-based model\cite{Aleta2020,Hinch2021a}---we review more
+literature in Supplementary Material~B.1---are physical contacts between heterogeneous
 agents (Figure~1a). Each contact between an infectious individual and somebody
@@
 showing symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms,
-requiring intensive care, and recovery or death similar to \cite{Aleta2020}. The
+requiring intensive care, and recovery or death (similar to \cite{Aleta2020}). The
 probabilities of transitioning between these states depend on age; their duration is
@@
 the test relative to the onset of infectiousness as in \cite{Smith2021}. We analyse
-robustness to different assumptions in Supplementary Material~A.12. After a phase-in
+robustness to different assumptions in Supplementary Material~B.12. After a phase-in
 period, all tests that are demanded will be performed. Figure~1c shows our model for
@@
 agent with an individual compliance parameter. This parameter determines whether she
-takes up rapid tests. (Positive test results or symptoms lead most individuals to reduce
-their contacts; this is why tests impact the actual contacts in Figure~1.)
+takes up rapid tests. \textcolor{red}{(Positive test results or symptoms lead most
+individuals to reduce their contacts; this is why tests impact the actual contacts in
+Figure~1.)}
 
@@
 that we can use a wide array of data to identify and calibrate the model's many
-parameters (see Supplementary Material~A for a complete description). Contact diaries
-yield pre-pandemic distributions of contacts for different contact types and their
-assortativity by age group. Mobility data is used to model the evolution of work
-contacts. School and daycare policies can be incorporated directly from official
+parameters \textcolor{red}{(see Supplementary Material~A for a complete description)}.
+Contact diaries yield pre-pandemic distributions of contacts for different contact types
+and their assortativity by age group. Mobility data is used to model the evolution of
+work contacts. School and daycare policies can be incorporated directly from official
 directives. Administrative records on the number of tests, vaccinations by age and
@@
 The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
-covering the period mid-September 2020 to the end of May 2021. Figure~2a describes the
+covering the period mid-September 2020 to the end of May 2021. Figure~2 describes the
 evolution of the pandemic and of its drivers. The black line in Figure~2a shows
@@
 is associated with the B.1.1.7 (Alpha) strain, which became dominant in March
-(Figure~2c)---B.1.617.2 (Delta) was first detected in Germany in April; at the end of
-our simulation period it accounted for less than 5\% of cases. In early March, some NPIs
-were relaxed; e.g., hairdressers and home improvement stores were allowed to open again
-to the public. There were many changes in details of regulations afterwards, but they
-did not change the overall stringency index.
+(Figure~2c). \textcolor{red}{B.1.617.2 (Delta) was first detected in Germany in April;
+at the end of our simulation period it accounted for less than 5\% of cases.} In early
+March, some NPIs were relaxed; e.g., hairdressers and home improvement stores were
+allowed to open again to the public. There were many changes in details of regulations
+afterwards, but they did not change the overall stringency index.
 
@@
 smallest effect, which is largely explained by schools not operating at full capacity
-during our period of study and the relatively small number of students. (18\% of our
-population are in the education sector, e.g., pupils, teachers; 46\% are workers outside
-the education sector.) Almost 40\% come from tests at the workplace. Despite the fact
-that rapid tests for private reasons are phased in only in mid-March, they make up for
-more than half of the total effect. The reason lies in the fact that a substantial share
-of these tests is driven by an elevated probability to carry the virus, i.e., showing
-symptoms of CoViD-19 or following up on a positive test of a household member. The
-latter is essentially a form of contact tracing, which has been shown to be very
-effective \cite{Contreras2021, Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in
-Supplementary Material~B.15 shows that the same amount of rapid tests administered
-randomly in the population would not have been nearly as effective.
+during our period of study and the relatively small number of students.
+\textcolor{red}{18\% of our population are in the education sector, e.g., pupils,
+teachers; 46\% are workers outside the education sector.} Almost 40\% come from tests at
+the workplace. Despite the fact that rapid tests for private reasons are phased in only
+in mid-March, they make up for more than half of the total effect. The reason lies in
+the fact that a substantial share of these tests is driven by an elevated probability to
+carry the virus, i.e., showing symptoms of CoViD-19 or following up on a positive test
+of a household member. The latter is essentially a form of contact tracing, which has
+been shown to be very effective \cite{Contreras2021, Fetzer2021,Kretzschmar2020}.
+Indeed, a deeper analysis in Supplementary Material~B.15 shows that the same amount of
+rapid tests administered randomly in the population would not have been nearly as
+effective.
 
@@
 
-\section*{Acknowledgements (not compulsory)}
+\section*{Acknowledgements}
 
@@
 
-To include, in this order: \textbf{Accession codes} (where applicable);
-
 \textbf{Competing interests}: The authors declare no competing interests.

### 2022-04-20 src/rev2_reply_comments.tex
Letter / rebuttal second round.

--- 2022-03-11 src/rev1_reply_comments.tex
@@
 
-\large{\textbf{\textcolor{red}{SUMMARY}}}
-
-\section*{Editors}
-
-\comment{ The title should have a clear, precise scientific meaning and should not
-contain a colon. Where possible, the title should be read as one concise sentence. }
-\vskip1ex \response{%
-    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2:
-    Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Testing, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic''
-}
+We would like to thank the editors and reviewers for their work. We comment briefly on the points made by Reviewer 1.
 
 \section*{Reviewer 1}
-\comment{The authors of the manuscript presented here have used mathematical models to
-examine the impact of different measures to contain the SARS-CoV-2 pandemic. Using
-epidemiological data from Germany, the contribution of rapid antigen testing is
-considered in particular.
+\comment{
+In the manuscript presented here, the authors describe model calculations on the quantitative influence of non-pharmaceutical measures, vaccination and rapid testing campaigns on the containment of the spread of SARS-CoV-2 in Germany.
+They also consider factors such as seasonality or virus variants with different transmission profiles.
+Modeling of the kind described here has now been carried out and published in large numbers by other groups.
+Unfortunately, the authors here refer exclusively to calculations of the spread of the SARS-CoV-2 alpha variant in Germany.
+In contrast, the much more extensive spread of the delta variant and the even more significant omicron spreads are not considered.
 
-The topic studied here is of great importance for the control of the coronavirus
-pandemic. The presentation of the scientific methodology is sufficiently detailed,
-especially with the supplementary materials. It allows the reader to follow the research
-process in detail. However, I noticed a significant deficiency when reviewing the
-submitted work. Therefore, the manuscript should definitely be revised again before
-publication: A crucial point of the authors is to point out the special benefit of a
-rapid test strategy. However, for the calculation of the corresponding influencing
-factors in their mathematical model, the authors simplify, specifically, the sensitivity
-of rapid antigen tests in an unacceptable way. In the development of the mathematical
-model, apparently only the knowledge available a few months ago on the performance of
-rapid antigen tests was taken into account. This is evident from literature references
-mentioned above. In part, incorrect references are also used, especially for the
-assumption of the high sensitivity of the rapid tests. For example, reference No. 74
-explicitly does not refer to rapid antigen tests, but to an immunological laboratory
-test method (Quidel SARS Sofia antigen FIA). Especially in light of the benefit of rapid
-antigen tests described by the authors, it should be considered that many of the
-commercially marketed tests show only extremely poor performance. This has recently been
-demonstrated by several publications:
-\begin{itemize}
-    \item Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic tests
-          for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer,
-          Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and
-          Corman, Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange,
-          Constanze and Emmerich, Petra and Müller, Michael and Knauer, Olivia and
-          Nübling, C Micha, Eurosurveillance, 26, 2100441 (2021),
-          \url{https://doi.org/10.2807/1560-7917.ES.2021.26.44.2100441}
-    \item Assessment of SARS-CoV-2 rapid antigen tests, Özcürümez, Mustafa, Katsounas,
-          Antonios, Holdenrieder, Stefan, von Meyer, Alexander, Renz, Harald and Wölfel,
-          Roman, Journal of Laboratory Medicine, vol. 45, no. 3, 2021, pp. 143-148.
-          \url{https://doi.org/10.1515/labmed-2021-0036}
-\end{itemize}
+This is a major weakness of this manuscript.
+Although the descriptions of the methods used seem adequate from my perspective, the analysis adds very little value.
+Expanding the manuscript to include data on the delta and omicron variants would greatly enhance its significance.
 
-Especially in the case of a mass application of antigen tests, it is therefore to be
-expected that products with lowest sensitivity will also be used in the general
-population. The authors should at least model this with an alternative calculation and
-discuss it accordingly. Such an improvement would increase the importance of this
-manuscript enormously.
-
+Despite this considerable weakness, I see no obstacles to publication, provided that the authors nevertheless wish to share the limited significance of their study with the scientific community.
 \vskip1ex
@@
 \response{%
-    We would like to thank the reviewer for these insightful comments. First of all, we
-    agree that the references we used have become outdated in the meantime. Note that we
-    submitted the original manuscript in early September 2021; the analysis had been
-    finalised in June of last year. Indeed, in the line of the additional evidence, the
-    numbers we use for test sensitivity appear to be in the upper part of plausible
-    values. We have followed your suggestion and report detailed robustness analyses in
-    the paper and its Appendix.
+    We would like to thank the Reviewer for his/her comments. We would like to make a couple of points:
 
-    Before describing our changes, we would like to thank you for pointing out that the
-    test analysed in \citet{Smith2021} is not a rapid test. While this is true, the
-    crucial feature in \citet{Smith2021} is that the paper reports the sensitivity over
-    the course of an infection. We have not found a comparable study. Getting this
-    profile right is important particularly in the beginning of an infection, as we
-    detail shortly. At that point, the differences between different methods to assess
-    test sensitivity by timing within an infection are relatively small.
+    \begin{itemize}
+        \item The manuscript was finalized in June 2021 and submitted to \textit{Scientific Reports} in August 2021.
+At that time, $\delta$ played a minor role in Germany and omicron was not known yet.
+        \item The time period includes the wild type and the $\alpha$ variant. The framework for adding different virus types is there.
+        \item In our view, our scholarship lies in the construction of the model and estimating its parameters using available data, thereby demonstrating the effectiveness of different strategies at a particular point in time.
+The latter may well be different for other types and omicron in particular, where vaccines loose much of their power regarding infection dynamics.
+        \item Extending the analysis in such a way would be a straightforward task with our model: The code is all open source, well documented, and available free of charge.
+All features required to extend the period of analysis to further variants are included in the model.
+Anyone who wants to collect the additional data and estimate the model is free to do so.
+    \end{itemize}
 
-    On page~\pageref{r.1.a} of the paper, we cite the references you pointed out
-    (numbering of references relates to this document) and note that we conduct a
-    robustness analysis:\\[1ex]
-
-    \textit{\Paste{r.1.a}}
-
-    \vskip1ex
-
-    We followed your suggestion and ran additional analyses with more conservative
-    assumptions on the sensitivity of rapid tests. The results are very robust. We write
-    in the main text on page~\pageref{r.1.b}:\\[1ex]
-
-    \textit{\Paste{r.1.b}}
-
-    \vskip1ex
-
-    The robustness is due to the fact that over the course of an infection, the largest
-    differences between the parameters we use in our main analysis and those based on
-    imputing values from other studies are found during the later stages of an
-    infection. Uncovering an infection that was previously undetected at that stage does
-    not have a large effect on subsequent infection dynamics. Our whole analysis in
-    Appendix~\ref{subsec:robustness_rapid_test_sensitivity} reads as follows:\\[1ex]
-
-    \textit{\Paste{r.1.b-appendix}}
-
-}
-
-~\\[4ex]
-
-\FloatBarrier
-
-Furthermore, I recommend including a declaration of potential conflicts of interest in
-the manuscript. This may be particularly important in light of the significance of the
-findings from the study for the continued operation of industrial companies.\\[2ex]
-
-\response{We have added such a declaration at the end of the paper:\\[2ex]
-
-    \textit{\Paste{r.1.c}}
+    In the light of these points, we have refrained from re-estimating the model. We have not changed the manuscript.
 }
@@
 }
-
-\clearpage
-\section*{Reviewer 2}
-
-\comment{ This is a comprehensive and important study that highlights the relevance of
-testing to stop the spread of the pandemic.
-
-I cannot comment on the validity of the underlaying mathematical analysis, but I would
-like to comment on some assumptions made, which I think require further clarification or
-discussion.
-
-\begin{enumerate}
-    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful
-          if behavior of people is already changed by the fact there is a pandemic?
-
-          \response{%
-
-              Many thanks for this comment. Ideally, we would of course use real-time
-              data. For example, several studies have used smartphone tracking data for
-              this purpose. However, any such data are not detailed enough for our
-              requirements because they do not provide any information about
-              \textit{who} is meeting with each other. These dimensions make up our
-              contact networks, which are absolutely crucial.
-
-              Furthermore, note that we only use the pre-pandemic diary data as the
-              starting point. Contact restrictions and individual behavior adjustments
-              reduce contacts. These are active throughout the period of analysis and
-              they vary with the strictness of measures. In order to do so, it is crucial
-              to take the type of contact into account; see for example our analysis of
-              potential school closures on Page~\pageref{ref.2.a}. Closed schools mean
-              that children have far fewer and parents have somewhat fewer contacts than
-              without school closures.
-
-              Finally, the very good fit of our model to official infection rates across
-              age groups increases our confidence that we manage to match actual meeting
-              patterns. If we were far off for, for example, the number of work
-              contacts, we would not be able to match either the infections of
-              individuals during working age (35-59) or of those that mostly do not work
-              anymore (60-79).
-          }
-    \item They state that susceptibility is dependent on age. This requires references
-          and needs to be further explained. Also, younger populations are highly susceptible,
-          they may just not contribute transmission as much. On that same note, infectiousness
-          may not be independent of age
-
-          Edward Goldstein, Marc Lipsitch, Muge Cevik, On the Effect of Age on the
-          Transmission of SARS-CoV-2 in Households, Schools, and the Community, The
-          Journal of Infectious Diseases, Volume 223, Issue 3, 1 February 2021, Pages
-          362–369, \url{https://doi.org/10.1093/infdis/jiaa691}
-
-          \response{%
-              Many thanks for pointing this out, indeed we had not cited our key
-              reference for susceptibilities by age group \citep{Davies2020} in the main
-              text. In fact, the paper you suggested \citep{Goldstein2021} comes to
-              similar conclusions: \textit{``We found evidence that compared to
-                  younger/middle-aged adults, children younger than 10 years have
-                  significantly lower estimated susceptibility to SARS-CoV-2 infection,
-                  while adults older than 60 years have elevated susceptibility to
-                  infection.''} The current version cites both papers on Page~\ref{r.2.b}.
-
-              We base our assumption on infectiousness being independent of age on
-              \citet{Jones2021}. That study was the only one we could find that was able
-              to differentiate between age differences in susceptibility (given a risk
-              contact) vs. age differences in contact patterns. Moreover, they had fine
-              grained age bins. That paper reports only slight differences in viral
-              loads between German adults and children: ``The least infectious youngest
-              children have 78\% (61, 94) of the peak culture probability of adults aged
-              45 to 55''. Furthermore, they caution that these lower numbers may be due
-              to different sample taking---such as the usage of smaller pediatric swab
-              devices---rather than actual lower viral loads and conclude that the
-              differences, if existent, are likely not clinically relevant. In fact,
-              \citet{Goldstein2021}  reach similar conclusions, writing: \textit{``There
-                  is limited evidence in the literature regarding age-related differences in
-                  infectivity, although point estimates in several studies suggest that
-                  infectivity may increase somewhat with age.''} We are thus confident that
-              different assumptions in the range of plausible values would not
-              change our results in a quantitatively meaningful way.
-
-          }
-    \item It is not accurate that the vaccines stop transmission. This has to be adapted
-          accordingly. Also, the authors should state which vaccines they based this
-          analysis on, as in which ones were in use in Germany at that time and how does
-          that affect the relative transmissibility as it cannot be assumed that all
-          vaccines will have the same effect.
-
-          \response{%
-              Many thanks, it is of course true that vaccines do not stop transmission
-              completely. We do not assume that, either. However, we have not been able
-              to find data on the way in which this affects different population groups
-              for the wild type and the $\alpha$-variant.
-
-              In our model, we thus simplify the way vaccines reduce transmission risk.
-              In particular, we assume that that 75\% of individuals achieve sterile
-              immunity through a vaccination while the remaining 25\% remain fully
-              susceptible. Firstly, the existing evidence supports the
-              assumption that all four vaccines approved in the EU (Biontech/Pfizer,
-              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
-              protection in the first few months after vaccination against the wild type
-              and the $\alpha$ variant (for details see Appendix~\ref{r.2.c}, starting
-              on Page~\pageref{r.2.c}), which is the only time period and variant our
-              simulations cover. Secondly, given the surprisingly good performance of
-              rapid tests we tried to err on the side of being too optimistic with our
-              modelling of vaccine performance. This is why we chose the perfect
-              protection from vaccination for a large part of the population. Thirdly,
-              reinfections are not a big issue for the time period that we look at. The
-              incidences were small and the chance of more than one exposure to Covid
-              are negligibly small. Thus, the value added by a probabilistic model of
-              vaccination protection is very small for the period that we cover. }
-\end{enumerate}
-}
+~\\[4ex]
 

### 2022-04-20 task_compile_documents.py

@@ -111,6 +111,28 @@ def task_copy_reply_to_root(depends_on, produces):
         "-f",
     ]
 )
+@pytask.mark.depends_on([SRC / "rev2_reply_comments.tex", BLD / "paper.pdf"])
+@pytask.mark.produces(BLD / "rev2_reply_comments.pdf")
+def task_compile_reply2():
+    pass
+
+
+@pytask.mark.depends_on(BLD / "rev2_reply_comments.pdf")
+@pytask.mark.produces(ROOT / "rev2_reply_comments.pdf")
+def task_copy_reply2_to_root(depends_on, produces):
+    shutil.copy(depends_on, produces)
+
+
+@pytask.mark.latex(
+    [
+        "--pdf",
+        "--interaction=nonstopmode",
+        "--synctex=1",
+        "--cd",
+        "--shell-escape",
+        "-f",
+    ]
+)
 @pytask.mark.depends_on(SRC / "cover-letter.tex")
 @pytask.mark.produces(BLD / "cover-letter.pdf")
 def task_compile_cover_letter():
@@ -131,7 +153,8 @@ def task_copy_cover_letter_to_root(depends_on, produces):
         "--cd",
         "--shell-escape",
         "-f",
-        "-e", "$bibtex_fudge=1"
+        "-e",
+        "$bibtex_fudge=1",
     ]
 )
 @pytask.mark.depends_on(SRC / "scientific_reports_submission" / "article.tex")
